### 2020-07-20 paper.tex
Before cleaning up

%-----------------------------------------------------------------------
% Beginning of article-template.tex
%-----------------------------------------------------------------------
%
%    This is a template file for proceedings articles prepared with AMS
%    author packages, for use with AMS-LaTeX.
%
%    Templates for various common text, math and figure elements are
%    given following the \end{document} line.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%    Remove any commented or uncommented macros you do not use.

%    Replace amsproc by the name of the author package.
\documentclass{amsproc}

%    If you need symbols beyond the basic set, uncomment this command.
%\usepackage{amssymb}

%    If your article includes graphics, uncomment this command.
%\usepackage{graphicx}

%    If the article includes commutative diagrams, ...
%\usepackage[cmtip,all]{xy}

%    Include other referenced packages here.
\input{definitions}

%    Update the information and uncomment if AMS is not the copyright
%    holder.
%\copyrightinfo{2009}{American Mathematical Society}

%% \newtheorem{theorem}{Theorem}[section]
%% \newtheorem{lemma}[theorem]{Lemma}

%% \theoremstyle{definition}
%% \newtheorem{definition}[theorem]{Definition}
%% \newtheorem{example}[theorem]{Example}
%% \newtheorem{xca}[theorem]{Exercise}

%% \theoremstyle{remark}
%% \newtheorem{remark}[theorem]{Remark}

\numberwithin{equation}{section}

\begin{document}

% \title[short text for running head]{full title}
\title{A tractable model for D-optimal design in infinite dimensional
  Bayesian inverse problems.}

%    Only \author and \address are required; other information is
%    optional.  Remove any unused author tags.

%    author one information
% \author[short version for running head]{name for top of paper}
\author{Yair Daon}
\address{Courant Institute}
%% \curraddr{}
%% \email{yair.daon@gmail.com}
%% \thanks{}

%    The 2010 edition of the Mathematics Subject Classification is
%    the current definitive version.
\subjclass{Primary: 
  62F15, % Statistics - bayesian inference
  35R30, % PDE - Inverse problems
  65C50; % Numerical analysis - other computational problems in probability
  Secondary:
  28C20, % Measure and integration - set functions and measures and integrals in
  %infinite-dimensional spaces (Wiener measure, Gaussian measure, etc.
}

\date{\today}

\begin{abstract}
  We investigate the problem of sensor clusterization in optimal
  experimental design for infinite-dimensional Bayesian inverse
  problems. We suggest an analytically tractable model for such
  designs and reason how it may lead to sensor clusterization in the
  case of iid measurement noise. We also show that in the case of
  spatially correlated measurement error clusterization does not
  occur. Our analysis suggests a novel method for finding optimal
  designs. As a part of the analysis we prove a matrix determinant
  lemma analog in infinite dimensions, as well as a lemma for
  calculating derivatives of $\log \det$ of operators.
\end{abstract}

\maketitle

\section{Introduction}\label{section:OED intro}
Experimental design is an important part of many scientific
investigations. When considering an inverse problem, one can often
specify sensor locations (e.g.\ in geophysics and oceanography
applications), certain wavelengths (e.g.\ in MRI) or wave reflections
from the ground (e.g.\ searching for oil or using a radar). Whatever
the allowed set of measurements is, one should select the optimal
measurements to take, in order to increase accuracy, reduce costs, or
both.

Designing experiments is usually done by optimizing some \emph{design
  criterion} and this is true both for frequentists
\cite{Silvey13,Ucinski05} as well as for Bayesians
\cite{ChalonerVerdinelli95}, and the analogy between the two
approaches was investigated in \cite{ChalonerVerdinelli95}. Although
there is a plethora of design criteria, in the current work we will
focus on just one of these, commonly referred to as \emph{D-optimal
  design}. It has a simple and appealing motivation in the Bayesian
context as explained in \cite{ChalonerVerdinelli95}: for a linear
model, a D-optimal design seeks to maximize the expected information
gain (KL divergence \cite{KullbackLeibler51,CoverThomas12}) between
posterior and prior. In finite dimensions, maximizing the expected
information gain amounts to minimizing the determinant of the
posterior covariance matrix. In a frequentist settings, an optimal
design minimizes the volume of the uncertainty ellipsoid \cite[page
  16]{Ucinski05}, but this is done for the Fisher information matrix
and not the posterior covariance. However, \cite{ChalonerVerdinelli95}
shows that the latter is just a regularized version of the former.

The previous discussion is classical for experimental design when
inference is to take place over a finite (not too large) number of
parameters. The subject of optimal experimental design for function
inference in a Bayesian context was pioneered by
\cite{AlexanderianGloorGhattas14,AlexanderianPetraStadlerEtAl16,
  AlexanderianPetraStadlerEtAl14}. Similarly to the finite dimensional
case, it can be shown that a D-optimal design arises naturally for
linear models when one wishes to maximize the KL divergence between
posterior and prior and this amounts to minimizing the determinant of
the posterior covariance operator (understood as a product of its
eigenvalues). Some difficulties arise in the process, but remedies can
be found as shown in \cite{AlexanderianGloorGhattas14}.

It seems counter intuitive that when one computes an optimal design
using the D-optimal criterion (and others that will not be
investigated here), the optimization process results in measurement
locations are very similar. Specifically, if a measurement is thought
of as measuring some function value at $\x \in \Omega \subseteq \R^d,
d=1,2,3$ (with added error) then the optimization procedure sometimes
places sensors in close proximity to each other (as can be seen in
figure \ref{fig:clusterization illustration}). Following
\cite{Ucinski05}, we refer to this phenomenon as \emph{sensor
  clusterization}.

\subsection{Related Work}
The phenomenon of sensor clusterization seems to be known in several
different contexts. In a frequentist and finite-dimensional context,
\cite{Fedorov96} and \cite[chapter 2.4.3]{Ucinski05} discuss this
phenomenon and suggest an approach called clusterization-free design,
where the user enforces measurement locations to be far from each
other. One way to do this is by introducing correlated error which,
philosophically, accounts for both measurement error and model error.
Another method considered is imposing distance constraints between
measurements. A somewhat different approach is suggested in \cite[page
  49]{Fedorov12}, where close by design measurements are merged --- a
procedure which obviously does not avoid clusterization. The same problem
arises in time-series analysis for pharmacokinetic experiments. The
authors of \cite{Hooker09} suggest modeling auto-correlation time in
the noise model, which is equivalent to the correlated errors
mentioned above.

Any of the above mentioned approaches might serve as a remedy and push
sensors away from each other. Yet, none offers any insight as to why
clusterization occurs. Also, as better models are employed, the model
error decreases and the clusterization phenomenon will reappear. While
these approaches are practical and help us avoid the problem, they do
not provide insight as to why sensors are clustering.

\subsection{Contribution}
In this work we propose and thoroughly study an analytically tractable
model for understanding D-optimal designs. The initial motivation was
to understand the causes of sensor-clusterization and the end result
is interesting in its own. Specifically, in section \ref{subsec:unit
  length} we suggest a somewhat simplified model for experimental
design that is analytically tractable. In section \ref{section:D and
  grad} we derive a constrained optimization problem for D-optimal
design using Lagrange multipliers. In the case of no model error we
show this constrained optimization problem is in fact a set of
eigenvector-eigenvalue problems \eqref{eq:eigenproblem} and suggest a
natural way for understanding sensor-clusterization in section
\ref{subsub:first attempt}. We explain why this approach fails and
show the reason is that (perhaps surprisingly), all eigenvalues are
equal --- this is done in section \ref{subsub:characterization}. An
important consequence is that we know the value of the best design
criterion without any optimization and which measurements give rise to
it. Practically, we can use this information to seek a design in the
real problem we are interested in and we outline this in section
\ref{section:conclusion}.

In the process we generalize several lemmas from linear algebra to
infinite-dimensional settings --- the Matrix Determinant Lemma and a
lemma for calculating the derivative of determinants. We also provide
other tools for understanding D-optimal designs in
infinite-dimensional Bayesian inverse problems, including lemmas for
calculating the increase in the design criterion per observation ---
Lemmas \ref{lemma:design increase} and Corollary \ref{cor:zero mod
  err}. We also show that in our model with independent measurement
noise, an interesting phenomenon arises, in which the eigenvectors of
the measurement operator (denoted below $\obs^*\obs$) are the same as
the eigenvectors of the prior covariance in observation space, which
helps arrive at the results outlined in the previous paragraph. The
arguments presented are completely generic and apply equally to any of
the scenarios mentioned above.

Based on the analysis presented, we show why clusterization is a
reasonable thing to expect in the case of independent measurements.
We further show how considering an observation model that is spatially
correlated makes an optimal design clusterization-free.

\subsection{Limitations}\label{subsec:limitations}
There are two main drawbacks of the arguments presented here. First,
one can argue that we allow considerably less restrictive observations
than any real-life problem does. This is because our observation model
allows taking measurements that are arbitrarily weighted linear
combinations of eigenvectors. The second drawback is that we do not
show rigorously that clusterization occurs in our model.

\subsection{Illustration and Example}\label{subsec:example}
\begin{figure}
  \begin{tikzpicture}[thick, scale=1.3, every node/.style={scale=0.99}]
    \begin{axis}
      [
      xmin = 0,
      xmax = 3.14,
      xlabel = {$x$},
      ylabel = posterior std,
      ymin   = 0,
      %compat = 1.3,
      % ymax   = 130,
      % ytick = \empty,
      legend cell align=left,
      % legend style={at={(0.45,0.2)}}
      legend pos= outer north east 
      ]
      % \draw[black!30!white, thin] (50,0) -- (50,130);
      % 
      \addplot [thin, black, mark=none] table{data/stdv-heat-sens1-var1.txt};
      \addlegendentry{1 sensors};

      \addplot [thin, blue, mark=none] table{data/stdv-heat-sens2-var1.txt};
      \addlegendentry{2 sensors};

      \addplot [thin, red, mark=none] table{data/stdv-heat-sens3-var1.txt};
      \addlegendentry{3 sensors};

      \addplot [thin, green, mark=none] table{data/stdv-heat-sens4-var1.txt};
      \addlegendentry{4 sensors};

      \addplot [thin, purple, mark=none] table{data/stdv-heat-sens5-var1.txt};
      \addlegendentry{5 sensors};

      \addplot [thin, cyan, mark=none] table{data/stdv-heat-sens6-var1.txt};
      \addlegendentry{6 sensors};

      \addplot [black,  only marks, mark=x, mark size=1.5] 
      table{data/locs-heat-sens1-var1.txt}; 
      \addplot [blue,   only marks, mark=x, mark size=1.5]
      table{data/locs-heat-sens2-var1.txt}; 
      \addplot [red,    only marks, mark=x, mark size=1.5]
      table{data/locs-heat-sens3-var1.txt};
      \addplot [green,  only marks, mark=x, mark size=1.5] 
      table{data/locs-heat-sens4-var1.txt}; 
      \addplot [purple, only marks, mark=x, mark size=1.5] 
      table{data/locs-heat-sens5-var1.txt}; 
      \addplot [cyan,   only marks, mark=x, mark size=1.5] 
      table{data/locs-heat-sens6-var1.txt};

    \end{axis}
  \end{tikzpicture}
  \caption{Shown are pointwise posterior standard deviations for
    various numbers of sensors, and location of these sensors for the
    1D heat equation with homogeneous Dirichlet boundary described in
    section \ref{subsec:example}. Note that the observation data does
    not influence the pointwise variances, so there is no point in
    showing the posterior mean.}
  \label{fig:clusterization illustration}
\end{figure}
In section \ref{section:abstract OED} we present a more abstract and
general formulation of the problem, but for the purpose of
illustration, we present the problem via a toy model --- the 1D heat
equation in $[0,\pi]$ with homogeneous Dirichlet boundary
conditions. The 1D heat equation is:
\begin{subequations}\label{eq:heat equation}
  \begin{alignat}{2}
    u_t &= \Delta u &&\qquad \text{in } [0,\pi] \times [0,\infty),\\
      u &= 0 &&\qquad \text{on } \{0, \pi\} \times [0,\infty),\\
        u &= u_0 &&\qquad \text{on }[0,\pi] \times \{0\}.
  \end{alignat}
\end{subequations}
We would like to infer the initial condition $u_0$. For that purpose,
we measure $u$ at some set of locations $\x_j \in [0,\pi], j=1,...,m$
and a final time $T > 0$. We assume centered Gaussian measurement
error, so we can observe $v(\x_j,T) = u(\x_j,T) + \eps(\x_j)$ with
$\eps(\x_j) \sim \normal(0, \sigma^2), \sigma > 0$ iid.  We model the
initial condition as $u_0 \sim \normal(0,\prcov)$, for $\prcov =
(-\Delta)^{-1}$ with homogeneous Dirichlet boundary condition. It is
well known \cite{Tarantola05} that for linear problems, with Gaussian
prior and error, the posterior is also Gaussian with a covariance that
does not depend on the observed data. The posterior covariance
$\postcov$ is known to have a closed form formula, even in infinite
dimensions\cite{Stuart10}. If we denote by $\fwd$ the dynamics
operator, so that $u( \cdot,T) = \fwd u_0$, and the observation
operator $\obs$ so that $u(\x_j,T) = (\obs u)_j, j=1,...,m$, then the
posterior covariance is known and depends on $\prcov, \fwd, \obs$ and
$\sigma^2$ (see section \ref{section:abstract OED} and
\eqref{eq:postcov} specifically).

We will consider the information-theoretic design criterion presented
in section \ref{section:OED intro}. Thus, we choose $\x_j,j=1,...,m$ so
as to minimize an expression analogous to the log-determinant of the
posterior covariance operator (corresponding to a D-optimal
design). We can see the clusterization effect in figure
\ref{fig:clusterization illustration}. For low numbers of sensors, the
design criterion suggests placing sensors in locations that are
close.

\subsection{Overview}\label{subsec:plan}
Half of this paper is devoted to finding the Lagrange multipliers
problem and culminates in \eqref{eq:conditions}. The reader may skip
directly there if they are willing to take the rest on faith. The more
interesting stuff may be found in section \ref{section:analysis}.

Seeking an optimal design amounts to maximizing $\tar(\obs)$ in
$\obs$, subject to constraints posed on $\obs$. These enter as
constraints on the allowed set of measurement vectors we can choose
from --- the rows of $\obs$. These constraints are determined by the
problem structure --- for example, in the 1D heat equation example
considered in section \ref{subsec:example}, we were constrained to
choose $\obs = (\meas_1,...,\meas_m)^t$ with $\meas_j = \delta_{\x_j}$
for some $\x_j \in [0,\pi]$. We present an approach to understanding
optimal experiment design based on the first-order necessary condition for
optimal design, where the constraints are incorporated using Lagrange
multipliers. This condition is used to understand what causes
clusterization. One problem with such an approach is that the set we
can choose $\obs$ from is complicated and difficult to work with. For
example, if $\ban = C(\Omega)$ for some compact $\Omega$, then the
allowed set of measurement vectors may only be point evaluations --- a
set that is not easy to describe as a level set of some simple
function, a description crucial for the use of Lagrange multipliers.
In order to facilitate analysis, we consider an easier constraint in
section \ref{subsec:unit length}. Reverting to this simpler constraint
is a limitation of our analysis but it provides some insight
into the cause of clusterization.

\section{Preliminaries}

The theoretical foundations for inverse problems over function spaces
can be found in \cite{Stuart10}. At this point, it is useful to
record a known result regarding the posterior covariance operator of
our inverse problem, namely
\begin{align}\label{eq:postcov precursor}
  \postcov = (\prcov^{-1} + \fwd^* \obs^* \Sigma^{-1} \obs \fwd )^{-1}.
\end{align}

\subsection{bayesian inverse problems}\label{section:abstract OED}
Let $\hil$ be a Hilbert space and $\ban$ a Banach space, both
infinite-dimensional. Consider $\fwd: \hil \to \ban$, a linear
operator (the ``forward operator''). We are interested in forward
operators that are strongly smoothing (have fast decaying singular
values --- the heat operator from section \ref{subsec:example} is a
prime example).

We want to infer $\param \in \hil$, some parameter of the dynamics
given noisy observations of $\fwd \param$. We model $\param \sim
\normal(0,\prcov)$ with some appropriate covariance operator $\prcov$
\cite{Stuart10}. Observations (measurements) are taken via the linear
observation operator $\obs \in ( \ban^* )^m$, where $m$ is the number
of observations allowed (e.g. sensors at our disposal). In an analogy
with linear algebra (where row vectors are thought of as linear
functionals), we think of $\obs$ as having ``rows'' $\meas_j,
j=1,...,m$:
\begin{equation}\label{eq:O}
  \obs = (\meas_1,..., \meas_m)^t, \meas_j \in \ban^*, j = 1,...,m.
\end{equation}
This way, for $u \in \ban$ we have $\obs u = (\meas_1(u), ...,
\meas_m(u) )^t \in \R^m$.
%% but we may drop the parentheses and write $\meas_j u = \meas_j(u)$.
A few observations regarding $\obs$ are in order. First, it is good to
keep in mind that $( \ban^* )^m$ is a Banach space with norm $\| \obs
\| = \sum_{j=1}^m \|\meas_j\|$. Next, for $u \in \ban$ and $v\in
\R^m$:
\begin{align*}
  \big (\obs^*v \big ) (u) &= \langle v, \obs u \rangle_{\R^m} = \sum_{j=1}^m  v_j \meas_j(u)
  = v^t \left ( \obs u \right ) = (v^t \obs) (u),
\end{align*}
and thus:
\begin{align}\label{eq:obs*}
  \obs^*v &= \sum_{j=1}^m v_j \meas_j = v^t \obs.
\end{align}

\begin{observation}
  It is best to think of $\meas_j,j=1,...,m$ as row vectors and of
  $\obs$ as a matrix with rows $\meas_j$. 
\end{observation}

Each measurement consists of a linear functional, chosen from an
allowed set of linear functionals in $\ban^{*}$. Data is acquired via
noisy measurements
\begin{align*}
  \data := \obs (\fwd \param + \eps') + \eps = \obs \fwd \param + \obs \eps' + \eps
\end{align*}
with $\data \in \R^m$ and $\eps, \eps'$ defined next. Measurement
error consists of two parts. First, there is spatially correlated
error $\eps' \sim \normal(0,\modcov)$, modeled as a centered Gaussian
measure on $\ban$ with covariance operator $\modcov: \ban^* \to \ban$
(see \cite[section 6]{Stuart10} for some details on Gaussian measures
on Banach spaces). Then there is measurement / independent error
$\eps \sim \normal(0, \sigma^2 I_m)$, with $I_m$ the $m \times m$
identity. Both error terms and the prior are assumed independent of
each other. Let us now understand the terms present in the above
expression.

\subsubsection{Dynamics and Observation Operators}\label{subsec:dynamics}
One can think of $(\obs \fwd \param)_j$ as pointwise evaluations of a
continuous function if $\ban$ is some appropriate function space,
e.g.\ $C(\Omega)$, the Banach space of real-valued continuous
functions over a compact $\Omega$ with the $\| \cdot \|_{\infty}$
norm. These measurements can be represented using the linear
functionals $\delta_{\x_j} \in C(\Omega)^*$. In a different setting,
we may be able to measure other quantities, e.g.\ some Fourier
coefficients as is the case in MRI applications. Either way, the rows
of $\obs$ are linear functionals and we refer to these as {\it
  measurement vector}s. Note that we cannot choose every measurement
vector we want, though. For example, we may be restricted only to
pointwise evaluations of $\fwd \param$ by the sensors at our disposal,
in which case measuring the mean $\ell(u) = \int_{\Omega}u$ (or any
non-atomic measure) of a function is not possible.

\subsubsection{Error Terms}
Considering the sum of the error terms it is easy to see that $\obs
\eps' + \eps \in \R^m$. It is a centered Gaussian random vector and
its covariance matrix is
\begin{align}\label{eq:Sigma}
  \begin{split}
    \Sigma(\obs) :&= \mathbb{E}[ (\obs \eps' + \eps)  (\obs \eps' + \eps)^t ] 
    % 
    % 
    = \obs \modcov \obs^* + \sigma^2I , 
  \end{split}
\end{align}
where
\begin{equation}\label{eq:modcov explained}
  [\obs \modcov \obs^*]_{ij} = e_i^t \obs \modcov \obs^* e_j = \meas_i (\modcov \meas_j),
\end{equation}
where the first equality in \ref{eq:modcov explained} holds by
definition and the second by \eqref{eq:obs*}. The explicit dependence
on $\obs$ will be mostly dropped for notational convenience, so
$\Sigma(\obs) = \Sigma$. Thus, for a fixed $\obs$ (i.e.\ a fixed set
of measurements) we can equivalently write $\data = \obs \fwd \param +
\bar{\eps}$ with $\bar{\eps} \sim \normal(0,\Sigma)$. At this point,
it is useful to rewrite the posterior covariance operator of our
inverse problem from \eqref{eq:postcov precursor}
\begin{align}\label{eq:postcov}
  \postcov = (\prcov^{-1} + \fwd^* \obs^* \Sigma^{-1} \obs \fwd
  )^{-1}.
\end{align}

Usually, this is written using $\tmp := \obs \fwd$ and $\tmp^*$ but
since we need to separate the observation operator $\obs$ from the
forward operator $\fwd$ we write the explicit expression
\eqref{eq:postcov}. Taking $\modcov = 0$ is common practice
\cite{Tarantola05,KaipioSomersalo05,Vogel02} and then we are reduced
to the case where we take iid observations and $\Sigma = \sigma^2I$ is
simply a scalar matrix that does not depend on $\obs$. Note that
taking an error model with a non-scalar covariance as we do here
allows us to consider model error (modeled by $\modcov$) as well as
measurement error (modeled by $\sigma^2$). For example, say we believe
our forward model does not capture some small scale phenomenon.  Then
we may express this belief by saying $\tru = \fwd + \err$, with $\fwd$
depending on $\param$ and $\err$ depending on $\sspar$, and $\param
\perp \sspar$. We do not know much about this effect but it is
reasonable to assume it changes continuously in our domain. We (may
choose to) model it as $\normal (0, \modcov)$ and take $\modcov$ to
reflect the spatial (or other) variability we imagine $\err \sspar$
has. Such small scale phenomenon can arise as a modeling issue, where
we might not model the system in its entirety. It can also arise from
a numerical source, where our discretization of the system is not fine
enough to capture all small scale phenomena. We will see that assuming
some correlation in the error may reduce the clusterization
phenomenon, as is also reported in the literature \cite{Ucinski05}.

\subsection{D-Optimal Design in Infinite Dimensions} 
As pointed out before, we consider D-optimal experimental design. The
meaning of this notion in infinite-dimensional Hilbert spaces was
investigated in \cite{AlexanderianGloorGhattas14}. Note that the
authors make assumptions that amount to $\Sigma=I$ (implied by
$\modcov = 0,\sigma^2=1$), but we choose not to take these
simplification here. This is because $\Sigma$ can determine ``how
much'' clusterization we see. As stated
\cite[pp. 681]{AlexanderianGloorGhattas14}, the results hold for more
general covariance matrices. The derivation of the design criterion
from information theoretic considerations can be verified similarly.
The conclusion is that in infinite-dimensions, a D-optimal design is
well-defined as maximizing the expected KL divergence between
posterior and prior and is summarized, using our notation, in the
following theorem:
\begin{theorem}[Slightly modified Theorem 1 from \cite{AlexanderianGloorGhattas14}]
  Let $\pr = \normal(0,\prcov)$ be a centered Gaussian prior
  on $\hil$   and let $\post = \normal(\postmean,\postcov)$ 
  the posterior measure on $\ban$ for the Bayesian linear
  inverse problem $\data =  \obs \fwd\param + \obs \eps' + \eps$ discussed
  above. Then 
  \begin{align}\label{eq:objective}
    \begin{split}
      \tar( \obs) :&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
      % 
      % 
      % 
      &= \frac12 \log \det 
      ( I + \prcov^{1/2}  \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}).
    \end{split}
  \end{align}
\end{theorem}

\section{The Constrained Optimization Problem of D-Optimal Design}\label{section:D and grad}
From \eqref{eq:objective} we wish to characterize solution(s) of the
following optimization problem for $\tar$. %%: (\ban^*)^m \to \R$:
\begin{align*}
  \obs^{\star} := \argmax_{\obs} \tar( \obs ) 
  = \argmax_{\obs} \frac12 \log \det 
  (I + \prcov^{1/2} \fwd^*\obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}),
\end{align*}
where $\obs$ is constrained to some allowed set of measurements
(e.g.\ pointwise evaluations on $[0,\pi]$)
%%, for the example in section \ref{subsec:example}).
We seek a formulation of the optimal design using Lagrange
multipliers. We first find gradients in function spaces.

\subsection{The Objective and its Gradient}\label{section:objective}
In order to use Lagrange multipliers, we need to find the gradient of
$\tar$. This section is mostly technical and devoted to calculating
said gradient. The result is recorded in \eqref{eq:tar grad}.

We start by calculating the first variation. Some calculations are
delegated to Lemma \ref{lemma:aux calc} in the appendix. Also, we
use the following lemma whose proof is also delegated to the appendix:
\begin{lemma}[Generalized from \cite{Lax97}]\label{lemma:lax}
  Let $Y(t)$ be a differentiable operator-valued function. Assume 
  $I+Y(t)$ is invertible, $Y(t)$ self-adjoint and trace-class. Then
  \begin{equation*}
    \frac{\der \log \det (I+Y(t))}{\der t} = \tr (I+Y(t))^{-1} \dot{Y}(t).
  \end{equation*}
\end{lemma}

First variation of $\tar (\obs)$ in the direction $V$ is:
\begin{align*}%%\label{eq:tar var}
  \begin{split}
    \delta \tar(\obs) V 
    :&= \frac{\der}{\der\tau} \Big |_{\tau=0} \tar(\obs + \tau V)\  \text{ (by definition of variation)}\\
    % 
    % 
    % 
    &= \frac12 \frac{\der}{\der \tau} \Big |_{\tau=0} \log \det 
    (I + \prcov^{1/2} \fwd^* T(\obs+\tau V)\fwd \prcov^{1/2} ) \text{ (by definition \eqref{eq:objective})}\\
    % 
    % 
    % 
    &= \frac12 \tr ( I + \prcov^{1/2} \fwd^* \obs^* \Sigma^{-1}
    \obs\fwd \prcov^{1/2} )^{-1}
    \frac{\der}{\der \tau} \Big |_{\tau=0}
    \prcov^{1/2} \fwd^* T(\obs+\tau V) \fwd \prcov^{1/2}\ \text{ (by \ref{lemma:lax})}\\
    % 
    % 
    % 
    &= \frac12 \tr \postcov \fwd^* (
      V^* \Sigma^{-1} \obs 
      - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \text{ (by \ref{lemma:aux calc})}\\
      &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs
      + \obs^* \Sigma^{-1} V ) \fwd \\
%%   \end{split}
%% \end{align} 
%% Taking adjoints and cyclic permutations we see that
%% \begin{align*}
%%   \tr \postcov \fwd^* V^* \Sigma^{-1} \obs \fwd 
%%   &= \tr \postcov \fwd^* \obs^* \Sigma^{-1} V \fwd,
%% \end{align*}
%% and
%% \begin{align*}
%%   \tr \postcov \fwd^* \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \fwd
%%   &= \tr \postcov \fwd^*  \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs \fwd.
%% \end{align*}
%% We use these in the first equality of the following derivation,
%% which continues the one in \eqref{eq:precursor}.
%% %
%% \begin{align} \label{eq:tar var}
%%   \begin{split}
      %% \delta \tar(\obs) V 
    &= \tr \postcov \fwd^* ( \obs^* \Sigma^{-1} V 
    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs ) \fwd \\
    %
    %
    % 
    &= \tr \postcov \fwd^* \obs^* \Sigma^{-1} V 
    ( I - \modcov \obs^* \Sigma^{-1}\obs ) \fwd \\
    % 
    %
    %
    &= \tr V ( I - \modcov \obs^* \Sigma^{-1}\obs )
    \fwd \postcov \fwd^* \obs^* \Sigma^{-1}.
  \end{split}
\end{align*} 
Denote
\begin{align}\label{eq:tar grad}
  \nabla \tar(\obs) &:= (I - \modcov \obs^* \Sigma^{-1} \obs) \fwd
  \postcov \fwd^* \obs^*\Sigma^{-1},
\end{align}
the gradient of $\tar (\obs )$ and we now justify this definition.
Since the trace of $A := V \nabla \tar(\obs) \in \R^{m \times m}$ is
just $\tr A = \sum_{j=1}^m e_j^t A e_j$ (with $e_j$ the $j$th standard
basis vector), we see that
\begin{align*}
  \delta \tar(\obs)V = \tr V \nabla \tar(\obs) = \sum_{j=1}^m
  V_j(\nabla \tar(\obs)_j),
\end{align*}
with $V_j \in \ban^*$ and $\nabla \tar(\obs)_j \in \ban^{**}\subseteq
\ban, j=1,...,m$. Thus, $\nabla \tar( \obs ) \in (\ban^{**})^m
\subseteq \ban^m$ is indeed the correct gradient and the notation
\eqref{eq:tar grad} is justified. Note that by the comment following
\eqref{eq:obs*}, we view $V$ (defined on the same space as $\obs$) as
a column vector. Then the gradient $\nabla \tar(\obs)$ should be
viewed as a row vector, as the product $V \nabla \tar$ is in $\R^{m
  \times m}$. This will prove important in section
\ref{subsec:necessary}.

\subsection{Unit Length Constraints and their Gradients}\label{subsec:unit length}
In this section we suggest relaxed constraints in
\eqref{eq:constraints} and derive their gradient in \eqref{eq:grad
  constraints}, for the purpose of using them in a Lagrange
multipliers problem.

As mentioned before, we cannot choose any $\obs$ when maximizing
$\tar(\obs)$. Recall $\obs = (\meas_1,...,\meas_m)^t \in (\ban^*)^m$
and each $\meas_j,j=1,...,m$ must be chosen from some allowed set of
functionals in $\ban^*$. This set differs based on the kind of sensors
we have at our disposal and the properties of $\ban$, as discussed in
section \ref{subsec:dynamics}. The following proposition will give us
better understanding of the constraints. 
\begin{proposition}\label{prop:bigger better}
  Let $\obs = (\meas_1,...,\meas_m)^t$, $j \in \{1,...,m\}$, $\sigma^2
  > 0$ and $\lambda > 1$. Then $\tar(\obs)$ increases if we use
  $\lambda \meas_j$ in $\obs$ instead of $\meas_j$.
\end{proposition}
Proof of Proposition \ref{prop:bigger better} is delegated to the
appendix but the idea behind it is simple: making rows of $\obs$ large
is equivalent to making $\sigma^2$ small. This can be understood
easily from the formulation of the inverse problem $\data = \obs \fwd
\param + \obs \eps' + \eps$, where $\eps'$ can be taken as model error
and $\eps$ is iid measurement error.

We now suggest a more convenient-to-work-with set of constraints. By
Proposition \ref{prop:bigger better}, we cannot just take all $\ban^*$
to be the allowed set of measurement vectors, since this would
effectively eliminate measurement error entirely. We note that for
point evaluations, the norm of the measurement is always one:
\begin{align*}
  \| \delta_{\x} \| = \sup_{0 \neq u \in C(\Omega)} \frac{
    |\int_{\Omega}u(\y) \delta_{\x}(\y) \der \y| 
  }{
    \sup|u|}
  = \sup_{0 \neq u \in C(\Omega)} \frac{|u(\x)|}{ \sup|u|} = 1,
  \forall \x \in \Omega.
\end{align*}
Following this, we also restrict the norm of our measurement vectors
to be one. But the norm on functionals is hard to work with. Instead,
we assume there exists a Hilbert space $\bar{\hil} \subseteq \ban^*$
and take measurements $\meas_j \in \bar{\hil}^* = \bar{\hil}$ such
that $\| \meas_j\| = 1, j=1,...,m$. Recall that we think of $\ban$ as
a somewhat restricted space ($C(\bar{\Omega})$ with the sup-norm over
a bounded domain $\Omega$, for example) and so $\ban^*$ can be thought
of as large. Then $\ban^*$ can accommodate a reasonable Hilbert space
$\bar{\hil}$. For example, if $\ban = C(\bar{\Omega})$, we take
$\bar{\hil} = L^2(\Omega) \subseteq \ban^*$.
%% (we can ignore the
%% boundary since is has zero measure).
Then
%
\begin{equation*}
  \ban \subseteq \bar{\hil} \subseteq \ban^*,
\end{equation*}
since members of $C(\bar{\Omega})$ are bounded, hence square integrable. The
unit norm constraint can be written using \eqref{eq:obs*} as a series of $m$ equality
constraints (one for each measurement) on $\obs$ as
\begin{align}\label{eq:constraints}
  \phi_j(\obs) :=\frac12 \| \obs^* e_j\|_{\bar{\hil}}^2 - \frac12 = 0,j=1,...,m
\end{align}
with $e_j \in \R^m$ the $j$th standard basis vector. The first variations are:
\begin{align*}
  \delta \phi_j(\obs)V  
  &= \frac12\lim_{\tau \to 0}\tau^{-1}
  ( \|(\obs + \tau V)^*e_j \|_{\bar{\hil}}^2 - \|\obs ^*e_j \|_{\bar{\hil}}^2  ) \\
  %
  %
  %
  &= \frac12\lim_{\tau \to 0}\tau^{-1}
  ( \langle (\obs + \tau V)^*e_j, (\obs + \tau V)^*e_j \rangle_{\bar{\hil}} - 
  \langle \obs^*e_j, \obs^*e_j \rangle_{\bar{\hil}} ) \\
  % 
  % 
  %
  &= \frac12\lim_{\tau \to 0}\tau^{-1}
  (2\tau \langle \obs^*e_j,V^*e_j \rangle_{\bar{\hil}} 
  +\tau^2 \langle V^*e_j, V^*e_j \rangle_{\bar{\hil}} ) \\
  %
  %
  % 
  &= \langle \obs^*e_j,V^*e_j \rangle_{\bar{\hil}} \\
  %
  %
  % 
  &= \langle V \obs^*e_j,e_j \rangle_{\R^m} \\
  %
  %
  %
  &= e_j^t V \obs^* e_j \\
  % 
  %
  %
  &= \tr V \obs^* e_je_j^t.
\end{align*}
Using the same arguments we used to justify the definition of
\eqref{eq:tar grad}, we conclude that
\begin{align}\label{eq:grad constraints}
\nabla \phi_j(\obs) = \obs^* e_j e_j^t = \meas_j e_j^t , j=1,...,m,
\end{align}
with $\nabla \phi_j(\obs) \in \bar{\hil}^m \subseteq \ban^m$ and where the
last equality follows since $\obs^*e_j = \meas_j$ by \eqref{eq:obs*}. As
we noted at the end of section \ref{section:D and grad}, the
gradient $\nabla \phi_j(\obs)$ is a row vector.

\subsection{Necessary Conditions for Optimal Design}\label{subsec:necessary}
Necessary conditions for optimality are found using Lagrange
multipliers:
\begin{align}
  %\begin{split}
  &\nabla \tar(\obs) = \sum_{j=1}^m \xi_j \nabla \phi_j (\obs)
  \label{eq:Lagrange mult1} \\
    &\phi_j(\obs) = 0, j = 1,...,m. \label{eq:Lagrange mult2}
  %\end{split}
\end{align}
%% This means that any optimal design satisfies \eqref{eq:Lagrange mult1}
%% and \eqref{eq:Lagrange mult2}.
Recall where the objects involved are defined: the observation operator,
$\obs \in \bar{\hil}^m \subseteq (\ban^*)^m$. The objective $\tar:
(\ban^*)^m \to \R$ and its gradient is $\nabla \tar(\obs) \in
\ban^m$. The constraints are defined similarly --- $\phi_j: (\ban^*)^m
\to \R$ and $\nabla \phi_j(\obs) \in \ban^m$.

Focusing on \eqref{eq:Lagrange mult1} and using the gradients of the
objective and constraints \eqref{eq:tar grad} and \eqref{eq:grad
  constraints} we have:
\begin{equation*}
  (I - \modcov \obs^* \Sigma^{-1} \obs) \fwd \postcov \fwd^* \obs^*\Sigma^{-1}
  = \sum_{j=1}^m \xi_j \obs^* e_je_j^t = (\xi_1 \meas_1,...,\xi_m \meas_m).
\end{equation*}
If we let $\Xi = \diag(\xi_j)$, then with some abuse of
notation, we can write this more compactly as:
\begin{equation}\label{eq:conditions}
  ( I - \modcov \obs^* \Sigma^{-1} \obs) \fwd \postcov \fwd^* \obs^*  \Sigma^{-1}
  = \obs^* \Xi.
\end{equation}

\section{Analysis of Optimal Designs --- Vanishing Model Error}
When $\modcov = 0$, then $\Sigma= \sigma^2I$. The necessary
first-order condition for optimality \eqref{eq:conditions} becomes
\begin{equation}\label{eq:eigenproblem}
  \sigma^{-2}\fwd \postcov \fwd^* \obs^* = \obs^* \Xi. 
\end{equation}  
This is an eigenvalue problem for the self-adjoint operator
$\sigma^{-2}\fwd \postcov \fwd^*$ where the rows of $\obs$, namely
$\meas_j,j=1,...m$, are the eigenvectors. If $\fwd$ is invertible,
then we may write:
\begin{align}\label{eq:mod conditions}
  \begin{split}
  \obs^* \Xi &= \sigma^{-2}\fwd \postcov \fwd^* \obs^*  \\
  %
  %
  %
  &= \sigma^{-2} \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* \obs^* \\
  %
  %
  %
  &= \sigma^{-2} \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1} \obs^*,
  \end{split}
\end{align}  
which presents the optimal design problem as a problem on $\ban$ and
not on $\hil$ --- the term $(\fwd \prcov \fwd^*)^{-1}$ is the prior
precision in $\ban$ and the big parentheses enclose the posterior
precision in $\ban$. This expression makes sense even if $\fwd$ is not
invertible --- consider $\fwd: \ker \fwd^{\perp} \to \text{Im} \fwd
\subseteq \ban$ and this does not change the optimal design problem at
all, since we will never take measurements that have a component in
$\fwd^{\perp}$ (this is intuitive by Occam's razor but can also be
understood from Corollary \ref{cor:zero mod err}).

\subsection{Characterizing D-Optimal Designs}\label{subsub:characterization}
In light of \eqref{eq:mod conditions}, let us denote eigenvalues of
$\fwd \prcov \fwd^*$ by $\lambda_1 \geq \lambda_2 \geq ...$ and its
eigenvectors $\{\ev_i\}_{i=1}^{\infty}$. First, let us present a lemma:

\begin{lemma}\label{lemma:sim diag}
  Let $C:\hil \to \hil$ self-adjoint and let $\func_1,...,\func_m \in
  \hil$. Denote $\func^*$ the element $\func$ acting as a linear
  functional. If
  \begin{equation*}
   (C + \sum_{j=1}^m \func_j\func_j^*) \func_k = \xi_k \func_k, k = 1,...,m
  \end{equation*}
  then $C$ and $\sum_{j=1}^m \func_j \func_j^*$ are simultaneously
  diagonalizable.
\end{lemma}

Thus, if we take $\func_j^{*} = \meas_j$ and $C := (\fwd \prcov
\fwd^*)^{-1}$ in Lemma \ref{lemma:sim diag}, we conclude from the
eigenproblem \eqref{eq:mod conditions} that $\obs^*\obs$ has the same
eigenvectors as $\fwd \prcov \fwd^*$, namely
$\{\ev_i\}_{i=1}^{\infty}$ (infinitely many of them with zero
eigenvalue, since $\obs^*\obs$ is finite-rank). Denote the
corresponding non-zero eigenvalues of $\obs^*\obs$ by
$\{\eta_i\}_{i=1}^{k}$ and let $\eta_i = 0$ for $i \geq k+1$. By the argument
in the paragraph following \eqref{eq:mod conditions}:
\begin{align}\label{eq:pre manipulations}
  \begin{split}
    I + \sigma^{-2} \prcov^{1/2}  \fwd ^* \obs^* \obs \fwd \prcov^{1/2} &= 
    \prcov^{1/2} ( \prcov^{-1} + \sigma^{-2} \fwd^*\obs^*  \obs \fwd ) \prcov^{1/2}\\
    %
    %
    %
    &=\prcov^{1/2} \fwd^* ( \fwd^{-*} \prcov^{-1}\fwd^{-1} + \sigma^{-2} \obs^* \obs  ) \fwd \prcov^{1/2}\\
    %
    %
    %
    &= \prcov^{1/2} \fwd^* \left ( (\fwd \prcov \fwd^* )^{-1} +\sigma^{-2} \obs^* \obs  \right ) \fwd \prcov^{1/2}.
  \end{split}
\end{align}
By definition of the objective \eqref{eq:objective} and using
\eqref{eq:pre manipulations}:
\begin{align}\label{eq:mod objective}
  \begin{split}
    \tar ( \obs ) &=\frac12 \log \det \left( I + \sigma^{-2}
    \prcov^{1/2} \fwd ^* \obs^* \obs \fwd \prcov^{1/2} \right )\\
    %
    %
    %
    &=\frac12 \log \det
    \left ( (\fwd \prcov \fwd^* )^{-1} + \sigma^{-2}\obs^* \obs  \right ) \fwd \prcov \fwd^* \\ 
    %
    %
    %
    &=\frac12 \log \left (\prod_{i=1}^{\infty} ( \lambda_i^{-1} + \sigma^{-2} \eta_i ) \prod_{i=1}^{\infty} \lambda_i \right ) \\
    %
    %
    %
    &=\frac12 \log \left ( \prod_{i=1}^{k} ( \lambda_i^{-1} + \sigma^{-2} \eta_i )\prod_{i=1}^{k} \lambda_i \right ). 
  \end{split}
\end{align}
Therefore, an optimal design maximizes
\begin{equation}
  \sum_{i=1}^{k}\log (\lambda_i^{-1} + \sigma^{-2} \eta_i),
\end{equation}
with the norm constraints on $\meas_j,j=1,...,m$. We show in the
appedix (Lemma \ref{lemma:free}) we can find $\meas_1,...,\meas_m$
such that $\{\eta_i\}_{i=1}^k$ have any value we desire, subject only
to the restriction that we do not change the trace of $\obs^*\obs$, so
that $\sum_{i=1}^k \eta_i = m$. By concavity of $\log$, the highest
increase in the design criterion is gained by increasing
$\lambda_i^{-1} + \sigma^{-2}\eta_i$ where it is small. So, no weight
should be given to (i.e. measurement taken of) the $k$th eigenvector
before $\lambda_i^{-1} + \sigma^{-2}\eta_i \geq \lambda_k^{-1} \forall
i \neq k$. Thus, we should choose $\{\eta_i\}_{i=1}^k$ as follows. We
would increase $\lambda_1^{-1} + \sigma^{-2} \eta_1$ until it equals
$\lambda_2^{-1}$. Then, $\lambda_i^{-1} + \sigma^{-2} \eta_i,i=1,2$ is
increased until $\lambda_i^{-1} + \sigma^{-2} \eta_i =
\lambda_3^{-1},i=1,2$ and so forth.

The argument above has the important consequence that the eigenvalues
in \eqref{eq:eigenproblem} are all equal! This makes perfect
sense. First, we know the Lagrange multipliers tell us how much we
gain by relaxing the constraints and there is no reason for any
measurement to give us more than any other (otherwise we would put
more weight on the better measurement, as can be understood from the
discussion in the following paragraph). We are also used to symmetric
optimization problems having symmetric solutions \cite{Waterhouse83},
so symmetric and identical Lagrange multipliers definitely make sense
in this context.

\subsection{Sensor Clusterization}\label{subsec:clusterization}
The symmetry mentioned in last paragraph of the previous section may
very well break in real life problems. If this happens and eigenvalues
of \eqref{eq:eigenproblem} are all different, then showing
sensors do cluster proceeds as follows. First, we show there
are only so many different equal, by the pigeonhole principle. In
order to make this argument work, one needs to show this eigenproblem
does, indeed, have only finitely many eigenvalues. This is done by
first recalling that $\fwd$ is strongly smoothing (so $\fwd \postcov
\fwd^*$ has fast decaying eigenvalues) and then showing that small
eigenvalues arise from un-informative measurements. We use the
following corollary, whose proof may be found in the appendix.
%
\begin{corollary}[Gain for No Model Error]\label{cor:zero mod err}
  If $\modcov = 0$, then
  \begin{equation*}
    \tar( \obs ) - \tar (\obsm )
    = -\frac12 \log (1 - \sigma^{-2} \langle \fwd \postcov \fwd^* \meas_m, \meas_m \rangle ).
  \end{equation*}
\end{corollary}
%
Suppose $\meas_j$ is a measurement with a small eigenvalue, i.e.\
\begin{equation}\label{eq:small eigenvalue}
  \sigma^{-2} \fwd \postcov \fwd^* \meas_j = \xi_j \meas_j,\ \xi_j \ll 1.
\end{equation}
By Corollary \ref{cor:zero mod err}, the increase in the design
criterion is
\begin{align*}
  \tar(\obs) - \tar(\obsm) &=
  -\frac12 \log (1- \sigma^{-2} \langle \fwd
  \postcovm \fwd^* \meas_j, \meas_j \rangle) \\
  %
  %
  %
  &= - \frac12 \log (1 - \xi_j) \text{ since $\|\meas_j\|=1$ }\\
  %
  %
  %
  &\approx \frac{\xi_j}{2}.
\end{align*}
Thus, negligibly small eigenvalues give negligible increase in the
design criterion and will not be a part of a truly optimal
design. Since the eigenvalues of $\fwd \postcov \fwd^*$ decay faster
than those of $\fwd\prcov\fwd^*$, the latter will determine how many
non-negligible eigenvalues $\fwd \postcov \fwd^*$ has. Thus, in the
eigenproblem \eqref{eq:eigenproblem}, we need only consider a finite
number of eigenvectors.

At this point, one would like to invoke the pigeonhole principle and
claim that eigenvectors cluster. This, however, is not as
straightforward as one may hope. If $\fwd \postcov \fwd^*$ only has
simple eigenvalues, then we may use the pigeonhole principle and
conclude that some eigenvectors are clustered (up to a sign). However,
if $\fwd \postcov \fwd^*$ has non-simple eigenvalues, one may hope to
somehow prove measurement vectors are either orthogonal or parallel,
which suffices for proving clusterization but that is not true.

\subsection{Non-Vanishing Model Error}
We can build on Lemma \ref{lemma:design increase} to find the gain in design
for identical measurements.
\begin{corollary}[Gain for Identical Measurement]\label{cor:same meas}
  If $\meas_m = \meas_j$ for some $1 \leq j \leq m-1$, then
  \begin{equation*}
    \tar(\obs) - \tar(\obsm) =
    \log \left ( 1 + \frac{\sigma^2
      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
      \obsm^* \Sigmam^{-1}e_j \rangle
    }{
      2 - \sigma^2 e_j^t\Sigmam^{-1}e_j 
    }       
    \right ).
  \end{equation*}
\end{corollary}

In contrast to the previous case of vanishing model error, we will see
that if $\modcov \neq 0$ clustering will not occur. From Corollary
\ref{cor:same meas} we can observe that as $\sigma^2 \to 0$, the
increase in the design criterion gained by taking a measurement
identical to a previous one approaches zero. We may conclude that for
small measurement error levels, the clusterization effect will be
canceled by the presence of a non-zero model error. Since the design
criterion is not defined for $\sigma^2 = 0$ and identical
measurements, we cannot make a statement regarding $\sigma^2 = 0$,
except in the limiting sense described above.

\section{Conclusion}\label{section:conclusion}
Our relaxed model gives us insight to D-optimal designs. We see that
in an ideal settings, we should not reduce uncertainty in any
direction before it is the largest uncertainty present.

We believe the model presented is not too far from reality. After all,
taking any measurement (point evaluation or other) is just recording
the value of a linear combination of individual eigenvectors of the
prior. In light of this, we believe the reason for clusterization is close
to the argument made in section \ref{subsub:first attempt}. The reason
we observe clusterization in real life problems is probably due to
symmetry breaking, although we cannot show that rigorously.

An important consequence of the argument in \ref{section:vanishing} is
that we can find an approximately optimal design by means of linear
algebra. Here is how this should be done.
\begin{enumerate}

   \item Find eigenvalues $\lambda_1 \geq \lambda_2 \geq ...$ and
     corresponding eigenvectors $\ev_i$ of $\fwd \prcov \fwd^*$.
   \item Denote the (assumed known) variance of the measurement noise
     $\sigma^2$.
   \item Decide how many measurements you can take and denote that
     number $m$. This is the total precision one can allocate to the
     different eigenvectors. Next, we see how to allocate these.
    \item Find $k$ and $\{\eta_i\}_{i=1}^k$, $\eta_i > 0, i=1,...,k$
      such that:
      \begin{itemize}
        \item $\sum_{i=1}^k \eta_i = m$ --- the total measurement precision available.
      \item $\lambda_i^{-1} + \sigma^{-2}\eta_i = \lambda_k^{-1} +
        \sigma^{-2}\eta_k$ for $i=1,...,k-1$ --- the posterior
        precision is equal for all $k$ eigenvectors.
      \item $\lambda_{k+1}^{-1} \geq \lambda_k^{-1} + \eta_k$ --- no
        information is gained for the rest of the eigenvectors.
      \end{itemize}
      \item Use the construction in Lemma \ref{lemma:free} to find the
        optimal measurements in terms of the eigenvectors $\ev_i$ ---
        without \emph{any} optimization.
      \item Find point measurements that approximate those in the domain. 
\end{enumerate}

\appendix
\section{Widely Applicable Lemmas}

The following lemma is generalized from \cite[Chapter 9,
  Theorem 4, pp. 127]{Lax97}
\begin{lemma}\label{lemma:lax}
  Let $Y(t)$ be a differentiable operator-valued function. Assume 
  $I+Y(t)$ is invertible, $Y(t)$ self-adjoint and trace-class. Then
  \begin{equation*}
    \frac{\der \log \det (I+Y(t))}{\der t} = \tr (I+Y(t))^{-1} \dot{Y}(t).
  \end{equation*}
\end{lemma}
\begin{proof}
  Consider a differentiable operator-valued function $X(t)$ such that
  $X(0) = 0$ and $X(t)$ is positive, self-adjoint and trace-class for
  every $t\in \R$. We denote the eigenvalues of this operator by
  $\lambda_k(X(t))$ and sometimes drop the dependence on $X(t)$, so
  $\lambda_k = \lambda_k(X(t))$.  Then $\det (I+X(t)) =
  \prod_{k=1}^{\infty} (1+\lambda_k) < \infty$ where the bound holds
  by the arguments given in \cite{AlexanderianGloorGhattas14}. The
  full derivative is
  \begin{align*}
    \frac{\der \det (I+X(t))}{\der t} 
    % 
    % 
    % 
    &= \sum_{k=1}^{\infty} 
    \frac{\partial \det (I+X(s))}{\partial (1+\lambda_k)}\Big |_{s=t}
    \frac{\der (1+\lambda_k)}{\der t} \\
    % 
    %
    %
    &= \sum_{k=1}^{\infty} \frac{\partial \prod_{l=1}^{\infty}
      (1+\lambda_l(s))}{\partial (1+\lambda_k)}\Big |_{s=t}
    \frac{\der (1+\lambda_k)}{\der t} \\
    %
    %
    %
    &= \sum_{k=1}^{\infty} \frac{\prod_{l=1}^{\infty}
      (1+\lambda_l(s))}{(1+\lambda_k)}\Big |_{s=t}
    \dot{\lambda_k}(X(t)) \\
    % 
    % 
    % 
    &= \sum_{k=1}^{\infty} \frac{\det (I+X(t))}{1 +\lambda_k} \dot{\lambda_k}(X(t)).
  \end{align*}
  The assumption $X(0) = 0$ means $\lambda_k(X(0)) = 0,\ \forall k \geq 1$. Thus:
  \begin{align*}
    \frac{\der (I+\det X(t))}{\der t}\Big |_{t=0} 
    = \sum_{k=1}^{\infty} \dot{\lambda_k}(X(0)) 
    = \frac{\der }{\der t}\tr X(0)
    = \tr \dot{X}(0),
  \end{align*}
  where the second equality follows by monotone convergence. 
  Let $Y(t)$ a trace-class self-adjoint operator such that 
  $I+Y(t)$ is invertible.
  Define $X(t)$ via $I+X(t) = (I+Y(0))^{-1/2} (I+Y(t)) (I+Y(0))^{-1/2}$. 
  We show $X(t)$ satisfies the conditions above. It is trace-class:
  \begin{align*}
    \tr X(t) = \tr \big ( (I+Y(0))^{-1} (I+Y(t)) - I \big )
    \leq \tr \big( I+Y(t) - I \big ) < \infty,
  \end{align*}
  since $Y(t)$ is trace-class. It is also clear that
  $X(0) = 0$ and $X(t)$ is self-adjoint.
  $I+Y(t) = (I+Y(0))^{1/2}(I+X(t))(I+Y(0))^{1/2}$, so
  \begin{align*}
    \frac{\der \det (I+Y(t))}{\der t}|_{t=0} 
    &= \det (I+Y(0))\frac{\der \det (I+X(t))}{\der t}\Big |_{t=0} \\
    % 
    % 
    % 
    &= \det (I+Y(0)) \tr \dot{X}(0) \\
    % 
    % 
    % 
    &= \det (I+Y(0)) \tr (I+Y(0))^{-1} \dot{Y}(0).
  \end{align*}
  Consequently, by the one-variable chain rule:
  \begin{align*}
    \frac{\der \log \det (I+Y(t))}{\der t}\Big |_{t=0} &=
    % 
    % 
    % 
    \frac{1}{\det (I+Y(0))}\frac{\der \det (I+Y(t))}{\der t}\Big |_{t=0} \\ 
    % 
    % 
    % 
    &= \tr (I+Y(t))^{-1} \dot{Y}(t) \big |_{t=0}.
  \end{align*}
  Since there is nothing special about $t=0$, the relation holds for all $t$.
\end{proof}

\begin{lemma}[Matrix Determinant Lemma in Hilbert Spaces]\label{lemma:MDL}
  Let $\hil$ a separable Hilbert space, $u,v\in \hil$ and $A: \hil \to
  \hil$ an invertible linear operator such that $\tr A-I <
  \infty$. Then $\det A$ and $\det A + uv^*$ are well defined and
  \begin{equation*}
    \det (A + uv^*) = (1 + \langle A^{-1} u, v \rangle ) \det A,
  \end{equation*}
  where $(A + uv^*)w := Aw + \langle v,w \rangle u$.
\end{lemma}
\begin{proof}
  In this proof we rely on definitions and results from
  \cite{Simon77}. First, consider $B := I + xy^*$ for some $x,y \in
  \hil$. We construct an eigenbasis for $B$ and use that to show $\det
  B = 1 + \langle x, y \rangle$. First let $x_1 := x$.  Now, if $x
  \parallel y$, take $\{x_n \}_{n=2}^{\infty}$ an orthogonal basis for
  $span\{x_1\} ^{\perp}$. If, on the other hand, $x \nparallel y$, let
  \begin{equation*}
    x_2 := x - \frac{ \langle x, y\rangle}{\|y\|^2}y
  \end{equation*}
  and it is easy to verify that $x_2 \perp y$ and $span \{x,y\} = span
  \{x_1,x_2\}$. Take $\{x_n \}_{n=3}^{\infty}$ an orthogonal basis for
  $span\{x_1,x_2\} ^{\perp}$. In both cases,
  \begin{equation*}
    B x_n =
    \begin{cases}
      (1 + \langle x, y \rangle) x_n & n = 1 \\
      x_n                            & n \neq 1,
    \end{cases}
  \end{equation*}
  and so $\det B = 1 + \langle x, y \rangle$.

  It is easy to verify that $uv^*$ is trace-class and since $\tr A - I
  < \infty$, also $\tr A + uv^* - I < \infty$ (sum of two trace-class
  operators is trace-class). Thus $\det A$ and $\det (A+uv^*)$ are
  well defined. Let $x:=A^{-1}u$ and $y := v$:
  \begin{equation*}
    \det (A + uv^*) = \det A \ \det(I+A^{-1}uv^*) =
    (1 + \langle A^{-1}u, v \rangle) \det A .
  \end{equation*}
\end{proof}

\section{Specific Lemmas}

\begin{lemma}[Increase due to a Measurement]\label{lemma:design increase}
  Let $\obs = (\meas_1,...,\meas_m)^t$ and $\obsm := (\meas_1,...,\meas_{m-1})^t$. Then
  \begin{align*}
    \tar( \obs ) - \tar (\obsm ) &=
    \frac12 \log \left ( 1 + \frac{
      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m,
      (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m \rangle
    }{
      \sigma^2 + \meas_m \modcov \meas_m - \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m 
    }       
    \right ).
  \end{align*}
\end{lemma}
\begin{proof}
  We use the Schur complement to write one inverse in terms of the other and
  introduce notations to make the derivation cleaner. Note that we think of
  $\obsm$ and $\obsm^*$ as column and row vectors (respectively).
  \begin{align*}
    \Sigma( \obs ) &= \Sigma = 
    \begin{bmatrix}
      \Sigma (\obsm )           & \obsm \modcov \meas_m \\
      \meas_m \modcov \obsm^*   & \sigma^2 + \meas_m \modcov \meas_m
    \end{bmatrix}
    =
    \begin{bmatrix}
      \Sigmam   & w \\
      w^t       & c
    \end{bmatrix}\\
    %
    %
    %
    \Sigma^{-1} &=
    \begin{bmatrix}
      \Sigmam^{-1} + \Sigmam^{-1} w ( c - w^t \Sigmam^{-1} w)^{-1} w^t \Sigmam^{-1} & - \Sigmam^{-1} w ( c - w^t \Sigmam^{-1} w)^{-1} \\
      -( c - w^t \Sigmam^{-1} w)^{-1} w^t \Sigmam^{-1}                            &  ( c - w^t \Sigmam^{-1} w)^{-1}
    \end{bmatrix} \\
    &=
    \begin{bmatrix}
      \Sigmam^{-1} & 0 \\
      0           & 0 
    \end{bmatrix}
    + (c -w^t \Sigmam^{-1} w )^{-1}
    \begin{bmatrix}
      \Sigmam^{-1} w \\
      -1
    \end{bmatrix}
    \begin{bmatrix}
      w^t \Sigmam^{-1} & -1 
    \end{bmatrix}
  \end{align*}
  %
  Further, define
  %
  \begin{align*}
    \M (\obs ):&= \prcov^{\frac12}\fwd^* \obs^* \Sigma^{-1} \obs \fwd
    \prcov^{\frac12}    
  \end{align*}
  %
  and note that, using our understanding of what is a column vector and
  what is a row vector:
  %
  \begin{align*}
    \M(\obs) &= \prcov^{1/2} \fwd^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2} \\
    %
    %
    %
    &= \prcov^{1/2} \fwd^* \obs^* \left \{
    \begin{bmatrix}
      \Sigmam^{-1} & 0 \\
      0           & 0 
    \end{bmatrix}
    + (c -w^t \Sigmam^{-1} w )^{-1}
    \begin{bmatrix}
      \Sigmam^{-1} w \\
      -1
    \end{bmatrix}
    \begin{bmatrix}
      w^t \Sigmam^{-1} & -1 
    \end{bmatrix} 
    \right \} \obs \fwd \prcov^{1/2} \\
    %
    %
    %
    &= \M (\obsm) + (c -w^t \Sigmam^{-1} w )^{-1}
    \prcov^{1/2} \fwd^* \obs^*
    \begin{bmatrix}
      \Sigmam^{-1} w \\
      -1
    \end{bmatrix}
    \begin{bmatrix}
      w^t \Sigmam^{-1} & -1 
    \end{bmatrix} 
    \obs \fwd \prcov^{1/2}
  \end{align*}
  %
  Now, denote:
  %
  \begin{align}\label{eq:u}
    \begin{split}
      u :&= (c -w^t \Sigmam^{-1} w )^{-1/2}
      \prcov^{1/2} \fwd^* \obs^* 
      \begin{bmatrix}
        \Sigmam^{-1} w \\
        -1 
      \end{bmatrix} \\
      %
      %
      %
      & = (c -w^t \Sigmam^{-1} w )^{-1/2} ( \prcov^{1/2}\fwd^* \obsm^* \Sigmam^{-1} \obsm  \modcov \meas_m - \prcov^{1/2} \fwd^* \meas_m )\\
      %
      %
      %
      u^* :&=  (c -w^t \Sigmam^{-1} w )^{-1/2} (\meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \fwd \prcov^{1/2} - \meas_m \fwd \prcov^{1/2} ),
    \end{split}
  \end{align}
  %
  so that
  %
  \begin{equation}\label{eq:M plus I}
    I + \M( \obs ) = I + \M (\obsm ) + uu^*.
  \end{equation}
  %
  Note that
  \begin{equation}\label{eq:M postcov}
    \prcov^{1/2} \left (I + \M( \obsm ) \right )^{-1} \prcov^{1/2} = \postcovm.
  \end{equation}
  The increase in the design criterion gained by including $\meas_m$
  in the design is found by using Lemma \ref{lemma:MDL} the above
  results:
  %
  \begin{align*}
    \tar( \obs ) - \tar( \obsm )
    %
    %
    %
    &= \frac12 \log \det \Big ( I + \M ( \obs ) \Big ) / \det \Big ( I + \M (\obsm) \Big ) \\
    %
    %
    %
    &= \frac12  \log \det \left ( I + \M(\obsm) + uu^* \right ) / \det \Big ( I + \M (\obsm) \Big ) \\
    %
    %
    %
    &= \frac12 \log \left ( 1 + \left \langle \left ( I+\M(\obsm) \right )^{-1} u, u  \right \rangle \right ).
  \end{align*}
  Using \eqref{eq:u}:
  \begin{align*}
    &\left \langle \left (I+\M (\obsm)\right )^{-1}u, u \right \rangle\\
    &= \frac{
      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m,
      (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m \rangle
    }{
      c- w^t \Sigmam^{-1} w
    }\\
    %
    %
    %
    &= 
    \frac{
      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m,
      (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m \rangle
    }{
      \sigma^2 + \meas_m \modcov \meas_m - \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m 
    }
  \end{align*}
  and the conclusion follows.
\end{proof}

Lemma \ref{lemma:design increase} implies the following two corollaries.
\begin{corollary}[Gain for No Model Error]\label{cor:zero mod err}
  If $\modcov = 0$, then
  \begin{equation*}
    \tar( \obs ) - \tar (\obsm )
    = -\frac12 \log (1 - \sigma^{-2} \langle \fwd \postcov \fwd^* \meas_m, \meas_m \rangle ).
  \end{equation*}
\end{corollary}
\begin{proof}
  Note that this is not immediate by substituting $\modcov = 0$ in the
  conclusion of Lemma \ref{lemma:design increase}, since we make a
  claim for $\postcov$, and not $\postcovm$. Let us first review
  \eqref{eq:u} and note that since $\modcov = 0$ the covariance
  $\Sigmam = \sigma^2I_{m-1}$ and $w = 0$, so $c - w^t\Sigmam^{-1}w =
  \sigma^2$:
  \begin{align*}
    u :&=-\sigma^{-1}\prcov^{1/2} \fwd^* \meas_m\\
    %
    %
    u^* :&= -\sigma^{-1} \meas_m \fwd \prcov^{1/2}.
  \end{align*}
  From \eqref{eq:M plus I}:
  \begin{equation*}
    I + \M(\obsm) = I +\M(\obs) - uu^*,
  \end{equation*}
  and thus:
  \begin{equation*}
    \left \langle \left ( I +\M(\obs) \right )^{-1} u, u \right \rangle
    = \sigma^{-2} \langle \fwd \postcov \fwd^* \meas_m, \meas_m \rangle.
  \end{equation*}
  Analogously to \eqref{eq:M postcov} we note that
  \begin{equation*}
    \prcov^{1/2} \left ( I + \M(\obs) \right )^{-1} \prcov^{1/2} = \postcov.
  \end{equation*}
  Using Lemma \ref{lemma:MDL} we conclude
  \begin{align*}
    \tar( \obs ) - \tar( \obs )
    &= \frac12 \log \det \left (I +\M(\obs) \right ) / \det \left (I + \M(\obsm) \right ) \\
    %
    %
    %
    &= \frac12 \log \det \left (I +\M(\obs) \right ) / \det \left (I + \M(\obs) - uu^* \right ) \\
    %
    %
    %
    &=-\frac12 \log (1 - \langle (I+\M(\obs))^{-1}u, u \rangle \\
    %
    %
    %
    &= -\frac12 \log (1 - \sigma^{-2} \langle \fwd \postcov \fwd^* \meas_m, \meas_m \rangle ).
  \end{align*}
\end{proof}

\begin{corollary}[Gain for Identical Measurement]\label{cor:same meas}
  If $\meas_m = \meas_j$ for some $1 \leq j \leq m-1$, then
  \begin{equation*}
    \tar(\obs) - \tar(\obsm) =
    \log \left ( 1 + \frac{\sigma^2
      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
      \obsm^* \Sigmam^{-1}e_j \rangle
    }{
      2 - \sigma^2 e_j^t\Sigmam^{-1}e_j 
    }       
    \right ).
  \end{equation*}
\end{corollary}
\begin{proof}
  Denote $A:= \obs \modcov \obs^*$ and $v_j$ the $j$th column of $A$.
  Note that $v_j = \obsm \modcov \meas_m$, since $(\obsm \modcov
  \obsm^*)_{ij} = \meas_i(\modcov \meas_j)$, as explained in
  \eqref{eq:modcov explained}. One can now verify that
  \begin{equation}\label{eq:observation}
    \Sigmam^{-1} \obsm \modcov \meas_m = \Sigmam^{-1}v_j = (A +\sigma^2I_{m-1})^{-1} v_j =
    e_j -\sigma^2 \Sigmam^{-1}e_j.
  \end{equation}
  %
  Using \eqref{eq:observation}:
  \begin{align}\label{eq:denominator}
    \begin{split}
      \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m
      &= \meas_m \modcov \obsm^* ( e^j - \sigma^2 \Sigmam^{-1} e_j )\\
      %
      %
      %
      &= \meas_m \modcov \meas_j - \sigma^2 \meas_m \modcov \obsm^* \Sigmam^{-1}e_j \\
      %
      %
      %
      &= \meas_m \modcov \meas_j -\sigma^2 (e_j - \sigma^2 \Sigmam^{-1}e_j)^t e_j \\
      %
      %
      %
      &= \meas_m \modcov \meas_m -\sigma^2 + \sigma^4 e_j^t\Sigmam^{-1}e_j.
    \end{split}
  \end{align}
  We use \eqref{eq:observation} to simplify the terms in the enumerator of
  the conclusion of Lemma \ref{lemma:design increase}:
  \begin{align}\label{eq:enumerator}
    \begin{split}
      (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m
      &= \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m - \meas_m \\
      %
      %
      %
      &= \obsm^* (e_j - \sigma^2 \Sigmam^{-1} e_j) -\meas_j \\ 
      %
      %
      %
      &= -\sigma^2 \obsm^* \Sigma^{-1}e_j. 
    \end{split}
  \end{align}
  %
  Substitute \eqref{eq:enumerator} and \eqref{eq:denominator} to
  the enumerator and denominator (respectively) of the conclusion of
  Lemma \ref{lemma:design increase}:
  %
  \begin{align*}
    \tar( \obs ) - \tar (\obsm ) &=
    \log \left ( 1 + \frac{
      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m,
      (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m \rangle
    }{
      \sigma^2 + \meas_m \modcov \meas_m - \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m 
    }       
    \right ) \\
    %
    %
    %
    &= \log \left ( 1 + \frac{\sigma^4
      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
      \obsm^* \Sigmam^{-1}e_j \rangle
    }{
      2\sigma^2 - \sigma^4 e_j^t\Sigmam^{-1}e_j 
    }       
    \right ) \\
    %
    %
    %
    &= \log \left ( 1 + \frac{\sigma^2
      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
      \obsm^* \Sigmam^{-1}e_j \rangle
    }{
      2 - \sigma^2 e_j^t\Sigmam^{-1}e_j 
    }       
    \right ).
  \end{align*}
\end{proof}

\begin{lemma}\label{lemma:free}
  Let $M \in \R^{k \times k}$ symmetric positive definite with $\tr M =
  m$, $m > k$. Then we can find $\func_j \in \R^k,j=1,...,m$ with
  $\|\func_j\|=1$ and $A = (\func_1,...,\func_m)$ such that $AA^t =
  M$.
\end{lemma}
\begin{proof}
  Let us diagonalize $M$, so that $M = U D U^t$ with $D =
  \diag(d_1,...,d_k)$ and $U \in \R^{k \times k }$ orthogonal. Let $S
  \in \R^{k \times m}$ with $S_{ii} = \sqrt{d_{i}}$ and zeros
  otherwise. Define $A:= U S V^t$, where $V \in \R^{m \times m}$ is
  orthogonal and will be further restricted later. Then $AA^t = U
  SV^tVS^t U^t = UDU^t$, so $AA^t$ has the required eigenvalues and
  eigenvectors by construction. If we can choose $V$ such that $A$
  also satisfies the unit norm constraints we are done. These
  constraints are, for $j=1,...,m$:
  \begin{equation}\label{eq:V constraints}
   1 = [A^tA]_{jj} = [V S^tS V^t]_{jj},
  \end{equation}
  and we can expect to do this since we assumed $\tr D = m$.

  Define $C = S^tS - I \in \R^{m \times m}$. Note that $\tr C = 0$ and
  $C$ is diagonal with non-zero entries $d_i-1,i=1,...,k$. It suffices
  to find $V$ orthogonal such that $V C V^t$ has zero diagonal. We
  construct such $V$ by sequentially inserting zeros in the diagonal
  and not destroying zeros we already introduced, starting from the
  last diagonal entry and moving to the first. If $c_{kk} \neq 0$ ,
  let $p < k$ such that $c_{pp}c_{kk} < 0$ (such $p$ exists because
  the trace is zero) and let $\theta \in (0,\pi)$. Define a Givens
  rotation $R^{(k)} \in \R^{m \times m}$ by
  \begin{equation*}
    r^{(k)}_{ab} :=
    \begin{cases}
      1 & a = b \neq p \text{ or } a = b \neq k \\
      \cos \theta & a = b = p  \\
     -\sin \theta & a = p, b = k\\
      \cos \theta & a = b = k \\
      \sin \theta & a = k, b = p \\ 
      0 & o.w
    \end{cases}
  \end{equation*}
  Note that conjugating a matrix by $R^{(k)}$ changes only its $k$ and
  $p$ rows and columns. Specifically, it leaves rows and columns
  $k+1,...,m$ unchanged. We want to choose $\theta$ such that
  \begin{equation}\label{eq:mm}
    0 = [R^{(k)} C (R^{(k)})^t]_{kk} = \cos^2 \theta c_{kk} + 2\cos \theta \sin
    \theta c_{kp} + \sin^2\theta c_{pp},
  \end{equation}
  and it suffices to choose $\theta$ such that
  \begin{equation*}
    c_{kk} \cot^2 \theta + 2 c_{kp} \cot \theta + c_{pp} = 0.
  \end{equation*}
  This quadratic in $\cot\theta$ has a real solution, since
  $c_{pp}c_{kk} < 0$ by assumption and we can find $\theta \in
  (0,\pi)$ such that \eqref{eq:mm} is satisfied. We continue to find
  $R^{(k-1)}$ that leaves rows and columns $k,...,m$ unchanged and
  continue introducing zeros to the diagonal. The assumption $\tr D =
  m \Rightarrow \tr C = 0$ guarantees we can do that. Taking $V:=
  R^{(1)} R^{(2)} \dots R^{(k-1)}R^{(k)}$ completes the proof.
\end{proof}

\begin{lemma}[Auxilliary Calculations]\label{lemma:aux calc}
  Let $T(\obs) := \obs^* \Sigma^{-1}(\obs)\obs$, with $\Sigma(\obs)$
  defined as in \eqref{eq:Sigma}. Then:
  \begin{align*}
    \delta T(\obs)V &= V^* \Sigma^{-1} \obs 
    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
    &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs
    + \obs^* \Sigma^{-1} V.
  \end{align*}
\end{lemma}

\begin{proof}
  We need a few supplementary calculations. First:
  \begin{align}\label{eq:der sig}
    \begin{split}
      \frac{\der}{\der \tau} \Big |_{\tau=0} \Sigma( \obs + \tau V )
      &= \frac{\der}{\der \tau} \Big |_{\tau=0} 
      (\obs + \tau V ) \modcov (\obs + \tau V )^*  + \sigma^2I\\
      % 
      % 
      % 
      &= V \modcov \obs^* + \obs \modcov V^*.
    \end{split}
  \end{align}
  By the standard trick for the derivative of an operator: 
  \begin{align*}
    0 &= \frac{\der}{\der \tau} \Big |_{\tau=0} I \\
    % 
    % 
    % 
    &= \frac{\der}{\der \tau} \Big |_{\tau=0}
    \left (\Sigma(\obs+\tau V)^{-1} \Sigma(\obs+\tau V) \right ) \\
    % 
    % 
    % 
    &= \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0} \Sigma+
    \Sigma^{-1} \frac{\der \Sigma(\obs+\tau V)}{\der \tau} \Big |_{\tau=0}\\  
    %
    %
    %
    &= \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0} \Sigma+
    \Sigma^{-1} (V\modcov \obs^* + \obs \modcov V^*) 
    \text{, by \eqref{eq:der sig}. }
  \end{align*}
  Now we can write the variation of $\Sigma^{-1}$:
  \begin{align}\label{eq:der sig inv}
    \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0}  
      &= -\Sigma^{-1} (V \modcov \obs^* + \obs \modcov V^*) \Sigma^{-1}.
    \end{align}
  % 
  Finally, we can do the calculation we need to. By Leibnitz (product) rule
  and \eqref{eq:der sig inv}:
  % 
  \begin{align*}
    \delta T(\obs) V 
    &= \frac{\der T(\obs + \tau V)}{\der \tau} \Big |_{\tau=0} \\
    %
    %
    %
    &= V^* \Sigma^{-1} \obs 
    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
    &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs
    + \obs^* \Sigma^{-1} V,
  \end{align*}
  as required.
\end{proof}

\begin{lemma}\label{lemma:sim diag}
  Let $C:\hil \to \hil$ self-adjoint and let $\func_1,...,\func_m \in
  \hil$. Denote $\func^*$ the element $\func$ acting as a
  functional. If
  \begin{equation*}
   (C + \sum_{j=1}^m \func_j\func_j^*) \func_k = \xi_k \func_k, k = 1,...,m
  \end{equation*}
  then $C$ and $\sum_{j=1}^m \func_j \func_j^*$ are simultaneously
  diagonalizable.
\end{lemma}
\begin{proof}
  First, enumerate the eigenvalues of $C + \sum_{j=1}^m
  \func_j\func_j^*$ as $\xi_1,...,\xi_\ell$. Denote the
  indices of the eigenvectors corresponding to $\xi_i$
  \begin{equation*}
    S_i := \{ 1 \leq k \leq m | (C + \sum_{j=1}^m \func_j\func_j^* )\func_k = \xi_i \func_k \}.
  \end{equation*}
  Define further
  \begin{equation*}
    A_i := \sum_{k \in S_i} \func_k \func_k^*,
  \end{equation*}
  which is self-adjoint. Two observations are in order. First,
  $\sum_{j=1}^m \func_j\func_j^* = \sum_{i=1}^\ell A_i$. Second, $A_i
  \func_k = 0$ if $k\not \in S_i$, since eigenvectors of different
  eigenvalue are orthogonal. For $k \in S_i$
  \begin{equation}\label{eq:on vi}
    \xi_i \func_k = (C + \sum_{j=1}^m \func_j \func_j^* ) \func_k = (C + A_i) \func_k.
  \end{equation}
  Let $V_i := span \{\func_k \}_{k\in S_i}$. Observe that $V_i$ is
  invariant under $A_i$, by definition, and under $C$, by \eqref{eq:on
    vi}. Using \eqref{eq:on vi} again, we conclude that on $V_i$, $A_i
  = \xi_iI - C$. This immediately implied $A_i$ and $C$ have the
  same eigenvectors on $V_i$. This holds for every $1 \leq i \leq
  \ell$ and we conclude that $C$ and $A$ have the same eigenvectors.
\end{proof}

\begin{proposition}\label{prop:bigger better}
  Let $\obs = (\meas_1,...,\meas_m)^t$, $j \in \{1,...,m\}$, $\sigma^2
  > 0$ and $\lambda > 1$. Then $\tar(\obs)$ increases if we use
  $\lambda \meas_j$ in $\obs$ instead of $\meas_j$.
\end{proposition}
\begin{proof} 
  Fix an arbitrary $j=1,...,m$ and take $V:= e_j e_j^t \obs$. We see
  that for $u \in \ban$
  \begin{equation*}
    Vu = e_je_j^t (\meas_1(u),...,\meas_m(u) )^t = e_j \meas_j(u)
    = (0,...,0,\meas_j(u),0,...,0)^t.
  \end{equation*}
  This way, $V$ has the same $j$th entry as $\obs$ while the rest are
  set to zero. We calculate the variation in this direction and show
  it is positive --- $\delta \tar(\obs)V > 0$. This means that
  increasing the magnitude of the $j$th measurement functional
  increases $\tar(\obs)$. We start with the last line of \eqref{eq:tar
    var}, and denote $\tmp : = \fwd \postcov \fwd^*: \ban^* \to \ban$:
  \begin{align*}
     \delta \tar(\obs) V 
    &= \tr V ( I - \modcov \obs^*\Sigma^{-1}\obs) \tmp \obs^* \Sigma^{-1} \\
    % 
    %
    %
    &= \tr e_je_j^t \obs ( I - \modcov \obs^*\Sigma^{-1}\obs) \tmp \obs^* \Sigma^{-1} \\
    %
    % 
    %
    &= e_j^t \obs ( I - \modcov \obs^*\Sigma^{-1}\obs) \tmp \obs^* \Sigma^{-1}e_j \\
    %
    % 
    %
    &= e_j^t ( I - \obs \modcov \obs^*\Sigma^{-1})\obs \tmp \obs^* \Sigma^{-1}e_j \\  
    % 
    %
    %
    &=  e_j^t(\Sigma-\obs \modcov \obs^*) \Sigma^{-1}\obs \tmp \obs^* \Sigma^{-1}e_j \\
    %
    %
    %
    &=\sigma^2 e_j^t \Sigma^{-1}\obs \tmp \obs^* \Sigma^{-1}e_j
    \text{ by \eqref{eq:Sigma} }\\
    %
    % 
    %
    &=\sigma^2 e_j^t \Sigma^{-1}\obs \fwd \postcov \fwd^* \obs^* \Sigma^{-1}e_j.
  \end{align*} 
  Since $\postcov$ is positive definite, $\delta \tar(\obs) V > 0$.
\end{proof}

%% HERE
%% Since a measurement that takes into account $\ker \fwd$ is
%% necessarily sub-optimal, we may assume $\fwd$ is in fact
%% invertible (it is, on $\ker \fwd ^{\perp}$) and rewrite the
%% necessary condition \eqref{eq:eigenproblem} using
%% \eqref{eq:postcov} as \begin{equation*} \sigma^{-2} \left ( (\fwd
%% \prcov \fwd^*)^{-1} + \sigma^{-2}\obs\obs^* \right )^{-1} =
%% \obs^* \Xi.  \end{equation*} By Lemma \ref{lemma:sim diag},
%% taking $C = (\fwd \prcov \fwd^*)^{-1}$ we observe that $\meas_j,
%% j=1,...,m$ are eigenvectors of $\fwd \prcov \fwd^*$. If they have
%% different eigenvalue, they are orthogonal. If not, this means, by
%% the common interpretation of the Lagrange multipliers, that an
%% infinitesimal increase in one will cause the same increase in
%% design criterion as others with the same eigenvalue. So if these
%% eigenvectors do not have the same angle between every pair, an
%% infinitesimal increase in one will give more increase than
%% others, which is a contradiction. So in every subspace, vectors
%% will be restricted to these forms and if there will be more
%% vectors than the dimensionality of the subspace, clusterization
%% will occur.

%% \begin{figure}
%%   \begin{tikzpicture}[thick, scale=1.4, every node/.style={scale=0.99}]
%%     \begin{axis}
%%       [
%%         xmin = 0,
%%         xmax = 3.14,
%%         xlabel = {$x$},
%%         ylabel = posterior std,
%%         ymin   = 0,
%%         %compat = 1.3,
%%         % ymax   = 130,
%%         % ytick = \empty,
%%         legend cell align=left,
%%         % legend style={at={(0.45,0.2)}}
%%         legend pos= outer north east 
%%       ]
%%       % \draw[black!30!white, thin] (50,0) -- (50,130);
%%       % 
%%       \addplot [thin, black, mark=none] table{data/stdv-trunc-sens1-var1.txt};
%%       \addlegendentry{1 sensor};

%%       \addplot [thin, blue, mark=none] table{data/stdv-trunc-sens2-var1.txt};
%%       \addlegendentry{2 sensors};

%%       \addplot [thin, red, mark=none] table{data/stdv-trunc-sens3-var1.txt};
%%       \addlegendentry{3 sensors};

%%       \addplot [thin, green, mark=none] table{data/stdv-trunc-sens4-var1.txt};
%%       \addlegendentry{4 sensors};

%%       \addplot [thin, purple, mark=none] table{data/stdv-trunc-sens5-var1.txt};
%%       \addlegendentry{5 sensors};

%%       \addplot [thin, cyan, mark=none] table{data/stdv-trunc-sens6-var1.txt};
%%       \addlegendentry{6 sensors};

%%       \addplot [black,  only marks, mark=x, mark size=1.5] 
%%       table{data/locs-trunc-sens1-var1.txt}; 
%%       \addplot [blue,   only marks, mark=x, mark size=1.5]
%%       table{data/locs-trunc-sens2-var1.txt}; 
%%       \addplot [red,    only marks, mark=x, mark size=1.5]
%%       table{data/locs-trunc-sens3-var1.txt};
%%       \addplot [green,  only marks, mark=x, mark size=1.5] 
%%       table{data/locs-trunc-sens4-var1.txt}; 
%%       \addplot [purple, only marks, mark=x, mark size=1.5] 
%%       table{data/locs-trunc-sens5-var1.txt}; 
%%       \addplot [cyan,   only marks, mark=x, mark size=1.5] 
%%       table{data/locs-trunc-sens6-var1.txt}; 
%%     \end{axis}
%%   \end{tikzpicture}
%%   \caption{Analogously to figure \ref{fig:clusterization
%%       illustration}, shown are posterior pointwise standard
%%     deviations. Note the exact clustering. Here the dynamics is not
%%     the one governed by the 1D heat equation but a dynamics that sets
%%     to zero all modes except the first three. This dynamics simulates
%%     a finite-rank operator as per the conditions of Theorem
%%     \ref{theorem:finite rank}. Although the conditions of the theorem
%%     allow the use of a different constraint on measurements (see
%%     section \ref{subsec:unit length}), the conclusion of the theorem
%%     still seems to hold.}
%%   \label{fig:finite rank illustration}
%% \end{figure}

%% \section{Counterexample}\label{section:counterexample}
%% Applying Lemma \ref{lemma:sim diag} to
%% \begin{equation*}
%%   (\fwd \postcov \fwd^*)^{-1} = (\fwd \prcov \fwd^*)^{-1} +
%%   \sigma^{-2}\obs^*\obs
%% \end{equation*}
%% implies $(\fwd \prcov \fwd^*)^{-1}$ and $\obs^*\obs$ have the same
%% eigenspaces. Since $\meas_j$ are eigenvectors of $\fwd \postcov
%% \fwd^*$, we may expect this implies that $\meas_j$ are also
%% eigenvectors of $(\fwd \prcov \fwd^*)^{-1}$ and assuming $(\fwd \prcov
%% \fwd^*)^{-1}$ has simple eigenvalues completes the proof. Our
%% expectations, however, are let down.

%% Here is a counterexample. Let $e_1, e_2$ denote two
%% eigenvectors of $C := (\fwd \prcov \fwd^*)^{-1}$ of different
%% eigenvalues such that $C + \obs^*\obs$ is scalar on span
%% $V:=\{e_1,e_2\}$ (we took $\sigma^2 = 1$ for simplicity). Then,
%% on $V$, we have  
%% \begin{equation*}
%%   (\fwd \prcov \fwd^*)^{-1} = diag( d_1, d_2),
%% \end{equation*}
%% and we assume $d_1 > d_2$ without loss of generality. For $\meas \in
%% \hil$, we denote $\meas^*$ its Riesz representer, making a departure
%% from the standard notation employed, where $\meas$ is usually taken to
%% be a linear functional on $\hil$. Let $\obs = (\meas_1^*, \meas_2^*,
%% \meas_3^*)^t$. In this setup, we would like that $\meas_j,j=1,2,3$ are
%% either orthogonal or parallel. We now outline construction that shows
%% this is not necessarily true.

%% We know, from Corollary \ref{cor:sim diag} that $\meas_j \in V$, so
%% let
%% \begin{align*}
%%   \meas_1 &=  c_1e_1 + c_2e_2 \\
%%   \meas_2 &= -c_1e_1 + c_2e_2 \\
%%   \meas_3 &=     e_1.
%% \end{align*}
%% Then
%% \begin{align*}
%%   \meas_1 \meas_1^*
%%   &= c_1^2e_1e_1^* + c_1c_2( e_1e_2^* + e_2e_1^*) +
%%   c_2^2 e_2e_2^* \\
%%   \meas_2 \meas_2^* &= c_1^2e_1e_1^* - c_1c_2( e_1e_2^* +
%%   e_2e_1^*) + c_2^2 e_2e_2^*.
%% \end{align*}
%% Note that
%% \begin{equation*}
%%   \obs^* \obs = \meas_1 \meas_1^* + \meas_2\meas_2^* +
%%   \meas_3\meas_3^* = (2c_1^2+1) e_1e_1^* + 2c_2^2e_2e_2^*,
%% \end{equation*}
%% is diaonal in $\{e_1,e_2\}$. In order for $\meas_j,j=1,2,3$ to be
%% eigenvectors of $C + \obs^*\obs$ with the same eigenvalue, $C +
%% \obs^*\obs$ must be a scalar operator on $V$. Thus the following
%% must hold:
%% \begin{align*}
%%   d_1 + 2c_1^2 + 1 &= d_2 + 2c_2^2 \\
%%   c_1^2 + c_2^2 &= 1 \text{ (arises from $\| \meas_i \| = 1$).}
%% \end{align*}
%% Equivalently,
%% \begin{equation*}
%%   \begin{bmatrix}
%%     2 & -2 \\
%%     1 & 1
%%   \end{bmatrix}
%%   \begin{bmatrix}
%%     c_1^2 \\
%%     c_2^2
%%   \end{bmatrix}
%%   =
%%   \begin{bmatrix}
%%     d_2 - d_1 - 1 \\
%%     1
%%   \end{bmatrix}.
%% \end{equation*}
%% We can always find a solution for this system, for any $d_1,d_2$.
%% However, it is necessary that either $c_1^2 = 1$ or $c_2^2 = 1$ for
%% $\meas_j, j=1,2,3$ to all be either orthogonal or parallel. The first
%% conditiion implies $d_2 - d_1 = 3$, which is impossible, since $d_1 >
%% d_2$. The second requires $d_1 - d_2 = 1$, a condition that does not
%% hold generically.

%    Text of article.

%    Bibliographies can be prepared with BibTeX using amsplain,
%    amsalpha, or (for "historical" overviews) natbib style.
\bibliographystyle{amsplain}
\bibliography{refs.bib,georg_refs}

\end{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%    Templates for common elements of an article; for additional
%    information, see the file Author_Handbook_ProcColl.pdf, included
%    in every AMS author package, and the amsthm user's guide, linked
%    from http://www.ams.org/tex/amslatex.html .

%    Section headings
\section{}
\subsection{}

%    Ordinary theorem and proof
\begin{theorem}[Optional addition to theorem head]
% text of theorem
\end{theorem}

\begin{proof}[Optional replacement proof heading]
% text of proof
\end{proof}

%    Figure insertion; default placement is top; if the figure occupies
%    more than 75% of a page, the [p] option should be specified.
\begin{figure}
\includegraphics{filename}
\caption{text of caption}
\label{}
\end{figure}

%    Mathematical displays; for additional information, see the amsmath
%    user's guide:  texdoc amamath  or
%  http://mirror.ctan.org/tex-archive/macros/latex/required/amsmath/amsldoc.pdf

% Numbered equation
\begin{equation}
\end{equation}

% Unnumbered equation
\begin{equation*}
\end{equation*}

% Aligned equations
\begin{align}
  &  \\
  &
\end{align}

%-----------------------------------------------------------------------
% End of article-template.tex
%-----------------------------------------------------------------------

### 2020-07-20 paper.tex
Add tikz image

@@
 $\fwd \prcov \fwd^*$ by $\lambda_1 \geq \lambda_2 \geq ...$ and its
-eigenvectors $\{\ev_i\}_{i=1}^{\infty}$. First, let us present a lemma:
+eigenvectors $\{\ev_i\}_{i=1}^{\infty}$. The following lemma will be useful:
 
@@
 \end{equation}
+
+\pgfplotstableread{
+  Label          prior  $a_1$  $a_2$   $a_3$   $a_4$   $a_5$         $a_6$    topper
+  $\ev_1$         0.2    0.8    0.5     0.5    0.4    0.33333333  0.3333333  0.001
+  $\ev_2$         0.8    0.2    0.5     0.5    0.4    0.33333333  0.3333333  0.001
+  $\ev_3$         2.2    0      0       0      0.2    0.33333333  0.3333333  0.001
+  $\ev_4$         3.5    0      0       0      0      0           0          0.001
+}\testdata
+
+\begin{figure}
+  \begin{tikzpicture}
+    \begin{axis}[
+        ybar stacked,
+        ymin=0,
+        ymax=4,
+        xtick=data,
+        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
+        reverse legend=false, % set to false to get correct display, but I'd like to have this true
+        xticklabels from table={\testdata}{Label},
+        xticklabel style={text width=2cm,align=center},
+        legend plot pos=right,
+        ylabel=prior and posterior precision,
+        xlabel=eigenvector,
+      ]
+
+      \addplot [fill=green!80]  table [y=prior, meta=Label, x expr=\coordindex] {\testdata};
+      \addplot [fill=blue!60]   table [y=$a_1$, meta=Label, x expr=\coordindex] {\testdata};
+      \addplot [fill=red!60]    table [y=$a_2$, meta=Label, x expr=\coordindex] {\testdata};
+      \addplot [fill=black!60]  table [y=$a_3$, meta=Label, x expr=\coordindex] {\testdata};
+      \addplot [fill=orange!60] table [y=$a_4$, meta=Label, x expr=\coordindex] {\testdata};
+      \addplot [fill=cyan!60]   table [y=$a_5$, meta=Label, x expr=\coordindex] {\testdata};
+      \addplot [fill=purple!60] table [y=$a_6$, meta=Label, x expr=\coordindex] {\testdata};
+
+      \addlegendentry{prior}
+      \addlegendentry{$a_1$}
+      \addlegendentry{$a_2$}
+      \addlegendentry{$a_3$}
+      \addlegendentry{$a_4$}
+      \addlegendentry{$a_5$}
+      \addlegendentry{$a_6$}   
+    \end{axis}
+  \end{tikzpicture}
+  \caption{Posterior precision per eigenvector. Each measurement
+    $\func_i$ increases precision uniformly across the lowest precision
+    eigenvectors. We see that $\func_2$ and $\func_3$ are repeated
+    measurements, as well as $\func_5$ and $\func_6$.}
+  \label{fig:clusterization}
+\end{figure}
 with the norm constraints on $\meas_j,j=1,...,m$. We show in the
@@
 \subsection{Sensor Clusterization}\label{subsec:clusterization}
-The symmetry mentioned in last paragraph of the previous section may
-very well break in real life problems. If this happens and eigenvalues
-of \eqref{eq:eigenproblem} are all different, then showing
-sensors do cluster proceeds as follows. First, we show there
-are only so many different equal, by the pigeonhole principle. In
-order to make this argument work, one needs to show this eigenproblem
-does, indeed, have only finitely many eigenvalues. This is done by
-first recalling that $\fwd$ is strongly smoothing (so $\fwd \postcov
-\fwd^*$ has fast decaying eigenvalues) and then showing that small
-eigenvalues arise from un-informative measurements. We use the
-following corollary, whose proof may be found in the appendix.
-%
-\begin{corollary}[Gain for No Model Error]\label{cor:zero mod err}
-  If $\modcov = 0$, then
-  \begin{equation*}
-    \tar( \obs ) - \tar (\obsm )
-    = -\frac12 \log (1 - \sigma^{-2} \langle \fwd \postcov \fwd^* \meas_m, \meas_m \rangle ).
-  \end{equation*}
-\end{corollary}
-%
-Suppose $\meas_j$ is a measurement with a small eigenvalue, i.e.\
-\begin{equation}\label{eq:small eigenvalue}
-  \sigma^{-2} \fwd \postcov \fwd^* \meas_j = \xi_j \meas_j,\ \xi_j \ll 1.
-\end{equation}
-By Corollary \ref{cor:zero mod err}, the increase in the design
-criterion is
-\begin{align*}
-  \tar(\obs) - \tar(\obsm) &=
-  -\frac12 \log (1- \sigma^{-2} \langle \fwd
-  \postcovm \fwd^* \meas_j, \meas_j \rangle) \\
-  %
-  %
-  %
-  &= - \frac12 \log (1 - \xi_j) \text{ since $\|\meas_j\|=1$ }\\
-  %
-  %
-  %
-  &\approx \frac{\xi_j}{2}.
-\end{align*}
-Thus, negligibly small eigenvalues give negligible increase in the
-design criterion and will not be a part of a truly optimal
-design. Since the eigenvalues of $\fwd \postcov \fwd^*$ decay faster
-than those of $\fwd\prcov\fwd^*$, the latter will determine how many
-non-negligible eigenvalues $\fwd \postcov \fwd^*$ has. Thus, in the
-eigenproblem \eqref{eq:eigenproblem}, we need only consider a finite
-number of eigenvectors.
-
-At this point, one would like to invoke the pigeonhole principle and
-claim that eigenvectors cluster. This, however, is not as
-straightforward as one may hope. If $\fwd \postcov \fwd^*$ only has
-simple eigenvalues, then we may use the pigeonhole principle and
-conclude that some eigenvectors are clustered (up to a sign). However,
-if $\fwd \postcov \fwd^*$ has non-simple eigenvalues, one may hope to
-somehow prove measurement vectors are either orthogonal or parallel,
-which suffices for proving clusterization but that is not true.
-
-\subsection{Non-Vanishing Model Error}
+The way clusterization can occur is best understood from figure
+\ref{fig:clusterization}.
+
+%% The symmetry mentioned in last paragraph of the previous section may
+%% very well break in real life problems. If this happens and eigenvalues
+%% of \eqref{eq:eigenproblem} are all different, then showing
+%% sensors do cluster proceeds as follows. First, we show there
+%% are only so many different equal, by the pigeonhole principle. In
+%% order to make this argument work, one needs to show this eigenproblem
+%% does, indeed, have only finitely many eigenvalues. This is done by
+%% first recalling that $\fwd$ is strongly smoothing (so $\fwd \postcov
+%% \fwd^*$ has fast decaying eigenvalues) and then showing that small
+%% eigenvalues arise from un-informative measurements. We use the
+%% following corollary, whose proof may be found in the appendix.
+%% %
+%% \begin{corollary}[Gain for No Model Error]\label{cor:zero mod err}
+%%   If $\modcov = 0$, then
+%%   \begin{equation*}
+%%     \tar( \obs ) - \tar (\obsm )
+%%     = -\frac12 \log (1 - \sigma^{-2} \langle \fwd \postcov \fwd^* \meas_m, \meas_m \rangle ).
+%%   \end{equation*}
+%% \end{corollary}
+%% %
+%% Suppose $\meas_j$ is a measurement with a small eigenvalue, i.e.\
+%% \begin{equation}\label{eq:small eigenvalue}
+%%   \sigma^{-2} \fwd \postcov \fwd^* \meas_j = \xi_j \meas_j,\ \xi_j \ll 1.
+%% \end{equation}
+%% By Corollary \ref{cor:zero mod err}, the increase in the design
+%% criterion is
+%% \begin{align*}
+%%   \tar(\obs) - \tar(\obsm) &=
+%%   -\frac12 \log (1- \sigma^{-2} \langle \fwd
+%%   \postcovm \fwd^* \meas_j, \meas_j \rangle) \\
+%%   %
+%%   %
+%%   %
+%%   &= - \frac12 \log (1 - \xi_j) \text{ since $\|\meas_j\|=1$ }\\
+%%   %
+%%   %
+%%   %
+%%   &\approx \frac{\xi_j}{2}.
+%% \end{align*}
+%% Thus, negligibly small eigenvalues give negligible increase in the
+%% design criterion and will not be a part of a truly optimal
+%% design. Since the eigenvalues of $\fwd \postcov \fwd^*$ decay faster
+%% than those of $\fwd\prcov\fwd^*$, the latter will determine how many
+%% non-negligible eigenvalues $\fwd \postcov \fwd^*$ has. Thus, in the
+%% eigenproblem \eqref{eq:eigenproblem}, we need only consider a finite
+%% number of eigenvectors.
+
+%% At this point, one would like to invoke the pigeonhole principle and
+%% claim that eigenvectors cluster. This, however, is not as
+%% straightforward as one may hope. If $\fwd \postcov \fwd^*$ only has
+%% simple eigenvalues, then we may use the pigeonhole principle and
+%% conclude that some eigenvectors are clustered (up to a sign). However,
+%% if $\fwd \postcov \fwd^*$ has non-simple eigenvalues, one may hope to
+%% somehow prove measurement vectors are either orthogonal or parallel,
+%% which suffices for proving clusterization but that is not true.
+
+\section{Analysis of Optimal Designs --- Non-Vanishing Model Error}
 We can build on Lemma \ref{lemma:design increase} to find the gain in design
@@
 small measurement error levels, the clusterization effect will be
-canceled by the presence of a non-zero model error. Since the design
+mitigated by the presence of a non-zero model error. Since the design
 criterion is not defined for $\sigma^2 = 0$ and identical
@@
 in an ideal settings, we should not reduce uncertainty in any
-direction before it is the largest uncertainty present.
-
-We believe the model presented is not too far from reality. After all,
-taking any measurement (point evaluation or other) is just recording
-the value of a linear combination of individual eigenvectors of the
-prior. In light of this, we believe the reason for clusterization is close
-to the argument made in section \ref{subsub:first attempt}. The reason
-we observe clusterization in real life problems is probably due to
-symmetry breaking, although we cannot show that rigorously.
+direction (eigenvector) before it is the largest uncertainty present.
+
+%% We believe the model presented is not too far from reality. After all,
+%% taking any measurement (point evaluation or other) is just recording
+%% the value of a linear combination of individual eigenvectors of the
+%% prior. In light of this, we believe the reason for clusterization is close
+%% to the argument made in section \ref{subsub:first attempt}. The reason
+%% we observe clusterization in real life problems is probably due to
+%% symmetry breaking, although we cannot show that rigorously.
 
@@
 \end{enumerate}
+
+\section{Acknowledgements}
+This study was conducted during when the author was a PhD student of
+Prof. Georg Stadler in New York University's Courant Institute.
 

### 2020-07-20 paper.tex
Could it be? Is it done???

@@
 
-\section{Analysis of Optimal Designs --- Vanishing Model Error}
+\section{Analysis of Optimal Designs --- Vanishing Model Error}\label{section:vanishing}
 When $\modcov = 0$, then $\Sigma= \sigma^2I$. The necessary
@@
   \caption{Posterior precision per eigenvector. Each measurement
-    $\func_i$ increases precision uniformly across the lowest precision
+    icreases precision uniformly across the lowest precision
     eigenvectors. We see that $\func_2$ and $\func_3$ are repeated
@@
 direction (eigenvector) before it is the largest uncertainty present.
-
+We show how and why repeated measurements can give rise to a D-optimal
+design. There is more work to be done in understanding exactly what
+causes such designs to be strictly better than others.
 %% We believe the model presented is not too far from reality. After all,
@@
 
-An important consequence of the argument in \ref{section:vanishing} is
+A consequence of the argument in section \ref{section:vanishing} is
 that we can find an approximately optimal design by means of linear
@@
      $\sigma^2$.
-   \item Decide how many measurements you can take and denote that
-     number $m$. This is the total precision one can allocate to the
-     different eigenvectors. Next, we see how to allocate these.
+   \item Decide how many measurements can be taken and denote that
+     number $m$. This is the total precision that can be allocated to
+     the different eigenvectors.
     \item Find $k$ and $\{\eta_i\}_{i=1}^k$, $\eta_i > 0, i=1,...,k$
@@
 \section{Acknowledgements}
-This study was conducted during when the author was a PhD student of
-Prof. Georg Stadler in New York University's Courant Institute.
+This study was conducted during when I was a PhD student of
+Prof. Georg Stadler in New York University's Courant Institute. I
+would like to thank him for his great mentorship.
 

### 2020-07-22 paper.tex
Before I delete a bunch of stuff

@@
-%-----------------------------------------------------------------------
-% Beginning of article-template.tex
-%-----------------------------------------------------------------------
-%
-%    This is a template file for proceedings articles prepared with AMS
-%    author packages, for use with AMS-LaTeX.
-%
-%    Templates for various common text, math and figure elements are
-%    given following the \end{document} line.
-%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%    Remove any commented or uncommented macros you do not use.
-
-%    Replace amsproc by the name of the author package.
-\documentclass{amsproc}
-
-%    If you need symbols beyond the basic set, uncomment this command.
-%\usepackage{amssymb}
-
-%    If your article includes graphics, uncomment this command.
-%\usepackage{graphicx}
-
-%    If the article includes commutative diagrams, ...
-%\usepackage[cmtip,all]{xy}
-
-%    Include other referenced packages here.
+\documentclass{amsart}
 \input{definitions}
 
-%    Update the information and uncomment if AMS is not the copyright
-%    holder.
-%\copyrightinfo{2009}{American Mathematical Society}
-
-%% \newtheorem{theorem}{Theorem}[section]
-%% \newtheorem{lemma}[theorem]{Lemma}
-
-%% \theoremstyle{definition}
-%% \newtheorem{definition}[theorem]{Definition}
-%% \newtheorem{example}[theorem]{Example}
-%% \newtheorem{xca}[theorem]{Exercise}
-
-%% \theoremstyle{remark}
-%% \newtheorem{remark}[theorem]{Remark}
-
 \numberwithin{equation}{section}
@@
 
-% \title[short text for running head]{full title}
-\title{A tractable model for D-optimal design in infinite dimensional
-  Bayesian inverse problems.}
-
-%    Only \author and \address are required; other information is
-%    optional.  Remove any unused author tags.
-
-%    author one information
-% \author[short version for running head]{name for top of paper}
+\title[Sensor Clusterization in D-optimal design in infinite
+  dimensions]{Sensor Clusterization in D-optimal Design in Infinite
+  Dimensional Bayesian Inverse Problems}
+
 \author{Yair Daon}
@@
 show this constrained optimization problem is in fact a set of
-eigenvector-eigenvalue problems \eqref{eq:eigenproblem} and suggest a
-natural way for understanding sensor-clusterization in section
-\ref{subsub:first attempt}. We explain why this approach fails and
-show the reason is that (perhaps surprisingly), all eigenvalues are
+eigenvector-eigenvalue problems \eqref{eq:eigenproblem}. all eigenvalues are
 equal --- this is done in section \ref{subsub:characterization}. An
@@
   Label    prior  $a_1$  $a_2$   $a_3$   $a_4$   $a_5$         $a_6$    topper
-  $\ev_1$         0.2    0.8    0.5     0.5    0.4    0.33333333  0.3333333  0.001
-  $\ev_2$         0.8    0.2    0.5     0.5    0.4    0.33333333  0.3333333  0.001
-  $\ev_3$         2.2    0      0       0      0.2    0.33333333  0.3333333  0.001
-  $\ev_4$         3.5    0      0       0      0      0           0          0.001
+  1         0.2    0.8    0.5     0.5    0.4    0.33333333  0.3333333  0.001
+  2         0.8    0.2    0.5     0.5    0.4    0.33333333  0.3333333  0.001
+  3         2.2    0      0       0      0.2    0.33333333  0.3333333  0.001
+  4         3.5    0      0       0      0      0           0          0.001
 }\testdata
@@
         legend plot pos=right,
-        ylabel=prior and posterior precision,
+        ylabel=precision --- prior and posterior,
         xlabel=eigenvector,
@@
   \end{tikzpicture}
-  \caption{Posterior precision per eigenvector. Each measurement
+  \caption{Posterior precision per eigenvector. Each measurement $\func_i$
     icreases precision uniformly across the lowest precision
@@
 The way clusterization can occur is best understood from figure
-\ref{fig:clusterization}.
+\ref{fig:clusterization}. We see that each measurement increases
+precision (reduces uncertainty) only for the eigenvectors of lowest
+precision. If the precision in these is lower than the precision for
+the next eigenvector, a repeated measurement is taken.
 
@@
 \section{Acknowledgements}
-This study was conducted during when I was a PhD student of
-Prof. Georg Stadler in New York University's Courant Institute. I
-would like to thank him for his great mentorship.
+This study is a part of my PhD thesis \cite{mine} which was written
+under the instruction of Prof. Georg Stadler in New York University's
+Courant Institute. I would like to thank him for his great mentorship.
 

### 2020-07-22 paper.tex
spell check

@@
 \cite{Silvey13,Ucinski05} as well as for Bayesians
-\cite{ChalonerVerdinelli95}, and the analogy between the two
-approaches was investigated in \cite{ChalonerVerdinelli95}. Although
+\cite{ChalonerVerdinelli95}, see \cite{ChalonerVerdinelli95} for an
+investigation of the analogy between the two approaches. Although
 there is a plethora of design criteria, in the current work we will
@@
 using the D-optimal criterion (and others that will not be
-investigated here), the optimization process results in measurement
-locations are very similar. Specifically, if a measurement is thought
-of as measuring some function value at $\x \in \Omega \subseteq \R^d,
+investigated here), the optimization process results in measurements
+that are very similar. For example, if a measurement is thought of as
+measuring some function value at $\x \in \Omega \subseteq \R^d,
 d=1,2,3$ (with added error) then the optimization procedure sometimes
@@
 where the user enforces measurement locations to be far from each
-other. One way to do this is by introducing correlated error which,
+other. One way to do this is by introducing correlated errors which,
 philosophically, accounts for both measurement error and model error.
@@
 \subsection{Contribution}
-In this work we propose and thoroughly study an analytically tractable
-model for understanding D-optimal designs. The initial motivation was
-to understand the causes of sensor-clusterization and the end result
-is interesting in its own. Specifically, in section \ref{subsec:unit
-  length} we suggest a somewhat simplified model for experimental
-design that is analytically tractable. In section \ref{section:D and
-  grad} we derive a constrained optimization problem for D-optimal
-design using Lagrange multipliers. In the case of no model error we
-show this constrained optimization problem is in fact a set of
-eigenvector-eigenvalue problems \eqref{eq:eigenproblem}. all eigenvalues are
-equal --- this is done in section \ref{subsub:characterization}. An
-important consequence is that we know the value of the best design
-criterion without any optimization and which measurements give rise to
-it. Practically, we can use this information to seek a design in the
-real problem we are interested in and we outline this in section
-\ref{section:conclusion}.
+We propose and thoroughly study an analytically tractable model for
+understanding D-optimal designs. In this model, an optimal design
+arises as a solution to a constrained optimization problem. This
+problem is solved using Lagrange multipliers (section \ref{section:D
+  and grad}). In the case of no model error we show this problem is in
+fact a set of eigenvector-eigenvalue problems (equation
+\eqref{eq:eigenproblem}). We then characterize an optimal design as
+one that reduces the highest uncertainties, and show how that can give
+rise to repeated measurements (figure \ref{fig:clusterization}).
+
+The solution to the this could be used to seek a design in the real
+problem we are interested in and we outline this in section
+\ref{section:conclusion}. The analysis that culminates in equation
+\eqref{eq:true target} also gives us an upper bound for the design
+criterion (denoted $\tar$) below.
 
@@
 
-Based on the analysis presented, we show why clusterization is a
-reasonable thing to expect in the case of independent measurements.
-We further show how considering an observation model that is spatially
-correlated makes an optimal design clusterization-free.
-
-\subsection{Limitations}\label{subsec:limitations}
-There are two main drawbacks of the arguments presented here. First,
-one can argue that we allow considerably less restrictive observations
-than any real-life problem does. This is because our observation model
-allows taking measurements that are arbitrarily weighted linear
-combinations of eigenvectors. The second drawback is that we do not
-show rigorously that clusterization occurs in our model.
-
-\subsection{Illustration and Example}\label{subsec:example}
+%% \subsection{Limitations}\label{subsec:limitations}
+%% There are two main drawbacks of the arguments presented here. First,
+%% one can argue that we allow considerably less restrictive observations
+%% than any real-life problem does. This is because our observation model
+%% allows taking measurements that are arbitrarily weighted linear
+%% combinations of eigenvectors. The second drawback is that we do not
+%% show rigorously that clusterization necessarily occurs in our model,
+%% we only show that it can occur.
+
+\subsection{An Example of Clusterization}\label{subsec:example}
 \begin{figure}
@@
 equation in $[0,\pi]$ with homogeneous Dirichlet boundary
-conditions. The 1D heat equation is:
+conditions.
+
+The 1D heat equation is:
 \begin{subequations}\label{eq:heat equation}
@@
 \end{subequations}
+
 We would like to infer the initial condition $u_0$. For that purpose,
@@
 We will consider the information-theoretic design criterion presented
-in section \ref{section:OED intro}. Thus, we choose $\x_j,j=1,...,m$ so
-as to minimize an expression analogous to the log-determinant of the
-posterior covariance operator (corresponding to a D-optimal
+in section \ref{section:OED intro}. Thus, we choose $\x_j,j=1,...,m$
+so as to minimize an expression analogous to the log-determinant of
+the posterior covariance operator (corresponding to a D-optimal
 design). We can see the clusterization effect in figure
-\ref{fig:clusterization illustration}. For low numbers of sensors, the
-design criterion suggests placing sensors in locations that are
-close.
-
-\subsection{Overview}\label{subsec:plan}
-Half of this paper is devoted to finding the Lagrange multipliers
-problem and culminates in \eqref{eq:conditions}. The reader may skip
-directly there if they are willing to take the rest on faith. The more
-interesting stuff may be found in section \ref{section:analysis}.
-
-Seeking an optimal design amounts to maximizing $\tar(\obs)$ in
-$\obs$, subject to constraints posed on $\obs$. These enter as
-constraints on the allowed set of measurement vectors we can choose
-from --- the rows of $\obs$. These constraints are determined by the
-problem structure --- for example, in the 1D heat equation example
-considered in section \ref{subsec:example}, we were constrained to
-choose $\obs = (\meas_1,...,\meas_m)^t$ with $\meas_j = \delta_{\x_j}$
-for some $\x_j \in [0,\pi]$. We present an approach to understanding
-optimal experiment design based on the first-order necessary condition for
-optimal design, where the constraints are incorporated using Lagrange
-multipliers. This condition is used to understand what causes
-clusterization. One problem with such an approach is that the set we
-can choose $\obs$ from is complicated and difficult to work with. For
-example, if $\ban = C(\Omega)$ for some compact $\Omega$, then the
-allowed set of measurement vectors may only be point evaluations --- a
-set that is not easy to describe as a level set of some simple
-function, a description crucial for the use of Lagrange multipliers.
-In order to facilitate analysis, we consider an easier constraint in
-section \ref{subsec:unit length}. Reverting to this simpler constraint
-is a limitation of our analysis but it provides some insight
-into the cause of clusterization.
-
-\section{Preliminaries}
+\ref{fig:clusterization illustration} --- the design criterion
+suggests placing sensors in locations that are extremely close.
+
+\subsection{Remarks to the Reader}\label{subsec:plan}
+The Lagrange multipliers problem for iid measurement noise is
+discussed in section \ref{section:vanishing}. In that section, it is
+shown how sensor clusterization occurs. It is advised that a reader
+less comfortable with Hilbert and Banach spaces mostly ignore these
+terms and consider $\hil$ and $\ban$ finite dimensional vector spaces.
+
+%% Seeking an optimal design amounts to maximizing $\tar(\obs)$ in
+%% $\obs$, subject to constraints posed on $\obs$. These enter as
+%% constraints on the allowed set of measurement vectors we can choose
+%% from --- the rows of $\obs$. These constraints are determined by the
+%% problem structure --- for example, in the 1D heat equation example
+%% considered in section \ref{subsec:example}, we were constrained to
+%% choose $\obs = (\meas_1,...,\meas_m)^t$ with $\meas_j = \delta_{\x_j}$
+%% for some $\x_j \in [0,\pi]$. We present an approach to understanding
+%% optimal experiment design based on the first-order necessary condition
+%% for optimal design, where the constraints are incorporated using
+%% Lagrange multipliers. This condition is used to understand what causes
+%% clusterization. One problem with such an approach is that the set we
+%% can choose $\obs$ from is complicated and difficult to work with. For
+%% example, if $\ban = C(\Omega)$ for some compact $\Omega$, then the
+%% allowed set of measurement vectors may only be point evaluations --- a
+%% set that is not easy to describe as a level set of some simple
+%% function, a description crucial for the use of Lagrange multipliers.
+%% In order to facilitate analysis, we consider an easier constraint in
+%% section \ref{subsec:unit length}. Reverting to this simpler constraint
+%% is a limitation of our analysis but it provides some insight into the
+%% cause of clusterization.
+
+\section{Preliminaries and Notation}\label{section:prelim}
 
@@
 
-\subsection{bayesian inverse problems}\label{section:abstract OED}
+\subsection{Bayesian Inverse Problems}\label{section:abstract OED}
 Let $\hil$ be a Hilbert space and $\ban$ a Banach space, both
@@
 \end{equation}
-where the first equality in \ref{eq:modcov explained} holds by
+where the first equality in \eqref{eq:modcov explained} holds by
 definition and the second by \eqref{eq:obs*}. The explicit dependence
@@
 \begin{equation}\label{eq:eigenproblem}
-  \sigma^{-2}\fwd \postcov \fwd^* \obs^* = \obs^* \Xi. 
+  \sigma^{-2}\fwd \postcov \fwd^* \obs^* = \obs^* \Xi, 
 \end{equation}
-This is an eigenvalue problem for the self-adjoint operator
-$\sigma^{-2}\fwd \postcov \fwd^*$ where the rows of $\obs$, namely
-$\meas_j,j=1,...m$, are the eigenvectors. If $\fwd$ is invertible,
-then we may write:
+with $\Xi$ diagonal. This is an eigenvalue problem for the
+self-adjoint operator $\sigma^{-2}\fwd \postcov \fwd^*$ where the rows
+of $\obs$, namely $\meas_j,j=1,...m$, are the eigenvectors. If $\fwd$
+is invertible, then we may write:
 \begin{align}\label{eq:mod conditions}
@@
 precision in $\ban$. This expression makes sense even if $\fwd$ is not
-invertible --- consider $\fwd: \ker \fwd^{\perp} \to \text{Im} \fwd
-\subseteq \ban$ and this does not change the optimal design problem at
+invertible --- consider instead $\fwd$ restricted to $\ker
+\fwd^{\perp}$. This does not change the optimal design problem at
 all, since we will never take measurements that have a component in
-$\fwd^{\perp}$ (this is intuitive by Occam's razor but can also be
+$\ker \fwd$ (this is intuitive by Occam's razor but can also be
 understood from Corollary \ref{cor:zero mod err}).
@@
 Therefore, an optimal design maximizes
-\begin{equation}
+
+\begin{equation}\label{eq:true target}
   \sum_{i=1}^{k}\log (\lambda_i^{-1} + \sigma^{-2} \eta_i),
@@
   \caption{Posterior precision per eigenvector. Each measurement $\func_i$
-    icreases precision uniformly across the lowest precision
+    increases precision uniformly across the lowest precision
     eigenvectors. We see that $\func_2$ and $\func_3$ are repeated
@@
 with the norm constraints on $\meas_j,j=1,...,m$. We show in the
-appedix (Lemma \ref{lemma:free}) we can find $\meas_1,...,\meas_m$
+appendix (Lemma \ref{lemma:free}) we can find $\meas_1,...,\meas_m$
 such that $\{\eta_i\}_{i=1}^k$ have any value we desire, subject only

### 2020-07-23 paper.tex
Submitted to arxivgit add paper.tex git add paper.tex git add paper.tex git add paper.tex git add paper.tex git add paper.tex git add paper.tex git add paper.tex !

@@
 \documentclass{amsart}
+
 \input{definitions}
@@
 \author{Yair Daon}
-\address{Courant Institute}
+\address{Porter School of the Environment and Earth
+  Sciences, Tel Aviv University\\ Tel Aviv, Israel}
+%% \curraddr{Porter School of Environment Studies, Tel Aviv University\\ Tel Aviv, Israel}
 %% \curraddr{}
-%% \email{yair.daon@gmail.com}
+\email{yair.daon@gmail.com}
 %% \thanks{}
@@
   35R30, % PDE - Inverse problems
-  65C50; % Numerical analysis - other computational problems in probability
   Secondary:
@@
 }
-
+%% Primary: 62F15, 35R30; Secondary: 28C20
 \date{\today}
@@
   spatially correlated measurement error clusterization does not
-  occur. Our analysis suggests a novel method for finding optimal
-  designs. As a part of the analysis we prove a matrix determinant
+  occur. As a part of the analysis we prove a matrix determinant
   lemma analog in infinite dimensions, as well as a lemma for
@@
 d=1,2,3$ (with added error) then the optimization procedure sometimes
-places sensors in close proximity to each other (as can be seen in
-figure \ref{fig:clusterization illustration}). Following
+places sensors in close proximity to each other (as can be seen
+in figure \ref{fig:clusterization illustration}). Following
 \cite{Ucinski05}, we refer to this phenomenon as \emph{sensor
@@
 sensors away from each other. Yet, none offers any insight as to why
-clusterization occurs. Also, as better models are employed, the model
-error decreases and the clusterization phenomenon will reappear. While
-these approaches are practical and help us avoid the problem, they do
-not provide insight as to why sensors are clustering.
+clusterization occurs. Also, as better models are employed, model
+error is decreased and the clusterization phenomenon will eventually
+reappear. While these approaches are practical and help us avoid the
+problem, they do not provide insight as to why sensors are clustering.
 
@@
 understanding D-optimal designs. In this model, an optimal design
-arises as a solution to a constrained optimization problem. This
-problem is solved using Lagrange multipliers (section \ref{section:D
-  and grad}). In the case of no model error we show this problem is in
-fact a set of eigenvector-eigenvalue problems (equation
-\eqref{eq:eigenproblem}). We then characterize an optimal design as
-one that reduces the highest uncertainties, and show how that can give
-rise to repeated measurements (figure \ref{fig:clusterization}).
-
-The solution to the this could be used to seek a design in the real
-problem we are interested in and we outline this in section
-\ref{section:conclusion}. The analysis that culminates in equation
-\eqref{eq:true target} also gives us an upper bound for the design
-criterion (denoted $\tar$) below.
+arises as a solution to a constrained optimization problem, formulated
+using Lagrange multipliers (section \ref{section:D and grad}). In the
+case of no model error we show this problem is in fact a set of
+eigenvector-eigenvalue problems (equation \eqref{eq:eigenproblem}). We
+then characterize an optimal design as one that reduces the highest
+uncertainties, and show how that can give rise to repeated
+measurements (section \ref{subsec:clusterization} and figure
+\ref{fig:clusterization}).
 
@@
 Lemmas \ref{lemma:design increase} and Corollary \ref{cor:zero mod
-  err}. We also show that in our model with independent measurement
-noise, an interesting phenomenon arises, in which the eigenvectors of
-the measurement operator (denoted below $\obs^*\obs$) are the same as
-the eigenvectors of the prior covariance in observation space, which
-helps arrive at the results outlined in the previous paragraph. The
-arguments presented are completely generic and apply equally to any of
-the scenarios mentioned above.
+  err}. %%  We also show that in our model with independent measurement
+%% noise, an interesting phenomenon arises, in which the eigenvectors
+%% of the measurement operator (denoted below $\obs^*\obs$) are the
+%% same as the eigenvectors of the prior covariance in observation
+%% space, which helps arrive at the results outlined in the previous
+%% paragraph.
+The arguments presented are completely generic and apply equally to
+any of the scenarios mentioned in the introduction.
 
@@
       % 
-      \addplot [thin, black, mark=none] table{data/stdv-heat-sens1-var1.txt};
-      \addlegendentry{1 sensors};
-
-      \addplot [thin, blue, mark=none] table{data/stdv-heat-sens2-var1.txt};
-      \addlegendentry{2 sensors};
-
-      \addplot [thin, red, mark=none] table{data/stdv-heat-sens3-var1.txt};
-      \addlegendentry{3 sensors};
-
-      \addplot [thin, green, mark=none] table{data/stdv-heat-sens4-var1.txt};
+      %% \addplot [thin, black, mark=none] table{stdv-heat-sens1-var1.txt};
+      %% \addlegendentry{1 sensors};
+
+      %% \addplot [thin, blue, mark=none] table{stdv-heat-sens2-var1.txt};
+      %% \addlegendentry{2 sensors};
+
+      %% \addplot [thin, red, mark=none] table{stdv-heat-sens3-var1.txt};
+      %% \addlegendentry{3 sensors};
+
+      \addplot [thin, green, mark=none] table{stdv-heat-sens4-var1.txt};
       \addlegendentry{4 sensors};
 
-      \addplot [thin, purple, mark=none] table{data/stdv-heat-sens5-var1.txt};
+      \addplot [thin, purple, mark=none] table{stdv-heat-sens5-var1.txt};
       \addlegendentry{5 sensors};
 
-      \addplot [thin, cyan, mark=none] table{data/stdv-heat-sens6-var1.txt};
+      \addplot [thin, cyan, mark=none] table{stdv-heat-sens6-var1.txt};
       \addlegendentry{6 sensors};
 
-      \addplot [black,  only marks, mark=x, mark size=1.5] 
-      table{data/locs-heat-sens1-var1.txt}; 
-      \addplot [blue,   only marks, mark=x, mark size=1.5]
-      table{data/locs-heat-sens2-var1.txt}; 
-      \addplot [red,    only marks, mark=x, mark size=1.5]
-      table{data/locs-heat-sens3-var1.txt};
+      %% \addplot [black,  only marks, mark=x, mark size=1.5] 
+      %% table{locs-heat-sens1-var1.txt}; 
+      %% \addplot [blue,   only marks, mark=x, mark size=1.5]
+      %% table{locs-heat-sens2-var1.txt}; 
+      %% \addplot [red,    only marks, mark=x, mark size=1.5]
+      %% table{locs-heat-sens3-var1.txt};
       \addplot [green,  only marks, mark=x, mark size=1.5] 
-      table{data/locs-heat-sens4-var1.txt}; 
+      table{locs-heat-sens4-var1.txt}; 
       \addplot [purple, only marks, mark=x, mark size=1.5] 
-      table{data/locs-heat-sens5-var1.txt}; 
+      table{locs-heat-sens5-var1.txt}; 
       \addplot [cyan,   only marks, mark=x, mark size=1.5] 
-      table{data/locs-heat-sens6-var1.txt};
+      table{locs-heat-sens6-var1.txt};
 
@@
     1D heat equation with homogeneous Dirichlet boundary described in
-    section \ref{subsec:example}. Note that the observation data does
-    not influence the pointwise variances, so there is no point in
-    showing the posterior mean.}
-  \label{fig:clusterization illustration}
-\end{figure}
+    section \ref{subsec:example}. Note the clusterization that occurs
+    when 6 sensors are used --- only 4 measurement locations are
+    included. Observed data does not influence the pointwise
+    variances, hence it is omitted.}
+\end{figure}\label{fig:clusterization illustration}
 In section \ref{section:abstract OED} we present a more abstract and
@@
 We would like to infer the initial condition $u_0$. For that purpose,
-we measure $u$ at some set of locations $\x_j \in [0,\pi], j=1,...,m$
+we measure $u$ at some set of locations $\x_j \in [0,\pi], j=1, \dots,m$
 and a final time $T > 0$. We assume centered Gaussian measurement
@@
 operator, so that $u( \cdot,T) = \fwd u_0$, and the observation
-operator $\obs$ so that $u(\x_j,T) = (\obs u)_j, j=1,...,m$, then the
+operator $\obs$ so that $u(\x_j,T) = (\obs u)_j, j=1,\dots,m$, then the
 posterior covariance is known and depends on $\prcov, \fwd, \obs$ and
@@
 
-We will consider the information-theoretic design criterion presented
-in section \ref{section:OED intro}. Thus, we choose $\x_j,j=1,...,m$
-so as to minimize an expression analogous to the log-determinant of
-the posterior covariance operator (corresponding to a D-optimal
-design). We can see the clusterization effect in figure
+We will consider generalization of the information-theoretic design
+criterion presented in section \ref{section:OED intro} to infinite
+dimensions (section \ref{subsec:D optimal design}). We choose
+$\x_j,j=1,\dots,m$ so as to minimize an expression analogous to the
+log-determinant of the posterior covariance operator (corresponding to
+a D-optimal design). We can see the clusterization effect in figure
 \ref{fig:clusterization illustration} --- the design criterion
@@
 discussed in section \ref{section:vanishing}. In that section, it is
-shown how sensor clusterization occurs. It is advised that a reader
-less comfortable with Hilbert and Banach spaces mostly ignore these
-terms and consider $\hil$ and $\ban$ finite dimensional vector spaces.
+shown how sensor clusterization can occur. The reader may skip if they
+are constrained by time. It is advised that a reader less comfortable
+with Hilbert and Banach spaces mostly ignore these terms and consider
+$\hil$ and $\ban$ finite dimensional vector spaces.
 
@@
 %% considered in section \ref{subsec:example}, we were constrained to
-%% choose $\obs = (\meas_1,...,\meas_m)^t$ with $\meas_j = \delta_{\x_j}$
+%% choose $\obs = (\meas_1,\dots,\meas_m)^t$ with $\meas_j = \delta_{\x_j}$
 %% for some $\x_j \in [0,\pi]$. We present an approach to understanding
@@
 The theoretical foundations for inverse problems over function spaces
-can be found in \cite{Stuart10}. At this point, it is useful to
-record a known result regarding the posterior covariance operator of
-our inverse problem, namely
+can be found in \cite{Stuart10}. This section is mostly devoted to
+setting notation. At this point, it is useful to record a known result
+regarding the posterior covariance operator of our inverse problem,
+namely
 \begin{align}\label{eq:postcov precursor}
@@
 functionals), we think of $\obs$ as having ``rows'' $\meas_j,
-j=1,...,m$:
+j=1,\dots,m$:
 \begin{equation}\label{eq:O}
-  \obs = (\meas_1,..., \meas_m)^t, \meas_j \in \ban^*, j = 1,...,m.
+  \obs = (\meas_1,\dots, \meas_m)^t, \meas_j \in \ban^*, j = 1,\dots,m.
 \end{equation}
-This way, for $u \in \ban$ we have $\obs u = (\meas_1(u), ...,
+This way, for $u \in \ban$ we have $\obs u = (\meas_1(u), \dots,
 \meas_m(u) )^t \in \R^m$.
@@
 \begin{observation}
-  It is best to think of $\meas_j,j=1,...,m$ as row vectors and of
+  It is best to think of $\meas_j,j=1,\dots,m$ as row vectors and of
   $\obs$ as a matrix with rows $\meas_j$. 
@@
 
-Usually, this is written using $\tmp := \obs \fwd$ and $\tmp^*$ but
-since we need to separate the observation operator $\obs$ from the
-forward operator $\fwd$ we write the explicit expression
-\eqref{eq:postcov}. Taking $\modcov = 0$ is common practice
+%% Usually, this is written using $\tmp := \obs \fwd$ and $\tmp^*$ but
+%% since we need to separate the observation operator $\obs$ from the
+%% forward operator $\fwd$ we write the explicit expression
+%% \eqref{eq:postcov}.
+Taking $\modcov = 0$ is common practice
 \cite{Tarantola05,KaipioSomersalo05,Vogel02} and then we are reduced
@@
 
-\subsection{D-Optimal Design in Infinite Dimensions} 
-As pointed out before, we consider D-optimal experimental design. The
-meaning of this notion in infinite-dimensional Hilbert spaces was
-investigated in \cite{AlexanderianGloorGhattas14}. Note that the
+\subsection{D-Optimal Design in Infinite Dimensions}\label{subsec:D optimal design} 
+The meaning of D-optimal design in infinite-dimensional Hilbert spaces
+was investigated in \cite{AlexanderianGloorGhattas14}. Note that the
 authors make assumptions that amount to $\Sigma=I$ (implied by
@@
 \cite[pp. 681]{AlexanderianGloorGhattas14}, the results hold for more
-general covariance matrices. The derivation of the design criterion
-from information theoretic considerations can be verified similarly.
-The conclusion is that in infinite-dimensions, a D-optimal design is
-well-defined as maximizing the expected KL divergence between
-posterior and prior and is summarized, using our notation, in the
-following theorem:
+general covariance matrices. The conclusion is that in
+infinite-dimensions, a D-optimal design is well-defined as maximizing
+the expected KL divergence between posterior and prior and is
+summarized, using our notation, in the following theorem:
 \begin{theorem}[Slightly modified Theorem 1 from \cite{AlexanderianGloorGhattas14}]
@@
 following optimization problem for $\tar$. %%: (\ban^*)^m \to \R$:
-\begin{align*}
+\begin{align}\label{eq:optimization}
   \obs^{\star} := \argmax_{\obs} \tar( \obs ) 
@@
   (I + \prcov^{1/2} \fwd^*\obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}),
-\end{align*}
+\end{align}
 where $\obs$ is constrained to some allowed set of measurements
@@
 We seek a formulation of the optimal design using Lagrange
-multipliers. We first find gradients in function spaces.
+multipliers. In section \ref{section:objective} we find the gradient
+of $\tar$.  In section \ref{subsec:unit length} we relax the
+constraints on $\obs$ so that the optimization problem in
+\eqref{eq:optimization} and find gradients for the new constraints.
 
@@
 with $V_j \in \ban^*$ and $\nabla \tar(\obs)_j \in \ban^{**}\subseteq
-\ban, j=1,...,m$. Thus, $\nabla \tar( \obs ) \in (\ban^{**})^m
+\ban, j=1,\dots,m$. Thus, $\nabla \tar( \obs ) \in (\ban^{**})^m
 \subseteq \ban^m$ is indeed the correct gradient and the notation
@@
 As mentioned before, we cannot choose any $\obs$ when maximizing
-$\tar(\obs)$. Recall $\obs = (\meas_1,...,\meas_m)^t \in (\ban^*)^m$
-and each $\meas_j,j=1,...,m$ must be chosen from some allowed set of
+$\tar(\obs)$. Recall $\obs = (\meas_1,\dots,\meas_m)^t \in (\ban^*)^m$
+and each $\meas_j,j=1,\dots,m$ must be chosen from some allowed set of
 functionals in $\ban^*$. This set differs based on the kind of sensors
@@
 \begin{proposition}\label{prop:bigger better}
-  Let $\obs = (\meas_1,...,\meas_m)^t$, $j \in \{1,...,m\}$, $\sigma^2
+  Let $\obs = (\meas_1,\dots,\meas_m)^t$, $j \in \{1,\dots,m\}$, $\sigma^2
   > 0$ and $\lambda > 1$. Then $\tar(\obs)$ increases if we use
@@
 and take measurements $\meas_j \in \bar{\hil}^* = \bar{\hil}$ such
-that $\| \meas_j\| = 1, j=1,...,m$. Recall that we think of $\ban$ as
+that $\| \meas_j\| = 1, j=1,\dots,m$. Recall that we think of $\ban$ as
 a somewhat restricted space ($C(\bar{\Omega})$ with the sup-norm over
@@
 \begin{align}\label{eq:constraints}
-  \phi_j(\obs) :=\frac12 \| \obs^* e_j\|_{\bar{\hil}}^2 - \frac12 = 0,j=1,...,m
+  \phi_j(\obs) :=\frac12 \| \obs^* e_j\|_{\bar{\hil}}^2 - \frac12 = 0,j=1,\dots,m
 \end{align}
@@
 \begin{align}\label{eq:grad constraints}
-\nabla \phi_j(\obs) = \obs^* e_j e_j^t = \meas_j e_j^t , j=1,...,m,
+\nabla \phi_j(\obs) = \obs^* e_j e_j^t = \meas_j e_j^t , j=1,\dots,m,
 \end{align}
@@
   \label{eq:Lagrange mult1} \\
-    &\phi_j(\obs) = 0, j = 1,...,m. \label{eq:Lagrange mult2}
+    &\phi_j(\obs) = 0, j = 1,\dots,m. \label{eq:Lagrange mult2}
   %\end{split}
@@
   (I - \modcov \obs^* \Sigma^{-1} \obs) \fwd \postcov \fwd^* \obs^*\Sigma^{-1}
-  = \sum_{j=1}^m \xi_j \obs^* e_je_j^t = (\xi_1 \meas_1,...,\xi_m \meas_m).
+  = \sum_{j=1}^m \xi_j \obs^* e_je_j^t = (\xi_1 \meas_1,\dots,\xi_m \meas_m).
 \end{equation*}
@@
 self-adjoint operator $\sigma^{-2}\fwd \postcov \fwd^*$ where the rows
-of $\obs$, namely $\meas_j,j=1,...m$, are the eigenvectors. If $\fwd$
+of $\obs$, namely $\meas_j,j=1,\dots, m$, are the eigenvectors. If $\fwd$
 is invertible, then we may write:
@@
 invertible --- consider instead $\fwd$ restricted to $\ker
-\fwd^{\perp}$. This does not change the optimal design problem at
-all, since we will never take measurements that have a component in
-$\ker \fwd$ (this is intuitive by Occam's razor but can also be
-understood from Corollary \ref{cor:zero mod err}).
+\fwd^{\perp}$. This does not change the optimal design problem at all,
+since we will never take measurements that have a component in $\ker
+\fwd$ (this can be understood from Corollary \ref{cor:zero mod err}).
 
@@
 In light of \eqref{eq:mod conditions}, let us denote eigenvalues of
-$\fwd \prcov \fwd^*$ by $\lambda_1 \geq \lambda_2 \geq ...$ and its
+$\fwd \prcov \fwd^*$ by $\lambda_1 \geq \lambda_2 \geq \dots$ and its
 eigenvectors $\{\ev_i\}_{i=1}^{\infty}$. The following lemma will be useful:
 
-\begin{lemma}\label{lemma:sim diag}
-  Let $C:\hil \to \hil$ self-adjoint and let $\func_1,...,\func_m \in
+\begin{lemma}[Proved as lemma \ref{lemma:sim diag} in the appendix]
+  Let $C:\hil \to \hil$ self-adjoint and let $\func_1,\dots,\func_m \in
   \hil$. Denote $\func^*$ the element $\func$ acting as a linear
@@
   \begin{equation*}
-   (C + \sum_{j=1}^m \func_j\func_j^*) \func_k = \xi_k \func_k, k = 1,...,m
+   (C + \sum_{j=1}^m \func_j\func_j^*) \func_k = \xi_k \func_k, k = 1,\dots,m
   \end{equation*}
@@
 Thus, if we take $\func_j^{*} = \meas_j$ and $C := (\fwd \prcov
-\fwd^*)^{-1}$ in Lemma \ref{lemma:sim diag}, we conclude from the
-eigenproblem \eqref{eq:mod conditions} that $\obs^*\obs$ has the same
-eigenvectors as $\fwd \prcov \fwd^*$, namely
-$\{\ev_i\}_{i=1}^{\infty}$ (infinitely many of them with zero
-eigenvalue, since $\obs^*\obs$ is finite-rank). Denote the
-corresponding non-zero eigenvalues of $\obs^*\obs$ by
-$\{\eta_i\}_{i=1}^{k}$ and let $\eta_i = 0$ for $i \geq k+1$. By the argument
-in the paragraph following \eqref{eq:mod conditions}:
+\fwd^*)^{-1}$ in the above lemma, we conclude from the eigenproblem
+\eqref{eq:mod conditions} that $\obs^*\obs$ has the same eigenvectors
+as $\fwd \prcov \fwd^*$, namely $\{\ev_i\}_{i=1}^{\infty}$ (infinitely
+many of them with zero eigenvalue, since $\obs^*\obs$ is
+finite-rank). Denote the corresponding non-zero eigenvalues of
+$\obs^*\obs$ by $\{\eta_i\}_{i=1}^{k}$ and let $\eta_i = 0$ for $i
+\geq k+1$. By the argument in the paragraph following \eqref{eq:mod
+  conditions}:
 \begin{align}\label{eq:pre manipulations}
@@
 \pgfplotstableread{
-  Label    prior  $a_1$  $a_2$   $a_3$   $a_4$   $a_5$         $a_6$    topper
+  Label    prior  $o_1$  $o_2$   $o_3$   $o_4$   $o_5$         $o_6$    topper
   1         0.2    0.8    0.5     0.5    0.4    0.33333333  0.3333333  0.001
@@
       \addplot [fill=green!80]  table [y=prior, meta=Label, x expr=\coordindex] {\testdata};
-      \addplot [fill=blue!60]   table [y=$a_1$, meta=Label, x expr=\coordindex] {\testdata};
-      \addplot [fill=red!60]    table [y=$a_2$, meta=Label, x expr=\coordindex] {\testdata};
-      \addplot [fill=black!60]  table [y=$a_3$, meta=Label, x expr=\coordindex] {\testdata};
-      \addplot [fill=orange!60] table [y=$a_4$, meta=Label, x expr=\coordindex] {\testdata};
-      \addplot [fill=cyan!60]   table [y=$a_5$, meta=Label, x expr=\coordindex] {\testdata};
-      \addplot [fill=purple!60] table [y=$a_6$, meta=Label, x expr=\coordindex] {\testdata};
+      \addplot [fill=blue!60]   table [y=$o_1$, meta=Label, x expr=\coordindex] {\testdata};
+      \addplot [fill=red!60]    table [y=$o_2$, meta=Label, x expr=\coordindex] {\testdata};
+      \addplot [fill=black!60]  table [y=$o_3$, meta=Label, x expr=\coordindex] {\testdata};
+      \addplot [fill=orange!60] table [y=$o_4$, meta=Label, x expr=\coordindex] {\testdata};
+      \addplot [fill=cyan!60]   table [y=$o_5$, meta=Label, x expr=\coordindex] {\testdata};
+      \addplot [fill=purple!60] table [y=$o_6$, meta=Label, x expr=\coordindex] {\testdata};
 
       \addlegendentry{prior}
-      \addlegendentry{$a_1$}
-      \addlegendentry{$a_2$}
-      \addlegendentry{$a_3$}
-      \addlegendentry{$a_4$}
-      \addlegendentry{$a_5$}
-      \addlegendentry{$a_6$}   
+      \addlegendentry{$o_1$}
+      \addlegendentry{$o_2$}
+      \addlegendentry{$o_3$}
+      \addlegendentry{$o_4$}
+      \addlegendentry{$o_5$}
+      \addlegendentry{$o_6$}   
     \end{axis}
   \end{tikzpicture}
-  \caption{Posterior precision per eigenvector. Each measurement $\func_i$
-    increases precision uniformly across the lowest precision
-    eigenvectors. We see that $\func_2$ and $\func_3$ are repeated
-    measurements, as well as $\func_5$ and $\func_6$.}
+  \caption{Posterior precision per eigenvector, after taking a
+    measurement $\meas_i$. Each $\meas_i$ increases precision
+    uniformly across the lowest precision eigenvectors. We see that
+    $\meas_2$ and $\meas_3$ are repeated measurements, as well as
+    $\meas_5$ and $\meas_6$.}
   \label{fig:clusterization}
 \end{figure}
-with the norm constraints on $\meas_j,j=1,...,m$. We show in the
-appendix (Lemma \ref{lemma:free}) we can find $\meas_1,...,\meas_m$
+with the norm constraints on $\meas_j,j=1,\dots,m$. We show in the
+appendix (Lemma \ref{lemma:free}) we can find $\meas_1,\dots,\meas_m$
 such that $\{\eta_i\}_{i=1}^k$ have any value we desire, subject only
 to the restriction that we do not change the trace of $\obs^*\obs$, so
-that $\sum_{i=1}^k \eta_i = m$. By concavity of $\log$, the highest
-increase in the design criterion is gained by increasing
-$\lambda_i^{-1} + \sigma^{-2}\eta_i$ where it is small. So, no weight
-should be given to (i.e. measurement taken of) the $k$th eigenvector
-before $\lambda_i^{-1} + \sigma^{-2}\eta_i \geq \lambda_k^{-1} \forall
-i \neq k$. Thus, we should choose $\{\eta_i\}_{i=1}^k$ as follows. We
-would increase $\lambda_1^{-1} + \sigma^{-2} \eta_1$ until it equals
-$\lambda_2^{-1}$. Then, $\lambda_i^{-1} + \sigma^{-2} \eta_i,i=1,2$ is
-increased until $\lambda_i^{-1} + \sigma^{-2} \eta_i =
-\lambda_3^{-1},i=1,2$ and so forth.
+that $\sum_{i=1}^k \eta_i = m$. By concavity of $\log$, the fastest
+increase in the design criterion (equation \eqref{eq:true target}) is
+gained by increasing $\lambda_i^{-1} + \sigma^{-2}\eta_i$ where it is
+smallest. So, no weight should be given to (i.e. measurement taken of)
+the $k$th eigenvector before $\lambda_i^{-1} + \sigma^{-2}\eta_i \geq
+\lambda_k^{-1},\ \forall i \neq k$. Thus, we should choose
+$\{\eta_i\}_{i=1}^k$ as follows. We would increase $\lambda_1^{-1} +
+\sigma^{-2} \eta_1$ until it equals $\lambda_2^{-1}$. Then,
+$\lambda_i^{-1} + \sigma^{-2} \eta_i,i=1,2$ are increased until
+$\lambda_i^{-1} + \sigma^{-2} \eta_i = \lambda_3^{-1},i=1,2$ and so
+forth.
 
@@
 precision. If the precision in these is lower than the precision for
-the next eigenvector, a repeated measurement is taken.
-
-%% The symmetry mentioned in last paragraph of the previous section may
-%% very well break in real life problems. If this happens and eigenvalues
-%% of \eqref{eq:eigenproblem} are all different, then showing
-%% sensors do cluster proceeds as follows. First, we show there
-%% are only so many different equal, by the pigeonhole principle. In
-%% order to make this argument work, one needs to show this eigenproblem
-%% does, indeed, have only finitely many eigenvalues. This is done by
-%% first recalling that $\fwd$ is strongly smoothing (so $\fwd \postcov
-%% \fwd^*$ has fast decaying eigenvalues) and then showing that small
-%% eigenvalues arise from un-informative measurements. We use the
-%% following corollary, whose proof may be found in the appendix.
-%% %
-%% \begin{corollary}[Gain for No Model Error]\label{cor:zero mod err}
-%%   If $\modcov = 0$, then
-%%   \begin{equation*}
-%%     \tar( \obs ) - \tar (\obsm )
-%%     = -\frac12 \log (1 - \sigma^{-2} \langle \fwd \postcov \fwd^* \meas_m, \meas_m \rangle ).
-%%   \end{equation*}
-%% \end{corollary}
-%% %
-%% Suppose $\meas_j$ is a measurement with a small eigenvalue, i.e.\
-%% \begin{equation}\label{eq:small eigenvalue}
-%%   \sigma^{-2} \fwd \postcov \fwd^* \meas_j = \xi_j \meas_j,\ \xi_j \ll 1.
-%% \end{equation}
-%% By Corollary \ref{cor:zero mod err}, the increase in the design
-%% criterion is
-%% \begin{align*}
-%%   \tar(\obs) - \tar(\obsm) &=
-%%   -\frac12 \log (1- \sigma^{-2} \langle \fwd
-%%   \postcovm \fwd^* \meas_j, \meas_j \rangle) \\
-%%   %
-%%   %
-%%   %
-%%   &= - \frac12 \log (1 - \xi_j) \text{ since $\|\meas_j\|=1$ }\\
-%%   %
-%%   %
-%%   %
-%%   &\approx \frac{\xi_j}{2}.
-%% \end{align*}
-%% Thus, negligibly small eigenvalues give negligible increase in the
-%% design criterion and will not be a part of a truly optimal
-%% design. Since the eigenvalues of $\fwd \postcov \fwd^*$ decay faster
-%% than those of $\fwd\prcov\fwd^*$, the latter will determine how many
-%% non-negligible eigenvalues $\fwd \postcov \fwd^*$ has. Thus, in the
-%% eigenproblem \eqref{eq:eigenproblem}, we need only consider a finite
-%% number of eigenvectors.
-
-%% At this point, one would like to invoke the pigeonhole principle and
-%% claim that eigenvectors cluster. This, however, is not as
-%% straightforward as one may hope. If $\fwd \postcov \fwd^*$ only has
-%% simple eigenvalues, then we may use the pigeonhole principle and
-%% conclude that some eigenvectors are clustered (up to a sign). However,
-%% if $\fwd \postcov \fwd^*$ has non-simple eigenvalues, one may hope to
-%% somehow prove measurement vectors are either orthogonal or parallel,
-%% which suffices for proving clusterization but that is not true.
+the next eigenvector, a repeated measurement is optimal.
 
 \section{Analysis of Optimal Designs --- Non-Vanishing Model Error}
-We can build on Lemma \ref{lemma:design increase} to find the gain in design
-for identical measurements.
-\begin{corollary}[Gain for Identical Measurement]\label{cor:same meas}
+Denote $\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm :=
+(\meas_1,\dots,\meas_{m-1})^t$. The following corollary is proved in
+the appendix.
+\begin{corollary}[Proved as corollary \ref{cor:same meas} in the appendix]
   If $\meas_m = \meas_j$ for some $1 \leq j \leq m-1$, then
@@
 Our relaxed model gives us insight to D-optimal designs. We see that
-in an ideal settings, we should not reduce uncertainty in any
-direction (eigenvector) before it is the largest uncertainty present.
-We show how and why repeated measurements can give rise to a D-optimal
+in our setting, uncertainty is not reduced in any direction
+(eigenvector) before it is the largest uncertainty present. We show
+how and why repeated measurements can give rise to a D-optimal
 design. There is more work to be done in understanding exactly what
 causes such designs to be strictly better than others.
-%% We believe the model presented is not too far from reality. After all,
-%% taking any measurement (point evaluation or other) is just recording
-%% the value of a linear combination of individual eigenvectors of the
-%% prior. In light of this, we believe the reason for clusterization is close
-%% to the argument made in section \ref{subsub:first attempt}. The reason
-%% we observe clusterization in real life problems is probably due to
-%% symmetry breaking, although we cannot show that rigorously.
-
-A consequence of the argument in section \ref{section:vanishing} is
-that we can find an approximately optimal design by means of linear
-algebra. Here is how this should be done.
-\begin{enumerate}
-
-   \item Find eigenvalues $\lambda_1 \geq \lambda_2 \geq ...$ and
-     corresponding eigenvectors $\ev_i$ of $\fwd \prcov \fwd^*$.
-   \item Denote the (assumed known) variance of the measurement noise
-     $\sigma^2$.
-   \item Decide how many measurements can be taken and denote that
-     number $m$. This is the total precision that can be allocated to
-     the different eigenvectors.
-    \item Find $k$ and $\{\eta_i\}_{i=1}^k$, $\eta_i > 0, i=1,...,k$
-      such that:
-      \begin{itemize}
-        \item $\sum_{i=1}^k \eta_i = m$ --- the total measurement precision available.
-      \item $\lambda_i^{-1} + \sigma^{-2}\eta_i = \lambda_k^{-1} +
-        \sigma^{-2}\eta_k$ for $i=1,...,k-1$ --- the posterior
-        precision is equal for all $k$ eigenvectors.
-      \item $\lambda_{k+1}^{-1} \geq \lambda_k^{-1} + \eta_k$ --- no
-        information is gained for the rest of the eigenvectors.
-      \end{itemize}
-      \item Use the construction in Lemma \ref{lemma:free} to find the
-        optimal measurements in terms of the eigenvectors $\ev_i$ ---
-        without \emph{any} optimization.
-      \item Find point measurements that approximate those in the domain. 
-\end{enumerate}
 
@@
 \begin{lemma}[Increase due to a Measurement]\label{lemma:design increase}
-  Let $\obs = (\meas_1,...,\meas_m)^t$ and $\obsm := (\meas_1,...,\meas_{m-1})^t$. Then
+  Let $\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm := (\meas_1,\dots,\meas_{m-1})^t$. Then
   \begin{align*}
@@
   Let $M \in \R^{k \times k}$ symmetric positive definite with $\tr M =
-  m$, $m > k$. Then we can find $\func_j \in \R^k,j=1,...,m$ with
-  $\|\func_j\|=1$ and $A = (\func_1,...,\func_m)$ such that $AA^t =
+  m$, $m > k$. Then we can find $\func_j \in \R^k,j=1,\dots,m$ with
+  $\|\func_j\|=1$ and $A = (\func_1,\dots,\func_m)$ such that $AA^t =
   M$.
@@
   Let us diagonalize $M$, so that $M = U D U^t$ with $D =
-  \diag(d_1,...,d_k)$ and $U \in \R^{k \times k }$ orthogonal. Let $S
+  \diag(d_1,\dots,d_k)$ and $U \in \R^{k \times k }$ orthogonal. Let $S
   \in \R^{k \times m}$ with $S_{ii} = \sqrt{d_{i}}$ and zeros
@@
   also satisfies the unit norm constraints we are done. These
-  constraints are, for $j=1,...,m$:
+  constraints are, for $j=1,\dots,m$:
   \begin{equation}\label{eq:V constraints}
@@
   Define $C = S^tS - I \in \R^{m \times m}$. Note that $\tr C = 0$ and
-  $C$ is diagonal with non-zero entries $d_i-1,i=1,...,k$. It suffices
+  $C$ is diagonal with non-zero entries $d_i-1,i=1,\dots,k$. It suffices
   to find $V$ orthogonal such that $V C V^t$ has zero diagonal. We
@@
   $p$ rows and columns. Specifically, it leaves rows and columns
-  $k+1,...,m$ unchanged. We want to choose $\theta$ such that
+  $k+1,\dots,m$ unchanged. We want to choose $\theta$ such that
   \begin{equation}\label{eq:mm}
@@
   (0,\pi)$ such that \eqref{eq:mm} is satisfied. We continue to find
-  $R^{(k-1)}$ that leaves rows and columns $k,...,m$ unchanged and
+  $R^{(k-1)}$ that leaves rows and columns $k,\dots,m$ unchanged and
   continue introducing zeros to the diagonal. The assumption $\tr D =
@@
 \begin{lemma}\label{lemma:sim diag}
-  Let $C:\hil \to \hil$ self-adjoint and let $\func_1,...,\func_m \in
+  Let $C:\hil \to \hil$ self-adjoint and let $\func_1,\dots,\func_m \in
   \hil$. Denote $\func^*$ the element $\func$ acting as a
@@
   First, enumerate the eigenvalues of $C + \sum_{j=1}^m
-  \func_j\func_j^*$ as $\xi_1,...,\xi_\ell$. Denote the
+  \func_j\func_j^*$ as $\xi_1,\dots,\xi_\ell$. Denote the
   indices of the eigenvectors corresponding to $\xi_i$
@@
 \begin{proof} 
-  Fix an arbitrary $j=1,...,m$ and take $V:= e_j e_j^t \obs$. We see
+  Fix an arbitrary $j=1,\dots,m$ and take $V:= e_j e_j^t \obs$. We see
   that for $u \in \ban$
   \begin{equation*}
-    Vu = e_je_j^t (\meas_1(u),...,\meas_m(u) )^t = e_j \meas_j(u)
-    = (0,...,0,\meas_j(u),0,...,0)^t.
+    Vu = e_je_j^t (\meas_1(u),\dots,\meas_m(u) )^t = e_j \meas_j(u)
+    = (0,\dots,0,\meas_j(u),0,\dots,0)^t.
   \end{equation*}
@@
 %% taking $C = (\fwd \prcov \fwd^*)^{-1}$ we observe that $\meas_j,
-%% j=1,...,m$ are eigenvectors of $\fwd \prcov \fwd^*$. If they have
+%% j=1,\dots,m$ are eigenvectors of $\fwd \prcov \fwd^*$. If they have
 %% different eigenvalue, they are orthogonal. If not, this means, by

### 2020-09-22 changes_highlighted.tex
some of the reply is finished

--- 2020-07-23 paper.tex
@@
   dimensions]{Sensor Clusterization in D-optimal Design in Infinite
-  Dimensional Bayesian Inverse Problems}
+  Dimensional \reftwo{Linear} Bayesian Inverse Problems}
 
@@
   We investigate the problem of sensor clusterization in optimal
-  experimental design for infinite-dimensional Bayesian inverse
-  problems. We suggest an analytically tractable model for such
-  designs and reason how it may lead to sensor clusterization in the
-  case of iid measurement noise. We also show that in the case of
-  spatially correlated measurement error clusterization does not
-  occur. As a part of the analysis we prove a matrix determinant
-  lemma analog in infinite dimensions, as well as a lemma for
-  calculating derivatives of $\log \det$ of operators.
+  experimental design for infinite-dimensional Bayesian
+  \reftwo{linear} inverse problems. We suggest an analytically
+  tractable model for such designs and reason how it may lead to
+  sensor clusterization in the case of iid measurement noise. We also
+  show that in the case of spatially correlated measurement error
+  clusterization does not occur. As a part of the analysis we prove a
+  matrix determinant lemma analog in infinite dimensions, as well as a
+  lemma for calculating derivatives of $\log \det$ of operators.
 \end{abstract}
@@
 information gain amounts to minimizing the determinant of the
-posterior covariance matrix. In a frequentist settings, an optimal
+posterior covariance matrix. In a frequentist setting\reftwo{\sout{s}}, an optimal
 design minimizes the volume of the uncertainty ellipsoid \cite[page
@@
 
-In the process we generalize several lemmas from linear algebra to
+In the process\reftwo{,} we generalize several lemmas from linear algebra to
 infinite-dimensional settings --- the Matrix Determinant Lemma and a
@@
     various numbers of sensors, and location of these sensors for the
-    1D heat equation with homogeneous Dirichlet boundary described in
+    1D heat equation with \reftwo{a} homogeneous Dirichlet boundary described in
     section \ref{subsec:example}. Note the clusterization that occurs
@@
 illustration, we present the problem via a toy model --- the 1D heat
-equation in $[0,\pi]$ with homogeneous Dirichlet boundary
+equation in $[0,\pi]$ with \reftwo{a} homogeneous Dirichlet boundary
 conditions.
@@
 initial condition as $u_0 \sim \normal(0,\prcov)$, for $\prcov =
-(-\Delta)^{-1}$ with homogeneous Dirichlet boundary condition. It is
+(-\Delta)^{-1}$ with \reftwo{a} homogeneous Dirichlet boundary condition. It is
 well known \cite{Tarantola05} that for linear problems, with Gaussian
@@
 operator $\obs$ so that $u(\x_j,T) = (\obs u)_j, j=1,\dots,m$, then the
-posterior covariance is known and depends on $\prcov, \fwd, \obs$ and
+posterior covariance is known and depends only on $\prcov, \fwd, \obs$ and
 $\sigma^2$ (see section \ref{section:abstract OED} and
@@
 can be found in \cite{Stuart10}. This section is mostly devoted to
-setting notation. At this point, it is useful to record a known result
+setting notation. \reftwo{\sout{At this point, it is useful to record a known result
 regarding the posterior covariance operator of our inverse problem,
-namely
+namely}
 \begin{align}\label{eq:postcov precursor}
-  \postcov = (\prcov^{-1} + \fwd^* \obs^* \Sigma^{-1} \obs \fwd )^{-1}.
+  \stkout{\postcov = (\prcov^{-1} + \fwd^* \obs^* \Sigma^{-1} \obs \fwd )^{-1}.}
 \end{align}
-
-\subsection{Bayesian Inverse Problems}\label{section:abstract OED}
+}
+
+\subsection{Bayesian \reftwo{Linear} Inverse Problems}\label{section:abstract OED}
 Let $\hil$ be a Hilbert space and $\ban$ a Banach space, both
@@
 values --- the heat operator from section \ref{subsec:example} is a
-prime example).
-
-We want to infer $\param \in \hil$, some parameter of the dynamics
-given noisy observations of $\fwd \param$. We model $\param \sim
-\normal(0,\prcov)$ with some appropriate covariance operator $\prcov$
-\cite{Stuart10}. Observations (measurements) are taken via the linear
-observation operator $\obs \in ( \ban^* )^m$, where $m$ is the number
-of observations allowed (e.g. sensors at our disposal). In an analogy
-with linear algebra (where row vectors are thought of as linear
-functionals), we think of $\obs$ as having ``rows'' $\meas_j,
-j=1,\dots,m$:
+prime example). We want to infer $\param \in \hil$, some parameter of
+the dynamics given noisy observations of $\fwd \param$. \reftwo{We
+  take a zero mean Gaussian prior $\pr$ for $\param$ (otherwise the
+  mean can just be subtracted everywhere). Hence, \sout{We model}}
+$\param \sim \pr = \normal(0,\prcov)$ with some appropriate covariance
+operator $\prcov$ \cite{Stuart10}. Observations (measurements) are
+taken via the linear observation operator $\obs \in ( \ban^* )^m$,
+where $m$ is the number of observations allowed (e.g. sensors at our
+disposal). In an analogy with linear algebra (where row vectors are
+thought of as linear functionals), we think of $\obs$ as having
+``rows'' $\meas_j, j=1,\dots,m$:
 \begin{equation}\label{eq:O}
@@
 \subsubsection{Error Terms}
-Considering the sum of the error terms it is easy to see that $\obs
+Considering the sum of the error terms\reftwo{,} it is easy to see that $\obs
 \eps' + \eps \in \R^m$. It is a centered Gaussian random vector and
@@
 \bar{\eps}$ with $\bar{\eps} \sim \normal(0,\Sigma)$. At this point,
-it is useful to rewrite the posterior covariance operator of our
-inverse problem from \eqref{eq:postcov precursor}
+it is useful to \reftwo{\sout{rewrite the posterior covariance operator of our
+inverse problem from \eqref{eq:postcov precursor}} record a known result
+regarding the posterior covariance operator of our inverse problem,
+namely}
 \begin{align}\label{eq:postcov}
@@
 \end{align}
-
-%% Usually, this is written using $\tmp := \obs \fwd$ and $\tmp^*$ but
-%% since we need to separate the observation operator $\obs$ from the
-%% forward operator $\fwd$ we write the explicit expression
-%% \eqref{eq:postcov}.
+\refthree{
+Usually, this is written using $\tmp := \obs \fwd$ and $\tmp^*$ but
+since we need to separate the observation operator $\obs$ from the
+forward operator $\fwd$ we write the explicit expression
+\eqref{eq:postcov}.}
+
 Taking $\modcov = 0$ is common practice
@@
 infinite-dimensions, a D-optimal design is well-defined as maximizing
-the expected KL divergence between posterior and prior and is
-summarized, using our notation, in the following theorem:
+the expected KL divergence between posterior and prior
+\reftwo{(denoted $\tar$ below and throughout the paper)} and is summarized, using our notation,
+in the following theorem:
 \begin{theorem}[Slightly modified Theorem 1 from \cite{AlexanderianGloorGhattas14}]
@@
 where $\obs$ is constrained to some allowed set of measurements
-(e.g.\ pointwise evaluations on $[0,\pi]$)
+(e.g.\ pointwise evaluations on $[0,\pi]$).
 %%, for the example in section \ref{subsec:example}).
@@
 constraints on $\obs$ so that the optimization problem in
-\eqref{eq:optimization} and find gradients for the new constraints.
+\eqref{eq:optimization} \reftwo{is analytically tractable} and find
+gradients for the new constraints.
 
@@
 We start by calculating the first variation. Some calculations are
-delegated to Lemma \ref{lemma:aux calc} in the appendix. Also, we
-use the following lemma whose proof is also delegated to the appendix:
-\begin{lemma}[Generalized from \cite{Lax97}]\label{lemma:lax}
+delegated to Lemma \ref{lemma:aux calc} in the appendix. Also, we use
+the following lemma whose proof is also delegated to the appendix
+\ref{lemma:lax}:
+\begin{restatable*}[Generalized from \cite{Lax97}]{lemma}{lax}\label{lemma:lax}
   Let $Y(t)$ be a differentiable operator-valued function. Assume 
@@
   \begin{equation*}
-    \frac{\der \log \det (I+Y(t))}{\der t} = \tr (I+Y(t))^{-1} \dot{Y}(t).
+    \frac{\der \log \det (I+Y(t))}{\der t} = \tr{(I+Y(t))^{-1} \dot{Y}(t)}.
   \end{equation*}
-\end{lemma}
+\end{restatable*}
 
@@
     % 
-    &= \frac12 \tr ( I + \prcov^{1/2} \fwd^* \obs^* \Sigma^{-1}
+    &= \frac12 \tr{( I + \prcov^{1/2} \fwd^* \obs^* \Sigma^{-1}
     \obs\fwd \prcov^{1/2} )^{-1}
     \frac{\der}{\der \tau} \Big |_{\tau=0}
-    \prcov^{1/2} \fwd^* T(\obs+\tau V) \fwd \prcov^{1/2}\ \text{ (by \ref{lemma:lax})}\\
-    % 
-    % 
-    % 
-    &= \frac12 \tr \postcov \fwd^* (
-      V^* \Sigma^{-1} \obs 
-      - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \text{ (by \ref{lemma:aux calc})}\\
+    \prcov^{1/2} \fwd^* T(\obs+\tau V) \fwd \prcov^{1/2}}\ \text{ (by \ref{lemma:lax})}\\
+    % 
+    % 
+    % 
+    &= \frac12 \ttr{\postcov \fwd^* (V^* \Sigma^{-1} \obs 
+      - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
       &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs 
-      + \obs^* \Sigma^{-1} V ) \fwd \\
+      + \obs^* \Sigma^{-1} V ) \fwd}  \text{ (by \ref{lemma:aux calc})}\\
 %%   \end{split}
@@
       %% \delta \tar(\obs) V 
-    &= \tr \postcov \fwd^* ( \obs^* \Sigma^{-1} V 
-    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs ) \fwd \\
-    %
-    %
-    % 
-    &= \tr \postcov \fwd^* \obs^* \Sigma^{-1} V 
-    ( I - \modcov \obs^* \Sigma^{-1}\obs ) \fwd \\
-    % 
-    %
-    %
-    &= \tr V ( I - \modcov \obs^* \Sigma^{-1}\obs )
-    \fwd \postcov \fwd^* \obs^* \Sigma^{-1}.
+    &= \tr{\postcov \fwd^* ( \obs^* \Sigma^{-1} V 
+    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs ) \fwd} \\
+    %
+    %
+    % 
+    &= \tr{\postcov \fwd^* \obs^* \Sigma^{-1} V 
+    ( I - \modcov \obs^* \Sigma^{-1}\obs ) \fwd} \\
+    % 
+    %
+    %
+    &= \tr{V ( I - \modcov \obs^* \Sigma^{-1}\obs )
+    \fwd \postcov \fwd^* \obs^* \Sigma^{-1}}.
   \end{split}
@@
 Since the trace of $A := V \nabla \tar(\obs) \in \R^{m \times m}$ is
-just $\tr A = \sum_{j=1}^m e_j^t A e_j$ (with $e_j$ the $j$th standard
+just $\tr{A} = \sum_{j=1}^m e_j^t A e_j$ (with $e_j$ the $j$th standard
 basis vector), we see that
 \begin{align*}
-  \delta \tar(\obs)V = \tr V \nabla \tar(\obs) = \sum_{j=1}^m
+  \delta \tar(\obs)V = \tr{V \nabla \tar(\obs)} = \sum_{j=1}^m
   V_j(\nabla \tar(\obs)_j),
@@
 better understanding of the constraints. 
-\begin{proposition}\label{prop:bigger better}
+\begin{restatable*}{proposition}{biggerbetter}\label{prop:bigger better}
   Let $\obs = (\meas_1,\dots,\meas_m)^t$, $j \in \{1,\dots,m\}$, $\sigma^2
@@
   $\lambda \meas_j$ in $\obs$ instead of $\meas_j$.
-\end{proposition}
+\end{restatable*}
 Proof of Proposition \ref{prop:bigger better} is delegated to the
@@
   %
-  &= \tr V \obs^* e_je_j^t.
+  &= \tr{V \obs^* e_je_j^t}.
 \end{align*}
@@
 self-adjoint operator $\sigma^{-2}\fwd \postcov \fwd^*$ where the rows
-of $\obs$, namely $\meas_j,j=1,\dots, m$, are the eigenvectors. If $\fwd$
-is invertible, then we may write:
+of $\obs$, namely $\meas_j,j=1,\dots, m$, are the
+eigenvectors. \refthree{Below, we use \eqref{eq:postcov} in the second
+  equality.} If $\fwd$ is invertible, \refthree{\sout{then we may
+    write} the third equality is trivial. Otherwise, we use some
+  calculations delegated to lemma \ref{lemma:twice woodbury} in the
+  appendix}:
 \begin{align}\label{eq:mod conditions}
@@
   %
-  &= \sigma^{-2} \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1} \obs^*,
+  &= \sigma^{-2} \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1} \obs^*
   \end{split}
 \end{align}  
-which presents the optimal design problem as a problem on $\ban$ and
-not on $\hil$ --- the term $(\fwd \prcov \fwd^*)^{-1}$ is the prior
-precision in $\ban$ and the big parentheses enclose the posterior
-precision in $\ban$. This expression makes sense even if $\fwd$ is not
-invertible --- consider instead $\fwd$ restricted to $\ker
-\fwd^{\perp}$. This does not change the optimal design problem at all,
-since we will never take measurements that have a component in $\ker
-\fwd$ (this can be understood from Corollary \ref{cor:zero mod err}).
+\refthree{\sout{which} This} presents the optimal design problem as a
+problem on $\ban$ and not on $\hil$ --- the term $(\fwd \prcov
+\fwd^*)^{-1}$ is the prior precision in $\ban$ and \reftwo{\sout{the
+    big parentheses enclose} $(\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}
+  \obs^* \obs$ is} the posterior precision in
+$\ban$. \refthree{\sout{This expression makes sense even if $\fwd$ is
+    not invertible --- consider instead $\fwd$ restricted to $\ker
+    \fwd^{\perp}$. This does not change the optimal design problem at
+    all, since we will never take measurements that have a component
+    in $\ker \fwd$ (this can be understood from Corollary
+    \ref{cor:zero mod err}).}}
 
@@
 $\fwd \prcov \fwd^*$ by $\lambda_1 \geq \lambda_2 \geq \dots$ and its
-eigenvectors $\{\ev_i\}_{i=1}^{\infty}$. The following lemma will be useful:
-
-\begin{lemma}[Proved as lemma \ref{lemma:sim diag} in the appendix]
+eigenvectors $\{\ev_i\}_{i=1}^{\infty}$. The following lemma (proved
+in the appendix \ref{lemma:sim diag}) will be useful:
+
+\begin{restatable*}{lemma}{simdiag}\label{lemma:sim diag}
   Let $C:\hil \to \hil$ self-adjoint and let $\func_1,\dots,\func_m \in
@@
   diagonalizable.
-\end{lemma}
+\end{restatable*}
 
@@
 $\obs^*\obs$ by $\{\eta_i\}_{i=1}^{k}$ and let $\eta_i = 0$ for $i
-\geq k+1$. By the argument in the paragraph following \eqref{eq:mod
-  conditions}:
+\geq k+1$. \refthree{\stkout{By the argument in the paragraph following
+    \eqref{eq:mod conditions}:}}
+
+\refthree{
+\begin{align}%\label{eq:old pre manipulations}
+  \cancel{
+  \begin{split}
+    I + \sigma^{-2} \prcov^{1/2} \fwd ^* \obs^*
+    \obs \fwd \prcov^{1/2}\right ) &= \prcov^{1/2} ( \prcov^{-1} +
+    \sigma^{-2} \fwd^*\obs^* \obs \fwd ) \prcov^{1/2} \\
+    %
+    %
+    %
+    &=\prcov^{1/2} \fwd^* ( \fwd^{-*}
+    \prcov^{-1}\fwd^{-1} + \sigma^{-2} \obs^* \obs ) \fwd \prcov^{1/2}
+    (\fwd\prcov\fwd^*)^{-1}\\
+    %
+    %
+    %
+    &= \prcov^{1/2} \fwd^* \left ( (\fwd \prcov \fwd^* )^{-1}
+    +\sigma^{-2} \obs^* \obs \right ) \fwd \prcov^{1/2}.
+  \end{split}
+  }
+\end{align}
+}
+
+\refthree{
 \begin{align}\label{eq:pre manipulations}
   \begin{split}
-    I + \sigma^{-2} \prcov^{1/2}  \fwd ^* \obs^* \obs \fwd \prcov^{1/2} &= 
-    \prcov^{1/2} ( \prcov^{-1} + \sigma^{-2} \fwd^*\obs^*  \obs \fwd ) \prcov^{1/2}\\
-    %
-    %
-    %
-    &=\prcov^{1/2} \fwd^* ( \fwd^{-*} \prcov^{-1}\fwd^{-1} + \sigma^{-2} \obs^* \obs  ) \fwd \prcov^{1/2}\\
-    %
-    %
-    %
-    &= \prcov^{1/2} \fwd^* \left ( (\fwd \prcov \fwd^* )^{-1} +\sigma^{-2} \obs^* \obs  \right ) \fwd \prcov^{1/2}.
+    \det \left (I + \sigma^{-2} \prcov^{1/2} \fwd ^* \obs^*
+    \obs \fwd \prcov^{1/2}\right ) &=
+    %
+    \det \left (  \fwd \prcov^{1/2}(I + \sigma^{-2} \prcov^{1/2} \fwd ^* \obs^*
+    \obs \fwd \prcov^{1/2}) \prcov^{1/2} \fwd^* (\fwd \prcov \fwd^*)^{-1} \right ) \\
+    %
+    %
+    %
+    &= \det \left ( (\fwd \prcov \fwd^* + \sigma^{-2}\fwd \prcov
+    \fwd^* \obs^*\obs \fwd \prcov \fwd^*) (\fwd \prcov \fwd^*)^{-1}
+    \right )\\
+    %
+    %
+    %
+    &= \det \left ( ((\fwd \prcov \fwd^*)^{-1} + \sigma^{-2}
+    \obs^*\obs ) (\fwd \prcov \fwd^*) \right )\\
   \end{split}
 \end{align}
+}
+
 By definition of the objective \eqref{eq:objective} and using
@@
     %
-    &=\frac12 \log \det
-    \left ( (\fwd \prcov \fwd^* )^{-1} + \sigma^{-2}\obs^* \obs  \right ) \fwd \prcov \fwd^* \\ 
+    %
+    &=\frac12 \log \det \Big ( \left ( (\fwd \prcov \fwd^* )^{-1} +
+    \sigma^{-2}\obs^* \obs \right ) \fwd \prcov \fwd^* \Big )\\
     %
@@
 more weight on the better measurement, as can be understood from the
-discussion in the following paragraph). We are also used to symmetric
-optimization problems having symmetric solutions \cite{Waterhouse83},
-so symmetric and identical Lagrange multipliers definitely make sense
-in this context.
+discussion in the following paragraph).%%  We are also used to symmetric
+%% optimization problems having symmetric solutions \cite{Waterhouse83},
+%% so symmetric and identical Lagrange multipliers definitely make sense
+%% in this context.
 
@@
 
+\reftwo{Results of section \ref{subsub:characterization} specifically,
+  the discussion at the end of said section, along with figure
+  \ref{figure:clusterization} and eqaution \eqref{eq:true target} show
+  how clusterization arises when an optimal design is saught.
+
+  }
 \section{Analysis of Optimal Designs --- Non-Vanishing Model Error}
@@
 (\meas_1,\dots,\meas_{m-1})^t$. The following corollary is proved in
-the appendix.
-\begin{corollary}[Proved as corollary \ref{cor:same meas} in the appendix]
+appendix \ref{cor:same meas proof}.
+\begin{restatable*}{corollary}{samemeas}\label{cor:same meas}.
   If $\meas_m = \meas_j$ for some $1 \leq j \leq m-1$, then
@@
   \end{equation*}
-\end{corollary}
+\end{restatable*}
 
@@
   Theorem 4, pp. 127]{Lax97}
-\begin{lemma}\label{lemma:lax}
-  Let $Y(t)$ be a differentiable operator-valued function. Assume 
-  $I+Y(t)$ is invertible, $Y(t)$ self-adjoint and trace-class. Then
-  \begin{equation*}
-    \frac{\der \log \det (I+Y(t))}{\der t} = \tr (I+Y(t))^{-1} \dot{Y}(t).
-  \end{equation*}
-\end{lemma}
+\lax
 \begin{proof}
@@
     = \sum_{k=1}^{\infty} \dot{\lambda_k}(X(0)) 
-    = \frac{\der }{\der t}\tr X(0)
-    = \tr \dot{X}(0),
+    = \frac{\der }{\der t}\tr{X(0)}
+    = \tr{\dot{X}(0)},
   \end{align*}
@@
   \begin{align*}
-    \tr X(t) = \tr \big ( (I+Y(0))^{-1} (I+Y(t)) - I \big )
-    \leq \tr \big( I+Y(t) - I \big ) < \infty,
+    \tr{X(t)} = \tr{(I+Y(0))^{-1} (I+Y(t)) - I}
+    \leq \tr{I+Y(t) - I}< \infty,
   \end{align*}
@@
     % 
-    &= \det (I+Y(0)) \tr \dot{X}(0) \\
-    % 
-    % 
-    % 
-    &= \det (I+Y(0)) \tr (I+Y(0))^{-1} \dot{Y}(0).
+    &= \det (I+Y(0)) \tr{\dot{X}(0)} \\
+    % 
+    % 
+    % 
+    &= \det (I+Y(0)) \tr{(I+Y(0))^{-1} \dot{Y}(0)}.
   \end{align*}
@@
     % 
-    &= \tr (I+Y(t))^{-1} \dot{Y}(t) \big |_{t=0}.
+    &= \tr{ (I+Y(t))^{-1} \dot{Y}(t)} \big |_{t=0}.
   \end{align*}
@@
   Let $\hil$ a separable Hilbert space, $u,v\in \hil$ and $A: \hil \to
-  \hil$ an invertible linear operator such that $\tr A-I <
+  \hil$ an invertible linear operator such that $\tr{A-I} <
   \infty$. Then $\det A$ and $\det A + uv^*$ are well defined and
@@
 
-  It is easy to verify that $uv^*$ is trace-class and since $\tr A - I
-  < \infty$, also $\tr A + uv^* - I < \infty$ (sum of two trace-class
+  It is easy to verify that $uv^*$ is trace-class and since $\tr{A-I}
+  < \infty$, also $\tr{A + uv^* - I} < \infty$ (sum of two trace-class
   operators is trace-class). Thus $\det A$ and $\det (A+uv^*)$ are
@@
 \section{Specific Lemmas}
-
+\refthree{
+\begin{lemma}\label{lemma:twice woodbury}
+  Assume $\fwd \prcov \fwd^*$ is invertible. Then
+\begin{align*}
+  \begin{split*}
+    \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* 
+    %
+    %
+    = \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1},
+  \end{split*}
+\end{align*}  
+\end{lemma}
+\begin{proof}
+  The proof is boring and amounts to using Woodbury's matrix identity
+  twice and a regularization trick. The standard proof for this
+  identity works in infinite dimensions, as long as all terms are well
+  defined. Unfortunately, $\obs^*\obs$ is not invertible, so we force
+  it to be. Throughout, we let $X := \fwd\prcov \fwd^*$ and $Y_{\eps}
+  := \obs^*\obs + \eps I$, for some small $\eps > 0$. Note that now
+  $Y_{\eps}$ is invertible. Recall Woodbury's matrix identity:
+  $$
+  (A + UCV)^{-1} = A^{-1} + A^{-1}U(C^{-1} + VA^{-1}U)^{-1}VA^{-1}.
+  $$
+
+  Taking $A := \prcov^{-1}, U := \fwd^*, V := \fwd$ and $C :=
+  \sigma^{-2} (\obs^*\obs+\eps I)$ we get:
+  \begin{align*}
+    \begin{split}
+      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* &=
+      \fwd ( \prcov - \prcov \fwd^* ( \sigma^2(\obs^*\obs + \eps I)^{-1} + \fwd \prcov \fwd^* )^{-1} \fwd \prcov ) \fwd^* \\
+      %
+      %
+      &= X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X
+    \end{split}
+  \end{align*}
+  Note that $X + \sigma^2 Y_{\eps}^{-1}$ is invertible, as the sum of
+  a two positive definite operators. Now, taking $A := X^{-1}, C :=
+  \sigma^2Y_{\eps}^{-1}, U := I$ and $V := I$ and using Wodbury's matrix
+  identity in reverse order:
+  \begin{align*}
+    \begin{split}
+      X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X &= (X^{-1} + \sigma^{-2}Y_{\eps})^{-1} \\
+      %
+      %
+      %
+      &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \eps I))^{-1}.
+    \end{split}
+  \end{align*}
+
+  We conclude that $\forall \eps > 0$
+  \begin{align*}
+    \begin{split}
+      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* 
+     &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \eps I))^{-1}.
+    \end{split}
+  \end{align*}
+  Letting $\eps \to 0$ completes the proof.
+\end{proof}
+}
 \begin{lemma}[Increase due to a Measurement]\label{lemma:design increase}
@@
 
-\begin{corollary}[Gain for Identical Measurement]\label{cor:same meas}
-  If $\meas_m = \meas_j$ for some $1 \leq j \leq m-1$, then
-  \begin{equation*}
-    \tar(\obs) - \tar(\obsm) =
-    \log \left ( 1 + \frac{\sigma^2
-      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
-      \obsm^* \Sigmam^{-1}e_j \rangle
-    }{
-      2 - \sigma^2 e_j^t\Sigmam^{-1}e_j 
-    }       
-    \right ).
-  \end{equation*}
-\end{corollary}
-\begin{proof}
+\samemeas
+\begin{proof} \label{cor:same meas proof}
   Denote $A:= \obs \modcov \obs^*$ and $v_j$ the $j$th column of $A$.
@@
 \begin{lemma}\label{lemma:free}
-  Let $M \in \R^{k \times k}$ symmetric positive definite with $\tr M =
+  Let $M \in \R^{k \times k}$ symmetric positive definite with $\tr{M} =
   m$, $m > k$. Then we can find $\func_j \in \R^k,j=1,\dots,m$ with
@@
   \end{equation}
-  and we can expect to do this since we assumed $\tr D = m$.
-
-  Define $C = S^tS - I \in \R^{m \times m}$. Note that $\tr C = 0$ and
+  and we can expect to do this since we assumed $\tr{D} = m$.
+
+  Define $C = S^tS - I \in \R^{m \times m}$. Note that $\tr{C} = 0$ and
   $C$ is diagonal with non-zero entries $d_i-1,i=1,\dots,k$. It suffices
@@
   $R^{(k-1)}$ that leaves rows and columns $k,\dots,m$ unchanged and
-  continue introducing zeros to the diagonal. The assumption $\tr D =
-  m \Rightarrow \tr C = 0$ guarantees we can do that. Taking $V:=
+  continue introducing zeros to the diagonal. The assumption $\tr{D} =
+  m \Rightarrow \tr{C} = 0$ guarantees we can do that. Taking $V:=
   R^{(1)} R^{(2)} \dots R^{(k-1)}R^{(k)}$ completes the proof.
@@
 
-\begin{lemma}\label{lemma:sim diag}
-  Let $C:\hil \to \hil$ self-adjoint and let $\func_1,\dots,\func_m \in
-  \hil$. Denote $\func^*$ the element $\func$ acting as a
-  functional. If
-  \begin{equation*}
-   (C + \sum_{j=1}^m \func_j\func_j^*) \func_k = \xi_k \func_k, k = 1,\dots,m
-  \end{equation*}
-  then $C$ and $\sum_{j=1}^m \func_j \func_j^*$ are simultaneously
-  diagonalizable.
-\end{lemma}
+\simdiag
 \begin{proof}
@@
 
-\begin{proposition}\label{prop:bigger better}
-  Let $\obs = (\meas_1,\dots,\meas_m)^t$, $j \in \{1,\dots,m\}$, $\sigma^2
-  > 0$ and $\lambda > 1$. Then $\tar(\obs)$ increases if we use
-  $\lambda \meas_j$ in $\obs$ instead of $\meas_j$.
-\end{proposition}
+%% \begin{proposition}\label{prop:bigger better}
+%%   Let $\obs = (\meas_1,\dots,\meas_m)^t$, $j \in \{1,\dots,m\}$, $\sigma^2
+%%   > 0$ and $\lambda > 1$. Then $\tar(\obs)$ increases if we use
+%%   $\lambda \meas_j$ in $\obs$ instead of $\meas_j$.
+%% \end{proposition}
+\biggerbetter
 \begin{proof} 
@@
      \delta \tar(\obs) V 
-    &= \tr V ( I - \modcov \obs^*\Sigma^{-1}\obs) \tmp \obs^* \Sigma^{-1} \\
-    % 
-    %
-    %
-    &= \tr e_je_j^t \obs ( I - \modcov \obs^*\Sigma^{-1}\obs) \tmp \obs^* \Sigma^{-1} \\
+    &= \tr{V ( I - \modcov \obs^*\Sigma^{-1}\obs) \tmp \obs^* \Sigma^{-1}} \\
+    % 
+    %
+    %
+    &= \tr{e_je_j^t \obs ( I - \modcov \obs^*\Sigma^{-1}\obs) \tmp \obs^* \Sigma^{-1}} \\
     %

### 2020-09-22 reply.tex
some of the reply is finished

\documentclass{amsart}
\input{definitions}

\title{Corrections}

\begin{document}

\maketitle

\section{Referee 2}
\subsection{Summary}
In this article, the author considers what causes sensor clustering in
optimal experimental design of Bayesian inverse problems. They show
how D-optimal designs may lead to sensor clustering in the case of iid
measurement noise. They also show that clustering does not occur for
spatially correlated measurement error. The paper is somewhat
difficult to follow due to writing and organizations corrections that
should be made, but does present original work. The article is
mathematically rigorous and significant in that the work is a
necessary step toward developing a deeper understanding of sensor
clustering. I suggest a major review and resubmission after reviewing
the comments below.

\subsection{General Comments}

\begin{enumerate}
\item It should be stated in the Abstract whether the work in this
  paper pertains to only linear inverse problems, or if it is
  applicable in general to non-linear inverse problems as well.
  \ydnote{Title and abstract now reflect the fact that the paper only
    pertains to linear inverse problems.}

\item It is unclear from the contribution section of the article what
exactly the primary contribution is. This should be stated very
explicitly and emphasized clearly as the contribution of the article.

\item In some equations (as seen below in the line comments) operators
are not clearly defined or described before they are given in a
proposition or theorem. Please clearly define all operators both with
mathematical definitions and descriptions (when applicable) before
using them. \ydnote{Corrected, thanks!}

\item Overall, I find the article difficult to follow. I think this can
be improved by properly defining all operators and including section
overviews that give a description of the purpose of each section, and
a brief summary of how this is being accomplished.

\item On page 10, figure 2 and section 4.2 deliver the main conclusion of
the article. These areas are deficient with only brief descriptions,
and should be much more thoroughly explained. In particular page 10,
lines 38-39 "If the precision in these is lower than the precision for
the next eigenvector, a repeated measurement is optimal" needs to be
elaborated on. How do you come to this conclusion? Additionally, the
wording of the statement can lead to confusion. From figure 2, it
looks like all measurements grant less precision than that of the
fourth eigenvector, so I first interpret this as all measurements
should be repeated. I believe these issues can be fixed by significant
elaboration in section 4.2.
\end{enumerate}

\subsection{Line Comments}

\begin{itemize}
\item Page 1, Line 39: "In a frequentist settings" should be changed
  to "in a frequentist setting" or "in frequentist
  settings". \ydnote{changed in text}

\item Page 2, Line 36: Add a comma: "In the process, we generalize..."
  \ydnote{changed in text}

\item Page 3, Line 37: "with homogeneous Dirichlet boundary condition"
  should be changed to "with a homogeneous Dirichlet boundary
  condiiton". \ydnote{changed}

\item Page 4, Line 14: In equation (2.1), $\Sigma$ has not been
  defined. Please formally define this before giving the equation.
  \ydnote{Moved the equation after $\Sigma$ is defined}

\item Page 4, Line 21: The prior distribution $\mathcal{N}(0,
  \Gamma_{pr})$ assumes the prior mean is 0. This is worthy of some
  comment. Why must the prior mean be 0? \ydnote{It does not have to
    be, we can subtract the mean. An explanation was added}

\item Page 5, Line 15: Add a comma: "Considering the sum of the error
  terms, it is easy..."  \ydnote{changed in text}

\item Page 6, Line 7: Please describe what $\psi$ is in the
  theorem. It is defined mathematically but not described. Only much
  later is it stated that this is the objective
  function. \ydnote{added comment before statment of theorem where
    $\tar$ first appears}

\item Page 6, Line 19-20: "In section 3.2 we relax the constraints on
  $\mathcal{O}$ so that the optimization problem in (3.1) and find
  gradients for the new constraints." This sentence is unclear and
  should be reworded. \ydnote{added short explanation}

\item Page 7, Line 24, 27, 31: The proposition is labeled as
  Proposition 1, but it is refered to as Proposition 2, and relabeled
  Proposition 2 in the appendix. \ydnote{all such errors were fixed}

\item Page 9, Line 6: Please be more exact when refering to "the big
  parentheses" to avoid any chance of confusion. Give the explicit
  formula here as you do for another term earlier in the sentence.
  \ydnote{explicit expression given}

\item Page 18, Line 6: This proposition was previously labeled
  Proposition 1. \ydnote{all such errors were fixed}
\end{itemize}

\section{Referee 3}

\subsection{Overview}

This article discusses D-optimal experimental design for infinite
dimensional Bayesian linear inverse problems, and attempts to provide
a mathematical foundation to understanding the issue of clusterization
in such scenario, i.e., finding D-optimal designs.  As the author
mentioned in the Acknowledgments Section, the paper is paper is part
of the Ph.D. dissertation, and in general provides a good prospective
to sensor placement for Bayesian linear inverse problems formulated
using infinite dimensional approach.  I regret, however, to say that
the paper requires a major revision before being considered for
publication.  Below, I provide some comments, that hopefully will help
improve the manuscript.  Also, provided are some recent references,
the author may find useful.

\subsection{General comments}

The author should clearly state why one should consider
infinite-dimensional formulation of the observation operator? One
could just merge the observation operator into the forward simulation
operator, and follow standard formulations adopted by the works in the
references below. Moreover, when it comes to sensor placement,
observations are almost surely pointwise in space/time. In this case,
I don't see the benefit of consider infinite-dimensional formulation.

\ydnote{I was not able to arrive at the same results when the forward
  and observation operators were merged. A short remark on this was
  added. It is placed after equation (2.6) in the highlighted text
  (equation (2.5) in the new clean text. Specifically, we need it in
  equation (3.1).}

The contributions detailed in Subsection 1.2 (page 2) are very much
appreciated, however they raise some concerns:

\begin{enumerate}
\item line 30: the author claims to present an analytically tractable
  model for understanding D-optimal designs. By reading the paper, I
  couldn't figure out how this reflects to any practical scenario. I'm
  not sure how this can be utilized for example in spatial sensor
  placement problems. An algorithmic description would be very useful
  here.

\item The statement made on page 2, lines 40-41 (... apply equally to
  any of the scenarios...) is rather strong and unjustified.
\end{enumerate}

On page 8, Equation 4.2, which is crucial in this study, assumes that
the forward operator is invertible. This is clearly impossible in
practical Bayesian inverse problems, otherwise, it reduces to a
standard regression problem. This is understandable for clarification,
however, the following argument on Page 9 that extends this assumption
is not well-justified. \ydnote{This is a great point! The flawed
  explanation was removed. I was able to remove that assumption.}

In Equation 4.3, not sure what $\fwd^{-*}$ stands for. Moreover, an
inverse of the forward operator is not well-defined even if
$\mathcal{F}\Gamma_{prior}\mathcal{F}^*$ is invertible! This is very
limiting, and can't be used in practical inverse problems. This should
be clearly discussed for the reader to appreciate the valuable effort
made. \ydnote{Again, great point. $ \fwd^{-*} = (\fwd^*)^{-1}$ but now
  I do not assume $\fwd$ is invertible so this is removed}

A numerical example with the setup studied in the paper, should help
clarify several confusions that a reader can fall for easily.

The title should reflect the fact that the study is restricted to
linear inverse problems. Also, this should be clarified in the
abstract, in addition to the fact that observations errors are
uncorrelated with fixed variance $\sigma^2$. \ydnote{The title and
  abstract were changed to reflect the fact that linear problems are concerned.
The treatment of correlated errors was emphasized.}

\subsection{Other Comments}

In the abstract (line 20, page 1), and on page 5, lines 30-40, the
author mentions that the effect of observation-correlations on sensor
clusterization will be considered later, but I failed to find any
relevant discussion to that issue which I find interesting. Even if
the model error covariance $\Gamma_{model}$ is known, this doesn't
amount for observation correlations, .e.g,
single-instrument-observations, because the observation operator
$\mathcal{O}$ is non invertible. This issue should addressed and
clarified clearly.

I found it difficult to trace the referenced Lemmas, Corollaries,
etc., mainly because bad referencing in the appendix, and the links do
not works properly.  For example Corollary 2 should be Corollary B.x,
etc. \ydnote{All such bad references were fixed}

The trace operator should be followed with some sort of brackets to
make it easier to follow and validated the formulae
\ydnote{curly brackets added everywhere}

Sensor clusterization in OED is central to this study. Figure 2 is
used to illustrate the phenomena, however, it is underrepresented, and
is hard to decipher. Specifically, Section 4.2 (Page 10, lines 36-39),
and caption of Figure 2 should be rewritten with clear explanation of
the axes, and the plot-generation process.

One page 10, lines 30-35, the argument doesn't seem to be quite
accurate. It's not always true that one associates a higher
weight/value to ``better'' observation. If one defines ``better'' as
lower uncertainty, then, variable observation variances must be
considered. On the other hand, if being ``better'' corresponds to a
more accurate realization of the observation (true snapshot of
reality), then the whole statement would be wrong, because optimal
design relies only on the uncertainty levels in linear Bayesian
settings, as manifested by the posterior covariance matrix.

The last statement of the proof of Lemma A.1 (see page 12, line 24),
stating that ``there is nothing special about $t=0$, and the relation
holds for all t''. This is rather odd. In fact, It is assumed that
$X(t=0)$ explicitly (see line 32 on page 11). A clarification is
required here.

\subsection{Minor Comments}

Page 1, line 22; ... ``calculating derivative of log det of
oeprators''. Please clarify for the reader, which operators you're
referring to.

Page 2, line 11; wrong reference to Figure 1.3?

Page 6; in equation 2.7, don't you need an expectation over all priors
to achieve this result? Please clarify.

Page 6, line 13; the optimization problem is solved for $\mathcal{O}$,
not $\Psi$ ?

Page 6, lines 19; the statement made about relaxing the constraints on
the observation operator is unclear.

Page 6, line 24, I believe the reference to (3.2) should be replaced
with (3.1).

Page 6, line 26, proof is provided in which Appendix? Please state
clearly.

page 6, In the proof of Lemma 3.1, the author utilized the operator
$T$, which is only introduced in the appendix (Proof of Lemma
A.1). Maybe state that clearly here. Also, in the proof, from step 4
to 5, the first and third terms vanished; why?

Page 7; Proposition 1 should be replaced with something like
Proposition 3.2 or something similar to match previous referencing
(e.g., Lemma 3.1).

Page 7, line 27 (and also line 31); the reference to Proposition 2
seems to be incorrect. Also, state clearly which appendix the proof is
delegated to.

Page 10, Corollary 1 should be renamed (see the comment on Proposition
1 above). Same applies to Corollary 3 in the appendix. Better also to
clarify which appendix. Finally, the author utilizes
$\widehat{\Gamma_{post}}$ int the corollary, which is defined only in
the appendix, under the proof.

Page 11, in the proof of Lemma A.1 (lines 39-43), the partial
derivative vanished from step 2 to step 3; please explain!

References:

* - Koval, Karina, Alen Alexanderian, and Georg
Stadler. "Optimal experimental design under irreducible uncertainty
for linear inverse problems governed by PDEs." Inverse Problems
(2020).

* - Attia, Ahmed, and Emil Constantinescu. "Optimal Experimental
Design for Inverse Problems in the Presence of Observation
Correlations." arXiv preprint arXiv:2007.14476 (2020).

Letter reference: DSMa01

\end{document}

### 2020-09-30 changes_highlighted.tex
Save changes mid way

@@
 \section{Preliminaries and Notation}\label{section:prelim}
-
+\reftwo{
+\subsection{Overview}
 The theoretical foundations for inverse problems over function spaces
-can be found in \cite{Stuart10}. This section is mostly devoted to
-setting notation. \reftwo{\sout{At this point, it is useful to record a known result
-regarding the posterior covariance operator of our inverse problem,
-namely}
+can be found in \cite{Stuart10}. In this section we define notations that will be
+  used throughout the paper.
+%% A reader mostly familiar with the
+%% literature may initially want to skip this section and only consult
+%% the list below, which summarizes the notation used. We denote:
+%% \begin{itemize}
+%%   \item Parameter space $\hil$.
+%%   \item Data space $\ban$.
+%%   \item Forward dynamics operator $\fwd: \hil \to \ban$.
+%%   \item Observation operator $\obs: \ban \to \mathbb{R}^m$.
+%%   \item Prior measure $\pr = \mathcal{N}(0, \prcov)$.
+%%   \item Data $\data \in \mathbb{R}^m$.
+%%   \item Observation and model errors, captured in $\Sigma(\obs)$.
+%% \end{itemize}
+%% The design objective was derived in \cite{AlexanderianGloorGhattas14}
+%% and recorded in \eqref{eq:objective} below.
+\sout{At this point, it is useful to record a known result regarding
+  the posterior covariance operator of our inverse problem, namely}
 \begin{align}\label{eq:postcov precursor}
@@
 \section{The Constrained Optimization Problem of D-Optimal Design}\label{section:D and grad}
+
+\subsection{\reftwo{Overview}}
 From \eqref{eq:objective} we wish to characterize solution(s) of the
@@
 multipliers. In section \ref{section:objective} we find the gradient
-of $\tar$. In section \ref{subsec:unit length} we relax the
-constraints on $\obs$ so that the optimization problem in
-\eqref{eq:optimization} \reftwo{is analytically tractable} and find
-gradients for the new constraints.
+of $\tar$ and record it in \eqref{eq:tar grad}. In section
+\ref{subsec:unit length} we relax the constraints on $\obs$ so that
+the optimization problem in \eqref{eq:optimization} \reftwo{is
+  analytically tractable}.  We find gradients for the new constraints
+\reftwo{and record them in \eqref{eq:grad constraints}. Finally, in
+  section \ref{subsec:necessary} we use these to formulate the optimal
+  design problem as a Lagrange multipliers problem in equation
+  \eqref{eq:conditions}.}
 
@@
 
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% SECTION Analysis of Optimal Designs --- Vanishing Model Error
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Analysis of Optimal Designs --- Vanishing Model Error}\label{section:vanishing}
+\subsection{\reftwo{Overview}}
+\reftwo{In this section we analyze optimal designs when no model error
+  is present. We show how sensor clusterization arises in section
+  \ref{subsec:clusterization}. We prepare the ground by first deriving
+  an eigenvalue problem in $\obs$ for an optimal design in section
+  \ref{subsec:eigenproblem}. This eigenvalue problem is recorded in
+  \eqref{eq:mod conditions}. Next, we characterize optimal designs as
+  ones that make all eigenvalues in equal in the above mentioned
+  eigenproblem. This is crucial for clusterization to occur. The
+  derivation is presented in section \ref{subsec:characterization}. }
+
+\subsection{\reftwo{The Eigenvalue Problem}}\label{subsec:eigenproblem}
 When $\modcov = 0$, then $\Sigma= \sigma^2I$. The necessary
@@
 
-\subsection{Characterizing D-Optimal Designs}\label{subsub:characterization}
+\subsection{Characterizing D-Optimal Designs}\label{subsec:characterization}
 In light of \eqref{eq:mod conditions}, let us denote eigenvalues of
@@
         I + \sigma^{-2} \prcov^{1/2} \fwd ^* \obs^*
-    \obs \fwd \prcov^{1/2}\right ) &= \prcov^{1/2} ( \prcov^{-1} +
+        \obs \fwd \prcov^{1/2} &= \prcov^{1/2} ( \prcov^{-1} +
         \sigma^{-2} \fwd^*\obs^* \obs \fwd ) \prcov^{1/2} \\
@@
         \prcov^{-1}\fwd^{-1} + \sigma^{-2} \obs^* \obs ) \fwd \prcov^{1/2}
-    (\fwd\prcov\fwd^*)^{-1}\\
+        (\fwd\prcov\fwd^*)^{-1} \\
         %
@@
     %
-    \det \left (  \fwd \prcov^{1/2}(I + \sigma^{-2} \prcov^{1/2} \fwd ^* \obs^*
-    \obs \fwd \prcov^{1/2}) \prcov^{1/2} \fwd^* (\fwd \prcov \fwd^*)^{-1} \right ) \\
-    %
-    %
-    %
-    &= \det \left ( (\fwd \prcov \fwd^* + \sigma^{-2}\fwd \prcov
-    \fwd^* \obs^*\obs \fwd \prcov \fwd^*) (\fwd \prcov \fwd^*)^{-1}
-    \right )\\
-    %
-    %
-    %
-    &= \det \left ( ((\fwd \prcov \fwd^*)^{-1} + \sigma^{-2}
-    \obs^*\obs ) (\fwd \prcov \fwd^*) \right )\\
+    \det \left (I + \sigma^{-2} \obs^*
+    \obs \fwd \prcov\fwd^* \right ) \text{(Sylvester's determinant theorem)}\\
+    %
+    %
+    %
+    &= \det \left \{ \left ( (\fwd \prcov \fwd^*)^{-1} +
+    \sigma^{-2}\obs^*\obs \right ) (\fwd \prcov \fwd^*)\right \}\\
   \end{split}
@@
 The argument above has the important consequence that the eigenvalues
-in \eqref{eq:eigenproblem} are all equal! This makes perfect
+in \eqref{eq:eigenproblem} are all equal This makes perfect
 sense. First, we know the Lagrange multipliers tell us how much we
@@
 
-\reftwo{Results of section \ref{subsub:characterization} specifically,
+\reftwo{Results of section \ref{subsec:characterization} specifically,
   the discussion at the end of said section, along with figure
-  \ref{figure:clusterization} and eqaution \eqref{eq:true target} show
+  \ref{figure:clusterization} and equation \eqref{eq:true target} show
   how clusterization arises when an optimal design is saught.
@@
 }
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% SECTION Analysis of Optimal Designs --- Non-Vanishing Model Error
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Analysis of Optimal Designs --- Non-Vanishing Model Error}
-Denote $\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm :=
-(\meas_1,\dots,\meas_{m-1})^t$. The following corollary is proved in
-appendix \ref{cor:same meas proof}.
+\subsection{\reftwo{Overview}}
+In this section we show how clusterization is avoided when model
+error does not vanish.
+
+\subsection{\reftwo{Derivation}}
+We rely on corollary \ref{cor:same meas} below, which is proved in the
+appendix. Denote $\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm :=
+(\meas_1,\dots,\meas_{m-1})^t$.
 \begin{restatable*}{corollary}{samemeas}\label{cor:same meas}.
@@
 \end{restatable*}
-
-In contrast to the previous case of vanishing model error, we will see
+ In
+contrast to the previous case of vanishing model error, we will see
 that if $\modcov \neq 0$ clustering will not occur. From Corollary
-\ref{cor:same meas} we can observe that as $\sigma^2 \to 0$, the
-increase in the design criterion gained by taking a measurement
-identical to a previous one approaches zero. We may conclude that for
-small measurement error levels, the clusterization effect will be
-mitigated by the presence of a non-zero model error. Since the design
-criterion is not defined for $\sigma^2 = 0$ and identical
-measurements, we cannot make a statement regarding $\sigma^2 = 0$,
-except in the limiting sense described above.
+\ref{cor:same meas} we can observe that
+\refthree{
+$$
+\lim_{\sigma^2 \to 0} \tar(\obs) -\tar(\obsm) = 0
+$$ \sout{as $\sigma^2 \to 0$, the increase in the design criterion
+  gained by taking a measurement identical to a previous one
+  approaches zero}. Hence no increase in the design criterion is
+achieved. This is clearly sub-optimal and $\obsm$ cannot be an optimal
+design}. We may conclude that for small measurement error levels, the
+clusterization effect will be mitigated by the presence of a non-zero
+model error. Since the design criterion is not defined for $\sigma^2 =
+0$ and identical measurements, we cannot make a statement regarding
+$\sigma^2 = 0$, except in the limiting sense described above.
 
@@
 \begin{align*}
-  \begin{split*}
+  \begin{split}
     \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* 
@@
     = \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1},
-  \end{split*}
+  \end{split}
 \end{align*}  

### 2020-09-30 reply.tex
Save changes mid way

@@
 \maketitle
+
+Both referees made great comments that I feel improve the manuscript
+substantially, and I would like to thank them for that. I followed up
+on every point made. The referees' comments and my replies are listed
+below. Below, my replies are colored \ydnote{cyan}. In the revised
+manuscript, replies to referee two are in \reftwo{green} and replies
+to referee three are in \refthree{red}. Again, thank you for taking
+the time and helping me improve this manuscript.
+
+\section{Referee 1}
+Preliminary Assessment, Editorial Board Membe.r The manuscript can be
+sent to reviewers.
 
@@
   applicable in general to non-linear inverse problems as well.
-  \ydnote{Title and abstract now reflect the fact that the paper only
+  \ydnote{\item Title and abstract now reflect the fact that the paper only
     pertains to linear inverse problems.}
@@
 mathematical definitions and descriptions (when applicable) before
-using them. \ydnote{Corrected, thanks!}
+using them.
+\ydnote{\item Corrected, thanks!}
 
@@
 a brief summary of how this is being accomplished.
+\ydnote{\item Added section overviews and operator definitions.}
 
@@
   to "in a frequentist setting" or "in frequentist
-  settings". \ydnote{changed in text}
+  settings".
+  \ydnote{\item Changed in text}
 
@@
   should be changed to "with a homogeneous Dirichlet boundary
-  condiiton". \ydnote{changed}
+  condiiton". \ydnote{\item Changed}
 
@@
   defined. Please formally define this before giving the equation.
-  \ydnote{Moved the equation after $\Sigma$ is defined}
+  \ydnote{\item Moved the equation after $\Sigma$ is defined}
 
@@
   \Gamma_{pr})$ assumes the prior mean is 0. This is worthy of some
-  comment. Why must the prior mean be 0? \ydnote{It does not have to
-    be, we can subtract the mean. An explanation was added}
+  comment. Why must the prior mean be 0?
+  \ydnote{\item It does not have to be, we can subtract the mean. An explanation was added}
 
 \item Page 5, Line 15: Add a comma: "Considering the sum of the error
-  terms, it is easy..."  \ydnote{changed in text}
+  terms, it is easy..."
+  \ydnote{\item Changed in text}
 
@@
   later is it stated that this is the objective
-  function. \ydnote{added comment before statment of theorem where
+  function.
+  \ydnote{\item Added comment before statment of theorem where
     $\tar$ first appears}
@@
   gradients for the new constraints." This sentence is unclear and
-  should be reworded. \ydnote{added short explanation}
+  should be reworded.
+  \ydnote{\item Added short explanation}
 
@@
   Proposition 1, but it is refered to as Proposition 2, and relabeled
-  Proposition 2 in the appendix. \ydnote{all such errors were fixed}
+  Proposition 2 in the appendix.
+  \ydnote{\item All such errors were fixed}
 
@@
   formula here as you do for another term earlier in the sentence.
-  \ydnote{explicit expression given}
+  \ydnote{\item Explicit expression given}
 
 \item Page 18, Line 6: This proposition was previously labeled
-  Proposition 1. \ydnote{all such errors were fixed}
+  Proposition 1. \ydnote{\item All such errors were fixed}
 \end{itemize}

### 2020-09-30 new_paper.tex
backup, trying to diff

--- 2020-09-30 changes_highlighted.tex
@@
   dimensions]{Sensor Clusterization in D-optimal Design in Infinite
-  Dimensional \reftwo{Linear} Bayesian Inverse Problems}
+  Dimensional Linear Bayesian Inverse Problems}
 
@@
   experimental design for infinite-dimensional Bayesian
-  \reftwo{linear} inverse problems. We suggest an analytically
+  linear inverse problems. We suggest an analytically
   tractable model for such designs and reason how it may lead to
@@
 information gain amounts to minimizing the determinant of the
-posterior covariance matrix. In a frequentist setting\reftwo{\sout{s}}, an optimal
+posterior covariance matrix. In a frequentist setting, an optimal
 design minimizes the volume of the uncertainty ellipsoid \cite[page
@@
 
-In the process\reftwo{,} we generalize several lemmas from linear algebra to
+In the process we generalize several lemmas from linear algebra to
 infinite-dimensional settings --- the Matrix Determinant Lemma and a
@@
     various numbers of sensors, and location of these sensors for the
-    1D heat equation with \reftwo{a} homogeneous Dirichlet boundary described in
+    1D heat equation with homogeneous Dirichlet boundary described in
     section \ref{subsec:example}. Note the clusterization that occurs
@@
 illustration, we present the problem via a toy model --- the 1D heat
-equation in $[0,\pi]$ with \reftwo{a} homogeneous Dirichlet boundary
+equation in $[0,\pi]$ with homogeneous Dirichlet boundary
 conditions.
@@
 initial condition as $u_0 \sim \normal(0,\prcov)$, for $\prcov =
-(-\Delta)^{-1}$ with \reftwo{a} homogeneous Dirichlet boundary condition. It is
+(-\Delta)^{-1}$ with homogeneous Dirichlet boundary condition. It is
 well known \cite{Tarantola05} that for linear problems, with Gaussian
@@
 \section{Preliminaries and Notation}\label{section:prelim}
-\reftwo{
+
 \subsection{Overview}
@@
 %% and recorded in \eqref{eq:objective} below.
-\sout{At this point, it is useful to record a known result regarding
-  the posterior covariance operator of our inverse problem, namely}
-\begin{align}\label{eq:postcov precursor}
-  \stkout{\postcov = (\prcov^{-1} + \fwd^* \obs^* \Sigma^{-1} \obs \fwd )^{-1}.}
-\end{align}
-}
-
-\subsection{Bayesian \reftwo{Linear} Inverse Problems}\label{section:abstract OED}
+
+\subsection{Bayesian Linear Inverse Problems}\label{section:abstract OED}
 Let $\hil$ be a Hilbert space and $\ban$ a Banach space, both
@@
 prime example). We want to infer $\param \in \hil$, some parameter of
-the dynamics given noisy observations of $\fwd \param$. \reftwo{We
-  take a zero mean Gaussian prior $\pr$ for $\param$ (otherwise the
-  mean can just be subtracted everywhere). Hence, \sout{We model}}
-$\param \sim \pr = \normal(0,\prcov)$ with some appropriate covariance
-operator $\prcov$ \cite{Stuart10}. Observations (measurements) are
-taken via the linear observation operator $\obs \in ( \ban^* )^m$,
-where $m$ is the number of observations allowed (e.g. sensors at our
-disposal). In an analogy with linear algebra (where row vectors are
-thought of as linear functionals), we think of $\obs$ as having
-``rows'' $\meas_j, j=1,\dots,m$:
+the dynamics given noisy observations of $\fwd \param$. We take a zero
+mean Gaussian prior $\pr$ for $\param$ (otherwise the mean can just be
+subtracted everywhere). Hence, $\param \sim \pr = \normal(0,\prcov)$
+with some appropriate covariance operator $\prcov$
+\cite{Stuart10}. Observations (measurements) are taken via the linear
+observation operator $\obs \in ( \ban^* )^m$, where $m$ is the number
+of observations allowed (e.g. sensors at our disposal). In an analogy
+with linear algebra (where row vectors are thought of as linear
+functionals), we think of $\obs$ as having ``rows'' $\meas_j,
+j=1,\dots,m$:
 \begin{equation}\label{eq:O}
@@
 \subsubsection{Error Terms}
-Considering the sum of the error terms\reftwo{,} it is easy to see that $\obs
+Considering the sum of the error terms it is easy to see that $\obs
 \eps' + \eps \in \R^m$. It is a centered Gaussian random vector and
@@
 \bar{\eps}$ with $\bar{\eps} \sim \normal(0,\Sigma)$. At this point,
-it is useful to \reftwo{\sout{rewrite the posterior covariance operator of our
-inverse problem from \eqref{eq:postcov precursor}} record a known result
-regarding the posterior covariance operator of our inverse problem,
-namely}
+it is useful to record a known result regarding the posterior
+covariance operator of our inverse problem, namely
 \begin{align}\label{eq:postcov}
@@
 \end{align}
-\refthree{
 Usually, this is written using $\tmp := \obs \fwd$ and $\tmp^*$ but
@@
 forward operator $\fwd$ we write the explicit expression
-\eqref{eq:postcov}.}
+\eqref{eq:postcov}.
 
@@
 infinite-dimensions, a D-optimal design is well-defined as maximizing
-the expected KL divergence between posterior and prior
-\reftwo{(denoted $\tar$ below and throughout the paper)} and is summarized, using our notation,
+the expected KL divergence between posterior and prior (denoted $\tar$
+below and throughout the paper) and is summarized, using our notation,
 in the following theorem:
@@
 
-\subsection{\reftwo{Overview}}
+\subsection{Overview}
 From \eqref{eq:objective} we wish to characterize solution(s) of the
@@
 \ref{subsec:unit length} we relax the constraints on $\obs$ so that
-the optimization problem in \eqref{eq:optimization} \reftwo{is
-  analytically tractable}.  We find gradients for the new constraints
-\reftwo{and record them in \eqref{eq:grad constraints}. Finally, in
-  section \ref{subsec:necessary} we use these to formulate the optimal
-  design problem as a Lagrange multipliers problem in equation
-  \eqref{eq:conditions}.}
+the optimization problem in \eqref{eq:optimization} is analytically
+tractable.  We find gradients for the new constraints and record them
+in \eqref{eq:grad constraints}. Finally, in section
+\ref{subsec:necessary} we use these to formulate the optimal design
+problem as a Lagrange multipliers problem in equation
+\eqref{eq:conditions}.
 
@@
 \section{Analysis of Optimal Designs --- Vanishing Model Error}\label{section:vanishing}
-\subsection{\reftwo{Overview}}
-\reftwo{In this section we analyze optimal designs when no model error
-  is present. We show how sensor clusterization arises in section
+\subsection{Overview}
+In this section we analyze optimal designs when no model error is
+present. We show how sensor clusterization arises in section
 \ref{subsec:clusterization}. We prepare the ground by first deriving
@@
 eigenproblem. This is crucial for clusterization to occur. The
-  derivation is presented in section \ref{subsec:characterization}. }
-
-\subsection{\reftwo{The Eigenvalue Problem}}\label{subsec:eigenproblem}
+derivation is presented in section \ref{subsec:characterization}.
+
+\subsection{The Eigenvalue Problem}\label{subsec:eigenproblem}
 When $\modcov = 0$, then $\Sigma= \sigma^2I$. The necessary
@@
 self-adjoint operator $\sigma^{-2}\fwd \postcov \fwd^*$ where the rows
-of $\obs$, namely $\meas_j,j=1,\dots, m$, are the
-eigenvectors. \refthree{Below, we use \eqref{eq:postcov} in the second
-  equality.} If $\fwd$ is invertible, \refthree{\sout{then we may
-    write} the third equality is trivial. Otherwise, we use some
+of $\obs$, namely $\meas_j,j=1,\dots, m$, are the eigenvectors. Below,
+we use \eqref{eq:postcov} in the second equality. If $\fwd$ is
+invertible, the third equality is trivial. Otherwise, we use some
 calculations delegated to lemma \ref{lemma:twice woodbury} in the
-  appendix}:
+appendix:
 \begin{align}\label{eq:mod conditions}
@@
 \end{align}  
-\refthree{\sout{which} This} presents the optimal design problem as a
-problem on $\ban$ and not on $\hil$ --- the term $(\fwd \prcov
-\fwd^*)^{-1}$ is the prior precision in $\ban$ and \reftwo{\sout{the
-    big parentheses enclose} $(\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}
-  \obs^* \obs$ is} the posterior precision in
-$\ban$. \refthree{\sout{This expression makes sense even if $\fwd$ is
-    not invertible --- consider instead $\fwd$ restricted to $\ker
-    \fwd^{\perp}$. This does not change the optimal design problem at
-    all, since we will never take measurements that have a component
-    in $\ker \fwd$ (this can be understood from Corollary
-    \ref{cor:zero mod err}).}}
+This presents the optimal design problem as a problem on $\ban$ and
+not on $\hil$ --- the term $(\fwd \prcov \fwd^*)^{-1}$ is the prior
+precision in $\ban$ and $(\fwd\prcov\fwd^*)^{-1} + \sigma^{-2} \obs^*
+\obs$ is the posterior precision in $\ban$.
 
@@
 $\obs^*\obs$ by $\{\eta_i\}_{i=1}^{k}$ and let $\eta_i = 0$ for $i
-\geq k+1$. \refthree{\stkout{By the argument in the paragraph following
-    \eqref{eq:mod conditions}:}}
-
-\refthree{
-  \begin{align}%\label{eq:old pre manipulations}
-    \cancel{
-      \begin{split}
-        I + \sigma^{-2} \prcov^{1/2} \fwd ^* \obs^*
-        \obs \fwd \prcov^{1/2} &= \prcov^{1/2} ( \prcov^{-1} +
-        \sigma^{-2} \fwd^*\obs^* \obs \fwd ) \prcov^{1/2} \\
-        %
-        %
-        %
-        &=\prcov^{1/2} \fwd^* ( \fwd^{-*}
-        \prcov^{-1}\fwd^{-1} + \sigma^{-2} \obs^* \obs ) \fwd \prcov^{1/2}
-        (\fwd\prcov\fwd^*)^{-1} \\
-        %
-        %
-        %
-        &= \prcov^{1/2} \fwd^* \left ( (\fwd \prcov \fwd^* )^{-1}
-        +\sigma^{-2} \obs^* \obs \right ) \fwd \prcov^{1/2}.
-      \end{split}
-    }
-\end{align}
-}
-
-\refthree{
+\geq k+1$.
+
 \begin{align}\label{eq:pre manipulations}
@@
 \end{align}
-}
 
@@
 
-\reftwo{Results of section \ref{subsec:characterization} specifically,
-  the discussion at the end of said section, along with figure
+Results of section \ref{subsec:characterization} specifically, the
+discussion at the end of said section, along with figure
 \ref{figure:clusterization} and equation \eqref{eq:true target} show
 how clusterization arises when an optimal design is saught.
-
-}
 
@@
 \section{Analysis of Optimal Designs --- Non-Vanishing Model Error}
-\subsection{\reftwo{Overview}}
+\subsection{Overview}
 In this section we show how clusterization is avoided when model
@@
 
-\subsection{\reftwo{Derivation}}
+\subsection{Derivation}
 We rely on corollary \ref{cor:same meas} below, which is proved in the
@@
 \ref{cor:same meas} we can observe that
-\refthree{
+
 $$
 \lim_{\sigma^2 \to 0} \tar(\obs) -\tar(\obsm) = 0
-$$ \sout{as $\sigma^2 \to 0$, the increase in the design criterion
-  gained by taking a measurement identical to a previous one
-  approaches zero}. Hence no increase in the design criterion is
+$$
+Hence no increase in the design criterion is
 achieved. This is clearly sub-optimal and $\obsm$ cannot be an optimal
-design}. We may conclude that for small measurement error levels, the
+design. We may conclude that for small measurement error levels, the
 clusterization effect will be mitigated by the presence of a non-zero
@@
 \section{Specific Lemmas}
-\refthree{
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{lemma}\label{lemma:twice woodbury}
@@
 \end{proof}
-}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{lemma}[Increase due to a Measurement]\label{lemma:design increase}

### 2020-10-01 paper.tex
Now old paper compiles
backup

@@
   \begin{equation*}
-    \frac{\der \log \det (I+Y(t))}{\der t} = \tr (I+Y(t))^{-1} \dot{Y}(t).
+    \frac{\der \log \det (I+Y(t))}{\der t} = \tr{(I+Y(t))^{-1} \dot{Y}(t)}.
   \end{equation*}
@@
 Since the trace of $A := V \nabla \tar(\obs) \in \R^{m \times m}$ is
-just $\tr A = \sum_{j=1}^m e_j^t A e_j$ (with $e_j$ the $j$th standard
+just $\ttr A = \sum_{j=1}^m e_j^t A e_j$ (with $e_j$ the $j$th standard
 basis vector), we see that
@@
   \begin{equation*}
-    \frac{\der \log \det (I+Y(t))}{\der t} = \tr (I+Y(t))^{-1} \dot{Y}(t).
+    \frac{\der \log \det (I+Y(t))}{\der t} = \tr{ (I+Y(t))^{-1} \dot{Y}(t)}.
   \end{equation*}
@@
     = \frac{\der }{\der t}\tr X(0)
-    = \tr \dot{X}(0),
+    = \ttr \dot{X}(0),
   \end{align*}
@@
   \begin{align*}
-    \tr X(t) = \tr \big ( (I+Y(0))^{-1} (I+Y(t)) - I \big )
-    \leq \tr \big( I+Y(t) - I \big ) < \infty,
+    \ttr X(t) = \tr{(I+Y(0))^{-1} (I+Y(t)) - I}
+    \leq \tr{I+Y(t) - I} < \infty,
   \end{align*}
@@
     % 
-    &= \det (I+Y(0)) \tr \dot{X}(0) \\
+    &= \det (I+Y(0)) \ttr \dot{X}(0) \\
     % 

### 2020-10-01 reply.tex
backup

@@
   settings".
-  \ydnote{\item Changed in text}
+  \ydnote{Changed in text}
 
 \item Page 2, Line 36: Add a comma: "In the process, we generalize..."
-  \ydnote{changed in text}
+  \ydnote{Changed in text}
 
@@
   should be changed to "with a homogeneous Dirichlet boundary
-  condiiton". \ydnote{\item Changed}
+  condiiton". \ydnote{Changed}
 
@@
   defined. Please formally define this before giving the equation.
-  \ydnote{\item Moved the equation after $\Sigma$ is defined}
+  \ydnote{Moved the equation after $\Sigma$ is defined}
 
@@
   comment. Why must the prior mean be 0?
-  \ydnote{\item It does not have to be, we can subtract the mean. An explanation was added}
+  \ydnote{It does not have to be, we can subtract the mean. An explanation was added}
 
@@
   terms, it is easy..."
-  \ydnote{\item Changed in text}
+  \ydnote{Changed in text}
 
@@
   function.
-  \ydnote{\item Added comment before statment of theorem where
+  \ydnote{Added comment before statment of theorem where
     $\tar$ first appears}
@@
   should be reworded.
-  \ydnote{\item Added short explanation}
+  \ydnote{Added short explanation}
 
@@
   Proposition 2 in the appendix.
-  \ydnote{\item All such errors were fixed}
+  \ydnote{All such errors were fixed}
 
@@
   formula here as you do for another term earlier in the sentence.
-  \ydnote{\item Explicit expression given}
+  \ydnote{Explicit expression given}
 
 \item Page 18, Line 6: This proposition was previously labeled
-  Proposition 1. \ydnote{\item All such errors were fixed}
+  Proposition 1. \ydnote{All such errors were fixed}
 \end{itemize}
@@
   here.
+  \ydnote{Unfortunately, this work does not have
+    immediate applications to (e.g.) spatial sensor placement
+    problems. I offer a toy model for understanding sensor
+    clusterization. I plan to use some of the conclusions in
+    subsequent work but I find it out of scope for the work under
+    consideration here.}
 
@@
 however, the following argument on Page 9 that extends this assumption
-is not well-justified. \ydnote{This is a great point! The flawed
-  explanation was removed. I was able to remove that assumption.}
+is not well-justified.
+\ydnote{This is a great point! The flawed explanation was removed. I
+  was able to remove that assumption.}
 
@@
 be clearly discussed for the reader to appreciate the valuable effort
-made. \ydnote{Again, great point. $ \fwd^{-*} = (\fwd^*)^{-1}$ but now
-  I do not assume $\fwd$ is invertible so this is removed}
+made.
+\ydnote{Again, great point. $\fwd^{-*} = (\fwd^*)^{-1}$ but now I do
+  not assume $\fwd$ is invertible so this is removed}
 
@@
 abstract, in addition to the fact that observations errors are
-uncorrelated with fixed variance $\sigma^2$. \ydnote{The title and
+uncorrelated with fixed variance $\sigma^2$.
+\ydnote{The title and
   abstract were changed to reflect the fact that linear problems are concerned.
@@
 not works properly.  For example Corollary 2 should be Corollary B.x,
-etc. \ydnote{All such bad references were fixed}
+etc.
+\ydnote{All such bad references were fixed.}
 
@@
 make it easier to follow and validated the formulae
-\ydnote{curly brackets added everywhere}
+\ydnote{Curly brackets added where needed.}
 

### 2020-10-02 new_paper.tex
Actually tested the algorithm for finding AAt with unit norm columns. It worksgit add zeros.py  Flawlesslygit add zeros.py !

@@
 infinite-dimensions, a D-optimal design is well-defined as maximizing
-the expected KL divergence between posterior and prior (denoted $\tar$
-below and throughout the paper) and is summarized, using our notation,
+the expected KL divergence between posterior and prior and is summarized, using our notation,
 in the following theorem:
@@
 \end{theorem}
+A D-optimal design seeks to maximize $\tar(\obs)$.
 
@@
 Since the trace of $A := V \nabla \tar(\obs) \in \R^{m \times m}$ is
-just $\tr{A} = \sum_{j=1}^m e_j^t A e_j$ (with $e_j$ the $j$th standard
+just $\ttr A = \sum_{j=1}^m e_j^t A e_j$ (with $e_j$ the $j$th standard
 basis vector), we see that
@@
 of $\obs$, namely $\meas_j,j=1,\dots, m$, are the eigenvectors. Below,
-we use \eqref{eq:postcov} in the second equality. If $\fwd$ is
-invertible, the third equality is trivial. Otherwise, we use some
-calculations delegated to lemma \ref{lemma:twice woodbury} in the
-appendix:
+we use the expression for the posterior covariance \eqref{eq:postcov}
+in the second equality. If $\fwd$ is invertible, the third equality is
+trivial. Otherwise, we use some calculations delegated to lemma
+\ref{lemma:twice woodbury} in the appendix:
 \begin{align}\label{eq:mod conditions}
@@
 In light of \eqref{eq:mod conditions}, let us denote eigenvalues of
-$\fwd \prcov \fwd^*$ by $\lambda_1 \geq \lambda_2 \geq \dots$ and its
+$\fwd \prcov \fwd^*$ by $\{\lambda_i\}_{i=1}^{\infty}$ and its
 eigenvectors $\{\ev_i\}_{i=1}^{\infty}$. The following lemma (proved
-in the appendix \ref{lemma:sim diag}) will be useful:
+in the appendix \ref{lemma:sim diag}) is crucial for the rest of the
+analysis:
 
@@
 \eqref{eq:mod conditions} that $\obs^*\obs$ has the same eigenvectors
-as $\fwd \prcov \fwd^*$, namely $\{\ev_i\}_{i=1}^{\infty}$ (infinitely
-many of them with zero eigenvalue, since $\obs^*\obs$ is
+as $\fwd \prcov \fwd^*$, namely $\{\ev_i\}_{i=1}^{\infty}$ (all but
+finitely many of them with zero eigenvalue, since $\obs^*\obs$ is
 finite-rank). Denote the corresponding non-zero eigenvalues of
 $\obs^*\obs$ by $\{\eta_i\}_{i=1}^{k}$ and let $\eta_i = 0$ for $i
-\geq k+1$.
+\geq k+1$. Using Sylvester's determinant theorem we get:
 
@@
     \det \left (I + \sigma^{-2} \obs^*
-    \obs \fwd \prcov\fwd^* \right ) \text{(Sylvester's determinant theorem)}\\
+    \obs \fwd \prcov\fwd^* \right ) \\
     %
@@
     %
-    &=\frac12 \log \left ( \prod_{i=1}^{k} ( \lambda_i^{-1} + \sigma^{-2} \eta_i )\prod_{i=1}^{k} \lambda_i \right ). 
+    &=\frac12 \log \left ( \prod_{i=1}^{k} ( \lambda_i^{-1} + \sigma^{-2} \eta_i )\prod_{i=1}^{k} \lambda_i \right )\\
+    %
+    %
+    %
+    &=\frac12 \sum_{i=1}^{k} \log (\sigma^2\lambda_i^{-1} + \eta_i), \text{ up to an additive constant}. 
   \end{split}
 \end{align}
+
 Therefore, an optimal design maximizes
@@
 \begin{equation}\label{eq:true target}
-  \sum_{i=1}^{k}\log (\lambda_i^{-1} + \sigma^{-2} \eta_i),
+  \sum_{i=1}^{k}\log (\sigma^2 \lambda_i^{-1} + \eta_i),
 \end{equation}
+
+\pgfplotstableread{
+  Label     prior  optimal  sub-optimal 
+  1         0.2    1.3           1.4
+  2         0.8    0.7           0.3
+  3         2.2    0             0.2
+  4         3.5    0             0.1
+}\optimalvsnot
+
+\begin{figure}
+  \begin{tikzpicture}[scale=0.85]
+    \begin{axis}[
+        ybar stacked,
+        ymin=0,
+        ymax=4,
+        xtick=data,
+        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
+        reverse legend=false, % set to false to get correct display, but I'd like to have this true
+        xticklabels from table={\optimalvsnot}{Label},
+        xticklabel style={text width=2cm,align=center},
+        legend plot pos=right,
+        ylabel=precision --- prior and posterior,
+        xlabel=eigenvector $i$,
+      ]
+
+      \addplot [fill=green!80]  table [y=prior,   meta=Label, x expr=\coordindex] {\optimalvsnot};
+      \addplot [fill=blue!60]   table [y=optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};
+
+      \addlegendentry{$\sigma^2\lambda_i^{-1}$}
+      \addlegendentry{$\eta_i$ from an optimal design}
+
+    \end{axis}
+  \end{tikzpicture}
+  \qquad
+  \begin{tikzpicture}[scale=0.85]
+    \begin{axis}[
+        ybar stacked,
+        ymin=0,
+        ymax=4,
+        xtick=data,
+        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
+        reverse legend=false, % set to false to get correct display, but I'd like to have this true
+        xticklabels from table={\optimalvsnot}{Label},
+        xticklabel style={text width=2cm,align=center},
+        legend plot pos=right,
+        ylabel=precision --- prior and posterior,
+        xlabel=eigenvector $i$,
+      ]
+
+      \addplot [fill=green!80]  table [y=prior,       meta=Label, x expr=\coordindex] {\optimalvsnot};
+      \addplot [fill=blue!60]   table [y=sub-optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};
+
+      \addlegendentry{$\sigma^2\lambda_i^{-1}$}
+      \addlegendentry{$\eta_i$ from a sub-optimal desgin}
+
+    \end{axis}
+  \end{tikzpicture}
+  \caption{Characterizing an optimal design, compared to a
+    sub-optimal design. Both designs are allowed two measurements,
+    which means the blue area has accumulated height of $2$ in both
+    plots. An optimal design (left) increases precision (reduces
+    uncertainty) where it is lowest. A sub-optimal design (right) does
+    not. In this example, a measurement is taken partly in the direction of the
+    fourth eigenvector, where precision is highest.}
+  \label{fig:optimal vs not}
+\end{figure}
 
@@
 \end{figure}
+
 with the norm constraints on $\meas_j,j=1,\dots,m$. We show in the
@@
 increase in the design criterion (equation \eqref{eq:true target}) is
-gained by increasing $\lambda_i^{-1} + \sigma^{-2}\eta_i$ where it is
-smallest. So, no weight should be given to (i.e. measurement taken of)
-the $k$th eigenvector before $\lambda_i^{-1} + \sigma^{-2}\eta_i \geq
-\lambda_k^{-1},\ \forall i \neq k$. Thus, we should choose
-$\{\eta_i\}_{i=1}^k$ as follows. We would increase $\lambda_1^{-1} +
-\sigma^{-2} \eta_1$ until it equals $\lambda_2^{-1}$. Then,
-$\lambda_i^{-1} + \sigma^{-2} \eta_i,i=1,2$ are increased until
-$\lambda_i^{-1} + \sigma^{-2} \eta_i = \lambda_3^{-1},i=1,2$ and so
-forth.
-
-The argument above has the important consequence that the eigenvalues
-in \eqref{eq:eigenproblem} are all equal This makes perfect
-sense. First, we know the Lagrange multipliers tell us how much we
-gain by relaxing the constraints and there is no reason for any
-measurement to give us more than any other (otherwise we would put
+gained by increasing $\sigma^2\lambda_i^{-1} + \eta_i$ where it is
+smallest. So, no weight should be given to (i.e. measurement taken in
+direction of) the $k$th eigenvector before $\sigma^2\lambda_i^{-1} +
+\eta_i \geq \sigma^2\lambda_k^{-1},\ \forall i \neq k$. Thus, we
+should choose $\{\eta_i\}_{i=1}^k$ as follows. First, order the
+eigenvalues $\lambda_1 \geq \lambda_2 \geq \dots$. We should increase
+$\sigma^2 \lambda_1^{-1} + \eta_1$ until it equals
+$\sigma^2\lambda_2^{-1}$. Then, $\sigma^2 \lambda_i^{-1} + \eta_i,i=1,2$ are
+increased until $\sigma^2 \lambda_i^{-1} + \eta_i =
+\sigma^2\lambda_3^{-1},i=1,2$ and so forth.
+
+The argument above has the important consequence (illustrated in
+figure \ref{fig:optimal vs not}) that the eigenvalues in the
+eigenvalue problem \eqref{eq:eigenproblem} are all equal! This makes
+perfect sense. First, we know the Lagrange multipliers tell us how
+much we gain by relaxing the constraints and there is no reason for
+any measurement to give us more than any other (otherwise we would put
 more weight on the better measurement, as can be understood from the
-discussion in the following paragraph).%%  We are also used to symmetric
-%% optimization problems having symmetric solutions \cite{Waterhouse83},
-%% so symmetric and identical Lagrange multipliers definitely make sense
-%% in this context.
-
-\subsection{Sensor Clusterization}\label{subsec:clusterization}
-The way clusterization can occur is best understood from figure
-\ref{fig:clusterization}. We see that each measurement increases
-precision (reduces uncertainty) only for the eigenvectors of lowest
-precision. If the precision in these is lower than the precision for
-the next eigenvector, a repeated measurement is optimal.
+discussion in the following paragraph). This result of the analysis
+above is summarized in the theorem below.
+
+\begin{theorem}[Characterizing D-optimal designs]\label{thm:char}
+In a D-optimal design $\eta_i, i=1,\dots,k$ are chosen such that
+$\sigma^2\lambda_i^{-1} + \eta_i < \sigma^2\lambda_k^{-1}$ and
+$\sigma^2\lambda_i^{-1} + \eta_i$ are all equal for $1 \leq i \leq k$
+for some $k$. For $i \geq k+1$, however, $\eta_i = 0$.
+\end{theorem}
+
+%% We are also used to symmetric optimization problems having
+%% symmetric solutions \cite{Waterhouse83}, so symmetric and identical
+%% Lagrange multipliers definitely make sense in this context.
+
+\section{Sensor Clusterization --- Vanishing Model Error}\label{section:clusterization}
+\subsection{Overview}
+In this section we explain how sensor clusterization may arise. We
+first consider a sequential optimal design, where measurements are
+chosen sequentially. We then consider the case where all measurements
+are chosen at once. Finally, we hypothesize how this may apply to a
+real-life inverse problem.
+
+\subsection{Sequential Design}
+The way clusterization can occur is illustrated in figure
+\ref{fig:clusterization} and understood using theorem
+\ref{thm:char}. The first measurement $\meas_1$ is optimal, since
+$\eta_i, i=1,2$ are chosen to increase and equalize
+$\sigma^2\lambda_i^{-1} +\eta_i, i=1,2$, which are lowest. In this
+case, $\meas_1$ has no component in the direction of $\ev_i, i\geq
+3$. We see that each measurement increases precision (reduces
+uncertainty) only for the eigenvectors of lowest precision. Consider a
+sequential optimization scenario. We first seek a measurement to
+maximize our our utility, given the prior. From , the posterior
+covariance does not depend on data observed, so we know it after
+deciding on the first measurement.
+
+the best
+first measuremen, thIf the precision in these is lower than the
+precision for the next eigenvector, a repeated measurement is optimal.
 
@@
 \subsection{Overview}
-In this section we show how clusterization is avoided when model
-error does not vanish.
+In this section we show the effect model error has on the
+clusterization phenomenon. In contrast to the previous case of
+vanishing model error, we will see that if $\modcov \neq 0$ clustering
+will not occur.
 
@@
 (\meas_1,\dots,\meas_{m-1})^t$.
-\begin{restatable*}{corollary}{samemeas}\label{cor:same meas}.
+\begin{restatable*}{corollary}{samemeas}\label{cor:same meas}
   If $\meas_m = \meas_j$ for some $1 \leq j \leq m-1$, then
@@
 \end{restatable*}
- In
-contrast to the previous case of vanishing model error, we will see
-that if $\modcov \neq 0$ clustering will not occur. From Corollary
-\ref{cor:same meas} we can observe that
+From Corollary \ref{cor:same meas} we can deduce that
 
@@
 \begin{lemma}\label{lemma:free}
-  Let $M \in \R^{k \times k}$ symmetric positive definite with $\tr{M} =
+  Let $M \in \R^{k \times k}$ symmetric positive definite with $\ttr M =
   m$, $m > k$. Then we can find $\func_j \in \R^k,j=1,\dots,m$ with
@@
     vi}. Using \eqref{eq:on vi} again, we conclude that on $V_i$, $A_i
-  = \xi_iI - C$. This immediately implied $A_i$ and $C$ have the
+  = \xi_iI - C$. This immediately implies $A_i$ and $C$ have the
   same eigenvectors on $V_i$. This holds for every $1 \leq i \leq

### 2020-10-02 zeros.py

@@ -1,0 +1,175 @@
+import numpy as np
+from numpy import sin, cos
+from scipy.stats import ortho_group
+import pdb
+
+ortho = ortho_group.rvs
+cot = lambda x: cos(x)/ sin(x)
+def conj(D, U):
+    ''' Returns UDU^t'''
+    assert U.shape == D.shape
+    return np.einsum('ij, jk, lk -> il', U, D, U)
+
+
+def givens(theta, dims, lower, upper):
+    assert theta >= 0 and theta <= np.pi
+    assert lower < upper
+    assert upper <= dims
+    R = np.zeros((dims,dims))
+    for a in range(dims):
+        for b in range(dims):
+            if a == b: 
+                if a == b == lower:
+                    R[a,b] = cos(theta)
+                elif a == b == upper:
+                    R[a,b] = cos(theta)
+                else:
+                    R[a,b] = 1
+            elif a == lower and b == upper:
+                R[a,b] = -sin(theta)
+            elif a == upper and b == lower:
+                R[a,b] = sin(theta)            
+    return R
+
+
+def MDUS(m, k):
+    assert m > k
+    
+    D = np.abs(np.random.uniform(size=k))
+    D = D * m / np.sum(D)
+    assert abs(np.sum(D) - m) < 1e-12 
+    D = np.diag(D)
+    assert D.shape == (k,k)
+
+    U = ortho(k)
+    assert np.all(np.abs(np.dot(U,U.T) - np.identity(k)) < 1e-12)
+
+    M = conj(D, U)
+
+    S = np.zeros((k,m))
+    for i in range(k):
+        S[i,i] = np.sqrt(D[i,i])
+
+    assert M.shape == D.shape == U.shape == (k,k)
+    assert S.shape == (k,m)
+    
+    return M, D, U, S
+
+def get_theta(C, upper):
+    assert upper < C.shape[0]
+    assert C.shape[0] == C.shape[1]
+    if upper == 0:
+        raise ValueError(f'Why upper == 0 and C.shape == {C.shape}?')
+    if C[upper,upper] == 0:
+        return None, None
+
+    success = False
+    for lower in reversed(range(upper)):
+        ckk = C[upper,upper]
+        cpp = C[lower,lower]
+        if ckk * cpp < 0:
+            success = True
[…]

### 2020-10-02 zeros.py
Modified the code to find A a bit

@@ -3,8 +3,11 @@ from numpy import sin, cos
 from scipy.stats import ortho_group
 import pdb
 
+EPS = 1e-14
 ortho = ortho_group.rvs
 cot = lambda x: cos(x)/ sin(x)
+
+
 def conj(D, U):
     ''' Returns UDU^t'''
     assert U.shape == D.shape
@@ -37,12 +40,12 @@ def MDUS(m, k):
     
     D = np.abs(np.random.uniform(size=k))
     D = D * m / np.sum(D)
-    assert abs(np.sum(D) - m) < 1e-12 
+    assert abs(np.sum(D) - m) < EPS 
     D = np.diag(D)
     assert D.shape == (k,k)
 
     U = ortho(k)
-    assert np.all(np.abs(np.dot(U,U.T) - np.identity(k)) < 1e-12)
+    assert np.all(np.abs(np.dot(U,U.T) - np.identity(k)) < EPS)
 
     M = conj(D, U)
 
@@ -96,10 +99,31 @@ def get_theta(C, upper):
     
 def main(m, k):
     assert m > k
-    M, D, U, S = MDUS(m, k)
+    M, _, _, _ = MDUS(m, k)
+
+    A = get_A(M, m)
+
+    ## Check that AAt = M, as stated.
+    AAt = np.dot(A,A.T)
+    assert np.all(np.abs(AAt - M) < EPS)
+    
+    ## Check that A does have unit norm columns, as stated.
+    norms = np.linalg.norm(A, axis=0)
+    assert np.all(np.abs(norms - 1) < EPS) 
+
+    
+def get_A(M, m):
+    k = M.shape[0]
+    assert abs(m - np.trace(M)) < EPS
+    
+    D, U = np.linalg.eig(M)
+    S = np.zeros((k,m))
+    for i in range(k):
+        S[i,i] = np.sqrt(D[i])
+
     
     C = np.dot(S.T, S) - np.identity(m)
-    assert abs(np.trace(C)) < 1e-14
+    assert abs(np.trace(C)) < EPS
 
     V = np.identity(m)
     for upper in reversed(range(1,k+1)):
@@ -107,23 +131,17 @@ def main(m, k):
         R = givens(theta=theta, dims=m, lower=lower, upper=upper)
         C = conj(C, R)
         V = np.dot(R, V)
-  
+
+    A = np.einsum('ij, jk, lk ->il', U, S, V)
+
[…]

### 2020-10-02 zeros.py
Small error, now test pass uniformly and consistently

@@ -3,7 +3,7 @@ from numpy import sin, cos
 from scipy.stats import ortho_group
 import pdb
 
-EPS = 1e-14
+EPS = 1e-13
 ortho = ortho_group.rvs
 cot = lambda x: cos(x)/ sin(x)
 
@@ -73,19 +73,22 @@ def get_theta(C, upper):
         if ckk * cpp < 0:
             success = True
             break
-    assert success
+    assert success, f'ckk = {ckk:2.3f}, cpp = {cpp:2.3f}, sum = {ckk+cpp:2.3f}' 
     
     ## Comment refers to common names in quadratic equation lingo
-    a = C[upper,upper] 
-    b = 2*C[upper,lower]
-    c = C[lower,lower] 
-    assert a * c < 0
+    c2 = C[upper,upper] 
+    c1 = 2*C[upper,lower]
+    c0 = C[lower,lower] 
+    assert c2 * c0 < 0
     
     ## Solving the quadratic to get cot(theta), which will soon
     ## be inverted to find theta.
-    cot1 = (-b + np.sqrt(b**2 - 4*a*c))/ 2 / a
-    cot2 = (-b - np.sqrt(b**2 - 4*a*c))/ 2 / a ## Second solution
-
+    disc = c1**2 - 4*c2*c0 # Discriminant
+    assert disc >= 0
+    cot1 = (-c1 + np.sqrt(disc))/ 2 / c2
+    cot2 = (-c1 - np.sqrt(disc))/ 2 / c2 ## Second solution
+    
+    
     theta1 = np.arctan(1/cot1)
     theta2 = np.arctan(1/cot2)
     if 0 <= theta1 <= np.pi:
@@ -105,11 +108,11 @@ def main(m, k):
 
     ## Check that AAt = M, as stated.
     AAt = np.dot(A,A.T)
-    assert np.all(np.abs(AAt - M) < EPS)
+    assert np.all(np.abs(AAt - M) < 1e-10)
     
     ## Check that A does have unit norm columns, as stated.
     norms = np.linalg.norm(A, axis=0)
-    assert np.all(np.abs(norms - 1) < EPS) 
+    assert np.all(np.abs(norms - 1) < 1e-10) 
 
     
 def get_A(M, m):
@@ -123,10 +126,9 @@ def get_A(M, m):
 
     
     C = np.dot(S.T, S) - np.identity(m)
-    assert abs(np.trace(C)) < EPS
-
     V = np.identity(m)
[…]

### 2020-10-03 new_paper.tex
Need to revise the introduction, write clear contribution and weaknesses and decide what to do with sylvester's determinant lemma

@@
     variances, hence it is omitted.}
-\end{figure}\label{fig:clusterization illustration}
+  \label{fig:clusterization illustration}
+\end{figure}
 In section \ref{section:abstract OED} we present a more abstract and
@@
 
-\subsection{Bayesian Linear Inverse Problems}\label{section:abstract OED}
+\subsection{Bayesian Linear Inverse Problems}\label{subsec:abstract OED}
 Let $\hil$ be a Hilbert space and $\ban$ a Banach space, both
@@
 
-\section{The Constrained Optimization Problem of D-Optimal Design}\label{section:D and grad}
-
-\subsection{Overview}
+\subsection{Sequential vs Simultaneous Optimization}\label{subsec:seq vs sim}
 From \eqref{eq:objective} we wish to characterize solution(s) of the
@@
 where $\obs$ is constrained to some allowed set of measurements
-(e.g.\ pointwise evaluations on $[0,\pi]$).
+(e.g.\ pointwise evaluations on $[0,\pi]$). We call this simultaneous
+optimization.
+
+For computational reasons, one may prefer to find the best
+measurements in a sequential manner. Denote
+\begin{equation}\label{eq:def obs_k}
+  \obs_k := \sum_{i=1}^k \meas_i\meas^*_i.
+\end{equation}
+Sequential optimal design proceeds as follows. Find $\meas_1$ by
+maximizing $\tar(\obs_1)$. Then, keeping $\meas_1$ fixed --- find
+$\meas_2$ as the maximizer of $\tar(\obs_2)$. Then, find $\meas_3$ by
+keeping $\meas_1,\meas_2$ fixed and taking $\meas_3$ the maximizer of
+$\tar(\obs_3)$. Continue in this fashion until $\obs_m = \obs$ is
+found, where $m$ is the number of available measurements. It is
+important to notice that this scheme does not require actually
+observing data --- in \eqref{eq:objective} data is averaged out.
+
+This paper is mostly focused with the more general simultaneous
+optimization case. The sequential optimization case is dealt with in
+section \ref{subsec:seq clus}. It is important to note, however that
+all conclusions we make on the simultaneous case easily specialize to
+the sequential case by considering the posterior as the next
+sequential step's prior.
+
+\section{The Constrained Optimization Problem of D-Optimal Design}\label{section:D and grad}
+
+\subsection{Overview}
 %%, for the example in section \ref{subsec:example}).
-We seek a formulation of the optimal design using Lagrange
-multipliers. In section \ref{section:objective} we find the gradient
-of $\tar$ and record it in \eqref{eq:tar grad}. In section
-\ref{subsec:unit length} we relax the constraints on $\obs$ so that
-the optimization problem in \eqref{eq:optimization} is analytically
-tractable.  We find gradients for the new constraints and record them
-in \eqref{eq:grad constraints}. Finally, in section
-\ref{subsec:necessary} we use these to formulate the optimal design
-problem as a Lagrange multipliers problem in equation
-\eqref{eq:conditions}.
+We seek a formulation of the optimal design problem
+\eqref{eq:optimization} using Lagrange multipliers. In section
+\ref{section:objective} we find the gradient of $\tar$ and record it
+in \eqref{eq:tar grad}. In section \ref{subsec:unit length} we relax
+the constraints on $\obs$ so that the optimization problem in
+\eqref{eq:optimization} is analytically tractable.  We find gradients
+for the new constraints and record them in \eqref{eq:grad
+  constraints}. Finally, in section \ref{subsec:necessary} we use
+these to formulate the optimal design problem as a Lagrange
+multipliers problem in equation \eqref{eq:conditions}.
 
@@
 
-Thus, if we take $\func_j^{*} = \meas_j$ and $C := (\fwd \prcov
-\fwd^*)^{-1}$ in the above lemma, we conclude from the eigenproblem
-\eqref{eq:mod conditions} that $\obs^*\obs$ has the same eigenvectors
-as $\fwd \prcov \fwd^*$, namely $\{\ev_i\}_{i=1}^{\infty}$ (all but
-finitely many of them with zero eigenvalue, since $\obs^*\obs$ is
-finite-rank). Denote the corresponding non-zero eigenvalues of
-$\obs^*\obs$ by $\{\eta_i\}_{i=1}^{k}$ and let $\eta_i = 0$ for $i
-\geq k+1$. Using Sylvester's determinant theorem we get:
+This allow us to arrive at the following:
+\begin{corollary}[Eigenvectors of a D-optimal design]\label{cor:same ev}
+  $\obs^*\obs$ has the same eigenvectors as $\fwd \prcov \fwd^*$.
+  %namely $\{\ev_i\}_{i=1}^{\infty}$.
+\end{corollary}
+\begin{proof}
+In \eqref{eq:mod conditions}, take $\func_j^{*} = \meas_j$ and $C :=
+(\fwd \prcov \fwd^*)^{-1}$ and use lemma \ref{lemma:sim diag}.
+\end{proof}
+Note that all but finitely many of these eigenvectors have zero
+eigenvalue, since $\obs^*\obs$ is finite-rank. Denote the
+corresponding non-zero eigenvalues of $\obs^*\obs$ by
+$\{\eta_i\}_{i=1}^{k}$ and let $\eta_i = 0$ for $i \geq k+1$. Using
+Sylvester's determinant theorem we get:
 
@@
 
+with the norm constraints on $\meas_j,j=1,\dots,m$. We show in the
+appendix (Lemma \ref{lemma:free}) we can find $\meas_1,\dots,\meas_m$
+such that $\{\eta_i\}_{i=1}^k$ have any value we desire, subject only
+to the restriction that we do not change the trace of $\obs^*\obs$, so
+that $\sum_{i=1}^k \eta_i = m$. By concavity of $\log$, the fastest
+increase in the design criterion (equation \eqref{eq:true target}) is
+gained by increasing $\sigma^2\lambda_i^{-1} + \eta_i$ where it is
+smallest. So, no weight should be given to (i.e. measurement taken in
+direction of) the $k$th eigenvector before $\sigma^2\lambda_i^{-1} +
+\eta_i \geq \sigma^2\lambda_k^{-1},\ \forall i \neq k$. Thus, we
+should choose $\{\eta_i\}_{i=1}^k$ as follows. First, order the
+eigenvalues $\lambda_1 \geq \lambda_2 \geq \dots$. We should increase
+$\sigma^2 \lambda_1^{-1} + \eta_1$ until it equals
+$\sigma^2\lambda_2^{-1}$. Then, $\sigma^2 \lambda_i^{-1} + \eta_i,i=1,2$ are
+increased until $\sigma^2 \lambda_i^{-1} + \eta_i =
+\sigma^2\lambda_3^{-1},i=1,2$ and so forth.
+
+\begin{figure}%{r}{0.25\textwidth} 
+    %\centering
+    \includegraphics[width=5cm, height=5cm]{cylinders.jpg}
+    \caption{Three graduated lab cylinders, corresponding to three
+      eigenvectors. Prior not shown.}
+    \label{fig:cylinder}
+\end{figure}
+
+The argument above has the important consequence (illustrated in
+figure \ref{fig:optimal vs not}) that the eigenvalues in the
+eigenvalue problem \eqref{eq:eigenproblem} are all equal! This makes
+perfect sense. First, we know the Lagrange multipliers tell us how
+much we gain by relaxing the constraints and there is no reason for
+any measurement to give us more than any other (otherwise we would put
+more weight on the better measurement, as can be understood from the
+discussion in the following paragraph).
+
+One can understand optimal designs as follows: Imagine each
+eigenvector is a lab cylinder (figure \ref{fig:cylinder}). Each
+cylinder $i$ is filled, a-priori, with green liquid
+$\sigma^2\lambda_i^{-1}$ volume units (this is the prior precision,
+weighted by obsevation noise). An optimal design with (e.g.) four
+measurements adds blue liquid of four volume units. It does so by
+adding each drop to whatever cylinder has the lowest level of liquid
+in it (lowest precision). The result of such a procedure is
+illustrated in figure \ref{fig:optimal vs not}. The analysis
+conducted above is summarized in theorem \ref{thm:char} below.
+
 \pgfplotstableread{
   Label     prior  optimal  sub-optimal 
-  1         0.2    1.3           1.4
-  2         0.8    0.7           0.3
-  3         2.2    0             0.2
-  4         3.5    0             0.1
+  1         0.2    1.8           1.7
+  2         0.8    1.2           0.8
+  3         2.2    0             0.5
+  4         3.5    0             0.0
 }\optimalvsnot
@@
   \caption{Characterizing an optimal design, compared to a
-    sub-optimal design. Both designs are allowed two measurements,
-    which means the blue area has accumulated height of $2$ in both
+    sub-optimal design. Both designs are allowed three measurements,
+    which means the blue area has accumulated height of $3$ in both
     plots. An optimal design (left) increases precision (reduces
@@
   4         3.5    0      0       0      0      0           0          0.001
-}\testdata
+}\clusterization
+
+\pgfplotstableread{
+  Label    prior  $o_1$  $o_2$   $o_3$   $o_4$   $o_5$         $o_6$    topper
+  1         0.2    0.8    0.7     0.3    0.4    0.2         0.4666666  0.001
+  2         0.8    0.2    0.3     0.7    0.4    0.46666666  0.2        0.001
+  3         2.2    0      0       0      0.2    0.33333333  0.3333333  0.001
+  4         3.5    0      0       0      0      0           0          0.001
+}\noclusterization
+
+\begin{theorem}[Characterizing D-optimal designs]\label{thm:char}
+In a D-optimal design $\eta_i, i=1,\dots,k$ are chosen such that
+$\sigma^2\lambda_i^{-1} + \eta_i < \sigma^2\lambda_k^{-1}$ and
+$\sigma^2\lambda_i^{-1} + \eta_i$ are all equal for $1 \leq i \leq k$
+for some $k$. For $i \geq k+1$, however, $\eta_i = 0$.
+\end{theorem}
+
+%% We are also used to symmetric optimization problems having
+%% symmetric solutions \cite{Waterhouse83}, so symmetric and identical
+%% Lagrange multipliers definitely make sense in this context.
+
+\section{Sensor Clusterization --- Vanishing Model Error}\label{section:clusterization}
+\subsection{Overview}
+In this section we explain how sensor clusterization may arise. We
+first consider a sequential optimal design, where measurements are
+chosen sequentially. We then consider the case where all measurements
+are chosen at once. Finally, we hypothesize how this may apply to a
+real-life inverse problem.
+
+%% illustrated in figure \ref{fig:clusterization} and
+
+\subsection{Sequential Design}
+Recall the sequential optimal design scheme from section
+\ref{subsec:seq vs sim} and the definition of $\obs_k$ from
+\eqref{eq:def obs_k}. The way clusterization can occur is understood
+using results of the previous section, specialized to one measurement
+each time. By corollary \ref{cor:ame ev}, the first measurement
+$\meas_1$ has to be (parallel to) an eigenvector of
+$\fwd\prcov\fwd^*$. By theorem \ref{thm:char}, it will be taken in the direction of the
+first eigenvector --- $\meas_1 \parallel \ev_1$. As hinted in
+section \ref{subsec:seq vs sim}, the posterior becomes the prior for
+the next step. Similar reasoning leads to:
+\begin{align*}
+  \begin{split}
+    \meas_2 \parallel \ev_1  &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 < \sigma^2\lambda^{-1}_2 \\
+    \meas_2 \parallel \ev_2  &\text{ o.w. }
+  \end{split}
+\end{align*}
+In the first case, we get clusterization. In the second, we consider $\meas_3$:
+\begin{align*}
+  \begin{split}
+    \meas_3 \parallel \ev_1  &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 < \sigma^2\lambda^{-1}_3 \text{ (note $\lambda^{-1}_1 < \lambda^{-1}_2$)}  \\
+    %\meas_3 \parallel  \ev_2  &\text{ if } \\
+    \meas_3 \parallel  \ev_3  &\text{ o.w. }
+  \end{split}
+\end{align*}
+Again, the first case results in clusterization. The sequential design
+proceeds in this fashion, where every $\meas_i$ is taken in the
+direction of the eigenvector of smallest eigenvalue of the previous
+posterior precision operator. Taking enough measurements we will
+surely arrive at clusterization, since the prior precision increases
+quickly --- in section \ref{subsec:abstract OED} we assumed  $\fwd$
+is strongly smoothing, which typpically means eigenvalues of $(\fwd
+\prcov\fwd^*)^{-1}$ grow exponentially.
+
+\subsection{Simultaneous Design}
+In a simultaneous design scenario, the result is not as clear cut as
+in the sequential design case. We still rely on previous results ---
+namely corollary \ref{cor:same ev} and theorem \ref{thm:char}. An
+illustration of two D-optimal designs scenarios for the same problem
+are presented in figure \ref{fig:clusterization} --- the first
+exhibits clusterization and second does not. As was discussed earlier
+and illustrated in figure \ref{fig:optimal vs not}, an optimal design
+makes the precision ``flat'' where it is lowest. Clusterization may
+arise when two or more eigenvectors have equal precision. This
+happens, for example for eigenvectors $1$ and $2$ when considering
+measurements $\meas_i,i=1,2,3$. Taking three measurements, it is
+possible that $\meas_1$ is such that it equalizes the precision of
+eigenvectors $1$ and $2$. In this case, taking mesurements $\meas_2 =
+\meas_3 = \frac{1}{2}(\ev_1 + \ev_2)$ leads to an optimal design that
+exhibits clusterization. Similarly, taking $\meas_5 = \meas_6 =
+\frac{1}{3}(\ev_1 + \ev_2 + \ev_3)$ also gives rise to an optimal
+design that exhibits clusterization. It is important to note that
+there are (infinitely) many optimal designs that actually do not
+exhibit clusterization --- e.g. the second design presented in figure
+\ref{fig:clusterization}. In the following section we postulate what
+causes such designs to be observed in realistic experimental design
+scenarios.
 
 \begin{figure}
-  \begin{tikzpicture}
+  \begin{tikzpicture}[scale=0.85]
     \begin{axis}[
@@
         reverse legend=false, % set to false to get correct display, but I'd like to have this true
-        xticklabels from table={\testdata}{Label},
+        xticklabels from table={\clusterization}{Label},
         xticklabel style={text width=2cm,align=center},
@@
 
-      \addplot [fill=green!80]  table [y=prior, meta=Label, x expr=\coordindex] {\testdata};
-      \addplot [fill=blue!60]   table [y=$o_1$, meta=Label, x expr=\coordindex] {\testdata};
-      \addplot [fill=red!60]    table [y=$o_2$, meta=Label, x expr=\coordindex] {\testdata};
-      \addplot [fill=black!60]  table [y=$o_3$, meta=Label, x expr=\coordindex] {\testdata};
-      \addplot [fill=orange!60] table [y=$o_4$, meta=Label, x expr=\coordindex] {\testdata};
-      \addplot [fill=cyan!60]   table [y=$o_5$, meta=Label, x expr=\coordindex] {\testdata};
-      \addplot [fill=purple!60] table [y=$o_6$, meta=Label, x expr=\coordindex] {\testdata};
+      \addplot [fill=green!80]  table [y=prior, meta=Label, x expr=\coordindex] {\clusterization};
+      \addplot [fill=blue!60]   table [y=$o_1$, meta=Label, x expr=\coordindex] {\clusterization};
+      \addplot [fill=red!60]    table [y=$o_2$, meta=Label, x expr=\coordindex] {\clusterization};
+      \addplot [fill=black!60]  table [y=$o_3$, meta=Label, x expr=\coordindex] {\clusterization};
+      \addplot [fill=orange!60] table [y=$o_4$, meta=Label, x expr=\coordindex] {\clusterization};
+      \addplot [fill=cyan!60]   table [y=$o_5$, meta=Label, x expr=\coordindex] {\clusterization};
+      \addplot [fill=purple!60] table [y=$o_6$, meta=Label, x expr=\coordindex] {\clusterization};
 
@@
   \end{tikzpicture}
-  \caption{Posterior precision per eigenvector, after taking a
+  \qquad
+  \begin{tikzpicture}[scale=0.85]
+    \begin{axis}[
+        ybar stacked,
+        ymin=0,
+        ymax=4,
+        xtick=data,
+        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
+        reverse legend=false, % set to false to get correct display, but I'd like to have this true
+        xticklabels from table={\noclusterization}{Label},
+        xticklabel style={text width=2cm,align=center},
+        legend plot pos=right,
+        ylabel=precision --- prior and posterior,
+        xlabel=eigenvector,
+      ]
+
+      \addplot [fill=green!80]  table [y=prior, meta=Label, x expr=\coordindex] {\noclusterization};
+      \addplot [fill=blue!60]   table [y=$o_1$, meta=Label, x expr=\coordindex] {\noclusterization};
+      \addplot [fill=red!60]    table [y=$o_2$, meta=Label, x expr=\coordindex] {\noclusterization};
+      \addplot [fill=black!60]  table [y=$o_3$, meta=Label, x expr=\coordindex] {\noclusterization};
+      \addplot [fill=orange!60] table [y=$o_4$, meta=Label, x expr=\coordindex] {\noclusterization};
+      \addplot [fill=cyan!60]   table [y=$o_5$, meta=Label, x expr=\coordindex] {\noclusterization};
+      \addplot [fill=purple!60] table [y=$o_6$, meta=Label, x expr=\coordindex] {\noclusterization};
+
+      \addlegendentry{prior}
+      \addlegendentry{$o_1$}
+      \addlegendentry{$o_2$}
+      \addlegendentry{$o_3$}
+      \addlegendentry{$o_4$}
+      \addlegendentry{$o_5$}
+      \addlegendentry{$o_6$}   
+    \end{axis}
+  \end{tikzpicture}
+  \caption{Clusterization and non-clusterization in simultaneous
+    design. Posterior precision per eigenvector, after taking a
     measurement $\meas_i$. Each $\meas_i$ increases precision
@@
 
-with the norm constraints on $\meas_j,j=1,\dots,m$. We show in the
-appendix (Lemma \ref{lemma:free}) we can find $\meas_1,\dots,\meas_m$
-such that $\{\eta_i\}_{i=1}^k$ have any value we desire, subject only
-to the restriction that we do not change the trace of $\obs^*\obs$, so
-that $\sum_{i=1}^k \eta_i = m$. By concavity of $\log$, the fastest
-increase in the design criterion (equation \eqref{eq:true target}) is
-gained by increasing $\sigma^2\lambda_i^{-1} + \eta_i$ where it is
-smallest. So, no weight should be given to (i.e. measurement taken in
-direction of) the $k$th eigenvector before $\sigma^2\lambda_i^{-1} +
-\eta_i \geq \sigma^2\lambda_k^{-1},\ \forall i \neq k$. Thus, we
-should choose $\{\eta_i\}_{i=1}^k$ as follows. First, order the
-eigenvalues $\lambda_1 \geq \lambda_2 \geq \dots$. We should increase
-$\sigma^2 \lambda_1^{-1} + \eta_1$ until it equals
-$\sigma^2\lambda_2^{-1}$. Then, $\sigma^2 \lambda_i^{-1} + \eta_i,i=1,2$ are
-increased until $\sigma^2 \lambda_i^{-1} + \eta_i =
-\sigma^2\lambda_3^{-1},i=1,2$ and so forth.
-
-The argument above has the important consequence (illustrated in
-figure \ref{fig:optimal vs not}) that the eigenvalues in the
-eigenvalue problem \eqref{eq:eigenproblem} are all equal! This makes
-perfect sense. First, we know the Lagrange multipliers tell us how
-much we gain by relaxing the constraints and there is no reason for
-any measurement to give us more than any other (otherwise we would put
-more weight on the better measurement, as can be understood from the
-discussion in the following paragraph). This result of the analysis
-above is summarized in the theorem below.
-
-\begin{theorem}[Characterizing D-optimal designs]\label{thm:char}
-In a D-optimal design $\eta_i, i=1,\dots,k$ are chosen such that
-$\sigma^2\lambda_i^{-1} + \eta_i < \sigma^2\lambda_k^{-1}$ and
-$\sigma^2\lambda_i^{-1} + \eta_i$ are all equal for $1 \leq i \leq k$
-for some $k$. For $i \geq k+1$, however, $\eta_i = 0$.
-\end{theorem}
-
-%% We are also used to symmetric optimization problems having
-%% symmetric solutions \cite{Waterhouse83}, so symmetric and identical
-%% Lagrange multipliers definitely make sense in this context.
-
-\section{Sensor Clusterization --- Vanishing Model Error}\label{section:clusterization}
-\subsection{Overview}
-In this section we explain how sensor clusterization may arise. We
-first consider a sequential optimal design, where measurements are
-chosen sequentially. We then consider the case where all measurements
-are chosen at once. Finally, we hypothesize how this may apply to a
-real-life inverse problem.
-
-\subsection{Sequential Design}
-The way clusterization can occur is illustrated in figure
-\ref{fig:clusterization} and understood using theorem
-\ref{thm:char}. The first measurement $\meas_1$ is optimal, since
-$\eta_i, i=1,2$ are chosen to increase and equalize
-$\sigma^2\lambda_i^{-1} +\eta_i, i=1,2$, which are lowest. In this
-case, $\meas_1$ has no component in the direction of $\ev_i, i\geq
-3$. We see that each measurement increases precision (reduces
-uncertainty) only for the eigenvectors of lowest precision. Consider a
-sequential optimization scenario. We first seek a measurement to
-maximize our our utility, given the prior. From , the posterior
-covariance does not depend on data observed, so we know it after
-deciding on the first measurement.
-
-the best
-first measuremen, thIf the precision in these is lower than the
-precision for the next eigenvector, a repeated measurement is optimal.
-
-Results of section \ref{subsec:characterization} specifically, the
-discussion at the end of said section, along with figure
-\ref{figure:clusterization} and equation \eqref{eq:true target} show
-how clusterization arises when an optimal design is saught.
+\subsection{Why do Realistic Designs Exhibit Clusterization?}
+It is unclear from the previous analysis what makes realistic designs
+exhibit clusterization, such as the one observed in figure
+\ref{fig:clusterization illustration}. We can only hypothesize.
 
@@
 \end{restatable*}
-From Corollary \ref{cor:same meas} we can deduce that
+Recall from \eqref{eq:Sigma} that $\Sigma(\obs) = \obs \modcov \obs^*
++ \sigma^2I$. From Corollary \ref{cor:same meas} we can deduce that
 
@@
 $$
-Hence no increase in the design criterion is
-achieved. This is clearly sub-optimal and $\obsm$ cannot be an optimal
-design. We may conclude that for small measurement error levels, the
+
+Hence no increase in the design criterion is achieved by taking a
+repeated measurement. This is clearly sub-optimal and $\obsm$ cannot
+be an optimal design. Thus, for small measurement error levels, the
 clusterization effect will be mitigated by the presence of a non-zero
@@
 
-\section{Conclusion}\label{section:conclusion}
-Our relaxed model gives us insight to D-optimal designs. We see that
-in our setting, uncertainty is not reduced in any direction
-(eigenvector) before it is the largest uncertainty present. We show
-how and why repeated measurements can give rise to a D-optimal
-design. There is more work to be done in understanding exactly what
-causes such designs to be strictly better than others.
+%% \section{Conclusion and Future Research Directions}\label{section:conclusion}
+%% Our relaxed model gives us insight to D-optimal designs. We see that
+%% in our setting, uncertainty is not reduced in any direction
+%% (eigenvector) before it is the largest uncertainty present. We show
+%% how and why repeated measurements can give rise to a D-optimal
+%% design. There is more work to be done in understanding exactly what
+%% causes such designs to be strictly better than others.
 
@@
   Let $M \in \R^{k \times k}$ symmetric positive definite with $\ttr M =
-  m$, $m > k$. Then we can find $\func_j \in \R^k,j=1,\dots,m$ with
+  m$, $m \geq k$. Then we can find $\func_j \in \R^k,j=1,\dots,m$ with
   $\|\func_j\|=1$ and $A = (\func_1,\dots,\func_m)$ such that $AA^t =
@@
   Let us diagonalize $M$, so that $M = U D U^t$ with $D =
-  \diag(d_1,\dots,d_k)$ and $U \in \R^{k \times k }$ orthogonal. Let $S
-  \in \R^{k \times m}$ with $S_{ii} = \sqrt{d_{i}}$ and zeros
+  \diag(d_1,\dots,d_k)$ and $U \in \R^{k \times k }$ orthogonal. Let
+  $S \in \R^{k \times m}$ with $S_{ii} = \sqrt{d_{i}}$ and zeros
   otherwise. Define $A:= U S V^t$, where $V \in \R^{m \times m}$ is
@@
   \end{equation}
-  and we can expect to do this since we assumed $\tr{D} = m$.
-
-  Define $C = S^tS - I \in \R^{m \times m}$. Note that $\tr{C} = 0$ and
+  and we can expect to do this since we assumed $\ttr D = m$.
+
+  Define $C = S^tS - I \in \R^{m \times m}$. Note that $\ttr C = 0$ and
   $C$ is diagonal with non-zero entries $d_i-1,i=1,\dots,k$. It suffices
@@
   and not destroying zeros we already introduced, starting from the
-  last diagonal entry and moving to the first. If $c_{kk} \neq 0$ ,
-  let $p < k$ such that $c_{pp}c_{kk} < 0$ (such $p$ exists because
+  last diagonal entry and moving to the first. If $c_{mm} \neq 0$ ,
+  let $p < m$ such that $c_{pp}c_{mm} < 0$ (such $p$ exists because
   the trace is zero) and let $\theta \in (0,\pi)$. Define a Givens
-  rotation $R^{(k)} \in \R^{m \times m}$ by
+  rotation $R^{(m)} \in \R^{m \times m}$ by
   \begin{equation*}
-    r^{(k)}_{ab} :=
+    r^{(m)}_{ab} :=
     \begin{cases}
-      1 & a = b \neq p \text{ or } a = b \neq k \\
+      1 & a = b \neq p \text{ or } a = b \neq m \\
       \cos \theta & a = b = p  \\
-     -\sin \theta & a = p, b = k\\
-      \cos \theta & a = b = k \\
-      \sin \theta & a = k, b = p \\ 
+     -\sin \theta & a = p, b = m\\
+      \cos \theta & a = b = m \\
+      \sin \theta & a = m, b = p \\ 
       0 & o.w
@@
   \end{equation*}
-  Note that conjugating a matrix by $R^{(k)}$ changes only its $k$ and
-  $p$ rows and columns. Specifically, it leaves rows and columns
-  $k+1,\dots,m$ unchanged. We want to choose $\theta$ such that
+  Note that conjugating a matrix by $R^{(m)}$ changes only its $m$ and
+  $p$ rows and columns. We want to choose $\theta$ such that
   \begin{equation}\label{eq:mm}
-    0 = [R^{(k)} C (R^{(k)})^t]_{kk} = \cos^2 \theta c_{kk} + 2\cos \theta \sin
-    \theta c_{kp} + \sin^2\theta c_{pp},
+    0 = [R^{(m)} C (R^{(m)})^t]_{mm} = \cos^2 \theta c_{mm} + 2\cos \theta \sin
+    \theta c_{mp} + \sin^2\theta c_{pp},
   \end{equation}
@@
   \begin{equation*}
-    c_{kk} \cot^2 \theta + 2 c_{kp} \cot \theta + c_{pp} = 0.
+    c_{mm} \cot^2 \theta + 2 c_{mp} \cot \theta + c_{pp} = 0.
   \end{equation*}
   This quadratic in $\cot\theta$ has a real solution, since
-  $c_{pp}c_{kk} < 0$ by assumption and we can find $\theta \in
+  $c_{pp}c_{mm} < 0$ by assumption and we can find $\theta \in
   (0,\pi)$ such that \eqref{eq:mm} is satisfied. We continue to find
-  $R^{(k-1)}$ that leaves rows and columns $k,\dots,m$ unchanged and
-  continue introducing zeros to the diagonal. The assumption $\tr{D} =
-  m \Rightarrow \tr{C} = 0$ guarantees we can do that. Taking $V:=
-  R^{(1)} R^{(2)} \dots R^{(k-1)}R^{(k)}$ completes the proof.
+  $R^{(m-1)}$ that leaves row and column $m$ unchanged and
+  continue introducing zeros to the diagonal. The assumption $\ttr D =
+  m \Rightarrow \ttr C = 0$ guarantees we can do that. Taking $V:=
+  R^{(1)} R^{(2)} \dots R^{(m-1)}R^{(m)}$ completes the proof.
 \end{proof}

### 2020-10-03 zeros.py

@@ -1,9 +1,22 @@
+## This here code implements (and tests) the algorithm outlined in
+## Lemma B.5 in the paper.
+##
+## Given M symmetric positive definite in R^{k X k} such that
+##
+## tr M = m > k, (m an integer),
+## 
+## it finds A such that:
+##
+## A's columns have unit norm, and
+## AA^t = M. 
+
 import numpy as np
+from pandas import isnull
 from numpy import sin, cos
 from scipy.stats import ortho_group
 import pdb
 
-EPS = 1e-13
+EPS = 1e-10
 ortho = ortho_group.rvs
 cot = lambda x: cos(x)/ sin(x)
 
@@ -14,7 +27,7 @@ def conj(D, U):
     return np.einsum('ij, jk, lk -> il', U, D, U)
 
 
-def givens(theta, dims, lower, upper):
+def givens(theta, dims, lower, upper):    
     assert theta >= 0 and theta <= np.pi
     assert lower < upper
     assert upper <= dims
@@ -36,19 +49,25 @@ def givens(theta, dims, lower, upper):
 
 
 def MDUS(m, k):
-    assert m > k
+    assert m >= k
     
-    D = np.abs(np.random.uniform(size=k))
+    # D = np.abs(np.random.uniform(size=k))
+    D = np.random.lognormal(mean=5, sigma=1, size=k)
+    # D = np.array([np.random.lognormal(mean=i, sigma=1) for i in range(k)]) 
     D = D * m / np.sum(D)
     assert abs(np.sum(D) - m) < EPS 
+    assert np.all(D > 0)
     D = np.diag(D)
     assert D.shape == (k,k)
+    assert np.all(~isnull(D))
+    
 
     U = ortho(k)
     assert np.all(np.abs(np.dot(U,U.T) - np.identity(k)) < EPS)
 
     M = conj(D, U)
-
+    M = (M + M.T)/2 ## Ensure it is symmetric
+    
     S = np.zeros((k,m))
     for i in range(k):
         S[i,i] = np.sqrt(D[i,i])
@@ -62,8 +81,8 @@ def get_theta(C, upper):
     assert upper < C.shape[0]
     assert C.shape[0] == C.shape[1]
     if upper == 0:
-        raise ValueError(f'Why upper == 0 and C.shape == {C.shape}?')
-    if C[upper,upper] == 0:
+        raise ValueError(f'Why upper == 0? C.shape == {C.shape}.')
+    if abs(C[upper,upper]) < EPS:
         return None, None
 
     success = False
[…]

### 2020-10-06 new_paper.tex
Restated the theorem, looks much much bettergit status !

@@
 uncertainties, and show how that can give rise to repeated
-measurements (section \ref{subsec:clusterization} and figure
+measurements (section \ref{section:clusterization} and figure
 \ref{fig:clusterization}).
@@
 %% paragraph.
-The arguments presented are completely generic and apply equally to
-any of the scenarios mentioned in the introduction.
-
-%% \subsection{Limitations}\label{subsec:limitations}
-%% There are two main drawbacks of the arguments presented here. First,
-%% one can argue that we allow considerably less restrictive observations
-%% than any real-life problem does. This is because our observation model
-%% allows taking measurements that are arbitrarily weighted linear
-%% combinations of eigenvectors. The second drawback is that we do not
-%% show rigorously that clusterization necessarily occurs in our model,
-%% we only show that it can occur.
+
+\subsection{Limitations}\label{subsec:limitations}
+There are two main drawbacks of the arguments presented here. First,
+one can argue that we allow considerably less restrictive observations
+than any real-life problem does. This is because our observation model
+allows taking measurements that are arbitrarily weighted linear
+combinations of eigenvectors. The second drawback is that we do not
+show rigorously that clusterization necessarily occurs in our model,
+we only show that there is no reason to be surprised when it does
+occur.
 
@@
 
-\subsection{Remarks to the Reader}\label{subsec:plan}
-The Lagrange multipliers problem for iid measurement noise is
-discussed in section \ref{section:vanishing}. In that section, it is
-shown how sensor clusterization can occur. The reader may skip if they
-are constrained by time. It is advised that a reader less comfortable
-with Hilbert and Banach spaces mostly ignore these terms and consider
-$\hil$ and $\ban$ finite dimensional vector spaces.
-
-%% Seeking an optimal design amounts to maximizing $\tar(\obs)$ in
-%% $\obs$, subject to constraints posed on $\obs$. These enter as
-%% constraints on the allowed set of measurement vectors we can choose
-%% from --- the rows of $\obs$. These constraints are determined by the
-%% problem structure --- for example, in the 1D heat equation example
-%% considered in section \ref{subsec:example}, we were constrained to
-%% choose $\obs = (\meas_1,\dots,\meas_m)^t$ with $\meas_j = \delta_{\x_j}$
-%% for some $\x_j \in [0,\pi]$. We present an approach to understanding
-%% optimal experiment design based on the first-order necessary condition
-%% for optimal design, where the constraints are incorporated using
-%% Lagrange multipliers. This condition is used to understand what causes
-%% clusterization. One problem with such an approach is that the set we
-%% can choose $\obs$ from is complicated and difficult to work with. For
-%% example, if $\ban = C(\Omega)$ for some compact $\Omega$, then the
-%% allowed set of measurement vectors may only be point evaluations --- a
-%% set that is not easy to describe as a level set of some simple
-%% function, a description crucial for the use of Lagrange multipliers.
-%% In order to facilitate analysis, we consider an easier constraint in
-%% section \ref{subsec:unit length}. Reverting to this simpler constraint
-%% is a limitation of our analysis but it provides some insight into the
-%% cause of clusterization.
-
 \section{Preliminaries and Notation}\label{section:prelim}
@@
 \subsection{Overview}
-The theoretical foundations for inverse problems over function spaces
-can be found in \cite{Stuart10}. In this section we define notations that will be
-  used throughout the paper.
+In this section we define notations that will be used throughout the
+paper. The theoretical foundations for inverse problems over function spaces
+can be found in \cite{Stuart10}.
 %% A reader mostly familiar with the
@@
 %% \begin{itemize}
-%%   \item Parameter space $\hil$.
-%%   \item Data space $\ban$.
-%%   \item Forward dynamics operator $\fwd: \hil \to \ban$.
-%%   \item Observation operator $\obs: \ban \to \mathbb{R}^m$.
+%%   \item Parameter space $\hilp$.
+%%   \item Data space $\hilo$.
+%%   \item Forward dynamics operator $\fwd: \hilp \to \hilo$.
+%%   \item Observation operator $\obs: \hilo \to \mathbb{R}^m$.
 %%   \item Prior measure $\pr = \mathcal{N}(0, \prcov)$.
@@
 \subsection{Bayesian Linear Inverse Problems}\label{subsec:abstract OED}
-Let $\hil$ be a Hilbert space and $\ban$ a Banach space, both
-infinite-dimensional. Consider $\fwd: \hil \to \ban$, a linear
-operator (the ``forward operator''). We are interested in forward
-operators that are strongly smoothing (have fast decaying singular
-values --- the heat operator from section \ref{subsec:example} is a
-prime example). We want to infer $\param \in \hil$, some parameter of
-the dynamics given noisy observations of $\fwd \param$. We take a zero
-mean Gaussian prior $\pr$ for $\param$ (otherwise the mean can just be
+Let $\hilp$ and $\hilo$ be seperable Hilbert spaces, both
+infinite-dimensional (the subscripts p and o are for ``parameter'' and
+``observation'', respectfully). We denote, $\hilo^*$ the Hilbert space
+of all linear functionals on $\hilo$. From standard functional
+analysis, it is known that $\hilp$ and $\hilo$ are isomorphic (and in
+practice may be equal) and that $\hilo = \hilo^*$. We will still use
+such notation, to keep track of where a certain object is
+defined. Consider $\fwd: \hilp \to \hilo$, a linear operator (the
+``forward operator''). We are interested in forward operators that are
+strongly smoothing (have fast decaying singular values --- the heat
+operator from section \ref{subsec:example} is a prime example). We
+want to infer $\param \in \hilp$, some parameter of the dynamics given
+noisy observations of $\fwd \param$. We take a zero mean Gaussian
+prior $\pr$ for $\param$ (otherwise the non-zero mean can just be
 subtracted everywhere). Hence, $\param \sim \pr = \normal(0,\prcov)$
-with some appropriate covariance operator $\prcov$
-\cite{Stuart10}. Observations (measurements) are taken via the linear
-observation operator $\obs \in ( \ban^* )^m$, where $m$ is the number
-of observations allowed (e.g. sensors at our disposal). In an analogy
-with linear algebra (where row vectors are thought of as linear
-functionals), we think of $\obs$ as having ``rows'' $\meas_j,
-j=1,\dots,m$:
+with some appropriate covariance operator $\prcov$ \cite{Stuart10}. It
+is important to note that $\fwd \prcov \fwd^*$ is the prior covariance
+in $\hilo$ \cite{Stuart10}. As such, it is assumed invertible --- an
+assumption we rely heavily upon. Observations (measurements) are taken
+via the linear observation operator $\obs \in ( \hilo^* )^m$, where
+$m$ is the number of observations allowed (e.g. sensors at our
+disposal). In an analogy with linear algebra (where row vectors are
+thought of as linear functionals), we think of $\obs$ as having
+``rows'' $\meas_j, j=1,\dots,m$:
 \begin{equation}\label{eq:O}
-  \obs = (\meas_1,\dots, \meas_m)^t, \meas_j \in \ban^*, j = 1,\dots,m.
+  \obs = (\meas_1,\dots, \meas_m)^t, \meas_j \in \hilo^*, j = 1,\dots,m.
 \end{equation}
-This way, for $u \in \ban$ we have $\obs u = (\meas_1(u), \dots,
+This way, for $u \in \hilo$ we have $\obs u = (\meas_1(u), \dots,
 \meas_m(u) )^t \in \R^m$.
@@
 A few observations regarding $\obs$ are in order. First, it is good to
-keep in mind that $( \ban^* )^m$ is a Banach space with norm $\| \obs
-\| = \sum_{j=1}^m \|\meas_j\|$. Next, for $u \in \ban$ and $v\in
+keep in mind that $( \hilo^* )^m$ is a Hilbert space with norm $\| \obs
+\| = \sum_{j=1}^m \|\meas_j\|$. Next, for $u \in \hilo$ and $v\in
 \R^m$:
@@
 Each measurement consists of a linear functional, chosen from an
-allowed set of linear functionals in $\ban^{*}$. Data is acquired via
+allowed set of linear functionals in $\hilo^{*}$. Data is acquired via
 noisy measurements
@@
 \end{align*}
-with $\data \in \R^m$ and $\eps, \eps'$ defined next. Measurement
-error consists of two parts. First, there is spatially correlated
-error $\eps' \sim \normal(0,\modcov)$, modeled as a centered Gaussian
-measure on $\ban$ with covariance operator $\modcov: \ban^* \to \ban$
-(see \cite[section 6]{Stuart10} for some details on Gaussian measures
-on Banach spaces). Then there is measurement / independent error
-$\eps \sim \normal(0, \sigma^2 I_m)$, with $I_m$ the $m \times m$
-identity. Both error terms and the prior are assumed independent of
-each other. Let us now understand the terms present in the above
-expression.
+with $\data \in \R^m$ and $\eps, \eps'$ defined next. We consider two
+types of noise / error terms. First, there is spatially correlated
+model error $\eps' \sim \normal(0,\modcov)$, modeled as a centered
+Gaussian measure on $\hilo$ with covariance operator $\modcov$. Then
+there is measurement / independent error $\eps \sim \normal(0,
+\sigma^2 I_m)$, with $I_m$ the $m \times m$ identity. Both error terms
+and the prior are assumed independent of each other. Let us now
+understand the terms present in the above expression.
+
+At this point, it is useful to record a known result regarding the
+posterior covariance operator of our inverse problem, namely
+\begin{align}\label{eq:postcov}
+  \postcov = (\prcov^{-1} + \fwd^* \obs^* \Sigma^{-1} \obs \fwd
+  )^{-1}.
+\end{align}
+Usually, this is written using $\tmp := \obs \fwd$ and $\tmp^*$ but
+since we need to separate the observation operator $\obs$ from the
+forward operator $\fwd$ we write the explicit expression
+\eqref{eq:postcov}.
 
 \subsubsection{Dynamics and Observation Operators}\label{subsec:dynamics}
-One can think of $(\obs \fwd \param)_j$ as pointwise evaluations of a
-continuous function if $\ban$ is some appropriate function space,
-e.g.\ $C(\Omega)$, the Banach space of real-valued continuous
-functions over a compact $\Omega$ with the $\| \cdot \|_{\infty}$
-norm. These measurements can be represented using the linear
-functionals $\delta_{\x_j} \in C(\Omega)^*$. In a different setting,
-we may be able to measure other quantities, e.g.\ some Fourier
-coefficients as is the case in MRI applications. Either way, the rows
-of $\obs$ are linear functionals and we refer to these as {\it
-  measurement vector}s. Note that we cannot choose every measurement
-vector we want, though. For example, we may be restricted only to
-pointwise evaluations of $\fwd \param$ by the sensors at our disposal,
-in which case measuring the mean $\ell(u) = \int_{\Omega}u$ (or any
-non-atomic measure) of a function is not possible.
+One can \textit{think} of $(\obs \fwd \param)_j$ as pointwise
+evaluations of a continuous function\footnote{Of course, this is
+  impossible if, e.g., $\hilo = L^2(\Omega)$ for some $\Omega$. Still,
+  this approach is in line with previous analyses
+  \cite{Karina,AlexanderianGloorGhattas14} that consider sensor
+  placement but formulate the inverse problem on Hilbert spaces.}. Such
+measurements can be represented using the linear functionals
+$\delta_{\x_j} \in C(\Omega)^*$. In a different setting, we may be
+able to measure other quantities, e.g.\ some Fourier coefficients as
+is the case in MRI applications. Either way, the rows of $\obs$ are
+linear functionals and we refer to these as {\it measurement
+  vector}s. %% Note that we cannot choose every measurement vector we
+%% want, though. For example, we may be restricted only to pointwise
+%% evaluations of $\fwd \param$ by the sensors at our disposal, in which
+%% case measuring the mean $\ell(u) = \int_{\Omega}u$ (or any non-atomic
+%% measure) of a function is not possible.
 
@@
 of measurements) we can equivalently write $\data = \obs \fwd \param +
-\bar{\eps}$ with $\bar{\eps} \sim \normal(0,\Sigma)$. At this point,
-it is useful to record a known result regarding the posterior
-covariance operator of our inverse problem, namely
-\begin{align}\label{eq:postcov}
-  \postcov = (\prcov^{-1} + \fwd^* \obs^* \Sigma^{-1} \obs \fwd
-  )^{-1}.
-\end{align}
-Usually, this is written using $\tmp := \obs \fwd$ and $\tmp^*$ but
-since we need to separate the observation operator $\obs$ from the
-forward operator $\fwd$ we write the explicit expression
-\eqref{eq:postcov}.
-
-Taking $\modcov = 0$ is common practice
-\cite{Tarantola05,KaipioSomersalo05,Vogel02} and then we are reduced
-to the case where we take iid observations and $\Sigma = \sigma^2I$ is
-simply a scalar matrix that does not depend on $\obs$. Note that
-taking an error model with a non-scalar covariance as we do here
-allows us to consider model error (modeled by $\modcov$) as well as
-measurement error (modeled by $\sigma^2$). For example, say we believe
-our forward model does not capture some small scale phenomenon.  Then
-we may express this belief by saying $\tru = \fwd + \err$, with $\fwd$
+\bar{\eps}$ with $\bar{\eps} \sim \normal(0,\Sigma)$. Taking $\modcov
+= 0$ is common practice
+\cite{Tarantola05,KaipioSomersalo05,Vogel02}. This reduces to the case
+where we take iid observations. Then $\Sigma = \sigma^2I$ is simply a
+scalar matrix that does not depend on $\obs$. Note that taking an
+error model with a non-scalar covariance as we do here allows us to
+consider model error (modeled by $\modcov$) as well as measurement
+error (modeled by $\sigma^2$). For example, say we believe our forward
+model does not capture some small scale phenomenon.  Then we may
+express this belief by saying $\tru = \fwd + \err$, with $\fwd$
 depending on $\param$ and $\err$ depending on $\sspar$, and $\param
@@
 a numerical source, where our discretization of the system is not fine
-enough to capture all small scale phenomena. We will see that assuming
-some correlation in the error may reduce the clusterization
-phenomenon, as is also reported in the literature \cite{Ucinski05}.
+enough to capture all small scale phenomena. In section
+\ref{section:non vanishing}, we will see that assuming some
+correlation in the error mitigates the clusterization phenomenon, as
+is also reported in the literature \cite{Ucinski05}.
 
@@
 The meaning of D-optimal design in infinite-dimensional Hilbert spaces
-was investigated in \cite{AlexanderianGloorGhattas14}. Note that the
+was investigated in \cite{AlexanderianGloorGhattas14}. There, the
 authors make assumptions that amount to $\Sigma=I$ (implied by
@@
 infinite-dimensions, a D-optimal design is well-defined as maximizing
-the expected KL divergence between posterior and prior and is summarized, using our notation,
-in the following theorem:
+the expected KL divergence between posterior and prior and is
+summarized, using our notation, in the following theorem:
 \begin{theorem}[Slightly modified Theorem 1 from \cite{AlexanderianGloorGhattas14}]
   Let $\pr = \normal(0,\prcov)$ be a centered Gaussian prior
-  on $\hil$   and let $\post = \normal(\postmean,\postcov)$ 
-  the posterior measure on $\ban$ for the Bayesian linear
+  on $\hilp$   and let $\post = \normal(\postmean,\postcov)$ 
+  the posterior measure on $\hilo$ for the Bayesian linear
   inverse problem $\data =  \obs \fwd\param + \obs \eps' + \eps$ discussed
@@
 \end{theorem}
-A D-optimal design seeks to maximize $\tar(\obs)$.
+\begin{definition}\label{def:d optimality}
+  A design $\obs^{\star}$ is said to be D-optimal if $\obs^{\star} = \argmax_{\obs} \tar(\obs)$.
+\end{definition}
 
 \subsection{Sequential vs Simultaneous Optimization}\label{subsec:seq vs sim}
-From \eqref{eq:objective} we wish to characterize solution(s) of the
-following optimization problem for $\tar$. %%: (\ban^*)^m \to \R$:
+From \eqref{def:d optimality} we wish to characterize solution(s) of the
+following optimization problem for $\tar$. %%: (\hilo^*)^m \to \R$:
 \begin{align}\label{eq:optimization}
@@
 \end{align}
-where $\obs$ is constrained to some allowed set of measurements
-(e.g.\ pointwise evaluations on $[0,\pi]$). We call this simultaneous
-optimization.
+where $\obs$ is constrained to some allowed set of measurements (we
+may \textit{think} of these as pointwise evaluations on $[0,\pi]$,
+even though these are not in $\hilo$). We call this problem
+``simultaneous optimization'', since all measurements are decided on
+simulatneously.
 
@@
 \begin{equation}\label{eq:def obs_k}
-  \obs_k := \sum_{i=1}^k \meas_i\meas^*_i.
+  \obs_k := (\meas_1,\dots, \meas_k)^t,  k\leq m.
 \end{equation}
@@
 
-This paper is mostly focused with the more general simultaneous
+The analysis in this paper is conducted for the general simultaneous
 optimization case. The sequential optimization case is dealt with in
 section \ref{subsec:seq clus}. It is important to note, however that
-all conclusions we make on the simultaneous case easily specialize to
-the sequential case by considering the posterior as the next
-sequential step's prior.
+all conclusions we arrive at for the simultaneous case easily
+specialize to the sequential case by considering the posterior as the
+next sequential step's prior.
 
@@
 \subsection{Overview}
-%%, for the example in section \ref{subsec:example}).
 We seek a formulation of the optimal design problem
@@
 \end{align*}
-with $V_j \in \ban^*$ and $\nabla \tar(\obs)_j \in \ban^{**}\subseteq
-\ban, j=1,\dots,m$. Thus, $\nabla \tar( \obs ) \in (\ban^{**})^m
-\subseteq \ban^m$ is indeed the correct gradient and the notation
-\eqref{eq:tar grad} is justified. Note that by the comment following
-\eqref{eq:obs*}, we view $V$ (defined on the same space as $\obs$) as
-a column vector. Then the gradient $\nabla \tar(\obs)$ should be
-viewed as a row vector, as the product $V \nabla \tar$ is in $\R^{m
-  \times m}$. This will prove important in section
-\ref{subsec:necessary}.
+with $V_j \in \hilo^*$ and $\nabla \tar(\obs)_j \in \hilo^{**} =
+\hilo, j=1,\dots,m$. Thus, $\nabla \tar( \obs ) \in \hilo^m$ is indeed
+the correct gradient and the notation \eqref{eq:tar grad} is
+justified. Note that by the comment following \eqref{eq:obs*}, we view
+$V$ (defined on the same space as $\obs$) as a column vector. Then the
+gradient $\nabla \tar(\obs)$ should be viewed as a row vector, as the
+product $V \nabla \tar$ is in $\R^{m \times m}$. This will prove
+important in section \ref{subsec:necessary}.
 
@@
 As mentioned before, we cannot choose any $\obs$ when maximizing
-$\tar(\obs)$. Recall $\obs = (\meas_1,\dots,\meas_m)^t \in (\ban^*)^m$
-and each $\meas_j,j=1,\dots,m$ must be chosen from some allowed set of
-functionals in $\ban^*$. This set differs based on the kind of sensors
-we have at our disposal and the properties of $\ban$, as discussed in
-section \ref{subsec:dynamics}. The following proposition will give us
-better understanding of the constraints. 
+$\tar(\obs)$. Recall $\obs = (\meas_1,\dots,\meas_m)^t \in
+(\hilo^*)^m$ and each $\meas_j,j=1,\dots,m$ must be chosen from some
+allowed set of functionals in $\hilo^*$. This set differs based on the
+kind of sensors we have at our disposal and the properties of $\hilo$,
+as discussed in section \ref{subsec:dynamics}. The following
+proposition will gives better understanding of the constraints.
 \begin{restatable*}{proposition}{biggerbetter}\label{prop:bigger better}
@@
 appendix but the idea behind it is simple: making rows of $\obs$ large
-is equivalent to making $\sigma^2$ small. This can be understood
-easily from the formulation of the inverse problem $\data = \obs \fwd
-\param + \obs \eps' + \eps$, where $\eps'$ can be taken as model error
-and $\eps$ is iid measurement error.
-
-We now suggest a more convenient-to-work-with set of constraints. By
-Proposition \ref{prop:bigger better}, we cannot just take all $\ban^*$
-to be the allowed set of measurement vectors, since this would
-effectively eliminate measurement error entirely. We note that for
-point evaluations, the norm of the measurement is always one:
+is equivalent to making $\sigma^2$ small. %%  Viewed another way, it
+%% increases the Signal to Noise Ratio (SNR) --- which is always
+%% desirable when taking any measurement prone to error.
+This can be understood easily from the formulation of the inverse
+problem $\data = \obs \fwd \param + \obs \eps' + \eps$, where $\eps'$
+can be taken as model error and $\eps$ is iid measurement error.
+
+We now suggest a set of constraints that will make the model
+analytically tractable. By Proposition \ref{prop:bigger better}, we
+cannot just take all $\hilo^*$ to be the allowed set of measurement
+vectors, since this would effectively eliminate measurement error
+entirely. We note that for point evaluations, the norm of the
+measurement is always one:
 \begin{align*}
@@
 Following this, we also restrict the norm of our measurement vectors
-to be one. But the norm on functionals is hard to work with. Instead,
-we assume there exists a Hilbert space $\bar{\hil} \subseteq \ban^*$
-and take measurements $\meas_j \in \bar{\hil}^* = \bar{\hil}$ such
-that $\| \meas_j\| = 1, j=1,\dots,m$. Recall that we think of $\ban$ as
-a somewhat restricted space ($C(\bar{\Omega})$ with the sup-norm over
-a bounded domain $\Omega$, for example) and so $\ban^*$ can be thought
-of as large. Then $\ban^*$ can accommodate a reasonable Hilbert space
-$\bar{\hil}$. For example, if $\ban = C(\bar{\Omega})$, we take
-$\bar{\hil} = L^2(\Omega) \subseteq \ban^*$.
-%% (we can ignore the
-%% boundary since is has zero measure).
-Then
-%
-\begin{equation*}
-  \ban \subseteq \bar{\hil} \subseteq \ban^*,
-\end{equation*}
-since members of $C(\bar{\Omega})$ are bounded, hence square integrable. The
-unit norm constraint can be written using \eqref{eq:obs*} as a series of $m$ equality
-constraints (one for each measurement) on $\obs$ as
+to be one. The unit norm constraints can be written using
+\eqref{eq:obs*} as a series of $m$ equality constraints (one for each
+measurement) on $\obs$ as
 \begin{align}\label{eq:constraints}
-  \phi_j(\obs) :=\frac12 \| \obs^* e_j\|_{\bar{\hil}}^2 - \frac12 = 0,j=1,\dots,m
+  \phi_j(\obs) :=\frac12 \| \obs^* e_j\|_{\bar{\hilp}}^2 - \frac12 = 0,j=1,\dots,m
 \end{align}
@@
   &= \frac12\lim_{\tau \to 0}\tau^{-1}
-  ( \|(\obs + \tau V)^*e_j \|_{\bar{\hil}}^2 - \|\obs ^*e_j \|_{\bar{\hil}}^2  ) \\
+  ( \|(\obs + \tau V)^*e_j \|_{\bar{\hilp}}^2 - \|\obs ^*e_j \|_{\bar{\hilp}}^2  ) \\
   %
@@
   &= \frac12\lim_{\tau \to 0}\tau^{-1}
-  ( \langle (\obs + \tau V)^*e_j, (\obs + \tau V)^*e_j \rangle_{\bar{\hil}} - 
-  \langle \obs^*e_j, \obs^*e_j \rangle_{\bar{\hil}} ) \\
+  ( \langle (\obs + \tau V)^*e_j, (\obs + \tau V)^*e_j \rangle_{\bar{\hilp}} - 
+  \langle \obs^*e_j, \obs^*e_j \rangle_{\bar{\hilp}} ) \\
   % 
@@
   &= \frac12\lim_{\tau \to 0}\tau^{-1}
-  (2\tau \langle \obs^*e_j,V^*e_j \rangle_{\bar{\hil}} 
-  +\tau^2 \langle V^*e_j, V^*e_j \rangle_{\bar{\hil}} ) \\
-  %
-  %
-  % 
-  &= \langle \obs^*e_j,V^*e_j \rangle_{\bar{\hil}} \\
+  (2\tau \langle \obs^*e_j,V^*e_j \rangle_{\bar{\hilp}} 
+  +\tau^2 \langle V^*e_j, V^*e_j \rangle_{\bar{\hilp}} ) \\
+  %
+  %
+  % 
+  &= \langle \obs^*e_j,V^*e_j \rangle_{\bar{\hilp}} \\
   %
@@
 \end{align}
-with $\nabla \phi_j(\obs) \in \bar{\hil}^m \subseteq \ban^m$ and where the
+with $\nabla \phi_j(\obs) \in \hilo^m$ and where the
 last equality follows since $\obs^*e_j = \meas_j$ by \eqref{eq:obs*}. As
-we noted at the end of section \ref{section:D and grad}, the
+we noted at the end of section \ref{section:objective}, the
 gradient $\nabla \phi_j(\obs)$ is a row vector.
@@
 Recall where the objects involved are defined: the observation operator,
-$\obs \in \bar{\hil}^m \subseteq (\ban^*)^m$. The objective $\tar:
-(\ban^*)^m \to \R$ and its gradient is $\nabla \tar(\obs) \in
-\ban^m$. The constraints are defined similarly --- $\phi_j: (\ban^*)^m
-\to \R$ and $\nabla \phi_j(\obs) \in \ban^m$.
+$\obs \in (\hilo^*)^m$. The objective $\tar:
+(\hilo^*)^m \to \R$ and its gradient is $\nabla \tar(\obs) \in
+\hilo^m$. The constraints are defined similarly --- $\phi_j: (\hilo^*)^m
+\to \R$ and $\nabla \phi_j(\obs) \in \hilo^m$.
 
@@
 In this section we analyze optimal designs when no model error is
-present. We show how sensor clusterization arises in section
-\ref{subsec:clusterization}. We prepare the ground by first deriving
-an eigenvalue problem in $\obs$ for an optimal design in section
-\ref{subsec:eigenproblem}. This eigenvalue problem is recorded in
-\eqref{eq:mod conditions}. Next, we characterize optimal designs as
-ones that make all eigenvalues in equal in the above mentioned
-eigenproblem. This is crucial for clusterization to occur. The
-derivation is presented in section \ref{subsec:characterization}.
-
-\subsection{The Eigenvalue Problem}\label{subsec:eigenproblem}
-When $\modcov = 0$, then $\Sigma= \sigma^2I$. The necessary
-first-order condition for optimality \eqref{eq:conditions} becomes
+present. We discuss sensor clusterization in section
+\ref{section:clusterization}. We prepare the ground by first deriving
+a ``weird'' eigenvalue problem in $\obs$ for an optimal design in
+section \ref{subsec:eigenproblem}. This ``weird'' eigenvalue problem
+is recorded in \eqref{eq:mod conditions}. Next, we characterize
+optimal designs as ones that make all eigenvalues equal in the above
+mentioned weird eigenvalue problem. The derivation is presented in
+section \ref{subsec:characterization}.
+
+\subsection{The Weird Eigenvalue Problem}\label{subsec:eigenproblem}
+$\modcov = 0$ implies $\Sigma= \sigma^2I$. The necessary first-order
+condition for optimality \eqref{eq:conditions} becomes
 \begin{equation}\label{eq:eigenproblem}
@@
 \end{equation}
-with $\Xi$ diagonal. This is an eigenvalue problem for the
-self-adjoint operator $\sigma^{-2}\fwd \postcov \fwd^*$ where the rows
-of $\obs$, namely $\meas_j,j=1,\dots, m$, are the eigenvectors. Below,
-we use the expression for the posterior covariance \eqref{eq:postcov}
-in the second equality. If $\fwd$ is invertible, the third equality is
-trivial. Otherwise, we use some calculations delegated to lemma
-\ref{lemma:twice woodbury} in the appendix:
+with $\Xi$ diagonal. This looks like an eigenvalue problem for the
+self-adjoint operator $\sigma^{-2}\fwd \postcov \fwd^*$. The rows of
+$\obs$, namely $\meas_j,j=1,\dots, m$, are the eigenvectors. Note,
+however, that $\postcov$ depends on $\obs$ --- which makes
+\eqref{eq:eigenproblem} a ``weird'' eigenvalue problem.
+
+Below, the expression for the posterior covariance \eqref{eq:postcov}
+justifies the second equality. Some calculations delegated to lemma
+\ref{lemma:twice woodbury} in the appendix justify the third (trivial
+if $\fwd$ is invertible):
 \begin{align}\label{eq:mod conditions}
@@
 \end{align}  
-This presents the optimal design problem as a problem on $\ban$ and
-not on $\hil$ --- the term $(\fwd \prcov \fwd^*)^{-1}$ is the prior
-precision in $\ban$ and $(\fwd\prcov\fwd^*)^{-1} + \sigma^{-2} \obs^*
-\obs$ is the posterior precision in $\ban$.
+The term $(\fwd \prcov \fwd^*)^{-1}$ is the prior precision in $\hilo$
+(and is invertible, as mentioned in section
+\ref{section:prelim}). $(\fwd\prcov\fwd^*)^{-1} + \sigma^{-2} \obs^*
+\obs$ is the posterior precision in $\hilo$.
 
 \subsection{Characterizing D-Optimal Designs}\label{subsec:characterization}
-In light of \eqref{eq:mod conditions}, let us denote eigenvalues of
-$\fwd \prcov \fwd^*$ by $\{\lambda_i\}_{i=1}^{\infty}$ and its
-eigenvectors $\{\ev_i\}_{i=1}^{\infty}$. The following lemma (proved
-in the appendix \ref{lemma:sim diag}) is crucial for the rest of the
-analysis:
+Denote eigenvalues of $\fwd \prcov \fwd^*$ by
+$\{\lambda_i\}_{i=1}^{\infty}$. These are all positive. Denote
+corresponding $\{\ev_i\}_{i=1}^{\infty}$. Lemma \ref{lemma:sim diag}
+(proved in the appendix) is crucial for the rest of the analysis:
 
 \begin{restatable*}{lemma}{simdiag}\label{lemma:sim diag}
-  Let $C:\hil \to \hil$ self-adjoint and let $\func_1,\dots,\func_m \in
-  \hil$. Denote $\func^*$ the element $\func$ acting as a linear
-  functional. If
+  Let $\hil$ seperable Hilbert space, $C:\hil \to \hil$ self-adjoint
+  and $\func_1,\dots,\func_m \in \hil$. Denote $\func^*$ the element
+  $\func$ acting as a linear functional. If
   \begin{equation*}
-   (C + \sum_{j=1}^m \func_j\func_j^*) \func_k = \xi_k \func_k, k = 1,\dots,m
+   (C + \sum_{j=1}^m \func_j\func_j^*) \func_l = \xi_l \func_l, l = 1,\dots,m
   \end{equation*}
@@
 
-This allow us to arrive at the following:
 \begin{corollary}[Eigenvectors of a D-optimal design]\label{cor:same ev}
@@
 \end{proof}
-Note that all but finitely many of these eigenvectors have zero
-eigenvalue, since $\obs^*\obs$ is finite-rank. Denote the
-corresponding non-zero eigenvalues of $\obs^*\obs$ by
-$\{\eta_i\}_{i=1}^{k}$ and let $\eta_i = 0$ for $i \geq k+1$. Using
-Sylvester's determinant theorem we get:
+Only finitely many of the eigenvectors of $\obs^*\obs$ have positive
+eigenvalue. This happens because $k := \rank \obs^*\obs \leq
+m$. Denote the corresponding non-zero eigenvalues of $\obs^*\obs$ by
+$\{\eta_i\}_{i=1}^{k}$ and let $\eta_i = 0$ for $i \geq k+1$. No
+assumption is made regarding the magnitudes the corresponding
+$\{\lambda_i\}_{i=1}^{\infty}$. They are to be thought of as fixed but
+unknown. Using Sylvester's determinant theorem (the standard proof
+works for Hilbert spaces as well) we get:
 
@@
     %
-    &=\frac12 \sum_{i=1}^{k} \log (\sigma^2\lambda_i^{-1} + \eta_i), \text{ up to an additive constant}. 
+    &=\frac12 \sum_{i=1}^{k} \log (\sigma^2\lambda_i^{-1} + \eta_i), 
   \end{split}
 \end{align}
-
-Therefore, an optimal design maximizes
+up to an additive constant independent of $\obs$. Therefore, an
+optimal design maximizes
 
@@
 
-with the norm constraints on $\meas_j,j=1,\dots,m$. We show in the
-appendix (Lemma \ref{lemma:free}) we can find $\meas_1,\dots,\meas_m$
-such that $\{\eta_i\}_{i=1}^k$ have any value we desire, subject only
-to the restriction that we do not change the trace of $\obs^*\obs$, so
-that $\sum_{i=1}^k \eta_i = m$. By concavity of $\log$, the fastest
-increase in the design criterion (equation \eqref{eq:true target}) is
-gained by increasing $\sigma^2\lambda_i^{-1} + \eta_i$ where it is
-smallest. So, no weight should be given to (i.e. measurement taken in
-direction of) the $k$th eigenvector before $\sigma^2\lambda_i^{-1} +
-\eta_i \geq \sigma^2\lambda_k^{-1},\ \forall i \neq k$. Thus, we
-should choose $\{\eta_i\}_{i=1}^k$ as follows. First, order the
-eigenvalues $\lambda_1 \geq \lambda_2 \geq \dots$. We should increase
-$\sigma^2 \lambda_1^{-1} + \eta_1$ until it equals
-$\sigma^2\lambda_2^{-1}$. Then, $\sigma^2 \lambda_i^{-1} + \eta_i,i=1,2$ are
-increased until $\sigma^2 \lambda_i^{-1} + \eta_i =
+with the norm constraints on $\meas_j,j=1,\dots,m$. This implies that
+$\meas_i \in span\{\ev_j\}_{j=1}^k,i=1,\dots, m$. The problem of
+D-optimal design has now been cast to finite dimensions. This
+justifies using linear algebra. Lemma \ref{lemma:free} (proved in the
+appendix) does exactly that.
+\begin{restatable*}[Unit norm decomposition]{lemma}{free}\label{lemma:free}
+  Let $M \in \R^{k \times k}$ symmetric positive definite with $\ttr M
+  = m$, $m \geq k$. We can find $\func_j \in \R^k,j=1,\dots,m$
+  with $\|\func_j\|=1$ and $A = (\func_1,\dots,\func_m)$ such that
+  $AA^t = M$.
+\end{restatable*}
+
+Take any $\{\eta_i\}_{i=1}^{k}$, with $\sum_{i=1}^k \eta_i = m$. Let
+$M = \diag(\eta_1,\dots,\eta_k)$. The lemma gives us $A$ with unit
+norm columns such that $AA^T = M$. Think of $A^t$ as representing
+$\obs$, so we take $\meas_i :=\sum_{j=1}^k A_{ji}\ev_j^*$. The
+constraints $\|\meas_i\| = 1, i=1,\dots,m$ are satisified by
+construction. Verifying $\obs^*\obs \ev_i = \eta_i \ev_i$ is not too
+difficult:
+
+\begin{align*}
+  \begin{split}
+    \obs \ev_i &= (\meas_1,\dots,\meas_k)^t\ev_i\\
+    &= (\sum_{j=1}^kA_{j1}\ev_j^*,\dots,\sum_{j=1}^kA_{jk}\ev_j^*)^t\ev_i\\
+    &= (\sum_{j=1}^kA_{j1}\ev_j^*\ev_i,\dots,\sum_{j=1}^kA_{jk}\ev_j^*\ev_i)^t\\
+    &= (A_{i1},\dots,A_{ik})^t.
+  \end{split}
+\end{align*}
+
+Now,
+
+\begin{align*}
+  \begin{split}
+    \obs^*\obs \ev_i &= \obs^*(A_{i1},\dots,A_{ik})^t\\
+    &=(A_{i1},\dots,A_{ik}) \obs \text{ (by \eqref{eq:obs*})}\\
+    &=(A_{i1},\dots,A_{ik}) (\sum_{j=1}^kA_{j1}\ev_j^*,\dots,\sum_{j=1}^kA_{jk}\ev_j^*)^t\\
+    &=\sum_{l=1}^k A_{il} \sum_{j=1}^kA_{jl}\ev_j^* \\
+    %&=\sum_{j=1}^k  \sum_{l=1}^kA_{il}A_{jl}\ev_j^* \\
+    &=\sum_{j=1}^k  AA^t_{ij}\ev_j^* \\
+    %&=\eta_i\ev_i^*\\
+    &=\eta_i\ev_i. 
+  \end{split}
+\end{align*}
+
+By concavity of $\log$, the fastest increase in the design criterion
+(equation \eqref{eq:true target}) is gained by increasing
+$\sigma^2\lambda_i^{-1} + \eta_i$ where it is smallest. So, no weight
+should be given to (i.e. measurement taken in direction of) the $k$th
+eigenvector before $\sigma^2\lambda_i^{-1} + \eta_i \geq
+\sigma^2\lambda_k^{-1},\ \forall i \neq k$. Thus, we should choose
+$\{\eta_i\}_{i=1}^k$ as follows. First, order the eigenvalues
+$\lambda_1 \geq \lambda_2 \geq \dots$. We should increase $\sigma^2
+\lambda_1^{-1} + \eta_1$ until it equals
+$\sigma^2\lambda_2^{-1}$. Then, $\sigma^2 \lambda_i^{-1} +
+\eta_i,i=1,2$ are increased until $\sigma^2 \lambda_i^{-1} + \eta_i =
 \sigma^2\lambda_3^{-1},i=1,2$ and so forth.
@@
     \caption{Three graduated lab cylinders, corresponding to three
-      eigenvectors. Prior not shown.}
+      eigenvectors. Prior eigenvalues not shown.}
     \label{fig:cylinder}
@@
 The argument above has the important consequence (illustrated in
-figure \ref{fig:optimal vs not}) that the eigenvalues in the
+figure \ref{fig:optimal vs not}) that the eigenvalues in the ``weird''
 eigenvalue problem \eqref{eq:eigenproblem} are all equal! This makes
-perfect sense. First, we know the Lagrange multipliers tell us how
-much we gain by relaxing the constraints and there is no reason for
-any measurement to give us more than any other (otherwise we would put
-more weight on the better measurement, as can be understood from the
-discussion in the following paragraph).
+perfect sense. We know the Lagrange multipliers $\xi_i,i=1,\dots,k$
+tell us how much we gain by relaxing the constraints and there is no
+reason for any measurement to give us more than any other (otherwise
+we would put more weight on the better measurement, as can be
+understood from the discussion in the following paragraph).
 
 One can understand optimal designs as follows: Imagine each
-eigenvector is a lab cylinder (figure \ref{fig:cylinder}). Each
-cylinder $i$ is filled, a-priori, with green liquid
-$\sigma^2\lambda_i^{-1}$ volume units (this is the prior precision,
-weighted by obsevation noise). An optimal design with (e.g.) four
-measurements adds blue liquid of four volume units. It does so by
-adding each drop to whatever cylinder has the lowest level of liquid
-in it (lowest precision). The result of such a procedure is
-illustrated in figure \ref{fig:optimal vs not}. The analysis
-conducted above is summarized in theorem \ref{thm:char} below.
+eigenvector is a graduated lab cylinder (figure
+\ref{fig:cylinder}). Each cylinder $i$ is filled, a-priori, with green
+liquid of $\sigma^2\lambda_i^{-1}$ volume units. This liquid is the
+prior precision, weighted by obsevation noise. The volume of the green
+liquid is increasing in $i$. An optimal design with $m$ measurements
+ignores cylinders indexed $m+1$ and above. It has blue liquid of $m$
+volume units at its disposal. It distributes it by repeatedly adding a
+drop to whatever cylinder currently has the lowest level of liquid in
+it (lowest precision). The result of such a procedure is illustrated
+in figure \ref{fig:optimal vs not}. The analysis conducted in this
+section is summarized in theorem \ref{thm:char} below.
 
@@
 
-\begin{theorem}[Characterizing D-optimal designs]\label{thm:char}
-In a D-optimal design $\eta_i, i=1,\dots,k$ are chosen such that
-$\sigma^2\lambda_i^{-1} + \eta_i < \sigma^2\lambda_k^{-1}$ and
-$\sigma^2\lambda_i^{-1} + \eta_i$ are all equal for $1 \leq i \leq k$
-for some $k$. For $i \geq k+1$, however, $\eta_i = 0$.
-\end{theorem}
+\begin{restatable}[Characterizing D-optimal designs]{theorem}{main}\label{thm:char}
+Denote $\{\lambda_i, \ev_i\}_{i=1}^\infty$ the eigenvalues and
+eigenvectors of $\fwd\prcov\fwd^*$, in decreasing eigenvalue order.
+Denote $\{\eta_i\}_{i=1}^\infty$ eigenvalues of $\obs^*\obs$, chosen
+to maximize $\tar(\obs)$. Denote $m$ number of measurements taken and
+$\sigma^2$ observation noise variance. Then:
+\begin{enumerate}
+\item $\obs^*\obs$ and $\fwd\prcov\fwd^*$ are simultaneously diagonalizable.
+\item There exists $k \leq m$ such that $\eta_i = 0$ for $i\geq k+1$.
+\item $\sigma^2\lambda_i^{-1} + \eta_i \leq \sigma^2\lambda_{k+1}^{-1}, i=1,\dots,k$.
+\item $\sigma^2\lambda_i^{-1} + \eta_i$ are equal for all $1 \leq i \leq k$.
+\end{enumerate}
+\end{restatable}
 
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Analysis of Optimal Designs --- Non-Vanishing Model Error}
+\section{Analysis of Optimal Designs --- Non-Vanishing Model Error}\label{section:non vanishing}
 \subsection{Overview}
@@
 \begin{lemma}[Matrix Determinant Lemma in Hilbert Spaces]\label{lemma:MDL}
-  Let $\hil$ a separable Hilbert space, $u,v\in \hil$ and $A: \hil \to
-  \hil$ an invertible linear operator such that $\tr{A-I} <
+  Let $\hilp$ a separable Hilbert space, $u,v\in \hilp$ and $A: \hilp \to
+  \hilp$ an invertible linear operator such that $\tr{A-I} <
   \infty$. Then $\det A$ and $\det A + uv^*$ are well defined and
@@
   \cite{Simon77}. First, consider $B := I + xy^*$ for some $x,y \in
-  \hil$. We construct an eigenbasis for $B$ and use that to show $\det
+  \hilp$. We construct an eigenbasis for $B$ and use that to show $\det
   B = 1 + \langle x, y \rangle$. First let $x_1 := x$.  Now, if $x
@@
 
-\begin{lemma}\label{lemma:free}
-  Let $M \in \R^{k \times k}$ symmetric positive definite with $\ttr M =
-  m$, $m \geq k$. Then we can find $\func_j \in \R^k,j=1,\dots,m$ with
-  $\|\func_j\|=1$ and $A = (\func_1,\dots,\func_m)$ such that $AA^t =
-  M$.
-\end{lemma}
+\free
 \begin{proof}
@@
   Fix an arbitrary $j=1,\dots,m$ and take $V:= e_j e_j^t \obs$. We see
-  that for $u \in \ban$
+  that for $u \in \hilo$
   \begin{equation*}
@@
   increases $\tar(\obs)$. We start with the last line of \eqref{eq:tar
-    var}, and denote $\tmp : = \fwd \postcov \fwd^*: \ban^* \to \ban$:
+    var}, and denote $\tmp : = \fwd \postcov \fwd^*: \hilo^* \to \hilo$:
   \begin{align*}
@@
 %% and we assume $d_1 > d_2$ without loss of generality. For $\meas \in
-%% \hil$, we denote $\meas^*$ its Riesz representer, making a departure
+%% \hilp$, we denote $\meas^*$ its Riesz representer, making a departure
 %% from the standard notation employed, where $\meas$ is usually taken to
-%% be a linear functional on $\hil$. Let $\obs = (\meas_1^*, \meas_2^*,
+%% be a linear functional on $\hilp$. Let $\obs = (\meas_1^*, \meas_2^*,
 %% \meas_3^*)^t$. In this setup, we would like that $\meas_j,j=1,2,3$ are

### 2020-10-06 zeros.py

@@ -124,14 +124,16 @@ def main(m, k):
     M, _, _, _ = MDUS(m, k)
 
     A = get_A(M, m)
-
+    assert A.shape == (k,m)
+    
     ## Check that AAt = M, as stated.
     AAt = np.dot(A,A.T)
-    assert np.all(np.abs(AAt - M) < 1e-10)
+    assert np.all(np.abs(AAt - M) < EPS)
     
     ## Check that A does have unit norm columns, as stated.
     norms = np.linalg.norm(A, axis=0)
-    assert np.all(np.abs(norms - 1) < 1e-10) 
+    assert len(norms) == m
+    assert np.all(np.abs(norms - 1) < EPS) 
 
     
 def get_A(M, m):

### 2020-10-06 new_paper.tex
Fails because of ell wtf

@@
 evaluations of a continuous function\footnote{Of course, this is
-  impossible if, e.g., $\hilo = L^2(\Omega)$ for some $\Omega$. Still,
-  this approach is in line with previous analyses
-  \cite{Karina,AlexanderianGloorGhattas14} that consider sensor
-  placement but formulate the inverse problem on Hilbert spaces.}. Such
+  impossible if, e.g., $\hilo = L^2(\Omega)$ for some $\Omega$.}. Such
 measurements can be represented using the linear functionals
@@
 
-\begin{restatable}[Characterizing D-optimal designs]{theorem}{main}\label{thm:char}
-Denote $\{\lambda_i, \ev_i\}_{i=1}^\infty$ the eigenvalues and
-eigenvectors of $\fwd\prcov\fwd^*$, in decreasing eigenvalue order.
-Denote $\{\eta_i\}_{i=1}^\infty$ eigenvalues of $\obs^*\obs$, chosen
-to maximize $\tar(\obs)$. Denote $m$ number of measurements taken and
-$\sigma^2$ observation noise variance. Then:
+\begin{restatable}[Characterizing D-optimal designs]{theorem}{char}\label{thm:char}
+  Let:
+  \begin{enumerate}
+  \item $\fwd,\prcov,\obs$ the forward, prior covariance and observation operators, respectively.
+  \item $\{\lambda_i\}_{i=1}^\infty$ the eigenvalues of
+    $\fwd\prcov\fwd^*$ in decreasing order.
+    \item $\{\ev_i\}_{i=1}^\infty$ the corresponding eigenvectors.
+  \item $\{\eta_i\}_{i=1}^\infty$ eigenvalues of $\obs^*\obs$, where
+    $\obs$ maximizes $\tar(\obs)$.
+  \item $m$ the number of measurements taken.
+  \item $\sigma^2$ observation noise variance.
+  \end{enumerate}
+  Then:
   \begin{enumerate}
   \item $\obs^*\obs$ and $\fwd\prcov\fwd^*$ are simultaneously diagonalizable.
-\item There exists $k \leq m$ such that $\eta_i = 0$ for $i\geq k+1$.
-\item $\sigma^2\lambda_i^{-1} + \eta_i \leq \sigma^2\lambda_{k+1}^{-1}, i=1,\dots,k$.
+  \item There exists $0 < k \leq m$ such that $\eta_i = 0$ for $i\geq k+1$.
   \item $\sigma^2\lambda_i^{-1} + \eta_i$ are equal for all $1 \leq i \leq k$.
+  \item If $k < m$, then $\sigma^2\lambda_i^{-1} + \eta_i \leq
+    \sigma^2\lambda_{k+1}^{-1}, i=1,\dots,k$.
   \end{enumerate}
@@
 
-%% We are also used to symmetric optimization problems having
-%% symmetric solutions \cite{Waterhouse83}, so symmetric and identical
-%% Lagrange multipliers definitely make sense in this context.
-
 \section{Sensor Clusterization --- Vanishing Model Error}\label{section:clusterization}
 \subsection{Overview}
-In this section we explain how sensor clusterization may arise. We
-first consider a sequential optimal design, where measurements are
-chosen sequentially. We then consider the case where all measurements
-are chosen at once. Finally, we hypothesize how this may apply to a
-real-life inverse problem.
-
-%% illustrated in figure \ref{fig:clusterization} and
+In this section we use theorem \ref{thm:char} to explain how sensor
+clusterization may arise in various scenarios. We first consider a
+sequential optimal design. We then consider the case where all
+measurements are chosen at once. Finally, we hypothesize how this may
+apply to a real-life inverse problem.
 
@@
 \ref{subsec:seq vs sim} and the definition of $\obs_k$ from
-\eqref{eq:def obs_k}. The way clusterization can occur is understood
-using results of the previous section, specialized to one measurement
-each time. By corollary \ref{cor:ame ev}, the first measurement
-$\meas_1$ has to be (parallel to) an eigenvector of
-$\fwd\prcov\fwd^*$. By theorem \ref{thm:char}, it will be taken in the direction of the
-first eigenvector --- $\meas_1 \parallel \ev_1$. As hinted in
-section \ref{subsec:seq vs sim}, the posterior becomes the prior for
-the next step. Similar reasoning leads to:
+\eqref{eq:def obs_k}. The way clusterization can occur in this
+scenario is understood using thorem \ref{thm:char}, specialized to one
+measurement each time. In this case $m=1$, so $k=1$, so $\meas_1
+\parallel \ev_1$. As hinted in section \ref{subsec:seq vs sim}, the
+posterior becomes the prior for the next step. Similar reasoning leads
+to:
 \begin{align*}
@@
   \begin{split}
-    \meas_3 \parallel \ev_1  &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 < \sigma^2\lambda^{-1}_3 \text{ (note $\lambda^{-1}_1 < \lambda^{-1}_2$)}  \\
+    %\meas_3 \parallel \ev_1  &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 < \sigma^2\lambda^{-1}_3 \text{ (note $\lambda^{-1}_1 < \lambda^{-1}_2$)}  \\
     %\meas_3 \parallel  \ev_2  &\text{ if } \\

### 2020-10-14 new_paper.tex
looking better and better

@@
 Designing experiments is usually done by optimizing some \emph{design
-  criterion} and this is true both for frequentists
+  criterion}. This is true both for frequentists
 \cite{Silvey13,Ucinski05} as well as for Bayesians
-\cite{ChalonerVerdinelli95}, see \cite{ChalonerVerdinelli95} for an
+\cite{ChalonerVerdinelli95}. See \cite{ChalonerVerdinelli95} for an
 investigation of the analogy between the two approaches. Although
-there is a plethora of design criteria, in the current work we will
-focus on just one of these, commonly referred to as \emph{D-optimal
-  design}. It has a simple and appealing motivation in the Bayesian
-context as explained in \cite{ChalonerVerdinelli95}: for a linear
-model, a D-optimal design seeks to maximize the expected information
-gain (KL divergence \cite{KullbackLeibler51,CoverThomas12}) between
-posterior and prior. In finite dimensions, maximizing the expected
-information gain amounts to minimizing the determinant of the
-posterior covariance matrix. In a frequentist setting, an optimal
-design minimizes the volume of the uncertainty ellipsoid \cite[page
-  16]{Ucinski05}, but this is done for the Fisher information matrix
-and not the posterior covariance. However, \cite{ChalonerVerdinelli95}
-shows that the latter is just a regularized version of the former.
+there is a plethora of design criteria, we focus on just one of these,
+commonly referred to as \emph{D-optimal design}. It has a simple and
+appealing motivation in the Bayesian context as explained in
+\cite{ChalonerVerdinelli95}: for a linear model, a D-optimal design
+seeks to maximize the expected information gain (KL divergence
+\cite{KullbackLeibler51,CoverThomas12}) between posterior and
+prior. In finite dimensions, maximizing the expected information gain
+amounts to minimizing the determinant of the posterior covariance
+matrix. In a frequentist setting, a D-optimal design minimizes the
+volume of the uncertainty ellipsoid \cite[page 16]{Ucinski05}, but
+this is done for the Fisher information matrix and not the posterior
+covariance. However, \cite{ChalonerVerdinelli95} show that the latter
+is just a regularized version of the former.
 
@@
 inference in a Bayesian context was pioneered by
-\cite{AlexanderianGloorGhattas14,AlexanderianPetraStadlerEtAl16,
+\cite{AlexanderianGloorGhattas14, AlexanderianPetraStadlerEtAl16,
   AlexanderianPetraStadlerEtAl14}. Similarly to the finite dimensional
@@
 linear models when one wishes to maximize the KL divergence between
-posterior and prior and this amounts to minimizing the determinant of
-the posterior covariance operator (understood as a product of its
+posterior and prior. This amounts to minimizing the determinant of the
+posterior covariance operator (understood as a product of its
 eigenvalues). Some difficulties arise in the process, but remedies can
@@
 d=1,2,3$ (with added error) then the optimization procedure sometimes
-places sensors in close proximity to each other (as can be seen
+places sensors in very close proximity to each other (as can be seen
 in figure \ref{fig:clusterization illustration}). Following
@@
 \cite{Fedorov96} and \cite[chapter 2.4.3]{Ucinski05} discuss this
-phenomenon and suggest an approach called clusterization-free design,
-where the user enforces measurement locations to be far from each
-other. One way to do this is by introducing correlated errors which,
-philosophically, accounts for both measurement error and model error.
-Another method considered is imposing distance constraints between
-measurements. A somewhat different approach is suggested in \cite[page
-  49]{Fedorov12}, where close by design measurements are merged --- a
-procedure which obviously does not avoid clusterization. The same problem
-arises in time-series analysis for pharmacokinetic experiments. The
-authors of \cite{Hooker09} suggest modeling auto-correlation time in
-the noise model, which is equivalent to the correlated errors
-mentioned above.
+phenomenon and suggest an approach called clusterization-free design.
+In such designs, the user enforces measurement locations to be far
+from each other. One way to do this is by introducing correlated
+errors which, philosophically, accounts for both measurement error and
+model error. Another method considered is imposing distance
+constraints between measurements. A somewhat different approach is
+suggested in \cite[page 49]{Fedorov12}, where close by design
+measurements are merged --- a procedure which obviously does not avoid
+clusterization. The same problem arises in time-series analysis for
+pharmacokinetic experiments. The authors of \cite{Hooker09} suggest
+modeling auto-correlation time in the noise model, which is equivalent
+to the correlated errors mentioned above.
 
@@
 arises as a solution to a constrained optimization problem, formulated
-using Lagrange multipliers (section \ref{section:D and grad}). In the
-case of no model error we show this problem is in fact a set of
-eigenvector-eigenvalue problems (equation \eqref{eq:eigenproblem}). We
-then characterize an optimal design as one that reduces the highest
-uncertainties, and show how that can give rise to repeated
-measurements (section \ref{section:clusterization} and figure
-\ref{fig:clusterization}).
+using Lagrange multipliers (section \ref{section:D and grad}). Using
+this model we rigorously show how model error mitigates clusterization
+(section \ref{section:non vanishing}). We show how clusterization
+occurs in a sequential design scenario (section
+\ref{subsec:clusterization sequential}). Then, we show that designs
+that exhibit clusterization are just as optimal as clusterization free
+designs in a simultaneous D-optimal designs (section
+\ref{subsec:clusterization simultaneous}).
+
+A beautiful mathematical structure arises in D-optimal designs when no
+model error is present. The Lagrange multipliers problems is in fact a
+``weird'' eigenvalue problem (``weird'' because the operator depends
+on its eigenvectors, see \eqref{eq:eigenproblem}). This helps us shed
+light into D-optimal designs. We characterize these designs using
+theorem \ref{thm:char}. One key and intuituve insight we prove is that
+a D-optimal design reduces uncertainties where they are highest
+first. More interestig phenomena arise, but they require setting
+notation, so these are discussed later.
 
 In the process we generalize several lemmas from linear algebra to
-infinite-dimensional settings --- the Matrix Determinant Lemma and a
-lemma for calculating the derivative of determinants. We also provide
-other tools for understanding D-optimal designs in
-infinite-dimensional Bayesian inverse problems, including lemmas for
-calculating the increase in the design criterion per observation ---
-Lemmas \ref{lemma:design increase} and Corollary \ref{cor:zero mod
-  err}. %%  We also show that in our model with independent measurement
-%% noise, an interesting phenomenon arises, in which the eigenvectors
-%% of the measurement operator (denoted below $\obs^*\obs$) are the
-%% same as the eigenvectors of the prior covariance in observation
-%% space, which helps arrive at the results outlined in the previous
-%% paragraph.
+infinite-dimensional settings. A matrix determinant emma and a lemma
+due to Lax \cite{lax97} for calculating the derivative of determinants
+of certain operators (\ref{lemma:MDL} and \ref{lemma:lax},
+respectfully). We also prove some non-trivial lemmas in linear
+algebra. One of which (\ref{lemma:free}) gives a decomposition $M =
+AA^t$ where $A$ has unit norm columns. Another (\ref{lemma:sim diag})
+shows simultaneous diagonizability for the ``weird'' eigenvalue problem
+referenced above. We also provide other tools for understanding
+D-optimal designs in infinite-dimensional Bayesian inverse problems,
+including lemmas for calculating the increase in the design criterion
+per observation --- Lemmas \ref{lemma:design increase} and Corollary
+\ref{cor:zero mod err}.
 
 \subsection{Limitations}\label{subsec:limitations}
-There are two main drawbacks of the arguments presented here. First,
-one can argue that we allow considerably less restrictive observations
-than any real-life problem does. This is because our observation model
-allows taking measurements that are arbitrarily weighted linear
-combinations of eigenvectors. The second drawback is that we do not
-show rigorously that clusterization necessarily occurs in our model,
-we only show that there is no reason to be surprised when it does
-occur.
+There are two main drawbacks of the study presented here. Our relaxed
+model does not consider any specific set of allowed
+measurements. Rather, we take measurements in the unit ball in some
+Hilbert space. This allows considerably less restrictive observations
+than any real-life problem does. The second drawback is that we do not
+show rigorously that clusterization necessarily occurs in a
+simultaneous design. We only show that it is as reasonable as no
+clusterization.
 
@@
 
+\begin{theorem}[Characterizing D-optimal designs]\label{thm:char}
+  Let:
+  \begin{enumerate}
+  \item $\fwd, \prcov, \obs$ the forward, prior covariance and
+    observation operators, respectively.
+  \item $\{\lambda_i\}_{i=1}^\infty$ the eigenvalues of
+    $\fwd\prcov\fwd^*$ in decreasing order.
+  \item $\{\ev_i\}_{i=1}^\infty$ the corresponding eigenvectors.
+  \item $\{\eta_i\}_{i=1}^\infty$ eigenvalues of $\obs^*\obs$, where
+    $\obs$ maximizes $\tar(\obs)$.
+  \item $m$ the number of measurements taken.
+  \item $\sigma^2$ observation noise variance.
+  \end{enumerate}
+  Then:
+  \begin{enumerate}
+  \item $\obs^*\obs$ and $\fwd\prcov\fwd^*$ are simultaneously diagonalizable.
+  \item There exists $0 < k \leq m$ such that $\eta_i = 0$ for $i\geq k+1$.
+  \item $\sigma^2\lambda_i^{-1} + \eta_i$ are equal for all $1 \leq i \leq k$.
+  \item If $k < m$, then $\sigma^2\lambda_i^{-1} + \eta_i \leq
+    \sigma^2\lambda_{k+1}^{-1}, i=1,\dots,k$.
+  \end{enumerate}
+\end{theorem}
+
 \pgfplotstableread{
@@
 
-\begin{restatable}[Characterizing D-optimal designs]{theorem}{char}\label{thm:char}
-  Let:
-  \begin{enumerate}
-  \item $\fwd,\prcov,\obs$ the forward, prior covariance and observation operators, respectively.
-  \item $\{\lambda_i\}_{i=1}^\infty$ the eigenvalues of
-    $\fwd\prcov\fwd^*$ in decreasing order.
-    \item $\{\ev_i\}_{i=1}^\infty$ the corresponding eigenvectors.
-  \item $\{\eta_i\}_{i=1}^\infty$ eigenvalues of $\obs^*\obs$, where
-    $\obs$ maximizes $\tar(\obs)$.
-  \item $m$ the number of measurements taken.
-  \item $\sigma^2$ observation noise variance.
-  \end{enumerate}
-  Then:
-  \begin{enumerate}
-  \item $\obs^*\obs$ and $\fwd\prcov\fwd^*$ are simultaneously diagonalizable.
-  \item There exists $0 < k \leq m$ such that $\eta_i = 0$ for $i\geq k+1$.
-  \item $\sigma^2\lambda_i^{-1} + \eta_i$ are equal for all $1 \leq i \leq k$.
-  \item If $k < m$, then $\sigma^2\lambda_i^{-1} + \eta_i \leq
-    \sigma^2\lambda_{k+1}^{-1}, i=1,\dots,k$.
-  \end{enumerate}
-\end{restatable}
-
 \section{Sensor Clusterization --- Vanishing Model Error}\label{section:clusterization}
@@
 clusterization may arise in various scenarios. We first consider a
-sequential optimal design. We then consider the case where all
-measurements are chosen at once. Finally, we hypothesize how this may
-apply to a real-life inverse problem.
-
-\subsection{Sequential Design}
+sequential optimal design. Then we consider simultaneous
+designs.
+
+\subsection{Sequential Design}\label{subsec:clusterization sequential}
 Recall the sequential optimal design scheme from section
 \ref{subsec:seq vs sim} and the definition of $\obs_k$ from
-\eqref{eq:def obs_k}. The way clusterization can occur in this
-scenario is understood using thorem \ref{thm:char}, specialized to one
-measurement each time. In this case $m=1$, so $k=1$, so $\meas_1
-\parallel \ev_1$. As hinted in section \ref{subsec:seq vs sim}, the
-posterior becomes the prior for the next step. Similar reasoning leads
-to:
+\eqref{eq:def obs_k}. The way clusterization can occur in sequential
+design scenario is understood by specializing thorem \ref{thm:char} to
+one measurement each time. In this case $m=1$, so $k=1$, $\meas_1
+\parallel \ev_1$ and $\eta_1 = 1$. As hinted in section
+\ref{subsec:seq vs sim}, the posterior becomes the prior for the next
+step. Similar reasoning leads to:
 \begin{align*}
@@
   \begin{split}
-    %\meas_3 \parallel \ev_1  &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 < \sigma^2\lambda^{-1}_3 \text{ (note $\lambda^{-1}_1 < \lambda^{-1}_2$)}  \\
+    \meas_3 \parallel \ev_1  &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 < \sigma^2\lambda^{-1}_3 \\
     %\meas_3 \parallel  \ev_2  &\text{ if } \\
@@
 Again, the first case results in clusterization. The sequential design
-proceeds in this fashion, where every $\meas_i$ is taken in the
-direction of the eigenvector of smallest eigenvalue of the previous
-posterior precision operator. Taking enough measurements we will
-surely arrive at clusterization, since the prior precision increases
-quickly --- in section \ref{subsec:abstract OED} we assumed  $\fwd$
-is strongly smoothing, which typpically means eigenvalues of $(\fwd
-\prcov\fwd^*)^{-1}$ grow exponentially.
-
-\subsection{Simultaneous Design}
+proceeds in this fashion. Every $\meas_i$ is taken in the direction of
+the eigenvector of smallest eigenvalue of the previous posterior
+precision operator. Taking enough measurements we will surely arrive
+at clusterization, since the prior precision increases quickly --- in
+section \ref{subsec:abstract OED} we assumed $\fwd$ is strongly
+smoothing, which means eigenvalues of $(\fwd \prcov\fwd^*)^{-1}$ grow
+quickly.
+
+\subsection{Simultaneous Design}\label{subsec:clusterization simultaneous}
 In a simultaneous design scenario, the result is not as clear cut as
@@
 
-\subsection{Why do Realistic Designs Exhibit Clusterization?}
-It is unclear from the previous analysis what makes realistic designs
-exhibit clusterization, such as the one observed in figure
-\ref{fig:clusterization illustration}. We can only hypothesize.
-
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 \section{Analysis of Optimal Designs --- Non-Vanishing Model Error}\label{section:non vanishing}
-\subsection{Overview}
 In this section we show the effect model error has on the
-clusterization phenomenon. In contrast to the previous case of
-vanishing model error, we will see that if $\modcov \neq 0$ clustering
-will not occur.
-
-\subsection{Derivation}
-We rely on corollary \ref{cor:same meas} below, which is proved in the
-appendix. Denote $\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm :=
-(\meas_1,\dots,\meas_{m-1})^t$.
+clusterization phenomenon. We will see that if $\modcov \neq 0$
+clustering will not occur. This is contrary to the previous case of
+vanishing model error.
+
+Corollary \ref{cor:same meas} below is proved in the appendix. Denote
+$\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm :=
+(\meas_1,\dots,\meas_{m-1})^t$. Every symbol with a wide hat
+$\ \widehat{}\ $ refers to the marked object for $\obsm$. E.g. $\Sigmam =
+\Sigma(\obsm)$.
 \begin{restatable*}{corollary}{samemeas}\label{cor:same meas}
@@
 Recall from \eqref{eq:Sigma} that $\Sigma(\obs) = \obs \modcov \obs^*
-+ \sigma^2I$. From Corollary \ref{cor:same meas} we can deduce that
++ \sigma^2I$. Then
 
@@
 Hence no increase in the design criterion is achieved by taking a
-repeated measurement. This is clearly sub-optimal and $\obsm$ cannot
-be an optimal design. Thus, for small measurement error levels, the
-clusterization effect will be mitigated by the presence of a non-zero
-model error. Since the design criterion is not defined for $\sigma^2 =
-0$ and identical measurements, we cannot make a statement regarding
-$\sigma^2 = 0$, except in the limiting sense described above.
+repeated measurement (in the limit $\sigma^2 \to 0$). This is clearly
+sub-optimal and $\obsm$ cannot be an optimal design. Thus, for small
+measurement error levels, the clusterization effect will be mitigated
+by the presence of a non-zero model error. Since the design criterion
+is not defined for $\sigma^2 = 0$ and identical measurements, we
+cannot make a statement regarding $\sigma^2 = 0$, except in the
+limiting sense described above.
 
@@
 \section{Acknowledgements}
-This study is a part of my PhD thesis \cite{mine} which was written
-under the instruction of Prof. Georg Stadler in New York University's
-Courant Institute. I would like to thank him for his great mentorship.
+This study is a part of my PhD thesis. It was written under the
+instruction of Prof. Georg Stadler in New York University's Courant
+Institute. I would like to thank him for his great mentorship.
 
@@
 \begin{lemma}[Matrix Determinant Lemma in Hilbert Spaces]\label{lemma:MDL}
-  Let $\hilp$ a separable Hilbert space, $u,v\in \hilp$ and $A: \hilp \to
-  \hilp$ an invertible linear operator such that $\tr{A-I} <
+  Let $\hil$ a separable Hilbert space, $u,v\in \hil$ and $A: \hil \to
+  \hil$ an invertible linear operator such that $\tr{A-I} <
   \infty$. Then $\det A$ and $\det A + uv^*$ are well defined and
@@
   \cite{Simon77}. First, consider $B := I + xy^*$ for some $x,y \in
-  \hilp$. We construct an eigenbasis for $B$ and use that to show $\det
+  \hil$. We construct an eigenbasis for $B$ and use that to show $\det
   B = 1 + \langle x, y \rangle$. First let $x_1 := x$.  Now, if $x
@@
   and not destroying zeros we already introduced, starting from the
-  last diagonal entry and moving to the first. If $c_{mm} \neq 0$ ,
+  last diagonal entry and moving to the first. Since $c_{mm} \neq 0$ ,
   let $p < m$ such that $c_{pp}c_{mm} < 0$ (such $p$ exists because
@@
 
-%% \begin{proposition}\label{prop:bigger better}
-%%   Let $\obs = (\meas_1,\dots,\meas_m)^t$, $j \in \{1,\dots,m\}$, $\sigma^2
-%%   > 0$ and $\lambda > 1$. Then $\tar(\obs)$ increases if we use
-%%   $\lambda \meas_j$ in $\obs$ instead of $\meas_j$.
-%% \end{proposition}
 \biggerbetter
@@
   increases $\tar(\obs)$. We start with the last line of \eqref{eq:tar
-    var}, and denote $\tmp : = \fwd \postcov \fwd^*: \hilo^* \to \hilo$:
+    var}, and denote $\tmp : = \fwd \postcov \fwd^*$.
   \begin{align*}
@@
 \bibliographystyle{amsplain}
-\bibliography{refs.bib,georg_refs}
+\bibliography{refs,georg_refs}
 

### 2020-10-14 differ.sh

@@ -1,0 +1,8 @@
+#!/bin/bash
+#latexdiff --config="PICTUREENV=(?:picture|DIFnomarkup|align|tabular)[\w\d*@]*" paper.tex new_paper.tex > diff.tex
+# latexdiff --exclude-textcmd=”section,subsection,subsubsection” --config=”PICTUREENV=(?:picture|DIFnomarkup|align|tabular)[\w\d*@]*” paper.tex new_paper.tex > diff.tex
+# latexdiff paper.tex new_paper.tex > diff.tex
+#latexmk -pdf -f diff
+latexmk -pdf paper
+latexmk -pdf new_paper
+latexmk -pdf reply

### 2020-10-22 new_paper.tex
getting closer

@@
 appealing motivation in the Bayesian context as explained in
-\cite{ChalonerVerdinelli95}: for a linear model, a D-optimal design
-seeks to maximize the expected information gain (KL divergence
+\cite{ChalonerVerdinelli95}: A D-optimal design seeks to maximize the
+expected information gain (KL divergence
 \cite{KullbackLeibler51,CoverThomas12}) between posterior and
-prior. In finite dimensions, maximizing the expected information gain
-amounts to minimizing the determinant of the posterior covariance
+prior. Consider a linear model in finite dimensions, with Gaussian
+prior and noise. In this model, maximizing the expected information
+gain amounts to minimizing the determinant of the posterior covariance
 matrix. In a frequentist setting, a D-optimal design minimizes the
@@
 We propose and thoroughly study an analytically tractable model for
-understanding D-optimal designs. In this model, an optimal design
-arises as a solution to a constrained optimization problem, formulated
+understanding D-optimal designs. In this model, D-optimal design
+are solutions of a constrained optimization problem, formulated
 using Lagrange multipliers (section \ref{section:D and grad}). Using
@@
 model error is present. The Lagrange multipliers problems is in fact a
-``weird'' eigenvalue problem (``weird'' because the operator depends
-on its eigenvectors, see \eqref{eq:eigenproblem}). This helps us shed
-light into D-optimal designs. We characterize these designs using
-theorem \ref{thm:char}. One key and intuituve insight we prove is that
-a D-optimal design reduces uncertainties where they are highest
-first. More interestig phenomena arise, but they require setting
-notation, so these are discussed later.
+nonlinear eigenvalue problem for the observations and prior
+covariance. The operator for which eigenvectors and eigenvalues are
+sought is a sum of two operators. The first is the prior
+covariance. The second is an outer product of the observations. See
+\eqref{eq:eigenproblem} for details and exact statement. This
+structure helps us shed light into D-optimal designs. We characterize
+these designs using theorem \ref{thm:char}. One insight we prove in
+said theorem, is that a D-optimal design reduces uncertainties where
+they are highest first. Other interestig phenomena arise, but they
+require setting notation, and are discussed later.
 
 In the process we generalize several lemmas from linear algebra to
-infinite-dimensional settings. A matrix determinant emma and a lemma
-due to Lax \cite{lax97} for calculating the derivative of determinants
-of certain operators (\ref{lemma:MDL} and \ref{lemma:lax},
-respectfully). We also prove some non-trivial lemmas in linear
-algebra. One of which (\ref{lemma:free}) gives a decomposition $M =
-AA^t$ where $A$ has unit norm columns. Another (\ref{lemma:sim diag})
-shows simultaneous diagonizability for the ``weird'' eigenvalue problem
-referenced above. We also provide other tools for understanding
-D-optimal designs in infinite-dimensional Bayesian inverse problems,
-including lemmas for calculating the increase in the design criterion
-per observation --- Lemmas \ref{lemma:design increase} and Corollary
-\ref{cor:zero mod err}.
+infinite-dimensional settings. We prove a matrix determinant lemma in
+\ref{lemma:MDL}. We generalize a lemma due to Lax \cite{lax97} for
+calculating $\frac{\der}{\der t} \log \det (I + X(t))$, for an
+operator valued function $X(t)$ in \ref{lemma:lax}. We also prove
+lemmas in linear algebra. One gives a decomposition $M = AA^t$ where
+$A$ has unit norm columns \ref{lemma:free}. Another shows simultaneous
+diagonizability of the prior and outer product in the nonlinear
+eigenvalue problem referenced above \ref{lemma:sim diag}. We also
+provide other tools for understanding D-optimal designs in
+infinite-dimensional Bayesian inverse problems, including lemmas for
+calculating the increase in the design criterion per observation ---
+Lemmas \ref{lemma:design increase} and Corollary \ref{cor:zero mod
+  err}.
 
@@
       [
+      title={Posterior Pointwise Standard Deviations and D-Optimal Sensor Locations},  
       xmin = 0,
@@
       %% table{locs-heat-sens3-var1.txt};
-      \addplot [green,  only marks, mark=x, mark size=1.5] 
+      \addplot [green,  only marks, mark=*, mark size=1.5] 
       table{locs-heat-sens4-var1.txt}; 
-      \addplot [purple, only marks, mark=x, mark size=1.5] 
+      \addplot [purple, only marks, mark=*, mark size=1.5] 
       table{locs-heat-sens5-var1.txt}; 
-      \addplot [cyan,   only marks, mark=x, mark size=1.5] 
+      \addplot [cyan,   only marks, mark=*, mark size=1.5] 
       table{locs-heat-sens6-var1.txt};
@@
   \end{tikzpicture}
-  \caption{Shown are pointwise posterior standard deviations for
-    various numbers of sensors, and location of these sensors for the
-    1D heat equation with homogeneous Dirichlet boundary described in
-    section \ref{subsec:example}. Note the clusterization that occurs
-    when 6 sensors are used --- only 4 measurement locations are
-    included. Observed data does not influence the pointwise
-    variances, hence it is omitted.}
+  \caption{The clusterization effect for the 1D heat equation
+    described in section \ref{subsec:example}. Posterior pointwise
+    standard deviations are plotted over the domain $[0, \pi]$, for
+    varying numbers of sensors (lines). Sensor locations (circles)
+    were chosen to minimize (an expression analogous to) the
+    determinant of the posterior covariance. The clusterization effect
+    can be clearly seen for six sensors. Only four measurement
+    locations are used --- two pairs of sensors are so close they are
+    indistinguishable.}
   \label{fig:clusterization illustration}
 \end{figure}
-In section \ref{section:abstract OED} we present a more abstract and
-general formulation of the problem, but for the purpose of
-illustration, we present the problem via a toy model --- the 1D heat
-equation in $[0,\pi]$ with homogeneous Dirichlet boundary
-conditions.
+In section \ref{section:prelim} we present a more abstract and general
+formulation of the problem, but for the purpose of illustration, we
+present the problem via a toy model --- the 1D heat equation in
+$[0,\pi]$ with homogeneous Dirichlet boundary conditions.
 
@@
 posterior covariance is known and depends only on $\prcov, \fwd, \obs$ and
-$\sigma^2$ (see section \ref{section:abstract OED} and
+$\sigma^2$ (see section \ref{section:prelim} and
 \eqref{eq:postcov} specifically).
 
-We will consider generalization of the information-theoretic design
-criterion presented in section \ref{section:OED intro} to infinite
-dimensions (section \ref{subsec:D optimal design}). We choose
-$\x_j,j=1,\dots,m$ so as to minimize an expression analogous to the
-log-determinant of the posterior covariance operator (corresponding to
-a D-optimal design). We can see the clusterization effect in figure
-\ref{fig:clusterization illustration} --- the design criterion
-suggests placing sensors in locations that are extremely close.
+We consider generalization of the information-theoretic design
+criterion presented in the introduction to infinite dimensions
+(section \ref{subsec:D optimal design} below). We choose
+$\x_j,j=1,\dots,m$ to minimize (an expression analogous to) the
+log-determinant of the posterior covariance operator. We will see
+later how this corresponds to maximizing expected information gain.
+
+The clusterization effect is illustrated in figure
+\ref{fig:clusterization illustration}. Posterior pointwise standard
+deviations are plotted over the domain $[0, \pi]$. For a Bayesian
+linear inverse problem with Gaussian prior and noise the posterior
+covariance does not depend on data \cite{Stuart10}. Thus, the plot has
+no reference to actual data observed. The posterior covariance does,
+however, depend on the measurement taken. In figure
+\ref{fig:clusterization illustration}, measurement locations are
+marked by circles. These were chosen to minimize (an expression
+analogous to) the log-determinant of the posterior covariance. The
+clusterization effect can be clearly seen for six sensors. It looks
+like only four measurements were taken. The reason is that two pairs
+of sensors are so close they are indistinguishable.
 
@@
 \subsection{Bayesian Linear Inverse Problems}\label{subsec:abstract OED}
-Let $\hilp$ and $\hilo$ be seperable Hilbert spaces, both
-infinite-dimensional (the subscripts p and o are for ``parameter'' and
-``observation'', respectfully). We denote, $\hilo^*$ the Hilbert space
-of all linear functionals on $\hilo$. From standard functional
-analysis, it is known that $\hilp$ and $\hilo$ are isomorphic (and in
-practice may be equal) and that $\hilo = \hilo^*$. We will still use
-such notation, to keep track of where a certain object is
-defined. Consider $\fwd: \hilp \to \hilo$, a linear operator (the
+Let $\hilp$ and $\hilo$ be seperable Hilbert spaces (the subscripts p
+and o are for ``parameter'' and ``observation'', respectfully). We
+denote $\hilo^*$ the Hilbert space of all linear functionals on
+$\hilo$. Consider $\fwd: \hilp \to \hilo$, a linear operator (the
 ``forward operator''). We are interested in forward operators that are
-strongly smoothing (have fast decaying singular values --- the heat
-operator from section \ref{subsec:example} is a prime example). We
-want to infer $\param \in \hilp$, some parameter of the dynamics given
+strongly smoothing (have fast decaying modes --- the heat operator
+from section \ref{subsec:example} is a prime example). Our goal is to
+infer $\param \in \hilp$ --- some parameter of the dynamics --- given
 noisy observations of $\fwd \param$. We take a zero mean Gaussian
@@
 in $\hilo$ \cite{Stuart10}. As such, it is assumed invertible --- an
-assumption we rely heavily upon. Observations (measurements) are taken
-via the linear observation operator $\obs \in ( \hilo^* )^m$, where
-$m$ is the number of observations allowed (e.g. sensors at our
-disposal). In an analogy with linear algebra (where row vectors are
-thought of as linear functionals), we think of $\obs$ as having
+assumption we rely heavily upon. Throughout this paper, we denote $m$
+the number of measurements we are allowed to take. Observations
+(measurements) are taken via the linear observation operator $\obs \in
+( \hilo^* )^m$. In an analogy with linear algebra (where row vectors
+are thought of as linear functionals), we think of $\obs$ as having
 ``rows'' $\meas_j, j=1,\dots,m$:
@@
 %% but we may drop the parentheses and write $\meas_j u = \meas_j(u)$.
-A few observations regarding $\obs$ are in order. First, it is good to
-keep in mind that $( \hilo^* )^m$ is a Hilbert space with norm $\| \obs
-\| = \sum_{j=1}^m \|\meas_j\|$. Next, for $u \in \hilo$ and $v\in
-\R^m$:
+%% A few observations regarding $\obs$ are in order.
+%% First, it is good to keep in mind that $( \hilo^* )^m$ is a Hilbert
+%% space with norm $\| \obs \| = \sum_{j=1}^m \|\meas_j\|$.
+For $u \in \hilo$ and $v\in \R^m$:
 \begin{align*}
@@
 
-\begin{observation}
-  It is best to think of $\meas_j,j=1,\dots,m$ as row vectors and of
-  $\obs$ as a matrix with rows $\meas_j$. 
-\end{observation}
-
-Each measurement consists of a linear functional, chosen from an
-allowed set of linear functionals in $\hilo^{*}$. Data is acquired via
-noisy measurements
-\begin{align*}
-  \data := \obs (\fwd \param + \eps') + \eps = \obs \fwd \param + \obs \eps' + \eps
+%% \begin{observation}
+%%   It is best to think of $\meas_j,j=1,\dots,m$ as row vectors and of
+%%   $\obs$ as a matrix with rows $\meas_j$. 
+%% \end{observation}
+
+Each measurement is a linear functional, chosen from some subset of
+$\hilo^{*}$. Data is acquired via noisy measurements
+\begin{align*}
+  \data := \obs (\fwd \param + \eps') + \eps = \obs \fwd \param + \obs \eps' + \eps,
 \end{align*}
@@
 \sigma^2 I_m)$, with $I_m$ the $m \times m$ identity. Both error terms
-and the prior are assumed independent of each other. Let us now
-understand the terms present in the above expression.
+and the prior are assumed independent of each other.
 
@@
 \end{align}
-Usually, this is written using $\tmp := \obs \fwd$ and $\tmp^*$ but
-since we need to separate the observation operator $\obs$ from the
-forward operator $\fwd$ we write the explicit expression
-\eqref{eq:postcov}.
+Usually, this is written using $\tmp := \obs \fwd$ and $\tmp^*$. Since
+we need to separate the observation operator $\obs$ from the forward
+operator $\fwd$ we write the explicit expression
+\eqref{eq:postcov}. We now turn to understand the terms present in the
+above expression.
 
@@
 linear functionals and we refer to these as {\it measurement
-  vector}s. %% Note that we cannot choose every measurement vector we
-%% want, though. For example, we may be restricted only to pointwise
-%% evaluations of $\fwd \param$ by the sensors at our disposal, in which
-%% case measuring the mean $\ell(u) = \int_{\Omega}u$ (or any non-atomic
-%% measure) of a function is not possible.
+  vector}s. Note that we cannot choose every measurement vector we
+want. For example, we may be restricted only to pointwise evaluations
+of $\fwd \param$ by the sensors at our disposal, in which case
+measuring the mean $\ell(u) = \int_{\Omega}u$ (or any non-atomic
+measure) of a function is not possible.
 
 \subsubsection{Error Terms}
-Considering the sum of the error terms it is easy to see that $\obs
-\eps' + \eps \in \R^m$. It is a centered Gaussian random vector and
-its covariance matrix is
+Considering the sum of the error terms it is easy to see that
+$\bar{\eps} := \obs \eps' + \eps \in \R^m$. It is a centered Gaussian
+random vector and its covariance matrix is
 \begin{align}\label{eq:Sigma}
@@
 \end{equation}
-where the first equality in \eqref{eq:modcov explained} holds by
-definition and the second by \eqref{eq:obs*}. The explicit dependence
-on $\obs$ will be mostly dropped for notational convenience, so
-$\Sigma(\obs) = \Sigma$. Thus, for a fixed $\obs$ (i.e.\ a fixed set
-of measurements) we can equivalently write $\data = \obs \fwd \param +
-\bar{\eps}$ with $\bar{\eps} \sim \normal(0,\Sigma)$. Taking $\modcov
-= 0$ is common practice
-\cite{Tarantola05,KaipioSomersalo05,Vogel02}. This reduces to the case
-where we take iid observations. Then $\Sigma = \sigma^2I$ is simply a
-scalar matrix that does not depend on $\obs$. Note that taking an
-error model with a non-scalar covariance as we do here allows us to
-consider model error (modeled by $\modcov$) as well as measurement
-error (modeled by $\sigma^2$). For example, say we believe our forward
-model does not capture some small scale phenomenon.  Then we may
-express this belief by saying $\tru = \fwd + \err$, with $\fwd$
+with $e_j \in \R^m$ the $j$th standard basis vector. The first
+equality in \eqref{eq:modcov explained} holds by definition and the
+second by \eqref{eq:obs*}. The explicit dependence on $\obs$ will be
+mostly dropped for notational convenience, so $\Sigma(\obs) =
+\Sigma$. Thus, for a fixed $\obs$ (i.e.\ a fixed set of measurements)
+we can equivalently write $\data = \obs \fwd \param + \bar{\eps}$ with
+$\bar{\eps} \sim \normal(0,\Sigma)$. Taking $\modcov = 0$ is common
+practice \cite{Tarantola05,KaipioSomersalo05,Vogel02}. This reduces to
+the case where we take iid observations. Then $\Sigma = \sigma^2I$ is
+simply a scalar matrix that does not depend on $\obs$. Note that
+taking an error model with a non-scalar covariance as we do here
+allows us to consider model error (modeled by $\modcov$) as well as
+measurement error (modeled by $\sigma^2$). For example, say we believe
+our forward model does not capture some small scale phenomenon.  Then
+we may express this belief by saying $\tru = \fwd + \err$, with $\fwd$
 depending on $\param$ and $\err$ depending on $\sspar$, and $\param
@@
 correlation in the error mitigates the clusterization phenomenon, as
-is also reported in the literature \cite{Ucinski05}.
-
-\subsection{D-Optimal Design in Infinite Dimensions}\label{subsec:D optimal design} 
+reported in the literature \cite{Ucinski05}.
+
+\subsection{D-Optimal Designs in Infinite Dimensions}\label{subsec:D optimal design} 
+A D-optimal design maximizes expected KL divergence between posterior
+and prior. It is useful to recall the definition of KL divergence for
+an arbitrary prior measure:
+$$
+KL(\post^\data||\pr) = \int \log \frac{\der \post^\data}{\der \pr}(x) \der \post(x).
+$$
+
 The meaning of D-optimal design in infinite-dimensional Hilbert spaces
@@
 infinite-dimensions, a D-optimal design is well-defined as maximizing
-the expected KL divergence between posterior and prior and is
-summarized, using our notation, in the following theorem:
+the expected KL divergence between posterior and prior. The main
+result is summarized, using our notation, in the following theorem:
 \begin{theorem}[Slightly modified Theorem 1 from \cite{AlexanderianGloorGhattas14}]
@@
 
+The intuition behind definition \ref{def:d optimality} is
+straightforward. From \eqref{eq:postcov}:
+\begin{align*}
+  \begin{split}
+    ( I + \prcov^{1/2}  \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}) &= \prcov ( \prcov^{-1} + \fwd ^* \obs^* \Sigma^{-1} \obs \fwd) \\
+    &= \prcov \postcov^{-1}.
+  \end{split}
+\end{align*}
+Since $\prcov$ is constant, a D-optimal design minimizes the posterior
+covariance determinant, analogously to the finite-dimensional case.
+
 \subsection{Sequential vs Simultaneous Optimization}\label{subsec:seq vs sim}
-From \eqref{def:d optimality} we wish to characterize solution(s) of the
+From defintion \ref{def:d optimality} we wish to characterize solution(s) of the
 following optimization problem for $\tar$. %%: (\hilo^*)^m \to \R$:
@@
 where $\obs$ is constrained to some allowed set of measurements (we
-may \textit{think} of these as pointwise evaluations on $[0,\pi]$,
-even though these are not in $\hilo$). We call this problem
-``simultaneous optimization'', since all measurements are decided on
-simulatneously.
+may \textit{think} of these as pointwise evaluations, even though
+these might not be in $\hilo$). We call this problem ``simultaneous
+optimization'', since all measurements are decided on simulatneously.
 
@@
 optimization case. The sequential optimization case is dealt with in
-section \ref{subsec:seq clus}. It is important to note, however that
+section \ref{subsec:seq vs sim}. It is important to note, however that
 all conclusions we arrive at for the simultaneous case easily
@@
 We start by calculating the first variation. Some calculations are
-delegated to Lemma \ref{lemma:aux calc} in the appendix. Also, we use
-the following lemma whose proof is also delegated to the appendix
-\ref{lemma:lax}:
+delegated to Lemma \ref{lemma:aux calc} in the appendix. We use
+the following lemma:
 \begin{restatable*}[Generalized from \cite{Lax97}]{lemma}{lax}\label{lemma:lax}
@@
 
-First variation of $\tar (\obs)$ in the direction $V$ is:
+The proof of \ref{lemma:lax} is delegated to the appendix. Let
+$T(\obs) := \obs^*\Sigma(\obs)\obs$. The first variation of $\tar
+(\obs)$ in the direction $V$ is:
 \begin{align*}%%\label{eq:tar var}
@@
 as discussed in section \ref{subsec:dynamics}. The following
-proposition will gives better understanding of the constraints.
+proposition gives better understanding of the constraints.
 \begin{restatable*}{proposition}{biggerbetter}\label{prop:bigger better}
@@
 \begin{align}\label{eq:constraints}
-  \phi_j(\obs) :=\frac12 \| \obs^* e_j\|_{\bar{\hilp}}^2 - \frac12 = 0,j=1,\dots,m
+  \phi_j(\obs) :=\frac12 \| \obs^* e_j\|_{\hilp}^2 - \frac12 = 0,j=1,\dots,m.
 \end{align}
-with $e_j \in \R^m$ the $j$th standard basis vector. The first variations are:
+The first variations are:
 \begin{align*}
@@
 \ref{section:clusterization}. We prepare the ground by first deriving
-a ``weird'' eigenvalue problem in $\obs$ for an optimal design in
+a ``weird'' eigenvalue problem in $\obs$ for D-optimal designs in
 section \ref{subsec:eigenproblem}. This ``weird'' eigenvalue problem
@@
 
-\subsection{The Weird Eigenvalue Problem}\label{subsec:eigenproblem}
+\subsection{The ``Weird'' Eigenvalue Problem}\label{subsec:eigenproblem}
 $\modcov = 0$ implies $\Sigma= \sigma^2I$. The necessary first-order
@@
 $\{\lambda_i\}_{i=1}^{\infty}$. These are all positive. Denote
-corresponding $\{\ev_i\}_{i=1}^{\infty}$. Lemma \ref{lemma:sim diag}
-(proved in the appendix) is crucial for the rest of the analysis:
-
-\begin{restatable*}{lemma}{simdiag}\label{lemma:sim diag}
+corresponding eigenvectors $\{\ev_i\}_{i=1}^{\infty}$. Lemma
+\ref{lemma:sim diag} (proved in the appendix) is crucial for the rest
+of the analysis:
+
+\begin{restatable*}[Simultaneous diagonizability for the nonlinear eigenvalue problem]{lemma}{simdiag}\label{lemma:sim diag}
   Let $\hil$ seperable Hilbert space, $C:\hil \to \hil$ self-adjoint
@@
 \end{align}
-up to an additive constant independent of $\obs$. Therefore, an
-optimal design maximizes
+up to an additive constant independent of $\obs$. Therefore, a
+D-optimal design maximizes
 
@@
     %&=\sum_{j=1}^k  \sum_{l=1}^kA_{il}A_{jl}\ev_j^* \\
-    &=\sum_{j=1}^k  AA^t_{ij}\ev_j^* \\
+    &=\sum_{j=1}^k  [AA^t]_{ij}\ev_j^* \\
     %&=\eta_i\ev_i^*\\
-    &=\eta_i\ev_i. 
+    &=\eta_i\ev_i^*. 
   \end{split}
@@
 $\sigma^2\lambda_i^{-1} + \eta_i$ where it is smallest. So, no weight
-should be given to (i.e. measurement taken in direction of) the $k$th
+should be given to (i.e. measurement taken in direction of) the $n$th
 eigenvector before $\sigma^2\lambda_i^{-1} + \eta_i \geq
-\sigma^2\lambda_k^{-1},\ \forall i \neq k$. Thus, we should choose
+\sigma^2\lambda_n^{-1},\ \forall i \neq n$. Thus, we should choose
 $\{\eta_i\}_{i=1}^k$ as follows. First, order the eigenvalues
@@
 tell us how much we gain by relaxing the constraints and there is no
-reason for any measurement to give us more than any other (otherwise
-we would put more weight on the better measurement, as can be
+reason for any eigenvector to give us more than any other (otherwise
+we would put more weight on the better eigenvector, as can be
 understood from the discussion in the following paragraph).
 
-One can understand optimal designs as follows: Imagine each
+One can understand D-optimal designs as follows: Imagine each
 eigenvector is a graduated lab cylinder (figure
@@
 liquid of $\sigma^2\lambda_i^{-1}$ volume units. This liquid is the
-prior precision, weighted by obsevation noise. The volume of the green
-liquid is increasing in $i$. An optimal design with $m$ measurements
-ignores cylinders indexed $m+1$ and above. It has blue liquid of $m$
-volume units at its disposal. It distributes it by repeatedly adding a
-drop to whatever cylinder currently has the lowest level of liquid in
-it (lowest precision). The result of such a procedure is illustrated
-in figure \ref{fig:optimal vs not}. The analysis conducted in this
-section is summarized in theorem \ref{thm:char} below.
+prior precision, weighted by observation noise. The volume of the
+green liquid is increasing in $i$. A D-optimal design with $m$
+measurements ignores cylinders indexed $m+1$ and above. It has blue
+liquid of $m$ volume units at its disposal. It distributes said liquid
+by repeatedly adding a drop to whatever cylinder currently has the
+lowest level of liquid in it (lowest precision). The result of such a
+procedure is illustrated in figure \ref{fig:optimal vs not}. The
+analysis conducted in this section is summarized in theorem
+\ref{thm:char} below.
 
@@
       \addlegendentry{$\sigma^2\lambda_i^{-1}$}
-      \addlegendentry{$\eta_i$ from an optimal design}
+      \addlegendentry{optimal $\eta_i$s}
 
@@
       \addlegendentry{$\sigma^2\lambda_i^{-1}$}
-      \addlegendentry{$\eta_i$ from a sub-optimal desgin}
+      \addlegendentry{sub-optimal $\eta_i$s}
 
@@
   \end{tikzpicture}
-  \caption{Characterizing an optimal design, compared to a
-    sub-optimal design. Both designs are allowed three measurements,
-    which means the blue area has accumulated height of $3$ in both
-    plots. An optimal design (left) increases precision (reduces
-    uncertainty) where it is lowest. A sub-optimal design (right) does
-    not. In this example, a measurement is taken partly in the direction of the
-    fourth eigenvector, where precision is highest.}
+  \caption{Characterizing an optimal design, compared to a sub-optimal
+    design. Both designs are allowed three measurements, which means
+    the blue area has accumulated height of $3$ in both plots. An
+    optimal design (left) increases precision (reduces uncertainty)
+    where it is lowest. A sub-optimal design (right) does not. In this
+    example, a measurement is taken partly in the direction of the
+    third eigenvector, where precision is bigger even compared to the
+    posterior precision of the optimal design on the left.}
   \label{fig:optimal vs not}
@@
 In this section we use theorem \ref{thm:char} to explain how sensor
-clusterization may arise in various scenarios. We first consider a
-sequential optimal design. Then we consider simultaneous
-designs.
+clusterization may arise in various scenarios. We first consider
+sequential designs, and then simultaneous designs.
 
@@
 $\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm :=
-(\meas_1,\dots,\meas_{m-1})^t$. Every symbol with a wide hat
-$\ \widehat{}\ $ refers to the marked object for $\obsm$. E.g. $\Sigmam =
-\Sigma(\obsm)$.
+(\meas_1,\dots,\meas_{m-1})^t$. Denote $\Sigmam := \Sigma (\obsm)$ and
+$\postcovm$ the posterior covariance that arises when we use $\obsm$.
 \begin{restatable*}{corollary}{samemeas}\label{cor:same meas}
@@
 \appendix
-\section{Widely Applicable Lemmas}
-
-The following lemma is generalized from \cite[Chapter 9,
-  Theorem 4, pp. 127]{Lax97}
+\section{General Lemmas}
+
+The following lemma is generalized from \cite[Chapter 9, Theorem 4, pp. 127]{Lax97}
 \lax
@@
     %
+    &= \sum_{k=1}^{\infty} \prod_{l=1, l\neq k}^{\infty}
+      (1+\lambda_l(s)) \frac{\partial (1+\lambda_k(s))}{\partial (1+\lambda_k)}\Big |_{s=t}
+    \frac{\der (1+\lambda_k)}{\der t} \\
+    %
+    %
+    %    
     &= \sum_{k=1}^{\infty} \frac{\prod_{l=1}^{\infty}
@@
   \end{align*}
-  Since there is nothing special about $t=0$, the relation holds for all $t$.
+  There is nothing special about $t_0 = 0$ --- we could have chosen
+  any other $t_0$ instead. Thus, the relation holds for all $t$.
 \end{proof}
@@
   \end{equation*}
-\end{proof}
-
-\section{Specific Lemmas}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{lemma}\label{lemma:twice woodbury}
-  Assume $\fwd \prcov \fwd^*$ is invertible. Then
-\begin{align*}
-  \begin{split}
-    \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* 
-    %
-    %
-    = \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1},
-  \end{split}
-\end{align*}  
-\end{lemma}
-\begin{proof}
-  The proof is boring and amounts to using Woodbury's matrix identity
-  twice and a regularization trick. The standard proof for this
-  identity works in infinite dimensions, as long as all terms are well
-  defined. Unfortunately, $\obs^*\obs$ is not invertible, so we force
-  it to be. Throughout, we let $X := \fwd\prcov \fwd^*$ and $Y_{\eps}
-  := \obs^*\obs + \eps I$, for some small $\eps > 0$. Note that now
-  $Y_{\eps}$ is invertible. Recall Woodbury's matrix identity:
-  $$
-  (A + UCV)^{-1} = A^{-1} + A^{-1}U(C^{-1} + VA^{-1}U)^{-1}VA^{-1}.
-  $$
-
-  Taking $A := \prcov^{-1}, U := \fwd^*, V := \fwd$ and $C :=
-  \sigma^{-2} (\obs^*\obs+\eps I)$ we get:
-  \begin{align*}
-    \begin{split}
-      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* &=
-      \fwd ( \prcov - \prcov \fwd^* ( \sigma^2(\obs^*\obs + \eps I)^{-1} + \fwd \prcov \fwd^* )^{-1} \fwd \prcov ) \fwd^* \\
-      %
-      %
-      &= X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X
-    \end{split}
-  \end{align*}
-  Note that $X + \sigma^2 Y_{\eps}^{-1}$ is invertible, as the sum of
-  a two positive definite operators. Now, taking $A := X^{-1}, C :=
-  \sigma^2Y_{\eps}^{-1}, U := I$ and $V := I$ and using Wodbury's matrix
-  identity in reverse order:
-  \begin{align*}
-    \begin{split}
-      X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X &= (X^{-1} + \sigma^{-2}Y_{\eps})^{-1} \\
-      %
-      %
-      %
-      &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \eps I))^{-1}.
-    \end{split}
-  \end{align*}
-
-  We conclude that $\forall \eps > 0$
-  \begin{align*}
-    \begin{split}
-      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* 
-     &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \eps I))^{-1}.
-    \end{split}
-  \end{align*}
-  Letting $\eps \to 0$ completes the proof.
-\end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{lemma}[Increase due to a Measurement]\label{lemma:design increase}
-  Let $\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm := (\meas_1,\dots,\meas_{m-1})^t$. Then
-  \begin{align*}
-    \tar( \obs ) - \tar (\obsm ) &=
-    \frac12 \log \left ( 1 + \frac{
-      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m,
-      (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m \rangle
-    }{
-      \sigma^2 + \meas_m \modcov \meas_m - \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m 
-    }       
-    \right ).
-  \end{align*}
-\end{lemma}
-\begin{proof}
-  We use the Schur complement to write one inverse in terms of the other and
-  introduce notations to make the derivation cleaner. Note that we think of
-  $\obsm$ and $\obsm^*$ as column and row vectors (respectively).
-  \begin{align*}
-    \Sigma( \obs ) &= \Sigma = 
-    \begin{bmatrix}
-      \Sigma (\obsm )           & \obsm \modcov \meas_m \\
-      \meas_m \modcov \obsm^*   & \sigma^2 + \meas_m \modcov \meas_m
-    \end{bmatrix}
-    =
-    \begin{bmatrix}
-      \Sigmam   & w \\
-      w^t       & c
-    \end{bmatrix}\\
-    %
-    %
-    %
-    \Sigma^{-1} &=
-    \begin{bmatrix}
-      \Sigmam^{-1} + \Sigmam^{-1} w ( c - w^t \Sigmam^{-1} w)^{-1} w^t \Sigmam^{-1} & - \Sigmam^{-1} w ( c - w^t \Sigmam^{-1} w)^{-1} \\
-      -( c - w^t \Sigmam^{-1} w)^{-1} w^t \Sigmam^{-1}                            &  ( c - w^t \Sigmam^{-1} w)^{-1}
-    \end{bmatrix} \\
-    &=
-    \begin{bmatrix}
-      \Sigmam^{-1} & 0 \\
-      0           & 0 
-    \end{bmatrix}
-    + (c -w^t \Sigmam^{-1} w )^{-1}
-    \begin{bmatrix}
-      \Sigmam^{-1} w \\
-      -1
-    \end{bmatrix}
-    \begin{bmatrix}
-      w^t \Sigmam^{-1} & -1 
-    \end{bmatrix}
-  \end{align*}
-  %
-  Further, define
-  %
-  \begin{align*}
-    \M (\obs ):&= \prcov^{\frac12}\fwd^* \obs^* \Sigma^{-1} \obs \fwd
-    \prcov^{\frac12}    
-  \end{align*}
-  %
-  and note that, using our understanding of what is a column vector and
-  what is a row vector:
-  %
-  \begin{align*}
-    \M(\obs) &= \prcov^{1/2} \fwd^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2} \\
-    %
-    %
-    %
-    &= \prcov^{1/2} \fwd^* \obs^* \left \{
-    \begin{bmatrix}
-      \Sigmam^{-1} & 0 \\
-      0           & 0 
-    \end{bmatrix}
-    + (c -w^t \Sigmam^{-1} w )^{-1}
-    \begin{bmatrix}
-      \Sigmam^{-1} w \\
-      -1
-    \end{bmatrix}
-    \begin{bmatrix}
-      w^t \Sigmam^{-1} & -1 
-    \end{bmatrix} 
-    \right \} \obs \fwd \prcov^{1/2} \\
-    %
-    %
-    %
-    &= \M (\obsm) + (c -w^t \Sigmam^{-1} w )^{-1}
-    \prcov^{1/2} \fwd^* \obs^*
-    \begin{bmatrix}
-      \Sigmam^{-1} w \\
-      -1
-    \end{bmatrix}
-    \begin{bmatrix}
-      w^t \Sigmam^{-1} & -1 
-    \end{bmatrix} 
-    \obs \fwd \prcov^{1/2}
-  \end{align*}
-  %
-  Now, denote:
-  %
-  \begin{align}\label{eq:u}
-    \begin{split}
-      u :&= (c -w^t \Sigmam^{-1} w )^{-1/2}
-      \prcov^{1/2} \fwd^* \obs^* 
-      \begin{bmatrix}
-        \Sigmam^{-1} w \\
-        -1 
-      \end{bmatrix} \\
-      %
-      %
-      %
-      & = (c -w^t \Sigmam^{-1} w )^{-1/2} ( \prcov^{1/2}\fwd^* \obsm^* \Sigmam^{-1} \obsm  \modcov \meas_m - \prcov^{1/2} \fwd^* \meas_m )\\
-      %
-      %
-      %
-      u^* :&=  (c -w^t \Sigmam^{-1} w )^{-1/2} (\meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \fwd \prcov^{1/2} - \meas_m \fwd \prcov^{1/2} ),
-    \end{split}
-  \end{align}
-  %
-  so that
-  %
-  \begin{equation}\label{eq:M plus I}
-    I + \M( \obs ) = I + \M (\obsm ) + uu^*.
-  \end{equation}
-  %
-  Note that
-  \begin{equation}\label{eq:M postcov}
-    \prcov^{1/2} \left (I + \M( \obsm ) \right )^{-1} \prcov^{1/2} = \postcovm.
-  \end{equation}
-  The increase in the design criterion gained by including $\meas_m$
-  in the design is found by using Lemma \ref{lemma:MDL} the above
-  results:
-  %
-  \begin{align*}
-    \tar( \obs ) - \tar( \obsm )
-    %
-    %
-    %
-    &= \frac12 \log \det \Big ( I + \M ( \obs ) \Big ) / \det \Big ( I + \M (\obsm) \Big ) \\
-    %
-    %
-    %
-    &= \frac12  \log \det \left ( I + \M(\obsm) + uu^* \right ) / \det \Big ( I + \M (\obsm) \Big ) \\
-    %
-    %
-    %
-    &= \frac12 \log \left ( 1 + \left \langle \left ( I+\M(\obsm) \right )^{-1} u, u  \right \rangle \right ).
-  \end{align*}
-  Using \eqref{eq:u}:
-  \begin{align*}
-    &\left \langle \left (I+\M (\obsm)\right )^{-1}u, u \right \rangle\\
-    &= \frac{
-      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m,
-      (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m \rangle
-    }{
-      c- w^t \Sigmam^{-1} w
-    }\\
-    %
-    %
-    %
-    &= 
-    \frac{
-      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m,
-      (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m \rangle
-    }{
-      \sigma^2 + \meas_m \modcov \meas_m - \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m 
-    }
-  \end{align*}
-  and the conclusion follows.
-\end{proof}
-
-Lemma \ref{lemma:design increase} implies the following two corollaries.
-\begin{corollary}[Gain for No Model Error]\label{cor:zero mod err}
-  If $\modcov = 0$, then
-  \begin{equation*}
-    \tar( \obs ) - \tar (\obsm )
-    = -\frac12 \log (1 - \sigma^{-2} \langle \fwd \postcov \fwd^* \meas_m, \meas_m \rangle ).
-  \end{equation*}
-\end{corollary}
-\begin{proof}
-  Note that this is not immediate by substituting $\modcov = 0$ in the
-  conclusion of Lemma \ref{lemma:design increase}, since we make a
-  claim for $\postcov$, and not $\postcovm$. Let us first review
-  \eqref{eq:u} and note that since $\modcov = 0$ the covariance
-  $\Sigmam = \sigma^2I_{m-1}$ and $w = 0$, so $c - w^t\Sigmam^{-1}w =
-  \sigma^2$:
-  \begin{align*}
-    u :&=-\sigma^{-1}\prcov^{1/2} \fwd^* \meas_m\\
-    %
-    %
-    u^* :&= -\sigma^{-1} \meas_m \fwd \prcov^{1/2}.
-  \end{align*}
-  From \eqref{eq:M plus I}:
-  \begin{equation*}
-    I + \M(\obsm) = I +\M(\obs) - uu^*,
-  \end{equation*}
-  and thus:
-  \begin{equation*}
-    \left \langle \left ( I +\M(\obs) \right )^{-1} u, u \right \rangle
-    = \sigma^{-2} \langle \fwd \postcov \fwd^* \meas_m, \meas_m \rangle.
-  \end{equation*}
-  Analogously to \eqref{eq:M postcov} we note that
-  \begin{equation*}
-    \prcov^{1/2} \left ( I + \M(\obs) \right )^{-1} \prcov^{1/2} = \postcov.
-  \end{equation*}
-  Using Lemma \ref{lemma:MDL} we conclude
-  \begin{align*}
-    \tar( \obs ) - \tar( \obs )
-    &= \frac12 \log \det \left (I +\M(\obs) \right ) / \det \left (I + \M(\obsm) \right ) \\
-    %
-    %
-    %
-    &= \frac12 \log \det \left (I +\M(\obs) \right ) / \det \left (I + \M(\obs) - uu^* \right ) \\
-    %
-    %
-    %
-    &=-\frac12 \log (1 - \langle (I+\M(\obs))^{-1}u, u \rangle \\
-    %
-    %
-    %
-    &= -\frac12 \log (1 - \sigma^{-2} \langle \fwd \postcov \fwd^* \meas_m, \meas_m \rangle ).
-  \end{align*}
-\end{proof}
-
-\samemeas
-\begin{proof} \label{cor:same meas proof}
-  Denote $A:= \obs \modcov \obs^*$ and $v_j$ the $j$th column of $A$.
-  Note that $v_j = \obsm \modcov \meas_m$, since $(\obsm \modcov
-  \obsm^*)_{ij} = \meas_i(\modcov \meas_j)$, as explained in
-  \eqref{eq:modcov explained}. One can now verify that
-  \begin{equation}\label{eq:observation}
-    \Sigmam^{-1} \obsm \modcov \meas_m = \Sigmam^{-1}v_j = (A +\sigma^2I_{m-1})^{-1} v_j =
-    e_j -\sigma^2 \Sigmam^{-1}e_j.
-  \end{equation}
-  %
-  Using \eqref{eq:observation}:
-  \begin{align}\label{eq:denominator}
-    \begin{split}
-      \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m
-      &= \meas_m \modcov \obsm^* ( e^j - \sigma^2 \Sigmam^{-1} e_j )\\
-      %
-      %
-      %
-      &= \meas_m \modcov \meas_j - \sigma^2 \meas_m \modcov \obsm^* \Sigmam^{-1}e_j \\
-      %
-      %
-      %
-      &= \meas_m \modcov \meas_j -\sigma^2 (e_j - \sigma^2 \Sigmam^{-1}e_j)^t e_j \\
-      %
-      %
-      %
-      &= \meas_m \modcov \meas_m -\sigma^2 + \sigma^4 e_j^t\Sigmam^{-1}e_j.
-    \end{split}
-  \end{align}
-  We use \eqref{eq:observation} to simplify the terms in the enumerator of
-  the conclusion of Lemma \ref{lemma:design increase}:
-  \begin{align}\label{eq:enumerator}
-    \begin{split}
-      (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m
-      &= \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m - \meas_m \\
-      %
-      %
-      %
-      &= \obsm^* (e_j - \sigma^2 \Sigmam^{-1} e_j) -\meas_j \\ 
-      %
-      %
-      %
-      &= -\sigma^2 \obsm^* \Sigma^{-1}e_j. 
-    \end{split}
-  \end{align}
-  %
-  Substitute \eqref{eq:enumerator} and \eqref{eq:denominator} to
-  the enumerator and denominator (respectively) of the conclusion of
-  Lemma \ref{lemma:design increase}:
-  %
-  \begin{align*}
-    \tar( \obs ) - \tar (\obsm ) &=
-    \log \left ( 1 + \frac{
-      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m,
-      (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m \rangle
-    }{
-      \sigma^2 + \meas_m \modcov \meas_m - \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m 
-    }       
-    \right ) \\
-    %
-    %
-    %
-    &= \log \left ( 1 + \frac{\sigma^4
-      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
-      \obsm^* \Sigmam^{-1}e_j \rangle
-    }{
-      2\sigma^2 - \sigma^4 e_j^t\Sigmam^{-1}e_j 
-    }       
-    \right ) \\
-    %
-    %
-    %
-    &= \log \left ( 1 + \frac{\sigma^2
-      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
-      \obsm^* \Sigmam^{-1}e_j \rangle
-    }{
-      2 - \sigma^2 e_j^t\Sigmam^{-1}e_j 
-    }       
-    \right ).
-  \end{align*}
 \end{proof}
@@
 
-\begin{lemma}[Auxilliary Calculations]\label{lemma:aux calc}
-  Let $T(\obs) := \obs^* \Sigma^{-1}(\obs)\obs$, with $\Sigma(\obs)$
-  defined as in \eqref{eq:Sigma}. Then:
-  \begin{align*}
-    \delta T(\obs)V &= V^* \Sigma^{-1} \obs 
-    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
-    &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs
-    + \obs^* \Sigma^{-1} V.
-  \end{align*}
-\end{lemma}
-
-\begin{proof}
-  We need a few supplementary calculations. First:
-  \begin{align}\label{eq:der sig}
-    \begin{split}
-      \frac{\der}{\der \tau} \Big |_{\tau=0} \Sigma( \obs + \tau V )
-      &= \frac{\der}{\der \tau} \Big |_{\tau=0} 
-      (\obs + \tau V ) \modcov (\obs + \tau V )^*  + \sigma^2I\\
-      % 
-      % 
-      % 
-      &= V \modcov \obs^* + \obs \modcov V^*.
-    \end{split}
-  \end{align}
-  By the standard trick for the derivative of an operator: 
-  \begin{align*}
-    0 &= \frac{\der}{\der \tau} \Big |_{\tau=0} I \\
-    % 
-    % 
-    % 
-    &= \frac{\der}{\der \tau} \Big |_{\tau=0}
-    \left (\Sigma(\obs+\tau V)^{-1} \Sigma(\obs+\tau V) \right ) \\
-    % 
-    % 
-    % 
-    &= \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0} \Sigma+
-    \Sigma^{-1} \frac{\der \Sigma(\obs+\tau V)}{\der \tau} \Big |_{\tau=0}\\  
-    %
-    %
-    %
-    &= \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0} \Sigma+
-    \Sigma^{-1} (V\modcov \obs^* + \obs \modcov V^*) 
-    \text{, by \eqref{eq:der sig}. }
-  \end{align*}
-  Now we can write the variation of $\Sigma^{-1}$:
-  \begin{align}\label{eq:der sig inv}
-    \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0}  
-      &= -\Sigma^{-1} (V \modcov \obs^* + \obs \modcov V^*) \Sigma^{-1}.
-    \end{align}
-  % 
-  Finally, we can do the calculation we need to. By Leibnitz (product) rule
-  and \eqref{eq:der sig inv}:
-  % 
-  \begin{align*}
-    \delta T(\obs) V 
-    &= \frac{\der T(\obs + \tau V)}{\der \tau} \Big |_{\tau=0} \\
-    %
-    %
-    %
-    &= V^* \Sigma^{-1} \obs 
-    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
-    &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs
-    + \obs^* \Sigma^{-1} V,
-  \end{align*}
-  as required.
-\end{proof}
-
 \simdiag
@@
   \ell$ and we conclude that $C$ and $A$ have the same eigenvectors.
+\end{proof}
+
+\section{Specific Lemmas}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{lemma}\label{lemma:twice woodbury}
+  Assume $\fwd \prcov \fwd^*$ is invertible. Then
+\begin{align*}
+  \begin{split}
+    \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* 
+    %
+    %
+    = \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1},
+  \end{split}
+\end{align*}  
+\end{lemma}
+\begin{proof}
+  The proof is boring and amounts to using Woodbury's matrix identity
+  twice and a regularization trick. The standard proof for this
+  identity works in infinite dimensions, as long as all terms are well
+  defined. Unfortunately, $\obs^*\obs$ is not invertible, so we force
+  it to be. Throughout, we let $X := \fwd\prcov \fwd^*$ and $Y_{\eps}
+  := \obs^*\obs + \eps I$, for some small $\eps > 0$. Note that now
+  $Y_{\eps}$ is invertible. Recall Woodbury's matrix identity:
+  $$
+  (A + UCV)^{-1} = A^{-1} + A^{-1}U(C^{-1} + VA^{-1}U)^{-1}VA^{-1}.
+  $$
+
+  Taking $A := \prcov^{-1}, U := \fwd^*, V := \fwd$ and $C :=
+  \sigma^{-2} (\obs^*\obs+\eps I)$ we get:
+  \begin{align*}
+    \begin{split}
+      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* &=
+      \fwd ( \prcov - \prcov \fwd^* ( \sigma^2(\obs^*\obs + \eps I)^{-1} + \fwd \prcov \fwd^* )^{-1} \fwd \prcov ) \fwd^* \\
+      %
+      %
+      &= X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X
+    \end{split}
+  \end{align*}
+  Note that $X + \sigma^2 Y_{\eps}^{-1}$ is invertible, as the sum of
+  a two positive definite operators. Now, taking $A := X^{-1}, C :=
+  \sigma^2Y_{\eps}^{-1}, U := I$ and $V := I$ and using Wodbury's matrix
+  identity in reverse order:
+  \begin{align*}
+    \begin{split}
+      X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X &= (X^{-1} + \sigma^{-2}Y_{\eps})^{-1} \\
+      %
+      %
+      %
+      &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \eps I))^{-1}.
+    \end{split}
+  \end{align*}
+
+  We conclude that $\forall \eps > 0$
+  \begin{align*}
+    \begin{split}
+      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* 
+     &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \eps I))^{-1}.
+    \end{split}
+  \end{align*}
+  Letting $\eps \to 0$ completes the proof.
+\end{proof}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{lemma}[Increase due to a Measurement]\label{lemma:design increase}
+  Let $\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm := (\meas_1,\dots,\meas_{m-1})^t$. Then
+  \begin{align*}
+    \tar( \obs ) - \tar (\obsm ) &=
+    \frac12 \log \left ( 1 + \frac{
+      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m,
+      (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m \rangle
+    }{
+      \sigma^2 + \meas_m \modcov \meas_m - \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m 
+    }       
+    \right ).
+  \end{align*}
+\end{lemma}
+\begin{proof}
+  We use the Schur complement to write one inverse in terms of the other and
+  introduce notations to make the derivation cleaner. Note that we think of
+  $\obsm$ and $\obsm^*$ as column and row vectors (respectively).
+  \begin{align*}
+    \Sigma( \obs ) &= \Sigma = 
+    \begin{bmatrix}
+      \Sigma (\obsm )           & \obsm \modcov \meas_m \\
+      \meas_m \modcov \obsm^*   & \sigma^2 + \meas_m \modcov \meas_m
+    \end{bmatrix}
+    =
+    \begin{bmatrix}
+      \Sigmam   & w \\
+      w^t       & c
+    \end{bmatrix}\\
+    %
+    %
+    %
+    \Sigma^{-1} &=
+    \begin{bmatrix}
+      \Sigmam^{-1} + \Sigmam^{-1} w ( c - w^t \Sigmam^{-1} w)^{-1} w^t \Sigmam^{-1} & - \Sigmam^{-1} w ( c - w^t \Sigmam^{-1} w)^{-1} \\
+      -( c - w^t \Sigmam^{-1} w)^{-1} w^t \Sigmam^{-1}                            &  ( c - w^t \Sigmam^{-1} w)^{-1}
+    \end{bmatrix} \\
+    &=
+    \begin{bmatrix}
+      \Sigmam^{-1} & 0 \\
+      0           & 0 
+    \end{bmatrix}
+    + (c -w^t \Sigmam^{-1} w )^{-1}
+    \begin{bmatrix}
+      \Sigmam^{-1} w \\
+      -1
+    \end{bmatrix}
+    \begin{bmatrix}
+      w^t \Sigmam^{-1} & -1 
+    \end{bmatrix}
+  \end{align*}
+  %
+  Further, define
+  %
+  \begin{align*}
+    \M (\obs ):&= \prcov^{\frac12}\fwd^* \obs^* \Sigma^{-1} \obs \fwd
+    \prcov^{\frac12}    
+  \end{align*}
+  %
+  and note that, using our understanding of what is a column vector and
+  what is a row vector:
+  %
+  \begin{align*}
+    \M(\obs) &= \prcov^{1/2} \fwd^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2} \\
+    %
+    %
+    %
+    &= \prcov^{1/2} \fwd^* \obs^* \left \{
+    \begin{bmatrix}
+      \Sigmam^{-1} & 0 \\
+      0           & 0 
+    \end{bmatrix}
+    + (c -w^t \Sigmam^{-1} w )^{-1}
+    \begin{bmatrix}
+      \Sigmam^{-1} w \\
+      -1
+    \end{bmatrix}
+    \begin{bmatrix}
+      w^t \Sigmam^{-1} & -1 
+    \end{bmatrix} 
+    \right \} \obs \fwd \prcov^{1/2} \\
+    %
+    %
+    %
+    &= \M (\obsm) + (c -w^t \Sigmam^{-1} w )^{-1}
+    \prcov^{1/2} \fwd^* \obs^*
+    \begin{bmatrix}
+      \Sigmam^{-1} w \\
+      -1
+    \end{bmatrix}
+    \begin{bmatrix}
+      w^t \Sigmam^{-1} & -1 
+    \end{bmatrix} 
+    \obs \fwd \prcov^{1/2}
+  \end{align*}
+  %
+  Now, denote:
+  %
+  \begin{align}\label{eq:u}
+    \begin{split}
+      u :&= (c -w^t \Sigmam^{-1} w )^{-1/2}
+      \prcov^{1/2} \fwd^* \obs^* 
+      \begin{bmatrix}
+        \Sigmam^{-1} w \\
+        -1 
+      \end{bmatrix} \\
+      %
+      %
+      %
+      & = (c -w^t \Sigmam^{-1} w )^{-1/2} ( \prcov^{1/2}\fwd^* \obsm^* \Sigmam^{-1} \obsm  \modcov \meas_m - \prcov^{1/2} \fwd^* \meas_m )\\
+      %
+      %
+      %
+      u^* :&=  (c -w^t \Sigmam^{-1} w )^{-1/2} (\meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \fwd \prcov^{1/2} - \meas_m \fwd \prcov^{1/2} ),
+    \end{split}
+  \end{align}
+  %
+  so that
+  %
+  \begin{equation}\label{eq:M plus I}
+    I + \M( \obs ) = I + \M (\obsm ) + uu^*.
+  \end{equation}
+  %
+  Note that
+  \begin{equation}\label{eq:M postcov}
+    \prcov^{1/2} \left (I + \M( \obsm ) \right )^{-1} \prcov^{1/2} = \postcovm.
+  \end{equation}
+  The increase in the design criterion gained by including $\meas_m$
+  in the design is found by using Lemma \ref{lemma:MDL} the above
+  results:
+  %
+  \begin{align*}
+    \tar( \obs ) - \tar( \obsm )
+    %
+    %
+    %
+    &= \frac12 \log \det \Big ( I + \M ( \obs ) \Big ) / \det \Big ( I + \M (\obsm) \Big ) \\
+    %
+    %
+    %
+    &= \frac12  \log \det \left ( I + \M(\obsm) + uu^* \right ) / \det \Big ( I + \M (\obsm) \Big ) \\
+    %
+    %
+    %
+    &= \frac12 \log \left ( 1 + \left \langle \left ( I+\M(\obsm) \right )^{-1} u, u  \right \rangle \right ).
+  \end{align*}
+  Using \eqref{eq:u}:
+  \begin{align*}
+    &\left \langle \left (I+\M (\obsm)\right )^{-1}u, u \right \rangle\\
+    &= \frac{
+      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m,
+      (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m \rangle
+    }{
+      c- w^t \Sigmam^{-1} w
+    }\\
+    %
+    %
+    %
+    &= 
+    \frac{
+      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m,
+      (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m \rangle
+    }{
+      \sigma^2 + \meas_m \modcov \meas_m - \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m 
+    }
+  \end{align*}
+  and the conclusion follows.
+\end{proof}
+
+Lemma \ref{lemma:design increase} implies the following two corollaries.
+\begin{corollary}[Gain for No Model Error]\label{cor:zero mod err}
+  If $\modcov = 0$, then
+  \begin{equation*}
+    \tar( \obs ) - \tar (\obsm )
+    = -\frac12 \log (1 - \sigma^{-2} \langle \fwd \postcov \fwd^* \meas_m, \meas_m \rangle ).
+  \end{equation*}
+\end{corollary}
+\begin{proof}
+  Note that this is not immediate by substituting $\modcov = 0$ in the
+  conclusion of Lemma \ref{lemma:design increase}, since we make a
+  claim for $\postcov$, and not $\postcovm$. Let us first review
+  \eqref{eq:u} and note that since $\modcov = 0$ the covariance
+  $\Sigmam = \sigma^2I_{m-1}$ and $w = 0$, so $c - w^t\Sigmam^{-1}w =
+  \sigma^2$:
+  \begin{align*}
+    u :&=-\sigma^{-1}\prcov^{1/2} \fwd^* \meas_m\\
+    %
+    %
+    u^* :&= -\sigma^{-1} \meas_m \fwd \prcov^{1/2}.
+  \end{align*}
+  From \eqref{eq:M plus I}:
+  \begin{equation*}
+    I + \M(\obsm) = I +\M(\obs) - uu^*,
+  \end{equation*}
+  and thus:
+  \begin{equation*}
+    \left \langle \left ( I +\M(\obs) \right )^{-1} u, u \right \rangle
+    = \sigma^{-2} \langle \fwd \postcov \fwd^* \meas_m, \meas_m \rangle.
+  \end{equation*}
+  Analogously to \eqref{eq:M postcov} we note that
+  \begin{equation*}
+    \prcov^{1/2} \left ( I + \M(\obs) \right )^{-1} \prcov^{1/2} = \postcov.
+  \end{equation*}
+  Using Lemma \ref{lemma:MDL} we conclude
+  \begin{align*}
+    \tar( \obs ) - \tar( \obs )
+    &= \frac12 \log \det \left (I +\M(\obs) \right ) / \det \left (I + \M(\obsm) \right ) \\
+    %
+    %
+    %
+    &= \frac12 \log \det \left (I +\M(\obs) \right ) / \det \left (I + \M(\obs) - uu^* \right ) \\
+    %
+    %
+    %
+    &=-\frac12 \log (1 - \langle (I+\M(\obs))^{-1}u, u \rangle \\
+    %
+    %
+    %
+    &= -\frac12 \log (1 - \sigma^{-2} \langle \fwd \postcov \fwd^* \meas_m, \meas_m \rangle ).
+  \end{align*}
+\end{proof}
+
+\samemeas
+\begin{proof} \label{cor:same meas proof}
+  Denote $A:= \obs \modcov \obs^*$ and $v_j$ the $j$th column of $A$.
+  Note that $v_j = \obsm \modcov \meas_m$, since $(\obsm \modcov
+  \obsm^*)_{ij} = \meas_i(\modcov \meas_j)$, as explained in
+  \eqref{eq:modcov explained}. One can now verify that
+  \begin{equation}\label{eq:observation}
+    \Sigmam^{-1} \obsm \modcov \meas_m = \Sigmam^{-1}v_j = (A +\sigma^2I_{m-1})^{-1} v_j =
+    e_j -\sigma^2 \Sigmam^{-1}e_j.
+  \end{equation}
+  %
+  Using \eqref{eq:observation}:
+  \begin{align}\label{eq:denominator}
+    \begin{split}
+      \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m
+      &= \meas_m \modcov \obsm^* ( e^j - \sigma^2 \Sigmam^{-1} e_j )\\
+      %
+      %
+      %
+      &= \meas_m \modcov \meas_j - \sigma^2 \meas_m \modcov \obsm^* \Sigmam^{-1}e_j \\
+      %
+      %
+      %
+      &= \meas_m \modcov \meas_j -\sigma^2 (e_j - \sigma^2 \Sigmam^{-1}e_j)^t e_j \\
+      %
+      %
+      %
+      &= \meas_m \modcov \meas_m -\sigma^2 + \sigma^4 e_j^t\Sigmam^{-1}e_j.
+    \end{split}
+  \end{align}
+  We use \eqref{eq:observation} to simplify the terms in the enumerator of
+  the conclusion of Lemma \ref{lemma:design increase}:
+  \begin{align}\label{eq:enumerator}
+    \begin{split}
+      (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m
+      &= \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m - \meas_m \\
+      %
+      %
+      %
+      &= \obsm^* (e_j - \sigma^2 \Sigmam^{-1} e_j) -\meas_j \\ 
+      %
+      %
+      %
+      &= -\sigma^2 \obsm^* \Sigma^{-1}e_j. 
+    \end{split}
+  \end{align}
+  %
+  Substitute \eqref{eq:enumerator} and \eqref{eq:denominator} to
+  the enumerator and denominator (respectively) of the conclusion of
+  Lemma \ref{lemma:design increase}:
+  %
+  \begin{align*}
+    \tar( \obs ) - \tar (\obsm ) &=
+    \log \left ( 1 + \frac{
+      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m,
+      (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m \rangle
+    }{
+      \sigma^2 + \meas_m \modcov \meas_m - \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m 
+    }       
+    \right ) \\
+    %
+    %
+    %
+    &= \log \left ( 1 + \frac{\sigma^4
+      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
+      \obsm^* \Sigmam^{-1}e_j \rangle
+    }{
+      2\sigma^2 - \sigma^4 e_j^t\Sigmam^{-1}e_j 
+    }       
+    \right ) \\
+    %
+    %
+    %
+    &= \log \left ( 1 + \frac{\sigma^2
+      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
+      \obsm^* \Sigmam^{-1}e_j \rangle
+    }{
+      2 - \sigma^2 e_j^t\Sigmam^{-1}e_j 
+    }       
+    \right ).
+  \end{align*}
+\end{proof}
+
+\begin{lemma}[Auxilliary Calculations]\label{lemma:aux calc}
+  Let $T(\obs) := \obs^* \Sigma^{-1}(\obs)\obs$, with $\Sigma(\obs)$
+  defined as in \eqref{eq:Sigma}. Then:
+  \begin{align*}
+    \delta T(\obs)V &= V^* \Sigma^{-1} \obs 
+    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
+    &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs
+    + \obs^* \Sigma^{-1} V.
+  \end{align*}
+\end{lemma}
+
+\begin{proof}
+  We need a few supplementary calculations. First:
+  \begin{align}\label{eq:der sig}
+    \begin{split}
+      \frac{\der}{\der \tau} \Big |_{\tau=0} \Sigma( \obs + \tau V )
+      &= \frac{\der}{\der \tau} \Big |_{\tau=0} 
+      (\obs + \tau V ) \modcov (\obs + \tau V )^*  + \sigma^2I\\
+      % 
+      % 
+      % 
+      &= V \modcov \obs^* + \obs \modcov V^*.
+    \end{split}
+  \end{align}
+  By the standard trick for the derivative of an operator: 
+  \begin{align*}
+    0 &= \frac{\der}{\der \tau} \Big |_{\tau=0} I \\
+    % 
+    % 
+    % 
+    &= \frac{\der}{\der \tau} \Big |_{\tau=0}
+    \left (\Sigma(\obs+\tau V)^{-1} \Sigma(\obs+\tau V) \right ) \\
+    % 
+    % 
+    % 
+    &= \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0} \Sigma+
+    \Sigma^{-1} \frac{\der \Sigma(\obs+\tau V)}{\der \tau} \Big |_{\tau=0}\\  
+    %
+    %
+    %
+    &= \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0} \Sigma+
+    \Sigma^{-1} (V\modcov \obs^* + \obs \modcov V^*) 
+    \text{, by \eqref{eq:der sig}. }
+  \end{align*}
+  Now we can write the variation of $\Sigma^{-1}$:
+  \begin{align}\label{eq:der sig inv}
+    \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0}  
+      &= -\Sigma^{-1} (V \modcov \obs^* + \obs \modcov V^*) \Sigma^{-1}.
+    \end{align}
+  % 
+  Finally, we can do the calculation we need to. By Leibnitz (product) rule
+  and \eqref{eq:der sig inv}:
+  % 
+  \begin{align*}
+    \delta T(\obs) V 
+    &= \frac{\der T(\obs + \tau V)}{\der \tau} \Big |_{\tau=0} \\
+    %
+    %
+    %
+    &= V^* \Sigma^{-1} \obs 
+    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
+    &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs
+    + \obs^* \Sigma^{-1} V,
+  \end{align*}
+  as required.
 \end{proof}

### 2020-10-22 reply.tex
getting closer

@@
 on every point made. The referees' comments and my replies are listed
-below. Below, my replies are colored \ydnote{cyan}. In the revised
-manuscript, replies to referee two are in \reftwo{green} and replies
-to referee three are in \refthree{red}. Again, thank you for taking
-the time and helping me improve this manuscript.
+below. My replies are colored \textcolor{cyan}{cyan}. Again, thank you
+for taking the time and helping me improve this manuscript.
 
 \section{Referee 1}
-Preliminary Assessment, Editorial Board Membe.r The manuscript can be
+Preliminary Assessment, Editorial Board Member The manuscript can be
 sent to reviewers.
@@
 \begin{enumerate}
-\item It should be stated in the Abstract whether the work in this
+\item It should be stated in the abstract whether the work in this
   paper pertains to only linear inverse problems, or if it is
   applicable in general to non-linear inverse problems as well.
-  \ydnote{\item Title and abstract now reflect the fact that the paper only
+  \ydnote{Title and abstract now reflect the fact that the paper only
     pertains to linear inverse problems.}
@@
 explicitly and emphasized clearly as the contribution of the article.
+\ydnote{The contribution section was extended.}
 
@@
 using them.
-\ydnote{\item Corrected, thanks!}
+\ydnote{Corrected, thanks!}
 
@@
 a brief summary of how this is being accomplished.
-\ydnote{\item Added section overviews and operator definitions.}
+\ydnote{Added section overviews and operator definitions.}
 
@@
 I don't see the benefit of consider infinite-dimensional formulation.
-
 \ydnote{I was not able to arrive at the same results when the forward
@@
   any of the scenarios...) is rather strong and unjustified.
+  \ydnote{Said statement was removed.}
 \end{enumerate}
@@
 however, the following argument on Page 9 that extends this assumption
-is not well-justified.
-\ydnote{This is a great point! The flawed explanation was removed. I
-  was able to remove that assumption.}
+is not well-justified.  \ydnote{This is a great point! The flawed
+  explanation was removed. I was able to remove that assumption. But
+  it required further changes to the manuscript.}
 
@@
 be clearly discussed for the reader to appreciate the valuable effort
-made.
-\ydnote{Again, great point. $\fwd^{-*} = (\fwd^*)^{-1}$ but now I do
-  not assume $\fwd$ is invertible so this is removed}
+made. \ydnote{Again, great point. $\fwd^{-*} = (\fwd^*)^{-1}$ but now
+  I do not assume $\fwd$ is invertible so this was removed.}
 
@@
 clarify several confusions that a reader can fall for easily.
+\ydnote{I added several plots that I believe make the arguments and
+  conclusions clearer. I also added an illustration story with
+  graduated lab cylinders.}
 
@@
 abstract, in addition to the fact that observations errors are
-uncorrelated with fixed variance $\sigma^2$.
-\ydnote{The title and
-  abstract were changed to reflect the fact that linear problems are concerned.
-The treatment of correlated errors was emphasized.}
+uncorrelated with fixed variance $\sigma^2$.  \ydnote{The title and
+  abstract were changed to reflect the fact that linear problems are
+  concerned.  The treatment of correlated errors was emphasized.}
 
@@
 clusterization will be considered later, but I failed to find any
-relevant discussion to that issue which I find interesting. Even if
-the model error covariance $\Gamma_{model}$ is known, this doesn't
-amount for observation correlations, .e.g,
+relevant discussion to that issue which I find interesting.
+\ydnote{The discussion on observation correlation
+  and how it mitigates clusterization can be found in the last
+  section.}
+
+Even if the model error covariance $\Gamma_{model}$ is known, this
+doesn't amount for observation correlations, .e.g,
 single-instrument-observations, because the observation operator
 $\mathcal{O}$ is non invertible. This issue should addressed and
-clarified clearly.
+clarified clearly.  \ydnote{I am not sure I understand this
+  point. $\modcov$ can model anything you want, including correlated
+  observation instruments. It is true that $\obs$ is not invertible
+  but I fail to understand why this is problematic.}
 
@@
 not works properly.  For example Corollary 2 should be Corollary B.x,
-etc.
-\ydnote{All such bad references were fixed.}
+etc.  \ydnote{All such bad references and links were fixed.}
 
@@
 and caption of Figure 2 should be rewritten with clear explanation of
-the axes, and the plot-generation process.
+the axes, and the plot-generation process.  \ydnote{I extended the
+  discussion relating to clusterization, added another plot and
+  explained how they arise.}
 
-One page 10, lines 30-35, the argument doesn't seem to be quite
+On page 10, lines 30-35, the argument doesn't seem to be quite
 accurate. It's not always true that one associates a higher
@@
 lower uncertainty, then, variable observation variances must be
-considered. On the other hand, if being ``better'' corresponds to a
-more accurate realization of the observation (true snapshot of
-reality), then the whole statement would be wrong, because optimal
-design relies only on the uncertainty levels in linear Bayesian
-settings, as manifested by the posterior covariance matrix.
+considered. \ydnote{Unfortunately, I am again not sure I understand
+  fully. Variable observation variances are fixed to $\sigma^2$
+  througout this section.}
+
+On the other hand, if being ``better'' corresponds to a more accurate
+realization of the observation (true snapshot of reality), then the
+whole statement would be wrong, because optimal design relies only on
+the uncertainty levels in linear Bayesian settings, as manifested by
+the posterior covariance matrix.  \ydnote{I think the referee implies
+  that the D-optimal design is only as good as the assumptions it
+  relies on. I could not agree more. However, as long as we do believe
+  the assumptions made --- a D-optimal design is indeed optimal (in a
+  precise and quantifiable way). I think the source of
+  misunderstanding is that I used the term ``measurement'' where I
+  should have used ``eigenvector''. I changed the paragraph
+  accordingly. The transition from thinking of ``measurements'' to
+  thinking of ``eigenvectors'' is confusing and I believe it is better
+  explained in the updated version of the manusctipt.}
 
@@
 $X(t=0)$ explicitly (see line 32 on page 11). A clarification is
-required here.
+required here.  \ydnote{Clarification was added at the end of the
+  proof. ``There is nothing special about $t=0$'' means we could have
+  chosen any other value $t_0$ instead.}
 
@@
 referring to.
+\ydnote{Made explicit.}
 
-Page 2, line 11; wrong reference to Figure 1.3?
+Page 2, line 11; wrong reference to Figure 1.3?  \ydnote{Reference now
+  fixed.}
 
@@
 to achieve this result? Please clarify.
+\ydnote{The expectation / integration over the prior space happens
+  inside the KL divergence:
+  $$
+  \mathbb{E}_{\data}[KL(\post^\data||\pr)] = \int KL(\post^\data||\pr) \der \mu(\data) = \int \int \log \frac{\der \post^\data}{\der \pr}(x) \der \post(x) \der \mu(\data)
+  $$ where $\mu(\data)$ is the distribution of data (usually called
+  evidence). I improved the notation to reflect the dependence of the
+  posterior on the data $\data$. I added above definition of KL
+  divergence to the manuscript.}
 
 Page 6, line 13; the optimization problem is solved for $\mathcal{O}$,
-not $\Psi$ ?
+not $\Psi$ ?  \ydnote{This line was unclear. I extended the
+  explanation. The mauscript now reflects the claim that the optimal
+  $\obs^{\star}$ is the maximizer of $\tar(\obs)$.}
 
 Page 6, lines 19; the statement made about relaxing the constraints on
-the observation operator is unclear.
+the observation operator is unclear.  \ydnote{This sentence was poorly
+  written. It is now rephrased.}
 
 Page 6, line 24, I believe the reference to (3.2) should be replaced
-with (3.1).
+with (3.1). \ydnote{This was actually OK. The gradient was recorded in
+  (3.2).}
 
@@
 clearly.
+\ydnote{Fixed.}
 
@@
 A.1). Maybe state that clearly here. Also, in the proof, from step 4
-to 5, the first and third terms vanished; why?
+to 5, the first and third terms vanished; why?  \ydnote{This point is
+  spot on. I added the definition of $T$ before the derivation. This
+  is not the proof of Lemma 3.1 --- it is the first variation of the
+  objective. I reworded a few sentences to make this clearer.}
 
@@
 Proposition 3.2 or something similar to match previous referencing
-(e.g., Lemma 3.1).
+(e.g., Lemma 3.1). \ydnote{Fixed.}
 
@@
 seems to be incorrect. Also, state clearly which appendix the proof is
-delegated to.
+delegated to.  \ydnote{Fixed.}
 
@@
 $\widehat{\Gamma_{post}}$ int the corollary, which is defined only in
-the appendix, under the proof.
+the appendix, under the proof. \ydnote{All comments on numbering and
+  references are great and correct. All were fixed. I added an
+  explanation of what $\widehat{\Gamma_{post}}$ means where it
+  appears.}
 
@@
 derivative vanished from step 2 to step 3; please explain!
+\ydnote{I added an extra equality line for explanation.}
 

### 2020-10-22 compiler.sh

@@ -1,0 +1,8 @@
+#!/bin/bash
+latexdiff --config="PICTUREENV=(?:picture|DIFnomarkup|figure|)[\w\d*@]*" paper.tex new_paper.tex > diff.tex
+#latexdiff --exclude-textcmd=”section,subsection,subsubsection” --config=”PICTUREENV=(?:picture|DIFnomarkup|align|tabular)[\w\d*@]*” paper.tex new_paper.tex > diff.tex
+#latexdiff paper.tex new_paper.tex > diff.tex
+latexmk -pdf diff
+#latexmk -pdf paper
+latexmk -pdf new_paper
+latexmk -pdf reply

### 2020-10-22 differ.sh

@@ -1,8 +1,0 @@
-#!/bin/bash
-#latexdiff --config="PICTUREENV=(?:picture|DIFnomarkup|align|tabular)[\w\d*@]*" paper.tex new_paper.tex > diff.tex
-# latexdiff --exclude-textcmd=”section,subsection,subsubsection” --config=”PICTUREENV=(?:picture|DIFnomarkup|align|tabular)[\w\d*@]*” paper.tex new_paper.tex > diff.tex
-# latexdiff paper.tex new_paper.tex > diff.tex
-#latexmk -pdf -f diff
-latexmk -pdf paper
-latexmk -pdf new_paper
-latexmk -pdf reply

### 2020-10-25 new_paper.tex
revised

@@
   dimensions]{Sensor Clusterization in D-optimal Design in Infinite
-  Dimensional Linear Bayesian Inverse Problems}
+  Dimensional Bayesian Linear Inverse Problems}
 
@@
   We investigate the problem of sensor clusterization in optimal
-  experimental design for infinite-dimensional Bayesian
-  linear inverse problems. We suggest an analytically
-  tractable model for such designs and reason how it may lead to
-  sensor clusterization in the case of iid measurement noise. We also
-  show that in the case of spatially correlated measurement error
-  clusterization does not occur. As a part of the analysis we prove a
-  matrix determinant lemma analog in infinite dimensions, as well as a
-  lemma for calculating derivatives of $\log \det$ of operators.
+  experimental design for infinite-dimensional Bayesian linear inverse
+  problems. We suggest an analytically tractable model for such
+  designs and reason how it may lead to sensor clusterization in the
+  case of iid measurement noise. We also show that in the case of
+  spatially correlated measurement error clusterization does not
+  occur. As a part of the analysis we prove a matrix determinant lemma
+  analog in infinite dimensions and a lemma for calculating
+  $\frac{\der \log \det (I + X(t))}{\der t}$ for certain
+  operator-valued functions. We also show how to decompose $M = AA^t$
+  with $A$ constrained to have unit norm columns.
 \end{abstract}
@@
 \cite{KullbackLeibler51,CoverThomas12}) between posterior and
-prior. Consider a linear model in finite dimensions, with Gaussian
-prior and noise. In this model, maximizing the expected information
-gain amounts to minimizing the determinant of the posterior covariance
-matrix. In a frequentist setting, a D-optimal design minimizes the
-volume of the uncertainty ellipsoid \cite[page 16]{Ucinski05}, but
-this is done for the Fisher information matrix and not the posterior
-covariance. However, \cite{ChalonerVerdinelli95} show that the latter
-is just a regularized version of the former.
+prior. For example, consider a linear model in finite dimensions, with
+Gaussian prior and noise. In this model, maximizing the expected
+information gain amounts to minimizing the determinant of the
+posterior covariance matrix. In a frequentist setting, a D-optimal
+design minimizes the volume of the uncertainty ellipsoid \cite[page
+  16]{Ucinski05}, but this is done for the Fisher information matrix
+and not the posterior covariance. However, \cite{ChalonerVerdinelli95}
+show that the latter is just a regularized version of the former.
 
@@
 It seems counter intuitive that when one computes an optimal design
-using the D-optimal criterion (and others that will not be
-investigated here), the optimization process results in measurements
-that are very similar. For example, if a measurement is thought of as
-measuring some function value at $\x \in \Omega \subseteq \R^d,
-d=1,2,3$ (with added error) then the optimization procedure sometimes
-places sensors in very close proximity to each other (as can be seen
-in figure \ref{fig:clusterization illustration}). Following
-\cite{Ucinski05}, we refer to this phenomenon as \emph{sensor
-  clusterization}.
+using the D-optimal criterion, the optimization process results in
+measurements that are very similar. For example, if a measurement is
+thought of as measuring some function value at $\x \in \Omega
+\subseteq \R^d, d=1,2,3$ (with added error) then the optimization
+procedure sometimes places sensors in very close proximity to each
+other (as can be seen in figure \ref{fig:clusterization
+  illustration}). Following \cite{Ucinski05}, we refer to this
+phenomenon as \emph{sensor clusterization}.
 
@@
 \subsection{Contribution}
-We propose and thoroughly study an analytically tractable model for
-understanding D-optimal designs. In this model, D-optimal design
-are solutions of a constrained optimization problem, formulated
+We propose and thoroughly study a relaxed and analytically tractable
+model for understanding D-optimal designs. In this model, D-optimal
+design are solutions of a constrained optimization problem, formulated
 using Lagrange multipliers (section \ref{section:D and grad}). Using
@@
 sought is a sum of two operators. The first is the prior
-covariance. The second is an outer product of the observations. See
-\eqref{eq:eigenproblem} for details and exact statement. This
+covariance. The second is an outer product of the observations (see
+\eqref{eq:eigenproblem} for details and exact statement). This
 structure helps us shed light into D-optimal designs. We characterize
@@
 present the problem via a toy model --- the 1D heat equation in
-$[0,\pi]$ with homogeneous Dirichlet boundary conditions.
+$[0,\pi]$ with a homogeneous Dirichlet boundary conditions.
 
@@
 We would like to infer the initial condition $u_0$. For that purpose,
-we measure $u$ at some set of locations $\x_j \in [0,\pi], j=1, \dots,m$
-and a final time $T > 0$. We assume centered Gaussian measurement
-error, so we can observe $v(\x_j,T) = u(\x_j,T) + \eps(\x_j)$ with
-$\eps(\x_j) \sim \normal(0, \sigma^2), \sigma > 0$ iid. We model the
-initial condition as $u_0 \sim \normal(0,\prcov)$, for $\prcov =
-(-\Delta)^{-1}$ with homogeneous Dirichlet boundary condition. It is
-well known \cite{Tarantola05} that for linear problems, with Gaussian
-prior and error, the posterior is also Gaussian with a covariance that
-does not depend on the observed data. The posterior covariance
-$\postcov$ is known to have a closed form formula, even in infinite
-dimensions\cite{Stuart10}. If we denote by $\fwd$ the dynamics
-operator, so that $u( \cdot,T) = \fwd u_0$, and the observation
-operator $\obs$ so that $u(\x_j,T) = (\obs u)_j, j=1,\dots,m$, then the
-posterior covariance is known and depends only on $\prcov, \fwd, \obs$ and
-$\sigma^2$ (see section \ref{section:prelim} and
-\eqref{eq:postcov} specifically).
+we measure $u$ at some set of locations $\x_j \in [0,\pi], j=1,
+\dots,m$ and a final time $T > 0$. We assume centered Gaussian
+measurement error, so we can observe $v(\x_j,T) = u(\x_j,T) +
+\eps(\x_j)$ with $\eps(\x_j) \sim \normal(0, \sigma^2), \sigma > 0$
+iid. We model the initial condition as $u_0 \sim \normal(0,\prcov)$,
+for $\prcov = (-\Delta)^{-1}$ with a homogeneous Dirichlet boundary
+condition. It is well known \cite{Tarantola05} that for linear
+problems, with Gaussian prior and error, the posterior is also
+Gaussian with a covariance that does not depend on the observed
+data. The posterior covariance $\postcov$ is known to have a closed
+form formula, even in infinite dimensions\cite{Stuart10}. We denote by
+$\fwd$ the dynamics operator, so that $u( \cdot,T) = \fwd u_0$, and
+the observation operator $\obs$ so that $u(\x_j,T) = (\obs u)_j,
+j=1,\dots,m$. The posterior covariance is known and depends only on
+$\prcov, \fwd, \obs$ and $\sigma^2$ (see section \ref{section:prelim}
+and \eqref{eq:postcov} specifically).
 
@@
 $\x_j,j=1,\dots,m$ to minimize (an expression analogous to) the
-log-determinant of the posterior covariance operator. We will see
-later how this corresponds to maximizing expected information gain.
+determinant of the posterior covariance operator. We will see later
+how this corresponds to maximizing expected information gain.
 
@@
 \ref{fig:clusterization illustration}. Posterior pointwise standard
-deviations are plotted over the domain $[0, \pi]$. For a Bayesian
-linear inverse problem with Gaussian prior and noise the posterior
-covariance does not depend on data \cite{Stuart10}. Thus, the plot has
-no reference to actual data observed. The posterior covariance does,
-however, depend on the measurement taken. In figure
+deviations are plotted over the domain $[0, \pi]$. Since the posterior
+covariance does not depend on data, the plot has no reference to
+actual data observed. The posterior covariance does, however, depend
+on location of the measurement taken. In figure
 \ref{fig:clusterization illustration}, measurement locations are
 marked by circles. These were chosen to minimize (an expression
-analogous to) the log-determinant of the posterior covariance. The
+analogous to) the determinant of the posterior covariance. The
 clusterization effect can be clearly seen for six sensors. It looks
@@
 can be found in \cite{Stuart10}.
-%% A reader mostly familiar with the
-%% literature may initially want to skip this section and only consult
-%% the list below, which summarizes the notation used. We denote:
-%% \begin{itemize}
-%%   \item Parameter space $\hilp$.
-%%   \item Data space $\hilo$.
-%%   \item Forward dynamics operator $\fwd: \hilp \to \hilo$.
-%%   \item Observation operator $\obs: \hilo \to \mathbb{R}^m$.
-%%   \item Prior measure $\pr = \mathcal{N}(0, \prcov)$.
-%%   \item Data $\data \in \mathbb{R}^m$.
-%%   \item Observation and model errors, captured in $\Sigma(\obs)$.
-%% \end{itemize}
-%% The design objective was derived in \cite{AlexanderianGloorGhattas14}
-%% and recorded in \eqref{eq:objective} below.
 
 \subsection{Bayesian Linear Inverse Problems}\label{subsec:abstract OED}
-Let $\hilp$ and $\hilo$ be seperable Hilbert spaces (the subscripts p
+Let $\hilp$ and $\hilo$ be separable Hilbert spaces (the subscripts p
 and o are for ``parameter'' and ``observation'', respectfully). We
@@
 infer $\param \in \hilp$ --- some parameter of the dynamics --- given
-noisy observations of $\fwd \param$. We take a zero mean Gaussian
-prior $\pr$ for $\param$ (otherwise the non-zero mean can just be
-subtracted everywhere). Hence, $\param \sim \pr = \normal(0,\prcov)$
+noisy observations of $\fwd \param$. We take a Gaussian prior $\pr$
+for $\param$. Hence, $\param \sim \pr = \normal(\param_0 ,\prcov)$
 with some appropriate covariance operator $\prcov$ \cite{Stuart10}. It
@@
 in $\hilo$ \cite{Stuart10}. As such, it is assumed invertible --- an
-assumption we rely heavily upon. Throughout this paper, we denote $m$
-the number of measurements we are allowed to take. Observations
+assumption we rely on below. Throughout this paper, we denote $m$ the
+number of measurements we are allowed to take. Observations
 (measurements) are taken via the linear observation operator $\obs \in
@@
 
-At this point, it is useful to record a known result regarding the
-posterior covariance operator of our inverse problem, namely
-\begin{align}\label{eq:postcov}
-  \postcov = (\prcov^{-1} + \fwd^* \obs^* \Sigma^{-1} \obs \fwd
-  )^{-1}.
-\end{align}
 Usually, this is written using $\tmp := \obs \fwd$ and $\tmp^*$. Since
@@
 
+Finally, it is useful to record a known result regarding the posterior
+covariance operator of our inverse problem, namely
+\begin{align}\label{eq:postcov}
+  \postcov = (\prcov^{-1} + \fwd^* \obs^* \Sigma^{-1} \obs \fwd
+  )^{-1}.
+\end{align}
+
 \subsection{D-Optimal Designs in Infinite Dimensions}\label{subsec:D optimal design} 
@@
 $$
-KL(\post^\data||\pr) = \int \log \frac{\der \post^\data}{\der \pr}(x) \der \post(x).
+KL(\post||\pr) = \int \log \frac{\der \post}{\der \pr}(x) \der \post(x).
 $$
@@
 \begin{theorem}[Slightly modified Theorem 1 from \cite{AlexanderianGloorGhattas14}]
-  Let $\pr = \normal(0,\prcov)$ be a centered Gaussian prior
-  on $\hilp$   and let $\post = \normal(\postmean,\postcov)$ 
-  the posterior measure on $\hilo$ for the Bayesian linear
-  inverse problem $\data =  \obs \fwd\param + \obs \eps' + \eps$ discussed
-  above. Then 
+  Let $\pr = \normal(\param_0,\prcov)$ be a Gaussian prior on $\hilp$
+  and let $\post = \normal(\postmean,\postcov)$ the posterior measure
+  on $\hilo$ for the Bayesian linear inverse problem $\data = \obs
+  \fwd\param + \obs \eps' + \eps$ discussed above. Then
   \begin{align}\label{eq:objective}
@@
   \begin{split}
-    ( I + \prcov^{1/2}  \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}) &= \prcov ( \prcov^{-1} + \fwd ^* \obs^* \Sigma^{-1} \obs \fwd) \\
-    &= \prcov \postcov^{-1}.
+    \det ( I + \prcov^{1/2}  \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}) &= \det \Big( \prcov ( \prcov^{-1} + \fwd ^* \obs^* \Sigma^{-1} \obs \fwd) \Big )\\
+    &= \det \prcov \det \postcov^{-1}.
   \end{split}
@@
 may \textit{think} of these as pointwise evaluations, even though
-these might not be in $\hilo$). We call this problem ``simultaneous
+these might not be in $\hilo^*$). We call this problem ``simultaneous
 optimization'', since all measurements are decided on simulatneously.
@@
 \ref{section:objective} we find the gradient of $\tar$ and record it
-in \eqref{eq:tar grad}. In section \ref{subsec:unit length} we relax
-the constraints on $\obs$ so that the optimization problem in
-\eqref{eq:optimization} is analytically tractable.  We find gradients
-for the new constraints and record them in \eqref{eq:grad
+in \eqref{eq:tar grad}. In section \ref{subsec:unit length} we suggest
+using unit-norm constraints on rows of $\obs$ so that the optimization
+problem in \eqref{eq:optimization} is analytically tractable. We find
+gradients for the new constraints and record them in \eqref{eq:grad
   constraints}. Finally, in section \ref{subsec:necessary} we use
@@
 (\obs)$ in the direction $V$ is:
-\begin{align*}%%\label{eq:tar var}
+\begin{align}\label{eq:tar var}
   \begin{split}
@@
     % 
-    &= \frac12 \ttr{\postcov \fwd^* (V^* \Sigma^{-1} \obs 
+    &= \frac12 \ttr\Big \{ \postcov \fwd^* (V^* \Sigma^{-1} \obs 
       - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
       &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs 
-      + \obs^* \Sigma^{-1} V ) \fwd}  \text{ (by \ref{lemma:aux calc})}\\
+      + \obs^* \Sigma^{-1} V ) \fwd \Big \}  \text{ (by \ref{lemma:aux calc})}\\
 %%   \end{split}
@@
     &= \tr{\postcov \fwd^* ( \obs^* \Sigma^{-1} V 
-    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs ) \fwd} \\
+    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs ) \fwd} \text{ (trace rules, $\frac{1}{2}$ prefactor disappears)} \\
     %
@@
   \end{split}
-\end{align*} 
+\end{align} 
 Denote
@@
 \ref{section:clusterization}. We prepare the ground by first deriving
-a ``weird'' eigenvalue problem in $\obs$ for D-optimal designs in
-section \ref{subsec:eigenproblem}. This ``weird'' eigenvalue problem
+a nonlinear eigenvalue problem in $\obs$ for D-optimal designs in
+section \ref{subsec:eigenproblem}. This nonlienar eigenvalue problem
 is recorded in \eqref{eq:mod conditions}. Next, we characterize
 optimal designs as ones that make all eigenvalues equal in the above
-mentioned weird eigenvalue problem. The derivation is presented in
+mentioned nonlinear eigenvalue problem. The derivation is presented in
 section \ref{subsec:characterization}.
 
-\subsection{The ``Weird'' Eigenvalue Problem}\label{subsec:eigenproblem}
+\subsection{The nonlinear Eigenvalue Problem}\label{subsec:eigenproblem}
 $\modcov = 0$ implies $\Sigma= \sigma^2I$. The necessary first-order
@@
 however, that $\postcov$ depends on $\obs$ --- which makes
-\eqref{eq:eigenproblem} a ``weird'' eigenvalue problem.
+\eqref{eq:eigenproblem} a nonlinear eigenvalue problem.
 
@@
 \begin{restatable*}[Simultaneous diagonizability for the nonlinear eigenvalue problem]{lemma}{simdiag}\label{lemma:sim diag}
-  Let $\hil$ seperable Hilbert space, $C:\hil \to \hil$ self-adjoint
+  Let $\hil$ separable Hilbert space, $C:\hil \to \hil$ self-adjoint
   and $\func_1,\dots,\func_m \in \hil$. Denote $\func^*$ the element
@@
 \end{proof}
-Only finitely many of the eigenvectors of $\obs^*\obs$ have positive
-eigenvalue. This happens because $k := \rank \obs^*\obs \leq
-m$. Denote the corresponding non-zero eigenvalues of $\obs^*\obs$ by
+At most $m$ of the eigenvectors of $\obs^*\obs$ have positive
+eigenvalue, because $k := \rank \obs^*\obs \leq m$. Denote the
+corresponding non-zero eigenvalues of $\obs^*\obs$ by
 $\{\eta_i\}_{i=1}^{k}$ and let $\eta_i = 0$ for $i \geq k+1$. No
-assumption is made regarding the magnitudes the corresponding
+assumption is made (yet) regarding the magnitudes of the corresponding
 $\{\lambda_i\}_{i=1}^{\infty}$. They are to be thought of as fixed but
@@
 
-with the norm constraints on $\meas_j,j=1,\dots,m$. This implies that
-$\meas_i \in span\{\ev_j\}_{j=1}^k,i=1,\dots, m$. The problem of
-D-optimal design has now been cast to finite dimensions. This
-justifies using linear algebra. Lemma \ref{lemma:free} (proved in the
-appendix) does exactly that.
+with the norm constraints on $\meas_j,j=1,\dots,m$.
+%%  This implies that $\meas_i \in span\{\ev_j\}_{j=1}^k,i=1,\dots,
+%% m$.
+The problem of D-optimal design has now been cast to finite
+dimensions. This justifies using linear algebra. Lemma
+\ref{lemma:free} (proved in the appendix) does exactly that.
 \begin{restatable*}[Unit norm decomposition]{lemma}{free}\label{lemma:free}
@@
 Take any $\{\eta_i\}_{i=1}^{k}$, with $\sum_{i=1}^k \eta_i = m$. Let
-$M = \diag(\eta_1,\dots,\eta_k)$. The lemma gives us $A$ with unit
-norm columns such that $AA^T = M$. Think of $A^t$ as representing
-$\obs$, so we take $\meas_i :=\sum_{j=1}^k A_{ji}\ev_j^*$. The
-constraints $\|\meas_i\| = 1, i=1,\dots,m$ are satisified by
-construction. Verifying $\obs^*\obs \ev_i = \eta_i \ev_i$ is not too
-difficult:
+$M = \diag(\eta_1,\dots,\eta_k)$. Lemma \ref{lemma:free} gives us $A$
+with unit norm columns such that $AA^T = M$. Take $\meas_i
+:=\sum_{j=1}^k A_{ji}\ev_j^*$. The constraints $\|\meas_i\| = 1,
+i=1,\dots,m$ are satisified by construction. Verifying $\obs^*\obs
+\ev_i = \eta_i \ev_i$ is not too difficult:
 
@@
 
-By concavity of $\log$, the fastest increase in the design criterion
-(equation \eqref{eq:true target}) is gained by increasing
-$\sigma^2\lambda_i^{-1} + \eta_i$ where it is smallest. So, no weight
-should be given to (i.e. measurement taken in direction of) the $n$th
-eigenvector before $\sigma^2\lambda_i^{-1} + \eta_i \geq
-\sigma^2\lambda_n^{-1},\ \forall i \neq n$. Thus, we should choose
-$\{\eta_i\}_{i=1}^k$ as follows. First, order the eigenvalues
-$\lambda_1 \geq \lambda_2 \geq \dots$. We should increase $\sigma^2
-\lambda_1^{-1} + \eta_1$ until it equals
+We have shown that we can realize any $\{\eta_i\}_{i=1}^k$, as long as
+$\sum+{i=1}^k \eta_i = m$. We now turn to understand which choice of
+$\{\eta_i\}_{i=1}^k$ is optimal. By concavity of $\log$, the fastest
+increase in the design criterion (equation \eqref{eq:true target}) is
+gained by increasing $\sigma^2\lambda_i^{-1} + \eta_i$ where it is
+smallest. So, no weight should be given to (i.e. measurement taken in
+direction of) the $n$th eigenvector before $\sigma^2\lambda_i^{-1} +
+\eta_i \geq \sigma^2\lambda_n^{-1},\ \forall i \neq n$. Thus, we
+should choose $\{\eta_i\}_{i=1}^k$ as follows. First, order the
+eigenvalues $\lambda_1 \geq \lambda_2 \geq \dots$. We should increase
+$\sigma^2 \lambda_1^{-1} + \eta_1$ until it equals
 $\sigma^2\lambda_2^{-1}$. Then, $\sigma^2 \lambda_i^{-1} +
@@
 The argument above has the important consequence (illustrated in
-figure \ref{fig:optimal vs not}) that the eigenvalues in the ``weird''
+figure \ref{fig:optimal vs not}) that the eigenvalues in the nonlinear
 eigenvalue problem \eqref{eq:eigenproblem} are all equal! This makes
@@
 One can understand D-optimal designs as follows: Imagine each
-eigenvector is a graduated lab cylinder (figure
+eigenvector corresponds to a graduated lab cylinder (figure
 \ref{fig:cylinder}). Each cylinder $i$ is filled, a-priori, with green
-liquid of $\sigma^2\lambda_i^{-1}$ volume units. This liquid is the
-prior precision, weighted by observation noise. The volume of the
-green liquid is increasing in $i$. A D-optimal design with $m$
-measurements ignores cylinders indexed $m+1$ and above. It has blue
-liquid of $m$ volume units at its disposal. It distributes said liquid
-by repeatedly adding a drop to whatever cylinder currently has the
-lowest level of liquid in it (lowest precision). The result of such a
-procedure is illustrated in figure \ref{fig:optimal vs not}. The
-analysis conducted in this section is summarized in theorem
-\ref{thm:char} below.
-
-\begin{theorem}[Characterizing D-optimal designs]\label{thm:char}
+liquid of $\sigma^2\lambda_i^{-1}$ volume units. This liquid
+represnets the prior precision, weighted by observation noise. The
+volume of the green liquid is increasing in $i$. A D-optimal design
+with $m$ measurements ignores cylinders indexed $m+1$ and above. It
+has blue liquid of $m$ volume units at its disposal. It distributes
+said liquid by repeatedly adding a drop to whatever cylinder currently
+has the lowest level of liquid in it (lowest precision). The result of
+such a procedure is illustrated in figure \ref{fig:optimal vs
+  not}. The analysis conducted in this section is summarized in
+theorem \ref{thm:char} below.
+
+\begin{theorem}[Characterizing D-optimal designs with vanishing model error]\label{thm:char}
   Let:
   \begin{enumerate}
-  \item $\fwd, \prcov, \obs$ the forward, prior covariance and
-    observation operators, respectively.
+  \item $\fwd, \prcov$ the forward and prior covariance operators, respectively.
+  \item $m, \sigma^2$ the number of measurements taken and observation
+    noise variance, respectively.
+  \item $\obs$ a D-optimal design operator.
   \item $\{\lambda_i\}_{i=1}^\infty$ the eigenvalues of
@@
     $\obs$ maximizes $\tar(\obs)$.
-  \item $m$ the number of measurements taken.
-  \item $\sigma^2$ observation noise variance.
   \end{enumerate}
@@
 Again, the first case results in clusterization. The sequential design
-proceeds in this fashion. Every $\meas_i$ is taken in the direction of
-the eigenvector of smallest eigenvalue of the previous posterior
-precision operator. Taking enough measurements we will surely arrive
-at clusterization, since the prior precision increases quickly --- in
-section \ref{subsec:abstract OED} we assumed $\fwd$ is strongly
-smoothing, which means eigenvalues of $(\fwd \prcov\fwd^*)^{-1}$ grow
-quickly.
+proceeds in this fashion. Every $\meas_i$ is taken in the direction
+some eigenvector of the prior $\fwd \prcov\fwd^*$. This eigenvector
+has the smallest eigenvalue of the previous step's posterior precision
+operator. If we do not encounter clusterization after $m$ such steps,
+this means that measurement $\meas_k$ was taken in the direction
+$\ev_k$ for $k=1,\dots,m$. Thus $\sigma^2\lambda_i^{-1} + 1 \geq
+\sigma^2\lambda_k^{-1}$ for $i=1,\dots,k-1$. This is will not happen if
+$\lambda_k^{-1}$ increase quickly --- in section \ref{subsec:abstract
+  OED} we assumed $\fwd$ is strongly smoothing, which means
+eigenvalues of $(\fwd \prcov\fwd^*)^{-1}$ grow quickly.
 
@@
 exhibit clusterization --- e.g. the second design presented in figure
-\ref{fig:clusterization}. In the following section we postulate what
-causes such designs to be observed in realistic experimental design
-scenarios.
+\ref{fig:clusterization}.
 
@@
         ylabel=precision --- prior and posterior,
-        xlabel=eigenvector,
+        xlabel=eigenvector $i$,
       ]
@@
   \end{tikzpicture}
-  \caption{Clusterization and non-clusterization in simultaneous
-    design. Posterior precision per eigenvector, after taking a
-    measurement $\meas_i$. Each $\meas_i$ increases precision
-    uniformly across the lowest precision eigenvectors. We see that
-    $\meas_2$ and $\meas_3$ are repeated measurements, as well as
-    $\meas_5$ and $\meas_6$.}
+  \caption{Clusterization (left) and non-clusterization (right) in
+    simultaneous D-optimal designs. Posterior precision per
+    eigenvector, after taking a measurement $\meas_i$ is plotted. An
+    optimal design reduces uncertainty where it is highest first. This
+    can be achieved by taking repeated measurements --- $\meas_2 =
+    \meas_3$ and $\meas_5 = \meas_6$ in the left design. This is not
+    necessary though, as can be seen in the right design. Both designs
+    achieve the same levels of uncertainty. The left one exhibits
+    clusterization. The right one does not.}
   \label{fig:clusterization}
@@
   increasing the magnitude of the $j$th measurement functional
-  increases $\tar(\obs)$. We start with the last line of \eqref{eq:tar
-    var}, and denote $\tmp : = \fwd \postcov \fwd^*$.
+  increases $\tar(\obs)$. We start with the last line of \eqref{eq:tar var}, and denote $\tmp : = \fwd \postcov \fwd^*$.
   \begin{align*}
@@
 
-%% HERE
-%% Since a measurement that takes into account $\ker \fwd$ is
-%% necessarily sub-optimal, we may assume $\fwd$ is in fact
-%% invertible (it is, on $\ker \fwd ^{\perp}$) and rewrite the
-%% necessary condition \eqref{eq:eigenproblem} using
-%% \eqref{eq:postcov} as \begin{equation*} \sigma^{-2} \left ( (\fwd
-%% \prcov \fwd^*)^{-1} + \sigma^{-2}\obs\obs^* \right )^{-1} =
-%% \obs^* \Xi.  \end{equation*} By Lemma \ref{lemma:sim diag},
-%% taking $C = (\fwd \prcov \fwd^*)^{-1}$ we observe that $\meas_j,
-%% j=1,\dots,m$ are eigenvectors of $\fwd \prcov \fwd^*$. If they have
-%% different eigenvalue, they are orthogonal. If not, this means, by
-%% the common interpretation of the Lagrange multipliers, that an
-%% infinitesimal increase in one will cause the same increase in
-%% design criterion as others with the same eigenvalue. So if these
-%% eigenvectors do not have the same angle between every pair, an
-%% infinitesimal increase in one will give more increase than
-%% others, which is a contradiction. So in every subspace, vectors
-%% will be restricted to these forms and if there will be more
-%% vectors than the dimensionality of the subspace, clusterization
-%% will occur.
-
-%% \begin{figure}
-%%   \begin{tikzpicture}[thick, scale=1.4, every node/.style={scale=0.99}]
-%%     \begin{axis}
-%%       [
-%%         xmin = 0,
-%%         xmax = 3.14,
-%%         xlabel = {$x$},
-%%         ylabel = posterior std,
-%%         ymin   = 0,
-%%         %compat = 1.3,
-%%         % ymax   = 130,
-%%         % ytick = \empty,
-%%         legend cell align=left,
-%%         % legend style={at={(0.45,0.2)}}
-%%         legend pos= outer north east 
-%%       ]
-%%       % \draw[black!30!white, thin] (50,0) -- (50,130);
-%%       % 
-%%       \addplot [thin, black, mark=none] table{data/stdv-trunc-sens1-var1.txt};
-%%       \addlegendentry{1 sensor};
-
-%%       \addplot [thin, blue, mark=none] table{data/stdv-trunc-sens2-var1.txt};
-%%       \addlegendentry{2 sensors};
-
-%%       \addplot [thin, red, mark=none] table{data/stdv-trunc-sens3-var1.txt};
-%%       \addlegendentry{3 sensors};
-
-%%       \addplot [thin, green, mark=none] table{data/stdv-trunc-sens4-var1.txt};
-%%       \addlegendentry{4 sensors};
-
-%%       \addplot [thin, purple, mark=none] table{data/stdv-trunc-sens5-var1.txt};
-%%       \addlegendentry{5 sensors};
-
-%%       \addplot [thin, cyan, mark=none] table{data/stdv-trunc-sens6-var1.txt};
-%%       \addlegendentry{6 sensors};
-
-%%       \addplot [black,  only marks, mark=x, mark size=1.5] 
-%%       table{data/locs-trunc-sens1-var1.txt}; 
-%%       \addplot [blue,   only marks, mark=x, mark size=1.5]
-%%       table{data/locs-trunc-sens2-var1.txt}; 
-%%       \addplot [red,    only marks, mark=x, mark size=1.5]
-%%       table{data/locs-trunc-sens3-var1.txt};
-%%       \addplot [green,  only marks, mark=x, mark size=1.5] 
-%%       table{data/locs-trunc-sens4-var1.txt}; 
-%%       \addplot [purple, only marks, mark=x, mark size=1.5] 
-%%       table{data/locs-trunc-sens5-var1.txt}; 
-%%       \addplot [cyan,   only marks, mark=x, mark size=1.5] 
-%%       table{data/locs-trunc-sens6-var1.txt}; 
-%%     \end{axis}
-%%   \end{tikzpicture}
-%%   \caption{Analogously to figure \ref{fig:clusterization
-%%       illustration}, shown are posterior pointwise standard
-%%     deviations. Note the exact clustering. Here the dynamics is not
-%%     the one governed by the 1D heat equation but a dynamics that sets
-%%     to zero all modes except the first three. This dynamics simulates
-%%     a finite-rank operator as per the conditions of Theorem
-%%     \ref{theorem:finite rank}. Although the conditions of the theorem
-%%     allow the use of a different constraint on measurements (see
-%%     section \ref{subsec:unit length}), the conclusion of the theorem
-%%     still seems to hold.}
-%%   \label{fig:finite rank illustration}
-%% \end{figure}
-
-%% \section{Counterexample}\label{section:counterexample}
-%% Applying Lemma \ref{lemma:sim diag} to
-%% \begin{equation*}
-%%   (\fwd \postcov \fwd^*)^{-1} = (\fwd \prcov \fwd^*)^{-1} +
-%%   \sigma^{-2}\obs^*\obs
-%% \end{equation*}
-%% implies $(\fwd \prcov \fwd^*)^{-1}$ and $\obs^*\obs$ have the same
-%% eigenspaces. Since $\meas_j$ are eigenvectors of $\fwd \postcov
-%% \fwd^*$, we may expect this implies that $\meas_j$ are also
-%% eigenvectors of $(\fwd \prcov \fwd^*)^{-1}$ and assuming $(\fwd \prcov
-%% \fwd^*)^{-1}$ has simple eigenvalues completes the proof. Our
-%% expectations, however, are let down.
-
-%% Here is a counterexample. Let $e_1, e_2$ denote two
-%% eigenvectors of $C := (\fwd \prcov \fwd^*)^{-1}$ of different
-%% eigenvalues such that $C + \obs^*\obs$ is scalar on span
-%% $V:=\{e_1,e_2\}$ (we took $\sigma^2 = 1$ for simplicity). Then,
-%% on $V$, we have  
-%% \begin{equation*}
-%%   (\fwd \prcov \fwd^*)^{-1} = diag( d_1, d_2),
-%% \end{equation*}
-%% and we assume $d_1 > d_2$ without loss of generality. For $\meas \in
-%% \hilp$, we denote $\meas^*$ its Riesz representer, making a departure
-%% from the standard notation employed, where $\meas$ is usually taken to
-%% be a linear functional on $\hilp$. Let $\obs = (\meas_1^*, \meas_2^*,
-%% \meas_3^*)^t$. In this setup, we would like that $\meas_j,j=1,2,3$ are
-%% either orthogonal or parallel. We now outline construction that shows
-%% this is not necessarily true.
-
-%% We know, from Corollary \ref{cor:sim diag} that $\meas_j \in V$, so
-%% let
-%% \begin{align*}
-%%   \meas_1 &=  c_1e_1 + c_2e_2 \\
-%%   \meas_2 &= -c_1e_1 + c_2e_2 \\
-%%   \meas_3 &=     e_1.
-%% \end{align*}
-%% Then
-%% \begin{align*}
-%%   \meas_1 \meas_1^*
-%%   &= c_1^2e_1e_1^* + c_1c_2( e_1e_2^* + e_2e_1^*) +
-%%   c_2^2 e_2e_2^* \\
-%%   \meas_2 \meas_2^* &= c_1^2e_1e_1^* - c_1c_2( e_1e_2^* +
-%%   e_2e_1^*) + c_2^2 e_2e_2^*.
-%% \end{align*}
-%% Note that
-%% \begin{equation*}
-%%   \obs^* \obs = \meas_1 \meas_1^* + \meas_2\meas_2^* +
-%%   \meas_3\meas_3^* = (2c_1^2+1) e_1e_1^* + 2c_2^2e_2e_2^*,
-%% \end{equation*}
-%% is diaonal in $\{e_1,e_2\}$. In order for $\meas_j,j=1,2,3$ to be
-%% eigenvectors of $C + \obs^*\obs$ with the same eigenvalue, $C +
-%% \obs^*\obs$ must be a scalar operator on $V$. Thus the following
-%% must hold:
-%% \begin{align*}
-%%   d_1 + 2c_1^2 + 1 &= d_2 + 2c_2^2 \\
-%%   c_1^2 + c_2^2 &= 1 \text{ (arises from $\| \meas_i \| = 1$).}
-%% \end{align*}
-%% Equivalently,
-%% \begin{equation*}
-%%   \begin{bmatrix}
-%%     2 & -2 \\
-%%     1 & 1
-%%   \end{bmatrix}
-%%   \begin{bmatrix}
-%%     c_1^2 \\
-%%     c_2^2
-%%   \end{bmatrix}
-%%   =
-%%   \begin{bmatrix}
-%%     d_2 - d_1 - 1 \\
-%%     1
-%%   \end{bmatrix}.
-%% \end{equation*}
-%% We can always find a solution for this system, for any $d_1,d_2$.
-%% However, it is necessary that either $c_1^2 = 1$ or $c_2^2 = 1$ for
-%% $\meas_j, j=1,2,3$ to all be either orthogonal or parallel. The first
-%% conditiion implies $d_2 - d_1 = 3$, which is impossible, since $d_1 >
-%% d_2$. The second requires $d_1 - d_2 = 1$, a condition that does not
-%% hold generically.
-
-%    Text of article.
-
-%    Bibliographies can be prepared with BibTeX using amsplain,
-%    amsalpha, or (for "historical" overviews) natbib style.
 \bibliographystyle{amsplain}
@@
 \end{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%    Templates for common elements of an article; for additional
-%    information, see the file Author_Handbook_ProcColl.pdf, included
-%    in every AMS author package, and the amsthm user's guide, linked
-%    from http://www.ams.org/tex/amslatex.html .
-
-%    Section headings
-\section{}
-\subsection{}
-
-%    Ordinary theorem and proof
-\begin{theorem}[Optional addition to theorem head]
-% text of theorem
-\end{theorem}
-
-\begin{proof}[Optional replacement proof heading]
-% text of proof
-\end{proof}
-
-%    Figure insertion; default placement is top; if the figure occupies
-%    more than 75% of a page, the [p] option should be specified.
-\begin{figure}
-\includegraphics{filename}
-\caption{text of caption}
-\label{}
-\end{figure}
-
-%    Mathematical displays; for additional information, see the amsmath
-%    user's guide:  texdoc amamath  or
-%  http://mirror.ctan.org/tex-archive/macros/latex/required/amsmath/amsldoc.pdf
-
-% Numbered equation
-\begin{equation}
-\end{equation}
-
-% Unnumbered equation
-\begin{equation*}
-\end{equation*}
-
-% Aligned equations
-\begin{align}
-  &  \\
-  &
-\end{align}
-
-%-----------------------------------------------------------------------
-% End of article-template.tex
-%-----------------------------------------------------------------------

### 2020-10-25 reply.tex
revised

@@
 on every point made. The referees' comments and my replies are listed
-below. My replies are colored \textcolor{cyan}{cyan}. Again, thank you
-for taking the time and helping me improve this manuscript.
+below. Again, thank you for taking the time and helping me improve
+this manuscript.
 
@@
 using them.
-\ydnote{Corrected, thanks!}
+\ydnote{Fixed.}
 
@@
   settings".
-  \ydnote{Changed in text}
+  \ydnote{Changed.}
 
 \item Page 2, Line 36: Add a comma: "In the process, we generalize..."
-  \ydnote{Changed in text}
+  \ydnote{Changed.}
 
@@
   should be changed to "with a homogeneous Dirichlet boundary
-  condiiton". \ydnote{Changed}
+  condiiton". \ydnote{Changed.}
 
@@
   defined. Please formally define this before giving the equation.
-  \ydnote{Moved the equation after $\Sigma$ is defined}
+  \ydnote{Moved the equation after $\Sigma$ is defined.}
 
@@
   \Gamma_{pr})$ assumes the prior mean is 0. This is worthy of some
-  comment. Why must the prior mean be 0?
-  \ydnote{It does not have to be, we can subtract the mean. An explanation was added}
+  comment. Why must the prior mean be 0?  \ydnote{Assuming zero mean
+    is hardly necessary, so I removed this statement.}
 
@@
   terms, it is easy..."
-  \ydnote{Changed in text}
+  \ydnote{Changed.}
 
@@
   function.
-  \ydnote{Added comment before statment of theorem where
-    $\tar$ first appears}
+  \ydnote{Added comment before statment of theorem where $\tar$ first
+    appears. Also, I added a formal definition of D-optimal design
+    just after the theorem.}
 
@@
   should be reworded.
-  \ydnote{Added short explanation}
+  \ydnote{Rephrased and added a short explanation.}
 
@@
   Proposition 2 in the appendix.
-  \ydnote{All such errors were fixed}
+  \ydnote{All such errors were fixed.}
 
@@
   formula here as you do for another term earlier in the sentence.
-  \ydnote{Explicit expression given}
+  \ydnote{Explicit expression given.}
 
 \item Page 18, Line 6: This proposition was previously labeled
-  Proposition 1. \ydnote{All such errors were fixed}
+  Proposition 1. \ydnote{All such errors were fixed.}
 \end{itemize}
@@
   and observation operators were merged. A short remark on this was
-  added. It is placed after equation (2.6) in the highlighted text
-  (equation (2.5) in the new clean text. Specifically, we need it in
-  equation (3.1).}
+  added. It is placed after equation (2.5). Specifically, we need it
+  in equation (3.1).}
 
@@
   here.
-  \ydnote{Unfortunately, this work does not have
-    immediate applications to (e.g.) spatial sensor placement
-    problems. I offer a toy model for understanding sensor
-    clusterization. I plan to use some of the conclusions in
-    subsequent work but I find it out of scope for the work under
-    consideration here.}
+  \ydnote{This work does have some potential applications to (e.g.)
+    spatial sensor placement problems. I believe they are out of scope
+    for the current paper and I plan on studying those in the
+    future. Here I only offer a toy model for understanding sensor
+    clusterization.}
 
@@
 however, the following argument on Page 9 that extends this assumption
-is not well-justified.  \ydnote{This is a great point! The flawed
-  explanation was removed. I was able to remove that assumption. But
-  it required further changes to the manuscript.}
+is not well-justified.
+\ydnote{This is a great point! The flawed explanation was removed. I
+  was able to remove that assumption, at the cost of changing several
+  claims in the paper. Mainly, it forced me to work with Hilbert
+  spaces everywhere. I believe this is for the best, because
+  considering Banach spaces did not add much.}
 
@@
 be clearly discussed for the reader to appreciate the valuable effort
-made. \ydnote{Again, great point. $\fwd^{-*} = (\fwd^*)^{-1}$ but now
-  I do not assume $\fwd$ is invertible so this was removed.}
+made.
+\ydnote{Again, great point. $\fwd^{-*} = (\fwd^*)^{-1}$ but now I do
+  not assume $\fwd$ is invertible so this was removed.}
 
@@
   conclusions clearer. I also added an illustration story with
-  graduated lab cylinders.}
+  graduated lab cylinders. I also added a theorem summarizing the
+  results for vanishing model error.}
 
@@
 abstract, in addition to the fact that observations errors are
-uncorrelated with fixed variance $\sigma^2$.  \ydnote{The title and
-  abstract were changed to reflect the fact that linear problems are
-  concerned.  The treatment of correlated errors was emphasized.}
+uncorrelated with fixed variance $\sigma^2$.
+\ydnote{The title and abstract were changed to reflect the fact that
+  linear problems are concerned. The treatment of correlated errors
+  was emphasized.}
 
@@
 relevant discussion to that issue which I find interesting.
-\ydnote{The discussion on observation correlation
-  and how it mitigates clusterization can be found in the last
-  section.}
+\ydnote{The discussion on observation correlation and how it mitigates
+  clusterization can be found in the last section. It is now stated
+  clearly in the contributions section.}
 
@@
 The trace operator should be followed with some sort of brackets to
-make it easier to follow and validated the formulae
+make it easier to follow and validated the formulae.
 \ydnote{Curly brackets added where needed.}
@@
 and caption of Figure 2 should be rewritten with clear explanation of
-the axes, and the plot-generation process.  \ydnote{I extended the
+the axes, and the plot-generation process.
+%
+\ydnote{I extended the
   discussion relating to clusterization, added another plot and
@@
 lower uncertainty, then, variable observation variances must be
-considered. \ydnote{Unfortunately, I am again not sure I understand
+considered.
+%
+\ydnote{Unfortunately, I am again not sure I understand
   fully. Variable observation variances are fixed to $\sigma^2$
-  througout this section.}
+  througout this section. The eigenvalues are always weigted by the
+  obeservation variance $sigma^2$. Before, $\sigma^{-2}$ was
+  multiplying $\eta_i$. Now $\sigma^2$ multiplies $\lambda_i^{-1}$. In
+  any case, obesevation error is taken into account.}
 
@@
 the uncertainty levels in linear Bayesian settings, as manifested by
-the posterior covariance matrix.  \ydnote{I think the referee implies
-  that the D-optimal design is only as good as the assumptions it
-  relies on. I could not agree more. However, as long as we do believe
-  the assumptions made --- a D-optimal design is indeed optimal (in a
-  precise and quantifiable way). I think the source of
-  misunderstanding is that I used the term ``measurement'' where I
-  should have used ``eigenvector''. I changed the paragraph
-  accordingly. The transition from thinking of ``measurements'' to
-  thinking of ``eigenvectors'' is confusing and I believe it is better
-  explained in the updated version of the manusctipt.}
+the posterior covariance matrix.
+%
+\ydnote{I think the referee implies that the D-optimal design is only
+  as good as the assumptions it relies on. I could not agree
+  more. However, as long as we do believe the assumptions made --- a
+  D-optimal design is indeed optimal, in a precise and quantifiable
+  way. I think the source of misunderstanding is that I used the term
+  ``measurement'' where I should have used ``eigenvector''. I changed
+  the paragraph accordingly. The transition from thinking of
+  ``measurements'' to thinking of ``eigenvectors'' is confusing and I
+  believe it is better explained in the updated version of the
+  manusctipt.}
 
@@
 $X(t=0)$ explicitly (see line 32 on page 11). A clarification is
-required here.  \ydnote{Clarification was added at the end of the
+required here.
+%
+\ydnote{Clarification was added at the end of the
   proof. ``There is nothing special about $t=0$'' means we could have
@@
 referring to.
+%
 \ydnote{Made explicit.}
 
-Page 2, line 11; wrong reference to Figure 1.3?  \ydnote{Reference now
-  fixed.}
+Page 2, line 11; wrong reference to Figure 1.3?
+%
+\ydnote{Fixed.}
 
@@
 to achieve this result? Please clarify.
+%
 \ydnote{The expectation / integration over the prior space happens
@@
   $$
-  \mathbb{E}_{\data}[KL(\post^\data||\pr)] = \int KL(\post^\data||\pr) \der \mu(\data) = \int \int \log \frac{\der \post^\data}{\der \pr}(x) \der \post(x) \der \mu(\data)
-  $$ where $\mu(\data)$ is the distribution of data (usually called
+  \mathbb{E}_{\data}[KL(\post||\pr)] = \int KL(\post||\pr) \der
+  \mu(\data) = \int \int \log \frac{\der \post}{\der \pr}(x) \der
+  \post(x) \der \mu(\data)
+  $$
+  where $\mu(\data)$ is the distribution of data (usually called
   evidence). I improved the notation to reflect the dependence of the
@@
 Page 6, lines 19; the statement made about relaxing the constraints on
-the observation operator is unclear.  \ydnote{This sentence was poorly
+the observation operator is unclear.
+%
+\ydnote{This sentence was poorly
   written. It is now rephrased.}
@@
 Page 6, line 24, I believe the reference to (3.2) should be replaced
-with (3.1). \ydnote{This was actually OK. The gradient was recorded in
-  (3.2).}
+with (3.1).
+%
+\ydnote{This is actually OK. The gradient was recorded in (3.2).}
 
@@
 clearly.
+%
 \ydnote{Fixed.}
@@
 A.1). Maybe state that clearly here. Also, in the proof, from step 4
-to 5, the first and third terms vanished; why?  \ydnote{This point is
-  spot on. I added the definition of $T$ before the derivation. This
-  is not the proof of Lemma 3.1 --- it is the first variation of the
-  objective. I reworded a few sentences to make this clearer.}
+to 5, the first and third terms vanished; why?
+%
+\ydnote{This point is spot on. I added the definition of $T$ before
+  the derivation. The proof referee 2 refers to is not the proof of
+  Lemma 3.1. It is the first variation of the objective. I reworded a
+  few sentences to make this clearer. Regarding the vanishing terms
+  --- this is done using trace rules (notice the $\frac{1}{2}$
+  prefactor disappears). I added a comment explaining this.}
 

### 2020-10-26 new_paper.tex
Cite karina and ahmed, changed propostions, theorems, lemmas to upper/lowercase, discarded the requirement for zero mean

@@
   spatially correlated measurement error clusterization does not
-  occur. As a part of the analysis we prove a matrix determinant lemma
+  occur. As a part of the analysis we prove a Matrix Determinant Lemma
   analog in infinite dimensions and a lemma for calculating
@@
 
+In the inverse problems community, work is mostly computational and
+less theoretic. Model errors were considered in \cite{Attia20,
+  Koval20}. The former study is focused on inferring of a Quantity of
+Interest (QoI). The focus of the latter is reducing forward solves,
+using randomized linear algebra. Both studies present numerical
+techniques for finding optimal designs when model error is
+present. Both are restricted to linear inverse problems (although in
+the latter the authors use their method on a nonlinear problem with a
+Laplace approximation for the posterior). Both find an optimal design
+by first solving a continuous problem for sensor weights. Said
+solution is then sparsified to give a binary design. Both studies are
+successful in the task of Bayesian inversion. However, neither of
+these studies mentions any effect model errors can have on sensor
+clusterization. The current study is mostly theoretical and aims to
+fill the theoretical gap of understanding sensor clusterization.
+
 \subsection{Contribution}
@@
 structure helps us shed light into D-optimal designs. We characterize
-these designs using theorem \ref{thm:char}. One insight we prove in
+these designs using Theorem \ref{thm:char}. One insight we prove in
 said theorem, is that a D-optimal design reduces uncertainties where
@@
 In the process we generalize several lemmas from linear algebra to
-infinite-dimensional settings. We prove a matrix determinant lemma in
+infinite-dimensional settings. We prove a Matrix Determinant Lemma in
 \ref{lemma:MDL}. We generalize a lemma due to Lax \cite{lax97} for
@@
 calculating the increase in the design criterion per observation ---
-Lemmas \ref{lemma:design increase} and Corollary \ref{cor:zero mod
+Lemma \ref{lemma:design increase} and Corollary \ref{cor:zero mod
   err}.
@@
 noisy observations of $\fwd \param$. We take a Gaussian prior $\pr$
-for $\param$. Hence, $\param \sim \pr = \normal(\param_0 ,\prcov)$
+for $\param$. Hence, $\param \sim \pr = \normal(\prmean ,\prcov)$
 with some appropriate covariance operator $\prcov$ \cite{Stuart10}. It
@@
 \begin{theorem}[Slightly modified Theorem 1 from \cite{AlexanderianGloorGhattas14}]
-  Let $\pr = \normal(\param_0,\prcov)$ be a Gaussian prior on $\hilp$
+  Let $\pr = \normal(\prmean,\prcov)$ be a Gaussian prior on $\hilp$
   and let $\post = \normal(\postmean,\postcov)$ the posterior measure
@@
 Below, the expression for the posterior covariance \eqref{eq:postcov}
-justifies the second equality. Some calculations delegated to lemma
+justifies the second equality. Some calculations delegated to Lemma
 \ref{lemma:twice woodbury} in the appendix justify the third (trivial
@@
 In \eqref{eq:mod conditions}, take $\func_j^{*} = \meas_j$ and $C :=
-(\fwd \prcov \fwd^*)^{-1}$ and use lemma \ref{lemma:sim diag}.
+(\fwd \prcov \fwd^*)^{-1}$ and use Lemma \ref{lemma:sim diag}.
 \end{proof}
@@
 $\{\lambda_i\}_{i=1}^{\infty}$. They are to be thought of as fixed but
-unknown. Using Sylvester's determinant theorem (the standard proof
+unknown. Using Sylvester's Determinant Theorem (the standard proof
 works for Hilbert spaces as well) we get:
@@
   not}. The analysis conducted in this section is summarized in
-theorem \ref{thm:char} below.
+Theorem \ref{thm:char} below.
 
@@
 \subsection{Overview}
-In this section we use theorem \ref{thm:char} to explain how sensor
+In this section we use Theorem \ref{thm:char} to explain how sensor
 clusterization may arise in various scenarios. We first consider
@@
 in the sequential design case. We still rely on previous results ---
-namely corollary \ref{cor:same ev} and theorem \ref{thm:char}. An
+namely Corollary \ref{cor:same ev} and Theorem \ref{thm:char}. An
 illustration of two D-optimal designs scenarios for the same problem

### 2020-10-26 new_paper.tex
Add refs inside the reply tex

@@
 present. Both are restricted to linear inverse problems (although in
-the latter the authors use their method on a nonlinear problem with a
-Laplace approximation for the posterior). Both find an optimal design
-by first solving a continuous problem for sensor weights. Said
-solution is then sparsified to give a binary design. Both studies are
-successful in the task of Bayesian inversion. However, neither of
-these studies mentions any effect model errors can have on sensor
-clusterization. The current study is mostly theoretical and aims to
-fill the theoretical gap of understanding sensor clusterization.
+the latter the authors use their method on a nonlinear problem by
+taking a Laplace approximation for the posterior). Both find an
+optimal design by first solving a continuous problem for sensor
+weights. Said solution is then sparsified to give a binary
+design. Both studies are successful in the task of Bayesian
+inversion. However, neither of these studies mentions any effect model
+errors can have on sensor clusterization. The current study is mostly
+theoretical and aims to fill the theoretical gap of understanding
+sensor clusterization.
 
@@
 $$
-KL(\post||\pr) = \int \log \frac{\der \post}{\der \pr}(x) \der \post(x).
+KL(\post||\pr) = \int \log \frac{\der \post}{\der \pr}(\param) \der \post(\param).
 $$
@@
 \section{Acknowledgements}
-This study is a part of my PhD thesis. It was written under the
-instruction of Prof. Georg Stadler in New York University's Courant
-Institute. I would like to thank him for his great mentorship.
+This study is a part of my PhD thesis. Most of it was written under
+the instruction of Prof. Georg Stadler in New York University's
+Courant Institute. I would like to thank him for his great
+mentorship. This work was supported by The Raymond and Beverly Sackler
+Post-Doctoral Scholarship.
 

### 2020-10-26 reply.tex
Cite karina and ahmed, changed propostions, theorems, lemmas to upper/lowercase, discarded the requirement for zero mean
Add refs inside the reply tex

@@
 \input{definitions}
-
+\usepackage{xr}
+\externaldocument{new_paper}
 \title{Corrections}
@@
 below. Again, thank you for taking the time and helping me improve
-this manuscript.
+this manuscript. Below, when I refer to an item - it refers to the new
+version of the paper.
 
@@
 elaboration in section 4.2.
+%
+\ydnote{I added figure \ref{fig:optimal vs not} prior to figure
+  \ref{fig:clusterization}. I also added Theorem \ref{thm:char}, which
+  neatly summarizes the results of section
+  \ref{section:vanishing}. This should help understand section
+  \ref{section:clusterization}, which was extended and modified
+  heavily.}
 \end{enumerate}
@@
   defined. Please formally define this before giving the equation.
-  \ydnote{Moved the equation after $\Sigma$ is defined.}
+  \ydnote{Moved this equation to after $\Sigma$ is defined.}
 
@@
   \Gamma_{pr})$ assumes the prior mean is 0. This is worthy of some
-  comment. Why must the prior mean be 0?  \ydnote{Assuming zero mean
-    is hardly necessary, so I removed this statement.}
+  comment. Why must the prior mean be 0?
+  %
+  \ydnote{Great point. Assuming zero mean is hardly necessary, so I
+  removed this statement.}
 
@@
   theorem. It is defined mathematically but not described. Only much
-  later is it stated that this is the objective
-  function.
+  later is it stated that this is the objective function.
   \ydnote{Added comment before statment of theorem where $\tar$ first
-    appears. Also, I added a formal definition of D-optimal design
-    just after the theorem.}
+    appears. I also added a definition of KL divergence in the
+    beginning of section \ref{subsec:D optimal design}. I added
+    Definition \ref{def:d optimality} which formally defines
+    D-optimality.}
 
@@
   should be reworded.
-  \ydnote{Rephrased and added a short explanation.}
+  \ydnote{Rephrased and added an explanation.}
 
@@
 \item Page 18, Line 6: This proposition was previously labeled
-  Proposition 1. \ydnote{All such errors were fixed.}
+  Proposition 1.
+  \ydnote{All such errors were fixed.}
 \end{itemize}
@@
 the author may find useful.
+%
+\ydnote{The references are much appreciated. I extended the ``related
+  work'' section accordingly.}
 
@@
 I don't see the benefit of consider infinite-dimensional formulation.
+%
 \ydnote{I was not able to arrive at the same results when the forward
   and observation operators were merged. A short remark on this was
-  added. It is placed after equation (2.5). Specifically, we need it
-  in equation (3.1).}
+  added. It is placed just before section
+  \eqref{subsec:dynamics}. Specifically, we need it in (e.g.)
+  \eqref{eq:mod conditions}.}
 
@@
 is not well-justified.
+%
 \ydnote{This is a great point! The flawed explanation was removed. I
@@
   spaces everywhere. I believe this is for the best, because
-  considering Banach spaces did not add much.}
+  considering Banach spaces did not add much. The new derivations
+  appear in \eqref{eq:pre manipulations} and \eqref{eq:mod
+    objective}. Instead, I use the much more reaonable assumption that
+  $\fwd \prcov \fwd^*$ is invertible.}
 
@@
 \ydnote{I added several plots that I believe make the arguments and
-  conclusions clearer. I also added an illustration story with
-  graduated lab cylinders. I also added a theorem summarizing the
-  results for vanishing model error.}
+  conclusions clearer. I added Theorem \ref{thm:char}, summarizing the
+  results for vanishing model error. I also added an illustration
+  story with graduated lab cylinders, just before the statement of
+  Theorem \ref{thm:char} (see also figure \ref{fig:cylinder}).}
 
@@
 relevant discussion to that issue which I find interesting.
+%
 \ydnote{The discussion on observation correlation and how it mitigates
-  clusterization can be found in the last section. It is now stated
-  clearly in the contributions section.}
+  clusterization can be found in section \ref{section:non
+    vanishing}. It is now stated clearly in the contributions
+  section.}
 
@@
 not works properly.  For example Corollary 2 should be Corollary B.x,
-etc.  \ydnote{All such bad references and links were fixed.}
+etc.
+%
+\ydnote{All such bad references and links were fixed.}
 
@@
 %
-\ydnote{I extended the
-  discussion relating to clusterization, added another plot and
-  explained how they arise.}
+\ydnote{I extended the discussion relating to clusterization. I added
+  figure \ref{fig:optimal vs not} to illustrate how the posterior of a
+  D-optimal design behaves. Then I extend the discussion to section
+  \ref{section:clusterization}.}
 
@@
 %
-\ydnote{Clarification was added at the end of the
-  proof. ``There is nothing special about $t=0$'' means we could have
-  chosen any other value $t_0$ instead.}
+\ydnote{Clarification was added at the end of the proof. ``There is
+  nothing special about $t=0$'' means we could have chosen any other
+  value $t_0$ instead.}
 
@@
   \mathbb{E}_{\data}[KL(\post||\pr)] = \int KL(\post||\pr) \der
-  \mu(\data) = \int \int \log \frac{\der \post}{\der \pr}(x) \der
-  \post(x) \der \mu(\data)
+  \mu(\data) = \int \int \log \frac{\der \post}{\der \pr}(\param) \der
+  \post(\param) \der \mu(\data)
   $$
@@
   posterior on the data $\data$. I added above definition of KL
-  divergence to the manuscript.}
+  divergence to the beginning of section \ref{subsec:D optimal design}.}
 
@@
   explanation. The mauscript now reflects the claim that the optimal
-  $\obs^{\star}$ is the maximizer of $\tar(\obs)$.}
+  $\obs^{\star}$ is the maximizer of $\tar(\obs)$. This is recorded
+  explicitly in Definition \ref{def:d optimality}.}
 
@@
 %
-\ydnote{This sentence was poorly
-  written. It is now rephrased.}
+\ydnote{This sentence was poorly written. It is now rephrased.}
 
@@
 \ydnote{This point is spot on. I added the definition of $T$ before
-  the derivation. The proof referee 2 refers to is not the proof of
+  the derivation, just after the statment of Lemma
+  \ref{lemma:lax}. The proof referee 2 refers to is not the proof of
   Lemma 3.1. It is the first variation of the objective. I reworded a
@@
 Proposition 3.2 or something similar to match previous referencing
-(e.g., Lemma 3.1). \ydnote{Fixed.}
+(e.g., Lemma 3.1).
+%
+\ydnote{Fixed.}
 
@@
 seems to be incorrect. Also, state clearly which appendix the proof is
-delegated to.  \ydnote{Fixed.}
+delegated to.
+%
+\ydnote{Fixed.}
 
@@
 $\widehat{\Gamma_{post}}$ int the corollary, which is defined only in
-the appendix, under the proof. \ydnote{All comments on numbering and
-  references are great and correct. All were fixed. I added an
-  explanation of what $\widehat{\Gamma_{post}}$ means where it
-  appears.}
+the appendix, under the proof.
+%
+\ydnote{All comments on numbering and references are great and
+  correct. All were fixed. I added an explanation of what
+  $\widehat{\Gamma_{post}}$ means before Corollary \ref{cor:same
+    meas}.}
 
@@
 derivative vanished from step 2 to step 3; please explain!
-\ydnote{I added an extra equality line for explanation.}
+%
+\ydnote{I added an extra equality line for explanation in the proof of
+  lemma \ref{lemma:lax}.}
 

### 2020-10-26 new_paper.tex
I think I am donegit add reply.tex git add reply.tex

@@
 \begin{abstract}
-  We investigate the problem of sensor clusterization in optimal
+  We investigate the problem of sensor clusterization in D-optimal
   experimental design for infinite-dimensional Bayesian linear inverse
@@
   occur. As a part of the analysis we prove a Matrix Determinant Lemma
-  analog in infinite dimensions and a lemma for calculating
-  $\frac{\der \log \det (I + X(t))}{\der t}$ for certain
-  operator-valued functions. We also show how to decompose $M = AA^t$
-  with $A$ constrained to have unit norm columns.
+  in Hilbert spaces and a lemma for calculating $\frac{\der}{\der t}
+  \log \det (I + X(t))$ for certain operator-valued functions. We also
+  show how to decompose $M = AA^t$ with $A$ constrained to have unit
+  norm columns.
 \end{abstract}
@@
 It seems counter intuitive that when one computes an optimal design
-using the D-optimal criterion, the optimization process results in
+using the D criterion, the optimization process results in
 measurements that are very similar. For example, if a measurement is
@@
 occurs in a sequential design scenario (section
-\ref{subsec:clusterization sequential}). Then, we show that designs
-that exhibit clusterization are just as optimal as clusterization free
-designs in a simultaneous D-optimal designs (section
-\ref{subsec:clusterization simultaneous}).
+\ref{subsec:clusterization sequential}). Then, we consider a
+simultaneous D-optimal design scenario (section
+\ref{subsec:clusterization simultaneous}). We show that designs that
+exhibit clusterization are just as optimal as clusterization free
+designs.
 
@@
 \subsection{Overview}
-We seek a formulation of the optimal design problem
+We seek a formulation of the D-optimal design problem
 \eqref{eq:optimization} using Lagrange multipliers. In section
@@
   constraints}. Finally, in section \ref{subsec:necessary} we use
-these to formulate the optimal design problem as a Lagrange
+these to formulate the D-optimal design problem as a Lagrange
 multipliers problem in equation \eqref{eq:conditions}.
@@
     + \obs^* \Sigma^{-1} V ) \fwd \Big \}  \text{ (by \ref{lemma:aux calc})}\\
-%%   \end{split}
-%% \end{align} 
-%% Taking adjoints and cyclic permutations we see that
-%% \begin{align*}
-%%   \tr \postcov \fwd^* V^* \Sigma^{-1} \obs \fwd 
-%%   &= \tr \postcov \fwd^* \obs^* \Sigma^{-1} V \fwd,
-%% \end{align*}
-%% and
-%% \begin{align*}
-%%   \tr \postcov \fwd^* \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \fwd
-%%   &= \tr \postcov \fwd^*  \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs \fwd.
-%% \end{align*}
-%% We use these in the first equality of the following derivation,
-%% which continues the one in \eqref{eq:precursor}.
-%% %
-%% \begin{align} \label{eq:tar var}
-%%   \begin{split}
-      %% \delta \tar(\obs) V 
-    &= \tr{\postcov \fwd^* ( \obs^* \Sigma^{-1} V 
-    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs ) \fwd} \text{ (trace rules, $\frac{1}{2}$ prefactor disappears)} \\
+    %
+    %
+    %
+    &= \tr{\postcov \fwd^* ( \obs^* \Sigma^{-1} V -
+    \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs ) \fwd} \text{
+        (cyclic property of trace)} \\
     %
@@
 In this section we suggest relaxed constraints in
-\eqref{eq:constraints} and derive their gradient in \eqref{eq:grad
-  constraints}, for the purpose of using them in a Lagrange
-multipliers problem.
+\eqref{eq:constraints}. We then find their gradient in \eqref{eq:grad
+  constraints}. This is for using them in a Lagrange multipliers
+problem.
 
@@
 \subsection{Necessary Conditions for Optimal Design}\label{subsec:necessary}
-Necessary conditions for optimality are found using Lagrange
+Necessary conditions for D-optimality are found using Lagrange
 multipliers:
 \begin{align}
-  %\begin{split}
   &\nabla \tar(\obs) = \sum_{j=1}^m \xi_j \nabla \phi_j (\obs)
@@
     &\phi_j(\obs) = 0, j = 1,\dots,m. \label{eq:Lagrange mult2}
-  %\end{split}
 \end{align}
-%% This means that any optimal design satisfies \eqref{eq:Lagrange mult1}
-%% and \eqref{eq:Lagrange mult2}.
+
 Recall where the objects involved are defined: the observation operator,
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Analysis of Optimal Designs --- Vanishing Model Error}\label{section:vanishing}
+\section{Analysis of D-Optimal Designs --- Vanishing Model Error}\label{section:vanishing}
 \subsection{Overview}
-In this section we analyze optimal designs when no model error is
+In this section we analyze D-optimal designs when no model error is
 present. We discuss sensor clusterization in section
@@
 is recorded in \eqref{eq:mod conditions}. Next, we characterize
-optimal designs as ones that make all eigenvalues equal in the above
+D-optimal designs as ones that make all eigenvalues equal in the above
 mentioned nonlinear eigenvalue problem. The derivation is presented in
@@
 
-\subsection{The nonlinear Eigenvalue Problem}\label{subsec:eigenproblem}
+\subsection{The Nonlinear Eigenvalue Problem}\label{subsec:eigenproblem}
 $\modcov = 0$ implies $\Sigma= \sigma^2I$. The necessary first-order
-condition for optimality \eqref{eq:conditions} becomes
+condition for D-optimality \eqref{eq:conditions} becomes
 \begin{equation}\label{eq:eigenproblem}
@@
 $\obs$, namely $\meas_j,j=1,\dots, m$, are the eigenvectors. Note,
-however, that $\postcov$ depends on $\obs$ --- which makes
+however, that $\postcov$ depends on $\obs$. This makes
 \eqref{eq:eigenproblem} a nonlinear eigenvalue problem.
@@
 \end{proof}
-At most $m$ of the eigenvectors of $\obs^*\obs$ have positive
+At most $m$ of the eigenvectors of $\obs^*\obs$ have a positive
 eigenvalue, because $k := \rank \obs^*\obs \leq m$. Denote the
@@
 \end{align}
-up to an additive constant independent of $\obs$. Therefore, a
-D-optimal design maximizes
+and the last equality holds up to an additive constant independent of
+$\obs$. Therefore, a D-optimal design maximizes
 
@@
 
-with the norm constraints on $\meas_j,j=1,\dots,m$.
-%%  This implies that $\meas_i \in span\{\ev_j\}_{j=1}^k,i=1,\dots,
-%% m$.
-The problem of D-optimal design has now been cast to finite
-dimensions. This justifies using linear algebra. Lemma
-\ref{lemma:free} (proved in the appendix) does exactly that.
+with the norm constraints on $\meas_j,j=1,\dots,m$. The problem of
+D-optimal design has now been cast to finite dimensions. This
+justifies using linear algebra. Lemma \ref{lemma:free} (given a
+constructive proof in the appendix) does exactly that.
 \begin{restatable*}[Unit norm decomposition]{lemma}{free}\label{lemma:free}
@@
 i=1,\dots,m$ are satisified by construction. Verifying $\obs^*\obs
-\ev_i = \eta_i \ev_i$ is not too difficult:
-
-\begin{align*}
-  \begin{split}
-    \obs \ev_i &= (\meas_1,\dots,\meas_k)^t\ev_i\\
-    &= (\sum_{j=1}^kA_{j1}\ev_j^*,\dots,\sum_{j=1}^kA_{jk}\ev_j^*)^t\ev_i\\
-    &= (\sum_{j=1}^kA_{j1}\ev_j^*\ev_i,\dots,\sum_{j=1}^kA_{jk}\ev_j^*\ev_i)^t\\
-    &= (A_{i1},\dots,A_{ik})^t.
-  \end{split}
-\end{align*}
-
-Now,
-
-\begin{align*}
-  \begin{split}
-    \obs^*\obs \ev_i &= \obs^*(A_{i1},\dots,A_{ik})^t\\
-    &=(A_{i1},\dots,A_{ik}) \obs \text{ (by \eqref{eq:obs*})}\\
-    &=(A_{i1},\dots,A_{ik}) (\sum_{j=1}^kA_{j1}\ev_j^*,\dots,\sum_{j=1}^kA_{jk}\ev_j^*)^t\\
-    &=\sum_{l=1}^k A_{il} \sum_{j=1}^kA_{jl}\ev_j^* \\
-    %&=\sum_{j=1}^k  \sum_{l=1}^kA_{il}A_{jl}\ev_j^* \\
-    &=\sum_{j=1}^k  [AA^t]_{ij}\ev_j^* \\
-    %&=\eta_i\ev_i^*\\
-    &=\eta_i\ev_i^*. 
-  \end{split}
-\end{align*}
+\ev_i = \eta_i \ev_i$ is not too difficult.
+
+%% \begin{align*}
+%%   \begin{split}
+%%     \obs \ev_i &= (\meas_1,\dots,\meas_k)^t\ev_i\\
+%%     &= (\sum_{j=1}^kA_{j1}\ev_j^*,\dots,\sum_{j=1}^kA_{jk}\ev_j^*)^t\ev_i\\
+%%     &= (\sum_{j=1}^kA_{j1}\ev_j^*\ev_i,\dots,\sum_{j=1}^kA_{jk}\ev_j^*\ev_i)^t\\
+%%     &= (A_{i1},\dots,A_{ik})^t.
+%%   \end{split}
+%% \end{align*}
+
+%% Now,
+
+%% \begin{align*}
+%%   \begin{split}
+%%     \obs^*\obs \ev_i &= \obs^*(A_{i1},\dots,A_{ik})^t\\
+%%     &=(A_{i1},\dots,A_{ik}) \obs \text{ (by \eqref{eq:obs*})}\\
+%%     &=(A_{i1},\dots,A_{ik}) (\sum_{j=1}^kA_{j1}\ev_j^*,\dots,\sum_{j=1}^kA_{jk}\ev_j^*)^t\\
+%%     &=\sum_{l=1}^k A_{il} \sum_{j=1}^kA_{jl}\ev_j^* \\
+%%     %&=\sum_{j=1}^k  \sum_{l=1}^kA_{il}A_{jl}\ev_j^* \\
+%%     &=\sum_{j=1}^k  [AA^t]_{ij}\ev_j^* \\
+%%     %&=\eta_i\ev_i^*\\
+%%     &=\eta_i\ev_i^*. 
+%%   \end{split}
+%% \end{align*}
 
 We have shown that we can realize any $\{\eta_i\}_{i=1}^k$, as long as
-$\sum+{i=1}^k \eta_i = m$. We now turn to understand which choice of
+$\sum_{i=1}^k \eta_i = m$. We now turn to understand which choice of
 $\{\eta_i\}_{i=1}^k$ is optimal. By concavity of $\log$, the fastest
-increase in the design criterion (equation \eqref{eq:true target}) is
-gained by increasing $\sigma^2\lambda_i^{-1} + \eta_i$ where it is
-smallest. So, no weight should be given to (i.e. measurement taken in
-direction of) the $n$th eigenvector before $\sigma^2\lambda_i^{-1} +
-\eta_i \geq \sigma^2\lambda_n^{-1},\ \forall i \neq n$. Thus, we
-should choose $\{\eta_i\}_{i=1}^k$ as follows. First, order the
-eigenvalues $\lambda_1 \geq \lambda_2 \geq \dots$. We should increase
-$\sigma^2 \lambda_1^{-1} + \eta_1$ until it equals
+increase in the design criterion \eqref{eq:true target} is gained by
+increasing $\sigma^2\lambda_i^{-1} + \eta_i$ where it is smallest. So,
+no weight should be given to (i.e. measurement taken in direction of)
+the $n$th eigenvector before $\sigma^2\lambda_i^{-1} + \eta_i \geq
+\sigma^2\lambda_n^{-1},\ \forall i \neq n$. Thus, we should choose
+$\{\eta_i\}_{i=1}^k$ as follows. First, order the eigenvalues
+$\lambda_1 \geq \lambda_2 \geq \dots$. We should increase $\sigma^2
+\lambda_1^{-1} + \eta_1$ until it equals
 $\sigma^2\lambda_2^{-1}$. Then, $\sigma^2 \lambda_i^{-1} +
@@
 \sigma^2\lambda_3^{-1},i=1,2$ and so forth.
-
-\begin{figure}%{r}{0.25\textwidth} 
-    %\centering
-    \includegraphics[width=5cm, height=5cm]{cylinders.jpg}
-    \caption{Three graduated lab cylinders, corresponding to three
-      eigenvectors. Prior eigenvalues not shown.}
-    \label{fig:cylinder}
-\end{figure}
-
-The argument above has the important consequence (illustrated in
-figure \ref{fig:optimal vs not}) that the eigenvalues in the nonlinear
-eigenvalue problem \eqref{eq:eigenproblem} are all equal! This makes
-perfect sense. We know the Lagrange multipliers $\xi_i,i=1,\dots,k$
-tell us how much we gain by relaxing the constraints and there is no
-reason for any eigenvector to give us more than any other (otherwise
-we would put more weight on the better eigenvector, as can be
-understood from the discussion in the following paragraph).
-
-One can understand D-optimal designs as follows: Imagine each
-eigenvector corresponds to a graduated lab cylinder (figure
-\ref{fig:cylinder}). Each cylinder $i$ is filled, a-priori, with green
-liquid of $\sigma^2\lambda_i^{-1}$ volume units. This liquid
-represnets the prior precision, weighted by observation noise. The
-volume of the green liquid is increasing in $i$. A D-optimal design
-with $m$ measurements ignores cylinders indexed $m+1$ and above. It
-has blue liquid of $m$ volume units at its disposal. It distributes
-said liquid by repeatedly adding a drop to whatever cylinder currently
-has the lowest level of liquid in it (lowest precision). The result of
-such a procedure is illustrated in figure \ref{fig:optimal vs
-  not}. The analysis conducted in this section is summarized in
-Theorem \ref{thm:char} below.
-
-\begin{theorem}[Characterizing D-optimal designs with vanishing model error]\label{thm:char}
-  Let:
-  \begin{enumerate}
-  \item $\fwd, \prcov$ the forward and prior covariance operators, respectively.
-  \item $m, \sigma^2$ the number of measurements taken and observation
-    noise variance, respectively.
-  \item $\obs$ a D-optimal design operator.
-  \item $\{\lambda_i\}_{i=1}^\infty$ the eigenvalues of
-    $\fwd\prcov\fwd^*$ in decreasing order.
-  \item $\{\ev_i\}_{i=1}^\infty$ the corresponding eigenvectors.
-  \item $\{\eta_i\}_{i=1}^\infty$ eigenvalues of $\obs^*\obs$, where
-    $\obs$ maximizes $\tar(\obs)$.
-  \end{enumerate}
-  Then:
-  \begin{enumerate}
-  \item $\obs^*\obs$ and $\fwd\prcov\fwd^*$ are simultaneously diagonalizable.
-  \item There exists $0 < k \leq m$ such that $\eta_i = 0$ for $i\geq k+1$.
-  \item $\sigma^2\lambda_i^{-1} + \eta_i$ are equal for all $1 \leq i \leq k$.
-  \item If $k < m$, then $\sigma^2\lambda_i^{-1} + \eta_i \leq
-    \sigma^2\lambda_{k+1}^{-1}, i=1,\dots,k$.
-  \end{enumerate}
-\end{theorem}
 
@@
   \end{tikzpicture}
-  \caption{Characterizing an optimal design, compared to a sub-optimal
-    design. Both designs are allowed three measurements, which means
-    the blue area has accumulated height of $3$ in both plots. An
-    optimal design (left) increases precision (reduces uncertainty)
-    where it is lowest. A sub-optimal design (right) does not. In this
-    example, a measurement is taken partly in the direction of the
-    third eigenvector, where precision is bigger even compared to the
-    posterior precision of the optimal design on the left.}
+  \caption{Characterizing a D-optimal design, compared to a
+    sub-optimal design. Both designs are allowed three measurements,
+    which means the blue area has accumulated height of $3$ in both
+    plots. A D-optimal design (left) increases precision (reduces
+    uncertainty) where it is lowest. A sub-optimal design (right) does
+    not. In this example, a measurement is taken partly in the
+    direction of the third eigenvector. The precision for this
+    eigenvector is bigger even than the posterior precision of the
+    D-optimal design on the left.}
   \label{fig:optimal vs not}
@@
 }\noclusterization
+
+The argument above has the important consequence (illustrated in
+figure \ref{fig:optimal vs not}) that the eigenvalues in the nonlinear
+eigenvalue problem \eqref{eq:eigenproblem} are all equal! This makes
+perfect sense. We know the Lagrange multipliers $\xi_i,i=1,\dots,k$
+tell us how much we gain by relaxing the constraints and there is no
+reason for any eigenvector to give us more than any other (otherwise
+we would put more weight on the better eigenvector, as can be
+understood from the discussion in the following paragraph).
+
+One can understand D-optimal designs as follows: Imagine each
+eigenvector corresponds to a graduated lab cylinder (figure
+\ref{fig:cylinder}). Each cylinder $i$ is filled, a-priori, with green
+liquid of $\sigma^2\lambda_i^{-1}$ volume units. This liquid
+represnets the prior precision, weighted by observation noise. The
+volume of the green liquid is increasing in $i$. A D-optimal design
+with $m$ measurements ignores cylinders indexed $m+1$ and above. It
+has blue liquid of $m$ volume units at its disposal. It distributes
+said liquid by repeatedly adding a drop to whatever cylinder currently
+has the lowest level of liquid in it (lowest precision). The result of
+such a procedure is illustrated in figure \ref{fig:optimal vs
+  not}. The analysis conducted in this section is summarized in
+Theorem \ref{thm:char} below.
+
+\begin{theorem}[Characterizing D-optimal designs with vanishing model error]\label{thm:char}
+  Let:
+  \begin{enumerate}
+  \item $\fwd, \prcov$ the forward and prior covariance operators, respectively.
+  \item $m, \sigma^2$ the number of measurements taken and observation
+    noise variance, respectively.
+  \item $\obs$ a D-optimal design operator.
+  \item $\tar$ as definded in \eqref{eq:objective}.
+  \item $\{\lambda_i\}_{i=1}^\infty$ eigenvalues of $\fwd\prcov\fwd^*$
+    in decreasing order.
+  \item $\{\ev_i\}_{i=1}^\infty$ their corresponding eigenvectors.
+  \item $\{\eta_i\}_{i=1}^\infty$ eigenvalues of $\obs^*\obs$, where
+    $\obs$ maximizes $\tar(\obs)$.
+  \end{enumerate}
+  Then:
+  \begin{enumerate}
+  \item $\obs^*\obs$ and $\fwd\prcov\fwd^*$ are simultaneously diagonalizable.
+  \item There exists $0 < k \leq m$ such that $\eta_i = 0$ for $i\geq
+    k+1$ and $k = \rank \obs^*\obs$.
+  \item $\sigma^2\lambda_i^{-1} + \eta_i$ are equal for all $1 \leq i \leq k$.
+  \item If $k < m$, then $\sigma^2\lambda_i^{-1} + \eta_i \leq
+    \sigma^2\lambda_{k+1}^{-1}, i=1,\dots,k$.
+  \end{enumerate}
+\end{theorem}
+
+\begin{figure}%{r}{0.25\textwidth} 
+    %\centering
+    \includegraphics[width=5cm, height=5cm]{cylinders.jpg}
+    \caption{Three graduated lab cylinders, corresponding to three
+      eigenvectors. Prior eigenvalues not shown.}
+    \label{fig:cylinder}
+\end{figure}
 
@@
 \ref{subsec:seq vs sim}, the posterior becomes the prior for the next
-step. Similar reasoning leads to:
+step. The eigenvectors of the new posterior are the same as for the
+previous one. A new measurement will be taken in the direction of
+eigenvector with smalles eigenvalue. This leads to:
 \begin{align*}
@@
 \end{align*}
-In the first case, we get clusterization. In the second, we consider $\meas_3$:
+In the first case, we get clusterization immediately. In the second,
+we consider $\meas_3$. Same reasoning leads to the conclusion that the
+eigenvectors of the new prior precision are the same. The eigenvalues,
+however, are $\lambda^{-1}_1 + \sigma^{-2}, \lambda^{-1}_2 +
+\sigma^{-2}, \lambda_3^{-1}, \dots$. A new measurement is taken in the
+direction of eigenvector of smallest precision (note that
+$\lambda^{-1}_1 + \sigma^{-2} < \lambda^{-1}_2 + \sigma^{-2}$):
 \begin{align*}
   \begin{split}
-    \meas_3 \parallel \ev_1  &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 < \sigma^2\lambda^{-1}_3 \\
-    %\meas_3 \parallel  \ev_2  &\text{ if } \\
+    \meas_3 \parallel \ev_1 &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 <
+    \sigma^2\lambda^{-1}_3 \\
     \meas_3 \parallel  \ev_3  &\text{ o.w. }
@@
 this means that measurement $\meas_k$ was taken in the direction
-$\ev_k$ for $k=1,\dots,m$. Thus $\sigma^2\lambda_i^{-1} + 1 \geq
-\sigma^2\lambda_k^{-1}$ for $i=1,\dots,k-1$. This is will not happen if
-$\lambda_k^{-1}$ increase quickly --- in section \ref{subsec:abstract
-  OED} we assumed $\fwd$ is strongly smoothing, which means
-eigenvalues of $(\fwd \prcov\fwd^*)^{-1}$ grow quickly.
+$\ev_k$ for $k=1,\dots,m$. Thus:
+\begin{align}\label{eq:failure}
+  \begin{split}
+    \sigma^2\lambda_{k-1}^{-1} + 1 \geq \sigma^2\lambda_k^{-1}, \text{ for } 1\leq k \leq m
+    \Longrightarrow \lambda_m^{-1} - \lambda_{m-1}^{-1} \leq \sigma^{-2}.
+  \end{split}
+\end{align}
+
+But \eqref{eq:failure} implies $\{\lambda_i^{-1}\}_{i=1}^{\infty}$
+grows at most linearly. In section \ref{subsec:abstract OED} we
+assumed $\fwd$ is strongly smoothing, which means eigenvalues of
+$(\fwd \prcov\fwd^*)^{-1}$ grow quickly. Thus, if we take $m$ large
+enough, \eqref{eq:failure} will eventually fail and we will observe
+sensor clusterization.
 
 \subsection{Simultaneous Design}\label{subsec:clusterization simultaneous}
-In a simultaneous design scenario, the result is not as clear cut as
-in the sequential design case. We still rely on previous results ---
-namely Corollary \ref{cor:same ev} and Theorem \ref{thm:char}. An
-illustration of two D-optimal designs scenarios for the same problem
-are presented in figure \ref{fig:clusterization} --- the first
-exhibits clusterization and second does not. As was discussed earlier
-and illustrated in figure \ref{fig:optimal vs not}, an optimal design
-makes the precision ``flat'' where it is lowest. Clusterization may
-arise when two or more eigenvectors have equal precision. This
-happens, for example for eigenvectors $1$ and $2$ when considering
-measurements $\meas_i,i=1,2,3$. Taking three measurements, it is
-possible that $\meas_1$ is such that it equalizes the precision of
-eigenvectors $1$ and $2$. In this case, taking mesurements $\meas_2 =
-\meas_3 = \frac{1}{2}(\ev_1 + \ev_2)$ leads to an optimal design that
-exhibits clusterization. Similarly, taking $\meas_5 = \meas_6 =
-\frac{1}{3}(\ev_1 + \ev_2 + \ev_3)$ also gives rise to an optimal
-design that exhibits clusterization. It is important to note that
-there are (infinitely) many optimal designs that actually do not
-exhibit clusterization --- e.g. the second design presented in figure
-\ref{fig:clusterization}.
+In a simultaneous design scenario, the result is not simple as in the
+sequential design case. We now see that a design that exhibits
+clusterization is as optimal as one that does not. Corollary
+\ref{cor:same ev} tells us that we can realize any design we want
+(encoded in $\{\eta_i\}$), given the trace constraint for
+$\obs^*\obs$. Theorem \ref{thm:char} tells us that for D-optimality,
+we only need to make the posterior flat where we take measurements
+(recall figure \ref{fig:optimal vs not}). We are left with the freedom
+of assigning any part of $\eta_i$ to any measurement $\meas_j$. Two
+possible assignments for the same problem are illustrated in figure
+\ref{fig:clusterization}. The first exhibits clusterization and second
+does not. It is important to note that neither arises as a result of
+some numerical optimization scheme. Both designs are give the same
+design criterion $\tar$, because their posteriors are identical. By
+Theorem \ref{thm:char}, both are (possible) maximizers to the same
+D-optimal design problem. One exhibits clusterization, and one does
+not. It is possible that a real-life problem is more restricted than
+the model we consider here. Exactly in what way this happens is not
+clear and should be the subject of further research.
 
@@
     simultaneous D-optimal designs. Posterior precision per
-    eigenvector, after taking a measurement $\meas_i$ is plotted. An
-    optimal design reduces uncertainty where it is highest first. This
-    can be achieved by taking repeated measurements --- $\meas_2 =
-    \meas_3$ and $\meas_5 = \meas_6$ in the left design. This is not
-    necessary though, as can be seen in the right design. Both designs
-    achieve the same levels of uncertainty. The left one exhibits
-    clusterization. The right one does not.}
+    eigenvector after taking measurements $\meas_i$ is plotted. Both
+    designs are identical for all practical matters --- their
+    posteriors are identical. In the left panel we see how an optimal
+    design is achieved with repeated measurements. $\meas_2 = \meas_3$
+    and $\meas_5 = \meas_6$. This is not necessary though, as can be
+    seen in the right panel. Both designs achieve the same levels of
+    uncertainty. The left exhibits clusterization. The right does not.}
   \label{fig:clusterization}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Analysis of Optimal Designs --- Non-Vanishing Model Error}\label{section:non vanishing}
+\section{Analysis of D-Optimal Designs --- Non-Vanishing Model Error}\label{section:non vanishing}
 In this section we show the effect model error has on the
@@
 repeated measurement (in the limit $\sigma^2 \to 0$). This is clearly
-sub-optimal and $\obsm$ cannot be an optimal design. Thus, for small
+sub-optimal and $\obsm$ cannot be a D-optimal design. Thus, for small
 measurement error levels, the clusterization effect will be mitigated
@@
 limiting sense described above.
-
-%% \section{Conclusion and Future Research Directions}\label{section:conclusion}
-%% Our relaxed model gives us insight to D-optimal designs. We see that
-%% in our setting, uncertainty is not reduced in any direction
-%% (eigenvector) before it is the largest uncertainty present. We show
-%% how and why repeated measurements can give rise to a D-optimal
-%% design. There is more work to be done in understanding exactly what
-%% causes such designs to be strictly better than others.
 

### 2020-10-26 compiler.sh

@@ -2,7 +2,10 @@
 latexdiff --config="PICTUREENV=(?:picture|DIFnomarkup|figure|)[\w\d*@]*" paper.tex new_paper.tex > diff.tex
 #latexdiff --exclude-textcmd=”section,subsection,subsubsection” --config=”PICTUREENV=(?:picture|DIFnomarkup|align|tabular)[\w\d*@]*” paper.tex new_paper.tex > diff.tex
 #latexdiff paper.tex new_paper.tex > diff.tex
-latexmk -pdf diff
 #latexmk -pdf paper
-latexmk -pdf new_paper
 latexmk -pdf reply
+latexmk -pdf reply
+latexmk -pdf new_paper
+latexmk -pdf new_paper
+latexmk -pdf diff
+latexmk -pdf diff

### 2020-11-22 paper.tex
Ready for new bound paper

@@
   dimensions]{Sensor Clusterization in D-optimal Design in Infinite
-  Dimensional Bayesian Inverse Problems}
+  Dimensional Bayesian Linear Inverse Problems}
 
@@
 \begin{abstract}
-  We investigate the problem of sensor clusterization in optimal
-  experimental design for infinite-dimensional Bayesian inverse
-  problems. We suggest an analytically tractable model for such
+  We investigate the problem of sensor clusterization in D-optimal
+  experimental design for infinite-dimensional Bayesian linear inverse
+  problems. We suggest an analytically tractable model for D-optimal
   designs and reason how it may lead to sensor clusterization in the
-  case of iid measurement noise. We also show that in the case of
-  spatially correlated measurement error clusterization does not
-  occur. As a part of the analysis we prove a matrix determinant
-  lemma analog in infinite dimensions, as well as a lemma for
-  calculating derivatives of $\log \det$ of operators.
+  case of iid measurement noise. We also spatially correlated
+  measurement and model errors mitigate the clusterization
+  phenomenon. As a part of the analysis we prove a Matrix Determinant
+  Lemma in Hilbert spaces and a lemma for calculating
+  $\frac{\der}{\der t} \log \det (I + X(t))$ for certain
+  operator-valued functions. We also show how to decompose $M = AA^t$
+  with $A$ constrained to have unit norm columns.
 \end{abstract}
@@
 Designing experiments is usually done by optimizing some \emph{design
-  criterion} and this is true both for frequentists
+  criterion}. This is true both for frequentists
 \cite{Silvey13,Ucinski05} as well as for Bayesians
-\cite{ChalonerVerdinelli95}, see \cite{ChalonerVerdinelli95} for an
+\cite{ChalonerVerdinelli95}. See \cite{ChalonerVerdinelli95} for an
 investigation of the analogy between the two approaches. Although
-there is a plethora of design criteria, in the current work we will
-focus on just one of these, commonly referred to as \emph{D-optimal
-  design}. It has a simple and appealing motivation in the Bayesian
-context as explained in \cite{ChalonerVerdinelli95}: for a linear
-model, a D-optimal design seeks to maximize the expected information
-gain (KL divergence \cite{KullbackLeibler51,CoverThomas12}) between
-posterior and prior. In finite dimensions, maximizing the expected
-information gain amounts to minimizing the determinant of the
-posterior covariance matrix. In a frequentist settings, an optimal
-design minimizes the volume of the uncertainty ellipsoid \cite[page
-  16]{Ucinski05}, but this is done for the Fisher information matrix
-and not the posterior covariance. However, \cite{ChalonerVerdinelli95}
-shows that the latter is just a regularized version of the former.
+there is a plethora of design criteria, we focus on just one of these,
+commonly referred to as \emph{D-optimal design}. It has a simple and
+appealing motivation in the Bayesian context as explained in
+\cite{ChalonerVerdinelli95}: Consider a linear model in finite
+dimensions, with Gaussian prior and noise. In this model, maximizing
+the expected information gain (KL divergence
+\cite{KullbackLeibler51,CoverThomas12}) between posterior and prior
+amounts to minimizing the determinant of the posterior covariance
+matrix. In a frequentist setting, a D-optimal design minimizes the
+volume of the uncertainty ellipsoid \cite[page 16]{Ucinski05}, but
+this is done for the Fisher information matrix and not the posterior
+covariance. However, \cite{ChalonerVerdinelli95} shows that the latter
+is just a regularized version of the former.
 
 The previous discussion is classical for experimental design when
-inference is to take place over a finite (not too large) number of
+inference takes place over a finite (not too large) number of
 parameters. The subject of optimal experimental design for function
 inference in a Bayesian context was pioneered by
-\cite{AlexanderianGloorGhattas14,AlexanderianPetraStadlerEtAl16,
+\cite{AlexanderianGloorGhattas14, AlexanderianPetraStadlerEtAl16,
   AlexanderianPetraStadlerEtAl14}. Similarly to the finite dimensional
@@
 It seems counter intuitive that when one computes an optimal design
-using the D-optimal criterion (and others that will not be
-investigated here), the optimization process results in measurements
-that are very similar. For example, if a measurement is thought of as
-measuring some function value at $\x \in \Omega \subseteq \R^d,
-d=1,2,3$ (with added error) then the optimization procedure sometimes
-places sensors in close proximity to each other (as can be seen
-in figure \ref{fig:clusterization illustration}). Following
-\cite{Ucinski05}, we refer to this phenomenon as \emph{sensor
-  clusterization}.
+using the D criterion, the optimization process results in
+measurements that are very similar. For example, if a measurement is
+thought of as measuring some function value at $\x \in \Omega
+\subseteq \R^d, d=1,2,3$ (with added error) then the optimization
+procedure sometimes places sensors in very close proximity to each
+other (as can be seen in figure \ref{fig:clusterization
+  illustration}). Following \cite{Ucinski05}, we refer to this
+phenomenon as \emph{sensor clusterization}.
 
@@
 
+Similarly, in the inverse problems community, work is mostly
+computational and less theoretic. Model errors were considered in
+\cite{Attia20, Koval20}. The former study is focused on inferring a
+Quantity of Interest (QoI). The focus of the latter is reducing
+forward solves, using randomized linear algebra. Both studies present
+numerical techniques for finding optimal designs when model error is
+present. Both are restricted to linear inverse problems (although in
+the latter the authors use their method on a nonlinear problem by
+taking a Laplace approximation for the posterior). Both find an
+optimal design by first solving a continuous problem for sensor
+weights. Said solution is then sparsified to give a binary
+design. Both studies are successful in the task of Bayesian
+inversion. However, neither of these studies mention any effect model
+errors can have on sensor clusterization. The current study is mostly
+theoretical and aims to fill the gap of understanding sensor
+clusterization.
+
 \subsection{Contribution}
-We propose and thoroughly study an analytically tractable model for
-understanding D-optimal designs. In this model, an optimal design
-arises as a solution to a constrained optimization problem, formulated
-using Lagrange multipliers (section \ref{section:D and grad}). In the
-case of no model error we show this problem is in fact a set of
-eigenvector-eigenvalue problems (equation \eqref{eq:eigenproblem}). We
-then characterize an optimal design as one that reduces the highest
-uncertainties, and show how that can give rise to repeated
-measurements (section \ref{subsec:clusterization} and figure
-\ref{fig:clusterization}).
+We propose and thoroughly study a relaxed and analytically tractable
+model for understanding D-optimal designs. In this model, D-optimal
+design are solutions of a constrained optimization problem, formulated
+using Lagrange multipliers (section \ref{section:D and grad}). Using
+this model we rigorously show how model error mitigates clusterization
+(section \ref{section:non vanishing}). We show how clusterization
+occurs in a sequential design scenario (section
+\ref{subsec:clusterization sequential}). Then, we consider a
+simultaneous D-optimal design scenario (section
+\ref{subsec:clusterization simultaneous}). We show that designs that
+exhibit clusterization are just as optimal as clusterization free
+designs.
+
+A beautiful mathematical structure arises in D-optimal designs when no
+model error is present. The Lagrange multipliers problems is in fact a
+nonlinear eigenvalue problem for the observations and prior
+covariance. The operator for which eigenvectors and eigenvalues are
+sought is a sum of two operators. The first is the prior
+covariance. The second is an outer product of the observations (see
+section \ref{section:vanishing} for details and exact statement). This
+structure helps us shed light on D-optimal designs. We characterize
+these designs in Theorem \ref{thm:char}. One insight we prove is that
+a D-optimal design reduces uncertainties where they are highest
+first. Other interestig phenomena arise, but they require setting
+notation, and are discussed later.
 
 In the process we generalize several lemmas from linear algebra to
-infinite-dimensional settings --- the Matrix Determinant Lemma and a
-lemma for calculating the derivative of determinants. We also provide
-other tools for understanding D-optimal designs in
-infinite-dimensional Bayesian inverse problems, including lemmas for
-calculating the increase in the design criterion per observation ---
-Lemmas \ref{lemma:design increase} and Corollary \ref{cor:zero mod
-  err}. %%  We also show that in our model with independent measurement
-%% noise, an interesting phenomenon arises, in which the eigenvectors
-%% of the measurement operator (denoted below $\obs^*\obs$) are the
-%% same as the eigenvectors of the prior covariance in observation
-%% space, which helps arrive at the results outlined in the previous
-%% paragraph.
-The arguments presented are completely generic and apply equally to
-any of the scenarios mentioned in the introduction.
-
-%% \subsection{Limitations}\label{subsec:limitations}
-%% There are two main drawbacks of the arguments presented here. First,
-%% one can argue that we allow considerably less restrictive observations
-%% than any real-life problem does. This is because our observation model
-%% allows taking measurements that are arbitrarily weighted linear
-%% combinations of eigenvectors. The second drawback is that we do not
-%% show rigorously that clusterization necessarily occurs in our model,
-%% we only show that it can occur.
+infinite-dimensional settings. We prove a Matrix Determinant Lemma in
+\ref{lemma:MDL}. We generalize a lemma due to Lax \cite{lax97} for
+calculating $\frac{\der}{\der t} \log \det (I + X(t))$, for an
+operator valued function $X(t)$ in \ref{lemma:lax}. We also prove
+lemmas in linear algebra. One constructs a decomposition $M = AA^t$
+where $A$ has unit norm columns \ref{lemma:free}. Another shows
+simultaneous diagonizability of the prior and outer product in the
+nonlinear eigenvalue problem referenced above \ref{lemma:sim diag}. We
+also provide other tools for understanding D-optimal designs in
+infinite-dimensional Bayesian inverse problems. These includet tools
+for calculating the increase in the design criterion per observation
+--- Lemma \ref{lemma:design increase} and Corollary \ref{cor:zero mod
+  err}.
+
+\subsection{Limitations}\label{subsec:limitations}
+There are two main drawbacks of the study presented here. Our relaxed
+model does not consider any specific set of allowed
+measurements. Rather, we take measurements in the unit ball in some
+Hilbert space. This allows considerably less restrictive observations
+than any real-life problem does. The second drawback is that we do not
+show rigorously that clusterization necessarily occurs in a
+simultaneous design. We only show that it is as reasonable as no
+clusterization.
 
@@
   \end{tikzpicture}
-  \caption{Shown are pointwise posterior standard deviations for
-    various numbers of sensors, and location of these sensors for the
-    1D heat equation with homogeneous Dirichlet boundary described in
-    section \ref{subsec:example}. Note the clusterization that occurs
-    when 6 sensors are used --- only 4 measurement locations are
-    included. Observed data does not influence the pointwise
-    variances, hence it is omitted.}
-\end{figure}\label{fig:clusterization illustration}
-In section \ref{section:abstract OED} we present a more abstract and
-general formulation of the problem, but for the purpose of
-illustration, we present the problem via a toy model --- the 1D heat
-equation in $[0,\pi]$ with homogeneous Dirichlet boundary
-conditions.
+  \caption{The clusterization effect for the 1D heat equation
+    described in section \ref{subsec:example}. Posterior pointwise
+    standard deviations (lines) are plotted over the domain $[0,
+      \pi]$, for varying numbers of sensors. Sensor locations
+    (circles) were chosen to minimize (an expression analogous to) the
+    determinant of the posterior covariance. The clusterization effect
+    can be clearly seen for six sensors. Only four measurement
+    locations are used --- two pairs of sensors are so close they are
+    indistinguishable.}
+  \label{fig:clusterization illustration}
+\end{figure}
+
+In section \ref{section:prelim} we present a more abstract and general
+formulation of the inverse problem we consider. But, for the purpose
+of illustration, we present clusterization via a toy model --- the 1D
+heat equation in $[0,\pi]$ with a homogeneous Dirichlet boundary
+condition.
 
@@
 We would like to infer the initial condition $u_0$. For that purpose,
-we measure $u$ at some set of locations $\x_j \in [0,\pi], j=1, \dots,m$
-and a final time $T > 0$. We assume centered Gaussian measurement
-error, so we can observe $v(\x_j,T) = u(\x_j,T) + \eps(\x_j)$ with
-$\eps(\x_j) \sim \normal(0, \sigma^2), \sigma > 0$ iid.  We model the
-initial condition as $u_0 \sim \normal(0,\prcov)$, for $\prcov =
-(-\Delta)^{-1}$ with homogeneous Dirichlet boundary condition. It is
+we measure $u$ at some set of locations $\x_j \in [0,\pi], j=1,
+\dots,m$ and a final time $T > 0$. We assume centered Gaussian
+measurement error, so we observe $v(\x_j,T) = u(\x_j,T) + \eps(\x_j)$
+with $\eps(\x_j) \sim \normal(0, \sigma^2), \sigma > 0$ iid. We model
+the initial condition as $u_0 \sim \normal(0,\prcov)$, for $\prcov =
+(-\Delta)^{-1}$ with a homogeneous Dirichlet boundary condition. It is
 well known \cite{Tarantola05} that for linear problems, with Gaussian
@@
 $\postcov$ is known to have a closed form formula, even in infinite
-dimensions\cite{Stuart10}. If we denote by $\fwd$ the dynamics
-operator, so that $u( \cdot,T) = \fwd u_0$, and the observation
-operator $\obs$ so that $u(\x_j,T) = (\obs u)_j, j=1,\dots,m$, then the
-posterior covariance is known and depends on $\prcov, \fwd, \obs$ and
-$\sigma^2$ (see section \ref{section:abstract OED} and
-\eqref{eq:postcov} specifically).
-
-We will consider generalization of the information-theoretic design
-criterion presented in section \ref{section:OED intro} to infinite
-dimensions (section \ref{subsec:D optimal design}). We choose
-$\x_j,j=1,\dots,m$ so as to minimize an expression analogous to the
-log-determinant of the posterior covariance operator (corresponding to
-a D-optimal design). We can see the clusterization effect in figure
-\ref{fig:clusterization illustration} --- the design criterion
-suggests placing sensors in locations that are extremely close.
-
-\subsection{Remarks to the Reader}\label{subsec:plan}
-The Lagrange multipliers problem for iid measurement noise is
-discussed in section \ref{section:vanishing}. In that section, it is
-shown how sensor clusterization can occur. The reader may skip if they
-are constrained by time. It is advised that a reader less comfortable
-with Hilbert and Banach spaces mostly ignore these terms and consider
-$\hil$ and $\ban$ finite dimensional vector spaces.
-
-%% Seeking an optimal design amounts to maximizing $\tar(\obs)$ in
-%% $\obs$, subject to constraints posed on $\obs$. These enter as
-%% constraints on the allowed set of measurement vectors we can choose
-%% from --- the rows of $\obs$. These constraints are determined by the
-%% problem structure --- for example, in the 1D heat equation example
-%% considered in section \ref{subsec:example}, we were constrained to
-%% choose $\obs = (\meas_1,\dots,\meas_m)^t$ with $\meas_j = \delta_{\x_j}$
-%% for some $\x_j \in [0,\pi]$. We present an approach to understanding
-%% optimal experiment design based on the first-order necessary condition
-%% for optimal design, where the constraints are incorporated using
-%% Lagrange multipliers. This condition is used to understand what causes
-%% clusterization. One problem with such an approach is that the set we
-%% can choose $\obs$ from is complicated and difficult to work with. For
-%% example, if $\ban = C(\Omega)$ for some compact $\Omega$, then the
-%% allowed set of measurement vectors may only be point evaluations --- a
-%% set that is not easy to describe as a level set of some simple
-%% function, a description crucial for the use of Lagrange multipliers.
-%% In order to facilitate analysis, we consider an easier constraint in
-%% section \ref{subsec:unit length}. Reverting to this simpler constraint
-%% is a limitation of our analysis but it provides some insight into the
-%% cause of clusterization.
+dimensions\cite{Stuart10}. We denote by $\fwd$ the dynamics operator,
+so that $u( \cdot,T) = \fwd u_0$, and the observation operator $\obs$
+so that $u(\x_j,T) = (\obs u)_j, j=1,\dots,m$. The posterior
+covariance is known and depends only on $\prcov, \fwd, \obs$ and
+$\sigma^2$ (see section \ref{section:prelim} and \eqref{eq:postcov}
+specifically).
+
+We consider generalization of the information-theoretic design
+criterion presented in the introduction to infinite dimensions
+(section \ref{subsec:D optimal design} below). We choose
+$\x_j,j=1,\dots,m$ to minimize (an expression analogous to) the
+determinant of the posterior covariance operator. We will see later
+how this corresponds to maximizing expected information gain.
+
+The clusterization effect is illustrated in figure
+\ref{fig:clusterization illustration}. Posterior pointwise standard
+deviations are plotted over the domain $[0, \pi]$. Since the posterior
+covariance does not depend on data, the plot has no reference to
+actual data observed. The posterior covariance does, however, depend
+on location of the measurement taken. In figure
+\ref{fig:clusterization illustration}, measurement locations are
+marked by circles. These were chosen to minimize (an expression
+analogous to) the determinant of the posterior covariance. The
+clusterization effect can be clearly seen for six sensors. It looks
+like only four measurements were taken. The reason is that two pairs
+of sensors are so close they are indistinguishable.
 
@@
 
-The theoretical foundations for inverse problems over function spaces
-can be found in \cite{Stuart10}. This section is mostly devoted to
-setting notation. At this point, it is useful to record a known result
-regarding the posterior covariance operator of our inverse problem,
-namely
-\begin{align}\label{eq:postcov precursor}
-  \postcov = (\prcov^{-1} + \fwd^* \obs^* \Sigma^{-1} \obs \fwd )^{-1}.
-\end{align}
-
-\subsection{Bayesian Inverse Problems}\label{section:abstract OED}
-Let $\hil$ be a Hilbert space and $\ban$ a Banach space, both
-infinite-dimensional. Consider $\fwd: \hil \to \ban$, a linear
-operator (the ``forward operator''). We are interested in forward
-operators that are strongly smoothing (have fast decaying singular
-values --- the heat operator from section \ref{subsec:example} is a
-prime example).
-
-We want to infer $\param \in \hil$, some parameter of the dynamics
-given noisy observations of $\fwd \param$. We model $\param \sim
-\normal(0,\prcov)$ with some appropriate covariance operator $\prcov$
-\cite{Stuart10}. Observations (measurements) are taken via the linear
-observation operator $\obs \in ( \ban^* )^m$, where $m$ is the number
-of observations allowed (e.g. sensors at our disposal). In an analogy
-with linear algebra (where row vectors are thought of as linear
-functionals), we think of $\obs$ as having ``rows'' $\meas_j,
-j=1,\dots,m$:
+\subsection{Overview}
+In this section we define notations that will be used throughout the
+paper. The theoretical foundations for inverse problems over function spaces
+can be found in \cite{Stuart10}.
+
+\subsection{Bayesian Linear Inverse Problems}\label{subsec:abstract OED}
+Let $\hilp$ and $\hilo$ be separable Hilbert spaces (the subscripts p
+and o are for ``parameter'' and ``observation'', respectfully). We
+denote $\hilo^*$ the Hilbert space of all linear functionals on
+$\hilo$. Consider $\fwd: \hilp \to \hilo$, a linear operator (the
+``forward operator''). We are interested in forward operators that are
+strongly smoothing (have fast decaying modes --- the heat operator
+from section \ref{subsec:example} is a prime example). Our goal is to
+infer $\param \in \hilp$ --- some parameter of the dynamics --- given
+noisy observations of $\fwd \param$. We take a Gaussian prior $\pr$
+for $\param$. Hence, $\param \sim \pr = \normal(\prmean ,\prcov)$ with
+some appropriate covariance operator $\prcov$ \cite{Stuart10}. It is
+important to note that $\fwd \prcov \fwd^*$ is the prior covariance in
+$\hilo$ \cite{Stuart10}. As such, it is assumed invertible --- an
+assumption we rely on below. Throughout this paper, we denote $m$ the
+number of measurements we are allowed to take. Observations
+(measurements) are taken via the linear observation operator $\obs \in
+( \hilo^* )^m$ (sometimes called a design). In an analogy with linear
+algebra (where row vectors are thought of as linear functionals), we
+think of $\obs$ as having ``rows'' $\meas_j, j=1,\dots,m$:
 \begin{equation}\label{eq:O}
-  \obs = (\meas_1,\dots, \meas_m)^t, \meas_j \in \ban^*, j = 1,\dots,m.
+  \obs = (\meas_1,\dots, \meas_m)^t, \meas_j \in \hilo^*, j = 1,\dots,m.
 \end{equation}
-This way, for $u \in \ban$ we have $\obs u = (\meas_1(u), \dots,
+This way, for $u \in \hilo$ we have $\obs u = (\meas_1(u), \dots,
 \meas_m(u) )^t \in \R^m$.
 %% but we may drop the parentheses and write $\meas_j u = \meas_j(u)$.
-A few observations regarding $\obs$ are in order. First, it is good to
-keep in mind that $( \ban^* )^m$ is a Banach space with norm $\| \obs
-\| = \sum_{j=1}^m \|\meas_j\|$. Next, for $u \in \ban$ and $v\in
-\R^m$:
+%% A few observations regarding $\obs$ are in order.
+%% First, it is good to keep in mind that $( \hilo^* )^m$ is a Hilbert
+%% space with norm $\| \obs \| = \sum_{j=1}^m \|\meas_j\|$.
+For $u \in \hilo$ and $v\in \R^m$:
 \begin{align*}
@@
 
-\begin{observation}
-  It is best to think of $\meas_j,j=1,\dots,m$ as row vectors and of
-  $\obs$ as a matrix with rows $\meas_j$. 
-\end{observation}
-
-Each measurement consists of a linear functional, chosen from an
-allowed set of linear functionals in $\ban^{*}$. Data is acquired via
-noisy measurements
-\begin{align*}
-  \data := \obs (\fwd \param + \eps') + \eps = \obs \fwd \param + \obs \eps' + \eps
-\end{align*}
-with $\data \in \R^m$ and $\eps, \eps'$ defined next. Measurement
-error consists of two parts. First, there is spatially correlated
-error $\eps' \sim \normal(0,\modcov)$, modeled as a centered Gaussian
-measure on $\ban$ with covariance operator $\modcov: \ban^* \to \ban$
-(see \cite[section 6]{Stuart10} for some details on Gaussian measures
-on Banach spaces). Then there is measurement / independent error
-$\eps \sim \normal(0, \sigma^2 I_m)$, with $I_m$ the $m \times m$
-identity. Both error terms and the prior are assumed independent of
-each other. Let us now understand the terms present in the above
-expression.
+%% \begin{observation}
+%%   It is best to think of $\meas_j,j=1,\dots,m$ as row vectors and of
+%%   $\obs$ as a matrix with rows $\meas_j$. 
+%% \end{observation}
+
+Each measurement is a linear functional, chosen from some subset of
+$\hilo^{*}$. Data is acquired via noisy measurements
+\begin{align}\label{eq:inverse problem}
+  \data := \obs (\fwd \param + \eps') + \eps = \obs \fwd \param + \obs \eps' + \eps,
+\end{align}
+with $\data \in \R^m$ and $\eps, \eps'$ defined next. We consider two
+types of noise / error terms. First, there is spatially correlated
+model error $\eps' \sim \normal(0,\modcov)$, modeled as a centered
+Gaussian measure on $\hilo$ with covariance operator $\modcov$. Then
+there is measurement / independent error $\eps \sim \normal(0,
+\sigma^2 I_m)$, with $I_m$ the $m \times m$ identity. Both error terms
+and the prior are assumed independent of each other. The observation
+and forward operators are merged $\tmp := \obs \fwd$. Since we need to
+separate the observation operator $\obs$ from the forward operator
+$\fwd$ we write this explicitly in \eqref{eq:inverse problem}. We now
+turn to understand the terms present in \eqref{eq:inverse problem}.
 
 \subsubsection{Dynamics and Observation Operators}\label{subsec:dynamics}
-One can think of $(\obs \fwd \param)_j$ as pointwise evaluations of a
-continuous function if $\ban$ is some appropriate function space,
-e.g.\ $C(\Omega)$, the Banach space of real-valued continuous
-functions over a compact $\Omega$ with the $\| \cdot \|_{\infty}$
-norm. These measurements can be represented using the linear
-functionals $\delta_{\x_j} \in C(\Omega)^*$. In a different setting,
-we may be able to measure other quantities, e.g.\ some Fourier
-coefficients as is the case in MRI applications. Either way, the rows
-of $\obs$ are linear functionals and we refer to these as {\it
-  measurement vector}s. Note that we cannot choose every measurement
-vector we want, though. For example, we may be restricted only to
-pointwise evaluations of $\fwd \param$ by the sensors at our disposal,
-in which case measuring the mean $\ell(u) = \int_{\Omega}u$ (or any
-non-atomic measure) of a function is not possible.
+One can \textit{think} of $(\obs \fwd \param)_j$ as pointwise
+evaluations of a continuous function\footnote{Of course, this is
+  impossible if, e.g., $\hilo = L^2(\Omega)$ for some $\Omega$.}. Such
+measurements can be represented using the linear functionals
+$\delta_{\x_j} \in C(\Omega)^*$. In a different setting, we may be
+able to measure other quantities, e.g.\ some Fourier coefficients as
+is the case in MRI applications. Either way, the rows of $\obs$ are
+linear functionals and we refer to these as {\it measurement
+  vector}s. Note that we cannot choose every measurement vector we
+want. For example, we may be restricted only to pointwise evaluations
+of $\fwd \param$ by the sensors at our disposal, in which case
+measuring, e.g., the mean $\ell(u) = \int_{\Omega}u$ of a function is
+not possible.
 
 \subsubsection{Error Terms}
-Considering the sum of the error terms it is easy to see that $\obs
-\eps' + \eps \in \R^m$. It is a centered Gaussian random vector and
-its covariance matrix is
+Considering the sum of the error terms it is easy to see that
+$\bar{\eps} := \obs \eps' + \eps \in \R^m$. It is a centered Gaussian
+random vector and its covariance matrix is
 \begin{align}\label{eq:Sigma}
@@
 \end{equation}
-where the first equality in \eqref{eq:modcov explained} holds by
-definition and the second by \eqref{eq:obs*}. The explicit dependence
-on $\obs$ will be mostly dropped for notational convenience, so
-$\Sigma(\obs) = \Sigma$. Thus, for a fixed $\obs$ (i.e.\ a fixed set
-of measurements) we can equivalently write $\data = \obs \fwd \param +
-\bar{\eps}$ with $\bar{\eps} \sim \normal(0,\Sigma)$. At this point,
-it is useful to rewrite the posterior covariance operator of our
-inverse problem from \eqref{eq:postcov precursor}
-\begin{align}\label{eq:postcov}
-  \postcov = (\prcov^{-1} + \fwd^* \obs^* \Sigma^{-1} \obs \fwd
-  )^{-1}.
-\end{align}
-
-%% Usually, this is written using $\tmp := \obs \fwd$ and $\tmp^*$ but
-%% since we need to separate the observation operator $\obs$ from the
-%% forward operator $\fwd$ we write the explicit expression
-%% \eqref{eq:postcov}.
-Taking $\modcov = 0$ is common practice
-\cite{Tarantola05,KaipioSomersalo05,Vogel02} and then we are reduced
-to the case where we take iid observations and $\Sigma = \sigma^2I$ is
+with $e_j \in \R^m$ the $j$th standard basis vector. The first
+equality in \eqref{eq:modcov explained} holds by definition and the
+second by \eqref{eq:obs*}. The explicit dependence on $\obs$ will be
+mostly dropped for notational convenience, so $\Sigma(\obs) =
+\Sigma$. Thus, for a fixed $\obs$ (i.e.\ a fixed set of measurements)
+we can equivalently write $\data = \obs \fwd \param + \bar{\eps}$ with
+$\bar{\eps} \sim \normal(0,\Sigma)$. Taking $\modcov = 0$ is common
+practice \cite{Tarantola05,KaipioSomersalo05,Vogel02}. This reduces to
+the case where we take iid observations. Then $\Sigma = \sigma^2I$ is
 simply a scalar matrix that does not depend on $\obs$. Note that
@@
 we may express this belief by saying $\tru = \fwd + \err$, with $\fwd$
-depending on $\param$ and $\err$ depending on $\sspar$, and $\param
+depending on $\param$ and $\err$ depending on $\sspar$, with $\param
 \perp \sspar$. We do not know much about this effect but it is
@@
 a numerical source, where our discretization of the system is not fine
-enough to capture all small scale phenomena. We will see that assuming
-some correlation in the error may reduce the clusterization
-phenomenon, as is also reported in the literature \cite{Ucinski05}.
-
-\subsection{D-Optimal Design in Infinite Dimensions}\label{subsec:D optimal design} 
+enough to capture all small scale phenomena. In section
+\ref{section:non vanishing}, we will see that assuming some
+correlation in the error mitigates the clusterization phenomenon, as
+reported in the literature \cite{Ucinski05}.
+
+Finally, it is useful to record a known result regarding the posterior
+covariance operator of our inverse problem:
+\begin{align}\label{eq:postcov}
+  \postcov = (\prcov^{-1} + \fwd^* \obs^* \Sigma^{-1} \obs \fwd
+  )^{-1}.
+\end{align}
+
+\subsection{D-Optimal Designs in Infinite Dimensions}\label{subsec:D optimal design} 
+A D-optimal design maximizes expected KL divergence between posterior
+and prior. It is useful to recall the definition of KL divergence for
+an arbitrary prior measure:
+$$
+KL(\post||\pr) = \int \log \frac{\der \post}{\der \pr}(\param) \der \post(\param).
+$$
+
 The meaning of D-optimal design in infinite-dimensional Hilbert spaces
-was investigated in \cite{AlexanderianGloorGhattas14}. Note that the
+was investigated in \cite{AlexanderianGloorGhattas14}. There, the
 authors make assumptions that amount to $\Sigma=I$ (implied by
@@
 infinite-dimensions, a D-optimal design is well-defined as maximizing
-the expected KL divergence between posterior and prior and is
-summarized, using our notation, in the following theorem:
+the expected KL divergence between posterior and prior. The main
+result is summarized, using our notation, in the following theorem:
 \begin{theorem}[Slightly modified Theorem 1 from \cite{AlexanderianGloorGhattas14}]
-  Let $\pr = \normal(0,\prcov)$ be a centered Gaussian prior
-  on $\hil$   and let $\post = \normal(\postmean,\postcov)$ 
-  the posterior measure on $\ban$ for the Bayesian linear
-  inverse problem $\data =  \obs \fwd\param + \obs \eps' + \eps$ discussed
-  above. Then 
+  Let $\pr = \normal(\prmean,\prcov)$ be a Gaussian prior on $\hilp$
+  and let $\post = \normal(\postmean,\postcov)$ the posterior measure
+  on $\hilo$ for the Bayesian linear inverse problem $\data = \obs
+  \fwd\param + \obs \eps' + \eps$ discussed above. Then
   \begin{align}\label{eq:objective}
@@
 \end{theorem}
-
-\section{The Constrained Optimization Problem of D-Optimal Design}\label{section:D and grad}
-From \eqref{eq:objective} we wish to characterize solution(s) of the
-following optimization problem for $\tar$. %%: (\ban^*)^m \to \R$:
+\begin{definition}\label{def:d optimality}
+  A design $\obs^{\star}$ is said to be D-optimal if $\obs^{\star} = \argmax_{\obs} \tar(\obs)$.
+\end{definition}
+
+The intuition behind definition \ref{def:d optimality} is
+straightforward. From \eqref{eq:objective} and \eqref{eq:objective}:
+\begin{align*}
+  \begin{split}
+    \det ( I + \prcov^{1/2}  \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}) &= \det \Big( \prcov ( \prcov^{-1} + \fwd ^* \obs^* \Sigma^{-1} \obs \fwd) \Big )\\
+    &= \det \prcov \det \postcov^{-1}.
+  \end{split}
+\end{align*}
+Since $\prcov$ is constant, a D-optimal design minimizes the posterior
+covariance determinant, analogously to the finite-dimensional case.
+
+\subsection{Sequential vs Simultaneous Optimization}\label{subsec:seq vs sim}
+From defintion \ref{def:d optimality} we wish to characterize solution(s) of the
+following optimization problem for $\tar$. %%: (\hilo^*)^m \to \R$:
 \begin{align}\label{eq:optimization}
@@
 \end{align}
-where $\obs$ is constrained to some allowed set of measurements
-(e.g.\ pointwise evaluations on $[0,\pi]$)
-%%, for the example in section \ref{subsec:example}).
-We seek a formulation of the optimal design using Lagrange
-multipliers. In section \ref{section:objective} we find the gradient
-of $\tar$.  In section \ref{subsec:unit length} we relax the
-constraints on $\obs$ so that the optimization problem in
-\eqref{eq:optimization} and find gradients for the new constraints.
+where $\obs$ is constrained to some allowed set of measurements (we
+may \textit{think} of these as pointwise evaluations, even though
+these might not be in $\hilo^*$). We call this problem ``simultaneous
+optimization'', since all measurements are decided on simulatneously.
+
+For computational reasons, one may prefer to find the best
+measurements in a sequential manner. Denote
+\begin{equation}\label{eq:def obs_k}
+  \obs_k := (\meas_1,\dots, \meas_k)^t,  k\leq m.
+\end{equation}
+Sequential optimal design proceeds as follows. Find $\meas_1$ by
+maximizing $\tar(\obs_1)$. Then, keeping $\meas_1$ fixed --- find
+$\meas_2$ as the maximizer of $\tar(\obs_2)$. Then, find $\meas_3$ by
+keeping $\meas_1,\meas_2$ fixed and taking $\meas_3$ as the maximizer
+of $\tar(\obs_3)$. Continue this way until $\obs_m = \obs$ is
+found, where $m$ is the number of available measurements. It is
+important to notice that this scheme does not require actually
+observing data --- in \eqref{eq:objective} data is averaged out.
+
+The analysis in this paper is conducted for the general simultaneous
+optimization case. The sequential optimization case is dealt with in
+section \ref{subsec:clusterization sequential}. It is important to
+note, however that all conclusions we arrive at for the simultaneous
+case easily specialize to the sequential case by considering the
+posterior as the next sequential step's prior.
+
+\section{The Constrained Optimization Problem of D-Optimal Design}\label{section:D and grad}
+
+\subsection{Overview}
+We seek a formulation of the D-optimal design problem
+\eqref{eq:optimization} using Lagrange multipliers. In section
+\ref{section:objective} we find the gradient of $\tar$ and record it
+in \eqref{eq:tar grad}. In section \ref{subsec:unit length} we suggest
+using unit-norm constraints on rows of $\obs$ so that the optimization
+problem \eqref{eq:optimization} is analytically tractable. We find
+gradients for the new constraints and record them in \eqref{eq:grad
+  constraints}. Finally, in section \ref{subsec:necessary} we use
+these to formulate the D-optimal design problem as a Lagrange
+multipliers problem in equation \eqref{eq:conditions}.
 
@@
 We start by calculating the first variation. Some calculations are
-delegated to Lemma \ref{lemma:aux calc} in the appendix. Also, we
-use the following lemma whose proof is also delegated to the appendix:
-\begin{lemma}[Generalized from \cite{Lax97}]\label{lemma:lax}
+delegated to Lemma \ref{lemma:aux calc} in the appendix. We use
+the following lemma:
+\begin{restatable*}[Generalized from \cite{Lax97}]{lemma}{lax}\label{lemma:lax}
   Let $Y(t)$ be a differentiable operator-valued function. Assume 
@@
   \end{equation*}
-\end{lemma}
-
-First variation of $\tar (\obs)$ in the direction $V$ is:
-\begin{align*}%%\label{eq:tar var}
+\end{restatable*}
+
+The proof of \ref{lemma:lax} is delegated to the appendix. Let
+$T(\obs) := \obs^*\Sigma(\obs)\obs$. The first variation of $\tar
+(\obs)$ in the direction $V$ is:
+\begin{align}\label{eq:tar var}
   \begin{split}
@@
     &= \frac12 \frac{\der}{\der \tau} \Big |_{\tau=0} \log \det 
-    (I + \prcov^{1/2} \fwd^* T(\obs+\tau V)\fwd \prcov^{1/2} ) \text{ (by definition \eqref{eq:objective})}\\
-    % 
-    % 
-    % 
-    &= \frac12 \tr ( I + \prcov^{1/2} \fwd^* \obs^* \Sigma^{-1}
+    (I + \prcov^{1/2} \fwd^* T(\obs+\tau V)\fwd \prcov^{1/2} ) \text{ (by \eqref{eq:objective})}\\
+    % 
+    % 
+    % 
+    &= \frac12 \tr{( I + \prcov^{1/2} \fwd^* \obs^* \Sigma^{-1}
     \obs\fwd \prcov^{1/2} )^{-1}
     \frac{\der}{\der \tau} \Big |_{\tau=0}
-    \prcov^{1/2} \fwd^* T(\obs+\tau V) \fwd \prcov^{1/2}\ \text{ (by \ref{lemma:lax})}\\
-    % 
-    % 
-    % 
-    &= \frac12 \tr \postcov \fwd^* (
-      V^* \Sigma^{-1} \obs 
-      - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \text{ (by \ref{lemma:aux calc})}\\
+    \prcov^{1/2} \fwd^* T(\obs+\tau V) \fwd \prcov^{1/2}}\ \text{ (by \ref{lemma:lax})}\\
+    % 
+    % 
+    % 
+    &= \frac12 \ttr\Big \{ \postcov \fwd^* (V^* \Sigma^{-1} \obs 
+    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
     &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs 
-      + \obs^* \Sigma^{-1} V ) \fwd \\
-%%   \end{split}
-%% \end{align} 
-%% Taking adjoints and cyclic permutations we see that
-%% \begin{align*}
-%%   \tr \postcov \fwd^* V^* \Sigma^{-1} \obs \fwd 
-%%   &= \tr \postcov \fwd^* \obs^* \Sigma^{-1} V \fwd,
-%% \end{align*}
-%% and
-%% \begin{align*}
-%%   \tr \postcov \fwd^* \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \fwd
-%%   &= \tr \postcov \fwd^*  \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs \fwd.
-%% \end{align*}
-%% We use these in the first equality of the following derivation,
-%% which continues the one in \eqref{eq:precursor}.
-%% %
-%% \begin{align} \label{eq:tar var}
-%%   \begin{split}
-      %% \delta \tar(\obs) V 
-    &= \tr \postcov \fwd^* ( \obs^* \Sigma^{-1} V 
-    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs ) \fwd \\
-    %
-    %
-    % 
-    &= \tr \postcov \fwd^* \obs^* \Sigma^{-1} V 
-    ( I - \modcov \obs^* \Sigma^{-1}\obs ) \fwd \\
-    % 
-    %
-    %
-    &= \tr V ( I - \modcov \obs^* \Sigma^{-1}\obs )
-    \fwd \postcov \fwd^* \obs^* \Sigma^{-1}.
+    + \obs^* \Sigma^{-1} V ) \fwd \Big \}  \text{ (by \ref{lemma:aux calc})}\\
+    %
+    %
+    %
+    &= \tr{\postcov \fwd^* ( \obs^* \Sigma^{-1} V -
+    \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs ) \fwd} \text{
+        (cyclic property of trace)} \\
+    %
+    %
+    % 
+    &= \tr{\postcov \fwd^* \obs^* \Sigma^{-1} V 
+    ( I - \modcov \obs^* \Sigma^{-1}\obs ) \fwd} \\
+    % 
+    %
+    %
+    &= \tr{V ( I - \modcov \obs^* \Sigma^{-1}\obs )
+    \fwd \postcov \fwd^* \obs^* \Sigma^{-1}}.
   \end{split}
-\end{align*} 
+\end{align} 
 Denote
@@
 \begin{align*}
-  \delta \tar(\obs)V = \tr V \nabla \tar(\obs) = \sum_{j=1}^m
+  \delta \tar(\obs)V = \tr{V \nabla \tar(\obs)} = \sum_{j=1}^m
   V_j(\nabla \tar(\obs)_j),
 \end{align*}
-with $V_j \in \ban^*$ and $\nabla \tar(\obs)_j \in \ban^{**}\subseteq
-\ban, j=1,\dots,m$. Thus, $\nabla \tar( \obs ) \in (\ban^{**})^m
-\subseteq \ban^m$ is indeed the correct gradient and the notation
-\eqref{eq:tar grad} is justified. Note that by the comment following
-\eqref{eq:obs*}, we view $V$ (defined on the same space as $\obs$) as
-a column vector. Then the gradient $\nabla \tar(\obs)$ should be
-viewed as a row vector, as the product $V \nabla \tar$ is in $\R^{m
-  \times m}$. This will prove important in section
-\ref{subsec:necessary}.
+with $V_j \in \hilo^*$ and $\nabla \tar(\obs)_j \in \hilo^{**} =
+\hilo, j=1,\dots,m$. Thus, $\nabla \tar( \obs ) \in \hilo^m$ is indeed
+the correct gradient and the notation \eqref{eq:tar grad} is
+justified. Note that by the comment following \eqref{eq:obs*}, we view
+$V$ (defined on the same space as $\obs$) as a column vector. Then the
+gradient $\nabla \tar(\obs)$ should be viewed as a row vector, as the
+product $V \nabla \tar$ is in $\R^{m \times m}$. This will prove
+important in section \ref{subsec:necessary}.
 
@@
 In this section we suggest relaxed constraints in
-\eqref{eq:constraints} and derive their gradient in \eqref{eq:grad
-  constraints}, for the purpose of using them in a Lagrange
-multipliers problem.
+\eqref{eq:constraints}. We then find their gradient in \eqref{eq:grad
+  constraints}. These are use in the Lagrange multipliers
+problem in section \ref{subsec:necessary}.
 
 As mentioned before, we cannot choose any $\obs$ when maximizing
-$\tar(\obs)$. Recall $\obs = (\meas_1,\dots,\meas_m)^t \in (\ban^*)^m$
-and each $\meas_j,j=1,\dots,m$ must be chosen from some allowed set of
-functionals in $\ban^*$. This set differs based on the kind of sensors
-we have at our disposal and the properties of $\ban$, as discussed in
-section \ref{subsec:dynamics}. The following proposition will give us
-better understanding of the constraints. 
-\begin{proposition}\label{prop:bigger better}
+$\tar(\obs)$. Recall $\obs = (\meas_1,\dots,\meas_m)^t \in
+(\hilo^*)^m$ and each $\meas_j,j=1,\dots,m$ must be chosen from some
+allowed set of functionals in $\hilo^*$. This set differs based on the
+kind of sensors we have at our disposal and the properties of $\hilo$,
+as discussed in section \ref{subsec:dynamics}. The following
+proposition gives better understanding of the constraints.
+\begin{restatable*}{proposition}{biggerbetter}\label{prop:bigger better}
   Let $\obs = (\meas_1,\dots,\meas_m)^t$, $j \in \{1,\dots,m\}$, $\sigma^2
@@
   $\lambda \meas_j$ in $\obs$ instead of $\meas_j$.
-\end{proposition}
+\end{restatable*}
 Proof of Proposition \ref{prop:bigger better} is delegated to the
 appendix but the idea behind it is simple: making rows of $\obs$ large
-is equivalent to making $\sigma^2$ small. This can be understood
-easily from the formulation of the inverse problem $\data = \obs \fwd
-\param + \obs \eps' + \eps$, where $\eps'$ can be taken as model error
-and $\eps$ is iid measurement error.
-
-We now suggest a more convenient-to-work-with set of constraints. By
-Proposition \ref{prop:bigger better}, we cannot just take all $\ban^*$
-to be the allowed set of measurement vectors, since this would
-effectively eliminate measurement error entirely. We note that for
-point evaluations, the norm of the measurement is always one:
+is equivalent to making $\sigma^2$ small. %%  Viewed another way, it
+%% increases the Signal to Noise Ratio (SNR) --- which is always
+%% desirable when taking any measurement prone to error.
+This can be understood easily from the formulation of the inverse
+problem $\data = \obs \fwd \param + \obs \eps' + \eps$, where $\eps'$
+can be taken as model error and $\eps$ is iid measurement error.
+
+We now suggest a set of constraints that will make the model
+analytically tractable. By Proposition \ref{prop:bigger better}, we
+cannot just take all $\hilo^*$ to be the allowed set of measurement
+vectors, since this would effectively eliminate measurement error
+entirely. We note that for point evaluations, the norm of the
+measurement is always one:
 \begin{align*}
@@
 Following this, we also restrict the norm of our measurement vectors
-to be one. But the norm on functionals is hard to work with. Instead,
-we assume there exists a Hilbert space $\bar{\hil} \subseteq \ban^*$
-and take measurements $\meas_j \in \bar{\hil}^* = \bar{\hil}$ such
-that $\| \meas_j\| = 1, j=1,\dots,m$. Recall that we think of $\ban$ as
-a somewhat restricted space ($C(\bar{\Omega})$ with the sup-norm over
-a bounded domain $\Omega$, for example) and so $\ban^*$ can be thought
-of as large. Then $\ban^*$ can accommodate a reasonable Hilbert space
-$\bar{\hil}$. For example, if $\ban = C(\bar{\Omega})$, we take
-$\bar{\hil} = L^2(\Omega) \subseteq \ban^*$.
-%% (we can ignore the
-%% boundary since is has zero measure).
-Then
-%
-\begin{equation*}
-  \ban \subseteq \bar{\hil} \subseteq \ban^*,
-\end{equation*}
-since members of $C(\bar{\Omega})$ are bounded, hence square integrable. The
-unit norm constraint can be written using \eqref{eq:obs*} as a series of $m$ equality
-constraints (one for each measurement) on $\obs$ as
+to be one. The unit norm constraints can be written using
+\eqref{eq:obs*} as a series of $m$ equality constraints (one for each
+measurement) on $\obs$ as
 \begin{align}\label{eq:constraints}
-  \phi_j(\obs) :=\frac12 \| \obs^* e_j\|_{\bar{\hil}}^2 - \frac12 = 0,j=1,\dots,m
+  \phi_j(\obs) :=\frac12 \| \obs^* e_j\|_{\hilp}^2 - \frac12 = 0,j=1,\dots,m.
 \end{align}
-with $e_j \in \R^m$ the $j$th standard basis vector. The first variations are:
+The first variations are:
 \begin{align*}
@@
   &= \frac12\lim_{\tau \to 0}\tau^{-1}
-  ( \|(\obs + \tau V)^*e_j \|_{\bar{\hil}}^2 - \|\obs ^*e_j \|_{\bar{\hil}}^2  ) \\
+  ( \|(\obs + \tau V)^*e_j \|_{\hilp}^2 - \|\obs ^*e_j \|_{\hilp}^2  ) \\
   %
@@
   &= \frac12\lim_{\tau \to 0}\tau^{-1}
-  ( \langle (\obs + \tau V)^*e_j, (\obs + \tau V)^*e_j \rangle_{\bar{\hil}} - 
-  \langle \obs^*e_j, \obs^*e_j \rangle_{\bar{\hil}} ) \\
+  ( \langle (\obs + \tau V)^*e_j, (\obs + \tau V)^*e_j \rangle_{\hilp} - 
+  \langle \obs^*e_j, \obs^*e_j \rangle_{\hilp} ) \\
   % 
@@
   &= \frac12\lim_{\tau \to 0}\tau^{-1}
-  (2\tau \langle \obs^*e_j,V^*e_j \rangle_{\bar{\hil}} 
-  +\tau^2 \langle V^*e_j, V^*e_j \rangle_{\bar{\hil}} ) \\
-  %
-  %
-  % 
-  &= \langle \obs^*e_j,V^*e_j \rangle_{\bar{\hil}} \\
+  (2\tau \langle \obs^*e_j,V^*e_j \rangle_{\hilp} 
+  +\tau^2 \langle V^*e_j, V^*e_j \rangle_{\hilp} ) \\
+  %
+  %
+  % 
+  &= \langle \obs^*e_j,V^*e_j \rangle_{\hilp} \\
   %
@@
     &\phi_j(\obs) = 0, j = 1,\dots,m. \label{eq:Lagrange mult2}
-  %\end{split}
 \end{align}
-%% This means that any optimal design satisfies \eqref{eq:Lagrange mult1}
-%% and \eqref{eq:Lagrange mult2}.
+
 Recall where the objects involved are defined: the observation operator,
-$\obs \in \bar{\hil}^m \subseteq (\ban^*)^m$. The objective $\tar:
-(\ban^*)^m \to \R$ and its gradient is $\nabla \tar(\obs) \in
-\ban^m$. The constraints are defined similarly --- $\phi_j: (\ban^*)^m
-\to \R$ and $\nabla \phi_j(\obs) \in \ban^m$.
-
-Focusing on \eqref{eq:Lagrange mult1} and using the gradients of the
-objective and constraints \eqref{eq:tar grad} and \eqref{eq:grad
-  constraints} we have:
+$\obs \in (\hilo^*)^m$. The objective $\tar:
+(\hilo^*)^m \to \R$ and its gradient is $\nabla \tar(\obs) \in
+\hilo^m$. The constraints are defined similarly --- $\phi_j: (\hilo^*)^m
+\to \R$ and $\nabla \phi_j(\obs) \in \hilo^m$.
+
+Substituting \eqref{eq:tar grad} and \eqref{eq:grad constraints} in
+\eqref{eq:Lagrange mult1} we have:
 \begin{equation*}
@@
 \end{equation*}
-If we let $\Xi = \diag(\xi_j)$, then with some abuse of
+If we let $\Xi := \diag(\xi_j)$, then with some abuse of
 notation, we can write this more compactly as:
@@
 
-\section{Analysis of Optimal Designs --- Vanishing Model Error}\label{section:vanishing}
-When $\modcov = 0$, then $\Sigma= \sigma^2I$. The necessary
-first-order condition for optimality \eqref{eq:conditions} becomes
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% SECTION Analysis of Optimal Designs --- Vanishing Model Error
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Analysis of D-Optimal Designs --- Vanishing Model Error}\label{section:vanishing}
+\subsection{Overview}
+In this section we analyze D-optimal designs when no model error is
+present. Sensor clusterization is discussed in section
+\ref{section:clusterization}. We prepare the ground by first deriving
+a nonlinear eigenvalue problem in $\obs$ for D-optimal designs in
+section \ref{subsec:eigenproblem}. This nonlienar eigenvalue problem
+is recorded in \eqref{eq:mod conditions}. Next, we characterize
+D-optimal designs as ones that make all eigenvalues equal in the above
+mentioned nonlinear eigenvalue problem. The derivation is presented in
+section \ref{subsec:characterization} and is summarized in Theorem
+\ref{thm:char}.
+
+\subsection{The Nonlinear Eigenvalue Problem}\label{subsec:eigenproblem}
+$\modcov = 0$ implies $\Sigma= \sigma^2I$. The necessary first-order
+condition for D-optimality \eqref{eq:conditions} becomes
 \begin{equation}\label{eq:eigenproblem}
@@
 \end{equation}
-with $\Xi$ diagonal. This is an eigenvalue problem for the
-self-adjoint operator $\sigma^{-2}\fwd \postcov \fwd^*$ where the rows
-of $\obs$, namely $\meas_j,j=1,\dots, m$, are the eigenvectors. If $\fwd$
-is invertible, then we may write:
+with $\Xi$ diagonal. This looks like an eigenvalue problem for the
+self-adjoint operator $\sigma^{-2}\fwd \postcov \fwd^*$. The rows of
+$\obs$, namely $\meas_j,j=1,\dots, m$, are the eigenvectors. Note,
+however, that $\postcov$ depends on $\obs$. This makes
+\eqref{eq:eigenproblem} a nonlinear eigenvalue problem.
+
+Below, the expression for the posterior covariance \eqref{eq:postcov}
+justifies the second equality. Some calculations delegated to Lemma
+\ref{lemma:twice woodbury} in the appendix justify the third (it is
+trivial if $\fwd$ is invertible):
 \begin{align}\label{eq:mod conditions}
@@
   %
-  &= \sigma^{-2} \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1} \obs^*,
+  &= \sigma^{-2} \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1} \obs^*
   \end{split}
 \end{align}  
-which presents the optimal design problem as a problem on $\ban$ and
-not on $\hil$ --- the term $(\fwd \prcov \fwd^*)^{-1}$ is the prior
-precision in $\ban$ and the big parentheses enclose the posterior
-precision in $\ban$. This expression makes sense even if $\fwd$ is not
-invertible --- consider instead $\fwd$ restricted to $\ker
-\fwd^{\perp}$. This does not change the optimal design problem at all,
-since we will never take measurements that have a component in $\ker
-\fwd$ (this can be understood from Corollary \ref{cor:zero mod err}).
-
-\subsection{Characterizing D-Optimal Designs}\label{subsub:characterization}
-In light of \eqref{eq:mod conditions}, let us denote eigenvalues of
-$\fwd \prcov \fwd^*$ by $\lambda_1 \geq \lambda_2 \geq \dots$ and its
-eigenvectors $\{\ev_i\}_{i=1}^{\infty}$. The following lemma will be useful:
-
-\begin{lemma}[Proved as lemma \ref{lemma:sim diag} in the appendix]
-  Let $C:\hil \to \hil$ self-adjoint and let $\func_1,\dots,\func_m \in
-  \hil$. Denote $\func^*$ the element $\func$ acting as a linear
-  functional. If
+The term $(\fwd \prcov \fwd^*)^{-1}$ is the prior precision in $\hilo$
+($\fwd\prcov\fwd^*$ is invertible, as mentioned in section
+\ref{section:prelim}). $(\fwd\prcov\fwd^*)^{-1} + \sigma^{-2} \obs^*
+\obs$ is the posterior precision in $\hilo$.
+
+\subsection{Characterizing D-Optimal Designs}\label{subsec:characterization}
+Denote eigenvalues of $\fwd \prcov \fwd^*$ by
+$\{\lambda_i\}_{i=1}^{\infty}$. These are all positive. Denote
+corresponding eigenvectors $\{\ev_i\}_{i=1}^{\infty}$. Lemma
+\ref{lemma:sim diag} (proved in the appendix) is crucial for the rest
+of the analysis:
+
+\begin{restatable*}[Simultaneous diagonizability for the nonlinear eigenvalue problem]{lemma}{simdiag}\label{lemma:sim diag}
+  Let $\hil$ separable Hilbert space, $C:\hil \to \hil$ self-adjoint
+  and $\func_1,\dots,\func_m \in \hil$. Denote $\func^*$ the element
+  $\func$ acting as a linear functional. If
   \begin{equation*}
-   (C + \sum_{j=1}^m \func_j\func_j^*) \func_k = \xi_k \func_k, k = 1,\dots,m
+   (C + \sum_{j=1}^m \func_j\func_j^*) \func_l = \xi_l \func_l, l = 1,\dots,m
   \end{equation*}
@@
   diagonalizable.
-\end{lemma}
-
-Thus, if we take $\func_j^{*} = \meas_j$ and $C := (\fwd \prcov
-\fwd^*)^{-1}$ in the above lemma, we conclude from the eigenproblem
-\eqref{eq:mod conditions} that $\obs^*\obs$ has the same eigenvectors
-as $\fwd \prcov \fwd^*$, namely $\{\ev_i\}_{i=1}^{\infty}$ (infinitely
-many of them with zero eigenvalue, since $\obs^*\obs$ is
-finite-rank). Denote the corresponding non-zero eigenvalues of
-$\obs^*\obs$ by $\{\eta_i\}_{i=1}^{k}$ and let $\eta_i = 0$ for $i
-\geq k+1$. By the argument in the paragraph following \eqref{eq:mod
-  conditions}:
+\end{restatable*}
+
+\begin{corollary}[Eigenvectors of a D-optimal design]\label{cor:same ev}
+  $\obs^*\obs$ has the same eigenvectors as $\fwd \prcov \fwd^*$.
+  %namely $\{\ev_i\}_{i=1}^{\infty}$.
+\end{corollary}
+\begin{proof}
+In \eqref{eq:mod conditions}, take $\func_j^{*} = \meas_j$ and $C :=
+(\fwd \prcov \fwd^*)^{-1}$ and use Lemma \ref{lemma:sim diag}.
+\end{proof}
+At most $m$ of the eigenvectors of $\obs^*\obs$ have a positive
+eigenvalue, because $k := \rank \obs^*\obs \leq m$. Denote the
+corresponding non-zero eigenvalues of $\obs^*\obs$ by
+$\{\eta_i\}_{i=1}^{k}$ and let $\eta_i = 0$ for $i \geq k+1$. No
+assumption is made (yet) regarding the magnitudes of the corresponding
+$\{\lambda_i\}_{i=1}^{\infty}$. They are to be thought of as fixed but
+unknown. Using Sylvester's Determinant Theorem (the standard proof
+works for Hilbert spaces as well) we get:
+
 \begin{align}\label{eq:pre manipulations}
   \begin{split}
-    I + \sigma^{-2} \prcov^{1/2}  \fwd ^* \obs^* \obs \fwd \prcov^{1/2} &= 
-    \prcov^{1/2} ( \prcov^{-1} + \sigma^{-2} \fwd^*\obs^*  \obs \fwd ) \prcov^{1/2}\\
-    %
-    %
-    %
-    &=\prcov^{1/2} \fwd^* ( \fwd^{-*} \prcov^{-1}\fwd^{-1} + \sigma^{-2} \obs^* \obs  ) \fwd \prcov^{1/2}\\
-    %
-    %
-    %
-    &= \prcov^{1/2} \fwd^* \left ( (\fwd \prcov \fwd^* )^{-1} +\sigma^{-2} \obs^* \obs  \right ) \fwd \prcov^{1/2}.
+    \det \left (I + \sigma^{-2} \prcov^{1/2} \fwd ^* \obs^*
+    \obs \fwd \prcov^{1/2}\right ) &=
+    %
+    \det \left (I + \sigma^{-2} \obs^*
+    \obs \fwd \prcov\fwd^* \right ) \\
+    %
+    %
+    %
+    &= \det \left \{ \left ( (\fwd \prcov \fwd^*)^{-1} +
+    \sigma^{-2}\obs^*\obs \right ) (\fwd \prcov \fwd^*)\right \}.\\
   \end{split}
 \end{align}
-By definition of the objective \eqref{eq:objective} and using
-\eqref{eq:pre manipulations}:
+
+By definition of the objective \eqref{eq:objective}, \eqref{eq:pre
+  manipulations} and Corollary \ref{cor:samve ev}:
 \begin{align}\label{eq:mod objective}
@@
     %
-    &=\frac12 \log \left ( \prod_{i=1}^{k} ( \lambda_i^{-1} + \sigma^{-2} \eta_i )\prod_{i=1}^{k} \lambda_i \right ). 
+    &=\frac12 \log \left ( \prod_{i=1}^{k} ( \lambda_i^{-1} + \sigma^{-2} \eta_i )\prod_{i=1}^{k} \lambda_i \right )\\
+    %
+    %
+    %
+    &=\frac12 \sum_{i=1}^{k} \log (\sigma^2\lambda_i^{-1} + \eta_i), 
   \end{split}
 \end{align}
-Therefore, an optimal design maximizes
+and the last equality holds up to an additive constant independent of
+$\obs$. Therefore, a D-optimal design maximizes
 
 \begin{equation}\label{eq:true target}
-  \sum_{i=1}^{k}\log (\lambda_i^{-1} + \sigma^{-2} \eta_i),
+  \sum_{i=1}^{k}\log (\sigma^2 \lambda_i^{-1} + \eta_i),
 \end{equation}
 
+with the norm constraints on $\meas_j,j=1,\dots,m$. We would like to
+make all $\eta_i$ as large as possible.  The problem of D-optimal
+design has now been cast to finite dimensions. This justifies using
+linear algebra. Lemma \ref{lemma:free} (a constructive proof is given
+in the appendix) does exactly that.
+\begin{restatable*}[Unit norm decomposition]{lemma}{free}\label{lemma:free}
+  Let $M \in \R^{k \times k}$ symmetric positive definite with $\ttr M
+  = m$, $m \geq k$. We can find $\func_j \in \R^k,j=1,\dots,m$
+  with $\|\func_j\|=1$ and $A = (\func_1,\dots,\func_m)$ such that
+  $AA^t = M$.
+\end{restatable*}
+
+Take any $\{\eta_i\}_{i=1}^{k}$, with $\sum_{i=1}^k \eta_i = m$. Let
+$M = \diag(\eta_1,\dots,\eta_k)$. Lemma \ref{lemma:free} gives us $A$
+with unit norm columns such that $AA^T = M$. Take $\meas_i
+:=\sum_{j=1}^k A_{ji}\ev_j^*$. The constraints $\|\meas_i\| = 1,
+i=1,\dots,m$ are satisified by construction. Verifying $\obs^*\obs
+\ev_i = \eta_i \ev_i$ is not too difficult.
+
+%% \begin{align*}
+%%   \begin{split}
+%%     \obs \ev_i &= (\meas_1,\dots,\meas_k)^t\ev_i\\
+%%     &= (\sum_{j=1}^kA_{j1}\ev_j^*,\dots,\sum_{j=1}^kA_{jk}\ev_j^*)^t\ev_i\\
+%%     &= (\sum_{j=1}^kA_{j1}\ev_j^*\ev_i,\dots,\sum_{j=1}^kA_{jk}\ev_j^*\ev_i)^t\\
+%%     &= (A_{i1},\dots,A_{ik})^t.
+%%   \end{split}
+%% \end{align*}
+
+%% Now,
+
+%% \begin{align*}
+%%   \begin{split}
+%%     \obs^*\obs \ev_i &= \obs^*(A_{i1},\dots,A_{ik})^t\\
+%%     &=(A_{i1},\dots,A_{ik}) \obs \text{ (by \eqref{eq:obs*})}\\
+%%     &=(A_{i1},\dots,A_{ik}) (\sum_{j=1}^kA_{j1}\ev_j^*,\dots,\sum_{j=1}^kA_{jk}\ev_j^*)^t\\
+%%     &=\sum_{l=1}^k A_{il} \sum_{j=1}^kA_{jl}\ev_j^* \\
+%%     %&=\sum_{j=1}^k  \sum_{l=1}^kA_{il}A_{jl}\ev_j^* \\
+%%     &=\sum_{j=1}^k  [AA^t]_{ij}\ev_j^* \\
+%%     %&=\eta_i\ev_i^*\\
+%%     &=\eta_i\ev_i^*. 
+%%   \end{split}
+%% \end{align*}
+
+We have shown that we can construct $\obs$ such that it has
+eigenvectors and eigenvalues $\{\ev_i, \eta_i\}_{i=1}^k$, as long as
+$\sum_{i=1}^k \eta_i = m$. We now turn to understand which choice of
+$\{\eta_i\}_{i=1}^k$ is optimal. By concavity of $\log$, the fastest
+increase in the design criterion \eqref{eq:true target} is gained by
+increasing $\sigma^2\lambda_i^{-1} + \eta_i$ where it is smallest. So,
+no weight should be given to (i.e. measurement taken in direction of)
+the $n$th eigenvector before $\sigma^2\lambda_i^{-1} + \eta_i \geq
+\sigma^2\lambda_n^{-1},\ \forall i \neq n$. Thus, we should choose
+$\{\eta_i\}_{i=1}^k$ as follows. First, order the eigenvalues
+$\lambda_1 \geq \lambda_2 \geq \dots$. We should increase $\sigma^2
+\lambda_1^{-1} + \eta_1$ until it equals
+$\sigma^2\lambda_2^{-1}$. Then, $\sigma^2 \lambda_i^{-1} +
+\eta_i,i=1,2$ are increased until $\sigma^2 \lambda_i^{-1} + \eta_i =
+\sigma^2\lambda_3^{-1},i=1,2$ and so forth.
+
 \pgfplotstableread{
-  Label    prior  $o_1$  $o_2$   $o_3$   $o_4$   $o_5$         $o_6$    topper
-  1         0.2    0.8    0.5     0.5    0.4    0.33333333  0.3333333  0.001
-  2         0.8    0.2    0.5     0.5    0.4    0.33333333  0.3333333  0.001
-  3         2.2    0      0       0      0.2    0.33333333  0.3333333  0.001
-  4         3.5    0      0       0      0      0           0          0.001
-}\testdata
+  Label     prior  optimal  sub-optimal 
+  1         0.2    1.8           1.7
+  2         0.8    1.2           0.8
+  3         2.2    0             0.5
+  4         3.5    0             0.0
+}\optimalvsnot
 
 \begin{figure}
-  \begin{tikzpicture}
+  \begin{tikzpicture}[scale=0.85]
     \begin{axis}[
@@
         reverse legend=false, % set to false to get correct display, but I'd like to have this true
-        xticklabels from table={\testdata}{Label},
+        xticklabels from table={\optimalvsnot}{Label},
         xticklabel style={text width=2cm,align=center},
@@
         ylabel=precision --- prior and posterior,
-        xlabel=eigenvector,
+        xlabel=eigenvector $i$,
       ]
 
-      \addplot [fill=green!80]  table [y=prior, meta=Label, x expr=\coordindex] {\testdata};
-      \addplot [fill=blue!60]   table [y=$o_1$, meta=Label, x expr=\coordindex] {\testdata};
-      \addplot [fill=red!60]    table [y=$o_2$, meta=Label, x expr=\coordindex] {\testdata};
-      \addplot [fill=black!60]  table [y=$o_3$, meta=Label, x expr=\coordindex] {\testdata};
-      \addplot [fill=orange!60] table [y=$o_4$, meta=Label, x expr=\coordindex] {\testdata};
-      \addplot [fill=cyan!60]   table [y=$o_5$, meta=Label, x expr=\coordindex] {\testdata};
-      \addplot [fill=purple!60] table [y=$o_6$, meta=Label, x expr=\coordindex] {\testdata};
+      \addplot [fill=green!80]  table [y=prior,   meta=Label, x expr=\coordindex] {\optimalvsnot};
+      \addplot [fill=blue!60]   table [y=optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};
+
+      \addlegendentry{$\sigma^2\lambda_i^{-1}$}
+      \addlegendentry{optimal $\eta_i$s}
+
+    \end{axis}
+  \end{tikzpicture}
+  \qquad
+  \begin{tikzpicture}[scale=0.85]
+    \begin{axis}[
+        ybar stacked,
+        ymin=0,
+        ymax=4,
+        xtick=data,
+        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
+        reverse legend=false, % set to false to get correct display, but I'd like to have this true
+        xticklabels from table={\optimalvsnot}{Label},
+        xticklabel style={text width=2cm,align=center},
+        legend plot pos=right,
+        ylabel=precision --- prior and posterior,
+        xlabel=eigenvector $i$,
+      ]
+
+      \addplot [fill=green!80]  table [y=prior,       meta=Label, x expr=\coordindex] {\optimalvsnot};
+      \addplot [fill=blue!60]   table [y=sub-optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};
+
+      \addlegendentry{$\sigma^2\lambda_i^{-1}$}
+      \addlegendentry{sub-optimal $\eta_i$s}
+
+    \end{axis}
+  \end{tikzpicture}
+  \caption{Characterizing a D-optimal design, compared to a
+    sub-optimal design. Both designs are allowed three measurements,
+    which means the blue area has accumulated height of $3$ in both
+    plots. A D-optimal design (left) increases precision (reduces
+    uncertainty) where it is lowest. A sub-optimal design (right) does
+    not. In this example, a measurement is taken partly in the
+    direction of the third eigenvector. The precision for this
+    eigenvector is bigger even than the posterior precision of the
+    D-optimal design on the left.}
+  \label{fig:optimal vs not}
+\end{figure}
+
+\pgfplotstableread{
+  Label    prior  $o_1$  $o_2$   $o_3$   $o_4$   $o_5$         $o_6$    topper
+  1         0.2    0.8    0.5     0.5    0.4    0.33333333  0.3333333  0.001
+  2         0.8    0.2    0.5     0.5    0.4    0.33333333  0.3333333  0.001
+  3         2.2    0      0       0      0.2    0.33333333  0.3333333  0.001
+  4         3.5    0      0       0      0      0           0          0.001
+}\clusterization
+
+\pgfplotstableread{
+  Label    prior  $o_1$  $o_2$   $o_3$   $o_4$   $o_5$         $o_6$    topper
+  1         0.2    0.8    0.7     0.3    0.4    0.2         0.4666666  0.001
+  2         0.8    0.2    0.3     0.7    0.4    0.46666666  0.2        0.001
+  3         2.2    0      0       0      0.2    0.33333333  0.3333333  0.001
+  4         3.5    0      0       0      0      0           0          0.001
+}\noclusterization
+
+The argument above has the important consequence (illustrated in
+figure \ref{fig:optimal vs not}) that the eigenvalues in the nonlinear
+eigenvalue problem \eqref{eq:eigenproblem} are all equal! This makes
+perfect sense. We know the Lagrange multipliers $\xi_i,i=1,\dots,k$
+tell us how much we gain by relaxing the constraints and there is no
+reason for any eigenvector to give us more than any other (otherwise
+we would put more weight on the better eigenvector, as can be
+understood from the discussion in the following paragraph).
+
+One can understand D-optimal designs as follows: Imagine each
+eigenvector corresponds to a graduated lab cylinder (figure
+\ref{fig:cylinder}). Each cylinder $i$ is filled, a-priori, with green
+liquid of $\sigma^2\lambda_i^{-1}$ volume units. This liquid
+represnets the prior precision, weighted by observation noise. The
+volume of the green liquid is increasing in $i$. A D-optimal design
+with $m$ measurements ignores cylinders indexed $m+1$ and above. It
+has blue liquid of $m$ volume units at its disposal. It distributes
+said liquid by repeatedly adding a drop to whatever cylinder currently
+has the lowest level of liquid in it (lowest precision). The result of
+such a procedure is illustrated in figure \ref{fig:optimal vs
+  not}. The analysis conducted in this section is summarized in
+Theorem \ref{thm:char} below.
+
+\begin{theorem}[Characterizing D-optimal designs with vanishing model error]\label{thm:char}
+  Let:
+  \begin{enumerate}
+  \item $\fwd, \prcov$ the forward and prior covariance operators, respectively.
+  \item $m, \sigma^2$ the number of measurements taken and observation
+    noise variance, respectively.
+  \item $\obs$ a D-optimal design operator.
+  \item $\tar$ as definded in \eqref{eq:objective}.
+  \item $\{\lambda_i\}_{i=1}^\infty$ eigenvalues of $\fwd\prcov\fwd^*$
+    in decreasing order.
+  \item $\{\ev_i\}_{i=1}^\infty$ their corresponding eigenvectors.
+  \item $\{\eta_i\}_{i=1}^\infty$ eigenvalues of $\obs^*\obs$, where
+    $\obs$ maximizes $\tar(\obs)$.
+  \end{enumerate}
+  Then:
+  \begin{enumerate}
+  \item $\obs^*\obs$ and $\fwd\prcov\fwd^*$ are simultaneously diagonalizable.
+  \item There exists $0 < k \leq m$ such that $\eta_i = 0$ for $i\geq
+    k+1$ and $k = \rank \obs^*\obs$.
+  \item $\sigma^2\lambda_i^{-1} + \eta_i$ are equal for all $1 \leq i \leq k$.
+  \item If $k < m$, then $\sigma^2\lambda_i^{-1} + \eta_i \leq
+    \sigma^2\lambda_{k+1}^{-1}, i=1,\dots,k$.
+  \end{enumerate}
+\end{theorem}
+
+\begin{figure}%{r}{0.25\textwidth} 
+    %\centering
+    \includegraphics[width=5cm, height=5cm]{cylinders.jpg}
+    \caption{Three graduated lab cylinders, corresponding to three
+      eigenvectors. Prior eigenvalues not shown.}
+    \label{fig:cylinder}
+\end{figure}
+
+\section{Sensor Clusterization --- Vanishing Model Error}\label{section:clusterization}
+\subsection{Overview}
+In this section we use Theorem \ref{thm:char} to explain how sensor
+clusterization may arise in various scenarios. We first consider
+sequential designs, and then simultaneous designs.
+
+\subsection{Sequential Design}\label{subsec:clusterization sequential}
+Recall the sequential optimal design scheme from section
+\ref{subsec:seq vs sim} and the definition of $\obs_k$ from
+\eqref{eq:def obs_k}. The way clusterization can occur in sequential
+design scenario is understood by specializing thorem \ref{thm:char} to
+one measurement each time. In this case $m=1$, so $k=1$, $\meas_1
+\parallel \ev_1$ and $\eta_1 = 1$. As hinted in section
+\ref{subsec:seq vs sim}, the posterior becomes the prior for the next
+step. The eigenvectors of the new posterior are the same as for the
+previous one. A new measurement will be taken in the direction of
+eigenvector with smalles eigenvalue. This leads to:
+\begin{align*}
+  \begin{split}
+    \meas_2 \parallel \ev_1  &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 < \sigma^2\lambda^{-1}_2 \\
+    \meas_2 \parallel \ev_2  &\text{ o.w. }
+  \end{split}
+\end{align*}
+In the first case, we get clusterization immediately. In the second,
+we consider $\meas_3$. Same reasoning leads to the conclusion that the
+eigenvectors of the new prior precision are the same. The eigenvalues,
+however, are $\lambda^{-1}_1 + \sigma^{-2}, \lambda^{-1}_2 +
+\sigma^{-2}, \lambda_3^{-1}, \dots$. A new measurement is taken in the
+direction of eigenvector of smallest precision (note that
+$\lambda^{-1}_1 + \sigma^{-2} < \lambda^{-1}_2 + \sigma^{-2}$):
+\begin{align*}
+  \begin{split}
+    \meas_3 \parallel \ev_1 &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 <
+    \sigma^2\lambda^{-1}_3 \\
+    \meas_3 \parallel  \ev_3  &\text{ o.w. }
+  \end{split}
+\end{align*}
+Again, the first case results in clusterization. The sequential design
+proceeds in this fashion. Every $\meas_i$ is taken in the direction
+some eigenvector of the prior $\fwd \prcov\fwd^*$. This eigenvector
+has the smallest eigenvalue of the previous step's posterior precision
+operator. If we do not encounter clusterization after $m$ such steps,
+this means that measurement $\meas_k$ was taken in the direction
+$\ev_k$ for $k=1,\dots,m$. Thus:
+\begin{align}\label{eq:failure}
+  \begin{split}
+    \sigma^2\lambda_{k-1}^{-1} + 1 \geq \sigma^2\lambda_k^{-1}, \text{ for } 1\leq k \leq m
+    \Longrightarrow \lambda_m^{-1} - \lambda_{m-1}^{-1} \leq \sigma^{-2}.
+  \end{split}
+\end{align}
+
+But \eqref{eq:failure} implies $\{\lambda_i^{-1}\}_{i=1}^{\infty}$
+grows at most linearly. In section \ref{subsec:abstract OED} we
+assumed $\fwd$ is strongly smoothing, which means eigenvalues of
+$(\fwd \prcov\fwd^*)^{-1}$ grow quickly. Thus, if we take $m$ large
+enough, \eqref{eq:failure} will eventually fail and we will observe
+sensor clusterization.
+
+\subsection{Simultaneous Design}\label{subsec:clusterization simultaneous}
+In a simultaneous design scenario, the result is not simple as in the
+sequential design case. We now see that a design that exhibits
+clusterization is as optimal as one that does not. Corollary
+\ref{cor:same ev} tells us that we can realize any design we want
+(encoded in $\{\eta_i\}$), given the trace constraint for
+$\obs^*\obs$. Theorem \ref{thm:char} tells us that for D-optimality,
+we only need to make the posterior flat where we take measurements
+(recall figure \ref{fig:optimal vs not}). We are left with the freedom
+of assigning any part of $\eta_i$ to any measurement $\meas_j$. Two
+possible assignments for the same problem are illustrated in figure
+\ref{fig:clusterization}. The first exhibits clusterization, the
+second does not. It is important to note that neither arises as a
+result of some numerical optimization scheme. Both designs give the
+same design criterion $\tar$, because their posteriors are
+identical. By Theorem \ref{thm:char}, both are maximizers of the same
+D-optimal design problem. One exhibits clusterization, and one does
+not. It is likely that a real-life problem is more restricted than the
+model we consider here. Such restriction is the likely cause of
+clusterization obseved in, e.g., the toy model of section
+\ref{subsec:example}. Exactly how this happens is not clear. At this
+point it is the subject of speculation and should be further studied
+in the future.
+
+\begin{figure}
+  \begin{tikzpicture}[scale=0.85]
+    \begin{axis}[
+        ybar stacked,
+        ymin=0,
+        ymax=4,
+        xtick=data,
+        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
+        reverse legend=false, % set to false to get correct display, but I'd like to have this true
+        xticklabels from table={\clusterization}{Label},
+        xticklabel style={text width=2cm,align=center},
+        legend plot pos=right,
+        ylabel=precision --- prior and posterior,
+        xlabel=eigenvector $i$,
+      ]
+
+      \addplot [fill=green!80]  table [y=prior, meta=Label, x expr=\coordindex] {\clusterization};
+      \addplot [fill=blue!60]   table [y=$o_1$, meta=Label, x expr=\coordindex] {\clusterization};
+      \addplot [fill=red!60]    table [y=$o_2$, meta=Label, x expr=\coordindex] {\clusterization};
+      \addplot [fill=black!60]  table [y=$o_3$, meta=Label, x expr=\coordindex] {\clusterization};
+      \addplot [fill=orange!60] table [y=$o_4$, meta=Label, x expr=\coordindex] {\clusterization};
+      \addplot [fill=cyan!60]   table [y=$o_5$, meta=Label, x expr=\coordindex] {\clusterization};
+      \addplot [fill=purple!60] table [y=$o_6$, meta=Label, x expr=\coordindex] {\clusterization};
 
@@
   \end{tikzpicture}
-  \caption{Posterior precision per eigenvector, after taking a
-    measurement $\meas_i$. Each $\meas_i$ increases precision
-    uniformly across the lowest precision eigenvectors. We see that
-    $\meas_2$ and $\meas_3$ are repeated measurements, as well as
-    $\meas_5$ and $\meas_6$.}
+  \qquad
+  \begin{tikzpicture}[scale=0.85]
+    \begin{axis}[
+        ybar stacked,
+        ymin=0,
+        ymax=4,
+        xtick=data,
+        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
+        reverse legend=false, % set to false to get correct display, but I'd like to have this true
+        xticklabels from table={\noclusterization}{Label},
+        xticklabel style={text width=2cm,align=center},
+        legend plot pos=right,
+        ylabel=precision --- prior and posterior,
+        xlabel=eigenvector $i$,
+      ]
+
+      \addplot [fill=green!80]  table [y=prior, meta=Label, x expr=\coordindex] {\noclusterization};
+      \addplot [fill=blue!60]   table [y=$o_1$, meta=Label, x expr=\coordindex] {\noclusterization};
+      \addplot [fill=red!60]    table [y=$o_2$, meta=Label, x expr=\coordindex] {\noclusterization};
+      \addplot [fill=black!60]  table [y=$o_3$, meta=Label, x expr=\coordindex] {\noclusterization};
+      \addplot [fill=orange!60] table [y=$o_4$, meta=Label, x expr=\coordindex] {\noclusterization};
+      \addplot [fill=cyan!60]   table [y=$o_5$, meta=Label, x expr=\coordindex] {\noclusterization};
+      \addplot [fill=purple!60] table [y=$o_6$, meta=Label, x expr=\coordindex] {\noclusterization};
+
+      \addlegendentry{prior}
+      \addlegendentry{$o_1$}
+      \addlegendentry{$o_2$}
+      \addlegendentry{$o_3$}
+      \addlegendentry{$o_4$}
+      \addlegendentry{$o_5$}
+      \addlegendentry{$o_6$}   
+    \end{axis}
+  \end{tikzpicture}
+  \caption{Clusterization (left) and non-clusterization (right) in
+    simultaneous D-optimal designs. Posterior precision per
+    eigenvector after taking measurements $\meas_i$ is plotted. Both
+    designs are identical for all practical matters --- their
+    posteriors are identical. In the left panel we see how an optimal
+    design is achieved with repeated measurements. $\meas_2 = \meas_3$
+    and $\meas_5 = \meas_6$. This is not necessary though, as can be
+    seen in the right panel. Both designs achieve the same levels of
+    uncertainty. The left exhibits clusterization. The right does not.}
   \label{fig:clusterization}
 \end{figure}
-with the norm constraints on $\meas_j,j=1,\dots,m$. We show in the
-appendix (Lemma \ref{lemma:free}) we can find $\meas_1,\dots,\meas_m$
-such that $\{\eta_i\}_{i=1}^k$ have any value we desire, subject only
-to the restriction that we do not change the trace of $\obs^*\obs$, so
-that $\sum_{i=1}^k \eta_i = m$. By concavity of $\log$, the fastest
-increase in the design criterion (equation \eqref{eq:true target}) is
-gained by increasing $\lambda_i^{-1} + \sigma^{-2}\eta_i$ where it is
-smallest. So, no weight should be given to (i.e. measurement taken of)
-the $k$th eigenvector before $\lambda_i^{-1} + \sigma^{-2}\eta_i \geq
-\lambda_k^{-1},\ \forall i \neq k$. Thus, we should choose
-$\{\eta_i\}_{i=1}^k$ as follows. We would increase $\lambda_1^{-1} +
-\sigma^{-2} \eta_1$ until it equals $\lambda_2^{-1}$. Then,
-$\lambda_i^{-1} + \sigma^{-2} \eta_i,i=1,2$ are increased until
-$\lambda_i^{-1} + \sigma^{-2} \eta_i = \lambda_3^{-1},i=1,2$ and so
-forth.
-
-The argument above has the important consequence that the eigenvalues
-in \eqref{eq:eigenproblem} are all equal! This makes perfect
-sense. First, we know the Lagrange multipliers tell us how much we
-gain by relaxing the constraints and there is no reason for any
-measurement to give us more than any other (otherwise we would put
-more weight on the better measurement, as can be understood from the
-discussion in the following paragraph). We are also used to symmetric
-optimization problems having symmetric solutions \cite{Waterhouse83},
-so symmetric and identical Lagrange multipliers definitely make sense
-in this context.
-
-\subsection{Sensor Clusterization}\label{subsec:clusterization}
-The way clusterization can occur is best understood from figure
-\ref{fig:clusterization}. We see that each measurement increases
-precision (reduces uncertainty) only for the eigenvectors of lowest
-precision. If the precision in these is lower than the precision for
-the next eigenvector, a repeated measurement is optimal.
-
-\section{Analysis of Optimal Designs --- Non-Vanishing Model Error}
-Denote $\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm :=
-(\meas_1,\dots,\meas_{m-1})^t$. The following corollary is proved in
-the appendix.
-\begin{corollary}[Proved as corollary \ref{cor:same meas} in the appendix]
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% SECTION Analysis of Optimal Designs --- Non-Vanishing Model Error
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\section{Analysis of D-Optimal Designs --- Non-Vanishing Model Error}\label{section:non vanishing}
+In this section we show the effect model error has on the
+clusterization phenomenon. We will see that if $\modcov \neq 0$
+clusterization will not occur. This is contrary to the previous case
+of vanishing model error.
+
+Corollary \ref{cor:same meas} below is proved in the appendix. Denote
+$\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm :=
+(\meas_1,\dots,\meas_{m-1})^t$. Also, denote $\Sigmam := \Sigma (\obsm)$ and
+$\postcovm$ the posterior covariance that arises when we use $\obsm$.
+\begin{restatable*}{corollary}{samemeas}\label{cor:same meas}
   If $\meas_m = \meas_j$ for some $1 \leq j \leq m-1$, then
@@
   \end{equation*}
-\end{corollary}
-
-In contrast to the previous case of vanishing model error, we will see
-that if $\modcov \neq 0$ clustering will not occur. From Corollary
-\ref{cor:same meas} we can observe that as $\sigma^2 \to 0$, the
-increase in the design criterion gained by taking a measurement
-identical to a previous one approaches zero. We may conclude that for
-small measurement error levels, the clusterization effect will be
+\end{restatable*}
+Recall from \eqref{eq:Sigma} that $\Sigma(\obs) = \obs \modcov \obs^*
++ \sigma^2I$. Then
+
+$$
+\lim_{\sigma^2 \to 0} \tar(\obs) -\tar(\obsm) = 0
+$$
+
+Hence no increase in the design criterion is achieved by taking a
+repeated measurement (in the limit $\sigma^2 \to 0$). This is clearly
+sub-optimal and $\obs$ cannot give rise to a D-optimal design. Thus,
+for small measurement error levels, the clusterization effect will be
 mitigated by the presence of a non-zero model error. Since the design
@@
 
-\section{Conclusion}\label{section:conclusion}
-Our relaxed model gives us insight to D-optimal designs. We see that
-in our setting, uncertainty is not reduced in any direction
-(eigenvector) before it is the largest uncertainty present. We show
-how and why repeated measurements can give rise to a D-optimal
-design. There is more work to be done in understanding exactly what
-causes such designs to be strictly better than others.
-
 \section{Acknowledgements}
-This study is a part of my PhD thesis \cite{mine} which was written
-under the instruction of Prof. Georg Stadler in New York University's
-Courant Institute. I would like to thank him for his great mentorship.
+This study is a part of my PhD thesis. Most of it was written under
+the instruction of Prof. Georg Stadler in New York University's
+Courant Institute. I would like to thank him for his great
+mentorship. This work was supported by The Raymond and Beverly Sackler
+Post-Doctoral Scholarship.
 
 \appendix
-\section{Widely Applicable Lemmas}
-
-The following lemma is generalized from \cite[Chapter 9,
-  Theorem 4, pp. 127]{Lax97}
-\begin{lemma}\label{lemma:lax}
-  Let $Y(t)$ be a differentiable operator-valued function. Assume 
-  $I+Y(t)$ is invertible, $Y(t)$ self-adjoint and trace-class. Then
-  \begin{equation*}
-    \frac{\der \log \det (I+Y(t))}{\der t} = \tr{ (I+Y(t))^{-1} \dot{Y}(t)}.
-  \end{equation*}
-\end{lemma}
+\section{General Lemmas}
+
+The following lemma is generalized from \cite[Chapter 9, Theorem 4, pp. 127]{Lax97}
+\lax
 \begin{proof}
@@
     = \sum_{k=1}^{\infty} \dot{\lambda_k}(X(0)) 
-    = \frac{\der }{\der t}\tr X(0)
-    = \ttr \dot{X}(0),
+    = \frac{\der }{\der t}\tr{X(0)}
+    = \tr{\dot{X}(0)},
   \end{align*}
@@
   \begin{align*}
-    \ttr X(t) = \tr{(I+Y(0))^{-1} (I+Y(t)) - I}
-    \leq \tr{I+Y(t) - I} < \infty,
+    \tr{X(t)} = \tr{(I+Y(0))^{-1} (I+Y(t)) - I}
+    \leq \tr{I+Y(t) - I}< \infty,
   \end{align*}
@@
     % 
-    &= \det (I+Y(0)) \ttr \dot{X}(0) \\
-    % 
-    % 
-    % 
-    &= \det (I+Y(0)) \tr (I+Y(0))^{-1} \dot{Y}(0).
+    &= \det (I+Y(0)) \tr{\dot{X}(0)} \\
+    % 
+    % 
+    % 
+    &= \det (I+Y(0)) \tr{(I+Y(0))^{-1} \dot{Y}(0)}.
   \end{align*}
@@
     % 
-    &= \tr (I+Y(t))^{-1} \dot{Y}(t) \big |_{t=0}.
-  \end{align*}
-  Since there is nothing special about $t=0$, the relation holds for all $t$.
+    &= \tr{ (I+Y(t))^{-1} \dot{Y}(t)} \big |_{t=0}.
+  \end{align*}
+  There is nothing special about $t_0 = 0$ --- we could have chosen
+  any other $t_0$ instead. Thus, the relation holds for all $t$.
 \end{proof}
@@
 
-\section{Specific Lemmas}
-
-\begin{lemma}[Increase due to a Measurement]\label{lemma:design increase}
-  Let $\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm := (\meas_1,\dots,\meas_{m-1})^t$. Then
-  \begin{align*}
-    \tar( \obs ) - \tar (\obsm ) &=
-    \frac12 \log \left ( 1 + \frac{
-      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m,
-      (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m \rangle
-    }{
-      \sigma^2 + \meas_m \modcov \meas_m - \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m 
-    }       
-    \right ).
-  \end{align*}
-\end{lemma}
-\begin{proof}
-  We use the Schur complement to write one inverse in terms of the other and
-  introduce notations to make the derivation cleaner. Note that we think of
-  $\obsm$ and $\obsm^*$ as column and row vectors (respectively).
-  \begin{align*}
-    \Sigma( \obs ) &= \Sigma = 
-    \begin{bmatrix}
-      \Sigma (\obsm )           & \obsm \modcov \meas_m \\
-      \meas_m \modcov \obsm^*   & \sigma^2 + \meas_m \modcov \meas_m
-    \end{bmatrix}
-    =
-    \begin{bmatrix}
-      \Sigmam   & w \\
-      w^t       & c
-    \end{bmatrix}\\
-    %
-    %
-    %
-    \Sigma^{-1} &=
-    \begin{bmatrix}
-      \Sigmam^{-1} + \Sigmam^{-1} w ( c - w^t \Sigmam^{-1} w)^{-1} w^t \Sigmam^{-1} & - \Sigmam^{-1} w ( c - w^t \Sigmam^{-1} w)^{-1} \\
-      -( c - w^t \Sigmam^{-1} w)^{-1} w^t \Sigmam^{-1}                            &  ( c - w^t \Sigmam^{-1} w)^{-1}
-    \end{bmatrix} \\
-    &=
-    \begin{bmatrix}
-      \Sigmam^{-1} & 0 \\
-      0           & 0 
-    \end{bmatrix}
-    + (c -w^t \Sigmam^{-1} w )^{-1}
-    \begin{bmatrix}
-      \Sigmam^{-1} w \\
-      -1
-    \end{bmatrix}
-    \begin{bmatrix}
-      w^t \Sigmam^{-1} & -1 
-    \end{bmatrix}
-  \end{align*}
-  %
-  Further, define
-  %
-  \begin{align*}
-    \M (\obs ):&= \prcov^{\frac12}\fwd^* \obs^* \Sigma^{-1} \obs \fwd
-    \prcov^{\frac12}    
-  \end{align*}
-  %
-  and note that, using our understanding of what is a column vector and
-  what is a row vector:
-  %
-  \begin{align*}
-    \M(\obs) &= \prcov^{1/2} \fwd^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2} \\
-    %
-    %
-    %
-    &= \prcov^{1/2} \fwd^* \obs^* \left \{
-    \begin{bmatrix}
-      \Sigmam^{-1} & 0 \\
-      0           & 0 
-    \end{bmatrix}
-    + (c -w^t \Sigmam^{-1} w )^{-1}
-    \begin{bmatrix}
-      \Sigmam^{-1} w \\
-      -1
-    \end{bmatrix}
-    \begin{bmatrix}
-      w^t \Sigmam^{-1} & -1 
-    \end{bmatrix} 
-    \right \} \obs \fwd \prcov^{1/2} \\
-    %
-    %
-    %
-    &= \M (\obsm) + (c -w^t \Sigmam^{-1} w )^{-1}
-    \prcov^{1/2} \fwd^* \obs^*
-    \begin{bmatrix}
-      \Sigmam^{-1} w \\
-      -1
-    \end{bmatrix}
-    \begin{bmatrix}
-      w^t \Sigmam^{-1} & -1 
-    \end{bmatrix} 
-    \obs \fwd \prcov^{1/2}
-  \end{align*}
-  %
-  Now, denote:
-  %
-  \begin{align}\label{eq:u}
-    \begin{split}
-      u :&= (c -w^t \Sigmam^{-1} w )^{-1/2}
-      \prcov^{1/2} \fwd^* \obs^* 
-      \begin{bmatrix}
-        \Sigmam^{-1} w \\
-        -1 
-      \end{bmatrix} \\
-      %
-      %
-      %
-      & = (c -w^t \Sigmam^{-1} w )^{-1/2} ( \prcov^{1/2}\fwd^* \obsm^* \Sigmam^{-1} \obsm  \modcov \meas_m - \prcov^{1/2} \fwd^* \meas_m )\\
-      %
-      %
-      %
-      u^* :&=  (c -w^t \Sigmam^{-1} w )^{-1/2} (\meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \fwd \prcov^{1/2} - \meas_m \fwd \prcov^{1/2} ),
-    \end{split}
-  \end{align}
-  %
-  so that
-  %
-  \begin{equation}\label{eq:M plus I}
-    I + \M( \obs ) = I + \M (\obsm ) + uu^*.
-  \end{equation}
-  %
-  Note that
-  \begin{equation}\label{eq:M postcov}
-    \prcov^{1/2} \left (I + \M( \obsm ) \right )^{-1} \prcov^{1/2} = \postcovm.
-  \end{equation}
-  The increase in the design criterion gained by including $\meas_m$
-  in the design is found by using Lemma \ref{lemma:MDL} the above
-  results:
-  %
-  \begin{align*}
-    \tar( \obs ) - \tar( \obsm )
-    %
-    %
-    %
-    &= \frac12 \log \det \Big ( I + \M ( \obs ) \Big ) / \det \Big ( I + \M (\obsm) \Big ) \\
-    %
-    %
-    %
-    &= \frac12  \log \det \left ( I + \M(\obsm) + uu^* \right ) / \det \Big ( I + \M (\obsm) \Big ) \\
-    %
-    %
-    %
-    &= \frac12 \log \left ( 1 + \left \langle \left ( I+\M(\obsm) \right )^{-1} u, u  \right \rangle \right ).
-  \end{align*}
-  Using \eqref{eq:u}:
-  \begin{align*}
-    &\left \langle \left (I+\M (\obsm)\right )^{-1}u, u \right \rangle\\
-    &= \frac{
-      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m,
-      (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m \rangle
-    }{
-      c- w^t \Sigmam^{-1} w
-    }\\
-    %
-    %
-    %
-    &= 
-    \frac{
-      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m,
-      (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m \rangle
-    }{
-      \sigma^2 + \meas_m \modcov \meas_m - \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m 
-    }
-  \end{align*}
-  and the conclusion follows.
-\end{proof}
-
-Lemma \ref{lemma:design increase} implies the following two corollaries.
-\begin{corollary}[Gain for No Model Error]\label{cor:zero mod err}
-  If $\modcov = 0$, then
-  \begin{equation*}
-    \tar( \obs ) - \tar (\obsm )
-    = -\frac12 \log (1 - \sigma^{-2} \langle \fwd \postcov \fwd^* \meas_m, \meas_m \rangle ).
-  \end{equation*}
-\end{corollary}
-\begin{proof}
-  Note that this is not immediate by substituting $\modcov = 0$ in the
-  conclusion of Lemma \ref{lemma:design increase}, since we make a
-  claim for $\postcov$, and not $\postcovm$. Let us first review
-  \eqref{eq:u} and note that since $\modcov = 0$ the covariance
-  $\Sigmam = \sigma^2I_{m-1}$ and $w = 0$, so $c - w^t\Sigmam^{-1}w =
-  \sigma^2$:
-  \begin{align*}
-    u :&=-\sigma^{-1}\prcov^{1/2} \fwd^* \meas_m\\
-    %
-    %
-    u^* :&= -\sigma^{-1} \meas_m \fwd \prcov^{1/2}.
-  \end{align*}
-  From \eqref{eq:M plus I}:
-  \begin{equation*}
-    I + \M(\obsm) = I +\M(\obs) - uu^*,
-  \end{equation*}
-  and thus:
-  \begin{equation*}
-    \left \langle \left ( I +\M(\obs) \right )^{-1} u, u \right \rangle
-    = \sigma^{-2} \langle \fwd \postcov \fwd^* \meas_m, \meas_m \rangle.
-  \end{equation*}
-  Analogously to \eqref{eq:M postcov} we note that
-  \begin{equation*}
-    \prcov^{1/2} \left ( I + \M(\obs) \right )^{-1} \prcov^{1/2} = \postcov.
-  \end{equation*}
-  Using Lemma \ref{lemma:MDL} we conclude
-  \begin{align*}
-    \tar( \obs ) - \tar( \obs )
-    &= \frac12 \log \det \left (I +\M(\obs) \right ) / \det \left (I + \M(\obsm) \right ) \\
-    %
-    %
-    %
-    &= \frac12 \log \det \left (I +\M(\obs) \right ) / \det \left (I + \M(\obs) - uu^* \right ) \\
-    %
-    %
-    %
-    &=-\frac12 \log (1 - \langle (I+\M(\obs))^{-1}u, u \rangle \\
-    %
-    %
-    %
-    &= -\frac12 \log (1 - \sigma^{-2} \langle \fwd \postcov \fwd^* \meas_m, \meas_m \rangle ).
-  \end{align*}
-\end{proof}
-
-\begin{corollary}[Gain for Identical Measurement]\label{cor:same meas}
-  If $\meas_m = \meas_j$ for some $1 \leq j \leq m-1$, then
-  \begin{equation*}
-    \tar(\obs) - \tar(\obsm) =
-    \log \left ( 1 + \frac{\sigma^2
-      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
-      \obsm^* \Sigmam^{-1}e_j \rangle
-    }{
-      2 - \sigma^2 e_j^t\Sigmam^{-1}e_j 
-    }       
-    \right ).
-  \end{equation*}
-\end{corollary}
-\begin{proof}
-  Denote $A:= \obs \modcov \obs^*$ and $v_j$ the $j$th column of $A$.
-  Note that $v_j = \obsm \modcov \meas_m$, since $(\obsm \modcov
-  \obsm^*)_{ij} = \meas_i(\modcov \meas_j)$, as explained in
-  \eqref{eq:modcov explained}. One can now verify that
-  \begin{equation}\label{eq:observation}
-    \Sigmam^{-1} \obsm \modcov \meas_m = \Sigmam^{-1}v_j = (A +\sigma^2I_{m-1})^{-1} v_j =
-    e_j -\sigma^2 \Sigmam^{-1}e_j.
-  \end{equation}
-  %
-  Using \eqref{eq:observation}:
-  \begin{align}\label{eq:denominator}
-    \begin{split}
-      \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m
-      &= \meas_m \modcov \obsm^* ( e^j - \sigma^2 \Sigmam^{-1} e_j )\\
-      %
-      %
-      %
-      &= \meas_m \modcov \meas_j - \sigma^2 \meas_m \modcov \obsm^* \Sigmam^{-1}e_j \\
-      %
-      %
-      %
-      &= \meas_m \modcov \meas_j -\sigma^2 (e_j - \sigma^2 \Sigmam^{-1}e_j)^t e_j \\
-      %
-      %
-      %
-      &= \meas_m \modcov \meas_m -\sigma^2 + \sigma^4 e_j^t\Sigmam^{-1}e_j.
-    \end{split}
-  \end{align}
-  We use \eqref{eq:observation} to simplify the terms in the enumerator of
-  the conclusion of Lemma \ref{lemma:design increase}:
-  \begin{align}\label{eq:enumerator}
-    \begin{split}
-      (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m
-      &= \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m - \meas_m \\
-      %
-      %
-      %
-      &= \obsm^* (e_j - \sigma^2 \Sigmam^{-1} e_j) -\meas_j \\ 
-      %
-      %
-      %
-      &= -\sigma^2 \obsm^* \Sigma^{-1}e_j. 
-    \end{split}
-  \end{align}
-  %
-  Substitute \eqref{eq:enumerator} and \eqref{eq:denominator} to
-  the enumerator and denominator (respectively) of the conclusion of
-  Lemma \ref{lemma:design increase}:
-  %
-  \begin{align*}
-    \tar( \obs ) - \tar (\obsm ) &=
-    \log \left ( 1 + \frac{
-      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m,
-      (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m \rangle
-    }{
-      \sigma^2 + \meas_m \modcov \meas_m - \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m 
-    }       
-    \right ) \\
-    %
-    %
-    %
-    &= \log \left ( 1 + \frac{\sigma^4
-      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
-      \obsm^* \Sigmam^{-1}e_j \rangle
-    }{
-      2\sigma^2 - \sigma^4 e_j^t\Sigmam^{-1}e_j 
-    }       
-    \right ) \\
-    %
-    %
-    %
-    &= \log \left ( 1 + \frac{\sigma^2
-      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
-      \obsm^* \Sigmam^{-1}e_j \rangle
-    }{
-      2 - \sigma^2 e_j^t\Sigmam^{-1}e_j 
-    }       
-    \right ).
-  \end{align*}
-\end{proof}
-
-\begin{lemma}\label{lemma:free}
-  Let $M \in \R^{k \times k}$ symmetric positive definite with $\tr M =
-  m$, $m > k$. Then we can find $\func_j \in \R^k,j=1,\dots,m$ with
-  $\|\func_j\|=1$ and $A = (\func_1,\dots,\func_m)$ such that $AA^t =
-  M$.
-\end{lemma}
+\free
 \begin{proof}
   Let us diagonalize $M$, so that $M = U D U^t$ with $D =
-  \diag(d_1,\dots,d_k)$ and $U \in \R^{k \times k }$ orthogonal. Let $S
-  \in \R^{k \times m}$ with $S_{ii} = \sqrt{d_{i}}$ and zeros
+  \diag(d_1,\dots,d_k)$ and $U \in \R^{k \times k }$ orthogonal. Let
+  $S \in \R^{k \times m}$ with $S_{ii} = \sqrt{d_{i}}$ and zeros
   otherwise. Define $A:= U S V^t$, where $V \in \R^{m \times m}$ is
@@
   \end{equation}
-  and we can expect to do this since we assumed $\tr D = m$.
-
-  Define $C = S^tS - I \in \R^{m \times m}$. Note that $\tr C = 0$ and
+  and we can expect to do this since we assumed $\ttr D = m$.
+
+  Define $C = S^tS - I \in \R^{m \times m}$. Note that $\ttr C = 0$ and
   $C$ is diagonal with non-zero entries $d_i-1,i=1,\dots,k$. It suffices
@@
   and not destroying zeros we already introduced, starting from the
-  last diagonal entry and moving to the first. If $c_{kk} \neq 0$ ,
-  let $p < k$ such that $c_{pp}c_{kk} < 0$ (such $p$ exists because
+  last diagonal entry and moving to the first. Since $c_{mm} \neq 0$ ,
+  let $p < m$ such that $c_{pp}c_{mm} < 0$ (such $p$ exists because
   the trace is zero) and let $\theta \in (0,\pi)$. Define a Givens
-  rotation $R^{(k)} \in \R^{m \times m}$ by
+  rotation $R^{(m)} \in \R^{m \times m}$ by
   \begin{equation*}
-    r^{(k)}_{ab} :=
+    r^{(m)}_{ab} :=
     \begin{cases}
-      1 & a = b \neq p \text{ or } a = b \neq k \\
+      1 & a = b \neq p \text{ or } a = b \neq m \\
       \cos \theta & a = b = p  \\
-     -\sin \theta & a = p, b = k\\
-      \cos \theta & a = b = k \\
-      \sin \theta & a = k, b = p \\ 
+     -\sin \theta & a = p, b = m\\
+      \cos \theta & a = b = m \\
+      \sin \theta & a = m, b = p \\ 
       0 & o.w
@@
   \begin{equation*}
-    c_{kk} \cot^2 \theta + 2 c_{kp} \cot \theta + c_{pp} = 0.
+    c_{mm} \cot^2 \theta + 2 c_{mp} \cot \theta + c_{pp} = 0.
   \end{equation*}
   This quadratic in $\cot\theta$ has a real solution, since
-  $c_{pp}c_{kk} < 0$ by assumption and we can find $\theta \in
+  $c_{pp}c_{mm} < 0$ by assumption and we can find $\theta \in
   (0,\pi)$ such that \eqref{eq:mm} is satisfied. We continue to find
-  $R^{(k-1)}$ that leaves rows and columns $k,\dots,m$ unchanged and
-  continue introducing zeros to the diagonal. The assumption $\tr D =
-  m \Rightarrow \tr C = 0$ guarantees we can do that. Taking $V:=
-  R^{(1)} R^{(2)} \dots R^{(k-1)}R^{(k)}$ completes the proof.
+  $R^{(m-1)}$ that leaves row and column $m$ unchanged and
+  continue introducing zeros to the diagonal. The assumption $\ttr D =
+  m \Rightarrow \ttr C = 0$ guarantees we can do that. Taking $V:=
+  R^{(1)} R^{(2)} \dots R^{(m-1)}R^{(m)}$ completes the proof.
 \end{proof}
 
-\begin{lemma}[Auxilliary Calculations]\label{lemma:aux calc}
-  Let $T(\obs) := \obs^* \Sigma^{-1}(\obs)\obs$, with $\Sigma(\obs)$
-  defined as in \eqref{eq:Sigma}. Then:
-  \begin{align*}
-    \delta T(\obs)V &= V^* \Sigma^{-1} \obs 
-    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
-    &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs
-    + \obs^* \Sigma^{-1} V.
-  \end{align*}
-\end{lemma}
-
-\begin{proof}
-  We need a few supplementary calculations. First:
-  \begin{align}\label{eq:der sig}
-    \begin{split}
-      \frac{\der}{\der \tau} \Big |_{\tau=0} \Sigma( \obs + \tau V )
-      &= \frac{\der}{\der \tau} \Big |_{\tau=0} 
-      (\obs + \tau V ) \modcov (\obs + \tau V )^*  + \sigma^2I\\
-      % 
-      % 
-      % 
-      &= V \modcov \obs^* + \obs \modcov V^*.
-    \end{split}
-  \end{align}
-  By the standard trick for the derivative of an operator: 
-  \begin{align*}
-    0 &= \frac{\der}{\der \tau} \Big |_{\tau=0} I \\
-    % 
-    % 
-    % 
-    &= \frac{\der}{\der \tau} \Big |_{\tau=0}
-    \left (\Sigma(\obs+\tau V)^{-1} \Sigma(\obs+\tau V) \right ) \\
-    % 
-    % 
-    % 
-    &= \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0} \Sigma+
-    \Sigma^{-1} \frac{\der \Sigma(\obs+\tau V)}{\der \tau} \Big |_{\tau=0}\\  
-    %
-    %
-    %
-    &= \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0} \Sigma+
-    \Sigma^{-1} (V\modcov \obs^* + \obs \modcov V^*) 
-    \text{, by \eqref{eq:der sig}. }
-  \end{align*}
-  Now we can write the variation of $\Sigma^{-1}$:
-  \begin{align}\label{eq:der sig inv}
-    \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0}  
-      &= -\Sigma^{-1} (V \modcov \obs^* + \obs \modcov V^*) \Sigma^{-1}.
-    \end{align}
-  % 
-  Finally, we can do the calculation we need to. By Leibnitz (product) rule
-  and \eqref{eq:der sig inv}:
-  % 
-  \begin{align*}
-    \delta T(\obs) V 
-    &= \frac{\der T(\obs + \tau V)}{\der \tau} \Big |_{\tau=0} \\
-    %
-    %
-    %
-    &= V^* \Sigma^{-1} \obs 
-    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
-    &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs
-    + \obs^* \Sigma^{-1} V,
-  \end{align*}
-  as required.
-\end{proof}
-
-\begin{lemma}\label{lemma:sim diag}
-  Let $C:\hil \to \hil$ self-adjoint and let $\func_1,\dots,\func_m \in
-  \hil$. Denote $\func^*$ the element $\func$ acting as a
-  functional. If
-  \begin{equation*}
-   (C + \sum_{j=1}^m \func_j\func_j^*) \func_k = \xi_k \func_k, k = 1,\dots,m
-  \end{equation*}
-  then $C$ and $\sum_{j=1}^m \func_j \func_j^*$ are simultaneously
-  diagonalizable.
-\end{lemma}
+\simdiag
 \begin{proof}
@@
 
-\begin{proposition}\label{prop:bigger better}
-  Let $\obs = (\meas_1,\dots,\meas_m)^t$, $j \in \{1,\dots,m\}$, $\sigma^2
-  > 0$ and $\lambda > 1$. Then $\tar(\obs)$ increases if we use
-  $\lambda \meas_j$ in $\obs$ instead of $\meas_j$.
-\end{proposition}
+\section{Specific Lemmas}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{lemma}\label{lemma:twice woodbury}
+  Assume $\fwd \prcov \fwd^*$ is invertible. Then
+\begin{align*}
+  \begin{split}
+    \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* 
+    %
+    %
+    = \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1},
+  \end{split}
+\end{align*}  
+\end{lemma}
+\begin{proof}
+  The proof is boring and amounts to using Woodbury's matrix identity
+  twice and a regularization trick. The standard proof for this
+  identity works in infinite dimensions, as long as all terms are well
+  defined. Unfortunately, $\obs^*\obs$ is not invertible, so we force
+  it to be. Throughout, we let $X := \fwd\prcov \fwd^*$ and $Y_{\eps}
+  := \obs^*\obs + \eps I$, for some small $\eps > 0$. Note that now
+  $Y_{\eps}$ is invertible. Recall Woodbury's matrix identity:
+  $$
+  (A + UCV)^{-1} = A^{-1} + A^{-1}U(C^{-1} + VA^{-1}U)^{-1}VA^{-1}.
+  $$
+
+  Taking $A := \prcov^{-1}, U := \fwd^*, V := \fwd$ and $C :=
+  \sigma^{-2} (\obs^*\obs+\eps I)$ we get:
+  \begin{align*}
+    \begin{split}
+      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* &=
+      \fwd ( \prcov - \prcov \fwd^* ( \sigma^2(\obs^*\obs + \eps I)^{-1} + \fwd \prcov \fwd^* )^{-1} \fwd \prcov ) \fwd^* \\
+      %
+      %
+      &= X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X
+    \end{split}
+  \end{align*}
+  Note that $X + \sigma^2 Y_{\eps}^{-1}$ is invertible, as the sum of
+  a two positive definite operators. Now, taking $A := X^{-1}, C :=
+  \sigma^2Y_{\eps}^{-1}, U := I$ and $V := I$ and using Wodbury's matrix
+  identity in reverse order:
+  \begin{align*}
+    \begin{split}
+      X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X &= (X^{-1} + \sigma^{-2}Y_{\eps})^{-1} \\
+      %
+      %
+      %
+      &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \eps I))^{-1}.
+    \end{split}
+  \end{align*}
+
+  We conclude that $\forall \eps > 0$
+  \begin{align*}
+    \begin{split}
+      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* 
+     &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \eps I))^{-1}.
+    \end{split}
+  \end{align*}
+  Letting $\eps \to 0$ completes the proof.
+\end{proof}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\begin{lemma}[Increase due to a Measurement]\label{lemma:design increase}
+  Let $\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm := (\meas_1,\dots,\meas_{m-1})^t$. Then
+  \begin{align*}
+    \tar( \obs ) - \tar (\obsm ) &=
+    \frac12 \log \left ( 1 + \frac{
+      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m,
+      (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m \rangle
+    }{
+      \sigma^2 + \meas_m \modcov \meas_m - \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m 
+    }       
+    \right ).
+  \end{align*}
+\end{lemma}
+\begin{proof}
+  We use the Schur complement to write one inverse in terms of the other and
+  introduce notations to make the derivation cleaner. Note that we think of
+  $\obsm$ and $\obsm^*$ as column and row vectors (respectively).
+  \begin{align*}
+    \Sigma( \obs ) &= \Sigma = 
+    \begin{bmatrix}
+      \Sigma (\obsm )           & \obsm \modcov \meas_m \\
+      \meas_m \modcov \obsm^*   & \sigma^2 + \meas_m \modcov \meas_m
+    \end{bmatrix}
+    =
+    \begin{bmatrix}
+      \Sigmam   & w \\
+      w^t       & c
+    \end{bmatrix}\\
+    %
+    %
+    %
+    \Sigma^{-1} &=
+    \begin{bmatrix}
+      \Sigmam^{-1} + \Sigmam^{-1} w ( c - w^t \Sigmam^{-1} w)^{-1} w^t \Sigmam^{-1} & - \Sigmam^{-1} w ( c - w^t \Sigmam^{-1} w)^{-1} \\
+      -( c - w^t \Sigmam^{-1} w)^{-1} w^t \Sigmam^{-1}                            &  ( c - w^t \Sigmam^{-1} w)^{-1}
+    \end{bmatrix} \\
+    &=
+    \begin{bmatrix}
+      \Sigmam^{-1} & 0 \\
+      0           & 0 
+    \end{bmatrix}
+    + (c -w^t \Sigmam^{-1} w )^{-1}
+    \begin{bmatrix}
+      \Sigmam^{-1} w \\
+      -1
+    \end{bmatrix}
+    \begin{bmatrix}
+      w^t \Sigmam^{-1} & -1 
+    \end{bmatrix}
+  \end{align*}
+  %
+  Further, define
+  %
+  \begin{align*}
+    \M (\obs ):&= \prcov^{\frac12}\fwd^* \obs^* \Sigma^{-1} \obs \fwd
+    \prcov^{\frac12}    
+  \end{align*}
+  %
+  and note that, using our understanding of what is a column vector and
+  what is a row vector:
+  %
+  \begin{align*}
+    \M(\obs) &= \prcov^{1/2} \fwd^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2} \\
+    %
+    %
+    %
+    &= \prcov^{1/2} \fwd^* \obs^* \left \{
+    \begin{bmatrix}
+      \Sigmam^{-1} & 0 \\
+      0           & 0 
+    \end{bmatrix}
+    + (c -w^t \Sigmam^{-1} w )^{-1}
+    \begin{bmatrix}
+      \Sigmam^{-1} w \\
+      -1
+    \end{bmatrix}
+    \begin{bmatrix}
+      w^t \Sigmam^{-1} & -1 
+    \end{bmatrix} 
+    \right \} \obs \fwd \prcov^{1/2} \\
+    %
+    %
+    %
+    &= \M (\obsm) + (c -w^t \Sigmam^{-1} w )^{-1}
+    \prcov^{1/2} \fwd^* \obs^*
+    \begin{bmatrix}
+      \Sigmam^{-1} w \\
+      -1
+    \end{bmatrix}
+    \begin{bmatrix}
+      w^t \Sigmam^{-1} & -1 
+    \end{bmatrix} 
+    \obs \fwd \prcov^{1/2}
+  \end{align*}
+  %
+  Now, denote:
+  %
+  \begin{align}\label{eq:u}
+    \begin{split}
+      u :&= (c -w^t \Sigmam^{-1} w )^{-1/2}
+      \prcov^{1/2} \fwd^* \obs^* 
+      \begin{bmatrix}
+        \Sigmam^{-1} w \\
+        -1 
+      \end{bmatrix} \\
+      %
+      %
+      %
+      & = (c -w^t \Sigmam^{-1} w )^{-1/2} ( \prcov^{1/2}\fwd^* \obsm^* \Sigmam^{-1} \obsm  \modcov \meas_m - \prcov^{1/2} \fwd^* \meas_m )\\
+      %
+      %
+      %
+      u^* :&=  (c -w^t \Sigmam^{-1} w )^{-1/2} (\meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \fwd \prcov^{1/2} - \meas_m \fwd \prcov^{1/2} ),
+    \end{split}
+  \end{align}
+  %
+  so that
+  %
+  \begin{equation}\label{eq:M plus I}
+    I + \M( \obs ) = I + \M (\obsm ) + uu^*.
+  \end{equation}
+  %
+  Note that
+  \begin{equation}\label{eq:M postcov}
+    \prcov^{1/2} \left (I + \M( \obsm ) \right )^{-1} \prcov^{1/2} = \postcovm.
+  \end{equation}
+  The increase in the design criterion gained by including $\meas_m$
+  in the design is found by using Lemma \ref{lemma:MDL} the above
+  results:
+  %
+  \begin{align*}
+    \tar( \obs ) - \tar( \obsm )
+    %
+    %
+    %
+    &= \frac12 \log \det \Big ( I + \M ( \obs ) \Big ) / \det \Big ( I + \M (\obsm) \Big ) \\
+    %
+    %
+    %
+    &= \frac12  \log \det \left ( I + \M(\obsm) + uu^* \right ) / \det \Big ( I + \M (\obsm) \Big ) \\
+    %
+    %
+    %
+    &= \frac12 \log \left ( 1 + \left \langle \left ( I+\M(\obsm) \right )^{-1} u, u  \right \rangle \right ).
+  \end{align*}
+  Using \eqref{eq:u}:
+  \begin{align*}
+    &\left \langle \left (I+\M (\obsm)\right )^{-1}u, u \right \rangle\\
+    &= \frac{
+      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m,
+      (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m \rangle
+    }{
+      c- w^t \Sigmam^{-1} w
+    }\\
+    %
+    %
+    %
+    &= 
+    \frac{
+      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m,
+      (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m \rangle
+    }{
+      \sigma^2 + \meas_m \modcov \meas_m - \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m 
+    }
+  \end{align*}
+  and the conclusion follows.
+\end{proof}
+
+Lemma \ref{lemma:design increase} implies the following two corollaries.
+\begin{corollary}[Gain for No Model Error]\label{cor:zero mod err}
+  If $\modcov = 0$, then
+  \begin{equation*}
+    \tar( \obs ) - \tar (\obsm )
+    = -\frac12 \log (1 - \sigma^{-2} \langle \fwd \postcov \fwd^* \meas_m, \meas_m \rangle ).
+  \end{equation*}
+\end{corollary}
+\begin{proof}
+  Note that this is not immediate by substituting $\modcov = 0$ in the
+  conclusion of Lemma \ref{lemma:design increase}, since we make a
+  claim for $\postcov$, and not $\postcovm$. Let us first review
+  \eqref{eq:u} and note that since $\modcov = 0$ the covariance
+  $\Sigmam = \sigma^2I_{m-1}$ and $w = 0$, so $c - w^t\Sigmam^{-1}w =
+  \sigma^2$:
+  \begin{align*}
+    u :&=-\sigma^{-1}\prcov^{1/2} \fwd^* \meas_m\\
+    %
+    %
+    u^* :&= -\sigma^{-1} \meas_m \fwd \prcov^{1/2}.
+  \end{align*}
+  From \eqref{eq:M plus I}:
+  \begin{equation*}
+    I + \M(\obsm) = I +\M(\obs) - uu^*,
+  \end{equation*}
+  and thus:
+  \begin{equation*}
+    \left \langle \left ( I +\M(\obs) \right )^{-1} u, u \right \rangle
+    = \sigma^{-2} \langle \fwd \postcov \fwd^* \meas_m, \meas_m \rangle.
+  \end{equation*}
+  Analogously to \eqref{eq:M postcov} we note that
+  \begin{equation*}
+    \prcov^{1/2} \left ( I + \M(\obs) \right )^{-1} \prcov^{1/2} = \postcov.
+  \end{equation*}
+  Using Lemma \ref{lemma:MDL} we conclude
+  \begin{align*}
+    \tar( \obs ) - \tar( \obs )
+    &= \frac12 \log \det \left (I +\M(\obs) \right ) / \det \left (I + \M(\obsm) \right ) \\
+    %
+    %
+    %
+    &= \frac12 \log \det \left (I +\M(\obs) \right ) / \det \left (I + \M(\obs) - uu^* \right ) \\
+    %
+    %
+    %
+    &=-\frac12 \log (1 - \langle (I+\M(\obs))^{-1}u, u \rangle \\
+    %
+    %
+    %
+    &= -\frac12 \log (1 - \sigma^{-2} \langle \fwd \postcov \fwd^* \meas_m, \meas_m \rangle ).
+  \end{align*}
+\end{proof}
+
+\samemeas
+\begin{proof} \label{cor:same meas proof}
+  Denote $A:= \obs \modcov \obs^*$ and $v_j$ the $j$th column of $A$.
+  Note that $v_j = \obsm \modcov \meas_m$, since $(\obsm \modcov
+  \obsm^*)_{ij} = \meas_i(\modcov \meas_j)$, as explained in
+  \eqref{eq:modcov explained}. One can now verify that
+  \begin{equation}\label{eq:observation}
+    \Sigmam^{-1} \obsm \modcov \meas_m = \Sigmam^{-1}v_j = (A +\sigma^2I_{m-1})^{-1} v_j =
+    e_j -\sigma^2 \Sigmam^{-1}e_j.
+  \end{equation}
+  %
+  Using \eqref{eq:observation}:
+  \begin{align}\label{eq:denominator}
+    \begin{split}
+      \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m
+      &= \meas_m \modcov \obsm^* ( e^j - \sigma^2 \Sigmam^{-1} e_j )\\
+      %
+      %
+      %
+      &= \meas_m \modcov \meas_j - \sigma^2 \meas_m \modcov \obsm^* \Sigmam^{-1}e_j \\
+      %
+      %
+      %
+      &= \meas_m \modcov \meas_j -\sigma^2 (e_j - \sigma^2 \Sigmam^{-1}e_j)^t e_j \\
+      %
+      %
+      %
+      &= \meas_m \modcov \meas_m -\sigma^2 + \sigma^4 e_j^t\Sigmam^{-1}e_j.
+    \end{split}
+  \end{align}
+  We use \eqref{eq:observation} to simplify the terms in the enumerator of
+  the conclusion of Lemma \ref{lemma:design increase}:
+  \begin{align}\label{eq:enumerator}
+    \begin{split}
+      (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m
+      &= \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m - \meas_m \\
+      %
+      %
+      %
+      &= \obsm^* (e_j - \sigma^2 \Sigmam^{-1} e_j) -\meas_j \\ 
+      %
+      %
+      %
+      &= -\sigma^2 \obsm^* \Sigma^{-1}e_j. 
+    \end{split}
+  \end{align}
+  %
+  Substitute \eqref{eq:enumerator} and \eqref{eq:denominator} to
+  the enumerator and denominator (respectively) of the conclusion of
+  Lemma \ref{lemma:design increase}:
+  %
+  \begin{align*}
+    \tar( \obs ) - \tar (\obsm ) &=
+    \log \left ( 1 + \frac{
+      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m,
+      (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m \rangle
+    }{
+      \sigma^2 + \meas_m \modcov \meas_m - \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m 
+    }       
+    \right ) \\
+    %
+    %
+    %
+    &= \log \left ( 1 + \frac{\sigma^4
+      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
+      \obsm^* \Sigmam^{-1}e_j \rangle
+    }{
+      2\sigma^2 - \sigma^4 e_j^t\Sigmam^{-1}e_j 
+    }       
+    \right ) \\
+    %
+    %
+    %
+    &= \log \left ( 1 + \frac{\sigma^2
+      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
+      \obsm^* \Sigmam^{-1}e_j \rangle
+    }{
+      2 - \sigma^2 e_j^t\Sigmam^{-1}e_j 
+    }       
+    \right ).
+  \end{align*}
+\end{proof}
+
+\begin{lemma}[Auxilliary Calculations]\label{lemma:aux calc}
+  Let $T(\obs) := \obs^* \Sigma^{-1}(\obs)\obs$, with $\Sigma(\obs)$
+  defined as in \eqref{eq:Sigma}. Then:
+  \begin{align*}
+    \delta T(\obs)V &= V^* \Sigma^{-1} \obs 
+    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
+    &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs
+    + \obs^* \Sigma^{-1} V.
+  \end{align*}
+\end{lemma}
+
+\begin{proof}
+  We need a few supplementary calculations. First:
+  \begin{align}\label{eq:der sig}
+    \begin{split}
+      \frac{\der}{\der \tau} \Big |_{\tau=0} \Sigma( \obs + \tau V )
+      &= \frac{\der}{\der \tau} \Big |_{\tau=0} 
+      (\obs + \tau V ) \modcov (\obs + \tau V )^*  + \sigma^2I\\
+      % 
+      % 
+      % 
+      &= V \modcov \obs^* + \obs \modcov V^*.
+    \end{split}
+  \end{align}
+  By the standard trick for the derivative of an operator: 
+  \begin{align*}
+    0 &= \frac{\der}{\der \tau} \Big |_{\tau=0} I \\
+    % 
+    % 
+    % 
+    &= \frac{\der}{\der \tau} \Big |_{\tau=0}
+    \left (\Sigma(\obs+\tau V)^{-1} \Sigma(\obs+\tau V) \right ) \\
+    % 
+    % 
+    % 
+    &= \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0} \Sigma+
+    \Sigma^{-1} \frac{\der \Sigma(\obs+\tau V)}{\der \tau} \Big |_{\tau=0}\\  
+    %
+    %
+    %
+    &= \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0} \Sigma+
+    \Sigma^{-1} (V\modcov \obs^* + \obs \modcov V^*) 
+    \text{, by \eqref{eq:der sig}. }
+  \end{align*}
+  Now we can write the variation of $\Sigma^{-1}$:
+  \begin{align}\label{eq:der sig inv}
+    \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0}  
+      &= -\Sigma^{-1} (V \modcov \obs^* + \obs \modcov V^*) \Sigma^{-1}.
+    \end{align}
+  % 
+  Finally, we can do the calculation we need to. By Leibnitz (product) rule
+  and \eqref{eq:der sig inv}:
+  % 
+  \begin{align*}
+    \delta T(\obs) V 
+    &= \frac{\der T(\obs + \tau V)}{\der \tau} \Big |_{\tau=0} \\
+    %
+    %
+    %
+    &= V^* \Sigma^{-1} \obs 
+    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
+    &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs
+    + \obs^* \Sigma^{-1} V,
+  \end{align*}
+  as required.
+\end{proof}
+
+\biggerbetter
 \begin{proof} 
   Fix an arbitrary $j=1,\dots,m$ and take $V:= e_j e_j^t \obs$. We see
-  that for $u \in \ban$
+  that for $u \in \hilo$
   \begin{equation*}
@@
   increasing the magnitude of the $j$th measurement functional
-  increases $\tar(\obs)$. We start with the last line of \eqref{eq:tar
-    var}, and denote $\tmp : = \fwd \postcov \fwd^*: \ban^* \to \ban$:
+  increases $\tar(\obs)$. We start with the last line of \eqref{eq:tar var}, and denote $\tmp : = \fwd \postcov \fwd^*$.
   \begin{align*}
      \delta \tar(\obs) V 
-    &= \tr V ( I - \modcov \obs^*\Sigma^{-1}\obs) \tmp \obs^* \Sigma^{-1} \\
-    % 
-    %
-    %
-    &= \tr e_je_j^t \obs ( I - \modcov \obs^*\Sigma^{-1}\obs) \tmp \obs^* \Sigma^{-1} \\
+    &= \tr{V ( I - \modcov \obs^*\Sigma^{-1}\obs) \tmp \obs^* \Sigma^{-1}} \\
+    % 
+    %
+    %
+    &= \tr{e_je_j^t \obs ( I - \modcov \obs^*\Sigma^{-1}\obs) \tmp \obs^* \Sigma^{-1}} \\
     %
@@
 
-%% HERE
-%% Since a measurement that takes into account $\ker \fwd$ is
-%% necessarily sub-optimal, we may assume $\fwd$ is in fact
-%% invertible (it is, on $\ker \fwd ^{\perp}$) and rewrite the
-%% necessary condition \eqref{eq:eigenproblem} using
-%% \eqref{eq:postcov} as \begin{equation*} \sigma^{-2} \left ( (\fwd
-%% \prcov \fwd^*)^{-1} + \sigma^{-2}\obs\obs^* \right )^{-1} =
-%% \obs^* \Xi.  \end{equation*} By Lemma \ref{lemma:sim diag},
-%% taking $C = (\fwd \prcov \fwd^*)^{-1}$ we observe that $\meas_j,
-%% j=1,\dots,m$ are eigenvectors of $\fwd \prcov \fwd^*$. If they have
-%% different eigenvalue, they are orthogonal. If not, this means, by
-%% the common interpretation of the Lagrange multipliers, that an
-%% infinitesimal increase in one will cause the same increase in
-%% design criterion as others with the same eigenvalue. So if these
-%% eigenvectors do not have the same angle between every pair, an
-%% infinitesimal increase in one will give more increase than
-%% others, which is a contradiction. So in every subspace, vectors
-%% will be restricted to these forms and if there will be more
-%% vectors than the dimensionality of the subspace, clusterization
-%% will occur.
-
-%% \begin{figure}
-%%   \begin{tikzpicture}[thick, scale=1.4, every node/.style={scale=0.99}]
-%%     \begin{axis}
-%%       [
-%%         xmin = 0,
-%%         xmax = 3.14,
-%%         xlabel = {$x$},
-%%         ylabel = posterior std,
-%%         ymin   = 0,
-%%         %compat = 1.3,
-%%         % ymax   = 130,
-%%         % ytick = \empty,
-%%         legend cell align=left,
-%%         % legend style={at={(0.45,0.2)}}
-%%         legend pos= outer north east 
-%%       ]
-%%       % \draw[black!30!white, thin] (50,0) -- (50,130);
-%%       % 
-%%       \addplot [thin, black, mark=none] table{data/stdv-trunc-sens1-var1.txt};
-%%       \addlegendentry{1 sensor};
-
-%%       \addplot [thin, blue, mark=none] table{data/stdv-trunc-sens2-var1.txt};
-%%       \addlegendentry{2 sensors};
-
-%%       \addplot [thin, red, mark=none] table{data/stdv-trunc-sens3-var1.txt};
-%%       \addlegendentry{3 sensors};
-
-%%       \addplot [thin, green, mark=none] table{data/stdv-trunc-sens4-var1.txt};
-%%       \addlegendentry{4 sensors};
-
-%%       \addplot [thin, purple, mark=none] table{data/stdv-trunc-sens5-var1.txt};
-%%       \addlegendentry{5 sensors};
-
-%%       \addplot [thin, cyan, mark=none] table{data/stdv-trunc-sens6-var1.txt};
-%%       \addlegendentry{6 sensors};
-
-%%       \addplot [black,  only marks, mark=x, mark size=1.5] 
-%%       table{data/locs-trunc-sens1-var1.txt}; 
-%%       \addplot [blue,   only marks, mark=x, mark size=1.5]
-%%       table{data/locs-trunc-sens2-var1.txt}; 
-%%       \addplot [red,    only marks, mark=x, mark size=1.5]
-%%       table{data/locs-trunc-sens3-var1.txt};
-%%       \addplot [green,  only marks, mark=x, mark size=1.5] 
-%%       table{data/locs-trunc-sens4-var1.txt}; 
-%%       \addplot [purple, only marks, mark=x, mark size=1.5] 
-%%       table{data/locs-trunc-sens5-var1.txt}; 
-%%       \addplot [cyan,   only marks, mark=x, mark size=1.5] 
-%%       table{data/locs-trunc-sens6-var1.txt}; 
-%%     \end{axis}
-%%   \end{tikzpicture}
-%%   \caption{Analogously to figure \ref{fig:clusterization
-%%       illustration}, shown are posterior pointwise standard
-%%     deviations. Note the exact clustering. Here the dynamics is not
-%%     the one governed by the 1D heat equation but a dynamics that sets
-%%     to zero all modes except the first three. This dynamics simulates
-%%     a finite-rank operator as per the conditions of Theorem
-%%     \ref{theorem:finite rank}. Although the conditions of the theorem
-%%     allow the use of a different constraint on measurements (see
-%%     section \ref{subsec:unit length}), the conclusion of the theorem
-%%     still seems to hold.}
-%%   \label{fig:finite rank illustration}
-%% \end{figure}
-
-%% \section{Counterexample}\label{section:counterexample}
-%% Applying Lemma \ref{lemma:sim diag} to
-%% \begin{equation*}
-%%   (\fwd \postcov \fwd^*)^{-1} = (\fwd \prcov \fwd^*)^{-1} +
-%%   \sigma^{-2}\obs^*\obs
-%% \end{equation*}
-%% implies $(\fwd \prcov \fwd^*)^{-1}$ and $\obs^*\obs$ have the same
-%% eigenspaces. Since $\meas_j$ are eigenvectors of $\fwd \postcov
-%% \fwd^*$, we may expect this implies that $\meas_j$ are also
-%% eigenvectors of $(\fwd \prcov \fwd^*)^{-1}$ and assuming $(\fwd \prcov
-%% \fwd^*)^{-1}$ has simple eigenvalues completes the proof. Our
-%% expectations, however, are let down.
-
-%% Here is a counterexample. Let $e_1, e_2$ denote two
-%% eigenvectors of $C := (\fwd \prcov \fwd^*)^{-1}$ of different
-%% eigenvalues such that $C + \obs^*\obs$ is scalar on span
-%% $V:=\{e_1,e_2\}$ (we took $\sigma^2 = 1$ for simplicity). Then,
-%% on $V$, we have  
-%% \begin{equation*}
-%%   (\fwd \prcov \fwd^*)^{-1} = diag( d_1, d_2),
-%% \end{equation*}
-%% and we assume $d_1 > d_2$ without loss of generality. For $\meas \in
-%% \hil$, we denote $\meas^*$ its Riesz representer, making a departure
-%% from the standard notation employed, where $\meas$ is usually taken to
-%% be a linear functional on $\hil$. Let $\obs = (\meas_1^*, \meas_2^*,
-%% \meas_3^*)^t$. In this setup, we would like that $\meas_j,j=1,2,3$ are
-%% either orthogonal or parallel. We now outline construction that shows
-%% this is not necessarily true.
-
-%% We know, from Corollary \ref{cor:sim diag} that $\meas_j \in V$, so
-%% let
-%% \begin{align*}
-%%   \meas_1 &=  c_1e_1 + c_2e_2 \\
-%%   \meas_2 &= -c_1e_1 + c_2e_2 \\
-%%   \meas_3 &=     e_1.
-%% \end{align*}
-%% Then
-%% \begin{align*}
-%%   \meas_1 \meas_1^*
-%%   &= c_1^2e_1e_1^* + c_1c_2( e_1e_2^* + e_2e_1^*) +
-%%   c_2^2 e_2e_2^* \\
-%%   \meas_2 \meas_2^* &= c_1^2e_1e_1^* - c_1c_2( e_1e_2^* +
-%%   e_2e_1^*) + c_2^2 e_2e_2^*.
-%% \end{align*}
-%% Note that
-%% \begin{equation*}
-%%   \obs^* \obs = \meas_1 \meas_1^* + \meas_2\meas_2^* +
-%%   \meas_3\meas_3^* = (2c_1^2+1) e_1e_1^* + 2c_2^2e_2e_2^*,
-%% \end{equation*}
-%% is diaonal in $\{e_1,e_2\}$. In order for $\meas_j,j=1,2,3$ to be
-%% eigenvectors of $C + \obs^*\obs$ with the same eigenvalue, $C +
-%% \obs^*\obs$ must be a scalar operator on $V$. Thus the following
-%% must hold:
-%% \begin{align*}
-%%   d_1 + 2c_1^2 + 1 &= d_2 + 2c_2^2 \\
-%%   c_1^2 + c_2^2 &= 1 \text{ (arises from $\| \meas_i \| = 1$).}
-%% \end{align*}
-%% Equivalently,
-%% \begin{equation*}
-%%   \begin{bmatrix}
-%%     2 & -2 \\
-%%     1 & 1
-%%   \end{bmatrix}
-%%   \begin{bmatrix}
-%%     c_1^2 \\
-%%     c_2^2
-%%   \end{bmatrix}
-%%   =
-%%   \begin{bmatrix}
-%%     d_2 - d_1 - 1 \\
-%%     1
-%%   \end{bmatrix}.
-%% \end{equation*}
-%% We can always find a solution for this system, for any $d_1,d_2$.
-%% However, it is necessary that either $c_1^2 = 1$ or $c_2^2 = 1$ for
-%% $\meas_j, j=1,2,3$ to all be either orthogonal or parallel. The first
-%% conditiion implies $d_2 - d_1 = 3$, which is impossible, since $d_1 >
-%% d_2$. The second requires $d_1 - d_2 = 1$, a condition that does not
-%% hold generically.
-
-%    Text of article.
-
-%    Bibliographies can be prepared with BibTeX using amsplain,
-%    amsalpha, or (for "historical" overviews) natbib style.
 \bibliographystyle{amsplain}
-\bibliography{refs.bib,georg_refs}
+\bibliography{refs,georg_refs}
 
 \end{document}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%    Templates for common elements of an article; for additional
-%    information, see the file Author_Handbook_ProcColl.pdf, included
-%    in every AMS author package, and the amsthm user's guide, linked
-%    from http://www.ams.org/tex/amslatex.html .
-
-%    Section headings
-\section{}
-\subsection{}
-
-%    Ordinary theorem and proof
-\begin{theorem}[Optional addition to theorem head]
-% text of theorem
-\end{theorem}
-
-\begin{proof}[Optional replacement proof heading]
-% text of proof
-\end{proof}
-
-%    Figure insertion; default placement is top; if the figure occupies
-%    more than 75% of a page, the [p] option should be specified.
-\begin{figure}
-\includegraphics{filename}
-\caption{text of caption}
-\label{}
-\end{figure}
-
-%    Mathematical displays; for additional information, see the amsmath
-%    user's guide:  texdoc amamath  or
-%  http://mirror.ctan.org/tex-archive/macros/latex/required/amsmath/amsldoc.pdf
-
-% Numbered equation
-\begin{equation}
-\end{equation}
-
-% Unnumbered equation
-\begin{equation*}
-\end{equation*}
-
-% Aligned equations
-\begin{align}
-  &  \\
-  &
-\end{align}
-
-%-----------------------------------------------------------------------
-% End of article-template.tex
-%-----------------------------------------------------------------------

### 2020-11-22 zeros.py

@@ -135,6 +135,16 @@ def main(m, k):
     assert len(norms) == m
     assert np.all(np.abs(norms - 1) < EPS) 
 
+
+    for vec in A.T:
+        pos_ind = np.all(np.abs(A.T - vec) < EPS, axis=1)
+        neg_ind = np.all(np.abs(A.T + vec) < EPS, axis=1)
+        ind = np.logical_or(pos_ind, neg_ind).astype(int)
+        assert ind.shape == (A.shape[1],)
+        if np.sum(ind) > 1:
+            print(ind)
+            #pdb.set_trace()
+    
     
 def get_A(M, m):
     k = M.shape[0]
@@ -209,8 +219,8 @@ if __name__ == '__main__':
     try:
         test_givens()
         test_theta()
-        for m in range(0,30,3):
-            for k in range(2, m+1, 2):
+        for m in range(0,10,3):
+            for k in range(2, m+1, 1):
                 print(m, k) 
                 main(k=k, m=m)
                        

### 2020-12-30 complex.py
Restructure a bit

@@ -1,0 +1,130 @@
+## The complex case. This works, dont touch
+import numpy as np
+from matplotlib import pyplot as plt
+  
+plt.rcParams['figure.figsize'] = [16, 4]
+plt.rcParams['figure.dpi'] = 100 # 200 e.g. is really fine, but slower
+
+def heat_multiplier(N, L, T, alpha=1):
+    freqs = np.fft.fftfreq(N, d=L/N) 
+    return np.exp(-alpha * np.pi**2 * T * freqs**2)
+
+def change_basis(u):
+    return np.fft.fft(u)/u.size
+
+def heat(u0, T, L=1, alpha=1):
+    N = u0.size
+    u0_hat = change_basis(u0)
+    uT_hat = u0_hat * heat_multiplier(N, L, T, alpha=alpha)
+    uT = np.fft.ifft(uT_hat)*N 
+    return uT
+
+def coeff2u(coeff, x, L):
+    N = x.shape[0]
+    freqs = np.fft.fftfreq(N, d=L/N) 
+    return sum(coeff[i]*np.exp(2j*np.pi*freqs[i]*x) for i in np.where(coeff)[0])
+
+
+T = 1e-2
+N = 1000
+L = 9
+alpha = 1.4
+x = np.linspace(0, L, num=N, endpoint=False)
+multiplier = heat_multiplier(N, L, T, alpha=alpha)
+
+
+## Initial condition
+initial = np.random.randn(N, 2).view(np.complex128) 
+initial = np.squeeze(initial)
+initial = initial / (np.arange(N)+1)**1.5  
+u0 = coeff2u(initial, x, L)
+
+## Numeric solution
+uT_num = heat(u0, T, L, alpha)
+
+## Analytic
+final = initial * multiplier
+uT = coeff2u(final, x, L)
+    
+plt.plot(x, u0.imag, label='IC')
+plt.plot(x, uT.imag, label='Analytic FC')
+plt.plot(x, uT_num.imag, label='Numeric FC')
+plt.legend()
+plt.show()
+
+
+u0_hat = change_basis(u0)
+err = np.max(np.abs(uT - uT_num))
+assert  err < 1e-9, err
+
+err = np.max(np.abs(u0_hat - initial))
+assert  err < 1e-9, err
+
+
+
+
+# ## Another case which is mehhh
+# import numpy as np
+# from matplotlib import pyplot as plt
+  
[…]

### 2022-04-07 latex/paper.tex
Fixing problems so that bound actually works

--- 2020-11-22 paper.tex
@@
 and the prior are assumed independent of each other. The observation
-and forward operators are merged $\tmp := \obs \fwd$. Since we need to
-separate the observation operator $\obs$ from the forward operator
-$\fwd$ we write this explicitly in \eqref{eq:inverse problem}. We now
-turn to understand the terms present in \eqref{eq:inverse problem}.
+and forward operators are typically merged $\tmp := \obs \fwd$. Since
+we need to separate the observation operator $\obs$ from the forward
+operator $\fwd$ we write this explicitly in \eqref{eq:inverse
+  problem}. We now turn to understand the terms present in
+\eqref{eq:inverse problem}.
 
@@
 The intuition behind definition \ref{def:d optimality} is
-straightforward. From \eqref{eq:objective} and \eqref{eq:objective}:
+straightforward. From \eqref{eq:postcov} and \eqref{eq:objective}:
 \begin{align*}
@@
 $\obs$, namely $\meas_j,j=1,\dots, m$, are the eigenvectors. Note,
-however, that $\postcov$ depends on $\obs$. This makes
-\eqref{eq:eigenproblem} a nonlinear eigenvalue problem.
+however, that $\postcov$ depends on $\obs$, and so
+\eqref{eq:eigenproblem} is a nonlinear eigenvalue problem.
 
@@
   %
-  &= \sigma^{-2} \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1} \obs^*
+  &= \sigma^{-2} \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1} \obs^*.
   \end{split}
 \end{align}  
-The term $(\fwd \prcov \fwd^*)^{-1}$ is the prior precision in $\hilo$
-($\fwd\prcov\fwd^*$ is invertible, as mentioned in section
-\ref{section:prelim}). $(\fwd\prcov\fwd^*)^{-1} + \sigma^{-2} \obs^*
-\obs$ is the posterior precision in $\hilo$.
+The term $(\fwd \prcov \fwd^*)^{-1}$ is the prior precision in
+$\hilo$, and $\fwd\prcov\fwd^*$ is invertible, as mentioned in
+section \ref{section:prelim}. $(\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}
+\obs^* \obs$ is the posterior precision in $\hilo$.
 
@@
 By definition of the objective \eqref{eq:objective}, \eqref{eq:pre
-  manipulations} and Corollary \ref{cor:samve ev}:
+  manipulations} and Corollary \ref{cor:same ev}:
 \begin{align}\label{eq:mod objective}
@@
   The proof is boring and amounts to using Woodbury's matrix identity
-  twice and a regularization trick. The standard proof for this
-  identity works in infinite dimensions, as long as all terms are well
-  defined. Unfortunately, $\obs^*\obs$ is not invertible, so we force
-  it to be. Throughout, we let $X := \fwd\prcov \fwd^*$ and $Y_{\eps}
-  := \obs^*\obs + \eps I$, for some small $\eps > 0$. Note that now
-  $Y_{\eps}$ is invertible. Recall Woodbury's matrix identity:
+  twice and a regularization trick. The standard proof for Woodbury's
+  matrix identity works in infinite dimensions, as long as all terms
+  are well defined. Unfortunately, $\obs^*\obs$ is not invertible, so
+  we force it to be. Recall Woodbury's matrix identity:
   $$
-  (A + UCV)^{-1} = A^{-1} + A^{-1}U(C^{-1} + VA^{-1}U)^{-1}VA^{-1}.
+  (A + UCV)^{-1} = A^{-1} - A^{-1}U(C^{-1} + VA^{-1}U)^{-1}VA^{-1}.
   $$
@@
   \end{align*}
-  Note that $X + \sigma^2 Y_{\eps}^{-1}$ is invertible, as the sum of
-  a two positive definite operators. Now, taking $A := X^{-1}, C :=
-  \sigma^2Y_{\eps}^{-1}, U := I$ and $V := I$ and using Wodbury's matrix
-  identity in reverse order:
+  Let $X := \fwd\prcov \fwd^*$ and $Y_{\eps} := \obs^*\obs + \eps I$,
+  for some small $\eps > 0$, so that $Y_{\eps}$ is invertible.  Note
+  that $X + \sigma^2 Y_{\eps}^{-1}$ is invertible, as the sum of two
+  positive definite operators. Now, taking $A := X^{-1}, C :=
+  \sigma^2Y_{\eps}^{-1}, U := I$ and $V := I$ and using Wodbury's
+  matrix identity in reverse order:
   \begin{align*}
@@
     &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs
-    + \obs^* \Sigma^{-1} V,
-  \end{align*}
-  as required.
+    + \obs^* \Sigma^{-1} V.
+  \end{align*}
 \end{proof}

### 2022-04-07 main.py

@@ -4,7 +4,7 @@ import numpy as np
 import pandas as pd
 from joblib import Parallel, delayed
 from matplotlib import pyplot as plt
-from observations import PointObservation
+from src.observations import PointObservation
 from scipy.stats import gaussian_kde
 from src.forward import Heat
 from src.probability import Posterior, Prior
@@ -19,8 +19,8 @@ from src.probability import Posterior, Prior
 def main(n_jobs=6):
     L, N = np.pi, 800
     specs = {'N': N, 'L': L}
-    ms = [3, 5, 7] #, 37, 61]
-    transforms = ['dct', 'fft']
+    ms = [1, 2, 3, 5, 7, 10, 12]
+    transforms = ['fft']
     fig1, axes1 = plt.subplots(figsize=(30, 16), nrows=len(transforms), ncols=len(ms))
     fig2, axes2 = plt.subplots(figsize=(30, 16), nrows=len(transforms), ncols=len(ms))
     figs = [fig1, fig2]
@@ -38,65 +38,71 @@ def main(n_jobs=6):
 
             # Optimal diagonal
             post.make_optimal_diagonal(m)
-            diagonal_utility = post.diagonal_utility(post.optimal_diagonal)
+            diagonal_utility = post.diagonal_utility(post.optimal_diagonal_O)
             # ax.scatter(diagonal_utility, kde(diagonal_utility), label='optimal diagonal')
             print(f"utility: Diag={diagonal_utility:3.3f}", end=' ')
 
-            # Random point measurements
-            measurements = np.random.uniform(low=0, high=L, size=(100, m))
-            utilities = Parallel(n_jobs=n_jobs)(delayed(post.minimization_point)(x) for x in measurements)
-            utilities = -np.array(utilities)
-            distances = Parallel(n_jobs=n_jobs)(delayed(post.close2diagonal)(x) for x in measurements)
-            f = lambda x: np.sum(PointObservation(measurements=x, **post.specs).eigenvalues())
-            sum_eigenvalues = Parallel(n_jobs=n_jobs)(delayed(f)(x) for x in measurements)
-            random = pd.DataFrame({'utility': utilities,
-                                   'diag_dist': distances,
-                                   'target': 'random',
-                                   'm': m,
-                                   'transform': transform,
-                                   'L': L,
-                                   'N': N,
-                                   'sum_eigenvalues': sum_eigenvalues})
-            random['x'] = [x for x in measurements]
+            # # Random point measurements
[…]

### 2022-04-07 src/observations.py

@@ -14,7 +14,7 @@ class PointObservation(Observation):
         self.measurements = np.array(measurements)
         self.multiplier = np.zeros(self.shape, dtype=self.dtype)
         for k in range(self.N):
-                self.multiplier[:, k] = self.eigenfunction(k)(self.measurements)
+            self.multiplier[:, k] = self.eigenfunction(k)(self.measurements)
 
     def __str__(self):
         return 'Point observations at ' + ', '.join([f'{me:.4f}' for me in self.measurements])

### 2022-04-07 src/probability.py

@@ -70,28 +70,28 @@ class Posterior(FourierMultiplier):
         self.prior = Prior(**self.specs) if prior is None else prior
         self.sigSqr = sigSqr
         self.C_sqrt_fwd = np.sqrt(self.prior.multiplier) * self.fwd.multiplier
-        self.optimal_diagonal = None
         assert np.all(np.abs(self.C_sqrt_fwd.imag) < 1e-12)
         assert np.all(self.C_sqrt_fwd.real >= 0)
 
 
     def make_optimal_diagonal(self, m):
         """Plots and returns the optimal design multiplier"""
-        eigenvalues = self.prior.inv_mult * self.sigSqr
-        data = []
-        for k in range(1, m + 8):
+        eigenvalues = self.prior.inv_mult * self.sigSqr / np.abs(self.fwd.multiplier)**2
+        # import pdb; pdb.set_trace()
+        k = 1
+        while True:
             eigs = eigenvalues[:k]
             uniform = (np.sum(eigs) + m) / k
-            extra = uniform - eigs
-            # assert abs(np.sum(extra) - num_observations) < 1e-9
-            if np.any(extra < 0):
+            if np.any(eigs >= uniform):
                 break
-            data.append({'eigs': eigs, 'extra': extra})
+            k += 1
+            self.optimal_diagonal_O = np.sqrt(uniform - eigs)
+            
+        self.optimal_diagonal_O_matrix = np.zeros((m, self.N))
+        np.fill_diagonal(self.optimal_diagonal_O_matrix, self.optimal_diagonal_O)
 
-        self.optimal_diagonal = data[-1]['extra']
-        self.optimal_diagonal_matrix = np.zeros((m, self.N))
-        np.fill_diagonal(self.optimal_diagonal_matrix, self.optimal_diagonal)
-        assert abs(np.sum(self.optimal_diagonal) - m) < 1e-12
+        power = np.sum(self.optimal_diagonal_O**2)
+        #assert abs(power - m) < 1e-12, (power, m)
 
     def operators(self, obs):
         self.A = np.einsum('ij,j->ij', obs.multiplier, self.fwd.multiplier)
@@ -133,13 +133,16 @@ class Posterior(FourierMultiplier):
         return abs(utility[0]) * utility[1]
 
     def diagonal_utility(self, diag):
-        tmp = (self.C_sqrt_fwd[:diag.size] * diag) ** 2 / self.sigSqr + 1
+        #tmp = self.C_sqrt_fwd.copy()
[…]

### 2022-04-10 main.py
Changed a bit. Currently we do not succeed with the bound but from tests.py I see there is a normalization issue where i need to normalize something by sqrt(L)

@@ -17,7 +17,7 @@ from src.probability import Posterior, Prior
 # plt.rcParams['text.color'] = 'white'
 
 def main(n_jobs=6):
-    L, N = np.pi, 800
+    L, N = 1, 800
     specs = {'N': N, 'L': L}
     ms = [1, 2, 3, 5, 7, 10, 12]
     transforms = ['fft']

### 2022-04-10 src/probability.py

@@ -56,7 +56,7 @@ class Prior(FourierMultiplier):
     def invstr(self):
         return '$(-\Delta)^{' + str(-self.gamma) + '}$'
 
-
+        
 class Posterior(FourierMultiplier):
     ''' Based on Stuart 2.16a and 2.16b'''
 

### 2022-04-10 tests/examples.py

@@ -1,9 +1,10 @@
 import numpy as np
 import pytest
-from forward import Heat
-from multiplier import FourierMultiplier
-from observations import DiagObservation, PointObservation
-from probability import Prior, Posterior
+
+from src.forward import Heat
+from src.multiplier import FourierMultiplier
+from src.observations import DiagObservation, PointObservation
+from src.probability import Prior, Posterior
 
 N = 600
 L = 3

### 2022-04-10 tests/test_probability.py

@@ -2,9 +2,9 @@ import pandas as pd
 from joblib import delayed, Parallel
 from matplotlib import pyplot as plt
 from numpy.testing import assert_allclose
-from probability import Posterior
 from scipy.stats import gaussian_kde
 
+from src.probability import Posterior
 from tests.examples import *
 
 
@@ -74,7 +74,7 @@ def test_prior_pointwise_std(prior):
     plt.title(f'transform = {prior.transform}')
     plt.xlabel('x')
     plt.ylabel('prior pointwise standard deviation')
-    # plt.show()
+    plt.show()
 
 
 @pytest.mark.parametrize("transform", ['dct', 'fft', 'dst'])
@@ -129,7 +129,7 @@ def test_posterior(posterior, point_observation):
 @pytest.mark.parametrize("m", range(2, 7))
 def test_unique_optimal(posterior, m):
     n_runs = 34
-    res = Parallel(n_jobs=7)(delayed(posterior.optimal)(m=m) for _ in range(n_runs))
+    res = Parallel(n_jobs=7)(delayed(posterior.optimize)(m=m) for _ in range(n_runs))
     successes = [r for r in res if r['success']]
     failures = [r for r in res if not r['success']]
 
@@ -143,7 +143,7 @@ def test_unique_optimal(posterior, m):
 
 @pytest.mark.parametrize("transform", ['dct', 'fft', 'dst'])
 def test_optimal_n_iterations(posterior):
-    posterior.optimal(m=3, n_iterations=5)
+    posterior.optimize(m=3, n_iterations=5)
 
 
 def test_default_posterior():

### 2022-06-16 latex/paper.tex
Revised paper, write stuff as corrolary, proposition etc. If this gets erased I die.

@@
   designs and reason how it may lead to sensor clusterization in the
-  case of iid measurement noise. We also spatially correlated
-  measurement and model errors mitigate the clusterization
-  phenomenon. As a part of the analysis we prove a Matrix Determinant
-  Lemma in Hilbert spaces and a lemma for calculating
-  $\frac{\der}{\der t} \log \det (I + X(t))$ for certain
-  operator-valued functions. We also show how to decompose $M = AA^t$
-  with $A$ constrained to have unit norm columns.
+  case of iid observation noise. We also consider how spatially
+  correlated observation and model errors mitigate the clusterization
+  phenomenon.
+
+  %% As a part of the analysis we prove a Matrix Determinant
+  %% Lemma in Hilbert spaces and a lemma for calculating
+  %% $\frac{\der}{\der t} \log \det (I + X(t))$ for certain
+  %% operator-valued functions. We also show how to decompose $M = AA^t$
+  %% with $A$ constrained to have unit norm columns.
 \end{abstract}
@@
 from the ground (e.g.\ searching for oil or using a radar). Whatever
-the allowed set of measurements is, one should select the optimal
-measurements to take, in order to increase accuracy, reduce costs, or
+the allowed set of observations is, one should select the optimal
+observations to take, in order to increase accuracy, reduce costs, or
 both.
@@
 \cite{ChalonerVerdinelli95}: Consider a linear model in finite
-dimensions, with Gaussian prior and noise. In this model, maximizing
-the expected information gain (KL divergence
+dimensions, with Gaussian prior and noise. Under this model,
+maximizing the expected information gain (KL divergence
 \cite{KullbackLeibler51,CoverThomas12}) between posterior and prior
@@
 using the D criterion, the optimization process results in
-measurements that are very similar. For example, if a measurement is
+observations that are very similar. For example, if an observation is
 thought of as measuring some function value at $\x \in \Omega
@@
 phenomenon and suggest an approach called clusterization-free design.
-In such designs, the user enforces measurement locations to be far
+In such designs, the user enforces observation locations to be far
 from each other. One way to do this is by introducing correlated
-errors which, philosophically, accounts for both measurement error and
+errors which, philosophically, accounts for both observation error and
 model error. Another method considered is imposing distance
-constraints between measurements. A somewhat different approach is
+constraints between observations. A somewhat different approach is
 suggested in \cite[page 49]{Fedorov12}, where close by design
-measurements are merged --- a procedure which obviously does not avoid
+observations are merged --- a procedure which obviously does not avoid
 clusterization. The same problem arises in time-series analysis for
@@
 We propose and thoroughly study a relaxed and analytically tractable
-model for understanding D-optimal designs. In this model, D-optimal
-design are solutions of a constrained optimization problem, formulated
-using Lagrange multipliers (section \ref{section:D and grad}). Using
-this model we rigorously show how model error mitigates clusterization
-(section \ref{section:non vanishing}). We show how clusterization
-occurs in a sequential design scenario (section
+model for understanding D-optimal designs. Under this model, D-optimal
+designs are solutions of a constrained optimization problem,
+formulated using Lagrange multipliers (section \ref{section:D and
+  grad}). This allows us to rigorously show how model error mitigates
+clusterization (section \ref{section:non vanishing}). We show how
+clusterization occurs in a sequential design scenario (section
 \ref{subsec:clusterization sequential}). Then, we consider a
@@
 
-In the process we generalize several lemmas from linear algebra to
+In the process, we generalize several lemmas from linear algebra to
 infinite-dimensional settings. We prove a Matrix Determinant Lemma in
@@
 also provide other tools for understanding D-optimal designs in
-infinite-dimensional Bayesian inverse problems. These includet tools
+infinite-dimensional Bayesian inverse problems. These include tools
 for calculating the increase in the design criterion per observation
@@
 model does not consider any specific set of allowed
-measurements. Rather, we take measurements in the unit ball in some
+observations. Rather, we take observations in the unit ball in some
 Hilbert space. This allows considerably less restrictive observations
@@
     determinant of the posterior covariance. The clusterization effect
-    can be clearly seen for six sensors. Only four measurement
+    can be clearly seen for six sensors. Only four observation
     locations are used --- two pairs of sensors are so close they are
@@
 \dots,m$ and a final time $T > 0$. We assume centered Gaussian
-measurement error, so we observe $v(\x_j,T) = u(\x_j,T) + \eps(\x_j)$
+observation error, so we observe $v(\x_j,T) = u(\x_j,T) + \eps(\x_j)$
 with $\eps(\x_j) \sim \normal(0, \sigma^2), \sigma > 0$ iid. We model
 the initial condition as $u_0 \sim \normal(0,\prcov)$, for $\prcov =
-(-\Delta)^{-1}$ with a homogeneous Dirichlet boundary condition. It is
-well known \cite{Tarantola05} that for linear problems, with Gaussian
-prior and error, the posterior is also Gaussian with a covariance that
-does not depend on the observed data. The posterior covariance
-$\postcov$ is known to have a closed form formula, even in infinite
-dimensions\cite{Stuart10}. We denote by $\fwd$ the dynamics operator,
-so that $u( \cdot,T) = \fwd u_0$, and the observation operator $\obs$
-so that $u(\x_j,T) = (\obs u)_j, j=1,\dots,m$. The posterior
-covariance is known and depends only on $\prcov, \fwd, \obs$ and
-$\sigma^2$ (see section \ref{section:prelim} and \eqref{eq:postcov}
-specifically).
+(-\Delta)^{-1}$ with a homogeneous Dirichlet boundary condition. If It
+is well known \cite{Tarantola05} that for linear problems, with
+Gaussian prior and error, the posterior is also Gaussian with a
+covariance that does not depend on the observed data. The posterior
+covariance $\postcov$ is known to have a closed form formula, even in
+infinite dimensions\cite{Stuart10}. We denote by $\fwd$ the dynamics
+operator, so that $u( \cdot,T) = \fwd u_0$, and the observation
+operator $\obs$ so that $u(\x_j,T) = (\obs u)_j, j=1,\dots,m$. The
+posterior covariance is known and depends only on $\prcov, \fwd, \obs$
+and $\sigma^2$ (see section \ref{section:prelim} and
+\eqref{eq:postcov} specifically).
 
@@
 actual data observed. The posterior covariance does, however, depend
-on location of the measurement taken. In figure
-\ref{fig:clusterization illustration}, measurement locations are
+on location of the observation taken. In figure
+\ref{fig:clusterization illustration}, observation locations are
 marked by circles. These were chosen to minimize (an expression
@@
 clusterization effect can be clearly seen for six sensors. It looks
-like only four measurements were taken. The reason is that two pairs
+like only four observations were taken. The reason is that two pairs
 of sensors are so close they are indistinguishable.
@@
 assumption we rely on below. Throughout this paper, we denote $m$ the
-number of measurements we are allowed to take. Observations
-(measurements) are taken via the linear observation operator $\obs \in
-( \hilo^* )^m$ (sometimes called a design). In an analogy with linear
-algebra (where row vectors are thought of as linear functionals), we
+number of observations we are allowed to take. Observations are taken
+via the operator $\obs \in ( \hilo^* )^m$. In an analogy with linear
+algebra, where row vectors are thought of as linear functionals, we
 think of $\obs$ as having ``rows'' $\meas_j, j=1,\dots,m$:
@@
 
-Each measurement is a linear functional, chosen from some subset of
-$\hilo^{*}$. Data is acquired via noisy measurements
+Each observation is a linear functional, chosen from some subset of
+$\hilo^{*}$. Data is acquired via noisy observations
 \begin{align}\label{eq:inverse problem}
@@
 with $\data \in \R^m$ and $\eps, \eps'$ defined next. We consider two
-types of noise / error terms. First, there is spatially correlated
-model error $\eps' \sim \normal(0,\modcov)$, modeled as a centered
-Gaussian measure on $\hilo$ with covariance operator $\modcov$. Then
-there is measurement / independent error $\eps \sim \normal(0,
-\sigma^2 I_m)$, with $I_m$ the $m \times m$ identity. Both error terms
-and the prior are assumed independent of each other. The observation
-and forward operators are typically merged $\tmp := \obs \fwd$. Since
-we need to separate the observation operator $\obs$ from the forward
-operator $\fwd$ we write this explicitly in \eqref{eq:inverse
-  problem}. We now turn to understand the terms present in
-\eqref{eq:inverse problem}.
+types of error terms. First, there is spatially correlated model error
+$\eps' \sim \normal(0,\modcov)$, modeled as a centered Gaussian
+measure on $\hilo$ with covariance operator $\modcov$. Then there is
+observation error $\eps \sim \normal(0, \sigma^2 I_m)$, with $I_m$ the
+$m \times m$ identity. Both error terms and the prior are assumed
+independent of each other. The observation and forward operators are
+typically merged $\tmp := \obs \fwd$
+\cite{AlexanderianGloorGhattas14}. Since we need to separate the
+observation operator $\obs$ from the forward operator $\fwd$ in order
+to carry out the analysis in the following sections, we explicitly
+separate $\fwd$ and $\obs$ in \eqref{eq:inverse problem}. We now turn
+to understand the terms present in \eqref{eq:inverse problem}.
 
 \subsubsection{Dynamics and Observation Operators}\label{subsec:dynamics}
-One can \textit{think} of $(\obs \fwd \param)_j$ as pointwise
-evaluations of a continuous function\footnote{Of course, this is
-  impossible if, e.g., $\hilo = L^2(\Omega)$ for some $\Omega$.}. Such
-measurements can be represented using the linear functionals
-$\delta_{\x_j} \in C(\Omega)^*$. In a different setting, we may be
-able to measure other quantities, e.g.\ some Fourier coefficients as
-is the case in MRI applications. Either way, the rows of $\obs$ are
-linear functionals and we refer to these as {\it measurement
-  vector}s. Note that we cannot choose every measurement vector we
-want. For example, we may be restricted only to pointwise evaluations
-of $\fwd \param$ by the sensors at our disposal, in which case
-measuring, e.g., the mean $\ell(u) = \int_{\Omega}u$ of a function is
-not possible.
+Rows of $\obs$ are linear functionals and we refer to these as {\it
+  observation vector}s. Note that we cannot choose every observation
+vector we want. For example, we may be restricted only to pointwise
+evaluations of $\fwd \param$ by the sensors at our disposal, in which
+case measuring, e.g., the mean $\ell(u) = \int_{\Omega}u$ of a
+function is not possible. Thus, it is best to think of $(\obs \fwd
+\param)_j$ as approximations to pointwise evaluations. For example, if
+$\hilo = L^2([-\pi, \pi])$, consider $\ell_1(x) := (2\pi
+\eps)^{-\frac{1}{2}}\exp(-\frac{x^2}{2\eps^2})$ and $\ell_2(x)
+:=\eps^{-1}\mathbf{1}_{[-\frac{\eps}{2}, \frac{\eps}{2}]}$. Both
+$\ell_1$ and $\ell_2$ approximate a point measurement at $0$ for small
+$\eps$ and, crucially, $\ell_1, \ell_2 \in \hilo^*$. It is important
+to note that although $\|\ell\|_1 = \|\ell_2\|_1=1$, unfortunately
+$\|\ell_1\|_2^2 = (4\pi \eps)^{-\frac{1}{2}}$ and $\|\ell_2\|_2^2 =
+\eps^{-1}$.
 
 \subsubsection{Error Terms}
-Considering the sum of the error terms it is easy to see that
+Considering the sum of the error terms, it is easy to see that
 $\bar{\eps} := \obs \eps' + \eps \in \R^m$. It is a centered Gaussian
@@
 mostly dropped for notational convenience, so $\Sigma(\obs) =
-\Sigma$. Thus, for a fixed $\obs$ (i.e.\ a fixed set of measurements)
+\Sigma$. Thus, for a fixed $\obs$ (i.e.\ a fixed set of observations)
 we can equivalently write $\data = \obs \fwd \param + \bar{\eps}$ with
@@
 allows us to consider model error (modeled by $\modcov$) as well as
-measurement error (modeled by $\sigma^2$). For example, say we believe
+observation error (modeled by $\sigma^2$). For example, say we believe
 our forward model does not capture some small scale phenomenon.  Then
@@
 
-The intuition behind definition \ref{def:d optimality} is
+\noindent The intuition behind Definition \ref{def:d optimality} is
 straightforward. From \eqref{eq:postcov} and \eqref{eq:objective}:
@@
 
-\subsection{Sequential vs Simultaneous Optimization}\label{subsec:seq vs sim}
-From defintion \ref{def:d optimality} we wish to characterize solution(s) of the
-following optimization problem for $\tar$. %%: (\hilo^*)^m \to \R$:
-\begin{align}\label{eq:optimization}
-  \obs^{\star} := \argmax_{\obs} \tar( \obs ) 
-  = \argmax_{\obs} \frac12 \log \det 
-  (I + \prcov^{1/2} \fwd^*\obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}),
-\end{align}
-where $\obs$ is constrained to some allowed set of measurements (we
-may \textit{think} of these as pointwise evaluations, even though
-these might not be in $\hilo^*$). We call this problem ``simultaneous
-optimization'', since all measurements are decided on simulatneously.
-
-For computational reasons, one may prefer to find the best
-measurements in a sequential manner. Denote
-\begin{equation}\label{eq:def obs_k}
-  \obs_k := (\meas_1,\dots, \meas_k)^t,  k\leq m.
-\end{equation}
-Sequential optimal design proceeds as follows. Find $\meas_1$ by
-maximizing $\tar(\obs_1)$. Then, keeping $\meas_1$ fixed --- find
-$\meas_2$ as the maximizer of $\tar(\obs_2)$. Then, find $\meas_3$ by
-keeping $\meas_1,\meas_2$ fixed and taking $\meas_3$ as the maximizer
-of $\tar(\obs_3)$. Continue this way until $\obs_m = \obs$ is
-found, where $m$ is the number of available measurements. It is
-important to notice that this scheme does not require actually
-observing data --- in \eqref{eq:objective} data is averaged out.
-
-The analysis in this paper is conducted for the general simultaneous
-optimization case. The sequential optimization case is dealt with in
-section \ref{subsec:clusterization sequential}. It is important to
-note, however that all conclusions we arrive at for the simultaneous
-case easily specialize to the sequential case by considering the
-posterior as the next sequential step's prior.
+%% \subsection{Sequential vs Simultaneous Optimization}\label{subsec:seq vs sim}
+%% From defintion \ref{def:d optimality} we wish to characterize solution(s) of the
+%% following optimization problem for $\tar$. %%: (\hilo^*)^m \to \R$:
+%% \begin{align}\label{eq:optimization}
+%%   \obs^{\star} := \argmax_{\obs} \tar( \obs ) 
+%%   = \argmax_{\obs} \frac12 \log \det 
+%%   (I + \prcov^{1/2} \fwd^*\obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}),
+%% \end{align}
+%% where $\obs$ is constrained to some allowed set of observations. We
+%% call this problem ``simultaneous optimization'', since all
+%% observations are decided on simulatneously.
+
+%% For computational reasons, one may prefer to find the best
+%% observations in a sequential manner. Denote
+%% \begin{equation}\label{eq:def obs_k}
+%%   \obs_k := (\meas_1,\dots, \meas_k)^t,  k\leq m.
+%% \end{equation}
+%% Sequential optimal design proceeds as follows. Find $\meas_1$ by
+%% maximizing $\tar(\obs_1)$. Then, keeping $\meas_1$ fixed --- find
+%% $\meas_2$ as the maximizer of $\tar(\obs_2)$. Then, find $\meas_3$ by
+%% keeping $\meas_1,\meas_2$ fixed and taking $\meas_3$ as the maximizer
+%% of $\tar(\obs_3)$. Continue this way until $\obs_m = \obs$ is
+%% found, where $m$ is the number of available observations. %% It is
+%% %% important to notice that this scheme does not require actually
+%% %% observing data --- in \eqref{eq:objective} data is averaged out.
+
+%% The analysis in this paper is conducted for the general simultaneous
+%% optimization case. The sequential optimization case is dealt with in
+%% section \ref{subsec:clusterization sequential}. It is important to
+%% note, however that all conclusions we arrive at for the simultaneous
+%% case easily specialize to the sequential case by considering the
+%% posterior as the next sequential step's prior.
 
@@
 We seek a formulation of the D-optimal design problem
-\eqref{eq:optimization} using Lagrange multipliers. In section
+\eqref{eq:optimization} using Lagrange multipliers. In Section
 \ref{section:objective} we find the gradient of $\tar$ and record it
-in \eqref{eq:tar grad}. In section \ref{subsec:unit length} we suggest
+in \eqref{eq:tar grad}. In Section \ref{subsec:unit norm} we suggest
 using unit-norm constraints on rows of $\obs$ so that the optimization
@@
 gradients for the new constraints and record them in \eqref{eq:grad
-  constraints}. Finally, in section \ref{subsec:necessary} we use
-these to formulate the D-optimal design problem as a Lagrange
-multipliers problem in equation \eqref{eq:conditions}.
+  constraints}. Finally, in Section \ref{subsec:necessary} we
+formulate the D-optimal design problem as a Lagrange multipliers
+problem in equation \eqref{eq:conditions}.
 
@@
 In order to use Lagrange multipliers, we need to find the gradient of
-$\tar$. This section is mostly technical and devoted to calculating
-said gradient. The result is recorded in \eqref{eq:tar grad}.
-
-We start by calculating the first variation. Some calculations are
-delegated to Lemma \ref{lemma:aux calc} in the appendix. We use
-the following lemma:
+$\tar$. Said gradient is recorded in \eqref{eq:tar grad} below. We
+start by recording Lemmas \ref{lemma:lax}, \ref{lemma:aux calc},
+proved in the appendix:
+
 \begin{restatable*}[Generalized from \cite{Lax97}]{lemma}{lax}\label{lemma:lax}
@@
 
-The proof of \ref{lemma:lax} is delegated to the appendix. Let
-$T(\obs) := \obs^*\Sigma(\obs)\obs$. The first variation of $\tar
-(\obs)$ in the direction $V$ is:
+\begin{restatable*}[Auxilliary Calculations]{lemma}{aux}\label{lemma:aux calc}
+  Let $T(\obs) := \obs^* \Sigma^{-1}(\obs)\obs$, with $\Sigma(\obs)$
+  defined as in \eqref{eq:Sigma}. Then:
+  \begin{align*}
+    \delta T(\obs)V &= V^* \Sigma^{-1} \obs 
+    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
+    &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs
+    + \obs^* \Sigma^{-1} V.
+  \end{align*}
+\end{restatable*}
+
+The first variation of $\tar (\obs)$ in the direction $V$ is:
 \begin{align}\label{eq:tar var}
@@
 the correct gradient and the notation \eqref{eq:tar grad} is
-justified. Note that by the comment following \eqref{eq:obs*}, we view
-$V$ (defined on the same space as $\obs$) as a column vector. Then the
-gradient $\nabla \tar(\obs)$ should be viewed as a row vector, as the
-product $V \nabla \tar$ is in $\R^{m \times m}$. This will prove
-important in section \ref{subsec:necessary}.
-
-\subsection{Unit Length Constraints and their Gradients}\label{subsec:unit length}
+justified. The gradient $\nabla \tar(\obs)$ should be viewed as a row
+vector, as the product $V \nabla \tar$ is in $\R^{m \times m}$. This
+will prove important in section \ref{subsec:necessary}.
+
+\subsection{Unit Norm Constraints and their Gradients}\label{subsec:unit norm}
 In this section we suggest relaxed constraints in
-\eqref{eq:constraints}. We then find their gradient in \eqref{eq:grad
-  constraints}. These are use in the Lagrange multipliers
-problem in section \ref{subsec:necessary}.
+\eqref{eq:constraints}, and their gradient in \eqref{eq:grad
+  constraints}.
 
@@
 (\hilo^*)^m$ and each $\meas_j,j=1,\dots,m$ must be chosen from some
-allowed set of functionals in $\hilo^*$. This set differs based on the
-kind of sensors we have at our disposal and the properties of $\hilo$,
-as discussed in section \ref{subsec:dynamics}. The following
-proposition gives better understanding of the constraints.
+allowed set of functionals in $\hilo^*$. %% This set differs based on
+%% the kind of sensors we have at our disposal and the properties of
+%% $\hilo$, as discussed in section \ref{subsec:dynamics}.
+Proposition \ref{prop:bigger better} below (proof in the appendix)
+gives better understanding of the constraints.
 \begin{restatable*}{proposition}{biggerbetter}\label{prop:bigger better}
@@
 \end{restatable*}
-Proof of Proposition \ref{prop:bigger better} is delegated to the
-appendix but the idea behind it is simple: making rows of $\obs$ large
-is equivalent to making $\sigma^2$ small. %%  Viewed another way, it
-%% increases the Signal to Noise Ratio (SNR) --- which is always
-%% desirable when taking any measurement prone to error.
-This can be understood easily from the formulation of the inverse
-problem $\data = \obs \fwd \param + \obs \eps' + \eps$, where $\eps'$
-can be taken as model error and $\eps$ is iid measurement error.
-
-We now suggest a set of constraints that will make the model
-analytically tractable. By Proposition \ref{prop:bigger better}, we
-cannot just take all $\hilo^*$ to be the allowed set of measurement
-vectors, since this would effectively eliminate measurement error
-entirely. We note that for point evaluations, the norm of the
-measurement is always one:
+The idea behind Proposition \ref{prop:bigger better} is simple: making
+the norm of rows of $\obs$ large is equivalent to making $\sigma^2$
+small. %% Viewed another way, it increases the Signal to Noise Ratio
+%% (SNR) --- which is always desirable when taking any observation
+%% prone to error.  This can be understood easily from the formulation
+%% of the inverse problem $\data = \obs \fwd \param + \obs \eps' +
+%% \eps$, where $\eps'$ can be taken as model error and $\eps$ is iid
+%% observation error.
+
+Note that for point evaluations, the norm of the observation
+functional is always one:
 \begin{align*}
@@
 \end{align*}
-Following this, we also restrict the norm of our measurement vectors
-to be one. The unit norm constraints can be written using
-\eqref{eq:obs*} as a series of $m$ equality constraints (one for each
-measurement) on $\obs$ as
+Following this, and in light of Proposition \ref{prop:bigger better},
+we consider observation vectors in the unit ball in $\hilo^*$. The
+unit norm constraints can be written using \eqref{eq:obs*} as a series
+of $m$ equality constraints (one for each observation) on $\obs$ as
 \begin{align}\label{eq:constraints}
@@
 \end{align*}
-Using the same arguments we used to justify the definition of
-\eqref{eq:tar grad}, we conclude that
+The same arguments justifying \eqref{eq:tar grad} hold here, and thus, by \eqref{eq:obs*}:
 \begin{align}\label{eq:grad constraints}
@@
 \end{align}
-with $\nabla \phi_j(\obs) \in \hilo^m$ and where the
-last equality follows since $\obs^*e_j = \meas_j$ by \eqref{eq:obs*}. As
-we noted at the end of section \ref{section:objective}, the
-gradient $\nabla \phi_j(\obs)$ is a row vector.
+where $\nabla \phi_j(\obs) \in \hilo^m$. As noted at the end of
+Section \ref{section:objective}, the gradient $\nabla \phi_j(\obs)$ is
+best thought of as a row vector.
 
 \subsection{Necessary Conditions for Optimal Design}\label{subsec:necessary}
-Necessary conditions for D-optimality are found using Lagrange
+Necessary conditions for D-optimality are found from Lagrange
 multipliers:
@@
 
-Recall where the objects involved are defined: the observation operator,
-$\obs \in (\hilo^*)^m$. The objective $\tar:
-(\hilo^*)^m \to \R$ and its gradient is $\nabla \tar(\obs) \in
-\hilo^m$. The constraints are defined similarly --- $\phi_j: (\hilo^*)^m
-\to \R$ and $\nabla \phi_j(\obs) \in \hilo^m$.
+Recall where the objects involved are defined: the observation
+operator, $\obs \in (\hilo^*)^m$. The objective $\tar: (\hilo^*)^m \to
+\R$ and its gradient is $\nabla \tar(\obs) \in \hilo^m$. Similarly,
+$\phi_j: (\hilo^*)^m \to \R$ and $\nabla \phi_j(\obs) \in \hilo^m$.
 
 Substituting \eqref{eq:tar grad} and \eqref{eq:grad constraints} in
-\eqref{eq:Lagrange mult1} we have:
+\eqref{eq:Lagrange mult1}:
 \begin{equation*}
@@
 \end{equation*}
-If we let $\Xi := \diag(\xi_j)$, then with some abuse of
-notation, we can write this more compactly as:
+Let $\Xi := \diag(\xi_j)$. Then, with some abuse of notation, this can
+be written more compactly as:
 \begin{equation}\label{eq:conditions}
@@
 present. Sensor clusterization is discussed in section
-\ref{section:clusterization}. We prepare the ground by first deriving
-a nonlinear eigenvalue problem in $\obs$ for D-optimal designs in
-section \ref{subsec:eigenproblem}. This nonlienar eigenvalue problem
-is recorded in \eqref{eq:mod conditions}. Next, we characterize
-D-optimal designs as ones that make all eigenvalues equal in the above
-mentioned nonlinear eigenvalue problem. The derivation is presented in
-section \ref{subsec:characterization} and is summarized in Theorem
+\ref{section:clusterization}. We first derive a nonlinear eigenvalue
+problem in $\obs$ for D-optimal designs in Section
+\ref{subsec:eigenproblem}. In Section \ref{subsec:characterization} we
+characterize D-optimal designs as ones that make all eigenvalues equal
+in the nonlinear eigenvalue problem. Results are summarized in Theorem
 \ref{thm:char}.
@@
 \end{equation}
-with $\Xi$ diagonal. This looks like an eigenvalue problem for the
-self-adjoint operator $\sigma^{-2}\fwd \postcov \fwd^*$. The rows of
-$\obs$, namely $\meas_j,j=1,\dots, m$, are the eigenvectors. Note,
-however, that $\postcov$ depends on $\obs$, and so
-\eqref{eq:eigenproblem} is a nonlinear eigenvalue problem.
-
-Below, the expression for the posterior covariance \eqref{eq:postcov}
-justifies the second equality. Some calculations delegated to Lemma
-\ref{lemma:twice woodbury} in the appendix justify the third (it is
-trivial if $\fwd$ is invertible):
-\begin{align}\label{eq:mod conditions}
+with $\Xi$ diagonal. We think of \eqref{eq:eigenproblem} as an
+eigenvalue problem for the self-adjoint operator $\sigma^{-2}\fwd
+\postcov \fwd^*$, where rows of $\obs$, namely $\meas_j,j=1,\dots, m$,
+are eigenvectors. However, that $\postcov$ depends on $\obs$ and thus,
+we refer to \eqref{eq:eigenproblem} as a nonlinear eigenvalue problem.
+
+The following two lemmas allow us to gain better understanding of the
+nonlinear eigenvalue problem.
+
+\begin{restatable*}[Two applications of Woodbury Matrix Identity]{lemma}{woodbury}\label{lemma:twice woodbury}
+  Assume $\fwd \prcov \fwd^*$ is invertible. Then
+  \begin{align*}
     \begin{split}
-  \obs^* \Xi &= \sigma^{-2}\fwd \postcov \fwd^* \obs^*  \\
-  %
-  %
-  %
-  &= \sigma^{-2} \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* \obs^* \\
-  %
-  %
-  %
-  &= \sigma^{-2} \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1} \obs^*.
+      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* 
+      %
+      %
+      = \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1},
     \end{split}
-\end{align}  
-The term $(\fwd \prcov \fwd^*)^{-1}$ is the prior precision in
-$\hilo$, and $\fwd\prcov\fwd^*$ is invertible, as mentioned in
-section \ref{section:prelim}. $(\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}
-\obs^* \obs$ is the posterior precision in $\hilo$.
-
-\subsection{Characterizing D-Optimal Designs}\label{subsec:characterization}
-Denote eigenvalues of $\fwd \prcov \fwd^*$ by
-$\{\lambda_i\}_{i=1}^{\infty}$. These are all positive. Denote
-corresponding eigenvectors $\{\ev_i\}_{i=1}^{\infty}$. Lemma
-\ref{lemma:sim diag} (proved in the appendix) is crucial for the rest
-of the analysis:
+  \end{align*}  
+\end{restatable*}
+
+%% The term $(\fwd \prcov \fwd^*)^{-1}$ is the prior precision in
+%% $\hilo$, and $\fwd\prcov\fwd^*$ is invertible, as mentioned in
+%% Section \ref{section:prelim}. $(\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}
+%% \obs^* \obs$ is the posterior precision in $\hilo$.
+
+%% \subsection{Characterizing D-Optimal Designs}\label{subsec:characterization}
+%% %% Denote corresponding eigenvectors $\{\ev_i\}_{i=1}^{\infty}$.
+%% Denote eigenvalues of $\fwd \prcov \fwd^*$ by
+%% $\{\lambda_i\}_{i=1}^{\infty}$, and note they are positive. Lemma
+%% \ref{lemma:sim diag} is crucial for the rest of the analysis:
 
@@
   \begin{equation*}
-   (C + \sum_{j=1}^m \func_j\func_j^*) \func_l = \xi_l \func_l, l = 1,\dots,m
+   (C + \sum_{j=1}^m \func_j\func_j^*) \func_l = \xi_l \func_l,\ l = 1,\dots,m
   \end{equation*}
@@
 \begin{corollary}[Eigenvectors of a D-optimal design]\label{cor:same ev}
-  $\obs^*\obs$ has the same eigenvectors as $\fwd \prcov \fwd^*$.
-  %namely $\{\ev_i\}_{i=1}^{\infty}$.
+  Let $\obs$ satisfy the nonlinear eigenvalue problem
+  \eqref{eq:eigenproblem}. Then $\obs^*\obs$ has the same eigenvectors
+  as $\fwd \prcov \fwd^*$.
 \end{corollary}
 \begin{proof}
-In \eqref{eq:mod conditions}, take $\func_j^{*} = \meas_j$ and $C :=
-(\fwd \prcov \fwd^*)^{-1}$ and use Lemma \ref{lemma:sim diag}.
-\end{proof}
-At most $m$ of the eigenvectors of $\obs^*\obs$ have a positive
-eigenvalue, because $k := \rank \obs^*\obs \leq m$. Denote the
-corresponding non-zero eigenvalues of $\obs^*\obs$ by
-$\{\eta_i\}_{i=1}^{k}$ and let $\eta_i = 0$ for $i \geq k+1$. No
-assumption is made (yet) regarding the magnitudes of the corresponding
-$\{\lambda_i\}_{i=1}^{\infty}$. They are to be thought of as fixed but
-unknown. Using Sylvester's Determinant Theorem (the standard proof
-works for Hilbert spaces as well) we get:
-
-\begin{align}\label{eq:pre manipulations}
+  \begin{align}\label{eq:mod conditions}
     \begin{split}
-    \det \left (I + \sigma^{-2} \prcov^{1/2} \fwd ^* \obs^*
-    \obs \fwd \prcov^{1/2}\right ) &=
-    %
-    \det \left (I + \sigma^{-2} \obs^*
-    \obs \fwd \prcov\fwd^* \right ) \\
-    %
-    %
-    %
-    &= \det \left \{ \left ( (\fwd \prcov \fwd^*)^{-1} +
-    \sigma^{-2}\obs^*\obs \right ) (\fwd \prcov \fwd^*)\right \}.\\
+      \obs^* \Xi &= \sigma^{-2}\fwd \postcov \fwd^* \obs^*  \text{ (by \eqref{eq:eigenproblem})}\\
+      %
+      %
+      %
+      &= \sigma^{-2} \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* \obs^*  \text{ (by \eqref{eq:postcov})} \\
+      %
+      %
+      %
+      &= \sigma^{-2} \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1} \obs^* \text{ (by Lemma \ref{lemma:twice woodbury})}.
     \end{split}
@@
 
-By definition of the objective \eqref{eq:objective}, \eqref{eq:pre
-  manipulations} and Corollary \ref{cor:same ev}:
-\begin{align}\label{eq:mod objective}
+  Now take $\func_j^{*} = \meas_j$ and $C := (\fwd \prcov
+  \fwd^*)^{-1}$ and use Lemma \ref{lemma:sim diag}.
+\end{proof}
+
+%% Corollary \ref{cor:same ev} is important because it allows us to only think
+
+implies we only need to consider $\obs$ such that  
+At most $m$ of the eigenvectors of $\obs^*\obs$ have a positive
+eigenvalue, because $k := \rank \obs^*\obs \leq m$. Since we made no
+assumption regarding the ordering of $\{\lambda_i\}$, we can denote
+the corresponding non-zero eigenvalues of $\obs^*\obs$ by
+$\{\eta_i\}_{i=1}^{k}$ and let $\eta_i = 0$ for $i \geq k+1$.
+
+\begin{corollary}\label{cor:true target}
+  Let $\obs$ with $m$ observations satisfy the nonlinear eigenvalue
+  problem \eqref{eq:eigenproblem}. Let $\{\eta_i\}_{i=1}^{\infty}$
+  eigenvalues of $\obs^*\obs$ and $\{\lambda_i\}_{i=1}^{\infty}$ the
+  corresponding eigenvalues of $\fwd \prcov \fwd^*$. Then $k:=\rank
+  \obs^*\obs \leq m$ and $\obs^*\obs$ has exactly $k$ positive
+  eigenvalues.
+
+  Without loss of generality, let $\eta_i > 0$ for $i\leq k$ and
+  $\eta_i = 0$ for $i > k$. Then:
+  \begin{equation*}
+    \tar(\obs) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i).
+  \end{equation*}
+  where $C$ dependes only on $\sigma^2$ and on
+  $\{\lambda_i\}_{i=1}^{\infty}$.
+
+\end{corollary}
+\begin{proof}
+  The first part is trivial. Since $k \leq m$, then $\obs^*\obs$ has
+  at most $k$ nonzero eigenvalues $\eta_1, \dots, \eta_k$. Then:
+
+  \begin{align}
     \begin{split}
-    \tar ( \obs ) &=\frac12 \log \det \left( I + \sigma^{-2}
-    \prcov^{1/2} \fwd ^* \obs^* \obs \fwd \prcov^{1/2} \right )\\
-    %
-    %
-    %
-    %
-    &=\frac12 \log \det \Big ( \left ( (\fwd \prcov \fwd^* )^{-1} +
-    \sigma^{-2}\obs^* \obs \right ) \fwd \prcov \fwd^* \Big )\\
-    %
-    %
-    %
-    &=\frac12 \log \left (\prod_{i=1}^{\infty} ( \lambda_i^{-1} + \sigma^{-2} \eta_i ) \prod_{i=1}^{\infty} \lambda_i \right ) \\
-    %
-    %
-    %
-    &=\frac12 \log \left ( \prod_{i=1}^{k} ( \lambda_i^{-1} + \sigma^{-2} \eta_i )\prod_{i=1}^{k} \lambda_i \right )\\
-    %
-    %
-    %
-    &=\frac12 \sum_{i=1}^{k} \log (\sigma^2\lambda_i^{-1} + \eta_i), 
+      \tar(\obs) &= \frac12\log \det \left (I + \sigma^{-2} \prcov^{1/2} \fwd ^* \obs^*
+      \obs \fwd \prcov^{1/2}\right )\\% \text{ (by definition)}\\
+      %
+      &= \frac12 \log \det \left (I + \sigma^{-2} \obs^* \obs \fwd
+      \prcov\fwd^* \right ) \text{ (Sylvester's Determinant
+      Theorem)}\\
+      %
+      %
+      %
+      &=\frac12 \log \prod_{i=1}^{\infty} ( 1 + \sigma^{-2} \lambda_i\eta_i ) \text{ (Corollary \ref{cor:same ev})} \\
+      %
+      %
+      %
+      %&=\frac12 \log \left ( \prod_{i=1}^{k} ( \lambda_i^{-1} + \sigma^{-2} \eta_i )\prod_{i=1}^{k} \lambda_i \right )
+      %
+      %
+      %
+      &=\frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i).
+      %
+      %
+      %% &=\frac12 \sum_{i=1}^{k} \log(\sigma^2\lambda_i^{-1} + \eta_i) - k\log \sigma + \frac12\sum_{i=1}^k \log \lambda_i.
     \end{split}
   \end{align}
-and the last equality holds up to an additive constant independent of
-$\obs$. Therefore, a D-optimal design maximizes
-
-\begin{equation}\label{eq:true target}
-  \sum_{i=1}^{k}\log (\sigma^2 \lambda_i^{-1} + \eta_i),
-\end{equation}
-
-with the norm constraints on $\meas_j,j=1,\dots,m$. We would like to
-make all $\eta_i$ as large as possible.  The problem of D-optimal
-design has now been cast to finite dimensions. This justifies using
-linear algebra. Lemma \ref{lemma:free} (a constructive proof is given
-in the appendix) does exactly that.
+\end{proof}
+
+In light of Corollary \ref{cor:true target} we define:
+\begin{definition}
+  The \emph{utility proxy} $\bar{\tar}$ is defined for $\eta \in
+  \mathbb{R}_+^k$, $\lambda \in \mathbb{R}_+^k$ and $\sigma^2 >0$ as:
+  \begin{equation*}
+    \bar{\tar}(\eta, \lambda, \sigma^2) := \sum_{i=1}^k \log(1 + \sigma^{-2}\lambda_i\eta_i).
+  \end{equation*}
+\end{definition}
+
+\begin{observation}
+  The utility proxy $\bar{\tar}$ is increasing in $\lambda$. Thus, in
+  a D-optimal design $\lambda$ are the $k$ largest eigenvalues of
+  $\fwd \prcov \fwd^*$.
+\end{observation}
+
+\begin{proposition}
+  Fix $\lambda, \sigma^2$ and let
+  $$
+  \eta_i = \frac{1}{k} \left ( m - k \sigma^2 \lambda_i^{-1} + \sigma^2 \sum_{j=1}^k \lambda_j^{-1} \right ).
+  $$
+
+  If $\eta_i > 0,\ i=1,\dots, k$, then $\eta_1,\dots,\eta_k$ maximize
+  the utility proxy $\bar{\tar}(\eta; \lambda, \sigma^2)$.
+\end{proposition}
+\begin{proof}
+  Let $\Phi(\eta) = \sum \eta_i - m$. Using Lagrange multipliers:
+  \begin{align*}
+    \begin{split}
+      \frac{\der}{\der \eta_i} \bar{\tar}(\eta, \lambda, \sigma^2) &=
+      \frac{1}{\sigma^{2}\lambda_i^{-1} + \eta_i} \\
+      %
+      %
+      %
+      \frac{\der\Phi}{\der \eta_i} &= 1.
+    \end{split}
+  \end{align*}
+
+  Denote $c := \sigma^{2}\lambda_i^{-1} + \eta_i$. Summing,
+  substituting $\sum \eta_i = m$ and dividng by $k$:
+  $$
+  c = \frac{m}{k} + \sigma^2 \frac{1}{k} \sum_j \lambda_j^{-1}.
+  $$
+
+  Consequently:
+  \begin{align*}
+    \begin{split}
+      \eta_i &= c - \sigma^2 \lambda_i^{-1} \\
+      %
+      %
+      %
+      &= \frac{m}{k} + \sigma^2 \frac{1}{k} \sum_j \lambda_j^{-1} - \sigma^2 \lambda_i^{-1} \\
+      %
+      %
+      %
+      &= \frac{1}{k} \left ( m - k \sigma^2\lambda_i^{-1} + \sigma^2 \sum_{j=1}^k \lambda_j^{-1} \right ).
+    \end{split}
+  \end{align*}
+\end{proof}
+
+%% We now turn our attention to maximizing $\bar{\tar}$ in $\eta$. The only
+%% constraint on $\eta$ arises from the unit norm contraint on rows of
+%% $\obs$, which implies that $\tr{\obs^*\obs} = \sum_i \eta_i= m$.
+%% \begin{proposition}
+%%   Fix $\lambda \in \mathbb{R}_+^k, \sigma^2 > 0$. Then
+%%   $$
+%%   \eta = \arg\max_{\sum_i \eta_i = m} \Phi(\eta; \lambda, \sigma^2)
+%%   $$ 
+%% \end{proposition}
+
+%% \begin{proof}
+%%   Lagrange multipliers:
+%%   \begin{align*}
+%%     \begin{split}
+%%       \frac{d}{d\eta_i} \Phi(\eta; \lambda, \sigma^2) &= \frac{1}{\sigma^2\lambda_i^{-1} + \eta_i}, \\
+%%       %
+%%       %
+%%       %
+%%       \frac{d}{d\eta_i} (\sum_i \eta_i - m)  &= 1.
+%%     \end{split}
+%%   \end{align*}
+
+%%   Denote the Lagrange multiplier $c$. Then
+%%   \begin{align*}
+%%     \begin{split}
+%%       \sigma^2\lambda_i^{-1} + \eta_i &= \frac{1}{c}, \\
+%%       %
+%%       %
+%%       %
+%%       \sum_i \eta_i &= m.
+%%     \end{split}
+%%   \end{align*}
+
+%%   and so
+
+%%   \begin{align*}
+%%     \begin{split}
+%%       \sigma^2\sum_{i=1}^k\lambda_i^{-1} + m &= kc^{-1} \\
+%%       %
+%%     \end{split}
+%%   \end{align*}
+
+%% \end{proof}
+
+%% Corollary \ref{cor:same ev} implies that instead of considering
+%% $\obs$ when searching for a D-optimal design, we can restrict
+%% ourselves to $k \leq m$ eigenvalues of $\obs^*\obs$ corresponding
+%% to the first $k$ eigenvalues of $\fwd \prcov \fwd^*$.
+
+The nonlinear eigenvalue problem \eqref{eq:eigenproblem} gives rise to
+necessary conditions for D-optimality. Thus, in order to find the
+global maximizer of $\tar$, we need to choose $k$, then choose $k$
+eigenvalues $\lambda_1, \dots, \lambda_k$ and corresponding $\eta_1,
+\dots, \eta_k$ to maximize $\sum_i \log( \sigma^2\lambda_i^{-1} +
+\eta_i)$. It is not clear, however, if we can in fact choose
+observation vectors to manifest any $\eta_1, \dots, \eta_k$
+combination. The following lemma answers this in the
+affirmitive. Moreover, the proof is constructive.
+
 \begin{restatable*}[Unit norm decomposition]{lemma}{free}\label{lemma:free}
@@
 \end{restatable*}
+
+with the norm constraints on $\meas_j,j=1,\dots,m$. We would like to
+make all $\eta_i$ as large as possible.  The problem of D-optimal
+design has now been cast to finite dimensions. This justifies using
+linear algebra. Lemma \ref{lemma:free} (a constructive proof is given
+in the appendix) does exactly that.
 
@@
 increasing $\sigma^2\lambda_i^{-1} + \eta_i$ where it is smallest. So,
-no weight should be given to (i.e. measurement taken in direction of)
+no weight should be given to (i.e. observation taken in direction of)
 the $n$th eigenvector before $\sigma^2\lambda_i^{-1} + \eta_i \geq
@@
   \caption{Characterizing a D-optimal design, compared to a
-    sub-optimal design. Both designs are allowed three measurements,
+    sub-optimal design. Both designs are allowed three observations,
     which means the blue area has accumulated height of $3$ in both
@@
     uncertainty) where it is lowest. A sub-optimal design (right) does
-    not. In this example, a measurement is taken partly in the
+    not. In this example, a observation is taken partly in the
     direction of the third eigenvector. The precision for this
@@
 volume of the green liquid is increasing in $i$. A D-optimal design
-with $m$ measurements ignores cylinders indexed $m+1$ and above. It
+with $m$ observations ignores cylinders indexed $m+1$ and above. It
 has blue liquid of $m$ volume units at its disposal. It distributes
@@
   \item $\fwd, \prcov$ the forward and prior covariance operators, respectively.
-  \item $m, \sigma^2$ the number of measurements taken and observation
+  \item $m, \sigma^2$ the number of observations taken and observation
     noise variance, respectively.
-  \item $\obs$ a D-optimal design operator.
-  \item $\tar$ as definded in \eqref{eq:objective}.
+  \item A D-optimal design operator $\obs$:
+    $$
+    \obs = \arg\max_{\|\meas_j\| = 1, j=1,\dots,m}\tar(\obs)
+    $$ 
   \item $\{\lambda_i\}_{i=1}^\infty$ eigenvalues of $\fwd\prcov\fwd^*$
-    in decreasing order.
-  \item $\{\ev_i\}_{i=1}^\infty$ their corresponding eigenvectors.
-  \item $\{\eta_i\}_{i=1}^\infty$ eigenvalues of $\obs^*\obs$, where
-    $\obs$ maximizes $\tar(\obs)$.
+    in decreasing order of magnitude.
+  %% \item $\{\ev_i\}_{i=1}^\infty$ their corresponding eigenvectors.
+  \item $\{\eta_i\}_{i=1}^\infty$ eigenvalues of $\obs^*\obs$.
   \end{enumerate}
@@
   \item There exists $0 < k \leq m$ such that $\eta_i = 0$ for $i\geq
-    k+1$ and $k = \rank \obs^*\obs$.
+    k+1$ and $k\leq\rank \obs^*\obs$.
   \item $\sigma^2\lambda_i^{-1} + \eta_i$ are equal for all $1 \leq i \leq k$.
@@
 design scenario is understood by specializing thorem \ref{thm:char} to
-one measurement each time. In this case $m=1$, so $k=1$, $\meas_1
+one observation each time. In this case $m=1$, so $k=1$, $\meas_1
 \parallel \ev_1$ and $\eta_1 = 1$. As hinted in section
@@
 step. The eigenvectors of the new posterior are the same as for the
-previous one. A new measurement will be taken in the direction of
+previous one. A new observation will be taken in the direction of
 eigenvector with smalles eigenvalue. This leads to:
@@
 however, are $\lambda^{-1}_1 + \sigma^{-2}, \lambda^{-1}_2 +
-\sigma^{-2}, \lambda_3^{-1}, \dots$. A new measurement is taken in the
+\sigma^{-2}, \lambda_3^{-1}, \dots$. A new observation is taken in the
 direction of eigenvector of smallest precision (note that
@@
 operator. If we do not encounter clusterization after $m$ such steps,
-this means that measurement $\meas_k$ was taken in the direction
+this means that observation $\meas_k$ was taken in the direction
 $\ev_k$ for $k=1,\dots,m$. Thus:
@@
 $\obs^*\obs$. Theorem \ref{thm:char} tells us that for D-optimality,
-we only need to make the posterior flat where we take measurements
+we only need to make the posterior flat where we take observations
 (recall figure \ref{fig:optimal vs not}). We are left with the freedom
-of assigning any part of $\eta_i$ to any measurement $\meas_j$. Two
+of assigning any part of $\eta_i$ to any observation $\meas_j$. Two
 possible assignments for the same problem are illustrated in figure
@@
     simultaneous D-optimal designs. Posterior precision per
-    eigenvector after taking measurements $\meas_i$ is plotted. Both
+    eigenvector after taking observations $\meas_i$ is plotted. Both
     designs are identical for all practical matters --- their
     posteriors are identical. In the left panel we see how an optimal
-    design is achieved with repeated measurements. $\meas_2 = \meas_3$
+    design is achieved with repeated observations. $\meas_2 = \meas_3$
     and $\meas_5 = \meas_6$. This is not necessary though, as can be
@@
 Hence no increase in the design criterion is achieved by taking a
-repeated measurement (in the limit $\sigma^2 \to 0$). This is clearly
+repeated observation (in the limit $\sigma^2 \to 0$). This is clearly
 sub-optimal and $\obs$ cannot give rise to a D-optimal design. Thus,
-for small measurement error levels, the clusterization effect will be
+for small observation error levels, the clusterization effect will be
 mitigated by the presence of a non-zero model error. Since the design
 criterion is not defined for $\sigma^2 = 0$ and identical
-measurements, we cannot make a statement regarding $\sigma^2 = 0$,
+observations, we cannot make a statement regarding $\sigma^2 = 0$,
 except in the limiting sense described above.
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{lemma}\label{lemma:twice woodbury}
-  Assume $\fwd \prcov \fwd^*$ is invertible. Then
-\begin{align*}
-  \begin{split}
-    \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* 
-    %
-    %
-    = \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1},
-  \end{split}
-\end{align*}  
-\end{lemma}
+\woodbury
 \begin{proof}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{lemma}[Increase due to a Measurement]\label{lemma:design increase}
+\begin{lemma}[Increase due to a Observation]\label{lemma:design increase}
   Let $\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm := (\meas_1,\dots,\meas_{m-1})^t$. Then
@@
 
-\begin{lemma}[Auxilliary Calculations]\label{lemma:aux calc}
-  Let $T(\obs) := \obs^* \Sigma^{-1}(\obs)\obs$, with $\Sigma(\obs)$
-  defined as in \eqref{eq:Sigma}. Then:
-  \begin{align*}
-    \delta T(\obs)V &= V^* \Sigma^{-1} \obs 
-    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
-    &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs
-    + \obs^* \Sigma^{-1} V.
-  \end{align*}
-\end{lemma}
-
+\aux
 \begin{proof}
@@
   it is positive --- $\delta \tar(\obs)V > 0$. This means that
-  increasing the magnitude of the $j$th measurement functional
+  increasing the magnitude of the $j$th observation functional
   increases $\tar(\obs)$. We start with the last line of \eqref{eq:tar var}, and denote $\tmp : = \fwd \postcov \fwd^*$.

### 2023-01-03 latex/paper.tex
With corollary on pushforward/pullback and without a proper statement on k

@@
 \documentclass{amsart}
-
 \input{definitions}
@@
   err}.
+
+\begin{restatable}[D-optimal designs with vanishing model error]{theorem}{main}\label{thm:char}
+  Let:
+  \begin{enumerate}
+  \item $\fwd:\hilp \to \hilo$ a linear forward operator,
+    $\prcov:\hilp \to \hilp$ prior covariance operator, $\obs: \hilo
+    \to \mathbb{R}^m$ observation operator, where $m \in \mathbb{N}$
+    is the number of observations taken.
+    \item $\sigma^2 \in \mathbb{R}_{+}$ observation noise variance,
+      $\data = \obs \fwd \param + \eps$, where $\eps \in \mathbb{R}^m$
+      is iid $\mathcal{N}(0, \sigma^2)$ noise, $\pr \sim
+      \mathcal{N}(0, \prcov)$ prior measure, $\post$ the posterior
+      measure.
+  \item A D-optimality utility function
+    \cite{AlexanderianGloorGhattas14}:
+    \begin{align*}
+      \begin{split}
+        \tar(\obs) :&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
+        % 
+        % 
+        % 
+        &= \frac12 \log \det ( I + \sigma^{-2} \prcov^{1/2} \fwd ^*
+        \obs^* \obs \fwd \prcov^{1/2}).
+    \end{split}
+  \end{align*}
+  \item A D-optimal design operator $\obs$:
+    $$
+    \obs = \argmax_{\|\meas_j\| = 1, j=1,\dots,m}\tar(\obs)
+    $$ 
+  \item $\{\lambda_i\}_{i=1}^\infty$ eigenvalues of $\fwd\prcov\fwd^*$
+    in decreasing order of magnitude.
+  %% \item $\{\ev_i\}_{i=1}^\infty$ their corresponding eigenvectors.
+  \item $\{\eta_i\}_{i=1}^\infty$ eigenvalues of $\obs^*\obs$.
+  \end{enumerate}
+
+  Then:
+  \begin{enumerate}
+  \item $\obs^*\obs$ and $\fwd\prcov\fwd^*$ are simultaneously
+    diagonalizable.
+  \item $k := \rank \obs^*\obs \leq m$.
+  \item     
+    \begin{equation*}
+      \tar(\obs) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i). %= \frac12 \sum_{i=1}^{m} \log (1 + \sigma^{-2}\lambda_i\eta_i).
+    \end{equation*}
+  \item 
+    \begin{equation*}
+      k = \argmax \left \{ k:\lambda_k^{-1} < \sigma^{-2}\frac{m}{k} + \frac{1}{k} \sum_{j=1}^{k}
+      \lambda_j^{-1} \right \}.
+    \end{equation*}
+  \item
+    \begin{equation*}
+        \eta_i = \begin{cases}
+          \frac{m}{k} - \sigma^2 \lambda_i^{-1} + \sigma^2 \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1} & 1 \leq i \leq k \\
+          0 & i > k 
+        \end{cases}.
+    \end{equation*}
+  \end{enumerate}
+\end{restatable}
 
@@
 
+%% \begin{definition}
+%%   Let $\{\lambda_i\}_{i=1}^{\infty}$ eigenvalues of $\fwd \prcov
+%%   \fwd^*$ sorted in descending order. Let $\eta \in
+%%   \mathbb{R}_+^m$. The \emph{utility proxy} $\bar{\tar}$ is defined
+%%   as:
+%%   \begin{equation*}
+%%     \bar{\tar}(\eta) := \frac12 \sum_{i=1}^m \log(1 + \sigma^{-2}\lambda_i\eta_i).
+%%   \end{equation*}
+%% \end{definition}
+
 \begin{corollary}\label{cor:true target}
@@
   eigenvalues of $\obs^*\obs$ and $\{\lambda_i\}_{i=1}^{\infty}$ the
-  corresponding eigenvalues of $\fwd \prcov \fwd^*$. Then $k:=\rank
-  \obs^*\obs \leq m$ and $\obs^*\obs$ has exactly $k$ positive
+  corresponding eigenvalues of $\fwd \prcov \fwd^*$. Let $k:=\rank
+  \obs^*\obs$. Without loss of generality, let $\eta_i > 0$ for $i\leq
+  k$ and $\eta_i = 0$ for $i > k$. Then:
+  \begin{enumerate}
+    \item $k \leq m$ and $\obs^*\obs$ has exactly $k$ positive
       eigenvalues.
-
-  Without loss of generality, let $\eta_i > 0$ for $i\leq k$ and
-  $\eta_i = 0$ for $i > k$. Then:
+    \item
       \begin{equation*}
-    \tar(\obs) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i).
+        \tar(\obs) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i) = \frac12 \sum_{i=1}^{m} \log (1 + \sigma^{-2}\lambda_i\eta_i).
       \end{equation*}
-  where $C$ dependes only on $\sigma^2$ and on
-  $\{\lambda_i\}_{i=1}^{\infty}$.
-
+    \item Let $\obs$ a D-optimal design. Then $\eta_i > 0$ for
+      eigenvectors corresponding to the $k$ largest $\lambda_i$.
+  \end{enumerate}
 \end{corollary}
 \begin{proof}
-  The first part is trivial. Since $k \leq m$, then $\obs^*\obs$ has
-  at most $k$ nonzero eigenvalues $\eta_1, \dots, \eta_k$. Then:
-
+  Part (1) is trivial. To see part (2) holds: 
   \begin{align}
@@
   \end{align}
+  Part (3) holds since $\log$ is increasing and $\eta_i \geq 0$.
 \end{proof}
 
-In light of Corollary \ref{cor:true target} we define:
-\begin{definition}
-  The \emph{utility proxy} $\bar{\tar}$ is defined for $\eta \in
-  \mathbb{R}_+^k$, $\lambda \in \mathbb{R}_+^k$ and $\sigma^2 >0$ as:
-  \begin{equation*}
-    \bar{\tar}(\eta, \lambda, \sigma^2) := \sum_{i=1}^k \log(1 + \sigma^{-2}\lambda_i\eta_i).
-  \end{equation*}
-\end{definition}
-
-\begin{observation}
-  The utility proxy $\bar{\tar}$ is increasing in $\lambda$. Thus, in
-  a D-optimal design $\lambda$ are the $k$ largest eigenvalues of
-  $\fwd \prcov \fwd^*$.
-\end{observation}
-
-\begin{proposition}
-  Fix $\lambda, \sigma^2$ and let
-  $$
-  \eta_i = \frac{1}{k} \left ( m - k \sigma^2 \lambda_i^{-1} + \sigma^2 \sum_{j=1}^k \lambda_j^{-1} \right ).
-  $$
-
-  If $\eta_i > 0,\ i=1,\dots, k$, then $\eta_1,\dots,\eta_k$ maximize
-  the utility proxy $\bar{\tar}(\eta; \lambda, \sigma^2)$.
+\begin{proposition}\label{prop:optimizer}
+  Let $\{\lambda_i\}_{i=1}^{\infty}$ the sorted eigenvectors of $\fwd
+  \prcov\fwd^*$, $k\in \mathbb{N}^+$ and
+  %% Let
+  %%   \begin{equation}
+  %%   k = \argmax \left \{ k:\lambda_k^{-1} \leq \sigma^{-2}\frac{m}{k} + \frac{1}{k} \sum_{j=1}^{k}
+  %%   \lambda_j^{-1} \right \}.
+  %% \end{equation}
+
+  \begin{equation}
+  \eta_i = \begin{cases}
+    \frac{m}{k} - \sigma^2 \lambda_i^{-1} + \sigma^2 \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1} & i \leq k \\
+    0 & i > k 
+  \end{cases}.
+  \end{equation}
+
+  Let $\obs$ an observation operator such that $\obs^*\obs$ has the
+  same eigenvectors as $\fwd \prcov \fwd^*$. If $\eta_i > 0$ for
+  $i\leq k$ and $\obs^*\obs$ has eigenvalues $\eta_i$, then $\obs$
+  maximizes $\tar$ subject to $\sum_{i=1}^m \eta_i = m$ and $\rank
+  \obs = k$.
 \end{proposition}
 \begin{proof}
-  Let $\Phi(\eta) = \sum \eta_i - m$. Using Lagrange multipliers:
+  Below, we use a slight abuse of notation and cast the entire
+  optimization problem of $\tar$ to the first $m$ eigenvectors of
+  $\fwd\ \prcov \fwd^*$. We treat $\eta$ as if it is $m-$dimensional,
+  since by \ref{}, $\eta_i = 0, i > m$. This abuse of notation gives
+  us the freedom to think of the problem at hand in $m-$dimensions. We
+  will now find the optimal $\eta_i$'s.
+
+  First, let $A := \{i: \eta_i > 0\}$. A choice of $A$ where $\eta_i
+  =0, \eta_j >0$ for some $1 \leq i < j \leq m$ is not optimal (just
+  switch $\eta_j$ and $\eta_i$ and recall $\{\lambda_i\}$ are arranged
+  in descending order). Thus, since $\rank \obs = k$ we conclude that,
+  $\eta_i=0, i>k$.
+
+  Let $\Phi(\eta) = \sum_{i=1}^k \eta_i - m$ and $\Omega_j(\eta) =
+   -\eta_j$. Then
   \begin{align*}
     \begin{split}
-      \frac{\der}{\der \eta_i} \bar{\tar}(\eta, \lambda, \sigma^2) &=
-      \frac{1}{\sigma^{2}\lambda_i^{-1} + \eta_i} \\
-      %
-      %
-      %
-      \frac{\der\Phi}{\der \eta_i} &= 1.
+      \frac{\partial \tar}{\partial \eta_i}  &=
+       \frac12 \frac{\sigma^{-2}\lambda_i}{1 + \sigma^{-2} \lambda_i\eta_i} = \frac12 \frac{1}{\sigma^{2}\lambda_i^{-1} + \eta_i} \\
+      %
+      %
+      %
+      \frac{\partial\Phi}{\partial \eta_i} &= 1 \\
+      %
+      %
+      %
+      \frac{\partial \Omega_j}{\partial \eta_i} &= -\delta_{ij}      
     \end{split}
@@
 
-  Denote $c := \sigma^{2}\lambda_i^{-1} + \eta_i$. Summing,
-  substituting $\sum \eta_i = m$ and dividng by $k$:
+  From the KKT conditions, there are $\alpha, \beta_i$ such that for $i=1,\dots,m$:
+  \begin{align}
+    \begin{split}
+      -\frac12 \frac{1}{\sigma^{2}\lambda_i^{-1} + \eta_i} + \alpha - \beta_i  &= 0 \\
+      %
+      %
+      %
+      \eta_i &\geq 0\\
+      %
+      %
+      %
+      \beta_i &\geq 0\\
+      %
+      %
+      %
+      \beta_i \eta_i &= 0\\
+      %
+      %
+      %
+      \sum_{i=1}^m \eta_i &= m 
+    \end{split}
+  \end{align}
+
+  Then, for $1 \leq i \leq k$:
+  \begin{align*}
+    \begin{split}
+      \beta_i &= 0\\
+      %
+      %
+      %
+      \sigma^{2}\lambda_i^{-1} + \eta_i  &= \frac{1}{2\alpha}%  \text{ for }  1 \leq i \leq k.\\
+      %
+      %
+      %
+      %% \sigma^{-2} \lambda_i &= 2\alpha - 2\beta_i, i \in B \\
+    \end{split}
+  \end{align*}
+
+  Summing, substituting $\sum_{i=1}^m \eta_i = \sum_{i=1}^k \eta_i =
+  m$, and dividng by $k$:
   $$
-  c = \frac{m}{k} + \sigma^2 \frac{1}{k} \sum_j \lambda_j^{-1}.
+  \frac{1}{2\alpha} = \frac{m}{k} + \frac{\sigma^2}{k} \sum_{j=1}^k \lambda_j^{-1}.
+  %\Rightarrow \alpha = \left ( \frac{2m}{k} + \frac{2\sigma^2}{k}
+  %\sum_{i\in A} \lambda_i^{-1} \right )^{-1}.
   $$
@@
   Consequently:
-  \begin{align*}
+  \begin{align} \label{eq:etas}
     \begin{split}
-      \eta_i &= c - \sigma^2 \lambda_i^{-1} \\
-      %
-      %
-      %
-      &= \frac{m}{k} + \sigma^2 \frac{1}{k} \sum_j \lambda_j^{-1} - \sigma^2 \lambda_i^{-1} \\
-      %
-      %
-      %
-      &= \frac{1}{k} \left ( m - k \sigma^2\lambda_i^{-1} + \sigma^2 \sum_{j=1}^k \lambda_j^{-1} \right ).
+      \eta_i &= \frac{1}{2\alpha} - \sigma^2 \lambda_i^{-1}\\
+      %
+      %
+      %
+      &=\frac{m}{k} + \frac{\sigma^2}{k} \sum_{j=1}^k \lambda_j^{-1} - \sigma^2 \lambda_i^{-1}.%\\
+      %
+      %
+      %
+      %% &= \frac{m}{k} + \frac{\sigma^2}{k} \sum_{j \in A, j\neq i} \lambda_j^{-1} - \frac{k-1}{k}\sigma^2 \lambda_i^{-1}\\
+      %% ,\ i \in A.
     \end{split}
-  \end{align*}
+  \end{align}
 \end{proof}
-
-%% We now turn our attention to maximizing $\bar{\tar}$ in $\eta$. The only
-%% constraint on $\eta$ arises from the unit norm contraint on rows of
-%% $\obs$, which implies that $\tr{\obs^*\obs} = \sum_i \eta_i= m$.
-%% \begin{proposition}
-%%   Fix $\lambda \in \mathbb{R}_+^k, \sigma^2 > 0$. Then
-%%   $$
-%%   \eta = \arg\max_{\sum_i \eta_i = m} \Phi(\eta; \lambda, \sigma^2)
-%%   $$ 
-%% \end{proposition}
-
-%% \begin{proof}
-%%   Lagrange multipliers:
-%%   \begin{align*}
-%%     \begin{split}
-%%       \frac{d}{d\eta_i} \Phi(\eta; \lambda, \sigma^2) &= \frac{1}{\sigma^2\lambda_i^{-1} + \eta_i}, \\
-%%       %
-%%       %
-%%       %
-%%       \frac{d}{d\eta_i} (\sum_i \eta_i - m)  &= 1.
-%%     \end{split}
-%%   \end{align*}
-
-%%   Denote the Lagrange multiplier $c$. Then
-%%   \begin{align*}
-%%     \begin{split}
-%%       \sigma^2\lambda_i^{-1} + \eta_i &= \frac{1}{c}, \\
-%%       %
-%%       %
-%%       %
-%%       \sum_i \eta_i &= m.
-%%     \end{split}
-%%   \end{align*}
-
-%%   and so
-
-%%   \begin{align*}
-%%     \begin{split}
-%%       \sigma^2\sum_{i=1}^k\lambda_i^{-1} + m &= kc^{-1} \\
-%%       %
-%%     \end{split}
-%%   \end{align*}
-
-%% \end{proof}
-
-%% Corollary \ref{cor:same ev} implies that instead of considering
-%% $\obs$ when searching for a D-optimal design, we can restrict
-%% ourselves to $k \leq m$ eigenvalues of $\obs^*\obs$ corresponding
-%% to the first $k$ eigenvalues of $\fwd \prcov \fwd^*$.
-
-The nonlinear eigenvalue problem \eqref{eq:eigenproblem} gives rise to
-necessary conditions for D-optimality. Thus, in order to find the
-global maximizer of $\tar$, we need to choose $k$, then choose $k$
-eigenvalues $\lambda_1, \dots, \lambda_k$ and corresponding $\eta_1,
-\dots, \eta_k$ to maximize $\sum_i \log( \sigma^2\lambda_i^{-1} +
-\eta_i)$. It is not clear, however, if we can in fact choose
-observation vectors to manifest any $\eta_1, \dots, \eta_k$
-combination. The following lemma answers this in the
-affirmitive. Moreover, the proof is constructive.
-
+An immediate corollary from Proposition \ref{prop:} and the formula
+for posterior eigenvalues is:
+\begin{corollary}
+  If $\obs$ is optimal and $k := \rank \obs$, then:
+  \begin{enumerate}
+    \item The pushforwad/pullback (??) $\fwd \post$ covariance is
+      $((\fwd \prcov \fwd^*)^{-1} + \sigma^{\pm 2} \obs^*\obs)$.
+    \item The eigenvalues of said operator are
+      \begin{equation}
+        \eta_i = \begin{cases}
+          \frac{m}{k} + \sigma^2 \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1} & i \leq k \\
+          \lambda_i^{-1} & i > k 
+        \end{cases}.
+      \end{equation}
+  \end{enumerate}
+\end{corollary}
+%% Thus, for a given $A$, we know which $\eta_i$ are optimal.
+
+  %% Now, fix some $k$ and consider the constrained optimzation problem
+  %% of maximizing $\tar$ subject to $\sum_{i=1}^k\eta_i=m$ (without the
+  %% constraints $\eta_i \geq 0$). It is easy to verify via Lagrange
+  %% multipliers that \eqref{eq:etas} are still optimal. Returning to our
+  %% original constrained optimization problem, where $\eta_i \geq 0$,
+  %% the optimal $\eta_i$'s are defined by \eqref{eq:etas}, where $k$
+  %% takes the largest possible value for which \eqref{eq:etas} defines
+  %% nonnegative $\eta_i$'s.
+
+  %% By definition, $k$ satisfies $\eta_i > 0, i=1,\dots k$. Rearranging
+  %% \eqref{eq:etas} for $\eta_i > 0$, we get that the following must
+  %% hold:
+  %% \begin{equation*} \label{eq:k}
+  %% \lambda_i^{-1} < \sigma^{-2}\frac{m}{k} + \frac{1}{k}
+  %% \sum_{j=1}^{k} \lambda_j^{-1}, i=1,\dots, k.
+  %% \end{equation*}
+
+  %% Since $\lambda_i$ are arranged in descending order, it suffices to
+  %% verify that
+  %% \begin{equation*}% \label{eq:k}
+  %%   \lambda_k^{-1} < \sigma^{-2}\frac{m}{k} + \frac{1}{k} \sum_{j=1}^{k}
+  %%   \lambda_j^{-1}.
+  %% \end{equation*}
+
+  %% Thus:
+  %% \begin{equation*}
+  %%   k = \argmax \left \{ k:\lambda_k^{-1} < \sigma^{-2}\frac{m}{k} + \frac{1}{k} \sum_{j=1}^{k}
+  %%   \lambda_j^{-1} \right \}.
+  %% \end{equation*}
+
+We are almost ready to state and prove Theorem \ref{thm:char} that
+characterizes D-optimal designs. We just need one final lemma, whose
+proof is delegated to the appendix:
 \begin{restatable*}[Unit norm decomposition]{lemma}{free}\label{lemma:free}
@@
 
-with the norm constraints on $\meas_j,j=1,\dots,m$. We would like to
-make all $\eta_i$ as large as possible.  The problem of D-optimal
-design has now been cast to finite dimensions. This justifies using
-linear algebra. Lemma \ref{lemma:free} (a constructive proof is given
-in the appendix) does exactly that.
-
-Take any $\{\eta_i\}_{i=1}^{k}$, with $\sum_{i=1}^k \eta_i = m$. Let
-$M = \diag(\eta_1,\dots,\eta_k)$. Lemma \ref{lemma:free} gives us $A$
-with unit norm columns such that $AA^T = M$. Take $\meas_i
-:=\sum_{j=1}^k A_{ji}\ev_j^*$. The constraints $\|\meas_i\| = 1,
-i=1,\dots,m$ are satisified by construction. Verifying $\obs^*\obs
-\ev_i = \eta_i \ev_i$ is not too difficult.
-
-%% \begin{align*}
-%%   \begin{split}
-%%     \obs \ev_i &= (\meas_1,\dots,\meas_k)^t\ev_i\\
-%%     &= (\sum_{j=1}^kA_{j1}\ev_j^*,\dots,\sum_{j=1}^kA_{jk}\ev_j^*)^t\ev_i\\
-%%     &= (\sum_{j=1}^kA_{j1}\ev_j^*\ev_i,\dots,\sum_{j=1}^kA_{jk}\ev_j^*\ev_i)^t\\
-%%     &= (A_{i1},\dots,A_{ik})^t.
-%%   \end{split}
-%% \end{align*}
-
-%% Now,
-
-%% \begin{align*}
-%%   \begin{split}
-%%     \obs^*\obs \ev_i &= \obs^*(A_{i1},\dots,A_{ik})^t\\
-%%     &=(A_{i1},\dots,A_{ik}) \obs \text{ (by \eqref{eq:obs*})}\\
-%%     &=(A_{i1},\dots,A_{ik}) (\sum_{j=1}^kA_{j1}\ev_j^*,\dots,\sum_{j=1}^kA_{jk}\ev_j^*)^t\\
-%%     &=\sum_{l=1}^k A_{il} \sum_{j=1}^kA_{jl}\ev_j^* \\
-%%     %&=\sum_{j=1}^k  \sum_{l=1}^kA_{il}A_{jl}\ev_j^* \\
-%%     &=\sum_{j=1}^k  [AA^t]_{ij}\ev_j^* \\
-%%     %&=\eta_i\ev_i^*\\
-%%     &=\eta_i\ev_i^*. 
-%%   \end{split}
-%% \end{align*}
-
-We have shown that we can construct $\obs$ such that it has
-eigenvectors and eigenvalues $\{\ev_i, \eta_i\}_{i=1}^k$, as long as
-$\sum_{i=1}^k \eta_i = m$. We now turn to understand which choice of
-$\{\eta_i\}_{i=1}^k$ is optimal. By concavity of $\log$, the fastest
-increase in the design criterion \eqref{eq:true target} is gained by
-increasing $\sigma^2\lambda_i^{-1} + \eta_i$ where it is smallest. So,
-no weight should be given to (i.e. observation taken in direction of)
-the $n$th eigenvector before $\sigma^2\lambda_i^{-1} + \eta_i \geq
-\sigma^2\lambda_n^{-1},\ \forall i \neq n$. Thus, we should choose
-$\{\eta_i\}_{i=1}^k$ as follows. First, order the eigenvalues
-$\lambda_1 \geq \lambda_2 \geq \dots$. We should increase $\sigma^2
-\lambda_1^{-1} + \eta_1$ until it equals
-$\sigma^2\lambda_2^{-1}$. Then, $\sigma^2 \lambda_i^{-1} +
-\eta_i,i=1,2$ are increased until $\sigma^2 \lambda_i^{-1} + \eta_i =
-\sigma^2\lambda_3^{-1},i=1,2$ and so forth.
+\main
+
+\begin{proof}
+  Part (1) was proved in Corollary \ref{cor:same ev}. Parts (2) and
+  (3) were proved in Corollary \ref{cor:true target}. Parts (4) and
+  (5) were proved in \ref{prop:optimizer}. The only missing part is
+  showing that we can indeed find $\obs$ such that $\obs^*\obs$ has
+  the required eigenvalues. To this end, we utilize Lemma
+  \ref{lemma:free}: let $M = \diag(\eta_1, \dots, \eta_k)$, diagonal
+  with respect to the first $k$ eigenvectors of $\fwd \prcov
+  \fwd^*$. We take $\obs := A$ from the Lemma \ref{lemma:free}, and
+  this concludes the proof.
+\end{proof}
+\begin{corollary}
+  In a D-optimal design, the posterior precision for eigenvector $i$
+  is $\eta_i + \sigma^2 \lambda_i^{-1} = \frac{m}{k} + \sigma^2
+  \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1}$. The posterior precision is
+  constant for $1 \leq i \leq k$.
+\end{corollary}
 
@@
   \end{tikzpicture}
-  \caption{Characterizing a D-optimal design, compared to a
-    sub-optimal design. Both designs are allowed three observations,
-    which means the blue area has accumulated height of $3$ in both
-    plots. A D-optimal design (left) increases precision (reduces
-    uncertainty) where it is lowest. A sub-optimal design (right) does
-    not. In this example, a observation is taken partly in the
-    direction of the third eigenvector. The precision for this
-    eigenvector is bigger even than the posterior precision of the
-    D-optimal design on the left.}
+  \caption{As an illustration of Corollary \ref{cor:equal eigs}, a
+    D-optimal design is compared with a sub-optimal design. Both
+    designs are allowed three observations. Thus, the blue area has
+    accumulated height of $3$ in both plots. A D-optimal design (left)
+    increases precision (reduces uncertainty) where it is lowest. A
+    sub-optimal design (right) does not. In this example, an
+    observation is taken partly in the direction of the third
+    eigenvector. The precision for this eigenvector is bigger even
+    than the posterior precision of the D-optimal design on the left.}
   \label{fig:optimal vs not}
@@
 
-The argument above has the important consequence (illustrated in
-figure \ref{fig:optimal vs not}) that the eigenvalues in the nonlinear
-eigenvalue problem \eqref{eq:eigenproblem} are all equal! This makes
-perfect sense. We know the Lagrange multipliers $\xi_i,i=1,\dots,k$
-tell us how much we gain by relaxing the constraints and there is no
-reason for any eigenvector to give us more than any other (otherwise
-we would put more weight on the better eigenvector, as can be
-understood from the discussion in the following paragraph).
-
-One can understand D-optimal designs as follows: Imagine each
-eigenvector corresponds to a graduated lab cylinder (figure
-\ref{fig:cylinder}). Each cylinder $i$ is filled, a-priori, with green
-liquid of $\sigma^2\lambda_i^{-1}$ volume units. This liquid
-represnets the prior precision, weighted by observation noise. The
-volume of the green liquid is increasing in $i$. A D-optimal design
-with $m$ observations ignores cylinders indexed $m+1$ and above. It
-has blue liquid of $m$ volume units at its disposal. It distributes
-said liquid by repeatedly adding a drop to whatever cylinder currently
-has the lowest level of liquid in it (lowest precision). The result of
-such a procedure is illustrated in figure \ref{fig:optimal vs
-  not}. The analysis conducted in this section is summarized in
-Theorem \ref{thm:char} below.
-
-\begin{theorem}[Characterizing D-optimal designs with vanishing model error]\label{thm:char}
-  Let:
-  \begin{enumerate}
-  \item $\fwd, \prcov$ the forward and prior covariance operators, respectively.
-  \item $m, \sigma^2$ the number of observations taken and observation
-    noise variance, respectively.
-  \item A D-optimal design operator $\obs$:
-    $$
-    \obs = \arg\max_{\|\meas_j\| = 1, j=1,\dots,m}\tar(\obs)
-    $$ 
-  \item $\{\lambda_i\}_{i=1}^\infty$ eigenvalues of $\fwd\prcov\fwd^*$
-    in decreasing order of magnitude.
-  %% \item $\{\ev_i\}_{i=1}^\infty$ their corresponding eigenvectors.
-  \item $\{\eta_i\}_{i=1}^\infty$ eigenvalues of $\obs^*\obs$.
-  \end{enumerate}
-  Then:
-  \begin{enumerate}
-  \item $\obs^*\obs$ and $\fwd\prcov\fwd^*$ are simultaneously diagonalizable.
-  \item There exists $0 < k \leq m$ such that $\eta_i = 0$ for $i\geq
-    k+1$ and $k\leq\rank \obs^*\obs$.
-  \item $\sigma^2\lambda_i^{-1} + \eta_i$ are equal for all $1 \leq i \leq k$.
-  \item If $k < m$, then $\sigma^2\lambda_i^{-1} + \eta_i \leq
-    \sigma^2\lambda_{k+1}^{-1}, i=1,\dots,k$.
-  \end{enumerate}
-\end{theorem}
+Here is a way to understand D-optimal designs with vanishing model
+error: Imagine each eigenvector corresponds to a graduated lab
+cylinder (Figure \ref{fig:cylinder}). Each cylinder $i$ is filled,
+a-priori, with green liquid of $\sigma^2\lambda_i^{-1}$ volume
+units. This liquid represnets the prior precision, weighted by
+observation noise. The volume of the green liquid is increasing in
+$i$. A D-optimal design with $m$ observations ignores cylinders
+indexed $m+1$ and above. It has blue liquid of volume $m$ units at its
+disposal. It distributes said liquid by repeatedly adding a drop to
+whatever cylinder currently has the lowest level of liquid in it
+(lowest precision). The result of such a procedure is illustrated in
+Figure \ref{fig:optimal vs not}.
 
@@
 \subsection{Sequential Design}\label{subsec:clusterization sequential}
-Recall the sequential optimal design scheme from section
-\ref{subsec:seq vs sim} and the definition of $\obs_k$ from
-\eqref{eq:def obs_k}. The way clusterization can occur in sequential
-design scenario is understood by specializing thorem \ref{thm:char} to
-one observation each time. In this case $m=1$, so $k=1$, $\meas_1
-\parallel \ev_1$ and $\eta_1 = 1$. As hinted in section
-\ref{subsec:seq vs sim}, the posterior becomes the prior for the next
-step. The eigenvectors of the new posterior are the same as for the
-previous one. A new observation will be taken in the direction of
-eigenvector with smalles eigenvalue. This leads to:
+In a sequential optimal design problem, each measurement is taken
+after results of previous measurements are recorded. In each step
+$m=1$, so $k=1$. In this case, Theorem \ref{thm:char} implies each
+measurement is taken along the direction of the eigenvector of $\fwd
+\prcov \fwd^*$ with lowest posterior precision. If $\lambda_i$
+decrease fast enough (e.g. exponentially), the
+
+, $\meas_1 \parallel \ev_1$ and $\eta_1 = 1$. As hinted in
+section \ref{subsec:seq vs sim}, the posterior becomes the prior for
+the next step. The eigenvectors of the new posterior are the same as
+for the previous one. A new observation will be taken in the direction
+of eigenvector with smalles eigenvalue. This leads to:
 \begin{align*}
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{lemma}[Increase due to a Observation]\label{lemma:design increase}
-  Let $\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm := (\meas_1,\dots,\meas_{m-1})^t$. Then
+\begin{lemma}[Increase due to an observation]\label{lemma:design increase}
+  Let $\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm :=
+  (\meas_1,\dots,\meas_{m-1})^t$. Then
   \begin{align*}

### 2023-01-08 latex/paper.tex
Refined proposition statement on eigenvalues, present optimality as a corollary

@@
   $i\leq k$ and $\obs^*\obs$ has eigenvalues $\eta_i$, then $\obs$
-  maximizes $\tar$ subject to $\sum_{i=1}^m \eta_i = m$ and $\rank
+  maximizes $\tar$ subject to $\sum_{i=1}^m \eta_i = m$ and to $\rank
   \obs = k$.
@@
 \end{proof}
-An immediate corollary from Proposition \ref{prop:} and the formula
+An immediate corollary from Proposition \ref{prop:optimizer} and the formula
 for posterior eigenvalues is:
@@
   \begin{enumerate}
-    \item The pushforwad/pullback (??) $\fwd \post$ covariance is
-      $((\fwd \prcov \fwd^*)^{-1} + \sigma^{\pm 2} \obs^*\obs)$.
+    \item The covariance of the pushforwad $\fwd_{*} \post$ is $((\fwd
+      \prcov \fwd^*)^{-1} + \sigma^{-2} \obs^*\obs)^{-1}$.
     \item The eigenvalues of said operator are

### 2023-01-08 src/probability.py

@@ -73,26 +73,20 @@ class Posterior(FourierMultiplier):
         assert np.all(self.C_sqrt_fwd.real >= 0)
         
 
+    def make_diagonal(self, m, k):
+        eigs = self.sigSqr / self.C_sqrt_fwd[:k]**2
+        uniform = np.mean(eigs) + m / k
+        return uniform - eigs
+
     def make_optimal_diagonal(self, m):
-        """Plots and returns the optimal design multiplier"""
-        eigenvalues = self.sigSqr / self.C_sqrt_fwd**2
-        # ind = np.argsort(eigenvalues)
-        # eigenvalues = eigenvalues[ind]
-        # pdb.set_trace()
-        #plt.plot(eigenvalues[ind])
-        # plt.plot(np.minimum(eigenvalues, 10))
-        # plt.show()
-        #print(eigenvalues[:3], eigenvalues[-3:])
         k = 1
         while True:
-            eigs = eigenvalues[:k]
-            uniform = (np.sum(eigs) + m) / k
-            if np.any(eigs >= uniform):
+            eta = self.make_diagonal(m, k)
+            if np.any(eta <= 0):
                 break
-            self.optimal_diagonal_O = np.sqrt(uniform - eigs)
+            self.optimal_diagonal_O = np.sqrt(eta)
             k += 1
           
-
         self.optimal_diagonal_O_matrix = np.zeros((m, self.N))
         np.fill_diagonal(self.optimal_diagonal_O_matrix, self.optimal_diagonal_O)
 

### 2023-03-16 latex/paper.tex
Finished de proof

--- 2023-01-03 latex/paper.tex
@@
     \end{equation*}
-  \item 
-    \begin{equation*}
-      k = \argmax \left \{ k:\lambda_k^{-1} < \sigma^{-2}\frac{m}{k} + \frac{1}{k} \sum_{j=1}^{k}
-      \lambda_j^{-1} \right \}.
-    \end{equation*}
+  %% \item 
+  %%   \begin{equation*}
+  %%     k = \argmax \left \{ k:\lambda_k^{-1} < \sigma^{-2}\frac{m}{k} + \frac{1}{k} \sum_{j=1}^{k}
+  %%     \lambda_j^{-1} \right \}.
+  %%   \end{equation*}
   \item
@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SECTION Analysis of Optimal Designs --- Vanishing Model Error
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Analysis of D-Optimal Designs --- Vanishing Model Error}\label{section:vanishing}
-\subsection{Overview}
-In this section we analyze D-optimal designs when no model error is
-present. Sensor clusterization is discussed in section
-\ref{section:clusterization}. We first derive a nonlinear eigenvalue
-problem in $\obs$ for D-optimal designs in Section
-\ref{subsec:eigenproblem}. In Section \ref{subsec:characterization} we
-characterize D-optimal designs as ones that make all eigenvalues equal
-in the nonlinear eigenvalue problem. Results are summarized in Theorem
-\ref{thm:char}.
-
-\subsection{The Nonlinear Eigenvalue Problem}\label{subsec:eigenproblem}
-$\modcov = 0$ implies $\Sigma= \sigma^2I$. The necessary first-order
-condition for D-optimality \eqref{eq:conditions} becomes
-\begin{equation}\label{eq:eigenproblem}
-  \sigma^{-2}\fwd \postcov \fwd^* \obs^* = \obs^* \Xi,
-\end{equation}
-with $\Xi$ diagonal. We think of \eqref{eq:eigenproblem} as an
-eigenvalue problem for the self-adjoint operator $\sigma^{-2}\fwd
-\postcov \fwd^*$, where rows of $\obs$, namely $\meas_j,j=1,\dots, m$,
-are eigenvectors. However, that $\postcov$ depends on $\obs$ and thus,
-we refer to \eqref{eq:eigenproblem} as a nonlinear eigenvalue problem.
-
-The following two lemmas allow us to gain better understanding of the
-nonlinear eigenvalue problem.
-
-\begin{restatable*}[Two applications of Woodbury Matrix Identity]{lemma}{woodbury}\label{lemma:twice woodbury}
-  Assume $\fwd \prcov \fwd^*$ is invertible. Then
-  \begin{align*}
-    \begin{split}
-      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* 
-      %
-      %
-      = \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1},
-    \end{split}
-  \end{align*}  
-\end{restatable*}
-
-%% The term $(\fwd \prcov \fwd^*)^{-1}$ is the prior precision in
-%% $\hilo$, and $\fwd\prcov\fwd^*$ is invertible, as mentioned in
-%% Section \ref{section:prelim}. $(\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}
-%% \obs^* \obs$ is the posterior precision in $\hilo$.
-
-%% \subsection{Characterizing D-Optimal Designs}\label{subsec:characterization}
-%% %% Denote corresponding eigenvectors $\{\ev_i\}_{i=1}^{\infty}$.
-%% Denote eigenvalues of $\fwd \prcov \fwd^*$ by
-%% $\{\lambda_i\}_{i=1}^{\infty}$, and note they are positive. Lemma
-%% \ref{lemma:sim diag} is crucial for the rest of the analysis:
-
-\begin{restatable*}[Simultaneous diagonizability for the nonlinear eigenvalue problem]{lemma}{simdiag}\label{lemma:sim diag}
-  Let $\hil$ separable Hilbert space, $C:\hil \to \hil$ self-adjoint
-  and $\func_1,\dots,\func_m \in \hil$. Denote $\func^*$ the element
-  $\func$ acting as a linear functional. If
-  \begin{equation*}
-   (C + \sum_{j=1}^m \func_j\func_j^*) \func_l = \xi_l \func_l,\ l = 1,\dots,m
-  \end{equation*}
-  then $C$ and $\sum_{j=1}^m \func_j \func_j^*$ are simultaneously
-  diagonalizable.
-\end{restatable*}
-
-\begin{corollary}[Eigenvectors of a D-optimal design]\label{cor:same ev}
-  Let $\obs$ satisfy the nonlinear eigenvalue problem
-  \eqref{eq:eigenproblem}. Then $\obs^*\obs$ has the same eigenvectors
-  as $\fwd \prcov \fwd^*$.
-\end{corollary}
-\begin{proof}
-  \begin{align}\label{eq:mod conditions}
-    \begin{split}
-      \obs^* \Xi &= \sigma^{-2}\fwd \postcov \fwd^* \obs^*  \text{ (by \eqref{eq:eigenproblem})}\\
-      %
-      %
-      %
-      &= \sigma^{-2} \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* \obs^*  \text{ (by \eqref{eq:postcov})} \\
-      %
-      %
-      %
-      &= \sigma^{-2} \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1} \obs^* \text{ (by Lemma \ref{lemma:twice woodbury})}.
-    \end{split}
-  \end{align}
-
-  Now take $\func_j^{*} = \meas_j$ and $C := (\fwd \prcov
-  \fwd^*)^{-1}$ and use Lemma \ref{lemma:sim diag}.
-\end{proof}
-
-%% Corollary \ref{cor:same ev} is important because it allows us to only think
-
-implies we only need to consider $\obs$ such that  
-At most $m$ of the eigenvectors of $\obs^*\obs$ have a positive
-eigenvalue, because $k := \rank \obs^*\obs \leq m$. Since we made no
-assumption regarding the ordering of $\{\lambda_i\}$, we can denote
-the corresponding non-zero eigenvalues of $\obs^*\obs$ by
-$\{\eta_i\}_{i=1}^{k}$ and let $\eta_i = 0$ for $i \geq k+1$.
-
-%% \begin{definition}
-%%   Let $\{\lambda_i\}_{i=1}^{\infty}$ eigenvalues of $\fwd \prcov
-%%   \fwd^*$ sorted in descending order. Let $\eta \in
-%%   \mathbb{R}_+^m$. The \emph{utility proxy} $\bar{\tar}$ is defined
-%%   as:
-%%   \begin{equation*}
-%%     \bar{\tar}(\eta) := \frac12 \sum_{i=1}^m \log(1 + \sigma^{-2}\lambda_i\eta_i).
-%%   \end{equation*}
-%% \end{definition}
-
-\begin{corollary}\label{cor:true target}
-  Let $\obs$ with $m$ observations satisfy the nonlinear eigenvalue
-  problem \eqref{eq:eigenproblem}. Let $\{\eta_i\}_{i=1}^{\infty}$
-  eigenvalues of $\obs^*\obs$ and $\{\lambda_i\}_{i=1}^{\infty}$ the
-  corresponding eigenvalues of $\fwd \prcov \fwd^*$. Let $k:=\rank
-  \obs^*\obs$. Without loss of generality, let $\eta_i > 0$ for $i\leq
-  k$ and $\eta_i = 0$ for $i > k$. Then:
-  \begin{enumerate}
-    \item $k \leq m$ and $\obs^*\obs$ has exactly $k$ positive
-      eigenvalues.
-    \item
-      \begin{equation*}
-        \tar(\obs) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i) = \frac12 \sum_{i=1}^{m} \log (1 + \sigma^{-2}\lambda_i\eta_i).
-      \end{equation*}
-    \item Let $\obs$ a D-optimal design. Then $\eta_i > 0$ for
-      eigenvectors corresponding to the $k$ largest $\lambda_i$.
-  \end{enumerate}
-\end{corollary}
-\begin{proof}
-  Part (1) is trivial. To see part (2) holds: 
-  \begin{align}
-    \begin{split}
-      \tar(\obs) &= \frac12\log \det \left (I + \sigma^{-2} \prcov^{1/2} \fwd ^* \obs^*
-      \obs \fwd \prcov^{1/2}\right )\\% \text{ (by definition)}\\
-      %
-      &= \frac12 \log \det \left (I + \sigma^{-2} \obs^* \obs \fwd
-      \prcov\fwd^* \right ) \text{ (Sylvester's Determinant
-      Theorem)}\\
-      %
-      %
-      %
-      &=\frac12 \log \prod_{i=1}^{\infty} ( 1 + \sigma^{-2} \lambda_i\eta_i ) \text{ (Corollary \ref{cor:same ev})} \\
-      %
-      %
-      %
-      %&=\frac12 \log \left ( \prod_{i=1}^{k} ( \lambda_i^{-1} + \sigma^{-2} \eta_i )\prod_{i=1}^{k} \lambda_i \right )
-      %
-      %
-      %
-      &=\frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i). 
-      %
-      %
-      %% &=\frac12 \sum_{i=1}^{k} \log(\sigma^2\lambda_i^{-1} + \eta_i) - k\log \sigma + \frac12\sum_{i=1}^k \log \lambda_i.
-    \end{split}
-  \end{align}
-  Part (3) holds since $\log$ is increasing and $\eta_i \geq 0$.
-\end{proof}
-
-\begin{proposition}\label{prop:optimizer}
-  Let $\{\lambda_i\}_{i=1}^{\infty}$ the sorted eigenvectors of $\fwd
-  \prcov\fwd^*$, $k\in \mathbb{N}^+$ and
-  %% Let
-  %%   \begin{equation}
-  %%   k = \argmax \left \{ k:\lambda_k^{-1} \leq \sigma^{-2}\frac{m}{k} + \frac{1}{k} \sum_{j=1}^{k}
-  %%   \lambda_j^{-1} \right \}.
-  %% \end{equation}
-
-  \begin{equation}
-  \eta_i = \begin{cases}
-    \frac{m}{k} - \sigma^2 \lambda_i^{-1} + \sigma^2 \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1} & i \leq k \\
-    0 & i > k 
-  \end{cases}.
-  \end{equation}
-
-  Let $\obs$ an observation operator such that $\obs^*\obs$ has the
-  same eigenvectors as $\fwd \prcov \fwd^*$. If $\eta_i > 0$ for
-  $i\leq k$ and $\obs^*\obs$ has eigenvalues $\eta_i$, then $\obs$
-  maximizes $\tar$ subject to $\sum_{i=1}^m \eta_i = m$ and $\rank
-  \obs = k$.
-\end{proposition}
-\begin{proof}
-  Below, we use a slight abuse of notation and cast the entire
-  optimization problem of $\tar$ to the first $m$ eigenvectors of
-  $\fwd\ \prcov \fwd^*$. We treat $\eta$ as if it is $m-$dimensional,
-  since by \ref{}, $\eta_i = 0, i > m$. This abuse of notation gives
-  us the freedom to think of the problem at hand in $m-$dimensions. We
-  will now find the optimal $\eta_i$'s.
-
-  First, let $A := \{i: \eta_i > 0\}$. A choice of $A$ where $\eta_i
-  =0, \eta_j >0$ for some $1 \leq i < j \leq m$ is not optimal (just
-  switch $\eta_j$ and $\eta_i$ and recall $\{\lambda_i\}$ are arranged
-  in descending order). Thus, since $\rank \obs = k$ we conclude that,
-  $\eta_i=0, i>k$.
-
-  Let $\Phi(\eta) = \sum_{i=1}^k \eta_i - m$ and $\Omega_j(\eta) =
-   -\eta_j$. Then
-  \begin{align*}
-    \begin{split}
-      \frac{\partial \tar}{\partial \eta_i}  &=
-       \frac12 \frac{\sigma^{-2}\lambda_i}{1 + \sigma^{-2} \lambda_i\eta_i} = \frac12 \frac{1}{\sigma^{2}\lambda_i^{-1} + \eta_i} \\
-      %
-      %
-      %
-      \frac{\partial\Phi}{\partial \eta_i} &= 1 \\
-      %
-      %
-      %
-      \frac{\partial \Omega_j}{\partial \eta_i} &= -\delta_{ij}      
-    \end{split}
-  \end{align*}
-
-  From the KKT conditions, there are $\alpha, \beta_i$ such that for $i=1,\dots,m$:
-  \begin{align}
-    \begin{split}
-      -\frac12 \frac{1}{\sigma^{2}\lambda_i^{-1} + \eta_i} + \alpha - \beta_i  &= 0 \\
-      %
-      %
-      %
-      \eta_i &\geq 0\\
-      %
-      %
-      %
-      \beta_i &\geq 0\\
-      %
-      %
-      %
-      \beta_i \eta_i &= 0\\
-      %
-      %
-      %
-      \sum_{i=1}^m \eta_i &= m 
-    \end{split}
-  \end{align}
-
-  Then, for $1 \leq i \leq k$:
-  \begin{align*}
-    \begin{split}
-      \beta_i &= 0\\
-      %
-      %
-      %
-      \sigma^{2}\lambda_i^{-1} + \eta_i  &= \frac{1}{2\alpha}%  \text{ for }  1 \leq i \leq k.\\
-      %
-      %
-      %
-      %% \sigma^{-2} \lambda_i &= 2\alpha - 2\beta_i, i \in B \\
-    \end{split}
-  \end{align*}
-
-  Summing, substituting $\sum_{i=1}^m \eta_i = \sum_{i=1}^k \eta_i =
-  m$, and dividng by $k$:
-  $$
-  \frac{1}{2\alpha} = \frac{m}{k} + \frac{\sigma^2}{k} \sum_{j=1}^k \lambda_j^{-1}.
-  %\Rightarrow \alpha = \left ( \frac{2m}{k} + \frac{2\sigma^2}{k}
-  %\sum_{i\in A} \lambda_i^{-1} \right )^{-1}.
-  $$
-
-  Consequently:
-  \begin{align} \label{eq:etas}
-    \begin{split}
-      \eta_i &= \frac{1}{2\alpha} - \sigma^2 \lambda_i^{-1}\\
-      %
-      %
-      %
-      &=\frac{m}{k} + \frac{\sigma^2}{k} \sum_{j=1}^k \lambda_j^{-1} - \sigma^2 \lambda_i^{-1}.%\\
-      %
-      %
-      %
-      %% &= \frac{m}{k} + \frac{\sigma^2}{k} \sum_{j \in A, j\neq i} \lambda_j^{-1} - \frac{k-1}{k}\sigma^2 \lambda_i^{-1}\\
-      %% ,\ i \in A.
-    \end{split}
-  \end{align}
-\end{proof}
-An immediate corollary from Proposition \ref{prop:} and the formula
-for posterior eigenvalues is:
-\begin{corollary}
-  If $\obs$ is optimal and $k := \rank \obs$, then:
-  \begin{enumerate}
-    \item The pushforwad/pullback (??) $\fwd \post$ covariance is
-      $((\fwd \prcov \fwd^*)^{-1} + \sigma^{\pm 2} \obs^*\obs)$.
-    \item The eigenvalues of said operator are
-      \begin{equation}
-        \eta_i = \begin{cases}
-          \frac{m}{k} + \sigma^2 \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1} & i \leq k \\
-          \lambda_i^{-1} & i > k 
-        \end{cases}.
-      \end{equation}
-  \end{enumerate}
-\end{corollary}
-%% Thus, for a given $A$, we know which $\eta_i$ are optimal.
-
-  %% Now, fix some $k$ and consider the constrained optimzation problem
-  %% of maximizing $\tar$ subject to $\sum_{i=1}^k\eta_i=m$ (without the
-  %% constraints $\eta_i \geq 0$). It is easy to verify via Lagrange
-  %% multipliers that \eqref{eq:etas} are still optimal. Returning to our
-  %% original constrained optimization problem, where $\eta_i \geq 0$,
-  %% the optimal $\eta_i$'s are defined by \eqref{eq:etas}, where $k$
-  %% takes the largest possible value for which \eqref{eq:etas} defines
-  %% nonnegative $\eta_i$'s.
-
-  %% By definition, $k$ satisfies $\eta_i > 0, i=1,\dots k$. Rearranging
-  %% \eqref{eq:etas} for $\eta_i > 0$, we get that the following must
-  %% hold:
-  %% \begin{equation*} \label{eq:k}
-  %% \lambda_i^{-1} < \sigma^{-2}\frac{m}{k} + \frac{1}{k}
-  %% \sum_{j=1}^{k} \lambda_j^{-1}, i=1,\dots, k.
-  %% \end{equation*}
-
-  %% Since $\lambda_i$ are arranged in descending order, it suffices to
-  %% verify that
-  %% \begin{equation*}% \label{eq:k}
-  %%   \lambda_k^{-1} < \sigma^{-2}\frac{m}{k} + \frac{1}{k} \sum_{j=1}^{k}
-  %%   \lambda_j^{-1}.
-  %% \end{equation*}
-
-  %% Thus:
-  %% \begin{equation*}
-  %%   k = \argmax \left \{ k:\lambda_k^{-1} < \sigma^{-2}\frac{m}{k} + \frac{1}{k} \sum_{j=1}^{k}
-  %%   \lambda_j^{-1} \right \}.
-  %% \end{equation*}
-
-We are almost ready to state and prove Theorem \ref{thm:char} that
-characterizes D-optimal designs. We just need one final lemma, whose
-proof is delegated to the appendix:
-\begin{restatable*}[Unit norm decomposition]{lemma}{free}\label{lemma:free}
-  Let $M \in \R^{k \times k}$ symmetric positive definite with $\ttr M
-  = m$, $m \geq k$. We can find $\func_j \in \R^k,j=1,\dots,m$
-  with $\|\func_j\|=1$ and $A = (\func_1,\dots,\func_m)$ such that
-  $AA^t = M$.
-\end{restatable*}
-
-\main
-
-\begin{proof}
-  Part (1) was proved in Corollary \ref{cor:same ev}. Parts (2) and
-  (3) were proved in Corollary \ref{cor:true target}. Parts (4) and
-  (5) were proved in \ref{prop:optimizer}. The only missing part is
-  showing that we can indeed find $\obs$ such that $\obs^*\obs$ has
-  the required eigenvalues. To this end, we utilize Lemma
-  \ref{lemma:free}: let $M = \diag(\eta_1, \dots, \eta_k)$, diagonal
-  with respect to the first $k$ eigenvectors of $\fwd \prcov
-  \fwd^*$. We take $\obs := A$ from the Lemma \ref{lemma:free}, and
-  this concludes the proof.
-\end{proof}
-\begin{corollary}
-  In a D-optimal design, the posterior precision for eigenvector $i$
-  is $\eta_i + \sigma^2 \lambda_i^{-1} = \frac{m}{k} + \sigma^2
-  \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1}$. The posterior precision is
-  constant for $1 \leq i \leq k$.
-\end{corollary}
-
-\pgfplotstableread{
-  Label     prior  optimal  sub-optimal 
-  1         0.2    1.8           1.7
-  2         0.8    1.2           0.8
-  3         2.2    0             0.5
-  4         3.5    0             0.0
-}\optimalvsnot
-
-\begin{figure}
-  \begin{tikzpicture}[scale=0.85]
-    \begin{axis}[
-        ybar stacked,
-        ymin=0,
-        ymax=4,
-        xtick=data,
-        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
-        reverse legend=false, % set to false to get correct display, but I'd like to have this true
-        xticklabels from table={\optimalvsnot}{Label},
-        xticklabel style={text width=2cm,align=center},
-        legend plot pos=right,
-        ylabel=precision --- prior and posterior,
-        xlabel=eigenvector $i$,
-      ]
-
-      \addplot [fill=green!80]  table [y=prior,   meta=Label, x expr=\coordindex] {\optimalvsnot};
-      \addplot [fill=blue!60]   table [y=optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};
-
-      \addlegendentry{$\sigma^2\lambda_i^{-1}$}
-      \addlegendentry{optimal $\eta_i$s}
-
-    \end{axis}
-  \end{tikzpicture}
-  \qquad
-  \begin{tikzpicture}[scale=0.85]
-    \begin{axis}[
-        ybar stacked,
-        ymin=0,
-        ymax=4,
-        xtick=data,
-        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
-        reverse legend=false, % set to false to get correct display, but I'd like to have this true
-        xticklabels from table={\optimalvsnot}{Label},
-        xticklabel style={text width=2cm,align=center},
-        legend plot pos=right,
-        ylabel=precision --- prior and posterior,
-        xlabel=eigenvector $i$,
-      ]
-
-      \addplot [fill=green!80]  table [y=prior,       meta=Label, x expr=\coordindex] {\optimalvsnot};
-      \addplot [fill=blue!60]   table [y=sub-optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};
-
-      \addlegendentry{$\sigma^2\lambda_i^{-1}$}
-      \addlegendentry{sub-optimal $\eta_i$s}
-
-    \end{axis}
-  \end{tikzpicture}
-  \caption{As an illustration of Corollary \ref{cor:equal eigs}, a
-    D-optimal design is compared with a sub-optimal design. Both
-    designs are allowed three observations. Thus, the blue area has
-    accumulated height of $3$ in both plots. A D-optimal design (left)
-    increases precision (reduces uncertainty) where it is lowest. A
-    sub-optimal design (right) does not. In this example, an
-    observation is taken partly in the direction of the third
-    eigenvector. The precision for this eigenvector is bigger even
-    than the posterior precision of the D-optimal design on the left.}
-  \label{fig:optimal vs not}
-\end{figure}
-
-\pgfplotstableread{
-  Label    prior  $o_1$  $o_2$   $o_3$   $o_4$   $o_5$         $o_6$    topper
-  1         0.2    0.8    0.5     0.5    0.4    0.33333333  0.3333333  0.001
-  2         0.8    0.2    0.5     0.5    0.4    0.33333333  0.3333333  0.001
-  3         2.2    0      0       0      0.2    0.33333333  0.3333333  0.001
-  4         3.5    0      0       0      0      0           0          0.001
-}\clusterization
-
-\pgfplotstableread{
-  Label    prior  $o_1$  $o_2$   $o_3$   $o_4$   $o_5$         $o_6$    topper
-  1         0.2    0.8    0.7     0.3    0.4    0.2         0.4666666  0.001
-  2         0.8    0.2    0.3     0.7    0.4    0.46666666  0.2        0.001
-  3         2.2    0      0       0      0.2    0.33333333  0.3333333  0.001
-  4         3.5    0      0       0      0      0           0          0.001
-}\noclusterization
-
-Here is a way to understand D-optimal designs with vanishing model
-error: Imagine each eigenvector corresponds to a graduated lab
-cylinder (Figure \ref{fig:cylinder}). Each cylinder $i$ is filled,
-a-priori, with green liquid of $\sigma^2\lambda_i^{-1}$ volume
-units. This liquid represnets the prior precision, weighted by
-observation noise. The volume of the green liquid is increasing in
-$i$. A D-optimal design with $m$ observations ignores cylinders
-indexed $m+1$ and above. It has blue liquid of volume $m$ units at its
-disposal. It distributes said liquid by repeatedly adding a drop to
-whatever cylinder currently has the lowest level of liquid in it
-(lowest precision). The result of such a procedure is illustrated in
-Figure \ref{fig:optimal vs not}.
-
-\begin{figure}%{r}{0.25\textwidth} 
-    %\centering
-    \includegraphics[width=5cm, height=5cm]{cylinders.jpg}
-    \caption{Three graduated lab cylinders, corresponding to three
-      eigenvectors. Prior eigenvalues not shown.}
-    \label{fig:cylinder}
-\end{figure}
-
-\section{Sensor Clusterization --- Vanishing Model Error}\label{section:clusterization}
-\subsection{Overview}
-In this section we use Theorem \ref{thm:char} to explain how sensor
-clusterization may arise in various scenarios. We first consider
-sequential designs, and then simultaneous designs.
-
-\subsection{Sequential Design}\label{subsec:clusterization sequential}
-In a sequential optimal design problem, each measurement is taken
-after results of previous measurements are recorded. In each step
-$m=1$, so $k=1$. In this case, Theorem \ref{thm:char} implies each
-measurement is taken along the direction of the eigenvector of $\fwd
-\prcov \fwd^*$ with lowest posterior precision. If $\lambda_i$
-decrease fast enough (e.g. exponentially), the
-
-, $\meas_1 \parallel \ev_1$ and $\eta_1 = 1$. As hinted in
-section \ref{subsec:seq vs sim}, the posterior becomes the prior for
-the next step. The eigenvectors of the new posterior are the same as
-for the previous one. A new observation will be taken in the direction
-of eigenvector with smalles eigenvalue. This leads to:
-\begin{align*}
-  \begin{split}
-    \meas_2 \parallel \ev_1  &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 < \sigma^2\lambda^{-1}_2 \\
-    \meas_2 \parallel \ev_2  &\text{ o.w. }
-  \end{split}
-\end{align*}
-In the first case, we get clusterization immediately. In the second,
-we consider $\meas_3$. Same reasoning leads to the conclusion that the
-eigenvectors of the new prior precision are the same. The eigenvalues,
-however, are $\lambda^{-1}_1 + \sigma^{-2}, \lambda^{-1}_2 +
-\sigma^{-2}, \lambda_3^{-1}, \dots$. A new observation is taken in the
-direction of eigenvector of smallest precision (note that
-$\lambda^{-1}_1 + \sigma^{-2} < \lambda^{-1}_2 + \sigma^{-2}$):
-\begin{align*}
-  \begin{split}
-    \meas_3 \parallel \ev_1 &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 <
-    \sigma^2\lambda^{-1}_3 \\
-    \meas_3 \parallel  \ev_3  &\text{ o.w. }
-  \end{split}
-\end{align*}
-Again, the first case results in clusterization. The sequential design
-proceeds in this fashion. Every $\meas_i$ is taken in the direction
-some eigenvector of the prior $\fwd \prcov\fwd^*$. This eigenvector
-has the smallest eigenvalue of the previous step's posterior precision
-operator. If we do not encounter clusterization after $m$ such steps,
-this means that observation $\meas_k$ was taken in the direction
-$\ev_k$ for $k=1,\dots,m$. Thus:
-\begin{align}\label{eq:failure}
-  \begin{split}
-    \sigma^2\lambda_{k-1}^{-1} + 1 \geq \sigma^2\lambda_k^{-1}, \text{ for } 1\leq k \leq m
-    \Longrightarrow \lambda_m^{-1} - \lambda_{m-1}^{-1} \leq \sigma^{-2}.
-  \end{split}
-\end{align}
-
-But \eqref{eq:failure} implies $\{\lambda_i^{-1}\}_{i=1}^{\infty}$
-grows at most linearly. In section \ref{subsec:abstract OED} we
-assumed $\fwd$ is strongly smoothing, which means eigenvalues of
-$(\fwd \prcov\fwd^*)^{-1}$ grow quickly. Thus, if we take $m$ large
-enough, \eqref{eq:failure} will eventually fail and we will observe
-sensor clusterization.
-
-\subsection{Simultaneous Design}\label{subsec:clusterization simultaneous}
-In a simultaneous design scenario, the result is not simple as in the
-sequential design case. We now see that a design that exhibits
-clusterization is as optimal as one that does not. Corollary
-\ref{cor:same ev} tells us that we can realize any design we want
-(encoded in $\{\eta_i\}$), given the trace constraint for
-$\obs^*\obs$. Theorem \ref{thm:char} tells us that for D-optimality,
-we only need to make the posterior flat where we take observations
-(recall figure \ref{fig:optimal vs not}). We are left with the freedom
-of assigning any part of $\eta_i$ to any observation $\meas_j$. Two
-possible assignments for the same problem are illustrated in figure
-\ref{fig:clusterization}. The first exhibits clusterization, the
-second does not. It is important to note that neither arises as a
-result of some numerical optimization scheme. Both designs give the
-same design criterion $\tar$, because their posteriors are
-identical. By Theorem \ref{thm:char}, both are maximizers of the same
-D-optimal design problem. One exhibits clusterization, and one does
-not. It is likely that a real-life problem is more restricted than the
-model we consider here. Such restriction is the likely cause of
-clusterization obseved in, e.g., the toy model of section
-\ref{subsec:example}. Exactly how this happens is not clear. At this
-point it is the subject of speculation and should be further studied
-in the future.
-
-\begin{figure}
-  \begin{tikzpicture}[scale=0.85]
-    \begin{axis}[
-        ybar stacked,
-        ymin=0,
-        ymax=4,
-        xtick=data,
-        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
-        reverse legend=false, % set to false to get correct display, but I'd like to have this true
-        xticklabels from table={\clusterization}{Label},
-        xticklabel style={text width=2cm,align=center},
-        legend plot pos=right,
-        ylabel=precision --- prior and posterior,
-        xlabel=eigenvector $i$,
-      ]
-
-      \addplot [fill=green!80]  table [y=prior, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=blue!60]   table [y=$o_1$, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=red!60]    table [y=$o_2$, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=black!60]  table [y=$o_3$, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=orange!60] table [y=$o_4$, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=cyan!60]   table [y=$o_5$, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=purple!60] table [y=$o_6$, meta=Label, x expr=\coordindex] {\clusterization};
-
-      \addlegendentry{prior}
-      \addlegendentry{$o_1$}
-      \addlegendentry{$o_2$}
-      \addlegendentry{$o_3$}
-      \addlegendentry{$o_4$}
-      \addlegendentry{$o_5$}
-      \addlegendentry{$o_6$}   
-    \end{axis}
-  \end{tikzpicture}
-  \qquad
-  \begin{tikzpicture}[scale=0.85]
-    \begin{axis}[
-        ybar stacked,
-        ymin=0,
-        ymax=4,
-        xtick=data,
-        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
-        reverse legend=false, % set to false to get correct display, but I'd like to have this true
-        xticklabels from table={\noclusterization}{Label},
-        xticklabel style={text width=2cm,align=center},
-        legend plot pos=right,
-        ylabel=precision --- prior and posterior,
-        xlabel=eigenvector $i$,
-      ]
-
-      \addplot [fill=green!80]  table [y=prior, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=blue!60]   table [y=$o_1$, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=red!60]    table [y=$o_2$, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=black!60]  table [y=$o_3$, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=orange!60] table [y=$o_4$, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=cyan!60]   table [y=$o_5$, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=purple!60] table [y=$o_6$, meta=Label, x expr=\coordindex] {\noclusterization};
-
-      \addlegendentry{prior}
-      \addlegendentry{$o_1$}
-      \addlegendentry{$o_2$}
-      \addlegendentry{$o_3$}
-      \addlegendentry{$o_4$}
-      \addlegendentry{$o_5$}
-      \addlegendentry{$o_6$}   
-    \end{axis}
-  \end{tikzpicture}
-  \caption{Clusterization (left) and non-clusterization (right) in
-    simultaneous D-optimal designs. Posterior precision per
-    eigenvector after taking observations $\meas_i$ is plotted. Both
-    designs are identical for all practical matters --- their
-    posteriors are identical. In the left panel we see how an optimal
-    design is achieved with repeated observations. $\meas_2 = \meas_3$
-    and $\meas_5 = \meas_6$. This is not necessary though, as can be
-    seen in the right panel. Both designs achieve the same levels of
-    uncertainty. The left exhibits clusterization. The right does not.}
-  \label{fig:clusterization}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 %% SECTION Analysis of Optimal Designs --- Non-Vanishing Model Error
@@
 
-\appendix
-\section{General Lemmas}
-
-The following lemma is generalized from \cite[Chapter 9, Theorem 4, pp. 127]{Lax97}
-\lax
-\begin{proof}
-  Consider a differentiable operator-valued function $X(t)$ such that
-  $X(0) = 0$ and $X(t)$ is positive, self-adjoint and trace-class for
-  every $t\in \R$. We denote the eigenvalues of this operator by
-  $\lambda_k(X(t))$ and sometimes drop the dependence on $X(t)$, so
-  $\lambda_k = \lambda_k(X(t))$.  Then $\det (I+X(t)) =
-  \prod_{k=1}^{\infty} (1+\lambda_k) < \infty$ where the bound holds
-  by the arguments given in \cite{AlexanderianGloorGhattas14}. The
-  full derivative is
-  \begin{align*}
-    \frac{\der \det (I+X(t))}{\der t} 
-    % 
-    % 
-    % 
-    &= \sum_{k=1}^{\infty} 
-    \frac{\partial \det (I+X(s))}{\partial (1+\lambda_k)}\Big |_{s=t}
-    \frac{\der (1+\lambda_k)}{\der t} \\
-    % 
-    %
-    %
-    &= \sum_{k=1}^{\infty} \frac{\partial \prod_{l=1}^{\infty}
-      (1+\lambda_l(s))}{\partial (1+\lambda_k)}\Big |_{s=t}
-    \frac{\der (1+\lambda_k)}{\der t} \\
-    %
-    %
-    %
-    &= \sum_{k=1}^{\infty} \prod_{l=1, l\neq k}^{\infty}
-      (1+\lambda_l(s)) \frac{\partial (1+\lambda_k(s))}{\partial (1+\lambda_k)}\Big |_{s=t}
-    \frac{\der (1+\lambda_k)}{\der t} \\
-    %
-    %
-    %    
-    &= \sum_{k=1}^{\infty} \frac{\prod_{l=1}^{\infty}
-      (1+\lambda_l(s))}{(1+\lambda_k)}\Big |_{s=t}
-    \dot{\lambda_k}(X(t)) \\
-    % 
-    % 
-    % 
-    &= \sum_{k=1}^{\infty} \frac{\det (I+X(t))}{1 +\lambda_k} \dot{\lambda_k}(X(t)).
-  \end{align*}
-  The assumption $X(0) = 0$ means $\lambda_k(X(0)) = 0,\ \forall k \geq 1$. Thus:
-  \begin{align*}
-    \frac{\der (I+\det X(t))}{\der t}\Big |_{t=0} 
-    = \sum_{k=1}^{\infty} \dot{\lambda_k}(X(0)) 
-    = \frac{\der }{\der t}\tr{X(0)}
-    = \tr{\dot{X}(0)},
-  \end{align*}
-  where the second equality follows by monotone convergence. 
-  Let $Y(t)$ a trace-class self-adjoint operator such that 
-  $I+Y(t)$ is invertible.
-  Define $X(t)$ via $I+X(t) = (I+Y(0))^{-1/2} (I+Y(t)) (I+Y(0))^{-1/2}$. 
-  We show $X(t)$ satisfies the conditions above. It is trace-class:
-  \begin{align*}
-    \tr{X(t)} = \tr{(I+Y(0))^{-1} (I+Y(t)) - I}
-    \leq \tr{I+Y(t) - I}< \infty,
-  \end{align*}
-  since $Y(t)$ is trace-class. It is also clear that
-  $X(0) = 0$ and $X(t)$ is self-adjoint.
-  $I+Y(t) = (I+Y(0))^{1/2}(I+X(t))(I+Y(0))^{1/2}$, so
-  \begin{align*}
-    \frac{\der \det (I+Y(t))}{\der t}|_{t=0} 
-    &= \det (I+Y(0))\frac{\der \det (I+X(t))}{\der t}\Big |_{t=0} \\
-    % 
-    % 
-    % 
-    &= \det (I+Y(0)) \tr{\dot{X}(0)} \\
-    % 
-    % 
-    % 
-    &= \det (I+Y(0)) \tr{(I+Y(0))^{-1} \dot{Y}(0)}.
-  \end{align*}
-  Consequently, by the one-variable chain rule:
-  \begin{align*}
-    \frac{\der \log \det (I+Y(t))}{\der t}\Big |_{t=0} &=
-    % 
-    % 
-    % 
-    \frac{1}{\det (I+Y(0))}\frac{\der \det (I+Y(t))}{\der t}\Big |_{t=0} \\ 
-    % 
-    % 
-    % 
-    &= \tr{ (I+Y(t))^{-1} \dot{Y}(t)} \big |_{t=0}.
-  \end{align*}
-  There is nothing special about $t_0 = 0$ --- we could have chosen
-  any other $t_0$ instead. Thus, the relation holds for all $t$.
-\end{proof}
-
-\begin{lemma}[Matrix Determinant Lemma in Hilbert Spaces]\label{lemma:MDL}
-  Let $\hil$ a separable Hilbert space, $u,v\in \hil$ and $A: \hil \to
-  \hil$ an invertible linear operator such that $\tr{A-I} <
-  \infty$. Then $\det A$ and $\det A + uv^*$ are well defined and
-  \begin{equation*}
-    \det (A + uv^*) = (1 + \langle A^{-1} u, v \rangle ) \det A,
-  \end{equation*}
-  where $(A + uv^*)w := Aw + \langle v,w \rangle u$.
-\end{lemma}
-\begin{proof}
-  In this proof we rely on definitions and results from
-  \cite{Simon77}. First, consider $B := I + xy^*$ for some $x,y \in
-  \hil$. We construct an eigenbasis for $B$ and use that to show $\det
-  B = 1 + \langle x, y \rangle$. First let $x_1 := x$.  Now, if $x
-  \parallel y$, take $\{x_n \}_{n=2}^{\infty}$ an orthogonal basis for
-  $span\{x_1\} ^{\perp}$. If, on the other hand, $x \nparallel y$, let
-  \begin{equation*}
-    x_2 := x - \frac{ \langle x, y\rangle}{\|y\|^2}y
-  \end{equation*}
-  and it is easy to verify that $x_2 \perp y$ and $span \{x,y\} = span
-  \{x_1,x_2\}$. Take $\{x_n \}_{n=3}^{\infty}$ an orthogonal basis for
-  $span\{x_1,x_2\} ^{\perp}$. In both cases,
-  \begin{equation*}
-    B x_n =
-    \begin{cases}
-      (1 + \langle x, y \rangle) x_n & n = 1 \\
-      x_n                            & n \neq 1,
-    \end{cases}
-  \end{equation*}
-  and so $\det B = 1 + \langle x, y \rangle$.
-
-  It is easy to verify that $uv^*$ is trace-class and since $\tr{A-I}
-  < \infty$, also $\tr{A + uv^* - I} < \infty$ (sum of two trace-class
-  operators is trace-class). Thus $\det A$ and $\det (A+uv^*)$ are
-  well defined. Let $x:=A^{-1}u$ and $y := v$:
-  \begin{equation*}
-    \det (A + uv^*) = \det A \ \det(I+A^{-1}uv^*) =
-    (1 + \langle A^{-1}u, v \rangle) \det A .
-  \end{equation*}
-\end{proof}
-
-\free
-\begin{proof}
-  Let us diagonalize $M$, so that $M = U D U^t$ with $D =
-  \diag(d_1,\dots,d_k)$ and $U \in \R^{k \times k }$ orthogonal. Let
-  $S \in \R^{k \times m}$ with $S_{ii} = \sqrt{d_{i}}$ and zeros
-  otherwise. Define $A:= U S V^t$, where $V \in \R^{m \times m}$ is
-  orthogonal and will be further restricted later. Then $AA^t = U
-  SV^tVS^t U^t = UDU^t$, so $AA^t$ has the required eigenvalues and
-  eigenvectors by construction. If we can choose $V$ such that $A$
-  also satisfies the unit norm constraints we are done. These
-  constraints are, for $j=1,\dots,m$:
-  \begin{equation}\label{eq:V constraints}
-   1 = [A^tA]_{jj} = [V S^tS V^t]_{jj},
-  \end{equation}
-  and we can expect to do this since we assumed $\ttr D = m$.
-
-  Define $C = S^tS - I \in \R^{m \times m}$. Note that $\ttr C = 0$ and
-  $C$ is diagonal with non-zero entries $d_i-1,i=1,\dots,k$. It suffices
-  to find $V$ orthogonal such that $V C V^t$ has zero diagonal. We
-  construct such $V$ by sequentially inserting zeros in the diagonal
-  and not destroying zeros we already introduced, starting from the
-  last diagonal entry and moving to the first. Since $c_{mm} \neq 0$ ,
-  let $p < m$ such that $c_{pp}c_{mm} < 0$ (such $p$ exists because
-  the trace is zero) and let $\theta \in (0,\pi)$. Define a Givens
-  rotation $R^{(m)} \in \R^{m \times m}$ by
-  \begin{equation*}
-    r^{(m)}_{ab} :=
-    \begin{cases}
-      1 & a = b \neq p \text{ or } a = b \neq m \\
-      \cos \theta & a = b = p  \\
-     -\sin \theta & a = p, b = m\\
-      \cos \theta & a = b = m \\
-      \sin \theta & a = m, b = p \\ 
-      0 & o.w
-    \end{cases}
-  \end{equation*}
-  Note that conjugating a matrix by $R^{(m)}$ changes only its $m$ and
-  $p$ rows and columns. We want to choose $\theta$ such that
-  \begin{equation}\label{eq:mm}
-    0 = [R^{(m)} C (R^{(m)})^t]_{mm} = \cos^2 \theta c_{mm} + 2\cos \theta \sin
-    \theta c_{mp} + \sin^2\theta c_{pp},
-  \end{equation}
-  and it suffices to choose $\theta$ such that
-  \begin{equation*}
-    c_{mm} \cot^2 \theta + 2 c_{mp} \cot \theta + c_{pp} = 0.
-  \end{equation*}
-  This quadratic in $\cot\theta$ has a real solution, since
-  $c_{pp}c_{mm} < 0$ by assumption and we can find $\theta \in
-  (0,\pi)$ such that \eqref{eq:mm} is satisfied. We continue to find
-  $R^{(m-1)}$ that leaves row and column $m$ unchanged and
-  continue introducing zeros to the diagonal. The assumption $\ttr D =
-  m \Rightarrow \ttr C = 0$ guarantees we can do that. Taking $V:=
-  R^{(1)} R^{(2)} \dots R^{(m-1)}R^{(m)}$ completes the proof.
-\end{proof}
-
-\simdiag
-\begin{proof}
-  First, enumerate the eigenvalues of $C + \sum_{j=1}^m
-  \func_j\func_j^*$ as $\xi_1,\dots,\xi_\ell$. Denote the
-  indices of the eigenvectors corresponding to $\xi_i$
-  \begin{equation*}
-    S_i := \{ 1 \leq k \leq m | (C + \sum_{j=1}^m \func_j\func_j^* )\func_k = \xi_i \func_k \}.
-  \end{equation*}
-  Define further
-  \begin{equation*}
-    A_i := \sum_{k \in S_i} \func_k \func_k^*,
-  \end{equation*}
-  which is self-adjoint. Two observations are in order. First,
-  $\sum_{j=1}^m \func_j\func_j^* = \sum_{i=1}^\ell A_i$. Second, $A_i
-  \func_k = 0$ if $k\not \in S_i$, since eigenvectors of different
-  eigenvalue are orthogonal. For $k \in S_i$
-  \begin{equation}\label{eq:on vi}
-    \xi_i \func_k = (C + \sum_{j=1}^m \func_j \func_j^* ) \func_k = (C + A_i) \func_k.
-  \end{equation}
-  Let $V_i := span \{\func_k \}_{k\in S_i}$. Observe that $V_i$ is
-  invariant under $A_i$, by definition, and under $C$, by \eqref{eq:on
-    vi}. Using \eqref{eq:on vi} again, we conclude that on $V_i$, $A_i
-  = \xi_iI - C$. This immediately implies $A_i$ and $C$ have the
-  same eigenvectors on $V_i$. This holds for every $1 \leq i \leq
-  \ell$ and we conclude that $C$ and $A$ have the same eigenvectors.
-\end{proof}
-
-\section{Specific Lemmas}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\woodbury
-\begin{proof}
-  The proof is boring and amounts to using Woodbury's matrix identity
-  twice and a regularization trick. The standard proof for Woodbury's
-  matrix identity works in infinite dimensions, as long as all terms
-  are well defined. Unfortunately, $\obs^*\obs$ is not invertible, so
-  we force it to be. Recall Woodbury's matrix identity:
-  $$
-  (A + UCV)^{-1} = A^{-1} - A^{-1}U(C^{-1} + VA^{-1}U)^{-1}VA^{-1}.
-  $$
-
-  Taking $A := \prcov^{-1}, U := \fwd^*, V := \fwd$ and $C :=
-  \sigma^{-2} (\obs^*\obs+\eps I)$ we get:
-  \begin{align*}
-    \begin{split}
-      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* &=
-      \fwd ( \prcov - \prcov \fwd^* ( \sigma^2(\obs^*\obs + \eps I)^{-1} + \fwd \prcov \fwd^* )^{-1} \fwd \prcov ) \fwd^* \\
-      %
-      %
-      &= X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X
-    \end{split}
-  \end{align*}
-  Let $X := \fwd\prcov \fwd^*$ and $Y_{\eps} := \obs^*\obs + \eps I$,
-  for some small $\eps > 0$, so that $Y_{\eps}$ is invertible.  Note
-  that $X + \sigma^2 Y_{\eps}^{-1}$ is invertible, as the sum of two
-  positive definite operators. Now, taking $A := X^{-1}, C :=
-  \sigma^2Y_{\eps}^{-1}, U := I$ and $V := I$ and using Wodbury's
-  matrix identity in reverse order:
-  \begin{align*}
-    \begin{split}
-      X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X &= (X^{-1} + \sigma^{-2}Y_{\eps})^{-1} \\
-      %
-      %
-      %
-      &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \eps I))^{-1}.
-    \end{split}
-  \end{align*}
-
-  We conclude that $\forall \eps > 0$
-  \begin{align*}
-    \begin{split}
-      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* 
-     &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \eps I))^{-1}.
-    \end{split}
-  \end{align*}
-  Letting $\eps \to 0$ completes the proof.
-\end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{lemma}[Increase due to an observation]\label{lemma:design increase}
-  Let $\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm :=
-  (\meas_1,\dots,\meas_{m-1})^t$. Then
-  \begin{align*}
-    \tar( \obs ) - \tar (\obsm ) &=
-    \frac12 \log \left ( 1 + \frac{
-      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m,
-      (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m \rangle
-    }{
-      \sigma^2 + \meas_m \modcov \meas_m - \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m 
-    }       
-    \right ).
-  \end{align*}
-\end{lemma}
-\begin{proof}
-  We use the Schur complement to write one inverse in terms of the other and
-  introduce notations to make the derivation cleaner. Note that we think of
-  $\obsm$ and $\obsm^*$ as column and row vectors (respectively).
-  \begin{align*}
-    \Sigma( \obs ) &= \Sigma = 
-    \begin{bmatrix}
-      \Sigma (\obsm )           & \obsm \modcov \meas_m \\
-      \meas_m \modcov \obsm^*   & \sigma^2 + \meas_m \modcov \meas_m
-    \end{bmatrix}
-    =
-    \begin{bmatrix}
-      \Sigmam   & w \\
-      w^t       & c
-    \end{bmatrix}\\
-    %
-    %
-    %
-    \Sigma^{-1} &=
-    \begin{bmatrix}
-      \Sigmam^{-1} + \Sigmam^{-1} w ( c - w^t \Sigmam^{-1} w)^{-1} w^t \Sigmam^{-1} & - \Sigmam^{-1} w ( c - w^t \Sigmam^{-1} w)^{-1} \\
-      -( c - w^t \Sigmam^{-1} w)^{-1} w^t \Sigmam^{-1}                            &  ( c - w^t \Sigmam^{-1} w)^{-1}
-    \end{bmatrix} \\
-    &=
-    \begin{bmatrix}
-      \Sigmam^{-1} & 0 \\
-      0           & 0 
-    \end{bmatrix}
-    + (c -w^t \Sigmam^{-1} w )^{-1}
-    \begin{bmatrix}
-      \Sigmam^{-1} w \\
-      -1
-    \end{bmatrix}
-    \begin{bmatrix}
-      w^t \Sigmam^{-1} & -1 
-    \end{bmatrix}
-  \end{align*}
-  %
-  Further, define
-  %
-  \begin{align*}
-    \M (\obs ):&= \prcov^{\frac12}\fwd^* \obs^* \Sigma^{-1} \obs \fwd
-    \prcov^{\frac12}    
-  \end{align*}
-  %
-  and note that, using our understanding of what is a column vector and
-  what is a row vector:
-  %
-  \begin{align*}
-    \M(\obs) &= \prcov^{1/2} \fwd^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2} \\
-    %
-    %
-    %
-    &= \prcov^{1/2} \fwd^* \obs^* \left \{
-    \begin{bmatrix}
-      \Sigmam^{-1} & 0 \\
-      0           & 0 
-    \end{bmatrix}
-    + (c -w^t \Sigmam^{-1} w )^{-1}
-    \begin{bmatrix}
-      \Sigmam^{-1} w \\
-      -1
-    \end{bmatrix}
-    \begin{bmatrix}
-      w^t \Sigmam^{-1} & -1 
-    \end{bmatrix} 
-    \right \} \obs \fwd \prcov^{1/2} \\
-    %
-    %
-    %
-    &= \M (\obsm) + (c -w^t \Sigmam^{-1} w )^{-1}
-    \prcov^{1/2} \fwd^* \obs^*
-    \begin{bmatrix}
-      \Sigmam^{-1} w \\
-      -1
-    \end{bmatrix}
-    \begin{bmatrix}
-      w^t \Sigmam^{-1} & -1 
-    \end{bmatrix} 
-    \obs \fwd \prcov^{1/2}
-  \end{align*}
-  %
-  Now, denote:
-  %
-  \begin{align}\label{eq:u}
-    \begin{split}
-      u :&= (c -w^t \Sigmam^{-1} w )^{-1/2}
-      \prcov^{1/2} \fwd^* \obs^* 
-      \begin{bmatrix}
-        \Sigmam^{-1} w \\
-        -1 
-      \end{bmatrix} \\
-      %
-      %
-      %
-      & = (c -w^t \Sigmam^{-1} w )^{-1/2} ( \prcov^{1/2}\fwd^* \obsm^* \Sigmam^{-1} \obsm  \modcov \meas_m - \prcov^{1/2} \fwd^* \meas_m )\\
-      %
-      %
-      %
-      u^* :&=  (c -w^t \Sigmam^{-1} w )^{-1/2} (\meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \fwd \prcov^{1/2} - \meas_m \fwd \prcov^{1/2} ),
-    \end{split}
-  \end{align}
-  %
-  so that
-  %
-  \begin{equation}\label{eq:M plus I}
-    I + \M( \obs ) = I + \M (\obsm ) + uu^*.
-  \end{equation}
-  %
-  Note that
-  \begin{equation}\label{eq:M postcov}
-    \prcov^{1/2} \left (I + \M( \obsm ) \right )^{-1} \prcov^{1/2} = \postcovm.
-  \end{equation}
-  The increase in the design criterion gained by including $\meas_m$
-  in the design is found by using Lemma \ref{lemma:MDL} the above
-  results:
-  %
-  \begin{align*}
-    \tar( \obs ) - \tar( \obsm )
-    %
-    %
-    %
-    &= \frac12 \log \det \Big ( I + \M ( \obs ) \Big ) / \det \Big ( I + \M (\obsm) \Big ) \\
-    %
-    %
-    %
-    &= \frac12  \log \det \left ( I + \M(\obsm) + uu^* \right ) / \det \Big ( I + \M (\obsm) \Big ) \\
-    %
-    %
-    %
-    &= \frac12 \log \left ( 1 + \left \langle \left ( I+\M(\obsm) \right )^{-1} u, u  \right \rangle \right ).
-  \end{align*}
-  Using \eqref{eq:u}:
-  \begin{align*}
-    &\left \langle \left (I+\M (\obsm)\right )^{-1}u, u \right \rangle\\
-    &= \frac{
-      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m,
-      (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m \rangle
-    }{
-      c- w^t \Sigmam^{-1} w
-    }\\
-    %
-    %
-    %
-    &= 
-    \frac{
-      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m,
-      (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m \rangle
-    }{
-      \sigma^2 + \meas_m \modcov \meas_m - \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m 
-    }
-  \end{align*}
-  and the conclusion follows.
-\end{proof}
-
-Lemma \ref{lemma:design increase} implies the following two corollaries.
-\begin{corollary}[Gain for No Model Error]\label{cor:zero mod err}
-  If $\modcov = 0$, then
-  \begin{equation*}
-    \tar( \obs ) - \tar (\obsm )
-    = -\frac12 \log (1 - \sigma^{-2} \langle \fwd \postcov \fwd^* \meas_m, \meas_m \rangle ).
-  \end{equation*}
-\end{corollary}
-\begin{proof}
-  Note that this is not immediate by substituting $\modcov = 0$ in the
-  conclusion of Lemma \ref{lemma:design increase}, since we make a
-  claim for $\postcov$, and not $\postcovm$. Let us first review
-  \eqref{eq:u} and note that since $\modcov = 0$ the covariance
-  $\Sigmam = \sigma^2I_{m-1}$ and $w = 0$, so $c - w^t\Sigmam^{-1}w =
-  \sigma^2$:
-  \begin{align*}
-    u :&=-\sigma^{-1}\prcov^{1/2} \fwd^* \meas_m\\
-    %
-    %
-    u^* :&= -\sigma^{-1} \meas_m \fwd \prcov^{1/2}.
-  \end{align*}
-  From \eqref{eq:M plus I}:
-  \begin{equation*}
-    I + \M(\obsm) = I +\M(\obs) - uu^*,
-  \end{equation*}
-  and thus:
-  \begin{equation*}
-    \left \langle \left ( I +\M(\obs) \right )^{-1} u, u \right \rangle
-    = \sigma^{-2} \langle \fwd \postcov \fwd^* \meas_m, \meas_m \rangle.
-  \end{equation*}
-  Analogously to \eqref{eq:M postcov} we note that
-  \begin{equation*}
-    \prcov^{1/2} \left ( I + \M(\obs) \right )^{-1} \prcov^{1/2} = \postcov.
-  \end{equation*}
-  Using Lemma \ref{lemma:MDL} we conclude
-  \begin{align*}
-    \tar( \obs ) - \tar( \obs )
-    &= \frac12 \log \det \left (I +\M(\obs) \right ) / \det \left (I + \M(\obsm) \right ) \\
-    %
-    %
-    %
-    &= \frac12 \log \det \left (I +\M(\obs) \right ) / \det \left (I + \M(\obs) - uu^* \right ) \\
-    %
-    %
-    %
-    &=-\frac12 \log (1 - \langle (I+\M(\obs))^{-1}u, u \rangle \\
-    %
-    %
-    %
-    &= -\frac12 \log (1 - \sigma^{-2} \langle \fwd \postcov \fwd^* \meas_m, \meas_m \rangle ).
-  \end{align*}
-\end{proof}
-
-\samemeas
-\begin{proof} \label{cor:same meas proof}
-  Denote $A:= \obs \modcov \obs^*$ and $v_j$ the $j$th column of $A$.
-  Note that $v_j = \obsm \modcov \meas_m$, since $(\obsm \modcov
-  \obsm^*)_{ij} = \meas_i(\modcov \meas_j)$, as explained in
-  \eqref{eq:modcov explained}. One can now verify that
-  \begin{equation}\label{eq:observation}
-    \Sigmam^{-1} \obsm \modcov \meas_m = \Sigmam^{-1}v_j = (A +\sigma^2I_{m-1})^{-1} v_j =
-    e_j -\sigma^2 \Sigmam^{-1}e_j.
-  \end{equation}
-  %
-  Using \eqref{eq:observation}:
-  \begin{align}\label{eq:denominator}
-    \begin{split}
-      \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m
-      &= \meas_m \modcov \obsm^* ( e^j - \sigma^2 \Sigmam^{-1} e_j )\\
-      %
-      %
-      %
-      &= \meas_m \modcov \meas_j - \sigma^2 \meas_m \modcov \obsm^* \Sigmam^{-1}e_j \\
-      %
-      %
-      %
-      &= \meas_m \modcov \meas_j -\sigma^2 (e_j - \sigma^2 \Sigmam^{-1}e_j)^t e_j \\
-      %
-      %
-      %
-      &= \meas_m \modcov \meas_m -\sigma^2 + \sigma^4 e_j^t\Sigmam^{-1}e_j.
-    \end{split}
-  \end{align}
-  We use \eqref{eq:observation} to simplify the terms in the enumerator of
-  the conclusion of Lemma \ref{lemma:design increase}:
-  \begin{align}\label{eq:enumerator}
-    \begin{split}
-      (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m
-      &= \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m - \meas_m \\
-      %
-      %
-      %
-      &= \obsm^* (e_j - \sigma^2 \Sigmam^{-1} e_j) -\meas_j \\ 
-      %
-      %
-      %
-      &= -\sigma^2 \obsm^* \Sigma^{-1}e_j. 
-    \end{split}
-  \end{align}
-  %
-  Substitute \eqref{eq:enumerator} and \eqref{eq:denominator} to
-  the enumerator and denominator (respectively) of the conclusion of
-  Lemma \ref{lemma:design increase}:
-  %
-  \begin{align*}
-    \tar( \obs ) - \tar (\obsm ) &=
-    \log \left ( 1 + \frac{
-      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m,
-      (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m \rangle
-    }{
-      \sigma^2 + \meas_m \modcov \meas_m - \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m 
-    }       
-    \right ) \\
-    %
-    %
-    %
-    &= \log \left ( 1 + \frac{\sigma^4
-      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
-      \obsm^* \Sigmam^{-1}e_j \rangle
-    }{
-      2\sigma^2 - \sigma^4 e_j^t\Sigmam^{-1}e_j 
-    }       
-    \right ) \\
-    %
-    %
-    %
-    &= \log \left ( 1 + \frac{\sigma^2
-      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
-      \obsm^* \Sigmam^{-1}e_j \rangle
-    }{
-      2 - \sigma^2 e_j^t\Sigmam^{-1}e_j 
-    }       
-    \right ).
-  \end{align*}
-\end{proof}
-
-\aux
-\begin{proof}
-  We need a few supplementary calculations. First:
-  \begin{align}\label{eq:der sig}
-    \begin{split}
-      \frac{\der}{\der \tau} \Big |_{\tau=0} \Sigma( \obs + \tau V )
-      &= \frac{\der}{\der \tau} \Big |_{\tau=0} 
-      (\obs + \tau V ) \modcov (\obs + \tau V )^*  + \sigma^2I\\
-      % 
-      % 
-      % 
-      &= V \modcov \obs^* + \obs \modcov V^*.
-    \end{split}
-  \end{align}
-  By the standard trick for the derivative of an operator: 
-  \begin{align*}
-    0 &= \frac{\der}{\der \tau} \Big |_{\tau=0} I \\
-    % 
-    % 
-    % 
-    &= \frac{\der}{\der \tau} \Big |_{\tau=0}
-    \left (\Sigma(\obs+\tau V)^{-1} \Sigma(\obs+\tau V) \right ) \\
-    % 
-    % 
-    % 
-    &= \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0} \Sigma+
-    \Sigma^{-1} \frac{\der \Sigma(\obs+\tau V)}{\der \tau} \Big |_{\tau=0}\\  
-    %
-    %
-    %
-    &= \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0} \Sigma+
-    \Sigma^{-1} (V\modcov \obs^* + \obs \modcov V^*) 
-    \text{, by \eqref{eq:der sig}. }
-  \end{align*}
-  Now we can write the variation of $\Sigma^{-1}$:
-  \begin{align}\label{eq:der sig inv}
-    \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0}  
-      &= -\Sigma^{-1} (V \modcov \obs^* + \obs \modcov V^*) \Sigma^{-1}.
-    \end{align}
-  % 
-  Finally, we can do the calculation we need to. By Leibnitz (product) rule
-  and \eqref{eq:der sig inv}:
-  % 
-  \begin{align*}
-    \delta T(\obs) V 
-    &= \frac{\der T(\obs + \tau V)}{\der \tau} \Big |_{\tau=0} \\
-    %
-    %
-    %
-    &= V^* \Sigma^{-1} \obs 
-    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
-    &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs
-    + \obs^* \Sigma^{-1} V.
-  \end{align*}
-\end{proof}
-
-\biggerbetter
-\begin{proof} 
-  Fix an arbitrary $j=1,\dots,m$ and take $V:= e_j e_j^t \obs$. We see
-  that for $u \in \hilo$
-  \begin{equation*}
-    Vu = e_je_j^t (\meas_1(u),\dots,\meas_m(u) )^t = e_j \meas_j(u)
-    = (0,\dots,0,\meas_j(u),0,\dots,0)^t.
-  \end{equation*}
-  This way, $V$ has the same $j$th entry as $\obs$ while the rest are
-  set to zero. We calculate the variation in this direction and show
-  it is positive --- $\delta \tar(\obs)V > 0$. This means that
-  increasing the magnitude of the $j$th observation functional
-  increases $\tar(\obs)$. We start with the last line of \eqref{eq:tar var}, and denote $\tmp : = \fwd \postcov \fwd^*$.
-  \begin{align*}
-     \delta \tar(\obs) V 
-    &= \tr{V ( I - \modcov \obs^*\Sigma^{-1}\obs) \tmp \obs^* \Sigma^{-1}} \\
-    % 
-    %
-    %
-    &= \tr{e_je_j^t \obs ( I - \modcov \obs^*\Sigma^{-1}\obs) \tmp \obs^* \Sigma^{-1}} \\
-    %
-    % 
-    %
-    &= e_j^t \obs ( I - \modcov \obs^*\Sigma^{-1}\obs) \tmp \obs^* \Sigma^{-1}e_j \\
-    %
-    % 
-    %
-    &= e_j^t ( I - \obs \modcov \obs^*\Sigma^{-1})\obs \tmp \obs^* \Sigma^{-1}e_j \\  
-    % 
-    %
-    %
-    &=  e_j^t(\Sigma-\obs \modcov \obs^*) \Sigma^{-1}\obs \tmp \obs^* \Sigma^{-1}e_j \\
-    %
-    %
-    %
-    &=\sigma^2 e_j^t \Sigma^{-1}\obs \tmp \obs^* \Sigma^{-1}e_j
-    \text{ by \eqref{eq:Sigma} }\\
-    %
-    % 
-    %
-    &=\sigma^2 e_j^t \Sigma^{-1}\obs \fwd \postcov \fwd^* \obs^* \Sigma^{-1}e_j.
-  \end{align*} 
-  Since $\postcov$ is positive definite, $\delta \tar(\obs) V > 0$.
-\end{proof}
+\input{vanishing}
+\input{appendix}
 

### 2023-03-16 latex/vanishing.tex
Finished de proof

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% SECTION Analysis of Optimal Designs --- Vanishing Model Error
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Analysis of D-Optimal Designs --- Vanishing Model Error}\label{section:vanishing}
\subsection{Overview}
In this section we analyze D-optimal designs when no model error is
present. Sensor clusterization is discussed in section
\ref{section:clusterization}. We first derive a nonlinear eigenvalue
problem in $\obs$ for D-optimal designs in Section
\ref{subsec:eigenproblem}. In Section \ref{subsec:characterization} we
characterize D-optimal designs as ones that make all eigenvalues equal
in the nonlinear eigenvalue problem. Results are summarized in Theorem
\ref{thm:char}.

\subsection{The Nonlinear Eigenvalue Problem}\label{subsec:eigenproblem}
$\modcov = 0$ implies $\Sigma= \sigma^2I$. The necessary first-order
condition for D-optimality \eqref{eq:conditions} becomes
\begin{equation}\label{eq:eigenproblem}
  \sigma^{-2}\fwd \postcov \fwd^* \obs^* = \obs^* \Xi,
\end{equation}
with $\Xi$ diagonal. We think of \eqref{eq:eigenproblem} as an
eigenvalue problem for the self-adjoint operator $\sigma^{-2}\fwd
\postcov \fwd^*$, where rows of $\obs$, namely $\meas_j,j=1,\dots, m$,
are eigenvectors. However, that $\postcov$ depends on $\obs$ and thus,
we refer to \eqref{eq:eigenproblem} as a nonlinear eigenvalue problem.

The following two lemmas allow us to gain better understanding of the
nonlinear eigenvalue problem.

\begin{restatable*}[Two applications of Woodbury Matrix Identity]{lemma}{woodbury}\label{lemma:twice woodbury}
  Assume $\fwd \prcov \fwd^*$ is invertible. Then
  \begin{align*}
    \begin{split}
      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* 
      %
      %
      = \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1},
    \end{split}
  \end{align*}  
\end{restatable*}

%% The term $(\fwd \prcov \fwd^*)^{-1}$ is the prior precision in
%% $\hilo$, and $\fwd\prcov\fwd^*$ is invertible, as mentioned in
%% Section \ref{section:prelim}. $(\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}
%% \obs^* \obs$ is the posterior precision in $\hilo$.

%% \subsection{Characterizing D-Optimal Designs}\label{subsec:characterization}
%% %% Denote corresponding eigenvectors $\{\ev_i\}_{i=1}^{\infty}$.
%% Denote eigenvalues of $\fwd \prcov \fwd^*$ by
%% $\{\lambda_i\}_{i=1}^{\infty}$, and note they are positive. Lemma
%% \ref{lemma:sim diag} is crucial for the rest of the analysis:

\begin{restatable*}[Simultaneous diagonizability for the nonlinear eigenvalue problem]{lemma}{simdiag}\label{lemma:sim diag}
  Let $\hil$ separable Hilbert space, $C:\hil \to \hil$ self-adjoint
  and $\func_1,\dots,\func_m \in \hil$. Denote $\func^*$ the element
  $\func$ acting as a linear functional. If
  \begin{equation*}
   (C + \sum_{j=1}^m \func_j\func_j^*) \func_l = \xi_l \func_l,\ l = 1,\dots,m
  \end{equation*}
  then $C$ and $\sum_{j=1}^m \func_j \func_j^*$ are simultaneously
  diagonalizable.
\end{restatable*}

\begin{corollary}[Eigenvectors of a D-optimal design]\label{cor:same ev}
  Let $\obs$ satisfy the nonlinear eigenvalue problem
  \eqref{eq:eigenproblem}. Then $\obs^*\obs$ has the same eigenvectors
  as $\fwd \prcov \fwd^*$.
\end{corollary}
\begin{proof}
  \begin{align}\label{eq:mod conditions}
    \begin{split}
      \obs^* \Xi &= \sigma^{-2}\fwd \postcov \fwd^* \obs^*  \text{ (by \eqref{eq:eigenproblem})}\\
      %
      %
      %
      &= \sigma^{-2} \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* \obs^*  \text{ (by \eqref{eq:postcov})} \\
      %
      %
      %
      &= \sigma^{-2} \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1} \obs^* \text{ (by Lemma \ref{lemma:twice woodbury})}.
    \end{split}
  \end{align}

  Now take $\func_j^{*} = \meas_j$ and $C := (\fwd \prcov
  \fwd^*)^{-1}$ and use Lemma \ref{lemma:sim diag}.
\end{proof}

%% Corollary \ref{cor:same ev} is important because it allows us to only think

implies we only need to consider $\obs$ such that  
At most $m$ of the eigenvectors of $\obs^*\obs$ have a positive
eigenvalue, because $k := \rank \obs^*\obs \leq m$. Since we made no
assumption regarding the ordering of $\{\lambda_i\}$, we can denote
the corresponding non-zero eigenvalues of $\obs^*\obs$ by
$\{\eta_i\}_{i=1}^{k}$ and let $\eta_i = 0$ for $i \geq k+1$.

%% \begin{definition}
%%   Let $\{\lambda_i\}_{i=1}^{\infty}$ eigenvalues of $\fwd \prcov
%%   \fwd^*$ sorted in descending order. Let $\eta \in
%%   \mathbb{R}_+^m$. The \emph{utility proxy} $\bar{\tar}$ is defined
%%   as:
%%   \begin{equation*}
%%     \bar{\tar}(\eta) := \frac12 \sum_{i=1}^m \log(1 + \sigma^{-2}\lambda_i\eta_i).
%%   \end{equation*}
%% \end{definition}

\begin{corollary}\label{cor:true target}
  Let $\obs$ with $m$ observations satisfy the nonlinear eigenvalue
  problem \eqref{eq:eigenproblem}. Let $\{\eta_i\}_{i=1}^{\infty}$
  eigenvalues of $\obs^*\obs$ and $\{\lambda_i\}_{i=1}^{\infty}$ the
  corresponding eigenvalues of $\fwd \prcov \fwd^*$. Let $k:=\rank
  \obs^*\obs$. Without loss of generality, let $\eta_i > 0$ for $i\leq
  k$ and $\eta_i = 0$ for $i > k$. Then:
  \begin{enumerate}
    \item $k \leq m$ and $\obs^*\obs$ has exactly $k$ positive
      eigenvalues.
    \item
      \begin{equation*}
        \tar(\obs) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i) = \frac12 \sum_{i=1}^{m} \log (1 + \sigma^{-2}\lambda_i\eta_i).
      \end{equation*}
    \item Let $\obs$ a D-optimal design. Then $\eta_i > 0$ for
      eigenvectors corresponding to the $k$ largest $\lambda_i$.
  \end{enumerate}
\end{corollary}
\begin{proof}
  Part (1) is trivial. To see part (2) holds: 
  \begin{align}
    \begin{split}
      \tar(\obs) &= \frac12\log \det \left (I + \sigma^{-2} \prcov^{1/2} \fwd ^* \obs^*
      \obs \fwd \prcov^{1/2}\right )\\% \text{ (by definition)}\\
      %
      &= \frac12 \log \det \left (I + \sigma^{-2} \obs^* \obs \fwd
      \prcov\fwd^* \right ) \text{ (Sylvester's Determinant
      Theorem)}\\
      %
      %
      %
      &=\frac12 \log \prod_{i=1}^{\infty} ( 1 + \sigma^{-2} \lambda_i\eta_i ) \text{ (Corollary \ref{cor:same ev})} \\
      %
      %
      %
      %&=\frac12 \log \left ( \prod_{i=1}^{k} ( \lambda_i^{-1} + \sigma^{-2} \eta_i )\prod_{i=1}^{k} \lambda_i \right )
      %
      %
      %
      &=\frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i). 
      %
      %
      %% &=\frac12 \sum_{i=1}^{k} \log(\sigma^2\lambda_i^{-1} + \eta_i) - k\log \sigma + \frac12\sum_{i=1}^k \log \lambda_i.
    \end{split}
  \end{align}
  Part (3) holds since $\log$ is increasing and $\eta_i \geq 0$.
\end{proof}

\begin{proposition}\label{prop:kkt}
  Let $\tar: \mathbb{R}^m \to \mathbb{R}$, $\psi(\eta) =
  \frac{1}{2}\sum_{i=1}^m \log (1+\sigma^{-2}\lambda_i \eta_i)$, with
  $\lambda_i > 0$ and $\sigma^{2} > 0$. Then the maximum of $\tar$
  subject to $\eta_i \geq 0$ and $\sum\eta_i = m$ is obtained at
  \begin{equation}
  \eta_i = \begin{cases}
    \frac{m}{k} - \sigma^2 \lambda_i^{-1} + \sigma^2 \frac{1}{k} \sum_{j\in A} \lambda_j^{-1} & i \in A \\
    0 & i \in A^c
  \end{cases}
  \end{equation}
  where $A:= \{1\leq i \leq m: \eta_i > 0\}$ and $A^c = \{1,\dots, m\}
    \backslash A$, and $|A|$ is the cardinality of $A$.
\end{proposition}
\begin{proof}
%% \begin{proposition}\label{prop:optimizer}
%%   Let $\{\lambda_i\}_{i=1}^{\infty}$ the sorted eigenvectors of $\fwd
%%   \prcov\fwd^*$, $k\in \mathbb{N}^+$ and
%%   %% Let
%%   %%   \begin{equation}
%%   %%   k = \argmax \left \{ k:\lambda_k^{-1} \leq \sigma^{-2}\frac{m}{k} + \frac{1}{k} \sum_{j=1}^{k}
%%   %%   \lambda_j^{-1} \right \}.
%%   %% \end{equation}

%%   \begin{equation}
%%   \eta_i = \begin{cases}
%%     \frac{m}{k} - \sigma^2 \lambda_i^{-1} + \sigma^2 \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1} & i \leq k \\
%%     0 & i > k 
%%   \end{cases}.
%%   \end{equation}

%%   Let $\obs$ an observation operator such that $\obs^*\obs$ has the
%%   same eigenvectors as $\fwd \prcov \fwd^*$. If $\eta_i > 0$ for
%%   $i\leq k$ and $\obs^*\obs$ has eigenvalues $\eta_i$, then $\obs$
%%   maximizes $\tar$ subject to $\sum_{i=1}^m \eta_i = m$ and to $\rank
%%   \obs = k$.
%% \end{proposition}
  %% \begin{proof}
  %% et $A := \{i: \eta_i > 0\}$. A choice of $A$ where $\eta_i =0,
  %% \eta_j >0$ for some $1 \leq i < j \leq m$ is not optimal (just
  %% switch $\eta_j$ and $\eta_i$ and recall $\{\lambda_i\}$ are arranged
  %% in descending order). Thus, since $\rank \obs = k$ we conclude that,
  %% $\eta_i=0, i>k$.

  Let $\Phi(\eta) = \sum_{i=1}^k \eta_i - m$ and $\Omega_j(\eta) =
   -\eta_j$. Then
  \begin{align*}
    \begin{split}
      \frac{\partial \tar}{\partial \eta_i}  &=
       \frac12 \frac{\sigma^{-2}\lambda_i}{1 + \sigma^{-2} \lambda_i\eta_i} = \frac12 \frac{1}{\sigma^{2}\lambda_i^{-1} + \eta_i} \\
      %
      %
      %
      \frac{\partial\Phi}{\partial \eta_i} &= 1 \\
      %
      %
      %
      \frac{\partial \Omega_j}{\partial \eta_i} &= -\delta_{ij}      
    \end{split}
  \end{align*}

  From the KKT conditions, there are $\alpha, \beta_i$ such that for $i=1,\dots,m$:
  \begin{align}
    \begin{split}
      -\frac12 \frac{1}{\sigma^{2}\lambda_i^{-1} + \eta_i} + \alpha - \beta_i  &= 0 \\
      %
      %
      %
      \eta_i &\geq 0\\
      %
      %
      %
      \beta_i &\geq 0\\
      %
      %
      %
      \beta_i \eta_i &= 0\\
      %
      %
      %
      \sum_{i=1}^m \eta_i &= m 
    \end{split}
  \end{align}

  Then, for $i \in A$:
  \begin{align*}
    \begin{split}
      \beta_i &= 0\\
      %
      %
      %
      \sigma^{2}\lambda_i^{-1} + \eta_i  &= \frac{1}{2\alpha}%  \text{ for }  1 \leq i \leq k.\\
      %
      %
      %
      %% \sigma^{-2} \lambda_i &= 2\alpha - 2\beta_i, i \in B \\
    \end{split}
  \end{align*}

  Summing over $i \in A$, substituting $\sum_{i\in A} \eta_i =
  \sum_{i=1}^m \eta_i = m$, and dividng by $k:= |A|$:
  $$
  \frac{1}{2\alpha} = \frac{m}{k} + \frac{\sigma^2}{k} \sum_{j\in A} \lambda_j^{-1}.
  %\Rightarrow \alpha = \left ( \frac{2m}{k} + \frac{2\sigma^2}{k}
  %\sum_{i\in A} \lambda_i^{-1} \right )^{-1}.
  $$

  Consequently:
  \begin{align} \label{eq:etas}
    \begin{split}
      \eta_i &= \frac{1}{2\alpha} - \sigma^2 \lambda_i^{-1}\\
      %
      %
      %
      &=\frac{m}{k} + \frac{\sigma^2}{k} \sum_{j\in A} \lambda_j^{-1} - \sigma^2 \lambda_i^{-1}.%\\
      %
      %
      %
      %% &= \frac{m}{k} + \frac{\sigma^2}{k} \sum_{j \in A, j\neq i} \lambda_j^{-1} - \frac{k-1}{k}\sigma^2 \lambda_i^{-1}\\
      %% ,\ i \in A.
    \end{split}
  \end{align}
\end{proof}

\begin{corollary}\label{cor:descending}
  In Proposition \ref{prop:kkt}, if $\lambda_i$ are arranged in
  descending order, then $A= \{1,\dots,k\}$ for some $1 \leq k \leq
  m$.
\end{corollary}

\begin{corollary}\label{cor:eigenvalues}
  If $\obs$ is optimal and $k := \rank \obs$, then:
  \begin{enumerate}
    \item The covariance of the pushforwad $\fwd_{*} \post$ is $((\fwd
      \prcov \fwd^*)^{-1} + \sigma^{-2} \obs^*\obs)^{-1}$.
    \item The eigenvalues of said operator are
      \begin{equation}
        \eta_i = \begin{cases}
          \frac{m}{k} + \sigma^2 \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1} & i \leq k \\
          \lambda_i^{-1} & i > k 
        \end{cases}.
      \end{equation}
  \end{enumerate}
\end{corollary}

\begin{proof}
  Recall from \eqref{eq:postcov}, that the posterior covariance is
  $\postcov^{-1} = \prcov^{-1} +
  \sigma^{-2}\fwd^*\obs^*\obs\fwd$. Part (1) now follows from Lemma
  \ref{lemma:twice woodbury}. Part (2) follows from Propositin
  \ref{prop:kkt} and Corollary \ref{cor:descending}.
\end{proof}

We are almost ready to state and prove Theorem \ref{thm:char} that
characterizes D-optimal designs. We just need one final lemma, whose
proof is delegated to the appendix:
\begin{restatable*}[Unit norm decomposition]{lemma}{free}\label{lemma:free}
  Let $M \in \R^{k \times k}$ symmetric positive definite with $\ttr M
  = m$, $m \geq k$. We can find $\func_j \in \R^k,j=1,\dots,m$
  with $\|\func_j\|=1$ and $A = (\func_1,\dots,\func_m)$ such that
  $AA^t = M$.
\end{restatable*}

\main

\begin{proof}
  Part (1) was proved in Corollary \ref{cor:same ev}. Parts (2) and
  (3) were proved in Corollary \ref{cor:true target}. Part (4) is a
  consequence Corollariy \ref{cor:eigenvalues}, with the caveat that
  we did not show that finding $\obs$ so that $\obs^*\obs$ has the
  desired eigenvalues is feasible. To this end, we utilize Lemma
  \ref{lemma:free}: let $M = \diag(\eta_1, \dots, \eta_k)$, diagonal
  with respect to the first $k$ eigenvectors of $\fwd \prcov
  \fwd^*$. We take $\obs := A$ from the Lemma \ref{lemma:free}, and
  this concludes the proof.
\end{proof}

\begin{corollary}
  In a D-optimal design, the posterior precision for eigenvector $i$
  is $\eta_i + \sigma^2 \lambda_i^{-1} = \frac{m}{k} + \sigma^2
  \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1}$. I.e. the posterior
  precision in $\hil_p$ is constant for $1 \leq i \leq k$, where $k
  =\rank \obs$.
\end{corollary}

\pgfplotstableread{
  Label     prior  optimal  sub-optimal 
  1         0.2    1.8           1.7
  2         0.8    1.2           0.8
  3         2.2    0             0.5
  4         3.5    0             0.0
}\optimalvsnot

\begin{figure}
  \begin{tikzpicture}[scale=0.85]
    \begin{axis}[
        ybar stacked,
        ymin=0,
        ymax=4,
        xtick=data,
        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
        reverse legend=false, % set to false to get correct display, but I'd like to have this true
        xticklabels from table={\optimalvsnot}{Label},
        xticklabel style={text width=2cm,align=center},
        legend plot pos=right,
        ylabel=precision --- prior and posterior,
        xlabel=eigenvector $i$,
      ]

      \addplot [fill=green!80]  table [y=prior,   meta=Label, x expr=\coordindex] {\optimalvsnot};
      \addplot [fill=blue!60]   table [y=optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};

      \addlegendentry{$\sigma^2\lambda_i^{-1}$}
      \addlegendentry{optimal $\eta_i$s}

    \end{axis}
  \end{tikzpicture}
  \qquad
  \begin{tikzpicture}[scale=0.85]
    \begin{axis}[
        ybar stacked,
        ymin=0,
        ymax=4,
        xtick=data,
        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
        reverse legend=false, % set to false to get correct display, but I'd like to have this true
        xticklabels from table={\optimalvsnot}{Label},
        xticklabel style={text width=2cm,align=center},
        legend plot pos=right,
        ylabel=precision --- prior and posterior,
        xlabel=eigenvector $i$,
      ]

      \addplot [fill=green!80]  table [y=prior,       meta=Label, x expr=\coordindex] {\optimalvsnot};
      \addplot [fill=blue!60]   table [y=sub-optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};

      \addlegendentry{$\sigma^2\lambda_i^{-1}$}
      \addlegendentry{sub-optimal $\eta_i$s}

    \end{axis}
  \end{tikzpicture}
  \caption{As an illustration of Corollary \ref{cor:equal eigs}, a
    D-optimal design is compared with a sub-optimal design. Both
    designs are allowed three observations. Thus, the blue area has
    accumulated height of $3$ in both plots. A D-optimal design (left)
    increases precision (reduces uncertainty) where it is lowest. A
    sub-optimal design (right) does not. In this example, an
    observation is taken partly in the direction of the third
    eigenvector. The precision for this eigenvector is bigger even
    than the posterior precision of the D-optimal design on the left.}
  \label{fig:optimal vs not}
\end{figure}

\pgfplotstableread{
  Label    prior  $o_1$  $o_2$   $o_3$   $o_4$   $o_5$         $o_6$    topper
  1         0.2    0.8    0.5     0.5    0.4    0.33333333  0.3333333  0.001
  2         0.8    0.2    0.5     0.5    0.4    0.33333333  0.3333333  0.001
  3         2.2    0      0       0      0.2    0.33333333  0.3333333  0.001
  4         3.5    0      0       0      0      0           0          0.001
}\clusterization

\pgfplotstableread{
  Label    prior  $o_1$  $o_2$   $o_3$   $o_4$   $o_5$         $o_6$    topper
  1         0.2    0.8    0.7     0.3    0.4    0.2         0.4666666  0.001
  2         0.8    0.2    0.3     0.7    0.4    0.46666666  0.2        0.001
  3         2.2    0      0       0      0.2    0.33333333  0.3333333  0.001
  4         3.5    0      0       0      0      0           0          0.001
}\noclusterization

Here is a way to understand D-optimal designs with vanishing model
error: Imagine each eigenvector corresponds to a graduated lab
cylinder (Figure \ref{fig:cylinder}). Each cylinder $i$ is filled,
a-priori, with green liquid of $\sigma^2\lambda_i^{-1}$ volume
units. This liquid represnets the prior precision, weighted by
observation noise. The volume of the green liquid is increasing in
$i$. A D-optimal design with $m$ observations ignores cylinders
indexed $m+1$ and above. It has blue liquid of volume $m$ units at its
disposal. It distributes said liquid by repeatedly adding a drop to
whatever cylinder currently has the lowest level of liquid in it
(lowest precision). The result of such a procedure is illustrated in
Figure \ref{fig:optimal vs not}.

\begin{figure}%{r}{0.25\textwidth} 
    %\centering
    \includegraphics[width=5cm, height=5cm]{cylinders.jpg}
    \caption{Three graduated lab cylinders, corresponding to three
      eigenvectors. Prior eigenvalues not shown.}
    \label{fig:cylinder}
\end{figure}

\section{Sensor Clusterization --- Vanishing Model Error}\label{section:clusterization}
\subsection{Overview}
In this section we use Theorem \ref{thm:char} to explain how sensor
clusterization may arise in various scenarios. We first consider
sequential designs, and then simultaneous designs.

\subsection{Sequential Design}\label{subsec:clusterization sequential}
In a sequential optimal design problem, each measurement is taken
after results of previous measurements are recorded. In each step
$m=1$, so $k=1$. In this case, Theorem \ref{thm:char} implies each
measurement is taken along the direction of the eigenvector of $\fwd
\prcov \fwd^*$ with lowest posterior precision. If $\lambda_i$
decrease fast enough (e.g. exponentially), the

, $\meas_1 \parallel \ev_1$ and $\eta_1 = 1$. As hinted in
section \ref{subsec:seq vs sim}, the posterior becomes the prior for
the next step. The eigenvectors of the new posterior are the same as
for the previous one. A new observation will be taken in the direction
of eigenvector with smalles eigenvalue. This leads to:
\begin{align*}
  \begin{split}
    \meas_2 \parallel \ev_1  &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 < \sigma^2\lambda^{-1}_2 \\
    \meas_2 \parallel \ev_2  &\text{ o.w. }
  \end{split}
\end{align*}
In the first case, we get clusterization immediately. In the second,
we consider $\meas_3$. Same reasoning leads to the conclusion that the
eigenvectors of the new prior precision are the same. The eigenvalues,
however, are $\lambda^{-1}_1 + \sigma^{-2}, \lambda^{-1}_2 +
\sigma^{-2}, \lambda_3^{-1}, \dots$. A new observation is taken in the
direction of eigenvector of smallest precision (note that
$\lambda^{-1}_1 + \sigma^{-2} < \lambda^{-1}_2 + \sigma^{-2}$):
\begin{align*}
  \begin{split}
    \meas_3 \parallel \ev_1 &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 <
    \sigma^2\lambda^{-1}_3 \\
    \meas_3 \parallel  \ev_3  &\text{ o.w. }
  \end{split}
\end{align*}
Again, the first case results in clusterization. The sequential design
proceeds in this fashion. Every $\meas_i$ is taken in the direction
some eigenvector of the prior $\fwd \prcov\fwd^*$. This eigenvector
has the smallest eigenvalue of the previous step's posterior precision
operator. If we do not encounter clusterization after $m$ such steps,
this means that observation $\meas_k$ was taken in the direction
$\ev_k$ for $k=1,\dots,m$. Thus:
\begin{align}\label{eq:failure}
  \begin{split}
    \sigma^2\lambda_{k-1}^{-1} + 1 \geq \sigma^2\lambda_k^{-1}, \text{ for } 1\leq k \leq m
    \Longrightarrow \lambda_m^{-1} - \lambda_{m-1}^{-1} \leq \sigma^{-2}.
  \end{split}
\end{align}

But \eqref{eq:failure} implies $\{\lambda_i^{-1}\}_{i=1}^{\infty}$
grows at most linearly. In section \ref{subsec:abstract OED} we
assumed $\fwd$ is strongly smoothing, which means eigenvalues of
$(\fwd \prcov\fwd^*)^{-1}$ grow quickly. Thus, if we take $m$ large
enough, \eqref{eq:failure} will eventually fail and we will observe
sensor clusterization.

\subsection{Simultaneous Design}\label{subsec:clusterization simultaneous}
In a simultaneous design scenario, the result is not simple as in the
sequential design case. We now see that a design that exhibits
clusterization is as optimal as one that does not. Corollary
\ref{cor:same ev} tells us that we can realize any design we want
(encoded in $\{\eta_i\}$), given the trace constraint for
$\obs^*\obs$. Theorem \ref{thm:char} tells us that for D-optimality,
we only need to make the posterior flat where we take observations
(recall figure \ref{fig:optimal vs not}). We are left with the freedom
of assigning any part of $\eta_i$ to any observation $\meas_j$. Two
possible assignments for the same problem are illustrated in figure
\ref{fig:clusterization}. The first exhibits clusterization, the
second does not. It is important to note that neither arises as a
result of some numerical optimization scheme. Both designs give the
same design criterion $\tar$, because their posteriors are
identical. By Theorem \ref{thm:char}, both are maximizers of the same
D-optimal design problem. One exhibits clusterization, and one does
not. It is likely that a real-life problem is more restricted than the
model we consider here. Such restriction is the likely cause of
clusterization obseved in, e.g., the toy model of section
\ref{subsec:example}. Exactly how this happens is not clear. At this
point it is the subject of speculation and should be further studied
in the future.

\begin{figure}
  \begin{tikzpicture}[scale=0.85]
    \begin{axis}[
        ybar stacked,
        ymin=0,
        ymax=4,
        xtick=data,
        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
        reverse legend=false, % set to false to get correct display, but I'd like to have this true
        xticklabels from table={\clusterization}{Label},
        xticklabel style={text width=2cm,align=center},
        legend plot pos=right,
        ylabel=precision --- prior and posterior,
        xlabel=eigenvector $i$,
      ]

      \addplot [fill=green!80]  table [y=prior, meta=Label, x expr=\coordindex] {\clusterization};
      \addplot [fill=blue!60]   table [y=$o_1$, meta=Label, x expr=\coordindex] {\clusterization};
      \addplot [fill=red!60]    table [y=$o_2$, meta=Label, x expr=\coordindex] {\clusterization};
      \addplot [fill=black!60]  table [y=$o_3$, meta=Label, x expr=\coordindex] {\clusterization};
      \addplot [fill=orange!60] table [y=$o_4$, meta=Label, x expr=\coordindex] {\clusterization};
      \addplot [fill=cyan!60]   table [y=$o_5$, meta=Label, x expr=\coordindex] {\clusterization};
      \addplot [fill=purple!60] table [y=$o_6$, meta=Label, x expr=\coordindex] {\clusterization};

      \addlegendentry{prior}
      \addlegendentry{$o_1$}
      \addlegendentry{$o_2$}
      \addlegendentry{$o_3$}
      \addlegendentry{$o_4$}
      \addlegendentry{$o_5$}
      \addlegendentry{$o_6$}   
    \end{axis}
  \end{tikzpicture}
  \qquad
  \begin{tikzpicture}[scale=0.85]
    \begin{axis}[
        ybar stacked,
        ymin=0,
        ymax=4,
        xtick=data,
        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
        reverse legend=false, % set to false to get correct display, but I'd like to have this true
        xticklabels from table={\noclusterization}{Label},
        xticklabel style={text width=2cm,align=center},
        legend plot pos=right,
        ylabel=precision --- prior and posterior,
        xlabel=eigenvector $i$,
      ]

      \addplot [fill=green!80]  table [y=prior, meta=Label, x expr=\coordindex] {\noclusterization};
      \addplot [fill=blue!60]   table [y=$o_1$, meta=Label, x expr=\coordindex] {\noclusterization};
      \addplot [fill=red!60]    table [y=$o_2$, meta=Label, x expr=\coordindex] {\noclusterization};
      \addplot [fill=black!60]  table [y=$o_3$, meta=Label, x expr=\coordindex] {\noclusterization};
      \addplot [fill=orange!60] table [y=$o_4$, meta=Label, x expr=\coordindex] {\noclusterization};
      \addplot [fill=cyan!60]   table [y=$o_5$, meta=Label, x expr=\coordindex] {\noclusterization};
      \addplot [fill=purple!60] table [y=$o_6$, meta=Label, x expr=\coordindex] {\noclusterization};

      \addlegendentry{prior}
      \addlegendentry{$o_1$}
      \addlegendentry{$o_2$}
      \addlegendentry{$o_3$}
      \addlegendentry{$o_4$}
      \addlegendentry{$o_5$}
      \addlegendentry{$o_6$}   
    \end{axis}
  \end{tikzpicture}
  \caption{Clusterization (left) and non-clusterization (right) in
    simultaneous D-optimal designs. Posterior precision per
    eigenvector after taking observations $\meas_i$ is plotted. Both
    designs are identical for all practical matters --- their
    posteriors are identical. In the left panel we see how an optimal
    design is achieved with repeated observations. $\meas_2 = \meas_3$
    and $\meas_5 = \meas_6$. This is not necessary though, as can be
    seen in the right panel. Both designs achieve the same levels of
    uncertainty. The left exhibits clusterization. The right does not.}
  \label{fig:clusterization}
\end{figure}

%% \free
%% \begin{proof}
%%   Let us diagonalize $M$, so that $M = U D U^t$ with $D =
%%   \diag(d_1,\dots,d_k)$ and $U \in \R^{k \times k }$ orthogonal. Let
%%   $S \in \R^{k \times m}$ with $S_{ii} = \sqrt{d_{i}}$ and zeros
%%   otherwise. Define $A:= U S V^t$, where $V \in \R^{m \times m}$ is
%%   orthogonal and will be further restricted later. Then $AA^t = U
%%   SV^tVS^t U^t = UDU^t$, so $AA^t$ has the required eigenvalues and
%%   eigenvectors by construction. If we can choose $V$ such that $A$
%%   also satisfies the unit norm constraints we are done. These
%%   constraints are, for $j=1,\dots,m$:
%%   \begin{equation}\label{eq:V constraints}
%%    1 = [A^tA]_{jj} = [V S^tS V^t]_{jj},
%%   \end{equation}
%%   and we can expect to do this since we assumed $\ttr D = m$.

%%   Define $C = S^tS - I \in \R^{m \times m}$. Note that $\ttr C = 0$ and
%%   $C$ is diagonal with non-zero entries $d_i-1,i=1,\dots,k$. It suffices
%%   to find $V$ orthogonal such that $V C V^t$ has zero diagonal. We
%%   construct such $V$ by sequentially inserting zeros in the diagonal
%%   and not destroying zeros we already introduced, starting from the
%%   last diagonal entry and moving to the first. Since $c_{mm} \neq 0$ ,
%%   let $p < m$ such that $c_{pp}c_{mm} < 0$ (such $p$ exists because
%%   the trace is zero) and let $\theta \in (0,\pi)$. Define a Givens
%%   rotation $R^{(m)} \in \R^{m \times m}$ by
%%   \begin{equation*}
%%     r^{(m)}_{ab} :=
%%     \begin{cases}
%%       1 & a = b \neq p \text{ or } a = b \neq m \\
%%       \cos \theta & a = b = p  \\
%%      -\sin \theta & a = p, b = m\\
%%       \cos \theta & a = b = m \\
%%       \sin \theta & a = m, b = p \\ 
%%       0 & o.w
%%     \end{cases}
%%   \end{equation*}
%%   Note that conjugating a matrix by $R^{(m)}$ changes only its $m$ and
%%   $p$ rows and columns. We want to choose $\theta$ such that
%%   \begin{equation}\label{eq:mm}
%%     0 = [R^{(m)} C (R^{(m)})^t]_{mm} = \cos^2 \theta c_{mm} + 2\cos \theta \sin
%%     \theta c_{mp} + \sin^2\theta c_{pp},
%%   \end{equation}
%%   and it suffices to choose $\theta$ such that
%%   \begin{equation*}
%%     c_{mm} \cot^2 \theta + 2 c_{mp} \cot \theta + c_{pp} = 0.
%%   \end{equation*}
%%   This quadratic in $\cot\theta$ has a real solution, since
%%   $c_{pp}c_{mm} < 0$ by assumption and we can find $\theta \in
%%   (0,\pi)$ such that \eqref{eq:mm} is satisfied. We continue to find
%%   $R^{(m-1)}$ that leaves row and column $m$ unchanged and
%%   continue introducing zeros to the diagonal. The assumption $\ttr D =
%%   m \Rightarrow \ttr C = 0$ guarantees we can do that. Taking $V:=
%%   R^{(1)} R^{(2)} \dots R^{(m-1)}R^{(m)}$ completes the proof.
%% \end{proof}

%% \simdiag
%% \begin{proof}
%%   First, enumerate the eigenvalues of $C + \sum_{j=1}^m
%%   \func_j\func_j^*$ as $\xi_1,\dots,\xi_\ell$. Denote the
%%   indices of the eigenvectors corresponding to $\xi_i$
%%   \begin{equation*}
%%     S_i := \{ 1 \leq k \leq m | (C + \sum_{j=1}^m \func_j\func_j^* )\func_k = \xi_i \func_k \}.
%%   \end{equation*}
%%   Define further
%%   \begin{equation*}
%%     A_i := \sum_{k \in S_i} \func_k \func_k^*,
%%   \end{equation*}
%%   which is self-adjoint. Two observations are in order. First,
%%   $\sum_{j=1}^m \func_j\func_j^* = \sum_{i=1}^\ell A_i$. Second, $A_i
%%   \func_k = 0$ if $k\not \in S_i$, since eigenvectors of different
%%   eigenvalue are orthogonal. For $k \in S_i$
%%   \begin{equation}\label{eq:on vi}
%%     \xi_i \func_k = (C + \sum_{j=1}^m \func_j \func_j^* ) \func_k = (C + A_i) \func_k.
%%   \end{equation}
%%   Let $V_i := span \{\func_k \}_{k\in S_i}$. Observe that $V_i$ is
%%   invariant under $A_i$, by definition, and under $C$, by \eqref{eq:on
%%     vi}. Using \eqref{eq:on vi} again, we conclude that on $V_i$, $A_i
%%   = \xi_iI - C$. This immediately implies $A_i$ and $C$ have the
%%   same eigenvectors on $V_i$. This holds for every $1 \leq i \leq
%%   \ell$ and we conclude that $C$ and $A$ have the same eigenvectors.
%% \end{proof}

### 2023-03-29 latex/intro.tex
Paper almost donegit status !

--- 2023-03-16 latex/paper.tex
@@
-\documentclass{amsart}
-\input{definitions}
-
-\numberwithin{equation}{section}
-
-\begin{document}
-
-\title[Sensor Clusterization in D-optimal design in infinite
-  dimensions]{Sensor Clusterization in D-optimal Design in Infinite
-  Dimensional Bayesian Linear Inverse Problems}
-
-\author{Yair Daon}
-\address{Porter School of the Environment and Earth
-  Sciences, Tel Aviv University\\ Tel Aviv, Israel}
-%% \curraddr{Porter School of Environment Studies, Tel Aviv University\\ Tel Aviv, Israel}
-%% \curraddr{}
-\email{yair.daon@gmail.com}
-%% \thanks{}
-
-%    The 2010 edition of the Mathematics Subject Classification is
-%    the current definitive version.
-\subjclass{Primary: 
-  62F15, % Statistics - bayesian inference
-  35R30, % PDE - Inverse problems
-  Secondary:
-  28C20, % Measure and integration - set functions and measures and integrals in
-  %infinite-dimensional spaces (Wiener measure, Gaussian measure, etc.
-}
-%% Primary: 62F15, 35R30; Secondary: 28C20
-\date{\today}
-
-\begin{abstract}
-  We investigate the problem of sensor clusterization in D-optimal
-  experimental design for infinite-dimensional Bayesian linear inverse
-  problems. We suggest an analytically tractable model for D-optimal
-  designs and reason how it may lead to sensor clusterization in the
-  case of iid observation noise. We also consider how spatially
-  correlated observation and model errors mitigate the clusterization
-  phenomenon.
-
-  %% As a part of the analysis we prove a Matrix Determinant
-  %% Lemma in Hilbert spaces and a lemma for calculating
-  %% $\frac{\der}{\der t} \log \det (I + X(t))$ for certain
-  %% operator-valued functions. We also show how to decompose $M = AA^t$
-  %% with $A$ constrained to have unit norm columns.
-\end{abstract}
-
-\maketitle
-
 \section{Introduction}\label{section:OED intro}
@@
 criterion}. This is true both for frequentists
-\cite{Silvey13,Ucinski05} as well as for Bayesians
-\cite{ChalonerVerdinelli95}. See \cite{ChalonerVerdinelli95} for an
-investigation of the analogy between the two approaches. Although
-there is a plethora of design criteria, we focus on just one of these,
-commonly referred to as \emph{D-optimal design}. It has a simple and
-appealing motivation in the Bayesian context as explained in
-\cite{ChalonerVerdinelli95}: Consider a linear model in finite
-dimensions, with Gaussian prior and noise. Under this model,
-maximizing the expected information gain (KL divergence
-\cite{KullbackLeibler51,CoverThomas12}) between posterior and prior
-amounts to minimizing the determinant of the posterior covariance
-matrix. In a frequentist setting, a D-optimal design minimizes the
-volume of the uncertainty ellipsoid \cite[page 16]{Ucinski05}, but
-this is done for the Fisher information matrix and not the posterior
-covariance. However, \cite{ChalonerVerdinelli95} shows that the latter
-is just a regularized version of the former.
+\cite{silvey2013,Ucinski05} as well as for Bayesians
+\cite{Chaloner1995}. See \cite{Chaloner1995} for an investigation of
+the analogy between the two approaches. Although there is a plethora
+of design criteria, we focus on just one of these, commonly referred
+to as \emph{D-optimal design}. It has a simple and appealing
+motivation in the Bayesian context as explained in
+\cite{Chaloner1995}: Consider a linear model in finite dimensions,
+with Gaussian prior and noise. Under this model, maximizing the
+expected information gain (KL divergence \cite{CoverThomas91}) between
+posterior and prior amounts to minimizing the determinant of the
+posterior covariance matrix. In a frequentist setting, a D-optimal
+design minimizes the volume of the uncertainty ellipsoid \cite[page
+  16]{Ucinski05}, but this is done for the Fisher information matrix
+and not the posterior covariance. However, \cite{Chaloner1995} shows
+that the latter is just a regularized version of the former.
 
@@
 different contexts. In a frequentist and finite-dimensional context,
-\cite{Fedorov96} and \cite[chapter 2.4.3]{Ucinski05} discuss this
+\cite{fedorov1996} and \cite[chapter 2.4.3]{Ucinski05} discuss this
 phenomenon and suggest an approach called clusterization-free design.
@@
 constraints between observations. A somewhat different approach is
-suggested in \cite[page 49]{Fedorov12}, where close by design
+suggested in \cite[page 49]{fedorov2012}, where close by design
 observations are merged --- a procedure which obviously does not avoid
 clusterization. The same problem arises in time-series analysis for
-pharmacokinetic experiments. The authors of \cite{Hooker09} suggest
+pharmacokinetic experiments. The authors of \cite{hooker2009} suggest
 modeling auto-correlation time in the noise model, which is equivalent
@@
 computational and less theoretic. Model errors were considered in
-\cite{Attia20, Koval20}. The former study is focused on inferring a
+\cite{attia2020, koval2020}. The former study is focused on inferring a
 Quantity of Interest (QoI). The focus of the latter is reducing
@@
   grad}). This allows us to rigorously show how model error mitigates
-clusterization (section \ref{section:non vanishing}). We show how
-clusterization occurs in a sequential design scenario (section
-\ref{subsec:clusterization sequential}). Then, we consider a
-simultaneous D-optimal design scenario (section
-\ref{subsec:clusterization simultaneous}). We show that designs that
-exhibit clusterization are just as optimal as clusterization free
-designs.
+clusterization (section \ref{section:non vanishing}). We show that
+designs that exhibit clusterization are just as optimal as
+clusterization free designs.
 
@@
 infinite-dimensional settings. We prove a Matrix Determinant Lemma in
-\ref{lemma:MDL}. We generalize a lemma due to Lax \cite{lax97} for
+\ref{lemma:MDL}. We generalize a lemma due to Lax \cite{Lax07} for
 calculating $\frac{\der}{\der t} \log \det (I + X(t))$, for an
@@
 also provide other tools for understanding D-optimal designs in
-infinite-dimensional Bayesian inverse problems. These include tools
-for calculating the increase in the design criterion per observation
---- Lemma \ref{lemma:design increase} and Corollary \ref{cor:zero mod
-  err}.
+infinite-dimensional Bayesian inverse problems.%%  Among these are tools
+%% for calculating the increase in the design criterion per observation
+%% --- Lemma \ref{lemma:design increase} and Corollary \ref{cor:zero mod err}.
 
@@
 covariance $\postcov$ is known to have a closed form formula, even in
-infinite dimensions\cite{Stuart10}. We denote by $\fwd$ the dynamics
+infinite dimensions \cite{Stuart10}. We denote by $\fwd$ the dynamics
 operator, so that $u( \cdot,T) = \fwd u_0$, and the observation
@@
 of sensors are so close they are indistinguishable.
-
-\section{Preliminaries and Notation}\label{section:prelim}
-
-\subsection{Overview}
-In this section we define notations that will be used throughout the
-paper. The theoretical foundations for inverse problems over function spaces
-can be found in \cite{Stuart10}.
-
-\subsection{Bayesian Linear Inverse Problems}\label{subsec:abstract OED}
-Let $\hilp$ and $\hilo$ be separable Hilbert spaces (the subscripts p
-and o are for ``parameter'' and ``observation'', respectfully). We
-denote $\hilo^*$ the Hilbert space of all linear functionals on
-$\hilo$. Consider $\fwd: \hilp \to \hilo$, a linear operator (the
-``forward operator''). We are interested in forward operators that are
-strongly smoothing (have fast decaying modes --- the heat operator
-from section \ref{subsec:example} is a prime example). Our goal is to
-infer $\param \in \hilp$ --- some parameter of the dynamics --- given
-noisy observations of $\fwd \param$. We take a Gaussian prior $\pr$
-for $\param$. Hence, $\param \sim \pr = \normal(\prmean ,\prcov)$ with
-some appropriate covariance operator $\prcov$ \cite{Stuart10}. It is
-important to note that $\fwd \prcov \fwd^*$ is the prior covariance in
-$\hilo$ \cite{Stuart10}. As such, it is assumed invertible --- an
-assumption we rely on below. Throughout this paper, we denote $m$ the
-number of observations we are allowed to take. Observations are taken
-via the operator $\obs \in ( \hilo^* )^m$. In an analogy with linear
-algebra, where row vectors are thought of as linear functionals, we
-think of $\obs$ as having ``rows'' $\meas_j, j=1,\dots,m$:
-\begin{equation}\label{eq:O}
-  \obs = (\meas_1,\dots, \meas_m)^t, \meas_j \in \hilo^*, j = 1,\dots,m.
-\end{equation}
-This way, for $u \in \hilo$ we have $\obs u = (\meas_1(u), \dots,
-\meas_m(u) )^t \in \R^m$.
-%% but we may drop the parentheses and write $\meas_j u = \meas_j(u)$.
-%% A few observations regarding $\obs$ are in order.
-%% First, it is good to keep in mind that $( \hilo^* )^m$ is a Hilbert
-%% space with norm $\| \obs \| = \sum_{j=1}^m \|\meas_j\|$.
-For $u \in \hilo$ and $v\in \R^m$:
-\begin{align*}
-  \big (\obs^*v \big ) (u) &= \langle v, \obs u \rangle_{\R^m} = \sum_{j=1}^m  v_j \meas_j(u)
-  = v^t \left ( \obs u \right ) = (v^t \obs) (u),
-\end{align*}
-and thus:
-\begin{align}\label{eq:obs*}
-  \obs^*v &= \sum_{j=1}^m v_j \meas_j = v^t \obs.
-\end{align}
-
-%% \begin{observation}
-%%   It is best to think of $\meas_j,j=1,\dots,m$ as row vectors and of
-%%   $\obs$ as a matrix with rows $\meas_j$. 
-%% \end{observation}
-
-Each observation is a linear functional, chosen from some subset of
-$\hilo^{*}$. Data is acquired via noisy observations
-\begin{align}\label{eq:inverse problem}
-  \data := \obs (\fwd \param + \eps') + \eps = \obs \fwd \param + \obs \eps' + \eps,
-\end{align}
-with $\data \in \R^m$ and $\eps, \eps'$ defined next. We consider two
-types of error terms. First, there is spatially correlated model error
-$\eps' \sim \normal(0,\modcov)$, modeled as a centered Gaussian
-measure on $\hilo$ with covariance operator $\modcov$. Then there is
-observation error $\eps \sim \normal(0, \sigma^2 I_m)$, with $I_m$ the
-$m \times m$ identity. Both error terms and the prior are assumed
-independent of each other. The observation and forward operators are
-typically merged $\tmp := \obs \fwd$
-\cite{AlexanderianGloorGhattas14}. Since we need to separate the
-observation operator $\obs$ from the forward operator $\fwd$ in order
-to carry out the analysis in the following sections, we explicitly
-separate $\fwd$ and $\obs$ in \eqref{eq:inverse problem}. We now turn
-to understand the terms present in \eqref{eq:inverse problem}.
-
-\subsubsection{Dynamics and Observation Operators}\label{subsec:dynamics}
-Rows of $\obs$ are linear functionals and we refer to these as {\it
-  observation vector}s. Note that we cannot choose every observation
-vector we want. For example, we may be restricted only to pointwise
-evaluations of $\fwd \param$ by the sensors at our disposal, in which
-case measuring, e.g., the mean $\ell(u) = \int_{\Omega}u$ of a
-function is not possible. Thus, it is best to think of $(\obs \fwd
-\param)_j$ as approximations to pointwise evaluations. For example, if
-$\hilo = L^2([-\pi, \pi])$, consider $\ell_1(x) := (2\pi
-\eps)^{-\frac{1}{2}}\exp(-\frac{x^2}{2\eps^2})$ and $\ell_2(x)
-:=\eps^{-1}\mathbf{1}_{[-\frac{\eps}{2}, \frac{\eps}{2}]}$. Both
-$\ell_1$ and $\ell_2$ approximate a point measurement at $0$ for small
-$\eps$ and, crucially, $\ell_1, \ell_2 \in \hilo^*$. It is important
-to note that although $\|\ell\|_1 = \|\ell_2\|_1=1$, unfortunately
-$\|\ell_1\|_2^2 = (4\pi \eps)^{-\frac{1}{2}}$ and $\|\ell_2\|_2^2 =
-\eps^{-1}$.
-
-\subsubsection{Error Terms}
-Considering the sum of the error terms, it is easy to see that
-$\bar{\eps} := \obs \eps' + \eps \in \R^m$. It is a centered Gaussian
-random vector and its covariance matrix is
-\begin{align}\label{eq:Sigma}
-  \begin{split}
-    \Sigma(\obs) :&= \mathbb{E}[ (\obs \eps' + \eps)  (\obs \eps' + \eps)^t ] 
-    % 
-    % 
-    = \obs \modcov \obs^* + \sigma^2I , 
-  \end{split}
-\end{align}
-where
-\begin{equation}\label{eq:modcov explained}
-  [\obs \modcov \obs^*]_{ij} = e_i^t \obs \modcov \obs^* e_j = \meas_i (\modcov \meas_j),
-\end{equation}
-with $e_j \in \R^m$ the $j$th standard basis vector. The first
-equality in \eqref{eq:modcov explained} holds by definition and the
-second by \eqref{eq:obs*}. The explicit dependence on $\obs$ will be
-mostly dropped for notational convenience, so $\Sigma(\obs) =
-\Sigma$. Thus, for a fixed $\obs$ (i.e.\ a fixed set of observations)
-we can equivalently write $\data = \obs \fwd \param + \bar{\eps}$ with
-$\bar{\eps} \sim \normal(0,\Sigma)$. Taking $\modcov = 0$ is common
-practice \cite{Tarantola05,KaipioSomersalo05,Vogel02}. This reduces to
-the case where we take iid observations. Then $\Sigma = \sigma^2I$ is
-simply a scalar matrix that does not depend on $\obs$. Note that
-taking an error model with a non-scalar covariance as we do here
-allows us to consider model error (modeled by $\modcov$) as well as
-observation error (modeled by $\sigma^2$). For example, say we believe
-our forward model does not capture some small scale phenomenon.  Then
-we may express this belief by saying $\tru = \fwd + \err$, with $\fwd$
-depending on $\param$ and $\err$ depending on $\sspar$, with $\param
-\perp \sspar$. We do not know much about this effect but it is
-reasonable to assume it changes continuously in our domain. We (may
-choose to) model it as $\normal (0, \modcov)$ and take $\modcov$ to
-reflect the spatial (or other) variability we imagine $\err \sspar$
-has. Such small scale phenomenon can arise as a modeling issue, where
-we might not model the system in its entirety. It can also arise from
-a numerical source, where our discretization of the system is not fine
-enough to capture all small scale phenomena. In section
-\ref{section:non vanishing}, we will see that assuming some
-correlation in the error mitigates the clusterization phenomenon, as
-reported in the literature \cite{Ucinski05}.
-
-Finally, it is useful to record a known result regarding the posterior
-covariance operator of our inverse problem:
-\begin{align}\label{eq:postcov}
-  \postcov = (\prcov^{-1} + \fwd^* \obs^* \Sigma^{-1} \obs \fwd
-  )^{-1}.
-\end{align}
-
-\subsection{D-Optimal Designs in Infinite Dimensions}\label{subsec:D optimal design} 
-A D-optimal design maximizes expected KL divergence between posterior
-and prior. It is useful to recall the definition of KL divergence for
-an arbitrary prior measure:
-$$
-KL(\post||\pr) = \int \log \frac{\der \post}{\der \pr}(\param) \der \post(\param).
-$$
-
-The meaning of D-optimal design in infinite-dimensional Hilbert spaces
-was investigated in \cite{AlexanderianGloorGhattas14}. There, the
-authors make assumptions that amount to $\Sigma=I$ (implied by
-$\modcov = 0,\sigma^2=1$), but we choose not to take these
-simplification here. This is because $\Sigma$ can determine ``how
-much'' clusterization we see. As stated
-\cite[pp. 681]{AlexanderianGloorGhattas14}, the results hold for more
-general covariance matrices. The conclusion is that in
-infinite-dimensions, a D-optimal design is well-defined as maximizing
-the expected KL divergence between posterior and prior. The main
-result is summarized, using our notation, in the following theorem:
-\begin{theorem}[Slightly modified Theorem 1 from \cite{AlexanderianGloorGhattas14}]
-  Let $\pr = \normal(\prmean,\prcov)$ be a Gaussian prior on $\hilp$
-  and let $\post = \normal(\postmean,\postcov)$ the posterior measure
-  on $\hilo$ for the Bayesian linear inverse problem $\data = \obs
-  \fwd\param + \obs \eps' + \eps$ discussed above. Then
-  \begin{align}\label{eq:objective}
-    \begin{split}
-      \tar( \obs) :&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
-      % 
-      % 
-      % 
-      &= \frac12 \log \det 
-      ( I + \prcov^{1/2}  \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}).
-    \end{split}
-  \end{align}
-\end{theorem}
-\begin{definition}\label{def:d optimality}
-  A design $\obs^{\star}$ is said to be D-optimal if $\obs^{\star} = \argmax_{\obs} \tar(\obs)$.
-\end{definition}
-
-\noindent The intuition behind Definition \ref{def:d optimality} is
-straightforward. From \eqref{eq:postcov} and \eqref{eq:objective}:
-\begin{align*}
-  \begin{split}
-    \det ( I + \prcov^{1/2}  \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}) &= \det \Big( \prcov ( \prcov^{-1} + \fwd ^* \obs^* \Sigma^{-1} \obs \fwd) \Big )\\
-    &= \det \prcov \det \postcov^{-1}.
-  \end{split}
-\end{align*}
-Since $\prcov$ is constant, a D-optimal design minimizes the posterior
-covariance determinant, analogously to the finite-dimensional case.
-
-%% \subsection{Sequential vs Simultaneous Optimization}\label{subsec:seq vs sim}
-%% From defintion \ref{def:d optimality} we wish to characterize solution(s) of the
-%% following optimization problem for $\tar$. %%: (\hilo^*)^m \to \R$:
-%% \begin{align}\label{eq:optimization}
-%%   \obs^{\star} := \argmax_{\obs} \tar( \obs ) 
-%%   = \argmax_{\obs} \frac12 \log \det 
-%%   (I + \prcov^{1/2} \fwd^*\obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}),
-%% \end{align}
-%% where $\obs$ is constrained to some allowed set of observations. We
-%% call this problem ``simultaneous optimization'', since all
-%% observations are decided on simulatneously.
-
-%% For computational reasons, one may prefer to find the best
-%% observations in a sequential manner. Denote
-%% \begin{equation}\label{eq:def obs_k}
-%%   \obs_k := (\meas_1,\dots, \meas_k)^t,  k\leq m.
-%% \end{equation}
-%% Sequential optimal design proceeds as follows. Find $\meas_1$ by
-%% maximizing $\tar(\obs_1)$. Then, keeping $\meas_1$ fixed --- find
-%% $\meas_2$ as the maximizer of $\tar(\obs_2)$. Then, find $\meas_3$ by
-%% keeping $\meas_1,\meas_2$ fixed and taking $\meas_3$ as the maximizer
-%% of $\tar(\obs_3)$. Continue this way until $\obs_m = \obs$ is
-%% found, where $m$ is the number of available observations. %% It is
-%% %% important to notice that this scheme does not require actually
-%% %% observing data --- in \eqref{eq:objective} data is averaged out.
-
-%% The analysis in this paper is conducted for the general simultaneous
-%% optimization case. The sequential optimization case is dealt with in
-%% section \ref{subsec:clusterization sequential}. It is important to
-%% note, however that all conclusions we arrive at for the simultaneous
-%% case easily specialize to the sequential case by considering the
-%% posterior as the next sequential step's prior.
-
-\section{The Constrained Optimization Problem of D-Optimal Design}\label{section:D and grad}
-
-\subsection{Overview}
-We seek a formulation of the D-optimal design problem
-\eqref{eq:optimization} using Lagrange multipliers. In Section
-\ref{section:objective} we find the gradient of $\tar$ and record it
-in \eqref{eq:tar grad}. In Section \ref{subsec:unit norm} we suggest
-using unit-norm constraints on rows of $\obs$ so that the optimization
-problem \eqref{eq:optimization} is analytically tractable. We find
-gradients for the new constraints and record them in \eqref{eq:grad
-  constraints}. Finally, in Section \ref{subsec:necessary} we
-formulate the D-optimal design problem as a Lagrange multipliers
-problem in equation \eqref{eq:conditions}.
-
-\subsection{The Objective and its Gradient}\label{section:objective}
-In order to use Lagrange multipliers, we need to find the gradient of
-$\tar$. Said gradient is recorded in \eqref{eq:tar grad} below. We
-start by recording Lemmas \ref{lemma:lax}, \ref{lemma:aux calc},
-proved in the appendix:
-
-\begin{restatable*}[Generalized from \cite{Lax97}]{lemma}{lax}\label{lemma:lax}
-  Let $Y(t)$ be a differentiable operator-valued function. Assume 
-  $I+Y(t)$ is invertible, $Y(t)$ self-adjoint and trace-class. Then
-  \begin{equation*}
-    \frac{\der \log \det (I+Y(t))}{\der t} = \tr{(I+Y(t))^{-1} \dot{Y}(t)}.
-  \end{equation*}
-\end{restatable*}
-
-\begin{restatable*}[Auxilliary Calculations]{lemma}{aux}\label{lemma:aux calc}
-  Let $T(\obs) := \obs^* \Sigma^{-1}(\obs)\obs$, with $\Sigma(\obs)$
-  defined as in \eqref{eq:Sigma}. Then:
-  \begin{align*}
-    \delta T(\obs)V &= V^* \Sigma^{-1} \obs 
-    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
-    &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs
-    + \obs^* \Sigma^{-1} V.
-  \end{align*}
-\end{restatable*}
-
-The first variation of $\tar (\obs)$ in the direction $V$ is:
-\begin{align}\label{eq:tar var}
-  \begin{split}
-    \delta \tar(\obs) V 
-    :&= \frac{\der}{\der\tau} \Big |_{\tau=0} \tar(\obs + \tau V)\  \text{ (by definition of variation)}\\
-    % 
-    % 
-    % 
-    &= \frac12 \frac{\der}{\der \tau} \Big |_{\tau=0} \log \det 
-    (I + \prcov^{1/2} \fwd^* T(\obs+\tau V)\fwd \prcov^{1/2} ) \text{ (by \eqref{eq:objective})}\\
-    % 
-    % 
-    % 
-    &= \frac12 \tr{( I + \prcov^{1/2} \fwd^* \obs^* \Sigma^{-1}
-    \obs\fwd \prcov^{1/2} )^{-1}
-    \frac{\der}{\der \tau} \Big |_{\tau=0}
-    \prcov^{1/2} \fwd^* T(\obs+\tau V) \fwd \prcov^{1/2}}\ \text{ (by \ref{lemma:lax})}\\
-    % 
-    % 
-    % 
-    &= \frac12 \ttr\Big \{ \postcov \fwd^* (V^* \Sigma^{-1} \obs 
-    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
-    &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs 
-    + \obs^* \Sigma^{-1} V ) \fwd \Big \}  \text{ (by \ref{lemma:aux calc})}\\
-    %
-    %
-    %
-    &= \tr{\postcov \fwd^* ( \obs^* \Sigma^{-1} V -
-    \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs ) \fwd} \text{
-        (cyclic property of trace)} \\
-    %
-    %
-    % 
-    &= \tr{\postcov \fwd^* \obs^* \Sigma^{-1} V 
-    ( I - \modcov \obs^* \Sigma^{-1}\obs ) \fwd} \\
-    % 
-    %
-    %
-    &= \tr{V ( I - \modcov \obs^* \Sigma^{-1}\obs )
-    \fwd \postcov \fwd^* \obs^* \Sigma^{-1}}.
-  \end{split}
-\end{align} 
-Denote
-\begin{align}\label{eq:tar grad}
-  \nabla \tar(\obs) &:= (I - \modcov \obs^* \Sigma^{-1} \obs) \fwd
-  \postcov \fwd^* \obs^*\Sigma^{-1},
-\end{align}
-the gradient of $\tar (\obs )$ and we now justify this definition.
-Since the trace of $A := V \nabla \tar(\obs) \in \R^{m \times m}$ is
-just $\ttr A = \sum_{j=1}^m e_j^t A e_j$ (with $e_j$ the $j$th standard
-basis vector), we see that
-\begin{align*}
-  \delta \tar(\obs)V = \tr{V \nabla \tar(\obs)} = \sum_{j=1}^m
-  V_j(\nabla \tar(\obs)_j),
-\end{align*}
-with $V_j \in \hilo^*$ and $\nabla \tar(\obs)_j \in \hilo^{**} =
-\hilo, j=1,\dots,m$. Thus, $\nabla \tar( \obs ) \in \hilo^m$ is indeed
-the correct gradient and the notation \eqref{eq:tar grad} is
-justified. The gradient $\nabla \tar(\obs)$ should be viewed as a row
-vector, as the product $V \nabla \tar$ is in $\R^{m \times m}$. This
-will prove important in section \ref{subsec:necessary}.
-
-\subsection{Unit Norm Constraints and their Gradients}\label{subsec:unit norm}
-In this section we suggest relaxed constraints in
-\eqref{eq:constraints}, and their gradient in \eqref{eq:grad
-  constraints}.
-
-As mentioned before, we cannot choose any $\obs$ when maximizing
-$\tar(\obs)$. Recall $\obs = (\meas_1,\dots,\meas_m)^t \in
-(\hilo^*)^m$ and each $\meas_j,j=1,\dots,m$ must be chosen from some
-allowed set of functionals in $\hilo^*$. %% This set differs based on
-%% the kind of sensors we have at our disposal and the properties of
-%% $\hilo$, as discussed in section \ref{subsec:dynamics}.
-Proposition \ref{prop:bigger better} below (proof in the appendix)
-gives better understanding of the constraints.
-\begin{restatable*}{proposition}{biggerbetter}\label{prop:bigger better}
-  Let $\obs = (\meas_1,\dots,\meas_m)^t$, $j \in \{1,\dots,m\}$, $\sigma^2
-  > 0$ and $\lambda > 1$. Then $\tar(\obs)$ increases if we use
-  $\lambda \meas_j$ in $\obs$ instead of $\meas_j$.
-\end{restatable*}
-The idea behind Proposition \ref{prop:bigger better} is simple: making
-the norm of rows of $\obs$ large is equivalent to making $\sigma^2$
-small. %% Viewed another way, it increases the Signal to Noise Ratio
-%% (SNR) --- which is always desirable when taking any observation
-%% prone to error.  This can be understood easily from the formulation
-%% of the inverse problem $\data = \obs \fwd \param + \obs \eps' +
-%% \eps$, where $\eps'$ can be taken as model error and $\eps$ is iid
-%% observation error.
-
-Note that for point evaluations, the norm of the observation
-functional is always one:
-\begin{align*}
-  \| \delta_{\x} \| = \sup_{0 \neq u \in C(\Omega)} \frac{
-    |\int_{\Omega}u(\y) \delta_{\x}(\y) \der \y| 
-  }{
-    \sup|u|}
-  = \sup_{0 \neq u \in C(\Omega)} \frac{|u(\x)|}{ \sup|u|} = 1,
-  \forall \x \in \Omega.
-\end{align*}
-Following this, and in light of Proposition \ref{prop:bigger better},
-we consider observation vectors in the unit ball in $\hilo^*$. The
-unit norm constraints can be written using \eqref{eq:obs*} as a series
-of $m$ equality constraints (one for each observation) on $\obs$ as
-\begin{align}\label{eq:constraints}
-  \phi_j(\obs) :=\frac12 \| \obs^* e_j\|_{\hilp}^2 - \frac12 = 0,j=1,\dots,m.
-\end{align}
-The first variations are:
-\begin{align*}
-  \delta \phi_j(\obs)V  
-  &= \frac12\lim_{\tau \to 0}\tau^{-1}
-  ( \|(\obs + \tau V)^*e_j \|_{\hilp}^2 - \|\obs ^*e_j \|_{\hilp}^2  ) \\
-  %
-  %
-  %
-  &= \frac12\lim_{\tau \to 0}\tau^{-1}
-  ( \langle (\obs + \tau V)^*e_j, (\obs + \tau V)^*e_j \rangle_{\hilp} - 
-  \langle \obs^*e_j, \obs^*e_j \rangle_{\hilp} ) \\
-  % 
-  % 
-  %
-  &= \frac12\lim_{\tau \to 0}\tau^{-1}
-  (2\tau \langle \obs^*e_j,V^*e_j \rangle_{\hilp} 
-  +\tau^2 \langle V^*e_j, V^*e_j \rangle_{\hilp} ) \\
-  %
-  %
-  % 
-  &= \langle \obs^*e_j,V^*e_j \rangle_{\hilp} \\
-  %
-  %
-  % 
-  &= \langle V \obs^*e_j,e_j \rangle_{\R^m} \\
-  %
-  %
-  %
-  &= e_j^t V \obs^* e_j \\
-  % 
-  %
-  %
-  &= \tr{V \obs^* e_je_j^t}.
-\end{align*}
-The same arguments justifying \eqref{eq:tar grad} hold here, and thus, by \eqref{eq:obs*}:
-\begin{align}\label{eq:grad constraints}
-\nabla \phi_j(\obs) = \obs^* e_j e_j^t = \meas_j e_j^t , j=1,\dots,m,
-\end{align}
-where $\nabla \phi_j(\obs) \in \hilo^m$. As noted at the end of
-Section \ref{section:objective}, the gradient $\nabla \phi_j(\obs)$ is
-best thought of as a row vector.
-
-\subsection{Necessary Conditions for Optimal Design}\label{subsec:necessary}
-Necessary conditions for D-optimality are found from Lagrange
-multipliers:
-\begin{align}
-  &\nabla \tar(\obs) = \sum_{j=1}^m \xi_j \nabla \phi_j (\obs)
-  \label{eq:Lagrange mult1} \\
-    &\phi_j(\obs) = 0, j = 1,\dots,m. \label{eq:Lagrange mult2}
-\end{align}
-
-Recall where the objects involved are defined: the observation
-operator, $\obs \in (\hilo^*)^m$. The objective $\tar: (\hilo^*)^m \to
-\R$ and its gradient is $\nabla \tar(\obs) \in \hilo^m$. Similarly,
-$\phi_j: (\hilo^*)^m \to \R$ and $\nabla \phi_j(\obs) \in \hilo^m$.
-
-Substituting \eqref{eq:tar grad} and \eqref{eq:grad constraints} in
-\eqref{eq:Lagrange mult1}:
-\begin{equation*}
-  (I - \modcov \obs^* \Sigma^{-1} \obs) \fwd \postcov \fwd^* \obs^*\Sigma^{-1}
-  = \sum_{j=1}^m \xi_j \obs^* e_je_j^t = (\xi_1 \meas_1,\dots,\xi_m \meas_m).
-\end{equation*}
-Let $\Xi := \diag(\xi_j)$. Then, with some abuse of notation, this can
-be written more compactly as:
-\begin{equation}\label{eq:conditions}
-  ( I - \modcov \obs^* \Sigma^{-1} \obs) \fwd \postcov \fwd^* \obs^*  \Sigma^{-1}
-  = \obs^* \Xi.
-\end{equation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SECTION Analysis of Optimal Designs --- Non-Vanishing Model Error
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Analysis of D-Optimal Designs --- Non-Vanishing Model Error}\label{section:non vanishing}
-In this section we show the effect model error has on the
-clusterization phenomenon. We will see that if $\modcov \neq 0$
-clusterization will not occur. This is contrary to the previous case
-of vanishing model error.
-
-Corollary \ref{cor:same meas} below is proved in the appendix. Denote
-$\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm :=
-(\meas_1,\dots,\meas_{m-1})^t$. Also, denote $\Sigmam := \Sigma (\obsm)$ and
-$\postcovm$ the posterior covariance that arises when we use $\obsm$.
-\begin{restatable*}{corollary}{samemeas}\label{cor:same meas}
-  If $\meas_m = \meas_j$ for some $1 \leq j \leq m-1$, then
-  \begin{equation*}
-    \tar(\obs) - \tar(\obsm) =
-    \log \left ( 1 + \frac{\sigma^2
-      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
-      \obsm^* \Sigmam^{-1}e_j \rangle
-    }{
-      2 - \sigma^2 e_j^t\Sigmam^{-1}e_j 
-    }       
-    \right ).
-  \end{equation*}
-\end{restatable*}
-Recall from \eqref{eq:Sigma} that $\Sigma(\obs) = \obs \modcov \obs^*
-+ \sigma^2I$. Then
-
-$$
-\lim_{\sigma^2 \to 0} \tar(\obs) -\tar(\obsm) = 0
-$$
-
-Hence no increase in the design criterion is achieved by taking a
-repeated observation (in the limit $\sigma^2 \to 0$). This is clearly
-sub-optimal and $\obs$ cannot give rise to a D-optimal design. Thus,
-for small observation error levels, the clusterization effect will be
-mitigated by the presence of a non-zero model error. Since the design
-criterion is not defined for $\sigma^2 = 0$ and identical
-observations, we cannot make a statement regarding $\sigma^2 = 0$,
-except in the limiting sense described above.
-
-\section{Acknowledgements}
-This study is a part of my PhD thesis. Most of it was written under
-the instruction of Prof. Georg Stadler in New York University's
-Courant Institute. I would like to thank him for his great
-mentorship. This work was supported by The Raymond and Beverly Sackler
-Post-Doctoral Scholarship.
-
-\input{vanishing}
-\input{appendix}
-
-\bibliographystyle{amsplain}
-\bibliography{refs,georg_refs}
-
-\end{document}

### 2023-03-29 latex/paper.tex
Paper almost donegit status !

@@
 
-\title[Sensor Clusterization in D-optimal design in infinite
-  dimensions]{Sensor Clusterization in D-optimal Design in Infinite
-  Dimensional Bayesian Linear Inverse Problems}
+\title[Generic Properties of D-optimal Designs]{Generic Properties of D-optimal
+  Designs in Infinite Dimensional Bayesian Linear Inverse Problems}
 
@@
 
-\section{Introduction}\label{section:OED intro}
-Experimental design is an important part of many scientific
-investigations. When considering an inverse problem, one can often
-specify sensor locations (e.g.\ in geophysics and oceanography
-applications), certain wavelengths (e.g.\ in MRI) or wave reflections
-from the ground (e.g.\ searching for oil or using a radar). Whatever
-the allowed set of observations is, one should select the optimal
-observations to take, in order to increase accuracy, reduce costs, or
-both.
-
-Designing experiments is usually done by optimizing some \emph{design
-  criterion}. This is true both for frequentists
-\cite{Silvey13,Ucinski05} as well as for Bayesians
-\cite{ChalonerVerdinelli95}. See \cite{ChalonerVerdinelli95} for an
-investigation of the analogy between the two approaches. Although
-there is a plethora of design criteria, we focus on just one of these,
-commonly referred to as \emph{D-optimal design}. It has a simple and
-appealing motivation in the Bayesian context as explained in
-\cite{ChalonerVerdinelli95}: Consider a linear model in finite
-dimensions, with Gaussian prior and noise. Under this model,
-maximizing the expected information gain (KL divergence
-\cite{KullbackLeibler51,CoverThomas12}) between posterior and prior
-amounts to minimizing the determinant of the posterior covariance
-matrix. In a frequentist setting, a D-optimal design minimizes the
-volume of the uncertainty ellipsoid \cite[page 16]{Ucinski05}, but
-this is done for the Fisher information matrix and not the posterior
-covariance. However, \cite{ChalonerVerdinelli95} shows that the latter
-is just a regularized version of the former.
-
-The previous discussion is classical for experimental design when
-inference takes place over a finite (not too large) number of
-parameters. The subject of optimal experimental design for function
-inference in a Bayesian context was pioneered by
-\cite{AlexanderianGloorGhattas14, AlexanderianPetraStadlerEtAl16,
-  AlexanderianPetraStadlerEtAl14}. Similarly to the finite dimensional
-case, it can be shown that a D-optimal design arises naturally for
-linear models when one wishes to maximize the KL divergence between
-posterior and prior. This amounts to minimizing the determinant of the
-posterior covariance operator (understood as a product of its
-eigenvalues). Some difficulties arise in the process, but remedies can
-be found as shown in \cite{AlexanderianGloorGhattas14}.
-
-It seems counter intuitive that when one computes an optimal design
-using the D criterion, the optimization process results in
-observations that are very similar. For example, if an observation is
-thought of as measuring some function value at $\x \in \Omega
-\subseteq \R^d, d=1,2,3$ (with added error) then the optimization
-procedure sometimes places sensors in very close proximity to each
-other (as can be seen in figure \ref{fig:clusterization
-  illustration}). Following \cite{Ucinski05}, we refer to this
-phenomenon as \emph{sensor clusterization}.
-
-\subsection{Related Work}
-The phenomenon of sensor clusterization seems to be known in several
-different contexts. In a frequentist and finite-dimensional context,
-\cite{Fedorov96} and \cite[chapter 2.4.3]{Ucinski05} discuss this
-phenomenon and suggest an approach called clusterization-free design.
-In such designs, the user enforces observation locations to be far
-from each other. One way to do this is by introducing correlated
-errors which, philosophically, accounts for both observation error and
-model error. Another method considered is imposing distance
-constraints between observations. A somewhat different approach is
-suggested in \cite[page 49]{Fedorov12}, where close by design
-observations are merged --- a procedure which obviously does not avoid
-clusterization. The same problem arises in time-series analysis for
-pharmacokinetic experiments. The authors of \cite{Hooker09} suggest
-modeling auto-correlation time in the noise model, which is equivalent
-to the correlated errors mentioned above.
-
-Any of the above mentioned approaches might serve as a remedy and push
-sensors away from each other. Yet, none offers any insight as to why
-clusterization occurs. Also, as better models are employed, model
-error is decreased and the clusterization phenomenon will eventually
-reappear. While these approaches are practical and help us avoid the
-problem, they do not provide insight as to why sensors are clustering.
-
-Similarly, in the inverse problems community, work is mostly
-computational and less theoretic. Model errors were considered in
-\cite{Attia20, Koval20}. The former study is focused on inferring a
-Quantity of Interest (QoI). The focus of the latter is reducing
-forward solves, using randomized linear algebra. Both studies present
-numerical techniques for finding optimal designs when model error is
-present. Both are restricted to linear inverse problems (although in
-the latter the authors use their method on a nonlinear problem by
-taking a Laplace approximation for the posterior). Both find an
-optimal design by first solving a continuous problem for sensor
-weights. Said solution is then sparsified to give a binary
-design. Both studies are successful in the task of Bayesian
-inversion. However, neither of these studies mention any effect model
-errors can have on sensor clusterization. The current study is mostly
-theoretical and aims to fill the gap of understanding sensor
-clusterization.
-
-\subsection{Contribution}
-We propose and thoroughly study a relaxed and analytically tractable
-model for understanding D-optimal designs. Under this model, D-optimal
-designs are solutions of a constrained optimization problem,
-formulated using Lagrange multipliers (section \ref{section:D and
-  grad}). This allows us to rigorously show how model error mitigates
-clusterization (section \ref{section:non vanishing}). We show how
-clusterization occurs in a sequential design scenario (section
-\ref{subsec:clusterization sequential}). Then, we consider a
-simultaneous D-optimal design scenario (section
-\ref{subsec:clusterization simultaneous}). We show that designs that
-exhibit clusterization are just as optimal as clusterization free
-designs.
-
-A beautiful mathematical structure arises in D-optimal designs when no
-model error is present. The Lagrange multipliers problems is in fact a
-nonlinear eigenvalue problem for the observations and prior
-covariance. The operator for which eigenvectors and eigenvalues are
-sought is a sum of two operators. The first is the prior
-covariance. The second is an outer product of the observations (see
-section \ref{section:vanishing} for details and exact statement). This
-structure helps us shed light on D-optimal designs. We characterize
-these designs in Theorem \ref{thm:char}. One insight we prove is that
-a D-optimal design reduces uncertainties where they are highest
-first. Other interestig phenomena arise, but they require setting
-notation, and are discussed later.
-
-In the process, we generalize several lemmas from linear algebra to
-infinite-dimensional settings. We prove a Matrix Determinant Lemma in
-\ref{lemma:MDL}. We generalize a lemma due to Lax \cite{lax97} for
-calculating $\frac{\der}{\der t} \log \det (I + X(t))$, for an
-operator valued function $X(t)$ in \ref{lemma:lax}. We also prove
-lemmas in linear algebra. One constructs a decomposition $M = AA^t$
-where $A$ has unit norm columns \ref{lemma:free}. Another shows
-simultaneous diagonizability of the prior and outer product in the
-nonlinear eigenvalue problem referenced above \ref{lemma:sim diag}. We
-also provide other tools for understanding D-optimal designs in
-infinite-dimensional Bayesian inverse problems. These include tools
-for calculating the increase in the design criterion per observation
---- Lemma \ref{lemma:design increase} and Corollary \ref{cor:zero mod
-  err}.
-
-\begin{restatable}[D-optimal designs with vanishing model error]{theorem}{main}\label{thm:char}
-  Let:
-  \begin{enumerate}
-  \item $\fwd:\hilp \to \hilo$ a linear forward operator,
-    $\prcov:\hilp \to \hilp$ prior covariance operator, $\obs: \hilo
-    \to \mathbb{R}^m$ observation operator, where $m \in \mathbb{N}$
-    is the number of observations taken.
-    \item $\sigma^2 \in \mathbb{R}_{+}$ observation noise variance,
-      $\data = \obs \fwd \param + \eps$, where $\eps \in \mathbb{R}^m$
-      is iid $\mathcal{N}(0, \sigma^2)$ noise, $\pr \sim
-      \mathcal{N}(0, \prcov)$ prior measure, $\post$ the posterior
-      measure.
-  \item A D-optimality utility function
-    \cite{AlexanderianGloorGhattas14}:
-    \begin{align*}
-      \begin{split}
-        \tar(\obs) :&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
-        % 
-        % 
-        % 
-        &= \frac12 \log \det ( I + \sigma^{-2} \prcov^{1/2} \fwd ^*
-        \obs^* \obs \fwd \prcov^{1/2}).
-    \end{split}
-  \end{align*}
-  \item A D-optimal design operator $\obs$:
-    $$
-    \obs = \argmax_{\|\meas_j\| = 1, j=1,\dots,m}\tar(\obs)
-    $$ 
-  \item $\{\lambda_i\}_{i=1}^\infty$ eigenvalues of $\fwd\prcov\fwd^*$
-    in decreasing order of magnitude.
-  %% \item $\{\ev_i\}_{i=1}^\infty$ their corresponding eigenvectors.
-  \item $\{\eta_i\}_{i=1}^\infty$ eigenvalues of $\obs^*\obs$.
-  \end{enumerate}
-
-  Then:
-  \begin{enumerate}
-  \item $\obs^*\obs$ and $\fwd\prcov\fwd^*$ are simultaneously
-    diagonalizable.
-  \item $k := \rank \obs^*\obs \leq m$.
-  \item     
-    \begin{equation*}
-      \tar(\obs) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i). %= \frac12 \sum_{i=1}^{m} \log (1 + \sigma^{-2}\lambda_i\eta_i).
-    \end{equation*}
-  %% \item 
-  %%   \begin{equation*}
-  %%     k = \argmax \left \{ k:\lambda_k^{-1} < \sigma^{-2}\frac{m}{k} + \frac{1}{k} \sum_{j=1}^{k}
-  %%     \lambda_j^{-1} \right \}.
-  %%   \end{equation*}
-  \item
-    \begin{equation*}
-        \eta_i = \begin{cases}
-          \frac{m}{k} - \sigma^2 \lambda_i^{-1} + \sigma^2 \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1} & 1 \leq i \leq k \\
-          0 & i > k 
-        \end{cases}.
-    \end{equation*}
-  \end{enumerate}
-\end{restatable}
-
-\subsection{Limitations}\label{subsec:limitations}
-There are two main drawbacks of the study presented here. Our relaxed
-model does not consider any specific set of allowed
-observations. Rather, we take observations in the unit ball in some
-Hilbert space. This allows considerably less restrictive observations
-than any real-life problem does. The second drawback is that we do not
-show rigorously that clusterization necessarily occurs in a
-simultaneous design. We only show that it is as reasonable as no
-clusterization.
-
-\subsection{An Example of Clusterization}\label{subsec:example}
-\begin{figure}
-  \begin{tikzpicture}[thick, scale=1.3, every node/.style={scale=0.99}]
-    \begin{axis}
-      [
-      title={Posterior Pointwise Standard Deviations and D-Optimal Sensor Locations},  
-      xmin = 0,
-      xmax = 3.14,
-      xlabel = {$x$},
-      ylabel = posterior std,
-      ymin   = 0,
-      %compat = 1.3,
-      % ymax   = 130,
-      % ytick = \empty,
-      legend cell align=left,
-      % legend style={at={(0.45,0.2)}}
-      legend pos= outer north east 
-      ]
-      % \draw[black!30!white, thin] (50,0) -- (50,130);
-      % 
-      %% \addplot [thin, black, mark=none] table{stdv-heat-sens1-var1.txt};
-      %% \addlegendentry{1 sensors};
-
-      %% \addplot [thin, blue, mark=none] table{stdv-heat-sens2-var1.txt};
-      %% \addlegendentry{2 sensors};
-
-      %% \addplot [thin, red, mark=none] table{stdv-heat-sens3-var1.txt};
-      %% \addlegendentry{3 sensors};
-
-      \addplot [thin, green, mark=none] table{stdv-heat-sens4-var1.txt};
-      \addlegendentry{4 sensors};
-
-      \addplot [thin, purple, mark=none] table{stdv-heat-sens5-var1.txt};
-      \addlegendentry{5 sensors};
-
-      \addplot [thin, cyan, mark=none] table{stdv-heat-sens6-var1.txt};
-      \addlegendentry{6 sensors};
-
-      %% \addplot [black,  only marks, mark=x, mark size=1.5] 
-      %% table{locs-heat-sens1-var1.txt}; 
-      %% \addplot [blue,   only marks, mark=x, mark size=1.5]
-      %% table{locs-heat-sens2-var1.txt}; 
-      %% \addplot [red,    only marks, mark=x, mark size=1.5]
-      %% table{locs-heat-sens3-var1.txt};
-      \addplot [green,  only marks, mark=*, mark size=1.5] 
-      table{locs-heat-sens4-var1.txt}; 
-      \addplot [purple, only marks, mark=*, mark size=1.5] 
-      table{locs-heat-sens5-var1.txt}; 
-      \addplot [cyan,   only marks, mark=*, mark size=1.5] 
-      table{locs-heat-sens6-var1.txt};
-
-    \end{axis}
-  \end{tikzpicture}
-  \caption{The clusterization effect for the 1D heat equation
-    described in section \ref{subsec:example}. Posterior pointwise
-    standard deviations (lines) are plotted over the domain $[0,
-      \pi]$, for varying numbers of sensors. Sensor locations
-    (circles) were chosen to minimize (an expression analogous to) the
-    determinant of the posterior covariance. The clusterization effect
-    can be clearly seen for six sensors. Only four observation
-    locations are used --- two pairs of sensors are so close they are
-    indistinguishable.}
-  \label{fig:clusterization illustration}
-\end{figure}
-
-In section \ref{section:prelim} we present a more abstract and general
-formulation of the inverse problem we consider. But, for the purpose
-of illustration, we present clusterization via a toy model --- the 1D
-heat equation in $[0,\pi]$ with a homogeneous Dirichlet boundary
-condition.
-
-The 1D heat equation is:
-\begin{subequations}\label{eq:heat equation}
-  \begin{alignat}{2}
-    u_t &= \Delta u &&\qquad \text{in } [0,\pi] \times [0,\infty),\\
-      u &= 0 &&\qquad \text{on } \{0, \pi\} \times [0,\infty),\\
-        u &= u_0 &&\qquad \text{on }[0,\pi] \times \{0\}.
-  \end{alignat}
-\end{subequations}
-
-We would like to infer the initial condition $u_0$. For that purpose,
-we measure $u$ at some set of locations $\x_j \in [0,\pi], j=1,
-\dots,m$ and a final time $T > 0$. We assume centered Gaussian
-observation error, so we observe $v(\x_j,T) = u(\x_j,T) + \eps(\x_j)$
-with $\eps(\x_j) \sim \normal(0, \sigma^2), \sigma > 0$ iid. We model
-the initial condition as $u_0 \sim \normal(0,\prcov)$, for $\prcov =
-(-\Delta)^{-1}$ with a homogeneous Dirichlet boundary condition. If It
-is well known \cite{Tarantola05} that for linear problems, with
-Gaussian prior and error, the posterior is also Gaussian with a
-covariance that does not depend on the observed data. The posterior
-covariance $\postcov$ is known to have a closed form formula, even in
-infinite dimensions\cite{Stuart10}. We denote by $\fwd$ the dynamics
-operator, so that $u( \cdot,T) = \fwd u_0$, and the observation
-operator $\obs$ so that $u(\x_j,T) = (\obs u)_j, j=1,\dots,m$. The
-posterior covariance is known and depends only on $\prcov, \fwd, \obs$
-and $\sigma^2$ (see section \ref{section:prelim} and
-\eqref{eq:postcov} specifically).
-
-We consider generalization of the information-theoretic design
-criterion presented in the introduction to infinite dimensions
-(section \ref{subsec:D optimal design} below). We choose
-$\x_j,j=1,\dots,m$ to minimize (an expression analogous to) the
-determinant of the posterior covariance operator. We will see later
-how this corresponds to maximizing expected information gain.
-
-The clusterization effect is illustrated in figure
-\ref{fig:clusterization illustration}. Posterior pointwise standard
-deviations are plotted over the domain $[0, \pi]$. Since the posterior
-covariance does not depend on data, the plot has no reference to
-actual data observed. The posterior covariance does, however, depend
-on location of the observation taken. In figure
-\ref{fig:clusterization illustration}, observation locations are
-marked by circles. These were chosen to minimize (an expression
-analogous to) the determinant of the posterior covariance. The
-clusterization effect can be clearly seen for six sensors. It looks
-like only four observations were taken. The reason is that two pairs
-of sensors are so close they are indistinguishable.
-
-\section{Preliminaries and Notation}\label{section:prelim}
-
-\subsection{Overview}
-In this section we define notations that will be used throughout the
-paper. The theoretical foundations for inverse problems over function spaces
-can be found in \cite{Stuart10}.
-
-\subsection{Bayesian Linear Inverse Problems}\label{subsec:abstract OED}
-Let $\hilp$ and $\hilo$ be separable Hilbert spaces (the subscripts p
-and o are for ``parameter'' and ``observation'', respectfully). We
-denote $\hilo^*$ the Hilbert space of all linear functionals on
-$\hilo$. Consider $\fwd: \hilp \to \hilo$, a linear operator (the
-``forward operator''). We are interested in forward operators that are
-strongly smoothing (have fast decaying modes --- the heat operator
-from section \ref{subsec:example} is a prime example). Our goal is to
-infer $\param \in \hilp$ --- some parameter of the dynamics --- given
-noisy observations of $\fwd \param$. We take a Gaussian prior $\pr$
-for $\param$. Hence, $\param \sim \pr = \normal(\prmean ,\prcov)$ with
-some appropriate covariance operator $\prcov$ \cite{Stuart10}. It is
-important to note that $\fwd \prcov \fwd^*$ is the prior covariance in
-$\hilo$ \cite{Stuart10}. As such, it is assumed invertible --- an
-assumption we rely on below. Throughout this paper, we denote $m$ the
-number of observations we are allowed to take. Observations are taken
-via the operator $\obs \in ( \hilo^* )^m$. In an analogy with linear
-algebra, where row vectors are thought of as linear functionals, we
-think of $\obs$ as having ``rows'' $\meas_j, j=1,\dots,m$:
-\begin{equation}\label{eq:O}
-  \obs = (\meas_1,\dots, \meas_m)^t, \meas_j \in \hilo^*, j = 1,\dots,m.
-\end{equation}
-This way, for $u \in \hilo$ we have $\obs u = (\meas_1(u), \dots,
-\meas_m(u) )^t \in \R^m$.
-%% but we may drop the parentheses and write $\meas_j u = \meas_j(u)$.
-%% A few observations regarding $\obs$ are in order.
-%% First, it is good to keep in mind that $( \hilo^* )^m$ is a Hilbert
-%% space with norm $\| \obs \| = \sum_{j=1}^m \|\meas_j\|$.
-For $u \in \hilo$ and $v\in \R^m$:
-\begin{align*}
-  \big (\obs^*v \big ) (u) &= \langle v, \obs u \rangle_{\R^m} = \sum_{j=1}^m  v_j \meas_j(u)
-  = v^t \left ( \obs u \right ) = (v^t \obs) (u),
-\end{align*}
-and thus:
-\begin{align}\label{eq:obs*}
-  \obs^*v &= \sum_{j=1}^m v_j \meas_j = v^t \obs.
-\end{align}
-
-%% \begin{observation}
-%%   It is best to think of $\meas_j,j=1,\dots,m$ as row vectors and of
-%%   $\obs$ as a matrix with rows $\meas_j$. 
-%% \end{observation}
-
-Each observation is a linear functional, chosen from some subset of
-$\hilo^{*}$. Data is acquired via noisy observations
-\begin{align}\label{eq:inverse problem}
-  \data := \obs (\fwd \param + \eps') + \eps = \obs \fwd \param + \obs \eps' + \eps,
-\end{align}
-with $\data \in \R^m$ and $\eps, \eps'$ defined next. We consider two
-types of error terms. First, there is spatially correlated model error
-$\eps' \sim \normal(0,\modcov)$, modeled as a centered Gaussian
-measure on $\hilo$ with covariance operator $\modcov$. Then there is
-observation error $\eps \sim \normal(0, \sigma^2 I_m)$, with $I_m$ the
-$m \times m$ identity. Both error terms and the prior are assumed
-independent of each other. The observation and forward operators are
-typically merged $\tmp := \obs \fwd$
-\cite{AlexanderianGloorGhattas14}. Since we need to separate the
-observation operator $\obs$ from the forward operator $\fwd$ in order
-to carry out the analysis in the following sections, we explicitly
-separate $\fwd$ and $\obs$ in \eqref{eq:inverse problem}. We now turn
-to understand the terms present in \eqref{eq:inverse problem}.
-
-\subsubsection{Dynamics and Observation Operators}\label{subsec:dynamics}
-Rows of $\obs$ are linear functionals and we refer to these as {\it
-  observation vector}s. Note that we cannot choose every observation
-vector we want. For example, we may be restricted only to pointwise
-evaluations of $\fwd \param$ by the sensors at our disposal, in which
-case measuring, e.g., the mean $\ell(u) = \int_{\Omega}u$ of a
-function is not possible. Thus, it is best to think of $(\obs \fwd
-\param)_j$ as approximations to pointwise evaluations. For example, if
-$\hilo = L^2([-\pi, \pi])$, consider $\ell_1(x) := (2\pi
-\eps)^{-\frac{1}{2}}\exp(-\frac{x^2}{2\eps^2})$ and $\ell_2(x)
-:=\eps^{-1}\mathbf{1}_{[-\frac{\eps}{2}, \frac{\eps}{2}]}$. Both
-$\ell_1$ and $\ell_2$ approximate a point measurement at $0$ for small
-$\eps$ and, crucially, $\ell_1, \ell_2 \in \hilo^*$. It is important
-to note that although $\|\ell\|_1 = \|\ell_2\|_1=1$, unfortunately
-$\|\ell_1\|_2^2 = (4\pi \eps)^{-\frac{1}{2}}$ and $\|\ell_2\|_2^2 =
-\eps^{-1}$.
-
-\subsubsection{Error Terms}
-Considering the sum of the error terms, it is easy to see that
-$\bar{\eps} := \obs \eps' + \eps \in \R^m$. It is a centered Gaussian
-random vector and its covariance matrix is
-\begin{align}\label{eq:Sigma}
-  \begin{split}
-    \Sigma(\obs) :&= \mathbb{E}[ (\obs \eps' + \eps)  (\obs \eps' + \eps)^t ] 
-    % 
-    % 
-    = \obs \modcov \obs^* + \sigma^2I , 
-  \end{split}
-\end{align}
-where
-\begin{equation}\label{eq:modcov explained}
-  [\obs \modcov \obs^*]_{ij} = e_i^t \obs \modcov \obs^* e_j = \meas_i (\modcov \meas_j),
-\end{equation}
-with $e_j \in \R^m$ the $j$th standard basis vector. The first
-equality in \eqref{eq:modcov explained} holds by definition and the
-second by \eqref{eq:obs*}. The explicit dependence on $\obs$ will be
-mostly dropped for notational convenience, so $\Sigma(\obs) =
-\Sigma$. Thus, for a fixed $\obs$ (i.e.\ a fixed set of observations)
-we can equivalently write $\data = \obs \fwd \param + \bar{\eps}$ with
-$\bar{\eps} \sim \normal(0,\Sigma)$. Taking $\modcov = 0$ is common
-practice \cite{Tarantola05,KaipioSomersalo05,Vogel02}. This reduces to
-the case where we take iid observations. Then $\Sigma = \sigma^2I$ is
-simply a scalar matrix that does not depend on $\obs$. Note that
-taking an error model with a non-scalar covariance as we do here
-allows us to consider model error (modeled by $\modcov$) as well as
-observation error (modeled by $\sigma^2$). For example, say we believe
-our forward model does not capture some small scale phenomenon.  Then
-we may express this belief by saying $\tru = \fwd + \err$, with $\fwd$
-depending on $\param$ and $\err$ depending on $\sspar$, with $\param
-\perp \sspar$. We do not know much about this effect but it is
-reasonable to assume it changes continuously in our domain. We (may
-choose to) model it as $\normal (0, \modcov)$ and take $\modcov$ to
-reflect the spatial (or other) variability we imagine $\err \sspar$
-has. Such small scale phenomenon can arise as a modeling issue, where
-we might not model the system in its entirety. It can also arise from
-a numerical source, where our discretization of the system is not fine
-enough to capture all small scale phenomena. In section
-\ref{section:non vanishing}, we will see that assuming some
-correlation in the error mitigates the clusterization phenomenon, as
-reported in the literature \cite{Ucinski05}.
-
-Finally, it is useful to record a known result regarding the posterior
-covariance operator of our inverse problem:
-\begin{align}\label{eq:postcov}
-  \postcov = (\prcov^{-1} + \fwd^* \obs^* \Sigma^{-1} \obs \fwd
-  )^{-1}.
-\end{align}
-
-\subsection{D-Optimal Designs in Infinite Dimensions}\label{subsec:D optimal design} 
-A D-optimal design maximizes expected KL divergence between posterior
-and prior. It is useful to recall the definition of KL divergence for
-an arbitrary prior measure:
-$$
-KL(\post||\pr) = \int \log \frac{\der \post}{\der \pr}(\param) \der \post(\param).
-$$
-
-The meaning of D-optimal design in infinite-dimensional Hilbert spaces
-was investigated in \cite{AlexanderianGloorGhattas14}. There, the
-authors make assumptions that amount to $\Sigma=I$ (implied by
-$\modcov = 0,\sigma^2=1$), but we choose not to take these
-simplification here. This is because $\Sigma$ can determine ``how
-much'' clusterization we see. As stated
-\cite[pp. 681]{AlexanderianGloorGhattas14}, the results hold for more
-general covariance matrices. The conclusion is that in
-infinite-dimensions, a D-optimal design is well-defined as maximizing
-the expected KL divergence between posterior and prior. The main
-result is summarized, using our notation, in the following theorem:
-\begin{theorem}[Slightly modified Theorem 1 from \cite{AlexanderianGloorGhattas14}]
-  Let $\pr = \normal(\prmean,\prcov)$ be a Gaussian prior on $\hilp$
-  and let $\post = \normal(\postmean,\postcov)$ the posterior measure
-  on $\hilo$ for the Bayesian linear inverse problem $\data = \obs
-  \fwd\param + \obs \eps' + \eps$ discussed above. Then
-  \begin{align}\label{eq:objective}
-    \begin{split}
-      \tar( \obs) :&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
-      % 
-      % 
-      % 
-      &= \frac12 \log \det 
-      ( I + \prcov^{1/2}  \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}).
-    \end{split}
-  \end{align}
-\end{theorem}
-\begin{definition}\label{def:d optimality}
-  A design $\obs^{\star}$ is said to be D-optimal if $\obs^{\star} = \argmax_{\obs} \tar(\obs)$.
-\end{definition}
-
-\noindent The intuition behind Definition \ref{def:d optimality} is
-straightforward. From \eqref{eq:postcov} and \eqref{eq:objective}:
-\begin{align*}
-  \begin{split}
-    \det ( I + \prcov^{1/2}  \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}) &= \det \Big( \prcov ( \prcov^{-1} + \fwd ^* \obs^* \Sigma^{-1} \obs \fwd) \Big )\\
-    &= \det \prcov \det \postcov^{-1}.
-  \end{split}
-\end{align*}
-Since $\prcov$ is constant, a D-optimal design minimizes the posterior
-covariance determinant, analogously to the finite-dimensional case.
-
-%% \subsection{Sequential vs Simultaneous Optimization}\label{subsec:seq vs sim}
-%% From defintion \ref{def:d optimality} we wish to characterize solution(s) of the
-%% following optimization problem for $\tar$. %%: (\hilo^*)^m \to \R$:
-%% \begin{align}\label{eq:optimization}
-%%   \obs^{\star} := \argmax_{\obs} \tar( \obs ) 
-%%   = \argmax_{\obs} \frac12 \log \det 
-%%   (I + \prcov^{1/2} \fwd^*\obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}),
-%% \end{align}
-%% where $\obs$ is constrained to some allowed set of observations. We
-%% call this problem ``simultaneous optimization'', since all
-%% observations are decided on simulatneously.
-
-%% For computational reasons, one may prefer to find the best
-%% observations in a sequential manner. Denote
-%% \begin{equation}\label{eq:def obs_k}
-%%   \obs_k := (\meas_1,\dots, \meas_k)^t,  k\leq m.
-%% \end{equation}
-%% Sequential optimal design proceeds as follows. Find $\meas_1$ by
-%% maximizing $\tar(\obs_1)$. Then, keeping $\meas_1$ fixed --- find
-%% $\meas_2$ as the maximizer of $\tar(\obs_2)$. Then, find $\meas_3$ by
-%% keeping $\meas_1,\meas_2$ fixed and taking $\meas_3$ as the maximizer
-%% of $\tar(\obs_3)$. Continue this way until $\obs_m = \obs$ is
-%% found, where $m$ is the number of available observations. %% It is
-%% %% important to notice that this scheme does not require actually
-%% %% observing data --- in \eqref{eq:objective} data is averaged out.
-
-%% The analysis in this paper is conducted for the general simultaneous
-%% optimization case. The sequential optimization case is dealt with in
-%% section \ref{subsec:clusterization sequential}. It is important to
-%% note, however that all conclusions we arrive at for the simultaneous
-%% case easily specialize to the sequential case by considering the
-%% posterior as the next sequential step's prior.
-
-\section{The Constrained Optimization Problem of D-Optimal Design}\label{section:D and grad}
-
-\subsection{Overview}
-We seek a formulation of the D-optimal design problem
-\eqref{eq:optimization} using Lagrange multipliers. In Section
-\ref{section:objective} we find the gradient of $\tar$ and record it
-in \eqref{eq:tar grad}. In Section \ref{subsec:unit norm} we suggest
-using unit-norm constraints on rows of $\obs$ so that the optimization
-problem \eqref{eq:optimization} is analytically tractable. We find
-gradients for the new constraints and record them in \eqref{eq:grad
-  constraints}. Finally, in Section \ref{subsec:necessary} we
-formulate the D-optimal design problem as a Lagrange multipliers
-problem in equation \eqref{eq:conditions}.
-
-\subsection{The Objective and its Gradient}\label{section:objective}
-In order to use Lagrange multipliers, we need to find the gradient of
-$\tar$. Said gradient is recorded in \eqref{eq:tar grad} below. We
-start by recording Lemmas \ref{lemma:lax}, \ref{lemma:aux calc},
-proved in the appendix:
-
-\begin{restatable*}[Generalized from \cite{Lax97}]{lemma}{lax}\label{lemma:lax}
-  Let $Y(t)$ be a differentiable operator-valued function. Assume 
-  $I+Y(t)$ is invertible, $Y(t)$ self-adjoint and trace-class. Then
-  \begin{equation*}
-    \frac{\der \log \det (I+Y(t))}{\der t} = \tr{(I+Y(t))^{-1} \dot{Y}(t)}.
-  \end{equation*}
-\end{restatable*}
-
-\begin{restatable*}[Auxilliary Calculations]{lemma}{aux}\label{lemma:aux calc}
-  Let $T(\obs) := \obs^* \Sigma^{-1}(\obs)\obs$, with $\Sigma(\obs)$
-  defined as in \eqref{eq:Sigma}. Then:
-  \begin{align*}
-    \delta T(\obs)V &= V^* \Sigma^{-1} \obs 
-    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
-    &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs
-    + \obs^* \Sigma^{-1} V.
-  \end{align*}
-\end{restatable*}
-
-The first variation of $\tar (\obs)$ in the direction $V$ is:
-\begin{align}\label{eq:tar var}
-  \begin{split}
-    \delta \tar(\obs) V 
-    :&= \frac{\der}{\der\tau} \Big |_{\tau=0} \tar(\obs + \tau V)\  \text{ (by definition of variation)}\\
-    % 
-    % 
-    % 
-    &= \frac12 \frac{\der}{\der \tau} \Big |_{\tau=0} \log \det 
-    (I + \prcov^{1/2} \fwd^* T(\obs+\tau V)\fwd \prcov^{1/2} ) \text{ (by \eqref{eq:objective})}\\
-    % 
-    % 
-    % 
-    &= \frac12 \tr{( I + \prcov^{1/2} \fwd^* \obs^* \Sigma^{-1}
-    \obs\fwd \prcov^{1/2} )^{-1}
-    \frac{\der}{\der \tau} \Big |_{\tau=0}
-    \prcov^{1/2} \fwd^* T(\obs+\tau V) \fwd \prcov^{1/2}}\ \text{ (by \ref{lemma:lax})}\\
-    % 
-    % 
-    % 
-    &= \frac12 \ttr\Big \{ \postcov \fwd^* (V^* \Sigma^{-1} \obs 
-    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
-    &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs 
-    + \obs^* \Sigma^{-1} V ) \fwd \Big \}  \text{ (by \ref{lemma:aux calc})}\\
-    %
-    %
-    %
-    &= \tr{\postcov \fwd^* ( \obs^* \Sigma^{-1} V -
-    \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs ) \fwd} \text{
-        (cyclic property of trace)} \\
-    %
-    %
-    % 
-    &= \tr{\postcov \fwd^* \obs^* \Sigma^{-1} V 
-    ( I - \modcov \obs^* \Sigma^{-1}\obs ) \fwd} \\
-    % 
-    %
-    %
-    &= \tr{V ( I - \modcov \obs^* \Sigma^{-1}\obs )
-    \fwd \postcov \fwd^* \obs^* \Sigma^{-1}}.
-  \end{split}
-\end{align} 
-Denote
-\begin{align}\label{eq:tar grad}
-  \nabla \tar(\obs) &:= (I - \modcov \obs^* \Sigma^{-1} \obs) \fwd
-  \postcov \fwd^* \obs^*\Sigma^{-1},
-\end{align}
-the gradient of $\tar (\obs )$ and we now justify this definition.
-Since the trace of $A := V \nabla \tar(\obs) \in \R^{m \times m}$ is
-just $\ttr A = \sum_{j=1}^m e_j^t A e_j$ (with $e_j$ the $j$th standard
-basis vector), we see that
-\begin{align*}
-  \delta \tar(\obs)V = \tr{V \nabla \tar(\obs)} = \sum_{j=1}^m
-  V_j(\nabla \tar(\obs)_j),
-\end{align*}
-with $V_j \in \hilo^*$ and $\nabla \tar(\obs)_j \in \hilo^{**} =
-\hilo, j=1,\dots,m$. Thus, $\nabla \tar( \obs ) \in \hilo^m$ is indeed
-the correct gradient and the notation \eqref{eq:tar grad} is
-justified. The gradient $\nabla \tar(\obs)$ should be viewed as a row
-vector, as the product $V \nabla \tar$ is in $\R^{m \times m}$. This
-will prove important in section \ref{subsec:necessary}.
-
-\subsection{Unit Norm Constraints and their Gradients}\label{subsec:unit norm}
-In this section we suggest relaxed constraints in
-\eqref{eq:constraints}, and their gradient in \eqref{eq:grad
-  constraints}.
-
-As mentioned before, we cannot choose any $\obs$ when maximizing
-$\tar(\obs)$. Recall $\obs = (\meas_1,\dots,\meas_m)^t \in
-(\hilo^*)^m$ and each $\meas_j,j=1,\dots,m$ must be chosen from some
-allowed set of functionals in $\hilo^*$. %% This set differs based on
-%% the kind of sensors we have at our disposal and the properties of
-%% $\hilo$, as discussed in section \ref{subsec:dynamics}.
-Proposition \ref{prop:bigger better} below (proof in the appendix)
-gives better understanding of the constraints.
-\begin{restatable*}{proposition}{biggerbetter}\label{prop:bigger better}
-  Let $\obs = (\meas_1,\dots,\meas_m)^t$, $j \in \{1,\dots,m\}$, $\sigma^2
-  > 0$ and $\lambda > 1$. Then $\tar(\obs)$ increases if we use
-  $\lambda \meas_j$ in $\obs$ instead of $\meas_j$.
-\end{restatable*}
-The idea behind Proposition \ref{prop:bigger better} is simple: making
-the norm of rows of $\obs$ large is equivalent to making $\sigma^2$
-small. %% Viewed another way, it increases the Signal to Noise Ratio
-%% (SNR) --- which is always desirable when taking any observation
-%% prone to error.  This can be understood easily from the formulation
-%% of the inverse problem $\data = \obs \fwd \param + \obs \eps' +
-%% \eps$, where $\eps'$ can be taken as model error and $\eps$ is iid
-%% observation error.
-
-Note that for point evaluations, the norm of the observation
-functional is always one:
-\begin{align*}
-  \| \delta_{\x} \| = \sup_{0 \neq u \in C(\Omega)} \frac{
-    |\int_{\Omega}u(\y) \delta_{\x}(\y) \der \y| 
-  }{
-    \sup|u|}
-  = \sup_{0 \neq u \in C(\Omega)} \frac{|u(\x)|}{ \sup|u|} = 1,
-  \forall \x \in \Omega.
-\end{align*}
-Following this, and in light of Proposition \ref{prop:bigger better},
-we consider observation vectors in the unit ball in $\hilo^*$. The
-unit norm constraints can be written using \eqref{eq:obs*} as a series
-of $m$ equality constraints (one for each observation) on $\obs$ as
-\begin{align}\label{eq:constraints}
-  \phi_j(\obs) :=\frac12 \| \obs^* e_j\|_{\hilp}^2 - \frac12 = 0,j=1,\dots,m.
-\end{align}
-The first variations are:
-\begin{align*}
-  \delta \phi_j(\obs)V  
-  &= \frac12\lim_{\tau \to 0}\tau^{-1}
-  ( \|(\obs + \tau V)^*e_j \|_{\hilp}^2 - \|\obs ^*e_j \|_{\hilp}^2  ) \\
-  %
-  %
-  %
-  &= \frac12\lim_{\tau \to 0}\tau^{-1}
-  ( \langle (\obs + \tau V)^*e_j, (\obs + \tau V)^*e_j \rangle_{\hilp} - 
-  \langle \obs^*e_j, \obs^*e_j \rangle_{\hilp} ) \\
-  % 
-  % 
-  %
-  &= \frac12\lim_{\tau \to 0}\tau^{-1}
-  (2\tau \langle \obs^*e_j,V^*e_j \rangle_{\hilp} 
-  +\tau^2 \langle V^*e_j, V^*e_j \rangle_{\hilp} ) \\
-  %
-  %
-  % 
-  &= \langle \obs^*e_j,V^*e_j \rangle_{\hilp} \\
-  %
-  %
-  % 
-  &= \langle V \obs^*e_j,e_j \rangle_{\R^m} \\
-  %
-  %
-  %
-  &= e_j^t V \obs^* e_j \\
-  % 
-  %
-  %
-  &= \tr{V \obs^* e_je_j^t}.
-\end{align*}
-The same arguments justifying \eqref{eq:tar grad} hold here, and thus, by \eqref{eq:obs*}:
-\begin{align}\label{eq:grad constraints}
-\nabla \phi_j(\obs) = \obs^* e_j e_j^t = \meas_j e_j^t , j=1,\dots,m,
-\end{align}
-where $\nabla \phi_j(\obs) \in \hilo^m$. As noted at the end of
-Section \ref{section:objective}, the gradient $\nabla \phi_j(\obs)$ is
-best thought of as a row vector.
-
-\subsection{Necessary Conditions for Optimal Design}\label{subsec:necessary}
-Necessary conditions for D-optimality are found from Lagrange
-multipliers:
-\begin{align}
-  &\nabla \tar(\obs) = \sum_{j=1}^m \xi_j \nabla \phi_j (\obs)
-  \label{eq:Lagrange mult1} \\
-    &\phi_j(\obs) = 0, j = 1,\dots,m. \label{eq:Lagrange mult2}
-\end{align}
-
-Recall where the objects involved are defined: the observation
-operator, $\obs \in (\hilo^*)^m$. The objective $\tar: (\hilo^*)^m \to
-\R$ and its gradient is $\nabla \tar(\obs) \in \hilo^m$. Similarly,
-$\phi_j: (\hilo^*)^m \to \R$ and $\nabla \phi_j(\obs) \in \hilo^m$.
-
-Substituting \eqref{eq:tar grad} and \eqref{eq:grad constraints} in
-\eqref{eq:Lagrange mult1}:
-\begin{equation*}
-  (I - \modcov \obs^* \Sigma^{-1} \obs) \fwd \postcov \fwd^* \obs^*\Sigma^{-1}
-  = \sum_{j=1}^m \xi_j \obs^* e_je_j^t = (\xi_1 \meas_1,\dots,\xi_m \meas_m).
-\end{equation*}
-Let $\Xi := \diag(\xi_j)$. Then, with some abuse of notation, this can
-be written more compactly as:
-\begin{equation}\label{eq:conditions}
-  ( I - \modcov \obs^* \Sigma^{-1} \obs) \fwd \postcov \fwd^* \obs^*  \Sigma^{-1}
-  = \obs^* \Xi.
-\end{equation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SECTION Analysis of Optimal Designs --- Non-Vanishing Model Error
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Analysis of D-Optimal Designs --- Non-Vanishing Model Error}\label{section:non vanishing}
-In this section we show the effect model error has on the
-clusterization phenomenon. We will see that if $\modcov \neq 0$
-clusterization will not occur. This is contrary to the previous case
-of vanishing model error.
-
-Corollary \ref{cor:same meas} below is proved in the appendix. Denote
-$\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm :=
-(\meas_1,\dots,\meas_{m-1})^t$. Also, denote $\Sigmam := \Sigma (\obsm)$ and
-$\postcovm$ the posterior covariance that arises when we use $\obsm$.
-\begin{restatable*}{corollary}{samemeas}\label{cor:same meas}
-  If $\meas_m = \meas_j$ for some $1 \leq j \leq m-1$, then
-  \begin{equation*}
-    \tar(\obs) - \tar(\obsm) =
-    \log \left ( 1 + \frac{\sigma^2
-      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
-      \obsm^* \Sigmam^{-1}e_j \rangle
-    }{
-      2 - \sigma^2 e_j^t\Sigmam^{-1}e_j 
-    }       
-    \right ).
-  \end{equation*}
-\end{restatable*}
-Recall from \eqref{eq:Sigma} that $\Sigma(\obs) = \obs \modcov \obs^*
-+ \sigma^2I$. Then
-
-$$
-\lim_{\sigma^2 \to 0} \tar(\obs) -\tar(\obsm) = 0
-$$
-
-Hence no increase in the design criterion is achieved by taking a
-repeated observation (in the limit $\sigma^2 \to 0$). This is clearly
-sub-optimal and $\obs$ cannot give rise to a D-optimal design. Thus,
-for small observation error levels, the clusterization effect will be
-mitigated by the presence of a non-zero model error. Since the design
-criterion is not defined for $\sigma^2 = 0$ and identical
-observations, we cannot make a statement regarding $\sigma^2 = 0$,
-except in the limiting sense described above.
+\input{intro}
+\input{prelims}
+\input{constrained}
+\input{nonvanishing}
+\input{vanishing}
 
@@
 Courant Institute. I would like to thank him for his great
-mentorship. This work was supported by The Raymond and Beverly Sackler
-Post-Doctoral Scholarship.
+mentorship.
 
-\input{vanishing}
+This material is based upon work supported by the National Science
+Foundation Graduate Research Fellowship Program under Grant No. (NSF
+grant number). Any opinions, findings, and conclusions or
+recommendations expressed in this material are those of the author(s)
+and do not necessarily reflect the views of the National Science
+Foundation. This work was also supported by The Raymond and Beverly
+Sackler Post-Doctoral Scholarship.
+
 \input{appendix}
@@
 \bibliographystyle{amsplain}
-\bibliography{refs,georg_refs}
+\bibliography{/home/yair/projects/library.bib}
 

### 2023-03-29 latex/prelims.tex
Paper almost donegit status !

--- 2020-11-22 paper.tex
@@
-\documentclass{amsart}
-
-\input{definitions}
-
-\numberwithin{equation}{section}
-
-\begin{document}
-
-\title[Sensor Clusterization in D-optimal design in infinite
-  dimensions]{Sensor Clusterization in D-optimal Design in Infinite
-  Dimensional Bayesian Linear Inverse Problems}
-
-\author{Yair Daon}
-\address{Porter School of the Environment and Earth
-  Sciences, Tel Aviv University\\ Tel Aviv, Israel}
-%% \curraddr{Porter School of Environment Studies, Tel Aviv University\\ Tel Aviv, Israel}
-%% \curraddr{}
-\email{yair.daon@gmail.com}
-%% \thanks{}
-
-%    The 2010 edition of the Mathematics Subject Classification is
-%    the current definitive version.
-\subjclass{Primary: 
-  62F15, % Statistics - bayesian inference
-  35R30, % PDE - Inverse problems
-  Secondary:
-  28C20, % Measure and integration - set functions and measures and integrals in
-  %infinite-dimensional spaces (Wiener measure, Gaussian measure, etc.
-}
-%% Primary: 62F15, 35R30; Secondary: 28C20
-\date{\today}
-
-\begin{abstract}
-  We investigate the problem of sensor clusterization in D-optimal
-  experimental design for infinite-dimensional Bayesian linear inverse
-  problems. We suggest an analytically tractable model for D-optimal
-  designs and reason how it may lead to sensor clusterization in the
-  case of iid measurement noise. We also spatially correlated
-  measurement and model errors mitigate the clusterization
-  phenomenon. As a part of the analysis we prove a Matrix Determinant
-  Lemma in Hilbert spaces and a lemma for calculating
-  $\frac{\der}{\der t} \log \det (I + X(t))$ for certain
-  operator-valued functions. We also show how to decompose $M = AA^t$
-  with $A$ constrained to have unit norm columns.
-\end{abstract}
-
-\maketitle
-
-\section{Introduction}\label{section:OED intro}
-Experimental design is an important part of many scientific
-investigations. When considering an inverse problem, one can often
-specify sensor locations (e.g.\ in geophysics and oceanography
-applications), certain wavelengths (e.g.\ in MRI) or wave reflections
-from the ground (e.g.\ searching for oil or using a radar). Whatever
-the allowed set of measurements is, one should select the optimal
-measurements to take, in order to increase accuracy, reduce costs, or
-both.
-
-Designing experiments is usually done by optimizing some \emph{design
-  criterion}. This is true both for frequentists
-\cite{Silvey13,Ucinski05} as well as for Bayesians
-\cite{ChalonerVerdinelli95}. See \cite{ChalonerVerdinelli95} for an
-investigation of the analogy between the two approaches. Although
-there is a plethora of design criteria, we focus on just one of these,
-commonly referred to as \emph{D-optimal design}. It has a simple and
-appealing motivation in the Bayesian context as explained in
-\cite{ChalonerVerdinelli95}: Consider a linear model in finite
-dimensions, with Gaussian prior and noise. In this model, maximizing
-the expected information gain (KL divergence
-\cite{KullbackLeibler51,CoverThomas12}) between posterior and prior
-amounts to minimizing the determinant of the posterior covariance
-matrix. In a frequentist setting, a D-optimal design minimizes the
-volume of the uncertainty ellipsoid \cite[page 16]{Ucinski05}, but
-this is done for the Fisher information matrix and not the posterior
-covariance. However, \cite{ChalonerVerdinelli95} shows that the latter
-is just a regularized version of the former.
-
-The previous discussion is classical for experimental design when
-inference takes place over a finite (not too large) number of
-parameters. The subject of optimal experimental design for function
-inference in a Bayesian context was pioneered by
-\cite{AlexanderianGloorGhattas14, AlexanderianPetraStadlerEtAl16,
-  AlexanderianPetraStadlerEtAl14}. Similarly to the finite dimensional
-case, it can be shown that a D-optimal design arises naturally for
-linear models when one wishes to maximize the KL divergence between
-posterior and prior. This amounts to minimizing the determinant of the
-posterior covariance operator (understood as a product of its
-eigenvalues). Some difficulties arise in the process, but remedies can
-be found as shown in \cite{AlexanderianGloorGhattas14}.
-
-It seems counter intuitive that when one computes an optimal design
-using the D criterion, the optimization process results in
-measurements that are very similar. For example, if a measurement is
-thought of as measuring some function value at $\x \in \Omega
-\subseteq \R^d, d=1,2,3$ (with added error) then the optimization
-procedure sometimes places sensors in very close proximity to each
-other (as can be seen in figure \ref{fig:clusterization
-  illustration}). Following \cite{Ucinski05}, we refer to this
-phenomenon as \emph{sensor clusterization}.
-
-\subsection{Related Work}
-The phenomenon of sensor clusterization seems to be known in several
-different contexts. In a frequentist and finite-dimensional context,
-\cite{Fedorov96} and \cite[chapter 2.4.3]{Ucinski05} discuss this
-phenomenon and suggest an approach called clusterization-free design.
-In such designs, the user enforces measurement locations to be far
-from each other. One way to do this is by introducing correlated
-errors which, philosophically, accounts for both measurement error and
-model error. Another method considered is imposing distance
-constraints between measurements. A somewhat different approach is
-suggested in \cite[page 49]{Fedorov12}, where close by design
-measurements are merged --- a procedure which obviously does not avoid
-clusterization. The same problem arises in time-series analysis for
-pharmacokinetic experiments. The authors of \cite{Hooker09} suggest
-modeling auto-correlation time in the noise model, which is equivalent
-to the correlated errors mentioned above.
-
-Any of the above mentioned approaches might serve as a remedy and push
-sensors away from each other. Yet, none offers any insight as to why
-clusterization occurs. Also, as better models are employed, model
-error is decreased and the clusterization phenomenon will eventually
-reappear. While these approaches are practical and help us avoid the
-problem, they do not provide insight as to why sensors are clustering.
-
-Similarly, in the inverse problems community, work is mostly
-computational and less theoretic. Model errors were considered in
-\cite{Attia20, Koval20}. The former study is focused on inferring a
-Quantity of Interest (QoI). The focus of the latter is reducing
-forward solves, using randomized linear algebra. Both studies present
-numerical techniques for finding optimal designs when model error is
-present. Both are restricted to linear inverse problems (although in
-the latter the authors use their method on a nonlinear problem by
-taking a Laplace approximation for the posterior). Both find an
-optimal design by first solving a continuous problem for sensor
-weights. Said solution is then sparsified to give a binary
-design. Both studies are successful in the task of Bayesian
-inversion. However, neither of these studies mention any effect model
-errors can have on sensor clusterization. The current study is mostly
-theoretical and aims to fill the gap of understanding sensor
-clusterization.
-
-\subsection{Contribution}
-We propose and thoroughly study a relaxed and analytically tractable
-model for understanding D-optimal designs. In this model, D-optimal
-design are solutions of a constrained optimization problem, formulated
-using Lagrange multipliers (section \ref{section:D and grad}). Using
-this model we rigorously show how model error mitigates clusterization
-(section \ref{section:non vanishing}). We show how clusterization
-occurs in a sequential design scenario (section
-\ref{subsec:clusterization sequential}). Then, we consider a
-simultaneous D-optimal design scenario (section
-\ref{subsec:clusterization simultaneous}). We show that designs that
-exhibit clusterization are just as optimal as clusterization free
-designs.
-
-A beautiful mathematical structure arises in D-optimal designs when no
-model error is present. The Lagrange multipliers problems is in fact a
-nonlinear eigenvalue problem for the observations and prior
-covariance. The operator for which eigenvectors and eigenvalues are
-sought is a sum of two operators. The first is the prior
-covariance. The second is an outer product of the observations (see
-section \ref{section:vanishing} for details and exact statement). This
-structure helps us shed light on D-optimal designs. We characterize
-these designs in Theorem \ref{thm:char}. One insight we prove is that
-a D-optimal design reduces uncertainties where they are highest
-first. Other interestig phenomena arise, but they require setting
-notation, and are discussed later.
-
-In the process we generalize several lemmas from linear algebra to
-infinite-dimensional settings. We prove a Matrix Determinant Lemma in
-\ref{lemma:MDL}. We generalize a lemma due to Lax \cite{lax97} for
-calculating $\frac{\der}{\der t} \log \det (I + X(t))$, for an
-operator valued function $X(t)$ in \ref{lemma:lax}. We also prove
-lemmas in linear algebra. One constructs a decomposition $M = AA^t$
-where $A$ has unit norm columns \ref{lemma:free}. Another shows
-simultaneous diagonizability of the prior and outer product in the
-nonlinear eigenvalue problem referenced above \ref{lemma:sim diag}. We
-also provide other tools for understanding D-optimal designs in
-infinite-dimensional Bayesian inverse problems. These includet tools
-for calculating the increase in the design criterion per observation
---- Lemma \ref{lemma:design increase} and Corollary \ref{cor:zero mod
-  err}.
-
-\subsection{Limitations}\label{subsec:limitations}
-There are two main drawbacks of the study presented here. Our relaxed
-model does not consider any specific set of allowed
-measurements. Rather, we take measurements in the unit ball in some
-Hilbert space. This allows considerably less restrictive observations
-than any real-life problem does. The second drawback is that we do not
-show rigorously that clusterization necessarily occurs in a
-simultaneous design. We only show that it is as reasonable as no
-clusterization.
-
-\subsection{An Example of Clusterization}\label{subsec:example}
-\begin{figure}
-  \begin{tikzpicture}[thick, scale=1.3, every node/.style={scale=0.99}]
-    \begin{axis}
-      [
-      title={Posterior Pointwise Standard Deviations and D-Optimal Sensor Locations},  
-      xmin = 0,
-      xmax = 3.14,
-      xlabel = {$x$},
-      ylabel = posterior std,
-      ymin   = 0,
-      %compat = 1.3,
-      % ymax   = 130,
-      % ytick = \empty,
-      legend cell align=left,
-      % legend style={at={(0.45,0.2)}}
-      legend pos= outer north east 
-      ]
-      % \draw[black!30!white, thin] (50,0) -- (50,130);
-      % 
-      %% \addplot [thin, black, mark=none] table{stdv-heat-sens1-var1.txt};
-      %% \addlegendentry{1 sensors};
-
-      %% \addplot [thin, blue, mark=none] table{stdv-heat-sens2-var1.txt};
-      %% \addlegendentry{2 sensors};
-
-      %% \addplot [thin, red, mark=none] table{stdv-heat-sens3-var1.txt};
-      %% \addlegendentry{3 sensors};
-
-      \addplot [thin, green, mark=none] table{stdv-heat-sens4-var1.txt};
-      \addlegendentry{4 sensors};
-
-      \addplot [thin, purple, mark=none] table{stdv-heat-sens5-var1.txt};
-      \addlegendentry{5 sensors};
-
-      \addplot [thin, cyan, mark=none] table{stdv-heat-sens6-var1.txt};
-      \addlegendentry{6 sensors};
-
-      %% \addplot [black,  only marks, mark=x, mark size=1.5] 
-      %% table{locs-heat-sens1-var1.txt}; 
-      %% \addplot [blue,   only marks, mark=x, mark size=1.5]
-      %% table{locs-heat-sens2-var1.txt}; 
-      %% \addplot [red,    only marks, mark=x, mark size=1.5]
-      %% table{locs-heat-sens3-var1.txt};
-      \addplot [green,  only marks, mark=*, mark size=1.5] 
-      table{locs-heat-sens4-var1.txt}; 
-      \addplot [purple, only marks, mark=*, mark size=1.5] 
-      table{locs-heat-sens5-var1.txt}; 
-      \addplot [cyan,   only marks, mark=*, mark size=1.5] 
-      table{locs-heat-sens6-var1.txt};
-
-    \end{axis}
-  \end{tikzpicture}
-  \caption{The clusterization effect for the 1D heat equation
-    described in section \ref{subsec:example}. Posterior pointwise
-    standard deviations (lines) are plotted over the domain $[0,
-      \pi]$, for varying numbers of sensors. Sensor locations
-    (circles) were chosen to minimize (an expression analogous to) the
-    determinant of the posterior covariance. The clusterization effect
-    can be clearly seen for six sensors. Only four measurement
-    locations are used --- two pairs of sensors are so close they are
-    indistinguishable.}
-  \label{fig:clusterization illustration}
-\end{figure}
-
-In section \ref{section:prelim} we present a more abstract and general
-formulation of the inverse problem we consider. But, for the purpose
-of illustration, we present clusterization via a toy model --- the 1D
-heat equation in $[0,\pi]$ with a homogeneous Dirichlet boundary
-condition.
-
-The 1D heat equation is:
-\begin{subequations}\label{eq:heat equation}
-  \begin{alignat}{2}
-    u_t &= \Delta u &&\qquad \text{in } [0,\pi] \times [0,\infty),\\
-      u &= 0 &&\qquad \text{on } \{0, \pi\} \times [0,\infty),\\
-        u &= u_0 &&\qquad \text{on }[0,\pi] \times \{0\}.
-  \end{alignat}
-\end{subequations}
-
-We would like to infer the initial condition $u_0$. For that purpose,
-we measure $u$ at some set of locations $\x_j \in [0,\pi], j=1,
-\dots,m$ and a final time $T > 0$. We assume centered Gaussian
-measurement error, so we observe $v(\x_j,T) = u(\x_j,T) + \eps(\x_j)$
-with $\eps(\x_j) \sim \normal(0, \sigma^2), \sigma > 0$ iid. We model
-the initial condition as $u_0 \sim \normal(0,\prcov)$, for $\prcov =
-(-\Delta)^{-1}$ with a homogeneous Dirichlet boundary condition. It is
-well known \cite{Tarantola05} that for linear problems, with Gaussian
-prior and error, the posterior is also Gaussian with a covariance that
-does not depend on the observed data. The posterior covariance
-$\postcov$ is known to have a closed form formula, even in infinite
-dimensions\cite{Stuart10}. We denote by $\fwd$ the dynamics operator,
-so that $u( \cdot,T) = \fwd u_0$, and the observation operator $\obs$
-so that $u(\x_j,T) = (\obs u)_j, j=1,\dots,m$. The posterior
-covariance is known and depends only on $\prcov, \fwd, \obs$ and
-$\sigma^2$ (see section \ref{section:prelim} and \eqref{eq:postcov}
-specifically).
-
-We consider generalization of the information-theoretic design
-criterion presented in the introduction to infinite dimensions
-(section \ref{subsec:D optimal design} below). We choose
-$\x_j,j=1,\dots,m$ to minimize (an expression analogous to) the
-determinant of the posterior covariance operator. We will see later
-how this corresponds to maximizing expected information gain.
-
-The clusterization effect is illustrated in figure
-\ref{fig:clusterization illustration}. Posterior pointwise standard
-deviations are plotted over the domain $[0, \pi]$. Since the posterior
-covariance does not depend on data, the plot has no reference to
-actual data observed. The posterior covariance does, however, depend
-on location of the measurement taken. In figure
-\ref{fig:clusterization illustration}, measurement locations are
-marked by circles. These were chosen to minimize (an expression
-analogous to) the determinant of the posterior covariance. The
-clusterization effect can be clearly seen for six sensors. It looks
-like only four measurements were taken. The reason is that two pairs
-of sensors are so close they are indistinguishable.
-
 \section{Preliminaries and Notation}\label{section:prelim}
@@
 with $\data \in \R^m$ and $\eps, \eps'$ defined next. We consider two
-types of noise / error terms. First, there is spatially correlated
-model error $\eps' \sim \normal(0,\modcov)$, modeled as a centered
-Gaussian measure on $\hilo$ with covariance operator $\modcov$. Then
-there is measurement / independent error $\eps \sim \normal(0,
-\sigma^2 I_m)$, with $I_m$ the $m \times m$ identity. Both error terms
-and the prior are assumed independent of each other. The observation
-and forward operators are merged $\tmp := \obs \fwd$. Since we need to
-separate the observation operator $\obs$ from the forward operator
-$\fwd$ we write this explicitly in \eqref{eq:inverse problem}. We now
-turn to understand the terms present in \eqref{eq:inverse problem}.
+types of error terms. First, there is spatially correlated model error
+$\eps' \sim \normal(0,\modcov)$, modeled as a centered Gaussian
+measure on $\hilo$ with covariance operator $\modcov$. Then there is
+observation error $\eps \sim \normal(0, \sigma^2 I_m)$, with $I_m$ the
+$m \times m$ identity. Both error terms and the prior are assumed
+independent of each other. The observation and forward operators are
+typically merged $\tmp := \obs \fwd$
+\cite{AlexanderianGloorGhattas14}. Since we need to separate the
+observation operator $\obs$ from the forward operator $\fwd$ in order
+to carry out the analysis in the following sections, we explicitly
+separate $\fwd$ and $\obs$ in \eqref{eq:inverse problem}. We now turn
+to understand the terms present in \eqref{eq:inverse problem}.
 
 \subsubsection{Dynamics and Observation Operators}\label{subsec:dynamics}
-One can \textit{think} of $(\obs \fwd \param)_j$ as pointwise
-evaluations of a continuous function\footnote{Of course, this is
-  impossible if, e.g., $\hilo = L^2(\Omega)$ for some $\Omega$.}. Such
-measurements can be represented using the linear functionals
-$\delta_{\x_j} \in C(\Omega)^*$. In a different setting, we may be
-able to measure other quantities, e.g.\ some Fourier coefficients as
-is the case in MRI applications. Either way, the rows of $\obs$ are
-linear functionals and we refer to these as {\it measurement
-  vector}s. Note that we cannot choose every measurement vector we
-want. For example, we may be restricted only to pointwise evaluations
-of $\fwd \param$ by the sensors at our disposal, in which case
-measuring, e.g., the mean $\ell(u) = \int_{\Omega}u$ of a function is
-not possible.
+Rows of $\obs$ are linear functionals and we refer to these as {\it
+  observation vector}s. Note that we cannot choose every observation
+vector we want. For example, we may be restricted only to pointwise
+evaluations of $\fwd \param$ by the sensors at our disposal, in which
+case measuring, e.g., the mean $\ell(u) = \int_{\Omega}u$ of a
+function is not possible. Thus, it is best to think of $(\obs \fwd
+\param)_j$ as approximations to pointwise evaluations. For example, if
+$\hilo = L^2([-\pi, \pi])$, consider $\ell_1(x) := (2\pi
+\eps)^{-\frac{1}{2}}\exp(-\frac{x^2}{2\eps^2})$ and $\ell_2(x)
+:=\eps^{-1}\mathbf{1}_{[-\frac{\eps}{2}, \frac{\eps}{2}]}$. Both
+$\ell_1$ and $\ell_2$ approximate a point measurement at $0$ for small
+$\eps$ and, crucially, $\ell_1, \ell_2 \in \hilo^*$. It is important
+to note that although $\|\ell\|_1 = \|\ell_2\|_1=1$, unfortunately
+$\|\ell_1\|_2^2 = (4\pi \eps)^{-\frac{1}{2}}$ and $\|\ell_2\|_2^2 =
+\eps^{-1}$.
 
 \subsubsection{Error Terms}
-Considering the sum of the error terms it is easy to see that
+Considering the sum of the error terms, it is easy to see that
 $\bar{\eps} := \obs \eps' + \eps \in \R^m$. It is a centered Gaussian
@@
 mostly dropped for notational convenience, so $\Sigma(\obs) =
-\Sigma$. Thus, for a fixed $\obs$ (i.e.\ a fixed set of measurements)
+\Sigma$. Thus, for a fixed $\obs$ (i.e.\ a fixed set of observations)
 we can equivalently write $\data = \obs \fwd \param + \bar{\eps}$ with
 $\bar{\eps} \sim \normal(0,\Sigma)$. Taking $\modcov = 0$ is common
-practice \cite{Tarantola05,KaipioSomersalo05,Vogel02}. This reduces to
+practice \cite{Tarantola05,Kaipio2006,Vogel02}. This reduces to
 the case where we take iid observations. Then $\Sigma = \sigma^2I$ is
@@
 result is summarized, using our notation, in the following theorem:
-\begin{theorem}[Slightly modified Theorem 1 from \cite{AlexanderianGloorGhattas14}]
+\begin{theorem}[Slightly modified Theorem 1 from \cite{AlexanderianGloorGhattas14}]\label{thm:d optimality}
   Let $\pr = \normal(\prmean,\prcov)$ be a Gaussian prior on $\hilp$
@@
 
-The intuition behind definition \ref{def:d optimality} is
-straightforward. From \eqref{eq:objective} and \eqref{eq:objective}:
+\noindent The intuition behind Definition \ref{def:d optimality} is
+straightforward. From \eqref{eq:postcov} and \eqref{eq:objective}:
 \begin{align*}
@@
 
-\subsection{Sequential vs Simultaneous Optimization}\label{subsec:seq vs sim}
-From defintion \ref{def:d optimality} we wish to characterize solution(s) of the
-following optimization problem for $\tar$. %%: (\hilo^*)^m \to \R$:
-\begin{align}\label{eq:optimization}
-  \obs^{\star} := \argmax_{\obs} \tar( \obs ) 
-  = \argmax_{\obs} \frac12 \log \det 
-  (I + \prcov^{1/2} \fwd^*\obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}),
-\end{align}
-where $\obs$ is constrained to some allowed set of measurements (we
-may \textit{think} of these as pointwise evaluations, even though
-these might not be in $\hilo^*$). We call this problem ``simultaneous
-optimization'', since all measurements are decided on simulatneously.
+%% \subsection{Sequential vs Simultaneous Optimization}\label{subsec:seq vs sim}
+%% From defintion \ref{def:d optimality} we wish to characterize solution(s) of the
+%% following optimization problem for $\tar$. %%: (\hilo^*)^m \to \R$:
+%% \begin{align}\label{eq:optimization}
+%%   \obs^{\star} := \argmax_{\obs} \tar( \obs ) 
+%%   = \argmax_{\obs} \frac12 \log \det 
+%%   (I + \prcov^{1/2} \fwd^*\obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}),
+%% \end{align}
+%% where $\obs$ is constrained to some allowed set of observations. We
+%% call this problem ``simultaneous optimization'', since all
+%% observations are decided on simulatneously.
 
-For computational reasons, one may prefer to find the best
-measurements in a sequential manner. Denote
-\begin{equation}\label{eq:def obs_k}
-  \obs_k := (\meas_1,\dots, \meas_k)^t,  k\leq m.
-\end{equation}
-Sequential optimal design proceeds as follows. Find $\meas_1$ by
-maximizing $\tar(\obs_1)$. Then, keeping $\meas_1$ fixed --- find
-$\meas_2$ as the maximizer of $\tar(\obs_2)$. Then, find $\meas_3$ by
-keeping $\meas_1,\meas_2$ fixed and taking $\meas_3$ as the maximizer
-of $\tar(\obs_3)$. Continue this way until $\obs_m = \obs$ is
-found, where $m$ is the number of available measurements. It is
-important to notice that this scheme does not require actually
-observing data --- in \eqref{eq:objective} data is averaged out.
+%% For computational reasons, one may prefer to find the best
+%% observations in a sequential manner. Denote
+%% \begin{equation}\label{eq:def obs_k}
+%%   \obs_k := (\meas_1,\dots, \meas_k)^t,  k\leq m.
+%% \end{equation}
+%% Sequential optimal design proceeds as follows. Find $\meas_1$ by
+%% maximizing $\tar(\obs_1)$. Then, keeping $\meas_1$ fixed --- find
+%% $\meas_2$ as the maximizer of $\tar(\obs_2)$. Then, find $\meas_3$ by
+%% keeping $\meas_1,\meas_2$ fixed and taking $\meas_3$ as the maximizer
+%% of $\tar(\obs_3)$. Continue this way until $\obs_m = \obs$ is
+%% found, where $m$ is the number of available observations. %% It is
+%% %% important to notice that this scheme does not require actually
+%% %% observing data --- in \eqref{eq:objective} data is averaged out.
 
-The analysis in this paper is conducted for the general simultaneous
-optimization case. The sequential optimization case is dealt with in
-section \ref{subsec:clusterization sequential}. It is important to
-note, however that all conclusions we arrive at for the simultaneous
-case easily specialize to the sequential case by considering the
-posterior as the next sequential step's prior.
-
-\section{The Constrained Optimization Problem of D-Optimal Design}\label{section:D and grad}
-
-\subsection{Overview}
-We seek a formulation of the D-optimal design problem
-\eqref{eq:optimization} using Lagrange multipliers. In section
-\ref{section:objective} we find the gradient of $\tar$ and record it
-in \eqref{eq:tar grad}. In section \ref{subsec:unit length} we suggest
-using unit-norm constraints on rows of $\obs$ so that the optimization
-problem \eqref{eq:optimization} is analytically tractable. We find
-gradients for the new constraints and record them in \eqref{eq:grad
-  constraints}. Finally, in section \ref{subsec:necessary} we use
-these to formulate the D-optimal design problem as a Lagrange
-multipliers problem in equation \eqref{eq:conditions}.
-
-\subsection{The Objective and its Gradient}\label{section:objective}
-In order to use Lagrange multipliers, we need to find the gradient of
-$\tar$. This section is mostly technical and devoted to calculating
-said gradient. The result is recorded in \eqref{eq:tar grad}.
-
-We start by calculating the first variation. Some calculations are
-delegated to Lemma \ref{lemma:aux calc} in the appendix. We use
-the following lemma:
-\begin{restatable*}[Generalized from \cite{Lax97}]{lemma}{lax}\label{lemma:lax}
-  Let $Y(t)$ be a differentiable operator-valued function. Assume 
-  $I+Y(t)$ is invertible, $Y(t)$ self-adjoint and trace-class. Then
-  \begin{equation*}
-    \frac{\der \log \det (I+Y(t))}{\der t} = \tr{(I+Y(t))^{-1} \dot{Y}(t)}.
-  \end{equation*}
-\end{restatable*}
-
-The proof of \ref{lemma:lax} is delegated to the appendix. Let
-$T(\obs) := \obs^*\Sigma(\obs)\obs$. The first variation of $\tar
-(\obs)$ in the direction $V$ is:
-\begin{align}\label{eq:tar var}
-  \begin{split}
-    \delta \tar(\obs) V 
-    :&= \frac{\der}{\der\tau} \Big |_{\tau=0} \tar(\obs + \tau V)\  \text{ (by definition of variation)}\\
-    % 
-    % 
-    % 
-    &= \frac12 \frac{\der}{\der \tau} \Big |_{\tau=0} \log \det 
-    (I + \prcov^{1/2} \fwd^* T(\obs+\tau V)\fwd \prcov^{1/2} ) \text{ (by \eqref{eq:objective})}\\
-    % 
-    % 
-    % 
-    &= \frac12 \tr{( I + \prcov^{1/2} \fwd^* \obs^* \Sigma^{-1}
-    \obs\fwd \prcov^{1/2} )^{-1}
-    \frac{\der}{\der \tau} \Big |_{\tau=0}
-    \prcov^{1/2} \fwd^* T(\obs+\tau V) \fwd \prcov^{1/2}}\ \text{ (by \ref{lemma:lax})}\\
-    % 
-    % 
-    % 
-    &= \frac12 \ttr\Big \{ \postcov \fwd^* (V^* \Sigma^{-1} \obs 
-    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
-    &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs 
-    + \obs^* \Sigma^{-1} V ) \fwd \Big \}  \text{ (by \ref{lemma:aux calc})}\\
-    %
-    %
-    %
-    &= \tr{\postcov \fwd^* ( \obs^* \Sigma^{-1} V -
-    \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs ) \fwd} \text{
-        (cyclic property of trace)} \\
-    %
-    %
-    % 
-    &= \tr{\postcov \fwd^* \obs^* \Sigma^{-1} V 
-    ( I - \modcov \obs^* \Sigma^{-1}\obs ) \fwd} \\
-    % 
-    %
-    %
-    &= \tr{V ( I - \modcov \obs^* \Sigma^{-1}\obs )
-    \fwd \postcov \fwd^* \obs^* \Sigma^{-1}}.
-  \end{split}
-\end{align} 
-Denote
-\begin{align}\label{eq:tar grad}
-  \nabla \tar(\obs) &:= (I - \modcov \obs^* \Sigma^{-1} \obs) \fwd
-  \postcov \fwd^* \obs^*\Sigma^{-1},
-\end{align}
-the gradient of $\tar (\obs )$ and we now justify this definition.
-Since the trace of $A := V \nabla \tar(\obs) \in \R^{m \times m}$ is
-just $\ttr A = \sum_{j=1}^m e_j^t A e_j$ (with $e_j$ the $j$th standard
-basis vector), we see that
-\begin{align*}
-  \delta \tar(\obs)V = \tr{V \nabla \tar(\obs)} = \sum_{j=1}^m
-  V_j(\nabla \tar(\obs)_j),
-\end{align*}
-with $V_j \in \hilo^*$ and $\nabla \tar(\obs)_j \in \hilo^{**} =
-\hilo, j=1,\dots,m$. Thus, $\nabla \tar( \obs ) \in \hilo^m$ is indeed
-the correct gradient and the notation \eqref{eq:tar grad} is
-justified. Note that by the comment following \eqref{eq:obs*}, we view
-$V$ (defined on the same space as $\obs$) as a column vector. Then the
-gradient $\nabla \tar(\obs)$ should be viewed as a row vector, as the
-product $V \nabla \tar$ is in $\R^{m \times m}$. This will prove
-important in section \ref{subsec:necessary}.
-
-\subsection{Unit Length Constraints and their Gradients}\label{subsec:unit length}
-In this section we suggest relaxed constraints in
-\eqref{eq:constraints}. We then find their gradient in \eqref{eq:grad
-  constraints}. These are use in the Lagrange multipliers
-problem in section \ref{subsec:necessary}.
-
-As mentioned before, we cannot choose any $\obs$ when maximizing
-$\tar(\obs)$. Recall $\obs = (\meas_1,\dots,\meas_m)^t \in
-(\hilo^*)^m$ and each $\meas_j,j=1,\dots,m$ must be chosen from some
-allowed set of functionals in $\hilo^*$. This set differs based on the
-kind of sensors we have at our disposal and the properties of $\hilo$,
-as discussed in section \ref{subsec:dynamics}. The following
-proposition gives better understanding of the constraints.
-\begin{restatable*}{proposition}{biggerbetter}\label{prop:bigger better}
-  Let $\obs = (\meas_1,\dots,\meas_m)^t$, $j \in \{1,\dots,m\}$, $\sigma^2
-  > 0$ and $\lambda > 1$. Then $\tar(\obs)$ increases if we use
-  $\lambda \meas_j$ in $\obs$ instead of $\meas_j$.
-\end{restatable*}
-Proof of Proposition \ref{prop:bigger better} is delegated to the
-appendix but the idea behind it is simple: making rows of $\obs$ large
-is equivalent to making $\sigma^2$ small. %%  Viewed another way, it
-%% increases the Signal to Noise Ratio (SNR) --- which is always
-%% desirable when taking any measurement prone to error.
-This can be understood easily from the formulation of the inverse
-problem $\data = \obs \fwd \param + \obs \eps' + \eps$, where $\eps'$
-can be taken as model error and $\eps$ is iid measurement error.
-
-We now suggest a set of constraints that will make the model
-analytically tractable. By Proposition \ref{prop:bigger better}, we
-cannot just take all $\hilo^*$ to be the allowed set of measurement
-vectors, since this would effectively eliminate measurement error
-entirely. We note that for point evaluations, the norm of the
-measurement is always one:
-\begin{align*}
-  \| \delta_{\x} \| = \sup_{0 \neq u \in C(\Omega)} \frac{
-    |\int_{\Omega}u(\y) \delta_{\x}(\y) \der \y| 
-  }{
-    \sup|u|}
-  = \sup_{0 \neq u \in C(\Omega)} \frac{|u(\x)|}{ \sup|u|} = 1,
-  \forall \x \in \Omega.
-\end{align*}
-Following this, we also restrict the norm of our measurement vectors
-to be one. The unit norm constraints can be written using
-\eqref{eq:obs*} as a series of $m$ equality constraints (one for each
-measurement) on $\obs$ as
-\begin{align}\label{eq:constraints}
-  \phi_j(\obs) :=\frac12 \| \obs^* e_j\|_{\hilp}^2 - \frac12 = 0,j=1,\dots,m.
-\end{align}
-The first variations are:
-\begin{align*}
-  \delta \phi_j(\obs)V  
-  &= \frac12\lim_{\tau \to 0}\tau^{-1}
-  ( \|(\obs + \tau V)^*e_j \|_{\hilp}^2 - \|\obs ^*e_j \|_{\hilp}^2  ) \\
-  %
-  %
-  %
-  &= \frac12\lim_{\tau \to 0}\tau^{-1}
-  ( \langle (\obs + \tau V)^*e_j, (\obs + \tau V)^*e_j \rangle_{\hilp} - 
-  \langle \obs^*e_j, \obs^*e_j \rangle_{\hilp} ) \\
-  % 
-  % 
-  %
-  &= \frac12\lim_{\tau \to 0}\tau^{-1}
-  (2\tau \langle \obs^*e_j,V^*e_j \rangle_{\hilp} 
-  +\tau^2 \langle V^*e_j, V^*e_j \rangle_{\hilp} ) \\
-  %
-  %
-  % 
-  &= \langle \obs^*e_j,V^*e_j \rangle_{\hilp} \\
-  %
-  %
-  % 
-  &= \langle V \obs^*e_j,e_j \rangle_{\R^m} \\
-  %
-  %
-  %
-  &= e_j^t V \obs^* e_j \\
-  % 
-  %
-  %
-  &= \tr{V \obs^* e_je_j^t}.
-\end{align*}
-Using the same arguments we used to justify the definition of
-\eqref{eq:tar grad}, we conclude that
-\begin{align}\label{eq:grad constraints}
-\nabla \phi_j(\obs) = \obs^* e_j e_j^t = \meas_j e_j^t , j=1,\dots,m,
-\end{align}
-with $\nabla \phi_j(\obs) \in \hilo^m$ and where the
-last equality follows since $\obs^*e_j = \meas_j$ by \eqref{eq:obs*}. As
-we noted at the end of section \ref{section:objective}, the
-gradient $\nabla \phi_j(\obs)$ is a row vector.
-
-\subsection{Necessary Conditions for Optimal Design}\label{subsec:necessary}
-Necessary conditions for D-optimality are found using Lagrange
-multipliers:
-\begin{align}
-  &\nabla \tar(\obs) = \sum_{j=1}^m \xi_j \nabla \phi_j (\obs)
-  \label{eq:Lagrange mult1} \\
-    &\phi_j(\obs) = 0, j = 1,\dots,m. \label{eq:Lagrange mult2}
-\end{align}
-
-Recall where the objects involved are defined: the observation operator,
-$\obs \in (\hilo^*)^m$. The objective $\tar:
-(\hilo^*)^m \to \R$ and its gradient is $\nabla \tar(\obs) \in
-\hilo^m$. The constraints are defined similarly --- $\phi_j: (\hilo^*)^m
-\to \R$ and $\nabla \phi_j(\obs) \in \hilo^m$.
-
-Substituting \eqref{eq:tar grad} and \eqref{eq:grad constraints} in
-\eqref{eq:Lagrange mult1} we have:
-\begin{equation*}
-  (I - \modcov \obs^* \Sigma^{-1} \obs) \fwd \postcov \fwd^* \obs^*\Sigma^{-1}
-  = \sum_{j=1}^m \xi_j \obs^* e_je_j^t = (\xi_1 \meas_1,\dots,\xi_m \meas_m).
-\end{equation*}
-If we let $\Xi := \diag(\xi_j)$, then with some abuse of
-notation, we can write this more compactly as:
-\begin{equation}\label{eq:conditions}
-  ( I - \modcov \obs^* \Sigma^{-1} \obs) \fwd \postcov \fwd^* \obs^*  \Sigma^{-1}
-  = \obs^* \Xi.
-\end{equation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SECTION Analysis of Optimal Designs --- Vanishing Model Error
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Analysis of D-Optimal Designs --- Vanishing Model Error}\label{section:vanishing}
-\subsection{Overview}
-In this section we analyze D-optimal designs when no model error is
-present. Sensor clusterization is discussed in section
-\ref{section:clusterization}. We prepare the ground by first deriving
-a nonlinear eigenvalue problem in $\obs$ for D-optimal designs in
-section \ref{subsec:eigenproblem}. This nonlienar eigenvalue problem
-is recorded in \eqref{eq:mod conditions}. Next, we characterize
-D-optimal designs as ones that make all eigenvalues equal in the above
-mentioned nonlinear eigenvalue problem. The derivation is presented in
-section \ref{subsec:characterization} and is summarized in Theorem
-\ref{thm:char}.
-
-\subsection{The Nonlinear Eigenvalue Problem}\label{subsec:eigenproblem}
-$\modcov = 0$ implies $\Sigma= \sigma^2I$. The necessary first-order
-condition for D-optimality \eqref{eq:conditions} becomes
-\begin{equation}\label{eq:eigenproblem}
-  \sigma^{-2}\fwd \postcov \fwd^* \obs^* = \obs^* \Xi, 
-\end{equation}
-with $\Xi$ diagonal. This looks like an eigenvalue problem for the
-self-adjoint operator $\sigma^{-2}\fwd \postcov \fwd^*$. The rows of
-$\obs$, namely $\meas_j,j=1,\dots, m$, are the eigenvectors. Note,
-however, that $\postcov$ depends on $\obs$. This makes
-\eqref{eq:eigenproblem} a nonlinear eigenvalue problem.
-
-Below, the expression for the posterior covariance \eqref{eq:postcov}
-justifies the second equality. Some calculations delegated to Lemma
-\ref{lemma:twice woodbury} in the appendix justify the third (it is
-trivial if $\fwd$ is invertible):
-\begin{align}\label{eq:mod conditions}
-  \begin{split}
-  \obs^* \Xi &= \sigma^{-2}\fwd \postcov \fwd^* \obs^*  \\
-  %
-  %
-  %
-  &= \sigma^{-2} \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* \obs^* \\
-  %
-  %
-  %
-  &= \sigma^{-2} \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1} \obs^*
-  \end{split}
-\end{align}  
-The term $(\fwd \prcov \fwd^*)^{-1}$ is the prior precision in $\hilo$
-($\fwd\prcov\fwd^*$ is invertible, as mentioned in section
-\ref{section:prelim}). $(\fwd\prcov\fwd^*)^{-1} + \sigma^{-2} \obs^*
-\obs$ is the posterior precision in $\hilo$.
-
-\subsection{Characterizing D-Optimal Designs}\label{subsec:characterization}
-Denote eigenvalues of $\fwd \prcov \fwd^*$ by
-$\{\lambda_i\}_{i=1}^{\infty}$. These are all positive. Denote
-corresponding eigenvectors $\{\ev_i\}_{i=1}^{\infty}$. Lemma
-\ref{lemma:sim diag} (proved in the appendix) is crucial for the rest
-of the analysis:
-
-\begin{restatable*}[Simultaneous diagonizability for the nonlinear eigenvalue problem]{lemma}{simdiag}\label{lemma:sim diag}
-  Let $\hil$ separable Hilbert space, $C:\hil \to \hil$ self-adjoint
-  and $\func_1,\dots,\func_m \in \hil$. Denote $\func^*$ the element
-  $\func$ acting as a linear functional. If
-  \begin{equation*}
-   (C + \sum_{j=1}^m \func_j\func_j^*) \func_l = \xi_l \func_l, l = 1,\dots,m
-  \end{equation*}
-  then $C$ and $\sum_{j=1}^m \func_j \func_j^*$ are simultaneously
-  diagonalizable.
-\end{restatable*}
-
-\begin{corollary}[Eigenvectors of a D-optimal design]\label{cor:same ev}
-  $\obs^*\obs$ has the same eigenvectors as $\fwd \prcov \fwd^*$.
-  %namely $\{\ev_i\}_{i=1}^{\infty}$.
-\end{corollary}
-\begin{proof}
-In \eqref{eq:mod conditions}, take $\func_j^{*} = \meas_j$ and $C :=
-(\fwd \prcov \fwd^*)^{-1}$ and use Lemma \ref{lemma:sim diag}.
-\end{proof}
-At most $m$ of the eigenvectors of $\obs^*\obs$ have a positive
-eigenvalue, because $k := \rank \obs^*\obs \leq m$. Denote the
-corresponding non-zero eigenvalues of $\obs^*\obs$ by
-$\{\eta_i\}_{i=1}^{k}$ and let $\eta_i = 0$ for $i \geq k+1$. No
-assumption is made (yet) regarding the magnitudes of the corresponding
-$\{\lambda_i\}_{i=1}^{\infty}$. They are to be thought of as fixed but
-unknown. Using Sylvester's Determinant Theorem (the standard proof
-works for Hilbert spaces as well) we get:
-
-\begin{align}\label{eq:pre manipulations}
-  \begin{split}
-    \det \left (I + \sigma^{-2} \prcov^{1/2} \fwd ^* \obs^*
-    \obs \fwd \prcov^{1/2}\right ) &=
-    %
-    \det \left (I + \sigma^{-2} \obs^*
-    \obs \fwd \prcov\fwd^* \right ) \\
-    %
-    %
-    %
-    &= \det \left \{ \left ( (\fwd \prcov \fwd^*)^{-1} +
-    \sigma^{-2}\obs^*\obs \right ) (\fwd \prcov \fwd^*)\right \}.\\
-  \end{split}
-\end{align}
-
-By definition of the objective \eqref{eq:objective}, \eqref{eq:pre
-  manipulations} and Corollary \ref{cor:samve ev}:
-\begin{align}\label{eq:mod objective}
-  \begin{split}
-    \tar ( \obs ) &=\frac12 \log \det \left( I + \sigma^{-2}
-    \prcov^{1/2} \fwd ^* \obs^* \obs \fwd \prcov^{1/2} \right )\\
-    %
-    %
-    %
-    %
-    &=\frac12 \log \det \Big ( \left ( (\fwd \prcov \fwd^* )^{-1} +
-    \sigma^{-2}\obs^* \obs \right ) \fwd \prcov \fwd^* \Big )\\
-    %
-    %
-    %
-    &=\frac12 \log \left (\prod_{i=1}^{\infty} ( \lambda_i^{-1} + \sigma^{-2} \eta_i ) \prod_{i=1}^{\infty} \lambda_i \right ) \\
-    %
-    %
-    %
-    &=\frac12 \log \left ( \prod_{i=1}^{k} ( \lambda_i^{-1} + \sigma^{-2} \eta_i )\prod_{i=1}^{k} \lambda_i \right )\\
-    %
-    %
-    %
-    &=\frac12 \sum_{i=1}^{k} \log (\sigma^2\lambda_i^{-1} + \eta_i), 
-  \end{split}
-\end{align}
-and the last equality holds up to an additive constant independent of
-$\obs$. Therefore, a D-optimal design maximizes
-
-\begin{equation}\label{eq:true target}
-  \sum_{i=1}^{k}\log (\sigma^2 \lambda_i^{-1} + \eta_i),
-\end{equation}
-
-with the norm constraints on $\meas_j,j=1,\dots,m$. We would like to
-make all $\eta_i$ as large as possible.  The problem of D-optimal
-design has now been cast to finite dimensions. This justifies using
-linear algebra. Lemma \ref{lemma:free} (a constructive proof is given
-in the appendix) does exactly that.
-\begin{restatable*}[Unit norm decomposition]{lemma}{free}\label{lemma:free}
-  Let $M \in \R^{k \times k}$ symmetric positive definite with $\ttr M
-  = m$, $m \geq k$. We can find $\func_j \in \R^k,j=1,\dots,m$
-  with $\|\func_j\|=1$ and $A = (\func_1,\dots,\func_m)$ such that
-  $AA^t = M$.
-\end{restatable*}
-
-Take any $\{\eta_i\}_{i=1}^{k}$, with $\sum_{i=1}^k \eta_i = m$. Let
-$M = \diag(\eta_1,\dots,\eta_k)$. Lemma \ref{lemma:free} gives us $A$
-with unit norm columns such that $AA^T = M$. Take $\meas_i
-:=\sum_{j=1}^k A_{ji}\ev_j^*$. The constraints $\|\meas_i\| = 1,
-i=1,\dots,m$ are satisified by construction. Verifying $\obs^*\obs
-\ev_i = \eta_i \ev_i$ is not too difficult.
-
-%% \begin{align*}
-%%   \begin{split}
-%%     \obs \ev_i &= (\meas_1,\dots,\meas_k)^t\ev_i\\
-%%     &= (\sum_{j=1}^kA_{j1}\ev_j^*,\dots,\sum_{j=1}^kA_{jk}\ev_j^*)^t\ev_i\\
-%%     &= (\sum_{j=1}^kA_{j1}\ev_j^*\ev_i,\dots,\sum_{j=1}^kA_{jk}\ev_j^*\ev_i)^t\\
-%%     &= (A_{i1},\dots,A_{ik})^t.
-%%   \end{split}
-%% \end{align*}
-
-%% Now,
-
-%% \begin{align*}
-%%   \begin{split}
-%%     \obs^*\obs \ev_i &= \obs^*(A_{i1},\dots,A_{ik})^t\\
-%%     &=(A_{i1},\dots,A_{ik}) \obs \text{ (by \eqref{eq:obs*})}\\
-%%     &=(A_{i1},\dots,A_{ik}) (\sum_{j=1}^kA_{j1}\ev_j^*,\dots,\sum_{j=1}^kA_{jk}\ev_j^*)^t\\
-%%     &=\sum_{l=1}^k A_{il} \sum_{j=1}^kA_{jl}\ev_j^* \\
-%%     %&=\sum_{j=1}^k  \sum_{l=1}^kA_{il}A_{jl}\ev_j^* \\
-%%     &=\sum_{j=1}^k  [AA^t]_{ij}\ev_j^* \\
-%%     %&=\eta_i\ev_i^*\\
-%%     &=\eta_i\ev_i^*. 
-%%   \end{split}
-%% \end{align*}
-
-We have shown that we can construct $\obs$ such that it has
-eigenvectors and eigenvalues $\{\ev_i, \eta_i\}_{i=1}^k$, as long as
-$\sum_{i=1}^k \eta_i = m$. We now turn to understand which choice of
-$\{\eta_i\}_{i=1}^k$ is optimal. By concavity of $\log$, the fastest
-increase in the design criterion \eqref{eq:true target} is gained by
-increasing $\sigma^2\lambda_i^{-1} + \eta_i$ where it is smallest. So,
-no weight should be given to (i.e. measurement taken in direction of)
-the $n$th eigenvector before $\sigma^2\lambda_i^{-1} + \eta_i \geq
-\sigma^2\lambda_n^{-1},\ \forall i \neq n$. Thus, we should choose
-$\{\eta_i\}_{i=1}^k$ as follows. First, order the eigenvalues
-$\lambda_1 \geq \lambda_2 \geq \dots$. We should increase $\sigma^2
-\lambda_1^{-1} + \eta_1$ until it equals
-$\sigma^2\lambda_2^{-1}$. Then, $\sigma^2 \lambda_i^{-1} +
-\eta_i,i=1,2$ are increased until $\sigma^2 \lambda_i^{-1} + \eta_i =
-\sigma^2\lambda_3^{-1},i=1,2$ and so forth.
-
-\pgfplotstableread{
-  Label     prior  optimal  sub-optimal 
-  1         0.2    1.8           1.7
-  2         0.8    1.2           0.8
-  3         2.2    0             0.5
-  4         3.5    0             0.0
-}\optimalvsnot
-
-\begin{figure}
-  \begin{tikzpicture}[scale=0.85]
-    \begin{axis}[
-        ybar stacked,
-        ymin=0,
-        ymax=4,
-        xtick=data,
-        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
-        reverse legend=false, % set to false to get correct display, but I'd like to have this true
-        xticklabels from table={\optimalvsnot}{Label},
-        xticklabel style={text width=2cm,align=center},
-        legend plot pos=right,
-        ylabel=precision --- prior and posterior,
-        xlabel=eigenvector $i$,
-      ]
-
-      \addplot [fill=green!80]  table [y=prior,   meta=Label, x expr=\coordindex] {\optimalvsnot};
-      \addplot [fill=blue!60]   table [y=optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};
-
-      \addlegendentry{$\sigma^2\lambda_i^{-1}$}
-      \addlegendentry{optimal $\eta_i$s}
-
-    \end{axis}
-  \end{tikzpicture}
-  \qquad
-  \begin{tikzpicture}[scale=0.85]
-    \begin{axis}[
-        ybar stacked,
-        ymin=0,
-        ymax=4,
-        xtick=data,
-        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
-        reverse legend=false, % set to false to get correct display, but I'd like to have this true
-        xticklabels from table={\optimalvsnot}{Label},
-        xticklabel style={text width=2cm,align=center},
-        legend plot pos=right,
-        ylabel=precision --- prior and posterior,
-        xlabel=eigenvector $i$,
-      ]
-
-      \addplot [fill=green!80]  table [y=prior,       meta=Label, x expr=\coordindex] {\optimalvsnot};
-      \addplot [fill=blue!60]   table [y=sub-optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};
-
-      \addlegendentry{$\sigma^2\lambda_i^{-1}$}
-      \addlegendentry{sub-optimal $\eta_i$s}
-
-    \end{axis}
-  \end{tikzpicture}
-  \caption{Characterizing a D-optimal design, compared to a
-    sub-optimal design. Both designs are allowed three measurements,
-    which means the blue area has accumulated height of $3$ in both
-    plots. A D-optimal design (left) increases precision (reduces
-    uncertainty) where it is lowest. A sub-optimal design (right) does
-    not. In this example, a measurement is taken partly in the
-    direction of the third eigenvector. The precision for this
-    eigenvector is bigger even than the posterior precision of the
-    D-optimal design on the left.}
-  \label{fig:optimal vs not}
-\end{figure}
-
-\pgfplotstableread{
-  Label    prior  $o_1$  $o_2$   $o_3$   $o_4$   $o_5$         $o_6$    topper
-  1         0.2    0.8    0.5     0.5    0.4    0.33333333  0.3333333  0.001
-  2         0.8    0.2    0.5     0.5    0.4    0.33333333  0.3333333  0.001
-  3         2.2    0      0       0      0.2    0.33333333  0.3333333  0.001
-  4         3.5    0      0       0      0      0           0          0.001
-}\clusterization
-
-\pgfplotstableread{
-  Label    prior  $o_1$  $o_2$   $o_3$   $o_4$   $o_5$         $o_6$    topper
-  1         0.2    0.8    0.7     0.3    0.4    0.2         0.4666666  0.001
-  2         0.8    0.2    0.3     0.7    0.4    0.46666666  0.2        0.001
-  3         2.2    0      0       0      0.2    0.33333333  0.3333333  0.001
-  4         3.5    0      0       0      0      0           0          0.001
-}\noclusterization
-
-The argument above has the important consequence (illustrated in
-figure \ref{fig:optimal vs not}) that the eigenvalues in the nonlinear
-eigenvalue problem \eqref{eq:eigenproblem} are all equal! This makes
-perfect sense. We know the Lagrange multipliers $\xi_i,i=1,\dots,k$
-tell us how much we gain by relaxing the constraints and there is no
-reason for any eigenvector to give us more than any other (otherwise
-we would put more weight on the better eigenvector, as can be
-understood from the discussion in the following paragraph).
-
-One can understand D-optimal designs as follows: Imagine each
-eigenvector corresponds to a graduated lab cylinder (figure
-\ref{fig:cylinder}). Each cylinder $i$ is filled, a-priori, with green
-liquid of $\sigma^2\lambda_i^{-1}$ volume units. This liquid
-represnets the prior precision, weighted by observation noise. The
-volume of the green liquid is increasing in $i$. A D-optimal design
-with $m$ measurements ignores cylinders indexed $m+1$ and above. It
-has blue liquid of $m$ volume units at its disposal. It distributes
-said liquid by repeatedly adding a drop to whatever cylinder currently
-has the lowest level of liquid in it (lowest precision). The result of
-such a procedure is illustrated in figure \ref{fig:optimal vs
-  not}. The analysis conducted in this section is summarized in
-Theorem \ref{thm:char} below.
-
-\begin{theorem}[Characterizing D-optimal designs with vanishing model error]\label{thm:char}
-  Let:
-  \begin{enumerate}
-  \item $\fwd, \prcov$ the forward and prior covariance operators, respectively.
-  \item $m, \sigma^2$ the number of measurements taken and observation
-    noise variance, respectively.
-  \item $\obs$ a D-optimal design operator.
-  \item $\tar$ as definded in \eqref{eq:objective}.
-  \item $\{\lambda_i\}_{i=1}^\infty$ eigenvalues of $\fwd\prcov\fwd^*$
-    in decreasing order.
-  \item $\{\ev_i\}_{i=1}^\infty$ their corresponding eigenvectors.
-  \item $\{\eta_i\}_{i=1}^\infty$ eigenvalues of $\obs^*\obs$, where
-    $\obs$ maximizes $\tar(\obs)$.
-  \end{enumerate}
-  Then:
-  \begin{enumerate}
-  \item $\obs^*\obs$ and $\fwd\prcov\fwd^*$ are simultaneously diagonalizable.
-  \item There exists $0 < k \leq m$ such that $\eta_i = 0$ for $i\geq
-    k+1$ and $k = \rank \obs^*\obs$.
-  \item $\sigma^2\lambda_i^{-1} + \eta_i$ are equal for all $1 \leq i \leq k$.
-  \item If $k < m$, then $\sigma^2\lambda_i^{-1} + \eta_i \leq
-    \sigma^2\lambda_{k+1}^{-1}, i=1,\dots,k$.
-  \end{enumerate}
-\end{theorem}
-
-\begin{figure}%{r}{0.25\textwidth} 
-    %\centering
-    \includegraphics[width=5cm, height=5cm]{cylinders.jpg}
-    \caption{Three graduated lab cylinders, corresponding to three
-      eigenvectors. Prior eigenvalues not shown.}
-    \label{fig:cylinder}
-\end{figure}
-
-\section{Sensor Clusterization --- Vanishing Model Error}\label{section:clusterization}
-\subsection{Overview}
-In this section we use Theorem \ref{thm:char} to explain how sensor
-clusterization may arise in various scenarios. We first consider
-sequential designs, and then simultaneous designs.
-
-\subsection{Sequential Design}\label{subsec:clusterization sequential}
-Recall the sequential optimal design scheme from section
-\ref{subsec:seq vs sim} and the definition of $\obs_k$ from
-\eqref{eq:def obs_k}. The way clusterization can occur in sequential
-design scenario is understood by specializing thorem \ref{thm:char} to
-one measurement each time. In this case $m=1$, so $k=1$, $\meas_1
-\parallel \ev_1$ and $\eta_1 = 1$. As hinted in section
-\ref{subsec:seq vs sim}, the posterior becomes the prior for the next
-step. The eigenvectors of the new posterior are the same as for the
-previous one. A new measurement will be taken in the direction of
-eigenvector with smalles eigenvalue. This leads to:
-\begin{align*}
-  \begin{split}
-    \meas_2 \parallel \ev_1  &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 < \sigma^2\lambda^{-1}_2 \\
-    \meas_2 \parallel \ev_2  &\text{ o.w. }
-  \end{split}
-\end{align*}
-In the first case, we get clusterization immediately. In the second,
-we consider $\meas_3$. Same reasoning leads to the conclusion that the
-eigenvectors of the new prior precision are the same. The eigenvalues,
-however, are $\lambda^{-1}_1 + \sigma^{-2}, \lambda^{-1}_2 +
-\sigma^{-2}, \lambda_3^{-1}, \dots$. A new measurement is taken in the
-direction of eigenvector of smallest precision (note that
-$\lambda^{-1}_1 + \sigma^{-2} < \lambda^{-1}_2 + \sigma^{-2}$):
-\begin{align*}
-  \begin{split}
-    \meas_3 \parallel \ev_1 &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 <
-    \sigma^2\lambda^{-1}_3 \\
-    \meas_3 \parallel  \ev_3  &\text{ o.w. }
-  \end{split}
-\end{align*}
-Again, the first case results in clusterization. The sequential design
-proceeds in this fashion. Every $\meas_i$ is taken in the direction
-some eigenvector of the prior $\fwd \prcov\fwd^*$. This eigenvector
-has the smallest eigenvalue of the previous step's posterior precision
-operator. If we do not encounter clusterization after $m$ such steps,
-this means that measurement $\meas_k$ was taken in the direction
-$\ev_k$ for $k=1,\dots,m$. Thus:
-\begin{align}\label{eq:failure}
-  \begin{split}
-    \sigma^2\lambda_{k-1}^{-1} + 1 \geq \sigma^2\lambda_k^{-1}, \text{ for } 1\leq k \leq m
-    \Longrightarrow \lambda_m^{-1} - \lambda_{m-1}^{-1} \leq \sigma^{-2}.
-  \end{split}
-\end{align}
-
-But \eqref{eq:failure} implies $\{\lambda_i^{-1}\}_{i=1}^{\infty}$
-grows at most linearly. In section \ref{subsec:abstract OED} we
-assumed $\fwd$ is strongly smoothing, which means eigenvalues of
-$(\fwd \prcov\fwd^*)^{-1}$ grow quickly. Thus, if we take $m$ large
-enough, \eqref{eq:failure} will eventually fail and we will observe
-sensor clusterization.
-
-\subsection{Simultaneous Design}\label{subsec:clusterization simultaneous}
-In a simultaneous design scenario, the result is not simple as in the
-sequential design case. We now see that a design that exhibits
-clusterization is as optimal as one that does not. Corollary
-\ref{cor:same ev} tells us that we can realize any design we want
-(encoded in $\{\eta_i\}$), given the trace constraint for
-$\obs^*\obs$. Theorem \ref{thm:char} tells us that for D-optimality,
-we only need to make the posterior flat where we take measurements
-(recall figure \ref{fig:optimal vs not}). We are left with the freedom
-of assigning any part of $\eta_i$ to any measurement $\meas_j$. Two
-possible assignments for the same problem are illustrated in figure
-\ref{fig:clusterization}. The first exhibits clusterization, the
-second does not. It is important to note that neither arises as a
-result of some numerical optimization scheme. Both designs give the
-same design criterion $\tar$, because their posteriors are
-identical. By Theorem \ref{thm:char}, both are maximizers of the same
-D-optimal design problem. One exhibits clusterization, and one does
-not. It is likely that a real-life problem is more restricted than the
-model we consider here. Such restriction is the likely cause of
-clusterization obseved in, e.g., the toy model of section
-\ref{subsec:example}. Exactly how this happens is not clear. At this
-point it is the subject of speculation and should be further studied
-in the future.
-
-\begin{figure}
-  \begin{tikzpicture}[scale=0.85]
-    \begin{axis}[
-        ybar stacked,
-        ymin=0,
-        ymax=4,
-        xtick=data,
-        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
-        reverse legend=false, % set to false to get correct display, but I'd like to have this true
-        xticklabels from table={\clusterization}{Label},
-        xticklabel style={text width=2cm,align=center},
-        legend plot pos=right,
-        ylabel=precision --- prior and posterior,
-        xlabel=eigenvector $i$,
-      ]
-
-      \addplot [fill=green!80]  table [y=prior, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=blue!60]   table [y=$o_1$, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=red!60]    table [y=$o_2$, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=black!60]  table [y=$o_3$, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=orange!60] table [y=$o_4$, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=cyan!60]   table [y=$o_5$, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=purple!60] table [y=$o_6$, meta=Label, x expr=\coordindex] {\clusterization};
-
-      \addlegendentry{prior}
-      \addlegendentry{$o_1$}
-      \addlegendentry{$o_2$}
-      \addlegendentry{$o_3$}
-      \addlegendentry{$o_4$}
-      \addlegendentry{$o_5$}
-      \addlegendentry{$o_6$}   
-    \end{axis}
-  \end{tikzpicture}
-  \qquad
-  \begin{tikzpicture}[scale=0.85]
-    \begin{axis}[
-        ybar stacked,
-        ymin=0,
-        ymax=4,
-        xtick=data,
-        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
-        reverse legend=false, % set to false to get correct display, but I'd like to have this true
-        xticklabels from table={\noclusterization}{Label},
-        xticklabel style={text width=2cm,align=center},
-        legend plot pos=right,
-        ylabel=precision --- prior and posterior,
-        xlabel=eigenvector $i$,
-      ]
-
-      \addplot [fill=green!80]  table [y=prior, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=blue!60]   table [y=$o_1$, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=red!60]    table [y=$o_2$, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=black!60]  table [y=$o_3$, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=orange!60] table [y=$o_4$, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=cyan!60]   table [y=$o_5$, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=purple!60] table [y=$o_6$, meta=Label, x expr=\coordindex] {\noclusterization};
-
-      \addlegendentry{prior}
-      \addlegendentry{$o_1$}
-      \addlegendentry{$o_2$}
-      \addlegendentry{$o_3$}
-      \addlegendentry{$o_4$}
-      \addlegendentry{$o_5$}
-      \addlegendentry{$o_6$}   
-    \end{axis}
-  \end{tikzpicture}
-  \caption{Clusterization (left) and non-clusterization (right) in
-    simultaneous D-optimal designs. Posterior precision per
-    eigenvector after taking measurements $\meas_i$ is plotted. Both
-    designs are identical for all practical matters --- their
-    posteriors are identical. In the left panel we see how an optimal
-    design is achieved with repeated measurements. $\meas_2 = \meas_3$
-    and $\meas_5 = \meas_6$. This is not necessary though, as can be
-    seen in the right panel. Both designs achieve the same levels of
-    uncertainty. The left exhibits clusterization. The right does not.}
-  \label{fig:clusterization}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SECTION Analysis of Optimal Designs --- Non-Vanishing Model Error
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Analysis of D-Optimal Designs --- Non-Vanishing Model Error}\label{section:non vanishing}
-In this section we show the effect model error has on the
-clusterization phenomenon. We will see that if $\modcov \neq 0$
-clusterization will not occur. This is contrary to the previous case
-of vanishing model error.
-
-Corollary \ref{cor:same meas} below is proved in the appendix. Denote
-$\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm :=
-(\meas_1,\dots,\meas_{m-1})^t$. Also, denote $\Sigmam := \Sigma (\obsm)$ and
-$\postcovm$ the posterior covariance that arises when we use $\obsm$.
-\begin{restatable*}{corollary}{samemeas}\label{cor:same meas}
-  If $\meas_m = \meas_j$ for some $1 \leq j \leq m-1$, then
-  \begin{equation*}
-    \tar(\obs) - \tar(\obsm) =
-    \log \left ( 1 + \frac{\sigma^2
-      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
-      \obsm^* \Sigmam^{-1}e_j \rangle
-    }{
-      2 - \sigma^2 e_j^t\Sigmam^{-1}e_j 
-    }       
-    \right ).
-  \end{equation*}
-\end{restatable*}
-Recall from \eqref{eq:Sigma} that $\Sigma(\obs) = \obs \modcov \obs^*
-+ \sigma^2I$. Then
-
-$$
-\lim_{\sigma^2 \to 0} \tar(\obs) -\tar(\obsm) = 0
-$$
-
-Hence no increase in the design criterion is achieved by taking a
-repeated measurement (in the limit $\sigma^2 \to 0$). This is clearly
-sub-optimal and $\obs$ cannot give rise to a D-optimal design. Thus,
-for small measurement error levels, the clusterization effect will be
-mitigated by the presence of a non-zero model error. Since the design
-criterion is not defined for $\sigma^2 = 0$ and identical
-measurements, we cannot make a statement regarding $\sigma^2 = 0$,
-except in the limiting sense described above.
-
-\section{Acknowledgements}
-This study is a part of my PhD thesis. Most of it was written under
-the instruction of Prof. Georg Stadler in New York University's
-Courant Institute. I would like to thank him for his great
-mentorship. This work was supported by The Raymond and Beverly Sackler
-Post-Doctoral Scholarship.
-
-\appendix
-\section{General Lemmas}
-
-The following lemma is generalized from \cite[Chapter 9, Theorem 4, pp. 127]{Lax97}
-\lax
-\begin{proof}
-  Consider a differentiable operator-valued function $X(t)$ such that
-  $X(0) = 0$ and $X(t)$ is positive, self-adjoint and trace-class for
-  every $t\in \R$. We denote the eigenvalues of this operator by
-  $\lambda_k(X(t))$ and sometimes drop the dependence on $X(t)$, so
-  $\lambda_k = \lambda_k(X(t))$.  Then $\det (I+X(t)) =
-  \prod_{k=1}^{\infty} (1+\lambda_k) < \infty$ where the bound holds
-  by the arguments given in \cite{AlexanderianGloorGhattas14}. The
-  full derivative is
-  \begin{align*}
-    \frac{\der \det (I+X(t))}{\der t} 
-    % 
-    % 
-    % 
-    &= \sum_{k=1}^{\infty} 
-    \frac{\partial \det (I+X(s))}{\partial (1+\lambda_k)}\Big |_{s=t}
-    \frac{\der (1+\lambda_k)}{\der t} \\
-    % 
-    %
-    %
-    &= \sum_{k=1}^{\infty} \frac{\partial \prod_{l=1}^{\infty}
-      (1+\lambda_l(s))}{\partial (1+\lambda_k)}\Big |_{s=t}
-    \frac{\der (1+\lambda_k)}{\der t} \\
-    %
-    %
-    %
-    &= \sum_{k=1}^{\infty} \prod_{l=1, l\neq k}^{\infty}
-      (1+\lambda_l(s)) \frac{\partial (1+\lambda_k(s))}{\partial (1+\lambda_k)}\Big |_{s=t}
-    \frac{\der (1+\lambda_k)}{\der t} \\
-    %
-    %
-    %    
-    &= \sum_{k=1}^{\infty} \frac{\prod_{l=1}^{\infty}
-      (1+\lambda_l(s))}{(1+\lambda_k)}\Big |_{s=t}
-    \dot{\lambda_k}(X(t)) \\
-    % 
-    % 
-    % 
-    &= \sum_{k=1}^{\infty} \frac{\det (I+X(t))}{1 +\lambda_k} \dot{\lambda_k}(X(t)).
-  \end{align*}
-  The assumption $X(0) = 0$ means $\lambda_k(X(0)) = 0,\ \forall k \geq 1$. Thus:
-  \begin{align*}
-    \frac{\der (I+\det X(t))}{\der t}\Big |_{t=0} 
-    = \sum_{k=1}^{\infty} \dot{\lambda_k}(X(0)) 
-    = \frac{\der }{\der t}\tr{X(0)}
-    = \tr{\dot{X}(0)},
-  \end{align*}
-  where the second equality follows by monotone convergence. 
-  Let $Y(t)$ a trace-class self-adjoint operator such that 
-  $I+Y(t)$ is invertible.
-  Define $X(t)$ via $I+X(t) = (I+Y(0))^{-1/2} (I+Y(t)) (I+Y(0))^{-1/2}$. 
-  We show $X(t)$ satisfies the conditions above. It is trace-class:
-  \begin{align*}
-    \tr{X(t)} = \tr{(I+Y(0))^{-1} (I+Y(t)) - I}
-    \leq \tr{I+Y(t) - I}< \infty,
-  \end{align*}
-  since $Y(t)$ is trace-class. It is also clear that
-  $X(0) = 0$ and $X(t)$ is self-adjoint.
-  $I+Y(t) = (I+Y(0))^{1/2}(I+X(t))(I+Y(0))^{1/2}$, so
-  \begin{align*}
-    \frac{\der \det (I+Y(t))}{\der t}|_{t=0} 
-    &= \det (I+Y(0))\frac{\der \det (I+X(t))}{\der t}\Big |_{t=0} \\
-    % 
-    % 
-    % 
-    &= \det (I+Y(0)) \tr{\dot{X}(0)} \\
-    % 
-    % 
-    % 
-    &= \det (I+Y(0)) \tr{(I+Y(0))^{-1} \dot{Y}(0)}.
-  \end{align*}
-  Consequently, by the one-variable chain rule:
-  \begin{align*}
-    \frac{\der \log \det (I+Y(t))}{\der t}\Big |_{t=0} &=
-    % 
-    % 
-    % 
-    \frac{1}{\det (I+Y(0))}\frac{\der \det (I+Y(t))}{\der t}\Big |_{t=0} \\ 
-    % 
-    % 
-    % 
-    &= \tr{ (I+Y(t))^{-1} \dot{Y}(t)} \big |_{t=0}.
-  \end{align*}
-  There is nothing special about $t_0 = 0$ --- we could have chosen
-  any other $t_0$ instead. Thus, the relation holds for all $t$.
-\end{proof}
-
-\begin{lemma}[Matrix Determinant Lemma in Hilbert Spaces]\label{lemma:MDL}
-  Let $\hil$ a separable Hilbert space, $u,v\in \hil$ and $A: \hil \to
-  \hil$ an invertible linear operator such that $\tr{A-I} <
-  \infty$. Then $\det A$ and $\det A + uv^*$ are well defined and
-  \begin{equation*}
-    \det (A + uv^*) = (1 + \langle A^{-1} u, v \rangle ) \det A,
-  \end{equation*}
-  where $(A + uv^*)w := Aw + \langle v,w \rangle u$.
-\end{lemma}
-\begin{proof}
-  In this proof we rely on definitions and results from
-  \cite{Simon77}. First, consider $B := I + xy^*$ for some $x,y \in
-  \hil$. We construct an eigenbasis for $B$ and use that to show $\det
-  B = 1 + \langle x, y \rangle$. First let $x_1 := x$.  Now, if $x
-  \parallel y$, take $\{x_n \}_{n=2}^{\infty}$ an orthogonal basis for
-  $span\{x_1\} ^{\perp}$. If, on the other hand, $x \nparallel y$, let
-  \begin{equation*}
-    x_2 := x - \frac{ \langle x, y\rangle}{\|y\|^2}y
-  \end{equation*}
-  and it is easy to verify that $x_2 \perp y$ and $span \{x,y\} = span
-  \{x_1,x_2\}$. Take $\{x_n \}_{n=3}^{\infty}$ an orthogonal basis for
-  $span\{x_1,x_2\} ^{\perp}$. In both cases,
-  \begin{equation*}
-    B x_n =
-    \begin{cases}
-      (1 + \langle x, y \rangle) x_n & n = 1 \\
-      x_n                            & n \neq 1,
-    \end{cases}
-  \end{equation*}
-  and so $\det B = 1 + \langle x, y \rangle$.
-
-  It is easy to verify that $uv^*$ is trace-class and since $\tr{A-I}
-  < \infty$, also $\tr{A + uv^* - I} < \infty$ (sum of two trace-class
-  operators is trace-class). Thus $\det A$ and $\det (A+uv^*)$ are
-  well defined. Let $x:=A^{-1}u$ and $y := v$:
-  \begin{equation*}
-    \det (A + uv^*) = \det A \ \det(I+A^{-1}uv^*) =
-    (1 + \langle A^{-1}u, v \rangle) \det A .
-  \end{equation*}
-\end{proof}
-
-\free
-\begin{proof}
-  Let us diagonalize $M$, so that $M = U D U^t$ with $D =
-  \diag(d_1,\dots,d_k)$ and $U \in \R^{k \times k }$ orthogonal. Let
-  $S \in \R^{k \times m}$ with $S_{ii} = \sqrt{d_{i}}$ and zeros
-  otherwise. Define $A:= U S V^t$, where $V \in \R^{m \times m}$ is
-  orthogonal and will be further restricted later. Then $AA^t = U
-  SV^tVS^t U^t = UDU^t$, so $AA^t$ has the required eigenvalues and
-  eigenvectors by construction. If we can choose $V$ such that $A$
-  also satisfies the unit norm constraints we are done. These
-  constraints are, for $j=1,\dots,m$:
-  \begin{equation}\label{eq:V constraints}
-   1 = [A^tA]_{jj} = [V S^tS V^t]_{jj},
-  \end{equation}
-  and we can expect to do this since we assumed $\ttr D = m$.
-
-  Define $C = S^tS - I \in \R^{m \times m}$. Note that $\ttr C = 0$ and
-  $C$ is diagonal with non-zero entries $d_i-1,i=1,\dots,k$. It suffices
-  to find $V$ orthogonal such that $V C V^t$ has zero diagonal. We
-  construct such $V$ by sequentially inserting zeros in the diagonal
-  and not destroying zeros we already introduced, starting from the
-  last diagonal entry and moving to the first. Since $c_{mm} \neq 0$ ,
-  let $p < m$ such that $c_{pp}c_{mm} < 0$ (such $p$ exists because
-  the trace is zero) and let $\theta \in (0,\pi)$. Define a Givens
-  rotation $R^{(m)} \in \R^{m \times m}$ by
-  \begin{equation*}
-    r^{(m)}_{ab} :=
-    \begin{cases}
-      1 & a = b \neq p \text{ or } a = b \neq m \\
-      \cos \theta & a = b = p  \\
-     -\sin \theta & a = p, b = m\\
-      \cos \theta & a = b = m \\
-      \sin \theta & a = m, b = p \\ 
-      0 & o.w
-    \end{cases}
-  \end{equation*}
-  Note that conjugating a matrix by $R^{(m)}$ changes only its $m$ and
-  $p$ rows and columns. We want to choose $\theta$ such that
-  \begin{equation}\label{eq:mm}
-    0 = [R^{(m)} C (R^{(m)})^t]_{mm} = \cos^2 \theta c_{mm} + 2\cos \theta \sin
-    \theta c_{mp} + \sin^2\theta c_{pp},
-  \end{equation}
-  and it suffices to choose $\theta$ such that
-  \begin{equation*}
-    c_{mm} \cot^2 \theta + 2 c_{mp} \cot \theta + c_{pp} = 0.
-  \end{equation*}
-  This quadratic in $\cot\theta$ has a real solution, since
-  $c_{pp}c_{mm} < 0$ by assumption and we can find $\theta \in
-  (0,\pi)$ such that \eqref{eq:mm} is satisfied. We continue to find
-  $R^{(m-1)}$ that leaves row and column $m$ unchanged and
-  continue introducing zeros to the diagonal. The assumption $\ttr D =
-  m \Rightarrow \ttr C = 0$ guarantees we can do that. Taking $V:=
-  R^{(1)} R^{(2)} \dots R^{(m-1)}R^{(m)}$ completes the proof.
-\end{proof}
-
-\simdiag
-\begin{proof}
-  First, enumerate the eigenvalues of $C + \sum_{j=1}^m
-  \func_j\func_j^*$ as $\xi_1,\dots,\xi_\ell$. Denote the
-  indices of the eigenvectors corresponding to $\xi_i$
-  \begin{equation*}
-    S_i := \{ 1 \leq k \leq m | (C + \sum_{j=1}^m \func_j\func_j^* )\func_k = \xi_i \func_k \}.
-  \end{equation*}
-  Define further
-  \begin{equation*}
-    A_i := \sum_{k \in S_i} \func_k \func_k^*,
-  \end{equation*}
-  which is self-adjoint. Two observations are in order. First,
-  $\sum_{j=1}^m \func_j\func_j^* = \sum_{i=1}^\ell A_i$. Second, $A_i
-  \func_k = 0$ if $k\not \in S_i$, since eigenvectors of different
-  eigenvalue are orthogonal. For $k \in S_i$
-  \begin{equation}\label{eq:on vi}
-    \xi_i \func_k = (C + \sum_{j=1}^m \func_j \func_j^* ) \func_k = (C + A_i) \func_k.
-  \end{equation}
-  Let $V_i := span \{\func_k \}_{k\in S_i}$. Observe that $V_i$ is
-  invariant under $A_i$, by definition, and under $C$, by \eqref{eq:on
-    vi}. Using \eqref{eq:on vi} again, we conclude that on $V_i$, $A_i
-  = \xi_iI - C$. This immediately implies $A_i$ and $C$ have the
-  same eigenvectors on $V_i$. This holds for every $1 \leq i \leq
-  \ell$ and we conclude that $C$ and $A$ have the same eigenvectors.
-\end{proof}
-
-\section{Specific Lemmas}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{lemma}\label{lemma:twice woodbury}
-  Assume $\fwd \prcov \fwd^*$ is invertible. Then
-\begin{align*}
-  \begin{split}
-    \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* 
-    %
-    %
-    = \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1},
-  \end{split}
-\end{align*}  
-\end{lemma}
-\begin{proof}
-  The proof is boring and amounts to using Woodbury's matrix identity
-  twice and a regularization trick. The standard proof for this
-  identity works in infinite dimensions, as long as all terms are well
-  defined. Unfortunately, $\obs^*\obs$ is not invertible, so we force
-  it to be. Throughout, we let $X := \fwd\prcov \fwd^*$ and $Y_{\eps}
-  := \obs^*\obs + \eps I$, for some small $\eps > 0$. Note that now
-  $Y_{\eps}$ is invertible. Recall Woodbury's matrix identity:
-  $$
-  (A + UCV)^{-1} = A^{-1} + A^{-1}U(C^{-1} + VA^{-1}U)^{-1}VA^{-1}.
-  $$
-
-  Taking $A := \prcov^{-1}, U := \fwd^*, V := \fwd$ and $C :=
-  \sigma^{-2} (\obs^*\obs+\eps I)$ we get:
-  \begin{align*}
-    \begin{split}
-      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* &=
-      \fwd ( \prcov - \prcov \fwd^* ( \sigma^2(\obs^*\obs + \eps I)^{-1} + \fwd \prcov \fwd^* )^{-1} \fwd \prcov ) \fwd^* \\
-      %
-      %
-      &= X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X
-    \end{split}
-  \end{align*}
-  Note that $X + \sigma^2 Y_{\eps}^{-1}$ is invertible, as the sum of
-  a two positive definite operators. Now, taking $A := X^{-1}, C :=
-  \sigma^2Y_{\eps}^{-1}, U := I$ and $V := I$ and using Wodbury's matrix
-  identity in reverse order:
-  \begin{align*}
-    \begin{split}
-      X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X &= (X^{-1} + \sigma^{-2}Y_{\eps})^{-1} \\
-      %
-      %
-      %
-      &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \eps I))^{-1}.
-    \end{split}
-  \end{align*}
-
-  We conclude that $\forall \eps > 0$
-  \begin{align*}
-    \begin{split}
-      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* 
-     &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \eps I))^{-1}.
-    \end{split}
-  \end{align*}
-  Letting $\eps \to 0$ completes the proof.
-\end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{lemma}[Increase due to a Measurement]\label{lemma:design increase}
-  Let $\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm := (\meas_1,\dots,\meas_{m-1})^t$. Then
-  \begin{align*}
-    \tar( \obs ) - \tar (\obsm ) &=
-    \frac12 \log \left ( 1 + \frac{
-      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m,
-      (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m \rangle
-    }{
-      \sigma^2 + \meas_m \modcov \meas_m - \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m 
-    }       
-    \right ).
-  \end{align*}
-\end{lemma}
-\begin{proof}
-  We use the Schur complement to write one inverse in terms of the other and
-  introduce notations to make the derivation cleaner. Note that we think of
-  $\obsm$ and $\obsm^*$ as column and row vectors (respectively).
-  \begin{align*}
-    \Sigma( \obs ) &= \Sigma = 
-    \begin{bmatrix}
-      \Sigma (\obsm )           & \obsm \modcov \meas_m \\
-      \meas_m \modcov \obsm^*   & \sigma^2 + \meas_m \modcov \meas_m
-    \end{bmatrix}
-    =
-    \begin{bmatrix}
-      \Sigmam   & w \\
-      w^t       & c
-    \end{bmatrix}\\
-    %
-    %
-    %
-    \Sigma^{-1} &=
-    \begin{bmatrix}
-      \Sigmam^{-1} + \Sigmam^{-1} w ( c - w^t \Sigmam^{-1} w)^{-1} w^t \Sigmam^{-1} & - \Sigmam^{-1} w ( c - w^t \Sigmam^{-1} w)^{-1} \\
-      -( c - w^t \Sigmam^{-1} w)^{-1} w^t \Sigmam^{-1}                            &  ( c - w^t \Sigmam^{-1} w)^{-1}
-    \end{bmatrix} \\
-    &=
-    \begin{bmatrix}
-      \Sigmam^{-1} & 0 \\
-      0           & 0 
-    \end{bmatrix}
-    + (c -w^t \Sigmam^{-1} w )^{-1}
-    \begin{bmatrix}
-      \Sigmam^{-1} w \\
-      -1
-    \end{bmatrix}
-    \begin{bmatrix}
-      w^t \Sigmam^{-1} & -1 
-    \end{bmatrix}
-  \end{align*}
-  %
-  Further, define
-  %
-  \begin{align*}
-    \M (\obs ):&= \prcov^{\frac12}\fwd^* \obs^* \Sigma^{-1} \obs \fwd
-    \prcov^{\frac12}    
-  \end{align*}
-  %
-  and note that, using our understanding of what is a column vector and
-  what is a row vector:
-  %
-  \begin{align*}
-    \M(\obs) &= \prcov^{1/2} \fwd^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2} \\
-    %
-    %
-    %
-    &= \prcov^{1/2} \fwd^* \obs^* \left \{
-    \begin{bmatrix}
-      \Sigmam^{-1} & 0 \\
-      0           & 0 
-    \end{bmatrix}
-    + (c -w^t \Sigmam^{-1} w )^{-1}
-    \begin{bmatrix}
-      \Sigmam^{-1} w \\
-      -1
-    \end{bmatrix}
-    \begin{bmatrix}
-      w^t \Sigmam^{-1} & -1 
-    \end{bmatrix} 
-    \right \} \obs \fwd \prcov^{1/2} \\
-    %
-    %
-    %
-    &= \M (\obsm) + (c -w^t \Sigmam^{-1} w )^{-1}
-    \prcov^{1/2} \fwd^* \obs^*
-    \begin{bmatrix}
-      \Sigmam^{-1} w \\
-      -1
-    \end{bmatrix}
-    \begin{bmatrix}
-      w^t \Sigmam^{-1} & -1 
-    \end{bmatrix} 
-    \obs \fwd \prcov^{1/2}
-  \end{align*}
-  %
-  Now, denote:
-  %
-  \begin{align}\label{eq:u}
-    \begin{split}
-      u :&= (c -w^t \Sigmam^{-1} w )^{-1/2}
-      \prcov^{1/2} \fwd^* \obs^* 
-      \begin{bmatrix}
-        \Sigmam^{-1} w \\
-        -1 
-      \end{bmatrix} \\
-      %
-      %
-      %
-      & = (c -w^t \Sigmam^{-1} w )^{-1/2} ( \prcov^{1/2}\fwd^* \obsm^* \Sigmam^{-1} \obsm  \modcov \meas_m - \prcov^{1/2} \fwd^* \meas_m )\\
-      %
-      %
-      %
-      u^* :&=  (c -w^t \Sigmam^{-1} w )^{-1/2} (\meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \fwd \prcov^{1/2} - \meas_m \fwd \prcov^{1/2} ),
-    \end{split}
-  \end{align}
-  %
-  so that
-  %
-  \begin{equation}\label{eq:M plus I}
-    I + \M( \obs ) = I + \M (\obsm ) + uu^*.
-  \end{equation}
-  %
-  Note that
-  \begin{equation}\label{eq:M postcov}
-    \prcov^{1/2} \left (I + \M( \obsm ) \right )^{-1} \prcov^{1/2} = \postcovm.
-  \end{equation}
-  The increase in the design criterion gained by including $\meas_m$
-  in the design is found by using Lemma \ref{lemma:MDL} the above
-  results:
-  %
-  \begin{align*}
-    \tar( \obs ) - \tar( \obsm )
-    %
-    %
-    %
-    &= \frac12 \log \det \Big ( I + \M ( \obs ) \Big ) / \det \Big ( I + \M (\obsm) \Big ) \\
-    %
-    %
-    %
-    &= \frac12  \log \det \left ( I + \M(\obsm) + uu^* \right ) / \det \Big ( I + \M (\obsm) \Big ) \\
-    %
-    %
-    %
-    &= \frac12 \log \left ( 1 + \left \langle \left ( I+\M(\obsm) \right )^{-1} u, u  \right \rangle \right ).
-  \end{align*}
-  Using \eqref{eq:u}:
-  \begin{align*}
-    &\left \langle \left (I+\M (\obsm)\right )^{-1}u, u \right \rangle\\
-    &= \frac{
-      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m,
-      (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m \rangle
-    }{
-      c- w^t \Sigmam^{-1} w
-    }\\
-    %
-    %
-    %
-    &= 
-    \frac{
-      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m,
-      (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m \rangle
-    }{
-      \sigma^2 + \meas_m \modcov \meas_m - \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m 
-    }
-  \end{align*}
-  and the conclusion follows.
-\end{proof}
-
-Lemma \ref{lemma:design increase} implies the following two corollaries.
-\begin{corollary}[Gain for No Model Error]\label{cor:zero mod err}
-  If $\modcov = 0$, then
-  \begin{equation*}
-    \tar( \obs ) - \tar (\obsm )
-    = -\frac12 \log (1 - \sigma^{-2} \langle \fwd \postcov \fwd^* \meas_m, \meas_m \rangle ).
-  \end{equation*}
-\end{corollary}
-\begin{proof}
-  Note that this is not immediate by substituting $\modcov = 0$ in the
-  conclusion of Lemma \ref{lemma:design increase}, since we make a
-  claim for $\postcov$, and not $\postcovm$. Let us first review
-  \eqref{eq:u} and note that since $\modcov = 0$ the covariance
-  $\Sigmam = \sigma^2I_{m-1}$ and $w = 0$, so $c - w^t\Sigmam^{-1}w =
-  \sigma^2$:
-  \begin{align*}
-    u :&=-\sigma^{-1}\prcov^{1/2} \fwd^* \meas_m\\
-    %
-    %
-    u^* :&= -\sigma^{-1} \meas_m \fwd \prcov^{1/2}.
-  \end{align*}
-  From \eqref{eq:M plus I}:
-  \begin{equation*}
-    I + \M(\obsm) = I +\M(\obs) - uu^*,
-  \end{equation*}
-  and thus:
-  \begin{equation*}
-    \left \langle \left ( I +\M(\obs) \right )^{-1} u, u \right \rangle
-    = \sigma^{-2} \langle \fwd \postcov \fwd^* \meas_m, \meas_m \rangle.
-  \end{equation*}
-  Analogously to \eqref{eq:M postcov} we note that
-  \begin{equation*}
-    \prcov^{1/2} \left ( I + \M(\obs) \right )^{-1} \prcov^{1/2} = \postcov.
-  \end{equation*}
-  Using Lemma \ref{lemma:MDL} we conclude
-  \begin{align*}
-    \tar( \obs ) - \tar( \obs )
-    &= \frac12 \log \det \left (I +\M(\obs) \right ) / \det \left (I + \M(\obsm) \right ) \\
-    %
-    %
-    %
-    &= \frac12 \log \det \left (I +\M(\obs) \right ) / \det \left (I + \M(\obs) - uu^* \right ) \\
-    %
-    %
-    %
-    &=-\frac12 \log (1 - \langle (I+\M(\obs))^{-1}u, u \rangle \\
-    %
-    %
-    %
-    &= -\frac12 \log (1 - \sigma^{-2} \langle \fwd \postcov \fwd^* \meas_m, \meas_m \rangle ).
-  \end{align*}
-\end{proof}
-
-\samemeas
-\begin{proof} \label{cor:same meas proof}
-  Denote $A:= \obs \modcov \obs^*$ and $v_j$ the $j$th column of $A$.
-  Note that $v_j = \obsm \modcov \meas_m$, since $(\obsm \modcov
-  \obsm^*)_{ij} = \meas_i(\modcov \meas_j)$, as explained in
-  \eqref{eq:modcov explained}. One can now verify that
-  \begin{equation}\label{eq:observation}
-    \Sigmam^{-1} \obsm \modcov \meas_m = \Sigmam^{-1}v_j = (A +\sigma^2I_{m-1})^{-1} v_j =
-    e_j -\sigma^2 \Sigmam^{-1}e_j.
-  \end{equation}
-  %
-  Using \eqref{eq:observation}:
-  \begin{align}\label{eq:denominator}
-    \begin{split}
-      \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m
-      &= \meas_m \modcov \obsm^* ( e^j - \sigma^2 \Sigmam^{-1} e_j )\\
-      %
-      %
-      %
-      &= \meas_m \modcov \meas_j - \sigma^2 \meas_m \modcov \obsm^* \Sigmam^{-1}e_j \\
-      %
-      %
-      %
-      &= \meas_m \modcov \meas_j -\sigma^2 (e_j - \sigma^2 \Sigmam^{-1}e_j)^t e_j \\
-      %
-      %
-      %
-      &= \meas_m \modcov \meas_m -\sigma^2 + \sigma^4 e_j^t\Sigmam^{-1}e_j.
-    \end{split}
-  \end{align}
-  We use \eqref{eq:observation} to simplify the terms in the enumerator of
-  the conclusion of Lemma \ref{lemma:design increase}:
-  \begin{align}\label{eq:enumerator}
-    \begin{split}
-      (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m
-      &= \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m - \meas_m \\
-      %
-      %
-      %
-      &= \obsm^* (e_j - \sigma^2 \Sigmam^{-1} e_j) -\meas_j \\ 
-      %
-      %
-      %
-      &= -\sigma^2 \obsm^* \Sigma^{-1}e_j. 
-    \end{split}
-  \end{align}
-  %
-  Substitute \eqref{eq:enumerator} and \eqref{eq:denominator} to
-  the enumerator and denominator (respectively) of the conclusion of
-  Lemma \ref{lemma:design increase}:
-  %
-  \begin{align*}
-    \tar( \obs ) - \tar (\obsm ) &=
-    \log \left ( 1 + \frac{
-      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m,
-      (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m \rangle
-    }{
-      \sigma^2 + \meas_m \modcov \meas_m - \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m 
-    }       
-    \right ) \\
-    %
-    %
-    %
-    &= \log \left ( 1 + \frac{\sigma^4
-      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
-      \obsm^* \Sigmam^{-1}e_j \rangle
-    }{
-      2\sigma^2 - \sigma^4 e_j^t\Sigmam^{-1}e_j 
-    }       
-    \right ) \\
-    %
-    %
-    %
-    &= \log \left ( 1 + \frac{\sigma^2
-      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
-      \obsm^* \Sigmam^{-1}e_j \rangle
-    }{
-      2 - \sigma^2 e_j^t\Sigmam^{-1}e_j 
-    }       
-    \right ).
-  \end{align*}
-\end{proof}
-
-\begin{lemma}[Auxilliary Calculations]\label{lemma:aux calc}
-  Let $T(\obs) := \obs^* \Sigma^{-1}(\obs)\obs$, with $\Sigma(\obs)$
-  defined as in \eqref{eq:Sigma}. Then:
-  \begin{align*}
-    \delta T(\obs)V &= V^* \Sigma^{-1} \obs 
-    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
-    &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs
-    + \obs^* \Sigma^{-1} V.
-  \end{align*}
-\end{lemma}
-
-\begin{proof}
-  We need a few supplementary calculations. First:
-  \begin{align}\label{eq:der sig}
-    \begin{split}
-      \frac{\der}{\der \tau} \Big |_{\tau=0} \Sigma( \obs + \tau V )
-      &= \frac{\der}{\der \tau} \Big |_{\tau=0} 
-      (\obs + \tau V ) \modcov (\obs + \tau V )^*  + \sigma^2I\\
-      % 
-      % 
-      % 
-      &= V \modcov \obs^* + \obs \modcov V^*.
-    \end{split}
-  \end{align}
-  By the standard trick for the derivative of an operator: 
-  \begin{align*}
-    0 &= \frac{\der}{\der \tau} \Big |_{\tau=0} I \\
-    % 
-    % 
-    % 
-    &= \frac{\der}{\der \tau} \Big |_{\tau=0}
-    \left (\Sigma(\obs+\tau V)^{-1} \Sigma(\obs+\tau V) \right ) \\
-    % 
-    % 
-    % 
-    &= \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0} \Sigma+
-    \Sigma^{-1} \frac{\der \Sigma(\obs+\tau V)}{\der \tau} \Big |_{\tau=0}\\  
-    %
-    %
-    %
-    &= \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0} \Sigma+
-    \Sigma^{-1} (V\modcov \obs^* + \obs \modcov V^*) 
-    \text{, by \eqref{eq:der sig}. }
-  \end{align*}
-  Now we can write the variation of $\Sigma^{-1}$:
-  \begin{align}\label{eq:der sig inv}
-    \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0}  
-      &= -\Sigma^{-1} (V \modcov \obs^* + \obs \modcov V^*) \Sigma^{-1}.
-    \end{align}
-  % 
-  Finally, we can do the calculation we need to. By Leibnitz (product) rule
-  and \eqref{eq:der sig inv}:
-  % 
-  \begin{align*}
-    \delta T(\obs) V 
-    &= \frac{\der T(\obs + \tau V)}{\der \tau} \Big |_{\tau=0} \\
-    %
-    %
-    %
-    &= V^* \Sigma^{-1} \obs 
-    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
-    &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs
-    + \obs^* \Sigma^{-1} V,
-  \end{align*}
-  as required.
-\end{proof}
-
-\biggerbetter
-\begin{proof} 
-  Fix an arbitrary $j=1,\dots,m$ and take $V:= e_j e_j^t \obs$. We see
-  that for $u \in \hilo$
-  \begin{equation*}
-    Vu = e_je_j^t (\meas_1(u),\dots,\meas_m(u) )^t = e_j \meas_j(u)
-    = (0,\dots,0,\meas_j(u),0,\dots,0)^t.
-  \end{equation*}
-  This way, $V$ has the same $j$th entry as $\obs$ while the rest are
-  set to zero. We calculate the variation in this direction and show
-  it is positive --- $\delta \tar(\obs)V > 0$. This means that
-  increasing the magnitude of the $j$th measurement functional
-  increases $\tar(\obs)$. We start with the last line of \eqref{eq:tar var}, and denote $\tmp : = \fwd \postcov \fwd^*$.
-  \begin{align*}
-     \delta \tar(\obs) V 
-    &= \tr{V ( I - \modcov \obs^*\Sigma^{-1}\obs) \tmp \obs^* \Sigma^{-1}} \\
-    % 
-    %
-    %
-    &= \tr{e_je_j^t \obs ( I - \modcov \obs^*\Sigma^{-1}\obs) \tmp \obs^* \Sigma^{-1}} \\
-    %
-    % 
-    %
-    &= e_j^t \obs ( I - \modcov \obs^*\Sigma^{-1}\obs) \tmp \obs^* \Sigma^{-1}e_j \\
-    %
-    % 
-    %
-    &= e_j^t ( I - \obs \modcov \obs^*\Sigma^{-1})\obs \tmp \obs^* \Sigma^{-1}e_j \\  
-    % 
-    %
-    %
-    &=  e_j^t(\Sigma-\obs \modcov \obs^*) \Sigma^{-1}\obs \tmp \obs^* \Sigma^{-1}e_j \\
-    %
-    %
-    %
-    &=\sigma^2 e_j^t \Sigma^{-1}\obs \tmp \obs^* \Sigma^{-1}e_j
-    \text{ by \eqref{eq:Sigma} }\\
-    %
-    % 
-    %
-    &=\sigma^2 e_j^t \Sigma^{-1}\obs \fwd \postcov \fwd^* \obs^* \Sigma^{-1}e_j.
-  \end{align*} 
-  Since $\postcov$ is positive definite, $\delta \tar(\obs) V > 0$.
-\end{proof}
-
-\bibliographystyle{amsplain}
-\bibliography{refs,georg_refs}
-
-\end{document}
+%% The analysis in this paper is conducted for the general simultaneous
+%% optimization case. The sequential optimization case is dealt with in
+%% section \ref{subsec:clusterization sequential}. It is important to
+%% note, however that all conclusions we arrive at for the simultaneous
+%% case easily specialize to the sequential case by considering the
+%% posterior as the next sequential step's prior.

### 2023-03-29 latex/vanishing.tex
Paper almost donegit status !

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Analysis of D-Optimal Designs --- Vanishing Model Error}\label{section:vanishing}
-\subsection{Overview}
+\section{Characterization of D-Optimal Designs with Vanishing Model Error}\label{section:vanishing}
 In this section we analyze D-optimal designs when no model error is
-present. Sensor clusterization is discussed in section
-\ref{section:clusterization}. We first derive a nonlinear eigenvalue
-problem in $\obs$ for D-optimal designs in Section
-\ref{subsec:eigenproblem}. In Section \ref{subsec:characterization} we
-characterize D-optimal designs as ones that make all eigenvalues equal
-in the nonlinear eigenvalue problem. Results are summarized in Theorem
-\ref{thm:char}.
+present. We first derive a nonlinear eigenvalue problem in $\obs$ for
+D-optimal designs, and then we characterize D-optimal designs as ones
+that make all eigenvalues equal in the nonlinear eigenvalue
+problem. Results are summarized in Theorem \ref{thm:char}.
 
-\subsection{The Nonlinear Eigenvalue Problem}\label{subsec:eigenproblem}
+%% \subsection{The Nonlinear Eigenvalue Problem}\label{subsec:eigenproblem}
 $\modcov = 0$ implies $\Sigma= \sigma^2I$. The necessary first-order
@@
 
-%% Corollary \ref{cor:same ev} is important because it allows us to only think
-
-implies we only need to consider $\obs$ such that  
-At most $m$ of the eigenvectors of $\obs^*\obs$ have a positive
-eigenvalue, because $k := \rank \obs^*\obs \leq m$. Since we made no
-assumption regarding the ordering of $\{\lambda_i\}$, we can denote
-the corresponding non-zero eigenvalues of $\obs^*\obs$ by
+Corollary \ref{cor:same ev} implies we only need to consider $\obs$
+such that at most $m$ of the eigenvectors of $\obs^*\obs$ have a
+positive eigenvalue, because $k := \rank \obs^*\obs \leq m$. Since we
+made no assumption regarding the ordering of $\{\lambda_i\}$, we can
+denote the corresponding non-zero eigenvalues of $\obs^*\obs$ by
 $\{\eta_i\}_{i=1}^{k}$ and let $\eta_i = 0$ for $i \geq k+1$.
@@
   \sigma^{-2}\fwd^*\obs^*\obs\fwd$. Part (1) now follows from Lemma
-  \ref{lemma:twice woodbury}. Part (2) follows from Propositin
+  \ref{lemma:twice woodbury}. Part (2) follows from Proposition
   \ref{prop:kkt} and Corollary \ref{cor:descending}.
@@
   (3) were proved in Corollary \ref{cor:true target}. Part (4) is a
-  consequence Corollariy \ref{cor:eigenvalues}, with the caveat that
+  consequence of Corollary \ref{cor:eigenvalues}, with the caveat that
   we did not show that finding $\obs$ so that $\obs^*\obs$ has the
@@
 
-\begin{corollary}
-  In a D-optimal design, the posterior precision for eigenvector $i$
-  is $\eta_i + \sigma^2 \lambda_i^{-1} = \frac{m}{k} + \sigma^2
-  \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1}$. I.e. the posterior
-  precision in $\hil_p$ is constant for $1 \leq i \leq k$, where $k
-  =\rank \obs$.
+\begin{corollary}\label{cor:equal eigs}
+  In a D-optimal design, the posterior covariance of the pushforward
+  $\fwd_{*} \post$ is constant for eigenvectors $1 \leq i \leq k$
+  where $k =\rank \obs$.
 \end{corollary}
@@
 
-\pgfplotstableread{
-  Label    prior  $o_1$  $o_2$   $o_3$   $o_4$   $o_5$         $o_6$    topper
-  1         0.2    0.8    0.5     0.5    0.4    0.33333333  0.3333333  0.001
-  2         0.8    0.2    0.5     0.5    0.4    0.33333333  0.3333333  0.001
-  3         2.2    0      0       0      0.2    0.33333333  0.3333333  0.001
-  4         3.5    0      0       0      0      0           0          0.001
-}\clusterization
-
-\pgfplotstableread{
-  Label    prior  $o_1$  $o_2$   $o_3$   $o_4$   $o_5$         $o_6$    topper
-  1         0.2    0.8    0.7     0.3    0.4    0.2         0.4666666  0.001
-  2         0.8    0.2    0.3     0.7    0.4    0.46666666  0.2        0.001
-  3         2.2    0      0       0      0.2    0.33333333  0.3333333  0.001
-  4         3.5    0      0       0      0      0           0          0.001
-}\noclusterization
-
 Here is a way to understand D-optimal designs with vanishing model
@@
 
-\section{Sensor Clusterization --- Vanishing Model Error}\label{section:clusterization}
-\subsection{Overview}
-In this section we use Theorem \ref{thm:char} to explain how sensor
-clusterization may arise in various scenarios. We first consider
-sequential designs, and then simultaneous designs.
+\section{Clusterization poses no obstruction to D-optimality}\label{section:clusterization}
+Corollary \ref{cor:same ev} tells us that we can realize any design we
+want (encoded in $\{\eta_i\}$), given the trace constraint
+$\tr{\obs^*\obs} = m$ is satisfied. From Theorem \ref{thm:char} we
+know we only need to make the posterior flat where we take
+observations (recall figure \ref{fig:optimal vs not}). We are left
+with the freedom of assigning any part of $\eta_i$ to any observation
+$\meas_j$. Two possible assignments for the same problem are
+illustrated in figure \ref{fig:clusterization}. The first exhibits
+clusterization, the second does not. It is important to note that
+neither arises as a result of some numerical optimization scheme. Both
+designs give the same design criterion $\tar$, because their
+posteriors are identical. By Theorem \ref{thm:char}, both are
+maximizers of the same D-optimal design problem. One exhibits
+clusterization, and one does not. Of course, a real-life problem is
+more restricted than the model we consider here. Such restriction is
+the likely cause of clusterization obseved in, e.g., the toy model of
+section \ref{subsec:example}. Exactly how this happens is not
+clear. At this point it is the subject of speculation and should be
+further studied in the future.
 
-\subsection{Sequential Design}\label{subsec:clusterization sequential}
-In a sequential optimal design problem, each measurement is taken
-after results of previous measurements are recorded. In each step
-$m=1$, so $k=1$. In this case, Theorem \ref{thm:char} implies each
-measurement is taken along the direction of the eigenvector of $\fwd
-\prcov \fwd^*$ with lowest posterior precision. If $\lambda_i$
-decrease fast enough (e.g. exponentially), the
+A computer implementation of Lemma \ref{lemma:free} (see code in
+Supplementary material) generates the design presented in left panel
+Figure \ref{fig:clusterization}, which exhibits clusterization, over
+the clusterization-free design presented in the right panel of Figure
+\ref{fig:clusterization}. Modifying the numerical values of the prior
+does not result in a clusterization-free design. Thus, the preference
+to clusterization of Lemma \ref{lemma:free} is robust and
+clusterization is a generic property of the model presented in this
+study.
 
-, $\meas_1 \parallel \ev_1$ and $\eta_1 = 1$. As hinted in
-section \ref{subsec:seq vs sim}, the posterior becomes the prior for
-the next step. The eigenvectors of the new posterior are the same as
-for the previous one. A new observation will be taken in the direction
-of eigenvector with smalles eigenvalue. This leads to:
-\begin{align*}
-  \begin{split}
-    \meas_2 \parallel \ev_1  &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 < \sigma^2\lambda^{-1}_2 \\
-    \meas_2 \parallel \ev_2  &\text{ o.w. }
-  \end{split}
-\end{align*}
-In the first case, we get clusterization immediately. In the second,
-we consider $\meas_3$. Same reasoning leads to the conclusion that the
-eigenvectors of the new prior precision are the same. The eigenvalues,
-however, are $\lambda^{-1}_1 + \sigma^{-2}, \lambda^{-1}_2 +
-\sigma^{-2}, \lambda_3^{-1}, \dots$. A new observation is taken in the
-direction of eigenvector of smallest precision (note that
-$\lambda^{-1}_1 + \sigma^{-2} < \lambda^{-1}_2 + \sigma^{-2}$):
-\begin{align*}
-  \begin{split}
-    \meas_3 \parallel \ev_1 &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 <
-    \sigma^2\lambda^{-1}_3 \\
-    \meas_3 \parallel  \ev_3  &\text{ o.w. }
-  \end{split}
-\end{align*}
-Again, the first case results in clusterization. The sequential design
-proceeds in this fashion. Every $\meas_i$ is taken in the direction
-some eigenvector of the prior $\fwd \prcov\fwd^*$. This eigenvector
-has the smallest eigenvalue of the previous step's posterior precision
-operator. If we do not encounter clusterization after $m$ such steps,
-this means that observation $\meas_k$ was taken in the direction
-$\ev_k$ for $k=1,\dots,m$. Thus:
-\begin{align}\label{eq:failure}
-  \begin{split}
-    \sigma^2\lambda_{k-1}^{-1} + 1 \geq \sigma^2\lambda_k^{-1}, \text{ for } 1\leq k \leq m
-    \Longrightarrow \lambda_m^{-1} - \lambda_{m-1}^{-1} \leq \sigma^{-2}.
-  \end{split}
-\end{align}
+\pgfplotstableread{
+  Label    prior  $o_1o_2$  $o_3$   $o_4$   $o_5$    topper
+  1         0.1      0.55       1       0       1      0.001
+  2         0.2      1.45       0       1       0      0.001
+  3         3.5      0          0       0       0      0.001
+}\clusterization
 
-But \eqref{eq:failure} implies $\{\lambda_i^{-1}\}_{i=1}^{\infty}$
-grows at most linearly. In section \ref{subsec:abstract OED} we
-assumed $\fwd$ is strongly smoothing, which means eigenvalues of
-$(\fwd \prcov\fwd^*)^{-1}$ grow quickly. Thus, if we take $m$ large
-enough, \eqref{eq:failure} will eventually fail and we will observe
-sensor clusterization.
-
-\subsection{Simultaneous Design}\label{subsec:clusterization simultaneous}
-In a simultaneous design scenario, the result is not simple as in the
-sequential design case. We now see that a design that exhibits
-clusterization is as optimal as one that does not. Corollary
-\ref{cor:same ev} tells us that we can realize any design we want
-(encoded in $\{\eta_i\}$), given the trace constraint for
-$\obs^*\obs$. Theorem \ref{thm:char} tells us that for D-optimality,
-we only need to make the posterior flat where we take observations
-(recall figure \ref{fig:optimal vs not}). We are left with the freedom
-of assigning any part of $\eta_i$ to any observation $\meas_j$. Two
-possible assignments for the same problem are illustrated in figure
-\ref{fig:clusterization}. The first exhibits clusterization, the
-second does not. It is important to note that neither arises as a
-result of some numerical optimization scheme. Both designs give the
-same design criterion $\tar$, because their posteriors are
-identical. By Theorem \ref{thm:char}, both are maximizers of the same
-D-optimal design problem. One exhibits clusterization, and one does
-not. It is likely that a real-life problem is more restricted than the
-model we consider here. Such restriction is the likely cause of
-clusterization obseved in, e.g., the toy model of section
-\ref{subsec:example}. Exactly how this happens is not clear. At this
-point it is the subject of speculation and should be further studied
-in the future.
+\pgfplotstableread{
+  Label    prior  $o_1o_2$  $o_3o_4$   $o_5$     topper
+  1         0.1       0.95        0.6        1         0.001
+  2         0.2       1.0         1.4        0.        0.001
+  3         3.5        0          0          0         0.001
+}\noclusterization
 
@@
       \addplot [fill=green!80]  table [y=prior, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=blue!60]   table [y=$o_1$, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=red!60]    table [y=$o_2$, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=black!60]  table [y=$o_3$, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=orange!60] table [y=$o_4$, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=cyan!60]   table [y=$o_5$, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=purple!60] table [y=$o_6$, meta=Label, x expr=\coordindex] {\clusterization};
+      \addplot [fill=blue!60]   table [y=$o_1o_2$, meta=Label, x expr=\coordindex] {\clusterization};
+      \addplot [fill=red!60]    table [y=$o_3$, meta=Label, x expr=\coordindex] {\clusterization};
+      \addplot [fill=black!60]  table [y=$o_4$, meta=Label, x expr=\coordindex] {\clusterization};
+      \addplot [fill=orange!60] table [y=$o_5$, meta=Label, x expr=\coordindex] {\clusterization};
+      %% \addplot [fill=cyan!60]   table [y=$o_6$, meta=Label, x expr=\coordindex] {\clusterization};
+      %% \addplot [fill=purple!60] table [y=$o_7$, meta=Label, x expr=\coordindex] {\clusterization};
 
       \addlegendentry{prior}
-      \addlegendentry{$o_1$}
-      \addlegendentry{$o_2$}
+      \addlegendentry{$o_1o_2$}
       \addlegendentry{$o_3$}
@@
       \addlegendentry{$o_5$}
-      \addlegendentry{$o_6$}   
+      %% \addlegendentry{$o_6$}
+      %% \addlegendentry{$o_7$}   
     \end{axis}
@@
       \addplot [fill=green!80]  table [y=prior, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=blue!60]   table [y=$o_1$, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=red!60]    table [y=$o_2$, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=black!60]  table [y=$o_3$, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=orange!60] table [y=$o_4$, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=cyan!60]   table [y=$o_5$, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=purple!60] table [y=$o_6$, meta=Label, x expr=\coordindex] {\noclusterization};
+      \addplot [fill=blue!60]   table [y=$o_1o_2$, meta=Label, x expr=\coordindex] {\noclusterization};
+      \addplot [fill=red!60]    table [y=$o_3o_4$, meta=Label, x expr=\coordindex] {\noclusterization};
+      \addplot [fill=black!60]  table [y=$o_5$, meta=Label, x expr=\coordindex] {\noclusterization};
+      %% \addplot [fill=orange!60] table [y=$o_4$, meta=Label, x expr=\coordindex] {\noclusterization};
+      %% \addplot [fill=cyan!60]   table [y=$o_5$, meta=Label, x expr=\coordindex] {\noclusterization};
+      %% \addplot [fill=purple!60] table [y=$o_6$, meta=Label, x expr=\coordindex] {\noclusterization};
 
       \addlegendentry{prior}
-      \addlegendentry{$o_1$}
-      \addlegendentry{$o_2$}
-      \addlegendentry{$o_3$}
-      \addlegendentry{$o_4$}
+      \addlegendentry{$o_1o_2$}
+      \addlegendentry{$o_3o_4$}
       \addlegendentry{$o_5$}
-      \addlegendentry{$o_6$}   
+      %% \addlegendentry{$o_4$}
+      %% \addlegendentry{$o_5$}
+      %% \addlegendentry{$o_6$}   
     \end{axis}
@@
   \caption{Clusterization (left) and non-clusterization (right) in
-    simultaneous D-optimal designs. Posterior precision per
-    eigenvector after taking observations $\meas_i$ is plotted. Both
-    designs are identical for all practical matters --- their
-    posteriors are identical. In the left panel we see how an optimal
-    design is achieved with repeated observations. $\meas_2 = \meas_3$
-    and $\meas_5 = \meas_6$. This is not necessary though, as can be
-    seen in the right panel. Both designs achieve the same levels of
-    uncertainty. The left exhibits clusterization. The right does not.}
+    D-optimal designs. Posterior precision per eigenvector after
+    taking observations $\meas_i$ is plotted. Both designs are
+    identical for all practical matters --- their posteriors are
+    equal. In the left panel we see how an optimal design is achieved
+    with repeated observations. $\meas_3 = \meas_5$. This is not
+    necessary though, as can be seen in the right panel. Both designs
+    achieve the same levels of uncertainty. The left exhibits
+    clusterization. The right does not.}
   \label{fig:clusterization}
@@
 
-%% \free
-%% \begin{proof}
-%%   Let us diagonalize $M$, so that $M = U D U^t$ with $D =
-%%   \diag(d_1,\dots,d_k)$ and $U \in \R^{k \times k }$ orthogonal. Let
-%%   $S \in \R^{k \times m}$ with $S_{ii} = \sqrt{d_{i}}$ and zeros
-%%   otherwise. Define $A:= U S V^t$, where $V \in \R^{m \times m}$ is
-%%   orthogonal and will be further restricted later. Then $AA^t = U
-%%   SV^tVS^t U^t = UDU^t$, so $AA^t$ has the required eigenvalues and
-%%   eigenvectors by construction. If we can choose $V$ such that $A$
-%%   also satisfies the unit norm constraints we are done. These
-%%   constraints are, for $j=1,\dots,m$:
-%%   \begin{equation}\label{eq:V constraints}
-%%    1 = [A^tA]_{jj} = [V S^tS V^t]_{jj},
-%%   \end{equation}
-%%   and we can expect to do this since we assumed $\ttr D = m$.
+%% may arise in various scenarios. We first consider
+%% sequential designs, and then simultaneous designs.
 
-%%   Define $C = S^tS - I \in \R^{m \times m}$. Note that $\ttr C = 0$ and
-%%   $C$ is diagonal with non-zero entries $d_i-1,i=1,\dots,k$. It suffices
-%%   to find $V$ orthogonal such that $V C V^t$ has zero diagonal. We
-%%   construct such $V$ by sequentially inserting zeros in the diagonal
-%%   and not destroying zeros we already introduced, starting from the
-%%   last diagonal entry and moving to the first. Since $c_{mm} \neq 0$ ,
-%%   let $p < m$ such that $c_{pp}c_{mm} < 0$ (such $p$ exists because
-%%   the trace is zero) and let $\theta \in (0,\pi)$. Define a Givens
-%%   rotation $R^{(m)} \in \R^{m \times m}$ by
-%%   \begin{equation*}
-%%     r^{(m)}_{ab} :=
-%%     \begin{cases}
-%%       1 & a = b \neq p \text{ or } a = b \neq m \\
-%%       \cos \theta & a = b = p  \\
-%%      -\sin \theta & a = p, b = m\\
-%%       \cos \theta & a = b = m \\
-%%       \sin \theta & a = m, b = p \\ 
-%%       0 & o.w
-%%     \end{cases}
-%%   \end{equation*}
-%%   Note that conjugating a matrix by $R^{(m)}$ changes only its $m$ and
-%%   $p$ rows and columns. We want to choose $\theta$ such that
-%%   \begin{equation}\label{eq:mm}
-%%     0 = [R^{(m)} C (R^{(m)})^t]_{mm} = \cos^2 \theta c_{mm} + 2\cos \theta \sin
-%%     \theta c_{mp} + \sin^2\theta c_{pp},
-%%   \end{equation}
-%%   and it suffices to choose $\theta$ such that
-%%   \begin{equation*}
-%%     c_{mm} \cot^2 \theta + 2 c_{mp} \cot \theta + c_{pp} = 0.
-%%   \end{equation*}
-%%   This quadratic in $\cot\theta$ has a real solution, since
-%%   $c_{pp}c_{mm} < 0$ by assumption and we can find $\theta \in
-%%   (0,\pi)$ such that \eqref{eq:mm} is satisfied. We continue to find
-%%   $R^{(m-1)}$ that leaves row and column $m$ unchanged and
-%%   continue introducing zeros to the diagonal. The assumption $\ttr D =
-%%   m \Rightarrow \ttr C = 0$ guarantees we can do that. Taking $V:=
-%%   R^{(1)} R^{(2)} \dots R^{(m-1)}R^{(m)}$ completes the proof.
-%% \end{proof}
+%% \subsection{Sequential Design}\label{subsec:clusterization sequential}
+%% In a sequential optimal design problem, a decision is made on each
+%% measurement with all previous measurements fixed. In each step $m=1$,
+%% so $k=1$. In this case, Theorem \ref{thm:char} implies each
+%% measurement is taken along the direction of the eigenvector of $\fwd
+%% \prcov \fwd^*$ with lowest posterior precision. If $\lambda_i$
+%% decrease fast enough (e.g. exponentially), then $\lambda_i^{-1}$
+%% increase exponentially. Then, at some point, a measurement will be
+%% taken along an already measured eigenvector. This situation is
+%% measurement clusterization.
 
-%% \simdiag
-%% \begin{proof}
-%%   First, enumerate the eigenvalues of $C + \sum_{j=1}^m
-%%   \func_j\func_j^*$ as $\xi_1,\dots,\xi_\ell$. Denote the
-%%   indices of the eigenvectors corresponding to $\xi_i$
-%%   \begin{equation*}
-%%     S_i := \{ 1 \leq k \leq m | (C + \sum_{j=1}^m \func_j\func_j^* )\func_k = \xi_i \func_k \}.
-%%   \end{equation*}
-%%   Define further
-%%   \begin{equation*}
-%%     A_i := \sum_{k \in S_i} \func_k \func_k^*,
-%%   \end{equation*}
-%%   which is self-adjoint. Two observations are in order. First,
-%%   $\sum_{j=1}^m \func_j\func_j^* = \sum_{i=1}^\ell A_i$. Second, $A_i
-%%   \func_k = 0$ if $k\not \in S_i$, since eigenvectors of different
-%%   eigenvalue are orthogonal. For $k \in S_i$
-%%   \begin{equation}\label{eq:on vi}
-%%     \xi_i \func_k = (C + \sum_{j=1}^m \func_j \func_j^* ) \func_k = (C + A_i) \func_k.
-%%   \end{equation}
-%%   Let $V_i := span \{\func_k \}_{k\in S_i}$. Observe that $V_i$ is
-%%   invariant under $A_i$, by definition, and under $C$, by \eqref{eq:on
-%%     vi}. Using \eqref{eq:on vi} again, we conclude that on $V_i$, $A_i
-%%   = \xi_iI - C$. This immediately implies $A_i$ and $C$ have the
-%%   same eigenvectors on $V_i$. This holds for every $1 \leq i \leq
-%%   \ell$ and we conclude that $C$ and $A$ have the same eigenvectors.
-%% \end{proof}
+%% , $\meas_1 \parallel \ev_1$ and $\eta_1 = 1$. As hinted in
+%% section \ref{subsec:seq vs sim}, the posterior becomes the prior for
+%% the next step. The eigenvectors of the new posterior are the same as
+%% for the previous one. A new observation will be taken in the direction
+%% of eigenvector with smalles eigenvalue. This leads to:
+%% \begin{align*}
+%%   \begin{split}
+%%     \meas_2 \parallel \ev_1  &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 < \sigma^2\lambda^{-1}_2 \\
+%%     \meas_2 \parallel \ev_2  &\text{ o.w. }
+%%   \end{split}
+%% \end{align*}
+%% In the first case, we get clusterization immediately. In the second,
+%% we consider $\meas_3$. Same reasoning leads to the conclusion that the
+%% eigenvectors of the new prior precision are the same. The eigenvalues,
+%% however, are $\lambda^{-1}_1 + \sigma^{-2}, \lambda^{-1}_2 +
+%% \sigma^{-2}, \lambda_3^{-1}, \dots$. A new observation is taken in the
+%% direction of eigenvector of smallest precision (note that
+%% $\lambda^{-1}_1 + \sigma^{-2} < \lambda^{-1}_2 + \sigma^{-2}$):
+%% \begin{align*}
+%%   \begin{split}
+%%     \meas_3 \parallel \ev_1 &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 <
+%%     \sigma^2\lambda^{-1}_3 \\
+%%     \meas_3 \parallel  \ev_3  &\text{ o.w. }
+%%   \end{split}
+%% \end{align*}
+%% Again, the first case results in clusterization. The sequential design
+%% proceeds in this fashion. Every $\meas_i$ is taken in the direction
+%% some eigenvector of the prior $\fwd \prcov\fwd^*$. This eigenvector
+%% has the smallest eigenvalue of the previous step's posterior precision
+%% operator. If we do not encounter clusterization after $m$ such steps,
+%% this means that observation $\meas_k$ was taken in the direction
+%% $\ev_k$ for $k=1,\dots,m$. Thus:
+%% \begin{align}\label{eq:failure}
+%%   \begin{split}
+%%     \sigma^2\lambda_{k-1}^{-1} + 1 \geq \sigma^2\lambda_k^{-1}, \text{ for } 1\leq k \leq m
+%%     \Longrightarrow \lambda_m^{-1} - \lambda_{m-1}^{-1} \leq \sigma^{-2}.
+%%   \end{split}
+%% \end{align}
+
+%% But \eqref{eq:failure} implies $\{\lambda_i^{-1}\}_{i=1}^{\infty}$
+%% grows at most linearly. In section \ref{subsec:abstract OED} we
+%% assumed $\fwd$ is strongly smoothing, which means eigenvalues of
+%% $(\fwd \prcov\fwd^*)^{-1}$ grow quickly. Thus, if we take $m$ large
+%% enough, \eqref{eq:failure} will eventually fail and we will observe
+%% sensor clusterization.
+
+%% \subsection{Simultaneous Design}\label{subsec:clusterization simultaneous}
+%% If the eigenvalues $\lambda_i$ decrease quickly enough, we will
+%% observe clusterizaion.

### 2023-03-29 zeros.py

@@ -17,13 +17,17 @@ from pandas import isnull
 from numpy import sin, cos
 from scipy.stats import ortho_group
 
-EPS = 1e-10
+EPS = 1e-9
 ortho = ortho_group.rvs
 cot = lambda x: cos(x) / sin(x)
 
 
 @pytest.fixture
 def params():
+    """ generate random  m (number of observations) and k (rank of O*O).
+    """
+
+    
     params = []
     for m in np.random.randint(low=3, high=80, size=30):
         for k in np.random.randint(low=2, high=m + 1, size=1):
@@ -42,28 +46,32 @@ def givens(theta, dims, lower, upper):
     assert theta >= 0 and theta <= np.pi
     assert lower < upper
     assert upper <= dims
-    R = np.zeros((dims, dims))
-    for a in range(dims):
-        for b in range(dims):
-            if a == b:
-                if a == b == lower:
-                    R[a, b] = cos(theta)
-                elif a == b == upper:
-                    R[a, b] = cos(theta)
-                else:
-                    R[a, b] = 1
-            elif a == lower and b == upper:
-                R[a, b] = -sin(theta)
-            elif a == upper and b == lower:
-                R[a, b] = sin(theta)
+    
+    R = np.eye(dims)
+    R[lower, lower] = cos(theta)
+    R[upper, upper] = cos(theta)
+    R[lower, upper] = -sin(theta)
+    R[upper, lower] = sin(theta)
+
     return R
 
 
+
+    
 def MDUS(m, k):
+    """ Generate random:
+    M symmetric positive definite
+    D diagonal with decreasing diagonal entries
+    U orthogonal such that M = UDU^t
+    S of shape (k,m) with diagonal that is the square root of D
+    """
+
+    
     assert m >= k
 
     # D = np.abs(np.random.uniform(size=k))
     D = np.random.lognormal(mean=5, sigma=1, size=k)
+    D = np.sort(D)[::-1]
     # D = np.array([np.random.lognormal(mean=i, sigma=1) for i in range(k)]) 
     D = D * m / np.sum(D)
     assert abs(np.sum(D) - m) < EPS
@@ -138,7 +146,7 @@ def caller(m, k):
     assert A.shape == (k, m)
 
     ## Check that AAt = M, as stated.
-    AAt = np.dot(A, A.T)
+    AAt = A @ A.T
[…]

### 2023-03-29 latex/intro.tex
Fixin

@@
 \cite{silvey2013,Ucinski05} as well as for Bayesians
-\cite{Chaloner1995}. See \cite{Chaloner1995} for an investigation of
-the analogy between the two approaches. Although there is a plethora
-of design criteria, we focus on just one of these, commonly referred
-to as \emph{D-optimal design}. It has a simple and appealing
-motivation in the Bayesian context as explained in
-\cite{Chaloner1995}: Consider a linear model in finite dimensions,
-with Gaussian prior and noise. Under this model, maximizing the
-expected information gain (KL divergence \cite{CoverThomas91}) between
-posterior and prior amounts to minimizing the determinant of the
-posterior covariance matrix. In a frequentist setting, a D-optimal
-design minimizes the volume of the uncertainty ellipsoid \cite[page
-  16]{Ucinski05}, but this is done for the Fisher information matrix
-and not the posterior covariance. However, \cite{Chaloner1995} shows
-that the latter is just a regularized version of the former.
+\cite{Chaloner1995} (see \cite{Chaloner1995} for an investigation of
+the analogy between the two approaches). Among the plethora of design
+criteria, \emph{D-optimal design} have one of the simplest and most
+appealing motivation in the Bayesian context \cite{Chaloner1995}: a
+D-optimal design maximizes the expected information gain (KL
+divergence \cite{CoverThomas91}) between posterior and prior. For a
+linear model in finite dimensions, with Gaussian prior and noise, a
+D-optimal design minimizes the determinant of the posterior covariance
+matrix. In a frequentist setting, a D-optimal design minimizes the
+volume of the uncertainty ellipsoid \cite[page 16]{Ucinski05}, but
+this is done for the Fisher information matrix and not the posterior
+covariance. However, \cite{Chaloner1995} show that the latter is just
+a regularized version of the former.
 
@@
 inference takes place over a finite (not too large) number of
-parameters. The subject of optimal experimental design for function
-inference in a Bayesian context was pioneered by
+parameters. The subject of optimal experimental design on function
+spaces in a Bayesian context was pioneered by
 \cite{AlexanderianGloorGhattas14, AlexanderianPetraStadlerEtAl16,
   AlexanderianPetraStadlerEtAl14}. Similarly to the finite dimensional
-case, it can be shown that a D-optimal design arises naturally for
-linear models when one wishes to maximize the KL divergence between
-posterior and prior. This amounts to minimizing the determinant of the
-posterior covariance operator (understood as a product of its
-eigenvalues). Some difficulties arise in the process, but remedies can
-be found as shown in \cite{AlexanderianGloorGhattas14}.
+case, a D-optimal design still maximizes the KL divergence between
+posterior and prior. For a linear model with Gaussian prior and noise,
+this amounts to minimizing the determinant of the posterior covariance
+operator (understood as a product of its eigenvalues). Some
+difficulties arise in the process, but remedies can be found as shown
+in \cite{AlexanderianGloorGhattas14}.
 
 It seems counter intuitive that when one computes an optimal design
-using the D criterion, the optimization process results in
+using the D-optimality criterion, the optimization process results in
 observations that are very similar. For example, if an observation is
@@
 from each other. One way to do this is by introducing correlated
-errors which, philosophically, accounts for both observation error and
-model error. Another method considered is imposing distance
-constraints between observations. A somewhat different approach is
-suggested in \cite[page 49]{fedorov2012}, where close by design
-observations are merged --- a procedure which obviously does not avoid
-clusterization. The same problem arises in time-series analysis for
-pharmacokinetic experiments. The authors of \cite{hooker2009} suggest
-modeling auto-correlation time in the noise model, which is equivalent
-to the correlated errors mentioned above.
+errors which account for both observation error and model
+error. Another method considered is imposing distance constraints
+between observations. A somewhat different approach is suggested in
+\cite[page 49]{fedorov2012}, where close observations are merged --- a
+procedure which obviously does not avoid clusterization. The same
+problem arises in time-series analysis for pharmacokinetic
+experiments. The authors of \cite{hooker2009} suggest modeling
+auto-correlation time in the noise model, which is equivalent to the
+correlated errors mentioned above.
 
@@
 
-Similarly, in the inverse problems community, work is mostly
-computational and less theoretic. Model errors were considered in
-\cite{attia2020, koval2020}. The former study is focused on inferring a
-Quantity of Interest (QoI). The focus of the latter is reducing
-forward solves, using randomized linear algebra. Both studies present
-numerical techniques for finding optimal designs when model error is
+In the inverse problems community, work is mostly computational and
+less theoretic. Model errors were considered in \cite{attia2020,
+  koval2020}. The former study is focused on inferring a Quantity of
+Interest (QoI). The focus of the latter is reducing forward solves,
+using randomized linear algebra. Both studies present numerical
+techniques for finding optimal designs when model error is
 present. Both are restricted to linear inverse problems (although in
@@
 inversion. However, neither of these studies mention any effect model
-errors can have on sensor clusterization. The current study is mostly
+errors can have on sensor clusterization. This study is mostly
 theoretical and aims to fill the gap of understanding sensor
@@
 \subsection{Contribution}
-We propose and thoroughly study a relaxed and analytically tractable
-model for understanding D-optimal designs. Under this model, D-optimal
-designs are solutions of a constrained optimization problem,
-formulated using Lagrange multipliers (section \ref{section:D and
-  grad}). This allows us to rigorously show how model error mitigates
-clusterization (section \ref{section:non vanishing}). We show that
-designs that exhibit clusterization are just as optimal as
-clusterization free designs.
+I propose and study a relaxed and analytically tractable model for
+D-optimal designs. Under this model, D-optimal designs are solutions
+of a constrained optimization problem, formulated using Lagrange
+multipliers. This allows me to rigorously show how model error
+mitigates clusterization (section \ref{section:non vanishing}). I show
+that designs that exhibit clusterization are indeed optimal.
 
-A beautiful mathematical structure arises in D-optimal designs when no
-model error is present. The Lagrange multipliers problems is in fact a
-nonlinear eigenvalue problem for the observations and prior
-covariance. The operator for which eigenvectors and eigenvalues are
-sought is a sum of two operators. The first is the prior
-covariance. The second is an outer product of the observations (see
-section \ref{section:vanishing} for details and exact statement). This
-structure helps us shed light on D-optimal designs. We characterize
-these designs in Theorem \ref{thm:char}. One insight we prove is that
-a D-optimal design reduces uncertainties where they are highest
-first. Other interestig phenomena arise, but they require setting
-notation, and are discussed later.
+%% A beautiful mathematical structure arises in D-optimal designs
 
-In the process, we generalize several lemmas from linear algebra to
-infinite-dimensional settings. We prove a Matrix Determinant Lemma in
-\ref{lemma:MDL}. We generalize a lemma due to Lax \cite{Lax07} for
-calculating $\frac{\der}{\der t} \log \det (I + X(t))$, for an
-operator valued function $X(t)$ in \ref{lemma:lax}. We also prove
-lemmas in linear algebra. One constructs a decomposition $M = AA^t$
-where $A$ has unit norm columns \ref{lemma:free}. Another shows
-simultaneous diagonizability of the prior and outer product in the
-nonlinear eigenvalue problem referenced above \ref{lemma:sim diag}. We
-also provide other tools for understanding D-optimal designs in
-infinite-dimensional Bayesian inverse problems.%%  Among these are tools
-%% for calculating the increase in the design criterion per observation
-%% --- Lemma \ref{lemma:design increase} and Corollary \ref{cor:zero mod err}.
+When no model error is present, the Lagrange multipliers problem boils
+down to a nonlinear eigenvalue problem.
+%% The operator for which eigenvectors and eigenvalues are sought is a
+%% sum of two operators. The first is the prior covariance. The second
+%% is an outer product of the observations (see section
+%% \ref{section:vanishing} for details and exact statement).
+This structure helps me characterize D-optimal designs in Theorem
+\ref{thm:char}. The key idea is that a D-optimal design reduces
+uncertainties where they are highest first.
 
@@
 
+In the process, I generalize several lemmas from linear algebra to
+infinite-dimensional settings. I prove a Matrix Determinant Lemma in
+\ref{lemma:MDL}. I generalize a lemma due to Lax \cite{Lax07} for
+calculating $\frac{\der}{\der t} \log \det (I + X(t))$, for an
+operator valued function $X(t)$ in \ref{lemma:lax}. I show how to
+decompose a symmetric positive definite matrix $m$ as $M = AA^t$,
+where $A$ has unit norm columns in Lemma \ref{lemma:free}.
+
 \subsection{Limitations}\label{subsec:limitations}
-There are two main drawbacks of the study presented here. Our relaxed
+Two main drawbacks of the study presented here are: First, the relaxed
 model does not consider any specific set of allowed
-observations. Rather, we take observations in the unit ball in some
+observations. Rather, I take observations in the unit ball in some
 Hilbert space. This allows considerably less restrictive observations
-than any real-life problem does. The second drawback is that we do not
-show rigorously that clusterization necessarily occurs in a
-simultaneous design. We only show that it is as reasonable as no
-clusterization.
+than any real-life problem does. The second drawback is that I do not
+show rigorously that clusterization necessarily occurs, I only show
+that it is as reasonable as no clusterization.
 
@@
 
-In section \ref{section:prelim} we present a more abstract and general
-formulation of the inverse problem we consider. But, for the purpose
-of illustration, we present clusterization via a toy model --- the 1D
-heat equation in $[0,\pi]$ with a homogeneous Dirichlet boundary
-condition.
+In section \ref{section:prelim} I present a more abstract and general
+formulation of the inverse problem I consider. But, for the purpose of
+illustration, I present clusterization via a toy model --- the 1D heat
+equation in $[0,\pi]$ with a homogeneous Dirichlet boundary condition.
 
@@
 
-We would like to infer the initial condition $u_0$. For that purpose,
-we measure $u$ at some set of locations $\x_j \in [0,\pi], j=1,
-\dots,m$ and a final time $T > 0$. We assume centered Gaussian
-observation error, so we observe $v(\x_j,T) = u(\x_j,T) + \eps(\x_j)$
-with $\eps(\x_j) \sim \normal(0, \sigma^2), \sigma > 0$ iid. We model
+The goal is to infer the initial condition $u_0$. For that purpose, we
+measure $u$ at some set of locations $\x_j \in [0,\pi], j=1, \dots,m$
+and a final time $T > 0$. We assume centered Gaussian observation
+error, so we observe $v(\x_j,T) = u(\x_j,T) + \eps(\x_j)$ with
+$\eps(\x_j) \sim \normal(0, \sigma^2)$ iid, $\sigma^2 > 0$. We model
 the initial condition as $u_0 \sim \normal(0,\prcov)$, for $\prcov =
-(-\Delta)^{-1}$ with a homogeneous Dirichlet boundary condition. If It
-is well known \cite{Tarantola05} that for linear problems, with
-Gaussian prior and error, the posterior is also Gaussian with a
-covariance that does not depend on the observed data. The posterior
-covariance $\postcov$ is known to have a closed form formula, even in
-infinite dimensions \cite{Stuart10}. We denote by $\fwd$ the dynamics
-operator, so that $u( \cdot,T) = \fwd u_0$, and the observation
-operator $\obs$ so that $u(\x_j,T) = (\obs u)_j, j=1,\dots,m$. The
-posterior covariance is known and depends only on $\prcov, \fwd, \obs$
-and $\sigma^2$ (see section \ref{section:prelim} and
-\eqref{eq:postcov} specifically).
+(-\Delta)^{-1}$ with a homogeneous Dirichlet boundary condition. It is
+known \cite{Tarantola05} that for linear problems, with Gaussian prior
+and error, the posterior is also Gaussian with a covariance that does
+not depend on the observed data. The posterior covariance $\postcov$
+has a closed form formula, even in infinite dimensions
+\cite{Stuart10}. We denote by $\fwd$ the dynamics operator, so that
+$u( \cdot,T) = \fwd u_0$, and the observation operator $\obs$ so that
+$u(\x_j,T) = (\obs u)_j, j=1,\dots,m$. The posterior covariance is
+known and depends only on $\prcov, \fwd, \obs$ and $\sigma^2$ (see
+section \ref{section:prelim} and \eqref{eq:postcov} specifically).
 

### 2023-03-29 latex/paper.tex
Fixin

@@
 
-This material is based upon work supported by the National Science
-Foundation Graduate Research Fellowship Program under Grant No. (NSF
-grant number). Any opinions, findings, and conclusions or
-recommendations expressed in this material are those of the author(s)
-and do not necessarily reflect the views of the National Science
-Foundation. This work was also supported by The Raymond and Beverly
-Sackler Post-Doctoral Scholarship.
+This research was supported in part by an appointment with the
+National Science Foundation (NSF) Mathematical Sciences Graduate
+Internship (MSGI) Program sponsored by the NSF Division of
+Mathematical Sciences. This program is administered by the Oak Ridge
+Institute for Science and Education (ORISE) through an interagency
+agreement between the U.S. Department of Energy (DOE) and NSF. ORISE
+is managed for DOE by ORAU. All opinions expressed in this paper are
+the author's and do not necessarily reflect the policies and views of
+NSF, ORAU/ORISE, or DOE.  This work was also supported by The Raymond
+and Beverly Sackler Post-Doctoral Scholarship.
 

### 2023-04-02 latex/intro.tex
getting closegit add latex/vanishing.tex !

@@
 %% \ref{section:vanishing} for details and exact statement).
-This structure helps me characterize D-optimal designs in Theorem
-\ref{thm:char}. The key idea is that a D-optimal design reduces
-uncertainties where they are highest first.
+This structure helps me characterize D-optimal designs. The key idea
+is that a D-optimal design reduces uncertainties where they are
+highest first. This is the main conclusion of Theorem \ref{thm:char}
+stated below.
 
-\begin{restatable}[D-optimal designs with vanishing model error]{theorem}{main}\label{thm:char}
+\begin{restatable}[D-optimal designs with vanishing model error]{theorem}{char}\label{thm:char}
+%% \begin{theorem}[D-optimal designs with vanishing model error]\label{thm:char}
   Let:
-  \begin{enumerate}
-  \item $\fwd:\hilp \to \hilo$ a linear forward operator,
-    $\prcov:\hilp \to \hilp$ prior covariance operator, $\obs: \hilo
-    \to \mathbb{R}^m$ observation operator, where $m \in \mathbb{N}$
-    is the number of observations taken.
+  \begin{itemize}
+  \item $\hilp, \hilo$ Hilbert spaces.
+  \item $\fwd:\hilp \to \hilo$ a linear compact operator.
+  \item $\pr \sim \mathcal{N}(0, \prcov)$ prior Gaussian measure on $\hilp$,
+    where $\prcov:\hilp \to \hilp$ is the prior covariance operator.
+  \item $\obs: \hilo \to \mathbb{R}^m$ observation operator, where $m
+    \in \mathbb{N}$ is the number of observations taken. 
   \item $\sigma^2 \in \mathbb{R}_{+}$ observation noise variance,
     $\data = \obs \fwd \param + \eps$, where $\eps \in \mathbb{R}^m$
-      is iid $\mathcal{N}(0, \sigma^2)$ noise, $\pr \sim
-      \mathcal{N}(0, \prcov)$ prior measure, $\post$ the posterior
-      measure.
+    is iid $\mathcal{N}(0, \sigma^2)$ noise.
+  \item $\post$ the posterior measure, with covariance $\postcov$.
   \item A D-optimality utility function
@@
     \end{align*}
-  \item A D-optimal design operator $\obs$:
-    $$
-    \obs = \argmax_{\|\meas_j\| = 1, j=1,\dots,m}\tar(\obs)
-    $$ 
   \item $\{\lambda_i\}_{i=1}^\infty$ eigenvalues of $\fwd\prcov\fwd^*$
@@
   \item $\{\eta_i\}_{i=1}^\infty$ eigenvalues of $\obs^*\obs$.
-  \end{enumerate}
+  \end{itemize}
 
-  Then:
+  If $\obs$ is a D-optimal design operator
+    \begin{equation*}
+      \obs = \argmax_{\|\meas_j\| = 1, j=1,\dots,m}\tar(\obs), 
+    \end{equation*}
+    then:
   \begin{enumerate}
+  \item  $\tr{\obs^*\obs} = m$.
   \item $\obs^*\obs$ and $\fwd\prcov\fwd^*$ are simultaneously
     diagonalizable.
-  \item $k := \rank \obs^*\obs \leq m$.
-  \item     
+  \item $k := \rank \obs^*\obs \leq m$ and
     \begin{equation*}
@@
     \end{equation*}
+  \item The covariance of the pushforwad $\fwd_{*} \post$ is $\left (
+    (\fwd \prcov \fwd^*)^{-1} + \sigma^{-2} \obs^*\obs \right )^{-1}$
+    and has eigenvalues
+    \begin{equation*}
+      \theta_i =
+      \begin{cases}
+        \left(\frac{\sum_{j=1}^k \lambda_j^{-1} + \sigma^{-2}m}{k} \right )^{-1} & i \leq k \\
+        \lambda_i &  i > k 
+      \end{cases}
+    \end{equation*}
   \end{enumerate}
+%\end{theorem}
 \end{restatable}
+
+%% \begin{restatable*}[Flat posterior covariance]{corollary}{flatpost}\label{cor:equal eigs}
+%%  %\begin{corollary}\label{cor:eigenvalues}
+%%   If $\obs$ is D-optimal and $k := \rank \obs$, then:
+%%   \begin{enumerate}
+%%     \item The covariance of the pushforwad $\fwd_{*} \post$ is $\left ( (\fwd
+%%       \prcov \fwd^*)^{-1} + \sigma^{-2} \obs^*\obs \right )^{-1}$.
+%%     \item The eigenvalues of said operator are
+%%       \begin{equation}
+%%         \theta_i =
+%%         \begin{cases}
+%%           \left(\frac{1}{k} \sum_{j=1}^k \lambda_j^{-1} + \sigma^{-2}\frac{m}{k} \right )^{-1} & i \leq k \\
+%%           \lambda_i &  i > k 
+%%         \end{cases}
+%%       \end{equation}
+%%     \item The eigenvectors of said operator are the eigenvectors of
+%%       $\fwd \prcov \fwd^*$
+%%   \end{enumerate}
+%% \end{restatable*}%\end{corollary}
+
+%% \begin{restatable*}[Flat posterior covariance]{corollary}{flatpost}\label{cor:equal eigs}
+%%   Consider a D-optimal design $\obs$ for a Bayesian linear inverse
+%%   problem where $\fwd$ is the forward operator and $\pr$ is the
+%%   prior. Then the eigenvectors of the covariance of the pushforward of
+%%   the posterior $\fwd_{*} \post$ equal the eigenvectors of the
+%%   covariance of the pushforward of the prior. Moreover, the
+%%   eigenvalues are constant for eigenvectors $1 \leq i \leq k$ where $k
+%%   =\rank \obs$.
+%% \end{restatable*}
 
@@
 
-In section \ref{section:prelim} I present a more abstract and general
+In section \ref{section:prelim}, I present a more abstract and general
 formulation of the inverse problem I consider. But, for the purpose of

### 2023-04-02 latex/prelims.tex
getting closegit add latex/vanishing.tex !

@@
 \subsection{Overview}
-In this section we define notations that will be used throughout the
-paper. The theoretical foundations for inverse problems over function spaces
-can be found in \cite{Stuart10}.
+In this section I will define notations that will be used throughout
+the paper. The theoretical foundations for inverse problems over
+function spaces can be found in \cite{Stuart10}. Most of the notations
+are summarized in the conditions of Theorem \ref{thm:char}.
 
@@
 Let $\hilp$ and $\hilo$ be separable Hilbert spaces (the subscripts p
-and o are for ``parameter'' and ``observation'', respectfully). We
-denote $\hilo^*$ the Hilbert space of all linear functionals on
+and o are for ``parameter'' and ``observation'', respectfully). Denote
+$\hilo^* \simeq \hilo$ the Hilbert space of all linear functionals on
 $\hilo$. Consider $\fwd: \hilp \to \hilo$, a linear operator (the
@@
 $m \times m$ identity. Both error terms and the prior are assumed
-independent of each other. The observation and forward operators are
-typically merged $\tmp := \obs \fwd$
+independent of each other. In the literature, it is common for the
+observation and forward operators to be merged $\tmp := \obs \fwd$
 \cite{AlexanderianGloorGhattas14}. Since we need to separate the
@@
 $\ell_1$ and $\ell_2$ approximate a point measurement at $0$ for small
-$\eps$ and, crucially, $\ell_1, \ell_2 \in \hilo^*$. It is important
-to note that although $\|\ell\|_1 = \|\ell_2\|_1=1$, unfortunately
-$\|\ell_1\|_2^2 = (4\pi \eps)^{-\frac{1}{2}}$ and $\|\ell_2\|_2^2 =
-\eps^{-1}$.
+$\eps$ and, crucially, $\ell_1, \ell_2 \in \hilo^*$.
+%% It is important to note that although $\|\ell\|_1 =
+%% \|\ell_2\|_1=1$, unfortunately $\|\ell_1\|_2^2 = (4\pi
+%% \eps)^{-\frac{1}{2}}$ and $\|\ell_2\|_2^2 = \eps^{-1}$.
 
 \subsubsection{Error Terms}
-Considering the sum of the error terms, it is easy to see that
+Considering the sum of the error terms, it is easy to verify that
 $\bar{\eps} := \obs \eps' + \eps \in \R^m$. It is a centered Gaussian
@@
 a numerical source, where our discretization of the system is not fine
-enough to capture all small scale phenomena. In section
-\ref{section:non vanishing}, we will see that assuming some
-correlation in the error mitigates the clusterization phenomenon, as
-reported in the literature \cite{Ucinski05}.
+enough to capture all small scale phenomena.
+%%  In section \ref{section:non vanishing}, we will see that assuming
+%% some correlation in the error mitigates the clusterization
+%% phenomenon, as reported in the literature \cite{Ucinski05}.
 
@@
 A D-optimal design maximizes expected KL divergence between posterior
-and prior. It is useful to recall the definition of KL divergence for
-an arbitrary prior measure:
+and prior measures, denoted $\post$ and $\pr$, respecively. It is
+useful to recall the definition of KL divergence for an arbitrary
+prior measure:
 $$
@@
   and let $\post = \normal(\postmean,\postcov)$ the posterior measure
-  on $\hilo$ for the Bayesian linear inverse problem $\data = \obs
+  on $\hilp$ for the Bayesian linear inverse problem $\data = \obs
   \fwd\param + \obs \eps' + \eps$ discussed above. Then

### 2023-04-02 latex/vanishing.tex
Fixin
getting closegit add latex/vanishing.tex !

@@
 %% \subsection{The Nonlinear Eigenvalue Problem}\label{subsec:eigenproblem}
-$\modcov = 0$ implies $\Sigma= \sigma^2I$. The necessary first-order
-condition for D-optimality \eqref{eq:conditions} becomes
+The necessary first-order condition for D-optimality
+\eqref{eq:conditions}, when $\Sigma= \sigma^2I$ (i.e. when $\modcov =
+0$) becomes
 \begin{equation}\label{eq:eigenproblem}
@@
 \postcov \fwd^*$, where rows of $\obs$, namely $\meas_j,j=1,\dots, m$,
-are eigenvectors. However, that $\postcov$ depends on $\obs$ and thus,
-we refer to \eqref{eq:eigenproblem} as a nonlinear eigenvalue problem.
-
-The following two lemmas allow us to gain better understanding of the
-nonlinear eigenvalue problem.
+are eigenvectors. However, $\postcov$ depends on $\obs$ and thus, we
+refer to \eqref{eq:eigenproblem} as a \emph{nonlinear} eigenvalue
+problem. The following two lemmas allow us to gain better
+understanding of the nonlinear eigenvalue problem.
 
@@
 
-\begin{restatable*}[Simultaneous diagonizability for the nonlinear eigenvalue problem]{lemma}{simdiag}\label{lemma:sim diag}
+\begin{restatable*}[Simultaneous diagonizability]{lemma}{simdiag}\label{lemma:sim diag}
   Let $\hil$ separable Hilbert space, $C:\hil \to \hil$ self-adjoint
@@
 
-\begin{corollary}[Eigenvectors of a D-optimal design]\label{cor:same ev}
+\begin{proposition}[Eigenvectors of a D-optimal design]\label{prop:same ev}
   Let $\obs$ satisfy the nonlinear eigenvalue problem
@@
   as $\fwd \prcov \fwd^*$.
-\end{corollary}
+\end{proposition}
 \begin{proof}
@@
 
-Corollary \ref{cor:same ev} implies we only need to consider $\obs$
-such that at most $m$ of the eigenvectors of $\obs^*\obs$ have a
-positive eigenvalue, because $k := \rank \obs^*\obs \leq m$. Since we
-made no assumption regarding the ordering of $\{\lambda_i\}$, we can
-denote the corresponding non-zero eigenvalues of $\obs^*\obs$ by
-$\{\eta_i\}_{i=1}^{k}$ and let $\eta_i = 0$ for $i \geq k+1$.
+Since we made no assumption regarding the ordering of $\{\lambda_i\}$,
+we can denote the corresponding non-zero eigenvalues of $\obs^*\obs$
+by $\{\eta_i\}_{i=1}^{k}$ and let $\eta_i = 0$ for $i \geq k+1$.
 
-%% \begin{definition}
-%%   Let $\{\lambda_i\}_{i=1}^{\infty}$ eigenvalues of $\fwd \prcov
-%%   \fwd^*$ sorted in descending order. Let $\eta \in
-%%   \mathbb{R}_+^m$. The \emph{utility proxy} $\bar{\tar}$ is defined
-%%   as:
-%%   \begin{equation*}
-%%     \bar{\tar}(\eta) := \frac12 \sum_{i=1}^m \log(1 + \sigma^{-2}\lambda_i\eta_i).
-%%   \end{equation*}
-%% \end{definition}
-
-\begin{corollary}\label{cor:true target}
+\begin{proposition}\label{prop:true target}
   Let $\obs$ with $m$ observations satisfy the nonlinear eigenvalue
@@
   \end{enumerate}
-\end{corollary}
+\end{proposition}
 \begin{proof}
@@
       %
-      &=\frac12 \log \prod_{i=1}^{\infty} ( 1 + \sigma^{-2} \lambda_i\eta_i ) \text{ (Corollary \ref{cor:same ev})} \\
+      &=\frac12 \log \prod_{i=1}^{\infty} ( 1 + \sigma^{-2} \lambda_i\eta_i ) \text{ (Proposition \ref{prop:same ev})} \\
       %
@@
 \end{proposition}
+
 \begin{proof}
-%% \begin{proposition}\label{prop:optimizer}
-%%   Let $\{\lambda_i\}_{i=1}^{\infty}$ the sorted eigenvectors of $\fwd
-%%   \prcov\fwd^*$, $k\in \mathbb{N}^+$ and
-%%   %% Let
-%%   %%   \begin{equation}
-%%   %%   k = \argmax \left \{ k:\lambda_k^{-1} \leq \sigma^{-2}\frac{m}{k} + \frac{1}{k} \sum_{j=1}^{k}
-%%   %%   \lambda_j^{-1} \right \}.
-%%   %% \end{equation}
-
-%%   \begin{equation}
-%%   \eta_i = \begin{cases}
-%%     \frac{m}{k} - \sigma^2 \lambda_i^{-1} + \sigma^2 \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1} & i \leq k \\
-%%     0 & i > k 
-%%   \end{cases}.
-%%   \end{equation}
-
-%%   Let $\obs$ an observation operator such that $\obs^*\obs$ has the
-%%   same eigenvectors as $\fwd \prcov \fwd^*$. If $\eta_i > 0$ for
-%%   $i\leq k$ and $\obs^*\obs$ has eigenvalues $\eta_i$, then $\obs$
-%%   maximizes $\tar$ subject to $\sum_{i=1}^m \eta_i = m$ and to $\rank
-%%   \obs = k$.
-%% \end{proposition}
-  %% \begin{proof}
-  %% et $A := \{i: \eta_i > 0\}$. A choice of $A$ where $\eta_i =0,
-  %% \eta_j >0$ for some $1 \leq i < j \leq m$ is not optimal (just
-  %% switch $\eta_j$ and $\eta_i$ and recall $\{\lambda_i\}$ are arranged
-  %% in descending order). Thus, since $\rank \obs = k$ we conclude that,
-  %% $\eta_i=0, i>k$.
-
   Let $\Phi(\eta) = \sum_{i=1}^k \eta_i - m$ and $\Omega_j(\eta) =
@@
 
-\begin{corollary}\label{cor:descending}
-  In Proposition \ref{prop:kkt}, if $\lambda_i$ are arranged in
-  descending order, then $A= \{1,\dots,k\}$ for some $1 \leq k \leq
-  m$.
-\end{corollary}
-
-\begin{corollary}\label{cor:eigenvalues}
-  If $\obs$ is optimal and $k := \rank \obs$, then:
-  \begin{enumerate}
-    \item The covariance of the pushforwad $\fwd_{*} \post$ is $((\fwd
-      \prcov \fwd^*)^{-1} + \sigma^{-2} \obs^*\obs)^{-1}$.
-    \item The eigenvalues of said operator are
-      \begin{equation}
-        \eta_i = \begin{cases}
-          \frac{m}{k} + \sigma^2 \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1} & i \leq k \\
-          \lambda_i^{-1} & i > k 
-        \end{cases}.
-      \end{equation}
-  \end{enumerate}
-\end{corollary}
-
-\begin{proof}
-  Recall from \eqref{eq:postcov}, that the posterior covariance is
-  $\postcov^{-1} = \prcov^{-1} +
-  \sigma^{-2}\fwd^*\obs^*\obs\fwd$. Part (1) now follows from Lemma
-  \ref{lemma:twice woodbury}. Part (2) follows from Proposition
-  \ref{prop:kkt} and Corollary \ref{cor:descending}.
-\end{proof}
-
 We are almost ready to state and prove Theorem \ref{thm:char} that
@@
 
-\main
+\char*
 
 \begin{proof}
-  Part (1) was proved in Corollary \ref{cor:same ev}. Parts (2) and
-  (3) were proved in Corollary \ref{cor:true target}. Part (4) is a
-  consequence of Corollary \ref{cor:eigenvalues}, with the caveat that
-  we did not show that finding $\obs$ so that $\obs^*\obs$ has the
-  desired eigenvalues is feasible. To this end, we utilize Lemma
-  \ref{lemma:free}: let $M = \diag(\eta_1, \dots, \eta_k)$, diagonal
-  with respect to the first $k$ eigenvectors of $\fwd \prcov
-  \fwd^*$. We take $\obs := A$ from the Lemma \ref{lemma:free}, and
-  this concludes the proof.
+  Part (1) is immediate for any observation operator $\obs$ that
+  satisfies the unit norm constraint on observation vectors. Part (2)
+  was proved in Proposition \ref{prop:same ev}. Part (3) was proved in
+  Proposition \ref{prop:true target}.
+
+  Part (4) is a consequence of Propositions \ref{prop:true target} and
+  \ref{prop:kkt}, with the caveat that we did not show that finding
+  $\obs$ so that $\obs^*\obs$ has the desired eigenvalues is
+  feasible. To this end, we utilize Lemma \ref{lemma:free}: let $M =
+  \diag(\eta_1, \dots, \eta_k)$, diagonal with respect to the first $k$
+  eigenvectors of $\fwd \prcov \fwd^*$. We take $\obs := A$ from the
+  Lemma \ref{lemma:free}.
+
+  Recall from \eqref{eq:postcov}, that the posterior precision is
+  $\postcov^{-1} = \prcov^{-1} + \sigma^{-2}\fwd^*\obs^*\obs\fwd$. The
+  first statement in part (5) now follows from Lemma \ref{lemma:twice
+    woodbury}, while the second statement follows from parts (1) and
+  (4).
 \end{proof}
-
-\begin{corollary}\label{cor:equal eigs}
-  In a D-optimal design, the posterior covariance of the pushforward
-  $\fwd_{*} \post$ is constant for eigenvectors $1 \leq i \leq k$
-  where $k =\rank \obs$.
-\end{corollary}
 
@@
   \end{tikzpicture}
-  \caption{As an illustration of Corollary \ref{cor:equal eigs}, a
-    D-optimal design is compared with a sub-optimal design. Both
-    designs are allowed three observations. Thus, the blue area has
+  \caption{A comparison of the eigenvalues of the pushforward
+    posterior precision $(\fwd\prcov\fwd^*)^{-1}
+    + \sigma^{-2}\obs^*\obs$ for a D-optimal design (left) and a
+    sub-optimal design (right). Both designs are allowed three
+    observations. We assume $\sigma^2=1$ and thus, the blue area has
     accumulated height of $3$ in both plots. A D-optimal design (left)
     increases precision (reduces uncertainty) where it is lowest. A
-    sub-optimal design (right) does not. In this example, an
-    observation is taken partly in the direction of the third
-    eigenvector. The precision for this eigenvector is bigger even
-    than the posterior precision of the D-optimal design on the left.}
+    sub-optimal design (right) does not.
+    %% In this example, an observation is taken partly in the
+    %% direction of the third eigenvector. The precision for this
+    %% eigenvector is bigger even than the posterior precision of the
+    %% D-optimal design on the left.
+  }
   \label{fig:optimal vs not}
@@
 Here is a way to understand D-optimal designs with vanishing model
-error: Imagine each eigenvector corresponds to a graduated lab
-cylinder (Figure \ref{fig:cylinder}). Each cylinder $i$ is filled,
-a-priori, with green liquid of $\sigma^2\lambda_i^{-1}$ volume
-units. This liquid represnets the prior precision, weighted by
-observation noise. The volume of the green liquid is increasing in
-$i$. A D-optimal design with $m$ observations ignores cylinders
-indexed $m+1$ and above. It has blue liquid of volume $m$ units at its
-disposal. It distributes said liquid by repeatedly adding a drop to
-whatever cylinder currently has the lowest level of liquid in it
-(lowest precision). The result of such a procedure is illustrated in
-Figure \ref{fig:optimal vs not}.
+error: Imagine each eigenvector of $\fwd \prcov \fwd^*$ corresponds to
+a graduated lab cylinder (Figure \ref{fig:cylinder}). Each cylinder
+$i$ is filled, a-priori, with green liquid of $\lambda_i^{-1}$ volume
+units. This liquid represnets the prior precision, and the volume of
+the green liquid is increasing in $i$. A D-optimal design with $m$
+observations ignores cylinders indexed $m+1$ and above. It has blue
+liquid of volume $\sigma^{-2}m$ units at its disposal. It distributes
+said liquid by repeatedly adding a drop to whatever cylinder currently
+has the lowest level of liquid in it (lowest precision). The result of
+such a procedure is illustrated in Figure \ref{fig:optimal vs not}.
 
@@
 \section{Clusterization poses no obstruction to D-optimality}\label{section:clusterization}
-Corollary \ref{cor:same ev} tells us that we can realize any design we
-want (encoded in $\{\eta_i\}$), given the trace constraint
-$\tr{\obs^*\obs} = m$ is satisfied. From Theorem \ref{thm:char} we
-know we only need to make the posterior flat where we take
-observations (recall figure \ref{fig:optimal vs not}). We are left
-with the freedom of assigning any part of $\eta_i$ to any observation
-$\meas_j$. Two possible assignments for the same problem are
-illustrated in figure \ref{fig:clusterization}. The first exhibits
-clusterization, the second does not. It is important to note that
-neither arises as a result of some numerical optimization scheme. Both
-designs give the same design criterion $\tar$, because their
-posteriors are identical. By Theorem \ref{thm:char}, both are
-maximizers of the same D-optimal design problem. One exhibits
-clusterization, and one does not. Of course, a real-life problem is
-more restricted than the model we consider here. Such restriction is
-the likely cause of clusterization obseved in, e.g., the toy model of
-section \ref{subsec:example}. Exactly how this happens is not
-clear. At this point it is the subject of speculation and should be
-further studied in the future.
+We saw that the necessary conditions for a design $\obs$ with $m$
+observations to be D-optimal are (1), (2) and (4) of Theorem
+\ref{thm:char}. As long as these requirements are satisfied, $\obs$ is
+D-optimal. These requirements, however, still leave some freedom in
+choosing $\obs$. This freedom may give rise to clusterization, as
+illustrated in Figure \ref{fig:clusterization} and explained below.
+
+Consider an inverse problem with $\lambda_1^{-1} = 0.1, \lambda_2^{-1}
+= 0.2$ and $\sigma^2=1$. In \eqref{eq:two designs}, we suggest two
+D-optimal designs $\obs_1, \obs_2$ such that $\obs_1^*\obs_1 =
+\obs_2^*\obs_2$. The former exhibits clusterization ($\meas_3 =
+\meas_5 = \ev_1$), while the latter does not. Since both of these
+designs are D-optimal, clusterization does not pose any obstruction to
+D-optimality.
+
+\begin{equation}\label{eq:two designs}
+  \obs_1 =
+  \begin{pmatrix}
+    \sqrt{0.275}\ev_1 - \sqrt{0.725}\ev_2 \\
+    \sqrt{0.275}\ev_1 + \sqrt{0.725}\ev_2 \\
+    \ev_1 \\
+    \ev_2 \\
+    \ev_1 \\    
+  \end{pmatrix},
+  \obs_2 =
+  \begin{pmatrix}
+    \sqrt{\frac{3}{8}}\ev_1 - \sqrt{\frac{5}{8}}\ev_2 \\
+    \sqrt{\frac{3}{8}}\ev_1 + \sqrt{\frac{5}{8}}\ev_2 \\
+    \sqrt{\frac{2}{5}}\ev_1 - \sqrt{\frac{3}{5}}\ev_2 \\
+    \sqrt{\frac{2}{5}}\ev_1 + \sqrt{\frac{3}{5}}\ev_2 \\
+    \ev_1 \\    
+  \end{pmatrix}
+\end{equation}
 
 A computer implementation of Lemma \ref{lemma:free} (see code in
-Supplementary material) generates the design presented in left panel
-Figure \ref{fig:clusterization}, which exhibits clusterization, over
-the clusterization-free design presented in the right panel of Figure
-\ref{fig:clusterization}. Modifying the numerical values of the prior
-does not result in a clusterization-free design. Thus, the preference
-to clusterization of Lemma \ref{lemma:free} is robust and
-clusterization is a generic property of the model presented in this
-study.
+Supplementary material) generates $\obs_1$, which exhibits
+clusterization. Modifying the numerical values of the prior does not
+break the clusterization. Thus, the preference to clusterization of
+Lemma \ref{lemma:free} is robust. The vast majority of randomized
+numerical simulations (see code in Supplementary material) also give
+rise to D-optimal designs that exhibit clustrization. We conclude that
+clusterization is a generic property of the relaxed model presented in
+this study.
+
+%% Of course, a real-life problem is more restricted than the model we
+%% consider here. Such restriction is the likely cause of clusterization
+%% obseved in, e.g., the toy model of section
+%% \ref{subsec:example}. Exactly how this happens is not clear. At this
+%% point it is the subject of speculation and should be further studied
+%% in the future.
 
@@
   Label    prior  $o_1o_2$  $o_3o_4$   $o_5$     topper
-  1         0.1       0.95        0.6        1         0.001
-  2         0.2       1.0         1.4        0.        0.001
+  1         0.1    0.75     0.8        1         0.001
+  2         0.2    1.25     1.2        0         0.001
   3         3.5    0        0          0         0.001
@@
   \end{tikzpicture}
-  \caption{Clusterization (left) and non-clusterization (right) in
-    D-optimal designs. Posterior precision per eigenvector after
-    taking observations $\meas_i$ is plotted. Both designs are
-    identical for all practical matters --- their posteriors are
-    equal. In the left panel we see how an optimal design is achieved
-    with repeated observations. $\meas_3 = \meas_5$. This is not
-    necessary though, as can be seen in the right panel. Both designs
-    achieve the same levels of uncertainty. The left exhibits
-    clusterization. The right does not.}
+  \caption{Clusterization (left, $\obs_1$ in \eqref{eq:two designs})
+    and non-clusterization (right, $\obs_2$ in \eqref{eq:two designs})
+    in D-optimal designs. Posterior precision per eigenvector is
+    plotted for $\obs_1$ and $\obs_2$ from \eqref{eq:two
+      designs}. Both designs are identical for all practical matters
+    --- their posteriors are equal. The left panel shows a D-optimal
+    design with clusterization: $\meas_3 = \meas_5$. The right panel
+    shows a D-optimal design without clusterization. The contribution
+    to the posterior precision due to $\meas_1$ and $\meas_2$ is
+    denoted $\meas_1\meas_2$. These contributions are plotted together
+    because $\meas_1$ and $\meas_2$ are not in the direction of any
+    eigenvector. Similarly for $\meas_3\meas_4$.}
   \label{fig:clusterization}

### 2023-04-08 latex/intro.tex
Almost done?

@@
 procedure sometimes places sensors in very close proximity to each
-other (as can be seen in figure \ref{fig:clusterization
+other (as can be seen in Figure \ref{fig:clusterization
   illustration}). Following \cite{Ucinski05}, we refer to this
@@
 \subsection{Contribution}
-I propose and study a relaxed and analytically tractable model for
+We propose and study a relaxed and analytically tractable model for
 D-optimal designs. Under this model, D-optimal designs are solutions
@@
 multipliers. This allows me to rigorously show how model error
-mitigates clusterization (section \ref{section:non vanishing}). I show
-that designs that exhibit clusterization are indeed optimal.
+mitigates clusterization (Section \ref{section:non vanishing}). We
+show that designs that exhibit clusterization are indeed optimal.
 
@@
 %% \ref{section:vanishing} for details and exact statement).
-This structure helps me characterize D-optimal designs. The key idea
-is that a D-optimal design reduces uncertainties where they are
+This structure helps us to characterize D-optimal designs. The key
+idea is that a D-optimal design reduces uncertainties where they are
 highest first. This is the main conclusion of Theorem \ref{thm:char}
@@
   %% \begin{theorem}[D-optimal designs with vanishing model error]\label{thm:char}
+  %% Let Notation \ref{not:setup}. Further, let
+
   Let:
@@
   \item $\{\eta_i\}_{i=1}^\infty$ eigenvalues of $\obs^*\obs$.
+  \item $\obs$ a D-optimal design operator
+    \begin{equation*}
+      \obs = \argmax_{\|\meas_j\| = 1, j=1,\dots,m}\tar(\obs).
+    \end{equation*}
   \end{itemize}
 
-  If $\obs$ is a D-optimal design operator
-    \begin{equation*}
-      \obs = \argmax_{\|\meas_j\| = 1, j=1,\dots,m}\tar(\obs), 
-    \end{equation*}
-    then:
+  Then:
   \begin{enumerate}
@@
 
-In the process, I generalize several lemmas from linear algebra to
-infinite-dimensional settings. I prove a Matrix Determinant Lemma in
-\ref{lemma:MDL}. I generalize a lemma due to Lax \cite{Lax07} for
+In the process, We generalize several lemmas from linear algebra to
+infinite-dimensional settings. We prove a Matrix Determinant Lemma in
+\ref{lemma:MDL}. We generalize a lemma due to Lax \cite{Lax07} for
 calculating $\frac{\der}{\der t} \log \det (I + X(t))$, for an
-operator valued function $X(t)$ in \ref{lemma:lax}. I show how to
+operator valued function $X(t)$ in \ref{lemma:lax}. We show how to
 decompose a symmetric positive definite matrix $m$ as $M = AA^t$,
@@
 model does not consider any specific set of allowed
-observations. Rather, I take observations in the unit ball in some
+observations. Rather, we take observations in the unit ball in some
 Hilbert space. This allows considerably less restrictive observations
-than any real-life problem does. The second drawback is that I do not
-show rigorously that clusterization necessarily occurs, I only show
+than any real-life problem does. The second drawback is that we do not
+show rigorously that clusterization necessarily occurs, we only show
 that it is as reasonable as no clusterization.
@@
   \caption{The clusterization effect for the 1D heat equation
-    described in section \ref{subsec:example}. Posterior pointwise
+    described in Section \ref{subsec:example}. Posterior pointwise
     standard deviations (lines) are plotted over the domain $[0,
@@
 
-In section \ref{section:prelim}, I present a more abstract and general
-formulation of the inverse problem I consider. But, for the purpose of
-illustration, I present clusterization via a toy model --- the 1D heat
+In Section \ref{section:prelim}, we present a more abstract and general
+formulation of the inverse problem we consider. But, for the purpose of
+illustration, we present clusterization via a toy model --- the 1D heat
 equation in $[0,\pi]$ with a homogeneous Dirichlet boundary condition.
@@
 known and depends only on $\prcov, \fwd, \obs$ and $\sigma^2$ (see
-section \ref{section:prelim} and \eqref{eq:postcov} specifically).
+Section \ref{section:prelim} and \eqref{eq:postcov} specifically).
 
@@
 criterion presented in the introduction to infinite dimensions
-(section \ref{subsec:D optimal design} below). We choose
+(Section \ref{subsec:D optimal design} below). We choose
 $\x_j,j=1,\dots,m$ to minimize (an expression analogous to) the
@@
 
-The clusterization effect is illustrated in figure
+The clusterization effect is illustrated in Figure
 \ref{fig:clusterization illustration}. Posterior pointwise standard
@@
 actual data observed. The posterior covariance does, however, depend
-on location of the observation taken. In figure
+on location of the observation taken. In Figure
 \ref{fig:clusterization illustration}, observation locations are

### 2023-04-08 latex/prelims.tex
Almost done?

@@
 \subsection{Overview}
-In this section I will define notations that will be used throughout
-the paper. The theoretical foundations for inverse problems over
-function spaces can be found in \cite{Stuart10}. Most of the notations
-are summarized in the conditions of Theorem \ref{thm:char}.
+In this section we will define notations that will be used throughout
+the paper. Notations presented in the section are summarized in
+Section \ref{subsec:notation}. The theoretical foundations for inverse
+problems over function spaces can be found in \cite{Stuart10}.
 
@@
 \subsubsection{Error Terms}
-Considering the sum of the error terms, it is easy to verify that
-$\bar{\eps} := \obs \eps' + \eps \in \R^m$. It is a centered Gaussian
-random vector and its covariance matrix is
+It is easy to verify that $\bar{\eps} := \obs \eps' + \eps \in \R^m$
+is a centered Gaussian random vector with covariance
+matrix
+
 \begin{align}\label{eq:Sigma}
   \begin{split}
-    \Sigma(\obs) :&= \mathbb{E}[ (\obs \eps' + \eps)  (\obs \eps' + \eps)^t ] 
+    \Sigma(\obs) :&= \mathbb{E}[ (\obs \eps' + \eps) (\obs \eps' +
+      \eps)^t ]
     % 
@@
 \end{equation}
-with $e_j \in \R^m$ the $j$th standard basis vector. The first
+and $e_j \in \R^m$ is the $j$th standard basis vector. The first
 equality in \eqref{eq:modcov explained} holds by definition and the
@@
 $\bar{\eps} \sim \normal(0,\Sigma)$. Taking $\modcov = 0$ is common
-practice \cite{Tarantola05,Kaipio2006,Vogel02}. This reduces to
-the case where we take iid observations. Then $\Sigma = \sigma^2I$ is
+practice \cite{Tarantola05,Kaipio2006,Vogel02}. This reduces to the
+case where we take iid observations. Then $\Sigma = \sigma^2I$ is
 simply a scalar matrix that does not depend on $\obs$. Note that
@@
 
+\subsection{Notation Summary}\label{subsec:notation}
+  Our notation is summarized below. We let:
+  \begin{itemize}
+  \item $\hilp, \hilo$ Hilbert spaces.
+  \item $\fwd:\hilp \to \hilo$ a linear compact operator.
+  \item $\pr \sim \mathcal{N}(0, \prcov)$ prior Gaussian measure on
+    $\hilp$, with prior covariance operator $\prcov:\hilp \to \hilp$.
+  \item $\obs: \hilo \to \mathbb{R}^m$ observation operator, where $m
+    \in \mathbb{N}$ is the number of observations taken. 
+  \item $\sigma^2 \in \mathbb{R}_{+}$ observation noise variance.
+  \item $\modcov$ model error covariance operator.
+    %% $\data = \obs \fwd \param + \eps$, where $\eps \in
+    %% \mathbb{R}^m$ isiid $\mathcal{N}(0, \sigma^2)$ noise.
+  \item $\Sigma(\obs) = \obs \modcov \obs^* + \sigma^2I$. 
+  \item $\post$ the posterior measure, with covariance $\postcov$.
+  \item A D-optimality utility function
+    \cite{AlexanderianGloorGhattas14}:
+    \begin{align*}
+      \begin{split}
+        \tar(\obs) :&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
+        % 
+        % 
+        % 
+        &= \frac12 \log \det ( I + \prcov^{1/2} \fwd ^* \obs^* \Sigma(\obs)^{-1} \obs
+        \fwd \prcov^{1/2}).
+      \end{split}
+    \end{align*}
+  %% \item $\{\lambda_i\}_{i=1}^\infty$ eigenvalues of $\fwd\prcov\fwd^*$
+  %%   in decreasing order of magnitude.
+  %% %% \item $\{\ev_i\}_{i=1}^\infty$ their corresponding eigenvectors.
+  %% \item $\{\eta_i\}_{i=1}^\infty$ eigenvalues of $\obs^*\obs$.
+  \end{itemize}
+
 %% \subsection{Sequential vs Simultaneous Optimization}\label{subsec:seq vs sim}

### 2023-04-08 latex/vanishing.tex
Almost done?

@@
 \section{Characterization of D-Optimal Designs with Vanishing Model Error}\label{section:vanishing}
-In this section we analyze D-optimal designs when no model error is
-present. We first derive a nonlinear eigenvalue problem in $\obs$ for
-D-optimal designs, and then we characterize D-optimal designs as ones
-that make all eigenvalues equal in the nonlinear eigenvalue
-problem. Results are summarized in Theorem \ref{thm:char}.
+We now turn to analyze D-optimal designs when $\modcov = 0$. Our goal
+is to prove Theorem \ref{thm:char} --- a characterization of D-optimal
+designs.
 
 %% \subsection{The Nonlinear Eigenvalue Problem}\label{subsec:eigenproblem}
-The necessary first-order condition for D-optimality
-\eqref{eq:conditions}, when $\Sigma= \sigma^2I$ (i.e. when $\modcov =
-0$) becomes
+The necessary first-order condition for D-optimality summarized in
+Theorem \ref{thm:constrained}, when $\Sigma= \sigma^2I$ (i.e. when
+$\modcov = 0$) becomes
 \begin{equation}\label{eq:eigenproblem}
@@
 refer to \eqref{eq:eigenproblem} as a \emph{nonlinear} eigenvalue
-problem. The following two lemmas allow us to gain better
-understanding of the nonlinear eigenvalue problem.
+problem. The following Proposition and Lemma are proved in the
+appendix:
 
-\begin{restatable*}[Two applications of Woodbury Matrix Identity]{lemma}{woodbury}\label{lemma:twice woodbury}
+\begin{restatable*}[Two applications of Woodbury Matrix Identity]{proposition}{woodbury}\label{prop:twice woodbury}
   Assume $\fwd \prcov \fwd^*$ is invertible. Then
@@
 \end{restatable*}
-
-%% The term $(\fwd \prcov \fwd^*)^{-1}$ is the prior precision in
-%% $\hilo$, and $\fwd\prcov\fwd^*$ is invertible, as mentioned in
-%% Section \ref{section:prelim}. $(\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}
-%% \obs^* \obs$ is the posterior precision in $\hilo$.
-
-%% \subsection{Characterizing D-Optimal Designs}\label{subsec:characterization}
-%% %% Denote corresponding eigenvectors $\{\ev_i\}_{i=1}^{\infty}$.
-%% Denote eigenvalues of $\fwd \prcov \fwd^*$ by
-%% $\{\lambda_i\}_{i=1}^{\infty}$, and note they are positive. Lemma
-%% \ref{lemma:sim diag} is crucial for the rest of the analysis:
 
@@
       %
-      &= \sigma^{-2} \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1} \obs^* \text{ (by Lemma \ref{lemma:twice woodbury})}.
+      &= \sigma^{-2} \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1} \obs^* \text{ (by Proposition \ref{prop:twice woodbury})}.
     \end{split}
@@
 
-We are almost ready to state and prove Theorem \ref{thm:char} that
-characterizes D-optimal designs. We just need one final lemma, whose
-proof is delegated to the appendix:
+We are almost ready to prove Theorem \ref{thm:char} that characterizes
+D-optimal designs. We just need one final lemma, whose proof is
+delegated to the appendix:
 \begin{restatable*}[Unit norm decomposition]{lemma}{free}\label{lemma:free}
@@
   $\postcov^{-1} = \prcov^{-1} + \sigma^{-2}\fwd^*\obs^*\obs\fwd$. The
-  first statement in part (5) now follows from Lemma \ref{lemma:twice
+  first statement in part (5) now follows from Lemma \ref{prop:twice
     woodbury}, while the second statement follows from parts (1) and
@@
   \caption{A comparison of the eigenvalues of the pushforward
-    posterior precision $(\fwd\prcov\fwd^*)^{-1}
-    + \sigma^{-2}\obs^*\obs$ for a D-optimal design (left) and a
+    posterior precision $(\fwd\prcov\fwd^*)^{-1} +
+    \sigma^{-2}\obs^*\obs$ for a D-optimal design (left) and a
     sub-optimal design (right). Both designs are allowed three
     observations. We assume $\sigma^2=1$ and thus, the blue area has
-    accumulated height of $3$ in both plots. A D-optimal design (left)
-    increases precision (reduces uncertainty) where it is lowest. A
-    sub-optimal design (right) does not.
+    accumulated height of $3$ in both plots. The D-optimal design
+    (left) increases precision (equivalently, reduces uncertainty)
+    where it is lowest. The sub-optimal design (right) does not.
     %% In this example, an observation is taken partly in the
@@
 \obs_2^*\obs_2$. The former exhibits clusterization ($\meas_3 =
-\meas_5 = \ev_1$), while the latter does not. Since both of these
-designs are D-optimal, clusterization does not pose any obstruction to
+\meas_5 = \ev_1$), while the latter does not. Both of these designs
+are D-optimal, hence clusterization poses no obstruction to
 D-optimality.
@@
   \begin{pmatrix}
-    \sqrt{0.275}\ev_1 - \sqrt{0.725}\ev_2 \\
-    \sqrt{0.275}\ev_1 + \sqrt{0.725}\ev_2 \\
+    \sqrt{\frac{11}{40}}\ev_1 - \sqrt{\frac{29}{40}}\ev_2 \\
+    \sqrt{\frac{11}{40}}\ev_1 + \sqrt{\frac{29}{40}}\ev_2 \\
     \ev_1 \\
@@
 A computer implementation of Lemma \ref{lemma:free} (see code in
-Supplementary material) generates $\obs_1$, which exhibits
-clusterization. Modifying the numerical values of the prior does not
-break the clusterization. Thus, the preference to clusterization of
-Lemma \ref{lemma:free} is robust. The vast majority of randomized
-numerical simulations (see code in Supplementary material) also give
-rise to D-optimal designs that exhibit clustrization. We conclude that
-clusterization is a generic property of the relaxed model presented in
-this study.
+Supplementary material) generates $\obs_1$ as a solution to the
+D-optimal design problem mentioned in the previous
+paragraph. Modifying the numerical values of the prior does not break
+the clusterization. Thus, the preference to clusterization of Lemma
+\ref{lemma:free} is robust. The vast majority of randomized numerical
+simulations (see code in Supplementary material) also give rise to
+D-optimal designs that exhibit clustrization. We conclude that
+clusterization poses no obstruction to D-optimality in the relaxed
+model presented in this study. Moreover, sensor clusterization is
+generic, as far as one accepts the model presented in this study.
 
@@
       designs}. Both designs are identical for all practical matters
-    --- their posteriors are equal. The left panel shows a D-optimal
-    design with clusterization: $\meas_3 = \meas_5$. The right panel
-    shows a D-optimal design without clusterization. The contribution
-    to the posterior precision due to $\meas_1$ and $\meas_2$ is
-    denoted $\meas_1\meas_2$. These contributions are plotted together
-    because $\meas_1$ and $\meas_2$ are not in the direction of any
-    eigenvector. Similarly for $\meas_3\meas_4$.}
+    --- their posteriors are equal. In the left panel, a D-optimal
+    design with clusterization is illustrated: $\meas_3 = \meas_5$. In
+    the right panel, a D-optimal design without clusterization is
+    illustrated. The contribution to the posterior precision due to
+    $\meas_1$ and $\meas_2$ is denoted $\meas_1\meas_2$. These
+    contributions are plotted together because $\meas_1$ and $\meas_2$
+    are not in the direction of any eigenvector. Similarly for
+    $\meas_3\meas_4$.}
   \label{fig:clusterization}

### 2023-04-09 latex/intro.tex
Getting there

@@
 
-Designing experiments is usually done by optimizing some \emph{design
-criterion}. This is true both for frequentists
-\cite{silvey2013,Ucinski05} as well as for Bayesians
-\cite{Chaloner1995} (see \cite{Chaloner1995} for an investigation of
-the analogy between the two approaches). Among the plethora of design
-criteria, \emph{D-optimal design} have one of the simplest and most
-appealing motivation in the Bayesian context \cite{Chaloner1995}: a
-D-optimal design maximizes the expected information gain (KL
-divergence \cite{CoverThomas91}) between posterior and prior. For a
-linear model in finite dimensions, with Gaussian prior and noise, a
-D-optimal design minimizes the determinant of the posterior covariance
-matrix. In a frequentist setting, a D-optimal design minimizes the
-volume of the uncertainty ellipsoid \cite[page 16]{Ucinski05}, but
-this is done for the Fisher information matrix and not the posterior
-covariance. However, \cite{Chaloner1995} show that the latter is just
-a regularized version of the former.
+The problem of choosing the best set of measurements is known as the
+problem of \emph{optimal experimental design}. Optimal designs are
+found by optimizing some \emph{design criterion}. Many design criteria
+exist, and in this study we focus on the Bayesian \emph{D-optimality
+criterion}. The D-optimality criterion has a simple and appealing
+motivation \cite{Chaloner1995}: a D-optimal design maximizes the
+expected information gain (KL divergence \cite{CoverThomas91}) between
+posterior and prior. For a linear model in finite dimensions, with
+Gaussian prior and iid Gaussian noise, a D-optimal design minimizes
+the determinant of the posterior covariance matrix\footnote{In a
+frequentist setting, a D-optimal design minimizes the determinant of
+the Fisher information matrix \cite[page 16]{Ucinski05}, so the
+Bayesian criterion is just a regularized version of the frequentist
+criterion \cite{Chaloner1995}.}.
 
-The previous discussion is classical for experimental design when
-inference takes place over a finite (not too large) number of
-parameters. The subject of optimal experimental design on function
-spaces in a Bayesian context was pioneered by
-\cite{AlexanderianGloorGhattas14, AlexanderianPetraStadlerEtAl16,
+Bayesian D-optimal designs for inverse problems on function spaces
+were first studied by \cite{AlexanderianGloorGhattas14,
+  AlexanderianPetraStadlerEtAl16,
   AlexanderianPetraStadlerEtAl14}. Similarly to the finite dimensional
-case, a D-optimal design still maximizes the KL divergence between
-posterior and prior. For a linear model with Gaussian prior and noise,
-this amounts to minimizing the determinant of the posterior covariance
-operator (understood as a product of its eigenvalues). Some
-difficulties arise in the process, but remedies can be found as shown
-in \cite{AlexanderianGloorGhattas14}.
+case, a D-optimal design maximizes the KL divergence between posterior
+and prior measures. For a linear inverse problem with Gaussian prior
+and iid Gaussian noise, a D-optimal design minimizes an expression
+analogous to the determinant of the posterior covariance operator,
+understood as a product of its eigenvalues.
+%% Some difficulties arise in the process, but remedies can be found
+%% as shown in \cite{AlexanderianGloorGhattas14}.
 
-It seems counter intuitive that when one computes an optimal design
-using the D-optimality criterion, the optimization process results in
-observations that are very similar. For example, if an observation is
-thought of as measuring some function value at $\x \in \Omega
-\subseteq \R^d, d=1,2,3$ (with added error) then the optimization
-procedure sometimes places sensors in very close proximity to each
-other (as can be seen in Figure \ref{fig:clusterization
-  illustration}). Following \cite{Ucinski05}, we refer to this
-phenomenon as \emph{sensor clusterization}.
+When one computes an optimal design using the D-optimality criterion,
+the optimization process sometimes results in observations that are
+very similar. For example, if an observation is thought of as
+measuring some function value at $\x \in \Omega \subseteq \R^d,
+d=1,2,3$ (with added error) then the optimization procedure sometimes
+places sensors in very close proximity to each other (as can be seen
+in Figure \ref{fig:clusterization illustration}). Following
+\cite{Ucinski05}, we refer to this phenomenon as \emph{sensor
+clusterization}.
 
-\subsection{Related Work}
-The phenomenon of sensor clusterization seems to be known in several
-different contexts. In a frequentist and finite-dimensional context,
-\cite{fedorov1996} and \cite[chapter 2.4.3]{Ucinski05} discuss this
-phenomenon and suggest an approach called clusterization-free design.
-In such designs, the user enforces observation locations to be far
-from each other. One way to do this is by introducing correlated
-errors which account for both observation error and model
-error. Another method considered is imposing distance constraints
-between observations. A somewhat different approach is suggested in
-\cite[page 49]{fedorov2012}, where close observations are merged --- a
-procedure which obviously does not avoid clusterization. The same
-problem arises in time-series analysis for pharmacokinetic
-experiments. The authors of \cite{hooker2009} suggest modeling
-auto-correlation time in the noise model, which is equivalent to the
-correlated errors mentioned above.
+Sensor clusterization is intuitively undesirable. Selecting a small
+set of sensors may lead to incomplete or biased data, as certain parts
+of the system or environment may be overlooked or
+underrepresented. This can result in inaccurate or unreliable
+estimates of the unknown parameters of interest, potentially leading
+to incorrect conclusions or decisions.
 
-Any of the above mentioned approaches might serve as a remedy and push
-sensors away from each other. Yet, none offers any insight as to why
-clusterization occurs. Also, as better models are employed, model
-error is decreased and the clusterization phenomenon will eventually
-reappear. While these approaches are practical and help us avoid the
-problem, they do not provide insight as to why sensors are clustering.
+%% \subsection{Related Work}
+Several remedies are suggested in the literature to phenomenon of
+sensor clusterization. In a frequentist and finite-dimensional
+context, \cite{fedorov1996} and \cite[chapter 2.4.3]{Ucinski05}
+discuss the clusterization phenomenon and suggest an approach called
+clusterization-free design. In such designs, the user enforces
+observation locations to be far from each other. One way to do this is
+by introducing correlated errors which account for both observation
+error and model error. Another method considered is imposing distance
+constraints between observations. A somewhat different approach is
+suggested in \cite[page 49]{fedorov2012}, where close observations are
+merged --- a procedure which obviously does not avoid
+clusterization. The same problem arises in time-series analysis for
+pharmacokinetic experiments. The authors of \cite{hooker2009} suggest
+modeling auto-correlation time in the noise model, which is equivalent
+to the correlated errors mentioned above. Any of the above mentioned
+approaches might serve as a remedy and push sensors away from each
+other. Yet, none offers any insight as to why clusterization
+occurs. Also, as better models are employed, model error is decreased
+and the clusterization phenomenon will eventually reappear. While
+these approaches are practical and help us avoid sensor
+clusterization, they do not provide insight as to why sensors are
+clustering.
 
-In the inverse problems community, work is mostly computational and
-less theoretic. Model errors were considered in \cite{attia2020,
-  koval2020}. The former study is focused on inferring a Quantity of
-Interest (QoI). The focus of the latter is reducing forward solves,
-using randomized linear algebra. Both studies present numerical
-techniques for finding optimal designs when model error is
+Model errors were also considered for inverse problems in
+\cite{attia2020, koval2020}. The former study is focused on inferring
+a Quantity of Interest (QoI). The focus of the latter is reducing
+forward solves, using randomized linear algebra. Both studies present
+numerical techniques for finding optimal designs when model error is
 present. Both are restricted to linear inverse problems (although in
@@
 design. Both studies are successful in the task of Bayesian
-inversion. However, neither of these studies mention any effect model
-errors can have on sensor clusterization. This study is mostly
-theoretical and aims to fill the gap of understanding sensor
-clusterization.
+inversion. Neither \cite{attia2020}, nor \cite{koval2020} mention any
+effect model errors can have on sensor clusterization.
 
 \subsection{Contribution}
-We propose and study a relaxed and analytically tractable model for
-D-optimal designs. Under this model, D-optimal designs are solutions
-of a constrained optimization problem, formulated using Lagrange
-multipliers. This allows me to rigorously show how model error
-mitigates clusterization (Section \ref{section:non vanishing}). We
-show that designs that exhibit clusterization are indeed optimal.
+The current study aims to fill the gap of understanding sensor
+clusterization in infinte dimensional Bayesian linear inverse
+problems. We propose and study a relaxed and analytically tractable
+model for D-optimal designs. Under this model, D-optimal designs are
+solutions of a constrained optimization problem, formulated using
+Lagrange multipliers. The Lagrange multipliers formulation allows us
+to rigorously show how model error mitigates clusterization (Section
+\ref{section:non vanishing}), and why designs that exhibit
+clusterization are indeed optimal when no model error is considered.
 
-%% A beautiful mathematical structure arises in D-optimal designs
-
-When no model error is present, the Lagrange multipliers problem boils
-down to a nonlinear eigenvalue problem.
-%% The operator for which eigenvectors and eigenvalues are sought is a
-%% sum of two operators. The first is the prior covariance. The second
-%% is an outer product of the observations (see section
-%% \ref{section:vanishing} for details and exact statement).
-This structure helps us to characterize D-optimal designs. The key
-idea is that a D-optimal design reduces uncertainties where they are
-highest first. This is the main conclusion of Theorem \ref{thm:char}
-stated below.
+As a byproduct, the Lagrange multipliers problem for no model error
+results in a characterization of D-optimal designs. The key idea is
+that a D-optimal design reduces uncertainties where they are highest
+first. This is the conclusion of part (5) of Theorem \ref{thm:char},
+which we state below and prove in Section \ref{section:vanishing}:
 
@@
 
-%% \begin{restatable*}[Flat posterior covariance]{corollary}{flatpost}\label{cor:equal eigs}
-%%  %\begin{corollary}\label{cor:eigenvalues}
-%%   If $\obs$ is D-optimal and $k := \rank \obs$, then:
-%%   \begin{enumerate}
-%%     \item The covariance of the pushforwad $\fwd_{*} \post$ is $\left ( (\fwd
-%%       \prcov \fwd^*)^{-1} + \sigma^{-2} \obs^*\obs \right )^{-1}$.
-%%     \item The eigenvalues of said operator are
-%%       \begin{equation}
-%%         \theta_i =
-%%         \begin{cases}
-%%           \left(\frac{1}{k} \sum_{j=1}^k \lambda_j^{-1} + \sigma^{-2}\frac{m}{k} \right )^{-1} & i \leq k \\
-%%           \lambda_i &  i > k 
-%%         \end{cases}
-%%       \end{equation}
-%%     \item The eigenvectors of said operator are the eigenvectors of
-%%       $\fwd \prcov \fwd^*$
-%%   \end{enumerate}
-%% \end{restatable*}%\end{corollary}
-
-%% \begin{restatable*}[Flat posterior covariance]{corollary}{flatpost}\label{cor:equal eigs}
-%%   Consider a D-optimal design $\obs$ for a Bayesian linear inverse
-%%   problem where $\fwd$ is the forward operator and $\pr$ is the
-%%   prior. Then the eigenvectors of the covariance of the pushforward of
-%%   the posterior $\fwd_{*} \post$ equal the eigenvectors of the
-%%   covariance of the pushforward of the prior. Moreover, the
-%%   eigenvalues are constant for eigenvectors $1 \leq i \leq k$ where $k
-%%   =\rank \obs$.
-%% \end{restatable*}
-
-In the process, We generalize several lemmas from linear algebra to
-infinite-dimensional settings. We prove a Matrix Determinant Lemma in
-\ref{lemma:MDL}. We generalize a lemma due to Lax \cite{Lax07} for
-calculating $\frac{\der}{\der t} \log \det (I + X(t))$, for an
-operator valued function $X(t)$ in \ref{lemma:lax}. We show how to
-decompose a symmetric positive definite matrix $m$ as $M = AA^t$,
-where $A$ has unit norm columns in Lemma \ref{lemma:free}.
+In the process of proving Theorem \ref{thm:char}, we prove a novel
+lemma in linear algebra: We show how to decompose a symmetric positive
+definite matrix $M$ as $M = AA^t$, where $A$ has unit norm columns
+(Lemma \ref{lemma:lax}). We also generalize two lemmas from linear
+algebra to infinite-dimensional settings: In Lemma \ref{lemma:MDL} we
+prove a matrix determinant lemma in Hilbert spaces, and in Lemma
+\ref{lemma:lax} we generalize a lemma \cite{Lax07} for calculating
+$\frac{\der}{\der t} \log \det (I + X(t))$, where $X(t)$ is an
+operator valued function.
 
 \subsection{Limitations}\label{subsec:limitations}
-Two main drawbacks of the study presented here are: First, the relaxed
-model does not consider any specific set of allowed
-observations. Rather, we take observations in the unit ball in some
-Hilbert space. This allows considerably less restrictive observations
-than any real-life problem does. The second drawback is that we do not
-show rigorously that clusterization necessarily occurs, we only show
-that it is as reasonable as no clusterization.
+The study presented here is not free from drawbacks. First, our
+relaxed model does not correspond to any real-life problem. It is
+generic enough to be analytically tractable, but one may argue it is
+to far removed from any real application. To these claims I would
+argue that Mathematics has a long history of studying models that are
+bare-bones simplifications of real systems, e.g. the Ising model
+\cite{cipra1987} the Lorenz system \cite{brin}, the SIR model
+\cite{keeling2008}, to name a few. The second drawback of our model is
+that we do not show rigorously that clusterization necessarily occurs,
+we only show that it does not pose any obstruction to D-optimality.
 

### 2023-04-09 latex/paper.tex
Almost done?
Getting there

@@
 
-\title[Generic Properties of D-optimal Designs]{Generic Properties of D-optimal
-  Designs in Infinite Dimensional Bayesian Linear Inverse Problems}
+\title[Clusterization Properties of D-optimal Designs]{Clusterization
+  Properties of D-optimal Designs in Infinite Dimensional Bayesian
+  Linear Inverse Problems}
 

### 2023-04-10 latex/intro.tex
backup

@@
 to far removed from any real application. To these claims I would
-argue that Mathematics has a long history of studying models that are
-bare-bones simplifications of real systems, e.g. the Ising model
-\cite{cipra1987} the Lorenz system \cite{brin}, the SIR model
+argue that Mathematicians have a long history of studying models that
+are bare-bones simplifications of real systems, e.g. the Ising model
+\cite{cipra1987} the Lorenz system \cite{brin}, and the SIR model
 \cite{keeling2008}, to name a few. The second drawback of our model is
@@
   \end{tikzpicture}
-  \caption{The clusterization effect for the 1D heat equation
-    described in Section \ref{subsec:example}. Posterior pointwise
-    standard deviations (lines) are plotted over the domain $[0,
-      \pi]$, for varying numbers of sensors. Sensor locations
-    (circles) were chosen to minimize (an expression analogous to) the
-    determinant of the posterior covariance. The clusterization effect
-    can be clearly seen for six sensors. Only four observation
-    locations are used --- two pairs of sensors are so close they are
+  \caption{Sensor clusterization for inferring the initial condition
+    of the 1D heat equation. Posterior pointwise standard deviations
+    (lines) are plotted over the domain $[0, \pi]$, for varying
+    numbers of sensors. Sensor locations (circles) were chosen to
+    optimize the D-optimality criterion. Sensor clusterization can be
+    clearly seen for six sensors. Only four observation locations are
+    used --- two pairs of sensors are so close together they are
     indistinguishable.}
@@
 
-In Section \ref{section:prelim}, we present a more abstract and general
-formulation of the inverse problem we consider. But, for the purpose of
-illustration, we present clusterization via a toy model --- the 1D heat
+For the purpose of illustration, we present clusterization via a toy
+inverse problem --- inferring the initial condition of the 1D heat
 equation in $[0,\pi]$ with a homogeneous Dirichlet boundary condition.
@@
 
-The goal is to infer the initial condition $u_0$. For that purpose, we
-measure $u$ at some set of locations $\x_j \in [0,\pi], j=1, \dots,m$
-and a final time $T > 0$. We assume centered Gaussian observation
+Our goal here is to infer the initial condition $u_0$. We measure
+$u(\x_j, T)$ at some set of locations $\x_j \in [0,\pi], j=1, \dots,m$
+and a final time $T > 0$. We assume iid centered Gaussian observation
 error, so we observe $v(\x_j,T) = u(\x_j,T) + \eps(\x_j)$ with
@@
 has a closed form formula, even in infinite dimensions
-\cite{Stuart10}. We denote by $\fwd$ the dynamics operator, so that
-$u( \cdot,T) = \fwd u_0$, and the observation operator $\obs$ so that
+\cite{Stuart10}. We denote by $\fwd$ the forward operator, so that $u(
+\cdot,T) = \fwd u_0$, and the observation operator $\obs$ so that
 $u(\x_j,T) = (\obs u)_j, j=1,\dots,m$. The posterior covariance is
@@
 
-We consider generalization of the information-theoretic design
-criterion presented in the introduction to infinite dimensions
-(Section \ref{subsec:D optimal design} below). We choose
-$\x_j,j=1,\dots,m$ to minimize (an expression analogous to) the
-determinant of the posterior covariance operator. We will see later
-how this corresponds to maximizing expected information gain.
+We consider the generalization of the D-optimality criterion to
+infinite dimensions \cite{AlexanderianGloorGhattas14} (also, Section
+\ref{subsec:D optimal design} below). We choose $\x_j,j=1,\dots,m$ to
+minimize the determinant\footnote{Not exactly, see Theorem \ref{thm:}
+for a precise statement.} of the posterior covariance operator.
 
-The clusterization effect is illustrated in Figure
-\ref{fig:clusterization illustration}. Posterior pointwise standard
-deviations are plotted over the domain $[0, \pi]$. Since the posterior
-covariance does not depend on data, the plot has no reference to
-actual data observed. The posterior covariance does, however, depend
-on location of the observation taken. In Figure
-\ref{fig:clusterization illustration}, observation locations are
-marked by circles. These were chosen to minimize (an expression
-analogous to) the determinant of the posterior covariance. The
-clusterization effect can be clearly seen for six sensors. It looks
-like only four observations were taken. The reason is that two pairs
-of sensors are so close they are indistinguishable.
+Sensor clusterization is illustrated in Figure \ref{fig:clusterization
+  illustration}. Posterior pointwise standard deviations are plotted
+over the domain $[0, \pi]$. Since the posterior covariance does not
+depend on data, the plot has no reference to actual data observed. The
+posterior covariance does, however, depend on location of the
+observation taken. Observation locations were chosen to optimize the
+D-optimality criterion, and are marked by circles. Sensor
+clusterization can be clearly seen for six sensors; it looks like only
+four observations were taken. The reason is that two pairs of sensors
+are clustered so close together that they are indistinguishable from
+one another.

### 2023-04-10 latex/vanishing.tex
backup

@@
 %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Characterization of D-Optimal Designs with Vanishing Model Error}\label{section:vanishing}
+\section{D-Optimal Designs with Vanishing Model Error}\label{section:vanishing}
 We now turn to analyze D-optimal designs when $\modcov = 0$. Our goal
@@
 
-%% \subsection{The Nonlinear Eigenvalue Problem}\label{subsec:eigenproblem}
 The necessary first-order condition for D-optimality summarized in
@@
 
-\section{Clusterization poses no obstruction to D-optimality}\label{section:clusterization}
+\subsection{Clusterization poses no obstruction to D-optimality}\label{subsec:clusterization}
 We saw that the necessary conditions for a design $\obs$ with $m$
@@
 \end{figure}
-
-%% may arise in various scenarios. We first consider
-%% sequential designs, and then simultaneous designs.
-
-%% \subsection{Sequential Design}\label{subsec:clusterization sequential}
-%% In a sequential optimal design problem, a decision is made on each
-%% measurement with all previous measurements fixed. In each step $m=1$,
-%% so $k=1$. In this case, Theorem \ref{thm:char} implies each
-%% measurement is taken along the direction of the eigenvector of $\fwd
-%% \prcov \fwd^*$ with lowest posterior precision. If $\lambda_i$
-%% decrease fast enough (e.g. exponentially), then $\lambda_i^{-1}$
-%% increase exponentially. Then, at some point, a measurement will be
-%% taken along an already measured eigenvector. This situation is
-%% measurement clusterization.
-
-%% , $\meas_1 \parallel \ev_1$ and $\eta_1 = 1$. As hinted in
-%% section \ref{subsec:seq vs sim}, the posterior becomes the prior for
-%% the next step. The eigenvectors of the new posterior are the same as
-%% for the previous one. A new observation will be taken in the direction
-%% of eigenvector with smalles eigenvalue. This leads to:
-%% \begin{align*}
-%%   \begin{split}
-%%     \meas_2 \parallel \ev_1  &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 < \sigma^2\lambda^{-1}_2 \\
-%%     \meas_2 \parallel \ev_2  &\text{ o.w. }
-%%   \end{split}
-%% \end{align*}
-%% In the first case, we get clusterization immediately. In the second,
-%% we consider $\meas_3$. Same reasoning leads to the conclusion that the
-%% eigenvectors of the new prior precision are the same. The eigenvalues,
-%% however, are $\lambda^{-1}_1 + \sigma^{-2}, \lambda^{-1}_2 +
-%% \sigma^{-2}, \lambda_3^{-1}, \dots$. A new observation is taken in the
-%% direction of eigenvector of smallest precision (note that
-%% $\lambda^{-1}_1 + \sigma^{-2} < \lambda^{-1}_2 + \sigma^{-2}$):
-%% \begin{align*}
-%%   \begin{split}
-%%     \meas_3 \parallel \ev_1 &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 <
-%%     \sigma^2\lambda^{-1}_3 \\
-%%     \meas_3 \parallel  \ev_3  &\text{ o.w. }
-%%   \end{split}
-%% \end{align*}
-%% Again, the first case results in clusterization. The sequential design
-%% proceeds in this fashion. Every $\meas_i$ is taken in the direction
-%% some eigenvector of the prior $\fwd \prcov\fwd^*$. This eigenvector
-%% has the smallest eigenvalue of the previous step's posterior precision
-%% operator. If we do not encounter clusterization after $m$ such steps,
-%% this means that observation $\meas_k$ was taken in the direction
-%% $\ev_k$ for $k=1,\dots,m$. Thus:
-%% \begin{align}\label{eq:failure}
-%%   \begin{split}
-%%     \sigma^2\lambda_{k-1}^{-1} + 1 \geq \sigma^2\lambda_k^{-1}, \text{ for } 1\leq k \leq m
-%%     \Longrightarrow \lambda_m^{-1} - \lambda_{m-1}^{-1} \leq \sigma^{-2}.
-%%   \end{split}
-%% \end{align}
-
-%% But \eqref{eq:failure} implies $\{\lambda_i^{-1}\}_{i=1}^{\infty}$
-%% grows at most linearly. In section \ref{subsec:abstract OED} we
-%% assumed $\fwd$ is strongly smoothing, which means eigenvalues of
-%% $(\fwd \prcov\fwd^*)^{-1}$ grow quickly. Thus, if we take $m$ large
-%% enough, \eqref{eq:failure} will eventually fail and we will observe
-%% sensor clusterization.
-
-%% \subsection{Simultaneous Design}\label{subsec:clusterization simultaneous}
-%% If the eigenvalues $\lambda_i$ decrease quickly enough, we will
-%% observe clusterizaion.

### 2023-04-13 latex/intro.tex
Need to write intro. Tell how everybody thinks model error is the solution

@@
 
-For the purpose of illustration, we present clusterization via a toy
-inverse problem --- inferring the initial condition of the 1D heat
-equation in $[0,\pi]$ with a homogeneous Dirichlet boundary condition.
+For the purpose of illustration, consider a toy inverse problem ---
+inferring the initial condition of the 1D heat equation in $[0,\pi]$
+with a homogeneous Dirichlet boundary condition. Sensor clusterization
+is clearly evident, see Figure \ref{fig:clusterization}.
 
-The 1D heat equation is:
+The 1D heat equation with homogeneous Dirichlet boundary condition is:
 \begin{subequations}\label{eq:heat equation}
@@
 
-Our goal here is to infer the initial condition $u_0$. We measure
-$u(\x_j, T)$ at some set of locations $\x_j \in [0,\pi], j=1, \dots,m$
-and a final time $T > 0$. We assume iid centered Gaussian observation
-error, so we observe $v(\x_j,T) = u(\x_j,T) + \eps(\x_j)$ with
-$\eps(\x_j) \sim \normal(0, \sigma^2)$ iid, $\sigma^2 > 0$. We model
-the initial condition as $u_0 \sim \normal(0,\prcov)$, for $\prcov =
-(-\Delta)^{-1}$ with a homogeneous Dirichlet boundary condition. It is
-known \cite{Tarantola05} that for linear problems, with Gaussian prior
-and error, the posterior is also Gaussian with a covariance that does
-not depend on the observed data. The posterior covariance $\postcov$
-has a closed form formula, even in infinite dimensions
-\cite{Stuart10}. We denote by $\fwd$ the forward operator, so that $u(
-\cdot,T) = \fwd u_0$, and the observation operator $\obs$ so that
-$u(\x_j,T) = (\obs u)_j, j=1,\dots,m$. The posterior covariance is
-known and depends only on $\prcov, \fwd, \obs$ and $\sigma^2$ (see
-Section \ref{section:prelim} and \eqref{eq:postcov} specifically).
-
-We consider the generalization of the D-optimality criterion to
-infinite dimensions \cite{AlexanderianGloorGhattas14} (also, Section
-\ref{subsec:D optimal design} below). We choose $\x_j,j=1,\dots,m$ to
-minimize the determinant\footnote{Not exactly, see Theorem \ref{thm:}
-for a precise statement.} of the posterior covariance operator.
+The goal is to infer the initial condition $u_0$. Model the initial
+condition as $u_0 \sim \normal(0,\prcov)$, for $\prcov =
+(-\Delta)^{-1}$ with a homogeneous Dirichlet boundary condition. Fix
+some final time $T$ and denote by $\fwd$ the forward operator, so that
+$u(\cdot, T) = \fwd u_0$. Let $\obs$ the observation operator, so that
+$u(\x_j,T) = (\obs u)_j, j=1,\dots,m$. Assume iid centered Gaussian
+observation error, so observations are given by $u(\x_j,T) +
+\eps(\x_j)$ with $\eps(\x_j) \sim \normal(0, \sigma^2)$ iid, $\sigma^2
+> 0$. The posterior covariance is known and depends only on $\prcov,
+\fwd, \obs$ and $\sigma^2$ \cite{Stuart10} (see Section
+\ref{section:prelim} and \eqref{eq:postcov} below). Finally, choose
+$\x_j,j=1,\dots,m$ to optimize the D-optimality criterion
+\cite{AlexanderianGloorGhattas14}.
 
 Sensor clusterization is illustrated in Figure \ref{fig:clusterization
-  illustration}. Posterior pointwise standard deviations are plotted
-over the domain $[0, \pi]$. Since the posterior covariance does not
-depend on data, the plot has no reference to actual data observed. The
-posterior covariance does, however, depend on location of the
-observation taken. Observation locations were chosen to optimize the
-D-optimality criterion, and are marked by circles. Sensor
-clusterization can be clearly seen for six sensors; it looks like only
-four observations were taken. The reason is that two pairs of sensors
-are clustered so close together that they are indistinguishable from
-one another.
+  illustration}. The posterior covariance does not depend on data, so
+the plot has no reference to actual data observed. The posterior
+covariance does depend on location of the observation taken, which
+were chosen to optimize the D-optimality criterion. Sensor
+clusterization is clearly seen for six sensors; it looks like only
+four observations were taken. Two pairs of sensors are clustered so
+close together that they are indistinguishable from one another.
+
+%% A D-optimal design
+%% is found via a generalization of the D-optimality criterion to
+%% infinite dimension(also, Section
+%% \ref{subsec:D optimal design} below).

### 2023-04-13 latex/paper.tex
Need to write intro. Tell how everybody thinks model error is the solution

@@
   problems. We suggest an analytically tractable model for D-optimal
-  designs and reason how it may lead to sensor clusterization in the
-  case of iid observation noise. We also consider how spatially
-  correlated observation and model errors mitigate the clusterization
-  phenomenon.
+  designs. We show that sensor clusterization in the case of iid
+  observation noise is generic. We also show that model spatially
+  correlated errors mitigate the clusterization phenomenon.
 
@@
 This study is a part of my PhD thesis. Most of it was written under
-the instruction of Prof. Georg Stadler in New York University's
-Courant Institute. I would like to thank him for his great
-mentorship.
+the instruction of Prof. Georg Stadler in the Courant Institute of
+Mathematical Sciences in New York University. I would like to thank
+Prof. Stadler for his great mentorship, attention to details and kind
+behavior.
 

### 2023-04-13 latex/prelims.tex
Need to write intro. Tell how everybody thinks model error is the solution

@@
 \subsection{Overview}
-In this section we will define notations that will be used throughout
-the paper. Notations presented in the section are summarized in
-Section \ref{subsec:notation}. The theoretical foundations for inverse
+In this section I will present notations that will be used throughout
+this article. These notations are summarized in Section
+\ref{subsec:notation}. The theoretical foundations for inverse
 problems over function spaces can be found in \cite{Stuart10}.
@@
 $\hilo^* \simeq \hilo$ the Hilbert space of all linear functionals on
-$\hilo$. Consider $\fwd: \hilp \to \hilo$, a linear operator (the
-``forward operator''). We are interested in forward operators that are
-strongly smoothing (have fast decaying modes --- the heat operator
-from section \ref{subsec:example} is a prime example). Our goal is to
-infer $\param \in \hilp$ --- some parameter of the dynamics --- given
-noisy observations of $\fwd \param$. We take a Gaussian prior $\pr$
-for $\param$. Hence, $\param \sim \pr = \normal(\prmean ,\prcov)$ with
-some appropriate covariance operator $\prcov$ \cite{Stuart10}. It is
-important to note that $\fwd \prcov \fwd^*$ is the prior covariance in
-$\hilo$ \cite{Stuart10}. As such, it is assumed invertible --- an
-assumption we rely on below. Throughout this paper, we denote $m$ the
-number of observations we are allowed to take. Observations are taken
-via the operator $\obs \in ( \hilo^* )^m$. In an analogy with linear
-algebra, where row vectors are thought of as linear functionals, we
-think of $\obs$ as having ``rows'' $\meas_j, j=1,\dots,m$:
+$\hilo$. Let $\fwd: \hilp \to \hilo$ the \emph{forward operator}. The
+forward operator $\fwd$ is assumed linear and strongly smoothing (has
+fast decaying modes --- the heat operator from section
+\ref{subsec:example} is a prime example). Take a Gaussian prior
+$\param \sim \pr = \normal(\prmean ,\prcov)$ with some appropriate
+covariance operator $\prcov$ \cite{Stuart10} on $\hilp$. Note that
+$\fwd \prcov \fwd^*$ is the prior covariance in $\hilo$
+\cite{Stuart10}. As such, it is assumed invertible --- an assumption
+which will be used heavily below. Let $m$ the number of observations
+taken. Observations are taken via the \emph{observation
+operator}\footnote{It is common for the observation and forward
+operators to be merged $\tmp := \obs \fwd$
+\cite{AlexanderianGloorGhattas14}, but the analysis carried out in the
+following sections requires $\fwd$ and $\obs$ are explicitly
+separated.} $\obs \in ( \hilo^* )^m$. In an analogy with linear
+algebra, where row vectors are thought of as linear functionals,
+$\obs$ is best thought of as having ``rows'' $\meas_j, j=1,\dots,m$:
 \begin{equation}\label{eq:O}
@@
 \end{equation}
-This way, for $u \in \hilo$ we have $\obs u = (\meas_1(u), \dots,
-\meas_m(u) )^t \in \R^m$.
+This way, $\obs u = (\meas_1(u), \dots, \meas_m(u) )^t \in \R^m$ and
+$u \in \hilo$.
 %% but we may drop the parentheses and write $\meas_j u = \meas_j(u)$.
@@
 
-Each observation is a linear functional, chosen from some subset of
-$\hilo^{*}$. Data is acquired via noisy observations
+Each observation vector is a linear functional, chosen from (some
+subset of) $\hilo^{*}$. Data is acquired via noisy observations, with
+two types of error terms considered. Spatially correlated model error
+is denoted $\eps' \sim \normal(0,\modcov)$, and modeled as a centered
+Gaussian measure on $\hilo$ with covariance operator
+$\modcov$. Observation error is denoted $\eps \sim \normal(0, \sigma^2
+I_m)$, with $I_m \in \mathbb{R}^{m \times m}$ the identity. Both error
+terms and the prior are assumed independent of each other. Thus, data
+is acquired via
 \begin{align}\label{eq:inverse problem}
-  \data := \obs (\fwd \param + \eps') + \eps = \obs \fwd \param + \obs \eps' + \eps,
+  \data := \obs (\fwd \param + \eps') + \eps = \obs \fwd \param + \obs \eps' + \eps.
 \end{align}
-with $\data \in \R^m$ and $\eps, \eps'$ defined next. We consider two
-types of error terms. First, there is spatially correlated model error
-$\eps' \sim \normal(0,\modcov)$, modeled as a centered Gaussian
-measure on $\hilo$ with covariance operator $\modcov$. Then there is
-observation error $\eps \sim \normal(0, \sigma^2 I_m)$, with $I_m$ the
-$m \times m$ identity. Both error terms and the prior are assumed
-independent of each other. In the literature, it is common for the
-observation and forward operators to be merged $\tmp := \obs \fwd$
-\cite{AlexanderianGloorGhattas14}. Since we need to separate the
-observation operator $\obs$ from the forward operator $\fwd$ in order
-to carry out the analysis in the following sections, we explicitly
-separate $\fwd$ and $\obs$ in \eqref{eq:inverse problem}. We now turn
-to understand the terms present in \eqref{eq:inverse problem}.
-
-\subsubsection{Dynamics and Observation Operators}\label{subsec:dynamics}
-Rows of $\obs$ are linear functionals and we refer to these as {\it
-  observation vector}s. Note that we cannot choose every observation
-vector we want. For example, we may be restricted only to pointwise
-evaluations of $\fwd \param$ by the sensors at our disposal, in which
-case measuring, e.g., the mean $\ell(u) = \int_{\Omega}u$ of a
-function is not possible. Thus, it is best to think of $(\obs \fwd
-\param)_j$ as approximations to pointwise evaluations. For example, if
-$\hilo = L^2([-\pi, \pi])$, consider $\ell_1(x) := (2\pi
-\eps)^{-\frac{1}{2}}\exp(-\frac{x^2}{2\eps^2})$ and $\ell_2(x)
-:=\eps^{-1}\mathbf{1}_{[-\frac{\eps}{2}, \frac{\eps}{2}]}$. Both
-$\ell_1$ and $\ell_2$ approximate a point measurement at $0$ for small
-$\eps$ and, crucially, $\ell_1, \ell_2 \in \hilo^*$.
-%% It is important to note that although $\|\ell\|_1 =
-%% \|\ell_2\|_1=1$, unfortunately $\|\ell_1\|_2^2 = (4\pi
-%% \eps)^{-\frac{1}{2}}$ and $\|\ell_2\|_2^2 = \eps^{-1}$.
-
-\subsubsection{Error Terms}
-It is easy to verify that $\bar{\eps} := \obs \eps' + \eps \in \R^m$
-is a centered Gaussian random vector with covariance
-matrix
+%% \subsubsection{Error Terms}
+It is easy to verify that $\obs \eps' + \eps \in \R^m$ is a centered
+Gaussian random vector with covariance matrix
 
@@
 where
-\begin{equation}\label{eq:modcov explained}
-  [\obs \modcov \obs^*]_{ij} = e_i^t \obs \modcov \obs^* e_j = \meas_i (\modcov \meas_j),
-\end{equation}
-and $e_j \in \R^m$ is the $j$th standard basis vector. The first
-equality in \eqref{eq:modcov explained} holds by definition and the
-second by \eqref{eq:obs*}. The explicit dependence on $\obs$ will be
-mostly dropped for notational convenience, so $\Sigma(\obs) =
-\Sigma$. Thus, for a fixed $\obs$ (i.e.\ a fixed set of observations)
-we can equivalently write $\data = \obs \fwd \param + \bar{\eps}$ with
-$\bar{\eps} \sim \normal(0,\Sigma)$. Taking $\modcov = 0$ is common
-practice \cite{Tarantola05,Kaipio2006,Vogel02}. This reduces to the
-case where we take iid observations. Then $\Sigma = \sigma^2I$ is
-simply a scalar matrix that does not depend on $\obs$. Note that
-taking an error model with a non-scalar covariance as we do here
-allows us to consider model error (modeled by $\modcov$) as well as
-observation error (modeled by $\sigma^2$). For example, say we believe
-our forward model does not capture some small scale phenomenon.  Then
-we may express this belief by saying $\tru = \fwd + \err$, with $\fwd$
-depending on $\param$ and $\err$ depending on $\sspar$, with $\param
-\perp \sspar$. We do not know much about this effect but it is
-reasonable to assume it changes continuously in our domain. We (may
-choose to) model it as $\normal (0, \modcov)$ and take $\modcov$ to
-reflect the spatial (or other) variability we imagine $\err \sspar$
-has. Such small scale phenomenon can arise as a modeling issue, where
-we might not model the system in its entirety. It can also arise from
-a numerical source, where our discretization of the system is not fine
-enough to capture all small scale phenomena.
+\begin{align}\label{eq:modcov explained}
+  \begin{split}
+    [\obs \modcov \obs^*]_{ij} & = e_i^t \obs \modcov \obs^* e_j \\
+    %
+    %
+    %
+    &= \meas_i (\modcov \meas_j) \text{ (by \eqref{eq:obs*})}.
+  \end{split}
+\end{align}
+%% and $e_j \in \R^m$ is the $j$th standard basis vector.
+%% The explicit dependence on $\obs$ will be mostly dropped for
+%% convenience, so $\Sigma(\obs) = \Sigma$.  Thus, for a fixed $\obs$
+%% (i.e.\ a fixed set of observations), $\data = \obs \fwd \param +
+%% \bar{\eps}$ with $\bar{\eps} \sim \normal(0,\Sigma)$.
+Taking $\modcov = 0$ is a common practice
+\cite{Tarantola05,Kaipio2006,Vogel02} and then $\Sigma = \sigma^2I$ is
+simply a scalar matrix that does not depend on $\obs$.
+
+%% Taking an error model with a non-scalar covariance allows us to
+%% consider model error (modeled by $\modcov$) as well as observation
+%% error (modeled by $\sigma^2$). For example, say we believe our forward
+%% model does not capture some small scale phenomenon.  Then we may
+%% express this belief by saying $\tru = \fwd + \err$, with $\fwd$
+%% depending on $\param$ and $\err$ depending on $\sspar$, with $\param
+%% \perp \sspar$. We do not know much about this effect but it is
+%% reasonable to assume it changes continuously in our domain. We (may
+%% choose to) model it as $\normal (0, \modcov)$ and take $\modcov$ to
+%% reflect the spatial (or other) variability we imagine $\err \sspar$
+%% has. Such small scale phenomenon can arise as a modeling issue, where
+%% we might not model the system in its entirety. It can also arise from
+%% a numerical source, where our discretization of the system is not fine
+%% enough to capture all small scale phenomena.
 %%  In section \ref{section:non vanishing}, we will see that assuming
@@
 
+%% \subsubsection{Dynamics and Observation Operators}\label{subsec:dynamics}
+%% Rows of $\obs$ are linear functionals and we refer to these as {\it
+%%   observation vector}s. Note that we cannot choose every observation
+%% vector we want. For example, we may be restricted only to pointwise
+%% evaluations of $\fwd \param$ by the sensors at our disposal, in which
+%% case measuring, e.g., the mean $\ell(u) = \int_{\Omega}u$ of a
+%% function is not possible. Thus, it is best to think of $(\obs \fwd
+%% \param)_j$ as approximations to pointwise evaluations. For example, if
+%% $\hilo = L^2([-\pi, \pi])$, consider $\ell_1(x) := (2\pi
+%% \eps)^{-\frac{1}{2}}\exp(-\frac{x^2}{2\eps^2})$ and $\ell_2(x)
+%% :=\eps^{-1}\mathbf{1}_{[-\frac{\eps}{2}, \frac{\eps}{2}]}$. Both
+%% $\ell_1$ and $\ell_2$ approximate a point measurement at $0$ for small
+%% $\eps$ and, crucially, $\ell_1, \ell_2 \in \hilo^*$.
+%% %% It is important to note that although $\|\ell\|_1 =
+%% %% \|\ell_2\|_1=1$, unfortunately $\|\ell_1\|_2^2 = (4\pi
+%% %% \eps)^{-\frac{1}{2}}$ and $\|\ell_2\|_2^2 = \eps^{-1}$.
+
 Finally, it is useful to record a known result regarding the posterior
-covariance operator of our inverse problem:
+covariance operator:
 \begin{align}\label{eq:postcov}
@@
 
-The meaning of D-optimal design in infinite-dimensional Hilbert spaces
-was investigated in \cite{AlexanderianGloorGhattas14}. There, the
-authors make assumptions that amount to $\Sigma=I$ (implied by
-$\modcov = 0,\sigma^2=1$), but we choose not to take these
-simplification here. This is because $\Sigma$ can determine ``how
-much'' clusterization we see. As stated
-\cite[pp. 681]{AlexanderianGloorGhattas14}, the results hold for more
-general covariance matrices. The conclusion is that in
+D-optimal designs in infinite-dimensional Hilbert spaces were
+investigated in \cite{AlexanderianGloorGhattas14}. In
 infinite-dimensions, a D-optimal design is well-defined as maximizing
 the expected KL divergence between posterior and prior. The main
-result is summarized, using our notation, in the following theorem:
+result is summarized, in the notation presented above, in the
+following theorem\footnote{In \cite{AlexanderianGloorGhattas14},
+results are stated for $\Sigma=I$ (implied by $\modcov =
+0,\sigma^2=1$), but the authors also mention these results hold for
+more general covariance matrices (p. 681).}:
 \begin{theorem}[Slightly modified Theorem 1 from \cite{AlexanderianGloorGhattas14}]\label{thm:d optimality}
@@
 \begin{definition}\label{def:d optimality}
-  A design $\obs^{\star}$ is said to be D-optimal if $\obs^{\star} = \argmax_{\obs} \tar(\obs)$.
+  $\obs^{\star}$ is said to be \emph{D-optimal} if $\obs^{\star} =
+  \argmax_{\obs} \tar(\obs)$, where rows of $\obs$ are in some allowed
+  set of measurement.
 \end{definition}
 
-\noindent The intuition behind Definition \ref{def:d optimality} is
-straightforward. From \eqref{eq:postcov} and \eqref{eq:objective}:
+\noindent In order to undestand the intuition behind Definition
+\ref{def:d optimality}, note that from \eqref{eq:postcov} and
+\eqref{eq:objective}:
 \begin{align*}
   \begin{split}
-    \det ( I + \prcov^{1/2}  \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}) &= \det \Big( \prcov ( \prcov^{-1} + \fwd ^* \obs^* \Sigma^{-1} \obs \fwd) \Big )\\
+    \tar(\obs) &= \det ( I + \prcov^{1/2}  \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}) \\
+    &= \det \Big( \prcov ( \prcov^{-1} + \fwd ^* \obs^* \Sigma^{-1} \obs \fwd) \Big )\\
     &= \det \prcov \det \postcov^{-1}.

### 2023-04-13 latex/vanishing.tex
Need to write intro. Tell how everybody thinks model error is the solution

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SECTION Analysis of Optimal Designs --- Vanishing Model Error
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{D-Optimal Designs with Vanishing Model Error}\label{section:vanishing}
-We now turn to analyze D-optimal designs when $\modcov = 0$. Our goal
-is to prove Theorem \ref{thm:char} --- a characterization of D-optimal
-designs.
+\section{D-Optimal Designs when $\modcov = 0$}\label{section:vanishing}
+We now turn to characterize D-optimal designs when $\modcov = 0$. Our
+goal is to prove Theorem \ref{thm:char}.
 
-The necessary first-order condition for D-optimality summarized in
-Theorem \ref{thm:constrained}, when $\Sigma= \sigma^2I$ (i.e. when
-$\modcov = 0$) becomes
+The necessary first-order condition for D-optimality of Theorem
+\ref{thm:constrained} for $\modcov = 0$ become:
+
 \begin{equation}\label{eq:eigenproblem}
@@
 \end{equation}
-with $\Xi$ diagonal. We think of \eqref{eq:eigenproblem} as an
+with $\Xi$ diagonal. Equation \eqref{eq:eigenproblem} looks like an
 eigenvalue problem for the self-adjoint operator $\sigma^{-2}\fwd
@@
   Let $\obs$ satisfy the nonlinear eigenvalue problem
-  \eqref{eq:eigenproblem}. Then $\obs^*\obs$ has the same eigenvectors
-  as $\fwd \prcov \fwd^*$.
+  \eqref{eq:eigenproblem}. Then $\obs^*\obs$ and $\fwd \prcov \fwd^*$
+  are simultaneously diagonalizable.
 \end{proposition}
@@
 
-We are almost ready to prove Theorem \ref{thm:char} that characterizes
-D-optimal designs. We just need one final lemma, whose proof is
-delegated to the appendix:
+Before we prove Theorem \ref{thm:char} we record Lemma
+\ref{lemma:free}, whose proof is delegated to the appendix:
 \begin{restatable*}[Unit norm decomposition]{lemma}{free}\label{lemma:free}
@@
     \sigma^{-2}\obs^*\obs$ for a D-optimal design (left) and a
-    sub-optimal design (right). Both designs are allowed three
+    sub-optimal design (right). Both designs are allowed $m=3$
     observations. We assume $\sigma^2=1$ and thus, the blue area has
-    accumulated height of $3$ in both plots. The D-optimal design
-    (left) increases precision (equivalently, reduces uncertainty)
-    where it is lowest. The sub-optimal design (right) does not.
-    %% In this example, an observation is taken partly in the
-    %% direction of the third eigenvector. The precision for this
-    %% eigenvector is bigger even than the posterior precision of the
-    %% D-optimal design on the left.
-  }
+    accumulated height of $\sigma^{-2}m = 3$ in both plots. The
+    D-optimal design (left) increases precision (equivalently, reduces
+    uncertainty) where it is lowest. The sub-optimal design (right)
+    does not.}
   \label{fig:optimal vs not}
@@
 
-Here is a way to understand D-optimal designs with vanishing model
-error: Imagine each eigenvector of $\fwd \prcov \fwd^*$ corresponds to
-a graduated lab cylinder (Figure \ref{fig:cylinder}). Each cylinder
-$i$ is filled, a-priori, with green liquid of $\lambda_i^{-1}$ volume
-units. This liquid represnets the prior precision, and the volume of
-the green liquid is increasing in $i$. A D-optimal design with $m$
-observations ignores cylinders indexed $m+1$ and above. It has blue
-liquid of volume $\sigma^{-2}m$ units at its disposal. It distributes
-said liquid by repeatedly adding a drop to whatever cylinder currently
-has the lowest level of liquid in it (lowest precision). The result of
-such a procedure is illustrated in Figure \ref{fig:optimal vs not}.
+Theorem \ref{thm:char} gives us a way to understand D-optimal designs
+when $\modcov = 0$: Imagine each eigenvector of $\fwd \prcov \fwd^*$
+corresponds to a graduated lab cylinder (Figure
+\ref{fig:cylinder}). Each cylinder $i$ is filled, a-priori, with green
+liquid of $\lambda_i^{-1}$ volume units. We are allowed $m$
+observations, so we have blue liquid of volume $\sigma^{-2}m$ units at
+our disposal. When we seek a D-optimal design, we distribute the blue
+liquid by repeatedly adding a drop to whatever cylinder currently has
+the lowest level of liquid in it, as long as its index $i \leq m$. The
+result of such a procedure is illustrated in Figure \ref{fig:optimal
+  vs not}.
 
@@
 
-\subsection{Clusterization poses no obstruction to D-optimality}\label{subsec:clusterization}
-We saw that the necessary conditions for a design $\obs$ with $m$
-observations to be D-optimal are (1), (2) and (4) of Theorem
-\ref{thm:char}. As long as these requirements are satisfied, $\obs$ is
-D-optimal. These requirements, however, still leave some freedom in
-choosing $\obs$. This freedom may give rise to clusterization, as
-illustrated in Figure \ref{fig:clusterization} and explained below.
+\section{Clusterization poses no obstruction to D-optimality}\label{section:clusterization}
+The necessary conditions for D-optimality are (1), (2) and (4) of
+Theorem \ref{thm:char}. As long as these conditions are satisfied,
+$\obs$ is an optimizer of $\tar$. These conditions, however, still
+leave some freedom in choosing $\obs$. This freedom may give rise to
+clusterization, as illustrated in Figure \ref{fig:clusterization} and
+demonstrated below.
 
-Consider an inverse problem with $\lambda_1^{-1} = 0.1, \lambda_2^{-1}
-= 0.2$ and $\sigma^2=1$. In \eqref{eq:two designs}, we suggest two
-D-optimal designs $\obs_1, \obs_2$ such that $\obs_1^*\obs_1 =
-\obs_2^*\obs_2$. The former exhibits clusterization ($\meas_3 =
-\meas_5 = \ev_1$), while the latter does not. Both of these designs
-are D-optimal, hence clusterization poses no obstruction to
-D-optimality.
+Consider an inverse problem for which $\lambda_1^{-1} = 0.1,
+\lambda_2^{-1} = 0.2,\sigma^2=1$ and $\{\ev_i\}_{i=1}^\infty$ are
+eigenvectors of $\fwd\prcov\fwd^*$. Two D-optimal designs $\obs_1,
+\obs_2$ presented in \eqref{eq:two designs}. It is easy to verify that
+$\obs_1^*\obs_1 = \obs_2^*\obs_2$. The former exhibits clusterization
+($\meas_3 = \meas_5 = \ev_1$), while the latter does not. Since both
+of these designs are D-optimal, we conclude that clusterization poses
+no obstruction to D-optimality.
 
@@
 
-A computer implementation of Lemma \ref{lemma:free} (see code in
+A computer implementation of Lemma \ref{lemma:free} (code in
 Supplementary material) generates $\obs_1$ as a solution to the
@@
 paragraph. Modifying the numerical values of the prior does not break
-the clusterization. Thus, the preference to clusterization of Lemma
-\ref{lemma:free} is robust. The vast majority of randomized numerical
-simulations (see code in Supplementary material) also give rise to
-D-optimal designs that exhibit clustrization. We conclude that
-clusterization poses no obstruction to D-optimality in the relaxed
-model presented in this study. Moreover, sensor clusterization is
-generic, as far as one accepts the model presented in this study.
+the clusterization. Thus, the preference to clusterization, at least
+in Lemma \ref{lemma:free} is robust. The vast majority of randomized
+numerical simulations (see code in Supplementary material) also give
+rise to D-optimal designs that exhibit clustrization. We conclude that
+clusterization poses no obstruction to D-optimality in the model
+presented in this study. Moreover, sensor clusterization is generic,
+as far as the model presented in this study is concerned.
 

### 2023-04-14 latex/vanishing.tex
Moved proofs to main text

@@
 refer to \eqref{eq:eigenproblem} as a \emph{nonlinear} eigenvalue
-problem. The following Proposition and Lemma are proved in the
-appendix:
+problem.
 
-\begin{restatable*}[Two applications of Woodbury Matrix Identity]{proposition}{woodbury}\label{prop:twice woodbury}
+\begin{proposition}\label{prop:twice woodbury}
   Assume $\fwd \prcov \fwd^*$ is invertible. Then
@@
   \end{align*}  
-\end{restatable*}
+\end{proposition}
 
-\begin{restatable*}[Simultaneous diagonizability]{lemma}{simdiag}\label{lemma:sim diag}
+\begin{proof}
+  The proof amounts to using Woodbury's matrix identity twice and a
+  regularization trick. The standard proof for Woodbury's matrix
+  identity works in infinite dimensions, as long as all terms are well
+  defined. Unfortunately, $\obs^*\obs$ is not invertible, so we force
+  it to be. Recall Woodbury's matrix identity:
+  $$
+  (A + UCV)^{-1} = A^{-1} - A^{-1}U(C^{-1} + VA^{-1}U)^{-1}VA^{-1}.
+  $$
+
+  Denote $A := \prcov^{-1}$, $U := \fwd^*$, $V := \fwd$, $C :=
+  \sigma^{-2} (\obs^*\obs+\eps I)$ for some $\eps > 0$. Then:
+  %% \begin{align*}
+  %%   \begin{split}
+  \begin{equation}\label{eq:first}
+  \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* = \fwd ( \prcov - \prcov \fwd^* ( \sigma^2(\obs^*\obs + \eps I)^{-1} + \fwd \prcov \fwd^* )^{-1} \fwd \prcov ) \fwd^* \\
+      %
+      %
+      %% &= X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X
+  %%   \end{split}
+  %% \end{align*}
+  \end{equation}
+   Now denote $X := \fwd\prcov \fwd^*$ and $Y_{\eps} := \obs^*\obs +
+   \eps I$, and $Y_{\eps}$ is invertible. Then \eqref{eq:first}
+   becomes:
+   \begin{equation}\label{eq:second}
+     \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* = X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X
+   \end{equation}
+
+   Note that $X + \sigma^2 Y_{\eps}^{-1}$ is invertible, as the sum of
+   two positive definite operators. Now, taking $A := X^{-1}, C :=
+   \sigma^2Y_{\eps}^{-1}, U := I$ and $V := I$ and using Wodbury's
+   matrix identity in reverse order:
+  \begin{align*}
+    \begin{split}
+      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* &= X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X \\
+      %
+      %
+      %
+      &= (X^{-1} + \sigma^{-2}Y_{\eps})^{-1} \\
+      %
+      %
+      %
+      &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \eps I))^{-1}.
+    \end{split}
+  \end{align*}
+
+  We conclude that $\forall \eps > 0$
+  \begin{align*}
+    \begin{split}
+      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* 
+     &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \eps I))^{-1}.
+    \end{split}
+  \end{align*}
+  Letting $\eps \to 0$ completes the proof.
+\end{proof}
+
+\begin{lemma}[Simultaneous diagonizability]\label{lemma:sim diag}
   Let $\hil$ separable Hilbert space, $C:\hil \to \hil$ self-adjoint
@@
   diagonalizable.
-\end{restatable*}
+\end{lemma}
+
+\begin{proof}
+  First, enumerate the eigenvalues of $C + \sum_{j=1}^m
+  \func_j\func_j^*$ as $\xi_1,\dots,\xi_\ell$. Denote the
+  indices of the eigenvectors corresponding to $\xi_i$
+  \begin{equation*}
+    S_i := \{ 1 \leq k \leq m | (C + \sum_{j=1}^m \func_j\func_j^* )\func_k = \xi_i \func_k \}.
+  \end{equation*}
+  Define further
+  \begin{equation*}
+    A_i := \sum_{k \in S_i} \func_k \func_k^*,
+  \end{equation*}
+  which is self-adjoint. Two observations are in order. First,
+  $\sum_{j=1}^m \func_j\func_j^* = \sum_{i=1}^\ell A_i$. Second, $A_i
+  \func_k = 0$ if $k\not \in S_i$, since eigenvectors of different
+  eigenvalue are orthogonal. For $k \in S_i$
+  \begin{equation}\label{eq:on vi}
+    \xi_i \func_k = (C + \sum_{j=1}^m \func_j \func_j^* ) \func_k = (C + A_i) \func_k.
+  \end{equation}
+  Let $V_i := span \{\func_k \}_{k\in S_i}$. Observe that $V_i$ is
+  invariant under $A_i$, by definition, and under $C$, by \eqref{eq:on
+    vi}. Using \eqref{eq:on vi} again, we conclude that on $V_i$, $A_i
+  = \xi_iI - C$. This immediately implies $A_i$ and $C$ have the
+  same eigenvectors on $V_i$. This holds for every $1 \leq i \leq
+  \ell$ and we conclude that $C$ and $A$ have the same eigenvectors.
+\end{proof}
 
@@
 
-Before we prove Theorem \ref{thm:char} we record Lemma
-\ref{lemma:free}, whose proof is delegated to the appendix:
-\begin{restatable*}[Unit norm decomposition]{lemma}{free}\label{lemma:free}
+For the proof of Theorem \ref{thm:char} we require Lemma
+\ref{lemma:free}:
+
+\begin{lemma}[Unit norm decomposition]\label{lemma:free}
   Let $M \in \R^{k \times k}$ symmetric positive definite with $\ttr M
@@
   $AA^t = M$.
-\end{restatable*}
+\end{lemma}
+
+\begin{proof}
+  Let us diagonalize $M$, so that $M = U D U^t$ with $D =
+  \diag(d_1,\dots,d_k)$ and $U \in \R^{k \times k }$ orthogonal. Let
+  $S \in \R^{k \times m}$ with $S_{ii} = \sqrt{d_{i}}$ and zeros
+  otherwise. Define $A:= U S V^t$, where $V \in \R^{m \times m}$ is
+  orthogonal and will be further restricted later. Then $AA^t = U
+  SV^tVS^t U^t = UDU^t$, so $AA^t$ has the required eigenvalues and
+  eigenvectors by construction. If we can choose $V$ such that $A$
+  also satisfies the unit norm constraints we are done. These
+  constraints are, for $j=1,\dots,m$:
+  \begin{equation}\label{eq:V constraints}
+   1 = [A^tA]_{jj} = [V S^tS V^t]_{jj},
+  \end{equation}
+  and we can expect to do this since we assumed $\ttr D = m$.
+
+  Define $C = S^tS - I \in \R^{m \times m}$. Note that $\ttr C = 0$ and
+  $C$ is diagonal with non-zero entries $d_i-1,i=1,\dots,k$. It suffices
+  to find $V$ orthogonal such that $V C V^t$ has zero diagonal. We
+  construct such $V$ by sequentially inserting zeros in the diagonal
+  and not destroying zeros we already introduced, starting from the
+  last diagonal entry and moving to the first. Since $c_{mm} \neq 0$ ,
+  let $p < m$ such that $c_{pp}c_{mm} < 0$ (such $p$ exists because
+  the trace is zero) and let $\theta \in (0,\pi)$. Define a Givens
+  rotation $R^{(m)} \in \R^{m \times m}$ by
+  \begin{equation*}
+    r^{(m)}_{ab} :=
+    \begin{cases}
+      1 & a = b \neq p \text{ or } a = b \neq m \\
+      \cos \theta & a = b = p  \\
+     -\sin \theta & a = p, b = m\\
+      \cos \theta & a = b = m \\
+      \sin \theta & a = m, b = p \\ 
+      0 & o.w
+    \end{cases}
+  \end{equation*}
+  Note that conjugating a matrix by $R^{(m)}$ changes only its $m$ and
+  $p$ rows and columns. We want to choose $\theta$ such that
+  \begin{equation}\label{eq:mm}
+    0 = [R^{(m)} C (R^{(m)})^t]_{mm} = \cos^2 \theta c_{mm} + 2\cos \theta \sin
+    \theta c_{mp} + \sin^2\theta c_{pp},
+  \end{equation}
+  and it suffices to choose $\theta$ such that
+  \begin{equation*}
+    c_{mm} \cot^2 \theta + 2 c_{mp} \cot \theta + c_{pp} = 0.
+  \end{equation*}
+  This quadratic in $\cot\theta$ has a real solution, since
+  $c_{pp}c_{mm} < 0$ by assumption and we can find $\theta \in
+  (0,\pi)$ such that \eqref{eq:mm} is satisfied. We continue to find
+  $R^{(m-1)}$ that leaves row and column $m$ unchanged and
+  continue introducing zeros to the diagonal. The assumption $\ttr D =
+  m \Rightarrow \ttr C = 0$ guarantees we can do that. Taking $V:=
+  R^{(1)} R^{(2)} \dots R^{(m-1)}R^{(m)}$ completes the proof.
+\end{proof}
 

### 2023-04-15 latex/vanishing.tex
read proofs and stuff again. Now only need to fix introgit add latex/vanishing.tex !

@@
 \section{D-Optimal Designs when $\modcov = 0$}\label{section:vanishing}
-We now turn to characterize D-optimal designs when $\modcov = 0$. Our
-goal is to prove Theorem \ref{thm:char}.
-
-The necessary first-order condition for D-optimality of Theorem
-\ref{thm:constrained} for $\modcov = 0$ become:
+We will now prove Theorem \ref{thm:char} which caracterizes D-optimal
+designs when $\modcov = 0$. The necessary first-order condition for
+D-optimality of Theorem \ref{thm:constrained} for $\modcov = 0$
+become:
 
@@
 refer to \eqref{eq:eigenproblem} as a \emph{nonlinear} eigenvalue
-problem.
+problem.%%  Proposition \ref{prop:twice woodbury} and Lemma
+%% \ref{lemma:sim diag} required
 
@@
   defined. Unfortunately, $\obs^*\obs$ is not invertible, so we force
-  it to be. Recall Woodbury's matrix identity:
-  $$
-  (A + UCV)^{-1} = A^{-1} - A^{-1}U(C^{-1} + VA^{-1}U)^{-1}VA^{-1}.
-  $$
+  it to be. Recall Woodbury's matrix identity and denote:
+  \begin{align*}
+    (A + UCV)^{-1} &= A^{-1} - A^{-1}U(C^{-1} + VA^{-1}U)^{-1}VA^{-1} \\
+    A :&= \prcov^{-1} \\
+    U :&= \fwd^* \\
+    V :&= \fwd \\
+    C :&= \sigma^{-2} (\obs^*\obs+\eps I),\ \eps > 0
+  \end{align*}
 
-  Denote $A := \prcov^{-1}$, $U := \fwd^*$, $V := \fwd$, $C :=
-  \sigma^{-2} (\obs^*\obs+\eps I)$ for some $\eps > 0$. Then:
-  %% \begin{align*}
-  %%   \begin{split}
+  Then, by Woodbury's matrix identity:
   \begin{equation}\label{eq:first}
@@
   \end{equation}
-   Now denote $X := \fwd\prcov \fwd^*$ and $Y_{\eps} := \obs^*\obs +
-   \eps I$, and $Y_{\eps}$ is invertible. Then \eqref{eq:first}
-   becomes:
+  Now, denote:
+  \begin{align*}
+    X :&= \fwd\prcov \fwd^* \\
+    Y_{\eps} :&= \obs^*\obs + \eps I,
+  \end{align*}
+  and observe that $Y_{\eps}$ is invertible. Subsitute $X, Y_{\eps}$
+  into the RHS of \eqref{eq:first}:
    \begin{equation}\label{eq:second}
@@
    Note that $X + \sigma^2 Y_{\eps}^{-1}$ is invertible, as the sum of
-   two positive definite operators. Now, taking $A := X^{-1}, C :=
-   \sigma^2Y_{\eps}^{-1}, U := I$ and $V := I$ and using Wodbury's
-   matrix identity in reverse order:
+   two positive definite operators. Now, let
+   \begin{align*}
+     A :&= X^{-1}, \\
+     C :&= \sigma^2Y_{\eps}^{-1}, \\
+     U :&= I, \\
+     V :&= I.
+   \end{align*}
+   Apply Woodbury's matrix identity to the RHS of \eqref{eq:second}:
   \begin{align*}
@@
   First, enumerate the eigenvalues of $C + \sum_{j=1}^m
-  \func_j\func_j^*$ as $\xi_1,\dots,\xi_\ell$. Denote the
-  indices of the eigenvectors corresponding to $\xi_i$
+  \func_j\func_j^*$ as $\xi_1,\dots,\xi_\ell$. Denote the indices of
+  the eigenvectors corresponding to $\xi_i$
   \begin{equation*}
@@
   invariant under $A_i$, by definition, and under $C$, by \eqref{eq:on
-    vi}. Using \eqref{eq:on vi} again, we conclude that on $V_i$, $A_i
-  = \xi_iI - C$. This immediately implies $A_i$ and $C$ have the
-  same eigenvectors on $V_i$. This holds for every $1 \leq i \leq
-  \ell$ and we conclude that $C$ and $A$ have the same eigenvectors.
+    vi}. A second application of \eqref{eq:on vi} shows that $A_i =
+  \xi_iI - C$ on $V_i$. This immediately implies $A_i$ and $C$ are
+  simultaneously diagonalizable on $V_i$. This holds for every $1 \leq
+  i \leq \ell$ and we conclude that $C$ and $A$ are simultaneously
+  diagonalizable.
 \end{proof}
@@
 \begin{proposition}\label{prop:kkt}
-  Let $\tar: \mathbb{R}^m \to \mathbb{R}$, $\psi(\eta) =
+  Let $\tar: \mathbb{R}^m \to \mathbb{R}$, $\tar(\eta) =
   \frac{1}{2}\sum_{i=1}^m \log (1+\sigma^{-2}\lambda_i \eta_i)$, with
@@
 
-For the proof of Theorem \ref{thm:char} we require Lemma
-\ref{lemma:free}:
+We require Lemma \ref{lemma:free} for the proof of Theorem
+\ref{thm:char}:
 
@@
   $\postcov^{-1} = \prcov^{-1} + \sigma^{-2}\fwd^*\obs^*\obs\fwd$. The
-  first statement in part (5) now follows from Lemma \ref{prop:twice
-    woodbury}, while the second statement follows from parts (1) and
-  (4).
+  first statement in part (5) now follows from Proposition
+  \ref{prop:twice woodbury}, while the second statement follows from
+  parts (1) and (4).
 \end{proof}
@@
     accumulated height of $\sigma^{-2}m = 3$ in both plots. The
-    D-optimal design (left) increases precision (equivalently, reduces
-    uncertainty) where it is lowest. The sub-optimal design (right)
-    does not.}
+    D-optimal design (left) increases precision where it is
+    lowest. The sub-optimal design (right) does not.}
   \label{fig:optimal vs not}

### 2023-04-17 latex/intro.tex
get ready to create different branches for different journals

@@
 
-In the process of proving Theorem \ref{thm:char}, we prove a novel
-lemma in linear algebra: We show how to decompose a symmetric positive
+One byproduct of the proof of Theorem \ref{thm:char} is a novel lemma
+in linear algebra: We show how to decompose a symmetric positive
 definite matrix $M$ as $M = AA^t$, where $A$ has unit norm columns
@@
     numbers of sensors. Sensor locations (circles) were chosen to
-    optimize the D-optimality criterion. Sensor clusterization can be
-    clearly seen for six sensors. Only four observation locations are
-    used --- two pairs of sensors are so close together they are
+    optimize the D-optimality criterion. Sensor clusterization occurs
+    for six sensors. Only four observation locations are used --- two
+    pairs of sensors are so close together they are
     indistinguishable.}
@@
 four observations were taken. Two pairs of sensors are clustered so
-close together that they are indistinguishable from one another.
+close together that they are indistinguishable from each other.
 

### 2023-04-17 latex/paper.tex
get ready to create different branches for different journals

@@
   experimental design for infinite-dimensional Bayesian linear inverse
-  problems. We suggest an analytically tractable model for D-optimal
-  designs. We show that sensor clusterization in the case of iid
-  observation noise is generic. We also show that model spatially
-  correlated errors mitigate the clusterization phenomenon.
+  problems. Many authors seem to believe that the fact that spatial
+  correlations mitigateWe suggest an analytically tractable model for D-optimal
+  designs. We argue that sensor clusterization in the case of iid
+  observation noise is generic. We also prove that a model error term
+  mitigates the clusterization phenomenon.
 
@@
 \input{vanishing}
+
+%% \section{Open Problems}
+%% Some open problems to
+
+%% We two phenomena for a generic inverse problem with unit-norm
+%% constraints on obsevations: First, we showed that when model error is
+%% present and obsevation noise is small --- sensor clusterization will
+%% not occur. In contrast, we saw that when no model error is present ---
+%% sensor clusterization is quite generic and poses no obstruction to
+%% D-optimality.
 

### 2023-04-17 latex/prelims.tex
get ready to create different branches for different journals

@@
 \ref{subsec:notation}. The theoretical foundations for inverse
-problems over function spaces can be found in \cite{Stuart10}.
+problems over function spaces can be found in
+\cite{Stuart10}.
 
@@
 $\fwd \prcov \fwd^*$ is the prior covariance in $\hilo$
-\cite{Stuart10}. As such, it is assumed invertible --- an assumption
-which will be used heavily below. Let $m$ the number of observations
-taken. Observations are taken via the \emph{observation
+\cite{Stuart10}. As such, $\fwd \prcov \fwd^*$ is assumed invertible
+--- an assumption which will be used below. Let $m$ the number of
+observations taken. Observations are taken via the \emph{observation
 operator}\footnote{It is common for the observation and forward
@@
 following sections requires $\fwd$ and $\obs$ are explicitly
-separated.} $\obs \in ( \hilo^* )^m$. In an analogy with linear
-algebra, where row vectors are thought of as linear functionals,
-$\obs$ is best thought of as having ``rows'' $\meas_j, j=1,\dots,m$:
-\begin{equation}\label{eq:O}
-  \obs = (\meas_1,\dots, \meas_m)^t, \meas_j \in \hilo^*, j = 1,\dots,m.
-\end{equation}
-This way, $\obs u = (\meas_1(u), \dots, \meas_m(u) )^t \in \R^m$ and
-$u \in \hilo$.
-%% but we may drop the parentheses and write $\meas_j u = \meas_j(u)$.
-%% A few observations regarding $\obs$ are in order.
-%% First, it is good to keep in mind that $( \hilo^* )^m$ is a Hilbert
-%% space with norm $\| \obs \| = \sum_{j=1}^m \|\meas_j\|$.
-For $u \in \hilo$ and $v\in \R^m$:
+separated.} $\obs \in ( \hilo^* )^m$, so that:
+
+\begin{equation*}%\label{eq:O}
+  \obs u = (\meas_1(u), \dots, \meas_m(u) )^t \in \R^m,\ u \in \hilo,
+\end{equation*}
+%
+%
+and we refer to $\meas_j, j=1,\dots,m$ as \emph{observation
+vector}s. For $u \in \hilo$ and $v\in \R^m$:
+%
+%
 \begin{align*}
@@
 
-%% \begin{observation}
-%%   It is best to think of $\meas_j,j=1,\dots,m$ as row vectors and of
-%%   $\obs$ as a matrix with rows $\meas_j$. 
-%% \end{observation}
-
-Each observation vector is a linear functional, chosen from (some
-subset of) $\hilo^{*}$. Data is acquired via noisy observations, with
-two types of error terms considered. Spatially correlated model error
-is denoted $\eps' \sim \normal(0,\modcov)$, and modeled as a centered
-Gaussian measure on $\hilo$ with covariance operator
-$\modcov$. Observation error is denoted $\eps \sim \normal(0, \sigma^2
-I_m)$, with $I_m \in \mathbb{R}^{m \times m}$ the identity. Both error
-terms and the prior are assumed independent of each other. Thus, data
-is acquired via
+Data is acquired via noisy observations, and we consider two types of
+error terms: Spatially correlated model error is denoted $\eps' \sim
+\normal(0,\modcov)$, and modeled as a centered Gaussian measure on
+$\hilo$ with covariance operator $\modcov$. Observation error is
+denoted $\eps \sim \normal(0, \sigma^2 I_m)$, with $I_m \in
+\mathbb{R}^{m \times m}$ the identity. Both error terms and the prior
+are assumed independent of each other. Thus, data is acquired via
 \begin{align}\label{eq:inverse problem}
@@
     % 
-    = \obs \modcov \obs^* + \sigma^2I , 
+    = \obs \modcov \obs^* + \sigma^2I_m , 
   \end{split}
@@
 Taking $\modcov = 0$ is a common practice
-\cite{Tarantola05,Kaipio2006,Vogel02} and then $\Sigma = \sigma^2I$ is
-simply a scalar matrix that does not depend on $\obs$.
+\cite{Tarantola05,Kaipio2006,Vogel02} and then $\Sigma = \sigma^2I_m$
+is a scalar matrix and does not depend on $\obs$.
 
@@
 
-Finally, it is useful to record a known result regarding the posterior
-covariance operator:
+The posterior measure $\post$ is Gaussian in this setting and its
+covariance operator is:
 \begin{align}\label{eq:postcov}
@@
 A D-optimal design maximizes expected KL divergence between posterior
-and prior measures, denoted $\post$ and $\pr$, respecively. It is
-useful to recall the definition of KL divergence for an arbitrary
-prior measure:
+and prior measures, $\post$ and $\pr$. It is useful to first recall
+the definition of KL divergence:% for an arbitrary prior measure:
 $$
-KL(\post||\pr) = \int \log \frac{\der \post}{\der \pr}(\param) \der \post(\param).
+D_{KL}(\post||\pr) = \int \log \frac{\der \post}{\der \pr}(\param) \der \post(\param).
 $$
@@
 D-optimal designs in infinite-dimensional Hilbert spaces were
-investigated in \cite{AlexanderianGloorGhattas14}. In
-infinite-dimensions, a D-optimal design is well-defined as maximizing
-the expected KL divergence between posterior and prior. The main
-result is summarized, in the notation presented above, in the
-following theorem\footnote{In \cite{AlexanderianGloorGhattas14},
-results are stated for $\Sigma=I$ (implied by $\modcov =
-0,\sigma^2=1$), but the authors also mention these results hold for
-more general covariance matrices (p. 681).}:
+investigated in \cite{AlexanderianGloorGhattas14}. Their main result
+is summarized, in the notation presented above, in the following
+theorem\footnote{In \cite{AlexanderianGloorGhattas14}, results are
+stated for $\Sigma=I$ (implied by $\modcov = 0,\sigma^2=1$), but the
+authors also mention these results hold for more general covariance
+matrices (p. 681).}:
 \begin{theorem}[Slightly modified Theorem 1 from \cite{AlexanderianGloorGhattas14}]\label{thm:d optimality}
@@
   $\obs^{\star}$ is said to be \emph{D-optimal} if $\obs^{\star} =
-  \argmax_{\obs} \tar(\obs)$, where rows of $\obs$ are in some allowed
-  set of measurement.
+  \argmax_{\obs} \tar(\obs)$, where entries of $\obs \in (\hilo^*)^m$ are constrained to
+  some allowed set of measurement vectors in $\hilo^*$.
 \end{definition}
 
-\noindent In order to undestand the intuition behind Definition
-\ref{def:d optimality}, note that from \eqref{eq:postcov} and
+\noindent For an intuitive understanding of Definition \ref{def:d
+  optimality}, note that from \eqref{eq:postcov} and
 \eqref{eq:objective}:

### 2023-04-17 latex/vanishing.tex
get ready to create different branches for different journals

@@
   regularization trick. The standard proof for Woodbury's matrix
-  identity works in infinite dimensions, as long as all terms are well
-  defined. Unfortunately, $\obs^*\obs$ is not invertible, so we force
-  it to be. Recall Woodbury's matrix identity and denote:
-  \begin{align*}
-    (A + UCV)^{-1} &= A^{-1} - A^{-1}U(C^{-1} + VA^{-1}U)^{-1}VA^{-1} \\
+  identity works for separable Hilbert spaces, as long as all terms
+  are well defined. Unfortunately, $\obs^*\obs$ is not invertible, so
+  we force it to be. Recall Woodbury's matrix identity:
+  \begin{equation}\label{eq:WMI}
+    (A + UCV)^{-1} = A^{-1} - A^{-1}U(C^{-1} + VA^{-1}U)^{-1}VA^{-1}. 
+  \end{equation}
+
+  Denote:
+  \begin{align}\label{eq:notation}
+    \begin{split}
       A :&= \prcov^{-1} \\
@@
       C :&= \sigma^{-2} (\obs^*\obs+\eps I),\ \eps > 0
-  \end{align*}
+    \end{split}
+  \end{align}
 
-  Then, by Woodbury's matrix identity:
+  Then, by \eqref{eq:WMI}, with the notation \eqref{eq:notation}:
   \begin{equation}\label{eq:first}
@@
    two positive definite operators. Now, let
-   \begin{align*}
+   \begin{align}\label{eq:notation2}
+     \begin{split}
        A :&= X^{-1}, \\
@@
        V :&= I.
-   \end{align*}
-   Apply Woodbury's matrix identity to the RHS of \eqref{eq:second}:
+     \end{split}
+   \end{align}
+   Apply \eqref{eq:WMI} with the notation \eqref{eq:notation2} to the
+   RHS of \eqref{eq:second}:
   \begin{align*}
@@
 \end{lemma}
-
 \begin{proof}
@@
 
-\begin{proposition}[Eigenvectors of a D-optimal design]\label{prop:same ev}
+\begin{proposition}\label{prop:same ev}
   Let $\obs$ satisfy the nonlinear eigenvalue problem
@@
       \end{equation*}
-    \item Let $\obs$ a D-optimal design. Then $\eta_i > 0$ for
+    \item Furthermore, if $\obs$ isD-optimal, $\eta_i > 0$ for
       eigenvectors corresponding to the $k$ largest $\lambda_i$.
@@
   where $A:= \{1\leq i \leq m: \eta_i > 0\}$ and $A^c = \{1,\dots, m\}
-    \backslash A$, and $|A|$ is the cardinality of $A$.
+  \backslash A$, and $k = |A|$, the cardinality of $A$.
 \end{proposition}
@@
 
-We require Lemma \ref{lemma:free} for the proof of Theorem
-\ref{thm:char}:
+The final ingredient we require for the proof of Theorem
+\ref{thm:char} the following lemma:
 
@@
 
-Theorem \ref{thm:char} gives us a way to understand D-optimal designs
+Theorem \ref{thm:char} facilitates understanding of D-optimal designs
 when $\modcov = 0$: Imagine each eigenvector of $\fwd \prcov \fwd^*$
@@
 The necessary conditions for D-optimality are (1), (2) and (4) of
-Theorem \ref{thm:char}. As long as these conditions are satisfied,
-$\obs$ is an optimizer of $\tar$. These conditions, however, still
-leave some freedom in choosing $\obs$. This freedom may give rise to
+Theorem \ref{thm:char}. These conditions, however, still leave some
+freedom in choosing $\obs$. This freedom may give rise to
 clusterization, as illustrated in Figure \ref{fig:clusterization} and
@@
 eigenvectors of $\fwd\prcov\fwd^*$. Two D-optimal designs $\obs_1,
-\obs_2$ presented in \eqref{eq:two designs}. It is easy to verify that
-$\obs_1^*\obs_1 = \obs_2^*\obs_2$. The former exhibits clusterization
-($\meas_3 = \meas_5 = \ev_1$), while the latter does not. Since both
-of these designs are D-optimal, we conclude that clusterization poses
-no obstruction to D-optimality.
+\obs_2$ are presented in \eqref{eq:two designs}, and it is easy to
+verify that $\obs_1^*\obs_1 = \obs_2^*\obs_2$. However, $\obs_1$
+exhibits clusterization ($\meas_3 = \meas_5 = \ev_1$), while $\obs_2$
+does not. Both of $\obs_1$ and $\obs_2$ are D-optimal, and we conclude
+that clusterization poses no obstruction to D-optimality.
 
@@
 
-A computer implementation of Lemma \ref{lemma:free} (code in
+A computer implementation of Lemma \ref{thm:char} (code in
 Supplementary material) generates $\obs_1$ as a solution to the
 D-optimal design problem mentioned in the previous
-paragraph. Modifying the numerical values of the prior does not break
-the clusterization. Thus, the preference to clusterization, at least
-in Lemma \ref{lemma:free} is robust. The vast majority of randomized
-numerical simulations (see code in Supplementary material) also give
-rise to D-optimal designs that exhibit clustrization. We conclude that
-clusterization poses no obstruction to D-optimality in the model
-presented in this study. Moreover, sensor clusterization is generic,
-as far as the model presented in this study is concerned.
-
-%% Of course, a real-life problem is more restricted than the model we
-%% consider here. Such restriction is the likely cause of clusterization
-%% obseved in, e.g., the toy model of section
-%% \ref{subsec:example}. Exactly how this happens is not clear. At this
-%% point it is the subject of speculation and should be further studied
-%% in the future.
+paragraph. Randomized numerical simulations (see code in Supplementary
+material) give rise to D-optimal designs that exhibit clustrization
+more than 95\% of the time. We conclude that clusterization poses no
+obstruction to D-optimality for the model presented in this study and
+that sensor clusterization is generic.
 

### 2023-04-19 latex/intro.tex
Writin

@@
 investigations. When considering an inverse problem, one can often
-specify sensor locations (e.g.\ in geophysics and oceanography
-applications), certain wavelengths (e.g.\ in MRI) or wave reflections
-from the ground (e.g.\ searching for oil or using a radar). Whatever
-the allowed set of observations is, one should select the optimal
-observations to take, in order to increase accuracy, reduce costs, or
-both.
+specify sensor locations (e.g.\ in electricat impedance tomography
+\cite{horesh2010impedance}), certain wavelengths (e.g.\ in MRI
+\cite{horesh2008mri}), or location of sensors for waves reflecting
+from the ground (e.g.\ searching for oil \cite{horesh2008borehole} or
+using a radar). Whatever the allowed set of measurements is, one
+should select the optimal observations to take, in order to increase
+accuracy, reduce costs, or both.
 
-The problem of choosing the best set of measurements is known as the
-problem of \emph{optimal experimental design}. Optimal designs are
-found by optimizing some \emph{design criterion}. Many design criteria
-exist, and in this study we focus on the Bayesian \emph{D-optimality
-criterion}. The D-optimality criterion has a simple and appealing
-motivation \cite{Chaloner1995}: a D-optimal design maximizes the
-expected information gain (KL divergence \cite{CoverThomas91}) between
-posterior and prior. For a linear model in finite dimensions, with
-Gaussian prior and iid Gaussian noise, a D-optimal design minimizes
-the determinant of the posterior covariance matrix\footnote{In a
-frequentist setting, a D-optimal design minimizes the determinant of
-the Fisher information matrix \cite[page 16]{Ucinski05}, so the
-Bayesian criterion is just a regularized version of the frequentist
-criterion \cite{Chaloner1995}.}.
+A set of measurements is called a \emph{design}, and an \emph{optimal
+design} is found by optimizing some \emph{design
+criterion}. Surprisingly, optimal designs sometimes result in
+measuremets that are very similar. For example consider measurements
+taken by sensors that evaluate some function at $\x \in \Omega
+\subseteq \R^d, d=1,2,3$. An optimal design may place sensors in very
+close proximity to each other (Figure \ref{fig:clusterization
+  illustration}). Following \cite{Ucinski05}, we refer to this
+phenomenon as \emph{sensor clusterization}.
+
+There is a consensus among researchers that sensor clusterization is
+undesirable. Some authors acknowledge the problem of sensor
+clusterization and suggest remedies \cite{fedorov1996, hooker2009,
+  fedorov2012, Ucinski05, neitzel2019sparse}. Others bypass the
+problem by allowing sensor locations to be chosen from a coarse finite
+grid \cite{alexanderian2014sparsification,
+  AlexanderianPetraStadlerEtAl16, alexanderian2018efficient,
+  koval2020, attia2020}, thus eliminating the possibility of sensor
+clusterization. Sadly, there is no free lunch, and choosing sensor
+locations from a coarse finite grid requires solving a hard
+combinatorial optimization problem. The current study is the first, to
+my knowledge, that aims to gain a deep understanding of sensor
+clusterization.
+
+\subsection{D-Optimal Designs}
+In this study we focus on the Bayesian \emph{D-optimality} criterion
+\cite{Chaloner1995}: a D-optimal design maximizes the expected
+Kullback-Leibler divergence \cite{CoverThomas91} between posterior and
+prior. For a linear model in finite dimensions, with Gaussian prior
+and Gaussian noise, a D-optimal design minimizes the determinant of
+the posterior covariance matrix\footnote{The frequentist version of
+the D-optimality criterion aims to minimize the determinant of the
+Fisher information matrix \cite[page 16]{Ucinski05}. For a linear
+model with Gaussian noise, the Bayesian criterion is just
+a regularized version of the frequentist criterion
+\cite{Chaloner1995}.}.
 
@@
 and prior measures. For a linear inverse problem with Gaussian prior
-and iid Gaussian noise, a D-optimal design minimizes an expression
-analogous to the determinant of the posterior covariance operator,
-understood as a product of its eigenvalues.
-%% Some difficulties arise in the process, but remedies can be found
-%% as shown in \cite{AlexanderianGloorGhattas14}.
+and Gaussian noise, a D-optimal design minimizes an expression
+analogous to the determinant of the posterior covariance operator
+\cite{AlexanderianGloorGhattas14} (also \cref{thm:d optimality}
+below).
 
-When one computes an optimal design using the D-optimality criterion,
-the optimization process sometimes results in observations that are
-very similar. For example, if an observation is thought of as
-measuring some function value at $\x \in \Omega \subseteq \R^d,
-d=1,2,3$ (with added error) then the optimization procedure sometimes
-places sensors in very close proximity to each other (as can be seen
-in Figure \ref{fig:clusterization illustration}). Following
-\cite{Ucinski05}, we refer to this phenomenon as \emph{sensor
-clusterization}.
+Considerable effort focused on computational aspects of finding
+optimal designs in infinite-dimensional Bayesian inverse problems.
+Attia \cite{attia2020} focused on QoI, Koval considered model
+uncertainty and focused on reducing forward and adjoint solves, using
+randomized linear algebra.\cite{koval2020}.
 
-Sensor clusterization is intuitively undesirable. Selecting a small
-set of sensors may lead to incomplete or biased data, as certain parts
-of the system or environment may be overlooked or
-underrepresented. This can result in inaccurate or unreliable
-estimates of the unknown parameters of interest, potentially leading
-to incorrect conclusions or decisions.
+\subsection{Remedies to Sensor Clusterization}
+Several remedies to sensor clusterization were suggeted through the
+years, and we will now overview some. In \emph{clusterization-free
+design} locations are forced to be far from each other
+\cite{Ucinski05}. This is achieved in a frequentist and
+finite-dimensional setting, by either imposing distance constraints
+between observations, or by introducing correlated errors which
+account for both observation error and model error
+\cite{Ucinski05}. Merging close measurements is also possible
+\cite{fedorov2012}. Sensor clusterization is mitigated in time-series
+analysis for pharmacokinetic experiments by modeling auto-correlation
+time in the noise \cite{hooker2009}, and this approach is similar in
+spirit to imposing error correlations as in \cite{Ucinski05}.
 
-%% \subsection{Related Work}
-Several remedies are suggested in the literature to phenomenon of
-sensor clusterization. In a frequentist and finite-dimensional
-context, \cite{fedorov1996} and \cite[chapter 2.4.3]{Ucinski05}
-discuss the clusterization phenomenon and suggest an approach called
-clusterization-free design. In such designs, the user enforces
-observation locations to be far from each other. One way to do this is
-by introducing correlated errors which account for both observation
-error and model error. Another method considered is imposing distance
-constraints between observations. A somewhat different approach is
-suggested in \cite[page 49]{fedorov2012}, where close observations are
-merged --- a procedure which obviously does not avoid
-clusterization. The same problem arises in time-series analysis for
-pharmacokinetic experiments. The authors of \cite{hooker2009} suggest
-modeling auto-correlation time in the noise model, which is equivalent
-to the correlated errors mentioned above. Any of the above mentioned
-approaches might serve as a remedy and push sensors away from each
-other. Yet, none offers any insight as to why clusterization
-occurs. Also, as better models are employed, model error is decreased
-and the clusterization phenomenon will eventually reappear. While
-these approaches are practical and help us avoid sensor
-clusterization, they do not provide insight as to why sensors are
-clustering.
+Many authors bypass the sensor clusterization problem by restricting
+measurements to a coarse finite grid \cite{koval2020,
+  AlexanderianPetraStadlerEtAl16, alexanderian2021, attia2020}, but
+this practice has two flaws: First, using a coarse grid requires
+solving a hard combinatorial optimization problem. Second, restricting
+sensors to a such a small subset of the computatoinal domain will
+necessarily give rise to sub-optimal designs. Several remedies were
+offered for the first problem: Stadler suggested ell0
+sparsification. Others suggested other stuff. None were offered for
+the second.
 
-Model errors were also considered for inverse problems in
-\cite{attia2020, koval2020}. The former study is focused on inferring
-a Quantity of Interest (QoI). The focus of the latter is reducing
-forward solves, using randomized linear algebra. Both studies present
-numerical techniques for finding optimal designs when model error is
-present. Both are restricted to linear inverse problems (although in
-the latter the authors use their method on a nonlinear problem by
-taking a Laplace approximation for the posterior). Both find an
-optimal design by first solving a continuous problem for sensor
-weights. Said solution is then sparsified to give a binary
-design. Both studies are successful in the task of Bayesian
-inversion. Neither \cite{attia2020}, nor \cite{koval2020} mention any
-effect model errors can have on sensor clusterization.
+%% Today, including model errors are a bona-fide????  standard in the
+%% inverse problems community: \cite{attia2020}, \cite{koval2020},
+%% \cite{Alexanderian2021} all empower their models.  Both studies
+%% present numerical techniques for finding optimal designs when model
+%% error is present. Both are restricted to linear inverse problems
+%% (although in the latter the authors use their method on a nonlinear
+%% problem by taking a Laplace approximation for the posterior). Both
+%% find an optimal design by first solving a continuous problem for
+%% sensor weights. Said solution is then sparsified to give a binary
+%% design. Both studies are successful in the task of Bayesian
+%% inversion. Neither \cite{attia2020}, nor \cite{koval2020} mention any
+%% effect model errors can have on sensor clusterization.
 
 \subsection{Contribution}
+Any of the above mentioned approaches might serve as a remedy and push
+sensors away from each other. Yet, none offers any insight as to why
+clusterization occurs. Also, as better models are employed, model
+error is decreased and the clusterization phenomenon will eventually
+reappear. While these approaches are practical and help practitioners
+avoid sensor clusterization, they do not provide insight as to why
+sensors are clustering.
+
 The current study aims to fill the gap of understanding sensor
@@
 first. This is the conclusion of part (5) of Theorem \ref{thm:char},
-which we state below and prove in Section \ref{section:vanishing}:
+which we state below and prove in Section \ref{section:vanishing}.
 

### 2023-04-19 latex/paper.tex
Writin

@@
-\documentclass{amsart}
+\documentclass{siamonline220329}
+%\documentclass{amsart}
 \input{definitions}
 
-\numberwithin{equation}{section}
+%\numberwithin{equation}{section}
 
+\title{Sensor Clusterization in D-optimal Designs for Bayesian Linear
+  Inverse Problems}
+
+\author{Yair Daon\thanks{Porter School of the Environment and Earth
+    Sciences, Tel Aviv University, Tel Aviv, Israel
+    (\email{yair.daon@gmail.com}).}}
+%% \address{Porter School of the Environment and Earth
+%%   Sciences, Tel Aviv University\\ Tel Aviv, Israel}
+%% %% \curraddr{Porter School of Environment Studies, Tel Aviv University\\ Tel Aviv, Israel}
+%% %% \curraddr{}
+%% \email{yair.daon@gmail.com}
+% The 2010 edition of the Mathematics Subject Classification is the
+% current definitive version.
+%% \subjclass{Primary: 
+%%   62F15, % Statistics - bayesian inference
+%%   35R30, % PDE - Inverse problems
+%%   Secondary:
+%%   28C20,
+%%   % Measure and integration - set functions and measures and integrals
+%%          % in infinite-dimensional spaces (Wiener measure, Gaussian
+%%   % measure, etc.
+%% }
+
+\date{\today}
 \begin{document}
-
-\title[Clusterization Properties of D-optimal Designs]{Clusterization
-  Properties of D-optimal Designs in Infinite Dimensional Bayesian
-  Linear Inverse Problems}
-
-\author{Yair Daon}
-\address{Porter School of the Environment and Earth
-  Sciences, Tel Aviv University\\ Tel Aviv, Israel}
-%% \curraddr{Porter School of Environment Studies, Tel Aviv University\\ Tel Aviv, Israel}
-%% \curraddr{}
-\email{yair.daon@gmail.com}
-%% \thanks{}
-
-%    The 2010 edition of the Mathematics Subject Classification is
-%    the current definitive version.
-\subjclass{Primary: 
-  62F15, % Statistics - bayesian inference
-  35R30, % PDE - Inverse problems
-  Secondary:
-  28C20, % Measure and integration - set functions and measures and integrals in
-  %infinite-dimensional spaces (Wiener measure, Gaussian measure, etc.
-}
-%% Primary: 62F15, 35R30; Secondary: 28C20
-\date{\today}
-
+\maketitle
 \begin{abstract}
@@
 %% \section{Open Problems}
-%% Some open problems to
-
-%% We two phenomena for a generic inverse problem with unit-norm
-%% constraints on obsevations: First, we showed that when model error is
-%% present and obsevation noise is small --- sensor clusterization will
-%% not occur. In contrast, we saw that when no model error is present ---
-%% sensor clusterization is quite generic and poses no obstruction to
-%% D-optimality.
 
@@
 This study is a part of my PhD thesis. Most of it was written under
-the instruction of Prof. Georg Stadler in the Courant Institute of
-Mathematical Sciences in New York University. I would like to thank
-Prof. Stadler for his great mentorship, attention to details and kind
-behavior.
+the instruction of Prof. Georg Stadler of the Courant Institute of
+Mathematical Sciences. I would like to thank Prof. Stadler for his
+great mentorship, attention to details and kind behavior.
 
@@
 
-\bibliographystyle{amsplain}
+\bibliographystyle{siamplain}
 \bibliography{/home/yair/projects/library.bib}

### 2023-04-19 latex/prelims.tex
Writin

@@
 following sections requires $\fwd$ and $\obs$ are explicitly
-separated.} $\obs \in ( \hilo^* )^m$, so that:
+separated \cite{cvetkovic2023choosing}.} $\obs \in ( \hilo^* )^m$, so that:
 
@@
 is summarized, in the notation presented above, in the following
-theorem\footnote{In \cite{AlexanderianGloorGhattas14}, results are
-stated for $\Sigma=I$ (implied by $\modcov = 0,\sigma^2=1$), but the
-authors also mention these results hold for more general covariance
-matrices (p. 681).}:
+theorem\footnote{In \cite{AlexanderianGloorGhattas14}\footnote{A
+similar theorem appears in \cite[Theorem
+  1]{alexanderian2018efficient}.}, results are stated for $\Sigma=I$
+(implied by $\modcov = 0,\sigma^2=1$), but the authors also mention
+these results hold for more general covariance matrices (p. 681).}:
 \begin{theorem}[Slightly modified Theorem 1 from \cite{AlexanderianGloorGhattas14}]\label{thm:d optimality}
@@
     &= \det \Big( \prcov ( \prcov^{-1} + \fwd ^* \obs^* \Sigma^{-1} \obs \fwd) \Big )\\
-    &= \det \prcov \det \postcov^{-1}.
+    &= \det \prcov \postcov^{-1}.
   \end{split}

### 2023-04-27 latex/intro.tex
getting there

@@
 \cite{horesh2008mri}), or location of sensors for waves reflecting
-from the ground (e.g.\ searching for oil \cite{horesh2008borehole} or
-using a radar). Whatever the allowed set of measurements is, one
-should select the optimal observations to take, in order to increase
-accuracy, reduce costs, or both.
+from the ground (e.g.\ searching for oil
+\cite{horesh2008borehole}). Whatever the allowed set of measurements
+is, one should select the optimal measurements to take, in order to
+increase accuracy, reduce costs, or both. A set of measurements is
+called a \emph{design}, and an \emph{optimal design} is found by
+optimizing some \emph{design criterion}. The most popular desgin
+criteria are A- and D-optimality
 
-A set of measurements is called a \emph{design}, and an \emph{optimal
-design} is found by optimizing some \emph{design
-criterion}. Surprisingly, optimal designs sometimes result in
-measuremets that are very similar. For example consider measurements
-taken by sensors that evaluate some function at $\x \in \Omega
-\subseteq \R^d, d=1,2,3$. An optimal design may place sensors in very
-close proximity to each other (Figure \ref{fig:clusterization
-  illustration}). Following \cite{Ucinski05}, we refer to this
-phenomenon as \emph{sensor clusterization}.
+Surprisingly, optimal designs sometimes result in measuremets that are
+very similar \cite{fedorov1996, hooker2009, fedorov2012, Ucinski05,
+  neitzel2019sparse}. As an example, consider the inverse problem of
+inferring the initial condition of the 1D heat equation (details in
+Section \ref{subsec:example}). In Figure \ref{fig:clusterization
+  illustration} we see that an optimal design results in two sets of
+sensors that are almost indistinguishable from each other. Following
+\cite{Ucinski05}, we refer to this phenomenon as \emph{sensor
+clusterization}.
+
+\begin{figure}
+  \begin{tikzpicture}[thick, scale=1.3, every node/.style={scale=0.99}]
+    \begin{axis}
+      [
+      title={Posterior Pointwise Standard Deviations and D-Optimal Sensor Locations},  
+      xmin = 0,
+      xmax = 3.14,
+      xlabel = {$x \in \Omega$},
+      ylabel = posterior std,
+      ymin   = 0,
+      %compat = 1.3,
+      % ymax   = 130,
+      % ytick = \empty,
+      legend cell align=left,
+      % legend style={at={(0.45,0.2)}}
+      legend pos= outer north east 
+      ]
+      % \draw[black!30!white, thin] (50,0) -- (50,130);
+      % 
+      %% \addplot [thin, black, mark=none] table{stdv-heat-sens1-var1.txt};
+      %% \addlegendentry{1 sensors};
+
+      %% \addplot [thin, blue, mark=none] table{stdv-heat-sens2-var1.txt};
+      %% \addlegendentry{2 sensors};
+
+      %% \addplot [thin, red, mark=none] table{stdv-heat-sens3-var1.txt};
+      %% \addlegendentry{3 sensors};
+
+      \addplot [thin, green, mark=none] table{stdv-heat-sens4-var1.txt};
+      \addlegendentry{4 sensors};
+
+      \addplot [thin, purple, mark=none] table{stdv-heat-sens5-var1.txt};
+      \addlegendentry{5 sensors};
+
+      \addplot [thin, cyan, mark=none] table{stdv-heat-sens6-var1.txt};
+      \addlegendentry{6 sensors};
+
+      %% \addplot [black,  only marks, mark=x, mark size=1.5] 
+      %% table{locs-heat-sens1-var1.txt}; 
+      %% \addplot [blue,   only marks, mark=x, mark size=1.5]
+      %% table{locs-heat-sens2-var1.txt}; 
+      %% \addplot [red,    only marks, mark=x, mark size=1.5]
+      %% table{locs-heat-sens3-var1.txt};
+      \addplot [green,  only marks, mark=*, mark size=1.5] 
+      table{locs-heat-sens4-var1.txt}; 
+      \addplot [purple, only marks, mark=*, mark size=1.5] 
+      table{locs-heat-sens5-var1.txt}; 
+      \addplot [cyan,   only marks, mark=*, mark size=1.5] 
+      table{locs-heat-sens6-var1.txt};
+
+    \end{axis}
+  \end{tikzpicture}
+  \caption{Sensor clusterization for inferring the initial condition
+    of the 1D heat equation. Posterior pointwise standard deviations
+    (lines) are plotted over the computatoinal domain $\Omega = [0,
+      \pi]$, for varying numbers of sensors. Sensor locations
+    (circles) were chosen to optimize the D-optimality
+    criterion. Sensor clusterization occurs for six sensors: Only four
+    observation locations are visible, since two pairs of sensors are
+    so close together they are indistinguishable.}
+  \label{fig:clusterization illustration}
+\end{figure}
 
 There is a consensus among researchers that sensor clusterization is
-undesirable. Some authors acknowledge the problem of sensor
-clusterization and suggest remedies \cite{fedorov1996, hooker2009,
-  fedorov2012, Ucinski05, neitzel2019sparse}. Others bypass the
-problem by allowing sensor locations to be chosen from a coarse finite
-grid \cite{alexanderian2014sparsification,
-  AlexanderianPetraStadlerEtAl16, alexanderian2018efficient,
-  koval2020, attia2020}, thus eliminating the possibility of sensor
-clusterization. Sadly, there is no free lunch, and choosing sensor
-locations from a coarse finite grid requires solving a hard
-combinatorial optimization problem. The current study is the first, to
-my knowledge, that aims to gain a deep understanding of sensor
-clusterization.
+undesirable \cite{fedorov1996, hooker2009, fedorov2012, Ucinski05,
+  neitzel2019sparse} and several remedies to sensor clusterization
+were suggeted through the years. In a \emph{clusterization-free
+design}, sensor locations are forced to be far from each other
+\cite{Ucinski05}. In a frequentist and finite-dimensional setting,
+senesor clusterization is avoided by either imposing distance
+constraints between measurements, or by introducing correlated errors
+which account for both observation error and model error
+\cite{Ucinski05}. Merging close measurements is also possible
+\cite{fedorov2012}, but this practice ignores sensor clusterization
+altogether. In time-series analysis for pharmacokinetic experiments,
+sensor clusterization is mitigated by modeling auto-correlation time
+in the noise \cite{hooker2009}, and this approach is similar in spirit
+to assuming errors are correlatied as in \cite{Ucinski05}.
 
-\subsection{D-Optimal Designs}
-In this study we focus on the Bayesian \emph{D-optimality} criterion
-\cite{Chaloner1995}: a D-optimal design maximizes the expected
-Kullback-Leibler divergence \cite{CoverThomas91} between posterior and
-prior. For a linear model in finite dimensions, with Gaussian prior
-and Gaussian noise, a D-optimal design minimizes the determinant of
-the posterior covariance matrix\footnote{The frequentist version of
-the D-optimality criterion aims to minimize the determinant of the
-Fisher information matrix \cite[page 16]{Ucinski05}. For a linear
-model with Gaussian noise, the Bayesian criterion is just
-a regularized version of the frequentist criterion
-\cite{Chaloner1995}.}.
+%% Sometimes, imposing a correlation structure on observation noise is
+%% reasonable indeed \cite{koval2020, Alexanderian2021}. For example,
+%% Koval \cite{koval2020} imposes a Gaussian prior on ocean bathytermy
+%% when inferring BLA. However, in the (un)fortunate event that ocean
+%% bathytermy is known, there is no use to modeling it as an unknown
+%% and sensor clusterization will rear its ugly head.
 
-Bayesian D-optimal designs for inverse problems on function spaces
-were first studied by \cite{AlexanderianGloorGhattas14,
-  AlexanderianPetraStadlerEtAl16,
-  AlexanderianPetraStadlerEtAl14}. Similarly to the finite dimensional
-case, a D-optimal design maximizes the KL divergence between posterior
-and prior measures. For a linear inverse problem with Gaussian prior
-and Gaussian noise, a D-optimal design minimizes an expression
-analogous to the determinant of the posterior covariance operator
-\cite{AlexanderianGloorGhattas14} (also \cref{thm:d optimality}
-below).
+Many scientists bypass the sensor clusterization problem by
+restricting measurements to a coarse grid of the computational domain
+$\Omega \subseteq \mathbb{R}^d, d=1,2,3$ \cite{koval2020,
+  alexanderian2021, attia2020, alexanderian2014sparsification,
+  AlexanderianPetraStadlerEtAl16, alexanderian2018efficient}, but this
+practice requires solvilg a hard combinatorial optimization problem
+for choosing optimal sensor locations from a finite set. A solution
+typically involves relaxing the problem to that of choosing optimal
+measurement ``weight'' in $[0,1]$ for each potential sensor
+location. Some authors then add a sparsifying $\ell_1$ penalty term to
+the design criterion, which is subsequently thresholded
+\cite{horesh2008borehole}, thus resulting in a sparse binary
+solution. Others successively relax the $\ell_1$ penalty to a $\ell_0$
+penalty via a continuation method
+\cite{AlexanderianPetraStadlerEtAl16,
+  alexanderian2014sparsification}. A clever alternative is to cast the
+problem of finding optimal sensor weights as a stochastic optimization
+problem, which in turn is solved by stochastic gradient descent
+\cite{attia2022stochastic}. While all of the above mentioned methods
+are valid solutions to the prohibitive computational cost of finding
+an binary optimal design restricted to a coarse grid, none treats the
+elephant in the room: restricting sensors to a coarse grid in $\Omega
+\subseteq \mathbb{R}^d, d=1,2,3$ necessarily gives rise to a
+sub-optimal design.
 
-Considerable effort focused on computational aspects of finding
-optimal designs in infinite-dimensional Bayesian inverse problems.
-Attia \cite{attia2020} focused on QoI, Koval considered model
-uncertainty and focused on reducing forward and adjoint solves, using
-randomized linear algebra.\cite{koval2020}.
+Avoiding sensor clusterization is pragmatical: we feel sensor
+clusterization is bad (but do not know why), and try to avoid
+it. However, all of the abovementioned techniques that avoid sensor
+clusterization do so by somehow restricting sensor locations, thus
+fundamentaly altering the optimal design problem. To the best of my
+knowledge, no study tried to answer the following seemingly trivial
+and basic questions: why do sensor clusterization occurs? Is sensor
+clusterization really bad? Can we find optimal designs BLABLA?
 
-\subsection{Remedies to Sensor Clusterization}
-Several remedies to sensor clusterization were suggeted through the
-years, and we will now overview some. In \emph{clusterization-free
-design} locations are forced to be far from each other
-\cite{Ucinski05}. This is achieved in a frequentist and
-finite-dimensional setting, by either imposing distance constraints
-between observations, or by introducing correlated errors which
-account for both observation error and model error
-\cite{Ucinski05}. Merging close measurements is also possible
-\cite{fedorov2012}. Sensor clusterization is mitigated in time-series
-analysis for pharmacokinetic experiments by modeling auto-correlation
-time in the noise \cite{hooker2009}, and this approach is similar in
-spirit to imposing error correlations as in \cite{Ucinski05}.
+%% However, as far as I am aware, no study concerned the fact that
+%% restricting sensors to a coarse grid is, by definition,
+%% sub-optimal.
 
-Many authors bypass the sensor clusterization problem by restricting
-measurements to a coarse finite grid \cite{koval2020,
-  AlexanderianPetraStadlerEtAl16, alexanderian2021, attia2020}, but
-this practice has two flaws: First, using a coarse grid requires
-solving a hard combinatorial optimization problem. Second, restricting
-sensors to a such a small subset of the computatoinal domain will
-necessarily give rise to sub-optimal designs. Several remedies were
-offered for the first problem: Stadler suggested ell0
-sparsification. Others suggested other stuff. None were offered for
-the second.
+%% Attia \cite{attia2020} focused on QoI, Koval considered model
+%% uncertainty and focused on reducing forward and adjoint solves, using
+%%  randomized linear algebra \cite{koval2020}.
 
@@
 
+%% Considerable efforts focused on various computational aspects of
+%% finding optimal designs in infinite-dimensional Bayesian inverse
+%% problems \cite{attia2020, koval2020,
+%% AlexanderianPetraStadlerEtAl16, attia2023pyoed, koval2021optimal,
+%% attia2020stochastic}, the subject of sensor clusterization was less
+%% investigated, and
+
+%% Second, restricting sensors to a such a small subset of the computatoinal
+%% domain $\Omega \subseteq \mathbb{R}^d, d=1,2,3$ necessarily gives rise
+%% to sub-optimal designs.
+
 \subsection{Contribution}
-Any of the above mentioned approaches might serve as a remedy and push
-sensors away from each other. Yet, none offers any insight as to why
-clusterization occurs. Also, as better models are employed, model
-error is decreased and the clusterization phenomenon will eventually
-reappear. While these approaches are practical and help practitioners
-avoid sensor clusterization, they do not provide insight as to why
-sensors are clustering.
-
-The current study aims to fill the gap of understanding sensor
-clusterization in infinte dimensional Bayesian linear inverse
-problems. We propose and study a relaxed and analytically tractable
-model for D-optimal designs. Under this model, D-optimal designs are
-solutions of a constrained optimization problem, formulated using
-Lagrange multipliers. The Lagrange multipliers formulation allows us
-to rigorously show how model error mitigates clusterization (Section
-\ref{section:non vanishing}), and why designs that exhibit
-clusterization are indeed optimal when no model error is considered.
+In this study, we seeks to understand sensor clusterization. We focus
+our attention on the Bayesian \emph{D-optimality} criterion: a
+D-optimal design maximizes the expected Kullback-Leibler divergence
+\cite{CoverThomas91} between posterior and prior
+\cite{Chaloner1995}. We study Bayesian D-optimal designs in linear
+inverse problems that are (potentially) of infinte dimensions. We
+propose and study a relaxed and analytically tractable model for such
+D-optimal designs. Under this model, D-optimal designs are solutions
+of a constrained optimization problem, formulated using Lagrange
+multipliers. The Lagrange multipliers formulation allows us to
+rigorously show why model error mitigates clusterization (Section
+\ref{section:non vanishing}), like many authors note, and why designs
+that exhibit clusterization are indeed optimal when no model error is
+considered (Section \ref{section:vanishing}).
 
@@
   \item $\post$ the posterior measure, with covariance $\postcov$.
-  \item A D-optimality utility function
+  \item A D-optimality design criterion
     \cite{AlexanderianGloorGhattas14}:
@@
 
-One byproduct of the proof of Theorem \ref{thm:char} is a novel lemma
-in linear algebra: We show how to decompose a symmetric positive
+Another byproduct of the proof of Theorem \ref{thm:char} is a novel
+linear algebra lemma: We show how to decompose a symmetric positive
 definite matrix $M$ as $M = AA^t$, where $A$ has unit norm columns
@@
 \subsection{An Example of Clusterization}\label{subsec:example}
-\begin{figure}
-  \begin{tikzpicture}[thick, scale=1.3, every node/.style={scale=0.99}]
-    \begin{axis}
-      [
-      title={Posterior Pointwise Standard Deviations and D-Optimal Sensor Locations},  
-      xmin = 0,
-      xmax = 3.14,
-      xlabel = {$x$},
-      ylabel = posterior std,
-      ymin   = 0,
-      %compat = 1.3,
-      % ymax   = 130,
-      % ytick = \empty,
-      legend cell align=left,
-      % legend style={at={(0.45,0.2)}}
-      legend pos= outer north east 
-      ]
-      % \draw[black!30!white, thin] (50,0) -- (50,130);
-      % 
-      %% \addplot [thin, black, mark=none] table{stdv-heat-sens1-var1.txt};
-      %% \addlegendentry{1 sensors};
-
-      %% \addplot [thin, blue, mark=none] table{stdv-heat-sens2-var1.txt};
-      %% \addlegendentry{2 sensors};
-
-      %% \addplot [thin, red, mark=none] table{stdv-heat-sens3-var1.txt};
-      %% \addlegendentry{3 sensors};
-
-      \addplot [thin, green, mark=none] table{stdv-heat-sens4-var1.txt};
-      \addlegendentry{4 sensors};
-
-      \addplot [thin, purple, mark=none] table{stdv-heat-sens5-var1.txt};
-      \addlegendentry{5 sensors};
-
-      \addplot [thin, cyan, mark=none] table{stdv-heat-sens6-var1.txt};
-      \addlegendentry{6 sensors};
-
-      %% \addplot [black,  only marks, mark=x, mark size=1.5] 
-      %% table{locs-heat-sens1-var1.txt}; 
-      %% \addplot [blue,   only marks, mark=x, mark size=1.5]
-      %% table{locs-heat-sens2-var1.txt}; 
-      %% \addplot [red,    only marks, mark=x, mark size=1.5]
-      %% table{locs-heat-sens3-var1.txt};
-      \addplot [green,  only marks, mark=*, mark size=1.5] 
-      table{locs-heat-sens4-var1.txt}; 
-      \addplot [purple, only marks, mark=*, mark size=1.5] 
-      table{locs-heat-sens5-var1.txt}; 
-      \addplot [cyan,   only marks, mark=*, mark size=1.5] 
-      table{locs-heat-sens6-var1.txt};
-
-    \end{axis}
-  \end{tikzpicture}
-  \caption{Sensor clusterization for inferring the initial condition
-    of the 1D heat equation. Posterior pointwise standard deviations
-    (lines) are plotted over the domain $[0, \pi]$, for varying
-    numbers of sensors. Sensor locations (circles) were chosen to
-    optimize the D-optimality criterion. Sensor clusterization occurs
-    for six sensors. Only four observation locations are used --- two
-    pairs of sensors are so close together they are
-    indistinguishable.}
-  \label{fig:clusterization illustration}
-\end{figure}
-
 For the purpose of illustration, consider a toy inverse problem ---
-inferring the initial condition of the 1D heat equation in $[0,\pi]$
-with a homogeneous Dirichlet boundary condition. Sensor clusterization
-is clearly evident, see Figure \ref{fig:clusterization}.
+inferring the initial condition of the 1D heat equation in $\Omega :=
+[0,\pi]$ with a homogeneous Dirichlet boundary condition. Sensor
+clusterization is clearly evident, see Figure \ref{fig:clusterization
+  illustration}.
 
@@
 \ref{section:prelim} and \eqref{eq:postcov} below). Finally, choose
-$\x_j,j=1,\dots,m$ to optimize the D-optimality criterion
-\cite{AlexanderianGloorGhattas14}.
+$\x_j,j=1,\dots,m$ to optimize the D-optimality criterion, see Theorem
+\ref{thm:d optimality} \cite{AlexanderianGloorGhattas14}.
 

### 2023-04-27 latex/prelims.tex
getting there

@@
 \cite{AlexanderianGloorGhattas14}, but the analysis carried out in the
-following sections requires $\fwd$ and $\obs$ are explicitly
-separated \cite{cvetkovic2023choosing}.} $\obs \in ( \hilo^* )^m$, so that:
+following sections requires $\fwd$ and $\obs$ are explicitly separated
+similarly to \cite{attia2022stochastic, cvetkovic2023choosing}.} $\obs
+\in ( \hilo^* )^m$, so that:
 
@@
 
-D-optimal designs in infinite-dimensional Hilbert spaces were
-investigated in \cite{AlexanderianGloorGhattas14}. Their main result
-is summarized, in the notation presented above, in the following
-theorem\footnote{In \cite{AlexanderianGloorGhattas14}\footnote{A
-similar theorem appears in \cite[Theorem
-  1]{alexanderian2018efficient}.}, results are stated for $\Sigma=I$
-(implied by $\modcov = 0,\sigma^2=1$), but the authors also mention
-these results hold for more general covariance matrices (p. 681).}:
+For a Bayesian linear model in finite dimensions, with Gaussian prior
+and Gaussian noise, a D-optimal design minimizes the determinant of
+the posterior covariance matrix\footnote{The frequentist version of
+the D-optimality criterion aims to minimize the determinant of the
+Fisher information matrix \cite[page 16]{Ucinski05}. Thus, the
+Bayesian criterion is just a regularized version of the frequentist
+criterion \cite{Chaloner1995}.}. Bayesian D-optimal designs for
+inverse problems on function spaces were first studied by
+\cite{AlexanderianGloorGhattas14, AlexanderianPetraStadlerEtAl16,
+  AlexanderianPetraStadlerEtAl14}, and just like in finite dimensions,
+a D-optimal design maximizes the Kullback-Leibler divergence between
+posterior and prior measures.
+%% For a linear inverse problem with Gaussian prior and Gaussian
+%% noise, a D-optimal design minimizes an expression analogous to the
+%% determinant of the posterior covariance operator
+%% \cite{AlexanderianGloorGhattas14} (also \cref{thm:d optimality}
+%% below).
+D-optimal designs in infinite-dimensional
+Hilbert spaces were investigated in
+\cite{AlexanderianGloorGhattas14}. Their main result is summarized, in
+the notation presented above, in the following theorem\footnote{In
+\cite{AlexanderianGloorGhattas14, alexanderian2018efficient}, results
+are stated for $\Sigma=I$ (implied by $\modcov = 0,\sigma^2=1$), but
+the authors also mention these results hold for more general
+covariance matrices \cite[p. 681]{AlexanderianGloorGhattas14}.}:
 \begin{theorem}[Slightly modified Theorem 1 from \cite{AlexanderianGloorGhattas14}]\label{thm:d optimality}

### 2023-05-02 latex/intro.tex
Tryin to fix refs

@@
 \section{Introduction}\label{section:OED intro}
-Experimental design is an important part of many scientific
-investigations. When considering an inverse problem, one can often
-specify sensor locations (e.g.\ in electricat impedance tomography
+Measurements of physical processes are indespensable when considering
+an inverse problem, choosing an optimal set of measurements is of
+utmost importance. In applications, one may have to specify electrode
+locations on the skin (e.g.\ in electricat impedance tomography
 \cite{horesh2010impedance}), certain wavelengths (e.g.\ in MRI
@@
 called a \emph{design}, and an \emph{optimal design} is found by
-optimizing some \emph{design criterion}. The most popular desgin
-criteria are A- and D-optimality
+optimizing some \emph{design criterion}. Two of the most popular
+desgin criteria are A- and D-optimality (see \cite{Chaloner1995} for a
+thorough review).
 
@@
   neitzel2019sparse} and several remedies to sensor clusterization
-were suggeted through the years. In a \emph{clusterization-free
-design}, sensor locations are forced to be far from each other
+were suggeted through the years. Merging ``close'' measurements is
+possible, \cite{fedorov2012}, but this practice ignores sensor
+clusterization altogether. In a \emph{clusterization-free design},
+sensor locations are forced to be far from each other
 \cite{Ucinski05}. In a frequentist and finite-dimensional setting,
@@
 which account for both observation error and model error
-\cite{Ucinski05}. Merging close measurements is also possible
-\cite{fedorov2012}, but this practice ignores sensor clusterization
-altogether. In time-series analysis for pharmacokinetic experiments,
-sensor clusterization is mitigated by modeling auto-correlation time
-in the noise \cite{hooker2009}, and this approach is similar in spirit
-to assuming errors are correlatied as in \cite{Ucinski05}.
+\cite{Ucinski05}. In time-series analysis for pharmacokinetic
+experiments, sensor clusterization is mitigated by modeling
+auto-correlation time in the noise \cite{hooker2009}, and this
+approach is similar in spirit to assuming errors are correlatied
+\cite{Ucinski05}.
 
@@
 
-Many scientists bypass the sensor clusterization problem by
-restricting measurements to a coarse grid of the computational domain
-$\Omega \subseteq \mathbb{R}^d, d=1,2,3$ \cite{koval2020,
-  alexanderian2021, attia2020, alexanderian2014sparsification,
-  AlexanderianPetraStadlerEtAl16, alexanderian2018efficient}, but this
-practice requires solvilg a hard combinatorial optimization problem
-for choosing optimal sensor locations from a finite set. A solution
-typically involves relaxing the problem to that of choosing optimal
-measurement ``weight'' in $[0,1]$ for each potential sensor
-location. Some authors then add a sparsifying $\ell_1$ penalty term to
-the design criterion, which is subsequently thresholded
-\cite{horesh2008borehole}, thus resulting in a sparse binary
-solution. Others successively relax the $\ell_1$ penalty to a $\ell_0$
-penalty via a continuation method
-\cite{AlexanderianPetraStadlerEtAl16,
-  alexanderian2014sparsification}. A clever alternative is to cast the
-problem of finding optimal sensor weights as a stochastic optimization
-problem, which in turn is solved by stochastic gradient descent
-\cite{attia2022stochastic}. While all of the above mentioned methods
-are valid solutions to the prohibitive computational cost of finding
-an binary optimal design restricted to a coarse grid, none treats the
-elephant in the room: restricting sensors to a coarse grid in $\Omega
-\subseteq \mathbb{R}^d, d=1,2,3$ necessarily gives rise to a
-sub-optimal design.
+In spatial problems, where sensors are placed in some domain $\Omega
+\subseteq \mathbb{R}^d, d=1,2,3$, many scientists bypass the sensor
+clusterization problem by restricting measurements to a coarse grid in
+$\Omega$ \cite{koval2020, alexanderian2021, attia2020,
+  alexanderian2014, alexanderian2016,
+  alexanderian2018efficient}. Unfortunately, this practice requires
+solvilg a hard combinatorial optimization problem for choosing optimal
+sensor locations from a finite set. A solution typically involves
+relaxing the problem to choosing optimal measurement ``weight'' in
+$[0,1]$ for each potential sensor location. Some authors then add a
+sparsifying $\ell_1$ penalty term to the design criterion, which is
+subsequently thresholded \cite{horesh2008borehole}, thus resulting in
+the desired binary design. Others successively relax the $\ell_1$
+penalty to a $\ell_0$ penalty via a continuation method
+\cite{alexanderian2016, alexanderian2014}. A clever alternative is to
+cast the problem of finding optimal sensor weights as a stochastic
+optimization problem, which in turn is solved by stochastic gradient
+descent \cite{attia2022stochastic}????. While all of the above
+mentioned methods are valid solutions to the prohibitive computational
+cost of finding a binary optimal design restricted to a coarse grid,
+none treats the elephant in the room: restricting sensor locations to
+a coarse grid in $\Omega \subseteq \mathbb{R}^d, d=1,2,3$ necessarily
+gives rise to a sub-optimal design.
 
@@
 knowledge, no study tried to answer the following seemingly trivial
-and basic questions: why do sensor clusterization occurs? Is sensor
-clusterization really bad? Can we find optimal designs BLABLA?
+and basic questions: why do sensor clusterization occurs? Should
+sensor clusterization really be avoided? How imposing a correlation
+structure between obsevations mitigates sensor clusterization? Can we
+find truly optimal designs that do not exhibit sensor clusterization?
 
@@
 
-As a byproduct, the Lagrange multipliers problem for no model error
-results in a characterization of D-optimal designs. The key idea is
-that a D-optimal design reduces uncertainties where they are highest
+the Lagrange multipliers problem for vanishing model error results in
+a characterization of D-optimal designs. The key idea is that a
+D-optimal design reduces uncertainties where they are highest
 first. This is the conclusion of part (5) of Theorem \ref{thm:char},
-which we state below and prove in Section \ref{section:vanishing}.
+stated below and proved in Section \ref{section:vanishing}.
 
@@
 
-Another byproduct of the proof of Theorem \ref{thm:char} is a novel
-linear algebra lemma: We show how to decompose a symmetric positive
-definite matrix $M$ as $M = AA^t$, where $A$ has unit norm columns
-(Lemma \ref{lemma:lax}). We also generalize two lemmas from linear
-algebra to infinite-dimensional settings: In Lemma \ref{lemma:MDL} we
-prove a matrix determinant lemma in Hilbert spaces, and in Lemma
-\ref{lemma:lax} we generalize a lemma \cite{Lax07} for calculating
-$\frac{\der}{\der t} \log \det (I + X(t))$, where $X(t)$ is an
-operator valued function.
+An interesting byproduct of the proof of Theorem \ref{thm:char} is a
+novel lemma in linear algebra: For a symmetric positive definite
+matrix $M \in \mathbb{R}^{k \times k}$ with $\ttr M = m \in
+\mathbb{N}$, we decompose $M = AA^t$, where $A$ has unit norm columns
+(Lemma \ref{lemma:free}). We also generalize two lemmas from linear
+algebra to infinite-dimensional settings, see Lemmas \ref{lemma:MDL}
+and \ref{lemma:lax}.
+
+%%  we prove a matrix determinant lemma in Hilbert spaces, and in
+%%  Lemma \ref{lemma:lax} we generalize a lemma for calculating
+%%  $\frac{\der}{\der t} \log \det (I + X(t))$, where $X(t)$ is an
+%%  operator valued function \cite{Lax07}.
 
 \subsection{Limitations}\label{subsec:limitations}
-The study presented here is not free from drawbacks. First, our
-relaxed model does not correspond to any real-life problem. It is
-generic enough to be analytically tractable, but one may argue it is
-to far removed from any real application. To these claims I would
-argue that Mathematicians have a long history of studying models that
-are bare-bones simplifications of real systems, e.g. the Ising model
-\cite{cipra1987} the Lorenz system \cite{brin}, and the SIR model
-\cite{keeling2008}, to name a few. The second drawback of our model is
-that we do not show rigorously that clusterization necessarily occurs,
-we only show that it does not pose any obstruction to D-optimality.
+The main limitation of this study is that our generic model does not
+correspond to any specific real-life problem. It is generic enough to
+be analytically tractable, but one may argue it is too far removed
+from any real application. To these claims I would argue that
+Mathematicians have a long history of studying models that are
+bare-bones simplifications of real systems. E.g. the Ising model
+\cite{cipra1987}, the Lorenz system \cite{brin}, and the SIR model
+\cite{keeling2008}, to name a few.
+
+%% The second drawback of our model is that we do not show rigorously
+%% that clusterization necessarily occurs, we only show that it does
+%% not pose any obstruction to D-optimality.
 

### 2023-05-02 latex/prelims.tex
Tryin to fix refs

@@
 
-\subsection{Overview}
+%% \subsection{Overview}
 In this section I will present notations that will be used throughout
@@
 \ref{subsec:notation}. The theoretical foundations for inverse
-problems over function spaces can be found in
-\cite{Stuart10}.
+problems over function spaces can be found in \cite{Stuart10}.
 
@@
 inverse problems on function spaces were first studied by
-\cite{AlexanderianGloorGhattas14, AlexanderianPetraStadlerEtAl16,
-  AlexanderianPetraStadlerEtAl14}, and just like in finite dimensions,
+\cite{AlexanderianGloorGhattas14, alexanderian2018efficient}, and just like in finite dimensions,
 a D-optimal design maximizes the Kullback-Leibler divergence between

### 2023-05-05 latex/intro.tex
Moved example to appendix, added questions and answers in intro.

--- 2023-04-19 latex/intro.tex
@@
-\section{Introduction}\label{section:OED intro}
-Experimental design is an important part of many scientific
-investigations. When considering an inverse problem, one can often
-specify sensor locations (e.g.\ in electricat impedance tomography
-\cite{horesh2010impedance}), certain wavelengths (e.g.\ in MRI
-\cite{horesh2008mri}), or location of sensors for waves reflecting
-from the ground (e.g.\ searching for oil \cite{horesh2008borehole} or
-using a radar). Whatever the allowed set of measurements is, one
-should select the optimal observations to take, in order to increase
-accuracy, reduce costs, or both.
+\section{Introduction}\label{section:intro}
+Measurements play a fundamental role in generating observations and
+are indispensable to any inference process. In the realm of data
+science, selecting the optimal set of measurements holds particular
+significance when inferring parameters of a physical process. Unlike
+many other fields, where measurements are fixed or limited, in this
+context, we have the freedom to choose which measurements to take. We
+would like to take advantage of this freedom to choose measurements
+that enhance accuracy, reduce costs, or achieve both objectives
+simultaneously. Whether measurements involves specifying electrode
+locations on the skin in electric impedance tomography
+\cite{horesh2010impedance}, determining certain wavelengths in MRI
+\cite{horesh2008mri}, or positioning sensors for detecting
+ground-reflected waves in the search for oil
+\cite{horesh2008borehole}, the selection of optimal measurements,
+referred to as the problem of \emph{optimal design}, becomes
+crucial. To achieve this, researchers optimize specific \emph{design
+criteria}, with A- and D-optimality being two of the most widely
+recognized and extensively studied design criteria
+\cite{Chaloner1995}.
 
-A set of measurements is called a \emph{design}, and an \emph{optimal
-design} is found by optimizing some \emph{design
-criterion}. Surprisingly, optimal designs sometimes result in
-measuremets that are very similar. For example consider measurements
-taken by sensors that evaluate some function at $\x \in \Omega
-\subseteq \R^d, d=1,2,3$. An optimal design may place sensors in very
-close proximity to each other (Figure \ref{fig:clusterization
-  illustration}). Following \cite{Ucinski05}, we refer to this
-phenomenon as \emph{sensor clusterization}.
+Interestingly, A- and D-optimal designs have been observed to yield
+remarkably similar measurements in certain cases \cite{fedorov1996,
+  hooker2009, fedorov2012, Ucinski05, neitzel2019sparse}. To
+illustrate this point, we consider a toy inverse problem: inferring
+the initial condition of the 1D heat equation (details in Section
+\ref{section:example}). In Figure \ref{fig:clusterization
+  illustration} D-optimal measurement locations are shown for
+different numbers of allowed measurements. Notably, for six
+measurement locations, the D-optimal design yields two sets of
+measurements that are almost indistinguishable from one another. This
+intriguing phenomenon is known as \emph{measurement clusterization}
+\cite{Ucinski05}, and we refer to a design that exhibits measurement
+clusterization as a \emph{clustered design}.
 
-There is a consensus among researchers that sensor clusterization is
-undesirable. Some authors acknowledge the problem of sensor
-clusterization and suggest remedies \cite{fedorov1996, hooker2009,
-  fedorov2012, Ucinski05, neitzel2019sparse}. Others bypass the
-problem by allowing sensor locations to be chosen from a coarse finite
-grid \cite{alexanderian2014sparsification,
-  AlexanderianPetraStadlerEtAl16, alexanderian2018efficient,
-  koval2020, attia2020}, thus eliminating the possibility of sensor
-clusterization. Sadly, there is no free lunch, and choosing sensor
-locations from a coarse finite grid requires solving a hard
-combinatorial optimization problem. The current study is the first, to
-my knowledge, that aims to gain a deep understanding of sensor
-clusterization.
+\begin{figure}
+  \begin{tikzpicture}[thick, scale=1.3, every node/.style={scale=0.99}]
+    \begin{axis}
+      %% title={Posterior Pointwise Standard Deviations and D-Optimal
+      %% Measurement Locations}
+      [xmin = 0, xmax = 3.14, xlabel = {$x \in \Omega$}, ylabel =
+        posterior pointwise std, ymin = 0,
+      %compat = 1.3,
+      % ymax   = 130,
+      % ytick = \empty,
+      legend cell align=left,
+      % legend style={at={(0.45,0.2)}}
+      legend pos= outer north east 
+      ]
+      % \draw[black!30!white, thin] (50,0) -- (50,130);
+      % 
+      %% \addplot [thin, black, mark=none] table{stdv-heat-sens1-var1.txt};
+      %% \addlegendentry{1 measurements};
 
-\subsection{D-Optimal Designs}
-In this study we focus on the Bayesian \emph{D-optimality} criterion
-\cite{Chaloner1995}: a D-optimal design maximizes the expected
-Kullback-Leibler divergence \cite{CoverThomas91} between posterior and
-prior. For a linear model in finite dimensions, with Gaussian prior
-and Gaussian noise, a D-optimal design minimizes the determinant of
-the posterior covariance matrix\footnote{The frequentist version of
-the D-optimality criterion aims to minimize the determinant of the
-Fisher information matrix \cite[page 16]{Ucinski05}. For a linear
-model with Gaussian noise, the Bayesian criterion is just
-a regularized version of the frequentist criterion
-\cite{Chaloner1995}.}.
+      %% \addplot [thin, blue, mark=none] table{stdv-heat-sens2-var1.txt};
+      %% \addlegendentry{2 measurements};
 
-Bayesian D-optimal designs for inverse problems on function spaces
-were first studied by \cite{AlexanderianGloorGhattas14,
-  AlexanderianPetraStadlerEtAl16,
-  AlexanderianPetraStadlerEtAl14}. Similarly to the finite dimensional
-case, a D-optimal design maximizes the KL divergence between posterior
-and prior measures. For a linear inverse problem with Gaussian prior
-and Gaussian noise, a D-optimal design minimizes an expression
-analogous to the determinant of the posterior covariance operator
-\cite{AlexanderianGloorGhattas14} (also \cref{thm:d optimality}
-below).
+      %% \addplot [thin, red, mark=none] table{stdv-heat-sens3-var1.txt};
+      %% \addlegendentry{3 measurements};
 
-Considerable effort focused on computational aspects of finding
-optimal designs in infinite-dimensional Bayesian inverse problems.
-Attia \cite{attia2020} focused on QoI, Koval considered model
-uncertainty and focused on reducing forward and adjoint solves, using
-randomized linear algebra.\cite{koval2020}.
+      \addplot [thin, green, mark=none] table{stdv-heat-sens4-var1.txt};
+      \addlegendentry{4 measurements};
 
-\subsection{Remedies to Sensor Clusterization}
-Several remedies to sensor clusterization were suggeted through the
-years, and we will now overview some. In \emph{clusterization-free
-design} locations are forced to be far from each other
-\cite{Ucinski05}. This is achieved in a frequentist and
-finite-dimensional setting, by either imposing distance constraints
-between observations, or by introducing correlated errors which
-account for both observation error and model error
-\cite{Ucinski05}. Merging close measurements is also possible
-\cite{fedorov2012}. Sensor clusterization is mitigated in time-series
-analysis for pharmacokinetic experiments by modeling auto-correlation
-time in the noise \cite{hooker2009}, and this approach is similar in
-spirit to imposing error correlations as in \cite{Ucinski05}.
+      \addplot [thin, purple, mark=none] table{stdv-heat-sens5-var1.txt};
+      \addlegendentry{5 measurements};
 
-Many authors bypass the sensor clusterization problem by restricting
-measurements to a coarse finite grid \cite{koval2020,
-  AlexanderianPetraStadlerEtAl16, alexanderian2021, attia2020}, but
-this practice has two flaws: First, using a coarse grid requires
-solving a hard combinatorial optimization problem. Second, restricting
-sensors to a such a small subset of the computatoinal domain will
-necessarily give rise to sub-optimal designs. Several remedies were
-offered for the first problem: Stadler suggested ell0
-sparsification. Others suggested other stuff. None were offered for
-the second.
+      \addplot [thin, cyan, mark=none] table{stdv-heat-sens6-var1.txt};
+      \addlegendentry{6 measurements};
 
-%% Today, including model errors are a bona-fide????  standard in the
-%% inverse problems community: \cite{attia2020}, \cite{koval2020},
-%% \cite{Alexanderian2021} all empower their models.  Both studies
-%% present numerical techniques for finding optimal designs when model
-%% error is present. Both are restricted to linear inverse problems
-%% (although in the latter the authors use their method on a nonlinear
-%% problem by taking a Laplace approximation for the posterior). Both
-%% find an optimal design by first solving a continuous problem for
-%% sensor weights. Said solution is then sparsified to give a binary
-%% design. Both studies are successful in the task of Bayesian
-%% inversion. Neither \cite{attia2020}, nor \cite{koval2020} mention any
-%% effect model errors can have on sensor clusterization.
+      %% \addplot [black,  only marks, mark=x, mark size=1.5] 
+      %% table{locs-heat-sens1-var1.txt}; 
+      %% \addplot [blue,   only marks, mark=x, mark size=1.5]
+      %% table{locs-heat-sens2-var1.txt}; 
+      %% \addplot [red,    only marks, mark=x, mark size=1.5]
+      %% table{locs-heat-sens3-var1.txt};
+      \addplot [green,  only marks, mark=*, mark size=1.5] 
+      table{locs-heat-sens4-var1.txt}; 
+      \addplot [purple, only marks, mark=*, mark size=1.5] 
+      table{locs-heat-sens5-var1.txt}; 
+      \addplot [cyan,   only marks, mark=*, mark size=1.5] 
+      table{locs-heat-sens6-var1.txt};
+
+    \end{axis}
+  \end{tikzpicture}
+  \caption{Measurement clusterization when inferring the initial
+    condition of the 1D heat equation. Posterior pointwise standard
+    deviations (lines) are plotted over the computational domain
+    $\Omega = [0, \pi]$, for varying numbers of
+    measurements. Measurement locations (circles) were chosen to
+    according to the D-optimality criterion. Measurement
+    clusterization occurs for six measurements: Only four measurement
+    locations are visible, since two pairs of measurements are so
+    close together they are indistinguishable.}
+  \label{fig:clusterization illustration}
+\end{figure}
+
+Researchers widely agree that measurement clusterization is undesirable
+\cite{fedorov1996, hooker2009, fedorov2012, Ucinski05,
+  neitzel2019sparse}, prompting the exploration of various remedies to
+address this issue. One approach involves merging ``close''
+measurements \cite{fedorov2012}; however, this strategy merely
+overlooks the phenomenon of measurement clusterization. An alternative
+solution lies in \emph{clusterization-free design}s, where measurement
+locations are deliberately chosen to be distant from one another. This
+can be achieved by imposing distance constraints between measurements
+or by introducing correlated errors that account for both observation
+error and model misspecification \cite{Ucinski05}. For instance, in
+the context of time-series analysis for pharmacokinetic experiments,
+measurement clusterization can be mitigated by incorporating the modeling
+of auto-correlation time within the noise terms
+\cite{hooker2009}. %% This methodology shares a similar rationale with
+%% the earlier approach, which assumed correlation between errors
+%% \cite{Ucinski05}.
+
+In spatial problems involving choice of measuremnts within a domain
+$\Omega \subseteq \mathbb{R}^d, d=1,2,3$, many researchers circumvent
+the problem of measurement clusterization by choosing measurements from a
+coarse grid in $\Omega$ \cite{koval2020, alexanderian2021, attia2020,
+  alexanderian2014, alexanderian2016,
+  alexanderian2018efficient}. However, this approach incurs a
+significant computational cost as it requires solving a difficult
+combinatorial optimization problem to select optimal measurement
+locations from a finite set. Typically, the optimization problem is
+relaxed by first assigning optimal measurement "weights" in
+$\mathbb{R}_+$ to potential measurement locations. Some authors
+incorporate a sparsifying $\ell_1$ penalty term into the design
+criterion, which is subsequently thresholded to achieve the desired
+binary design \cite{horesh2008borehole}. Others progressively relax
+the $\ell_1$ penalty to an $\ell_0$ penalty via a continuation method
+\cite{alexanderian2016, alexanderian2014}. A clever alternative is to
+cast the problem of finding optimal measurement weights as a stochastic
+optimization problem \cite{attia2022stochastic}. All of the
+aforementioned methods may indeed find a binary optimal design
+restricted to a given coarse grid. However, none addresses one
+fundamental issue: the restriction of measurement locations to a coarse
+grid in $\Omega \subseteq \mathbb{R}^d, d=1,2,3$ inherently results in
+a sub-optimal design.
+
+Avoiding measurement clusterization is a pragmatic approach:
+intuitively, we recognize that measurement clusterization is
+undesirable, even though the underlying reasons may not be fully
+clear. Consequently, we strive to prevent it and devise various
+methodologies to avoid measurement clusterization. Yet each and every
+one of these methodologies achieves this objective by imposing
+restrictions on measurement locations, thereby fundamentally altering
+the optimal design problem. Surprisingly, to the best of my knowledge,
+no previous study has tried to address some seemingly simple yet
+fundamental questions:
+%
+Why does measurement clusterization occur?
+%
+Why does imposing correlations between observations alleviate
+measurement clusterization?
+%
+Should we aim to avoid measurement clusterization?
+%
+Can an optimal clustered design be relpaced with an equally optimal
+non-clustered design?design?
 
 \subsection{Contribution}
-Any of the above mentioned approaches might serve as a remedy and push
-sensors away from each other. Yet, none offers any insight as to why
-clusterization occurs. Also, as better models are employed, model
-error is decreased and the clusterization phenomenon will eventually
-reappear. While these approaches are practical and help practitioners
-avoid sensor clusterization, they do not provide insight as to why
-sensors are clustering.
+The primary objective of this study is to provide a comprehensive
+understanding of measurement clusterization by addressing the
+aforementioned questions. Our focus centers around investigating the
+Bayesian D-optimality criterion, which involves maximizing the
+expected Kullback-Leibler divergence between the posterior and prior
+measures \cite{CoverThomas91, Chaloner1995}. To delve deeper into this
+subject, we conduct an analysis of Bayesian D-optimal designs within
+the context of linear inverse problems in Hilbert spaces. We propose a
+novel relaxed model for D-optimality that maintains analytical
+tractability and enables the identification of D-optimal designs using
+Lagrange multipliers. This analytical framework facilitates the
+exploration of the following questions posed at the end of the
+previous paragraph:
 
-The current study aims to fill the gap of understanding sensor
-clusterization in infinte dimensional Bayesian linear inverse
-problems. We propose and study a relaxed and analytically tractable
-model for D-optimal designs. Under this model, D-optimal designs are
-solutions of a constrained optimization problem, formulated using
-Lagrange multipliers. The Lagrange multipliers formulation allows us
-to rigorously show how model error mitigates clusterization (Section
-\ref{section:non vanishing}), and why designs that exhibit
-clusterization are indeed optimal when no model error is considered.
+\begin{enumerate}
+\item \label{q:why} \textbf{Why does measurement clusterization
+  occur?} In Section \ref{section:vanishing}, we provide an insightful
+  explanation for the optimality of clustered designs when no model
+  error is present. We demonstrate that measurement clusterization
+  poses no obstruction to D-optimality and that clustered design are
+  preferred by numerical algorithms.
 
-As a byproduct, the Lagrange multipliers problem for no model error
-results in a characterization of D-optimal designs. The key idea is
-that a D-optimal design reduces uncertainties where they are highest
-first. This is the conclusion of part (5) of Theorem \ref{thm:char},
-which we state below and prove in Section \ref{section:vanishing}.
+  %% We conjecture that the prevalence of measurement clusterization
+  %% arises due to the ease of discovering clustered designs.
+
+\item \label{q:mitigate} \textbf{Why does imposing correlations
+  between observations alleviate measurement clusterization?} In
+  Section \ref{section:non vanishing}, we rigorously demonstrate the
+  role of model error in mitigating clusterization, thereby
+  corroborating earlier observations made by various researchers.
+
+\item \label{q:avoid} \textbf{Should we aim to avoid measurement
+  clusterization?} Based on the analysis conducted in this study, we
+  did not find any compelling reason to explicitly avoid optimal
+  clustered designs.
+
+\item \label{q:replace} \textbf{Can an optimal clustered design be
+  relpaced with an equally optimal non-clustered design?} In Section
+  \ref{section:vanishing}, we answer this question in the affirmative,
+  although we show that numerical experiments conducted using our
+  model indicate a strong preference for clustered designs.
+\end{enumerate}
+
+The cornerstone of our investigation into measurement clusterization is
+Theorem \ref{thm:char}, presented below and proven in Section
+\ref{section:vanishing}. The key insight of Theorem \ref{thm:char}
+lies in its fifth component, which highlights that a D-optimal design
+aims to uniformly reduce posterior uncertainties across all posterior
+covariance eigenvectors.
 
 \begin{restatable}[D-optimal designs with vanishing model error]{theorem}{char}\label{thm:char}
-  %% \begin{theorem}[D-optimal designs with vanishing model error]\label{thm:char}
-  %% Let Notation \ref{not:setup}. Further, let
-
   Let:
@@
     where $\prcov:\hilp \to \hilp$ is the prior covariance operator.
-  \item $\obs: \hilo \to \mathbb{R}^m$ observation operator, where $m
-    \in \mathbb{N}$ is the number of observations taken. 
+  \item $\obs: \hilo \to \mathbb{R}^m$ measurement operator, where $m
+    \in \mathbb{N}$ is the number of measurements taken.
   \item $\sigma^2 \in \mathbb{R}_{+}$ observation noise variance,
@@
 
-One byproduct of the proof of Theorem \ref{thm:char} is a novel lemma
-in linear algebra: We show how to decompose a symmetric positive
-definite matrix $M$ as $M = AA^t$, where $A$ has unit norm columns
-(Lemma \ref{lemma:lax}). We also generalize two lemmas from linear
-algebra to infinite-dimensional settings: In Lemma \ref{lemma:MDL} we
-prove a matrix determinant lemma in Hilbert spaces, and in Lemma
-\ref{lemma:lax} we generalize a lemma \cite{Lax07} for calculating
+Other contributions include three useful lemmas: In Lemma
+\ref{lemma:free} we decompose a symmetric positive definite matrix $M
+\in \mathbb{R}^{k \times k}$ with $\ttr M = m \in \mathbb{N}$ as $M =
+AA^t$, where $A$ has unit norm columns. In Lemma \ref{lemma:MDL} we
+prove a matrix determinant lemma in Hilbert spaces. Finally in Lemma
+\ref{lemma:lax} we generalize a lemma for calculating
 $\frac{\der}{\der t} \log \det (I + X(t))$, where $X(t)$ is an
-operator valued function.
+operator valued function \cite{Lax07}.
 
 \subsection{Limitations}\label{subsec:limitations}
-The study presented here is not free from drawbacks. First, our
-relaxed model does not correspond to any real-life problem. It is
-generic enough to be analytically tractable, but one may argue it is
-to far removed from any real application. To these claims I would
-argue that Mathematicians have a long history of studying models that
-are bare-bones simplifications of real systems, e.g. the Ising model
-\cite{cipra1987} the Lorenz system \cite{brin}, and the SIR model
-\cite{keeling2008}, to name a few. The second drawback of our model is
-that we do not show rigorously that clusterization necessarily occurs,
-we only show that it does not pose any obstruction to D-optimality.
+The main limitation of this study is that our generic model does not
+correspond to any specific real-life problem. It is generic enough to
+be analytically tractable, but one may argue it is too far removed
+from any real application. To these claims I would argue that
+scientists have a long history of studying models that are bare-bones
+simplifications of real systems, e.g. the Ising model
+\cite{cipra1987}, the Lorenz system \cite{brin}, the Lotka-Volterra
+equations \cite{logan2006}, the Carnot engine \cite{kardar2007}, and
+many others.
 
-\subsection{An Example of Clusterization}\label{subsec:example}
-\begin{figure}
-  \begin{tikzpicture}[thick, scale=1.3, every node/.style={scale=0.99}]
-    \begin{axis}
-      [
-      title={Posterior Pointwise Standard Deviations and D-Optimal Sensor Locations},  
-      xmin = 0,
-      xmax = 3.14,
-      xlabel = {$x$},
-      ylabel = posterior std,
-      ymin   = 0,
-      %compat = 1.3,
-      % ymax   = 130,
-      % ytick = \empty,
-      legend cell align=left,
-      % legend style={at={(0.45,0.2)}}
-      legend pos= outer north east 
-      ]
-      % \draw[black!30!white, thin] (50,0) -- (50,130);
-      % 
-      %% \addplot [thin, black, mark=none] table{stdv-heat-sens1-var1.txt};
-      %% \addlegendentry{1 sensors};
-
-      %% \addplot [thin, blue, mark=none] table{stdv-heat-sens2-var1.txt};
-      %% \addlegendentry{2 sensors};
-
-      %% \addplot [thin, red, mark=none] table{stdv-heat-sens3-var1.txt};
-      %% \addlegendentry{3 sensors};
-
-      \addplot [thin, green, mark=none] table{stdv-heat-sens4-var1.txt};
-      \addlegendentry{4 sensors};
-
-      \addplot [thin, purple, mark=none] table{stdv-heat-sens5-var1.txt};
-      \addlegendentry{5 sensors};
-
-      \addplot [thin, cyan, mark=none] table{stdv-heat-sens6-var1.txt};
-      \addlegendentry{6 sensors};
-
-      %% \addplot [black,  only marks, mark=x, mark size=1.5] 
-      %% table{locs-heat-sens1-var1.txt}; 
-      %% \addplot [blue,   only marks, mark=x, mark size=1.5]
-      %% table{locs-heat-sens2-var1.txt}; 
-      %% \addplot [red,    only marks, mark=x, mark size=1.5]
-      %% table{locs-heat-sens3-var1.txt};
-      \addplot [green,  only marks, mark=*, mark size=1.5] 
-      table{locs-heat-sens4-var1.txt}; 
-      \addplot [purple, only marks, mark=*, mark size=1.5] 
-      table{locs-heat-sens5-var1.txt}; 
-      \addplot [cyan,   only marks, mark=*, mark size=1.5] 
-      table{locs-heat-sens6-var1.txt};
-
-    \end{axis}
-  \end{tikzpicture}
-  \caption{Sensor clusterization for inferring the initial condition
-    of the 1D heat equation. Posterior pointwise standard deviations
-    (lines) are plotted over the domain $[0, \pi]$, for varying
-    numbers of sensors. Sensor locations (circles) were chosen to
-    optimize the D-optimality criterion. Sensor clusterization occurs
-    for six sensors. Only four observation locations are used --- two
-    pairs of sensors are so close together they are
-    indistinguishable.}
-  \label{fig:clusterization illustration}
-\end{figure}
-
-For the purpose of illustration, consider a toy inverse problem ---
-inferring the initial condition of the 1D heat equation in $[0,\pi]$
-with a homogeneous Dirichlet boundary condition. Sensor clusterization
-is clearly evident, see Figure \ref{fig:clusterization}.
-
-The 1D heat equation with homogeneous Dirichlet boundary condition is:
-\begin{subequations}\label{eq:heat equation}
-  \begin{alignat}{2}
-    u_t &= \Delta u &&\qquad \text{in } [0,\pi] \times [0,\infty),\\
-      u &= 0 &&\qquad \text{on } \{0, \pi\} \times [0,\infty),\\
-        u &= u_0 &&\qquad \text{on }[0,\pi] \times \{0\}.
-  \end{alignat}
-\end{subequations}
-
-The goal is to infer the initial condition $u_0$. Model the initial
-condition as $u_0 \sim \normal(0,\prcov)$, for $\prcov =
-(-\Delta)^{-1}$ with a homogeneous Dirichlet boundary condition. Fix
-some final time $T$ and denote by $\fwd$ the forward operator, so that
-$u(\cdot, T) = \fwd u_0$. Let $\obs$ the observation operator, so that
-$u(\x_j,T) = (\obs u)_j, j=1,\dots,m$. Assume iid centered Gaussian
-observation error, so observations are given by $u(\x_j,T) +
-\eps(\x_j)$ with $\eps(\x_j) \sim \normal(0, \sigma^2)$ iid, $\sigma^2
-> 0$. The posterior covariance is known and depends only on $\prcov,
-\fwd, \obs$ and $\sigma^2$ \cite{Stuart10} (see Section
-\ref{section:prelim} and \eqref{eq:postcov} below). Finally, choose
-$\x_j,j=1,\dots,m$ to optimize the D-optimality criterion
-\cite{AlexanderianGloorGhattas14}.
-
-Sensor clusterization is illustrated in Figure \ref{fig:clusterization
-  illustration}. The posterior covariance does not depend on data, so
-the plot has no reference to actual data observed. The posterior
-covariance does depend on location of the observation taken, which
-were chosen to optimize the D-optimality criterion. Sensor
-clusterization is clearly seen for six sensors; it looks like only
-four observations were taken. Two pairs of sensors are clustered so
-close together that they are indistinguishable from each other.
-
-%% A D-optimal design
-%% is found via a generalization of the D-optimality criterion to
-%% infinite dimension(also, Section
-%% \ref{subsec:D optimal design} below).
+%% \subsection{An Example of Measurement Clusterization}\label{subsec:example}

### 2023-05-05 latex/paper.tex
Tryin to fix refs
Moved example to appendix, added questions and answers in intro.

@@
 %\documentclass{amsart}
+%\documentclass{article}
 \input{definitions}
@@
 
-\title{Sensor Clusterization in D-optimal Designs for Bayesian Linear
+\title{Measurement Clusterization in D-optimal Designs for Bayesian Linear
   Inverse Problems}
@@
 \begin{abstract}
-  We investigate the problem of sensor clusterization in D-optimal
+  We investigate the problem of measurement clusterization in D-optimal
   experimental design for infinite-dimensional Bayesian linear inverse
@@
   correlations mitigateWe suggest an analytically tractable model for D-optimal
-  designs. We argue that sensor clusterization in the case of iid
+  designs. We argue that measurement clusterization in the case of iid
   observation noise is generic. We also prove that a model error term
@@
 \bibliographystyle{siamplain}
+%% %%\bibliographystyle{amsplain}
 \bibliography{/home/yair/projects/library.bib}

### 2023-05-05 latex/prelims.tex
Moved example to appendix, added questions and answers in intro.

@@
 Let $\hilp$ and $\hilo$ be separable Hilbert spaces (the subscripts p
-and o are for ``parameter'' and ``observation'', respectfully). Denote
+and o are for ``parameter'' and ``observation'', respectively). Denote
 $\hilo^* \simeq \hilo$ the Hilbert space of all linear functionals on
@@
 forward operator $\fwd$ is assumed linear and strongly smoothing (has
-fast decaying modes --- the heat operator from section
-\ref{subsec:example} is a prime example). Take a Gaussian prior
-$\param \sim \pr = \normal(\prmean ,\prcov)$ with some appropriate
-covariance operator $\prcov$ \cite{Stuart10} on $\hilp$. Note that
-$\fwd \prcov \fwd^*$ is the prior covariance in $\hilo$
-\cite{Stuart10}. As such, $\fwd \prcov \fwd^*$ is assumed invertible
---- an assumption which will be used below. Let $m$ the number of
-observations taken. Observations are taken via the \emph{observation
-operator}\footnote{It is common for the observation and forward
+fast decaying modes --- the heat operator of the example in Section
+\ref{section:intro} is a prime example). Take a Gaussian prior $\param
+\sim \pr = \normal(\prmean ,\prcov)$ with some appropriate covariance
+operator $\prcov$ on $\hilp$ \cite{Stuart10}. Note that $\fwd \prcov
+\fwd^*$ is the prior covariance in $\hilo$ \cite{Stuart10}. As such,
+$\fwd \prcov \fwd^*$ is assumed invertible --- an assumption which
+will be used below. Let $m$ the number of measurements
+taken. Measurements are taken via the \emph{measurement
+operator}\footnote{It is common for the measurement and forward
 operators to be merged $\tmp := \obs \fwd$
@@
 following sections requires $\fwd$ and $\obs$ are explicitly separated
-similarly to \cite{attia2022stochastic, cvetkovic2023choosing}.} $\obs
-\in ( \hilo^* )^m$, so that:
+as in \cite{attia2022stochastic, cvetkovic2023choosing}.} $\obs \in (
+\hilo^* )^m$, whose entries $\meas_j, j=1,\dots,m$ are called
+\emph{measurement vector}s:
 
 \begin{equation*}%\label{eq:O}
-  \obs u = (\meas_1(u), \dots, \meas_m(u) )^t \in \R^m,\ u \in \hilo,
+  \obs u = (\meas_1(u), \dots, \meas_m(u) )^t \in \R^m,\ u \in \hilo.
 \end{equation*}
@@
 %
-and we refer to $\meas_j, j=1,\dots,m$ as \emph{observation
-vector}s. For $u \in \hilo$ and $v\in \R^m$:
-%
-%
-\begin{align*}
-  \big (\obs^*v \big ) (u) &= \langle v, \obs u \rangle_{\R^m} = \sum_{j=1}^m  v_j \meas_j(u)
-  = v^t \left ( \obs u \right ) = (v^t \obs) (u),
-\end{align*}
-and thus:
-\begin{align}\label{eq:obs*}
-  \obs^*v &= \sum_{j=1}^m v_j \meas_j = v^t \obs.
-\end{align}
+
+%% For $u \in \hilo$ and $v\in \R^m$:
+%% %
+%% %
+%% \begin{align*}
+%%   \big (\obs^*v \big ) (u) &= \langle v, \obs u \rangle_{\R^m} = \sum_{j=1}^m  v_j \meas_j(u)
+%%   = v^t \left ( \obs u \right ) = (v^t \obs) (u),
+%% \end{align*}
+%% and thus:
+%% \begin{align}\label{eq:obs*}
+%%   \obs^*v &= \sum_{j=1}^m v_j \meas_j = v^t \obs.
+%% \end{align}
 
 Data is acquired via noisy observations, and we consider two types of
-error terms: Spatially correlated model error is denoted $\eps' \sim
-\normal(0,\modcov)$, and modeled as a centered Gaussian measure on
-$\hilo$ with covariance operator $\modcov$. Observation error is
-denoted $\eps \sim \normal(0, \sigma^2 I_m)$, with $I_m \in
-\mathbb{R}^{m \times m}$ the identity. Both error terms and the prior
-are assumed independent of each other. Thus, data is acquired via
+error terms: Spatially correlated model error $\eps' \sim
+\normal(0,\modcov)$ with $\modcov: \hilo \to \hilo$ a covariance
+operator. Observation error is denoted $\eps \sim \normal(0, \sigma^2
+I_m)$, with $I_m \in \mathbb{R}^{m \times m}$ the identity. Both error
+terms and the prior are assumed independent of each other. Thus, data
+is acquired via
 \begin{align}\label{eq:inverse problem}
@@
   \begin{split}
-    [\obs \modcov \obs^*]_{ij} & = e_i^t \obs \modcov \obs^* e_j \\
+    [\obs \modcov \obs^*]_{ij} & = e_i^t \obs \modcov \obs^* e_j 
     %
@@
     %
-    &= \meas_i (\modcov \meas_j) \text{ (by \eqref{eq:obs*})}.
+    = \meas_i (\modcov \meas_j).% \text{ (by \eqref{eq:obs*})}.
   \end{split}
@@
 %% consider model error (modeled by $\modcov$) as well as observation
-%% error (modeled by $\sigma^2$). For example, say we believe our forward
-%% model does not capture some small scale phenomenon.  Then we may
-%% express this belief by saying $\tru = \fwd + \err$, with $\fwd$
-%% depending on $\param$ and $\err$ depending on $\sspar$, with $\param
-%% \perp \sspar$. We do not know much about this effect but it is
-%% reasonable to assume it changes continuously in our domain. We (may
-%% choose to) model it as $\normal (0, \modcov)$ and take $\modcov$ to
-%% reflect the spatial (or other) variability we imagine $\err \sspar$
-%% has. Such small scale phenomenon can arise as a modeling issue, where
-%% we might not model the system in its entirety. It can also arise from
-%% a numerical source, where our discretization of the system is not fine
-%% enough to capture all small scale phenomena.
-%%  In section \ref{section:non vanishing}, we will see that assuming
-%% some correlation in the error mitigates the clusterization
-%% phenomenon, as reported in the literature \cite{Ucinski05}.
+%% error (modeled by $\sigma^2$). For example, say we believe our
+%% forward model does not capture some small scale phenomenon.  Then
+%% we may express this belief by saying $\tru = \fwd + \err$, with
+%% $\fwd$ depending on $\param$ and $\err$ depending on $\sspar$, with
+%% $\param \perp \sspar$. We do not know much about this effect but it
+%% is reasonable to assume it changes continuously in our domain. We
+%% (may choose to) model it as $\normal (0, \modcov)$ and take
+%% $\modcov$ to reflect the spatial (or other) variability we imagine
+%% $\err \sspar$ has. Such small scale phenomenon can arise as a
+%% modeling issue, where we might not model the system in its
+%% entirety. It can also arise from a numerical source, where our
+%% discretization of the system is not fine enough to capture all
+%% small scale phenomena.  In section \ref{section:non vanishing}, we
+%% will see that assuming some correlation in the error mitigates the
+%% clusterization phenomenon, as reported in the literature
+%% \cite{Ucinski05}.
 
-%% \subsubsection{Dynamics and Observation Operators}\label{subsec:dynamics}
-%% Rows of $\obs$ are linear functionals and we refer to these as {\it
-%%   observation vector}s. Note that we cannot choose every observation
-%% vector we want. For example, we may be restricted only to pointwise
-%% evaluations of $\fwd \param$ by the sensors at our disposal, in which
-%% case measuring, e.g., the mean $\ell(u) = \int_{\Omega}u$ of a
-%% function is not possible. Thus, it is best to think of $(\obs \fwd
-%% \param)_j$ as approximations to pointwise evaluations. For example, if
-%% $\hilo = L^2([-\pi, \pi])$, consider $\ell_1(x) := (2\pi
-%% \eps)^{-\frac{1}{2}}\exp(-\frac{x^2}{2\eps^2})$ and $\ell_2(x)
-%% :=\eps^{-1}\mathbf{1}_{[-\frac{\eps}{2}, \frac{\eps}{2}]}$. Both
-%% $\ell_1$ and $\ell_2$ approximate a point measurement at $0$ for small
-%% $\eps$ and, crucially, $\ell_1, \ell_2 \in \hilo^*$.
-%% %% It is important to note that although $\|\ell\|_1 =
-%% %% \|\ell_2\|_1=1$, unfortunately $\|\ell_1\|_2^2 = (4\pi
-%% %% \eps)^{-\frac{1}{2}}$ and $\|\ell_2\|_2^2 = \eps^{-1}$.
+%% \subsubsection{Dynamics and Observation
+%% Operators}\label{subsec:dynamics} Rows of $\obs$ are linear
+%% functionals and we refer to these as {\it observation
+%% vector}s. Note that we cannot choose every observation vector we
+%% want. For example, we may be restricted only to pointwise
+%% evaluations of $\fwd \param$ by the sensors at our disposal, in
+%% which case measuring, e.g., the mean $\ell(u) = \int_{\Omega}u$ of
+%% a function is not possible. Thus, it is best to think of $(\obs
+%% \fwd \param)_j$ as approximations to pointwise evaluations. For
+%% example, if $\hilo = L^2([-\pi, \pi])$, consider $\ell_1(x) :=
+%% (2\pi \eps)^{-\frac{1}{2}}\exp(-\frac{x^2}{2\eps^2})$ and
+%% $\ell_2(x) :=\eps^{-1}\mathbf{1}_{[-\frac{\eps}{2},
+%% \frac{\eps}{2}]}$. Both $\ell_1$ and $\ell_2$ approximate a point
+%% measurement at $0$ for small $\eps$ and, crucially, $\ell_1, \ell_2
+%% \in \hilo^*$.  %% It is important to note that although $\|\ell\|_1
+%% = %% \|\ell_2\|_1=1$, unfortunately $\|\ell_1\|_2^2 = (4\pi %%
+%% \eps)^{-\frac{1}{2}}$ and $\|\ell_2\|_2^2 = \eps^{-1}$.
 
@@
 
-\subsection{D-Optimal Designs in Infinite Dimensions}\label{subsec:D optimal design} 
-A D-optimal design maximizes expected KL divergence between posterior
-and prior measures, $\post$ and $\pr$. It is useful to first recall
-the definition of KL divergence:% for an arbitrary prior measure:
+\subsection{Bayesian D-Optimal Designs in Infinite Dimensions}\label{subsec:D optimal design} 
+A Bayesian D-optimal design maximizes expected KL divergence between
+posterior and prior measures, $\post$ and $\pr$. It is useful to first
+recall the definition of KL divergence:% for an arbitrary prior
+measure:
 $$
@@
 
-For a Bayesian linear model in finite dimensions, with Gaussian prior
-and Gaussian noise, a D-optimal design minimizes the determinant of
-the posterior covariance matrix\footnote{The frequentist version of
-the D-optimality criterion aims to minimize the determinant of the
-Fisher information matrix \cite[page 16]{Ucinski05}. Thus, the
-Bayesian criterion is just a regularized version of the frequentist
-criterion \cite{Chaloner1995}.}. Bayesian D-optimal designs for
-inverse problems on function spaces were first studied by
-\cite{AlexanderianGloorGhattas14, alexanderian2018efficient}, and just like in finite dimensions,
-a D-optimal design maximizes the Kullback-Leibler divergence between
-posterior and prior measures.
-%% For a linear inverse problem with Gaussian prior and Gaussian
-%% noise, a D-optimal design minimizes an expression analogous to the
-%% determinant of the posterior covariance operator
-%% \cite{AlexanderianGloorGhattas14} (also \cref{thm:d optimality}
-%% below).
-D-optimal designs in infinite-dimensional
-Hilbert spaces were investigated in
-\cite{AlexanderianGloorGhattas14}. Their main result is summarized, in
-the notation presented above, in the following theorem\footnote{In
+The study of D-optimal designs for Bayesian linear inverse problems in
+infinite dimensions was pioneered by \cite{AlexanderianGloorGhattas14,
+  alexanderian2018efficient}. The main result we will make use of is
+summarized (in our notation) in Theorem \ref{thm:d optimality}\footnote{In
 \cite{AlexanderianGloorGhattas14, alexanderian2018efficient}, results
@@
 covariance matrices \cite[p. 681]{AlexanderianGloorGhattas14}.}:
+
+%% , and just like in finite dimensions, a D-optimal design maximizes
+%% the Kullback-Leibler divergence between posterior and prior
+%% measures For a linear inverse problem with Gaussian prior and
+%% Gaussian noise, a D-optimal design minimizes an expression
+%% analogous to the determinant of the posterior covariance operator
+%% \cite{AlexanderianGloorGhattas14} (also \cref{thm:d optimality}
+%% below).  D-optimal designs in infinite-dimensional Hilbert spaces
+%% were investigated in \cite{AlexanderianGloorGhattas14}.
 \begin{theorem}[Slightly modified Theorem 1 from \cite{AlexanderianGloorGhattas14}]\label{thm:d optimality}
@@
 
-\noindent For an intuitive understanding of Definition \ref{def:d
-  optimality}, note that from \eqref{eq:postcov} and
-\eqref{eq:objective}:
+For a Bayesian linear model in finite dimensions, with Gaussian prior
+and Gaussian noise, a D-optimal design minimizes the determinant of
+the posterior covariance matrix\footnote{The frequentist version of
+the D-optimality criterion aims to minimize the determinant of the
+Fisher information matrix \cite[page 16]{Ucinski05}. Thus, the
+Bayesian criterion is just a regularized version of the frequentist
+criterion \cite{Chaloner1995}.}. Theorem \ref{thm:d optimality} and
+Definition \ref{def:d optimality} have a similar intuition:
 \begin{align*}
   \begin{split}
-    \tar(\obs) &= \det ( I + \prcov^{1/2}  \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}) \\
-    &= \det \Big( \prcov ( \prcov^{-1} + \fwd ^* \obs^* \Sigma^{-1} \obs \fwd) \Big )\\
-    &= \det \prcov \postcov^{-1}.
+    \tar(\obs) &= \frac12 \log \det ( I + \prcov^{1/2}  \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}) \text{ by \eqref{eq:objective}}\\
+    &= \frac12 \log \det \Big( \prcov ( \prcov^{-1} + \fwd ^* \obs^* \Sigma^{-1} \obs \fwd) \Big )\\
+    &= \frac12 \log \det \prcov \postcov^{-1} \text{ by \eqref{eq:postcov}} \\
+    &= \frac12 \log \det \prcov -\frac12 \log \det \postcov.
   \end{split}
@@
     $\hilp$, with prior covariance operator $\prcov:\hilp \to \hilp$.
-  \item $\obs: \hilo \to \mathbb{R}^m$ observation operator, where $m
-    \in \mathbb{N}$ is the number of observations taken. 
+  \item $\obs: \hilo \to \mathbb{R}^m$ measurement operator, where $m
+    \in \mathbb{N}$ is the number of measurements taken.
   \item $\sigma^2 \in \mathbb{R}_{+}$ observation noise variance.

### 2023-05-05 latex/vanishing.tex
Moved example to appendix, added questions and answers in intro.

@@
-\section{D-Optimal Designs when $\modcov = 0$}\label{section:vanishing}
+\section{Characterizing D-Optimal Designs when $\modcov = 0$}\label{section:vanishing}
 We will now prove Theorem \ref{thm:char} which caracterizes D-optimal
@@
 \begin{proposition}\label{prop:true target}
-  Let $\obs$ with $m$ observations satisfy the nonlinear eigenvalue
+  Let $\obs$ with $m$ measurements satisfy the nonlinear eigenvalue
   problem \eqref{eq:eigenproblem}. Let $\{\eta_i\}_{i=1}^{\infty}$
@@
 \begin{proof}
-  Part (1) is immediate for any observation operator $\obs$ that
-  satisfies the unit norm constraint on observation vectors. Part (2)
+  Part (1) is immediate for any measurement operator $\obs$ that
+  satisfies the unit norm constraint on measurement vectors. Part (2)
   was proved in Proposition \ref{prop:same ev}. Part (3) was proved in
@@
     sub-optimal design (right). Both designs are allowed $m=3$
-    observations. We assume $\sigma^2=1$ and thus, the blue area has
+    measurements. We assume $\sigma^2=1$ and thus, the blue area has
     accumulated height of $\sigma^{-2}m = 3$ in both plots. The
@@
 liquid of $\lambda_i^{-1}$ volume units. We are allowed $m$
-observations, so we have blue liquid of volume $\sigma^{-2}m$ units at
+measurement, so we have blue liquid of volume $\sigma^{-2}m$ units at
 our disposal. When we seek a D-optimal design, we distribute the blue
@@
 \begin{figure}%{r}{0.25\textwidth} 
-    %\centering
+    \centering
     \includegraphics[width=5cm, height=5cm]{cylinders.jpg}
@@
 
-\section{Clusterization poses no obstruction to D-optimality}\label{section:clusterization}
-The necessary conditions for D-optimality are (1), (2) and (4) of
-Theorem \ref{thm:char}. These conditions, however, still leave some
-freedom in choosing $\obs$. This freedom may give rise to
-clusterization, as illustrated in Figure \ref{fig:clusterization} and
-demonstrated below.
-
-Consider an inverse problem for which $\lambda_1^{-1} = 0.1,
-\lambda_2^{-1} = 0.2,\sigma^2=1$ and $\{\ev_i\}_{i=1}^\infty$ are
-eigenvectors of $\fwd\prcov\fwd^*$. Two D-optimal designs $\obs_1,
-\obs_2$ are presented in \eqref{eq:two designs}, and it is easy to
-verify that $\obs_1^*\obs_1 = \obs_2^*\obs_2$. However, $\obs_1$
-exhibits clusterization ($\meas_3 = \meas_5 = \ev_1$), while $\obs_2$
-does not. Both of $\obs_1$ and $\obs_2$ are D-optimal, and we conclude
-that clusterization poses no obstruction to D-optimality.
-
+\section{Clusterization poses no obstruction to D-optimality}\label{section:clusterization}T
+The necessary conditions for D-optimality of Theorem \ref{thm:char}
+leave some freedom in choosing $\obs$, which may or may not give rise
+clustered designs. Consider an inverse problem for which the first two
+eigenvalues of $\fwd \prcov\fwd^*$ are $\lambda_1^{-1} = 0.1,
+\lambda_2^{-1} = 0.2$, its eigenvectors are $\{\ev_i\}_{i=1}^\infty$
+and $\sigma^2=1$. Consider the following $\obs_1, \obs_2$ (see also
+Figure \ref{fig:clusterization}):
 \begin{equation}\label{eq:two designs}
@@
 \end{equation}
+\noindent It is easy to verify that $\obs_1^*\obs_1 = \obs_2^*\obs_2$
+and that both $\obs_1$ and $\obs_2$ are D-optimal. However, $\obs_1$
+is a clustered design ($\meas_3 = \meas_5 = \ev_1$), while $\obs_2$ is
+not.
 
-A computer implementation of Lemma \ref{thm:char} (code in
-Supplementary material) generates $\obs_1$ as a solution to the
-D-optimal design problem mentioned in the previous
-paragraph. Randomized numerical simulations (see code in Supplementary
-material) give rise to D-optimal designs that exhibit clustrization
-more than 95\% of the time. We conclude that clusterization poses no
-obstruction to D-optimality for the model presented in this study and
-that sensor clusterization is generic.
+We are now ready to answer some of the questions posed in the
+introduction:
+
+\textbf{Question \ref{q:why}}: Since $\obs_1$ is clustered and
+D-optimal, measurement clusterization poses no obstruction to
+D-optimality. Furthermore, a computer implementation of Lemma
+\ref{thm:char} also generates $\obs_1$ as a solution to the D-optimal
+design problem. Randomized numerical simulations give rise to
+D-optimal designs that exhibit clustrization more than 95\% of the
+time (see code in supplementary). Since our model is as generic as
+possible, we conclude that it is possible that measurement
+clusterization is a generic phenomenon that.
+
+\textbf{Question \ref{q:replace}}: Since $\obs_2$ is D-optimal but not
+clustered, we conclude that it is possible to substitute a clustered
+design for an equally optimal non-clustered design. Achieveing this in
+the setup of our model not hard, but in real life problems this may
+prove more difficult.
 

### 2023-05-07 latex/intro.tex
Add abstract (raw)

@@
 significance when inferring parameters of a physical process. Unlike
-many other fields, where measurements are fixed or limited, in this
-context, we have the freedom to choose which measurements to take. We
-would like to take advantage of this freedom to choose measurements
-that enhance accuracy, reduce costs, or achieve both objectives
-simultaneously. Whether measurements involves specifying electrode
+many other fields, where measurements are fixed, in this context, we
+have the freedom to choose which measurements to take. This freedom
+should be taken advantage of to choose measurements that enhance
+accuracy, reduce costs, or achieve both objectives
+simultaneously. Whether measurements involve specifying electrode
 locations on the skin in electric impedance tomography
@@
 
-Interestingly, A- and D-optimal designs have been observed to yield
+Surprisingly, A- and D-optimal designs have been observed to yield
 remarkably similar measurements in certain cases \cite{fedorov1996,
@@
   illustration} D-optimal measurement locations are shown for
-different numbers of allowed measurements. Notably, for six
-measurement locations, the D-optimal design yields two sets of
-measurements that are almost indistinguishable from one another. This
-intriguing phenomenon is known as \emph{measurement clusterization}
-\cite{Ucinski05}, and we refer to a design that exhibits measurement
-clusterization as a \emph{clustered design}.
+different numbers of measurements. Notably, for six measurements, a
+D-optimal design yields two sets of measurements that are almost
+indistinguishable from one another. This intriguing phenomenon is
+known as \emph{measurement clusterization} \cite{Ucinski05}, and we
+refer to a design that exhibits measurement clusterization as a
+\emph{clustered design}.
 
@@
   \end{tikzpicture}
-  \caption{Measurement clusterization when inferring the initial
-    condition of the 1D heat equation. Posterior pointwise standard
-    deviations (lines) are plotted over the computational domain
-    $\Omega = [0, \pi]$, for varying numbers of
-    measurements. Measurement locations (circles) were chosen to
+  \caption{Measurement clusterization for the inverse problem of
+    inferring the initial condition of the 1D heat equation. Posterior
+    pointwise standard deviations (lines) are plotted over the
+    computational domain $\Omega = [0, \pi]$, for varying numbers of
+    measurements. Measurement locations (circles) were chosen
     according to the D-optimality criterion. Measurement
@@
 
-Researchers widely agree that measurement clusterization is undesirable
-\cite{fedorov1996, hooker2009, fedorov2012, Ucinski05,
+Researchers widely agree that measurement clusterization is
+undesirable \cite{fedorov1996, hooker2009, fedorov2012, Ucinski05,
   neitzel2019sparse}, prompting the exploration of various remedies to
@@
 the context of time-series analysis for pharmacokinetic experiments,
-measurement clusterization can be mitigated by incorporating the modeling
-of auto-correlation time within the noise terms
-\cite{hooker2009}. %% This methodology shares a similar rationale with
-%% the earlier approach, which assumed correlation between errors
-%% \cite{Ucinski05}.
+measurement clusterization can be mitigated by incorporating the
+modeling of auto-correlation time within the noise terms
+\cite{hooker2009}.
 
@@
 $\Omega \subseteq \mathbb{R}^d, d=1,2,3$, many researchers circumvent
-the problem of measurement clusterization by choosing measurements from a
-coarse grid in $\Omega$ \cite{koval2020, alexanderian2021, attia2020,
-  alexanderian2014, alexanderian2016,
+the problem of measurement clusterization by choosing measurements
+from a coarse grid in $\Omega$ \cite{koval2020, alexanderian2021,
+  attia2020, alexanderian2014, alexanderian2016,
   alexanderian2018efficient}. However, this approach incurs a
 significant computational cost as it requires solving a difficult
-combinatorial optimization problem to select optimal measurement
-locations from a finite set. Typically, the optimization problem is
-relaxed by first assigning optimal measurement "weights" in
-$\mathbb{R}_+$ to potential measurement locations. Some authors
-incorporate a sparsifying $\ell_1$ penalty term into the design
-criterion, which is subsequently thresholded to achieve the desired
-binary design \cite{horesh2008borehole}. Others progressively relax
-the $\ell_1$ penalty to an $\ell_0$ penalty via a continuation method
+combinatorial optimization problem for choosing measurementst from a
+finite set. The combinatorial optimization problem is usually relaxed
+by first assigning optimal measurement "weights" in $\mathbb{R}_+$ to
+the potential measurement locations. Some researchers incorporate a
+sparsifying $\ell_1$ penalty term into the design criterion, which is
+subsequently thresholded to achieve the desired binary design over the
+coarse grid \cite{horesh2008borehole}. Others progressively relax the
+$\ell_1$ penalty to an $\ell_0$ penalty via a continuation method
 \cite{alexanderian2016, alexanderian2014}. A clever alternative is to
-cast the problem of finding optimal measurement weights as a stochastic
-optimization problem \cite{attia2022stochastic}. All of the
+cast the problem of finding optimal measurement weights as a
+stochastic optimization problem \cite{attia2022stochastic}. All of the
 aforementioned methods may indeed find a binary optimal design
 restricted to a given coarse grid. However, none addresses one
-fundamental issue: the restriction of measurement locations to a coarse
-grid in $\Omega \subseteq \mathbb{R}^d, d=1,2,3$ inherently results in
-a sub-optimal design.
+fundamental issue: the restriction of measurement locations to a
+coarse grid in $\Omega \subseteq \mathbb{R}^d, d=1,2,3$ inherently
+results in a sub-optimal design.
 
@@
 restrictions on measurement locations, thereby fundamentally altering
-the optimal design problem. Surprisingly, to the best of my knowledge,
-no previous study has tried to address some seemingly simple yet
-fundamental questions:
+the optimal design problem. To the best of my knowledge, no previous
+study has tried to address some seemingly simple yet fundamental
+questions:
 %
@@
 %
-Can an optimal clustered design be relpaced with an equally optimal
-non-clustered design?design?
+Is it possible to substitute an optimal clustered design with an
+equally optimal non-clustered design?
+%Can an optimal clustered design be relpaced with an equally optimal
+%non-clustered design?
 
@@
 expected Kullback-Leibler divergence between the posterior and prior
-measures \cite{CoverThomas91, Chaloner1995}. To delve deeper into this
-subject, we conduct an analysis of Bayesian D-optimal designs within
-the context of linear inverse problems in Hilbert spaces. We propose a
-novel relaxed model for D-optimality that maintains analytical
-tractability and enables the identification of D-optimal designs using
-Lagrange multipliers. This analytical framework facilitates the
-exploration of the following questions posed at the end of the
-previous paragraph:
+measures \cite{CoverThomas91, Chaloner1995}. We conduct an analysis of
+Bayesian D-optimal designs within the context of linear inverse
+problems in Hilbert spaces. We propose a novel relaxed model for
+D-optimality that maintains analytical tractability and enables the
+identification of D-optimal designs using Lagrange multipliers. This
+analytical framework facilitates the exploration of the questions
+posed at the end of the previous paragraph:
 
@@
 
-\item \label{q:replace} \textbf{Can an optimal clustered design be
-  relpaced with an equally optimal non-clustered design?} In Section
-  \ref{section:vanishing}, we answer this question in the affirmative,
-  although we show that numerical experiments conducted using our
-  model indicate a strong preference for clustered designs.
+\item \label{q:replace} \textbf{Is it possible to substitute an
+  optimal clustered design with an equally optimal non-clustered
+  design?} In Section \ref{section:vanishing}, we answer this question
+  in the affirmative, although we show that numerical experiments
+  conducted using our model indicate a strong preference for clustered
+  designs.
 \end{enumerate}
 
-The cornerstone of our investigation into measurement clusterization is
-Theorem \ref{thm:char}, presented below and proven in Section
+The cornerstone of our investigation into measurement clusterization
+is Theorem \ref{thm:char} proven in Section
 \ref{section:vanishing}. The key insight of Theorem \ref{thm:char}
@@
 aims to uniformly reduce posterior uncertainties across all posterior
-covariance eigenvectors.
+covariance eigenvectors. A similar conclusion was reached by Koval et
+al.~\cite{koval2020}, who showed that A-optimal designs should be
+constructed in the space of \emph{observations}.
 

### 2023-05-07 latex/paper.tex
Add abstract (raw)

@@
 \begin{abstract}
-  We investigate the problem of measurement clusterization in D-optimal
-  experimental design for infinite-dimensional Bayesian linear inverse
-  problems. Many authors seem to believe that the fact that spatial
-  correlations mitigateWe suggest an analytically tractable model for D-optimal
-  designs. We argue that measurement clusterization in the case of iid
-  observation noise is generic. We also prove that a model error term
-  mitigates the clusterization phenomenon.
-
-  %% As a part of the analysis we prove a Matrix Determinant
-  %% Lemma in Hilbert spaces and a lemma for calculating
-  %% $\frac{\der}{\der t} \log \det (I + X(t))$ for certain
-  %% operator-valued functions. We also show how to decompose $M = AA^t$
-  %% with $A$ constrained to have unit norm columns.
+  \textbf{Setup} Taking measurements for inversion for physical
+  processes requires expesive measurements.  \textbf{Problem}
+  Unfortunately, when optimizing measurement locations, measurements
+  get clustered.  \textbf{Why care} This a imortatn \textbf{What
+    others do} Many researchers tried to avoind measuremnt
+  clusterization, typically by modifying the optimization problem
+  \textbf{What we do} We consider an analytically tractable model for
+  Bayesian D-optimal design over Hilbert spaces. We prove that
+  measurement clusterization poses no obstruction to D-optimality. We
+  demonstrate that measuremnt clusterization is a generic property of
+  D-optimal designs, and prove that measurement correlation mitigates
+  the problem of measurement clusterization. We also give a full
+  characterization of D-optimal designs, which, surprisingly, shows
+  that D-optimal desings reduce uncertainty in the space of
+  observations.  \textbf{Test on data} We test our conclusions via
+  numeric implementation of our model
 \end{abstract}

### 2023-05-07 latex/prelims.tex
Add abstract (raw)

@@
 \cite{AlexanderianGloorGhattas14}, but the analysis carried out in the
-following sections requires $\fwd$ and $\obs$ are explicitly separated
-as in \cite{attia2022stochastic, cvetkovic2023choosing}.} $\obs \in (
-\hilo^* )^m$, whose entries $\meas_j, j=1,\dots,m$ are called
-\emph{measurement vector}s:
+following sections requires that $\fwd$ and $\obs$ are explicitly
+separated as in \cite{attia2022stochastic, cvetkovic2023choosing}.}
+$\obs \in ( \hilo^* )^m$, whose entries $\meas_j, j=1,\dots,m$ are
+called \emph{measurement vector}s:
 
@@
 
-The posterior measure $\post$ is Gaussian in this setting and its
-covariance operator is:
+It is useful to record that the posterior measure $\post$ is Gaussian
+in this setting and its covariance operator is \cite{Stuart10}:
 \begin{align}\label{eq:postcov}

### 2023-05-07 latex/vanishing.tex
Add abstract (raw)

@@
 
-We are now ready to answer some of the questions posed in the
+We are now ready to answer two of the questions posed in the
 introduction:
@@
     $\meas_1$ and $\meas_2$ is denoted $\meas_1\meas_2$. These
-    contributions are plotted together because $\meas_1$ and $\meas_2$
-    are not in the direction of any eigenvector. Similarly for
-    $\meas_3\meas_4$.}
+    contributions are plotted together because neither $\meas_1$ nor
+    $\meas_2$ are not in the direction of any eigenvector. Similarly
+    for $\meas_3\meas_4$.}
   \label{fig:clusterization}

### 2023-05-10 zeros.py
Generic into zeros

@@ -11,11 +11,13 @@
 # AA^t = M.
 
 import numpy as np
+import pandas as pd
 import pytest
 from joblib import delayed, Parallel
 from pandas import isnull
 from numpy import sin, cos
 from scipy.stats import ortho_group
+from scipy.spatial.distance import cdist
 
 EPS = 1e-9
 ortho = ortho_group.rvs
@@ -234,8 +236,75 @@ def test_givens(k=8, lower=4, upper=6):
 def test_all(params):
     Parallel(n_jobs=7)(delayed(caller)(**param) for param in params)
 
+
+def generic():
+    """ We run randomized simulations of D-optimal design with our relaxed model.
+    """
     
+    N = 1000
+    res = []
+    for m in range(6, 9):
+        for k in range(3, m-1):
+
+            ## Counts how many simulated designs are clustered
+            counter = 0 
+
+            # N == number of simulations
+            for i in range(N):
+
+                ## Random matrices such that:
+                ## M is symmetric positive definite
+                ## M = UDU^*
+                ## S = sqrd(D)
+                ## m = number of measurements
+                ## Number of nonzero entries in D
+                M, D, U, S = MDUS(m, k)
+
+                ## AA^t = M, and A has unit norm columns
+                A = get_A(D, m)
+
+                ## distances between columns of A
+                distances = cdist(A.T, A.T)
+
+                ## Diagonal has 0 entries. We do not want to count
+                ## them, so we fill diagonal with 1's
+                np.fill_diagonal(distances, 1)
+
+                ## Which distances are > 0 <==> which pairs of
+                ## measurements are ***not*** clustered
+                dis = distances > 1e-9
+
+                ## If all measurements are not clustered --- the
+                ## design does not exhibit sensor clusterization.
+                if dis.all():
+
+                    ## Increment counter of non clustered designs
+                    counter += 1
+
+                dic = dict(m=m, ## Number of measurements
+                           k=k, ## Rank of O
+                           cluster=1-counter/N ## Fraction of designs that ***are** clustered
+                           )
+                
+            res.append(dic)
+    res = pd.DataFrame(res)
+    res = res.assign(mk=res.m-res.k) 
+    # res.to_pickle("simulation.pickle") ## Save this if u take a long run
+    print("Parameters such that the probability of clusterization is < 1:\n", res.query("cluster < 1"))
+
+
 if __name__ == '__main__':
-    test_givens()
[…]

### 2023-05-10 latex/intro.tex
Improving

@@
 have the freedom to choose which measurements to take. This freedom
-should be taken advantage of to choose measurements that enhance
-accuracy, reduce costs, or achieve both objectives
-simultaneously. Whether measurements involve specifying electrode
-locations on the skin in electric impedance tomography
-\cite{horesh2010impedance}, determining certain wavelengths in MRI
-\cite{horesh2008mri}, or positioning sensors for detecting
+should be taken advantage of: measurements should be chosen to enhance
+accuracy, reduce costs, or both. Whether measurements involve
+specifying electrode locations on the skin in electric impedance
+tomography \cite{horesh2010impedance}, determining certain wavelengths
+in MRI \cite{horesh2008mri}, or positioning sensors for detecting
 ground-reflected waves in the search for oil
@@
 referred to as the problem of \emph{optimal design}, becomes
-crucial. To achieve this, researchers optimize specific \emph{design
-criteria}, with A- and D-optimality being two of the most widely
-recognized and extensively studied design criteria
+crucial. Optimal measurements are chosen according to specific
+\emph{design criteria}, with A- and D-optimality being two of the most
+widely recognized and extensively studied design criteria
 \cite{Chaloner1995}.
@@
 D-optimal design yields two sets of measurements that are almost
-indistinguishable from one another. This intriguing phenomenon is
-known as \emph{measurement clusterization} \cite{Ucinski05}, and we
-refer to a design that exhibits measurement clusterization as a
+indistinguishable from one another. We refer to this intriguing
+phenomenon as \emph{measurement clusterization} \cite{Ucinski05}, and
+we refer to a design that exhibits measurement clusterization as a
 \emph{clustered design}.
@@
   attia2020, alexanderian2014, alexanderian2016,
-  alexanderian2018efficient}. However, this approach incurs a
-significant computational cost as it requires solving a difficult
-combinatorial optimization problem for choosing measurementst from a
-finite set. The combinatorial optimization problem is usually relaxed
-by first assigning optimal measurement "weights" in $\mathbb{R}_+$ to
-the potential measurement locations. Some researchers incorporate a
+  alexanderian2018efficient}. This approach incurs a significant
+computational cost as it requires solving a difficult combinatorial
+optimization problem for measurements locations over a finite set. The
+combinatorial optimization problem is usually relaxed by first
+assigning optimal measurement ``weights'' in $\mathbb{R}_+$ to the
+potential measurement locations. Some researchers incorporate a
 sparsifying $\ell_1$ penalty term into the design criterion, which is
@@
 fundamental issue: the restriction of measurement locations to a
-coarse grid in $\Omega \subseteq \mathbb{R}^d, d=1,2,3$ inherently
-results in a sub-optimal design.
+coarse grid in $\Omega$ inherently results in a sub-optimal design.
 
@@
 methodologies to avoid measurement clusterization. Yet each and every
-one of these methodologies achieves this objective by imposing
+one of these methodologies achieves its objective by imposing
 restrictions on measurement locations, thereby fundamentally altering
@@
 Bayesian D-optimal designs within the context of linear inverse
-problems in Hilbert spaces. We propose a novel relaxed model for
+problems over Hilbert spaces. We propose a novel relaxed model for
 D-optimality that maintains analytical tractability and enables the
@@
   error is present. We demonstrate that measurement clusterization
-  poses no obstruction to D-optimality and that clustered design are
+  poses no obstruction to D-optimality and that clustered designs are
   preferred by numerical algorithms.
@@
 covariance eigenvectors. A similar conclusion was reached by Koval et
-al.~\cite{koval2020}, who showed that A-optimal designs should be
-constructed in the space of \emph{observations}.
+al.~\cite{koval2020}, who showed that A-optimal designs are best
+constructed in the space of observations.
 
@@
 AA^t$, where $A$ has unit norm columns. In Lemma \ref{lemma:MDL} we
-prove a matrix determinant lemma in Hilbert spaces. Finally in Lemma
+prove a matrix determinant lemma in Hilbert spaces. Finally, in Lemma
 \ref{lemma:lax} we generalize a lemma for calculating

### 2023-05-10 latex/paper.tex
Improving

--- 2023-05-05 latex/paper.tex
@@
 \title{Measurement Clusterization in D-optimal Designs for Bayesian Linear
-  Inverse Problems}
+  Inverse Problems over Hilbert Spaces}
 
@@
 \begin{abstract}
-  We investigate the problem of measurement clusterization in D-optimal
-  experimental design for infinite-dimensional Bayesian linear inverse
-  problems. Many authors seem to believe that the fact that spatial
-  correlations mitigateWe suggest an analytically tractable model for D-optimal
-  designs. We argue that measurement clusterization in the case of iid
-  observation noise is generic. We also prove that a model error term
-  mitigates the clusterization phenomenon.
-
-  %% As a part of the analysis we prove a Matrix Determinant
-  %% Lemma in Hilbert spaces and a lemma for calculating
-  %% $\frac{\der}{\der t} \log \det (I + X(t))$ for certain
-  %% operator-valued functions. We also show how to decompose $M = AA^t$
-  %% with $A$ constrained to have unit norm columns.
+  Inversion for parameteres of physical processes typically requires
+  taking expensive measurements. Finding an optimal set of
+  measurements is known as the optimal design problem. Surprisingly,
+  measurement locations in optimal design are sometimes extremely
+  clustered. Practically and intuitively, measurement clusterization
+  is a huge problem. Researchers usually avoid measurement
+  clusterization by somehow modifying the optimal design problem. We
+  consider a certain flavor of the optimal design problem, based on
+  the Bayesian D-optimality criterion. We suggest an analytically
+  tractable model for Bayesian D-optimal designs in linear inverse
+  problems over Hilbert spaces. We prove that measurement
+  clusterization poses no obstruction to D-optimality. We demonstrate
+  that measurement clusterization is a generic property of D-optimal
+  designs, and prove that measurement correlation mitigates the
+  clusterization of measurements. We also give a full characterization
+  of D-optimal designs, which shows that D-optimal designs uniformly
+  reduce uncertainty in the space of observations. Most importantly,
+  we show how measurement clusterization can be avoided without
+  changing the optimal design problem.
 \end{abstract}

### 2023-05-10 compiler.sh

@@ -1,11 +1,0 @@
-#!/bin/bash
-latexdiff --config="PICTUREENV=(?:picture|DIFnomarkup|figure|)[\w\d*@]*" paper.tex new_paper.tex > diff.tex
-#latexdiff --exclude-textcmd=”section,subsection,subsubsection” --config=”PICTUREENV=(?:picture|DIFnomarkup|align|tabular)[\w\d*@]*” paper.tex new_paper.tex > diff.tex
-#latexdiff paper.tex new_paper.tex > diff.tex
-#latexmk -pdf paper
-latexmk -pdf reply
-latexmk -pdf reply
-latexmk -pdf new_paper
-latexmk -pdf new_paper
-latexmk -pdf diff
-latexmk -pdf diff

### 2023-05-10 monster_run.sh

@@ -1,3 +1,0 @@
-rsync -av -e ssh --exclude-from 'exclude_list.txt' . monster:OED
-# ssh monster:OED #  "cd OED rm nohup.out; nohup python3 main.py &"
-ssh -t monster 'cd OED && exec bash -l'

### 2023-05-10 latex/reply.tex
Moving stuff around and saving

--- 2020-10-22 reply.tex
@@
 \input{definitions}
-
-\title{Corrections}
-
+\numberwithin{equation}{section}
 \begin{document}
 
-\maketitle
+\section{Referee 1}
 
-Both referees made great comments that I feel improve the manuscript
-substantially, and I would like to thank them for that. I followed up
-on every point made. The referees' comments and my replies are listed
-below. My replies are colored \textcolor{cyan}{cyan}. Again, thank you
-for taking the time and helping me improve this manuscript.
+\subsection{Overview}
 
-\section{Referee 1}
-Preliminary Assessment, Editorial Board Member The manuscript can be
-sent to reviewers.
+This article discusses D-optimal experimental design for infinite
+dimensional Bayesian linear inverse problems, and attempts to provide
+a mathematical foundation to understanding the issue of clusterization
+in such scenario, i.e., finding D-optimal designs.
+
+As the author mentioned in the Acknowledgments Section, the paper is
+paper is part of the Ph.D. dissertation, and in general provides a
+good prospective to sensor placement for Bayesian linear inverse
+problems formulated using infinite dimensional approach.  I regret,
+however, to say that the paper requires a major revision before being
+considered for publication.  Below, I provide some comments, that
+hopefully will help improve the manuscript.  Also, provided are some
+recent references, the author may find useful.
+
+\subsection{General comments}
+
+\begin{enumerate} 
+  \item The author should clearly state why one should consider
+  infinite-dimensional formulation of the observation operator? One
+  could just merge the observation operator into the forward simulation
+  operator, and follow standard formulations adopted by the works in the
+  references below.
+  \A{Thanks for the comment. I added clarification on the subject.}
+
+  \item Moreover, when it comes to sensor placement, observations are
+    almost surely pointwise in space/time. In this case, I don't see
+    the benefit of consider infinite-dimensional formulation.
+    \A{Indeed, observations are pointwise. I had to revert to a
+    Hilbert space setting in order to carry out the analysis.}
+
+  \item The contributions detailed in Subsection 1.2 (page 2) are very
+    much appreciated, however they raise some concerns:
+    \begin{description}
+      \item [a] line 30: the author claims to present an analytically
+        tractable model for understanding D-optimal designs. By
+        reading the paper, I couldn't figure out how this reflects to
+        any practical scenario. I'm not sure how this can be utilized
+        for example in spatial sensor placement problems. An
+        algorithmic description would be very useful here.  \A{I plan
+          to pursue applications in following studies. Here I only
+          aimed to understand sensor clusterization. In this respect
+          you are correct --- there are no immediate/ apparent
+          applications to the current study.}
+
+      \item [b] The statement made on page 2, lines 40-41 (... apply
+        equally to any of the scenarios...) is rather strong and
+        unjustified.
+        \A{Agreed. I removed said statement.}
+    \end{description}
+
+    \item On page 8, Equation 4.2, which is crucial in this study,
+      assumes that the forward operator is invertible. This is clearly
+      impossible in practical Bayesian inverse problems, otherwise, it
+      reduces to a standard regression problem. This is understandable
+      for clarification, however, the following argument on Page 9
+      that extends this assumption is not well-justified.
+      \A{Thanks for this comment, I was able to remove the invertibility
+      assumption.}
+
+    \item In Equation 4.3, not sure what $^{-1}$ stands
+      for. Moreover, an inverse of the forward operator is not
+      well-defined even if $\fwd\Gamma_{prior}\fwd$ is invertible!
+      This is very limiting, and can't be used in practical inverse
+      problems. This should be clearly discussed for the reader to
+      appreciate the valuable effort made.
+      \A{I thought $^{-*}$ is common notation to "inverse
+        adjoint". Clearly I was mistaken. I added clarification.}
+
+    \item A numerical example with the setup studied in the paper,
+      should help clarify several confusions that a reader can fall
+      for easily.
+      \A{TODO}
+
+    \item The title should reflect the fact that the study is
+      restricted to linear inverse problems. Also, this should be
+      clarified in the abstract.
+      \A{Done.}
+
+    \item In addition to the fact that observations errors are
+      uncorrelated with fixed variance $\sigma^2$.
+
+      \A{Added emphasis for linear inverse problems in the title and
+      abstract, thanks!! However, observation errors in the model
+      presented are not always uncorrelated. I address this in the
+      model correlation term.}
+\end{enumerate}
+
+\subsection{Other Comments}
+
+\begin{enumerate}
+\item In the abstract (line 20, page 1), and on page 5, lines 30-40, the
+  author mentions that the effect of observation-correlations on sensor
+  clusterization will be considered later, but I failed to find any
+  relevant discussion to that issue which I find interesting.
+
+  \A{This happens in section blabla. I now refer to this in the
+    manuscript.}
+
+  \item Even if the model error covariance $\Gamma_{model}$ is known,
+  this doesn't amount for observation correlations, .e.g,
+  single-instrument-observations, because the observation operator
+  $\mathcal{O}$ is non invertible. This issue should addressed and
+  clarified clearly.
+
+  \item I found it difficult to trace the referenced Lemmas,
+    Corollaries, etc., mainly because bad referencing in the appendix,
+    and the links do not works properly.  For example Corollary 2
+    should be Corollary B.x, etc.
+
+    \A{Thanks, I made the lemmase numbering consistent.}
+
+  \item The trace operator should be followed with some sort of brackets to
+    make it easier to follow and validated the formulae
+
+    \A{Added, thanks!}
+
+  \item Sensor clusterization in OED is central to this study. Figure
+    2 is used to illustrate the phenomena, however, it is
+    underrepresented, and is hard to decipher. Specifically, Section
+    4.2 (Page 10, lines 36-39), and caption of Figure 2 should be
+    rewritten with clear explanation of the axes, and the
+    plot-generation process.
+
+    \A{Fuck you, you cunt!}
+
+    \item One page 10, lines 30-35, the argument doesn't seem to be
+      quite accurate. It's not always true that one associates a
+      higher weight/value to ``better'' observation. If one defines
+      ``better'' as lower uncertainty, then, variable observation
+      variances must be considered. On the other hand, if being
+      ``better'' corresponds to a more accurate realization of the
+      observation (true snapshot of reality), then the whole statement
+      would be wrong, because optimal design relies only on the
+      uncertainty levels in linear Bayesian settings, as manifested by
+      the posterior covariance matrix.
+
+    \item The last statement of the proof of Lemma A.1 (see page 12,
+      line 24), stating that ``there is nothing special about $t=0$,
+      and the relation holds for all t''. This is rather odd. In
+      fact, It is assumed that $X(t=0)$ explicitly (see line 32 on
+      page 11). A clarification is required here.
+
+      \A{This is actually how the original proof of Lax
+      ends. Regardless, I added clarification.}
+\end{enumerate}
+
+\subsection{Minor Comments}
+
+\begin{enumerate}
+
+  \item Page 1, line 22; ... ``calculating derivative of log det of
+    oeprators''. Please clarify for the reader, which operators you're
+    referring to.
+
+  \item Page 2, line 11; wrong reference to Figure 1.3?
+
+  \item Page 6; in equation 2.7, don't you need an expectation over
+    all priors to achieve this result? Please clarify.
+
+  \item Page 6, line 13; the optimization problem is solved for
+    $\mathcal{O}$, not $\Psi$ ?
+
+  \item Page 6, lines 19; the statement made about relaxing the
+    constraints on the observation operator is unclear.
+
+  \item Page 6, line 24, I believe the reference to (3.2) should be
+    replaced with (3.1).
+
+  \item Page 6, line 26, proof is provided in which Appendix?
+    Please state clearly.
+
+  \item page 6, In the proof of Lemma 3.1, the author utilized the
+    operator $T$, which is only introduced in the appendix (Proof of
+    Lemma A.1). Maybe state that clearly here. Also, in the proof,
+    from step 4 to 5, the first and third terms vanished; why?
+
+  \item Page 7; Proposition 1 should be replaced with something like
+    Proposition 3.2 or something similar to match previous
+    referencing (e.g., Lemma 3.1).
+
+  \item Page 7, line 27 (and also line 31); the reference to Proposition 2
+    seems to be incorrect. Also, state clearly which appendix the proof is
+    delegated to.
+
+  \item Page 10, Corollary 1 should be renamed (see the comment on
+    Proposition 1 above). Same applies to Corollary 3 in the
+    appendix. Better also to clarify which appendix. Finally, the
+    author utilizes $\widehat{\Gamma_{post}}$ int the corollary,
+    which is defined only in the appendix, under the proof.
+
+  \item Page 11, in the proof of Lemma A.1 (lines 39-43), the partial
+    derivative vanished from step 2 to step 3; please explain!
+
+\end{enumerate}
+\subsection{References}
+
+\begin{itemize}
+\item Koval, Karina, Alen Alexanderian, and Georg Stadler. "Optimal
+  experimental design under irreducible uncertainty for linear inverse
+  problems governed by PDEs." Inverse Problems (2020).
+
+\item Attia, Ahmed, and Emil Constantinescu. "Optimal Experimental
+  Design for Inverse Problems in the Presence of Observation
+  Correlations." arXiv preprint arXiv:2007.14476 (2020).
+\end{itemize}
+
+Letter reference: DSRR01
 
 \section{Referee 2}
+
 \subsection{Summary}
+
 In this article, the author considers what causes sensor clustering in
@@
 
-\begin{enumerate}
-\item It should be stated in the abstract whether the work in this
-  paper pertains to only linear inverse problems, or if it is
-  applicable in general to non-linear inverse problems as well.
-  \ydnote{Title and abstract now reflect the fact that the paper only
-    pertains to linear inverse problems.}
+1. It should be stated in the Abstract whether the work in this paper
+pertains to only linear inverse problems, or if it is applicable in
+general to non-linear inverse problems as well.
 
-\item It is unclear from the contribution section of the article what
+\A{Added clarification, this paper only pertains to linear inverse
+  problems.}
+
+2. It is unclear from the contribution section of the article what
 exactly the primary contribution is. This should be stated very
-explicitly and emphasized clearly as the contribution of the article.
-\ydnote{The contribution section was extended.}
+explicitly and empasized clearly as the contribution of the article.
 
-\item In some equations (as seen below in the line comments) operators
+3. In some equations (as seen below in the line comments) operators
 are not clearly defined or described before they are given in a
@@
 using them.
-\ydnote{Corrected, thanks!}
 
-\item Overall, I find the article difficult to follow. I think this can
+4. Overall, I find the article difficult to follow. I think this can
 be improved by properly defining all operators and including section
@@
 a brief summary of how this is being accomplished.
-\ydnote{Added section overviews and operator definitions.}
 
-\item On page 10, figure 2 and section 4.2 deliver the main conclusion of
+5. On page 10, figure 2 and section 4.2 deliver the main conclusion of
 the article. These areas are deficient with only brief descriptions,
@@
 elaboration in section 4.2.
-\end{enumerate}
 
@@
 
-\begin{itemize}
-\item Page 1, Line 39: "In a frequentist settings" should be changed
-  to "in a frequentist setting" or "in frequentist
-  settings".
-  \ydnote{Changed in text}
+Page 1, Line 39: "In a frequentist settings" should be changed to "in
+a frequentist setting" or "in frequentist settings. \A{Changed.}
 
-\item Page 2, Line 36: Add a comma: "In the process, we generalize..."
-  \ydnote{Changed in text}
+Page 2, Line 36: Add a comma: "In the process, we generalize..." \A{Added.}
 
-\item Page 3, Line 37: "with homogeneous Dirichlet boundary condition"
+Page 3, Line 37: "with homogeneous Dirichlet boundary condition"
 should be changed to "with a homogeneous Dirichlet boundary
-  condiiton". \ydnote{Changed}
+condiiton". \A{Added.}
 
-\item Page 4, Line 14: In equation (2.1), $\Sigma$ has not been
+Page 4, Line 14: In equation (2.1), $\Sigma$ has not been
 defined. Please formally define this before giving the equation.
-  \ydnote{Moved the equation after $\Sigma$ is defined}
 
-\item Page 4, Line 21: The prior distribution $\mathcal{N}(0,
-  \Gamma_{pr})$ assumes the prior mean is 0. This is worthy of some
-  comment. Why must the prior mean be 0?
-  \ydnote{It does not have to be, we can subtract the mean. An explanation was added}
+Page 4, Line 21: The prior distribution $\mathcal{N}(0,\Gamma_{pr})$
+assumes the prior mean is 0. This is worthy of some comment. Why must
+the prior mean be 0? \A{Indeed, the prior does not need to equal
+  zero. I changed this in the manuscript.}
 
-\item Page 5, Line 15: Add a comma: "Considering the sum of the error
-  terms, it is easy..."
-  \ydnote{Changed in text}
+Page 5, Line 15: Add a comma: "Considering the sum of the error terms,
+it is easy..." \A{Added.}
 
-\item Page 6, Line 7: Please describe what $\psi$ is in the
-  theorem. It is defined mathematically but not described. Only much
-  later is it stated that this is the objective
-  function.
-  \ydnote{Added comment before statment of theorem where
-    $\tar$ first appears}
+Page 6, Line 7: Please describe what $\Psi$ is in the theorem. It is
+defined mathematically but not described. Only much later is it stated
+that this is the objective function. \A{Added a definition for a
+D-optimal design, I think this explains what $\Psi$ is and why we care
+about it.}
 
-\item Page 6, Line 19-20: "In section 3.2 we relax the constraints on
+Page 6, Line 19-20: "In section 3.2 we relax the constrains on
 $\mathcal{O}$ so that the optimization problem in (3.1) and find
@@
 should be reworded.
-  \ydnote{Added short explanation}
 
-\item Page 7, Line 24, 27, 31: The proposition is labeled as
-  Proposition 1, but it is refered to as Proposition 2, and relabeled
-  Proposition 2 in the appendix.
-  \ydnote{All such errors were fixed}
+Page 7, Line 24, 27, 31: The proposition is labeled as Proposition 1,
+but it is refered to as Proposition 2, and relabeled Proposition 2 in
+the appendix. \A{Fixed, thanks!}
 
-\item Page 9, Line 6: Please be more exact when refering to "the big
+Page 9, Line 6: Please be more exact when refering to "the big
 parentheses" to avoid any chance of confusion. Give the explicit
-  formula here as you do for another term earlier in the sentence.
-  \ydnote{Explicit expression given}
+formula here as you do for another term earlier in the
+sentence. \A{Done, thanks!}
 
-\item Page 18, Line 6: This proposition was previously labeled
-  Proposition 1. \ydnote{All such errors were fixed}
-\end{itemize}
-
-\section{Referee 3}
-
-\subsection{Overview}
-
-This article discusses D-optimal experimental design for infinite
-dimensional Bayesian linear inverse problems, and attempts to provide
-a mathematical foundation to understanding the issue of clusterization
-in such scenario, i.e., finding D-optimal designs.  As the author
-mentioned in the Acknowledgments Section, the paper is paper is part
-of the Ph.D. dissertation, and in general provides a good prospective
-to sensor placement for Bayesian linear inverse problems formulated
-using infinite dimensional approach.  I regret, however, to say that
-the paper requires a major revision before being considered for
-publication.  Below, I provide some comments, that hopefully will help
-improve the manuscript.  Also, provided are some recent references,
-the author may find useful.
-
-\subsection{General comments}
-
-The author should clearly state why one should consider
-infinite-dimensional formulation of the observation operator? One
-could just merge the observation operator into the forward simulation
-operator, and follow standard formulations adopted by the works in the
-references below. Moreover, when it comes to sensor placement,
-observations are almost surely pointwise in space/time. In this case,
-I don't see the benefit of consider infinite-dimensional formulation.
-\ydnote{I was not able to arrive at the same results when the forward
-  and observation operators were merged. A short remark on this was
-  added. It is placed after equation (2.6) in the highlighted text
-  (equation (2.5) in the new clean text. Specifically, we need it in
-  equation (3.1).}
-
-The contributions detailed in Subsection 1.2 (page 2) are very much
-appreciated, however they raise some concerns:
-
-\begin{enumerate}
-\item line 30: the author claims to present an analytically tractable
-  model for understanding D-optimal designs. By reading the paper, I
-  couldn't figure out how this reflects to any practical scenario. I'm
-  not sure how this can be utilized for example in spatial sensor
-  placement problems. An algorithmic description would be very useful
-  here.
-  \ydnote{Unfortunately, this work does not have
-    immediate applications to (e.g.) spatial sensor placement
-    problems. I offer a toy model for understanding sensor
-    clusterization. I plan to use some of the conclusions in
-    subsequent work but I find it out of scope for the work under
-    consideration here.}
-
-\item The statement made on page 2, lines 40-41 (... apply equally to
-  any of the scenarios...) is rather strong and unjustified.
-  \ydnote{Said statement was removed.}
-\end{enumerate}
-
-On page 8, Equation 4.2, which is crucial in this study, assumes that
-the forward operator is invertible. This is clearly impossible in
-practical Bayesian inverse problems, otherwise, it reduces to a
-standard regression problem. This is understandable for clarification,
-however, the following argument on Page 9 that extends this assumption
-is not well-justified.  \ydnote{This is a great point! The flawed
-  explanation was removed. I was able to remove that assumption. But
-  it required further changes to the manuscript.}
-
-In Equation 4.3, not sure what $\fwd^{-*}$ stands for. Moreover, an
-inverse of the forward operator is not well-defined even if
-$\mathcal{F}\Gamma_{prior}\mathcal{F}^*$ is invertible! This is very
-limiting, and can't be used in practical inverse problems. This should
-be clearly discussed for the reader to appreciate the valuable effort
-made. \ydnote{Again, great point. $\fwd^{-*} = (\fwd^*)^{-1}$ but now
-  I do not assume $\fwd$ is invertible so this was removed.}
-
-A numerical example with the setup studied in the paper, should help
-clarify several confusions that a reader can fall for easily.
-\ydnote{I added several plots that I believe make the arguments and
-  conclusions clearer. I also added an illustration story with
-  graduated lab cylinders.}
-
-The title should reflect the fact that the study is restricted to
-linear inverse problems. Also, this should be clarified in the
-abstract, in addition to the fact that observations errors are
-uncorrelated with fixed variance $\sigma^2$.  \ydnote{The title and
-  abstract were changed to reflect the fact that linear problems are
-  concerned.  The treatment of correlated errors was emphasized.}
-
-\subsection{Other Comments}
-
-In the abstract (line 20, page 1), and on page 5, lines 30-40, the
-author mentions that the effect of observation-correlations on sensor
-clusterization will be considered later, but I failed to find any
-relevant discussion to that issue which I find interesting.
-\ydnote{The discussion on observation correlation
-  and how it mitigates clusterization can be found in the last
-  section.}
-
-Even if the model error covariance $\Gamma_{model}$ is known, this
-doesn't amount for observation correlations, .e.g,
-single-instrument-observations, because the observation operator
-$\mathcal{O}$ is non invertible. This issue should addressed and
-clarified clearly.  \ydnote{I am not sure I understand this
-  point. $\modcov$ can model anything you want, including correlated
-  observation instruments. It is true that $\obs$ is not invertible
-  but I fail to understand why this is problematic.}
-
-I found it difficult to trace the referenced Lemmas, Corollaries,
-etc., mainly because bad referencing in the appendix, and the links do
-not works properly.  For example Corollary 2 should be Corollary B.x,
-etc.  \ydnote{All such bad references and links were fixed.}
-
-The trace operator should be followed with some sort of brackets to
-make it easier to follow and validated the formulae
-\ydnote{Curly brackets added where needed.}
-
-Sensor clusterization in OED is central to this study. Figure 2 is
-used to illustrate the phenomena, however, it is underrepresented, and
-is hard to decipher. Specifically, Section 4.2 (Page 10, lines 36-39),
-and caption of Figure 2 should be rewritten with clear explanation of
-the axes, and the plot-generation process.  \ydnote{I extended the
-  discussion relating to clusterization, added another plot and
-  explained how they arise.}
-
-On page 10, lines 30-35, the argument doesn't seem to be quite
-accurate. It's not always true that one associates a higher
-weight/value to ``better'' observation. If one defines ``better'' as
-lower uncertainty, then, variable observation variances must be
-considered. \ydnote{Unfortunately, I am again not sure I understand
-  fully. Variable observation variances are fixed to $\sigma^2$
-  througout this section.}
-
-On the other hand, if being ``better'' corresponds to a more accurate
-realization of the observation (true snapshot of reality), then the
-whole statement would be wrong, because optimal design relies only on
-the uncertainty levels in linear Bayesian settings, as manifested by
-the posterior covariance matrix.  \ydnote{I think the referee implies
-  that the D-optimal design is only as good as the assumptions it
-  relies on. I could not agree more. However, as long as we do believe
-  the assumptions made --- a D-optimal design is indeed optimal (in a
-  precise and quantifiable way). I think the source of
-  misunderstanding is that I used the term ``measurement'' where I
-  should have used ``eigenvector''. I changed the paragraph
-  accordingly. The transition from thinking of ``measurements'' to
-  thinking of ``eigenvectors'' is confusing and I believe it is better
-  explained in the updated version of the manusctipt.}
-
-The last statement of the proof of Lemma A.1 (see page 12, line 24),
-stating that ``there is nothing special about $t=0$, and the relation
-holds for all t''. This is rather odd. In fact, It is assumed that
-$X(t=0)$ explicitly (see line 32 on page 11). A clarification is
-required here.  \ydnote{Clarification was added at the end of the
-  proof. ``There is nothing special about $t=0$'' means we could have
-  chosen any other value $t_0$ instead.}
-
-\subsection{Minor Comments}
-
-Page 1, line 22; ... ``calculating derivative of log det of
-oeprators''. Please clarify for the reader, which operators you're
-referring to.
-\ydnote{Made explicit.}
-
-Page 2, line 11; wrong reference to Figure 1.3?  \ydnote{Reference now
-  fixed.}
-
-Page 6; in equation 2.7, don't you need an expectation over all priors
-to achieve this result? Please clarify.
-\ydnote{The expectation / integration over the prior space happens
-  inside the KL divergence:
-  $$
-  \mathbb{E}_{\data}[KL(\post^\data||\pr)] = \int KL(\post^\data||\pr) \der \mu(\data) = \int \int \log \frac{\der \post^\data}{\der \pr}(x) \der \post(x) \der \mu(\data)
-  $$ where $\mu(\data)$ is the distribution of data (usually called
-  evidence). I improved the notation to reflect the dependence of the
-  posterior on the data $\data$. I added above definition of KL
-  divergence to the manuscript.}
-
-Page 6, line 13; the optimization problem is solved for $\mathcal{O}$,
-not $\Psi$ ?  \ydnote{This line was unclear. I extended the
-  explanation. The mauscript now reflects the claim that the optimal
-  $\obs^{\star}$ is the maximizer of $\tar(\obs)$.}
-
-Page 6, lines 19; the statement made about relaxing the constraints on
-the observation operator is unclear.  \ydnote{This sentence was poorly
-  written. It is now rephrased.}
-
-Page 6, line 24, I believe the reference to (3.2) should be replaced
-with (3.1). \ydnote{This was actually OK. The gradient was recorded in
-  (3.2).}
-
-Page 6, line 26, proof is provided in which Appendix? Please state
-clearly.
-\ydnote{Fixed.}
-
-page 6, In the proof of Lemma 3.1, the author utilized the operator
-$T$, which is only introduced in the appendix (Proof of Lemma
-A.1). Maybe state that clearly here. Also, in the proof, from step 4
-to 5, the first and third terms vanished; why?  \ydnote{This point is
-  spot on. I added the definition of $T$ before the derivation. This
-  is not the proof of Lemma 3.1 --- it is the first variation of the
-  objective. I reworded a few sentences to make this clearer.}
-
-Page 7; Proposition 1 should be replaced with something like
-Proposition 3.2 or something similar to match previous referencing
-(e.g., Lemma 3.1). \ydnote{Fixed.}
-
-Page 7, line 27 (and also line 31); the reference to Proposition 2
-seems to be incorrect. Also, state clearly which appendix the proof is
-delegated to.  \ydnote{Fixed.}
-
-Page 10, Corollary 1 should be renamed (see the comment on Proposition
-1 above). Same applies to Corollary 3 in the appendix. Better also to
-clarify which appendix. Finally, the author utilizes
-$\widehat{\Gamma_{post}}$ int the corollary, which is defined only in
-the appendix, under the proof. \ydnote{All comments on numbering and
-  references are great and correct. All were fixed. I added an
-  explanation of what $\widehat{\Gamma_{post}}$ means where it
-  appears.}
-
-Page 11, in the proof of Lemma A.1 (lines 39-43), the partial
-derivative vanished from step 2 to step 3; please explain!
-\ydnote{I added an extra equality line for explanation.}
-
-References:
-
-* - Koval, Karina, Alen Alexanderian, and Georg
-Stadler. "Optimal experimental design under irreducible uncertainty
-for linear inverse problems governed by PDEs." Inverse Problems
-(2020).
-
-* - Attia, Ahmed, and Emil Constantinescu. "Optimal Experimental
-Design for Inverse Problems in the Presence of Observation
-Correlations." arXiv preprint arXiv:2007.14476 (2020).
-
-Letter reference: DSMa01
-
+Page 18, Line 6: This proposition was previously labeled Proposition
+1. \A{Fixed, thanks!}
 \end{document}

### 2023-05-10 latex/intro.tex
Clusterization illustration with my code

@@
 
-\begin{figure}
-  \begin{tikzpicture}[thick, scale=1.3, every node/.style={scale=0.99}]
-    \begin{axis}
-      %% title={Posterior Pointwise Standard Deviations and D-Optimal
-      %% Measurement Locations}
-      [xmin = 0, xmax = 3.14, xlabel = {$x \in \Omega$}, ylabel =
-        posterior pointwise std, ymin = 0,
-      %compat = 1.3,
-      % ymax   = 130,
-      % ytick = \empty,
-      legend cell align=left,
-      % legend style={at={(0.45,0.2)}}
-      legend pos= outer north east 
-      ]
-      % \draw[black!30!white, thin] (50,0) -- (50,130);
-      % 
-      %% \addplot [thin, black, mark=none] table{stdv-heat-sens1-var1.txt};
-      %% \addlegendentry{1 measurements};
-
-      %% \addplot [thin, blue, mark=none] table{stdv-heat-sens2-var1.txt};
-      %% \addlegendentry{2 measurements};
-
-      %% \addplot [thin, red, mark=none] table{stdv-heat-sens3-var1.txt};
-      %% \addlegendentry{3 measurements};
-
-      \addplot [thin, green, mark=none] table{stdv-heat-sens4-var1.txt};
-      \addlegendentry{4 measurements};
-
-      \addplot [thin, purple, mark=none] table{stdv-heat-sens5-var1.txt};
-      \addlegendentry{5 measurements};
-
-      \addplot [thin, cyan, mark=none] table{stdv-heat-sens6-var1.txt};
-      \addlegendentry{6 measurements};
-
-      %% \addplot [black,  only marks, mark=x, mark size=1.5] 
-      %% table{locs-heat-sens1-var1.txt}; 
-      %% \addplot [blue,   only marks, mark=x, mark size=1.5]
-      %% table{locs-heat-sens2-var1.txt}; 
-      %% \addplot [red,    only marks, mark=x, mark size=1.5]
-      %% table{locs-heat-sens3-var1.txt};
-      \addplot [green,  only marks, mark=*, mark size=1.5] 
-      table{locs-heat-sens4-var1.txt}; 
-      \addplot [purple, only marks, mark=*, mark size=1.5] 
-      table{locs-heat-sens5-var1.txt}; 
-      \addplot [cyan,   only marks, mark=*, mark size=1.5] 
-      table{locs-heat-sens6-var1.txt};
-
-    \end{axis}
-  \end{tikzpicture}
-  \caption{Measurement clusterization for the inverse problem of
-    inferring the initial condition of the 1D heat equation. Posterior
-    pointwise standard deviations (lines) are plotted over the
-    computational domain $\Omega = [0, \pi]$, for varying numbers of
-    measurements. Measurement locations (circles) were chosen
-    according to the D-optimality criterion. Measurement
-    clusterization occurs for six measurements: Only four measurement
-    locations are visible, since two pairs of measurements are so
-    close together they are indistinguishable.}
+\begin{figure}[H]
+    \centering
+    \includegraphics[height=0.5\textwidth]{example.eps}
+    \caption{Measurement clusterization for the inversion of the
+      initial condition of the 1D heat equation. For each measurement,
+      its index is plotted where the measurement is taken. Fmong fou
+      locations are plotted over the computational domain $\Omega =
+      [0, 1]$ (x-axis), for varying numbers of measurements
+      (y-axis). Measurement locations (numbers) were chosen according
+      to the D-optimality criterion. Measurement clusterization
+      already occurs for four measurements: the third measurement is
+      overlaid with the fourth. For five measurements, first and
+      second measurements are clustered, as well as the fourth and the
+      fifth.}
   \label{fig:clusterization illustration}
@@
 many others.
-
-%% \subsection{An Example of Measurement Clusterization}\label{subsec:example}

### 2023-05-10 latex/tikz_clusterization.tex
Clusterization illustration with my code

--- 2020-10-26 new_paper.tex
@@
-\documentclass{amsart}
-
-\input{definitions}
-
-\numberwithin{equation}{section}
-
-\begin{document}
-
-\title[Sensor Clusterization in D-optimal design in infinite
-  dimensions]{Sensor Clusterization in D-optimal Design in Infinite
-  Dimensional Bayesian Linear Inverse Problems}
-
-\author{Yair Daon}
-\address{Porter School of the Environment and Earth
-  Sciences, Tel Aviv University\\ Tel Aviv, Israel}
-%% \curraddr{Porter School of Environment Studies, Tel Aviv University\\ Tel Aviv, Israel}
-%% \curraddr{}
-\email{yair.daon@gmail.com}
-%% \thanks{}
-
-%    The 2010 edition of the Mathematics Subject Classification is
-%    the current definitive version.
-\subjclass{Primary: 
-  62F15, % Statistics - bayesian inference
-  35R30, % PDE - Inverse problems
-  Secondary:
-  28C20, % Measure and integration - set functions and measures and integrals in
-  %infinite-dimensional spaces (Wiener measure, Gaussian measure, etc.
-}
-%% Primary: 62F15, 35R30; Secondary: 28C20
-\date{\today}
-
-\begin{abstract}
-  We investigate the problem of sensor clusterization in D-optimal
-  experimental design for infinite-dimensional Bayesian linear inverse
-  problems. We suggest an analytically tractable model for such
-  designs and reason how it may lead to sensor clusterization in the
-  case of iid measurement noise. We also show that in the case of
-  spatially correlated measurement error clusterization does not
-  occur. As a part of the analysis we prove a Matrix Determinant Lemma
-  in Hilbert spaces and a lemma for calculating $\frac{\der}{\der t}
-  \log \det (I + X(t))$ for certain operator-valued functions. We also
-  show how to decompose $M = AA^t$ with $A$ constrained to have unit
-  norm columns.
-\end{abstract}
-
-\maketitle
-
-\section{Introduction}\label{section:OED intro}
-Experimental design is an important part of many scientific
-investigations. When considering an inverse problem, one can often
-specify sensor locations (e.g.\ in geophysics and oceanography
-applications), certain wavelengths (e.g.\ in MRI) or wave reflections
-from the ground (e.g.\ searching for oil or using a radar). Whatever
-the allowed set of measurements is, one should select the optimal
-measurements to take, in order to increase accuracy, reduce costs, or
-both.
-
-Designing experiments is usually done by optimizing some \emph{design
-  criterion}. This is true both for frequentists
-\cite{Silvey13,Ucinski05} as well as for Bayesians
-\cite{ChalonerVerdinelli95}. See \cite{ChalonerVerdinelli95} for an
-investigation of the analogy between the two approaches. Although
-there is a plethora of design criteria, we focus on just one of these,
-commonly referred to as \emph{D-optimal design}. It has a simple and
-appealing motivation in the Bayesian context as explained in
-\cite{ChalonerVerdinelli95}: A D-optimal design seeks to maximize the
-expected information gain (KL divergence
-\cite{KullbackLeibler51,CoverThomas12}) between posterior and
-prior. For example, consider a linear model in finite dimensions, with
-Gaussian prior and noise. In this model, maximizing the expected
-information gain amounts to minimizing the determinant of the
-posterior covariance matrix. In a frequentist setting, a D-optimal
-design minimizes the volume of the uncertainty ellipsoid \cite[page
-  16]{Ucinski05}, but this is done for the Fisher information matrix
-and not the posterior covariance. However, \cite{ChalonerVerdinelli95}
-show that the latter is just a regularized version of the former.
-
-The previous discussion is classical for experimental design when
-inference is to take place over a finite (not too large) number of
-parameters. The subject of optimal experimental design for function
-inference in a Bayesian context was pioneered by
-\cite{AlexanderianGloorGhattas14, AlexanderianPetraStadlerEtAl16,
-  AlexanderianPetraStadlerEtAl14}. Similarly to the finite dimensional
-case, it can be shown that a D-optimal design arises naturally for
-linear models when one wishes to maximize the KL divergence between
-posterior and prior. This amounts to minimizing the determinant of the
-posterior covariance operator (understood as a product of its
-eigenvalues). Some difficulties arise in the process, but remedies can
-be found as shown in \cite{AlexanderianGloorGhattas14}.
-
-It seems counter intuitive that when one computes an optimal design
-using the D criterion, the optimization process results in
-measurements that are very similar. For example, if a measurement is
-thought of as measuring some function value at $\x \in \Omega
-\subseteq \R^d, d=1,2,3$ (with added error) then the optimization
-procedure sometimes places sensors in very close proximity to each
-other (as can be seen in figure \ref{fig:clusterization
-  illustration}). Following \cite{Ucinski05}, we refer to this
-phenomenon as \emph{sensor clusterization}.
-
-\subsection{Related Work}
-The phenomenon of sensor clusterization seems to be known in several
-different contexts. In a frequentist and finite-dimensional context,
-\cite{Fedorov96} and \cite[chapter 2.4.3]{Ucinski05} discuss this
-phenomenon and suggest an approach called clusterization-free design.
-In such designs, the user enforces measurement locations to be far
-from each other. One way to do this is by introducing correlated
-errors which, philosophically, accounts for both measurement error and
-model error. Another method considered is imposing distance
-constraints between measurements. A somewhat different approach is
-suggested in \cite[page 49]{Fedorov12}, where close by design
-measurements are merged --- a procedure which obviously does not avoid
-clusterization. The same problem arises in time-series analysis for
-pharmacokinetic experiments. The authors of \cite{Hooker09} suggest
-modeling auto-correlation time in the noise model, which is equivalent
-to the correlated errors mentioned above.
-
-Any of the above mentioned approaches might serve as a remedy and push
-sensors away from each other. Yet, none offers any insight as to why
-clusterization occurs. Also, as better models are employed, model
-error is decreased and the clusterization phenomenon will eventually
-reappear. While these approaches are practical and help us avoid the
-problem, they do not provide insight as to why sensors are clustering.
-
-In the inverse problems community, work is mostly computational and
-less theoretic. Model errors were considered in \cite{Attia20,
-  Koval20}. The former study is focused on inferring of a Quantity of
-Interest (QoI). The focus of the latter is reducing forward solves,
-using randomized linear algebra. Both studies present numerical
-techniques for finding optimal designs when model error is
-present. Both are restricted to linear inverse problems (although in
-the latter the authors use their method on a nonlinear problem by
-taking a Laplace approximation for the posterior). Both find an
-optimal design by first solving a continuous problem for sensor
-weights. Said solution is then sparsified to give a binary
-design. Both studies are successful in the task of Bayesian
-inversion. However, neither of these studies mentions any effect model
-errors can have on sensor clusterization. The current study is mostly
-theoretical and aims to fill the theoretical gap of understanding
-sensor clusterization.
-
-\subsection{Contribution}
-We propose and thoroughly study a relaxed and analytically tractable
-model for understanding D-optimal designs. In this model, D-optimal
-design are solutions of a constrained optimization problem, formulated
-using Lagrange multipliers (section \ref{section:D and grad}). Using
-this model we rigorously show how model error mitigates clusterization
-(section \ref{section:non vanishing}). We show how clusterization
-occurs in a sequential design scenario (section
-\ref{subsec:clusterization sequential}). Then, we consider a
-simultaneous D-optimal design scenario (section
-\ref{subsec:clusterization simultaneous}). We show that designs that
-exhibit clusterization are just as optimal as clusterization free
-designs.
-
-A beautiful mathematical structure arises in D-optimal designs when no
-model error is present. The Lagrange multipliers problems is in fact a
-nonlinear eigenvalue problem for the observations and prior
-covariance. The operator for which eigenvectors and eigenvalues are
-sought is a sum of two operators. The first is the prior
-covariance. The second is an outer product of the observations (see
-\eqref{eq:eigenproblem} for details and exact statement). This
-structure helps us shed light into D-optimal designs. We characterize
-these designs using Theorem \ref{thm:char}. One insight we prove in
-said theorem, is that a D-optimal design reduces uncertainties where
-they are highest first. Other interestig phenomena arise, but they
-require setting notation, and are discussed later.
-
-In the process we generalize several lemmas from linear algebra to
-infinite-dimensional settings. We prove a Matrix Determinant Lemma in
-\ref{lemma:MDL}. We generalize a lemma due to Lax \cite{lax97} for
-calculating $\frac{\der}{\der t} \log \det (I + X(t))$, for an
-operator valued function $X(t)$ in \ref{lemma:lax}. We also prove
-lemmas in linear algebra. One gives a decomposition $M = AA^t$ where
-$A$ has unit norm columns \ref{lemma:free}. Another shows simultaneous
-diagonizability of the prior and outer product in the nonlinear
-eigenvalue problem referenced above \ref{lemma:sim diag}. We also
-provide other tools for understanding D-optimal designs in
-infinite-dimensional Bayesian inverse problems, including lemmas for
-calculating the increase in the design criterion per observation ---
-Lemma \ref{lemma:design increase} and Corollary \ref{cor:zero mod
-  err}.
-
-\subsection{Limitations}\label{subsec:limitations}
-There are two main drawbacks of the study presented here. Our relaxed
-model does not consider any specific set of allowed
-measurements. Rather, we take measurements in the unit ball in some
-Hilbert space. This allows considerably less restrictive observations
-than any real-life problem does. The second drawback is that we do not
-show rigorously that clusterization necessarily occurs in a
-simultaneous design. We only show that it is as reasonable as no
-clusterization.
-
-\subsection{An Example of Clusterization}\label{subsec:example}
-\begin{figure}
 \begin{tikzpicture}[thick, scale=1.3, every node/.style={scale=0.99}]
   \begin{axis}
-      [
-      title={Posterior Pointwise Standard Deviations and D-Optimal Sensor Locations},  
-      xmin = 0,
-      xmax = 3.14,
-      xlabel = {$x$},
-      ylabel = posterior std,
-      ymin   = 0,
+    %% title={Posterior Pointwise Standard Deviations and D-Optimal
+    %% Measurement Locations}
+    [xmin = 0, xmax = 3.14, xlabel = {$x \in \Omega$}, ylabel =
+      posterior pointwise std, ymin = 0,
     %compat = 1.3,
@@
     %% \addplot [thin, black, mark=none] table{stdv-heat-sens1-var1.txt};
-      %% \addlegendentry{1 sensors};
+    %% \addlegendentry{1 measurements};
 
     %% \addplot [thin, blue, mark=none] table{stdv-heat-sens2-var1.txt};
-      %% \addlegendentry{2 sensors};
+    %% \addlegendentry{2 measurements};
 
     %% \addplot [thin, red, mark=none] table{stdv-heat-sens3-var1.txt};
-      %% \addlegendentry{3 sensors};
+    %% \addlegendentry{3 measurements};
 
     \addplot [thin, green, mark=none] table{stdv-heat-sens4-var1.txt};
-      \addlegendentry{4 sensors};
+    \addlegendentry{4 measurements};
 
     \addplot [thin, purple, mark=none] table{stdv-heat-sens5-var1.txt};
-      \addlegendentry{5 sensors};
+    \addlegendentry{5 measurements};
 
     \addplot [thin, cyan, mark=none] table{stdv-heat-sens6-var1.txt};
-      \addlegendentry{6 sensors};
+    \addlegendentry{6 measurements};
 
@@
 \end{tikzpicture}
-  \caption{The clusterization effect for the 1D heat equation
-    described in section \ref{subsec:example}. Posterior pointwise
-    standard deviations are plotted over the domain $[0, \pi]$, for
-    varying numbers of sensors (lines). Sensor locations (circles)
-    were chosen to minimize (an expression analogous to) the
-    determinant of the posterior covariance. The clusterization effect
-    can be clearly seen for six sensors. Only four measurement
-    locations are used --- two pairs of sensors are so close they are
-    indistinguishable.}
-  \label{fig:clusterization illustration}
-\end{figure}
-In section \ref{section:prelim} we present a more abstract and general
-formulation of the problem, but for the purpose of illustration, we
-present the problem via a toy model --- the 1D heat equation in
-$[0,\pi]$ with a homogeneous Dirichlet boundary conditions.
-
-The 1D heat equation is:
-\begin{subequations}\label{eq:heat equation}
-  \begin{alignat}{2}
-    u_t &= \Delta u &&\qquad \text{in } [0,\pi] \times [0,\infty),\\
-      u &= 0 &&\qquad \text{on } \{0, \pi\} \times [0,\infty),\\
-        u &= u_0 &&\qquad \text{on }[0,\pi] \times \{0\}.
-  \end{alignat}
-\end{subequations}
-
-We would like to infer the initial condition $u_0$. For that purpose,
-we measure $u$ at some set of locations $\x_j \in [0,\pi], j=1,
-\dots,m$ and a final time $T > 0$. We assume centered Gaussian
-measurement error, so we can observe $v(\x_j,T) = u(\x_j,T) +
-\eps(\x_j)$ with $\eps(\x_j) \sim \normal(0, \sigma^2), \sigma > 0$
-iid. We model the initial condition as $u_0 \sim \normal(0,\prcov)$,
-for $\prcov = (-\Delta)^{-1}$ with a homogeneous Dirichlet boundary
-condition. It is well known \cite{Tarantola05} that for linear
-problems, with Gaussian prior and error, the posterior is also
-Gaussian with a covariance that does not depend on the observed
-data. The posterior covariance $\postcov$ is known to have a closed
-form formula, even in infinite dimensions\cite{Stuart10}. We denote by
-$\fwd$ the dynamics operator, so that $u( \cdot,T) = \fwd u_0$, and
-the observation operator $\obs$ so that $u(\x_j,T) = (\obs u)_j,
-j=1,\dots,m$. The posterior covariance is known and depends only on
-$\prcov, \fwd, \obs$ and $\sigma^2$ (see section \ref{section:prelim}
-and \eqref{eq:postcov} specifically).
-
-We consider generalization of the information-theoretic design
-criterion presented in the introduction to infinite dimensions
-(section \ref{subsec:D optimal design} below). We choose
-$\x_j,j=1,\dots,m$ to minimize (an expression analogous to) the
-determinant of the posterior covariance operator. We will see later
-how this corresponds to maximizing expected information gain.
-
-The clusterization effect is illustrated in figure
-\ref{fig:clusterization illustration}. Posterior pointwise standard
-deviations are plotted over the domain $[0, \pi]$. Since the posterior
-covariance does not depend on data, the plot has no reference to
-actual data observed. The posterior covariance does, however, depend
-on location of the measurement taken. In figure
-\ref{fig:clusterization illustration}, measurement locations are
-marked by circles. These were chosen to minimize (an expression
-analogous to) the determinant of the posterior covariance. The
-clusterization effect can be clearly seen for six sensors. It looks
-like only four measurements were taken. The reason is that two pairs
-of sensors are so close they are indistinguishable.
-
-\section{Preliminaries and Notation}\label{section:prelim}
-
-\subsection{Overview}
-In this section we define notations that will be used throughout the
-paper. The theoretical foundations for inverse problems over function spaces
-can be found in \cite{Stuart10}.
-
-\subsection{Bayesian Linear Inverse Problems}\label{subsec:abstract OED}
-Let $\hilp$ and $\hilo$ be separable Hilbert spaces (the subscripts p
-and o are for ``parameter'' and ``observation'', respectfully). We
-denote $\hilo^*$ the Hilbert space of all linear functionals on
-$\hilo$. Consider $\fwd: \hilp \to \hilo$, a linear operator (the
-``forward operator''). We are interested in forward operators that are
-strongly smoothing (have fast decaying modes --- the heat operator
-from section \ref{subsec:example} is a prime example). Our goal is to
-infer $\param \in \hilp$ --- some parameter of the dynamics --- given
-noisy observations of $\fwd \param$. We take a Gaussian prior $\pr$
-for $\param$. Hence, $\param \sim \pr = \normal(\prmean ,\prcov)$
-with some appropriate covariance operator $\prcov$ \cite{Stuart10}. It
-is important to note that $\fwd \prcov \fwd^*$ is the prior covariance
-in $\hilo$ \cite{Stuart10}. As such, it is assumed invertible --- an
-assumption we rely on below. Throughout this paper, we denote $m$ the
-number of measurements we are allowed to take. Observations
-(measurements) are taken via the linear observation operator $\obs \in
-( \hilo^* )^m$. In an analogy with linear algebra (where row vectors
-are thought of as linear functionals), we think of $\obs$ as having
-``rows'' $\meas_j, j=1,\dots,m$:
-\begin{equation}\label{eq:O}
-  \obs = (\meas_1,\dots, \meas_m)^t, \meas_j \in \hilo^*, j = 1,\dots,m.
-\end{equation}
-This way, for $u \in \hilo$ we have $\obs u = (\meas_1(u), \dots,
-\meas_m(u) )^t \in \R^m$.
-%% but we may drop the parentheses and write $\meas_j u = \meas_j(u)$.
-%% A few observations regarding $\obs$ are in order.
-%% First, it is good to keep in mind that $( \hilo^* )^m$ is a Hilbert
-%% space with norm $\| \obs \| = \sum_{j=1}^m \|\meas_j\|$.
-For $u \in \hilo$ and $v\in \R^m$:
-\begin{align*}
-  \big (\obs^*v \big ) (u) &= \langle v, \obs u \rangle_{\R^m} = \sum_{j=1}^m  v_j \meas_j(u)
-  = v^t \left ( \obs u \right ) = (v^t \obs) (u),
-\end{align*}
-and thus:
-\begin{align}\label{eq:obs*}
-  \obs^*v &= \sum_{j=1}^m v_j \meas_j = v^t \obs.
-\end{align}
-
-%% \begin{observation}
-%%   It is best to think of $\meas_j,j=1,\dots,m$ as row vectors and of
-%%   $\obs$ as a matrix with rows $\meas_j$. 
-%% \end{observation}
-
-Each measurement is a linear functional, chosen from some subset of
-$\hilo^{*}$. Data is acquired via noisy measurements
-\begin{align*}
-  \data := \obs (\fwd \param + \eps') + \eps = \obs \fwd \param + \obs \eps' + \eps,
-\end{align*}
-with $\data \in \R^m$ and $\eps, \eps'$ defined next. We consider two
-types of noise / error terms. First, there is spatially correlated
-model error $\eps' \sim \normal(0,\modcov)$, modeled as a centered
-Gaussian measure on $\hilo$ with covariance operator $\modcov$. Then
-there is measurement / independent error $\eps \sim \normal(0,
-\sigma^2 I_m)$, with $I_m$ the $m \times m$ identity. Both error terms
-and the prior are assumed independent of each other.
-
-Usually, this is written using $\tmp := \obs \fwd$ and $\tmp^*$. Since
-we need to separate the observation operator $\obs$ from the forward
-operator $\fwd$ we write the explicit expression
-\eqref{eq:postcov}. We now turn to understand the terms present in the
-above expression.
-
-\subsubsection{Dynamics and Observation Operators}\label{subsec:dynamics}
-One can \textit{think} of $(\obs \fwd \param)_j$ as pointwise
-evaluations of a continuous function\footnote{Of course, this is
-  impossible if, e.g., $\hilo = L^2(\Omega)$ for some $\Omega$.}. Such
-measurements can be represented using the linear functionals
-$\delta_{\x_j} \in C(\Omega)^*$. In a different setting, we may be
-able to measure other quantities, e.g.\ some Fourier coefficients as
-is the case in MRI applications. Either way, the rows of $\obs$ are
-linear functionals and we refer to these as {\it measurement
-  vector}s. Note that we cannot choose every measurement vector we
-want. For example, we may be restricted only to pointwise evaluations
-of $\fwd \param$ by the sensors at our disposal, in which case
-measuring the mean $\ell(u) = \int_{\Omega}u$ (or any non-atomic
-measure) of a function is not possible.
-
-\subsubsection{Error Terms}
-Considering the sum of the error terms it is easy to see that
-$\bar{\eps} := \obs \eps' + \eps \in \R^m$. It is a centered Gaussian
-random vector and its covariance matrix is
-\begin{align}\label{eq:Sigma}
-  \begin{split}
-    \Sigma(\obs) :&= \mathbb{E}[ (\obs \eps' + \eps)  (\obs \eps' + \eps)^t ] 
-    % 
-    % 
-    = \obs \modcov \obs^* + \sigma^2I , 
-  \end{split}
-\end{align}
-where
-\begin{equation}\label{eq:modcov explained}
-  [\obs \modcov \obs^*]_{ij} = e_i^t \obs \modcov \obs^* e_j = \meas_i (\modcov \meas_j),
-\end{equation}
-with $e_j \in \R^m$ the $j$th standard basis vector. The first
-equality in \eqref{eq:modcov explained} holds by definition and the
-second by \eqref{eq:obs*}. The explicit dependence on $\obs$ will be
-mostly dropped for notational convenience, so $\Sigma(\obs) =
-\Sigma$. Thus, for a fixed $\obs$ (i.e.\ a fixed set of measurements)
-we can equivalently write $\data = \obs \fwd \param + \bar{\eps}$ with
-$\bar{\eps} \sim \normal(0,\Sigma)$. Taking $\modcov = 0$ is common
-practice \cite{Tarantola05,KaipioSomersalo05,Vogel02}. This reduces to
-the case where we take iid observations. Then $\Sigma = \sigma^2I$ is
-simply a scalar matrix that does not depend on $\obs$. Note that
-taking an error model with a non-scalar covariance as we do here
-allows us to consider model error (modeled by $\modcov$) as well as
-measurement error (modeled by $\sigma^2$). For example, say we believe
-our forward model does not capture some small scale phenomenon.  Then
-we may express this belief by saying $\tru = \fwd + \err$, with $\fwd$
-depending on $\param$ and $\err$ depending on $\sspar$, and $\param
-\perp \sspar$. We do not know much about this effect but it is
-reasonable to assume it changes continuously in our domain. We (may
-choose to) model it as $\normal (0, \modcov)$ and take $\modcov$ to
-reflect the spatial (or other) variability we imagine $\err \sspar$
-has. Such small scale phenomenon can arise as a modeling issue, where
-we might not model the system in its entirety. It can also arise from
-a numerical source, where our discretization of the system is not fine
-enough to capture all small scale phenomena. In section
-\ref{section:non vanishing}, we will see that assuming some
-correlation in the error mitigates the clusterization phenomenon, as
-reported in the literature \cite{Ucinski05}.
-
-Finally, it is useful to record a known result regarding the posterior
-covariance operator of our inverse problem, namely
-\begin{align}\label{eq:postcov}
-  \postcov = (\prcov^{-1} + \fwd^* \obs^* \Sigma^{-1} \obs \fwd
-  )^{-1}.
-\end{align}
-
-\subsection{D-Optimal Designs in Infinite Dimensions}\label{subsec:D optimal design} 
-A D-optimal design maximizes expected KL divergence between posterior
-and prior. It is useful to recall the definition of KL divergence for
-an arbitrary prior measure:
-$$
-KL(\post||\pr) = \int \log \frac{\der \post}{\der \pr}(\param) \der \post(\param).
-$$
-
-The meaning of D-optimal design in infinite-dimensional Hilbert spaces
-was investigated in \cite{AlexanderianGloorGhattas14}. There, the
-authors make assumptions that amount to $\Sigma=I$ (implied by
-$\modcov = 0,\sigma^2=1$), but we choose not to take these
-simplification here. This is because $\Sigma$ can determine ``how
-much'' clusterization we see. As stated
-\cite[pp. 681]{AlexanderianGloorGhattas14}, the results hold for more
-general covariance matrices. The conclusion is that in
-infinite-dimensions, a D-optimal design is well-defined as maximizing
-the expected KL divergence between posterior and prior. The main
-result is summarized, using our notation, in the following theorem:
-\begin{theorem}[Slightly modified Theorem 1 from \cite{AlexanderianGloorGhattas14}]
-  Let $\pr = \normal(\prmean,\prcov)$ be a Gaussian prior on $\hilp$
-  and let $\post = \normal(\postmean,\postcov)$ the posterior measure
-  on $\hilo$ for the Bayesian linear inverse problem $\data = \obs
-  \fwd\param + \obs \eps' + \eps$ discussed above. Then
-  \begin{align}\label{eq:objective}
-    \begin{split}
-      \tar( \obs) :&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
-      % 
-      % 
-      % 
-      &= \frac12 \log \det 
-      ( I + \prcov^{1/2}  \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}).
-    \end{split}
-  \end{align}
-\end{theorem}
-\begin{definition}\label{def:d optimality}
-  A design $\obs^{\star}$ is said to be D-optimal if $\obs^{\star} = \argmax_{\obs} \tar(\obs)$.
-\end{definition}
-
-The intuition behind definition \ref{def:d optimality} is
-straightforward. From \eqref{eq:postcov}:
-\begin{align*}
-  \begin{split}
-    \det ( I + \prcov^{1/2}  \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}) &= \det \Big( \prcov ( \prcov^{-1} + \fwd ^* \obs^* \Sigma^{-1} \obs \fwd) \Big )\\
-    &= \det \prcov \det \postcov^{-1}.
-  \end{split}
-\end{align*}
-Since $\prcov$ is constant, a D-optimal design minimizes the posterior
-covariance determinant, analogously to the finite-dimensional case.
-
-\subsection{Sequential vs Simultaneous Optimization}\label{subsec:seq vs sim}
-From defintion \ref{def:d optimality} we wish to characterize solution(s) of the
-following optimization problem for $\tar$. %%: (\hilo^*)^m \to \R$:
-\begin{align}\label{eq:optimization}
-  \obs^{\star} := \argmax_{\obs} \tar( \obs ) 
-  = \argmax_{\obs} \frac12 \log \det 
-  (I + \prcov^{1/2} \fwd^*\obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}),
-\end{align}
-where $\obs$ is constrained to some allowed set of measurements (we
-may \textit{think} of these as pointwise evaluations, even though
-these might not be in $\hilo^*$). We call this problem ``simultaneous
-optimization'', since all measurements are decided on simulatneously.
-
-For computational reasons, one may prefer to find the best
-measurements in a sequential manner. Denote
-\begin{equation}\label{eq:def obs_k}
-  \obs_k := (\meas_1,\dots, \meas_k)^t,  k\leq m.
-\end{equation}
-Sequential optimal design proceeds as follows. Find $\meas_1$ by
-maximizing $\tar(\obs_1)$. Then, keeping $\meas_1$ fixed --- find
-$\meas_2$ as the maximizer of $\tar(\obs_2)$. Then, find $\meas_3$ by
-keeping $\meas_1,\meas_2$ fixed and taking $\meas_3$ the maximizer of
-$\tar(\obs_3)$. Continue in this fashion until $\obs_m = \obs$ is
-found, where $m$ is the number of available measurements. It is
-important to notice that this scheme does not require actually
-observing data --- in \eqref{eq:objective} data is averaged out.
-
-The analysis in this paper is conducted for the general simultaneous
-optimization case. The sequential optimization case is dealt with in
-section \ref{subsec:seq vs sim}. It is important to note, however that
-all conclusions we arrive at for the simultaneous case easily
-specialize to the sequential case by considering the posterior as the
-next sequential step's prior.
-
-\section{The Constrained Optimization Problem of D-Optimal Design}\label{section:D and grad}
-
-\subsection{Overview}
-We seek a formulation of the D-optimal design problem
-\eqref{eq:optimization} using Lagrange multipliers. In section
-\ref{section:objective} we find the gradient of $\tar$ and record it
-in \eqref{eq:tar grad}. In section \ref{subsec:unit length} we suggest
-using unit-norm constraints on rows of $\obs$ so that the optimization
-problem in \eqref{eq:optimization} is analytically tractable. We find
-gradients for the new constraints and record them in \eqref{eq:grad
-  constraints}. Finally, in section \ref{subsec:necessary} we use
-these to formulate the D-optimal design problem as a Lagrange
-multipliers problem in equation \eqref{eq:conditions}.
-
-\subsection{The Objective and its Gradient}\label{section:objective}
-In order to use Lagrange multipliers, we need to find the gradient of
-$\tar$. This section is mostly technical and devoted to calculating
-said gradient. The result is recorded in \eqref{eq:tar grad}.
-
-We start by calculating the first variation. Some calculations are
-delegated to Lemma \ref{lemma:aux calc} in the appendix. We use
-the following lemma:
-\begin{restatable*}[Generalized from \cite{Lax97}]{lemma}{lax}\label{lemma:lax}
-  Let $Y(t)$ be a differentiable operator-valued function. Assume 
-  $I+Y(t)$ is invertible, $Y(t)$ self-adjoint and trace-class. Then
-  \begin{equation*}
-    \frac{\der \log \det (I+Y(t))}{\der t} = \tr{(I+Y(t))^{-1} \dot{Y}(t)}.
-  \end{equation*}
-\end{restatable*}
-
-The proof of \ref{lemma:lax} is delegated to the appendix. Let
-$T(\obs) := \obs^*\Sigma(\obs)\obs$. The first variation of $\tar
-(\obs)$ in the direction $V$ is:
-\begin{align}\label{eq:tar var}
-  \begin{split}
-    \delta \tar(\obs) V 
-    :&= \frac{\der}{\der\tau} \Big |_{\tau=0} \tar(\obs + \tau V)\  \text{ (by definition of variation)}\\
-    % 
-    % 
-    % 
-    &= \frac12 \frac{\der}{\der \tau} \Big |_{\tau=0} \log \det 
-    (I + \prcov^{1/2} \fwd^* T(\obs+\tau V)\fwd \prcov^{1/2} ) \text{ (by definition \eqref{eq:objective})}\\
-    % 
-    % 
-    % 
-    &= \frac12 \tr{( I + \prcov^{1/2} \fwd^* \obs^* \Sigma^{-1}
-    \obs\fwd \prcov^{1/2} )^{-1}
-    \frac{\der}{\der \tau} \Big |_{\tau=0}
-    \prcov^{1/2} \fwd^* T(\obs+\tau V) \fwd \prcov^{1/2}}\ \text{ (by \ref{lemma:lax})}\\
-    % 
-    % 
-    % 
-    &= \frac12 \ttr\Big \{ \postcov \fwd^* (V^* \Sigma^{-1} \obs 
-    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
-    &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs 
-    + \obs^* \Sigma^{-1} V ) \fwd \Big \}  \text{ (by \ref{lemma:aux calc})}\\
-    %
-    %
-    %
-    &= \tr{\postcov \fwd^* ( \obs^* \Sigma^{-1} V -
-    \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs ) \fwd} \text{
-        (cyclic property of trace)} \\
-    %
-    %
-    % 
-    &= \tr{\postcov \fwd^* \obs^* \Sigma^{-1} V 
-    ( I - \modcov \obs^* \Sigma^{-1}\obs ) \fwd} \\
-    % 
-    %
-    %
-    &= \tr{V ( I - \modcov \obs^* \Sigma^{-1}\obs )
-    \fwd \postcov \fwd^* \obs^* \Sigma^{-1}}.
-  \end{split}
-\end{align} 
-Denote
-\begin{align}\label{eq:tar grad}
-  \nabla \tar(\obs) &:= (I - \modcov \obs^* \Sigma^{-1} \obs) \fwd
-  \postcov \fwd^* \obs^*\Sigma^{-1},
-\end{align}
-the gradient of $\tar (\obs )$ and we now justify this definition.
-Since the trace of $A := V \nabla \tar(\obs) \in \R^{m \times m}$ is
-just $\ttr A = \sum_{j=1}^m e_j^t A e_j$ (with $e_j$ the $j$th standard
-basis vector), we see that
-\begin{align*}
-  \delta \tar(\obs)V = \tr{V \nabla \tar(\obs)} = \sum_{j=1}^m
-  V_j(\nabla \tar(\obs)_j),
-\end{align*}
-with $V_j \in \hilo^*$ and $\nabla \tar(\obs)_j \in \hilo^{**} =
-\hilo, j=1,\dots,m$. Thus, $\nabla \tar( \obs ) \in \hilo^m$ is indeed
-the correct gradient and the notation \eqref{eq:tar grad} is
-justified. Note that by the comment following \eqref{eq:obs*}, we view
-$V$ (defined on the same space as $\obs$) as a column vector. Then the
-gradient $\nabla \tar(\obs)$ should be viewed as a row vector, as the
-product $V \nabla \tar$ is in $\R^{m \times m}$. This will prove
-important in section \ref{subsec:necessary}.
-
-\subsection{Unit Length Constraints and their Gradients}\label{subsec:unit length}
-In this section we suggest relaxed constraints in
-\eqref{eq:constraints}. We then find their gradient in \eqref{eq:grad
-  constraints}. This is for using them in a Lagrange multipliers
-problem.
-
-As mentioned before, we cannot choose any $\obs$ when maximizing
-$\tar(\obs)$. Recall $\obs = (\meas_1,\dots,\meas_m)^t \in
-(\hilo^*)^m$ and each $\meas_j,j=1,\dots,m$ must be chosen from some
-allowed set of functionals in $\hilo^*$. This set differs based on the
-kind of sensors we have at our disposal and the properties of $\hilo$,
-as discussed in section \ref{subsec:dynamics}. The following
-proposition gives better understanding of the constraints.
-\begin{restatable*}{proposition}{biggerbetter}\label{prop:bigger better}
-  Let $\obs = (\meas_1,\dots,\meas_m)^t$, $j \in \{1,\dots,m\}$, $\sigma^2
-  > 0$ and $\lambda > 1$. Then $\tar(\obs)$ increases if we use
-  $\lambda \meas_j$ in $\obs$ instead of $\meas_j$.
-\end{restatable*}
-Proof of Proposition \ref{prop:bigger better} is delegated to the
-appendix but the idea behind it is simple: making rows of $\obs$ large
-is equivalent to making $\sigma^2$ small. %%  Viewed another way, it
-%% increases the Signal to Noise Ratio (SNR) --- which is always
-%% desirable when taking any measurement prone to error.
-This can be understood easily from the formulation of the inverse
-problem $\data = \obs \fwd \param + \obs \eps' + \eps$, where $\eps'$
-can be taken as model error and $\eps$ is iid measurement error.
-
-We now suggest a set of constraints that will make the model
-analytically tractable. By Proposition \ref{prop:bigger better}, we
-cannot just take all $\hilo^*$ to be the allowed set of measurement
-vectors, since this would effectively eliminate measurement error
-entirely. We note that for point evaluations, the norm of the
-measurement is always one:
-\begin{align*}
-  \| \delta_{\x} \| = \sup_{0 \neq u \in C(\Omega)} \frac{
-    |\int_{\Omega}u(\y) \delta_{\x}(\y) \der \y| 
-  }{
-    \sup|u|}
-  = \sup_{0 \neq u \in C(\Omega)} \frac{|u(\x)|}{ \sup|u|} = 1,
-  \forall \x \in \Omega.
-\end{align*}
-Following this, we also restrict the norm of our measurement vectors
-to be one. The unit norm constraints can be written using
-\eqref{eq:obs*} as a series of $m$ equality constraints (one for each
-measurement) on $\obs$ as
-\begin{align}\label{eq:constraints}
-  \phi_j(\obs) :=\frac12 \| \obs^* e_j\|_{\hilp}^2 - \frac12 = 0,j=1,\dots,m.
-\end{align}
-The first variations are:
-\begin{align*}
-  \delta \phi_j(\obs)V  
-  &= \frac12\lim_{\tau \to 0}\tau^{-1}
-  ( \|(\obs + \tau V)^*e_j \|_{\bar{\hilp}}^2 - \|\obs ^*e_j \|_{\bar{\hilp}}^2  ) \\
-  %
-  %
-  %
-  &= \frac12\lim_{\tau \to 0}\tau^{-1}
-  ( \langle (\obs + \tau V)^*e_j, (\obs + \tau V)^*e_j \rangle_{\bar{\hilp}} - 
-  \langle \obs^*e_j, \obs^*e_j \rangle_{\bar{\hilp}} ) \\
-  % 
-  % 
-  %
-  &= \frac12\lim_{\tau \to 0}\tau^{-1}
-  (2\tau \langle \obs^*e_j,V^*e_j \rangle_{\bar{\hilp}} 
-  +\tau^2 \langle V^*e_j, V^*e_j \rangle_{\bar{\hilp}} ) \\
-  %
-  %
-  % 
-  &= \langle \obs^*e_j,V^*e_j \rangle_{\bar{\hilp}} \\
-  %
-  %
-  % 
-  &= \langle V \obs^*e_j,e_j \rangle_{\R^m} \\
-  %
-  %
-  %
-  &= e_j^t V \obs^* e_j \\
-  % 
-  %
-  %
-  &= \tr{V \obs^* e_je_j^t}.
-\end{align*}
-Using the same arguments we used to justify the definition of
-\eqref{eq:tar grad}, we conclude that
-\begin{align}\label{eq:grad constraints}
-\nabla \phi_j(\obs) = \obs^* e_j e_j^t = \meas_j e_j^t , j=1,\dots,m,
-\end{align}
-with $\nabla \phi_j(\obs) \in \hilo^m$ and where the
-last equality follows since $\obs^*e_j = \meas_j$ by \eqref{eq:obs*}. As
-we noted at the end of section \ref{section:objective}, the
-gradient $\nabla \phi_j(\obs)$ is a row vector.
-
-\subsection{Necessary Conditions for Optimal Design}\label{subsec:necessary}
-Necessary conditions for D-optimality are found using Lagrange
-multipliers:
-\begin{align}
-  &\nabla \tar(\obs) = \sum_{j=1}^m \xi_j \nabla \phi_j (\obs)
-  \label{eq:Lagrange mult1} \\
-    &\phi_j(\obs) = 0, j = 1,\dots,m. \label{eq:Lagrange mult2}
-\end{align}
-
-Recall where the objects involved are defined: the observation operator,
-$\obs \in (\hilo^*)^m$. The objective $\tar:
-(\hilo^*)^m \to \R$ and its gradient is $\nabla \tar(\obs) \in
-\hilo^m$. The constraints are defined similarly --- $\phi_j: (\hilo^*)^m
-\to \R$ and $\nabla \phi_j(\obs) \in \hilo^m$.
-
-Focusing on \eqref{eq:Lagrange mult1} and using the gradients of the
-objective and constraints \eqref{eq:tar grad} and \eqref{eq:grad
-  constraints} we have:
-\begin{equation*}
-  (I - \modcov \obs^* \Sigma^{-1} \obs) \fwd \postcov \fwd^* \obs^*\Sigma^{-1}
-  = \sum_{j=1}^m \xi_j \obs^* e_je_j^t = (\xi_1 \meas_1,\dots,\xi_m \meas_m).
-\end{equation*}
-If we let $\Xi = \diag(\xi_j)$, then with some abuse of
-notation, we can write this more compactly as:
-\begin{equation}\label{eq:conditions}
-  ( I - \modcov \obs^* \Sigma^{-1} \obs) \fwd \postcov \fwd^* \obs^*  \Sigma^{-1}
-  = \obs^* \Xi.
-\end{equation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SECTION Analysis of Optimal Designs --- Vanishing Model Error
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Analysis of D-Optimal Designs --- Vanishing Model Error}\label{section:vanishing}
-\subsection{Overview}
-In this section we analyze D-optimal designs when no model error is
-present. We discuss sensor clusterization in section
-\ref{section:clusterization}. We prepare the ground by first deriving
-a nonlinear eigenvalue problem in $\obs$ for D-optimal designs in
-section \ref{subsec:eigenproblem}. This nonlienar eigenvalue problem
-is recorded in \eqref{eq:mod conditions}. Next, we characterize
-D-optimal designs as ones that make all eigenvalues equal in the above
-mentioned nonlinear eigenvalue problem. The derivation is presented in
-section \ref{subsec:characterization}.
-
-\subsection{The Nonlinear Eigenvalue Problem}\label{subsec:eigenproblem}
-$\modcov = 0$ implies $\Sigma= \sigma^2I$. The necessary first-order
-condition for D-optimality \eqref{eq:conditions} becomes
-\begin{equation}\label{eq:eigenproblem}
-  \sigma^{-2}\fwd \postcov \fwd^* \obs^* = \obs^* \Xi, 
-\end{equation}
-with $\Xi$ diagonal. This looks like an eigenvalue problem for the
-self-adjoint operator $\sigma^{-2}\fwd \postcov \fwd^*$. The rows of
-$\obs$, namely $\meas_j,j=1,\dots, m$, are the eigenvectors. Note,
-however, that $\postcov$ depends on $\obs$. This makes
-\eqref{eq:eigenproblem} a nonlinear eigenvalue problem.
-
-Below, the expression for the posterior covariance \eqref{eq:postcov}
-justifies the second equality. Some calculations delegated to Lemma
-\ref{lemma:twice woodbury} in the appendix justify the third (trivial
-if $\fwd$ is invertible):
-\begin{align}\label{eq:mod conditions}
-  \begin{split}
-  \obs^* \Xi &= \sigma^{-2}\fwd \postcov \fwd^* \obs^*  \\
-  %
-  %
-  %
-  &= \sigma^{-2} \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* \obs^* \\
-  %
-  %
-  %
-  &= \sigma^{-2} \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1} \obs^*
-  \end{split}
-\end{align}  
-The term $(\fwd \prcov \fwd^*)^{-1}$ is the prior precision in $\hilo$
-(and is invertible, as mentioned in section
-\ref{section:prelim}). $(\fwd\prcov\fwd^*)^{-1} + \sigma^{-2} \obs^*
-\obs$ is the posterior precision in $\hilo$.
-
-\subsection{Characterizing D-Optimal Designs}\label{subsec:characterization}
-Denote eigenvalues of $\fwd \prcov \fwd^*$ by
-$\{\lambda_i\}_{i=1}^{\infty}$. These are all positive. Denote
-corresponding eigenvectors $\{\ev_i\}_{i=1}^{\infty}$. Lemma
-\ref{lemma:sim diag} (proved in the appendix) is crucial for the rest
-of the analysis:
-
-\begin{restatable*}[Simultaneous diagonizability for the nonlinear eigenvalue problem]{lemma}{simdiag}\label{lemma:sim diag}
-  Let $\hil$ separable Hilbert space, $C:\hil \to \hil$ self-adjoint
-  and $\func_1,\dots,\func_m \in \hil$. Denote $\func^*$ the element
-  $\func$ acting as a linear functional. If
-  \begin{equation*}
-   (C + \sum_{j=1}^m \func_j\func_j^*) \func_l = \xi_l \func_l, l = 1,\dots,m
-  \end{equation*}
-  then $C$ and $\sum_{j=1}^m \func_j \func_j^*$ are simultaneously
-  diagonalizable.
-\end{restatable*}
-
-\begin{corollary}[Eigenvectors of a D-optimal design]\label{cor:same ev}
-  $\obs^*\obs$ has the same eigenvectors as $\fwd \prcov \fwd^*$.
-  %namely $\{\ev_i\}_{i=1}^{\infty}$.
-\end{corollary}
-\begin{proof}
-In \eqref{eq:mod conditions}, take $\func_j^{*} = \meas_j$ and $C :=
-(\fwd \prcov \fwd^*)^{-1}$ and use Lemma \ref{lemma:sim diag}.
-\end{proof}
-At most $m$ of the eigenvectors of $\obs^*\obs$ have a positive
-eigenvalue, because $k := \rank \obs^*\obs \leq m$. Denote the
-corresponding non-zero eigenvalues of $\obs^*\obs$ by
-$\{\eta_i\}_{i=1}^{k}$ and let $\eta_i = 0$ for $i \geq k+1$. No
-assumption is made (yet) regarding the magnitudes of the corresponding
-$\{\lambda_i\}_{i=1}^{\infty}$. They are to be thought of as fixed but
-unknown. Using Sylvester's Determinant Theorem (the standard proof
-works for Hilbert spaces as well) we get:
-
-\begin{align}\label{eq:pre manipulations}
-  \begin{split}
-    \det \left (I + \sigma^{-2} \prcov^{1/2} \fwd ^* \obs^*
-    \obs \fwd \prcov^{1/2}\right ) &=
-    %
-    \det \left (I + \sigma^{-2} \obs^*
-    \obs \fwd \prcov\fwd^* \right ) \\
-    %
-    %
-    %
-    &= \det \left \{ \left ( (\fwd \prcov \fwd^*)^{-1} +
-    \sigma^{-2}\obs^*\obs \right ) (\fwd \prcov \fwd^*)\right \}\\
-  \end{split}
-\end{align}
-
-By definition of the objective \eqref{eq:objective} and using
-\eqref{eq:pre manipulations}:
-\begin{align}\label{eq:mod objective}
-  \begin{split}
-    \tar ( \obs ) &=\frac12 \log \det \left( I + \sigma^{-2}
-    \prcov^{1/2} \fwd ^* \obs^* \obs \fwd \prcov^{1/2} \right )\\
-    %
-    %
-    %
-    %
-    &=\frac12 \log \det \Big ( \left ( (\fwd \prcov \fwd^* )^{-1} +
-    \sigma^{-2}\obs^* \obs \right ) \fwd \prcov \fwd^* \Big )\\
-    %
-    %
-    %
-    &=\frac12 \log \left (\prod_{i=1}^{\infty} ( \lambda_i^{-1} + \sigma^{-2} \eta_i ) \prod_{i=1}^{\infty} \lambda_i \right ) \\
-    %
-    %
-    %
-    &=\frac12 \log \left ( \prod_{i=1}^{k} ( \lambda_i^{-1} + \sigma^{-2} \eta_i )\prod_{i=1}^{k} \lambda_i \right )\\
-    %
-    %
-    %
-    &=\frac12 \sum_{i=1}^{k} \log (\sigma^2\lambda_i^{-1} + \eta_i), 
-  \end{split}
-\end{align}
-and the last equality holds up to an additive constant independent of
-$\obs$. Therefore, a D-optimal design maximizes
-
-\begin{equation}\label{eq:true target}
-  \sum_{i=1}^{k}\log (\sigma^2 \lambda_i^{-1} + \eta_i),
-\end{equation}
-
-with the norm constraints on $\meas_j,j=1,\dots,m$. The problem of
-D-optimal design has now been cast to finite dimensions. This
-justifies using linear algebra. Lemma \ref{lemma:free} (given a
-constructive proof in the appendix) does exactly that.
-\begin{restatable*}[Unit norm decomposition]{lemma}{free}\label{lemma:free}
-  Let $M \in \R^{k \times k}$ symmetric positive definite with $\ttr M
-  = m$, $m \geq k$. We can find $\func_j \in \R^k,j=1,\dots,m$
-  with $\|\func_j\|=1$ and $A = (\func_1,\dots,\func_m)$ such that
-  $AA^t = M$.
-\end{restatable*}
-
-Take any $\{\eta_i\}_{i=1}^{k}$, with $\sum_{i=1}^k \eta_i = m$. Let
-$M = \diag(\eta_1,\dots,\eta_k)$. Lemma \ref{lemma:free} gives us $A$
-with unit norm columns such that $AA^T = M$. Take $\meas_i
-:=\sum_{j=1}^k A_{ji}\ev_j^*$. The constraints $\|\meas_i\| = 1,
-i=1,\dots,m$ are satisified by construction. Verifying $\obs^*\obs
-\ev_i = \eta_i \ev_i$ is not too difficult.
-
-%% \begin{align*}
-%%   \begin{split}
-%%     \obs \ev_i &= (\meas_1,\dots,\meas_k)^t\ev_i\\
-%%     &= (\sum_{j=1}^kA_{j1}\ev_j^*,\dots,\sum_{j=1}^kA_{jk}\ev_j^*)^t\ev_i\\
-%%     &= (\sum_{j=1}^kA_{j1}\ev_j^*\ev_i,\dots,\sum_{j=1}^kA_{jk}\ev_j^*\ev_i)^t\\
-%%     &= (A_{i1},\dots,A_{ik})^t.
-%%   \end{split}
-%% \end{align*}
-
-%% Now,
-
-%% \begin{align*}
-%%   \begin{split}
-%%     \obs^*\obs \ev_i &= \obs^*(A_{i1},\dots,A_{ik})^t\\
-%%     &=(A_{i1},\dots,A_{ik}) \obs \text{ (by \eqref{eq:obs*})}\\
-%%     &=(A_{i1},\dots,A_{ik}) (\sum_{j=1}^kA_{j1}\ev_j^*,\dots,\sum_{j=1}^kA_{jk}\ev_j^*)^t\\
-%%     &=\sum_{l=1}^k A_{il} \sum_{j=1}^kA_{jl}\ev_j^* \\
-%%     %&=\sum_{j=1}^k  \sum_{l=1}^kA_{il}A_{jl}\ev_j^* \\
-%%     &=\sum_{j=1}^k  [AA^t]_{ij}\ev_j^* \\
-%%     %&=\eta_i\ev_i^*\\
-%%     &=\eta_i\ev_i^*. 
-%%   \end{split}
-%% \end{align*}
-
-We have shown that we can realize any $\{\eta_i\}_{i=1}^k$, as long as
-$\sum_{i=1}^k \eta_i = m$. We now turn to understand which choice of
-$\{\eta_i\}_{i=1}^k$ is optimal. By concavity of $\log$, the fastest
-increase in the design criterion \eqref{eq:true target} is gained by
-increasing $\sigma^2\lambda_i^{-1} + \eta_i$ where it is smallest. So,
-no weight should be given to (i.e. measurement taken in direction of)
-the $n$th eigenvector before $\sigma^2\lambda_i^{-1} + \eta_i \geq
-\sigma^2\lambda_n^{-1},\ \forall i \neq n$. Thus, we should choose
-$\{\eta_i\}_{i=1}^k$ as follows. First, order the eigenvalues
-$\lambda_1 \geq \lambda_2 \geq \dots$. We should increase $\sigma^2
-\lambda_1^{-1} + \eta_1$ until it equals
-$\sigma^2\lambda_2^{-1}$. Then, $\sigma^2 \lambda_i^{-1} +
-\eta_i,i=1,2$ are increased until $\sigma^2 \lambda_i^{-1} + \eta_i =
-\sigma^2\lambda_3^{-1},i=1,2$ and so forth.
-
-\pgfplotstableread{
-  Label     prior  optimal  sub-optimal 
-  1         0.2    1.8           1.7
-  2         0.8    1.2           0.8
-  3         2.2    0             0.5
-  4         3.5    0             0.0
-}\optimalvsnot
-
-\begin{figure}
-  \begin{tikzpicture}[scale=0.85]
-    \begin{axis}[
-        ybar stacked,
-        ymin=0,
-        ymax=4,
-        xtick=data,
-        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
-        reverse legend=false, % set to false to get correct display, but I'd like to have this true
-        xticklabels from table={\optimalvsnot}{Label},
-        xticklabel style={text width=2cm,align=center},
-        legend plot pos=right,
-        ylabel=precision --- prior and posterior,
-        xlabel=eigenvector $i$,
-      ]
-
-      \addplot [fill=green!80]  table [y=prior,   meta=Label, x expr=\coordindex] {\optimalvsnot};
-      \addplot [fill=blue!60]   table [y=optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};
-
-      \addlegendentry{$\sigma^2\lambda_i^{-1}$}
-      \addlegendentry{optimal $\eta_i$s}
-
-    \end{axis}
-  \end{tikzpicture}
-  \qquad
-  \begin{tikzpicture}[scale=0.85]
-    \begin{axis}[
-        ybar stacked,
-        ymin=0,
-        ymax=4,
-        xtick=data,
-        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
-        reverse legend=false, % set to false to get correct display, but I'd like to have this true
-        xticklabels from table={\optimalvsnot}{Label},
-        xticklabel style={text width=2cm,align=center},
-        legend plot pos=right,
-        ylabel=precision --- prior and posterior,
-        xlabel=eigenvector $i$,
-      ]
-
-      \addplot [fill=green!80]  table [y=prior,       meta=Label, x expr=\coordindex] {\optimalvsnot};
-      \addplot [fill=blue!60]   table [y=sub-optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};
-
-      \addlegendentry{$\sigma^2\lambda_i^{-1}$}
-      \addlegendentry{sub-optimal $\eta_i$s}
-
-    \end{axis}
-  \end{tikzpicture}
-  \caption{Characterizing a D-optimal design, compared to a
-    sub-optimal design. Both designs are allowed three measurements,
-    which means the blue area has accumulated height of $3$ in both
-    plots. A D-optimal design (left) increases precision (reduces
-    uncertainty) where it is lowest. A sub-optimal design (right) does
-    not. In this example, a measurement is taken partly in the
-    direction of the third eigenvector. The precision for this
-    eigenvector is bigger even than the posterior precision of the
-    D-optimal design on the left.}
-  \label{fig:optimal vs not}
-\end{figure}
-
-\pgfplotstableread{
-  Label    prior  $o_1$  $o_2$   $o_3$   $o_4$   $o_5$         $o_6$    topper
-  1         0.2    0.8    0.5     0.5    0.4    0.33333333  0.3333333  0.001
-  2         0.8    0.2    0.5     0.5    0.4    0.33333333  0.3333333  0.001
-  3         2.2    0      0       0      0.2    0.33333333  0.3333333  0.001
-  4         3.5    0      0       0      0      0           0          0.001
-}\clusterization
-
-\pgfplotstableread{
-  Label    prior  $o_1$  $o_2$   $o_3$   $o_4$   $o_5$         $o_6$    topper
-  1         0.2    0.8    0.7     0.3    0.4    0.2         0.4666666  0.001
-  2         0.8    0.2    0.3     0.7    0.4    0.46666666  0.2        0.001
-  3         2.2    0      0       0      0.2    0.33333333  0.3333333  0.001
-  4         3.5    0      0       0      0      0           0          0.001
-}\noclusterization
-
-The argument above has the important consequence (illustrated in
-figure \ref{fig:optimal vs not}) that the eigenvalues in the nonlinear
-eigenvalue problem \eqref{eq:eigenproblem} are all equal! This makes
-perfect sense. We know the Lagrange multipliers $\xi_i,i=1,\dots,k$
-tell us how much we gain by relaxing the constraints and there is no
-reason for any eigenvector to give us more than any other (otherwise
-we would put more weight on the better eigenvector, as can be
-understood from the discussion in the following paragraph).
-
-One can understand D-optimal designs as follows: Imagine each
-eigenvector corresponds to a graduated lab cylinder (figure
-\ref{fig:cylinder}). Each cylinder $i$ is filled, a-priori, with green
-liquid of $\sigma^2\lambda_i^{-1}$ volume units. This liquid
-represnets the prior precision, weighted by observation noise. The
-volume of the green liquid is increasing in $i$. A D-optimal design
-with $m$ measurements ignores cylinders indexed $m+1$ and above. It
-has blue liquid of $m$ volume units at its disposal. It distributes
-said liquid by repeatedly adding a drop to whatever cylinder currently
-has the lowest level of liquid in it (lowest precision). The result of
-such a procedure is illustrated in figure \ref{fig:optimal vs
-  not}. The analysis conducted in this section is summarized in
-Theorem \ref{thm:char} below.
-
-\begin{theorem}[Characterizing D-optimal designs with vanishing model error]\label{thm:char}
-  Let:
-  \begin{enumerate}
-  \item $\fwd, \prcov$ the forward and prior covariance operators, respectively.
-  \item $m, \sigma^2$ the number of measurements taken and observation
-    noise variance, respectively.
-  \item $\obs$ a D-optimal design operator.
-  \item $\tar$ as definded in \eqref{eq:objective}.
-  \item $\{\lambda_i\}_{i=1}^\infty$ eigenvalues of $\fwd\prcov\fwd^*$
-    in decreasing order.
-  \item $\{\ev_i\}_{i=1}^\infty$ their corresponding eigenvectors.
-  \item $\{\eta_i\}_{i=1}^\infty$ eigenvalues of $\obs^*\obs$, where
-    $\obs$ maximizes $\tar(\obs)$.
-  \end{enumerate}
-  Then:
-  \begin{enumerate}
-  \item $\obs^*\obs$ and $\fwd\prcov\fwd^*$ are simultaneously diagonalizable.
-  \item There exists $0 < k \leq m$ such that $\eta_i = 0$ for $i\geq
-    k+1$ and $k = \rank \obs^*\obs$.
-  \item $\sigma^2\lambda_i^{-1} + \eta_i$ are equal for all $1 \leq i \leq k$.
-  \item If $k < m$, then $\sigma^2\lambda_i^{-1} + \eta_i \leq
-    \sigma^2\lambda_{k+1}^{-1}, i=1,\dots,k$.
-  \end{enumerate}
-\end{theorem}
-
-\begin{figure}%{r}{0.25\textwidth} 
-    %\centering
-    \includegraphics[width=5cm, height=5cm]{cylinders.jpg}
-    \caption{Three graduated lab cylinders, corresponding to three
-      eigenvectors. Prior eigenvalues not shown.}
-    \label{fig:cylinder}
-\end{figure}
-
-\section{Sensor Clusterization --- Vanishing Model Error}\label{section:clusterization}
-\subsection{Overview}
-In this section we use Theorem \ref{thm:char} to explain how sensor
-clusterization may arise in various scenarios. We first consider
-sequential designs, and then simultaneous designs.
-
-\subsection{Sequential Design}\label{subsec:clusterization sequential}
-Recall the sequential optimal design scheme from section
-\ref{subsec:seq vs sim} and the definition of $\obs_k$ from
-\eqref{eq:def obs_k}. The way clusterization can occur in sequential
-design scenario is understood by specializing thorem \ref{thm:char} to
-one measurement each time. In this case $m=1$, so $k=1$, $\meas_1
-\parallel \ev_1$ and $\eta_1 = 1$. As hinted in section
-\ref{subsec:seq vs sim}, the posterior becomes the prior for the next
-step. The eigenvectors of the new posterior are the same as for the
-previous one. A new measurement will be taken in the direction of
-eigenvector with smalles eigenvalue. This leads to:
-\begin{align*}
-  \begin{split}
-    \meas_2 \parallel \ev_1  &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 < \sigma^2\lambda^{-1}_2 \\
-    \meas_2 \parallel \ev_2  &\text{ o.w. }
-  \end{split}
-\end{align*}
-In the first case, we get clusterization immediately. In the second,
-we consider $\meas_3$. Same reasoning leads to the conclusion that the
-eigenvectors of the new prior precision are the same. The eigenvalues,
-however, are $\lambda^{-1}_1 + \sigma^{-2}, \lambda^{-1}_2 +
-\sigma^{-2}, \lambda_3^{-1}, \dots$. A new measurement is taken in the
-direction of eigenvector of smallest precision (note that
-$\lambda^{-1}_1 + \sigma^{-2} < \lambda^{-1}_2 + \sigma^{-2}$):
-\begin{align*}
-  \begin{split}
-    \meas_3 \parallel \ev_1 &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 <
-    \sigma^2\lambda^{-1}_3 \\
-    \meas_3 \parallel  \ev_3  &\text{ o.w. }
-  \end{split}
-\end{align*}
-Again, the first case results in clusterization. The sequential design
-proceeds in this fashion. Every $\meas_i$ is taken in the direction
-some eigenvector of the prior $\fwd \prcov\fwd^*$. This eigenvector
-has the smallest eigenvalue of the previous step's posterior precision
-operator. If we do not encounter clusterization after $m$ such steps,
-this means that measurement $\meas_k$ was taken in the direction
-$\ev_k$ for $k=1,\dots,m$. Thus:
-\begin{align}\label{eq:failure}
-  \begin{split}
-    \sigma^2\lambda_{k-1}^{-1} + 1 \geq \sigma^2\lambda_k^{-1}, \text{ for } 1\leq k \leq m
-    \Longrightarrow \lambda_m^{-1} - \lambda_{m-1}^{-1} \leq \sigma^{-2}.
-  \end{split}
-\end{align}
-
-But \eqref{eq:failure} implies $\{\lambda_i^{-1}\}_{i=1}^{\infty}$
-grows at most linearly. In section \ref{subsec:abstract OED} we
-assumed $\fwd$ is strongly smoothing, which means eigenvalues of
-$(\fwd \prcov\fwd^*)^{-1}$ grow quickly. Thus, if we take $m$ large
-enough, \eqref{eq:failure} will eventually fail and we will observe
-sensor clusterization.
-
-\subsection{Simultaneous Design}\label{subsec:clusterization simultaneous}
-In a simultaneous design scenario, the result is not simple as in the
-sequential design case. We now see that a design that exhibits
-clusterization is as optimal as one that does not. Corollary
-\ref{cor:same ev} tells us that we can realize any design we want
-(encoded in $\{\eta_i\}$), given the trace constraint for
-$\obs^*\obs$. Theorem \ref{thm:char} tells us that for D-optimality,
-we only need to make the posterior flat where we take measurements
-(recall figure \ref{fig:optimal vs not}). We are left with the freedom
-of assigning any part of $\eta_i$ to any measurement $\meas_j$. Two
-possible assignments for the same problem are illustrated in figure
-\ref{fig:clusterization}. The first exhibits clusterization and second
-does not. It is important to note that neither arises as a result of
-some numerical optimization scheme. Both designs are give the same
-design criterion $\tar$, because their posteriors are identical. By
-Theorem \ref{thm:char}, both are (possible) maximizers to the same
-D-optimal design problem. One exhibits clusterization, and one does
-not. It is possible that a real-life problem is more restricted than
-the model we consider here. Exactly in what way this happens is not
-clear and should be the subject of further research.
-
-\begin{figure}
-  \begin{tikzpicture}[scale=0.85]
-    \begin{axis}[
-        ybar stacked,
-        ymin=0,
-        ymax=4,
-        xtick=data,
-        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
-        reverse legend=false, % set to false to get correct display, but I'd like to have this true
-        xticklabels from table={\clusterization}{Label},
-        xticklabel style={text width=2cm,align=center},
-        legend plot pos=right,
-        ylabel=precision --- prior and posterior,
-        xlabel=eigenvector $i$,
-      ]
-
-      \addplot [fill=green!80]  table [y=prior, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=blue!60]   table [y=$o_1$, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=red!60]    table [y=$o_2$, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=black!60]  table [y=$o_3$, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=orange!60] table [y=$o_4$, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=cyan!60]   table [y=$o_5$, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=purple!60] table [y=$o_6$, meta=Label, x expr=\coordindex] {\clusterization};
-
-      \addlegendentry{prior}
-      \addlegendentry{$o_1$}
-      \addlegendentry{$o_2$}
-      \addlegendentry{$o_3$}
-      \addlegendentry{$o_4$}
-      \addlegendentry{$o_5$}
-      \addlegendentry{$o_6$}   
-    \end{axis}
-  \end{tikzpicture}
-  \qquad
-  \begin{tikzpicture}[scale=0.85]
-    \begin{axis}[
-        ybar stacked,
-        ymin=0,
-        ymax=4,
-        xtick=data,
-        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
-        reverse legend=false, % set to false to get correct display, but I'd like to have this true
-        xticklabels from table={\noclusterization}{Label},
-        xticklabel style={text width=2cm,align=center},
-        legend plot pos=right,
-        ylabel=precision --- prior and posterior,
-        xlabel=eigenvector $i$,
-      ]
-
-      \addplot [fill=green!80]  table [y=prior, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=blue!60]   table [y=$o_1$, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=red!60]    table [y=$o_2$, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=black!60]  table [y=$o_3$, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=orange!60] table [y=$o_4$, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=cyan!60]   table [y=$o_5$, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=purple!60] table [y=$o_6$, meta=Label, x expr=\coordindex] {\noclusterization};
-
-      \addlegendentry{prior}
-      \addlegendentry{$o_1$}
-      \addlegendentry{$o_2$}
-      \addlegendentry{$o_3$}
-      \addlegendentry{$o_4$}
-      \addlegendentry{$o_5$}
-      \addlegendentry{$o_6$}   
-    \end{axis}
-  \end{tikzpicture}
-  \caption{Clusterization (left) and non-clusterization (right) in
-    simultaneous D-optimal designs. Posterior precision per
-    eigenvector after taking measurements $\meas_i$ is plotted. Both
-    designs are identical for all practical matters --- their
-    posteriors are identical. In the left panel we see how an optimal
-    design is achieved with repeated measurements. $\meas_2 = \meas_3$
-    and $\meas_5 = \meas_6$. This is not necessary though, as can be
-    seen in the right panel. Both designs achieve the same levels of
-    uncertainty. The left exhibits clusterization. The right does not.}
-  \label{fig:clusterization}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SECTION Analysis of Optimal Designs --- Non-Vanishing Model Error
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Analysis of D-Optimal Designs --- Non-Vanishing Model Error}\label{section:non vanishing}
-In this section we show the effect model error has on the
-clusterization phenomenon. We will see that if $\modcov \neq 0$
-clustering will not occur. This is contrary to the previous case of
-vanishing model error.
-
-Corollary \ref{cor:same meas} below is proved in the appendix. Denote
-$\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm :=
-(\meas_1,\dots,\meas_{m-1})^t$. Denote $\Sigmam := \Sigma (\obsm)$ and
-$\postcovm$ the posterior covariance that arises when we use $\obsm$.
-\begin{restatable*}{corollary}{samemeas}\label{cor:same meas}
-  If $\meas_m = \meas_j$ for some $1 \leq j \leq m-1$, then
-  \begin{equation*}
-    \tar(\obs) - \tar(\obsm) =
-    \log \left ( 1 + \frac{\sigma^2
-      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
-      \obsm^* \Sigmam^{-1}e_j \rangle
-    }{
-      2 - \sigma^2 e_j^t\Sigmam^{-1}e_j 
-    }       
-    \right ).
-  \end{equation*}
-\end{restatable*}
-Recall from \eqref{eq:Sigma} that $\Sigma(\obs) = \obs \modcov \obs^*
-+ \sigma^2I$. Then
-
-$$
-\lim_{\sigma^2 \to 0} \tar(\obs) -\tar(\obsm) = 0
-$$
-
-Hence no increase in the design criterion is achieved by taking a
-repeated measurement (in the limit $\sigma^2 \to 0$). This is clearly
-sub-optimal and $\obsm$ cannot be a D-optimal design. Thus, for small
-measurement error levels, the clusterization effect will be mitigated
-by the presence of a non-zero model error. Since the design criterion
-is not defined for $\sigma^2 = 0$ and identical measurements, we
-cannot make a statement regarding $\sigma^2 = 0$, except in the
-limiting sense described above.
-
-\section{Acknowledgements}
-This study is a part of my PhD thesis. Most of it was written under
-the instruction of Prof. Georg Stadler in New York University's
-Courant Institute. I would like to thank him for his great
-mentorship. This work was supported by The Raymond and Beverly Sackler
-Post-Doctoral Scholarship.
-
-\appendix
-\section{General Lemmas}
-
-The following lemma is generalized from \cite[Chapter 9, Theorem 4, pp. 127]{Lax97}
-\lax
-\begin{proof}
-  Consider a differentiable operator-valued function $X(t)$ such that
-  $X(0) = 0$ and $X(t)$ is positive, self-adjoint and trace-class for
-  every $t\in \R$. We denote the eigenvalues of this operator by
-  $\lambda_k(X(t))$ and sometimes drop the dependence on $X(t)$, so
-  $\lambda_k = \lambda_k(X(t))$.  Then $\det (I+X(t)) =
-  \prod_{k=1}^{\infty} (1+\lambda_k) < \infty$ where the bound holds
-  by the arguments given in \cite{AlexanderianGloorGhattas14}. The
-  full derivative is
-  \begin{align*}
-    \frac{\der \det (I+X(t))}{\der t} 
-    % 
-    % 
-    % 
-    &= \sum_{k=1}^{\infty} 
-    \frac{\partial \det (I+X(s))}{\partial (1+\lambda_k)}\Big |_{s=t}
-    \frac{\der (1+\lambda_k)}{\der t} \\
-    % 
-    %
-    %
-    &= \sum_{k=1}^{\infty} \frac{\partial \prod_{l=1}^{\infty}
-      (1+\lambda_l(s))}{\partial (1+\lambda_k)}\Big |_{s=t}
-    \frac{\der (1+\lambda_k)}{\der t} \\
-    %
-    %
-    %
-    &= \sum_{k=1}^{\infty} \prod_{l=1, l\neq k}^{\infty}
-      (1+\lambda_l(s)) \frac{\partial (1+\lambda_k(s))}{\partial (1+\lambda_k)}\Big |_{s=t}
-    \frac{\der (1+\lambda_k)}{\der t} \\
-    %
-    %
-    %    
-    &= \sum_{k=1}^{\infty} \frac{\prod_{l=1}^{\infty}
-      (1+\lambda_l(s))}{(1+\lambda_k)}\Big |_{s=t}
-    \dot{\lambda_k}(X(t)) \\
-    % 
-    % 
-    % 
-    &= \sum_{k=1}^{\infty} \frac{\det (I+X(t))}{1 +\lambda_k} \dot{\lambda_k}(X(t)).
-  \end{align*}
-  The assumption $X(0) = 0$ means $\lambda_k(X(0)) = 0,\ \forall k \geq 1$. Thus:
-  \begin{align*}
-    \frac{\der (I+\det X(t))}{\der t}\Big |_{t=0} 
-    = \sum_{k=1}^{\infty} \dot{\lambda_k}(X(0)) 
-    = \frac{\der }{\der t}\tr{X(0)}
-    = \tr{\dot{X}(0)},
-  \end{align*}
-  where the second equality follows by monotone convergence. 
-  Let $Y(t)$ a trace-class self-adjoint operator such that 
-  $I+Y(t)$ is invertible.
-  Define $X(t)$ via $I+X(t) = (I+Y(0))^{-1/2} (I+Y(t)) (I+Y(0))^{-1/2}$. 
-  We show $X(t)$ satisfies the conditions above. It is trace-class:
-  \begin{align*}
-    \tr{X(t)} = \tr{(I+Y(0))^{-1} (I+Y(t)) - I}
-    \leq \tr{I+Y(t) - I}< \infty,
-  \end{align*}
-  since $Y(t)$ is trace-class. It is also clear that
-  $X(0) = 0$ and $X(t)$ is self-adjoint.
-  $I+Y(t) = (I+Y(0))^{1/2}(I+X(t))(I+Y(0))^{1/2}$, so
-  \begin{align*}
-    \frac{\der \det (I+Y(t))}{\der t}|_{t=0} 
-    &= \det (I+Y(0))\frac{\der \det (I+X(t))}{\der t}\Big |_{t=0} \\
-    % 
-    % 
-    % 
-    &= \det (I+Y(0)) \tr{\dot{X}(0)} \\
-    % 
-    % 
-    % 
-    &= \det (I+Y(0)) \tr{(I+Y(0))^{-1} \dot{Y}(0)}.
-  \end{align*}
-  Consequently, by the one-variable chain rule:
-  \begin{align*}
-    \frac{\der \log \det (I+Y(t))}{\der t}\Big |_{t=0} &=
-    % 
-    % 
-    % 
-    \frac{1}{\det (I+Y(0))}\frac{\der \det (I+Y(t))}{\der t}\Big |_{t=0} \\ 
-    % 
-    % 
-    % 
-    &= \tr{ (I+Y(t))^{-1} \dot{Y}(t)} \big |_{t=0}.
-  \end{align*}
-  There is nothing special about $t_0 = 0$ --- we could have chosen
-  any other $t_0$ instead. Thus, the relation holds for all $t$.
-\end{proof}
-
-\begin{lemma}[Matrix Determinant Lemma in Hilbert Spaces]\label{lemma:MDL}
-  Let $\hil$ a separable Hilbert space, $u,v\in \hil$ and $A: \hil \to
-  \hil$ an invertible linear operator such that $\tr{A-I} <
-  \infty$. Then $\det A$ and $\det A + uv^*$ are well defined and
-  \begin{equation*}
-    \det (A + uv^*) = (1 + \langle A^{-1} u, v \rangle ) \det A,
-  \end{equation*}
-  where $(A + uv^*)w := Aw + \langle v,w \rangle u$.
-\end{lemma}
-\begin{proof}
-  In this proof we rely on definitions and results from
-  \cite{Simon77}. First, consider $B := I + xy^*$ for some $x,y \in
-  \hil$. We construct an eigenbasis for $B$ and use that to show $\det
-  B = 1 + \langle x, y \rangle$. First let $x_1 := x$.  Now, if $x
-  \parallel y$, take $\{x_n \}_{n=2}^{\infty}$ an orthogonal basis for
-  $span\{x_1\} ^{\perp}$. If, on the other hand, $x \nparallel y$, let
-  \begin{equation*}
-    x_2 := x - \frac{ \langle x, y\rangle}{\|y\|^2}y
-  \end{equation*}
-  and it is easy to verify that $x_2 \perp y$ and $span \{x,y\} = span
-  \{x_1,x_2\}$. Take $\{x_n \}_{n=3}^{\infty}$ an orthogonal basis for
-  $span\{x_1,x_2\} ^{\perp}$. In both cases,
-  \begin{equation*}
-    B x_n =
-    \begin{cases}
-      (1 + \langle x, y \rangle) x_n & n = 1 \\
-      x_n                            & n \neq 1,
-    \end{cases}
-  \end{equation*}
-  and so $\det B = 1 + \langle x, y \rangle$.
-
-  It is easy to verify that $uv^*$ is trace-class and since $\tr{A-I}
-  < \infty$, also $\tr{A + uv^* - I} < \infty$ (sum of two trace-class
-  operators is trace-class). Thus $\det A$ and $\det (A+uv^*)$ are
-  well defined. Let $x:=A^{-1}u$ and $y := v$:
-  \begin{equation*}
-    \det (A + uv^*) = \det A \ \det(I+A^{-1}uv^*) =
-    (1 + \langle A^{-1}u, v \rangle) \det A .
-  \end{equation*}
-\end{proof}
-
-\free
-\begin{proof}
-  Let us diagonalize $M$, so that $M = U D U^t$ with $D =
-  \diag(d_1,\dots,d_k)$ and $U \in \R^{k \times k }$ orthogonal. Let
-  $S \in \R^{k \times m}$ with $S_{ii} = \sqrt{d_{i}}$ and zeros
-  otherwise. Define $A:= U S V^t$, where $V \in \R^{m \times m}$ is
-  orthogonal and will be further restricted later. Then $AA^t = U
-  SV^tVS^t U^t = UDU^t$, so $AA^t$ has the required eigenvalues and
-  eigenvectors by construction. If we can choose $V$ such that $A$
-  also satisfies the unit norm constraints we are done. These
-  constraints are, for $j=1,\dots,m$:
-  \begin{equation}\label{eq:V constraints}
-   1 = [A^tA]_{jj} = [V S^tS V^t]_{jj},
-  \end{equation}
-  and we can expect to do this since we assumed $\ttr D = m$.
-
-  Define $C = S^tS - I \in \R^{m \times m}$. Note that $\ttr C = 0$ and
-  $C$ is diagonal with non-zero entries $d_i-1,i=1,\dots,k$. It suffices
-  to find $V$ orthogonal such that $V C V^t$ has zero diagonal. We
-  construct such $V$ by sequentially inserting zeros in the diagonal
-  and not destroying zeros we already introduced, starting from the
-  last diagonal entry and moving to the first. Since $c_{mm} \neq 0$ ,
-  let $p < m$ such that $c_{pp}c_{mm} < 0$ (such $p$ exists because
-  the trace is zero) and let $\theta \in (0,\pi)$. Define a Givens
-  rotation $R^{(m)} \in \R^{m \times m}$ by
-  \begin{equation*}
-    r^{(m)}_{ab} :=
-    \begin{cases}
-      1 & a = b \neq p \text{ or } a = b \neq m \\
-      \cos \theta & a = b = p  \\
-     -\sin \theta & a = p, b = m\\
-      \cos \theta & a = b = m \\
-      \sin \theta & a = m, b = p \\ 
-      0 & o.w
-    \end{cases}
-  \end{equation*}
-  Note that conjugating a matrix by $R^{(m)}$ changes only its $m$ and
-  $p$ rows and columns. We want to choose $\theta$ such that
-  \begin{equation}\label{eq:mm}
-    0 = [R^{(m)} C (R^{(m)})^t]_{mm} = \cos^2 \theta c_{mm} + 2\cos \theta \sin
-    \theta c_{mp} + \sin^2\theta c_{pp},
-  \end{equation}
-  and it suffices to choose $\theta$ such that
-  \begin{equation*}
-    c_{mm} \cot^2 \theta + 2 c_{mp} \cot \theta + c_{pp} = 0.
-  \end{equation*}
-  This quadratic in $\cot\theta$ has a real solution, since
-  $c_{pp}c_{mm} < 0$ by assumption and we can find $\theta \in
-  (0,\pi)$ such that \eqref{eq:mm} is satisfied. We continue to find
-  $R^{(m-1)}$ that leaves row and column $m$ unchanged and
-  continue introducing zeros to the diagonal. The assumption $\ttr D =
-  m \Rightarrow \ttr C = 0$ guarantees we can do that. Taking $V:=
-  R^{(1)} R^{(2)} \dots R^{(m-1)}R^{(m)}$ completes the proof.
-\end{proof}
-
-\simdiag
-\begin{proof}
-  First, enumerate the eigenvalues of $C + \sum_{j=1}^m
-  \func_j\func_j^*$ as $\xi_1,\dots,\xi_\ell$. Denote the
-  indices of the eigenvectors corresponding to $\xi_i$
-  \begin{equation*}
-    S_i := \{ 1 \leq k \leq m | (C + \sum_{j=1}^m \func_j\func_j^* )\func_k = \xi_i \func_k \}.
-  \end{equation*}
-  Define further
-  \begin{equation*}
-    A_i := \sum_{k \in S_i} \func_k \func_k^*,
-  \end{equation*}
-  which is self-adjoint. Two observations are in order. First,
-  $\sum_{j=1}^m \func_j\func_j^* = \sum_{i=1}^\ell A_i$. Second, $A_i
-  \func_k = 0$ if $k\not \in S_i$, since eigenvectors of different
-  eigenvalue are orthogonal. For $k \in S_i$
-  \begin{equation}\label{eq:on vi}
-    \xi_i \func_k = (C + \sum_{j=1}^m \func_j \func_j^* ) \func_k = (C + A_i) \func_k.
-  \end{equation}
-  Let $V_i := span \{\func_k \}_{k\in S_i}$. Observe that $V_i$ is
-  invariant under $A_i$, by definition, and under $C$, by \eqref{eq:on
-    vi}. Using \eqref{eq:on vi} again, we conclude that on $V_i$, $A_i
-  = \xi_iI - C$. This immediately implies $A_i$ and $C$ have the
-  same eigenvectors on $V_i$. This holds for every $1 \leq i \leq
-  \ell$ and we conclude that $C$ and $A$ have the same eigenvectors.
-\end{proof}
-
-\section{Specific Lemmas}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{lemma}\label{lemma:twice woodbury}
-  Assume $\fwd \prcov \fwd^*$ is invertible. Then
-\begin{align*}
-  \begin{split}
-    \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* 
-    %
-    %
-    = \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1},
-  \end{split}
-\end{align*}  
-\end{lemma}
-\begin{proof}
-  The proof is boring and amounts to using Woodbury's matrix identity
-  twice and a regularization trick. The standard proof for this
-  identity works in infinite dimensions, as long as all terms are well
-  defined. Unfortunately, $\obs^*\obs$ is not invertible, so we force
-  it to be. Throughout, we let $X := \fwd\prcov \fwd^*$ and $Y_{\eps}
-  := \obs^*\obs + \eps I$, for some small $\eps > 0$. Note that now
-  $Y_{\eps}$ is invertible. Recall Woodbury's matrix identity:
-  $$
-  (A + UCV)^{-1} = A^{-1} + A^{-1}U(C^{-1} + VA^{-1}U)^{-1}VA^{-1}.
-  $$
-
-  Taking $A := \prcov^{-1}, U := \fwd^*, V := \fwd$ and $C :=
-  \sigma^{-2} (\obs^*\obs+\eps I)$ we get:
-  \begin{align*}
-    \begin{split}
-      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* &=
-      \fwd ( \prcov - \prcov \fwd^* ( \sigma^2(\obs^*\obs + \eps I)^{-1} + \fwd \prcov \fwd^* )^{-1} \fwd \prcov ) \fwd^* \\
-      %
-      %
-      &= X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X
-    \end{split}
-  \end{align*}
-  Note that $X + \sigma^2 Y_{\eps}^{-1}$ is invertible, as the sum of
-  a two positive definite operators. Now, taking $A := X^{-1}, C :=
-  \sigma^2Y_{\eps}^{-1}, U := I$ and $V := I$ and using Wodbury's matrix
-  identity in reverse order:
-  \begin{align*}
-    \begin{split}
-      X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X &= (X^{-1} + \sigma^{-2}Y_{\eps})^{-1} \\
-      %
-      %
-      %
-      &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \eps I))^{-1}.
-    \end{split}
-  \end{align*}
-
-  We conclude that $\forall \eps > 0$
-  \begin{align*}
-    \begin{split}
-      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* 
-     &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \eps I))^{-1}.
-    \end{split}
-  \end{align*}
-  Letting $\eps \to 0$ completes the proof.
-\end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{lemma}[Increase due to a Measurement]\label{lemma:design increase}
-  Let $\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm := (\meas_1,\dots,\meas_{m-1})^t$. Then
-  \begin{align*}
-    \tar( \obs ) - \tar (\obsm ) &=
-    \frac12 \log \left ( 1 + \frac{
-      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m,
-      (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m \rangle
-    }{
-      \sigma^2 + \meas_m \modcov \meas_m - \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m 
-    }       
-    \right ).
-  \end{align*}
-\end{lemma}
-\begin{proof}
-  We use the Schur complement to write one inverse in terms of the other and
-  introduce notations to make the derivation cleaner. Note that we think of
-  $\obsm$ and $\obsm^*$ as column and row vectors (respectively).
-  \begin{align*}
-    \Sigma( \obs ) &= \Sigma = 
-    \begin{bmatrix}
-      \Sigma (\obsm )           & \obsm \modcov \meas_m \\
-      \meas_m \modcov \obsm^*   & \sigma^2 + \meas_m \modcov \meas_m
-    \end{bmatrix}
-    =
-    \begin{bmatrix}
-      \Sigmam   & w \\
-      w^t       & c
-    \end{bmatrix}\\
-    %
-    %
-    %
-    \Sigma^{-1} &=
-    \begin{bmatrix}
-      \Sigmam^{-1} + \Sigmam^{-1} w ( c - w^t \Sigmam^{-1} w)^{-1} w^t \Sigmam^{-1} & - \Sigmam^{-1} w ( c - w^t \Sigmam^{-1} w)^{-1} \\
-      -( c - w^t \Sigmam^{-1} w)^{-1} w^t \Sigmam^{-1}                            &  ( c - w^t \Sigmam^{-1} w)^{-1}
-    \end{bmatrix} \\
-    &=
-    \begin{bmatrix}
-      \Sigmam^{-1} & 0 \\
-      0           & 0 
-    \end{bmatrix}
-    + (c -w^t \Sigmam^{-1} w )^{-1}
-    \begin{bmatrix}
-      \Sigmam^{-1} w \\
-      -1
-    \end{bmatrix}
-    \begin{bmatrix}
-      w^t \Sigmam^{-1} & -1 
-    \end{bmatrix}
-  \end{align*}
-  %
-  Further, define
-  %
-  \begin{align*}
-    \M (\obs ):&= \prcov^{\frac12}\fwd^* \obs^* \Sigma^{-1} \obs \fwd
-    \prcov^{\frac12}    
-  \end{align*}
-  %
-  and note that, using our understanding of what is a column vector and
-  what is a row vector:
-  %
-  \begin{align*}
-    \M(\obs) &= \prcov^{1/2} \fwd^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2} \\
-    %
-    %
-    %
-    &= \prcov^{1/2} \fwd^* \obs^* \left \{
-    \begin{bmatrix}
-      \Sigmam^{-1} & 0 \\
-      0           & 0 
-    \end{bmatrix}
-    + (c -w^t \Sigmam^{-1} w )^{-1}
-    \begin{bmatrix}
-      \Sigmam^{-1} w \\
-      -1
-    \end{bmatrix}
-    \begin{bmatrix}
-      w^t \Sigmam^{-1} & -1 
-    \end{bmatrix} 
-    \right \} \obs \fwd \prcov^{1/2} \\
-    %
-    %
-    %
-    &= \M (\obsm) + (c -w^t \Sigmam^{-1} w )^{-1}
-    \prcov^{1/2} \fwd^* \obs^*
-    \begin{bmatrix}
-      \Sigmam^{-1} w \\
-      -1
-    \end{bmatrix}
-    \begin{bmatrix}
-      w^t \Sigmam^{-1} & -1 
-    \end{bmatrix} 
-    \obs \fwd \prcov^{1/2}
-  \end{align*}
-  %
-  Now, denote:
-  %
-  \begin{align}\label{eq:u}
-    \begin{split}
-      u :&= (c -w^t \Sigmam^{-1} w )^{-1/2}
-      \prcov^{1/2} \fwd^* \obs^* 
-      \begin{bmatrix}
-        \Sigmam^{-1} w \\
-        -1 
-      \end{bmatrix} \\
-      %
-      %
-      %
-      & = (c -w^t \Sigmam^{-1} w )^{-1/2} ( \prcov^{1/2}\fwd^* \obsm^* \Sigmam^{-1} \obsm  \modcov \meas_m - \prcov^{1/2} \fwd^* \meas_m )\\
-      %
-      %
-      %
-      u^* :&=  (c -w^t \Sigmam^{-1} w )^{-1/2} (\meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \fwd \prcov^{1/2} - \meas_m \fwd \prcov^{1/2} ),
-    \end{split}
-  \end{align}
-  %
-  so that
-  %
-  \begin{equation}\label{eq:M plus I}
-    I + \M( \obs ) = I + \M (\obsm ) + uu^*.
-  \end{equation}
-  %
-  Note that
-  \begin{equation}\label{eq:M postcov}
-    \prcov^{1/2} \left (I + \M( \obsm ) \right )^{-1} \prcov^{1/2} = \postcovm.
-  \end{equation}
-  The increase in the design criterion gained by including $\meas_m$
-  in the design is found by using Lemma \ref{lemma:MDL} the above
-  results:
-  %
-  \begin{align*}
-    \tar( \obs ) - \tar( \obsm )
-    %
-    %
-    %
-    &= \frac12 \log \det \Big ( I + \M ( \obs ) \Big ) / \det \Big ( I + \M (\obsm) \Big ) \\
-    %
-    %
-    %
-    &= \frac12  \log \det \left ( I + \M(\obsm) + uu^* \right ) / \det \Big ( I + \M (\obsm) \Big ) \\
-    %
-    %
-    %
-    &= \frac12 \log \left ( 1 + \left \langle \left ( I+\M(\obsm) \right )^{-1} u, u  \right \rangle \right ).
-  \end{align*}
-  Using \eqref{eq:u}:
-  \begin{align*}
-    &\left \langle \left (I+\M (\obsm)\right )^{-1}u, u \right \rangle\\
-    &= \frac{
-      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m,
-      (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m \rangle
-    }{
-      c- w^t \Sigmam^{-1} w
-    }\\
-    %
-    %
-    %
-    &= 
-    \frac{
-      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m,
-      (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m \rangle
-    }{
-      \sigma^2 + \meas_m \modcov \meas_m - \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m 
-    }
-  \end{align*}
-  and the conclusion follows.
-\end{proof}
-
-Lemma \ref{lemma:design increase} implies the following two corollaries.
-\begin{corollary}[Gain for No Model Error]\label{cor:zero mod err}
-  If $\modcov = 0$, then
-  \begin{equation*}
-    \tar( \obs ) - \tar (\obsm )
-    = -\frac12 \log (1 - \sigma^{-2} \langle \fwd \postcov \fwd^* \meas_m, \meas_m \rangle ).
-  \end{equation*}
-\end{corollary}
-\begin{proof}
-  Note that this is not immediate by substituting $\modcov = 0$ in the
-  conclusion of Lemma \ref{lemma:design increase}, since we make a
-  claim for $\postcov$, and not $\postcovm$. Let us first review
-  \eqref{eq:u} and note that since $\modcov = 0$ the covariance
-  $\Sigmam = \sigma^2I_{m-1}$ and $w = 0$, so $c - w^t\Sigmam^{-1}w =
-  \sigma^2$:
-  \begin{align*}
-    u :&=-\sigma^{-1}\prcov^{1/2} \fwd^* \meas_m\\
-    %
-    %
-    u^* :&= -\sigma^{-1} \meas_m \fwd \prcov^{1/2}.
-  \end{align*}
-  From \eqref{eq:M plus I}:
-  \begin{equation*}
-    I + \M(\obsm) = I +\M(\obs) - uu^*,
-  \end{equation*}
-  and thus:
-  \begin{equation*}
-    \left \langle \left ( I +\M(\obs) \right )^{-1} u, u \right \rangle
-    = \sigma^{-2} \langle \fwd \postcov \fwd^* \meas_m, \meas_m \rangle.
-  \end{equation*}
-  Analogously to \eqref{eq:M postcov} we note that
-  \begin{equation*}
-    \prcov^{1/2} \left ( I + \M(\obs) \right )^{-1} \prcov^{1/2} = \postcov.
-  \end{equation*}
-  Using Lemma \ref{lemma:MDL} we conclude
-  \begin{align*}
-    \tar( \obs ) - \tar( \obs )
-    &= \frac12 \log \det \left (I +\M(\obs) \right ) / \det \left (I + \M(\obsm) \right ) \\
-    %
-    %
-    %
-    &= \frac12 \log \det \left (I +\M(\obs) \right ) / \det \left (I + \M(\obs) - uu^* \right ) \\
-    %
-    %
-    %
-    &=-\frac12 \log (1 - \langle (I+\M(\obs))^{-1}u, u \rangle \\
-    %
-    %
-    %
-    &= -\frac12 \log (1 - \sigma^{-2} \langle \fwd \postcov \fwd^* \meas_m, \meas_m \rangle ).
-  \end{align*}
-\end{proof}
-
-\samemeas
-\begin{proof} \label{cor:same meas proof}
-  Denote $A:= \obs \modcov \obs^*$ and $v_j$ the $j$th column of $A$.
-  Note that $v_j = \obsm \modcov \meas_m$, since $(\obsm \modcov
-  \obsm^*)_{ij} = \meas_i(\modcov \meas_j)$, as explained in
-  \eqref{eq:modcov explained}. One can now verify that
-  \begin{equation}\label{eq:observation}
-    \Sigmam^{-1} \obsm \modcov \meas_m = \Sigmam^{-1}v_j = (A +\sigma^2I_{m-1})^{-1} v_j =
-    e_j -\sigma^2 \Sigmam^{-1}e_j.
-  \end{equation}
-  %
-  Using \eqref{eq:observation}:
-  \begin{align}\label{eq:denominator}
-    \begin{split}
-      \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m
-      &= \meas_m \modcov \obsm^* ( e^j - \sigma^2 \Sigmam^{-1} e_j )\\
-      %
-      %
-      %
-      &= \meas_m \modcov \meas_j - \sigma^2 \meas_m \modcov \obsm^* \Sigmam^{-1}e_j \\
-      %
-      %
-      %
-      &= \meas_m \modcov \meas_j -\sigma^2 (e_j - \sigma^2 \Sigmam^{-1}e_j)^t e_j \\
-      %
-      %
-      %
-      &= \meas_m \modcov \meas_m -\sigma^2 + \sigma^4 e_j^t\Sigmam^{-1}e_j.
-    \end{split}
-  \end{align}
-  We use \eqref{eq:observation} to simplify the terms in the enumerator of
-  the conclusion of Lemma \ref{lemma:design increase}:
-  \begin{align}\label{eq:enumerator}
-    \begin{split}
-      (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m
-      &= \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m - \meas_m \\
-      %
-      %
-      %
-      &= \obsm^* (e_j - \sigma^2 \Sigmam^{-1} e_j) -\meas_j \\ 
-      %
-      %
-      %
-      &= -\sigma^2 \obsm^* \Sigma^{-1}e_j. 
-    \end{split}
-  \end{align}
-  %
-  Substitute \eqref{eq:enumerator} and \eqref{eq:denominator} to
-  the enumerator and denominator (respectively) of the conclusion of
-  Lemma \ref{lemma:design increase}:
-  %
-  \begin{align*}
-    \tar( \obs ) - \tar (\obsm ) &=
-    \log \left ( 1 + \frac{
-      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m,
-      (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m \rangle
-    }{
-      \sigma^2 + \meas_m \modcov \meas_m - \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m 
-    }       
-    \right ) \\
-    %
-    %
-    %
-    &= \log \left ( 1 + \frac{\sigma^4
-      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
-      \obsm^* \Sigmam^{-1}e_j \rangle
-    }{
-      2\sigma^2 - \sigma^4 e_j^t\Sigmam^{-1}e_j 
-    }       
-    \right ) \\
-    %
-    %
-    %
-    &= \log \left ( 1 + \frac{\sigma^2
-      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
-      \obsm^* \Sigmam^{-1}e_j \rangle
-    }{
-      2 - \sigma^2 e_j^t\Sigmam^{-1}e_j 
-    }       
-    \right ).
-  \end{align*}
-\end{proof}
-
-\begin{lemma}[Auxilliary Calculations]\label{lemma:aux calc}
-  Let $T(\obs) := \obs^* \Sigma^{-1}(\obs)\obs$, with $\Sigma(\obs)$
-  defined as in \eqref{eq:Sigma}. Then:
-  \begin{align*}
-    \delta T(\obs)V &= V^* \Sigma^{-1} \obs 
-    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
-    &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs
-    + \obs^* \Sigma^{-1} V.
-  \end{align*}
-\end{lemma}
-
-\begin{proof}
-  We need a few supplementary calculations. First:
-  \begin{align}\label{eq:der sig}
-    \begin{split}
-      \frac{\der}{\der \tau} \Big |_{\tau=0} \Sigma( \obs + \tau V )
-      &= \frac{\der}{\der \tau} \Big |_{\tau=0} 
-      (\obs + \tau V ) \modcov (\obs + \tau V )^*  + \sigma^2I\\
-      % 
-      % 
-      % 
-      &= V \modcov \obs^* + \obs \modcov V^*.
-    \end{split}
-  \end{align}
-  By the standard trick for the derivative of an operator: 
-  \begin{align*}
-    0 &= \frac{\der}{\der \tau} \Big |_{\tau=0} I \\
-    % 
-    % 
-    % 
-    &= \frac{\der}{\der \tau} \Big |_{\tau=0}
-    \left (\Sigma(\obs+\tau V)^{-1} \Sigma(\obs+\tau V) \right ) \\
-    % 
-    % 
-    % 
-    &= \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0} \Sigma+
-    \Sigma^{-1} \frac{\der \Sigma(\obs+\tau V)}{\der \tau} \Big |_{\tau=0}\\  
-    %
-    %
-    %
-    &= \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0} \Sigma+
-    \Sigma^{-1} (V\modcov \obs^* + \obs \modcov V^*) 
-    \text{, by \eqref{eq:der sig}. }
-  \end{align*}
-  Now we can write the variation of $\Sigma^{-1}$:
-  \begin{align}\label{eq:der sig inv}
-    \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0}  
-      &= -\Sigma^{-1} (V \modcov \obs^* + \obs \modcov V^*) \Sigma^{-1}.
-    \end{align}
-  % 
-  Finally, we can do the calculation we need to. By Leibnitz (product) rule
-  and \eqref{eq:der sig inv}:
-  % 
-  \begin{align*}
-    \delta T(\obs) V 
-    &= \frac{\der T(\obs + \tau V)}{\der \tau} \Big |_{\tau=0} \\
-    %
-    %
-    %
-    &= V^* \Sigma^{-1} \obs 
-    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
-    &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs
-    + \obs^* \Sigma^{-1} V,
-  \end{align*}
-  as required.
-\end{proof}
-
-\biggerbetter
-\begin{proof} 
-  Fix an arbitrary $j=1,\dots,m$ and take $V:= e_j e_j^t \obs$. We see
-  that for $u \in \hilo$
-  \begin{equation*}
-    Vu = e_je_j^t (\meas_1(u),\dots,\meas_m(u) )^t = e_j \meas_j(u)
-    = (0,\dots,0,\meas_j(u),0,\dots,0)^t.
-  \end{equation*}
-  This way, $V$ has the same $j$th entry as $\obs$ while the rest are
-  set to zero. We calculate the variation in this direction and show
-  it is positive --- $\delta \tar(\obs)V > 0$. This means that
-  increasing the magnitude of the $j$th measurement functional
-  increases $\tar(\obs)$. We start with the last line of \eqref{eq:tar var}, and denote $\tmp : = \fwd \postcov \fwd^*$.
-  \begin{align*}
-     \delta \tar(\obs) V 
-    &= \tr{V ( I - \modcov \obs^*\Sigma^{-1}\obs) \tmp \obs^* \Sigma^{-1}} \\
-    % 
-    %
-    %
-    &= \tr{e_je_j^t \obs ( I - \modcov \obs^*\Sigma^{-1}\obs) \tmp \obs^* \Sigma^{-1}} \\
-    %
-    % 
-    %
-    &= e_j^t \obs ( I - \modcov \obs^*\Sigma^{-1}\obs) \tmp \obs^* \Sigma^{-1}e_j \\
-    %
-    % 
-    %
-    &= e_j^t ( I - \obs \modcov \obs^*\Sigma^{-1})\obs \tmp \obs^* \Sigma^{-1}e_j \\  
-    % 
-    %
-    %
-    &=  e_j^t(\Sigma-\obs \modcov \obs^*) \Sigma^{-1}\obs \tmp \obs^* \Sigma^{-1}e_j \\
-    %
-    %
-    %
-    &=\sigma^2 e_j^t \Sigma^{-1}\obs \tmp \obs^* \Sigma^{-1}e_j
-    \text{ by \eqref{eq:Sigma} }\\
-    %
-    % 
-    %
-    &=\sigma^2 e_j^t \Sigma^{-1}\obs \fwd \postcov \fwd^* \obs^* \Sigma^{-1}e_j.
-  \end{align*} 
-  Since $\postcov$ is positive definite, $\delta \tar(\obs) V > 0$.
-\end{proof}
-
-\bibliographystyle{amsplain}
-\bibliography{refs,georg_refs}
-
-\end{document}

### 2023-05-11 latex/vanishing.tex
Some minor changes, plus an explanation of how to modify clustered to non clustered

@@
 
-\section{Clusterization poses no obstruction to D-optimality}\label{section:clusterization}T
+\section{Clusterization poses no obstruction to D-optimality}\label{section:clusterization}
 The necessary conditions for D-optimality of Theorem \ref{thm:char}
@@
 clustered designs. Consider an inverse problem for which the first two
-eigenvalues of $\fwd \prcov\fwd^*$ are $\lambda_1^{-1} = 0.1,
-\lambda_2^{-1} = 0.2$, its eigenvectors are $\{\ev_i\}_{i=1}^\infty$
-and $\sigma^2=1$. Consider the following $\obs_1, \obs_2$ (see also
-Figure \ref{fig:clusterization}):
+eigenvalues of the pushforward prior $\fwd \prcov\fwd^*$ are
+$\lambda_1 = 10, \lambda_2 = 5$, its eigenvectors are
+$\{\ev_i\}_{i=1}^\infty$ and $\sigma^2=1$. Consider the following
+$\obs, \pobs$ (see also Figure \ref{fig:clusterization}):
 \begin{equation}\label{eq:two designs}
-  \obs_1 =
+  \obs =
   \begin{pmatrix}
@@
   \end{pmatrix},
-  \obs_2 =
+  \pobs =
   \begin{pmatrix}
@@
 \end{equation}
-\noindent It is easy to verify that $\obs_1^*\obs_1 = \obs_2^*\obs_2$
-and that both $\obs_1$ and $\obs_2$ are D-optimal. However, $\obs_1$
-is a clustered design ($\meas_3 = \meas_5 = \ev_1$), while $\obs_2$ is
+\noindent It is easy to verify that $\obs^*\obs = \pobs^*\pobs$
+and that both $\obs$ and $\pobs$ are D-optimal. However, $\obs$
+is a clustered design ($\meas_3 = \meas_5 = \ev_1$), while $\pobs$ is
 not.
@@
 
-\textbf{Question \ref{q:why}}: Since $\obs_1$ is clustered and
+\textbf{Question \ref{q:why}}: Since $\obs$ is clustered and
 D-optimal, measurement clusterization poses no obstruction to
 D-optimality. Furthermore, a computer implementation of Lemma
-\ref{thm:char} also generates $\obs_1$ as a solution to the D-optimal
+\ref{thm:char} also generates $\obs$ as a solution to the D-optimal
 design problem. Randomized numerical simulations give rise to
@@
 
-\textbf{Question \ref{q:replace}}: Since $\obs_2$ is D-optimal but not
+\textbf{Question \ref{q:replace}}: Since $\pobs$ is D-optimal but not
 clustered, we conclude that it is possible to substitute a clustered
-design for an equally optimal non-clustered design. Achieveing this in
-the setup of our model not hard, but in real life problems this may
-prove more difficult.
+design for an equally optimal non-clustered design. It is easy to give
+some simple intuition on how the clustered design $\obs$ was modified
+to the non-clustered design $\pobs$. First let us introduce some
+notation: let $\obs_{12}$ be the design composed of the first and
+second rows of $\obs$ and use a similar definition for
+$\pobs_{12}$. Note that $\obs_{12}$ uniquely determines $\obs$ as a
+clustered design (up to permutations and signs). Thus, we wanted to
+slightly modify $\obs_{12}$. We replaced $\sqrt{11/40}$, the
+coefficent of $\ev_1$ in the first measurement of $\obs$, with
+$\sqrt{3/8}$ in the first measurement of $\pobs$. After that, we chose
+$\pobs_{12}$ such that its measuremtns have unit norm and
+$\pobs_{12}^* \pobs_{12}$ has the same eigenvectors as
+$\obs^*\obs$. We then considered a ``new'' inverse problem, for which
+the pushforward prior covariance is $(\fwd \prcov\fwd^*)^{-1} +
+\sigma^{-2} \pobs_{12}^*\pobs_{12}$. We then completed $\pobs$ to a
+D-optimal design via the construction detailed in Lemma
+\ref{lemma:free}. This is how $\pobs$ was constructed, but
+(infinitely) many other non-clustered D-optimal designs could be
+constructed this way and it is not clear why the numerical
+implementation of Lemma \ref{lemma:free} resulted in a clustered
+design. Giving a full algorithm for the procedure detailed above is
+out of scope for the current study, and I intend to pursue it in the
+future.
 
@@
   \end{tikzpicture}
-  \caption{Clusterization (left, $\obs_1$ in \eqref{eq:two designs})
-    and non-clusterization (right, $\obs_2$ in \eqref{eq:two designs})
+  \caption{Clusterization (left, $\obs$ in \eqref{eq:two designs})
+    and non-clusterization (right, $\pobs$ in \eqref{eq:two designs})
     in D-optimal designs. Posterior precision per eigenvector is
-    plotted for $\obs_1$ and $\obs_2$ from \eqref{eq:two
+    plotted for $\obs$ and $\pobs$ from \eqref{eq:two
       designs}. Both designs are identical for all practical matters

### 2023-05-14 latex/vanishing.tex
Moved appendix to supp material, rewrite, spellcheck

@@
 \section{Characterizing D-Optimal Designs when $\modcov = 0$}\label{section:vanishing}
-We will now prove Theorem \ref{thm:char} which caracterizes D-optimal
+We will now prove Theorem \ref{thm:char} which characterizes D-optimal
 designs when $\modcov = 0$. The necessary first-order condition for
@@
   \end{align*}
-  and observe that $Y_{\eps}$ is invertible. Subsitute $X, Y_{\eps}$
+  and observe that $Y_{\eps}$ is invertible. Substitute $X, Y_{\eps}$
   into the RHS of \eqref{eq:first}:
@@
       \end{equation*}
-    \item Furthermore, if $\obs$ isD-optimal, $\eta_i > 0$ for
+    \item Furthermore, if $\obs$ is D-optimal, $\eta_i > 0$ for
       eigenvectors corresponding to the $k$ largest $\lambda_i$.
@@
   Summing over $i \in A$, substituting $\sum_{i\in A} \eta_i =
-  \sum_{i=1}^m \eta_i = m$, and dividng by $k:= |A|$:
+  \sum_{i=1}^m \eta_i = m$, and dividing by $k:= |A|$:
   $$
@@
   Part (1) is immediate for any measurement operator $\obs$ that
-  satisfies the unit norm constraint on measurement vectors. Part (2)
+  satisfies the unit norm constraint on measurements. Part (2)
   was proved in Proposition \ref{prop:same ev}. Part (3) was proved in
@@
     measurements. We assume $\sigma^2=1$ and thus, the blue area has
-    accumulated height of $\sigma^{-2}m = 3$ in both plots. The
+    accumulated height of $\sigma^{-2}m = 3$ in both panels. The
     D-optimal design (left) increases precision where it is
@@
 when $\modcov = 0$: Imagine each eigenvector of $\fwd \prcov \fwd^*$
-corresponds to a graduated lab cylinder (Figure
-\ref{fig:cylinder}). Each cylinder $i$ is filled, a-priori, with green
-liquid of $\lambda_i^{-1}$ volume units. We are allowed $m$
-measurement, so we have blue liquid of volume $\sigma^{-2}m$ units at
-our disposal. When we seek a D-optimal design, we distribute the blue
-liquid by repeatedly adding a drop to whatever cylinder currently has
-the lowest level of liquid in it, as long as its index $i \leq m$. The
-result of such a procedure is illustrated in Figure \ref{fig:optimal
-  vs not}.
-
-\begin{figure}%{r}{0.25\textwidth} 
-    \centering
-    \includegraphics[width=5cm, height=5cm]{cylinders.jpg}
-    \caption{Three graduated lab cylinders, corresponding to three
-      eigenvectors. Prior eigenvalues not shown.}
-    \label{fig:cylinder}
-\end{figure}
-
+corresponds to a graduated lab cylinder.
+%% (Figure \ref{fig:cylinder})
+Each cylinder $i$ is filled, a-priori, with green liquid of
+$\lambda_i^{-1}$ volume units. We are allowed $m$ measurement, so we
+have blue liquid of volume $\sigma^{-2}m$ units at our disposal. When
+we seek a D-optimal design, we distribute the blue liquid by
+repeatedly adding a drop to whatever cylinder currently has the lowest
+level of liquid in it, as long as its index $i \leq m$. The result of
+such a procedure is illustrated in Figure \ref{fig:optimal vs not}.
+%% \begin{figure}%{r}{0.25\textwidth} 
+%%     \centering
+%%     \includegraphics[width=5cm, height=5cm]{cylinders.jpg}
+%%     \caption{Three graduated lab cylinders, corresponding to three
+%%       eigenvectors. Prior eigenvalues not shown.}
+%%     \label{fig:cylinder}
+%% \end{figure}
 \section{Clusterization poses no obstruction to D-optimality}\label{section:clusterization}
 The necessary conditions for D-optimality of Theorem \ref{thm:char}
-leave some freedom in choosing $\obs$, which may or may not give rise
-clustered designs. Consider an inverse problem for which the first two
-eigenvalues of the pushforward prior $\fwd \prcov\fwd^*$ are
-$\lambda_1 = 10, \lambda_2 = 5$, its eigenvectors are
+leave some freedom in choosing $\obs$, which might allow us to avoid a
+clustered design. As an example, consider an inverse problem for which
+the first two eigenvalues of the pushforward prior $\fwd \prcov\fwd^*$
+are $\lambda_1 = 10, \lambda_2 = 5$, with eigenvectors
 $\{\ev_i\}_{i=1}^\infty$ and $\sigma^2=1$. Consider the following
@@
 \ref{thm:char} also generates $\obs$ as a solution to the D-optimal
-design problem. Randomized numerical simulations give rise to
-D-optimal designs that exhibit clustrization more than 95\% of the
-time (see code in supplementary). Since our model is as generic as
-possible, we conclude that it is possible that measurement
-clusterization is a generic phenomenon that.
+design problem. Furthermore, randomized numerical simulations give
+rise to D-optimal designs that exhibit clusterization more than 95\% of
+the time (see code in supplementary). Given our model's genericity, we
+conclude that measurement clusterization may be a generic phenomenon.
 
-\textbf{Question \ref{q:replace}}: Since $\pobs$ is D-optimal but not
-clustered, we conclude that it is possible to substitute a clustered
-design for an equally optimal non-clustered design. It is easy to give
-some simple intuition on how the clustered design $\obs$ was modified
-to the non-clustered design $\pobs$. First let us introduce some
-notation: let $\obs_{12}$ be the design composed of the first and
-second rows of $\obs$ and use a similar definition for
-$\pobs_{12}$. Note that $\obs_{12}$ uniquely determines $\obs$ as a
-clustered design (up to permutations and signs). Thus, we wanted to
-slightly modify $\obs_{12}$. We replaced $\sqrt{11/40}$, the
-coefficent of $\ev_1$ in the first measurement of $\obs$, with
-$\sqrt{3/8}$ in the first measurement of $\pobs$. After that, we chose
-$\pobs_{12}$ such that its measuremtns have unit norm and
-$\pobs_{12}^* \pobs_{12}$ has the same eigenvectors as
-$\obs^*\obs$. We then considered a ``new'' inverse problem, for which
-the pushforward prior covariance is $(\fwd \prcov\fwd^*)^{-1} +
-\sigma^{-2} \pobs_{12}^*\pobs_{12}$. We then completed $\pobs$ to a
-D-optimal design via the construction detailed in Lemma
-\ref{lemma:free}. This is how $\pobs$ was constructed, but
-(infinitely) many other non-clustered D-optimal designs could be
-constructed this way and it is not clear why the numerical
-implementation of Lemma \ref{lemma:free} resulted in a clustered
-design. Giving a full algorithm for the procedure detailed above is
-out of scope for the current study, and I intend to pursue it in the
-future.
+\textbf{Question \ref{q:replace}}: We showed that it is possible to
+substitute a D-optimal clustered design (e.g. $\obs$) for a D-optimal
+non-clustered design (e.g. $\pobs$). Presenting a general algorithm
+for replacing a clustered D-optimal design with a non-clustered
+D-optimal design is out of the scope of the current study. We can,
+however, provide some intuition, utilizing $\obs, \pobs$ as a test
+case. Let $\obs_{12}$ the design composed of the first and second rows
+of $\obs$ (with a similar definition for $\pobs_{12}$), and note that,
+up to permutations and sign changes, $\obs_{12}$ uniquely determines
+$\obs$ as a clustered design. We find $\pobs_{12}$ by requiring that
+(a) measurements in $\pobs_{12}$ have unit norm, (b) $\pobs_{12}^*
+\pobs_{12}$ has the same eigenvectors as $\obs^*\obs$, (c)
+$\pobs^*\pobs \approx \obs^*\obs$, and (d) $\pobs^*\pobs \neq
+\obs^*\obs$. We then consider a ``new'' inverse problem, for which the
+pushforward prior covariance is the covariance of the posterior
+$\mu_{\textup{post}}^{\data, \pobs_{12}}$, namely
+
+$$
+\left(\left(\fwd \prcov\fwd^*\right)^{-1} + \sigma^{-2}
+\pobs_{12}^*\pobs_{12}\right)^{-1}.
+$$
+%
+Finally, we complete $\pobs$ to a D-optimal design with $m=5$
+measurements via the construction outlined in Lemma
+\ref{lemma:free}.
+
+Interestingly, part (c) in the process detailed above implies that
+there are actually infinitely many non-clustered D-optimal designs. It
+is not at all clear why the numerical implementation of Lemma
+\ref{lemma:free} results in a clustered design. An answer to this
+question may further help shed light on the measurement clusterization
+phenomenon.
 

### 2023-05-14 latex/intro.tex
Some minor changes, plus an explanation of how to modify clustered to non clustered
Done?

@@
 illustrate this point, we consider a toy inverse problem: inferring
-the initial condition of the 1D heat equation (details in Section
-\ref{section:example}). In Figure \ref{fig:clusterization
+the initial condition of the 1D heat equation (details in the
+supplementary material). In Figure \ref{fig:clusterization
   illustration} D-optimal measurement locations are shown for
@@
 
+%% first: blue, second: red, third: green, fourth: black,
+%% fifth:magenta
 \begin{figure}[H]
@@
     \includegraphics[height=0.5\textwidth]{example.eps}
-    \caption{Measurement clusterization for the inversion of the
-      initial condition of the 1D heat equation. For each measurement,
-      its index is plotted where the measurement is taken. Fmong fou
-      locations are plotted over the computational domain $\Omega =
-      [0, 1]$ (x-axis), for varying numbers of measurements
-      (y-axis). Measurement locations (numbers) were chosen according
-      to the D-optimality criterion. Measurement clusterization
-      already occurs for four measurements: the third measurement is
-      overlaid with the fourth. For five measurements, first and
-      second measurements are clustered, as well as the fourth and the
-      fifth.}
+    \caption{Measurement clusterization for optimal designs when
+      inverting for the initial condition of the 1D heat equation (see
+      supplementary material for details). Measurement locations were
+      chosen according to the Bayesian D-optimality criterion of
+      Theorem \ref{thm:d optimality}. Measurement locations are
+      plotted over the computational domain $\Omega = [0, 1]$
+      (x-axis), for varying numbers of measurements (y-axis). The
+      colored numbers are measurement indices, plotted for visual
+      clarity. Measurement clusterization already occurs for three
+      measurements: the second measurement (red) is overlaid with the
+      third (green). For five measurements, first (blue) and second
+      (red) measurements are clustered, as well as the fourth (black)
+      and the fifth (magenta).}
   \label{fig:clusterization illustration}
@@
   neitzel2019sparse}, prompting the exploration of various remedies to
-address this issue. One approach involves merging ``close''
-measurements \cite{fedorov2012}; however, this strategy merely
-overlooks the phenomenon of measurement clusterization. An alternative
-solution lies in \emph{clusterization-free design}s, where measurement
-locations are deliberately chosen to be distant from one another. This
-can be achieved by imposing distance constraints between measurements
-or by introducing correlated errors that account for both observation
-error and model misspecification \cite{Ucinski05}. For instance, in
-the context of time-series analysis for pharmacokinetic experiments,
+address this issue. One approach involves merging close measurements
+\cite{fedorov2012}; however, this strategy merely overlooks the
+phenomenon of measurement clusterization. An alternative solution lies
+in \emph{clusterization-free design}s, where measurement locations are
+deliberately chosen to be distant from one another. This can be
+achieved by imposing distance constraints between measurements or by
+introducing correlated errors that account for both observation error
+and model misspecification \cite{Ucinski05}. For instance, in the
+context of time-series analysis for pharmacokinetic experiments,
 measurement clusterization can be mitigated by incorporating the
@@
 
-In spatial problems involving choice of measuremnts within a domain
+In spatial problems involving choice of measurements within a domain
 $\Omega \subseteq \mathbb{R}^d, d=1,2,3$, many researchers circumvent
@@
 computational cost as it requires solving a difficult combinatorial
-optimization problem for measurements locations over a finite set. The
+optimization problem for measurement locations over a finite set. The
 combinatorial optimization problem is usually relaxed by first
-assigning optimal measurement ``weights'' in $\mathbb{R}_+$ to the
+assigning optimal measurement weights in $\mathbb{R}_+$ to the
 potential measurement locations. Some researchers incorporate a
@@
 fundamental issue: the restriction of measurement locations to a
-coarse grid in $\Omega$ inherently results in a sub-optimal design.
+coarse grid in $\Omega$ fundamentally changes the optimal design
+problem and thus results in a sub-optimal design.
 
 Avoiding measurement clusterization is a pragmatic approach:
-intuitively, we recognize that measurement clusterization is
+intuitively, researchers recognize that measurement clusterization is
 undesirable, even though the underlying reasons may not be fully
-clear. Consequently, we strive to prevent it and devise various
+clear. Consequently, they strive to prevent it and devise various
 methodologies to avoid measurement clusterization. Yet each and every
@@
 
-\begin{restatable}[D-optimal designs with vanishing model error]{theorem}{char}\label{thm:char}
+We first give some general definitions, and then state Theorem
+\ref{thm:char}: Let $\hilp, \hilo$ Hilbert spaces, $\fwd:\hilp \to
+\hilo$ a linear compact operator. Let $\obs: \hilo \to \mathbb{R}^m$ a
+linear measurement operator, where $m \in \mathbb{N}$ is the number of
+measurements taken. Let $\sigma^2 \in \mathbb{R}_{+}$ observation
+noise variance, $\data = \obs \fwd \param + \eps$, where $\eps \in
+\mathbb{R}^m$ is iid $\mathcal{N}(0, \sigma^2)$ noise. Let $\pr \sim
+\mathcal{N}(\prmean, \prcov)$ prior Gaussian measure on $\hilp$, where
+$\prcov$ is the prior covariance operator and let $\post$ the
+posterior (Gaussian) measure.
+
+\begin{restatable}{theorem}{char}\label{thm:char}
   Let:
   \begin{itemize}
-  \item $\hilp, \hilo$ Hilbert spaces.
-  \item $\fwd:\hilp \to \hilo$ a linear compact operator.
-  \item $\pr \sim \mathcal{N}(0, \prcov)$ prior Gaussian measure on $\hilp$,
-    where $\prcov:\hilp \to \hilp$ is the prior covariance operator.
-  \item $\obs: \hilo \to \mathbb{R}^m$ measurement operator, where $m
-    \in \mathbb{N}$ is the number of measurements taken.
-  \item $\sigma^2 \in \mathbb{R}_{+}$ observation noise variance,
-    $\data = \obs \fwd \param + \eps$, where $\eps \in \mathbb{R}^m$
-    is iid $\mathcal{N}(0, \sigma^2)$ noise.
-  \item $\post$ the posterior measure, with covariance $\postcov$.
-  \item A D-optimality design criterion
+    \item The D-optimality design criterion
     \cite{AlexanderianGloorGhattas14}:
@@
       \begin{split}
-        \tar(\obs) :&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
+        \tar(\obs) %:&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
         % 
@@
         &= \frac12 \log \det ( I + \sigma^{-2} \prcov^{1/2} \fwd ^*
-        \obs^* \obs \fwd \prcov^{1/2}).
+        \obs^* \obs \fwd \prcov^{1/2}), 
       \end{split}
     \end{align*}
-  \item $\{\lambda_i\}_{i=1}^\infty$ eigenvalues of $\fwd\prcov\fwd^*$
-    in decreasing order of magnitude.
+  \item $\obs$ a D-optimal design operator
+    \begin{equation*}
+      \obs = \argmax_{\|\meas_j\| = 1, j=1,\dots,m}\tar(\obs),
+    \end{equation*}
+  \item $\{\lambda_i\}_{i=1}^\infty$ eigenvalues of
+    $\fwd\prcov\fwd^*$ in decreasing order of magnitude.
     %% \item $\{\ev_i\}_{i=1}^\infty$ their corresponding eigenvectors.
   \item $\{\eta_i\}_{i=1}^\infty$ eigenvalues of $\obs^*\obs$.
-  \item $\obs$ a D-optimal design operator
-    \begin{equation*}
-      \obs = \argmax_{\|\meas_j\| = 1, j=1,\dots,m}\tar(\obs).
-    \end{equation*}
+
   \end{itemize}
@@
 
-Other contributions include three useful lemmas: In Lemma
-\ref{lemma:free} we decompose a symmetric positive definite matrix $M
-\in \mathbb{R}^{k \times k}$ with $\ttr M = m \in \mathbb{N}$ as $M =
-AA^t$, where $A$ has unit norm columns. In Lemma \ref{lemma:MDL} we
-prove a matrix determinant lemma in Hilbert spaces. Finally, in Lemma
-\ref{lemma:lax} we generalize a lemma for calculating
-$\frac{\der}{\der t} \log \det (I + X(t))$, where $X(t)$ is an
-operator valued function \cite{Lax07}.
+Lastly, in Lemma \ref{lemma:free} we decompose a symmetric positive
+definite matrix $M \in \mathbb{R}^{k \times k}$ with $\ttr M = m \in
+\mathbb{N}$ as $M = AA^t$, where $A$ has unit norm columns.
+
+%% Finally, in Lemma \ref{lemma:lax} we generalize a lemma for
+%% calculating $\frac{\der}{\der t} \log \det (I + X(t))$, where
+%% $X(t)$ is an operator valued function \cite{Lax07}.
 

### 2023-05-14 latex/paper.tex
Done?

@@
 
-\author{Yair Daon\thanks{Porter School of the Environment and Earth
-    Sciences, Tel Aviv University, Tel Aviv, Israel
-    (\email{yair.daon@gmail.com}).}}
+%% \author{Yair Daon\thanks{Porter School of the Environment and Earth
+%%     Sciences, Tel Aviv University, Tel Aviv, Israel
+%%     (\email{yair.daon@gmail.com}).}}
+\author{Yair Daon\thanks{Azrieli Faculty of Medicine, Bar Ilan
+    University, Safed, Israel.  (\email{yair.daon@gmail.com}).}}
+
 %% \address{Porter School of the Environment and Earth
@@
   Inversion for parameteres of physical processes typically requires
-  taking expensive measurements. Finding an optimal set of
+  taking expensive measurements; finding an optimal set of
   measurements is known as the optimal design problem. Surprisingly,
-  measurement locations in optimal design are sometimes extremely
-  clustered. Practically and intuitively, measurement clusterization
-  is a huge problem. Researchers usually avoid measurement
-  clusterization by somehow modifying the optimal design problem. We
-  consider a certain flavor of the optimal design problem, based on
-  the Bayesian D-optimality criterion. We suggest an analytically
-  tractable model for Bayesian D-optimal designs in linear inverse
-  problems over Hilbert spaces. We prove that measurement
-  clusterization poses no obstruction to D-optimality. We demonstrate
-  that measurement clusterization is a generic property of D-optimal
-  designs, and prove that measurement correlation mitigates the
-  clusterization of measurements. We also give a full characterization
-  of D-optimal designs, which shows that D-optimal designs uniformly
-  reduce uncertainty in the space of observations. Most importantly,
-  we show how measurement clusterization can be avoided without
-  changing the optimal design problem.
+  measurement locations in optimal designs are sometimes extremely
+  clustered. Since measurement clusterization is nonintuitive,
+  researchers often avoid it by modifying the optimal design
+  problem. We consider a certain flavor of the optimal design problem,
+  based on the Bayesian D-optimality criterion, and suggest an
+  analytically tractable model for Bayesian D-optimal designs in
+  linear inverse problems over Hilbert spaces. We prove that
+  measurement clusterization poses no obstruction to D-optimality. We
+  demonstrate that measurement clusterization is a generic property of
+  D-optimal designs, and prove that correlated noise between
+  measurements mitigates clusterization. We also give a full
+  characterization of D-optimal designs, which shows that D-optimal
+  designs uniformly reduce uncertainty in the space of
+  observations. Finally, we show how measurement clusterization can be
+  avoided without changing the optimal design problem.
 \end{abstract}
@@
 \section{Acknowledgements}
-This study is a part of my PhD thesis. Most of it was written under
-the instruction of Prof. Georg Stadler of the Courant Institute of
-Mathematical Sciences. I would like to thank Prof. Stadler for his
-great mentorship, attention to details and kind behavior.
+This study is a result of research I started during my PhD studies
+under the instruction of Prof. Georg Stadler of the Courant Institute
+of Mathematical Sciences. I would like to thank him for his great
+mentorship, attention to details and kind behavior. This work was
+supported by The Raymond and Beverly Sackler Post-Doctoral
+Scholarship.
 
@@
 the author's and do not necessarily reflect the policies and views of
-NSF, ORAU/ORISE, or DOE.  This work was also supported by The Raymond
-and Beverly Sackler Post-Doctoral Scholarship.
+NSF, ORAU/ORISE, or DOE.
 
-\input{appendix}
+%% \input{appendix}
 

### 2023-05-14 latex/prelims.tex
Some minor changes, plus an explanation of how to modify clustered to non clustered
Done?

@@
 %% \subsection{Overview}
-In this section I will present notations that will be used throughout
-this article. These notations are summarized in Section
-\ref{subsec:notation}. The theoretical foundations for inverse
+In this section we present notations that will be used throughout this
+article. These notations are summarized in the definition of Theorem \ref{thm:char}. The theoretical foundations for inverse
 problems over function spaces can be found in \cite{Stuart10}.
@@
 \subsection{Bayesian Linear Inverse Problems}\label{subsec:abstract OED}
-Let $\hilp$ and $\hilo$ be separable Hilbert spaces (the subscripts p
-and o are for ``parameter'' and ``observation'', respectively). Denote
-$\hilo^* \simeq \hilo$ the Hilbert space of all linear functionals on
-$\hilo$. Let $\fwd: \hilp \to \hilo$ the \emph{forward operator}. The
-forward operator $\fwd$ is assumed linear and strongly smoothing (has
-fast decaying modes --- the heat operator of the example in Section
-\ref{section:intro} is a prime example). Take a Gaussian prior $\param
-\sim \pr = \normal(\prmean ,\prcov)$ with some appropriate covariance
-operator $\prcov$ on $\hilp$ \cite{Stuart10}. Note that $\fwd \prcov
-\fwd^*$ is the prior covariance in $\hilo$ \cite{Stuart10}. As such,
-$\fwd \prcov \fwd^*$ is assumed invertible --- an assumption which
-will be used below. Let $m$ the number of measurements
-taken. Measurements are taken via the \emph{measurement
+Let $\hilp$ and $\hilo$ be separable Hilbert spaces,\footnote{The
+subscripts p and o are for ``parameter'' and ``observation'',
+respectively.} and denote $\hilo^* \simeq \hilo$ the Hilbert space of
+all linear functionals on $\hilo$. Let $\fwd: \hilp \to \hilo$ the
+\emph{forward operator}. The forward operator $\fwd$ is assumed linear
+and strongly smoothing (has fast decaying modes --- the heat operator
+of the 1D heat equation described in Section \ref{section:intro} is a
+prime example). Take a Gaussian prior $\param \sim \pr =
+\normal(\prmean ,\prcov)$ with some appropriate covariance operator
+$\prcov$ on $\hilp$ \cite{Stuart10}. Note that $\fwd \prcov \fwd^*$ is
+the prior covariance in $\hilo$ \cite{Stuart10}, and as such, $\fwd
+\prcov \fwd^*$ is assumed invertible --- an assumption which will be
+used below\footnote{If $\fwd$ has a nontrivial kernel we utilize
+Occam's Razor and ignore said kernel.}. Let $m$ the number of
+measurements taken. Measurements are taken via the \emph{measurement
 operator}\footnote{It is common for the measurement and forward
@@
 $\obs \in ( \hilo^* )^m$, whose entries $\meas_j, j=1,\dots,m$ are
-called \emph{measurement vector}s:
+called \emph{measurements}:
 
@@
 error terms: Spatially correlated model error $\eps' \sim
-\normal(0,\modcov)$ with $\modcov: \hilo \to \hilo$ a covariance
-operator. Observation error is denoted $\eps \sim \normal(0, \sigma^2
-I_m)$, with $I_m \in \mathbb{R}^{m \times m}$ the identity. Both error
-terms and the prior are assumed independent of each other. Thus, data
-is acquired via
+\normal(0,\modcov)$ with $\modcov$ a covariance operator. Observation
+error is denoted $\eps \sim \normal(0, \sigma^2 I_m)$, with $I_m \in
+\mathbb{R}^{m \times m}$ the identity. Both error terms and the prior
+are assumed independent of each other. Thus, data is acquired via
 \begin{align}\label{eq:inverse problem}
@@
 \cite{Tarantola05,Kaipio2006,Vogel02} and then $\Sigma = \sigma^2I_m$
-is a scalar matrix and does not depend on $\obs$.
+is a scalar matrix which does not depend on $\obs$.
 
@@
 \subsection{Bayesian D-Optimal Designs in Infinite Dimensions}\label{subsec:D optimal design} 
-A Bayesian D-optimal design maximizes expected KL divergence between
-posterior and prior measures, $\post$ and $\pr$. It is useful to first
-recall the definition of KL divergence:% for an arbitrary prior
-measure:
+A Bayesian D-optimal design maximizes expected Kullback-Leibler (KL)
+divergence between posterior $\post$ and prior measures $\pr$. It is
+useful to first recall the definition of KL divergence \cite{CoverThomas91}:
 $$
@@
 %% were investigated in \cite{AlexanderianGloorGhattas14}.
-\begin{theorem}[Slightly modified Theorem 1 from \cite{AlexanderianGloorGhattas14}]\label{thm:d optimality}
+\begin{theorem}[Alexanderian, Gloor, Ghattas \cite{AlexanderianGloorGhattas14}]\label{thm:d optimality}
   Let $\pr = \normal(\prmean,\prcov)$ be a Gaussian prior on $\hilp$
@@
 \end{theorem}
+%% It is important to note that since $\obs$ is finite-rank,
+%% $\prcov^{1/2} \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}$ is
+%% trace-class.
 \begin{definition}\label{def:d optimality}
   $\obs^{\star}$ is said to be \emph{D-optimal} if $\obs^{\star} =
-  \argmax_{\obs} \tar(\obs)$, where entries of $\obs \in (\hilo^*)^m$ are constrained to
-  some allowed set of measurement vectors in $\hilo^*$.
+  \argmax_{\obs} \tar(\obs)$, where entries of $\obs \in (\hilo^*)^m$
+  are constrained to some allowed set of measurements in $\hilo^*$.
 \end{definition}
@@
 criterion \cite{Chaloner1995}.}. Theorem \ref{thm:d optimality} and
-Definition \ref{def:d optimality} have a similar intuition:
+Definition \ref{def:d optimality} carry a similar intuition:
 \begin{align*}
@@
     &= \frac12 \log \det \Big( \prcov ( \prcov^{-1} + \fwd ^* \obs^* \Sigma^{-1} \obs \fwd) \Big )\\
-    &= \frac12 \log \det \prcov \postcov^{-1} \text{ by \eqref{eq:postcov}} \\
-    &= \frac12 \log \det \prcov -\frac12 \log \det \postcov.
+    &= \frac12 \log \det \prcov \postcov^{-1} \text{ by \eqref{eq:postcov}}.
+    %%  &= \frac12 \log \det \prcov -\frac12 \log \det \postcov.
   \end{split}
 \end{align*}
-Since $\prcov$ is constant, a D-optimal design minimizes the posterior
-covariance determinant, analogously to the finite-dimensional case.
+We think of $\prcov$ as constant, so a D-optimal design minimizes the
+posterior covariance determinant, analogously to the
+finite-dimensional case.
 
-\subsection{Notation Summary}\label{subsec:notation}
-  Our notation is summarized below. We let:
-  \begin{itemize}
-  \item $\hilp, \hilo$ Hilbert spaces.
-  \item $\fwd:\hilp \to \hilo$ a linear compact operator.
-  \item $\pr \sim \mathcal{N}(0, \prcov)$ prior Gaussian measure on
-    $\hilp$, with prior covariance operator $\prcov:\hilp \to \hilp$.
-  \item $\obs: \hilo \to \mathbb{R}^m$ measurement operator, where $m
-    \in \mathbb{N}$ is the number of measurements taken.
-  \item $\sigma^2 \in \mathbb{R}_{+}$ observation noise variance.
-  \item $\modcov$ model error covariance operator.
-    %% $\data = \obs \fwd \param + \eps$, where $\eps \in
-    %% \mathbb{R}^m$ isiid $\mathcal{N}(0, \sigma^2)$ noise.
-  \item $\Sigma(\obs) = \obs \modcov \obs^* + \sigma^2I$. 
-  \item $\post$ the posterior measure, with covariance $\postcov$.
-  \item A D-optimality design criterion
-    \cite{AlexanderianGloorGhattas14}:
-    \begin{align*}
-      \begin{split}
-        \tar(\obs) :&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
-        % 
-        % 
-        % 
-        &= \frac12 \log \det ( I + \prcov^{1/2} \fwd ^* \obs^* \Sigma(\obs)^{-1} \obs
-        \fwd \prcov^{1/2}).
-      \end{split}
-    \end{align*}
-  %% \item $\{\lambda_i\}_{i=1}^\infty$ eigenvalues of $\fwd\prcov\fwd^*$
-  %%   in decreasing order of magnitude.
-  %% %% \item $\{\ev_i\}_{i=1}^\infty$ their corresponding eigenvectors.
-  %% \item $\{\eta_i\}_{i=1}^\infty$ eigenvalues of $\obs^*\obs$.
-  \end{itemize}
+%% \subsection{Notation Summary}\label{subsec:notation}
+%%   Our notation is summarized below. We let:
+%%   \begin{itemize}
+%%   \item $\hilp, \hilo$ Hilbert spaces.
+%%   \item $\fwd:\hilp \to \hilo$ a linear compact operator.
+%%   \item $\pr \sim \mathcal{N}(0, \prcov)$ prior Gaussian measure on
+%%     $\hilp$, with prior covariance operator $\prcov:\hilp \to \hilp$.
+%%   \item $\obs: \hilo \to \mathbb{R}^m$ measurement operator, where $m
+%%     \in \mathbb{N}$ is the number of measurements taken.
+%%   \item $\sigma^2 \in \mathbb{R}_{+}$ observation noise variance.
+%%   \item $\modcov$ model error covariance operator.
+%%     %% $\data = \obs \fwd \param + \eps$, where $\eps \in
+%%     %% \mathbb{R}^m$ isiid $\mathcal{N}(0, \sigma^2)$ noise.
+%%   \item $\Sigma(\obs) = \obs \modcov \obs^* + \sigma^2I$. 
+%%   \item $\post$ the posterior measure, with covariance $\postcov$.
+%%   \item A D-optimality design criterion
+%%     \cite{AlexanderianGloorGhattas14}:
+%%     \begin{align*}
+%%       \begin{split}
+%%         \tar(\obs) :&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
+%%         % 
+%%         % 
+%%         % 
+%%         &= \frac12 \log \det ( I + \prcov^{1/2} \fwd ^* \obs^* \Sigma(\obs)^{-1} \obs
+%%         \fwd \prcov^{1/2}).
+%%       \end{split}
+%%     \end{align*}
+%%   %% \item $\{\lambda_i\}_{i=1}^\infty$ eigenvalues of $\fwd\prcov\fwd^*$
+%%   %%   in decreasing order of magnitude.
+%%   %% %% \item $\{\ev_i\}_{i=1}^\infty$ their corresponding eigenvectors.
+%%   %% \item $\{\eta_i\}_{i=1}^\infty$ eigenvalues of $\obs^*\obs$.
+%%   \end{itemize}
 

### 2023-05-19 latex/how.tex
add explanation for why clusterization

--- 2023-05-14 latex/vanishing.tex
@@
-\section{Characterizing D-Optimal Designs when $\modcov = 0$}\label{section:vanishing}
-We will now prove Theorem \ref{thm:char} which characterizes D-optimal
-designs when $\modcov = 0$. The necessary first-order condition for
-D-optimality of Theorem \ref{thm:constrained} for $\modcov = 0$
-become:
-
-\begin{equation}\label{eq:eigenproblem}
-  \sigma^{-2}\fwd \postcov \fwd^* \obs^* = \obs^* \Xi,
-\end{equation}
-with $\Xi$ diagonal. Equation \eqref{eq:eigenproblem} looks like an
-eigenvalue problem for the self-adjoint operator $\sigma^{-2}\fwd
-\postcov \fwd^*$, where rows of $\obs$, namely $\meas_j,j=1,\dots, m$,
-are eigenvectors. However, $\postcov$ depends on $\obs$ and thus, we
-refer to \eqref{eq:eigenproblem} as a \emph{nonlinear} eigenvalue
-problem.%%  Proposition \ref{prop:twice woodbury} and Lemma
-%% \ref{lemma:sim diag} required
-
-\begin{proposition}\label{prop:twice woodbury}
-  Assume $\fwd \prcov \fwd^*$ is invertible. Then
-  \begin{align*}
-    \begin{split}
-      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* 
+%\section{Clusterization poses no obstruction to D-optimality}\label{section:clusterization}
+\section{Substituting an optimal clustered design with an equally optimal non-clustered design}\label{section:how}
 %
-      %
-      = \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1},
-    \end{split}
-  \end{align*}  
-\end{proposition}
-
-\begin{proof}
-  The proof amounts to using Woodbury's matrix identity twice and a
-  regularization trick. The standard proof for Woodbury's matrix
-  identity works for separable Hilbert spaces, as long as all terms
-  are well defined. Unfortunately, $\obs^*\obs$ is not invertible, so
-  we force it to be. Recall Woodbury's matrix identity:
-  \begin{equation}\label{eq:WMI}
-    (A + UCV)^{-1} = A^{-1} - A^{-1}U(C^{-1} + VA^{-1}U)^{-1}VA^{-1}. 
-  \end{equation}
-
-  Denote:
-  \begin{align}\label{eq:notation}
-    \begin{split}
-      A :&= \prcov^{-1} \\
-      U :&= \fwd^* \\
-      V :&= \fwd \\
-      C :&= \sigma^{-2} (\obs^*\obs+\eps I),\ \eps > 0
-    \end{split}
-  \end{align}
-
-  Then, by \eqref{eq:WMI}, with the notation \eqref{eq:notation}:
-  \begin{equation}\label{eq:first}
-  \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* = \fwd ( \prcov - \prcov \fwd^* ( \sigma^2(\obs^*\obs + \eps I)^{-1} + \fwd \prcov \fwd^* )^{-1} \fwd \prcov ) \fwd^* \\
-      %
-      %
-      %% &= X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X
-  %%   \end{split}
-  %% \end{align*}
-  \end{equation}
-  Now, denote:
-  \begin{align*}
-    X :&= \fwd\prcov \fwd^* \\
-    Y_{\eps} :&= \obs^*\obs + \eps I,
-  \end{align*}
-  and observe that $Y_{\eps}$ is invertible. Substitute $X, Y_{\eps}$
-  into the RHS of \eqref{eq:first}:
-   \begin{equation}\label{eq:second}
-     \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* = X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X
-   \end{equation}
-
-   Note that $X + \sigma^2 Y_{\eps}^{-1}$ is invertible, as the sum of
-   two positive definite operators. Now, let
-   \begin{align}\label{eq:notation2}
-     \begin{split}
-       A :&= X^{-1}, \\
-       C :&= \sigma^2Y_{\eps}^{-1}, \\
-       U :&= I, \\
-       V :&= I.
-     \end{split}
-   \end{align}
-   Apply \eqref{eq:WMI} with the notation \eqref{eq:notation2} to the
-   RHS of \eqref{eq:second}:
-  \begin{align*}
-    \begin{split}
-      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* &= X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X \\
-      %
-      %
-      %
-      &= (X^{-1} + \sigma^{-2}Y_{\eps})^{-1} \\
-      %
-      %
-      %
-      &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \eps I))^{-1}.
-    \end{split}
-  \end{align*}
-
-  We conclude that $\forall \eps > 0$
-  \begin{align*}
-    \begin{split}
-      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* 
-     &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \eps I))^{-1}.
-    \end{split}
-  \end{align*}
-  Letting $\eps \to 0$ completes the proof.
-\end{proof}
-
-\begin{lemma}[Simultaneous diagonizability]\label{lemma:sim diag}
-  Let $\hil$ separable Hilbert space, $C:\hil \to \hil$ self-adjoint
-  and $\func_1,\dots,\func_m \in \hil$. Denote $\func^*$ the element
-  $\func$ acting as a linear functional. If
-  \begin{equation*}
-   (C + \sum_{j=1}^m \func_j\func_j^*) \func_l = \xi_l \func_l,\ l = 1,\dots,m
-  \end{equation*}
-  then $C$ and $\sum_{j=1}^m \func_j \func_j^*$ are simultaneously
-  diagonalizable.
-\end{lemma}
-\begin{proof}
-  First, enumerate the eigenvalues of $C + \sum_{j=1}^m
-  \func_j\func_j^*$ as $\xi_1,\dots,\xi_\ell$. Denote the indices of
-  the eigenvectors corresponding to $\xi_i$
-  \begin{equation*}
-    S_i := \{ 1 \leq k \leq m | (C + \sum_{j=1}^m \func_j\func_j^* )\func_k = \xi_i \func_k \}.
-  \end{equation*}
-  Define further
-  \begin{equation*}
-    A_i := \sum_{k \in S_i} \func_k \func_k^*,
-  \end{equation*}
-  which is self-adjoint. Two observations are in order. First,
-  $\sum_{j=1}^m \func_j\func_j^* = \sum_{i=1}^\ell A_i$. Second, $A_i
-  \func_k = 0$ if $k\not \in S_i$, since eigenvectors of different
-  eigenvalue are orthogonal. For $k \in S_i$
-  \begin{equation}\label{eq:on vi}
-    \xi_i \func_k = (C + \sum_{j=1}^m \func_j \func_j^* ) \func_k = (C + A_i) \func_k.
-  \end{equation}
-  Let $V_i := span \{\func_k \}_{k\in S_i}$. Observe that $V_i$ is
-  invariant under $A_i$, by definition, and under $C$, by \eqref{eq:on
-    vi}. A second application of \eqref{eq:on vi} shows that $A_i =
-  \xi_iI - C$ on $V_i$. This immediately implies $A_i$ and $C$ are
-  simultaneously diagonalizable on $V_i$. This holds for every $1 \leq
-  i \leq \ell$ and we conclude that $C$ and $A$ are simultaneously
-  diagonalizable.
-\end{proof}
-
-\begin{proposition}\label{prop:same ev}
-  Let $\obs$ satisfy the nonlinear eigenvalue problem
-  \eqref{eq:eigenproblem}. Then $\obs^*\obs$ and $\fwd \prcov \fwd^*$
-  are simultaneously diagonalizable.
-\end{proposition}
-\begin{proof}
-  \begin{align}\label{eq:mod conditions}
-    \begin{split}
-      \obs^* \Xi &= \sigma^{-2}\fwd \postcov \fwd^* \obs^*  \text{ (by \eqref{eq:eigenproblem})}\\
-      %
-      %
-      %
-      &= \sigma^{-2} \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* \obs^*  \text{ (by \eqref{eq:postcov})} \\
-      %
-      %
-      %
-      &= \sigma^{-2} \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1} \obs^* \text{ (by Proposition \ref{prop:twice woodbury})}.
-    \end{split}
-  \end{align}
-
-  Now take $\func_j^{*} = \meas_j$ and $C := (\fwd \prcov
-  \fwd^*)^{-1}$ and use Lemma \ref{lemma:sim diag}.
-\end{proof}
-
-Since we made no assumption regarding the ordering of $\{\lambda_i\}$,
-we can denote the corresponding non-zero eigenvalues of $\obs^*\obs$
-by $\{\eta_i\}_{i=1}^{k}$ and let $\eta_i = 0$ for $i \geq k+1$.
-
-\begin{proposition}\label{prop:true target}
-  Let $\obs$ with $m$ measurements satisfy the nonlinear eigenvalue
-  problem \eqref{eq:eigenproblem}. Let $\{\eta_i\}_{i=1}^{\infty}$
-  eigenvalues of $\obs^*\obs$ and $\{\lambda_i\}_{i=1}^{\infty}$ the
-  corresponding eigenvalues of $\fwd \prcov \fwd^*$. Let $k:=\rank
-  \obs^*\obs$. Without loss of generality, let $\eta_i > 0$ for $i\leq
-  k$ and $\eta_i = 0$ for $i > k$. Then:
-  \begin{enumerate}
-    \item $k \leq m$ and $\obs^*\obs$ has exactly $k$ positive
-      eigenvalues.
-    \item
-      \begin{equation*}
-        \tar(\obs) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i) = \frac12 \sum_{i=1}^{m} \log (1 + \sigma^{-2}\lambda_i\eta_i).
-      \end{equation*}
-    \item Furthermore, if $\obs$ is D-optimal, $\eta_i > 0$ for
-      eigenvectors corresponding to the $k$ largest $\lambda_i$.
-  \end{enumerate}
-\end{proposition}
-\begin{proof}
-  Part (1) is trivial. To see part (2) holds: 
-  \begin{align}
-    \begin{split}
-      \tar(\obs) &= \frac12\log \det \left (I + \sigma^{-2} \prcov^{1/2} \fwd ^* \obs^*
-      \obs \fwd \prcov^{1/2}\right )\\% \text{ (by definition)}\\
-      %
-      &= \frac12 \log \det \left (I + \sigma^{-2} \obs^* \obs \fwd
-      \prcov\fwd^* \right ) \text{ (Sylvester's Determinant
-      Theorem)}\\
-      %
-      %
-      %
-      &=\frac12 \log \prod_{i=1}^{\infty} ( 1 + \sigma^{-2} \lambda_i\eta_i ) \text{ (Proposition \ref{prop:same ev})} \\
-      %
-      %
-      %
-      %&=\frac12 \log \left ( \prod_{i=1}^{k} ( \lambda_i^{-1} + \sigma^{-2} \eta_i )\prod_{i=1}^{k} \lambda_i \right )
-      %
-      %
-      %
-      &=\frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i). 
-      %
-      %
-      %% &=\frac12 \sum_{i=1}^{k} \log(\sigma^2\lambda_i^{-1} + \eta_i) - k\log \sigma + \frac12\sum_{i=1}^k \log \lambda_i.
-    \end{split}
-  \end{align}
-  Part (3) holds since $\log$ is increasing and $\eta_i \geq 0$.
-\end{proof}
-
-\begin{proposition}\label{prop:kkt}
-  Let $\tar: \mathbb{R}^m \to \mathbb{R}$, $\tar(\eta) =
-  \frac{1}{2}\sum_{i=1}^m \log (1+\sigma^{-2}\lambda_i \eta_i)$, with
-  $\lambda_i > 0$ and $\sigma^{2} > 0$. Then the maximum of $\tar$
-  subject to $\eta_i \geq 0$ and $\sum\eta_i = m$ is obtained at
-  \begin{equation}
-  \eta_i = \begin{cases}
-    \frac{m}{k} - \sigma^2 \lambda_i^{-1} + \sigma^2 \frac{1}{k} \sum_{j\in A} \lambda_j^{-1} & i \in A \\
-    0 & i \in A^c
-  \end{cases}
-  \end{equation}
-  where $A:= \{1\leq i \leq m: \eta_i > 0\}$ and $A^c = \{1,\dots, m\}
-  \backslash A$, and $k = |A|$, the cardinality of $A$.
-\end{proposition}
-
-\begin{proof}
-  Let $\Phi(\eta) = \sum_{i=1}^k \eta_i - m$ and $\Omega_j(\eta) =
-   -\eta_j$. Then
-  \begin{align*}
-    \begin{split}
-      \frac{\partial \tar}{\partial \eta_i}  &=
-       \frac12 \frac{\sigma^{-2}\lambda_i}{1 + \sigma^{-2} \lambda_i\eta_i} = \frac12 \frac{1}{\sigma^{2}\lambda_i^{-1} + \eta_i} \\
-      %
-      %
-      %
-      \frac{\partial\Phi}{\partial \eta_i} &= 1 \\
-      %
-      %
-      %
-      \frac{\partial \Omega_j}{\partial \eta_i} &= -\delta_{ij}      
-    \end{split}
-  \end{align*}
-
-  From the KKT conditions, there are $\alpha, \beta_i$ such that for $i=1,\dots,m$:
-  \begin{align}
-    \begin{split}
-      -\frac12 \frac{1}{\sigma^{2}\lambda_i^{-1} + \eta_i} + \alpha - \beta_i  &= 0 \\
-      %
-      %
-      %
-      \eta_i &\geq 0\\
-      %
-      %
-      %
-      \beta_i &\geq 0\\
-      %
-      %
-      %
-      \beta_i \eta_i &= 0\\
-      %
-      %
-      %
-      \sum_{i=1}^m \eta_i &= m 
-    \end{split}
-  \end{align}
-
-  Then, for $i \in A$:
-  \begin{align*}
-    \begin{split}
-      \beta_i &= 0\\
-      %
-      %
-      %
-      \sigma^{2}\lambda_i^{-1} + \eta_i  &= \frac{1}{2\alpha}%  \text{ for }  1 \leq i \leq k.\\
-      %
-      %
-      %
-      %% \sigma^{-2} \lambda_i &= 2\alpha - 2\beta_i, i \in B \\
-    \end{split}
-  \end{align*}
-
-  Summing over $i \in A$, substituting $\sum_{i\in A} \eta_i =
-  \sum_{i=1}^m \eta_i = m$, and dividing by $k:= |A|$:
-  $$
-  \frac{1}{2\alpha} = \frac{m}{k} + \frac{\sigma^2}{k} \sum_{j\in A} \lambda_j^{-1}.
-  %\Rightarrow \alpha = \left ( \frac{2m}{k} + \frac{2\sigma^2}{k}
-  %\sum_{i\in A} \lambda_i^{-1} \right )^{-1}.
-  $$
-
-  Consequently:
-  \begin{align} \label{eq:etas}
-    \begin{split}
-      \eta_i &= \frac{1}{2\alpha} - \sigma^2 \lambda_i^{-1}\\
-      %
-      %
-      %
-      &=\frac{m}{k} + \frac{\sigma^2}{k} \sum_{j\in A} \lambda_j^{-1} - \sigma^2 \lambda_i^{-1}.%\\
-      %
-      %
-      %
-      %% &= \frac{m}{k} + \frac{\sigma^2}{k} \sum_{j \in A, j\neq i} \lambda_j^{-1} - \frac{k-1}{k}\sigma^2 \lambda_i^{-1}\\
-      %% ,\ i \in A.
-    \end{split}
-  \end{align}
-\end{proof}
-
-The final ingredient we require for the proof of Theorem
-\ref{thm:char} the following lemma:
-
-\begin{lemma}[Unit norm decomposition]\label{lemma:free}
-  Let $M \in \R^{k \times k}$ symmetric positive definite with $\ttr M
-  = m$, $m \geq k$. We can find $\func_j \in \R^k,j=1,\dots,m$
-  with $\|\func_j\|=1$ and $A = (\func_1,\dots,\func_m)$ such that
-  $AA^t = M$.
-\end{lemma}
-
-\begin{proof}
-  Let us diagonalize $M$, so that $M = U D U^t$ with $D =
-  \diag(d_1,\dots,d_k)$ and $U \in \R^{k \times k }$ orthogonal. Let
-  $S \in \R^{k \times m}$ with $S_{ii} = \sqrt{d_{i}}$ and zeros
-  otherwise. Define $A:= U S V^t$, where $V \in \R^{m \times m}$ is
-  orthogonal and will be further restricted later. Then $AA^t = U
-  SV^tVS^t U^t = UDU^t$, so $AA^t$ has the required eigenvalues and
-  eigenvectors by construction. If we can choose $V$ such that $A$
-  also satisfies the unit norm constraints we are done. These
-  constraints are, for $j=1,\dots,m$:
-  \begin{equation}\label{eq:V constraints}
-   1 = [A^tA]_{jj} = [V S^tS V^t]_{jj},
-  \end{equation}
-  and we can expect to do this since we assumed $\ttr D = m$.
-
-  Define $C = S^tS - I \in \R^{m \times m}$. Note that $\ttr C = 0$ and
-  $C$ is diagonal with non-zero entries $d_i-1,i=1,\dots,k$. It suffices
-  to find $V$ orthogonal such that $V C V^t$ has zero diagonal. We
-  construct such $V$ by sequentially inserting zeros in the diagonal
-  and not destroying zeros we already introduced, starting from the
-  last diagonal entry and moving to the first. Since $c_{mm} \neq 0$ ,
-  let $p < m$ such that $c_{pp}c_{mm} < 0$ (such $p$ exists because
-  the trace is zero) and let $\theta \in (0,\pi)$. Define a Givens
-  rotation $R^{(m)} \in \R^{m \times m}$ by
-  \begin{equation*}
-    r^{(m)}_{ab} :=
-    \begin{cases}
-      1 & a = b \neq p \text{ or } a = b \neq m \\
-      \cos \theta & a = b = p  \\
-     -\sin \theta & a = p, b = m\\
-      \cos \theta & a = b = m \\
-      \sin \theta & a = m, b = p \\ 
-      0 & o.w
-    \end{cases}
-  \end{equation*}
-  Note that conjugating a matrix by $R^{(m)}$ changes only its $m$ and
-  $p$ rows and columns. We want to choose $\theta$ such that
-  \begin{equation}\label{eq:mm}
-    0 = [R^{(m)} C (R^{(m)})^t]_{mm} = \cos^2 \theta c_{mm} + 2\cos \theta \sin
-    \theta c_{mp} + \sin^2\theta c_{pp},
-  \end{equation}
-  and it suffices to choose $\theta$ such that
-  \begin{equation*}
-    c_{mm} \cot^2 \theta + 2 c_{mp} \cot \theta + c_{pp} = 0.
-  \end{equation*}
-  This quadratic in $\cot\theta$ has a real solution, since
-  $c_{pp}c_{mm} < 0$ by assumption and we can find $\theta \in
-  (0,\pi)$ such that \eqref{eq:mm} is satisfied. We continue to find
-  $R^{(m-1)}$ that leaves row and column $m$ unchanged and
-  continue introducing zeros to the diagonal. The assumption $\ttr D =
-  m \Rightarrow \ttr C = 0$ guarantees we can do that. Taking $V:=
-  R^{(1)} R^{(2)} \dots R^{(m-1)}R^{(m)}$ completes the proof.
-\end{proof}
-
-\char*
-
-\begin{proof}
-  Part (1) is immediate for any measurement operator $\obs$ that
-  satisfies the unit norm constraint on measurements. Part (2)
-  was proved in Proposition \ref{prop:same ev}. Part (3) was proved in
-  Proposition \ref{prop:true target}.
-
-  Part (4) is a consequence of Propositions \ref{prop:true target} and
-  \ref{prop:kkt}, with the caveat that we did not show that finding
-  $\obs$ so that $\obs^*\obs$ has the desired eigenvalues is
-  feasible. To this end, we utilize Lemma \ref{lemma:free}: let $M =
-  \diag(\eta_1, \dots, \eta_k)$, diagonal with respect to the first $k$
-  eigenvectors of $\fwd \prcov \fwd^*$. We take $\obs := A$ from the
-  Lemma \ref{lemma:free}.
-
-  Recall from \eqref{eq:postcov}, that the posterior precision is
-  $\postcov^{-1} = \prcov^{-1} + \sigma^{-2}\fwd^*\obs^*\obs\fwd$. The
-  first statement in part (5) now follows from Proposition
-  \ref{prop:twice woodbury}, while the second statement follows from
-  parts (1) and (4).
-\end{proof}
-
-\pgfplotstableread{
-  Label     prior  optimal  sub-optimal 
-  1         0.2    1.8           1.7
-  2         0.8    1.2           0.8
-  3         2.2    0             0.5
-  4         3.5    0             0.0
-}\optimalvsnot
-
-\begin{figure}
-  \begin{tikzpicture}[scale=0.85]
-    \begin{axis}[
-        ybar stacked,
-        ymin=0,
-        ymax=4,
-        xtick=data,
-        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
-        reverse legend=false, % set to false to get correct display, but I'd like to have this true
-        xticklabels from table={\optimalvsnot}{Label},
-        xticklabel style={text width=2cm,align=center},
-        legend plot pos=right,
-        ylabel=precision --- prior and posterior,
-        xlabel=eigenvector $i$,
-      ]
-
-      \addplot [fill=green!80]  table [y=prior,   meta=Label, x expr=\coordindex] {\optimalvsnot};
-      \addplot [fill=blue!60]   table [y=optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};
-
-      \addlegendentry{$\sigma^2\lambda_i^{-1}$}
-      \addlegendentry{optimal $\eta_i$s}
-
-    \end{axis}
-  \end{tikzpicture}
-  \qquad
-  \begin{tikzpicture}[scale=0.85]
-    \begin{axis}[
-        ybar stacked,
-        ymin=0,
-        ymax=4,
-        xtick=data,
-        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
-        reverse legend=false, % set to false to get correct display, but I'd like to have this true
-        xticklabels from table={\optimalvsnot}{Label},
-        xticklabel style={text width=2cm,align=center},
-        legend plot pos=right,
-        ylabel=precision --- prior and posterior,
-        xlabel=eigenvector $i$,
-      ]
-
-      \addplot [fill=green!80]  table [y=prior,       meta=Label, x expr=\coordindex] {\optimalvsnot};
-      \addplot [fill=blue!60]   table [y=sub-optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};
-
-      \addlegendentry{$\sigma^2\lambda_i^{-1}$}
-      \addlegendentry{sub-optimal $\eta_i$s}
-
-    \end{axis}
-  \end{tikzpicture}
-  \caption{A comparison of the eigenvalues of the pushforward
-    posterior precision $(\fwd\prcov\fwd^*)^{-1} +
-    \sigma^{-2}\obs^*\obs$ for a D-optimal design (left) and a
-    sub-optimal design (right). Both designs are allowed $m=3$
-    measurements. We assume $\sigma^2=1$ and thus, the blue area has
-    accumulated height of $\sigma^{-2}m = 3$ in both panels. The
-    D-optimal design (left) increases precision where it is
-    lowest. The sub-optimal design (right) does not.}
-  \label{fig:optimal vs not}
-\end{figure}
-
-Theorem \ref{thm:char} facilitates understanding of D-optimal designs
-when $\modcov = 0$: Imagine each eigenvector of $\fwd \prcov \fwd^*$
-corresponds to a graduated lab cylinder.
-%% (Figure \ref{fig:cylinder})
-Each cylinder $i$ is filled, a-priori, with green liquid of
-$\lambda_i^{-1}$ volume units. We are allowed $m$ measurement, so we
-have blue liquid of volume $\sigma^{-2}m$ units at our disposal. When
-we seek a D-optimal design, we distribute the blue liquid by
-repeatedly adding a drop to whatever cylinder currently has the lowest
-level of liquid in it, as long as its index $i \leq m$. The result of
-such a procedure is illustrated in Figure \ref{fig:optimal vs not}.
-%% \begin{figure}%{r}{0.25\textwidth} 
-%%     \centering
-%%     \includegraphics[width=5cm, height=5cm]{cylinders.jpg}
-%%     \caption{Three graduated lab cylinders, corresponding to three
-%%       eigenvectors. Prior eigenvalues not shown.}
-%%     \label{fig:cylinder}
-%% \end{figure}
-\section{Clusterization poses no obstruction to D-optimality}\label{section:clusterization}
 The necessary conditions for D-optimality of Theorem \ref{thm:char}
@@
 
-We are now ready to answer two of the questions posed in the
-introduction:
-
-\textbf{Question \ref{q:why}}: Since $\obs$ is clustered and
-D-optimal, measurement clusterization poses no obstruction to
-D-optimality. Furthermore, a computer implementation of Lemma
-\ref{thm:char} also generates $\obs$ as a solution to the D-optimal
-design problem. Furthermore, randomized numerical simulations give
-rise to D-optimal designs that exhibit clusterization more than 95\% of
-the time (see code in supplementary). Given our model's genericity, we
-conclude that measurement clusterization may be a generic phenomenon.
-
-\textbf{Question \ref{q:replace}}: We showed that it is possible to
-substitute a D-optimal clustered design (e.g. $\obs$) for a D-optimal
-non-clustered design (e.g. $\pobs$). Presenting a general algorithm
-for replacing a clustered D-optimal design with a non-clustered
-D-optimal design is out of the scope of the current study. We can,
-however, provide some intuition, utilizing $\obs, \pobs$ as a test
-case. Let $\obs_{12}$ the design composed of the first and second rows
-of $\obs$ (with a similar definition for $\pobs_{12}$), and note that,
-up to permutations and sign changes, $\obs_{12}$ uniquely determines
-$\obs$ as a clustered design. We find $\pobs_{12}$ by requiring that
-(a) measurements in $\pobs_{12}$ have unit norm, (b) $\pobs_{12}^*
-\pobs_{12}$ has the same eigenvectors as $\obs^*\obs$, (c)
-$\pobs^*\pobs \approx \obs^*\obs$, and (d) $\pobs^*\pobs \neq
-\obs^*\obs$. We then consider a ``new'' inverse problem, for which the
-pushforward prior covariance is the covariance of the posterior
-$\mu_{\textup{post}}^{\data, \pobs_{12}}$, namely
+We may now answer \textbf{Question \ref{q:replace}}: We showed that it
+is possible to substitute a D-optimal clustered design (e.g. $\obs$)
+for a D-optimal non-clustered design (e.g. $\pobs$). Presenting a
+general algorithm for replacing a clustered D-optimal design with a
+non-clustered D-optimal design is out of the scope of the current
+study. We can, however, provide some intuition, utilizing $\obs,
+\pobs$ as a test case. Let $\obs_{12}$ the design composed of the
+first and second rows of $\obs$ (with a similar definition for
+$\pobs_{12}$), and note that, up to permutations and sign changes,
+$\obs_{12}$ uniquely determines $\obs$ as a clustered design. We find
+$\pobs_{12}$ by requiring that (a) measurements in $\pobs_{12}$ have
+unit norm, (b) $\pobs_{12}^* \pobs_{12}$ has the same eigenvectors as
+$\obs^*\obs$, (c) $\pobs^*\pobs \approx \obs^*\obs$, and (d)
+$\pobs^*\pobs \neq \obs^*\obs$. We then consider a ``new'' inverse
+problem, for which the pushforward prior covariance is the covariance
+of the posterior $\mu_{\textup{post}}^{\data, \pobs_{12}}$, namely
 

### 2023-05-19 latex/intro.tex
add explanation for why clusterization

@@
 remarkably similar measurements in certain cases \cite{fedorov1996,
-  hooker2009, fedorov2012, Ucinski05, neitzel2019sparse}. To
+  hooker2009, fedorov1997, Ucinski05, neitzel2019sparse}. To
 illustrate this point, we consider a toy inverse problem: inferring
@@
 Researchers widely agree that measurement clusterization is
-undesirable \cite{fedorov1996, hooker2009, fedorov2012, Ucinski05,
+undesirable \cite{fedorov1996, hooker2009, fedorov1997, Ucinski05,
   neitzel2019sparse}, prompting the exploration of various remedies to
 address this issue. One approach involves merging close measurements
-\cite{fedorov2012}; however, this strategy merely overlooks the
+\cite{fedorov1997}; however, this strategy merely overlooks the
 phenomenon of measurement clusterization. An alternative solution lies
@@
 from a coarse grid in $\Omega$ \cite{koval2020, alexanderian2021,
-  attia2020, alexanderian2014, alexanderian2016,
+  attia2022, alexanderian2014, alexanderian2016,
   alexanderian2018efficient}. This approach incurs a significant
@@
 %
-Why does measurement clusterization occur?
-%
 Why does imposing correlations between observations alleviate
@@
 %
-Should we aim to avoid measurement clusterization?
+Is measure clusterization a generic phenomenon?
 %
-Is it possible to substitute an optimal clustered design with an
-equally optimal non-clustered design?
+And, most importantly: Why does measurement clusterization occur?
+%
+%Should we aim to avoid measurement clusterization?
+%
+%Is it possible to substitute an optimal clustered design with an
+%equally optimal non-clustered design?
 %Can an optimal clustered design be relpaced with an equally optimal
@@
 \begin{enumerate}
-\item \label{q:why} \textbf{Why does measurement clusterization
-  occur?} In Section \ref{section:vanishing}, we provide an insightful
-  explanation for the optimality of clustered designs when no model
-  error is present. We demonstrate that measurement clusterization
-  poses no obstruction to D-optimality and that clustered designs are
-  preferred by numerical algorithms.
 
-  %% We conjecture that the prevalence of measurement clusterization
-  %% arises due to the ease of discovering clustered designs.
+\item \label{q:generic} \textbf{Is measurement clusterization a
+  generic phenomenon?}
+  %\subsection{An answer for Question \ref{q:generic}: Genericity of measurement clusterization}
+  %% Computer implementation of Lemma \ref{thm:char} for the inverse
+  %% problem outlined in Section \ref{section:how} also generates $\obs$
+  %% as a solution to the D-optimal design problem.  Furthermore,
+  Randomized numerical simulations of our model give rise to D-optimal
+  designs that exhibit clusterization more than 95\% of the time (see
+  code in supplementary material). Given our model's genericity, we
+  expect that measurement clusterization to be a generic phenomenon.
 
@@
 
-\item \label{q:avoid} \textbf{Should we aim to avoid measurement
-  clusterization?} Based on the analysis conducted in this study, we
-  did not find any compelling reason to explicitly avoid optimal
-  clustered designs.
+\item \label{q:why} \textbf{Why does measurement clusterization
+  occur?} In Section \ref{section:why}, we present a compelling
+  explanation for the optimality of clustered designs in the absence
+  of model error. Our analysis reveals that, in our model, a D-optimal
+  design focuses on a select set of prior eigenvectors, specifically
+  those with the highest power spectrum. In practical scenarios, the
+  number of locations where the relevant prior eigenvectors are
+  significantly large, while other eigenvectors have a power spectrum
+  of zero, is limited. Consequently, the clustering of measurements
+  arises as a natural consequence of the pigeonhole principle, as
+  there are more measurements available than there are locations
+  satisfying conditions (a) and (b).
 
-\item \label{q:replace} \textbf{Is it possible to substitute an
-  optimal clustered design with an equally optimal non-clustered
-  design?} In Section \ref{section:vanishing}, we answer this question
-  in the affirmative, although we show that numerical experiments
-  conducted using our model indicate a strong preference for clustered
-  designs.
+  %% In Section \ref{section:vanishing}, we provide an insightful
+  %% explanation for the optimality of clustered designs when no model
+  %% error is present. We demonstrate that for our model, a D-optimal
+  %% design measures only a small subset of prior eigenvectors, which are
+  %% the prior eigenvectors with largest power spectrum. In real-life
+  %% problems there are limited number of locations where: (a) the
+  %% relevant prior eigenvectors are large, and other prior eigenvectors
+  %% with large power in the prior spectrum are zero. Then measurement
+  %% clusterization is a result of the pigeonhole principle, where there
+  %% are more measurements than measurement locations satisfying (a) and
+  %% (b).
+
+  %% We conjecture that the prevalence of measurement clusterization
+  %% arises due to the ease of discovering clustered designs.
+
+%% \item \label{q:avoid} \textbf{Should we aim to avoid measurement
+%%   clusterization?} Based on the analysis conducted in this study, we
+%%   did not find any compelling reason to explicitly avoid optimal
+%%   clustered designs.
+
+%% \item \label{q:replace} \textbf{Is it possible to substitute an
+%%   optimal clustered design with an equally optimal non-clustered
+%%   design?} In Section \ref{section:vanishing}, we answer this question
+%%   in the affirmative, although we show that numerical experiments
+%%   conducted using our model indicate a strong preference for clustered
+%%   designs.
+
 \end{enumerate}
@@
 lies in its fifth component, which highlights that a D-optimal design
-aims to uniformly reduce posterior uncertainties across all posterior
-covariance eigenvectors. A similar conclusion was reached by Koval et
-al.~\cite{koval2020}, who showed that A-optimal designs are best
-constructed in the space of observations.
+aims to uniformly reduce posterior uncertainties for posterior
+covariance eigenvectors in observation space. A similar conclusion was
+reached by Koval et al.~\cite{koval2020}, who showed that A-optimal
+designs are best constructed in the space of observations.
 
-We first give some general definitions, and then state Theorem
+We first give some general definitions before we state Theorem
 \ref{thm:char}: Let $\hilp, \hilo$ Hilbert spaces, $\fwd:\hilp \to
@@
 correspond to any specific real-life problem. It is generic enough to
-be analytically tractable, but one may argue it is too far removed
-from any real application. To these claims I would argue that
+be analytically tractable, but one may argue our model is too far
+removed from any real application. To these claims I would answer that
 scientists have a long history of studying models that are bare-bones

### 2023-05-19 latex/paper.tex
add explanation for why clusterization

@@
 
-%% \author{Yair Daon\thanks{Porter School of the Environment and Earth
-%%     Sciences, Tel Aviv University, Tel Aviv, Israel
-%%     (\email{yair.daon@gmail.com}).}}
-\author{Yair Daon\thanks{Azrieli Faculty of Medicine, Bar Ilan
-    University, Safed, Israel.  (\email{yair.daon@gmail.com}).}}
+\author{Yair Daon\thanks{Porter School of the Environment and Earth
+    Sciences, Tel Aviv University, Tel Aviv, Israel
+    (\email{yair.daon@gmail.com}).}}
+%% \author{Yair Daon\thanks{Azrieli Faculty of Medicine, Bar Ilan
+%%     University, Safed, Israel.  (\email{yair.daon@gmail.com}).}}
 
@@
 \input{vanishing}
-
+\input{why}
+%% \input{how}
 %% \section{Open Problems}
@@
 This study is a result of research I started during my PhD studies
-under the instruction of Prof. Georg Stadler of the Courant Institute
+under the instruction of Prof. Georg Stadler at the Courant Institute
 of Mathematical Sciences. I would like to thank him for his great
-mentorship, attention to details and kind behavior. This work was
+mentorship, attention to details and kindness. This work was also
 supported by The Raymond and Beverly Sackler Post-Doctoral
@@
 %% %%\bibliographystyle{amsplain}
-\bibliography{/home/yair/projects/library.bib}
+\bibliography{/home/yair/projects/bibtex.bib}
 

### 2023-05-19 latex/prelims.tex
add explanation for why clusterization

@@
 
-%% \subsection{Overview}
-In this section we present notations that will be used throughout this
-article. These notations are summarized in the definition of Theorem \ref{thm:char}. The theoretical foundations for inverse
-problems over function spaces can be found in \cite{Stuart10}.
+In this section we present the setup that will be used throughout this
+article. The theoretical foundations for inverse problems over
+function spaces can be found in \cite{Stuart10} and will not be
+reviewed here.
 
@@
 subscripts p and o are for ``parameter'' and ``observation'',
-respectively.} and denote $\hilo^* \simeq \hilo$ the Hilbert space of
-all linear functionals on $\hilo$. Let $\fwd: \hilp \to \hilo$ the
-\emph{forward operator}. The forward operator $\fwd$ is assumed linear
-and strongly smoothing (has fast decaying modes --- the heat operator
-of the 1D heat equation described in Section \ref{section:intro} is a
-prime example). Take a Gaussian prior $\param \sim \pr =
-\normal(\prmean ,\prcov)$ with some appropriate covariance operator
-$\prcov$ on $\hilp$ \cite{Stuart10}. Note that $\fwd \prcov \fwd^*$ is
-the prior covariance in $\hilo$ \cite{Stuart10}, and as such, $\fwd
-\prcov \fwd^*$ is assumed invertible --- an assumption which will be
-used below\footnote{If $\fwd$ has a nontrivial kernel we utilize
-Occam's Razor and ignore said kernel.}. Let $m$ the number of
-measurements taken. Measurements are taken via the \emph{measurement
+respectively.} and let $\fwd: \hilp \to \hilo$ the \emph{forward
+operator}. The forward operator $\fwd$ is assumed linear and strongly
+smoothing (the heat operator of the 1D heat equation introduced in
+Section \ref{section:intro} is a prime example). Take a Gaussian prior
+$\param \sim \pr = \normal(\prmean ,\prcov)$ with some appropriate
+covariance operator $\prcov$ on $\hilp$ \cite{Stuart10}. Note that
+$\fwd \prcov \fwd^*$ is the prior covariance in $\hilo$
+\cite{Stuart10}, and as such, $\fwd \prcov \fwd^*$ is assumed
+invertible --- an assumption which will be used below\footnote{If
+$\fwd$ has a nontrivial kernel we utilize Occam's Razor and ignore
+said kernel.}. Measurements are taken via the \emph{measurement
 operator}\footnote{It is common for the measurement and forward
@@
 separated as in \cite{attia2022stochastic, cvetkovic2023choosing}.}
-$\obs \in ( \hilo^* )^m$, whose entries $\meas_j, j=1,\dots,m$ are
-called \emph{measurements}:
+$\obs \in ( \hilo^* )^m$, where $m$ is the number of measurements
+taken. Entries $\meas_j, j=1,\dots,m$ of the observation operator
+$\obs$ are called \emph{measurements}:
 
@@
 \end{equation*}
-%
-%
-
-%% For $u \in \hilo$ and $v\in \R^m$:
-%% %
-%% %
-%% \begin{align*}
-%%   \big (\obs^*v \big ) (u) &= \langle v, \obs u \rangle_{\R^m} = \sum_{j=1}^m  v_j \meas_j(u)
-%%   = v^t \left ( \obs u \right ) = (v^t \obs) (u),
-%% \end{align*}
-%% and thus:
-%% \begin{align}\label{eq:obs*}
-%%   \obs^*v &= \sum_{j=1}^m v_j \meas_j = v^t \obs.
-%% \end{align}
 
@@
 \end{align}
-%% \subsubsection{Error Terms}
+
 It is easy to verify that $\obs \eps' + \eps \in \R^m$ is a centered
@@
 \end{align}
-%% and $e_j \in \R^m$ is the $j$th standard basis vector.
-%% The explicit dependence on $\obs$ will be mostly dropped for
-%% convenience, so $\Sigma(\obs) = \Sigma$.  Thus, for a fixed $\obs$
-%% (i.e.\ a fixed set of observations), $\data = \obs \fwd \param +
-%% \bar{\eps}$ with $\bar{\eps} \sim \normal(0,\Sigma)$.
 Taking $\modcov = 0$ is a common practice
-\cite{Tarantola05,Kaipio2006,Vogel02} and then $\Sigma = \sigma^2I_m$
-is a scalar matrix which does not depend on $\obs$.
-
-%% Taking an error model with a non-scalar covariance allows us to
-%% consider model error (modeled by $\modcov$) as well as observation
-%% error (modeled by $\sigma^2$). For example, say we believe our
-%% forward model does not capture some small scale phenomenon.  Then
-%% we may express this belief by saying $\tru = \fwd + \err$, with
-%% $\fwd$ depending on $\param$ and $\err$ depending on $\sspar$, with
-%% $\param \perp \sspar$. We do not know much about this effect but it
-%% is reasonable to assume it changes continuously in our domain. We
-%% (may choose to) model it as $\normal (0, \modcov)$ and take
-%% $\modcov$ to reflect the spatial (or other) variability we imagine
-%% $\err \sspar$ has. Such small scale phenomenon can arise as a
-%% modeling issue, where we might not model the system in its
-%% entirety. It can also arise from a numerical source, where our
-%% discretization of the system is not fine enough to capture all
-%% small scale phenomena.  In section \ref{section:non vanishing}, we
-%% will see that assuming some correlation in the error mitigates the
-%% clusterization phenomenon, as reported in the literature
-%% \cite{Ucinski05}.
-
-%% \subsubsection{Dynamics and Observation
-%% Operators}\label{subsec:dynamics} Rows of $\obs$ are linear
-%% functionals and we refer to these as {\it observation
-%% vector}s. Note that we cannot choose every observation vector we
-%% want. For example, we may be restricted only to pointwise
-%% evaluations of $\fwd \param$ by the sensors at our disposal, in
-%% which case measuring, e.g., the mean $\ell(u) = \int_{\Omega}u$ of
-%% a function is not possible. Thus, it is best to think of $(\obs
-%% \fwd \param)_j$ as approximations to pointwise evaluations. For
-%% example, if $\hilo = L^2([-\pi, \pi])$, consider $\ell_1(x) :=
-%% (2\pi \eps)^{-\frac{1}{2}}\exp(-\frac{x^2}{2\eps^2})$ and
-%% $\ell_2(x) :=\eps^{-1}\mathbf{1}_{[-\frac{\eps}{2},
-%% \frac{\eps}{2}]}$. Both $\ell_1$ and $\ell_2$ approximate a point
-%% measurement at $0$ for small $\eps$ and, crucially, $\ell_1, \ell_2
-%% \in \hilo^*$.  %% It is important to note that although $\|\ell\|_1
-%% = %% \|\ell_2\|_1=1$, unfortunately $\|\ell_1\|_2^2 = (4\pi %%
-%% \eps)^{-\frac{1}{2}}$ and $\|\ell_2\|_2^2 = \eps^{-1}$.
+\cite{tarantola2005,Kaipio2005,Vogel02} and then $\Sigma =
+\sigma^2I_m$ is a scalar matrix which does not depend on $\obs$.
 
@@
 \end{align}
+Note that the poserior covariance $\postcov$ does not depend on data
+$\data$.
 
@@
 
-%% , and just like in finite dimensions, a D-optimal design maximizes
-%% the Kullback-Leibler divergence between posterior and prior
-%% measures For a linear inverse problem with Gaussian prior and
-%% Gaussian noise, a D-optimal design minimizes an expression
-%% analogous to the determinant of the posterior covariance operator
-%% \cite{AlexanderianGloorGhattas14} (also \cref{thm:d optimality}
-%% below).  D-optimal designs in infinite-dimensional Hilbert spaces
-%% were investigated in \cite{AlexanderianGloorGhattas14}.
 \begin{theorem}[Alexanderian, Gloor, Ghattas \cite{AlexanderianGloorGhattas14}]\label{thm:d optimality}
@@
 \end{theorem}
+
 %% It is important to note that since $\obs$ is finite-rank,

### 2023-05-19 latex/why.tex
add explanation for why clusterization

\section{An answer for Question \ref{q:why}: The cause of measurement clusterization}\label{section:why}
\begin{figure}[H]
    \centering
    \includegraphics[height=0.5\textwidth]{eigenvectors.eps}
    \caption{D-optimal measurement locations ($m=5$ measurements) and
      weighted eigenvectors for inversion of the initial condition of
      the 1D heat equation (see supplementary material for
      details). Measurement locations and weighted eigenvectors are
      plotted over the computational domain $\Omega = [0, 1]$
      (x-axis).  Measurement clusterization occurs approxmately at
      $0.31$ and $0.69$. These two locations are a compromise between
      zerosz of the irrelevant eigenvectors (third and up) and staying
      far from a zero of the first and second eigenvectors.}
  \label{fig:why}
\end{figure}

According to Theorem \ref{thm:char}, D-optimal designs aim to measure
a small subset of eigenvectors; namely, eigenvectors $1$ through
$k$. In our model, not measuring eigenvectors $k+1$ and above is
readily achieved, as the proof of Theorem \ref{thm:char} demonstrates.
Taking this understanding to spatial problems, we expect a D-optimal
design to prefer measurement locations where eigenvectors $k+1$ and
above are (close to) zero: either because their eigenvalue in the
spectrum of the prior is small, or because the value of the
eigenvector itself is (close to) zero. Figure \ref{fig:why}
demonstrates this preference for the 1D heat equation with homogeneous
Dirichlet boundary conditions (details in the supplementary
material). We plot four eigenvectors, scaled according to their prior
standard deviations. We ignore eigenvectors above the fourth, due to
their small prior eigenvalues. We can observe that measurements are
clustered close to zeros of the third and fourth eigenvectors. The
reason for this measurement clusterization is that there are only two
locations where the third and fourth eigenvectors are close to zero,
and the first and second eigenvectors are considerably larger than
zero. When we fit four measurements into these measurement locations,
we find that measurements cluster, in accordance with the pigeonhole
principle.

%% \textbf{Question \ref{q:why}}:

### 2023-05-19 code/main.py

@@ -1,0 +1,112 @@
+import os
+import numpy as np
+import pandas as pd
+from joblib import Parallel, delayed
+from matplotlib import pyplot as plt
+from src.observations import PointObservation
+from scipy.stats import gaussian_kde
+from pdb import set_trace
+
+from src.forward import Heat
+from src.probability import Posterior, Prior
+
+
+# plt.rcParams['figure.figsize'] = [8, 4]
+# plt.rcParams['figure.dpi'] = 200  # 200 e.g. is really fine, but slower
+# plt.rcParams['axes.facecolor'] = 'black'
+# plt.rcParams['figure.facecolor'] = 'black'
+# plt.rcParams['text.color'] = 'white'
+
+def main(n_jobs=6):
+    L, N = 1, 100
+    ms = [1,3, 20]
+    transforms = ['fft']
+    # transforms.append('dct')
+    # fig1, axes1 = plt.subplots(figsize=(30, 16), nrows=len(transforms), ncols=len(ms))
+    # fig2, axes2 = plt.subplots(figsize=(30, 16), nrows=len(transforms), ncols=len(ms))
+    # figs = [fig1, fig2]
+    # axess = [axes1, axes2]
+
+    for col, m in enumerate(ms):
+
+        for transform in transforms:#, fig, axes in zip(transforms, figs, axess):
+            print('\nm =', m, transform)
+
+            prior = Prior(gamma=-1.1, N=N, L=L, transform=transform)
+            forward = Heat(time=3e-2, alpha=0.6, N=N, L=L, transform=transform)
+            post = Posterior(fwd=forward, prior=prior, sigSqr=1, N=N, L=L, transform=transform)
+            # set_trace()
+            # sample = prior.sample().ravel()
+        
+            # Optimal diagonal
+            post.make_optimal_diagonal(m)
+            diagonal_utility = post.diag_utility(post.optimal_diagonal_O)
+            # ax.scatter(diagonal_utility, kde(diagonal_utility), label='optimal diagonal')
+            print(f"utility: Diag={diagonal_utility:3.3f}", end=' ')
+
+            # # Random point measurements
+            # measurements = np.random.uniform(low=0, high=L, size=(100, m))
+            # utilities = Parallel(n_jobs=n_jobs)(delayed(post.minimization_point)(x) for x in measurements)
+            # utilities = -np.array(utilities)
[…]

### 2023-05-19 code/playground.py

@@ -1,0 +1,94 @@
+import numpy as np
+from matplotlib import pyplot as plt
+from scipy.interpolate import interp1d
+from pdb import set_trace as stahp
+
+# from importlib import reload
+# import src.probability
+# reload(src.probability)
+# reload(src.multiplier)
+# reload(src.observations)
+from src.probability import Prior, Posterior
+from src.multiplier import FourierMultiplier
+from src.observations import PointObservation, DiagObservation
+from src.forward import Heat
+
+class TestPrior(Prior):
+    def __init__(self,
+                 eigenvalues,
+                 **kwargs):
+        super().__init__(**kwargs)
+        eigenvalues = np.array(eigenvalues)
+        self.gamma = None
+        self.delta = None
+        self.multiplier = np.zeros(self.N)
+        self.multiplier[:eigenvalues.size] = eigenvalues
+        self.inv_mult = np.full(self.N, np.inf)
+        self.inv_mult[:eigenvalues.size] = 1/eigenvalues
+        
+
+class TestForward(FourierMultiplier):
+    def __init__(self,
+                 eigenvalues,
+                 **kwargs):
+            
+        super().__init__(**kwargs)
+        eigenvalues = np.array(eigenvalues)
+        self.multiplier = np.zeros(self.N)
+        self.multiplier[:eigenvalues.size] = eigenvalues
+
+    def __str__(self):
+        return f'Test Forward operator'
+
+def main():
+    n = 3
+    N = 200
+    L = 1
+    sigSqr = 1e-1
+    prior_eigs = np.ones(n)#np.power(2., -np.arange(n))
+    fwd_eigs = np.ones(n)
+    ms = [1, 2, 3, 5, 7, 10, 12]
+    transforms = ['dct', 'fft']
+    for transform in transforms:
+        print('\n\n')
+        print(transform)
+        for m in ms:
+            print(m)
+            prior = TestPrior(N=N, L=L, transform=transform, eigenvalues=prior_eigs)
+            fwd = TestForward(N=N, L=L, transform=transform, eigenvalues=fwd_eigs)
+            post = Posterior(prior=prior, fwd=fwd, N=N, L=L, transform=transform, sigSqr=sigSqr)
+            print('yasss')
+            diagonal_O = post.make_optimal_diagonal(m)
+            print('naaa')
+            # print(diagonal_O)
+            eigs = prior.multiplier ** 2 * fwd.multiplier
[…]

### 2023-05-19 playground.py
with pix

@@ -1,89 +1,0 @@
-import numpy as np
-from matplotlib import pyplot as plt
-from scipy.interpolate import interp1d
-from pdb import set_trace as stahp
-
-# from importlib import reload
-# import src.probability
-# reload(src.probability)
-# reload(src.multiplier)
-# reload(src.observations)
-from src.probability import Prior, Posterior
-from src.multiplier import FourierMultiplier
-from src.observations import PointObservation, DiagObservation
-from src.forward import Heat
-
-class TestPrior(Prior):
-    def __init__(self,
-                 eigenvalues,
-                 **kwargs):
-        super().__init__(**kwargs)
-        eigenvalues = np.array(eigenvalues)
-        self.gamma = None
-        self.delta = None
-        self.multiplier = np.zeros(self.N)
-        self.multiplier[:eigenvalues.size] = eigenvalues
-        self.inv_mult = np.full(self.N, np.inf)
-        self.inv_mult[:eigenvalues.size] = 1/eigenvalues
-        
-
-class TestForward(FourierMultiplier):
-    def __init__(self,
-                 eigenvalues,
-                 **kwargs):
-            
-        super().__init__(**kwargs)
-        eigenvalues = np.array(eigenvalues)
-        self.multiplier = np.zeros(self.N)
-        self.multiplier[:eigenvalues.size] = eigenvalues
-
-    def __str__(self):
-        return f'Test Forward operator'
-
-def main():
-    n = 3
-    N = 1000
-    L = 1
-    sigSqr = 1e-1
-    prior_eigs = np.ones(n)#np.power(2., -np.arange(n))
-    fwd_eigs = np.ones(n)
-    ms = [1, 2, 3, 5, 7, 10, 12]
-    transforms = ['dct', 'fft']
-    for transform in transforms:
-        print('\n\n')
-        print(transform)
-        for m in ms:
-            prior = TestPrior(N=N, L=L, transform=transform, eigenvalues=prior_eigs)
-            fwd = TestForward(N=N, L=L, transform=transform, eigenvalues=fwd_eigs)
-            post = Posterior(prior=prior, fwd=fwd, N=N, L=L, transform=transform, sigSqr=sigSqr)
-            diagonal_O = post.make_optimal_diagonal(m)
-            # print(diagonal_O)
-            eigs = prior.multiplier ** 2 * fwd.multiplier
-            psi = sum(np.log(1 + eig * dia**2 /sigSqr) for eig, dia in zip(eigs, diagonal_O))
[…]

### 2023-05-21 latex/intro.tex
I think its donegit add why.tex

@@
   code in supplementary material). Given our model's genericity, we
-  expect that measurement clusterization to be a generic phenomenon.
+  expect measurement clusterization to be a generic phenomenon.
 
@@
   design focuses on a select set of prior eigenvectors, specifically
-  those with the highest power spectrum. In practical scenarios, the
-  number of locations where the relevant prior eigenvectors are
-  significantly large, while other eigenvectors have a power spectrum
-  of zero, is limited. Consequently, the clustering of measurements
-  arises as a natural consequence of the pigeonhole principle, as
-  there are more measurements available than there are locations
-  satisfying conditions (a) and (b).
+  those with the largest eigenvalues in the prior covariance
+  spectrum. In practical scenarios, the number of locations where the
+  relevant prior eigenvectors are significantly large, while other
+  eigenvectors have a power spectrum close to zero, is
+  limited. Consequently, the clustering of measurements arises as a
+  natural consequence of the pigeonhole principle, as there are more
+  measurements available than there are locations satisfying
+  conditions (a) and (b).
 
@@
 
-Lastly, in Lemma \ref{lemma:free} we decompose a symmetric positive
-definite matrix $M \in \mathbb{R}^{k \times k}$ with $\ttr M = m \in
-\mathbb{N}$ as $M = AA^t$, where $A$ has unit norm columns.
+In the proving Theorem \ref{thm:char} we prove and generalize several
+lemmas. Among those, is a Lemma \ref{lemma:free}, which is (to the the
+best of my knowledge) a novel lemma in linear algebra: We decompose a
+symmetric positive definite matrix $M \in \mathbb{R}^{k \times k}$
+with $\ttr M = m \in \mathbb{N}$ as $M = AA^t$, where $A$ has unit
+norm columns.
 

### 2023-05-21 latex/paper.tex
I think its donegit add why.tex

@@
 \begin{abstract}
-  Inversion for parameteres of physical processes typically requires
+  Inversion for parameters of physical processes typically requires
   taking expensive measurements; finding an optimal set of
@@
   analytically tractable model for Bayesian D-optimal designs in
-  linear inverse problems over Hilbert spaces. We prove that
-  measurement clusterization poses no obstruction to D-optimality. We
-  demonstrate that measurement clusterization is a generic property of
-  D-optimal designs, and prove that correlated noise between
-  measurements mitigates clusterization. We also give a full
-  characterization of D-optimal designs, which shows that D-optimal
-  designs uniformly reduce uncertainty in the space of
-  observations. Finally, we show how measurement clusterization can be
-  avoided without changing the optimal design problem.
+  linear inverse problems over Hilbert spaces. We demonstrate that
+  measurement clusterization is a generic property of D-optimal
+  designs, and prove that correlated noise between measurements
+  mitigates clusterization. We also give a full characterization of
+  D-optimal designs, which shows that D-optimal designs uniformly
+  reduce uncertainty in a select subset of the prior covariance
+  eigenvectors. Finally, we show how measurement clusterization is a
+  simple consequence of the characterization mentioned above and the
+  pigeonhole principle.
 \end{abstract}
@@
 of Mathematical Sciences. I would like to thank him for his great
-mentorship, attention to details and kindness. This work was also
-supported by The Raymond and Beverly Sackler Post-Doctoral
-Scholarship.
+mentorship, attention to details and kindness. I would also like to
+thank Christian Remling, who helped me find a proof for Lemma
+\ref{lemma:free} in
+\href{https://mathoverflow.net/questions/280168/redistribute-diagonal-entries-of-a-matrix/280203#280203c}{Mathoverflow}.
 
@@
 the author's and do not necessarily reflect the policies and views of
-NSF, ORAU/ORISE, or DOE.
+NSF, ORAU/ORISE, or DOE. This work was also supported by The Raymond
+and Beverly Sackler Post-Doctoral Scholarship.
 

### 2023-05-21 latex/why.tex
I think its donegit add why.tex

@@
       weighted eigenvectors for inversion of the initial condition of
-      the 1D heat equation (see supplementary material for
-      details). Measurement locations and weighted eigenvectors are
-      plotted over the computational domain $\Omega = [0, 1]$
-      (x-axis).  Measurement clusterization occurs approxmately at
-      $0.31$ and $0.69$. These two locations are a compromise between
-      zerosz of the irrelevant eigenvectors (third and up) and staying
-      far from a zero of the first and second eigenvectors.}
+      the 1D heat equation. Measurement locations and weighted
+      eigenvectors are plotted over the computational domain $\Omega =
+      [0, 1]$ (x-axis). Measurement clusterization occurs
+      approximately at $0.31$ and $0.69$. These two locations are a
+      compromise between zeros of eigenvectors a D-optimal design aims
+      to ignore (third and up) and staying far from a zero of the
+      first and second eigenvectors. Allocating $m=5$ measurements
+      into two locations results in clusterization, according to the
+      pigeonhole principle.}
   \label{fig:why}
@@
 
-According to Theorem \ref{thm:char}, D-optimal designs aim to measure
-a small subset of eigenvectors; namely, eigenvectors $1$ through
-$k$. In our model, not measuring eigenvectors $k+1$ and above is
-readily achieved, as the proof of Theorem \ref{thm:char} demonstrates.
-Taking this understanding to spatial problems, we expect a D-optimal
-design to prefer measurement locations where eigenvectors $k+1$ and
-above are (close to) zero: either because their eigenvalue in the
-spectrum of the prior is small, or because the value of the
-eigenvector itself is (close to) zero. Figure \ref{fig:why}
-demonstrates this preference for the 1D heat equation with homogeneous
-Dirichlet boundary conditions (details in the supplementary
-material). We plot four eigenvectors, scaled according to their prior
-standard deviations. We ignore eigenvectors above the fourth, due to
-their small prior eigenvalues. We can observe that measurements are
-clustered close to zeros of the third and fourth eigenvectors. The
-reason for this measurement clusterization is that there are only two
-locations where the third and fourth eigenvectors are close to zero,
-and the first and second eigenvectors are considerably larger than
-zero. When we fit four measurements into these measurement locations,
-we find that measurements cluster, in accordance with the pigeonhole
-principle.
-
-%% \textbf{Question \ref{q:why}}:
+According to Theorem \ref{thm:char}, D-optimal designs aim to capture
+a small subset of eigenvectors of the prior covariance, specifically
+the $k$ eigenvectors with the highest prior variance. Our model
+naturally achieves this objective by not measuring eigenvectors $k+1$
+and above, as proven in Theorem \ref{thm:char}. Translating this
+understanding to spatial problems, we anticipate that a D-optimal
+design would favor measurement locations where eigenvectors $k+1$ and
+above are either close to zero in value or possess small eigenvalues
+in the prior spectrum, for some $k > 0$. To illustrate this
+preference, Figure \ref{fig:why} depicts the scenario using the 1D
+heat equation with homogeneous Dirichlet boundary
+conditions\footnote{Additional details provided in the supplementary
+material.}. The plot showcases four eigenvectors, scaled according to
+their prior standard deviations. Since eigenvectors beyond the fourth
+have insignificant prior eigenvalues, we exclude them from
+consideration. Notably, we observe that measurements are clustered
+near the zeros of the third and fourth eigenvectors, so we conclude
+that $k=2$. Clusterization arises because there are only two locations
+where the third and fourth eigenvectors approach zero, while the first
+and second eigenvectors exhibit significantly non-zero
+values. Consequently, when we allocate four measurements to these two
+locations, they naturally cluster together, aligning with the
+pigeonhole principle.

### 2023-05-21 arxiv/prelims.tex
working on arxiv

--- 2023-05-19 latex/prelims.tex
@@
 \subsection{Bayesian Linear Inverse Problems}\label{subsec:abstract OED}
-Let $\hilp$ and $\hilo$ be separable Hilbert spaces,\footnote{The
-subscripts p and o are for ``parameter'' and ``observation'',
-respectively.} and let $\fwd: \hilp \to \hilo$ the \emph{forward
-operator}. The forward operator $\fwd$ is assumed linear and strongly
-smoothing (the heat operator of the 1D heat equation introduced in
-Section \ref{section:intro} is a prime example). Take a Gaussian prior
-$\param \sim \pr = \normal(\prmean ,\prcov)$ with some appropriate
-covariance operator $\prcov$ on $\hilp$ \cite{Stuart10}. Note that
-$\fwd \prcov \fwd^*$ is the prior covariance in $\hilo$
-\cite{Stuart10}, and as such, $\fwd \prcov \fwd^*$ is assumed
-invertible --- an assumption which will be used below\footnote{If
-$\fwd$ has a nontrivial kernel we utilize Occam's Razor and ignore
-said kernel.}. Measurements are taken via the \emph{measurement
-operator}\footnote{It is common for the measurement and forward
-operators to be merged $\tmp := \obs \fwd$
+Let $\hilp$ and $\hilo$ be separable Hilbert spaces (the subscripts p
+and o are for ``parameter'' and ``observation'', respectively), and
+let $\fwd: \hilp \to \hilo$ the \emph{forward operator}. The forward
+operator $\fwd$ is assumed linear and strongly smoothing (the heat
+operator of the 1D heat equation introduced in Section
+\ref{section:intro} is a prime example). Take a Gaussian prior $\param
+\sim \pr = \normal(\prmean ,\prcov)$ with some appropriate covariance
+operator $\prcov$ on $\hilp$ \cite{Stuart10}. Note that $\fwd \prcov
+\fwd^*$ is the prior covariance in $\hilo$ \cite{Stuart10}, and as
+such, $\fwd \prcov \fwd^*$ is assumed invertible --- an assumption
+which will be used below (and if $\fwd$ has a nontrivial kernel we
+utilize Occam's Razor and ignore said kernel). Measurements are taken
+via the \emph{measurement operator}. It is common for the measurement
+and forward operators to be merged $\tmp := \obs \fwd$
 \cite{AlexanderianGloorGhattas14}, but the analysis carried out in the
 following sections requires that $\fwd$ and $\obs$ are explicitly
-separated as in \cite{attia2022stochastic, cvetkovic2023choosing}.}
+separated as in \cite{attia2022stochastic, cvetkovic2023choosing}.
 $\obs \in ( \hilo^* )^m$, where $m$ is the number of measurements
@@
   alexanderian2018efficient}. The main result we will make use of is
-summarized (in our notation) in Theorem \ref{thm:d optimality}\footnote{In
-\cite{AlexanderianGloorGhattas14, alexanderian2018efficient}, results
-are stated for $\Sigma=I$ (implied by $\modcov = 0,\sigma^2=1$), but
-the authors also mention these results hold for more general
-covariance matrices \cite[p. 681]{AlexanderianGloorGhattas14}.}:
+summarized (in our notation) in Theorem \ref{thm:d optimality}:
 
@@
 
+Note that in \cite{AlexanderianGloorGhattas14,
+  alexanderian2018efficient}, results are stated for $\Sigma=I$
+(implied by $\modcov = 0,\sigma^2=1$), but the authors also mention
+these results hold for more general covariance matrices
+\cite[p. 681]{AlexanderianGloorGhattas14}.
+
 %% It is important to note that since $\obs$ is finite-rank,
@@
 and Gaussian noise, a D-optimal design minimizes the determinant of
-the posterior covariance matrix\footnote{The frequentist version of
-the D-optimality criterion aims to minimize the determinant of the
-Fisher information matrix \cite[page 16]{Ucinski05}. Thus, the
-Bayesian criterion is just a regularized version of the frequentist
-criterion \cite{Chaloner1995}.}. Theorem \ref{thm:d optimality} and
-Definition \ref{def:d optimality} carry a similar intuition:
+the posterior covariance matrix, and this turns out to be a
+regularized version of the frequentist criterion
+\cite{Chaloner1995}. Theorem \ref{thm:d optimality} and Definition
+\ref{def:d optimality} carry a similar intuition:
 \begin{align*}

### 2023-05-21 arxiv/vanishing.tex
working on arxiv

--- 2023-05-14 latex/vanishing.tex
@@
-\section{Characterizing D-Optimal Designs when $\modcov = 0$}\label{section:vanishing}
-We will now prove Theorem \ref{thm:char} which characterizes D-optimal
-designs when $\modcov = 0$. The necessary first-order condition for
-D-optimality of Theorem \ref{thm:constrained} for $\modcov = 0$
-become:
+\section{Characterization of D-Optimal Designs when $\modcov = 0$}\label{section:vanishing}
+Our goal in this section is to prove Theorem \ref{thm:char} which
+characterizes D-optimal designs when $\modcov = 0$. The necessary
+first-order condition for D-optimality of Theorem
+\ref{thm:constrained} for $\modcov = 0$ become:
 
@@
 %% \end{figure}
-\section{Clusterization poses no obstruction to D-optimality}\label{section:clusterization}
-The necessary conditions for D-optimality of Theorem \ref{thm:char}
-leave some freedom in choosing $\obs$, which might allow us to avoid a
-clustered design. As an example, consider an inverse problem for which
-the first two eigenvalues of the pushforward prior $\fwd \prcov\fwd^*$
-are $\lambda_1 = 10, \lambda_2 = 5$, with eigenvectors
-$\{\ev_i\}_{i=1}^\infty$ and $\sigma^2=1$. Consider the following
-$\obs, \pobs$ (see also Figure \ref{fig:clusterization}):
-\begin{equation}\label{eq:two designs}
-  \obs =
-  \begin{pmatrix}
-    \sqrt{\frac{11}{40}}\ev_1 - \sqrt{\frac{29}{40}}\ev_2 \\
-    \sqrt{\frac{11}{40}}\ev_1 + \sqrt{\frac{29}{40}}\ev_2 \\
-    \ev_1 \\
-    \ev_2 \\
-    \ev_1 \\    
-  \end{pmatrix},
-  \pobs =
-  \begin{pmatrix}
-    \sqrt{\frac{3}{8}}\ev_1 - \sqrt{\frac{5}{8}}\ev_2 \\
-    \sqrt{\frac{3}{8}}\ev_1 + \sqrt{\frac{5}{8}}\ev_2 \\
-    \sqrt{\frac{2}{5}}\ev_1 - \sqrt{\frac{3}{5}}\ev_2 \\
-    \sqrt{\frac{2}{5}}\ev_1 + \sqrt{\frac{3}{5}}\ev_2 \\
-    \ev_1 \\    
-  \end{pmatrix}
-\end{equation}
-\noindent It is easy to verify that $\obs^*\obs = \pobs^*\pobs$
-and that both $\obs$ and $\pobs$ are D-optimal. However, $\obs$
-is a clustered design ($\meas_3 = \meas_5 = \ev_1$), while $\pobs$ is
-not.
-
-We are now ready to answer two of the questions posed in the
-introduction:
-
-\textbf{Question \ref{q:why}}: Since $\obs$ is clustered and
-D-optimal, measurement clusterization poses no obstruction to
-D-optimality. Furthermore, a computer implementation of Lemma
-\ref{thm:char} also generates $\obs$ as a solution to the D-optimal
-design problem. Furthermore, randomized numerical simulations give
-rise to D-optimal designs that exhibit clusterization more than 95\% of
-the time (see code in supplementary). Given our model's genericity, we
-conclude that measurement clusterization may be a generic phenomenon.
-
-\textbf{Question \ref{q:replace}}: We showed that it is possible to
-substitute a D-optimal clustered design (e.g. $\obs$) for a D-optimal
-non-clustered design (e.g. $\pobs$). Presenting a general algorithm
-for replacing a clustered D-optimal design with a non-clustered
-D-optimal design is out of the scope of the current study. We can,
-however, provide some intuition, utilizing $\obs, \pobs$ as a test
-case. Let $\obs_{12}$ the design composed of the first and second rows
-of $\obs$ (with a similar definition for $\pobs_{12}$), and note that,
-up to permutations and sign changes, $\obs_{12}$ uniquely determines
-$\obs$ as a clustered design. We find $\pobs_{12}$ by requiring that
-(a) measurements in $\pobs_{12}$ have unit norm, (b) $\pobs_{12}^*
-\pobs_{12}$ has the same eigenvectors as $\obs^*\obs$, (c)
-$\pobs^*\pobs \approx \obs^*\obs$, and (d) $\pobs^*\pobs \neq
-\obs^*\obs$. We then consider a ``new'' inverse problem, for which the
-pushforward prior covariance is the covariance of the posterior
-$\mu_{\textup{post}}^{\data, \pobs_{12}}$, namely
-
-$$
-\left(\left(\fwd \prcov\fwd^*\right)^{-1} + \sigma^{-2}
-\pobs_{12}^*\pobs_{12}\right)^{-1}.
-$$
-%
-Finally, we complete $\pobs$ to a D-optimal design with $m=5$
-measurements via the construction outlined in Lemma
-\ref{lemma:free}.
-
-Interestingly, part (c) in the process detailed above implies that
-there are actually infinitely many non-clustered D-optimal designs. It
-is not at all clear why the numerical implementation of Lemma
-\ref{lemma:free} results in a clustered design. An answer to this
-question may further help shed light on the measurement clusterization
-phenomenon.
-
-\pgfplotstableread{
-  Label    prior  $o_1o_2$  $o_3$   $o_4$   $o_5$    topper
-  1         0.1      0.55       1       0       1      0.001
-  2         0.2      1.45       0       1       0      0.001
-  3         3.5      0          0       0       0      0.001
-}\clusterization
-
-\pgfplotstableread{
-  Label    prior  $o_1o_2$  $o_3o_4$   $o_5$     topper
-  1         0.1    0.75     0.8        1         0.001
-  2         0.2    1.25     1.2        0         0.001
-  3         3.5    0        0          0         0.001
-}\noclusterization
-
-\begin{figure}
-  \begin{tikzpicture}[scale=0.85]
-    \begin{axis}[
-        ybar stacked,
-        ymin=0,
-        ymax=4,
-        xtick=data,
-        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
-        reverse legend=false, % set to false to get correct display, but I'd like to have this true
-        xticklabels from table={\clusterization}{Label},
-        xticklabel style={text width=2cm,align=center},
-        legend plot pos=right,
-        ylabel=precision --- prior and posterior,
-        xlabel=eigenvector $i$,
-      ]
-
-      \addplot [fill=green!80]  table [y=prior, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=blue!60]   table [y=$o_1o_2$, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=red!60]    table [y=$o_3$, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=black!60]  table [y=$o_4$, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=orange!60] table [y=$o_5$, meta=Label, x expr=\coordindex] {\clusterization};
-      %% \addplot [fill=cyan!60]   table [y=$o_6$, meta=Label, x expr=\coordindex] {\clusterization};
-      %% \addplot [fill=purple!60] table [y=$o_7$, meta=Label, x expr=\coordindex] {\clusterization};
-
-      \addlegendentry{prior}
-      \addlegendentry{$o_1o_2$}
-      \addlegendentry{$o_3$}
-      \addlegendentry{$o_4$}
-      \addlegendentry{$o_5$}
-      %% \addlegendentry{$o_6$}
-      %% \addlegendentry{$o_7$}   
-    \end{axis}
-  \end{tikzpicture}
-  \qquad
-  \begin{tikzpicture}[scale=0.85]
-    \begin{axis}[
-        ybar stacked,
-        ymin=0,
-        ymax=4,
-        xtick=data,
-        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
-        reverse legend=false, % set to false to get correct display, but I'd like to have this true
-        xticklabels from table={\noclusterization}{Label},
-        xticklabel style={text width=2cm,align=center},
-        legend plot pos=right,
-        ylabel=precision --- prior and posterior,
-        xlabel=eigenvector $i$,
-      ]
-
-      \addplot [fill=green!80]  table [y=prior, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=blue!60]   table [y=$o_1o_2$, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=red!60]    table [y=$o_3o_4$, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=black!60]  table [y=$o_5$, meta=Label, x expr=\coordindex] {\noclusterization};
-      %% \addplot [fill=orange!60] table [y=$o_4$, meta=Label, x expr=\coordindex] {\noclusterization};
-      %% \addplot [fill=cyan!60]   table [y=$o_5$, meta=Label, x expr=\coordindex] {\noclusterization};
-      %% \addplot [fill=purple!60] table [y=$o_6$, meta=Label, x expr=\coordindex] {\noclusterization};
-
-      \addlegendentry{prior}
-      \addlegendentry{$o_1o_2$}
-      \addlegendentry{$o_3o_4$}
-      \addlegendentry{$o_5$}
-      %% \addlegendentry{$o_4$}
-      %% \addlegendentry{$o_5$}
-      %% \addlegendentry{$o_6$}   
-    \end{axis}
-  \end{tikzpicture}
-  \caption{Clusterization (left, $\obs$ in \eqref{eq:two designs})
-    and non-clusterization (right, $\pobs$ in \eqref{eq:two designs})
-    in D-optimal designs. Posterior precision per eigenvector is
-    plotted for $\obs$ and $\pobs$ from \eqref{eq:two
-      designs}. Both designs are identical for all practical matters
-    --- their posteriors are equal. In the left panel, a D-optimal
-    design with clusterization is illustrated: $\meas_3 = \meas_5$. In
-    the right panel, a D-optimal design without clusterization is
-    illustrated. The contribution to the posterior precision due to
-    $\meas_1$ and $\meas_2$ is denoted $\meas_1\meas_2$. These
-    contributions are plotted together because neither $\meas_1$ nor
-    $\meas_2$ are not in the direction of any eigenvector. Similarly
-    for $\meas_3\meas_4$.}
-  \label{fig:clusterization}
-\end{figure}

### 2023-05-21 arxiv/why.tex
working on arxiv

--- 2023-05-21 latex/why.tex
@@
-\section{An answer for Question \ref{q:why}: The cause of measurement clusterization}\label{section:why}
-\begin{figure}[H]
+\section{Answer to Question \ref{q:why}: The cause of measurement clusterization}\label{section:why}
+\begin{figure}
     \centering
-    \includegraphics[height=0.5\textwidth]{eigenvectors.eps}
+    \includegraphics[height=0.5\textwidth]{eigenvectors.pdf}
     \caption{D-optimal measurement locations ($m=5$ measurements) and
@@
 preference, Figure \ref{fig:why} depicts the scenario using the 1D
-heat equation with homogeneous Dirichlet boundary
-conditions\footnote{Additional details provided in the supplementary
-material.}. The plot showcases four eigenvectors, scaled according to
-their prior standard deviations. Since eigenvectors beyond the fourth
-have insignificant prior eigenvalues, we exclude them from
-consideration. Notably, we observe that measurements are clustered
-near the zeros of the third and fourth eigenvectors, so we conclude
-that $k=2$. Clusterization arises because there are only two locations
-where the third and fourth eigenvectors approach zero, while the first
-and second eigenvectors exhibit significantly non-zero
-values. Consequently, when we allocate four measurements to these two
-locations, they naturally cluster together, aligning with the
-pigeonhole principle.
+heat equation with homogeneous Dirichlet boundary conditions (details
+in the supplementary material). The plot showcases four eigenvectors,
+scaled according to their prior standard deviations. Since
+eigenvectors beyond the fourth have insignificant prior eigenvalues,
+we exclude them from consideration. Notably, we observe that
+measurements are clustered near the zeros of the third and fourth
+eigenvectors, so we conclude that $k=2$. Clusterization arises because
+there are only two locations where the third and fourth eigenvectors
+approach zero, while the first and second eigenvectors exhibit
+significantly non-zero values. Consequently, when we allocate four
+measurements to these two locations, they naturally cluster together,
+aligning with the pigeonhole principle.

### 2023-05-21 arxiv/intro.tex
working on arxiv
Ready for arxiv

--- 2023-05-21 latex/intro.tex
@@
 remarkably similar measurements in certain cases \cite{fedorov1996,
-  hooker2009, fedorov1997, Ucinski05, neitzel2019sparse}. To
+  nyberg2012, fedorov1997, Ucinski05, neitzel2019sparse}. To
 illustrate this point, we consider a toy inverse problem: inferring
@@
 %% fifth:magenta
-\begin{figure}[H]
+\begin{figure}
     \centering
-    \includegraphics[height=0.5\textwidth]{example.eps}
+    \includegraphics[height=0.5\textwidth]{example.pdf}
     \caption{Measurement clusterization for optimal designs when
@@
 Researchers widely agree that measurement clusterization is
-undesirable \cite{fedorov1996, hooker2009, fedorov1997, Ucinski05,
+undesirable \cite{fedorov1996, nyberg2012, fedorov1997, Ucinski05,
   neitzel2019sparse}, prompting the exploration of various remedies to
@@
 modeling of auto-correlation time within the noise terms
-\cite{hooker2009}.
+\cite{nyberg2012}.
 

### 2023-05-21 arxiv/ms.tex
working on arxiv
Ready for arxiv

--- 2023-05-21 latex/paper.tex
@@
-\documentclass{siamonline220329}
-%\documentclass{amsart}
-%\documentclass{article}
+%\documentclass{siamonline220329}
+\documentclass{amsart}
 \input{definitions}
@@
 
-\title{Measurement Clusterization in D-optimal Designs for Bayesian Linear
-  Inverse Problems over Hilbert Spaces}
+\title[Measurement Clusterization in D-optimal Designs]{Measurement
+  Clusterization in D-optimal Designs for Bayesian Linear Inverse
+  Problems over Hilbert Spaces}
 
-\author{Yair Daon\thanks{Porter School of the Environment and Earth
-    Sciences, Tel Aviv University, Tel Aviv, Israel
-    (\email{yair.daon@gmail.com}).}}
+\author{Yair Daon}
+
+%% \thanks{Porter School of the Environment and Earth
+%%     Sciences, Tel Aviv University, Tel Aviv, Israel
+%%     (\email{yair.daon@gmail.com}).}
 %% \author{Yair Daon\thanks{Azrieli Faculty of Medicine, Bar Ilan
@@
 
-%% \address{Porter School of the Environment and Earth
-%%   Sciences, Tel Aviv University\\ Tel Aviv, Israel}
+\address{Porter School of the Environment and Earth Sciences, Tel Aviv
+  University\\ Tel Aviv, Israel}
+\address{Sackler Faculty of Medicine,
+  Sciences, Tel Aviv University\\ Tel Aviv, Israel}
+
 %% %% \curraddr{Porter School of Environment Studies, Tel Aviv University\\ Tel Aviv, Israel}
 %% %% \curraddr{}
-%% \email{yair.daon@gmail.com}
-% The 2010 edition of the Mathematics Subject Classification is the
-% current definitive version.
-%% \subjclass{Primary: 
-%%   62F15, % Statistics - bayesian inference
-%%   35R30, % PDE - Inverse problems
-%%   Secondary:
-%%   28C20,
-%%   % Measure and integration - set functions and measures and integrals
-%%          % in infinite-dimensional spaces (Wiener measure, Gaussian
-%%   % measure, etc.
-%% }
+\email{yair.daon@gmail.com}
+%% The 2010 edition of the Mathematics Subject Classification is the
+%% current definitive version.
+
+\subjclass{Primary: 
+  62F15, % Statistics - bayesian inference
+  35R30, % PDE - Inverse problems
+  Secondary:
+  28C20.
+  % Measure and integration - set functions and measures and integrals
+         % in infinite-dimensional spaces (Wiener measure, Gaussian
+  % measure, etc.
+}
 
@@
 \begin{document}
-\maketitle
+
 \begin{abstract}
@@
 \end{abstract}
-
 \maketitle
@@
 
-\bibliographystyle{siamplain}
-%% %%\bibliographystyle{amsplain}
+%\bibliographystyle{siamplain}
+\bibliographystyle{amsplain}
 \bibliography{/home/yair/projects/bibtex.bib}

### 2023-05-21 arxiv/prelims.tex
Ready for arxiv

@@
 
-It is useful to record that the posterior measure $\post$ is Gaussian
-in this setting and its covariance operator is \cite{Stuart10}:
+Finally, it is useful to record that the posterior measure $\post$ is
+Gaussian in this setting and its covariance operator is does not
+depend on data $\data$ \cite{Stuart10}:
 \begin{align}\label{eq:postcov}
@@
 \end{align}
-Note that the poserior covariance $\postcov$ does not depend on data
-$\data$.
 
@@
   alexanderian2018efficient}. The main result we will make use of is
-summarized (in our notation) in Theorem \ref{thm:d optimality}:
+summarized in our notation below:
 
@@
   alexanderian2018efficient}, results are stated for $\Sigma=I$
-(implied by $\modcov = 0,\sigma^2=1$), but the authors also mention
-these results hold for more general covariance matrices
+(implied by $\modcov = 0,\sigma^2=1$), but the these results also hold
+for more general covariance matrices
 \cite[p. 681]{AlexanderianGloorGhattas14}.
@@
 \begin{definition}\label{def:d optimality}
-  $\obs^{\star}$ is said to be \emph{D-optimal} if $\obs^{\star} =
+  We say $\obs^{\star}$ is \emph{D-optimal} if $\obs^{\star} =
   \argmax_{\obs} \tar(\obs)$, where entries of $\obs \in (\hilo^*)^m$
@@
 \end{align*}
-We think of $\prcov$ as constant, so a D-optimal design minimizes the
-posterior covariance determinant, analogously to the
-finite-dimensional case.
+We think of $\prcov$ as constant, so a D-optimal design minimizes a
+quantity analogous to the posterior covariance determinant, similarly
+to the finite-dimensional case.
 

### 2023-05-21 arxiv/vanishing.tex
Ready for arxiv

@@
-\section{Characterization of D-Optimal Designs when $\modcov = 0$}\label{section:vanishing}
+\section{D-Optimal Designs Without Model Error}\label{section:vanishing}
 Our goal in this section is to prove Theorem \ref{thm:char} which
@@
 The final ingredient we require for the proof of Theorem
-\ref{thm:char} the following lemma:
+\ref{thm:char} is:
 
@@
 \begin{figure}
-  \begin{tikzpicture}[scale=0.85]
+  \centering
+  \begin{tikzpicture}[scale=0.735]
     \begin{axis}[
@@
         legend plot pos=right,
-        ylabel=precision --- prior and posterior,
-        xlabel=eigenvector $i$,
+        ylabel={\Large precision --- prior and posterior},
+        xlabel={\Large eigenvector $i$},
       ]
-
       \addplot [fill=green!80]  table [y=prior,   meta=Label, x expr=\coordindex] {\optimalvsnot};
       \addplot [fill=blue!60]   table [y=optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};     
-
-      \addlegendentry{$\sigma^2\lambda_i^{-1}$}
-      \addlegendentry{optimal $\eta_i$s}
-
+      \addlegendentry{\Large $\sigma^2\lambda_i^{-1}$}
+      \addlegendentry{\Large optimal $\eta_i$s}
     \end{axis}
   \end{tikzpicture}
-  \qquad
-  \begin{tikzpicture}[scale=0.85]
+  \begin{tikzpicture}[scale=0.735]
     \begin{axis}[
@@
         legend plot pos=right,
-        ylabel=precision --- prior and posterior,
-        xlabel=eigenvector $i$,
+        ylabel={\Large precision --- prior and posterior},
+        xlabel={\Large eigenvector $i$},
       ]   
-
       \addplot [fill=green!80]  table [y=prior,       meta=Label, x expr=\coordindex] {\optimalvsnot};
       \addplot [fill=blue!60]   table [y=sub-optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};
-
-      \addlegendentry{$\sigma^2\lambda_i^{-1}$}
-      \addlegendentry{sub-optimal $\eta_i$s}
-
+      \addlegendentry{\Large $\sigma^2\lambda_i^{-1}$}
+      \addlegendentry{\Large sub-optimal $\eta_i$s}
     \end{axis}

### 2023-05-21 arxiv/ms.tex
To ArXiV

@@
 \begin{abstract}
-  Inversion for parameters of physical processes typically requires
-  taking expensive measurements; finding an optimal set of
-  measurements is known as the optimal design problem. Surprisingly,
-  measurement locations in optimal designs are sometimes extremely
-  clustered. Since measurement clusterization is nonintuitive,
-  researchers often avoid it by modifying the optimal design
-  problem. We consider a certain flavor of the optimal design problem,
-  based on the Bayesian D-optimality criterion, and suggest an
-  analytically tractable model for Bayesian D-optimal designs in
+  Inversion for parameters of a physical process typically requires
+  taking expensive measurements, and the task of finding an optimal
+  set of measurements is known as the optimal design
+  problem. Surprisingly, measurement locations in optimal designs are
+  sometimes extremely clustered, and since measurement clusterization
+  is nonintuitive, researchers often avoid it by modifying the optimal
+  design problem. We consider a certain flavor of the optimal design
+  problem, based on the Bayesian D-optimality criterion, and suggest
+  an analytically tractable model for D-optimal designs in Bayesian
   linear inverse problems over Hilbert spaces. We demonstrate that
@@
   D-optimal designs, which shows that D-optimal designs uniformly
-  reduce uncertainty in a select subset of the prior covariance
+  reduce uncertainty in a select subset of prior covariance
   eigenvectors. Finally, we show how measurement clusterization is a
-  simple consequence of the characterization mentioned above and the
+  consequence of the characterization mentioned above and the
   pigeonhole principle.

### 2023-05-21 arxiv/vanishing.tex
To ArXiV

@@
         legend plot pos=right,
-        ylabel={\Large precision --- prior and posterior},
-        xlabel={\Large eigenvector $i$},
+        ylabel={\LARGE precision --- prior and posterior},
+        xlabel={\LARGE eigenvector},
       ]
@@
       \addplot [fill=blue!60]   table [y=optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};     
-      \addlegendentry{\Large $\sigma^2\lambda_i^{-1}$}
-      \addlegendentry{\Large optimal $\eta_i$s}
+      \addlegendentry[scale=1.4]{$\sigma^2\lambda_i^{-1}$}
+      \addlegendentry[scale=1.4]{optimal $\eta_i$s}
     \end{axis}
@@
         legend plot pos=right,
-        ylabel={\Large precision --- prior and posterior},
-        xlabel={\Large eigenvector $i$},
+        ylabel={\LARGE precision --- prior and posterior},
+        xlabel={\LARGE eigenvector} ,
       ]   
@@
       \addplot [fill=blue!60]   table [y=sub-optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};
-      \addlegendentry{\Large $\sigma^2\lambda_i^{-1}$}
-      \addlegendentry{\Large sub-optimal $\eta_i$s}
+      \addlegendentry[scale=1.4]{$\sigma^2\lambda_i^{-1}$}
+      \addlegendentry[scale=1.4]{sub-optimal $\eta_i$s}
     \end{axis}

### 2023-05-22 arxiv/ms.tex
With supp material in arxiv lib

@@
   University\\ Tel Aviv, Israel}
-\address{Sackler Faculty of Medicine,
-  Sciences, Tel Aviv University\\ Tel Aviv, Israel}
+%
+\address{Sackler Faculty of Medicine, Tel Aviv University\\ Tel Aviv,
+  Israel}
 
@@
 \begin{abstract}
-  Inversion for parameters of a physical process typically requires
-  taking expensive measurements, and the task of finding an optimal
-  set of measurements is known as the optimal design
-  problem. Surprisingly, measurement locations in optimal designs are
-  sometimes extremely clustered, and since measurement clusterization
-  is nonintuitive, researchers often avoid it by modifying the optimal
-  design problem. We consider a certain flavor of the optimal design
-  problem, based on the Bayesian D-optimality criterion, and suggest
-  an analytically tractable model for D-optimal designs in Bayesian
-  linear inverse problems over Hilbert spaces. We demonstrate that
-  measurement clusterization is a generic property of D-optimal
-  designs, and prove that correlated noise between measurements
-  mitigates clusterization. We also give a full characterization of
-  D-optimal designs, which shows that D-optimal designs uniformly
-  reduce uncertainty in a select subset of prior covariance
-  eigenvectors. Finally, we show how measurement clusterization is a
-  consequence of the characterization mentioned above and the
-  pigeonhole principle.
+  Inversion for parameters of a physical process typically requires taking expensive measurements, and the task of finding an optimal set of measurements is known as the optimal design problem.
+Surprisingly, measurement locations in optimal designs are sometimes extremely clustered, and researchers often avoid measurement clusterization by modifying the optimal design problem.
+We consider a certain flavor of the optimal design problem, based on the Bayesian D-optimality criterion, and suggest an analytically tractable model for D-optimal designs in Bayesian linear inverse problems over Hilbert spaces.
+We demonstrate that measurement clusterization is a generic property of D-optimal designs, and prove that correlated noise between measurements mitigates clusterization.
+We also give a full characterization of D-optimal designs under our model: We prove that D-optimal designs uniformly reduce uncertainty in a select subset of prior covariance eigenvectors.
+Finally, we show how measurement clusterization is a consequence of the characterization mentioned above and the pigeonhole principle.
 \end{abstract}

### 2023-05-22 arxiv/supplementary.tex
With supp material in arxiv lib

--- 2020-10-26 new_paper.tex
@@
+%\documentclass{siamonline220329}
 \documentclass{amsart}
-
+%%\documentclass{article}
 \input{definitions}
 
-\numberwithin{equation}{section}
+\title[Supplementary Material]{Supplementary Material for: Measurement
+  Clusterization in D-optimal Designs for Bayesian Linear Inverse
+  Problems over Hilbert Spaces}
 
 \begin{document}
-
-\title[Sensor Clusterization in D-optimal design in infinite
-  dimensions]{Sensor Clusterization in D-optimal Design in Infinite
-  Dimensional Bayesian Linear Inverse Problems}
-
-\author{Yair Daon}
-\address{Porter School of the Environment and Earth
-  Sciences, Tel Aviv University\\ Tel Aviv, Israel}
-%% \curraddr{Porter School of Environment Studies, Tel Aviv University\\ Tel Aviv, Israel}
-%% \curraddr{}
-\email{yair.daon@gmail.com}
-%% \thanks{}
-
-%    The 2010 edition of the Mathematics Subject Classification is
-%    the current definitive version.
-\subjclass{Primary: 
-  62F15, % Statistics - bayesian inference
-  35R30, % PDE - Inverse problems
-  Secondary:
-  28C20, % Measure and integration - set functions and measures and integrals in
-  %infinite-dimensional spaces (Wiener measure, Gaussian measure, etc.
-}
-%% Primary: 62F15, 35R30; Secondary: 28C20
-\date{\today}
-
-\begin{abstract}
-  We investigate the problem of sensor clusterization in D-optimal
-  experimental design for infinite-dimensional Bayesian linear inverse
-  problems. We suggest an analytically tractable model for such
-  designs and reason how it may lead to sensor clusterization in the
-  case of iid measurement noise. We also show that in the case of
-  spatially correlated measurement error clusterization does not
-  occur. As a part of the analysis we prove a Matrix Determinant Lemma
-  in Hilbert spaces and a lemma for calculating $\frac{\der}{\der t}
-  \log \det (I + X(t))$ for certain operator-valued functions. We also
-  show how to decompose $M = AA^t$ with $A$ constrained to have unit
-  norm columns.
-\end{abstract}
-
 \maketitle
 
-\section{Introduction}\label{section:OED intro}
-Experimental design is an important part of many scientific
-investigations. When considering an inverse problem, one can often
-specify sensor locations (e.g.\ in geophysics and oceanography
-applications), certain wavelengths (e.g.\ in MRI) or wave reflections
-from the ground (e.g.\ searching for oil or using a radar). Whatever
-the allowed set of measurements is, one should select the optimal
-measurements to take, in order to increase accuracy, reduce costs, or
-both.
-
-Designing experiments is usually done by optimizing some \emph{design
-  criterion}. This is true both for frequentists
-\cite{Silvey13,Ucinski05} as well as for Bayesians
-\cite{ChalonerVerdinelli95}. See \cite{ChalonerVerdinelli95} for an
-investigation of the analogy between the two approaches. Although
-there is a plethora of design criteria, we focus on just one of these,
-commonly referred to as \emph{D-optimal design}. It has a simple and
-appealing motivation in the Bayesian context as explained in
-\cite{ChalonerVerdinelli95}: A D-optimal design seeks to maximize the
-expected information gain (KL divergence
-\cite{KullbackLeibler51,CoverThomas12}) between posterior and
-prior. For example, consider a linear model in finite dimensions, with
-Gaussian prior and noise. In this model, maximizing the expected
-information gain amounts to minimizing the determinant of the
-posterior covariance matrix. In a frequentist setting, a D-optimal
-design minimizes the volume of the uncertainty ellipsoid \cite[page
-  16]{Ucinski05}, but this is done for the Fisher information matrix
-and not the posterior covariance. However, \cite{ChalonerVerdinelli95}
-show that the latter is just a regularized version of the former.
-
-The previous discussion is classical for experimental design when
-inference is to take place over a finite (not too large) number of
-parameters. The subject of optimal experimental design for function
-inference in a Bayesian context was pioneered by
-\cite{AlexanderianGloorGhattas14, AlexanderianPetraStadlerEtAl16,
-  AlexanderianPetraStadlerEtAl14}. Similarly to the finite dimensional
-case, it can be shown that a D-optimal design arises naturally for
-linear models when one wishes to maximize the KL divergence between
-posterior and prior. This amounts to minimizing the determinant of the
-posterior covariance operator (understood as a product of its
-eigenvalues). Some difficulties arise in the process, but remedies can
-be found as shown in \cite{AlexanderianGloorGhattas14}.
-
-It seems counter intuitive that when one computes an optimal design
-using the D criterion, the optimization process results in
-measurements that are very similar. For example, if a measurement is
-thought of as measuring some function value at $\x \in \Omega
-\subseteq \R^d, d=1,2,3$ (with added error) then the optimization
-procedure sometimes places sensors in very close proximity to each
-other (as can be seen in figure \ref{fig:clusterization
-  illustration}). Following \cite{Ucinski05}, we refer to this
-phenomenon as \emph{sensor clusterization}.
-
-\subsection{Related Work}
-The phenomenon of sensor clusterization seems to be known in several
-different contexts. In a frequentist and finite-dimensional context,
-\cite{Fedorov96} and \cite[chapter 2.4.3]{Ucinski05} discuss this
-phenomenon and suggest an approach called clusterization-free design.
-In such designs, the user enforces measurement locations to be far
-from each other. One way to do this is by introducing correlated
-errors which, philosophically, accounts for both measurement error and
-model error. Another method considered is imposing distance
-constraints between measurements. A somewhat different approach is
-suggested in \cite[page 49]{Fedorov12}, where close by design
-measurements are merged --- a procedure which obviously does not avoid
-clusterization. The same problem arises in time-series analysis for
-pharmacokinetic experiments. The authors of \cite{Hooker09} suggest
-modeling auto-correlation time in the noise model, which is equivalent
-to the correlated errors mentioned above.
-
-Any of the above mentioned approaches might serve as a remedy and push
-sensors away from each other. Yet, none offers any insight as to why
-clusterization occurs. Also, as better models are employed, model
-error is decreased and the clusterization phenomenon will eventually
-reappear. While these approaches are practical and help us avoid the
-problem, they do not provide insight as to why sensors are clustering.
-
-In the inverse problems community, work is mostly computational and
-less theoretic. Model errors were considered in \cite{Attia20,
-  Koval20}. The former study is focused on inferring of a Quantity of
-Interest (QoI). The focus of the latter is reducing forward solves,
-using randomized linear algebra. Both studies present numerical
-techniques for finding optimal designs when model error is
-present. Both are restricted to linear inverse problems (although in
-the latter the authors use their method on a nonlinear problem by
-taking a Laplace approximation for the posterior). Both find an
-optimal design by first solving a continuous problem for sensor
-weights. Said solution is then sparsified to give a binary
-design. Both studies are successful in the task of Bayesian
-inversion. However, neither of these studies mentions any effect model
-errors can have on sensor clusterization. The current study is mostly
-theoretical and aims to fill the theoretical gap of understanding
-sensor clusterization.
-
-\subsection{Contribution}
-We propose and thoroughly study a relaxed and analytically tractable
-model for understanding D-optimal designs. In this model, D-optimal
-design are solutions of a constrained optimization problem, formulated
-using Lagrange multipliers (section \ref{section:D and grad}). Using
-this model we rigorously show how model error mitigates clusterization
-(section \ref{section:non vanishing}). We show how clusterization
-occurs in a sequential design scenario (section
-\ref{subsec:clusterization sequential}). Then, we consider a
-simultaneous D-optimal design scenario (section
-\ref{subsec:clusterization simultaneous}). We show that designs that
-exhibit clusterization are just as optimal as clusterization free
-designs.
-
-A beautiful mathematical structure arises in D-optimal designs when no
-model error is present. The Lagrange multipliers problems is in fact a
-nonlinear eigenvalue problem for the observations and prior
-covariance. The operator for which eigenvectors and eigenvalues are
-sought is a sum of two operators. The first is the prior
-covariance. The second is an outer product of the observations (see
-\eqref{eq:eigenproblem} for details and exact statement). This
-structure helps us shed light into D-optimal designs. We characterize
-these designs using Theorem \ref{thm:char}. One insight we prove in
-said theorem, is that a D-optimal design reduces uncertainties where
-they are highest first. Other interestig phenomena arise, but they
-require setting notation, and are discussed later.
-
-In the process we generalize several lemmas from linear algebra to
-infinite-dimensional settings. We prove a Matrix Determinant Lemma in
-\ref{lemma:MDL}. We generalize a lemma due to Lax \cite{lax97} for
-calculating $\frac{\der}{\der t} \log \det (I + X(t))$, for an
-operator valued function $X(t)$ in \ref{lemma:lax}. We also prove
-lemmas in linear algebra. One gives a decomposition $M = AA^t$ where
-$A$ has unit norm columns \ref{lemma:free}. Another shows simultaneous
-diagonizability of the prior and outer product in the nonlinear
-eigenvalue problem referenced above \ref{lemma:sim diag}. We also
-provide other tools for understanding D-optimal designs in
-infinite-dimensional Bayesian inverse problems, including lemmas for
-calculating the increase in the design criterion per observation ---
-Lemma \ref{lemma:design increase} and Corollary \ref{cor:zero mod
-  err}.
-
-\subsection{Limitations}\label{subsec:limitations}
-There are two main drawbacks of the study presented here. Our relaxed
-model does not consider any specific set of allowed
-measurements. Rather, we take measurements in the unit ball in some
-Hilbert space. This allows considerably less restrictive observations
-than any real-life problem does. The second drawback is that we do not
-show rigorously that clusterization necessarily occurs in a
-simultaneous design. We only show that it is as reasonable as no
-clusterization.
-
-\subsection{An Example of Clusterization}\label{subsec:example}
-\begin{figure}
-  \begin{tikzpicture}[thick, scale=1.3, every node/.style={scale=0.99}]
-    \begin{axis}
-      [
-      title={Posterior Pointwise Standard Deviations and D-Optimal Sensor Locations},  
-      xmin = 0,
-      xmax = 3.14,
-      xlabel = {$x$},
-      ylabel = posterior std,
-      ymin   = 0,
-      %compat = 1.3,
-      % ymax   = 130,
-      % ytick = \empty,
-      legend cell align=left,
-      % legend style={at={(0.45,0.2)}}
-      legend pos= outer north east 
-      ]
-      % \draw[black!30!white, thin] (50,0) -- (50,130);
-      % 
-      %% \addplot [thin, black, mark=none] table{stdv-heat-sens1-var1.txt};
-      %% \addlegendentry{1 sensors};
-
-      %% \addplot [thin, blue, mark=none] table{stdv-heat-sens2-var1.txt};
-      %% \addlegendentry{2 sensors};
-
-      %% \addplot [thin, red, mark=none] table{stdv-heat-sens3-var1.txt};
-      %% \addlegendentry{3 sensors};
-
-      \addplot [thin, green, mark=none] table{stdv-heat-sens4-var1.txt};
-      \addlegendentry{4 sensors};
-
-      \addplot [thin, purple, mark=none] table{stdv-heat-sens5-var1.txt};
-      \addlegendentry{5 sensors};
-
-      \addplot [thin, cyan, mark=none] table{stdv-heat-sens6-var1.txt};
-      \addlegendentry{6 sensors};
-
-      %% \addplot [black,  only marks, mark=x, mark size=1.5] 
-      %% table{locs-heat-sens1-var1.txt}; 
-      %% \addplot [blue,   only marks, mark=x, mark size=1.5]
-      %% table{locs-heat-sens2-var1.txt}; 
-      %% \addplot [red,    only marks, mark=x, mark size=1.5]
-      %% table{locs-heat-sens3-var1.txt};
-      \addplot [green,  only marks, mark=*, mark size=1.5] 
-      table{locs-heat-sens4-var1.txt}; 
-      \addplot [purple, only marks, mark=*, mark size=1.5] 
-      table{locs-heat-sens5-var1.txt}; 
-      \addplot [cyan,   only marks, mark=*, mark size=1.5] 
-      table{locs-heat-sens6-var1.txt};
-
-    \end{axis}
-  \end{tikzpicture}
-  \caption{The clusterization effect for the 1D heat equation
-    described in section \ref{subsec:example}. Posterior pointwise
-    standard deviations are plotted over the domain $[0, \pi]$, for
-    varying numbers of sensors (lines). Sensor locations (circles)
-    were chosen to minimize (an expression analogous to) the
-    determinant of the posterior covariance. The clusterization effect
-    can be clearly seen for six sensors. Only four measurement
-    locations are used --- two pairs of sensors are so close they are
-    indistinguishable.}
-  \label{fig:clusterization illustration}
-\end{figure}
-In section \ref{section:prelim} we present a more abstract and general
-formulation of the problem, but for the purpose of illustration, we
-present the problem via a toy model --- the 1D heat equation in
-$[0,\pi]$ with a homogeneous Dirichlet boundary conditions.
-
-The 1D heat equation is:
+\section{The inverse problem of the 1D heat equation}\label{section:example}
+The mathematical foundations of inversion in function spaces can be
+found in \cite{Stuart10}. Our goal is to infer the initial condition of
+the one-dimensional heat equation with a homogeneous Dirichlet
+boundary condition:
 \begin{subequations}\label{eq:heat equation}
   \begin{alignat}{2}
-    u_t &= \Delta u &&\qquad \text{in } [0,\pi] \times [0,\infty),\\
-      u &= 0 &&\qquad \text{on } \{0, \pi\} \times [0,\infty),\\
-        u &= u_0 &&\qquad \text{on }[0,\pi] \times \{0\}.
+    u_t &= \Delta u &&\qquad \text{in } [0,1] \times [0,\infty),\\
+      u &= 0 &&\qquad \text{on } \{0, 1\} \times [0,\infty),\\
+        u &= u_0 &&\qquad \text{on }[0,1] \times \{0\}.
   \end{alignat}
@@
 
-We would like to infer the initial condition $u_0$. For that purpose,
-we measure $u$ at some set of locations $\x_j \in [0,\pi], j=1,
-\dots,m$ and a final time $T > 0$. We assume centered Gaussian
-measurement error, so we can observe $v(\x_j,T) = u(\x_j,T) +
-\eps(\x_j)$ with $\eps(\x_j) \sim \normal(0, \sigma^2), \sigma > 0$
-iid. We model the initial condition as $u_0 \sim \normal(0,\prcov)$,
-for $\prcov = (-\Delta)^{-1}$ with a homogeneous Dirichlet boundary
-condition. It is well known \cite{Tarantola05} that for linear
-problems, with Gaussian prior and error, the posterior is also
-Gaussian with a covariance that does not depend on the observed
-data. The posterior covariance $\postcov$ is known to have a closed
-form formula, even in infinite dimensions\cite{Stuart10}. We denote by
-$\fwd$ the dynamics operator, so that $u( \cdot,T) = \fwd u_0$, and
-the observation operator $\obs$ so that $u(\x_j,T) = (\obs u)_j,
-j=1,\dots,m$. The posterior covariance is known and depends only on
-$\prcov, \fwd, \obs$ and $\sigma^2$ (see section \ref{section:prelim}
-and \eqref{eq:postcov} specifically).
+Model the prior for the initial condition as $u_0 \sim
+\normal(0,\prcov)$, for $\prcov = (-\Delta)^{-1}$ with a homogeneous
+Dirichlet boundary condition. Let the final time $T = 0.03$ and denote
+$\fwd$ the forward operator, so that $u(\cdot, T) = \fwd u_0$. Let
+$\obs$ a measurement operator, so that $u(\x_j,T) = (\obs u)_j,
+j=1,\dots,m$, where $m$ is the number of allowed measurements. Assume
+iid centered Gaussian observation error, so observations are given by
+$u(\x_j,T) + \eps(\x_j)$ with $\eps(\x_j) \sim \normal(0, \sigma^2)$
+iid, $\sigma = 0.05$. Measurement locations $\x_j,j=1,\dots,m$ are
+chosen according to the D-optimality criterion of
+\cite{AlexanderianGloorGhattas14}.
 
-We consider generalization of the information-theoretic design
-criterion presented in the introduction to infinite dimensions
-(section \ref{subsec:D optimal design} below). We choose
-$\x_j,j=1,\dots,m$ to minimize (an expression analogous to) the
-determinant of the posterior covariance operator. We will see later
-how this corresponds to maximizing expected information gain.
+Implementation of this inverse problem can be found in the
+repository \url{https://github.com/yairdaon/OED}, with an optimization
+routine for finding D-optimal designs. Unfortunately, I was not able
+to install PyOED \cite{attia2023pyoed}, nor was I able to utilize
+hIPPYlib \cite{VillaPetraGhattas16, VillaPetraGhattas18,
+VillaPetraGhattas21}, hence the inverse problem is implemented from
+scratch.
 
-The clusterization effect is illustrated in figure
-\ref{fig:clusterization illustration}. Posterior pointwise standard
-deviations are plotted over the domain $[0, \pi]$. Since the posterior
-covariance does not depend on data, the plot has no reference to
-actual data observed. The posterior covariance does, however, depend
-on location of the measurement taken. In figure
-\ref{fig:clusterization illustration}, measurement locations are
-marked by circles. These were chosen to minimize (an expression
-analogous to) the determinant of the posterior covariance. The
-clusterization effect can be clearly seen for six sensors. It looks
-like only four measurements were taken. The reason is that two pairs
-of sensors are so close they are indistinguishable.
+\section{Generalizations of known lemmas}
 
-\section{Preliminaries and Notation}\label{section:prelim}
-
-\subsection{Overview}
-In this section we define notations that will be used throughout the
-paper. The theoretical foundations for inverse problems over function spaces
-can be found in \cite{Stuart10}.
-
-\subsection{Bayesian Linear Inverse Problems}\label{subsec:abstract OED}
-Let $\hilp$ and $\hilo$ be separable Hilbert spaces (the subscripts p
-and o are for ``parameter'' and ``observation'', respectfully). We
-denote $\hilo^*$ the Hilbert space of all linear functionals on
-$\hilo$. Consider $\fwd: \hilp \to \hilo$, a linear operator (the
-``forward operator''). We are interested in forward operators that are
-strongly smoothing (have fast decaying modes --- the heat operator
-from section \ref{subsec:example} is a prime example). Our goal is to
-infer $\param \in \hilp$ --- some parameter of the dynamics --- given
-noisy observations of $\fwd \param$. We take a Gaussian prior $\pr$
-for $\param$. Hence, $\param \sim \pr = \normal(\prmean ,\prcov)$
-with some appropriate covariance operator $\prcov$ \cite{Stuart10}. It
-is important to note that $\fwd \prcov \fwd^*$ is the prior covariance
-in $\hilo$ \cite{Stuart10}. As such, it is assumed invertible --- an
-assumption we rely on below. Throughout this paper, we denote $m$ the
-number of measurements we are allowed to take. Observations
-(measurements) are taken via the linear observation operator $\obs \in
-( \hilo^* )^m$. In an analogy with linear algebra (where row vectors
-are thought of as linear functionals), we think of $\obs$ as having
-``rows'' $\meas_j, j=1,\dots,m$:
-\begin{equation}\label{eq:O}
-  \obs = (\meas_1,\dots, \meas_m)^t, \meas_j \in \hilo^*, j = 1,\dots,m.
-\end{equation}
-This way, for $u \in \hilo$ we have $\obs u = (\meas_1(u), \dots,
-\meas_m(u) )^t \in \R^m$.
-%% but we may drop the parentheses and write $\meas_j u = \meas_j(u)$.
-%% A few observations regarding $\obs$ are in order.
-%% First, it is good to keep in mind that $( \hilo^* )^m$ is a Hilbert
-%% space with norm $\| \obs \| = \sum_{j=1}^m \|\meas_j\|$.
-For $u \in \hilo$ and $v\in \R^m$:
-\begin{align*}
-  \big (\obs^*v \big ) (u) &= \langle v, \obs u \rangle_{\R^m} = \sum_{j=1}^m  v_j \meas_j(u)
-  = v^t \left ( \obs u \right ) = (v^t \obs) (u),
-\end{align*}
-and thus:
-\begin{align}\label{eq:obs*}
-  \obs^*v &= \sum_{j=1}^m v_j \meas_j = v^t \obs.
-\end{align}
-
-%% \begin{observation}
-%%   It is best to think of $\meas_j,j=1,\dots,m$ as row vectors and of
-%%   $\obs$ as a matrix with rows $\meas_j$. 
-%% \end{observation}
-
-Each measurement is a linear functional, chosen from some subset of
-$\hilo^{*}$. Data is acquired via noisy measurements
-\begin{align*}
-  \data := \obs (\fwd \param + \eps') + \eps = \obs \fwd \param + \obs \eps' + \eps,
-\end{align*}
-with $\data \in \R^m$ and $\eps, \eps'$ defined next. We consider two
-types of noise / error terms. First, there is spatially correlated
-model error $\eps' \sim \normal(0,\modcov)$, modeled as a centered
-Gaussian measure on $\hilo$ with covariance operator $\modcov$. Then
-there is measurement / independent error $\eps \sim \normal(0,
-\sigma^2 I_m)$, with $I_m$ the $m \times m$ identity. Both error terms
-and the prior are assumed independent of each other.
-
-Usually, this is written using $\tmp := \obs \fwd$ and $\tmp^*$. Since
-we need to separate the observation operator $\obs$ from the forward
-operator $\fwd$ we write the explicit expression
-\eqref{eq:postcov}. We now turn to understand the terms present in the
-above expression.
-
-\subsubsection{Dynamics and Observation Operators}\label{subsec:dynamics}
-One can \textit{think} of $(\obs \fwd \param)_j$ as pointwise
-evaluations of a continuous function\footnote{Of course, this is
-  impossible if, e.g., $\hilo = L^2(\Omega)$ for some $\Omega$.}. Such
-measurements can be represented using the linear functionals
-$\delta_{\x_j} \in C(\Omega)^*$. In a different setting, we may be
-able to measure other quantities, e.g.\ some Fourier coefficients as
-is the case in MRI applications. Either way, the rows of $\obs$ are
-linear functionals and we refer to these as {\it measurement
-  vector}s. Note that we cannot choose every measurement vector we
-want. For example, we may be restricted only to pointwise evaluations
-of $\fwd \param$ by the sensors at our disposal, in which case
-measuring the mean $\ell(u) = \int_{\Omega}u$ (or any non-atomic
-measure) of a function is not possible.
-
-\subsubsection{Error Terms}
-Considering the sum of the error terms it is easy to see that
-$\bar{\eps} := \obs \eps' + \eps \in \R^m$. It is a centered Gaussian
-random vector and its covariance matrix is
-\begin{align}\label{eq:Sigma}
-  \begin{split}
-    \Sigma(\obs) :&= \mathbb{E}[ (\obs \eps' + \eps)  (\obs \eps' + \eps)^t ] 
-    % 
-    % 
-    = \obs \modcov \obs^* + \sigma^2I , 
-  \end{split}
-\end{align}
-where
-\begin{equation}\label{eq:modcov explained}
-  [\obs \modcov \obs^*]_{ij} = e_i^t \obs \modcov \obs^* e_j = \meas_i (\modcov \meas_j),
-\end{equation}
-with $e_j \in \R^m$ the $j$th standard basis vector. The first
-equality in \eqref{eq:modcov explained} holds by definition and the
-second by \eqref{eq:obs*}. The explicit dependence on $\obs$ will be
-mostly dropped for notational convenience, so $\Sigma(\obs) =
-\Sigma$. Thus, for a fixed $\obs$ (i.e.\ a fixed set of measurements)
-we can equivalently write $\data = \obs \fwd \param + \bar{\eps}$ with
-$\bar{\eps} \sim \normal(0,\Sigma)$. Taking $\modcov = 0$ is common
-practice \cite{Tarantola05,KaipioSomersalo05,Vogel02}. This reduces to
-the case where we take iid observations. Then $\Sigma = \sigma^2I$ is
-simply a scalar matrix that does not depend on $\obs$. Note that
-taking an error model with a non-scalar covariance as we do here
-allows us to consider model error (modeled by $\modcov$) as well as
-measurement error (modeled by $\sigma^2$). For example, say we believe
-our forward model does not capture some small scale phenomenon.  Then
-we may express this belief by saying $\tru = \fwd + \err$, with $\fwd$
-depending on $\param$ and $\err$ depending on $\sspar$, and $\param
-\perp \sspar$. We do not know much about this effect but it is
-reasonable to assume it changes continuously in our domain. We (may
-choose to) model it as $\normal (0, \modcov)$ and take $\modcov$ to
-reflect the spatial (or other) variability we imagine $\err \sspar$
-has. Such small scale phenomenon can arise as a modeling issue, where
-we might not model the system in its entirety. It can also arise from
-a numerical source, where our discretization of the system is not fine
-enough to capture all small scale phenomena. In section
-\ref{section:non vanishing}, we will see that assuming some
-correlation in the error mitigates the clusterization phenomenon, as
-reported in the literature \cite{Ucinski05}.
-
-Finally, it is useful to record a known result regarding the posterior
-covariance operator of our inverse problem, namely
-\begin{align}\label{eq:postcov}
-  \postcov = (\prcov^{-1} + \fwd^* \obs^* \Sigma^{-1} \obs \fwd
-  )^{-1}.
-\end{align}
-
-\subsection{D-Optimal Designs in Infinite Dimensions}\label{subsec:D optimal design} 
-A D-optimal design maximizes expected KL divergence between posterior
-and prior. It is useful to recall the definition of KL divergence for
-an arbitrary prior measure:
-$$
-KL(\post||\pr) = \int \log \frac{\der \post}{\der \pr}(\param) \der \post(\param).
-$$
-
-The meaning of D-optimal design in infinite-dimensional Hilbert spaces
-was investigated in \cite{AlexanderianGloorGhattas14}. There, the
-authors make assumptions that amount to $\Sigma=I$ (implied by
-$\modcov = 0,\sigma^2=1$), but we choose not to take these
-simplification here. This is because $\Sigma$ can determine ``how
-much'' clusterization we see. As stated
-\cite[pp. 681]{AlexanderianGloorGhattas14}, the results hold for more
-general covariance matrices. The conclusion is that in
-infinite-dimensions, a D-optimal design is well-defined as maximizing
-the expected KL divergence between posterior and prior. The main
-result is summarized, using our notation, in the following theorem:
-\begin{theorem}[Slightly modified Theorem 1 from \cite{AlexanderianGloorGhattas14}]
-  Let $\pr = \normal(\prmean,\prcov)$ be a Gaussian prior on $\hilp$
-  and let $\post = \normal(\postmean,\postcov)$ the posterior measure
-  on $\hilo$ for the Bayesian linear inverse problem $\data = \obs
-  \fwd\param + \obs \eps' + \eps$ discussed above. Then
-  \begin{align}\label{eq:objective}
-    \begin{split}
-      \tar( \obs) :&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
-      % 
-      % 
-      % 
-      &= \frac12 \log \det 
-      ( I + \prcov^{1/2}  \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}).
-    \end{split}
-  \end{align}
-\end{theorem}
-\begin{definition}\label{def:d optimality}
-  A design $\obs^{\star}$ is said to be D-optimal if $\obs^{\star} = \argmax_{\obs} \tar(\obs)$.
-\end{definition}
-
-The intuition behind definition \ref{def:d optimality} is
-straightforward. From \eqref{eq:postcov}:
-\begin{align*}
-  \begin{split}
-    \det ( I + \prcov^{1/2}  \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}) &= \det \Big( \prcov ( \prcov^{-1} + \fwd ^* \obs^* \Sigma^{-1} \obs \fwd) \Big )\\
-    &= \det \prcov \det \postcov^{-1}.
-  \end{split}
-\end{align*}
-Since $\prcov$ is constant, a D-optimal design minimizes the posterior
-covariance determinant, analogously to the finite-dimensional case.
-
-\subsection{Sequential vs Simultaneous Optimization}\label{subsec:seq vs sim}
-From defintion \ref{def:d optimality} we wish to characterize solution(s) of the
-following optimization problem for $\tar$. %%: (\hilo^*)^m \to \R$:
-\begin{align}\label{eq:optimization}
-  \obs^{\star} := \argmax_{\obs} \tar( \obs ) 
-  = \argmax_{\obs} \frac12 \log \det 
-  (I + \prcov^{1/2} \fwd^*\obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}),
-\end{align}
-where $\obs$ is constrained to some allowed set of measurements (we
-may \textit{think} of these as pointwise evaluations, even though
-these might not be in $\hilo^*$). We call this problem ``simultaneous
-optimization'', since all measurements are decided on simulatneously.
-
-For computational reasons, one may prefer to find the best
-measurements in a sequential manner. Denote
-\begin{equation}\label{eq:def obs_k}
-  \obs_k := (\meas_1,\dots, \meas_k)^t,  k\leq m.
-\end{equation}
-Sequential optimal design proceeds as follows. Find $\meas_1$ by
-maximizing $\tar(\obs_1)$. Then, keeping $\meas_1$ fixed --- find
-$\meas_2$ as the maximizer of $\tar(\obs_2)$. Then, find $\meas_3$ by
-keeping $\meas_1,\meas_2$ fixed and taking $\meas_3$ the maximizer of
-$\tar(\obs_3)$. Continue in this fashion until $\obs_m = \obs$ is
-found, where $m$ is the number of available measurements. It is
-important to notice that this scheme does not require actually
-observing data --- in \eqref{eq:objective} data is averaged out.
-
-The analysis in this paper is conducted for the general simultaneous
-optimization case. The sequential optimization case is dealt with in
-section \ref{subsec:seq vs sim}. It is important to note, however that
-all conclusions we arrive at for the simultaneous case easily
-specialize to the sequential case by considering the posterior as the
-next sequential step's prior.
-
-\section{The Constrained Optimization Problem of D-Optimal Design}\label{section:D and grad}
-
-\subsection{Overview}
-We seek a formulation of the D-optimal design problem
-\eqref{eq:optimization} using Lagrange multipliers. In section
-\ref{section:objective} we find the gradient of $\tar$ and record it
-in \eqref{eq:tar grad}. In section \ref{subsec:unit length} we suggest
-using unit-norm constraints on rows of $\obs$ so that the optimization
-problem in \eqref{eq:optimization} is analytically tractable. We find
-gradients for the new constraints and record them in \eqref{eq:grad
-  constraints}. Finally, in section \ref{subsec:necessary} we use
-these to formulate the D-optimal design problem as a Lagrange
-multipliers problem in equation \eqref{eq:conditions}.
-
-\subsection{The Objective and its Gradient}\label{section:objective}
-In order to use Lagrange multipliers, we need to find the gradient of
-$\tar$. This section is mostly technical and devoted to calculating
-said gradient. The result is recorded in \eqref{eq:tar grad}.
-
-We start by calculating the first variation. Some calculations are
-delegated to Lemma \ref{lemma:aux calc} in the appendix. We use
-the following lemma:
-\begin{restatable*}[Generalized from \cite{Lax97}]{lemma}{lax}\label{lemma:lax}
+The following lemma is generalized from \cite[Chapter 9, Theorem 4,
+p. 127]{Lax07}
+\begin{lemma}\label{lemma:lax}
   Let $Y(t)$ be a differentiable operator-valued function. Assume 
@@
   \end{equation*}
-\end{restatable*}
+\end{lemma}
 
-The proof of \ref{lemma:lax} is delegated to the appendix. Let
-$T(\obs) := \obs^*\Sigma(\obs)\obs$. The first variation of $\tar
-(\obs)$ in the direction $V$ is:
-\begin{align}\label{eq:tar var}
-  \begin{split}
-    \delta \tar(\obs) V 
-    :&= \frac{\der}{\der\tau} \Big |_{\tau=0} \tar(\obs + \tau V)\  \text{ (by definition of variation)}\\
-    % 
-    % 
-    % 
-    &= \frac12 \frac{\der}{\der \tau} \Big |_{\tau=0} \log \det 
-    (I + \prcov^{1/2} \fwd^* T(\obs+\tau V)\fwd \prcov^{1/2} ) \text{ (by definition \eqref{eq:objective})}\\
-    % 
-    % 
-    % 
-    &= \frac12 \tr{( I + \prcov^{1/2} \fwd^* \obs^* \Sigma^{-1}
-    \obs\fwd \prcov^{1/2} )^{-1}
-    \frac{\der}{\der \tau} \Big |_{\tau=0}
-    \prcov^{1/2} \fwd^* T(\obs+\tau V) \fwd \prcov^{1/2}}\ \text{ (by \ref{lemma:lax})}\\
-    % 
-    % 
-    % 
-    &= \frac12 \ttr\Big \{ \postcov \fwd^* (V^* \Sigma^{-1} \obs 
-    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
-    &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs 
-    + \obs^* \Sigma^{-1} V ) \fwd \Big \}  \text{ (by \ref{lemma:aux calc})}\\
-    %
-    %
-    %
-    &= \tr{\postcov \fwd^* ( \obs^* \Sigma^{-1} V -
-    \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs ) \fwd} \text{
-        (cyclic property of trace)} \\
-    %
-    %
-    % 
-    &= \tr{\postcov \fwd^* \obs^* \Sigma^{-1} V 
-    ( I - \modcov \obs^* \Sigma^{-1}\obs ) \fwd} \\
-    % 
-    %
-    %
-    &= \tr{V ( I - \modcov \obs^* \Sigma^{-1}\obs )
-    \fwd \postcov \fwd^* \obs^* \Sigma^{-1}}.
-  \end{split}
-\end{align} 
-Denote
-\begin{align}\label{eq:tar grad}
-  \nabla \tar(\obs) &:= (I - \modcov \obs^* \Sigma^{-1} \obs) \fwd
-  \postcov \fwd^* \obs^*\Sigma^{-1},
-\end{align}
-the gradient of $\tar (\obs )$ and we now justify this definition.
-Since the trace of $A := V \nabla \tar(\obs) \in \R^{m \times m}$ is
-just $\ttr A = \sum_{j=1}^m e_j^t A e_j$ (with $e_j$ the $j$th standard
-basis vector), we see that
-\begin{align*}
-  \delta \tar(\obs)V = \tr{V \nabla \tar(\obs)} = \sum_{j=1}^m
-  V_j(\nabla \tar(\obs)_j),
-\end{align*}
-with $V_j \in \hilo^*$ and $\nabla \tar(\obs)_j \in \hilo^{**} =
-\hilo, j=1,\dots,m$. Thus, $\nabla \tar( \obs ) \in \hilo^m$ is indeed
-the correct gradient and the notation \eqref{eq:tar grad} is
-justified. Note that by the comment following \eqref{eq:obs*}, we view
-$V$ (defined on the same space as $\obs$) as a column vector. Then the
-gradient $\nabla \tar(\obs)$ should be viewed as a row vector, as the
-product $V \nabla \tar$ is in $\R^{m \times m}$. This will prove
-important in section \ref{subsec:necessary}.
-
-\subsection{Unit Length Constraints and their Gradients}\label{subsec:unit length}
-In this section we suggest relaxed constraints in
-\eqref{eq:constraints}. We then find their gradient in \eqref{eq:grad
-  constraints}. This is for using them in a Lagrange multipliers
-problem.
-
-As mentioned before, we cannot choose any $\obs$ when maximizing
-$\tar(\obs)$. Recall $\obs = (\meas_1,\dots,\meas_m)^t \in
-(\hilo^*)^m$ and each $\meas_j,j=1,\dots,m$ must be chosen from some
-allowed set of functionals in $\hilo^*$. This set differs based on the
-kind of sensors we have at our disposal and the properties of $\hilo$,
-as discussed in section \ref{subsec:dynamics}. The following
-proposition gives better understanding of the constraints.
-\begin{restatable*}{proposition}{biggerbetter}\label{prop:bigger better}
-  Let $\obs = (\meas_1,\dots,\meas_m)^t$, $j \in \{1,\dots,m\}$, $\sigma^2
-  > 0$ and $\lambda > 1$. Then $\tar(\obs)$ increases if we use
-  $\lambda \meas_j$ in $\obs$ instead of $\meas_j$.
-\end{restatable*}
-Proof of Proposition \ref{prop:bigger better} is delegated to the
-appendix but the idea behind it is simple: making rows of $\obs$ large
-is equivalent to making $\sigma^2$ small. %%  Viewed another way, it
-%% increases the Signal to Noise Ratio (SNR) --- which is always
-%% desirable when taking any measurement prone to error.
-This can be understood easily from the formulation of the inverse
-problem $\data = \obs \fwd \param + \obs \eps' + \eps$, where $\eps'$
-can be taken as model error and $\eps$ is iid measurement error.
-
-We now suggest a set of constraints that will make the model
-analytically tractable. By Proposition \ref{prop:bigger better}, we
-cannot just take all $\hilo^*$ to be the allowed set of measurement
-vectors, since this would effectively eliminate measurement error
-entirely. We note that for point evaluations, the norm of the
-measurement is always one:
-\begin{align*}
-  \| \delta_{\x} \| = \sup_{0 \neq u \in C(\Omega)} \frac{
-    |\int_{\Omega}u(\y) \delta_{\x}(\y) \der \y| 
-  }{
-    \sup|u|}
-  = \sup_{0 \neq u \in C(\Omega)} \frac{|u(\x)|}{ \sup|u|} = 1,
-  \forall \x \in \Omega.
-\end{align*}
-Following this, we also restrict the norm of our measurement vectors
-to be one. The unit norm constraints can be written using
-\eqref{eq:obs*} as a series of $m$ equality constraints (one for each
-measurement) on $\obs$ as
-\begin{align}\label{eq:constraints}
-  \phi_j(\obs) :=\frac12 \| \obs^* e_j\|_{\hilp}^2 - \frac12 = 0,j=1,\dots,m.
-\end{align}
-The first variations are:
-\begin{align*}
-  \delta \phi_j(\obs)V  
-  &= \frac12\lim_{\tau \to 0}\tau^{-1}
-  ( \|(\obs + \tau V)^*e_j \|_{\bar{\hilp}}^2 - \|\obs ^*e_j \|_{\bar{\hilp}}^2  ) \\
-  %
-  %
-  %
-  &= \frac12\lim_{\tau \to 0}\tau^{-1}
-  ( \langle (\obs + \tau V)^*e_j, (\obs + \tau V)^*e_j \rangle_{\bar{\hilp}} - 
-  \langle \obs^*e_j, \obs^*e_j \rangle_{\bar{\hilp}} ) \\
-  % 
-  % 
-  %
-  &= \frac12\lim_{\tau \to 0}\tau^{-1}
-  (2\tau \langle \obs^*e_j,V^*e_j \rangle_{\bar{\hilp}} 
-  +\tau^2 \langle V^*e_j, V^*e_j \rangle_{\bar{\hilp}} ) \\
-  %
-  %
-  % 
-  &= \langle \obs^*e_j,V^*e_j \rangle_{\bar{\hilp}} \\
-  %
-  %
-  % 
-  &= \langle V \obs^*e_j,e_j \rangle_{\R^m} \\
-  %
-  %
-  %
-  &= e_j^t V \obs^* e_j \\
-  % 
-  %
-  %
-  &= \tr{V \obs^* e_je_j^t}.
-\end{align*}
-Using the same arguments we used to justify the definition of
-\eqref{eq:tar grad}, we conclude that
-\begin{align}\label{eq:grad constraints}
-\nabla \phi_j(\obs) = \obs^* e_j e_j^t = \meas_j e_j^t , j=1,\dots,m,
-\end{align}
-with $\nabla \phi_j(\obs) \in \hilo^m$ and where the
-last equality follows since $\obs^*e_j = \meas_j$ by \eqref{eq:obs*}. As
-we noted at the end of section \ref{section:objective}, the
-gradient $\nabla \phi_j(\obs)$ is a row vector.
-
-\subsection{Necessary Conditions for Optimal Design}\label{subsec:necessary}
-Necessary conditions for D-optimality are found using Lagrange
-multipliers:
-\begin{align}
-  &\nabla \tar(\obs) = \sum_{j=1}^m \xi_j \nabla \phi_j (\obs)
-  \label{eq:Lagrange mult1} \\
-    &\phi_j(\obs) = 0, j = 1,\dots,m. \label{eq:Lagrange mult2}
-\end{align}
-
-Recall where the objects involved are defined: the observation operator,
-$\obs \in (\hilo^*)^m$. The objective $\tar:
-(\hilo^*)^m \to \R$ and its gradient is $\nabla \tar(\obs) \in
-\hilo^m$. The constraints are defined similarly --- $\phi_j: (\hilo^*)^m
-\to \R$ and $\nabla \phi_j(\obs) \in \hilo^m$.
-
-Focusing on \eqref{eq:Lagrange mult1} and using the gradients of the
-objective and constraints \eqref{eq:tar grad} and \eqref{eq:grad
-  constraints} we have:
-\begin{equation*}
-  (I - \modcov \obs^* \Sigma^{-1} \obs) \fwd \postcov \fwd^* \obs^*\Sigma^{-1}
-  = \sum_{j=1}^m \xi_j \obs^* e_je_j^t = (\xi_1 \meas_1,\dots,\xi_m \meas_m).
-\end{equation*}
-If we let $\Xi = \diag(\xi_j)$, then with some abuse of
-notation, we can write this more compactly as:
-\begin{equation}\label{eq:conditions}
-  ( I - \modcov \obs^* \Sigma^{-1} \obs) \fwd \postcov \fwd^* \obs^*  \Sigma^{-1}
-  = \obs^* \Xi.
-\end{equation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SECTION Analysis of Optimal Designs --- Vanishing Model Error
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Analysis of D-Optimal Designs --- Vanishing Model Error}\label{section:vanishing}
-\subsection{Overview}
-In this section we analyze D-optimal designs when no model error is
-present. We discuss sensor clusterization in section
-\ref{section:clusterization}. We prepare the ground by first deriving
-a nonlinear eigenvalue problem in $\obs$ for D-optimal designs in
-section \ref{subsec:eigenproblem}. This nonlienar eigenvalue problem
-is recorded in \eqref{eq:mod conditions}. Next, we characterize
-D-optimal designs as ones that make all eigenvalues equal in the above
-mentioned nonlinear eigenvalue problem. The derivation is presented in
-section \ref{subsec:characterization}.
-
-\subsection{The Nonlinear Eigenvalue Problem}\label{subsec:eigenproblem}
-$\modcov = 0$ implies $\Sigma= \sigma^2I$. The necessary first-order
-condition for D-optimality \eqref{eq:conditions} becomes
-\begin{equation}\label{eq:eigenproblem}
-  \sigma^{-2}\fwd \postcov \fwd^* \obs^* = \obs^* \Xi, 
-\end{equation}
-with $\Xi$ diagonal. This looks like an eigenvalue problem for the
-self-adjoint operator $\sigma^{-2}\fwd \postcov \fwd^*$. The rows of
-$\obs$, namely $\meas_j,j=1,\dots, m$, are the eigenvectors. Note,
-however, that $\postcov$ depends on $\obs$. This makes
-\eqref{eq:eigenproblem} a nonlinear eigenvalue problem.
-
-Below, the expression for the posterior covariance \eqref{eq:postcov}
-justifies the second equality. Some calculations delegated to Lemma
-\ref{lemma:twice woodbury} in the appendix justify the third (trivial
-if $\fwd$ is invertible):
-\begin{align}\label{eq:mod conditions}
-  \begin{split}
-  \obs^* \Xi &= \sigma^{-2}\fwd \postcov \fwd^* \obs^*  \\
-  %
-  %
-  %
-  &= \sigma^{-2} \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* \obs^* \\
-  %
-  %
-  %
-  &= \sigma^{-2} \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1} \obs^*
-  \end{split}
-\end{align}  
-The term $(\fwd \prcov \fwd^*)^{-1}$ is the prior precision in $\hilo$
-(and is invertible, as mentioned in section
-\ref{section:prelim}). $(\fwd\prcov\fwd^*)^{-1} + \sigma^{-2} \obs^*
-\obs$ is the posterior precision in $\hilo$.
-
-\subsection{Characterizing D-Optimal Designs}\label{subsec:characterization}
-Denote eigenvalues of $\fwd \prcov \fwd^*$ by
-$\{\lambda_i\}_{i=1}^{\infty}$. These are all positive. Denote
-corresponding eigenvectors $\{\ev_i\}_{i=1}^{\infty}$. Lemma
-\ref{lemma:sim diag} (proved in the appendix) is crucial for the rest
-of the analysis:
-
-\begin{restatable*}[Simultaneous diagonizability for the nonlinear eigenvalue problem]{lemma}{simdiag}\label{lemma:sim diag}
-  Let $\hil$ separable Hilbert space, $C:\hil \to \hil$ self-adjoint
-  and $\func_1,\dots,\func_m \in \hil$. Denote $\func^*$ the element
-  $\func$ acting as a linear functional. If
-  \begin{equation*}
-   (C + \sum_{j=1}^m \func_j\func_j^*) \func_l = \xi_l \func_l, l = 1,\dots,m
-  \end{equation*}
-  then $C$ and $\sum_{j=1}^m \func_j \func_j^*$ are simultaneously
-  diagonalizable.
-\end{restatable*}
-
-\begin{corollary}[Eigenvectors of a D-optimal design]\label{cor:same ev}
-  $\obs^*\obs$ has the same eigenvectors as $\fwd \prcov \fwd^*$.
-  %namely $\{\ev_i\}_{i=1}^{\infty}$.
-\end{corollary}
-\begin{proof}
-In \eqref{eq:mod conditions}, take $\func_j^{*} = \meas_j$ and $C :=
-(\fwd \prcov \fwd^*)^{-1}$ and use Lemma \ref{lemma:sim diag}.
-\end{proof}
-At most $m$ of the eigenvectors of $\obs^*\obs$ have a positive
-eigenvalue, because $k := \rank \obs^*\obs \leq m$. Denote the
-corresponding non-zero eigenvalues of $\obs^*\obs$ by
-$\{\eta_i\}_{i=1}^{k}$ and let $\eta_i = 0$ for $i \geq k+1$. No
-assumption is made (yet) regarding the magnitudes of the corresponding
-$\{\lambda_i\}_{i=1}^{\infty}$. They are to be thought of as fixed but
-unknown. Using Sylvester's Determinant Theorem (the standard proof
-works for Hilbert spaces as well) we get:
-
-\begin{align}\label{eq:pre manipulations}
-  \begin{split}
-    \det \left (I + \sigma^{-2} \prcov^{1/2} \fwd ^* \obs^*
-    \obs \fwd \prcov^{1/2}\right ) &=
-    %
-    \det \left (I + \sigma^{-2} \obs^*
-    \obs \fwd \prcov\fwd^* \right ) \\
-    %
-    %
-    %
-    &= \det \left \{ \left ( (\fwd \prcov \fwd^*)^{-1} +
-    \sigma^{-2}\obs^*\obs \right ) (\fwd \prcov \fwd^*)\right \}\\
-  \end{split}
-\end{align}
-
-By definition of the objective \eqref{eq:objective} and using
-\eqref{eq:pre manipulations}:
-\begin{align}\label{eq:mod objective}
-  \begin{split}
-    \tar ( \obs ) &=\frac12 \log \det \left( I + \sigma^{-2}
-    \prcov^{1/2} \fwd ^* \obs^* \obs \fwd \prcov^{1/2} \right )\\
-    %
-    %
-    %
-    %
-    &=\frac12 \log \det \Big ( \left ( (\fwd \prcov \fwd^* )^{-1} +
-    \sigma^{-2}\obs^* \obs \right ) \fwd \prcov \fwd^* \Big )\\
-    %
-    %
-    %
-    &=\frac12 \log \left (\prod_{i=1}^{\infty} ( \lambda_i^{-1} + \sigma^{-2} \eta_i ) \prod_{i=1}^{\infty} \lambda_i \right ) \\
-    %
-    %
-    %
-    &=\frac12 \log \left ( \prod_{i=1}^{k} ( \lambda_i^{-1} + \sigma^{-2} \eta_i )\prod_{i=1}^{k} \lambda_i \right )\\
-    %
-    %
-    %
-    &=\frac12 \sum_{i=1}^{k} \log (\sigma^2\lambda_i^{-1} + \eta_i), 
-  \end{split}
-\end{align}
-and the last equality holds up to an additive constant independent of
-$\obs$. Therefore, a D-optimal design maximizes
-
-\begin{equation}\label{eq:true target}
-  \sum_{i=1}^{k}\log (\sigma^2 \lambda_i^{-1} + \eta_i),
-\end{equation}
-
-with the norm constraints on $\meas_j,j=1,\dots,m$. The problem of
-D-optimal design has now been cast to finite dimensions. This
-justifies using linear algebra. Lemma \ref{lemma:free} (given a
-constructive proof in the appendix) does exactly that.
-\begin{restatable*}[Unit norm decomposition]{lemma}{free}\label{lemma:free}
-  Let $M \in \R^{k \times k}$ symmetric positive definite with $\ttr M
-  = m$, $m \geq k$. We can find $\func_j \in \R^k,j=1,\dots,m$
-  with $\|\func_j\|=1$ and $A = (\func_1,\dots,\func_m)$ such that
-  $AA^t = M$.
-\end{restatable*}
-
-Take any $\{\eta_i\}_{i=1}^{k}$, with $\sum_{i=1}^k \eta_i = m$. Let
-$M = \diag(\eta_1,\dots,\eta_k)$. Lemma \ref{lemma:free} gives us $A$
-with unit norm columns such that $AA^T = M$. Take $\meas_i
-:=\sum_{j=1}^k A_{ji}\ev_j^*$. The constraints $\|\meas_i\| = 1,
-i=1,\dots,m$ are satisified by construction. Verifying $\obs^*\obs
-\ev_i = \eta_i \ev_i$ is not too difficult.
-
-%% \begin{align*}
-%%   \begin{split}
-%%     \obs \ev_i &= (\meas_1,\dots,\meas_k)^t\ev_i\\
-%%     &= (\sum_{j=1}^kA_{j1}\ev_j^*,\dots,\sum_{j=1}^kA_{jk}\ev_j^*)^t\ev_i\\
-%%     &= (\sum_{j=1}^kA_{j1}\ev_j^*\ev_i,\dots,\sum_{j=1}^kA_{jk}\ev_j^*\ev_i)^t\\
-%%     &= (A_{i1},\dots,A_{ik})^t.
-%%   \end{split}
-%% \end{align*}
-
-%% Now,
-
-%% \begin{align*}
-%%   \begin{split}
-%%     \obs^*\obs \ev_i &= \obs^*(A_{i1},\dots,A_{ik})^t\\
-%%     &=(A_{i1},\dots,A_{ik}) \obs \text{ (by \eqref{eq:obs*})}\\
-%%     &=(A_{i1},\dots,A_{ik}) (\sum_{j=1}^kA_{j1}\ev_j^*,\dots,\sum_{j=1}^kA_{jk}\ev_j^*)^t\\
-%%     &=\sum_{l=1}^k A_{il} \sum_{j=1}^kA_{jl}\ev_j^* \\
-%%     %&=\sum_{j=1}^k  \sum_{l=1}^kA_{il}A_{jl}\ev_j^* \\
-%%     &=\sum_{j=1}^k  [AA^t]_{ij}\ev_j^* \\
-%%     %&=\eta_i\ev_i^*\\
-%%     &=\eta_i\ev_i^*. 
-%%   \end{split}
-%% \end{align*}
-
-We have shown that we can realize any $\{\eta_i\}_{i=1}^k$, as long as
-$\sum_{i=1}^k \eta_i = m$. We now turn to understand which choice of
-$\{\eta_i\}_{i=1}^k$ is optimal. By concavity of $\log$, the fastest
-increase in the design criterion \eqref{eq:true target} is gained by
-increasing $\sigma^2\lambda_i^{-1} + \eta_i$ where it is smallest. So,
-no weight should be given to (i.e. measurement taken in direction of)
-the $n$th eigenvector before $\sigma^2\lambda_i^{-1} + \eta_i \geq
-\sigma^2\lambda_n^{-1},\ \forall i \neq n$. Thus, we should choose
-$\{\eta_i\}_{i=1}^k$ as follows. First, order the eigenvalues
-$\lambda_1 \geq \lambda_2 \geq \dots$. We should increase $\sigma^2
-\lambda_1^{-1} + \eta_1$ until it equals
-$\sigma^2\lambda_2^{-1}$. Then, $\sigma^2 \lambda_i^{-1} +
-\eta_i,i=1,2$ are increased until $\sigma^2 \lambda_i^{-1} + \eta_i =
-\sigma^2\lambda_3^{-1},i=1,2$ and so forth.
-
-\pgfplotstableread{
-  Label     prior  optimal  sub-optimal 
-  1         0.2    1.8           1.7
-  2         0.8    1.2           0.8
-  3         2.2    0             0.5
-  4         3.5    0             0.0
-}\optimalvsnot
-
-\begin{figure}
-  \begin{tikzpicture}[scale=0.85]
-    \begin{axis}[
-        ybar stacked,
-        ymin=0,
-        ymax=4,
-        xtick=data,
-        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
-        reverse legend=false, % set to false to get correct display, but I'd like to have this true
-        xticklabels from table={\optimalvsnot}{Label},
-        xticklabel style={text width=2cm,align=center},
-        legend plot pos=right,
-        ylabel=precision --- prior and posterior,
-        xlabel=eigenvector $i$,
-      ]
-
-      \addplot [fill=green!80]  table [y=prior,   meta=Label, x expr=\coordindex] {\optimalvsnot};
-      \addplot [fill=blue!60]   table [y=optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};
-
-      \addlegendentry{$\sigma^2\lambda_i^{-1}$}
-      \addlegendentry{optimal $\eta_i$s}
-
-    \end{axis}
-  \end{tikzpicture}
-  \qquad
-  \begin{tikzpicture}[scale=0.85]
-    \begin{axis}[
-        ybar stacked,
-        ymin=0,
-        ymax=4,
-        xtick=data,
-        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
-        reverse legend=false, % set to false to get correct display, but I'd like to have this true
-        xticklabels from table={\optimalvsnot}{Label},
-        xticklabel style={text width=2cm,align=center},
-        legend plot pos=right,
-        ylabel=precision --- prior and posterior,
-        xlabel=eigenvector $i$,
-      ]
-
-      \addplot [fill=green!80]  table [y=prior,       meta=Label, x expr=\coordindex] {\optimalvsnot};
-      \addplot [fill=blue!60]   table [y=sub-optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};
-
-      \addlegendentry{$\sigma^2\lambda_i^{-1}$}
-      \addlegendentry{sub-optimal $\eta_i$s}
-
-    \end{axis}
-  \end{tikzpicture}
-  \caption{Characterizing a D-optimal design, compared to a
-    sub-optimal design. Both designs are allowed three measurements,
-    which means the blue area has accumulated height of $3$ in both
-    plots. A D-optimal design (left) increases precision (reduces
-    uncertainty) where it is lowest. A sub-optimal design (right) does
-    not. In this example, a measurement is taken partly in the
-    direction of the third eigenvector. The precision for this
-    eigenvector is bigger even than the posterior precision of the
-    D-optimal design on the left.}
-  \label{fig:optimal vs not}
-\end{figure}
-
-\pgfplotstableread{
-  Label    prior  $o_1$  $o_2$   $o_3$   $o_4$   $o_5$         $o_6$    topper
-  1         0.2    0.8    0.5     0.5    0.4    0.33333333  0.3333333  0.001
-  2         0.8    0.2    0.5     0.5    0.4    0.33333333  0.3333333  0.001
-  3         2.2    0      0       0      0.2    0.33333333  0.3333333  0.001
-  4         3.5    0      0       0      0      0           0          0.001
-}\clusterization
-
-\pgfplotstableread{
-  Label    prior  $o_1$  $o_2$   $o_3$   $o_4$   $o_5$         $o_6$    topper
-  1         0.2    0.8    0.7     0.3    0.4    0.2         0.4666666  0.001
-  2         0.8    0.2    0.3     0.7    0.4    0.46666666  0.2        0.001
-  3         2.2    0      0       0      0.2    0.33333333  0.3333333  0.001
-  4         3.5    0      0       0      0      0           0          0.001
-}\noclusterization
-
-The argument above has the important consequence (illustrated in
-figure \ref{fig:optimal vs not}) that the eigenvalues in the nonlinear
-eigenvalue problem \eqref{eq:eigenproblem} are all equal! This makes
-perfect sense. We know the Lagrange multipliers $\xi_i,i=1,\dots,k$
-tell us how much we gain by relaxing the constraints and there is no
-reason for any eigenvector to give us more than any other (otherwise
-we would put more weight on the better eigenvector, as can be
-understood from the discussion in the following paragraph).
-
-One can understand D-optimal designs as follows: Imagine each
-eigenvector corresponds to a graduated lab cylinder (figure
-\ref{fig:cylinder}). Each cylinder $i$ is filled, a-priori, with green
-liquid of $\sigma^2\lambda_i^{-1}$ volume units. This liquid
-represnets the prior precision, weighted by observation noise. The
-volume of the green liquid is increasing in $i$. A D-optimal design
-with $m$ measurements ignores cylinders indexed $m+1$ and above. It
-has blue liquid of $m$ volume units at its disposal. It distributes
-said liquid by repeatedly adding a drop to whatever cylinder currently
-has the lowest level of liquid in it (lowest precision). The result of
-such a procedure is illustrated in figure \ref{fig:optimal vs
-  not}. The analysis conducted in this section is summarized in
-Theorem \ref{thm:char} below.
-
-\begin{theorem}[Characterizing D-optimal designs with vanishing model error]\label{thm:char}
-  Let:
-  \begin{enumerate}
-  \item $\fwd, \prcov$ the forward and prior covariance operators, respectively.
-  \item $m, \sigma^2$ the number of measurements taken and observation
-    noise variance, respectively.
-  \item $\obs$ a D-optimal design operator.
-  \item $\tar$ as definded in \eqref{eq:objective}.
-  \item $\{\lambda_i\}_{i=1}^\infty$ eigenvalues of $\fwd\prcov\fwd^*$
-    in decreasing order.
-  \item $\{\ev_i\}_{i=1}^\infty$ their corresponding eigenvectors.
-  \item $\{\eta_i\}_{i=1}^\infty$ eigenvalues of $\obs^*\obs$, where
-    $\obs$ maximizes $\tar(\obs)$.
-  \end{enumerate}
-  Then:
-  \begin{enumerate}
-  \item $\obs^*\obs$ and $\fwd\prcov\fwd^*$ are simultaneously diagonalizable.
-  \item There exists $0 < k \leq m$ such that $\eta_i = 0$ for $i\geq
-    k+1$ and $k = \rank \obs^*\obs$.
-  \item $\sigma^2\lambda_i^{-1} + \eta_i$ are equal for all $1 \leq i \leq k$.
-  \item If $k < m$, then $\sigma^2\lambda_i^{-1} + \eta_i \leq
-    \sigma^2\lambda_{k+1}^{-1}, i=1,\dots,k$.
-  \end{enumerate}
-\end{theorem}
-
-\begin{figure}%{r}{0.25\textwidth} 
-    %\centering
-    \includegraphics[width=5cm, height=5cm]{cylinders.jpg}
-    \caption{Three graduated lab cylinders, corresponding to three
-      eigenvectors. Prior eigenvalues not shown.}
-    \label{fig:cylinder}
-\end{figure}
-
-\section{Sensor Clusterization --- Vanishing Model Error}\label{section:clusterization}
-\subsection{Overview}
-In this section we use Theorem \ref{thm:char} to explain how sensor
-clusterization may arise in various scenarios. We first consider
-sequential designs, and then simultaneous designs.
-
-\subsection{Sequential Design}\label{subsec:clusterization sequential}
-Recall the sequential optimal design scheme from section
-\ref{subsec:seq vs sim} and the definition of $\obs_k$ from
-\eqref{eq:def obs_k}. The way clusterization can occur in sequential
-design scenario is understood by specializing thorem \ref{thm:char} to
-one measurement each time. In this case $m=1$, so $k=1$, $\meas_1
-\parallel \ev_1$ and $\eta_1 = 1$. As hinted in section
-\ref{subsec:seq vs sim}, the posterior becomes the prior for the next
-step. The eigenvectors of the new posterior are the same as for the
-previous one. A new measurement will be taken in the direction of
-eigenvector with smalles eigenvalue. This leads to:
-\begin{align*}
-  \begin{split}
-    \meas_2 \parallel \ev_1  &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 < \sigma^2\lambda^{-1}_2 \\
-    \meas_2 \parallel \ev_2  &\text{ o.w. }
-  \end{split}
-\end{align*}
-In the first case, we get clusterization immediately. In the second,
-we consider $\meas_3$. Same reasoning leads to the conclusion that the
-eigenvectors of the new prior precision are the same. The eigenvalues,
-however, are $\lambda^{-1}_1 + \sigma^{-2}, \lambda^{-1}_2 +
-\sigma^{-2}, \lambda_3^{-1}, \dots$. A new measurement is taken in the
-direction of eigenvector of smallest precision (note that
-$\lambda^{-1}_1 + \sigma^{-2} < \lambda^{-1}_2 + \sigma^{-2}$):
-\begin{align*}
-  \begin{split}
-    \meas_3 \parallel \ev_1 &\text{ if } \sigma^2 \lambda^{-1}_1 + 1 <
-    \sigma^2\lambda^{-1}_3 \\
-    \meas_3 \parallel  \ev_3  &\text{ o.w. }
-  \end{split}
-\end{align*}
-Again, the first case results in clusterization. The sequential design
-proceeds in this fashion. Every $\meas_i$ is taken in the direction
-some eigenvector of the prior $\fwd \prcov\fwd^*$. This eigenvector
-has the smallest eigenvalue of the previous step's posterior precision
-operator. If we do not encounter clusterization after $m$ such steps,
-this means that measurement $\meas_k$ was taken in the direction
-$\ev_k$ for $k=1,\dots,m$. Thus:
-\begin{align}\label{eq:failure}
-  \begin{split}
-    \sigma^2\lambda_{k-1}^{-1} + 1 \geq \sigma^2\lambda_k^{-1}, \text{ for } 1\leq k \leq m
-    \Longrightarrow \lambda_m^{-1} - \lambda_{m-1}^{-1} \leq \sigma^{-2}.
-  \end{split}
-\end{align}
-
-But \eqref{eq:failure} implies $\{\lambda_i^{-1}\}_{i=1}^{\infty}$
-grows at most linearly. In section \ref{subsec:abstract OED} we
-assumed $\fwd$ is strongly smoothing, which means eigenvalues of
-$(\fwd \prcov\fwd^*)^{-1}$ grow quickly. Thus, if we take $m$ large
-enough, \eqref{eq:failure} will eventually fail and we will observe
-sensor clusterization.
-
-\subsection{Simultaneous Design}\label{subsec:clusterization simultaneous}
-In a simultaneous design scenario, the result is not simple as in the
-sequential design case. We now see that a design that exhibits
-clusterization is as optimal as one that does not. Corollary
-\ref{cor:same ev} tells us that we can realize any design we want
-(encoded in $\{\eta_i\}$), given the trace constraint for
-$\obs^*\obs$. Theorem \ref{thm:char} tells us that for D-optimality,
-we only need to make the posterior flat where we take measurements
-(recall figure \ref{fig:optimal vs not}). We are left with the freedom
-of assigning any part of $\eta_i$ to any measurement $\meas_j$. Two
-possible assignments for the same problem are illustrated in figure
-\ref{fig:clusterization}. The first exhibits clusterization and second
-does not. It is important to note that neither arises as a result of
-some numerical optimization scheme. Both designs are give the same
-design criterion $\tar$, because their posteriors are identical. By
-Theorem \ref{thm:char}, both are (possible) maximizers to the same
-D-optimal design problem. One exhibits clusterization, and one does
-not. It is possible that a real-life problem is more restricted than
-the model we consider here. Exactly in what way this happens is not
-clear and should be the subject of further research.
-
-\begin{figure}
-  \begin{tikzpicture}[scale=0.85]
-    \begin{axis}[
-        ybar stacked,
-        ymin=0,
-        ymax=4,
-        xtick=data,
-        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
-        reverse legend=false, % set to false to get correct display, but I'd like to have this true
-        xticklabels from table={\clusterization}{Label},
-        xticklabel style={text width=2cm,align=center},
-        legend plot pos=right,
-        ylabel=precision --- prior and posterior,
-        xlabel=eigenvector $i$,
-      ]
-
-      \addplot [fill=green!80]  table [y=prior, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=blue!60]   table [y=$o_1$, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=red!60]    table [y=$o_2$, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=black!60]  table [y=$o_3$, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=orange!60] table [y=$o_4$, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=cyan!60]   table [y=$o_5$, meta=Label, x expr=\coordindex] {\clusterization};
-      \addplot [fill=purple!60] table [y=$o_6$, meta=Label, x expr=\coordindex] {\clusterization};
-
-      \addlegendentry{prior}
-      \addlegendentry{$o_1$}
-      \addlegendentry{$o_2$}
-      \addlegendentry{$o_3$}
-      \addlegendentry{$o_4$}
-      \addlegendentry{$o_5$}
-      \addlegendentry{$o_6$}   
-    \end{axis}
-  \end{tikzpicture}
-  \qquad
-  \begin{tikzpicture}[scale=0.85]
-    \begin{axis}[
-        ybar stacked,
-        ymin=0,
-        ymax=4,
-        xtick=data,
-        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
-        reverse legend=false, % set to false to get correct display, but I'd like to have this true
-        xticklabels from table={\noclusterization}{Label},
-        xticklabel style={text width=2cm,align=center},
-        legend plot pos=right,
-        ylabel=precision --- prior and posterior,
-        xlabel=eigenvector $i$,
-      ]
-
-      \addplot [fill=green!80]  table [y=prior, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=blue!60]   table [y=$o_1$, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=red!60]    table [y=$o_2$, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=black!60]  table [y=$o_3$, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=orange!60] table [y=$o_4$, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=cyan!60]   table [y=$o_5$, meta=Label, x expr=\coordindex] {\noclusterization};
-      \addplot [fill=purple!60] table [y=$o_6$, meta=Label, x expr=\coordindex] {\noclusterization};
-
-      \addlegendentry{prior}
-      \addlegendentry{$o_1$}
-      \addlegendentry{$o_2$}
-      \addlegendentry{$o_3$}
-      \addlegendentry{$o_4$}
-      \addlegendentry{$o_5$}
-      \addlegendentry{$o_6$}   
-    \end{axis}
-  \end{tikzpicture}
-  \caption{Clusterization (left) and non-clusterization (right) in
-    simultaneous D-optimal designs. Posterior precision per
-    eigenvector after taking measurements $\meas_i$ is plotted. Both
-    designs are identical for all practical matters --- their
-    posteriors are identical. In the left panel we see how an optimal
-    design is achieved with repeated measurements. $\meas_2 = \meas_3$
-    and $\meas_5 = \meas_6$. This is not necessary though, as can be
-    seen in the right panel. Both designs achieve the same levels of
-    uncertainty. The left exhibits clusterization. The right does not.}
-  \label{fig:clusterization}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% SECTION Analysis of Optimal Designs --- Non-Vanishing Model Error
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\section{Analysis of D-Optimal Designs --- Non-Vanishing Model Error}\label{section:non vanishing}
-In this section we show the effect model error has on the
-clusterization phenomenon. We will see that if $\modcov \neq 0$
-clustering will not occur. This is contrary to the previous case of
-vanishing model error.
-
-Corollary \ref{cor:same meas} below is proved in the appendix. Denote
-$\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm :=
-(\meas_1,\dots,\meas_{m-1})^t$. Denote $\Sigmam := \Sigma (\obsm)$ and
-$\postcovm$ the posterior covariance that arises when we use $\obsm$.
-\begin{restatable*}{corollary}{samemeas}\label{cor:same meas}
-  If $\meas_m = \meas_j$ for some $1 \leq j \leq m-1$, then
-  \begin{equation*}
-    \tar(\obs) - \tar(\obsm) =
-    \log \left ( 1 + \frac{\sigma^2
-      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
-      \obsm^* \Sigmam^{-1}e_j \rangle
-    }{
-      2 - \sigma^2 e_j^t\Sigmam^{-1}e_j 
-    }       
-    \right ).
-  \end{equation*}
-\end{restatable*}
-Recall from \eqref{eq:Sigma} that $\Sigma(\obs) = \obs \modcov \obs^*
-+ \sigma^2I$. Then
-
-$$
-\lim_{\sigma^2 \to 0} \tar(\obs) -\tar(\obsm) = 0
-$$
-
-Hence no increase in the design criterion is achieved by taking a
-repeated measurement (in the limit $\sigma^2 \to 0$). This is clearly
-sub-optimal and $\obsm$ cannot be a D-optimal design. Thus, for small
-measurement error levels, the clusterization effect will be mitigated
-by the presence of a non-zero model error. Since the design criterion
-is not defined for $\sigma^2 = 0$ and identical measurements, we
-cannot make a statement regarding $\sigma^2 = 0$, except in the
-limiting sense described above.
-
-\section{Acknowledgements}
-This study is a part of my PhD thesis. Most of it was written under
-the instruction of Prof. Georg Stadler in New York University's
-Courant Institute. I would like to thank him for his great
-mentorship. This work was supported by The Raymond and Beverly Sackler
-Post-Doctoral Scholarship.
-
-\appendix
-\section{General Lemmas}
-
-The following lemma is generalized from \cite[Chapter 9, Theorem 4, pp. 127]{Lax97}
-\lax
 \begin{proof}
@@
   $\lambda_k(X(t))$ and sometimes drop the dependence on $X(t)$, so
-  $\lambda_k = \lambda_k(X(t))$.  Then $\det (I+X(t)) =
-  \prod_{k=1}^{\infty} (1+\lambda_k) < \infty$ where the bound holds
-  by the arguments given in \cite{AlexanderianGloorGhattas14}. The
-  full derivative is
-  \begin{align*}
-    \frac{\der \det (I+X(t))}{\der t} 
+  $\lambda_k = \lambda_k(X(t))$.  Then $\det (I+X(t))
+  = \prod_{k=1}^{\infty} (1+\lambda_k) < \infty$ and this is finite by
+  the arguments given in \cite{AlexanderianGloorGhattas14}. The full
+  derivative is \begin{align*} \frac{\der \det (I+X(t))}{\der t}
     % 
@@
 
-\begin{lemma}[Matrix Determinant Lemma in Hilbert Spaces]\label{lemma:MDL}
+The following is a generalization of the Matrix Determinant Lemma to
+Hilbert spaces.
+\begin{lemma}%\label{lemma:MDL}
   Let $\hil$ a separable Hilbert space, $u,v\in \hil$ and $A: \hil \to
@@
   In this proof we rely on definitions and results from
-  \cite{Simon77}. First, consider $B := I + xy^*$ for some $x,y \in
+  \cite{simon1977}. First, consider $B := I + xy^*$ for some $x,y \in
   \hil$. We construct an eigenbasis for $B$ and use that to show $\det
@@
   \parallel y$, take $\{x_n \}_{n=2}^{\infty}$ an orthogonal basis for
-  $span\{x_1\} ^{\perp}$. If, on the other hand, $x \nparallel y$, let
+  $span\{x_1\} ^{\perp}$. If, on the other hand, $x \not \parallel y$, let
   \begin{equation*}
@@
 
-\free
-\begin{proof}
-  Let us diagonalize $M$, so that $M = U D U^t$ with $D =
-  \diag(d_1,\dots,d_k)$ and $U \in \R^{k \times k }$ orthogonal. Let
-  $S \in \R^{k \times m}$ with $S_{ii} = \sqrt{d_{i}}$ and zeros
-  otherwise. Define $A:= U S V^t$, where $V \in \R^{m \times m}$ is
-  orthogonal and will be further restricted later. Then $AA^t = U
-  SV^tVS^t U^t = UDU^t$, so $AA^t$ has the required eigenvalues and
-  eigenvectors by construction. If we can choose $V$ such that $A$
-  also satisfies the unit norm constraints we are done. These
-  constraints are, for $j=1,\dots,m$:
-  \begin{equation}\label{eq:V constraints}
-   1 = [A^tA]_{jj} = [V S^tS V^t]_{jj},
-  \end{equation}
-  and we can expect to do this since we assumed $\ttr D = m$.
-
-  Define $C = S^tS - I \in \R^{m \times m}$. Note that $\ttr C = 0$ and
-  $C$ is diagonal with non-zero entries $d_i-1,i=1,\dots,k$. It suffices
-  to find $V$ orthogonal such that $V C V^t$ has zero diagonal. We
-  construct such $V$ by sequentially inserting zeros in the diagonal
-  and not destroying zeros we already introduced, starting from the
-  last diagonal entry and moving to the first. Since $c_{mm} \neq 0$ ,
-  let $p < m$ such that $c_{pp}c_{mm} < 0$ (such $p$ exists because
-  the trace is zero) and let $\theta \in (0,\pi)$. Define a Givens
-  rotation $R^{(m)} \in \R^{m \times m}$ by
-  \begin{equation*}
-    r^{(m)}_{ab} :=
-    \begin{cases}
-      1 & a = b \neq p \text{ or } a = b \neq m \\
-      \cos \theta & a = b = p  \\
-     -\sin \theta & a = p, b = m\\
-      \cos \theta & a = b = m \\
-      \sin \theta & a = m, b = p \\ 
-      0 & o.w
-    \end{cases}
-  \end{equation*}
-  Note that conjugating a matrix by $R^{(m)}$ changes only its $m$ and
-  $p$ rows and columns. We want to choose $\theta$ such that
-  \begin{equation}\label{eq:mm}
-    0 = [R^{(m)} C (R^{(m)})^t]_{mm} = \cos^2 \theta c_{mm} + 2\cos \theta \sin
-    \theta c_{mp} + \sin^2\theta c_{pp},
-  \end{equation}
-  and it suffices to choose $\theta$ such that
-  \begin{equation*}
-    c_{mm} \cot^2 \theta + 2 c_{mp} \cot \theta + c_{pp} = 0.
-  \end{equation*}
-  This quadratic in $\cot\theta$ has a real solution, since
-  $c_{pp}c_{mm} < 0$ by assumption and we can find $\theta \in
-  (0,\pi)$ such that \eqref{eq:mm} is satisfied. We continue to find
-  $R^{(m-1)}$ that leaves row and column $m$ unchanged and
-  continue introducing zeros to the diagonal. The assumption $\ttr D =
-  m \Rightarrow \ttr C = 0$ guarantees we can do that. Taking $V:=
-  R^{(1)} R^{(2)} \dots R^{(m-1)}R^{(m)}$ completes the proof.
-\end{proof}
-
-\simdiag
-\begin{proof}
-  First, enumerate the eigenvalues of $C + \sum_{j=1}^m
-  \func_j\func_j^*$ as $\xi_1,\dots,\xi_\ell$. Denote the
-  indices of the eigenvectors corresponding to $\xi_i$
-  \begin{equation*}
-    S_i := \{ 1 \leq k \leq m | (C + \sum_{j=1}^m \func_j\func_j^* )\func_k = \xi_i \func_k \}.
-  \end{equation*}
-  Define further
-  \begin{equation*}
-    A_i := \sum_{k \in S_i} \func_k \func_k^*,
-  \end{equation*}
-  which is self-adjoint. Two observations are in order. First,
-  $\sum_{j=1}^m \func_j\func_j^* = \sum_{i=1}^\ell A_i$. Second, $A_i
-  \func_k = 0$ if $k\not \in S_i$, since eigenvectors of different
-  eigenvalue are orthogonal. For $k \in S_i$
-  \begin{equation}\label{eq:on vi}
-    \xi_i \func_k = (C + \sum_{j=1}^m \func_j \func_j^* ) \func_k = (C + A_i) \func_k.
-  \end{equation}
-  Let $V_i := span \{\func_k \}_{k\in S_i}$. Observe that $V_i$ is
-  invariant under $A_i$, by definition, and under $C$, by \eqref{eq:on
-    vi}. Using \eqref{eq:on vi} again, we conclude that on $V_i$, $A_i
-  = \xi_iI - C$. This immediately implies $A_i$ and $C$ have the
-  same eigenvectors on $V_i$. This holds for every $1 \leq i \leq
-  \ell$ and we conclude that $C$ and $A$ have the same eigenvectors.
-\end{proof}
-
-\section{Specific Lemmas}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{lemma}\label{lemma:twice woodbury}
-  Assume $\fwd \prcov \fwd^*$ is invertible. Then
-\begin{align*}
-  \begin{split}
-    \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* 
-    %
-    %
-    = \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1},
-  \end{split}
-\end{align*}  
-\end{lemma}
-\begin{proof}
-  The proof is boring and amounts to using Woodbury's matrix identity
-  twice and a regularization trick. The standard proof for this
-  identity works in infinite dimensions, as long as all terms are well
-  defined. Unfortunately, $\obs^*\obs$ is not invertible, so we force
-  it to be. Throughout, we let $X := \fwd\prcov \fwd^*$ and $Y_{\eps}
-  := \obs^*\obs + \eps I$, for some small $\eps > 0$. Note that now
-  $Y_{\eps}$ is invertible. Recall Woodbury's matrix identity:
-  $$
-  (A + UCV)^{-1} = A^{-1} + A^{-1}U(C^{-1} + VA^{-1}U)^{-1}VA^{-1}.
-  $$
-
-  Taking $A := \prcov^{-1}, U := \fwd^*, V := \fwd$ and $C :=
-  \sigma^{-2} (\obs^*\obs+\eps I)$ we get:
-  \begin{align*}
-    \begin{split}
-      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* &=
-      \fwd ( \prcov - \prcov \fwd^* ( \sigma^2(\obs^*\obs + \eps I)^{-1} + \fwd \prcov \fwd^* )^{-1} \fwd \prcov ) \fwd^* \\
-      %
-      %
-      &= X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X
-    \end{split}
-  \end{align*}
-  Note that $X + \sigma^2 Y_{\eps}^{-1}$ is invertible, as the sum of
-  a two positive definite operators. Now, taking $A := X^{-1}, C :=
-  \sigma^2Y_{\eps}^{-1}, U := I$ and $V := I$ and using Wodbury's matrix
-  identity in reverse order:
-  \begin{align*}
-    \begin{split}
-      X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X &= (X^{-1} + \sigma^{-2}Y_{\eps})^{-1} \\
-      %
-      %
-      %
-      &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \eps I))^{-1}.
-    \end{split}
-  \end{align*}
-
-  We conclude that $\forall \eps > 0$
-  \begin{align*}
-    \begin{split}
-      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* 
-     &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \eps I))^{-1}.
-    \end{split}
-  \end{align*}
-  Letting $\eps \to 0$ completes the proof.
-\end{proof}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{lemma}[Increase due to a Measurement]\label{lemma:design increase}
-  Let $\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm := (\meas_1,\dots,\meas_{m-1})^t$. Then
-  \begin{align*}
-    \tar( \obs ) - \tar (\obsm ) &=
-    \frac12 \log \left ( 1 + \frac{
-      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m,
-      (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m \rangle
-    }{
-      \sigma^2 + \meas_m \modcov \meas_m - \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m 
-    }       
-    \right ).
-  \end{align*}
-\end{lemma}
-\begin{proof}
-  We use the Schur complement to write one inverse in terms of the other and
-  introduce notations to make the derivation cleaner. Note that we think of
-  $\obsm$ and $\obsm^*$ as column and row vectors (respectively).
-  \begin{align*}
-    \Sigma( \obs ) &= \Sigma = 
-    \begin{bmatrix}
-      \Sigma (\obsm )           & \obsm \modcov \meas_m \\
-      \meas_m \modcov \obsm^*   & \sigma^2 + \meas_m \modcov \meas_m
-    \end{bmatrix}
-    =
-    \begin{bmatrix}
-      \Sigmam   & w \\
-      w^t       & c
-    \end{bmatrix}\\
-    %
-    %
-    %
-    \Sigma^{-1} &=
-    \begin{bmatrix}
-      \Sigmam^{-1} + \Sigmam^{-1} w ( c - w^t \Sigmam^{-1} w)^{-1} w^t \Sigmam^{-1} & - \Sigmam^{-1} w ( c - w^t \Sigmam^{-1} w)^{-1} \\
-      -( c - w^t \Sigmam^{-1} w)^{-1} w^t \Sigmam^{-1}                            &  ( c - w^t \Sigmam^{-1} w)^{-1}
-    \end{bmatrix} \\
-    &=
-    \begin{bmatrix}
-      \Sigmam^{-1} & 0 \\
-      0           & 0 
-    \end{bmatrix}
-    + (c -w^t \Sigmam^{-1} w )^{-1}
-    \begin{bmatrix}
-      \Sigmam^{-1} w \\
-      -1
-    \end{bmatrix}
-    \begin{bmatrix}
-      w^t \Sigmam^{-1} & -1 
-    \end{bmatrix}
-  \end{align*}
-  %
-  Further, define
-  %
-  \begin{align*}
-    \M (\obs ):&= \prcov^{\frac12}\fwd^* \obs^* \Sigma^{-1} \obs \fwd
-    \prcov^{\frac12}    
-  \end{align*}
-  %
-  and note that, using our understanding of what is a column vector and
-  what is a row vector:
-  %
-  \begin{align*}
-    \M(\obs) &= \prcov^{1/2} \fwd^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2} \\
-    %
-    %
-    %
-    &= \prcov^{1/2} \fwd^* \obs^* \left \{
-    \begin{bmatrix}
-      \Sigmam^{-1} & 0 \\
-      0           & 0 
-    \end{bmatrix}
-    + (c -w^t \Sigmam^{-1} w )^{-1}
-    \begin{bmatrix}
-      \Sigmam^{-1} w \\
-      -1
-    \end{bmatrix}
-    \begin{bmatrix}
-      w^t \Sigmam^{-1} & -1 
-    \end{bmatrix} 
-    \right \} \obs \fwd \prcov^{1/2} \\
-    %
-    %
-    %
-    &= \M (\obsm) + (c -w^t \Sigmam^{-1} w )^{-1}
-    \prcov^{1/2} \fwd^* \obs^*
-    \begin{bmatrix}
-      \Sigmam^{-1} w \\
-      -1
-    \end{bmatrix}
-    \begin{bmatrix}
-      w^t \Sigmam^{-1} & -1 
-    \end{bmatrix} 
-    \obs \fwd \prcov^{1/2}
-  \end{align*}
-  %
-  Now, denote:
-  %
-  \begin{align}\label{eq:u}
-    \begin{split}
-      u :&= (c -w^t \Sigmam^{-1} w )^{-1/2}
-      \prcov^{1/2} \fwd^* \obs^* 
-      \begin{bmatrix}
-        \Sigmam^{-1} w \\
-        -1 
-      \end{bmatrix} \\
-      %
-      %
-      %
-      & = (c -w^t \Sigmam^{-1} w )^{-1/2} ( \prcov^{1/2}\fwd^* \obsm^* \Sigmam^{-1} \obsm  \modcov \meas_m - \prcov^{1/2} \fwd^* \meas_m )\\
-      %
-      %
-      %
-      u^* :&=  (c -w^t \Sigmam^{-1} w )^{-1/2} (\meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \fwd \prcov^{1/2} - \meas_m \fwd \prcov^{1/2} ),
-    \end{split}
-  \end{align}
-  %
-  so that
-  %
-  \begin{equation}\label{eq:M plus I}
-    I + \M( \obs ) = I + \M (\obsm ) + uu^*.
-  \end{equation}
-  %
-  Note that
-  \begin{equation}\label{eq:M postcov}
-    \prcov^{1/2} \left (I + \M( \obsm ) \right )^{-1} \prcov^{1/2} = \postcovm.
-  \end{equation}
-  The increase in the design criterion gained by including $\meas_m$
-  in the design is found by using Lemma \ref{lemma:MDL} the above
-  results:
-  %
-  \begin{align*}
-    \tar( \obs ) - \tar( \obsm )
-    %
-    %
-    %
-    &= \frac12 \log \det \Big ( I + \M ( \obs ) \Big ) / \det \Big ( I + \M (\obsm) \Big ) \\
-    %
-    %
-    %
-    &= \frac12  \log \det \left ( I + \M(\obsm) + uu^* \right ) / \det \Big ( I + \M (\obsm) \Big ) \\
-    %
-    %
-    %
-    &= \frac12 \log \left ( 1 + \left \langle \left ( I+\M(\obsm) \right )^{-1} u, u  \right \rangle \right ).
-  \end{align*}
-  Using \eqref{eq:u}:
-  \begin{align*}
-    &\left \langle \left (I+\M (\obsm)\right )^{-1}u, u \right \rangle\\
-    &= \frac{
-      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m,
-      (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m \rangle
-    }{
-      c- w^t \Sigmam^{-1} w
-    }\\
-    %
-    %
-    %
-    &= 
-    \frac{
-      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m,
-      (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m \rangle
-    }{
-      \sigma^2 + \meas_m \modcov \meas_m - \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m 
-    }
-  \end{align*}
-  and the conclusion follows.
-\end{proof}
-
-Lemma \ref{lemma:design increase} implies the following two corollaries.
-\begin{corollary}[Gain for No Model Error]\label{cor:zero mod err}
-  If $\modcov = 0$, then
-  \begin{equation*}
-    \tar( \obs ) - \tar (\obsm )
-    = -\frac12 \log (1 - \sigma^{-2} \langle \fwd \postcov \fwd^* \meas_m, \meas_m \rangle ).
-  \end{equation*}
-\end{corollary}
-\begin{proof}
-  Note that this is not immediate by substituting $\modcov = 0$ in the
-  conclusion of Lemma \ref{lemma:design increase}, since we make a
-  claim for $\postcov$, and not $\postcovm$. Let us first review
-  \eqref{eq:u} and note that since $\modcov = 0$ the covariance
-  $\Sigmam = \sigma^2I_{m-1}$ and $w = 0$, so $c - w^t\Sigmam^{-1}w =
-  \sigma^2$:
-  \begin{align*}
-    u :&=-\sigma^{-1}\prcov^{1/2} \fwd^* \meas_m\\
-    %
-    %
-    u^* :&= -\sigma^{-1} \meas_m \fwd \prcov^{1/2}.
-  \end{align*}
-  From \eqref{eq:M plus I}:
-  \begin{equation*}
-    I + \M(\obsm) = I +\M(\obs) - uu^*,
-  \end{equation*}
-  and thus:
-  \begin{equation*}
-    \left \langle \left ( I +\M(\obs) \right )^{-1} u, u \right \rangle
-    = \sigma^{-2} \langle \fwd \postcov \fwd^* \meas_m, \meas_m \rangle.
-  \end{equation*}
-  Analogously to \eqref{eq:M postcov} we note that
-  \begin{equation*}
-    \prcov^{1/2} \left ( I + \M(\obs) \right )^{-1} \prcov^{1/2} = \postcov.
-  \end{equation*}
-  Using Lemma \ref{lemma:MDL} we conclude
-  \begin{align*}
-    \tar( \obs ) - \tar( \obs )
-    &= \frac12 \log \det \left (I +\M(\obs) \right ) / \det \left (I + \M(\obsm) \right ) \\
-    %
-    %
-    %
-    &= \frac12 \log \det \left (I +\M(\obs) \right ) / \det \left (I + \M(\obs) - uu^* \right ) \\
-    %
-    %
-    %
-    &=-\frac12 \log (1 - \langle (I+\M(\obs))^{-1}u, u \rangle \\
-    %
-    %
-    %
-    &= -\frac12 \log (1 - \sigma^{-2} \langle \fwd \postcov \fwd^* \meas_m, \meas_m \rangle ).
-  \end{align*}
-\end{proof}
-
-\samemeas
-\begin{proof} \label{cor:same meas proof}
-  Denote $A:= \obs \modcov \obs^*$ and $v_j$ the $j$th column of $A$.
-  Note that $v_j = \obsm \modcov \meas_m$, since $(\obsm \modcov
-  \obsm^*)_{ij} = \meas_i(\modcov \meas_j)$, as explained in
-  \eqref{eq:modcov explained}. One can now verify that
-  \begin{equation}\label{eq:observation}
-    \Sigmam^{-1} \obsm \modcov \meas_m = \Sigmam^{-1}v_j = (A +\sigma^2I_{m-1})^{-1} v_j =
-    e_j -\sigma^2 \Sigmam^{-1}e_j.
-  \end{equation}
-  %
-  Using \eqref{eq:observation}:
-  \begin{align}\label{eq:denominator}
-    \begin{split}
-      \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m
-      &= \meas_m \modcov \obsm^* ( e^j - \sigma^2 \Sigmam^{-1} e_j )\\
-      %
-      %
-      %
-      &= \meas_m \modcov \meas_j - \sigma^2 \meas_m \modcov \obsm^* \Sigmam^{-1}e_j \\
-      %
-      %
-      %
-      &= \meas_m \modcov \meas_j -\sigma^2 (e_j - \sigma^2 \Sigmam^{-1}e_j)^t e_j \\
-      %
-      %
-      %
-      &= \meas_m \modcov \meas_m -\sigma^2 + \sigma^4 e_j^t\Sigmam^{-1}e_j.
-    \end{split}
-  \end{align}
-  We use \eqref{eq:observation} to simplify the terms in the enumerator of
-  the conclusion of Lemma \ref{lemma:design increase}:
-  \begin{align}\label{eq:enumerator}
-    \begin{split}
-      (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m
-      &= \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m - \meas_m \\
-      %
-      %
-      %
-      &= \obsm^* (e_j - \sigma^2 \Sigmam^{-1} e_j) -\meas_j \\ 
-      %
-      %
-      %
-      &= -\sigma^2 \obsm^* \Sigma^{-1}e_j. 
-    \end{split}
-  \end{align}
-  %
-  Substitute \eqref{eq:enumerator} and \eqref{eq:denominator} to
-  the enumerator and denominator (respectively) of the conclusion of
-  Lemma \ref{lemma:design increase}:
-  %
-  \begin{align*}
-    \tar( \obs ) - \tar (\obsm ) &=
-    \log \left ( 1 + \frac{
-      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m,
-      (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m \rangle
-    }{
-      \sigma^2 + \meas_m \modcov \meas_m - \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m 
-    }       
-    \right ) \\
-    %
-    %
-    %
-    &= \log \left ( 1 + \frac{\sigma^4
-      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
-      \obsm^* \Sigmam^{-1}e_j \rangle
-    }{
-      2\sigma^2 - \sigma^4 e_j^t\Sigmam^{-1}e_j 
-    }       
-    \right ) \\
-    %
-    %
-    %
-    &= \log \left ( 1 + \frac{\sigma^2
-      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
-      \obsm^* \Sigmam^{-1}e_j \rangle
-    }{
-      2 - \sigma^2 e_j^t\Sigmam^{-1}e_j 
-    }       
-    \right ).
-  \end{align*}
-\end{proof}
-
-\begin{lemma}[Auxilliary Calculations]\label{lemma:aux calc}
-  Let $T(\obs) := \obs^* \Sigma^{-1}(\obs)\obs$, with $\Sigma(\obs)$
-  defined as in \eqref{eq:Sigma}. Then:
-  \begin{align*}
-    \delta T(\obs)V &= V^* \Sigma^{-1} \obs 
-    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
-    &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs
-    + \obs^* \Sigma^{-1} V.
-  \end{align*}
-\end{lemma}
-
-\begin{proof}
-  We need a few supplementary calculations. First:
-  \begin{align}\label{eq:der sig}
-    \begin{split}
-      \frac{\der}{\der \tau} \Big |_{\tau=0} \Sigma( \obs + \tau V )
-      &= \frac{\der}{\der \tau} \Big |_{\tau=0} 
-      (\obs + \tau V ) \modcov (\obs + \tau V )^*  + \sigma^2I\\
-      % 
-      % 
-      % 
-      &= V \modcov \obs^* + \obs \modcov V^*.
-    \end{split}
-  \end{align}
-  By the standard trick for the derivative of an operator: 
-  \begin{align*}
-    0 &= \frac{\der}{\der \tau} \Big |_{\tau=0} I \\
-    % 
-    % 
-    % 
-    &= \frac{\der}{\der \tau} \Big |_{\tau=0}
-    \left (\Sigma(\obs+\tau V)^{-1} \Sigma(\obs+\tau V) \right ) \\
-    % 
-    % 
-    % 
-    &= \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0} \Sigma+
-    \Sigma^{-1} \frac{\der \Sigma(\obs+\tau V)}{\der \tau} \Big |_{\tau=0}\\  
-    %
-    %
-    %
-    &= \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0} \Sigma+
-    \Sigma^{-1} (V\modcov \obs^* + \obs \modcov V^*) 
-    \text{, by \eqref{eq:der sig}. }
-  \end{align*}
-  Now we can write the variation of $\Sigma^{-1}$:
-  \begin{align}\label{eq:der sig inv}
-    \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0}  
-      &= -\Sigma^{-1} (V \modcov \obs^* + \obs \modcov V^*) \Sigma^{-1}.
-    \end{align}
-  % 
-  Finally, we can do the calculation we need to. By Leibnitz (product) rule
-  and \eqref{eq:der sig inv}:
-  % 
-  \begin{align*}
-    \delta T(\obs) V 
-    &= \frac{\der T(\obs + \tau V)}{\der \tau} \Big |_{\tau=0} \\
-    %
-    %
-    %
-    &= V^* \Sigma^{-1} \obs 
-    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
-    &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs
-    + \obs^* \Sigma^{-1} V,
-  \end{align*}
-  as required.
-\end{proof}
-
-\biggerbetter
-\begin{proof} 
-  Fix an arbitrary $j=1,\dots,m$ and take $V:= e_j e_j^t \obs$. We see
-  that for $u \in \hilo$
-  \begin{equation*}
-    Vu = e_je_j^t (\meas_1(u),\dots,\meas_m(u) )^t = e_j \meas_j(u)
-    = (0,\dots,0,\meas_j(u),0,\dots,0)^t.
-  \end{equation*}
-  This way, $V$ has the same $j$th entry as $\obs$ while the rest are
-  set to zero. We calculate the variation in this direction and show
-  it is positive --- $\delta \tar(\obs)V > 0$. This means that
-  increasing the magnitude of the $j$th measurement functional
-  increases $\tar(\obs)$. We start with the last line of \eqref{eq:tar var}, and denote $\tmp : = \fwd \postcov \fwd^*$.
-  \begin{align*}
-     \delta \tar(\obs) V 
-    &= \tr{V ( I - \modcov \obs^*\Sigma^{-1}\obs) \tmp \obs^* \Sigma^{-1}} \\
-    % 
-    %
-    %
-    &= \tr{e_je_j^t \obs ( I - \modcov \obs^*\Sigma^{-1}\obs) \tmp \obs^* \Sigma^{-1}} \\
-    %
-    % 
-    %
-    &= e_j^t \obs ( I - \modcov \obs^*\Sigma^{-1}\obs) \tmp \obs^* \Sigma^{-1}e_j \\
-    %
-    % 
-    %
-    &= e_j^t ( I - \obs \modcov \obs^*\Sigma^{-1})\obs \tmp \obs^* \Sigma^{-1}e_j \\  
-    % 
-    %
-    %
-    &=  e_j^t(\Sigma-\obs \modcov \obs^*) \Sigma^{-1}\obs \tmp \obs^* \Sigma^{-1}e_j \\
-    %
-    %
-    %
-    &=\sigma^2 e_j^t \Sigma^{-1}\obs \tmp \obs^* \Sigma^{-1}e_j
-    \text{ by \eqref{eq:Sigma} }\\
-    %
-    % 
-    %
-    &=\sigma^2 e_j^t \Sigma^{-1}\obs \fwd \postcov \fwd^* \obs^* \Sigma^{-1}e_j.
-  \end{align*} 
-  Since $\postcov$ is positive definite, $\delta \tar(\obs) V > 0$.
-\end{proof}
-
+%\bibliographystyle{siamplain}
 \bibliographystyle{amsplain}
-\bibliography{refs,georg_refs}
+\bibliography{/home/yair/projects/bibtex.bib}
 

### 2023-08-09 latex/intro.tex
Ready for SIAM applied math?

@@
       Theorem \ref{thm:d optimality}. Measurement locations are
-      plotted over the computational domain $\Omega = [0, 1]$
+      plotted over the computational domain \(\Omega = [0, 1]\)
       (x-axis), for varying numbers of measurements (y-axis). The
@@
 modeling of auto-correlation time within the noise terms
-\cite{hooker2009}.
+\cite{nyberg2012}.
 
 In spatial problems involving choice of measurements within a domain
-$\Omega \subseteq \mathbb{R}^d, d=1,2,3$, many researchers circumvent
+\(\Omega \subseteq \mathbb{R}^d, d=1,2,3\), many researchers circumvent
 the problem of measurement clusterization by choosing measurements
-from a coarse grid in $\Omega$ \cite{koval2020, alexanderian2021,
+from a coarse grid in \(\Omega\) \cite{koval2020, alexanderian2021,
   attia2022, alexanderian2014, alexanderian2016,
-  alexanderian2018efficient}. This approach incurs a significant
-computational cost as it requires solving a difficult combinatorial
-optimization problem for measurement locations over a finite set. The
-combinatorial optimization problem is usually relaxed by first
-assigning optimal measurement weights in $\mathbb{R}_+$ to the
-potential measurement locations. Some researchers incorporate a
-sparsifying $\ell_1$ penalty term into the design criterion, which is
+  alexanderian2018efficient, brunton2016}. This approach incurs a
+significant computational cost as it requires solving a difficult
+combinatorial optimization problem for measurement locations over a
+finite set. The combinatorial optimization problem is usually relaxed
+by first assigning optimal measurement weights in \(\mathbb{R}_+\) to
+the potential measurement locations. Some researchers incorporate a
+sparsifying \(\ell_1\) penalty term into the design criterion, which is
 subsequently thresholded to achieve the desired binary design over the
 coarse grid \cite{horesh2008borehole}. Others progressively relax the
-$\ell_1$ penalty to an $\ell_0$ penalty via a continuation method
+\(\ell_1\) penalty to an \(\ell_0\) penalty via a continuation method
 \cite{alexanderian2016, alexanderian2014}. A clever alternative is to
@@
 fundamental issue: the restriction of measurement locations to a
-coarse grid in $\Omega$ fundamentally changes the optimal design
+coarse grid in \(\Omega\) fundamentally changes the optimal design
 problem and thus results in a sub-optimal design.
@@
 %
-Is measure clusterization a generic phenomenon?
+Is measurement clusterization a generic phenomenon?
 %
@@
   %% Computer implementation of Lemma \ref{thm:char} for the inverse
-  %% problem outlined in Section \ref{section:how} also generates $\obs$
+  %% problem outlined in Section \ref{section:how} also generates \(\obs\)
   %% as a solution to the D-optimal design problem.  Furthermore,
@@
   those with the largest eigenvalues in the prior covariance
-  spectrum. In practical scenarios, the number of locations where the
-  relevant prior eigenvectors are significantly large, while other
-  eigenvectors have a power spectrum close to zero, is
-  limited. Consequently, the clustering of measurements arises as a
-  natural consequence of the pigeonhole principle, as there are more
+  spectrum. In practical scenarios, the number of locations where (a)
+  the relevant prior eigenvectors are significantly large, and (b)
+  other eigenvectors have a power spectrum close to zero, is
+  limited. Consequently, the clusterization of measurements arises as
+  a natural consequence of the pigeonhole principle, as there are more
   measurements available than there are locations satisfying
@@
 We first give some general definitions before we state Theorem
-\ref{thm:char}: Let $\hilp, \hilo$ Hilbert spaces, $\fwd:\hilp \to
-\hilo$ a linear compact operator. Let $\obs: \hilo \to \mathbb{R}^m$ a
-linear measurement operator, where $m \in \mathbb{N}$ is the number of
-measurements taken. Let $\sigma^2 \in \mathbb{R}_{+}$ observation
-noise variance, $\data = \obs \fwd \param + \eps$, where $\eps \in
-\mathbb{R}^m$ is iid $\mathcal{N}(0, \sigma^2)$ noise. Let $\pr \sim
-\mathcal{N}(\prmean, \prcov)$ prior Gaussian measure on $\hilp$, where
-$\prcov$ is the prior covariance operator and let $\post$ the
+\ref{thm:char}: Let \(\hilp, \hilo\) Hilbert spaces, \(\fwd:\hilp \to
+\hilo\) a linear compact operator. Let \(\obs: \hilo \to \mathbb{R}^m\) a
+linear measurement operator, where \(m \in \mathbb{N}\) is the number of
+measurements taken. Let \(\sigma^2 \in \mathbb{R}_{+}\) observation
+noise variance, \(\data = \obs \fwd \param + \eps\), where \(\eps \in
+\mathbb{R}^m\) is iid \(\mathcal{N}(0, \sigma^2)\) noise. Let \(\pr \sim
+\mathcal{N}(\prmean, \prcov)\) prior Gaussian measure on \(\hilp\), where
+\(\prcov\) is the prior covariance operator and let \(\post\) the
 posterior (Gaussian) measure.
@@
     \end{align*}
-  \item $\obs$ a D-optimal design operator
+  \item \(\obs\) a D-optimal design operator
     \begin{equation*}
@@
     \end{equation*}
-  \item $\{\lambda_i\}_{i=1}^\infty$ eigenvalues of
-    $\fwd\prcov\fwd^*$ in decreasing order of magnitude.
-    %% \item $\{\ev_i\}_{i=1}^\infty$ their corresponding eigenvectors.
-  \item $\{\eta_i\}_{i=1}^\infty$ eigenvalues of $\obs^*\obs$.
+  \item \(\{\lambda_i\}_{i=1}^\infty\) eigenvalues of
+    \(\fwd\prcov\fwd^*\) in decreasing order of magnitude.
+    %% \item \(\{\ev_i\}_{i=1}^\infty\) their corresponding eigenvectors.
+  \item \(\{\eta_i\}_{i=1}^\infty\) eigenvalues of \(\obs^*\obs\).
 
@@
   \begin{enumerate}
-  \item  $\tr{\obs^*\obs} = m$.
-  \item $\obs^*\obs$ and $\fwd\prcov\fwd^*$ are simultaneously
+  \item  \(\tr{\obs^*\obs} = m\).
+  \item \(\obs^*\obs\) and \(\fwd\prcov\fwd^*\) are simultaneously
     diagonalizable.
-  \item $k := \rank \obs^*\obs \leq m$ and
+  \item \(k := \rank \obs^*\obs \leq m\) and
     \begin{equation*}
@@
     \end{equation*}
-  \item The covariance of the pushforwad $\fwd_{*} \post$ is $\left (
-    (\fwd \prcov \fwd^*)^{-1} + \sigma^{-2} \obs^*\obs \right )^{-1}$
+  \item The covariance of the pushforwad \(\fwd_{*} \post\) is \(\left (
+    (\fwd \prcov \fwd^*)^{-1} + \sigma^{-2} \obs^*\obs \right )^{-1}\)
     and has eigenvalues
@@
 best of my knowledge) a novel lemma in linear algebra: We decompose a
-symmetric positive definite matrix $M \in \mathbb{R}^{k \times k}$
-with $\ttr M = m \in \mathbb{N}$ as $M = AA^t$, where $A$ has unit
+symmetric positive definite matrix \(M \in \mathbb{R}^{k \times k}\)
+with \(\ttr M = m \in \mathbb{N}\) as \(M = AA^t\), where \(A\) has unit
 norm columns.
@@
 %% Finally, in Lemma \ref{lemma:lax} we generalize a lemma for
-%% calculating $\frac{\der}{\der t} \log \det (I + X(t))$, where
-%% $X(t)$ is an operator valued function \cite{Lax07}.
+%% calculating \(\frac{\der}{\der t} \log \det (I + X(t))\), where
+%% \(X(t)\)is an operator valued function \cite{Lax07}.
 

### 2023-08-09 latex/ms.tex
Ready for SIAM applied math?

--- 2023-05-21 latex/paper.tex
@@
 %\documentclass{amsart}
-%\documentclass{article}
 \input{definitions}
-
 %\numberwithin{equation}{section}
 
-\title{Measurement Clusterization in D-optimal Designs for Bayesian Linear
-  Inverse Problems over Hilbert Spaces}
+\title{Measurement Clusterization in D-optimal Designs for Bayesian
+  Linear Inverse Problems over Hilbert Spaces}
 
-\author{Yair Daon\thanks{Porter School of the Environment and Earth
-    Sciences, Tel Aviv University, Tel Aviv, Israel
-    (\email{yair.daon@gmail.com}).}}
-%% \author{Yair Daon\thanks{Azrieli Faculty of Medicine, Bar Ilan
-%%     University, Safed, Israel.  (\email{yair.daon@gmail.com}).}}
+\author{Yair Daon \thanks{Azrieli Faculty of Medicine, Bar Ilan University, Safed,
+      Israel (\email{yair.daon@gmail.com}).}}
 
-%% \address{Porter School of the Environment and Earth
-%%   Sciences, Tel Aviv University\\ Tel Aviv, Israel}
-%% %% \curraddr{Porter School of Environment Studies, Tel Aviv University\\ Tel Aviv, Israel}
-%% %% \curraddr{}
-%% \email{yair.daon@gmail.com}
-% The 2010 edition of the Mathematics Subject Classification is the
-% current definitive version.
+%% The 2010 edition of the Mathematics Subject Classification is the
+%% current definitive version.
 %% \subjclass{Primary: 
@@
 %%   Secondary:
-%%   28C20,
+%%   28C20.
 %%   % Measure and integration - set functions and measures and integrals
@@
 
-\date{\today}
 \begin{document}
 \maketitle
+
 \begin{abstract}
-  Inversion for parameters of physical processes typically requires
-  taking expensive measurements; finding an optimal set of
-  measurements is known as the optimal design problem. Surprisingly,
-  measurement locations in optimal designs are sometimes extremely
-  clustered. Since measurement clusterization is nonintuitive,
-  researchers often avoid it by modifying the optimal design
-  problem. We consider a certain flavor of the optimal design problem,
-  based on the Bayesian D-optimality criterion, and suggest an
-  analytically tractable model for Bayesian D-optimal designs in
-  linear inverse problems over Hilbert spaces. We demonstrate that
-  measurement clusterization is a generic property of D-optimal
-  designs, and prove that correlated noise between measurements
-  mitigates clusterization. We also give a full characterization of
-  D-optimal designs, which shows that D-optimal designs uniformly
-  reduce uncertainty in a select subset of the prior covariance
-  eigenvectors. Finally, we show how measurement clusterization is a
-  simple consequence of the characterization mentioned above and the
-  pigeonhole principle.
+The estimation of parameters in physical processes often demands
+costly measurements, prompting the pursuit of an optimal measurement
+strategy. Finding such strategy is termed the problem of \emph{optimal
+experimental design}, abbreviated as optimal design. Remarkably,
+optimal designs can yield tightly clustered measurement locations,
+leading researchers to fundamentally revise the design problem just to
+circumvent this issue. One approach involves introducing error
+correlation among error terms that are initially independent. In
+another approach, where sensors placement is sought, sensor locations
+are restricted to a finite set of locations. While both approaches may
+prevent clusterization, they also fundamentally alter the optimal
+design problem.
+%
+In this study, we consider Bayesian D-optimal designs, defined as
+designs that maximize the expected Kullback-Leibler divergence between
+posterior and prior. We propose an analytically tractable model for
+D-optimal designs in Bayesian linear inverse problems over Hilbert
+spaces. In this framework, we make several key contributions:
+\textbf{(a)} We establish that measurement clusterization is a generic
+trait of D-optimal designs with independent Gaussian measurement
+errors, and prove that introducing correlations among measurement
+error terms mitigates clusterization. \textbf{(b)} We characterize
+D-optimal designs as reducing uncertainty across a subset of prior
+covariance eigenvectors. Finally, \textbf{(c)} We leverage this
+characterization to argue that measurement clusterization arises as a
+consequence of the pigeonhole principle: when more measurements are
+taken than there are locations where the select eigenvectors are large
+and others are small --- clusterization occurs.
+%
+%% In summary, in this study we shed light on an often ignored issue with
+%% Bayesian D-optimal designs. We characterize such designs and show why
+%% clusterization arisesit arisestheir inherent properties, propose
+%% strategies to address clusterization, and provide insights into the
+%% fundamental attributes underlying measurement design optimization.
 \end{abstract}
-
-\maketitle
 

### 2023-08-09 latex/prelims.tex
Ready for SIAM applied math?

@@
 \subsection{Bayesian Linear Inverse Problems}\label{subsec:abstract OED}
-Let $\hilp$ and $\hilo$ be separable Hilbert spaces,\footnote{The
-subscripts p and o are for ``parameter'' and ``observation'',
-respectively.} and let $\fwd: \hilp \to \hilo$ the \emph{forward
-operator}. The forward operator $\fwd$ is assumed linear and strongly
-smoothing (the heat operator of the 1D heat equation introduced in
-Section \ref{section:intro} is a prime example). Take a Gaussian prior
-$\param \sim \pr = \normal(\prmean ,\prcov)$ with some appropriate
-covariance operator $\prcov$ on $\hilp$ \cite{Stuart10}. Note that
-$\fwd \prcov \fwd^*$ is the prior covariance in $\hilo$
-\cite{Stuart10}, and as such, $\fwd \prcov \fwd^*$ is assumed
-invertible --- an assumption which will be used below\footnote{If
-$\fwd$ has a nontrivial kernel we utilize Occam's Razor and ignore
-said kernel.}. Measurements are taken via the \emph{measurement
-operator}\footnote{It is common for the measurement and forward
-operators to be merged $\tmp := \obs \fwd$
+Let \(\hilp\) and \(\hilo\) be separable Hilbert spaces (the
+subscripts p and o are for "parameter" and "observation",
+respectively), and let \(\fwd: \hilp \to \hilo\) the \emph{forward
+operator}. The forward operator \(\fwd\) is assumed linear and
+strongly smoothing (the heat operator of the 1D heat equation
+introduced in Section \ref{section:intro} is a prime example). Take a
+Gaussian prior \(\param \sim \pr = \normal(\prmean ,\prcov)\) with
+some appropriate covariance operator \(\prcov\) on \(\hilp\)
+\cite{Stuart10}. Note that \(\fwd \prcov \fwd^*\) is the prior
+covariance in \(\hilo\) \cite{Stuart10}, and as such, \(\fwd \prcov
+\fwd^*\) is assumed invertible --- an assumption which will be used
+below (if \(\fwd\) has a nontrivial kernel we utilize Occam's Razor
+and ignore said kernel altogether). Measurements are taken via the
+\emph{measurement operator}. It is common for the measurement and
+forward operators to be merged \(\tmp := \obs \fwd\)
 \cite{AlexanderianGloorGhattas14}, but the analysis carried out in the
-following sections requires that $\fwd$ and $\obs$ are explicitly
-separated as in \cite{attia2022stochastic, cvetkovic2023choosing}.}
-$\obs \in ( \hilo^* )^m$, where $m$ is the number of measurements
-taken. Entries $\meas_j, j=1,\dots,m$ of the observation operator
-$\obs$ are called \emph{measurements}:
+following sections requires that \(\fwd\) and \(\obs\) are explicitly
+separated as in \cite{attia2022stochastic, cvetkovic2023choosing}.
+\(\obs \in ( \hilo^* )^m\), where \(m\) is the number of measurements
+taken. Entries \(\meas_j, j=1,\dots,m\) of the observation operator
+\(\obs\) are called \emph{measurements}:
 
@@
 Data is acquired via noisy observations, and we consider two types of
-error terms: Spatially correlated model error $\eps' \sim
-\normal(0,\modcov)$ with $\modcov$ a covariance operator. Observation
-error is denoted $\eps \sim \normal(0, \sigma^2 I_m)$, with $I_m \in
-\mathbb{R}^{m \times m}$ the identity. Both error terms and the prior
+error terms: Spatially correlated model error \(\eps' \sim
+\normal(0,\modcov)\) with \(\modcov\) a covariance operator. Observation
+error is denoted \(\eps \sim \normal(0, \sigma^2 I_m)\), with \(I_m \in
+\mathbb{R}^{m \times m}\) the identity. Both error terms and the prior
 are assumed independent of each other. Thus, data is acquired via
@@
 
-It is easy to verify that $\obs \eps' + \eps \in \R^m$ is a centered
+It is easy to verify that \(\obs \eps' + \eps \in \R^m\) is a centered
 Gaussian random vector with covariance matrix
@@
 \end{align}
-Taking $\modcov = 0$ is a common practice
-\cite{tarantola2005,Kaipio2005,Vogel02} and then $\Sigma =
-\sigma^2I_m$ is a scalar matrix which does not depend on $\obs$.
+Taking \(\modcov = 0\) is a common practice
+\cite{tarantola2005,Kaipio2005,Vogel02} and then \(\Sigma =
+\sigma^2I_m\) is a scalar matrix which does not depend on \(\obs\).
 
-It is useful to record that the posterior measure $\post$ is Gaussian
-in this setting and its covariance operator is \cite{Stuart10}:
+Finally, it is useful to record that the posterior measure \(\post\) is
+Gaussian in this setting and its covariance operator is does not
+depend on data \(\data\) \cite{Stuart10}:
 \begin{align}\label{eq:postcov}
@@
 A Bayesian D-optimal design maximizes expected Kullback-Leibler (KL)
-divergence between posterior $\post$ and prior measures $\pr$. It is
+divergence between posterior \(\post\) and prior measures \(\pr\). It is
 useful to first recall the definition of KL divergence \cite{CoverThomas91}:
-$$
+\begin{equation*}
   D_{KL}(\post||\pr) = \int \log \frac{\der \post}{\der \pr}(\param) \der \post(\param).
-$$
+\end{equation*}
 
@@
   alexanderian2018efficient}. The main result we will make use of is
-summarized (in our notation) in Theorem \ref{thm:d optimality}\footnote{In
-\cite{AlexanderianGloorGhattas14, alexanderian2018efficient}, results
-are stated for $\Sigma=I$ (implied by $\modcov = 0,\sigma^2=1$), but
-the authors also mention these results hold for more general
-covariance matrices \cite[p. 681]{AlexanderianGloorGhattas14}.}:
+summarized in our notation below:
 
 \begin{theorem}[Alexanderian, Gloor, Ghattas \cite{AlexanderianGloorGhattas14}]\label{thm:d optimality}
-  Let $\pr = \normal(\prmean,\prcov)$ be a Gaussian prior on $\hilp$
-  and let $\post = \normal(\postmean,\postcov)$ the posterior measure
-  on $\hilp$ for the Bayesian linear inverse problem $\data = \obs
-  \fwd\param + \obs \eps' + \eps$ discussed above. Then
+  Let \(\pr = \normal(\prmean,\prcov)\) be a Gaussian prior on \(\hilp\)
+  and let \(\post = \normal(\postmean,\postcov)\) the posterior measure
+  on \(\hilp\) for the Bayesian linear inverse problem \(\data = \obs
+  \fwd\param + \obs \eps' + \eps\) discussed above. Then
   \begin{align}\label{eq:objective}
@@
 
-%% It is important to note that since $\obs$ is finite-rank,
-%% $\prcov^{1/2} \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}$ is
+Note that in \cite{AlexanderianGloorGhattas14,
+  alexanderian2018efficient}, results are stated for \(\Sigma=I\)
+(implied by \(\modcov = 0,\sigma^2=1\)), but the these results also hold
+for more general covariance matrices
+\cite[p. 681]{AlexanderianGloorGhattas14}.
+
+%% It is important to note that since \(\obs\) is finite-rank,
+%% \(\prcov^{1/2} \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}\) is
 %% trace-class.
 \begin{definition}\label{def:d optimality}
-  $\obs^{\star}$ is said to be \emph{D-optimal} if $\obs^{\star} =
-  \argmax_{\obs} \tar(\obs)$, where entries of $\obs \in (\hilo^*)^m$
-  are constrained to some allowed set of measurements in $\hilo^*$.
+  We say \(\obs^{\star}\) is \emph{D-optimal} if \(\obs^{\star} =
+  \argmax_{\obs} \tar(\obs)\), where entries of \(\obs \in (\hilo^*)^m\)
+  are constrained to some allowed set of measurements in \(\hilo^*\).
 \end{definition}
@@
 \end{align*}
-We think of $\prcov$ as constant, so a D-optimal design minimizes the
-posterior covariance determinant, analogously to the
-finite-dimensional case.
+We think of \(\prcov\) as constant, so a D-optimal design minimizes a
+quantity analogous to the posterior covariance determinant, similarly
+to the finite-dimensional case.
 
@@
 %%   \begin{itemize}
-%%   \item $\hilp, \hilo$ Hilbert spaces.
-%%   \item $\fwd:\hilp \to \hilo$ a linear compact operator.
-%%   \item $\pr \sim \mathcal{N}(0, \prcov)$ prior Gaussian measure on
-%%     $\hilp$, with prior covariance operator $\prcov:\hilp \to \hilp$.
-%%   \item $\obs: \hilo \to \mathbb{R}^m$ measurement operator, where $m
-%%     \in \mathbb{N}$ is the number of measurements taken.
-%%   \item $\sigma^2 \in \mathbb{R}_{+}$ observation noise variance.
-%%   \item $\modcov$ model error covariance operator.
-%%     %% $\data = \obs \fwd \param + \eps$, where $\eps \in
-%%     %% \mathbb{R}^m$ isiid $\mathcal{N}(0, \sigma^2)$ noise.
-%%   \item $\Sigma(\obs) = \obs \modcov \obs^* + \sigma^2I$. 
-%%   \item $\post$ the posterior measure, with covariance $\postcov$.
+%%   \item \(\hilp, \hilo\) Hilbert spaces.
+%%   \item \(\fwd:\hilp \to \hilo\) a linear compact operator.
+%%   \item \(\pr \sim \mathcal{N}(0, \prcov)\) prior Gaussian measure on
+%%     \(\hilp$, with prior covariance operator \(\prcov:\hilp \to \hilp$.
+%%   \item \(\obs: \hilo \to \mathbb{R}^m\) measurement operator, where \(m
+%%     \in \mathbb{N}\) is the number of measurements taken.
+%%   \item \(\sigma^2 \in \mathbb{R}_{+}\) observation noise variance.
+%%   \item \(\modcov\) model error covariance operator.
+%%     %% \(\data = \obs \fwd \param + \eps$, where \(\eps \in
+%%     %% \mathbb{R}^m\) isiid \(\mathcal{N}(0, \sigma^2)\) noise.
+%%   \item \(\Sigma(\obs) = \obs \modcov \obs^* + \sigma^2I$. 
+%%   \item \(\post\) the posterior measure, with covariance \(\postcov$.
 %%   \item A D-optimality design criterion
@@
 %%     \end{align*}
-%%   %% \item $\{\lambda_i\}_{i=1}^\infty$ eigenvalues of $\fwd\prcov\fwd^*$
+%%   %% \item \(\{\lambda_i\}_{i=1}^\infty\) eigenvalues of \(\fwd\prcov\fwd^*$
 %%   %%   in decreasing order of magnitude.
-%%   %% %% \item $\{\ev_i\}_{i=1}^\infty$ their corresponding eigenvectors.
-%%   %% \item $\{\eta_i\}_{i=1}^\infty$ eigenvalues of $\obs^*\obs$.
+%%   %% %% \item \(\{\ev_i\}_{i=1}^\infty\) their corresponding eigenvectors.
+%%   %% \item \(\{\eta_i\}_{i=1}^\infty\) eigenvalues of \(\obs^*\obs$.
 %%   \end{itemize}
@@
 %% From defintion \ref{def:d optimality} we wish to characterize solution(s) of the
-%% following optimization problem for $\tar$. %%: (\hilo^*)^m \to \R$:
+%% following optimization problem for \(\tar$. %%: (\hilo^*)^m \to \R$:
 %% \begin{align}\label{eq:optimization}
@@
 %% \end{align}
-%% where $\obs$ is constrained to some allowed set of observations. We
+%% where \(\obs\) is constrained to some allowed set of observations. We
 %% call this problem ``simultaneous optimization'', since all
@@
 %% \end{equation}
-%% Sequential optimal design proceeds as follows. Find $\meas_1$ by
-%% maximizing $\tar(\obs_1)$. Then, keeping $\meas_1$ fixed --- find
-%% $\meas_2$ as the maximizer of $\tar(\obs_2)$. Then, find $\meas_3$ by
-%% keeping $\meas_1,\meas_2$ fixed and taking $\meas_3$ as the maximizer
-%% of $\tar(\obs_3)$. Continue this way until $\obs_m = \obs$ is
-%% found, where $m$ is the number of available observations. %% It is
+%% Sequential optimal design proceeds as follows. Find \(\meas_1\) by
+%% maximizing \(\tar(\obs_1)$. Then, keeping \(\meas_1\) fixed --- find
+%% \(\meas_2\) as the maximizer of \(\tar(\obs_2)$. Then, find \(\meas_3\) by
+%% keeping \(\meas_1,\meas_2\) fixed and taking \(\meas_3\) as the maximizer
+%% of \(\tar(\obs_3)$. Continue this way until \(\obs_m = \obs\) is
+%% found, where \(m\) is the number of available observations. %% It is
 %% %% important to notice that this scheme does not require actually

### 2023-08-09 latex/vanishing.tex
Ready for SIAM applied math?

@@
-\section{Characterizing D-Optimal Designs when $\modcov = 0$}\label{section:vanishing}
-We will now prove Theorem \ref{thm:char} which characterizes D-optimal
-designs when $\modcov = 0$. The necessary first-order condition for
-D-optimality of Theorem \ref{thm:constrained} for $\modcov = 0$
-become:
+\section{D-Optimal Designs Without Model Error}\label{section:vanishing}
+Our goal in this section is to prove Theorem \ref{thm:char} which
+characterizes D-optimal designs when $\modcov = 0$. The necessary
+first-order condition for D-optimality of Theorem
+\ref{thm:constrained} for $\modcov = 0$ become:
 
@@
 \begin{figure}
+  \centering
   \begin{tikzpicture}[scale=0.85]
@@
         legend plot pos=right,
-        ylabel=precision --- prior and posterior,
-        xlabel=eigenvector $i$,
+        ylabel={\LARGE precision --- prior and posterior},
+        xlabel={\LARGE eigenvector},
       ]
-
       \addplot [fill=green!80]  table [y=prior,   meta=Label, x expr=\coordindex] {\optimalvsnot};
       \addplot [fill=blue!60]   table [y=optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};     
-
-      \addlegendentry{$\sigma^2\lambda_i^{-1}$}
-      \addlegendentry{optimal $\eta_i$s}
-
+      \addlegendentry[scale=1.4]{$\sigma^2\lambda_i^{-1}$}
+      \addlegendentry[scale=1.4]{optimal $\eta_i$s}
     \end{axis}
   \end{tikzpicture}
-  \qquad
   \begin{tikzpicture}[scale=0.85]
@@
         legend plot pos=right,
-        ylabel=precision --- prior and posterior,
-        xlabel=eigenvector $i$,
+        ylabel={\LARGE precision --- prior and posterior},
+        xlabel={\LARGE eigenvector} ,
       ]   
-
       \addplot [fill=green!80]  table [y=prior,       meta=Label, x expr=\coordindex] {\optimalvsnot};
       \addplot [fill=blue!60]   table [y=sub-optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};
-
-      \addlegendentry{$\sigma^2\lambda_i^{-1}$}
-      \addlegendentry{sub-optimal $\eta_i$s}
-
+      \addlegendentry[scale=1.4]{$\sigma^2\lambda_i^{-1}$}
+      \addlegendentry[scale=1.4]{sub-optimal $\eta_i$s}
     \end{axis}

### 2023-08-09 latex/why.tex
Ready for SIAM applied math?

@@
-\section{An answer for Question \ref{q:why}: The cause of measurement clusterization}\label{section:why}
-\begin{figure}[H]
+\section{Answer to Question \ref{q:why}: The cause of clusterization}\label{section:why}
+\begin{figure}
     \centering
-    \includegraphics[height=0.5\textwidth]{eigenvectors.eps}
+    \includegraphics[height=0.5\textwidth]{eigenvectors.pdf}
     \caption{D-optimal measurement locations ($m=5$ measurements) and

### 2023-08-10 latex/intro.tex
Submitted to SIAP

@@
 Measurements play a fundamental role in generating observations and
-are indispensable to any inference process. In the realm of data
-science, selecting the optimal set of measurements holds particular
+are indispensable to any inference process. In the practice of inverse
+problems, selecting the optimal set of measurements holds particular
 significance when inferring parameters of a physical process. Unlike
@@
 tomography \cite{horesh2010impedance}, determining certain wavelengths
-in MRI \cite{horesh2008mri}, or positioning sensors for detecting
+in MRI \cite{horesh2008mri}, positioning sensors for detecting
 ground-reflected waves in the search for oil
-\cite{horesh2008borehole}, the selection of optimal measurements,
-referred to as the problem of \emph{optimal design}, becomes
-crucial. Optimal measurements are chosen according to specific
-\emph{design criteria}, with A- and D-optimality being two of the most
-widely recognized and extensively studied design criteria
-\cite{Chaloner1995}.
+\cite{horesh2008borehole}, or optimizing sensor locations in a
+seismographic network \cite{rabinowitz1990, steinberg1995} --- the
+selection of optimal measurements, referred to as the problem of
+\emph{optimal design}, becomes crucial. Optimal measurements are
+chosen according to specific \emph{design criteria}, with A- and
+D-optimality being two of the most widely recognized and extensively
+studied design criteria \cite{Chaloner1995}.
 
@@
 the initial condition of the 1D heat equation (details in the
-supplementary material). In Figure \ref{fig:clusterization
+supplementary material). In Fig.~\ref{fig:clusterization
   illustration} D-optimal measurement locations are shown for
@@
 In spatial problems involving choice of measurements within a domain
-\(\Omega \subseteq \mathbb{R}^d, d=1,2,3\), many researchers circumvent
-the problem of measurement clusterization by choosing measurements
-from a coarse grid in \(\Omega\) \cite{koval2020, alexanderian2021,
-  attia2022, alexanderian2014, alexanderian2016,
+\(\Omega \subseteq \mathbb{R}^d, d=1,2,3\), many researchers
+circumvent the problem of measurement clusterization by choosing
+measurements from a coarse grid in \(\Omega\) \cite{koval2020,
+  alexanderian2021, attia2022, alexanderian2014, alexanderian2016,
   alexanderian2018efficient, brunton2016}. This approach incurs a
@@
 the potential measurement locations. Some researchers incorporate a
-sparsifying \(\ell_1\) penalty term into the design criterion, which is
-subsequently thresholded to achieve the desired binary design over the
-coarse grid \cite{horesh2008borehole}. Others progressively relax the
-\(\ell_1\) penalty to an \(\ell_0\) penalty via a continuation method
-\cite{alexanderian2016, alexanderian2014}. A clever alternative is to
-cast the problem of finding optimal measurement weights as a
-stochastic optimization problem \cite{attia2022stochastic}. All of the
+sparsifying \(\ell_1\) penalty term into the design criterion, which
+is subsequently thresholded to achieve the desired binary design over
+the coarse grid \cite{horesh2008borehole}. Others progressively relax
+the \(\ell_1\) penalty to an \(\ell_0\) penalty via a continuation
+method \cite{alexanderian2016, alexanderian2014}. Others cast the
+problem of finding optimal measurement weights as a stochastic
+optimization problem \cite{attia2022stochastic}. All of the
 aforementioned methods may indeed find a binary optimal design
@@
   the relevant prior eigenvectors are significantly large, and (b)
-  other eigenvectors have a power spectrum close to zero, is
-  limited. Consequently, the clusterization of measurements arises as
-  a natural consequence of the pigeonhole principle, as there are more
-  measurements available than there are locations satisfying
-  conditions (a) and (b).
+  other eigenvectors are close to zero, is limited. Consequently, the
+  clusterization of measurements arises as a natural consequence of
+  the pigeonhole principle, as there are more measurements available
+  than there are locations satisfying conditions (a) and (b).
 
@@
 
-We first give some general definitions before we state Theorem
-\ref{thm:char}: Let \(\hilp, \hilo\) Hilbert spaces, \(\fwd:\hilp \to
-\hilo\) a linear compact operator. Let \(\obs: \hilo \to \mathbb{R}^m\) a
-linear measurement operator, where \(m \in \mathbb{N}\) is the number of
+Before we state Theorem \ref{thm:char} we give some definitions: Let
+\(\hilp, \hilo\) Hilbert spaces, \(\fwd:\hilp \to \hilo\) a linear
+compact operator. Let \(\obs: \hilo \to \mathbb{R}^m\) a linear
+measurement operator, where \(m \in \mathbb{N}\) is the number of
 measurements taken. Let \(\sigma^2 \in \mathbb{R}_{+}\) observation
 noise variance, \(\data = \obs \fwd \param + \eps\), where \(\eps \in
-\mathbb{R}^m\) is iid \(\mathcal{N}(0, \sigma^2)\) noise. Let \(\pr \sim
-\mathcal{N}(\prmean, \prcov)\) prior Gaussian measure on \(\hilp\), where
-\(\prcov\) is the prior covariance operator and let \(\post\) the
-posterior (Gaussian) measure.
+\mathbb{R}^m\) is iid \(\mathcal{N}(0, \sigma^2)\) noise. Let \(\pr
+\sim \mathcal{N}(\prmean, \prcov)\) prior Gaussian measure on
+\(\hilp\), where \(\prcov\) is the prior covariance operator and let
+\(\post\) the posterior (Gaussian) measure.
 
@@
     \end{equation*}
-  \item The covariance of the pushforwad \(\fwd_{*} \post\) is \(\left (
-    (\fwd \prcov \fwd^*)^{-1} + \sigma^{-2} \obs^*\obs \right )^{-1}\)
-    and has eigenvalues
+  \item The covariance of the pushforwad \(\fwd_{*} \post\) is \(\left
+    ( (\fwd \prcov \fwd^*)^{-1} + \sigma^{-2} \obs^*\obs \right
+    )^{-1}\) and its eigenvalues are
     \begin{equation*}
@@
 
-In the proving Theorem \ref{thm:char} we prove and generalize several
-lemmas. Among those, is a Lemma \ref{lemma:free}, which is (to the the
-best of my knowledge) a novel lemma in linear algebra: We decompose a
-symmetric positive definite matrix \(M \in \mathbb{R}^{k \times k}\)
-with \(\ttr M = m \in \mathbb{N}\) as \(M = AA^t\), where \(A\) has unit
-norm columns.
+In the process of proving Theorem \ref{thm:char} we prove and
+generalize several lemmas. Among those, is Lemma \ref{lemma:free},
+which is (to the the best of my knowledge) a novel lemma in linear
+algebra: We decompose a symmetric positive definite matrix \(M \in
+\mathbb{R}^{k \times k}\) with \(\ttr M = m \in \mathbb{N}\) as \(M =
+AA^t\), where \(A\) has unit norm columns.
 

### 2023-08-10 latex/ms.tex
Submitted to SIAP

@@
-\documentclass{siamonline220329}
+\documentclass{siamart220329}
 %\documentclass{amsart}
@@
 %\numberwithin{equation}{section}
-
-\title{Measurement Clusterization in D-optimal Designs for Bayesian
-  Linear Inverse Problems over Hilbert Spaces}
-
-\author{Yair Daon \thanks{Azrieli Faculty of Medicine, Bar Ilan University, Safed,
-      Israel (\email{yair.daon@gmail.com}).}}
 
@@
 \begin{abstract}
-The estimation of parameters in physical processes often demands
-costly measurements, prompting the pursuit of an optimal measurement
-strategy. Finding such strategy is termed the problem of \emph{optimal
-experimental design}, abbreviated as optimal design. Remarkably,
-optimal designs can yield tightly clustered measurement locations,
-leading researchers to fundamentally revise the design problem just to
-circumvent this issue. One approach involves introducing error
-correlation among error terms that are initially independent. In
-another approach, where sensors placement is sought, sensor locations
-are restricted to a finite set of locations. While both approaches may
-prevent clusterization, they also fundamentally alter the optimal
-design problem.
-%
-In this study, we consider Bayesian D-optimal designs, defined as
-designs that maximize the expected Kullback-Leibler divergence between
-posterior and prior. We propose an analytically tractable model for
-D-optimal designs in Bayesian linear inverse problems over Hilbert
-spaces. In this framework, we make several key contributions:
-\textbf{(a)} We establish that measurement clusterization is a generic
-trait of D-optimal designs with independent Gaussian measurement
-errors, and prove that introducing correlations among measurement
-error terms mitigates clusterization. \textbf{(b)} We characterize
-D-optimal designs as reducing uncertainty across a subset of prior
-covariance eigenvectors. Finally, \textbf{(c)} We leverage this
-characterization to argue that measurement clusterization arises as a
-consequence of the pigeonhole principle: when more measurements are
-taken than there are locations where the select eigenvectors are large
-and others are small --- clusterization occurs.
+Estimation of parameters in physical processes often demands costly measurements, prompting the pursuit of an optimal measurement strategy.
+Finding such strategy is termed the problem of \emph{optimal experimental design}, abbreviated as optimal design.
+Remarkably, optimal designs can yield tightly clustered measurement locations, leading researchers to fundamentally revise the design problem just to circumvent this issue.
+Some authors introduce error correlation among error terms that are initially independent, while others restrict measurement locations to a finite set of locations.
+While both approaches may prevent clusterization, they also fundamentally alter the optimal design problem.
+
+In this study, we consider Bayesian D-optimal designs, i.e.~designs that maximize the expected Kullback-Leibler divergence between posterior and prior.
+We propose an analytically tractable model for D-optimal designs in Bayesian linear inverse problems over Hilbert spaces.
+In this framework, we make several key contributions: \textbf{(a)} We establish that measurement clusterization is a generic trait of D-optimal designs with independent Gaussian measurement errors, and prove that introducing correlations among measurement error terms mitigates clusterization.
+\textbf{(b)} We characterize D-optimal designs as reducing uncertainty across a subset of prior covariance eigenvectors.
+Finally, \textbf{(c)} We leverage this characterization to argue that measurement clusterization arises as a consequence of the pigeonhole principle: when more measurements are taken than there are locations where the select eigenvectors are large and others are small --- clusterization occurs.
 %
@@
 %% \input{appendix}
-
-%\bibliographystyle{siamplain}
-\bibliographystyle{amsplain}
+\bibliographystyle{siamplain}
+%\bibliographystyle{amsplain}
 \bibliography{/home/yair/projects/bibtex.bib}

### 2023-08-10 latex/prelims.tex
Submitted to SIAP

@@
 article. The theoretical foundations for inverse problems over
-function spaces can be found in \cite{Stuart10} and will not be
+function spaces can be found elsewhere \cite{Stuart10} and will not be
 reviewed here.
@@
 \subsection{Bayesian Linear Inverse Problems}\label{subsec:abstract OED}
-Let \(\hilp\) and \(\hilo\) be separable Hilbert spaces (the
-subscripts p and o are for "parameter" and "observation",
-respectively), and let \(\fwd: \hilp \to \hilo\) the \emph{forward
-operator}. The forward operator \(\fwd\) is assumed linear and
-strongly smoothing (the heat operator of the 1D heat equation
-introduced in Section \ref{section:intro} is a prime example). Take a
-Gaussian prior \(\param \sim \pr = \normal(\prmean ,\prcov)\) with
-some appropriate covariance operator \(\prcov\) on \(\hilp\)
-\cite{Stuart10}. Note that \(\fwd \prcov \fwd^*\) is the prior
-covariance in \(\hilo\) \cite{Stuart10}, and as such, \(\fwd \prcov
-\fwd^*\) is assumed invertible --- an assumption which will be used
-below (if \(\fwd\) has a nontrivial kernel we utilize Occam's Razor
-and ignore said kernel altogether). Measurements are taken via the
-\emph{measurement operator}. It is common for the measurement and
-forward operators to be merged \(\tmp := \obs \fwd\)
-\cite{AlexanderianGloorGhattas14}, but the analysis carried out in the
-following sections requires that \(\fwd\) and \(\obs\) are explicitly
-separated as in \cite{attia2022stochastic, cvetkovic2023choosing}.
-\(\obs \in ( \hilo^* )^m\), where \(m\) is the number of measurements
+Let \(\hilp\) and \(\hilo\) separable Hilbert spaces (the subscripts p
+and o are for "parameter" and "observation", respectively), and let
+\(\fwd: \hilp \to \hilo\) the \emph{forward operator}. The forward
+operator \(\fwd\) is assumed linear and strongly smoothing (the heat
+operator of the 1D heat equation introduced in Section
+\ref{section:intro} is a prime example). Take a Gaussian prior
+\(\param \sim \pr = \normal(\prmean ,\prcov)\) with some appropriate
+covariance operator \(\prcov\) on \(\hilp\) \cite{Stuart10}. Note that
+\(\fwd \prcov \fwd^*\) is the prior covariance in \(\hilo\)
+\cite{Stuart10}, and as such, \(\fwd \prcov \fwd^*\) is assumed
+invertible --- an assumption which will be used below (if \(\fwd\) has
+a nontrivial kernel we utilize Occam's Razor and ignore said kernel
+altogether). Measurements are taken via a \emph{measurement operator}
+\(\obs\). It is common for the measurement and forward operators to be
+merged \(\tmp := \obs \fwd\) \cite{AlexanderianGloorGhattas14}, but
+the analysis carried out in the following sections requires that
+\(\fwd\) and \(\obs\) are explicitly separated as in
+\cite{attia2022stochastic, cvetkovic2023choosing}.  \(\obs \in (
+\hilo^* )^m\), where \(m\) is the number of measurements
 taken. Entries \(\meas_j, j=1,\dots,m\) of the observation operator
@@
 
-Finally, it is useful to record that the posterior measure \(\post\) is
-Gaussian in this setting and its covariance operator is does not
+Finally, it is useful to record that the posterior measure \(\post\)
+is Gaussian in this setting and its covariance operator does not
 depend on data \(\data\) \cite{Stuart10}:
@@
   alexanderian2018efficient}. The main result we will make use of is
-summarized in our notation below:
+summarized (in our notation) below:
 
@@
   alexanderian2018efficient}, results are stated for \(\Sigma=I\)
-(implied by \(\modcov = 0,\sigma^2=1\)), but the these results also hold
+(implied by \(\modcov = 0,\sigma^2=1\)), but these results also hold
 for more general covariance matrices
@@
 
-For a Bayesian linear model in finite dimensions, with Gaussian prior
-and Gaussian noise, a D-optimal design minimizes the determinant of
-the posterior covariance matrix, and this turns out to be a
-regularized version of the frequentist criterion
+Intuition for Theorem \ref{thm:d optimality} can be gained by
+considering a Bayesian linear model in finite dimensions, with
+Gaussian prior and Gaussian noise. Then, a D-optimal design minimizes
+the determinant of the posterior covariance matrix, and this turns out
+to be a regularized version of the frequentist D-optimality criterion
 \cite{Chaloner1995}. Theorem \ref{thm:d optimality} and Definition

### 2023-08-10 latex/vanishing.tex
Submitted to SIAP

@@
 \postcov \fwd^*$, where rows of $\obs$, namely $\meas_j,j=1,\dots, m$,
-are eigenvectors. However, $\postcov$ depends on $\obs$ and thus, we
-refer to \eqref{eq:eigenproblem} as a \emph{nonlinear} eigenvalue
-problem.%%  Proposition \ref{prop:twice woodbury} and Lemma
-%% \ref{lemma:sim diag} required
+are eigenvectors. However, $\postcov$ depends on $\obs$, so we refer
+to \eqref{eq:eigenproblem} as a \emph{nonlinear} eigenvalue problem.
+%% Proposition \ref{prop:twice woodbury} and Lemma \ref{lemma:sim
+%% diag} required
 
@@
 \begin{proof}
-  The proof amounts to using Woodbury's matrix identity twice and a
+  The proof amounts to using Woodbury's matrix identity twice with a
   regularization trick. The standard proof for Woodbury's matrix
@@
 
-  Then, by \eqref{eq:WMI}, with the notation \eqref{eq:notation}:
-  \begin{equation}\label{eq:first}
-  \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* = \fwd ( \prcov - \prcov \fwd^* ( \sigma^2(\obs^*\obs + \eps I)^{-1} + \fwd \prcov \fwd^* )^{-1} \fwd \prcov ) \fwd^* \\
+  Then, \eqref{eq:WMI} with the notation \eqref{eq:notation} implies:
+  \begin{align}\label{eq:first}
+    \begin{split}
+    &\fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* \\
+    &\ \ \ \ \ = \fwd ( \prcov - \prcov \fwd^* ( \sigma^2(\obs^*\obs + \eps I)^{-1} + \fwd \prcov \fwd^* )^{-1} \fwd \prcov ) \fwd^*. \\
       %
@@
     %% \end{align*}
-  \end{equation}
+    \end{split}
+  \end{align}
   Now, denote:
   \begin{align*}
-    X :&= \fwd\prcov \fwd^* \\
+    X :&= \fwd\prcov \fwd^*, \\
     Y_{\eps} :&= \obs^*\obs + \eps I,
@@
   \centering
-  \begin{tikzpicture}[scale=0.85]
+  \begin{tikzpicture}[scale=0.75]
     \begin{axis}[
@@
       ]
-      \addplot [fill=green!80]  table [y=prior,   meta=Label, x expr=\coordindex] {\optimalvsnot};
-      \addplot [fill=blue!60]   table [y=optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};     
+      \addplot [fill=blue!60]  table [y=prior,   meta=Label, x expr=\coordindex] {\optimalvsnot};
+      \addplot [pattern=north east lines, pattern color=green!80]  table [y=optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};     
       \addlegendentry[scale=1.4]{$\sigma^2\lambda_i^{-1}$}
@@
   \end{tikzpicture}
-  \begin{tikzpicture}[scale=0.85]
+  \begin{tikzpicture}[scale=0.75]
     \begin{axis}[
@@
       ]   
-      \addplot [fill=green!80]  table [y=prior,       meta=Label, x expr=\coordindex] {\optimalvsnot};
-      \addplot [fill=blue!60]   table [y=sub-optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};
+      \addplot [fill=blue!60]  table [y=prior,       meta=Label, x expr=\coordindex] {\optimalvsnot};
+      \addplot [pattern=north east lines, pattern color=green!80]  table [y=sub-optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};
       \addlegendentry[scale=1.4]{$\sigma^2\lambda_i^{-1}$}
@@
 level of liquid in it, as long as its index $i \leq m$. The result of
-such a procedure is illustrated in Figure \ref{fig:optimal vs not}.
+such a procedure is illustrated in Fig.~\ref{fig:optimal vs not}.
 %% \begin{figure}%{r}{0.25\textwidth} 

### 2023-08-10 latex/why.tex
Submitted to SIAP

@@
 in the prior spectrum, for some $k > 0$. To illustrate this
-preference, Figure \ref{fig:why} depicts the scenario using the 1D
+preference, Fig.~\ref{fig:why} depicts the scenario using the 1D
 heat equation with homogeneous Dirichlet boundary conditions (details

### 2023-08-29 latex/ack.tex
Submitted to M3AS

--- 2023-05-21 latex/paper.tex
@@
-\documentclass{siamonline220329}
-%\documentclass{amsart}
-%\documentclass{article}
-\input{definitions}
-
-%\numberwithin{equation}{section}
-
-\title{Measurement Clusterization in D-optimal Designs for Bayesian Linear
-  Inverse Problems over Hilbert Spaces}
-
-\author{Yair Daon\thanks{Porter School of the Environment and Earth
-    Sciences, Tel Aviv University, Tel Aviv, Israel
-    (\email{yair.daon@gmail.com}).}}
-%% \author{Yair Daon\thanks{Azrieli Faculty of Medicine, Bar Ilan
-%%     University, Safed, Israel.  (\email{yair.daon@gmail.com}).}}
-
-%% \address{Porter School of the Environment and Earth
-%%   Sciences, Tel Aviv University\\ Tel Aviv, Israel}
-%% %% \curraddr{Porter School of Environment Studies, Tel Aviv University\\ Tel Aviv, Israel}
-%% %% \curraddr{}
-%% \email{yair.daon@gmail.com}
-% The 2010 edition of the Mathematics Subject Classification is the
-% current definitive version.
-%% \subjclass{Primary: 
-%%   62F15, % Statistics - bayesian inference
-%%   35R30, % PDE - Inverse problems
-%%   Secondary:
-%%   28C20,
-%%   % Measure and integration - set functions and measures and integrals
-%%          % in infinite-dimensional spaces (Wiener measure, Gaussian
-%%   % measure, etc.
-%% }
-
-\date{\today}
-\begin{document}
-\maketitle
-\begin{abstract}
-  Inversion for parameters of physical processes typically requires
-  taking expensive measurements; finding an optimal set of
-  measurements is known as the optimal design problem. Surprisingly,
-  measurement locations in optimal designs are sometimes extremely
-  clustered. Since measurement clusterization is nonintuitive,
-  researchers often avoid it by modifying the optimal design
-  problem. We consider a certain flavor of the optimal design problem,
-  based on the Bayesian D-optimality criterion, and suggest an
-  analytically tractable model for Bayesian D-optimal designs in
-  linear inverse problems over Hilbert spaces. We demonstrate that
-  measurement clusterization is a generic property of D-optimal
-  designs, and prove that correlated noise between measurements
-  mitigates clusterization. We also give a full characterization of
-  D-optimal designs, which shows that D-optimal designs uniformly
-  reduce uncertainty in a select subset of the prior covariance
-  eigenvectors. Finally, we show how measurement clusterization is a
-  simple consequence of the characterization mentioned above and the
-  pigeonhole principle.
-\end{abstract}
-
-\maketitle
-
-\input{intro}
-\input{prelims}
-\input{constrained}
-\input{nonvanishing}
-\input{vanishing}
-\input{why}
-%% \input{how}
-%% \section{Open Problems}
-
-\section{Acknowledgements}
 This study is a result of research I started during my PhD studies
@@
 \href{https://mathoverflow.net/questions/280168/redistribute-diagonal-entries-of-a-matrix/280203#280203c}{Mathoverflow}.
-
-This research was supported in part by an appointment with the
-National Science Foundation (NSF) Mathematical Sciences Graduate
-Internship (MSGI) Program sponsored by the NSF Division of
-Mathematical Sciences. This program is administered by the Oak Ridge
-Institute for Science and Education (ORISE) through an interagency
-agreement between the U.S. Department of Energy (DOE) and NSF. ORISE
-is managed for DOE by ORAU. All opinions expressed in this paper are
-the author's and do not necessarily reflect the policies and views of
-NSF, ORAU/ORISE, or DOE. This work was also supported by The Raymond
-and Beverly Sackler Post-Doctoral Scholarship.
-
-%% \input{appendix}
-
-\bibliographystyle{siamplain}
-%% %%\bibliographystyle{amsplain}
-\bibliography{/home/yair/projects/bibtex.bib}
-
-\end{document}

### 2023-08-29 latex/funding.tex
Submitted to M3AS

--- 2023-05-21 latex/paper.tex
@@
-\documentclass{siamonline220329}
-%\documentclass{amsart}
-%\documentclass{article}
-\input{definitions}
-
-%\numberwithin{equation}{section}
-
-\title{Measurement Clusterization in D-optimal Designs for Bayesian Linear
-  Inverse Problems over Hilbert Spaces}
-
-\author{Yair Daon\thanks{Porter School of the Environment and Earth
-    Sciences, Tel Aviv University, Tel Aviv, Israel
-    (\email{yair.daon@gmail.com}).}}
-%% \author{Yair Daon\thanks{Azrieli Faculty of Medicine, Bar Ilan
-%%     University, Safed, Israel.  (\email{yair.daon@gmail.com}).}}
-
-%% \address{Porter School of the Environment and Earth
-%%   Sciences, Tel Aviv University\\ Tel Aviv, Israel}
-%% %% \curraddr{Porter School of Environment Studies, Tel Aviv University\\ Tel Aviv, Israel}
-%% %% \curraddr{}
-%% \email{yair.daon@gmail.com}
-% The 2010 edition of the Mathematics Subject Classification is the
-% current definitive version.
-%% \subjclass{Primary: 
-%%   62F15, % Statistics - bayesian inference
-%%   35R30, % PDE - Inverse problems
-%%   Secondary:
-%%   28C20,
-%%   % Measure and integration - set functions and measures and integrals
-%%          % in infinite-dimensional spaces (Wiener measure, Gaussian
-%%   % measure, etc.
-%% }
-
-\date{\today}
-\begin{document}
-\maketitle
-\begin{abstract}
-  Inversion for parameters of physical processes typically requires
-  taking expensive measurements; finding an optimal set of
-  measurements is known as the optimal design problem. Surprisingly,
-  measurement locations in optimal designs are sometimes extremely
-  clustered. Since measurement clusterization is nonintuitive,
-  researchers often avoid it by modifying the optimal design
-  problem. We consider a certain flavor of the optimal design problem,
-  based on the Bayesian D-optimality criterion, and suggest an
-  analytically tractable model for Bayesian D-optimal designs in
-  linear inverse problems over Hilbert spaces. We demonstrate that
-  measurement clusterization is a generic property of D-optimal
-  designs, and prove that correlated noise between measurements
-  mitigates clusterization. We also give a full characterization of
-  D-optimal designs, which shows that D-optimal designs uniformly
-  reduce uncertainty in a select subset of the prior covariance
-  eigenvectors. Finally, we show how measurement clusterization is a
-  simple consequence of the characterization mentioned above and the
-  pigeonhole principle.
-\end{abstract}
-
-\maketitle
-
-\input{intro}
-\input{prelims}
-\input{constrained}
-\input{nonvanishing}
-\input{vanishing}
-\input{why}
-%% \input{how}
-%% \section{Open Problems}
-
-\section{Acknowledgements}
-This study is a result of research I started during my PhD studies
-under the instruction of Prof. Georg Stadler at the Courant Institute
-of Mathematical Sciences. I would like to thank him for his great
-mentorship, attention to details and kindness. I would also like to
-thank Christian Remling, who helped me find a proof for Lemma
-\ref{lemma:free} in
-\href{https://mathoverflow.net/questions/280168/redistribute-diagonal-entries-of-a-matrix/280203#280203c}{Mathoverflow}.
-
 This research was supported in part by an appointment with the
@@
 and Beverly Sackler Post-Doctoral Scholarship.
-
-%% \input{appendix}
-
-\bibliographystyle{siamplain}
-%% %%\bibliographystyle{amsplain}
-\bibliography{/home/yair/projects/bibtex.bib}
-
-\end{document}

### 2023-08-29 latex/intro.tex
Submitted to M3AS

@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=\textwidth]{eigenvectors.pdf}
+    \caption{D-optimal measurement locations ($m=4$ measurements) and
+      weighted eigenvectors for the inverse problem of finding the
+      initial condition of the 1D heat equation. Measurement locations
+      and weighted eigenvectors are plotted over the computational
+      domain $\Omega = [0, 1]$ (x-axis). Measurement clusterization
+      occurs approximately at $0.31$ and $0.69$. These two locations
+      are a compromise between zeros of eigenvectors a D-optimal
+      design aims to ignore (third and up) and staying far from zeros
+      of the eigenvectors a D-optimal design aims to measure (first
+      and second). Allocating $m=4$ measurements into two locations
+      results in clusterization, according to the pigeonhole
+      principle.}
+  \label{fig:why}
+\end{figure}
+
 \begin{enumerate}
@@
   the pigeonhole principle, as there are more measurements available
-  than there are locations satisfying conditions (a) and (b).
+  than there are locations satisfying conditions (a) and (b). See
+  Fig~\ref{fig:why}.
 

### 2023-08-29 latex/m3as.tex
Submitted to M3AS

\documentclass{ws-m3as}
\input{definitions}
\usepackage[super]{cite}
\usepackage{url}

\begin{document}

\markboth{Yair Daon}{Measurement Clusterization in D-optimal Designs}

%%%%%%%%%%%%%%%%%%% Publisher's Area please ignore %%%%%%%%%%%%%%%%%%%%%%%
%
\catchline{}{}{}{}{}
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Measurement Clusterization in D-optimal Designs for Bayesian
  Linear Inverse Problems over Hilbert Spaces}

\author{Yair Daon}

\address{Azrieli Faculty of Medicine, Bar-Ilan University, Safed, Israel\\
yair.daon@gmail.com}

\maketitle

\begin{history}
\received{(Day Month Year)}
\revised{(Day Month Year)}
%\accepted{(Day Month Year)}
\comby{(xxxxxxxxxx)}
\end{history}

\begin{abstract}
  \input{abstract}
\end{abstract}

\keywords{Optimal Design; Clusterization; Inverse Problem; Bayesian.}

\ccode{AMS Subject Classification: 62F15, 35R30, 28C20}

\input{intro}
\input{prelims}
\input{constrained}
\input{nonvanishing}
\input{vanishing}
\input{why}
%% \input{how}
%% \section{Open Problems}

\section{Acknowledgements}
\input{ack}

\input{funding}

\bibliographystyle{ws-m3as}
\bibliography{/home/yair/projects/bibtex.bib}

\end{document}

### 2023-08-29 latex/ms.tex
Submitted to M3AS

--- 2023-05-21 latex/paper.tex
@@
-\documentclass{siamonline220329}
-%\documentclass{amsart}
-%\documentclass{article}
+\documentclass{amsart}
 \input{definitions}
+
+\title{Measurement Clusterization in D-optimal Designs for Bayesian
+  Linear Inverse Problems over Hilbert Spaces
+}
+
+\author[1]{Yair Daon}
+
+\address[1]{Azrieli Faculty of Medicine, Bar Ilan University, Safed,
+  Israel}
+
+%\email{yair.daon@gmail.com}
 
@@
 
-\title{Measurement Clusterization in D-optimal Designs for Bayesian Linear
-  Inverse Problems over Hilbert Spaces}
-
-\author{Yair Daon\thanks{Porter School of the Environment and Earth
-    Sciences, Tel Aviv University, Tel Aviv, Israel
-    (\email{yair.daon@gmail.com}).}}
-%% \author{Yair Daon\thanks{Azrieli Faculty of Medicine, Bar Ilan
-%%     University, Safed, Israel.  (\email{yair.daon@gmail.com}).}}
-
-%% \address{Porter School of the Environment and Earth
-%%   Sciences, Tel Aviv University\\ Tel Aviv, Israel}
-%% %% \curraddr{Porter School of Environment Studies, Tel Aviv University\\ Tel Aviv, Israel}
-%% %% \curraddr{}
-%% \email{yair.daon@gmail.com}
-% The 2010 edition of the Mathematics Subject Classification is the
-% current definitive version.
+%% The 2010 edition of the Mathematics Subject Classification is the
+%% current definitive version.
 %% \subjclass{Primary: 
@@
 
-\date{\today}
 \begin{document}
 \maketitle
+
 \begin{abstract}
-  Inversion for parameters of physical processes typically requires
-  taking expensive measurements; finding an optimal set of
-  measurements is known as the optimal design problem. Surprisingly,
-  measurement locations in optimal designs are sometimes extremely
-  clustered. Since measurement clusterization is nonintuitive,
-  researchers often avoid it by modifying the optimal design
-  problem. We consider a certain flavor of the optimal design problem,
-  based on the Bayesian D-optimality criterion, and suggest an
-  analytically tractable model for Bayesian D-optimal designs in
-  linear inverse problems over Hilbert spaces. We demonstrate that
-  measurement clusterization is a generic property of D-optimal
-  designs, and prove that correlated noise between measurements
-  mitigates clusterization. We also give a full characterization of
-  D-optimal designs, which shows that D-optimal designs uniformly
-  reduce uncertainty in a select subset of the prior covariance
-  eigenvectors. Finally, we show how measurement clusterization is a
-  simple consequence of the characterization mentioned above and the
-  pigeonhole principle.
+  \input{abstract}
 \end{abstract}
-
-\maketitle
 
@@
 \section{Acknowledgements}
-This study is a result of research I started during my PhD studies
-under the instruction of Prof. Georg Stadler at the Courant Institute
-of Mathematical Sciences. I would like to thank him for his great
-mentorship, attention to details and kindness. I would also like to
-thank Christian Remling, who helped me find a proof for Lemma
-\ref{lemma:free} in
-\href{https://mathoverflow.net/questions/280168/redistribute-diagonal-entries-of-a-matrix/280203#280203c}{Mathoverflow}.
-
-This research was supported in part by an appointment with the
-National Science Foundation (NSF) Mathematical Sciences Graduate
-Internship (MSGI) Program sponsored by the NSF Division of
-Mathematical Sciences. This program is administered by the Oak Ridge
-Institute for Science and Education (ORISE) through an interagency
-agreement between the U.S. Department of Energy (DOE) and NSF. ORISE
-is managed for DOE by ORAU. All opinions expressed in this paper are
-the author's and do not necessarily reflect the policies and views of
-NSF, ORAU/ORISE, or DOE. This work was also supported by The Raymond
-and Beverly Sackler Post-Doctoral Scholarship.
+\input{ack}
+\input{funding}
 
 %% \input{appendix}
-
-\bibliographystyle{siamplain}
-%% %%\bibliographystyle{amsplain}
+\bibliographystyle{amsplain}
 \bibliography{/home/yair/projects/bibtex.bib}

### 2023-08-29 latex/why.tex
Submitted to M3AS

@@
 \section{Answer to Question \ref{q:why}: The cause of clusterization}\label{section:why}
-\begin{figure}
-    \centering
-    \includegraphics[height=0.5\textwidth]{eigenvectors.pdf}
-    \caption{D-optimal measurement locations ($m=5$ measurements) and
-      weighted eigenvectors for inversion of the initial condition of
-      the 1D heat equation. Measurement locations and weighted
-      eigenvectors are plotted over the computational domain $\Omega =
-      [0, 1]$ (x-axis). Measurement clusterization occurs
-      approximately at $0.31$ and $0.69$. These two locations are a
-      compromise between zeros of eigenvectors a D-optimal design aims
-      to ignore (third and up) and staying far from a zero of the
-      first and second eigenvectors. Allocating $m=5$ measurements
-      into two locations results in clusterization, according to the
-      pigeonhole principle.}
-  \label{fig:why}
-\end{figure}
-
 According to Theorem \ref{thm:char}, D-optimal designs aim to capture

### 2023-08-29 latex/abstract.tex
Sumbitted to M3AS

Estimation of parameters in physical processes often demands costly measurements, prompting the pursuit of an optimal measurement strategy. Finding such strategy is termed the problem of \emph{optimal experimental design}, abbreviated as optimal design. Remarkably, optimal designs can yield tightly clustered measurement locations, leading researchers to fundamentally revise the design problem just to circumvent this issue. Some authors introduce error correlation among error terms that are initially independent, while others restrict measurement locations to a finite set of locations. While both approaches may prevent clusterization, they also fundamentally alter the optimal design problem.

In this study, we consider Bayesian D-optimal designs, i.e.~designs that maximize the expected Kullback-Leibler divergence between posterior and prior. We propose an analytically tractable model for D-optimal designs in Bayesian linear inverse problems over Hilbert spaces. In this framework, we make several key contributions: \textbf{(a)} We establish that measurement clusterization is a generic trait of D-optimal designs with independent Gaussian measurement errors, and prove that introducing correlations among measurement error terms mitigates clusterization. \textbf{(b)} We characterize D-optimal designs as reducing uncertainty across a subset of prior covariance eigenvectors. Finally, \textbf{(c)} We leverage this characterization to argue that measurement clusterization arises as a consequence of the pigeonhole principle: when more measurements are taken than there are locations where the select eigenvectors are large and others are small --- clusterization occurs. 
%
%% In summary, in this study we shed light on an often ignored issue with
%% Bayesian D-optimal designs. We characterize such designs and show why
%% clusterization arisesit arisestheir inherent properties, propose
%% strategies to address clusterization, and provide insights into the
%% fundamental attributes underlying measurement design optimization.

### 2023-10-01 latex/SIAM/siam.tex
Submit to Bayesian Analysis

\documentclass{siamart220329}
\input{definitions}

% Sets running headers as well as PDF title and authors
\headers{Measurement Clusterization in D-optimal Designs}{Yair Daon}

% Title. If the supplement option is on, then "Supplementary Material"
% is automatically inserted before the title.

\title{Measurement Clusterization in D-optimal Designs for Bayesian
  Linear Inverse Problems over Hilbert Spaces
  \thanks{Submitted to the editors Aug. 29\textsuperscript{th} 2023}
  %% \funding{\input{funding}}
}

\author{Yair Daon \thanks{Azrieli Faculty of Medicine, Bar Ilan University, Safed,
      Israel (\email{yair.daon@gmail.com}).}}

\begin{document}
\maketitle

\begin{abstract}
  \input{abstract}
\end{abstract}

\input{intro}
\input{prelims}
\input{constrained}
\input{nonvanishing}
\input{vanishing}
\input{why}
%% \input{how}
%% \section{Open Problems}

\section{Acknowledgements}
\input{ack}

%% \input{appendix}
\bibliographystyle{siamplain}
%\bibliographystyle{amsplain}
\bibliography{/home/yair/projects/bibtex.bib}

\end{document}

### 2023-10-01 latex/ba.tex
Submit to Bayesian Analysis

% \documentclass[ba,preprint]{imsart}% use this for supplement article
\documentclass[ba]{imsart}
\input{definitions}
%
\pubyear{2021}
\volume{TBA}
\issue{TBA}
% \doi{0000}
%\arxiv{}
\firstpage{1}
\lastpage{1}

\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{natbib}
\usepackage[colorlinks,citecolor=blue,urlcolor=blue,filecolor=blue,backref=page]{hyperref}
\usepackage{graphicx}

\startlocaldefs
% ** Local definitions **
\endlocaldefs

\begin{document}

%% *** Frontmatter ***

\begin{frontmatter}
\title{Measurement Clusterization in Bayesian D-optimal Designs in Infinite Dimensions}

%\title{\support{}}
\runtitle{Measurement Clusterization in D-optimal Designs}
%\author{Yair Daon}
%\numberwithin{equation}{section}

\begin{aug}
\author{\fnms{Yair} \snm{Daon}\thanksref{addr1}\ead[label=e1]{yair.daon@gmail.com}}
\runauthor{Yair Daon}

\address[addr1]{Azrieli Faculty of Medicine, Bar-Ilan University, Safed, Israel
    \printead{e1} % print email address of "e1"
}

\end{aug}

\begin{abstract}
  \input{abstract}
\end{abstract}

%% ** Keywords **
\begin{keyword}[class=MSC]
\kwd[Primary ]{62F15}
\kwd{35R30}
\kwd[; secondary ]{28C20}
\end{keyword}

\end{frontmatter}

\input{intro}
\input{prelims}
\input{constrained}
\input{nonvanishing}
\input{vanishing}
\input{why}

\bibliographystyle{ba}
\bibliography{/home/yair/projects/bibtex.bib}

\begin{acks}[Acknowledgments]
\input{ack}

\input{funding}
\end{acks}

\end{document}

### 2023-10-01 latex/intro.tex
Submit to Bayesian Analysis

@@
 Measurements play a fundamental role in generating observations and
-are indispensable to any inference process. In the practice of inverse
-problems, selecting the optimal set of measurements holds particular
-significance when inferring parameters of a physical process. Unlike
-many other fields, where measurements are fixed, in this context, we
-have the freedom to choose which measurements to take. This freedom
-should be taken advantage of: measurements should be chosen to enhance
-accuracy, reduce costs, or both. Whether measurements involve
-specifying electrode locations on the skin in electric impedance
-tomography \cite{horesh2010impedance}, determining certain wavelengths
-in MRI \cite{horesh2008mri}, positioning sensors for detecting
-ground-reflected waves in the search for oil
-\cite{horesh2008borehole}, or optimizing sensor locations in a
+are indispensable to any inference process. Selecting the optimal set
+of measurements holds particular significance when inferring
+parameters of a physical process. Unlike many other fields, where
+measurements are fixed, in this context, we have the freedom to choose
+which measurements to take. This freedom should be taken advantage of:
+measurements should be chosen to enhance accuracy, reduce costs, or
+both. Whether measurements involve specifying electrode locations on
+the skin in electric impedance tomography \cite{horesh2010impedance},
+determining certain wavelengths in MRI \cite{horesh2008mri},
+positioning sensors for detecting ground-reflected waves in the search
+for oil \cite{horesh2008borehole}, or optimizing sensor locations in a
 seismographic network \cite{rabinowitz1990, steinberg1995} --- the
@@
 D-optimality being two of the most widely recognized and extensively
-studied design criteria \cite{Chaloner1995}.
+studied design criteria \cite{Chaloner1995,
+  AlexanderianGloorGhattas14}.
 
@@
   nyberg2012, fedorov1997, Ucinski05, neitzel2019sparse}. To
-illustrate this point, we consider a toy inverse problem: inferring
+illustrate this point, we consider a toy problem: inferring
 the initial condition of the 1D heat equation (details in the
@@
     \caption{D-optimal measurement locations ($m=4$ measurements) and
-      weighted eigenvectors for the inverse problem of finding the
-      initial condition of the 1D heat equation. Measurement locations
-      and weighted eigenvectors are plotted over the computational
-      domain $\Omega = [0, 1]$ (x-axis). Measurement clusterization
-      occurs approximately at $0.31$ and $0.69$. These two locations
-      are a compromise between zeros of eigenvectors a D-optimal
-      design aims to ignore (third and up) and staying far from zeros
-      of the eigenvectors a D-optimal design aims to measure (first
-      and second). Allocating $m=4$ measurements into two locations
+      weighted eigenvectors for finding the initial condition of
+      the 1D heat equation. Measurement locations and weighted
+      eigenvectors are plotted over the computational domain $\Omega =
+      [0, 1]$ (x-axis). Measurement clusterization occurs
+      approximately at $0.31$ and $0.69$. These two locations are a
+      compromise between zeros of eigenvectors a D-optimal design aims
+      to ignore (third and up) and staying far from zeros of the
+      eigenvectors a D-optimal design aims to measure (first and
+      second). Allocating $m=4$ measurements into two locations
       results in clusterization, according to the pigeonhole

### 2023-10-01 latex/ms.tex
Submit to Bayesian Analysis

@@
 
-\title{Measurement Clusterization in D-optimal Designs for Bayesian
-  Linear Inverse Problems over Hilbert Spaces
-}
+\title[Measurement Clusterization in D-optimal Designs]{Measurement Clusterization in Bayesian D-optimal Designs in Infinite Dimensions}
 
-\author[1]{Yair Daon}
-
-\address[1]{Azrieli Faculty of Medicine, Bar Ilan University, Safed,
-  Israel}
+\author{Yair Daon}
+\address{Azrieli Faculty of Medicine, Bar-Ilan University, Safed, Israel}
 
@@
 \input{why}
-%% \input{how}
-%% \section{Open Problems}
 
@@
 
-%% \input{appendix}
 \bibliographystyle{amsplain}
@@
 \end{document}
+
+%% \input{how}
+%% \section{Open Problems}

### 2024-06-05 latex/review.tex
Starting to work on review

\documentclass{amsart}
\input{definitions}

\begin{document}

\section{Editor in Chief}
Personally I like the paper, but it represents a somehow unusual
submission for Bayesian Analysis, and I am worried about it being of
sufficient general interest for the broad readership of the
journal. Still, in agreement with the AE, we decided to give the
author the possibility to revise the paper. However, I must stress
that there is no guarantee of a successful outcome, which will solely
depend on the quality of the revision. In particular:

\begin{description}
  \item[a] the Introduction is to be made accessible to the whole BA
    readership;
  \item[b] the motivation and the practical statistical implications are to be
    developed and discussed;
  \item[c] the contribution has to be better contextualized with
    respect to the literature in both inverse problems and D-optimal
    designs;
  \item[d] numerical evidence and code, mentioned in the paper
    but actually not provided, need to be made available;
  \item[e] the mathematical derivations have to be watertight and some
    clarifications are needed in this respect (see Report of Referee
    3).

\end{description}

\section{Associate Editor}
The three referees found merits and interest in your work, but also
pointed out several major issues to be addressed. Of particular
concern to me, the paper's exposition jumps too quickly into the
formulation / maths without giving a proper intuition and
context. This makes the paper less accessible to readers who aren't
already quite familiar with inverse problems (which are introduced
with very little preamble, and it's not even immediately explicit what
are the parameters being learned) and optimal experimental design.

A referee makes valid points as to whether clusterization of
measurements is a problem in general, when in fact in can be an
appealing feature in some settings, and another referee asks whether
D-optimal designs should be abandoned altogether. The exposition needs
to discuss these high-level issues, and introduce the related
literature and results appropriately, to give the work proper
context. Altogether, I think that your work has the potential to make
a good contribution to Bayesian analysis, but as it currently stands
significant work is needed.

\section{Reviewer 1}
This paper explains a common practical issue: the cause of
clusterization in Bayesian D-optimal designs in infinite dimensions
and why adding a correlated measurement process decluster the design
points. Its theoretical results are valuable to the field of inverse
problems/uncertainty quantification and the conclusions should be
useful to practitioners in this field.  However, I feel changing the
assumption from independent to correlated measurements is quite
artificial: people tweak a data generation process assumption to solve
an optimal design problem, and the author just takes the assumption as
it is. Since the author proves the clusterization issue under the
independent measurement assumption, I expect the author to go deeper
and explain why an ordinary data generation assumption + a common
optimal design leads to a “suboptimal” design? Not just by the
mathematical proof, but from an information theory point of view. For
example, does the redundancy points mean additional points does not
reduce the variance of the estimation at all? If that’s the case, how
does correlated measurement assumption “help” with it?

\section{Reviewer 2}
\subsection{Measurement Clusterization vs Repetition}
I will start by saying measurement clusterisation is not well-defined
in this paper. On page 1, it is referred to as two sets of
measurements that are almost indistinguishable from one
another. However, the setup for Corollary 11 implicitly defines
measurement clusterisation as repeated measurements. The latter
definition matches the references of Fedorov (1996) and Nyberg et al
(2012). So I will assume for the rest of this review that, measurement
clusterisation refers to repeated measurements.  Nearly all textbooks
on classical design of experiments (e.g. Morris, 2011, Section 1.2.4)
will feature a section with the three words randomisation, blocking
and repetition: the important principles of design of experiments.  In
these texts, repeated measurements are considered to be a beneficial
property of a design. Interestingly, when optimal designs are found,
they are often found to have repeated measurements.  The paper does
not explain why the author thinks repeated measurements (measurement
clusterisation) is an undesirable property. If one looks at, for
example, Fedorov (1996) and Nyberg et al (2012), it is an undesirable
property because they are interested in sensor location, and two or
more sensors cannot be in the same location. However, there are many
experiments where repeated measurements are possible, and are
therefore optimal (if the model is correct, which it won’t be). An
example, is a chemical engineering experiment where one wishes to
learn the relationship between a series of variables and yield of a
chemical reaction. The experiment will involve specifying the
variables, and measuring the yield (the observation) after a specified
period of time. This process is then duplicated with a potentially
different set of variables.  It is perfectly possible to have repeated
variables.
\begin{enumerate}
\item The author needs to make clear that they are considering
  experiments where repeated measurements are impossible. Admitedly,
  this consideration is implicit in the paper, for example, the
  examples given on page 1. However, it needs to be made explicit in
  the Abstract and the first paragraph of the paper.

\item The paper explains why repeated measurements can be optimal
  (answer to Question 3). The author should expand on this and draw
  the link with how repeated measurements are, typically, regarded as
  desirable in classical design of experiments for finite
  dimensions. See for example, the discussion of the relationship
  between Caratheodory’s Theorem and the number of support points in
  Pronzato \& Pazman (2013, page 139). The support points are the
  unique design points, so if this is less than m, then there are
  repeated design points.
\end{enumerate}

\subsection{Context}
The paper lacks discussion on context. I read the answers to the three
questions and thought “So what?”  What are the implications to
statistical practice, or experimental science, of these findings?  The
presentation of the paper does not help. The results are presented
with no explanation or intuition. It might be better to relegate much
of the mathematical detail to Supplementary Material and use the main
manuscript for exposition.

\subsection{Avoiding measurement clusterisation}
The approaches to avoid measurement clusterisation, described on page
3, are described as pragmatic and "fundamentally altering the optimal
design problem".

Firstly, this is not strictly true. Suppose measurements correspond to
time, it is not possible to take two of more observations at the same
time, and there should be a minimum time period between measurements.
Implementing this minimum time period is not pragmatic, rather it is
characterising the application correctly.  One could imagine a similar
thing occurs with the eletrode locations on the skin in impedence
tomography: the physical size of the sensor imposes a miniumum
distance between sensors.

Secondly, the paper seems to suggest “imposing correlations between
observations” as a solution. However, isn’t this imposition pragmatic
as well? I suppose it is argued that the $\obs \eps′$ term is
accounting for model error: but it is still pragmatic: the data
acquisition given by equation (1) is not the true data-generating
process.

\subsection{Minor comments}
\begin{itemize}
\item The author uses two different forms of asterisk: one for
  optimality of the design (e.g. Definition 3) and one for the
  adjoint. Suggest changing the one for optimality.
\item I do not like the use of the term measurements as used in the
  paper. The measurements are the $\data$’s since these are a result of
  measuring a physical quantity. However, the paper refers to the
  quantities that are controlled as the measurements.
\item Proof of Proposition 12 uses $\eps$ as part of the
  regularisation trick but this has been used previously for error
  (e.g. equation (1)).
\end{itemize}

\section{Reviewer 3}
The paper studies Bayesian D-optimal design (that is, the task of
allocating measurements in order to maximise the Kullback-Leibler
divergence between the prior and posterior disribution) in an inverse
regression type model in Hilbert spaces with Gaussian measurements
errors. For Gaussian priors, the paper provides a characterisation of
the optimality criterion and the optimal design. Some conclusions
about the phenomenon of clusterisation of measurements known in
literature are then drawn.

While the results in the paper provide some novel insights on the
topic, I believe the current version of the manuscript is far from the
standard required by Bayesian Analysis. Let me raise immediately my
main concern: on p.4 the author describes the generality of
measurements clus- terisation as the first main question being
explored, which is answered via, quoting, "randomized numerical
simulations that exhibit clusterization more than 95\% of the time
(see the code in sup- plementary material)". However, in the version
of the supplementary materials I had access to there is no further
mention of these numerical simulations and results; only a link to a
Github repository is provided, which is however not
functioning. Evidently, this issue needs to be fixed before any future
consideration can be given to the manuscript. For example, a detailed
description of the repeated experiments could be provided either in
the main article or in the supplement, with ta- bles indicating the
various values of the parameters for the simulations and the precise
percentages obtained. Also, the code needs to be publicly available
for reproducibility.

\subsection{Further comments}
I have a number of further comments in regards to the results and the overall presentation of
the manuscript.
\begin{itemize}
\item D-optimal design vs. statistical recovery rates: being more
  familiar with the statistical inverse problem literature than with
  D-optimal design, the clusterisation phenomenon demonstrated in this
  paper raises the question as to wether consistent recovery of the
  unknown (m in the notation of this paper) is possible under this
  choice. All the results I am aware of either assume white
  noise/equally spaced design (e.g. [1]) or measurement loctions
  sampled uni- formly at random (e.g. [2]). There are also some
  results with more general design under conditions on the
  fill-distance of the grid (e.g. [3]). All of these specifically
  prevent clusteri- sation, which is key to the statistical
  analysis. Hence, it is not clear to me even if D-optimal design
  should be pursued at all, if it indeed it leads to clusterisation as
  implied by the present 1paper. I think a discussion on this would be
  illuminating for the reader and help draw a connection to the
  broader statistical inverse problems literature.

\item Figure 1: It is hard to descern whether the measurements
  locations are perfectly overlapping or just very closely
  placed. Perhaps using dots with numbers here would help the
  readability.

\item Experiment with the 1D heat equation: the experiment showcases
  some interesting phenom- ena, but also raises many questions. For
  instance: I would expect the dimensionality of the working domain
  (here the unit interval [0, 1]) to play a role, since larger
  dimension allow for higher freedom in placing the design points.

\item Further, a dimensionality effect should also arise directly from
  Theorem 1 via the prior co- variance eigenvalues: in the experiment
  the prior covariance operator is set to be equal to the inverse
  Laplace operator $(−\Delta)^{−1}$ . In d-dimensional domains, its
  eigenvalues λj are known to follow Weyl’s asymptotics, growing as j
  2/d , which should then impact the index after which the eigenvalues
  are ‘thresholded’ by the procedure.

\item Also, it should be relatively easy here to also study empirically the effect of the correlated
error model in preventing clusterisation. Investigating all these issues in the context of the
presented numerical set up would considerably enlarge the breath of the present work.

\item I found very hard to understand Theorem 1 immediately at the end
  of Section 1, as all the necessary background is introduced only
  later. Since the same result is also stated and proved in Section 5,
  I would consider cutting it from Section 1. The paragraphs in p.4
  and 5 already do a good job presenting the results, and the
  repetition of the formal statement does not seem necessary here.

\item Denoting the optimal design in Theorem 1 by simply O is somewhat confusing, cf. the
equation in the second item. Perhaps a separate notation such as Ō would help readability.

\item Section 1.2 seems altogether unnecessary; the inverse regression model considered in the
paper are a modelisation of many real-world phenomena, and they are routinely studied in
statistical papers.

\item What does ‘strongly smoothning’ operator means in the thid line of Section 2.1?

\item Section 4: the conclusion reached at the end of the section on
  p.15 seem to imply that in the noiseless limit, repeating any
  measurement does not improve the objective criterion, if corre-
  lated error model are present. Is it clear however that adding any
  (non-reapted) measurement always strictly increases it? I.e., there
  could be situations where adding one measurement does not change the
  value of the optimised criterion? Also, can the conclusion drawn
  here be extended to the case of noisy observations σ > 0?

\item At the end of Section 4 on p.15 it is mentioned that $\tar$ is not
  defined for $\sigma^2= 0$, but in the latter case, could not a Gaussian
  posterior measure be still defined as in Gaussian process
  regression, e.g.[4]

\item Theorem 1 is expressed in terms of the prior covariance matrix,
  which is a user specified quantity. Hence, it seems to me that all
  sort of design behaviour under the D-optimal crite- rior can occur
  by engineering the prior. It would perhaps be informative to study
  in more details some representative example of inverse problem and
  Gaussian prior. For example, the recovery of the initial condition
  with the ‘Matérn-like’ Gaussian prior laid out in the sup- plement
  are good candidates. For these it is still not clear how the
  conclusion from Theorem 1 translates into the Figure 1, also because
  I am unsure about the applicability of the result to point
  evaluations. In particular, how the conclusion that with m = 4
  measurements the D-optimal procedure will aim to ignore the third
  and fourth eigenvalue/function was drawn?
\end{itemize}

\end{document}

### 2024-06-16 latex/intro.tex
Working towards revision. Modified intro, now working on refactoring code

@@
 \section{Introduction}\label{section:intro}
-Measurements play a fundamental role in generating observations and
-are indispensable to any inference process. Selecting the optimal set
-of measurements holds particular significance when inferring
-parameters of a physical process. Unlike many other fields, where
-measurements are fixed, in this context, we have the freedom to choose
-which measurements to take. This freedom should be taken advantage of:
-measurements should be chosen to enhance accuracy, reduce costs, or
-both. Whether measurements involve specifying electrode locations on
-the skin in electric impedance tomography \cite{horesh2010impedance},
-determining certain wavelengths in MRI \cite{horesh2008mri},
-positioning sensors for detecting ground-reflected waves in the search
-for oil \cite{horesh2008borehole}, or optimizing sensor locations in a
-seismographic network \cite{rabinowitz1990, steinberg1995} --- the
-selection of optimal measurements, referred to as the problem of
-\emph{optimal design}, becomes crucial. Optimal measurements are
-chosen according to specific \emph{design criteria}, with A- and
-D-optimality being two of the most widely recognized and extensively
-studied design criteria \cite{Chaloner1995,
-  AlexanderianGloorGhattas14}.
+The practice of experimental design entails specifying every detail of
+running an experiment \cite{chaloner}: at its simplest form,
+experimental design involves choosing a set of measurements to be
+taken. In a medical experiment, experimental design may involve
+choosing participants to whom some drug will be administered, and
+measuring their blood pressure. Experimental design also arises 
+process of searching for oil. Since measuremnts are expensive and
+require digging holes in the ground, only a limited number of
+measuremnts can be taken, hence carefully choosing measuremt locations
+is crucial.
 
-Surprisingly, A- and D-optimal designs have been observed to yield
-remarkably similar measurements in certain cases \cite{fedorov1996,
-  nyberg2012, fedorov1997, Ucinski05, neitzel2019sparse}. To
-illustrate this point, we consider a toy problem: inferring
-the initial condition of the 1D heat equation (details in the
-supplementary material). In Fig.~\ref{fig:clusterization
-  illustration} D-optimal measurement locations are shown for
-different numbers of measurements. Notably, for six measurements, a
-D-optimal design yields two sets of measurements that are almost
-indistinguishable from one another. We refer to this intriguing
-phenomenon as \emph{measurement clusterization} \cite{Ucinski05}, and
-we refer to a design that exhibits measurement clusterization as a
-\emph{clustered design}.
+Specifiyng the right set of measurements holds particular significance
+when solving an \emph{inverse problem} --- i.e. making inference of
+the physical world utilizing a physical model of some phenomenon of
+interest. In an inverse problem, we seek to infer physical quantities,
+based on observations and a model of the physical world, typically
+phrased in the language of ordinary or partial differential equations.
+
+Inverse problems are ubiquitous in science and engineering, and we
+will now give several illustrative examples. In electric impedance
+tomography (EIT) we seek to infer the impedance of the inside of a
+human body from observations of electrical current impedence measured
+at specific electrode locations on the skin, leveraging known
+equations of electric current flow
+\cite{horesh2010impedance}. Magnetic Resonance Imaging (MRI) also
+entails solving an inverse problem, where radio-frequency pulses are
+sent through the human body in the presence of a strong magnetic field
+and the response of our bodies' contents to such perutrbations is
+measured, allowing a radiologist to view the internals of a body in a
+noninvasive manner \cite{horesh2008mri}. Yet another example of an
+inverse problem arises in oil and gas exploration, where acoustic
+waves are sent through "boreholes" --- deep cylindrical wells drilled
+into the ground --- and data of travel and return times of said waves
+are recorded. Since travel and return time is influenced by the
+properties of the subsurface materials, such as density, elasticity,
+and the presence of fluids or voids, we can combine travel data with a
+geophisical models of contents of the earth's crust to facilitate
+reconstructing and inferring the structure of the subsurface in a
+process called \emph{borehole tomography}
+\cite{horesh2008borehole}. Inverse problems also arise in many other
+areas of seismology and geology \cite{rabinowitz1990, steinberg1995}
+and medical imaging \cite{tarantola}. Importantly, note that in the
+above examples our goal is to infer some function $f \in C(\Omega)$,
+where \(\Omega \subseteq \mathbb{R}^d, d=1,2,3\) is a spatial domain
+of interest.
+
+%% In EIT, electrode locations determine the quality of the
+%% reconstructed impedance field, while in MRI utilizng certain
+%% radio-frequency pulses may give more accurate information than
+%% others. In borehole tomogrpahy, only several boreholes could be dug
+%% and the choice of their locations is also crucial for oil and gas
+%% discovery. Thus, the selection of optimal measurements, referred to as
+%% the problem of \emph{optimal design}, becomes crucial.
+%% in a specific location (in a typical
+%% medical application to measure inverventions on, . These experimental
+%% units may be individual patients, but also Although common sense is
+%% useful in this process, it is often the case that naive intuition
+%% results in experiments that do not generate the maximal amount of
+%% information possible. To this end
+
+We will now introduce a toy model which we will use to define a toy
+inverse problem: the heat equation in one dimension with a homogeneous
+Dirichlet boundary condition. The equations are presented below for
+completeness, but readers less familiar with partial differential
+equations should think of the heat equation as follows: before us
+there is a perfectly insulated metal rod. At each point on the rod the
+temperature is different, and we call this temperature distribution
+the \emph{initial condition}. We wait for a short time and let the
+heat dissipate a bit inside the rod. The final heat distribution in
+the rod is dictated by the \emph{1D heat equation}:
+\begin{subequations}\label{eq:heat equation}
+  \begin{alignat}{2}
+    u_t &= \Delta u &&\qquad \text{in } [0,1] \times [0,\infty),\\ u
+      &= 0 &&\qquad \text{on } \{0, 1\} \times [0,\infty),\\ u &= u_0
+        &&\qquad \text{on }[0,1] \times \{0\}.
+  \end{alignat}
+\end{subequations}
+
+In the inverse problem, our goal is to infer the initial condition
+(i.e.~initial temperature distribution inside the rod). We are allowed
+to take some number of temperature readings on the rod once our
+waiting time has passed. The inverse problem of the 1D heat equation
+is difficult ("ill-posed") since heat spreads in a diffusive manner
+and any roughness in the initial condition is quickly smoothed, see
+Supplementary movies S1 and S2. We will give a fully Bayesian
+formulation of this inverse problem in Section \ref{section:prelims}.
+
+In all of the examples of inverse problems we have seen, as well as in
+many other applications, measurements should be chosen to most benefit
+the user. Hence, the user may consider some utility, called a
+\emph{design criterion} and take measurements that maximize this
+utility. Two of the most widely recognized and extensively studied
+design criteria are A- and D-optimality. We focus on the Bayesian
+D-optimality criterion, which carries a simple and intuitive meaning:
+measurements chosen according to the D-optimality criterion maximize
+the expected Kullback-Liebler divergence from posterior to prior
+\cite{Chaloner1995, AlexanderianGloorGhattas14}, see also precise
+definition in Section \ref{subsec:D optimal design}. We refer to a set
+of measuremnts that maximizes the Bayesian D-optimality criterion a
+\emph{D-optimal design}.
+
+Surprisingly, A- and D-optimal designs for inverse problems have been
+observed to yield remarkably similar measurements in certain cases
+\cite{fedorov1996, nyberg2012, fedorov1997, Ucinski05,
+  neitzel2019sparse}. This phenomenon is illustrated for our toy
+inverse problem in Fig.~\ref{fig:clusterization illustration}, where
+D-optimal measurement locations are shown for different numbers of
+measurements. Notably, for six measurements, a D-optimal design yields
+two sets of measurements that are almost indistinguishable from one
+another. We refer to this intriguing phenomenon as \emph{measurement
+clusterization} \cite{Ucinski05}. For our purposes, it suffices for
+measurements to be visibly close for clusterization to hold. We refer
+to a design that exhibits measurement clusterization as a
+\emph{clustered design},
 
@@
     \includegraphics[height=0.5\textwidth]{example.pdf}
-    \caption{Measurement clusterization for optimal designs when
-      inverting for the initial condition of the 1D heat equation (see
-      supplementary material for details). Measurement locations were
+    \caption{Measurement clusterization for D-optimal designs for the
+      inverse problem of the heat equation. Measurement locations were
       chosen according to the Bayesian D-optimality criterion of
@@
       clarity. Measurement clusterization already occurs for three
-      measurements: the second measurement (red) is overlaid with the
+      measurements: the second measurement (red) is overlaid on the
       third (green). For five measurements, first (blue) and second
@@
 
-Researchers widely agree that measurement clusterization is
-undesirable \cite{fedorov1996, nyberg2012, fedorov1997, Ucinski05,
-  neitzel2019sparse}, prompting the exploration of various remedies to
-address this issue. One approach involves merging close measurements
-\cite{fedorov1997}; however, this strategy merely overlooks the
-phenomenon of measurement clusterization. An alternative solution lies
-in \emph{clusterization-free design}s, where measurement locations are
+Clusterization should not be confused with either repetition nor with
+replication, which are commonly viewed as beneficial and even
+necessary aspects of an optimal design \cite{morris, fisher
+  schafer2001replication}. For example, in his famous milk and tea
+experiment, \cite[Section 11]{fisher} suggested that repetition is a
+"way of enlarging the experiment and, thereby, increasing its
+sensitiveness". In another example, \cite{fay2000rainfall} measured
+the effect of rainfall on grass growth in plots of land. The
+experiment involved fifteen "rainfall manipulation shelters", where
+"Four rainfall manipulation treatments (three replicates) then were
+assigned to 12 of the plots". It seems reasonable the researchers
+replicate the phenomenon they are trying to study. Clusterization
+seems like a whole different beast: a clustered design in the rainfall
+experiment would imply the researchers should take a repeated
+measurement \emph{on the same plot}, at the expense of measuring grass
+growth in other plots. In sharp contrast to repetition and
+replication, clusterization is highly nonintuitive and, as we will
+see, its origins run considerably deeper.
+
+Researchers of inverse problems widely agree that measurement
+clusterization is undesirable \cite{fedorov1996, nyberg2012,
+  fedorov1997, Ucinski05, neitzel2019sparse}, prompting the
+exploration of various remedies to address this issue. One approach
+involves merging close measurements \cite{fedorov1997}; however, this
+strategy merely overlooks the phenomenon of measurement
+clusterization. An alternative solution lies in
+\emph{clusterization-free design}s, where measurement locations are
 deliberately chosen to be distant from one another. This can be
@@
 
-The cornerstone of our investigation into measurement clusterization
-is Theorem \ref{thm:char} proven in Section
-\ref{section:vanishing}. The key insight of Theorem \ref{thm:char}
-lies in its fifth component, which highlights that a D-optimal design
-aims to uniformly reduce posterior uncertainties for posterior
-covariance eigenvectors in observation space. A similar conclusion was
-reached by Koval et al.~\cite{koval2020}, who showed that A-optimal
-designs are best constructed in the space of observations.
+The first insight our study gives is that a D-optimal design aims to
+uniformly reduce uncertainties not over space (i.e.~$\Omega$), but
+rather over \emph{eigenvectors}. This insight is a part of Theorem
+\ref{thm:char}, see Section \ref{section:vanishing} for precise
+statement and proof, and see Fig.~\ref{cluste} for an illustration.
 
-Before we state Theorem \ref{thm:char} we give some definitions: Let
-\(\hilp, \hilo\) Hilbert spaces, \(\fwd:\hilp \to \hilo\) a linear
-compact operator. Let \(\obs: \hilo \to \mathbb{R}^m\) a linear
-measurement operator, where \(m \in \mathbb{N}\) is the number of
-measurements taken. Let \(\sigma^2 \in \mathbb{R}_{+}\) observation
-noise variance, \(\data = \obs \fwd \param + \eps\), where \(\eps \in
-\mathbb{R}^m\) is iid \(\mathcal{N}(0, \sigma^2)\) noise. Let \(\pr
-\sim \mathcal{N}(\prmean, \prcov)\) prior Gaussian measure on
-\(\hilp\), where \(\prcov\) is the prior covariance operator and let
-\(\post\) the posterior (Gaussian) measure.
-
-\begin{restatable}{theorem}{char}\label{thm:char}
-  Let:
-  \begin{itemize}
-    \item The D-optimality design criterion
-    \cite{AlexanderianGloorGhattas14}:
-    \begin{align*}
-      \begin{split}
-        \tar(\obs) %:&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
-        % 
-        % 
-        % 
-        &= \frac12 \log \det ( I + \sigma^{-2} \prcov^{1/2} \fwd ^*
-        \obs^* \obs \fwd \prcov^{1/2}), 
-      \end{split}
-    \end{align*}
-  \item \(\obs\) a D-optimal design operator
-    \begin{equation*}
-      \obs = \argmax_{\|\meas_j\| = 1, j=1,\dots,m}\tar(\obs),
-    \end{equation*}
-  \item \(\{\lambda_i\}_{i=1}^\infty\) eigenvalues of
-    \(\fwd\prcov\fwd^*\) in decreasing order of magnitude.
-    %% \item \(\{\ev_i\}_{i=1}^\infty\) their corresponding eigenvectors.
-  \item \(\{\eta_i\}_{i=1}^\infty\) eigenvalues of \(\obs^*\obs\).
-
-  \end{itemize}
-
-  Then:
-  \begin{enumerate}
-  \item  \(\tr{\obs^*\obs} = m\).
-  \item \(\obs^*\obs\) and \(\fwd\prcov\fwd^*\) are simultaneously
-    diagonalizable.
-  \item \(k := \rank \obs^*\obs \leq m\) and
-    \begin{equation*}
-      \tar(\obs) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i). %= \frac12 \sum_{i=1}^{m} \log (1 + \sigma^{-2}\lambda_i\eta_i).
-    \end{equation*}
-  %% \item 
-  %%   \begin{equation*}
-  %%     k = \argmax \left \{ k:\lambda_k^{-1} < \sigma^{-2}\frac{m}{k} + \frac{1}{k} \sum_{j=1}^{k}
-  %%     \lambda_j^{-1} \right \}.
-  %%   \end{equation*}
-  \item
-    \begin{equation*}
-        \eta_i = \begin{cases}
-          \frac{m}{k} - \sigma^2 \lambda_i^{-1} + \sigma^2 \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1} & 1 \leq i \leq k \\
-          0 & i > k 
-        \end{cases}.
-    \end{equation*}
-  \item The covariance of the pushforwad \(\fwd_{*} \post\) is \(\left
-    ( (\fwd \prcov \fwd^*)^{-1} + \sigma^{-2} \obs^*\obs \right
-    )^{-1}\) and its eigenvalues are
-    \begin{equation*}
-      \theta_i =
-      \begin{cases}
-        \left(\frac{\sum_{j=1}^k \lambda_j^{-1} + \sigma^{-2}m}{k} \right )^{-1} & i \leq k \\
-        \lambda_i &  i > k 
-      \end{cases}
-    \end{equation*}
-  \end{enumerate}
-%\end{theorem}
-\end{restatable}
+In Theorem \ref{thm:char}, we also show that D-optimal designs are
+best understood in the sapce of \emph{observations}. Again, a precise
+statement could be found in Section \ref{section:vanishing}.  This
+result is in accordance with previous work by \cite{koval2020}, who
+showed that A-optimal designs are best constructed in the space of
+observations.
 
@@
 generalize several lemmas. Among those, is Lemma \ref{lemma:free},
-which is (to the the best of my knowledge) a novel lemma in linear
-algebra: We decompose a symmetric positive definite matrix \(M \in
-\mathbb{R}^{k \times k}\) with \(\ttr M = m \in \mathbb{N}\) as \(M =
-AA^t\), where \(A\) has unit norm columns.
+which is a lemma in linear algebra that (to the the best of my
+knowledge) is novel: We decompose a symmetric positive definite matrix
+\(M \in \mathbb{R}^{k \times k}\) with \(\ttr M = m \in \mathbb{N}\)
+as \(M = AA^t\), where \(A\in \mathbb{R}^{k \times m}\) has unit norm
+columns.
 
@@
 %% \(X(t)\)is an operator valued function \cite{Lax07}.
+
+Finally, we must stress that we do not suggest that clustered designs
+should be abandoned, nor do we support the use of clustered designs in
+any circumstances. We merely investigate an intriguing phenomenon of
+clusterization and seek a deeper understanding for it. Researchers in
+other disciplinesis not viewed as a problem. For example, \cite{lozan}
+suggest in the introduction section that repeated measurements are
+necessary. Our results align with that common wisdom and settle the
+seemed contradiction.
 

### 2024-06-16 latex/numerix.tex
Working towards revision. Modified intro, now working on refactoring code

--- 2023-05-22 arxiv/supplementary.tex
@@
-%\documentclass{siamonline220329}
-\documentclass{amsart}
-%%\documentclass{article}
-\input{definitions}
+\section{Numerical Experiments}
+We provide numerical evidence supporting our conclusions.
 
-\title[Supplementary Material]{Supplementary Material for: Measurement
-  Clusterization in D-optimal Designs for Bayesian Linear Inverse
-  Problems over Hilbert Spaces}
-
-\begin{document}
-\maketitle
-
-\section{The inverse problem of the 1D heat equation}\label{section:example}
-The mathematical foundations of inversion in function spaces can be
-found in \cite{Stuart10}. Our goal is to infer the initial condition of
+\subsection{Heat Equation}
+Our goal is to infer the initial condition of
 the one-dimensional heat equation with a homogeneous Dirichlet
@@
 iid, $\sigma = 0.05$. Measurement locations $\x_j,j=1,\dots,m$ are
-chosen according to the D-optimality criterion of
+chosen according to the D-optimality criterion
 \cite{AlexanderianGloorGhattas14}.
 
-Implementation of this inverse problem can be found in the
-repository \url{https://github.com/yairdaon/OED}, with an optimization
+Implementation of this inverse problem can be found in this 
+\href{https://github.com/yairdaon/OED}{repository}, including an optimization
 routine for finding D-optimal designs. Unfortunately, I was not able
@@
 
-\section{Generalizations of known lemmas}
+\subsection{Simulating Lemma ... }
+In theorem 17 we explicitly construct a D-optimal design in a Hilbert
+space where the base vectors diagonalize $\fwd \prcov\fwd^*$. We
+explicitly implement the algorithm outlined in Lemma
 
-The following lemma is generalized from \cite[Chapter 9, Theorem 4,
-p. 127]{Lax07}
-\begin{lemma}\label{lemma:lax}
-  Let $Y(t)$ be a differentiable operator-valued function. Assume 
-  $I+Y(t)$ is invertible, $Y(t)$ self-adjoint and trace-class. Then
-  \begin{equation*}
-    \frac{\der \log \det (I+Y(t))}{\der t} = \tr{(I+Y(t))^{-1} \dot{Y}(t)}.
-  \end{equation*}
-\end{lemma}
-
-\begin{proof}
-  Consider a differentiable operator-valued function $X(t)$ such that
-  $X(0) = 0$ and $X(t)$ is positive, self-adjoint and trace-class for
-  every $t\in \R$. We denote the eigenvalues of this operator by
-  $\lambda_k(X(t))$ and sometimes drop the dependence on $X(t)$, so
-  $\lambda_k = \lambda_k(X(t))$.  Then $\det (I+X(t))
-  = \prod_{k=1}^{\infty} (1+\lambda_k) < \infty$ and this is finite by
-  the arguments given in \cite{AlexanderianGloorGhattas14}. The full
-  derivative is \begin{align*} \frac{\der \det (I+X(t))}{\der t}
-    % 
-    % 
-    % 
-    &= \sum_{k=1}^{\infty} 
-    \frac{\partial \det (I+X(s))}{\partial (1+\lambda_k)}\Big |_{s=t}
-    \frac{\der (1+\lambda_k)}{\der t} \\
-    % 
-    %
-    %
-    &= \sum_{k=1}^{\infty} \frac{\partial \prod_{l=1}^{\infty}
-      (1+\lambda_l(s))}{\partial (1+\lambda_k)}\Big |_{s=t}
-    \frac{\der (1+\lambda_k)}{\der t} \\
-    %
-    %
-    %
-    &= \sum_{k=1}^{\infty} \prod_{l=1, l\neq k}^{\infty}
-      (1+\lambda_l(s)) \frac{\partial (1+\lambda_k(s))}{\partial (1+\lambda_k)}\Big |_{s=t}
-    \frac{\der (1+\lambda_k)}{\der t} \\
-    %
-    %
-    %    
-    &= \sum_{k=1}^{\infty} \frac{\prod_{l=1}^{\infty}
-      (1+\lambda_l(s))}{(1+\lambda_k)}\Big |_{s=t}
-    \dot{\lambda_k}(X(t)) \\
-    % 
-    % 
-    % 
-    &= \sum_{k=1}^{\infty} \frac{\det (I+X(t))}{1 +\lambda_k} \dot{\lambda_k}(X(t)).
-  \end{align*}
-  The assumption $X(0) = 0$ means $\lambda_k(X(0)) = 0,\ \forall k \geq 1$. Thus:
-  \begin{align*}
-    \frac{\der (I+\det X(t))}{\der t}\Big |_{t=0} 
-    = \sum_{k=1}^{\infty} \dot{\lambda_k}(X(0)) 
-    = \frac{\der }{\der t}\tr{X(0)}
-    = \tr{\dot{X}(0)},
-  \end{align*}
-  where the second equality follows by monotone convergence. 
-  Let $Y(t)$ a trace-class self-adjoint operator such that 
-  $I+Y(t)$ is invertible.
-  Define $X(t)$ via $I+X(t) = (I+Y(0))^{-1/2} (I+Y(t)) (I+Y(0))^{-1/2}$. 
-  We show $X(t)$ satisfies the conditions above. It is trace-class:
-  \begin{align*}
-    \tr{X(t)} = \tr{(I+Y(0))^{-1} (I+Y(t)) - I}
-    \leq \tr{I+Y(t) - I}< \infty,
-  \end{align*}
-  since $Y(t)$ is trace-class. It is also clear that
-  $X(0) = 0$ and $X(t)$ is self-adjoint.
-  $I+Y(t) = (I+Y(0))^{1/2}(I+X(t))(I+Y(0))^{1/2}$, so
-  \begin{align*}
-    \frac{\der \det (I+Y(t))}{\der t}|_{t=0} 
-    &= \det (I+Y(0))\frac{\der \det (I+X(t))}{\der t}\Big |_{t=0} \\
-    % 
-    % 
-    % 
-    &= \det (I+Y(0)) \tr{\dot{X}(0)} \\
-    % 
-    % 
-    % 
-    &= \det (I+Y(0)) \tr{(I+Y(0))^{-1} \dot{Y}(0)}.
-  \end{align*}
-  Consequently, by the one-variable chain rule:
-  \begin{align*}
-    \frac{\der \log \det (I+Y(t))}{\der t}\Big |_{t=0} &=
-    % 
-    % 
-    % 
-    \frac{1}{\det (I+Y(0))}\frac{\der \det (I+Y(t))}{\der t}\Big |_{t=0} \\ 
-    % 
-    % 
-    % 
-    &= \tr{ (I+Y(t))^{-1} \dot{Y}(t)} \big |_{t=0}.
-  \end{align*}
-  There is nothing special about $t_0 = 0$ --- we could have chosen
-  any other $t_0$ instead. Thus, the relation holds for all $t$.
-\end{proof}
-
-The following is a generalization of the Matrix Determinant Lemma to
-Hilbert spaces.
-\begin{lemma}%\label{lemma:MDL}
-  Let $\hil$ a separable Hilbert space, $u,v\in \hil$ and $A: \hil \to
-  \hil$ an invertible linear operator such that $\tr{A-I} <
-  \infty$. Then $\det A$ and $\det A + uv^*$ are well defined and
-  \begin{equation*}
-    \det (A + uv^*) = (1 + \langle A^{-1} u, v \rangle ) \det A,
-  \end{equation*}
-  where $(A + uv^*)w := Aw + \langle v,w \rangle u$.
-\end{lemma}
-\begin{proof}
-  In this proof we rely on definitions and results from
-  \cite{simon1977}. First, consider $B := I + xy^*$ for some $x,y \in
-  \hil$. We construct an eigenbasis for $B$ and use that to show $\det
-  B = 1 + \langle x, y \rangle$. First let $x_1 := x$.  Now, if $x
-  \parallel y$, take $\{x_n \}_{n=2}^{\infty}$ an orthogonal basis for
-  $span\{x_1\} ^{\perp}$. If, on the other hand, $x \not \parallel y$, let
-  \begin{equation*}
-    x_2 := x - \frac{ \langle x, y\rangle}{\|y\|^2}y
-  \end{equation*}
-  and it is easy to verify that $x_2 \perp y$ and $span \{x,y\} = span
-  \{x_1,x_2\}$. Take $\{x_n \}_{n=3}^{\infty}$ an orthogonal basis for
-  $span\{x_1,x_2\} ^{\perp}$. In both cases,
-  \begin{equation*}
-    B x_n =
-    \begin{cases}
-      (1 + \langle x, y \rangle) x_n & n = 1 \\
-      x_n                            & n \neq 1,
-    \end{cases}
-  \end{equation*}
-  and so $\det B = 1 + \langle x, y \rangle$.
-
-  It is easy to verify that $uv^*$ is trace-class and since $\tr{A-I}
-  < \infty$, also $\tr{A + uv^* - I} < \infty$ (sum of two trace-class
-  operators is trace-class). Thus $\det A$ and $\det (A+uv^*)$ are
-  well defined. Let $x:=A^{-1}u$ and $y := v$:
-  \begin{equation*}
-    \det (A + uv^*) = \det A \ \det(I+A^{-1}uv^*) =
-    (1 + \langle A^{-1}u, v \rangle) \det A .
-  \end{equation*}
-\end{proof}
-
-%\bibliographystyle{siamplain}
-\bibliographystyle{amsplain}
-\bibliography{/home/yair/projects/bibtex.bib}
-
-\end{document}
+\subsection{Correlated errors}
+We simulate the 1D heat eq

### 2024-06-16 latex/prelims.tex
Working towards revision. Modified intro, now working on refactoring code

@@
 \(\fwd: \hilp \to \hilo\) the \emph{forward operator}. The forward
-operator \(\fwd\) is assumed linear and strongly smoothing (the heat
-operator of the 1D heat equation introduced in Section
+operator \(\fwd\) is assumed linear with fast decaying singular values
+(the heat operator of the 1D heat equation introduced in Section
 \ref{section:intro} is a prime example). Take a Gaussian prior
@@
 \begin{definition}\label{def:d optimality}
-  We say \(\obs^{\star}\) is \emph{D-optimal} if \(\obs^{\star} =
+  We say \(\opt\) is \emph{D-optimal} if \(\opt =
   \argmax_{\obs} \tar(\obs)\), where entries of \(\obs \in (\hilo^*)^m\)

### 2024-06-16 latex/vanishing.tex
Working towards revision. Modified intro, now working on refactoring code

@@
       V :&= \fwd \\
-      C :&= \sigma^{-2} (\obs^*\obs+\eps I),\ \eps > 0
+      C :&= \sigma^{-2} (\obs^*\obs+\nu I),\ \nu > 0
     \end{split}
@@
     \begin{split}
-    &\fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* \\
-    &\ \ \ \ \ = \fwd ( \prcov - \prcov \fwd^* ( \sigma^2(\obs^*\obs + \eps I)^{-1} + \fwd \prcov \fwd^* )^{-1} \fwd \prcov ) \fwd^*. \\
+    &\fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\nu I) \fwd )^{-1}\fwd^* \\
+    &\ \ \ \ \ = \fwd ( \prcov - \prcov \fwd^* ( \sigma^2(\obs^*\obs + \nu I)^{-1} + \fwd \prcov \fwd^* )^{-1} \fwd \prcov ) \fwd^*. \\
       %
       %
-      %% &= X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X
+      %% &= X - X(\sigma^2Y_{\nu}^{-1} + X)^{-1}X
   %%   \end{split}
@@
     X :&= \fwd\prcov \fwd^*, \\
-    Y_{\eps} :&= \obs^*\obs + \eps I,
+    Y_{\nu} :&= \obs^*\obs + \nu I,
   \end{align*}
-  and observe that $Y_{\eps}$ is invertible. Substitute $X, Y_{\eps}$
+  and observe that $Y_{\nu}$ is invertible. Substitute $X, Y_{\nu}$
   into the RHS of \eqref{eq:first}:
    \begin{equation}\label{eq:second}
-     \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* = X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X
+     \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\nu I) \fwd )^{-1}\fwd^* = X - X(\sigma^2Y_{\nu}^{-1} + X)^{-1}X
    \end{equation}
 
-   Note that $X + \sigma^2 Y_{\eps}^{-1}$ is invertible, as the sum of
+   Note that $X + \sigma^2 Y_{\nu}^{-1}$ is invertible, as the sum of
    two positive definite operators. Now, let
@@
        A :&= X^{-1}, \\
-       C :&= \sigma^2Y_{\eps}^{-1}, \\
+       C :&= \sigma^2Y_{\nu}^{-1}, \\
        U :&= I, \\
@@
     \begin{split}
-      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* &= X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X \\
+      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\nu I) \fwd )^{-1}\fwd^* &= X - X(\sigma^2Y_{\nu}^{-1} + X)^{-1}X \\
       %
@@
       %
-      &= (X^{-1} + \sigma^{-2}Y_{\eps})^{-1} \\
+      &= (X^{-1} + \sigma^{-2}Y_{\nu})^{-1} \\
       %
@@
       %
-      &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \eps I))^{-1}.
+      &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \nu I))^{-1}.
     \end{split}
@@
 
-  We conclude that $\forall \eps > 0$
+  We conclude that $\forall \nu > 0$
   \begin{align*}
     \begin{split}
-      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* 
-     &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \eps I))^{-1}.
+      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\nu I) \fwd )^{-1}\fwd^* 
+     &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \nu I))^{-1}.
     \end{split}
   \end{align*}
-  Letting $\eps \to 0$ completes the proof.
+  Letting $\nu \to 0$ completes the proof.
 \end{proof}
@@
 
-\char*
+\begin{theorem}\label{thm:char}
+  Let:
+  \begin{itemize}
+    \item The D-optimality design criterion
+    \cite{AlexanderianGloorGhattas14}:
+    \begin{align*}
+      \begin{split}
+        \tar(\obs) %:&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
+        % 
+        % 
+        % 
+        &= \frac12 \log \det ( I + \sigma^{-2} \prcov^{1/2} \fwd ^*
+        \obs^* \obs \fwd \prcov^{1/2}), 
+      \end{split}
+    \end{align*}
+  \item \(\opt\) a D-optimal design operator
+    \begin{equation*}
+      \opt = \argmax_{\|\meas_j\| = 1, j=1,\dots,m}\tar(\obs),
+    \end{equation*}
+  \item \(\{\lambda_i\}_{i=1}^\infty\) eigenvalues of
+    \(\fwd\prcov\fwd^*\) in decreasing order of magnitude.
+    %% \item \(\{\ev_i\}_{i=1}^\infty\) their corresponding eigenvectors.
+  \item \(\{\eta_i\}_{i=1}^\infty\) eigenvalues of \(\opt^*\opt\).
+
+  \end{itemize}
+
+  Then:
+  \begin{enumerate}
+  \item  \(\tr{\opt^*\opt} = m\).
+  \item \(\opt^*\opt\) and \(\fwd\prcov\fwd^*\) are simultaneously
+    diagonalizable.
+  \item \(k := \rank \opt^*\opt \leq m\) and
+    \begin{equation*}
+      \tar(\opt) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i). %= \frac12 \sum_{i=1}^{m} \log (1 + \sigma^{-2}\lambda_i\eta_i).
+    \end{equation*}
+  %% \item 
+  %%   \begin{equation*}
+  %%     k = \argmax \left \{ k:\lambda_k^{-1} < \sigma^{-2}\frac{m}{k} + \frac{1}{k} \sum_{j=1}^{k}
+  %%     \lambda_j^{-1} \right \}.
+  %%   \end{equation*}
+  \item
+    \begin{equation*}
+        \eta_i = \begin{cases}
+          \frac{m}{k} - \sigma^2 \lambda_i^{-1} + \sigma^2 \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1} & 1 \leq i \leq k \\
+          0 & i > k 
+        \end{cases}.
+    \end{equation*}
+  \item The covariance of the pushforwad \(\fwd_{*} \postopt\) is \(\left
+    ( (\fwd \prcov \fwd^*)^{-1} + \sigma^{-2} \opt^*\opt \right
+    )^{-1}\) and its eigenvalues are
+    \begin{equation*}
+      \theta_i =
+      \begin{cases}
+        \left(\frac{\sum_{j=1}^k \lambda_j^{-1} + \sigma^{-2}m}{k} \right )^{-1} & i \leq k \\
+        \lambda_i &  i > k 
+      \end{cases}
+    \end{equation*}
+  \end{enumerate}
+\end{theorem}
 
@@
   \ref{prop:kkt}, with the caveat that we did not show that finding
-  $\obs$ so that $\obs^*\obs$ has the desired eigenvalues is
+  $\opt$ so that $\opt^*\opt$ has the desired eigenvalues is
   feasible. To this end, we utilize Lemma \ref{lemma:free}: let $M =
-  \diag(\eta_1, \dots, \eta_k)$, diagonal with respect to the first $k$
-  eigenvectors of $\fwd \prcov \fwd^*$. We take $\obs := A$ from the
+  \diag(\eta_1, \dots, \eta_k)$, diagonal with respect to the first
+  $k$ eigenvectors of $\fwd \prcov \fwd^*$. We take $\opt := A$ from
   Lemma \ref{lemma:free}.

### 2024-06-16 main.py

@@ -1,0 +1,84 @@
+import numpy as np
+from matplotlib import pyplot as plt
+
+from src.probability import Prior, Posterior
+from src.forward import Heat
+from src.observations import PointObservation, DiagObservation
+from src.multiplier import FourierMultiplier
+
+
+
+def main():
+
+    ## DST / DCT - discrete sine / cosine transform. 
+    ## Corresponds to a homogeneous Dirichlet/ Neumann boundary condition.
+    ## Clusterization also happens with Neumann boundary, I just chose to present 
+    ## Dirichlet boundary
+    transform = 'dst'
+    delta = 0. if transform == 'dst' else 0.5
+    boundary = 'Dirichlet' if transform == 'dst' else 'Neumann'
+    
+    N = 200
+    L = 1
+    time = 3e-2
+    alpha = 1.
+    gamma = -1.
+    model_error = True
+    sig = 5e-2
+    
+    fwd = Heat(N=N, L=L, transform=transform, alpha=alpha, time=time)
+    pr = Prior(N=N, L=L, transform=transform, gamma=gamma, delta=delta)
+    post = Posterior(fwd=fwd,
+                     prior=pr,
+                     sigSqr=sig**2,
+                     L=L,
+                     N=N,
+                     transform=transform,
+                     model_error=True)
+
+    dic = {}
+    print("model error", model_error, 'transform', transform)
+    for m in range(3, 8):
+        print(m, end=' ')
+        res = post.optimize(m=m, n_iterations=20, n_jobs=-1)
+        design = res['x']
+        dic[m] = design
+        print(design)
+
+
+    fs = 24
+    fig, ax = plt.subplots(figsize=(10,5))
+
+    for m, array in dic.items():
+        colors = iter(list("brgkmcy"))
+
+        vals = np.repeat(m, len(array))
+        ax.scatter(array, vals, s=0)
+        for i, val in enumerate(array):
+            ax.annotate(str(i+1),
+                        xy=(val, m),
+                        ha='center', 
+                        va='center',
+                        fontsize=fs,
+                        color=next(colors))
+            
+
+    ax.set_xlabel('Measurement Location', fontsize=fs)
+    ax.set_ylabel('No. of Measurements', fontsize=fs)
+    ax.set_yticks(list(dic.keys()))
+    ax.set_xlim(0,1)
[…]

### 2024-06-16 src/forward.py

@@ -3,15 +3,35 @@ import numpy as np
 from src.multiplier import FourierMultiplier
 
 class Heat(FourierMultiplier):
-    """Run forward heat equation for some time"""
+    """Run forward heat equation for some time.
+    The heat equation is defined as:
+        u_t = \alpha \Delta u
+
+    with various boundary conditions. We utilize the homogeneous Dirichlet or Neumann boundary.
+    Homogeneous Dirichlet boundary conditions require the "heat" at the edges to be zero at all times.
+    Homogeneous Neumann boundary conditions require the derivative of the "heat" in the direction orthogonal to
+    the boundary to be zero at all times.
+    """
 
     def __init__(self,
                  time=3e-3,
                  alpha=0.6,
                  **kwargs):
+
+        """
+        Parameters:
+
+        time: float
+            run time of the heat equation, in arbitrary units.
+        alpha: float
+            coefficient of the laplace operator in the heat equation.
+        """
         super().__init__(**kwargs)
         self.alpha = alpha
         self.time = time
+
+        ## In the frequency domain, running the heat equation for the given time amounts to
+        ## multiplying by this Fourier multiplier
         self.multiplier = np.exp(-self.alpha * np.pi ** 2 * self.time * self.freqs ** 2)
 
     def __str__(self):

### 2024-06-16 src/multiplier.py

@@ -7,7 +7,26 @@ from functools import partial
 
 
 class Operator(LinearOperator):
-    def __init__(self, N=500, L=1, size=None, dtype=None, **kwargs):
+    """An operator, that extends scipy's LinearOperator like most objects
+    we are dealing with.
+    """ 
+
+    def __init__(self,
+                 N=500,
+                 L=1,
+                 size=None,
+                 dtype=None,
+                 **kwargs):
+        """
+        Parameters
+        ----------
+        N : int
+        Number of spatial discretization points
+        L : float
+        Length of spatial domain
+        """
+        
+        
         shape = (N, N) if size is None else (size, N)
         super().__init__(dtype=np.dtype(dtype), shape=shape)
         self.L = L
@@ -17,12 +36,28 @@ class Operator(LinearOperator):
         self.sqrt_h = np.sqrt(self.h)
 
     def norm(self, v, axis=-1):
+        """Return the L2 norm of a *function* over the computational
+        domain. Hence the norm calculation is: squaring v, integrating
+        over domain and taking square root.
+
+        """
         return np.linalg.norm(v, axis=axis) * self.sqrt_h
 
 
 class FourierMultiplier(Operator):
+    """We solve the heat equation from scratch in its eigenspace. In this
+    space the Laplace operator is a Fourier multiplier, so applying
+    and inverting it are easy.
 
-    def __init__(self, transform='dct', **kwargs):
+    We implement three such eigenspaces, each based on a different transform:
+    fft - Fast Fourier Transform. Implements *periodic* boundary condition.
+    dst - Discrete Sine Transform. Implements a homogeneous Dirichlet boundary condition.
+    dct - Discrete Cosine Transform. Implements a homogeneous Neumann boundary condition.
+    """
+
+    def __init__(self,
+                 transform='dct',
+                 **kwargs):
         dtype = 'complex' if transform == 'fft' else 'float'
         super().__init__(**kwargs, dtype=dtype)
         self.multiplier = None # Need to implement this on particular case
@@ -37,9 +72,15 @@ class FourierMultiplier(Operator):
             
     @property
     def specs(self):
+        """Convenieve function to move copy around specifications of a
+        Fourier Multiplier"""
         return {"N": self.N, "L": self.L, "transform":self.transform}
 
     def to_freq_domain(self, x, axis=-1):
[…]

### 2024-06-16 src/observations.py

@@ -10,11 +10,14 @@ class Observation(FourierMultiplier):
 
     
 class PointObservation(Observation):
-    def __init__(self, measurements=[], **kwargs):
+    """A point obeservation, at arbitrary location, calculated via the frequency domain."""
+
+    def __init__(self,
+                 measurements=[],
+                 **kwargs):
         super().__init__(**kwargs, size=len(measurements))
         self.measurements = np.array(measurements)
         self.multiplier = self.normalized_block(self.measurements).T
-        
 
     def __str__(self):
         return 'Point observations at ' + ', '.join([f'{me:.4f}' for me in self.measurements])

### 2024-06-16 src/probability.py

@@ -12,8 +12,22 @@ from src.forward import Heat
 
 
 class Prior(FourierMultiplier):
-    """ Implement a negative Laplacian prior Delta^{gamma}, where gamma < 0"""
-    def __init__(self, gamma=-0.6, delta=0.5, **kwargs):
+    """ Implement a prior with the following *precision* \delta I +  \Delta^{-gamma}, where gamma < 0"""
+    def __init__(self,
+                 gamma=-0.6,
+                 delta=0.5,
+                 **kwargs):
+        """
+        Parameters:
+
+        gamma: float
+            the power of the Laplacian is the prior precision.
+        delta: float
+            a regularizing constant used for the case of Homogeneous Neumann boundary condition, so that
+            \delta I + \Delta^{-gamma} is invertible.
+
+        """
+
         assert gamma < 0
         super().__init__(**kwargs)
         inv_mult = np.power(np.pi**2 * self.freqs**2, -gamma) + delta
@@ -40,7 +54,7 @@ class Prior(FourierMultiplier):
         return u0
 
     def inverse(self, v):
-        ''' A bit of a hack - change multiplier to inverse multiplier and back'''
+        '''Implement the inverse of  A bit of a hack - change multiplier to inverse multiplier and back'''
         tmp = self.multiplier
         self.multiplier = self.inv_mult
         ret = self(v)
@@ -57,17 +71,27 @@ class Prior(FourierMultiplier):
 
         
 class Posterior(FourierMultiplier):
-    ''' Based on Andrew Stuart's Acta Numerica paper 2.16a and 2.16b'''
+    '''Posterior Gaussian. Formula for posterior covariance/ precision is based on
+    equations 2.16a and 2.16b from Andrew Stuart's Acta Numerica (2010) paper.'''
 
     def __init__(self,
                  fwd=None,
                  prior=None,
                  sigSqr=0.1,
+                 model_error=False,
                  **kwargs):
+        """
+        Parameters:
+        fwd: the forward operator. We implement the heat equation in 1D
+        prior: prior operator. We implement a Laplacian-like prior.
+
+        """
+
         super().__init__(**kwargs)
         self.fwd = Heat(**self.specs) if fwd is None else fwd
[…]

### 2024-06-16 src/observations.py
Removed lots of shit

@@ -38,56 +38,56 @@ class PointObservation(Observation):
         O = self.multiplier
         return np.linalg.eigh(O @ O.conjugate().T)[0] * self.h
 
-    def pow(self):
-        return np.sum(self.eigenvalues()**2)
+    # def pow(self):
+    #     return np.sum(self.eigenvalues()**2)
 
-class DiagObservation(Observation):
-    def __init__(self, multiplier, random_U=False, **kwargs):
-        super().__init__(**kwargs, size=len(multiplier))
-        self.multiplier = np.zeros(self.N)
-        self.multiplier[:len(multiplier)] = multiplier
-        # np.fill_diagonal(self.multiplier, multiplier)
-        assert np.all(self.multiplier.imag < 1e-14)
-        assert np.all(self.multiplier.real >= 0)
-
-        # if random_U:
-        #     # From https://stackoverflow.com/questions/38426349/how-to-create-random-orthonormal-matrix-in-python-numpy
-        #     H = np.random.randn(self.singular_values.shape[0], self.singular_values.shape[0])
-        #     Q, R = np.linalg.qr(H)
-        #     # Q = np.vdot(Q, np.diag(np.sign(np.diag(R))))
-        #     self.multiplier = np.dot(Q, self.multiplier)
-        #     self.U = Q
-        # else:
-        #     self.U = np.eye(len(singular_values))
-
-    def _matvec(self, v):
-        return self.to_freq_domain(v) * self.multiplier
-
-    def _matmat(self, M):
-        M_hat = self.to_freq_domain(M, axis=0)
-        return np.einsum('i, ij -> ij', self.multiplier, M_hat)
-
-    @property
-    def matrix(self):
-        return self(np.eye(self.N))
-
-    def singular_values(self):
-        return self.multiplier * self.sqrt_h
-
-    # def _matvec(self, v):
-    #     v_hat = self.to_freq_domain(v)
-    #     return v_hat * self.multiplier
-
-    # def _matmat(self, M):
-    #     """M = PDP*, where P* == to_freq_domain, D == (diagonal) multiplier
-    #     P == to_time_domain. We find OM """
-
-    #     M_hat = self.to_freq_domain(M, axis=0)
-    #     DM_hat = np.einsum('i, ij-> ij', self.multiplier, M_hat)
-    #     return DM_hat
-
-    
-    def __str__(self):
[…]

### 2024-06-16 src/probability.py

@@ -7,7 +7,7 @@ from matplotlib import pyplot as plt
 import pdb
 
 from src.multiplier import FourierMultiplier
-from src.observations import PointObservation, DiagObservation
+from src.observations import PointObservation#, DiagObservation
 from src.forward import Heat
 
 
@@ -78,7 +78,7 @@ class Posterior(FourierMultiplier):
                  fwd=None,
                  prior=None,
                  sigSqr=0.1,
-                 model_error=False,
+                 model_error=0,
                  **kwargs):
         """
         Parameters:
@@ -97,26 +97,26 @@ class Posterior(FourierMultiplier):
         assert np.all(self.C_sqrt_fwd.real >= 0)
         
 
-    def make_diagonal(self, m, k):
-        eigs = self.sigSqr / self.C_sqrt_fwd[:k]**2
-        uniform = np.mean(eigs) + m / k
-        return uniform - eigs
-
-    def make_optimal_diagonal(self, m):
-        k = 1
-        while True:
-            eta = self.make_diagonal(m, k)
-            if np.any(eta <= 0):
-                break
-            self.optimal_diagonal_O = np.sqrt(eta)
-            k += 1
-          
-        self.optimal_diagonal_O_matrix = np.zeros((m, self.N))
-        np.fill_diagonal(self.optimal_diagonal_O_matrix, self.optimal_diagonal_O)
-
-        power = np.sum(self.optimal_diagonal_O**2)
-        assert abs(power - m) < 1e-12, (power, m)
-        return self.optimal_diagonal_O
+    # def make_diagonal(self, m, k):
+    #     eigs = self.sigSqr / self.C_sqrt_fwd[:k]**2
+    #     uniform = np.mean(eigs) + m / k
+    #     return uniform - eigs
+    #
+    # def make_optimal_diagonal(self, m):
+    #     k = 1
+    #     while True:
+    #         eta = self.make_diagonal(m, k)
+    #         if np.any(eta <= 0):
+    #             break
+    #         self.optimal_diagonal_O = np.sqrt(eta)
+    #         k += 1
+    #
+    #     self.optimal_diagonal_O_matrix = np.zeros((m, self.N))
+    #     np.fill_diagonal(self.optimal_diagonal_O_matrix, self.optimal_diagonal_O)
+    #
+    #     power = np.sum(self.optimal_diagonal_O**2)
+    #     assert abs(power - m) < 1e-12, (power, m)
+    #     return self.optimal_diagonal_O
     #
[…]

### 2024-06-16 clusterization.py
Refactored a lot

@@ -1,0 +1,104 @@
+import numpy as np
+from matplotlib import pyplot as plt
+
+from src.probability import Prior, Posterior
+from src.forward import Heat
+from src.observations import PointObservation
+
+
+def main():
+
+    sig = 5e-2 ## Observation error amplitude
+    N = 200 ## Number of spatial discretizatio points
+    L = 1 ## Length of computational domain
+    time = 3e-2 ## Time for heat dissipation in arbitrary units.
+    alpha = 1. ## Coefficient of Laplacian in heat equation
+    gamma = -1. ## Exponent of Laplacian in prior. 
+
+    ## Choose one, comment out the other
+    transform = 'dct' ## dct - Discrete Cosine Transform. Corresponds
+                      ## to homogenoeus Neumann boundary condition.
+    transform = 'dst' ## dst - Discrete Sine Transform. Corresponds
+                      ## homogenoeus Dirichlet boundary condition.
+
+    ## We need to regularize the Laplacian so it is invertible and can
+    ## function as a useful prior. Hence we add the constant delta
+    ## when employing homogeneous Neumann boundary condition.
+    delta = 0. if transform == 'dst' else 0.5
+    
+    ## Avoid clusterization by increasing this to e.g. 4
+    model_error = 0
+
+    ## Forward model - the heat equation with BC chosen according to transform
+    fwd = Heat(N=N,
+               L=L,
+               transform=transform,
+               alpha=alpha,
+               time=time)
+
+    ## Prior - with precision \delta
+    pr = Prior(N=N,
+               L=L,
+               transform=transform,
+               gamma=gamma,
+               delta=delta)
+
+    ## Posterior. We don't care about data, so no data is involved,
+    ## only observation operators. This effectively implements
+    ## calculating the D-optimality criterion design and searching for
+    ## a D-optimal design.
+    post = Posterior(fwd=fwd,
+                     prior=pr,
+                     sigSqr=sig**2,
+                     L=L,
+                     N=N,
+                     transform=transform,
+                     model_error=model_error)
+
+
+    dic = {}
+
+    ## Form  number of observations in this range
+    for m in range(2, 6):
+
+        ## Calculate an optimal design
+        design = post.optimize(m=m, n_iterations=25)['x']
+
+        ## And store it in dic
+        dic[m] = design
+
+        ## Print it so we see
+        print(transform, m, design)
+        
+
[…]

### 2024-06-16 main.py

@@ -1,84 +1,0 @@
-import numpy as np
-from matplotlib import pyplot as plt
-
-from src.probability import Prior, Posterior
-from src.forward import Heat
-from src.observations import PointObservation, DiagObservation
-from src.multiplier import FourierMultiplier
-
-
-
-def main():
-
-    ## DST / DCT - discrete sine / cosine transform. 
-    ## Corresponds to a homogeneous Dirichlet/ Neumann boundary condition.
-    ## Clusterization also happens with Neumann boundary, I just chose to present 
-    ## Dirichlet boundary
-    transform = 'dst'
-    delta = 0. if transform == 'dst' else 0.5
-    boundary = 'Dirichlet' if transform == 'dst' else 'Neumann'
-    
-    N = 200
-    L = 1
-    time = 3e-2
-    alpha = 1.
-    gamma = -1.
-    model_error = True
-    sig = 5e-2
-    
-    fwd = Heat(N=N, L=L, transform=transform, alpha=alpha, time=time)
-    pr = Prior(N=N, L=L, transform=transform, gamma=gamma, delta=delta)
-    post = Posterior(fwd=fwd,
-                     prior=pr,
-                     sigSqr=sig**2,
-                     L=L,
-                     N=N,
-                     transform=transform,
-                     model_error=True)
-
-    dic = {}
-    print("model error", model_error, 'transform', transform)
-    for m in range(3, 8):
-        print(m, end=' ')
-        res = post.optimize(m=m, n_iterations=20, n_jobs=-1)
-        design = res['x']
-        dic[m] = design
-        print(design)
-
-
-    fs = 24
-    fig, ax = plt.subplots(figsize=(10,5))
-
-    for m, array in dic.items():
-        colors = iter(list("brgkmcy"))
-
-        vals = np.repeat(m, len(array))
-        ax.scatter(array, vals, s=0)
-        for i, val in enumerate(array):
-            ax.annotate(str(i+1),
-                        xy=(val, m),
-                        ha='center', 
-                        va='center',
-                        fontsize=fs,
-                        color=next(colors))
-            
-
-    ax.set_xlabel('Measurement Location', fontsize=fs)
-    ax.set_ylabel('No. of Measurements', fontsize=fs)
-    ax.set_yticks(list(dic.keys()))
-    ax.set_xlim(0,1)
[…]

### 2024-06-16 src/observations.py

@@ -10,7 +10,10 @@ class Observation(FourierMultiplier):
 
     
 class PointObservation(Observation):
-    """A point obeservation, at arbitrary location, calculated via the frequency domain."""
+    """A point obeservation, at arbitrary location not necessarily on the
+    computational grid. Calculations carried out in the frequency
+    domain.
+    """
 
     def __init__(self,
                  measurements=[],
@@ -30,64 +33,11 @@ class PointObservation(Observation):
         M_hat = self.to_freq_domain(M, axis=0)
         return np.einsum('ij, jk -> ik', self.multiplier, M_hat)
 
-    @property
-    def matrix(self):
-        return self.to_freq_domain_from_right(self.multiplier)
+    # @property
+    # def matrix(self):
+    #     return self.to_freq_domain_from_right(self.multiplier)
 
     def eigenvalues(self):
         O = self.multiplier
         return np.linalg.eigh(O @ O.conjugate().T)[0] * self.h
-
-    # def pow(self):
-    #     return np.sum(self.eigenvalues()**2)
-
-# class DiagObservation(Observation):
-#     def __init__(self, multiplier, random_U=False, **kwargs):
-#         super().__init__(**kwargs, size=len(multiplier))
-#         self.multiplier = np.zeros(self.N)
-#         self.multiplier[:len(multiplier)] = multiplier
-#         # np.fill_diagonal(self.multiplier, multiplier)
-#         assert np.all(self.multiplier.imag < 1e-14)
-#         assert np.all(self.multiplier.real >= 0)
-#
-#         # if random_U:
-#         #     # From https://stackoverflow.com/questions/38426349/how-to-create-random-orthonormal-matrix-in-python-numpy
-#         #     H = np.random.randn(self.singular_values.shape[0], self.singular_values.shape[0])
-#         #     Q, R = np.linalg.qr(H)
-#         #     # Q = np.vdot(Q, np.diag(np.sign(np.diag(R))))
-#         #     self.multiplier = np.dot(Q, self.multiplier)
-#         #     self.U = Q
-#         # else:
-#         #     self.U = np.eye(len(singular_values))
-#
-#     def _matvec(self, v):
-#         return self.to_freq_domain(v) * self.multiplier
-#
-#     def _matmat(self, M):
-#         M_hat = self.to_freq_domain(M, axis=0)
[…]

### 2024-06-16 src/probability.py

@@ -40,8 +40,10 @@ class Prior(FourierMultiplier):
         # self.ind = ind
 
     def sample(self, return_coeffs=False, n_sample=1):
-        """ Generate a sample and return its coefficients if needed. This effectively uses the
-        Karhunen-Loeve expansion"""
+        """Generate a sample and return its coefficients if needed. This
+        effectively uses the Karhunen-Loeve expansion
+
+        """
         assert n_sample > 0
         coeffs = self.normal(n_sample)
 
@@ -54,7 +56,8 @@ class Prior(FourierMultiplier):
         return u0
 
     def inverse(self, v):
-        '''Implement the inverse of  A bit of a hack - change multiplier to inverse multiplier and back'''
+        '''Implement the inverse of A bit of a hack - change multiplier to
+        inverse multiplier and back'''
         tmp = self.multiplier
         self.multiplier = self.inv_mult
         ret = self(v)
@@ -91,68 +94,21 @@ class Posterior(FourierMultiplier):
         self.fwd = Heat(**self.specs) if fwd is None else fwd
         self.prior = Prior(**self.specs) if prior is None else prior
         self.sigSqr = sigSqr
-        self.model_error = True
+        self.model_error = model_error
         self.C_sqrt_fwd = np.sqrt(self.prior.multiplier) * self.fwd.multiplier
         assert np.all(np.abs(self.C_sqrt_fwd.imag) < 1e-12)
         assert np.all(self.C_sqrt_fwd.real >= 0)
         
 
-    # def make_diagonal(self, m, k):
-    #     eigs = self.sigSqr / self.C_sqrt_fwd[:k]**2
-    #     uniform = np.mean(eigs) + m / k
-    #     return uniform - eigs
-    #
-    # def make_optimal_diagonal(self, m):
-    #     k = 1
-    #     while True:
-    #         eta = self.make_diagonal(m, k)
-    #         if np.any(eta <= 0):
-    #             break
-    #         self.optimal_diagonal_O = np.sqrt(eta)
-    #         k += 1
-    #
-    #     self.optimal_diagonal_O_matrix = np.zeros((m, self.N))
-    #     np.fill_diagonal(self.optimal_diagonal_O_matrix, self.optimal_diagonal_O)
-    #
-    #     power = np.sum(self.optimal_diagonal_O**2)
-    #     assert abs(power - m) < 1e-12, (power, m)
-    #     return self.optimal_diagonal_O
-    #
[…]

### 2024-06-16 clusterization.py
More refactoring, moved movies to a different file and added eigenvector plots to clusterization.py

@@ -98,7 +98,36 @@ def main():
     plt.tight_layout()
 
     plt.savefig(f"latex/{transform}_modelError{model_error}.pdf")
+    #plt.show()
+
+    ####################################
+    #### Plot scaled eigenvectors ######
+    ####################################
+    
+    ## Number of measurements
+    m = 4
+    design = dic[m]
+    
+    ## Show eigenvectors and design for last design
+    plt.close('all')
+    fig = plt.figure(figsize=(8,4))
+    fs = 18
+    lss = ['solid', 'dotted', 'dashed', 'dashdot']
+    vals = np.zeros(m)
+    plt.scatter(design, vals)
+    x = fwd.x
+    for i,ls in enumerate(lss):
+        print(i, ls)
+        ev = fwd.eigenvector(i)
+        lam = np.sqrt(fwd.multiplier[i])
+        plt.plot(x, ev * lam, label=i+1, ls=ls)
+    plt.plot([0,1], [0,0], ls='-', color='k', alpha=0.5)    
+    plt.xlabel(r"$x \in \Omega$", fontsize=fs)
+    plt.legend(title='eigenvector, weighted')
+    plt.tight_layout()
+    plt.savefig(f"latex/eigenvectors_{transform}.pdf")
     plt.show()
 
-    
+
 main()
+

### 2024-06-16 src/multiplier.py

@@ -7,8 +7,9 @@ from functools import partial
 
 
 class Operator(LinearOperator):
-    """An operator, that extends scipy's LinearOperator like most objects
-    we are dealing with.
+    """An operator class, that extends scipy's LinearOperator like most
+    objects we are dealing with.
+
     """ 
 
     def __init__(self,
@@ -45,9 +46,9 @@ class Operator(LinearOperator):
 
 
 class FourierMultiplier(Operator):
-    """We solve the heat equation from scratch in its eigenspace. In this
-    space the Laplace operator is a Fourier multiplier, so applying
-    and inverting it are easy.
+    """We solve the heat equation from scratch in frequency domain. In
+    this space the Laplace operator is a Fourier multiplier, so
+    applying and inverting it are easy.
 
     We implement three such eigenspaces, each based on a different transform:
     fft - Fast Fourier Transform. Implements *periodic* boundary condition.
@@ -72,13 +73,16 @@ class FourierMultiplier(Operator):
             
     @property
     def specs(self):
-        """Convenieve function to move copy around specifications of a
-        Fourier Multiplier"""
+        """Convenience function to move around specifications of a Fourier
+        Multiplier.
+
+        """
         return {"N": self.N, "L": self.L, "transform":self.transform}
 
     def to_freq_domain(self, x, axis=-1):
         """Change a spatial function to the frequency domain, depending on
         which transform we use (fft, dst or dct).
+
         """
         
         if self.transform == 'dct':
@@ -88,14 +92,11 @@ class FourierMultiplier(Operator):
         elif self.transform == 'dst':
              return fft.dst(x, norm='ortho', type=2, axis=axis) * self.sqrt_h
 
-         
-    def to_freq_domain_from_right(self, x):
-        return self.to_time_domain(x.conjugate(), axis=1).conjugate() * self.h
-
-    
+             
     def to_time_domain(self, x, axis=-1):
         """Change a frequency function to the time domain, depending on which
         transform we use (fft, dst or dct).
+
         """
         
         if self.transform == 'dct':
@@ -146,9 +147,7 @@ class FourierMultiplier(Operator):
         """ Create the ith eigenvector, as an array"""
         return self.eigenfunction(i)(self.x)
 
-    # def coeff2u(self, coeff):
-    #     return self.to_time_domain(coeff) / self.sqrt_h
-
+    
[…]

### 2024-06-16 code/main.py
Add README.md and a bunch of docs

@@ -1,112 +1,0 @@
-import os
-import numpy as np
-import pandas as pd
-from joblib import Parallel, delayed
-from matplotlib import pyplot as plt
-from src.observations import PointObservation
-from scipy.stats import gaussian_kde
-from pdb import set_trace
-
-from src.forward import Heat
-from src.probability import Posterior, Prior
-
-
-# plt.rcParams['figure.figsize'] = [8, 4]
-# plt.rcParams['figure.dpi'] = 200  # 200 e.g. is really fine, but slower
-# plt.rcParams['axes.facecolor'] = 'black'
-# plt.rcParams['figure.facecolor'] = 'black'
-# plt.rcParams['text.color'] = 'white'
-
-def main(n_jobs=6):
-    L, N = 1, 100
-    ms = [1,3, 20]
-    transforms = ['fft']
-    # transforms.append('dct')
-    # fig1, axes1 = plt.subplots(figsize=(30, 16), nrows=len(transforms), ncols=len(ms))
-    # fig2, axes2 = plt.subplots(figsize=(30, 16), nrows=len(transforms), ncols=len(ms))
-    # figs = [fig1, fig2]
-    # axess = [axes1, axes2]
-
-    for col, m in enumerate(ms):
-
-        for transform in transforms:#, fig, axes in zip(transforms, figs, axess):
-            print('\nm =', m, transform)
-
-            prior = Prior(gamma=-1.1, N=N, L=L, transform=transform)
-            forward = Heat(time=3e-2, alpha=0.6, N=N, L=L, transform=transform)
-            post = Posterior(fwd=forward, prior=prior, sigSqr=1, N=N, L=L, transform=transform)
-            # set_trace()
-            # sample = prior.sample().ravel()
-        
-            # Optimal diagonal
-            post.make_optimal_diagonal(m)
-            diagonal_utility = post.diag_utility(post.optimal_diagonal_O)
-            # ax.scatter(diagonal_utility, kde(diagonal_utility), label='optimal diagonal')
-            print(f"utility: Diag={diagonal_utility:3.3f}", end=' ')
-
-            # # Random point measurements
-            # measurements = np.random.uniform(low=0, high=L, size=(100, m))
-            # utilities = Parallel(n_jobs=n_jobs)(delayed(post.minimization_point)(x) for x in measurements)
-            # utilities = -np.array(utilities)
[…]

### 2024-06-16 code/playground.py

@@ -1,94 +1,0 @@
-import numpy as np
-from matplotlib import pyplot as plt
-from scipy.interpolate import interp1d
-from pdb import set_trace as stahp
-
-# from importlib import reload
-# import src.probability
-# reload(src.probability)
-# reload(src.multiplier)
-# reload(src.observations)
-from src.probability import Prior, Posterior
-from src.multiplier import FourierMultiplier
-from src.observations import PointObservation, DiagObservation
-from src.forward import Heat
-
-class TestPrior(Prior):
-    def __init__(self,
-                 eigenvalues,
-                 **kwargs):
-        super().__init__(**kwargs)
-        eigenvalues = np.array(eigenvalues)
-        self.gamma = None
-        self.delta = None
-        self.multiplier = np.zeros(self.N)
-        self.multiplier[:eigenvalues.size] = eigenvalues
-        self.inv_mult = np.full(self.N, np.inf)
-        self.inv_mult[:eigenvalues.size] = 1/eigenvalues
-        
-
-class TestForward(FourierMultiplier):
-    def __init__(self,
-                 eigenvalues,
-                 **kwargs):
-            
-        super().__init__(**kwargs)
-        eigenvalues = np.array(eigenvalues)
-        self.multiplier = np.zeros(self.N)
-        self.multiplier[:eigenvalues.size] = eigenvalues
-
-    def __str__(self):
-        return f'Test Forward operator'
-
-def main():
-    n = 3
-    N = 200
-    L = 1
-    sigSqr = 1e-1
-    prior_eigs = np.ones(n)#np.power(2., -np.arange(n))
-    fwd_eigs = np.ones(n)
-    ms = [1, 2, 3, 5, 7, 10, 12]
-    transforms = ['dct', 'fft']
-    for transform in transforms:
-        print('\n\n')
-        print(transform)
-        for m in ms:
-            print(m)
-            prior = TestPrior(N=N, L=L, transform=transform, eigenvalues=prior_eigs)
-            fwd = TestForward(N=N, L=L, transform=transform, eigenvalues=fwd_eigs)
-            post = Posterior(prior=prior, fwd=fwd, N=N, L=L, transform=transform, sigSqr=sigSqr)
-            print('yasss')
-            diagonal_O = post.make_optimal_diagonal(m)
-            print('naaa')
-            # print(diagonal_O)
-            eigs = prior.multiplier ** 2 * fwd.multiplier
[…]

### 2024-06-16 movies.py

@@ -1,0 +1,90 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.animation as animation
+from src.forward import Heat
+from src.probability import Prior
+
+def main():
+    np.random.seed(19)
+
+    n = 800
+    sig = 5e-2
+    N = 500
+    L = 1
+    delta_t = 1e-5
+    alpha = 1e-8
+    gamma = -1.2
+    acceleration = 3
+
+    pr = Prior(N=N, L=L, transform='dst', gamma=gamma, delta=0.)
+    u = pr.sample(n_sample=100)
+    ind = np.where(np.all(u > 0, axis=1))[0]
+    u = u[ind[0],:]
+
+
+    us = [u]
+    times = [0]
+    for k in range(n):
+        dt = delta_t*k**acceleration
+        fwd = Heat(N=N, L=L, transform='dst', alpha=alpha, time=dt)
+        u = fwd(u)
+        us.append(u)
+        times.append(times[-1] + dt)
+    us = np.vstack(us)
+    print('Done forward simulation')
+
+    fig, ax = plt.subplots()
+    line, = ax.plot(fwd.x, us[0, :], color='k')
+    ax.set_xlim(0, 1)
+    ax.set_ylim(0, 0.7)
+    #ax.set_ylim(np.min(u), np.max(u))
+    ax.set_xlabel('Position')
+    ax.set_ylabel('Temperature')
+    plt.savefig('forward_heat_equation.png')
+
+    time_template = 'Time elapsed {}' 
+    time_text = ax.text(0.95, 0.95, '', transform=ax.transAxes, ha='right', va='top')
+
+    def update(frame):
+        tt = times[frame]#tt = frame**acceleration * time    
+        tt = int(tt)
+        time_text.set_text(time_template.format(tt))
+        line.set_ydata(us[frame, :])
+        return line, time_text
+
+    ani = animation.FuncAnimation(fig, update, frames=n, blit=True)
+    ani.save('forward_heat_equation.mp4', writer='ffmpeg', fps=30)
+    print('Done forward animation')
+    
+    ######################################
+    ##### Reverse heat equation!!!  ######
+    ######################################
+
+    plt.close('all')
+    fig, ax = plt.subplots()
+    line, = ax.plot(fwd.x, us[-1, :], color='k')
+    ax.set_xlim(0, 1)
+    ax.set_ylim(0, 0.7)
+    ax.set_xlabel('Position')
+    ax.set_ylabel('Temperature')
[…]

### 2024-06-16 tests/test_zeros.py

@@ -1,0 +1,58 @@
+import numpy as np
+import pytest
+from joblib import delayed, Parallel
+
+from zeros import *
+
+
+@pytest.fixture
+def params():
+    """ generate random  m (number of observations) and k (rank of O*O).
+    """
+
+    
+    params = []
+    for m in np.random.randint(low=3, high=80, size=30):
+        for k in np.random.randint(low=2, high=m + 1, size=1):
+            params.append({'m': m, 'k': k})
+            params.append({'m': m, 'k': m})
+    return params
+
+
+def test_theta(big_dim=20, dim=9, upper=4):
+    assert big_dim > dim > upper
+    M, D, U, S = MDUS(big_dim, dim)
+
+    C = M - np.identity(dim) * np.trace(M) / dim
+    assert abs(np.trace(C)) < EPS
+
+    theta, lower = get_theta(C, upper)
+    R = givens(theta=theta, dims=dim, lower=lower, upper=upper)
+    T = conj(C, R)
+
+    cot = cos(theta) / sin(theta)
+    eqn = cot ** 2 * C[upper, upper] + 2 * cot * C[upper, lower] + C[lower, lower]
+    assert abs(eqn) < EPS
+    assert abs(T[upper, upper]) < EPS
+    #print("pass theta")
+
+    
+def test_givens(k=8, lower=4, upper=6):
+    M, D, U, S = MDUS(k + 2, k)
+
+    ## Testing Givens rotations R
+    R = givens(2.3, k, lower, upper)
+    C = conj(M, R)
+
+    res = np.abs(M - C)
+    res[lower, :] = 0
+    res[upper, :] = 0
+    res[:, lower] = 0
+    res[:, upper] = 0
+
+    assert np.all(np.abs(res) < EPS)
+    #$ print("pass givens")
+
+    
+def test_all(params):
+    Parallel(n_jobs=7)(delayed(caller)(**param) for param in params)

### 2024-06-16 zeros.py

@@ -1,5 +1,5 @@
-# This here code implements (and tests) the algorithm outlined in
-# Lemma B.5 in the paper.
+# The goal of this module is to implement (and test) the construction
+# outlined in the proof of Lemma B.5 in the paper.
 #
 # Given M symmetric positive definite in R^{k X k} such that
 #
@@ -9,11 +9,15 @@
 #
 # A's columns have unit norm, and
 # AA^t = M.
+#
+# The required functionality is found in method get_A below.
+
+# The method generic generates many PSD matrices, constructs their
+# corresponding A, verifies whether they display clusterization and
+# saves results to the file simulations.csv.
 
 import numpy as np
 import pandas as pd
-import pytest
-from joblib import delayed, Parallel
 from pandas import isnull
 from numpy import sin, cos
 from scipy.stats import ortho_group
@@ -24,20 +28,6 @@ ortho = ortho_group.rvs
 cot = lambda x: cos(x) / sin(x)
 
 
-@pytest.fixture
-def params():
-    """ generate random  m (number of observations) and k (rank of O*O).
-    """
-
-    
-    params = []
-    for m in np.random.randint(low=3, high=80, size=30):
-        for k in np.random.randint(low=2, high=m + 1, size=1):
-            params.append({'m': m, 'k': k})
-            params.append({'m': m, 'k': m})
-    return params
-
-
 def conj(D, U):
     ''' Returns UDU^t'''
     assert U.shape == D.shape
@@ -45,6 +35,10 @@ def conj(D, U):
 
 
 def givens(theta, dims, lower, upper):
+    """Implement a Givens rotation
+
+    """
+    
     assert theta >= 0 and theta <= np.pi
     assert lower < upper
     assert upper <= dims
@@ -57,9 +51,7 @@ def givens(theta, dims, lower, upper):
 
     return R
 
-
-
-    
+   
 def MDUS(m, k):
     """ Generate random:
     M symmetric positive definite
@@ -165,7 +157,21 @@ def caller(m, k):
     #         print(ind)
 
 
-def get_A(M, m):
+def get_A(M,
+          m):
+    """Implement the construction in Lemma ...
+
+    Parameters
+    M: PSD matrix 
+    m: integer
+
+    Returns 
+    
+    A: matrix of shape (M.shape[0], m) such that A has unit norm
+    columns, i.e. np.linalg.norm(A, axis=0) is all ones and A @ A.T ==
[…]

### 2024-06-23 latex/discussion.tex
changing paper and makin slides

\subsection{Discussion}
First of all, it is important to state that we do not view measurement
clusterization as undesirable, nor do we believe it should be avoided
at all. Clusterization is a peculiar phenomenon and it is perfectly
reasonable for someone to argue against the D-optimality criterion
based on the fact that it results in clustered designs. We have seen,
however that there is perfectly reasonable explanation for
clusterization, i.e.~that clusterization arises when we seek a
designWe have shown that clusterization is an innevitable conseqeunce
of having a problem with modes of higher uncertainty than
others. Clusterization manifests the true underlying objective

\subsubsection{Answer to Question \ref{q:why}}
Armed with Theorem \ref{thm:char} we can give a compelling explanation
of the measurement clusterization we obeserved in
Fig.~\ref{fig:}. From part (..) of the Theorem \ref{thm:char} and
Fig.~\ref{fig:tikz clusterization}, we know that a D-optimal design
aims to reduce uncertainty of eigenvectors of
$(\fwd\prcov\fwd^*)^{-1}$ where prior covariance eigenvalues
$\lambda_j$ are highest. D-optimal design will try to focus on the
leading eigenvectors of $(\fwd\prcov\fwd^*)^{-1}$ and take
measurements where other \emph{scaled} eigenvectors are close to
zero.

For concreteness, let us assume that in a certain problem, a D-optimal
design aims to measure only the two leading eigenvectors, i.e.~$\rank
\obs^*\obs =: k=2$. Thus, a D-optimal design will (try to) ignore all
eigenvectors corresponding to $j > 2$. Since $\lambda_j$ decay to
zero, a D-optimal design should only try to avoid eigenvectors for
which $\lambda_j$ is not already small --- e.g.~eigenvectors
$j=3,4$. It is possible that the geometry of the domain $\Omega$ and
the structure of the eigenvectors leave only a few locations in
$\Omega$ where the amplitude of eigenvectors $j=1,2$ is high while the
amplitude of eigenvectors $j=3,4$ is low. In such case, clusterization
will occur whenever the number of measurements $m$ is large enough, by
the pigeonhole principle. This scenario is illustrated for the inverse
problem of the 1D heat equation in Fig.~\ref{fig:eigenvectors}.

\begin{figure}\label{fig:eigenvectors}
    \centering
    \includegraphics[width=\textwidth]{eigenvectors_dst.pdf}
    \caption{D-optimal measurement locations ($m=4$ measurements) and
      weighted eigenvectors for finding the initial condition of
      the 1D heat equation. Measurement locations and weighted
      eigenvectors are plotted over the computational domain $\Omega =
      [0, 1]$ (x-axis). Measurement clusterization occurs
      approximately at $0.31$ and $0.69$. These two locations are a
      compromise between zeros of eigenvectors a D-optimal design aims
      to ignore (third and up) and staying far from zeros of the
      eigenvectors a D-optimal design aims to measure (first and
      second). Allocating $m=4$ measurements into two locations
      results in clusterization, according to the pigeonhole
      principle.}
  \label{fig:why}
\end{figure}

By Weyl's law, eigenvalues $\lambda_j$ of the Laplace operator
increase as $\lambda_j \sim j^{d/2}$, where $d = \dim \Omega$ is the
dimensionality of the domain. In accordance with Weyl's law,
\cite[Theorem 3.1]{Stuart10} suggests taking priors as
"Laplacian-like" operators to power smaller than $-\frac{d}{2}$, to
maintain regularity of posterior realizations. Thus, even though the
computational domain is getting larger and sparser as the dimension
$d$ grows, clusterization is not necessarily harder to achieve.

%% In our model we did not specify any specific prior, so it is possible
%% that one will be able to devise a prior that will not result in
%% clusterization. Indeed, if the rate of increase of prior precision
%% eigenvalues is small,

As mentioned in the Limitations section, our model does not accomodate
point evaluations. For example, in the inverse problem of the 1D heat
equation, it seems reasonable to take $\hilp = \hilo =
L^2(\Omega)$. Unfortunately, for any point evaluation $\delta_x \not
\in L^2(\Omega)$.are taken of the final heat distribution $u_T$. If we
consider the initial and final conditions $u_0$ and $u_T$ as elements
of $L^2([0,1])$, theheat equation as

### 2024-06-23 latex/intro.tex
changing paper and makin slides

@@
 areas of seismology and geology \cite{rabinowitz1990, steinberg1995}
-and medical imaging \cite{tarantola}. Importantly, note that in the
-above examples our goal is to infer some function $f \in C(\Omega)$,
-where \(\Omega \subseteq \mathbb{R}^d, d=1,2,3\) is a spatial domain
-of interest.
+and medical imaging \cite{tarantola}. While in many inverse problems
+the goal is to infer some parameter, in all of the above examples and
+in the formulation we consider here, the goal is to infer some
+\emph{function} over $ \Omega$, where \(\Omega \subseteq \mathbb{R}^d,
+d=1,2,3\) is the spatial domain of interest.
 
@@
 
-We will now introduce a toy model which we will use to define a toy
-inverse problem: the heat equation in one dimension with a homogeneous
-Dirichlet boundary condition. The equations are presented below for
-completeness, but readers less familiar with partial differential
+For concreteness, let us consider a toy model of an inverse problem:
+the \emph{heat equation} in one dimension (1D heat equation
+henceforth, see Section \ref{subsec:1d heat} for more details). The
+heat equation is a partial differential equation which is in fact a
+set of three equations: eqs.~\eqref{eq:heat1}, \eqref{eq:heat2}, and
+\eqref{eq:heat3}. Readers less familiar with partial differential
 equations should think of the heat equation as follows: before us
 there is a perfectly insulated metal rod. At each point on the rod the
-temperature is different, and we call this temperature distribution
-the \emph{initial condition}. We wait for a short time and let the
-heat dissipate a bit inside the rod. The final heat distribution in
-the rod is dictated by the \emph{1D heat equation}:
-\begin{subequations}\label{eq:heat equation}
+temperature is different and unknown, and we call this temperature
+distribution the \emph{initial condition}. We wait for a short time
+and let the heat dissipate a bit inside the rod. The final heat
+distribution in the rod is dictated by the following set of equations
+which defines the 1D heat equation:
+\begin{subequations}%\label{eq:heat equation}
   \begin{alignat}{2}
-    u_t &= \Delta u &&\qquad \text{in } [0,1] \times [0,\infty),\\ u
-      &= 0 &&\qquad \text{on } \{0, 1\} \times [0,\infty),\\ u &= u_0
-        &&\qquad \text{on }[0,1] \times \{0\}.
+    u_t &= \alpha \Delta u &&\qquad \text{in } [0,1] \times [0,\infty), \label{eq:heat1}\\
+    u &= 0 &&\qquad \text{on } \{0, 1\} \times [0,\infty), \label{eq:heat2}\\
+    u &= u_0 &&\qquad \text{on }[0,1] \times \{0\}. \label{eq:heat3}\\
   \end{alignat}
@@
 
-In the inverse problem, our goal is to infer the initial condition
-(i.e.~initial temperature distribution inside the rod). We are allowed
-to take some number of temperature readings on the rod once our
-waiting time has passed. The inverse problem of the 1D heat equation
-is difficult ("ill-posed") since heat spreads in a diffusive manner
-and any roughness in the initial condition is quickly smoothed, see
-Supplementary movies S1 and S2. We will give a fully Bayesian
-formulation of this inverse problem in Section \ref{section:prelims}.
+As time passes, heat disspates across the rod as hotter regions,
+e.g.~local maximas of the heat distribution, become cooler. This
+behavior is captured by eq.~\eqref{eq:heat1}: the Laplacian at local
+maxima is negative, so $u_t = \Delta u$ implies $u$ decreases at its
+maximas, and the reverse happens at local minima. Eq.~\eqref{eq:heat2}
+describes the \emph{boundary condition}, i.e.~the interaction of the
+heat inside the rod with the boundary of the domain $\Omega =
+[0,1]$. Here we consider a \emph{homogeneous Dirichlet boundary
+condition} which is best understood as an absorbing boundary: heat
+that interacts with the boundary immediately
+disappears. Eq.~\eqref{eq:heat3} describes the \emph{initial
+condition} of the heat, i.e.~the heat distribution at time $t=0$.
+
+So far, we described the \emph{forward} problem: how heat evolves in
+time. In the inverse problem, our goal is to infer the initial
+condition $u_0 = u(\cdot, 0)$ from observations of the final condition
+$u(\cdot, T)$. We are allowed to take some number of temperature
+measurements on the rod once our waiting time has passed. The inverse
+problem of the 1D heat equation is difficult ("ill-posed") since heat
+spreads in a diffusive manner and any roughness in the initial
+condition is quickly smoothed, see Supplementary movies S1 and S2. We
+will give a fully Bayesian formulation of a generic linear inverse
+problem in Section \ref{section:prelims}.
 
@@
 the expected Kullback-Liebler divergence from posterior to prior
-\cite{Chaloner1995, AlexanderianGloorGhattas14}, see also precise
-definition in Section \ref{subsec:D optimal design}. We refer to a set
-of measuremnts that maximizes the Bayesian D-optimality criterion a
-\emph{D-optimal design}.
+\cite{Chaloner1995, AlexanderianGloorGhattas14}, see Section
+\ref{subsec:D optimal design} for a precise definition. We refer to a
+set of measuremnts that maximizes the Bayesian D-optimality criterion
+a \emph{D-optimal design}.
 
@@
 another. We refer to this intriguing phenomenon as \emph{measurement
-clusterization} \cite{Ucinski05}. For our purposes, it suffices for
-measurements to be visibly close for clusterization to hold. We refer
-to a design that exhibits measurement clusterization as a
-\emph{clustered design},
+clusterization} \cite{Ucinski05}. We consider a design to be
+\emph{clustered} when two or more of its constituent measurements are
+identical.
 
@@
     \centering
-    \includegraphics[height=0.5\textwidth]{example.pdf}
+    \includegraphics[height=0.5\textwidth]{dst_modelError0.pdf}
     \caption{Measurement clusterization for D-optimal designs for the
@@
 combinatorial optimization problem for measurement locations over a
-finite set. The combinatorial optimization problem is usually relaxed
-by first assigning optimal measurement weights in \(\mathbb{R}_+\) to
-the potential measurement locations. Some researchers incorporate a
-sparsifying \(\ell_1\) penalty term into the design criterion, which
-is subsequently thresholded to achieve the desired binary design over
-the coarse grid \cite{horesh2008borehole}. Others progressively relax
-the \(\ell_1\) penalty to an \(\ell_0\) penalty via a continuation
-method \cite{alexanderian2016, alexanderian2014}. Others cast the
-problem of finding optimal measurement weights as a stochastic
-optimization problem \cite{attia2022stochastic}. All of the
-aforementioned methods may indeed find a binary optimal design
-restricted to a given coarse grid. However, none addresses one
-fundamental issue: the restriction of measurement locations to a
-coarse grid in \(\Omega\) fundamentally changes the optimal design
-problem and thus results in a sub-optimal design.
+discrete set. The combinatorial optimization problem is usually
+relaxed by first assigning optimal measurement weights in
+\(\mathbb{R}_+\) to the potential measurement locations. Some
+researchers incorporate a sparsifying \(\ell_1\) penalty term into the
+design criterion, which is subsequently thresholded to achieve the
+desired binary design over the coarse grid
+\cite{horesh2008borehole}. Others progressively relax the \(\ell_1\)
+penalty to an \(\ell_0\) penalty via a continuation method
+\cite{alexanderian2016, alexanderian2014}. Others cast the problem of
+finding optimal measurement weights as a stochastic optimization
+problem \cite{attia2022stochastic}. All of the aforementioned methods
+may indeed find a binary optimal design restricted to a given coarse
+grid. However, none addresses one fundamental issue: the restriction
+of measurement locations to a coarse grid in \(\Omega\) fundamentally
+changes the optimal design problem and thus results in a sub-optimal
+design.
 
@@
 aforementioned questions. Our focus centers around investigating the
-Bayesian D-optimality criterion, which involves maximizing the
-expected Kullback-Leibler divergence between the posterior and prior
-measures \cite{CoverThomas91, Chaloner1995}. We conduct an analysis of
-Bayesian D-optimal designs within the context of linear inverse
-problems over Hilbert spaces. We propose a novel relaxed model for
-D-optimality that maintains analytical tractability and enables the
-identification of D-optimal designs using Lagrange multipliers. This
-analytical framework facilitates the exploration of the questions
-posed at the end of the previous paragraph:
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{eigenvectors.pdf}
-    \caption{D-optimal measurement locations ($m=4$ measurements) and
-      weighted eigenvectors for finding the initial condition of
-      the 1D heat equation. Measurement locations and weighted
-      eigenvectors are plotted over the computational domain $\Omega =
-      [0, 1]$ (x-axis). Measurement clusterization occurs
-      approximately at $0.31$ and $0.69$. These two locations are a
-      compromise between zeros of eigenvectors a D-optimal design aims
-      to ignore (third and up) and staying far from zeros of the
-      eigenvectors a D-optimal design aims to measure (first and
-      second). Allocating $m=4$ measurements into two locations
-      results in clusterization, according to the pigeonhole
-      principle.}
-  \label{fig:why}
-\end{figure}
+Bayesian D-optimality criterion. We conduct an analysis of Bayesian
+D-optimal designs within the context of linear inverse problems over
+Hilbert spaces. We propose a novel relaxed model for D-optimality that
+maintains analytical tractability and enables the identification of
+D-optimal designs using Lagrange multipliers. This analytical
+framework facilitates the exploration of the questions posed at the
+end of the previous paragraph:
 
 \begin{enumerate}
-
 \item \label{q:generic} \textbf{Is measurement clusterization a
@@
   designs that exhibit clusterization more than 95\% of the time (see
-  code in supplementary material). Given our model's genericity, we
-  expect measurement clusterization to be a generic phenomenon.
+  Section \ref{subsec:lemma sims} for details). Given the genericity
+  of our model, we expect measurement clusterization to be a generic
+  and ubiquitous phenomenon.
 
@@
   role of model error in mitigating clusterization, thereby
-  corroborating earlier observations made by various researchers.
+  corroborating earlier observations made by various researchers. We
+  also show that adding to the inverse problem of the 1D heat
+  equationan error term corresponding to correlations between
+  measurements mitigates clusterization, see Section \ref{subsec:corr
+    errors sims}.
 
 \item \label{q:why} \textbf{Why does measurement clusterization
-  occur?} In Section \ref{section:why}, we present a compelling
-  explanation for the optimality of clustered designs in the absence
-  of model error. Our analysis reveals that, in our model, a D-optimal
-  design focuses on a select set of prior eigenvectors, specifically
-  those with the largest eigenvalues in the prior covariance
-  spectrum. In practical scenarios, the number of locations where (a)
-  the relevant prior eigenvectors are significantly large, and (b)
-  other eigenvectors are close to zero, is limited. Consequently, the
-  clusterization of measurements arises as a natural consequence of
-  the pigeonhole principle, as there are more measurements available
-  than there are locations satisfying conditions (a) and (b). See
-  Fig~\ref{fig:why}.
+  occur?} Answering this question is the main contribution of this
+  study, and it is arrived at in Section \ref{section:why}. There, we
+  present a compelling explanation for the optimality of clustered
+  designs in the absence of model error. Our analysis reveals that a
+  D-optimal design focuses on a select set of prior eigenvectors,
+  specifically those with the largest eigenvalues in the prior
+  covariance spectrum. In practical scenarios, the number of locations
+  where (a) the relevant prior eigenvectors are significantly large,
+  and (b) other eigenvectors are close to zero, is
+  limited. Consequently, the clusterization of measurements arises as
+  a natural consequence of the pigeonhole principle, as there are more
+  measurements available than there are locations satisfying
+  conditions (a) and (b). See Fig~\ref{fig:why}.
 
@@
 
-The first insight our study gives is that a D-optimal design aims to
-uniformly reduce uncertainties not over space (i.e.~$\Omega$), but
-rather over \emph{eigenvectors}. This insight is a part of Theorem
-\ref{thm:char}, see Section \ref{section:vanishing} for precise
-statement and proof, and see Fig.~\ref{cluste} for an illustration.
+The implication of the answers we give to Questions \ref{q:generic},
+\ref{q:mitigate}, and \ref{q:why} is two-fold:
 
+The implication of our answer to Question \ref{q:generic} is that
+encountering clusterization should be expected in many different
+inverse problems across many different scientific fields and
+disciplines. Thus, researchers that encounter clusterization should
+not be surprised or wary. Then, in our answer to Question \ref{q:why},
+we explain what is the source of clusterization. As it appears, the
+source is universal: a D-optimal design reduces uncertainty for a
+select set of prior covariance eigenvectors --- those with the most
+prior uncertainty, i.e.~those the practitioner cares about the most!
+Consequently, practitioners should not try to avoid measurement
+clusterization. Rather, practitioners should should take repeated
+measurements (e.g.~in MRI and borehole tomography), increase apparatus
+sensitivity (e.g.~in EIT), or take consecutive measurements (e.g.~in
+the 1D heat equation). Overall, we show that measurement
+clusterization is a natural and (almost) inevitable part of Bayesian
+D-optimal designs.
+
+\subsubsection{Further Contributions}
 In Theorem \ref{thm:char}, we also show that D-optimal designs are
@@
 The main limitation of this study is that our generic model does not
-correspond to any specific real-life problem. It is generic enough to
-be analytically tractable, but one may argue our model is too far
-removed from any real application. To these claims I would answer that
-scientists have a long history of studying models that are bare-bones
-simplifications of real systems, e.g. the Ising model
+correspond to any specific real-life problem. Specifically, in its
+current form, our model does not allow point evaluations, see the
+Discussion for a deeper analysis of this subject. Our model is generic
+enough to be analytically tractable, but one may argue our model is
+too far removed from any real application. To these claims I would
+answer that scientists have a long history of studying models that are
+bare-bones simplifications of real systems, e.g. the Ising model
 \cite{cipra1987}, the Lorenz system \cite{brin}, the Lotka-Volterra

### 2024-06-23 latex/numerix.tex
changing paper and makin slides

@@
 
-\subsection{Heat Equation}
-Our goal is to infer the initial condition of
-the one-dimensional heat equation with a homogeneous Dirichlet
-boundary condition:
-\begin{subequations}\label{eq:heat equation}
-  \begin{alignat}{2}
-    u_t &= \Delta u &&\qquad \text{in } [0,1] \times [0,\infty),\\
-      u &= 0 &&\qquad \text{on } \{0, 1\} \times [0,\infty),\\
-        u &= u_0 &&\qquad \text{on }[0,1] \times \{0\}.
-  \end{alignat}
-\end{subequations}
+\subsection{Simulating Lemma \ref{lemma:free}}\label{subsec:lemma sims}
+Lemma \ref{lemma:free} is used to construct D-optimal designs as a
+part of the proof of Theorem \ref{thm:char}. We implement the
+construction in Lemma \ref{lemma:free} (see module \texttt{zeros.py}
+in the accompanying
+\href{https://github.com/yairdaon/OED}{repository}). We iterate over
+the number of measurements $m \in \{4,\dots, 15\}$, and for every $m$
+we then iterate over $k:=\rank \obs^*\obs \in \{2,\dots, m-1\}$. For
+each pair $m,k$ we run $N=50000$ simulations according to the
+following steps:
+\begin{enumerate}
+\item Generate random diagonal $D\in \mathbb{R}^{k\times k}$ and
+  normalize it so that $\ttr D = m$.
+\item Conjugate $D$ by a random orthogonal matrix to form a positive
+  semi-definite $M := UDU^t \in \mathbb{R}^{k\times k}$.
+\item Apply the construction of Lemma \ref{lemma:free} to calculate
+  $A$ such that $AA^t = M$, where $A$ has unit norm columns.
+\item We call $A$ "clustered" if $A$ has two or more identical columns
+  (up to some numerical precision threshold).
+\end{enumerate}
+We then calculate the fraction of clustered designs of the $N=5000$
+simulations we ran, for each pair $m,k$. Full results in the
+\texttt{simulations.csv} file within the accompanying
+\href{https://github.com/yairdaon/OED}{repository}), below we give a
+brief description of the results.
 
-Model the prior for the initial condition as $u_0 \sim
-\normal(0,\prcov)$, for $\prcov = (-\Delta)^{-1}$ with a homogeneous
-Dirichlet boundary condition. Let the final time $T = 0.03$ and denote
-$\fwd$ the forward operator, so that $u(\cdot, T) = \fwd u_0$. Let
-$\obs$ a measurement operator, so that $u(\x_j,T) = (\obs u)_j,
-j=1,\dots,m$, where $m$ is the number of allowed measurements. Assume
-iid centered Gaussian observation error, so observations are given by
-$u(\x_j,T) + \eps(\x_j)$ with $\eps(\x_j) \sim \normal(0, \sigma^2)$
-iid, $\sigma = 0.05$. Measurement locations $\x_j,j=1,\dots,m$ are
-chosen according to the D-optimality criterion
-\cite{AlexanderianGloorGhattas14}.
+Clusteriztion occured at high rates ($>95\%$) whenever $m-k > 1$; see
+Fig.~\ref{fig:sim AAt}. Hence, in these simulations, clusterization is
+a generic property. However, when $m-k \leq 2$, clusterization does
+not occur. We do not why and further investigation into this
+phenomenon is out of scope for the current study.
 
-Implementation of this inverse problem can be found in this 
-\href{https://github.com/yairdaon/OED}{repository}, including an optimization
-routine for finding D-optimal designs. Unfortunately, I was not able
-to install PyOED \cite{attia2023pyoed}, nor was I able to utilize
-hIPPYlib \cite{VillaPetraGhattas16, VillaPetraGhattas18,
-  VillaPetraGhattas21}, hence the inverse problem is implemented from
-scratch.
+\begin{figure}
+    \centering
+    \includegraphics[height=0.5\textwidth]{simulations.png}
+    \caption{Fraction of clustered $A$ for $AA^t = M$ and $M$
+      generated randomly (see text and repository for details on
+      generating $M$). It is evident that when $m-k >3$ clusterization
+      is ubiquitous, whereas for lower $m-k$ clusterization does not
+      occur.}
+  \label{fig:sim AAt}
+\end{figure}
 
-\subsection{Simulating Lemma ... }
-In theorem 17 we explicitly construct a D-optimal design in a Hilbert
-space where the base vectors diagonalize $\fwd \prcov\fwd^*$. We
-explicitly implement the algorithm outlined in Lemma
+\subsection{Correlated errors}\label{subsec:corr errors sims}
+In order to verify the results of Section \ref{section:non vanishing},
+we run simulations of the inverse problem of the 1D heat equation
+\ref{subsec:1d heat} with nonvanishing model error. We take a model
+correlation of \(\modcov = \prcov^2 \). Indeed, adding model
+correlation pushes measurements apart as can be seen in
+Fig.~\ref{fig:corr errors}. Code generating the image below can be
+located at module \texttt{clusterization.py} in the accompanying
+\href{https://github.com/yairdaon/OED}{repository}).
 
-\subsection{Correlated errors}
-We simulate the 1D heat eq
+\begin{figure}
+    \centering
+    \includegraphics[height=0.5\textwidth]{dst_modelError4.pdf}
+    \caption{Model correlation mitigates clusterization. We add a
+      model correlation term to the error terms in the 1D heat
+      equation inverse problem. Lo and behold, measurements are not
+      close anymore and are pushed away thanks to the model error
+      term.}
+  \label{fig:corr errors}
+\end{figure}

### 2024-06-23 latex/prelims.tex
changing paper and makin slides

@@
 \section{Preliminaries and Notation}\label{section:prelim}
+Here we present the basics of Bayesian inverse problems over hilbert
+spaces. Since we are ultimately interested in inferring a function over some
+domain, we will keep in mind that these Hilbert spaces should really
+be thought of as function spaces. The theoretical foundations for
+inverse problems over function spaces can be found in \cite{Stuart10}
+and will be only briefly reviewed here.
 
-In this section we present the setup that will be used throughout this
-article. The theoretical foundations for inverse problems over
-function spaces can be found elsewhere \cite{Stuart10} and will not be
-reviewed here.
+\subsection{Forward Problems}\label{subsec:abstract OED}
+In this section we give definitions and notations for forward probles,
+which are an essential part of the inverse problems we discuss
+later. Consider a "parameter space" \(\hilp\) and an "observation
+space" \(\hilo\) --- both separable Hilbert spaces (the subscripts p
+and o are for "parameter" and "observation", respectively). The
+parameter space includes the quantity we seek to infer; in the inverse
+problem of the heat equation, the parameter space $\hilp$ is where the
+initial condition lives. The observation space $\hilo$, on the other
+hand, is the space from which we take measurements; in the example of
+the 1D heat equation, $u(\cdot, T) \in \hilo$.
 
-\subsection{Bayesian Linear Inverse Problems}\label{subsec:abstract OED}
-Let \(\hilp\) and \(\hilo\) separable Hilbert spaces (the subscripts p
-and o are for "parameter" and "observation", respectively), and let
-\(\fwd: \hilp \to \hilo\) the \emph{forward operator}. The forward
-operator \(\fwd\) is assumed linear with fast decaying singular values
-(the heat operator of the 1D heat equation introduced in Section
-\ref{section:intro} is a prime example). Take a Gaussian prior
-\(\param \sim \pr = \normal(\prmean ,\prcov)\) with some appropriate
-covariance operator \(\prcov\) on \(\hilp\) \cite{Stuart10}. Note that
-\(\fwd \prcov \fwd^*\) is the prior covariance in \(\hilo\)
-\cite{Stuart10}, and as such, \(\fwd \prcov \fwd^*\) is assumed
-invertible --- an assumption which will be used below (if \(\fwd\) has
-a nontrivial kernel we utilize Occam's Razor and ignore said kernel
-altogether). Measurements are taken via a \emph{measurement operator}
-\(\obs\). It is common for the measurement and forward operators to be
-merged \(\tmp := \obs \fwd\) \cite{AlexanderianGloorGhattas14}, but
-the analysis carried out in the following sections requires that
-\(\fwd\) and \(\obs\) are explicitly separated as in
-\cite{attia2022stochastic, cvetkovic2023choosing}.  \(\obs \in (
-\hilo^* )^m\), where \(m\) is the number of measurements
-taken. Entries \(\meas_j, j=1,\dots,m\) of the observation operator
-\(\obs\) are called \emph{measurements}:
+The connection between parameter and observation spaces is made by the
+\emph{forward operator} \(\fwd: \hilp \to \hilo\). We assume the
+forward operator \(\fwd\) is linear. In the inverse problem of the 1D
+heat, the forward operator is determined by the time evolution of the
+1D heat equation \eqref{eq:heat1}, \eqref{eq:heat2} and
+\eqref{eq:heat3}.
+
+Measurements are taken via a \emph{measurement operator} \(\obs\). It
+is common for the measurement and forward operators to be merged
+\(\tmp := \obs \fwd\) \cite{AlexanderianGloorGhattas14}, but the
+analysis carried out in the following sections requires that \(\fwd\)
+and \(\obs\) are explicitly separated (as in
+\cite{attia2022stochastic, cvetkovic2023choosing}). Note that \(\obs
+\in ( \hilo^* )^m\), where \(m\) is the number of measurements
+taken\footnote{The alert reader will likely ask how do we reconcile
+point measurements $\delta_x$ as suggested by the formulation of the
+1D heat equation. We don't. This entire study could be formulated for
+Banach spaces where point measurements are allowed. We avoid this
+technical difficulty because: (1) it does not add much to the study,
+(2) it is beyond the technical abilities of the author}. Entries
+\(\meas_j, j=1,\dots,m\) of the measurement operator \(\obs\) are
+called \emph{measurements}:
 
@@
 
-Finally, it is useful to record that the posterior measure \(\post\)
-is Gaussian in this setting and its covariance operator does not
-depend on data \(\data\) \cite{Stuart10}:
+\subsection{Bayesian Linear Inverse Problems}\label{subsec:bayesian inverse problems}
+In the previous section, we saw how a parameter $u\in \hilp$ is
+transported to the observation space via the forward operaror $\fwd u
+\in \hilo$, how observations are generated via $\obs \fwd u$ and how
+observations and noise give rise to data $\data$. It is time to
+formulate the process of inferring the parameter as a Bayesian inverse
+problem. The first ingredient we need is a prior.
+
+We take a Gaussian prior \(\param \sim \pr = \normal(\prmean
+,\prcov)\) with some appropriate covariance operator \(\prcov\) on
+\(\hilp\) \cite{Stuart10}. For the inverse problem of the 1D heat
+equation we choose $\prcov = -\Delta^{-1.1}$, see Section
+\ref{subsec:1d heat} for details and a short discussion of
+"Laplacian-like" covariance operators. Note that \(\fwd \prcov
+\fwd^*\) is the prior covariance in \(\hilo\) \cite{Stuart10}, and as
+such, \(\fwd \prcov \fwd^*\) is assumed invertible --- an assumption
+which will be used below (if \(\fwd\) has a nontrivial kernel we
+utilize Occam's Razor and ignore said kernel altogether).
+
+Since this inverse problem is linear and the prior is Gaussian --- the
+posterior measure \(\post\) is Gaussian as well. We do not utilize the
+posterior mean in this study and posterior covariance operator
+$\postcov$ is given by \cite{Stuart10}:
 \begin{align}\label{eq:postcov}
@@
 
-\subsection{Bayesian D-Optimal Designs in Infinite Dimensions}\label{subsec:D optimal design} 
+\subsection{Bayesian D-Optimal Designs}\label{subsec:D optimal design} 
 A Bayesian D-optimal design maximizes expected Kullback-Leibler (KL)

### 2024-06-23 latex/refs.tex
changing paper and makin slides

--- 2024-06-05 latex/review.tex
@@
 (2012). So I will assume for the rest of this review that, measurement
-clusterisation refers to repeated measurements.  Nearly all textbooks
-on classical design of experiments (e.g. Morris, 2011, Section 1.2.4)
-will feature a section with the three words randomisation, blocking
-and repetition: the important principles of design of experiments.  In
-these texts, repeated measurements are considered to be a beneficial
-property of a design. Interestingly, when optimal designs are found,
-they are often found to have repeated measurements.  The paper does
-not explain why the author thinks repeated measurements (measurement
-clusterisation) is an undesirable property. If one looks at, for
-example, Fedorov (1996) and Nyberg et al (2012), it is an undesirable
-property because they are interested in sensor location, and two or
-more sensors cannot be in the same location. However, there are many
-experiments where repeated measurements are possible, and are
-therefore optimal (if the model is correct, which it won’t be). An
-example, is a chemical engineering experiment where one wishes to
-learn the relationship between a series of variables and yield of a
-chemical reaction. The experiment will involve specifying the
-variables, and measuring the yield (the observation) after a specified
-period of time. This process is then duplicated with a potentially
-different set of variables.  It is perfectly possible to have repeated
-variables.
+clusterisation refers to repeated measurements.
+
+Nearly all textbooks on classical design of experiments (e.g. Morris,
+2011, Section 1.2.4) will feature a section with the three words
+randomisation, blocking and repetition: the important principles of
+design of experiments.  In these texts, repeated measurements are
+considered to be a beneficial property of a design. Interestingly,
+when optimal designs are found, they are often found to have repeated
+measurements.  The paper does not explain why the author thinks
+repeated measurements (measurement clusterisation) is an undesirable
+property. If one looks at, for example, Fedorov (1996) and Nyberg et
+al (2012), it is an undesirable property because they are interested
+in sensor location, and two or more sensors cannot be in the same
+location. However, there are many experiments where repeated
+measurements are possible, and are therefore optimal (if the model is
+correct, which it won’t be). An example, is a chemical engineering
+experiment where one wishes to learn the relationship between a series
+of variables and yield of a chemical reaction. The experiment will
+involve specifying the variables, and measuring the yield (the
+observation) after a specified period of time. This process is then
+duplicated with a potentially different set of variables.  It is
+perfectly possible to have repeated variables.
+
 \begin{enumerate}
+
 \item The author needs to make clear that they are considering
@@
 
+  Answer: It is impossible to visually discern the measurements, even
+  when resolution is increased.
+
 \item Experiment with the 1D heat equation: the experiment showcases
-  some interesting phenom- ena, but also raises many questions. For
+  some interesting phenomena, but also raises many questions. For
   instance: I would expect the dimensionality of the working domain
@@
 
+  Answer: That is a great point! I added the followng discussion:
+
 \item Further, a dimensionality effect should also arise directly from
-  Theorem 1 via the prior co- variance eigenvalues: in the experiment
+  Theorem 1 via the prior covariance eigenvalues: in the experiment
   the prior covariance operator is set to be equal to the inverse
   Laplace operator $(−\Delta)^{−1}$ . In d-dimensional domains, its
-  eigenvalues λj are known to follow Weyl’s asymptotics, growing as j
-  2/d , which should then impact the index after which the eigenvalues
-  are ‘thresholded’ by the procedure.
+  eigenvalues $\lambda_j$ are known to follow Weyl’s asymptotics,
+  growing as $j^{2/d}$, which should then impact the index after
+  which the eigenvalues are ‘thresholded’ by the procedure.
 
@@
 statistical papers.
+
+Answer: Thanks for this comment but I would rather keep this section
+here since...
 

### 2024-06-23 clusterization.py

@@ -13,7 +13,7 @@ def main():
     L = 1 ## Length of computational domain
     time = 3e-2 ## Time for heat dissipation in arbitrary units.
     alpha = 1. ## Coefficient of Laplacian in heat equation
-    gamma = -1. ## Exponent of Laplacian in prior. 
+    gamma = -1.1 ## Exponent of Laplacian in prior. 
 
     ## Choose one, comment out the other
     transform = 'dct' ## dct - Discrete Cosine Transform. Corresponds
@@ -59,10 +59,10 @@ def main():
     dic = {}
 
     ## Form  number of observations in this range
-    for m in range(2, 6):
+    for m in range(2, 7):
 
         ## Calculate an optimal design
-        design = post.optimize(m=m, n_iterations=25)['x']
+        design = post.optimize(m=m, n_iterations=500)['x']
 
         ## And store it in dic
         dic[m] = design
@@ -79,7 +79,7 @@ def main():
     ## at ordinate m, with different colors.
     for m, array in dic.items():
 
-        colors = iter(list("brgkm"))
+        colors = iter(list("brgkmc"))
         vals = np.repeat(m, len(array))
         ax.scatter(array, vals, s=0)
         for i, val in enumerate(array):
@@ -97,7 +97,7 @@ def main():
     ax.set_xlim(0,1)
     plt.tight_layout()
 
-    plt.savefig(f"latex/{transform}_modelError{model_error}.pdf")
+    plt.savefig(f"latex/{transform}_modelError{model_error}.png")
     #plt.show()
 
     ####################################
@@ -125,7 +125,7 @@ def main():
     plt.xlabel(r"$x \in \Omega$", fontsize=fs)
     plt.legend(title='eigenvector, weighted')
     plt.tight_layout()
-    plt.savefig(f"latex/eigenvectors_{transform}.pdf")
+    plt.savefig(f"latex/eigenvectors_{transform}.png")
     plt.show()
 
 

### 2024-06-23 movies.py

@@ -40,7 +40,7 @@ def main():
     #ax.set_ylim(np.min(u), np.max(u))
     ax.set_xlabel('Position')
     ax.set_ylabel('Temperature')
-    plt.savefig('forward_heat_equation.png')
+    plt.savefig('latex/forward_heat_equation.png')
 
     time_template = 'Time elapsed {}' 
     time_text = ax.text(0.95, 0.95, '', transform=ax.transAxes, ha='right', va='top')
@@ -53,7 +53,7 @@ def main():
         return line, time_text
 
     ani = animation.FuncAnimation(fig, update, frames=n, blit=True)
-    ani.save('forward_heat_equation.mp4', writer='ffmpeg', fps=30)
+    ani.save('latex/forward_heat_equation.mp4', writer='ffmpeg', fps=30)
     print('Done forward animation')
     
     ######################################
@@ -67,7 +67,7 @@ def main():
     ax.set_ylim(0, 0.7)
     ax.set_xlabel('Position')
     ax.set_ylabel('Temperature')
-    plt.savefig('reverse_heat_equation.png')
+    plt.savefig('latex/reverse_heat_equation.png')
 
     u_reversed = us[::-1]
     time_reversed = np.array(times)[::-1]
@@ -82,7 +82,7 @@ def main():
         return line, time_text
 
     ani = animation.FuncAnimation(fig, update, frames=n, blit=True)
-    ani.save('reverse_heat_equation.mp4', writer='ffmpeg', fps=30)
+    ani.save('latex/reverse_heat_equation.mp4', writer='ffmpeg', fps=30)
     print('Done backwards animation')
     plt.show()
 

### 2024-06-23 zeros.py

@@ -22,6 +22,10 @@ from pandas import isnull
 from numpy import sin, cos
 from scipy.stats import ortho_group
 from scipy.spatial.distance import cdist
+from matplotlib import pyplot as plt
+from joblib import Parallel, delayed
+from tqdm import tqdm
+import seaborn as sns
 
 EPS = 1e-9
 ortho = ortho_group.rvs
@@ -53,7 +57,7 @@ def givens(theta, dims, lower, upper):
 
    
 def MDUS(m, k):
-    """ Generate random:
+    """Generate and return random:
     M symmetric positive definite
     D diagonal with decreasing diagonal entries
     U orthogonal such that M = UDU^t
@@ -205,6 +209,47 @@ def get_A(M,
     return A
 
 
+def simulate(m, k):
+    """   
+    Generate a random D-optimal design, following Lemma from paper.
+    
+    Parameters
+
+    m: number of measurement points
+    k: Rank of O^*O
+
+    Returns
+    True if the design is clustered, False o.w.
+
+    """
+    
+    ## Random matrices such that:
+    ## M is symmetric positive definite
+    ## m = number of measurements
+    ## Number of nonzero entries in D
+    M, _, _, _ = MDUS(m, k)
+                
+    ## AA^t = M, and A has unit norm columns.
+    A = get_A(M, m)
+    
+    ## distances between columns of A
+    distances = cdist(A.T, A.T)
+    
+    ## Clusterization does not occur if all off-diagonal
+    ## distances are large. Diagonal has 0 entries, but
+    ## obviously that does not mean clusterization
+    ## occurs. Since we do not want to count the diagonal
+    ## entries as clustered, we fill diagonal with 1's
+    np.fill_diagonal(distances, 1)
+
+    ## Which distances are > 0 <==> which pairs of
+    ## measurements are ***not*** clustered
+    dis = distances > EPS
+    
+    ## If all distances are large, then measurements are not clustered
+    ## and the design does not exhibit sensor clusterization.
+    return {'m': m, 'k': k, 'cluster': not dis.all()}
+    
 
 def generic():
     """Run randomized simulations of D-optimal design with our model.
@@ -212,68 +257,49 @@ def generic():
     """
 
     ## Number of simulations per pair m, k.
-    N = 5000
-    
-    res = []
-    for m in range(4, 16):
-        print()
-        print(f'm={m}: k=', end=' ')
-        
-        for k in range(2, m-1):
-            print(k, end=' ')
-            
[…]

### 2024-06-25 latex/bayess.py
slides ready

@@ -1,0 +1,60 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from scipy.stats import norm
+
+plt.rcParams.update({'font.size': 16})
+
+# Set up the x-axis
+x = np.linspace(30, 100, 1000)
+
+# Prior distribution (blue)
+prior_mean, prior_std = 60, 10
+prior = norm.pdf(x, prior_mean, prior_std)
+
+# Single observation (green dot)
+observation = 75
+
+# Likelihood (green)
+likelihood_mean = observation
+likelihood_std = 5  # Adjust this value to change the width of the likelihood curve
+likelihood = norm.pdf(x, likelihood_mean, likelihood_std)
+
+# Posterior distribution (purple)
+posterior_mean = (prior_mean/prior_std**2 + observation/likelihood_std**2) / (1/prior_std**2 + 1/likelihood_std**2)
+posterior_std = np.sqrt(1 / (1/prior_std**2 + 1/likelihood_std**2))
+posterior = norm.pdf(x, posterior_mean, posterior_std)
+
+# Create the plot
+plt.figure(figsize=(10, 6))
+plt.plot(x, prior, 'b-', label='Prior (m)')
+plt.plot(x, likelihood, 'g-', label='Likelihood (d|m)')
+plt.plot(x, posterior, 'red', label='Posterior (m|d)')
+
+# Plot single observation as green dot on x-axis
+plt.plot(observation, 0, 'go', markersize=10, label='Observation (d)')
+
+# Customize the plot
+# plt.xlabel('μ')
+# plt.ylabel('Probability Density')
+# plt.title('Bayesian Inference with Single Observation')
+plt.legend()
+plt.grid(True, linestyle=':', alpha=0.7)
+
+# Set axis limits
+plt.ylim(-0.005, 0.12)
+plt.xlim(30, 100)
+
+
+plt.xticks([])
+plt.yticks([])
+
+# Remove x and y axis labels
+plt.xlabel('')
+plt.ylabel('')
+
+plt.box(False)
+    
+# Show the plot
+plt.tight_layout()
+plt.savefig('bayes.png')
+#plt.show()

### 2024-07-21 latex/ack.tex
before diff

@@
 This study is a result of research I started during my PhD studies
-under the instruction of Prof. Georg Stadler at the Courant Institute
+under the instruction of Prof.~Georg Stadler at the Courant Institute
 of Mathematical Sciences. I would like to thank him for his great
@@
 \href{https://mathoverflow.net/questions/280168/redistribute-diagonal-entries-of-a-matrix/280203#280203c}{Mathoverflow}.
+Last, but certainly not least, I would like to thank the three
+referees, associate editor and editor in chief for providing detailed
+and insightful reviewes that have made this manuscript a whole lot
+better.
+%Having worked on this manuscript alone for several years, they have
+%almost become my coautoh

### 2024-07-21 latex/discussion.tex
before diff

@@
 \subsubsection{Answer to Question \ref{q:why}}
-Armed with Theorem \ref{thm:char} we can give a compelling explanation
-of the measurement clusterization we obeserved in
-Fig.~\ref{fig:}. From part (..) of the Theorem \ref{thm:char} and
-Fig.~\ref{fig:tikz clusterization}, we know that a D-optimal design
-aims to reduce uncertainty of eigenvectors of
-$(\fwd\prcov\fwd^*)^{-1}$ where prior covariance eigenvalues
-$\lambda_j$ are highest. D-optimal design will try to focus on the
-leading eigenvectors of $(\fwd\prcov\fwd^*)^{-1}$ and take
-measurements where other \emph{scaled} eigenvectors are close to
-zero.
+Theorem \ref{thm:char} helps us give a compelling explanation of the
+measurement clusterization we obeserved for the inverse problem of the
+heat equation (Fig.~\ref{fig:}). Below, we take the following steps:
+We first utilize Theorem \ref{thm:char} to understand what a D-optimal
+design will look like for our relaxed model; we then argue that
+although not identical, a D-optimal design for the full inverse
+problem of the heat equation shares similar charachteristics to the
+relaxed design; finally, we show how ...
 
-For concreteness, let us assume that in a certain problem, a D-optimal
-design aims to measure only the two leading eigenvectors, i.e.~$\rank
-\obs^*\obs =: k=2$. Thus, a D-optimal design will (try to) ignore all
-eigenvectors corresponding to $j > 2$. Since $\lambda_j$ decay to
-zero, a D-optimal design should only try to avoid eigenvectors for
-which $\lambda_j$ is not already small --- e.g.~eigenvectors
-$j=3,4$. It is possible that the geometry of the domain $\Omega$ and
-the structure of the eigenvectors leave only a few locations in
-$\Omega$ where the amplitude of eigenvectors $j=1,2$ is high while the
-amplitude of eigenvectors $j=3,4$ is low. In such case, clusterization
-will occur whenever the number of measurements $m$ is large enough, by
-the pigeonhole principle. This scenario is illustrated for the inverse
-problem of the 1D heat equation in Fig.~\ref{fig:eigenvectors}.
+First, recall from Theorem \ref{thm:char} that in our relaxed model,
+for a D-optimal design $\opt$:
+\begin{enumerate}
+  \item $\opt^*\opt$ is simultaneously diagonalizable with $\fwd
+    \prcov \fwd^*$.
+  \item $\opt^*\opt$ has non-zero eigenvalues only for the $k:=\rank
+    \opt^*\opt$ smallest eigenvalues $\lambda_j$ of
+    $\fwd\prcov\fwd^*$.
+  \item $k$ is monotonic in $m$, the number of measurements we take.
+\end{enumerate}
+
+Now, consider $\fwd$ and $\prcov$ from the \emph{the inverse problem
+of the heat equation}. As before, we denote the eigenvalues of
+$\fwd\prcov\fwd^*$ by $\lambda_j$. We input these eigenvalues into our
+\emph{relaxed} model, and find a D-optimal design $\opt$. In our
+relaxed model, the measurements we take are best utilized in reducing
+uncertainty for the first $k$ eigenvectors. So, a D-optimal design
+arising from our \emph{relaxed model} completely avoids measuring
+eigenvectors $k+1$ and above.
+
+Of course, in a real life problem --- such as the inverse problem of
+the heat equatiion --- it is likely impossible to find measurement
+locations for which all eigenvectors $k+1$ and above are
+zero. However, we can still expect a D-optimal design to give as
+little weight as possible to eigenvectors $k+1$ and above. Thus a
+D-optimal design is a trade-off between measuring eigenvectors we want
+(eigenvectors for which $\lambda_j$ is large) and treating the rest as
+noise.
+
+We explore the abovementione trade-off for the inverse problem of the
+heat equation in Fig.~\ref{fig:eigs}. We allow $m=4$ measurements in
+$[0,1]$ and observe that D-optimal measurement locations are clustered
+at $x_1 = 0.31$ and $x_2 = 0.69$. Upon close inspection of the scaled
+eigenvectors of $\fwd \prcov \fwd^*$, we first observe that
+eigenvectors $3$ and above have negligible prior amplitude. Since we
+only have $m=4$ measurements at our disposal, Theorem \ref{thm:char}
+tells us we should only care about measuring the first and second
+eigenvectors. Then, we note the D-optimal $x_1,x_2$ present a
+compromise between the amplitude of the first and second
+eigenvectors. For example, a measurement at $x=0.5$ would have ignored
+the second eigenvector altogether, since its value is zero at $x=0.5$.
+Now we can understand measurement clusterization: attempting to
+measure what we want and ignoring the rest, along with the spatial
+limitations on measurements imposed by the structure of eigenvectors
+of $\fwd \prcov \fwd^*$ is the cause for measurement clusterization in
+D-optimal designs.
+
+%% In the inverse problem of the heat equation, \ev_3, \ev_4,\dots$
+%% implies we seek measurement locations for which $\lambda_1\ev_1,
+%% \lambda_2\ev_2$ are large, while $\lambda_3\ev_3,
+%% \lambda_3\ev_3,\dots$ are small. It turns out that there are two
+%% locations that give the best trade-off, and measurements are taken in
+%% those two locations. Once we have more than two measurements at our
+%% disposal, measurement locations becomde clustered, by the pigeonhole
+%% principle, see Fig.~\ref{fig:eigenvectors}.
+
+%% The way a D-optimal design tries to ignore all eigenvectors
+%% corresponding to $j > 2$. Since $\lambda_j$ decay to zero, a D-optimal
+%% design should only try to avoid eigenvectors for which $\lambda_j$ is
+%% not already small --- e.g.~eigenvectors $j=3,4$. It is possible that
+%% the geometry of the domain $\Omega$ and the structure of the
+%% eigenvectors leave only a few locations in $\Omega$ where the
+%% amplitude of eigenvectors $j=1,2$ is high while the amplitude of
+%% eigenvectors $j=3,4$ is low. In such case, clusterization will occur
+%% whenever the number of measurements $m$ is large enough, by the
+%% pigeonhole principle. This scenario is illustrated for the inverse
+%% problem of the 1D heat equation in Fig.~\ref{fig:eigenvectors}.
 
@@
     \centering
-    \includegraphics[width=\textwidth]{eigenvectors_dst.pdf}
+    \includegraphics[width=\textwidth]{eigenvectors_dst_scaled.png}
     \caption{D-optimal measurement locations ($m=4$ measurements) and
@@
 
-By Weyl's law, eigenvalues $\lambda_j$ of the Laplace operator
-increase as $\lambda_j \sim j^{d/2}$, where $d = \dim \Omega$ is the
-dimensionality of the domain. In accordance with Weyl's law,
-\cite[Theorem 3.1]{Stuart10} suggests taking priors as
-"Laplacian-like" operators to power smaller than $-\frac{d}{2}$, to
-maintain regularity of posterior realizations. Thus, even though the
-computational domain is getting larger and sparser as the dimension
-$d$ grows, clusterization is not necessarily harder to achieve.
-
-%% In our model we did not specify any specific prior, so it is possible
-%% that one will be able to devise a prior that will not result in
-%% clusterization. Indeed, if the rate of increase of prior precision
-%% eigenvalues is small,
-
 As mentioned in the Limitations section, our model does not accomodate

### 2024-07-21 latex/intro.tex
before diff

@@
 experimental design involves choosing a set of measurements to be
-taken. In a medical experiment, experimental design may involve
-choosing participants to whom some drug will be administered, and
-measuring their blood pressure. Experimental design also arises 
-process of searching for oil. Since measuremnts are expensive and
+taken. For example, in a medical experiment, experimental design may
+involve choosing participants to whom some drug will be administered,
+and measuring their blood pressure. Experimental design also arises in
+the process of searching for oil. Since measuremnts are expensive and
 require digging holes in the ground, only a limited number of
@@
 when solving an \emph{inverse problem} --- i.e. making inference of
-the physical world utilizing a physical model of some phenomenon of
-interest. In an inverse problem, we seek to infer physical quantities,
-based on observations and a model of the physical world, typically
-phrased in the language of ordinary or partial differential equations.
+the physical world utilizing a physical model a phenomenon of interest
+\cite{tarantola,somersalo}. In an inverse problem, we seek to infer
+physical quantities, based on observations and a model of the physical
+world. Models are typically phrased in the language of ordinary or
+partial differential equations.
 
 Inverse problems are ubiquitous in science and engineering, and we
-will now give several illustrative examples. In electric impedance
+will now give several illustrative examples. In electrical impedance
 tomography (EIT) we seek to infer the impedance of the inside of a
-human body from observations of electrical current impedence measured
+human body from observations of electrical current impedance measured
 at specific electrode locations on the skin, leveraging known
@@
 sent through the human body in the presence of a strong magnetic field
-and the response of our bodies' contents to such perutrbations is
+and the response of one's body contents to such perutrbations is
 measured, allowing a radiologist to view the internals of a body in a
@@
 into the ground --- and data of travel and return times of said waves
-are recorded. Since travel and return time is influenced by the
-properties of the subsurface materials, such as density, elasticity,
-and the presence of fluids or voids, we can combine travel data with a
-geophisical models of contents of the earth's crust to facilitate
-reconstructing and inferring the structure of the subsurface in a
-process called \emph{borehole tomography}
-\cite{horesh2008borehole}. Inverse problems also arise in many other
-areas of seismology and geology \cite{rabinowitz1990, steinberg1995}
-and medical imaging \cite{tarantola}. While in many inverse problems
-the goal is to infer some parameter, in all of the above examples and
-in the formulation we consider here, the goal is to infer some
-\emph{function} over $ \Omega$, where \(\Omega \subseteq \mathbb{R}^d,
-d=1,2,3\) is the spatial domain of interest.
+are recorded. Travel and return time is influenced by the properties
+of the subsurface materials, such as density, elasticity, and the
+presence of fluids or voids, combining travel data with a geophisical
+models of contents of the earth's crust to facilitate reconstructing
+and inferring the structure of the subsurface in a process called
+\emph{borehole tomography} \cite{horesh2008borehole}. Inverse problems
+also arise in many other areas of seismology and geology
+\cite{rabinowitz1990, steinberg1995} and medical imaging
+\cite{tarantola}. While in many inverse problems the goal is to infer
+some parameter, in all of the above examples and in the formulation we
+consider here, the goal is to infer some \emph{function} over $
+\Omega$, where (usually) \(\Omega \subseteq \mathbb{R}^d, d=1,2,3\) is
+the spatial domain of interest.
 
@@
 henceforth, see Section \ref{subsec:1d heat} for more details). The
-heat equation is a partial differential equation which is in fact a
-set of three equations: eqs.~\eqref{eq:heat1}, \eqref{eq:heat2}, and
+heat equation is a partial differential equation compose of three
+equations: eqs.~\eqref{eq:heat1}, \eqref{eq:heat2}, and
 \eqref{eq:heat3}. Readers less familiar with partial differential
@@
 distribution the \emph{initial condition}. We wait for a short time
-and let the heat dissipate a bit inside the rod. The final heat
+$T$ and let the heat dissipate a bit inside the rod. The final heat
 distribution in the rod is dictated by the following set of equations
@@
 condition} which is best understood as an absorbing boundary: heat
-that interacts with the boundary immediately
-disappears. Eq.~\eqref{eq:heat3} describes the \emph{initial
-condition} of the heat, i.e.~the heat distribution at time $t=0$.
+that interacts with the boundary is immediately
+absorbed. Eq.~\eqref{eq:heat3} describes the \emph{initial condition}
+of the heat, i.e.~the heat distribution at time $t=0$. Thus, we have
+described the \emph{forward} problem: how heat evolves in time,
+starting at $t=0$ up until $t=T$.
 
-So far, we described the \emph{forward} problem: how heat evolves in
-time. In the inverse problem, our goal is to infer the initial
-condition $u_0 = u(\cdot, 0)$ from observations of the final condition
-$u(\cdot, T)$. We are allowed to take some number of temperature
-measurements on the rod once our waiting time has passed. The inverse
-problem of the 1D heat equation is difficult ("ill-posed") since heat
-spreads in a diffusive manner and any roughness in the initial
-condition is quickly smoothed, see Supplementary movies S1 and S2. We
-will give a fully Bayesian formulation of a generic linear inverse
-problem in Section \ref{section:prelims}.
+In the inverse problem of the 1D heat equation, our goal is to infer
+the initial condition $u_0 = u(\cdot, 0)$ from observations of the
+final condition $u(\cdot, T)$. We are not able to measure $u(\cdot,
+T)$ at every point in the domain $\Omega$. Rather, we take some number
+of temperature measurements on the rod once our waiting time $T$ has
+passed and try to infer $u(\cdot, 0)$ from these measurements.
+
+The inverse problem of the 1D heat equation is difficult ("ill-posed")
+since heat spreads in a diffusive manner and any roughness in the
+initial condition is quickly smoothed, see Supplementary movies S1 and
+S2. For ill-posed problems regularization is in order, and we
+implement such regularization via a fully Bayesian formulation of the
+inverse problem (see Section \ref{section:prelims} for details).
 
@@
     \centering
-    \includegraphics[height=0.5\textwidth]{dst_modelError0.pdf}
+    \includegraphics[height=0.5\textwidth]{dst_modelError0.png}
     \caption{Measurement clusterization for D-optimal designs for the
@@
 assigned to 12 of the plots". It seems reasonable the researchers
-replicate the phenomenon they are trying to study. Clusterization
-seems like a whole different beast: a clustered design in the rainfall
-experiment would imply the researchers should take a repeated
-measurement \emph{on the same plot}, at the expense of measuring grass
-growth in other plots. In sharp contrast to repetition and
-replication, clusterization is highly nonintuitive and, as we will
-see, its origins run considerably deeper.
+replicate the phenomenon they are trying to study. Clusterization is
+different: a clustered design in the rainfall experiment would imply
+the researchers should take a repeated measurement \emph{on the same
+plot}, at the expense of measuring grass growth in other plots. In
+sharp contrast to repetition and replication, clusterization is highly
+nonintuitive and, as we will see, its origins run considerably deeper.
 
@@
 \subsection{Contribution}
-The primary objective of this study is to provide a comprehensive
-understanding of measurement clusterization by addressing the
-aforementioned questions. Our focus centers around investigating the
-Bayesian D-optimality criterion. We conduct an analysis of Bayesian
-D-optimal designs within the context of linear inverse problems over
-Hilbert spaces. We propose a novel relaxed model for D-optimality that
+The primary objective of this study is to provide a deep understanding
+of measurement clusterization by addressing the aforementioned
+questions. Our focus centers around investigating the Bayesian
+D-optimality criterion. We conduct an analysis of Bayesian D-optimal
+designs within the context of linear inverse problems over Hilbert
+spaces. We propose a novel relaxed model for D-optimality that
 maintains analytical tractability and enables the identification of
@@
 
-The implication of the answers we give to Questions \ref{q:generic},
-\ref{q:mitigate}, and \ref{q:why} is two-fold:
-
 The implication of our answer to Question \ref{q:generic} is that
@@
 not be surprised or wary. Then, in our answer to Question \ref{q:why},
-we explain what is the source of clusterization. As it appears, the
-source is universal: a D-optimal design reduces uncertainty for a
+we explain what is the cause of clusterization. As it appears, the
+cause is universal: a D-optimal design reduces uncertainty for a
 select set of prior covariance eigenvectors --- those with the most

### 2024-07-21 latex/presentation.tex
before diff

\documentclass[aspectratio=169]{beamer}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{amsmath}
\usepackage{amssymb}

\usetheme{CambridgeUS}
\usecolortheme{beetle}

\title{Measurement Clusterization in Bayesian D-optimal Designs in Infinite Dimensions}
\author{Yair Daon}
\institute{Azrieli Faculty of Medicine, Bar-Ilan University}
\date{\today}

\begin{document}

\begin{frame}
\titlepage
\end{frame}

\begin{frame}{Outline}
\tableofcontents
\end{frame}

\section{Introduction}

\begin{frame}{Motivation}
\begin{itemize}
    \item Selecting optimal measurements is crucial for inferring parameters of physical processes
    \item Measurements should enhance accuracy, reduce costs, or both
    \item A-optimal and D-optimal design criteria are widely studied
    \item Optimal designs can yield tightly clustered measurement locations (clusterization)
\end{itemize}
\end{frame}

\begin{frame}{Example: 1D Heat Equation}
\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{fig1.png}
    \caption{Measurement clusterization for optimal designs when inverting for the initial condition of the 1D heat equation.}
\end{figure}
\end{frame}

\begin{frame}{Existing Approaches to Mitigate Clusterization}
\begin{itemize}
    \item Merging close measurements
    \item Imposing distance constraints between measurements 
    \item Introducing correlated errors accounting for model misspecification
    \item Choosing measurements from a coarse grid
\end{itemize}
\textbf{Issue:} These approaches fundamentally alter the optimal design problem
\end{frame}

\begin{frame}{Key Questions}
\begin{enumerate}
    \item Why does imposing correlations between observations alleviate clusterization?
    \item Is measurement clusterization a generic phenomenon? 
    \item Why does measurement clusterization occur?
\end{enumerate}
\end{frame}

\section{Problem Formulation}

\begin{frame}{Bayesian Linear Inverse Problem}
Let $H_p$, $H_o$ be Hilbert spaces and $F: H_p \to H_o$ a linear compact forward operator.\\~\\
Measurement operator $O: H_o \to \mathbb{R}^m$ takes $m$ noisy measurements:
\begin{equation*}
    d = OFm + O\varepsilon' + \varepsilon
\end{equation*}
where $\varepsilon' \sim \mathcal{N}(0,\Gamma_\text{model})$ is model error and $\varepsilon \sim \mathcal{N}(0,\sigma^2I_m)$ is measurement noise.
\end{frame}

\begin{frame}{Prior and Posterior}
Prior on $m$ is Gaussian: $\mu_\text{pr} = \mathcal{N}(m_\text{pr},\Gamma_\text{pr})$\\~\\  
Posterior $\mu_\text{post}^{d,O}$ is also Gaussian with covariance:
\begin{equation*}
    \Gamma_\text{post} = \left( \Gamma_\text{pr}^{-1} + F^*O^*\Sigma^{-1}OF \right)^{-1}
\end{equation*}
where $\Sigma(O) = O\Gamma_\text{model}O^* + \sigma^2I_m$.
\end{frame}

\begin{frame}{Bayesian D-Optimal Design}
Maximizes expected KL divergence between posterior and prior:
\begin{equation*}
    \Psi(O) := \mathbb{E}_d\left[ D_\text{KL}(\mu_\text{post}^{d,O} \| \mu_\text{pr}) \right] = \frac{1}{2} \log\det\left(I + \Gamma_\text{pr}^{1/2}F^*O^*\Sigma^{-1}OF\Gamma_\text{pr}^{1/2}\right)
\end{equation*}
D-optimal design operator:
\begin{equation*}
    O = \arg\max_{\|o_j\|=1, j=1,\dots,m} \Psi(O)
\end{equation*}
\end{frame}

\section{Main Results}

\begin{frame}{Theorem 1}
Let $\{\lambda_i\}$ be the eigenvalues of $F\Gamma_\text{pr}F^*$ and $\{\eta_i\}$ the eigenvalues of $O^*O$. \\~\\
For a D-optimal design $O$: 
\begin{enumerate}
    \item $\text{tr}\{O^*O\} = m$
    \item $O^*O$ and $F\Gamma_\text{pr}F^*$ are simultaneously diagonalizable
    \item $\Psi(O) = \frac{1}{2} \sum_{i=1}^k \log(1 + \sigma^{-2}\lambda_i\eta_i)$ where $k=\text{rank}(O^*O) \leq m$
    \item $\eta_i = \begin{cases} 
          \frac{m}{k} - \sigma^2\lambda_i^{-1} + \sigma^2\frac{1}{k}\sum_{j=1}^k \lambda_j^{-1}, & 1 \leq i \leq k \\
          0, & i > k
       \end{cases}$
    \item Covariance of $F^*\mu_\text{post}^{d,O}$ is $\left((F\Gamma_\text{pr}F^*)^{-1} + \sigma^{-2}O^*O\right)^{-1}$ 
\end{enumerate}
\end{frame}

\begin{frame}{Model Error Mitigates Clusterization}
For $\Gamma_\text{model} \neq 0$, no increase in $\Psi$ is achieved by repeating a measurement as $\sigma^2 \to 0$. \\~\\
Designs exhibiting clusterization are not D-optimal.\\~\\
\textbf{Conclusion:} For small $\sigma^2$, clusterization is mitigated by non-zero model error $\Gamma_\text{model}$.
\end{frame}

\begin{frame}{D-Optimal Designs Without Model Error}
Necessary condition for D-optimality when $\Gamma_\text{model}=0$:
\begin{equation*}
    \sigma^{-2}F\Gamma_\text{post}F^*O^* = O^*\Xi
\end{equation*}
where $\Xi$ is diagonal. This is a nonlinear eigenvalue problem.\\~\\
Key results:
\begin{itemize}
    \item $O^*O$ and $F\Gamma_\text{pr}F^*$ are simultaneously diagonalizable
    \item D-optimal designs focus on a subset of prior covariance eigenvectors with largest eigenvalues
\end{itemize}
\end{frame}

\begin{frame}
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{fig2.png}
    \caption{D-optimal measurement locations and weighted eigenvectors for 1D heat equation with $m=4$ measurements.}
\end{figure}
\end{frame}

\begin{frame}{Intuition for Theorem 1}
\begin{columns}
\begin{column}{0.5\textwidth}
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{fig3a.png}
    \caption{D-optimal design}
\end{figure}
\end{column}
\begin{column}{0.5\textwidth}
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{fig3b.png}
    \caption{Sub-optimal design}
\end{figure}
\end{column}
\end{columns}
D-optimal designs increase precision where it is lowest in the prior.
\end{frame}

\section{Cause of Clusterization}

\begin{frame}{Why Clusterization Occurs}
\begin{itemize}
    \item D-optimal designs focus on eigenvectors with highest prior variance
    \item In spatial problems, designs favor locations where:
    \begin{enumerate}
        \item Select eigenvectors are large
        \item Other eigenvectors are small
    \end{enumerate}
    \item Number of such locations is limited
    \item Clusterization occurs due to pigeonhole principle - more measurements than optimal locations
\end{itemize}
\end{frame}

\section{Conclusions}

\begin{frame}{Summary}
\begin{itemize}
    \item Investigated measurement clusterization in Bayesian D-optimal designs
    \item Proposed analytically tractable model for D-optimality over Hilbert spaces
    \item Showed clusterization is generic for independent Gaussian errors
    \item Proven model error mitigates clusterization
    \item Characterized D-optimal designs as reducing uncertainty across subset of prior eigenvectors
    \item Argued clusterization occurs due to pigeonhole principle
\end{itemize}
\end{frame}

\begin{frame}{Limitations and Future Work}
\begin{itemize}
    \item Generic model does not correspond to a specific real problem
    \item Extend analysis to A-optimal designs
    \item Investigate clusterization in nonlinear inverse problems
    \item Develop computational methods to avoid clusterization without compromising optimality
\end{itemize}
\end{frame}

\begin{frame}{References}
\footnotesize
\bibliographystyle{plain}
\bibliography{references}
\end{frame}

\begin{frame}
\begin{center}
    {\huge Thank You!}\\~\\
    Questions?
\end{center}
\end{frame}

\end{document}

### 2024-07-21 latex/refs.tex
before diff

@@
 problems/uncertainty quantification and the conclusions should be
-useful to practitioners in this field.  However, I feel changing the
-assumption from independent to correlated measurements is quite
-artificial: people tweak a data generation process assumption to solve
-an optimal design problem, and the author just takes the assumption as
-it is. Since the author proves the clusterization issue under the
-independent measurement assumption, I expect the author to go deeper
-and explain why an ordinary data generation assumption + a common
-optimal design leads to a “suboptimal” design? Not just by the
-mathematical proof, but from an information theory point of view. For
-example, does the redundancy points mean additional points does not
-reduce the variance of the estimation at all? If that’s the case, how
-does correlated measurement assumption “help” with it?
+useful to practitioners in this field.
+
+\answer: I appreciate the kind words.
+
+However, I feel changing the assumption from independent to correlated
+measurements is quite artificial: people tweak a data generation
+process assumption to solve an optimal design problem, and the author
+just takes the assumption as it is. Since the author proves the
+clusterization issue under the independent measurement assumption, I
+expect the author to go deeper and explain why an ordinary data
+generation assumption + a common optimal design leads to a
+“suboptimal” design? Not just by the mathematical proof, but from an
+information theory point of view. For example, does the redundancy
+points mean additional points does not reduce the variance of the
+estimation at all? If that’s the case, how does correlated measurement
+assumption “help” with it?
+
+\answer: Thanks for your questions. I am not a proponent of using
+correlated noise, I just wanted to show the usefullness of my model by
+rigorously proving how correlated noise mitigates clusterization. I
+have added several lines clearing this issue.
 
@@
 
+\answer:
+
 Nearly all textbooks on classical design of experiments (e.g. Morris,
@@
 repeated measurements (measurement clusterisation) is an undesirable
-property. If one looks at, for example, Fedorov (1996) and Nyberg et
+property.
+
+\answer: I do not view clusterization as an undesired property, I only
+mentioned many authors consider it to be undesirable. I have added a
+section contrasting repetition and replication with clusterization and
+why I believe they may be different.
+
+If one looks at, for example, Fedorov (1996) and Nyberg et
 al (2012), it is an undesirable property because they are interested
@@
 
+\answer: again
+
 \begin{enumerate}
@@
   the Abstract and the first paragraph of the paper.
+
+  \answer: I do consider experiments where repeated measurements are
+  possible. I am giving insight on why such designs are useful, even
+  though repeated measurements may be non-intuitive in certain
+  settings.
 
@@
   repeated design points.
+
+  \answer: great comment
 \end{enumerate}

### 2024-07-21 latex/vanishing.tex
before diff

@@
 
-\begin{proof}
-  The proof amounts to using Woodbury's matrix identity twice with a
-  regularization trick. The standard proof for Woodbury's matrix
-  identity works for separable Hilbert spaces, as long as all terms
-  are well defined. Unfortunately, $\obs^*\obs$ is not invertible, so
-  we force it to be. Recall Woodbury's matrix identity:
-  \begin{equation}\label{eq:WMI}
-    (A + UCV)^{-1} = A^{-1} - A^{-1}U(C^{-1} + VA^{-1}U)^{-1}VA^{-1}. 
-  \end{equation}
-
-  Denote:
-  \begin{align}\label{eq:notation}
-    \begin{split}
-      A :&= \prcov^{-1} \\
-      U :&= \fwd^* \\
-      V :&= \fwd \\
-      C :&= \sigma^{-2} (\obs^*\obs+\nu I),\ \nu > 0
-    \end{split}
-  \end{align}
-
-  Then, \eqref{eq:WMI} with the notation \eqref{eq:notation} implies:
-  \begin{align}\label{eq:first}
-    \begin{split}
-    &\fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\nu I) \fwd )^{-1}\fwd^* \\
-    &\ \ \ \ \ = \fwd ( \prcov - \prcov \fwd^* ( \sigma^2(\obs^*\obs + \nu I)^{-1} + \fwd \prcov \fwd^* )^{-1} \fwd \prcov ) \fwd^*. \\
-      %
-      %
-      %% &= X - X(\sigma^2Y_{\nu}^{-1} + X)^{-1}X
-  %%   \end{split}
-    %% \end{align*}
-    \end{split}
-  \end{align}
-  Now, denote:
-  \begin{align*}
-    X :&= \fwd\prcov \fwd^*, \\
-    Y_{\nu} :&= \obs^*\obs + \nu I,
-  \end{align*}
-  and observe that $Y_{\nu}$ is invertible. Substitute $X, Y_{\nu}$
-  into the RHS of \eqref{eq:first}:
-   \begin{equation}\label{eq:second}
-     \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\nu I) \fwd )^{-1}\fwd^* = X - X(\sigma^2Y_{\nu}^{-1} + X)^{-1}X
-   \end{equation}
-
-   Note that $X + \sigma^2 Y_{\nu}^{-1}$ is invertible, as the sum of
-   two positive definite operators. Now, let
-   \begin{align}\label{eq:notation2}
-     \begin{split}
-       A :&= X^{-1}, \\
-       C :&= \sigma^2Y_{\nu}^{-1}, \\
-       U :&= I, \\
-       V :&= I.
-     \end{split}
-   \end{align}
-   Apply \eqref{eq:WMI} with the notation \eqref{eq:notation2} to the
-   RHS of \eqref{eq:second}:
-  \begin{align*}
-    \begin{split}
-      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\nu I) \fwd )^{-1}\fwd^* &= X - X(\sigma^2Y_{\nu}^{-1} + X)^{-1}X \\
-      %
-      %
-      %
-      &= (X^{-1} + \sigma^{-2}Y_{\nu})^{-1} \\
-      %
-      %
-      %
-      &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \nu I))^{-1}.
-    \end{split}
-  \end{align*}
-
-  We conclude that $\forall \nu > 0$
-  \begin{align*}
-    \begin{split}
-      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\nu I) \fwd )^{-1}\fwd^* 
-     &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \nu I))^{-1}.
-    \end{split}
-  \end{align*}
-  Letting $\nu \to 0$ completes the proof.
-\end{proof}
+As we mentioned in Section \ref{subsec:bayesian inverse problems},
+$\fwd \prcov \fwd^*$ is the prior covariance $\hilo$ and could be
+safely assumed invertible. Proof of Proposition \ref{prop:twice
+  woodbury} is delegated to the Supplementary. It amounts to using
+Woodbury's matrix identity twice. The standard proof for Woodbury's
+matrix identity works for separable Hilbert spaces, as long as all
+terms are well defined. Unfortunately, $\obs^*\obs$ is not invertible,
+so we utilize a regularization trick to force it to be.
 
@@
 \end{lemma}
-\begin{proof}
-  First, enumerate the eigenvalues of $C + \sum_{j=1}^m
-  \func_j\func_j^*$ as $\xi_1,\dots,\xi_\ell$. Denote the indices of
-  the eigenvectors corresponding to $\xi_i$
-  \begin{equation*}
-    S_i := \{ 1 \leq k \leq m | (C + \sum_{j=1}^m \func_j\func_j^* )\func_k = \xi_i \func_k \}.
-  \end{equation*}
-  Define further
-  \begin{equation*}
-    A_i := \sum_{k \in S_i} \func_k \func_k^*,
-  \end{equation*}
-  which is self-adjoint. Two observations are in order. First,
-  $\sum_{j=1}^m \func_j\func_j^* = \sum_{i=1}^\ell A_i$. Second, $A_i
-  \func_k = 0$ if $k\not \in S_i$, since eigenvectors of different
-  eigenvalue are orthogonal. For $k \in S_i$
-  \begin{equation}\label{eq:on vi}
-    \xi_i \func_k = (C + \sum_{j=1}^m \func_j \func_j^* ) \func_k = (C + A_i) \func_k.
-  \end{equation}
-  Let $V_i := span \{\func_k \}_{k\in S_i}$. Observe that $V_i$ is
-  invariant under $A_i$, by definition, and under $C$, by \eqref{eq:on
-    vi}. A second application of \eqref{eq:on vi} shows that $A_i =
-  \xi_iI - C$ on $V_i$. This immediately implies $A_i$ and $C$ are
-  simultaneously diagonalizable on $V_i$. This holds for every $1 \leq
-  i \leq \ell$ and we conclude that $C$ and $A$ are simultaneously
-  diagonalizable.
-\end{proof}
+The proof of Lemma \ref{lemma:sim diag} is delegated to the
+Supplementary.
 
@@
 
-\begin{proof}
-  Let $\Phi(\eta) = \sum_{i=1}^k \eta_i - m$ and $\Omega_j(\eta) =
-   -\eta_j$. Then
-  \begin{align*}
-    \begin{split}
-      \frac{\partial \tar}{\partial \eta_i}  &=
-       \frac12 \frac{\sigma^{-2}\lambda_i}{1 + \sigma^{-2} \lambda_i\eta_i} = \frac12 \frac{1}{\sigma^{2}\lambda_i^{-1} + \eta_i} \\
-      %
-      %
-      %
-      \frac{\partial\Phi}{\partial \eta_i} &= 1 \\
-      %
-      %
-      %
-      \frac{\partial \Omega_j}{\partial \eta_i} &= -\delta_{ij}      
-    \end{split}
-  \end{align*}
-
-  From the KKT conditions, there are $\alpha, \beta_i$ such that for $i=1,\dots,m$:
-  \begin{align}
-    \begin{split}
-      -\frac12 \frac{1}{\sigma^{2}\lambda_i^{-1} + \eta_i} + \alpha - \beta_i  &= 0 \\
-      %
-      %
-      %
-      \eta_i &\geq 0\\
-      %
-      %
-      %
-      \beta_i &\geq 0\\
-      %
-      %
-      %
-      \beta_i \eta_i &= 0\\
-      %
-      %
-      %
-      \sum_{i=1}^m \eta_i &= m 
-    \end{split}
-  \end{align}
-
-  Then, for $i \in A$:
-  \begin{align*}
-    \begin{split}
-      \beta_i &= 0\\
-      %
-      %
-      %
-      \sigma^{2}\lambda_i^{-1} + \eta_i  &= \frac{1}{2\alpha}%  \text{ for }  1 \leq i \leq k.\\
-      %
-      %
-      %
-      %% \sigma^{-2} \lambda_i &= 2\alpha - 2\beta_i, i \in B \\
-    \end{split}
-  \end{align*}
-
-  Summing over $i \in A$, substituting $\sum_{i\in A} \eta_i =
-  \sum_{i=1}^m \eta_i = m$, and dividing by $k:= |A|$:
-  $$
-  \frac{1}{2\alpha} = \frac{m}{k} + \frac{\sigma^2}{k} \sum_{j\in A} \lambda_j^{-1}.
-  %\Rightarrow \alpha = \left ( \frac{2m}{k} + \frac{2\sigma^2}{k}
-  %\sum_{i\in A} \lambda_i^{-1} \right )^{-1}.
-  $$
-
-  Consequently:
-  \begin{align} \label{eq:etas}
-    \begin{split}
-      \eta_i &= \frac{1}{2\alpha} - \sigma^2 \lambda_i^{-1}\\
-      %
-      %
-      %
-      &=\frac{m}{k} + \frac{\sigma^2}{k} \sum_{j\in A} \lambda_j^{-1} - \sigma^2 \lambda_i^{-1}.%\\
-      %
-      %
-      %
-      %% &= \frac{m}{k} + \frac{\sigma^2}{k} \sum_{j \in A, j\neq i} \lambda_j^{-1} - \frac{k-1}{k}\sigma^2 \lambda_i^{-1}\\
-      %% ,\ i \in A.
-    \end{split}
-  \end{align}
-\end{proof}
-
-The final ingredient we require for the proof of Theorem
-\ref{thm:char} is:
+The proof of Proposition \ref{prop:kkt} amounts to utilizing the
+Karush-Kuhn-Tucker conditions and is delegated to the
+Supplementary. The final ingredient we require for the proof of
+Theorem \ref{thm:char} is:
 
@@
 
-\begin{proof}
-  Let us diagonalize $M$, so that $M = U D U^t$ with $D =
-  \diag(d_1,\dots,d_k)$ and $U \in \R^{k \times k }$ orthogonal. Let
-  $S \in \R^{k \times m}$ with $S_{ii} = \sqrt{d_{i}}$ and zeros
-  otherwise. Define $A:= U S V^t$, where $V \in \R^{m \times m}$ is
-  orthogonal and will be further restricted later. Then $AA^t = U
-  SV^tVS^t U^t = UDU^t$, so $AA^t$ has the required eigenvalues and
-  eigenvectors by construction. If we can choose $V$ such that $A$
-  also satisfies the unit norm constraints we are done. These
-  constraints are, for $j=1,\dots,m$:
-  \begin{equation}\label{eq:V constraints}
-   1 = [A^tA]_{jj} = [V S^tS V^t]_{jj},
-  \end{equation}
-  and we can expect to do this since we assumed $\ttr D = m$.
-
-  Define $C = S^tS - I \in \R^{m \times m}$. Note that $\ttr C = 0$ and
-  $C$ is diagonal with non-zero entries $d_i-1,i=1,\dots,k$. It suffices
-  to find $V$ orthogonal such that $V C V^t$ has zero diagonal. We
-  construct such $V$ by sequentially inserting zeros in the diagonal
-  and not destroying zeros we already introduced, starting from the
-  last diagonal entry and moving to the first. Since $c_{mm} \neq 0$ ,
-  let $p < m$ such that $c_{pp}c_{mm} < 0$ (such $p$ exists because
-  the trace is zero) and let $\theta \in (0,\pi)$. Define a Givens
-  rotation $R^{(m)} \in \R^{m \times m}$ by
-  \begin{equation*}
-    r^{(m)}_{ab} :=
-    \begin{cases}
-      1 & a = b \neq p \text{ or } a = b \neq m \\
-      \cos \theta & a = b = p  \\
-     -\sin \theta & a = p, b = m\\
-      \cos \theta & a = b = m \\
-      \sin \theta & a = m, b = p \\ 
-      0 & o.w
-    \end{cases}
-  \end{equation*}
-  Note that conjugating a matrix by $R^{(m)}$ changes only its $m$ and
-  $p$ rows and columns. We want to choose $\theta$ such that
-  \begin{equation}\label{eq:mm}
-    0 = [R^{(m)} C (R^{(m)})^t]_{mm} = \cos^2 \theta c_{mm} + 2\cos \theta \sin
-    \theta c_{mp} + \sin^2\theta c_{pp},
-  \end{equation}
-  and it suffices to choose $\theta$ such that
-  \begin{equation*}
-    c_{mm} \cot^2 \theta + 2 c_{mp} \cot \theta + c_{pp} = 0.
-  \end{equation*}
-  This quadratic in $\cot\theta$ has a real solution, since
-  $c_{pp}c_{mm} < 0$ by assumption and we can find $\theta \in
-  (0,\pi)$ such that \eqref{eq:mm} is satisfied. We continue to find
-  $R^{(m-1)}$ that leaves row and column $m$ unchanged and
-  continue introducing zeros to the diagonal. The assumption $\ttr D =
-  m \Rightarrow \ttr C = 0$ guarantees we can do that. Taking $V:=
-  R^{(1)} R^{(2)} \dots R^{(m-1)}R^{(m)}$ completes the proof.
-\end{proof}
+The proof of Lemma is delegated to the Supplementary. It is however
+important to note that this proof is construcive. It will give us a
+way to construct D-optimal designs, once we fully characterize them in
+Theorem \ref{thm:char} below.
 
@@
 
-\begin{figure}
+\begin{figure}\label{fig:tikz clusterization}
   \centering

### 2024-07-21 bayes.py

@@ -1,0 +1,79 @@
+import numpy as np
+import matplotlib.pyplot as plt
+from scipy.stats import norm
+
+
+fs = 22
+def gaussian(x, mu, sigma):
+    return (1 / (sigma * np.sqrt(2 * np.pi))) * np.exp(-0.5 * ((x - mu) / sigma) ** 2)
+
+# Set up the x-axis
+x = np.linspace(-10, 10, 1000)
+
+# Prior
+prior_mu, prior_sigma = 0, 2
+prior = gaussian(x, prior_mu, prior_sigma)
+
+# Likelihood
+likelihood_mu, likelihood_sigma = 5, 1.5
+likelihood = gaussian(x, likelihood_mu, likelihood_sigma)
+
+# Posterior (unnormalized)
+posterior_unnorm = prior * likelihood
+
+# Calculate evidence (normalization factor)
+evidence = np.trapz(posterior_unnorm, x)
+
+# Normalize the posterior
+posterior = posterior_unnorm / evidence
+
+# Calculate means and variances
+posterior_mu = np.sum(x * posterior) / np.sum(posterior)
+posterior_var = np.sum((x - posterior_mu)**2 * posterior) / np.sum(posterior)
+posterior_sigma = np.sqrt(posterior_var)
+
+# Plot
+fig, ax = plt.subplots(figsize=(14, 10))
+ax.plot(x, prior, color='blue')
+ax.plot(x, likelihood, color='red')
+ax.plot(x, posterior, color='green')
+
+#ax.set_title("Visualization of Bayes' Rule with Gaussian Distributions", fontsize=fs)
+#ax.set_xlabel('x', fontsize=fs)
+#ax.set_ylabel('Probability Density', fontsize=fs)
+#ax.xtick_params([])#axis='both', which='major', labelsize=fs)
+ax.grid(True, alpha=0.3)
+
+# Add annotations with arrows
+ax.annotate(f'Prior (μ={prior_mu}, σ={prior_sigma})', 
+            xy=(prior_mu, gaussian(prior_mu, prior_mu, prior_sigma)), 
+            xytext=(prior_mu-3, 0.28),
+            arrowprops=dict(facecolor='blue', shrink=0.05),
+            color='blue', fontsize=fs)
+
+ax.annotate(f'Likelihood (μ={likelihood_mu}, σ={likelihood_sigma})', 
+            xy=(likelihood_mu, gaussian(likelihood_mu, likelihood_mu, likelihood_sigma)), 
[…]

### 2024-07-21 clusterization.py

@@ -5,6 +5,7 @@ from src.probability import Prior, Posterior
 from src.forward import Heat
 from src.observations import PointObservation
 
+FS = 24 ## Font size
 
 def main():
 
@@ -13,7 +14,7 @@ def main():
     L = 1 ## Length of computational domain
     time = 3e-2 ## Time for heat dissipation in arbitrary units.
     alpha = 1. ## Coefficient of Laplacian in heat equation
-    gamma = -1.1 ## Exponent of Laplacian in prior. 
+    gamma = -1. ## Exponent of Laplacian in prior. 
 
     ## Choose one, comment out the other
     transform = 'dct' ## dct - Discrete Cosine Transform. Corresponds
@@ -56,10 +57,42 @@ def main():
                      model_error=model_error)
 
 
+    
+    F_gamma_Fstar = fwd.multiplier**2 * pr.multiplier ## Prior covariance in H_o
+    precision = 1 / F_gamma_Fstar ## Prior precision
+    
+    k = 2 ## rank O^*O
+    m = 4 ## Number of measurements
+
+    ## See last part of Theorem in paper. This is the value the pusforward 
+    ## posterior precision takes for the first k eigenvalues
+    sigSqr = 2e-4
+    theta = ((np.sum(precision[:k]) + m /sigSqr ) / k) 
+
+    ## Plotting
+    n = 4 ## Truncate after this many eigevanlues, for visualization purposes
+    y = precision[:n]
+    x = np.arange(n, dtype=int) + 1
+    xtra = np.zeros(n)
+    xtra[:k] = theta - y[:k]
+   
+    plt.bar(x, y, label=r"prior")
+    plt.bar(x, xtra, bottom=y, label="measurements")
+    plt.yscale('log')
+    plt.xlabel("Eigenvector", fontsize=FS)
+    plt.ylabel("Precision", fontsize=FS)
+    plt.tight_layout()
+    plt.savefig(f"latex/FgammaFstar_modelError{model_error}.png")
+    
+    
+    ###################################
+    ## find and plot optimal designs ##
+    ###################################
+    
     dic = {}
 
     ## Form  number of observations in this range
-    for m in range(2, 7):
+    for m in range(1, 7):
 
         ## Calculate an optimal design
         design = post.optimize(m=m, n_iterations=500)['x']
@@ -69,10 +102,8 @@ def main():
 
         ## Print it so we see
[…]

### 2024-07-21 latex/bayess.py

@@ -1,6 +1,7 @@
 import numpy as np
 import matplotlib.pyplot as plt
 from scipy.stats import norm
+import math
 
 plt.rcParams.update({'font.size': 16})
 
@@ -34,16 +35,12 @@ plt.plot(x, posterior, 'red', label='Posterior (m|d)')
 plt.plot(observation, 0, 'go', markersize=10, label='Observation (d)')
 
 # Customize the plot
-# plt.xlabel('μ')
-# plt.ylabel('Probability Density')
-# plt.title('Bayesian Inference with Single Observation')
 plt.legend()
 plt.grid(True, linestyle=':', alpha=0.7)
 
 # Set axis limits
 plt.ylim(-0.005, 0.12)
 plt.xlim(30, 100)
-
 
 plt.xticks([])
 plt.yticks([])
@@ -53,8 +50,164 @@ plt.xlabel('')
 plt.ylabel('')
 
 plt.box(False)
-    
+
+# Calculate the y-values for the modes
+prior_mode_y = norm.pdf(prior_mean, prior_mean, prior_std)
+posterior_mode_y = norm.pdf(posterior_mean, posterior_mean, posterior_std)
+
+
+
 # Show the plot
 plt.tight_layout()
 plt.savefig('bayes.png')
-#plt.show()
+
+
+
+
+# Add black arrow from prior mode to posterior mode
+arrow = plt.annotate('', xy=(posterior_mean, posterior_mode_y), xytext=(prior_mean, prior_mode_y),
+             arrowprops=dict(facecolor='black', shrink=0.05, width=1, headwidth=8))
+
+# Calculate angle of the arrow
+dx = posterior_mean - prior_mean
+dy = posterior_mode_y - prior_mode_y
+angle = math.pi/4 +0.15#math.atan2(dy, dx)
+
+# Calculate the perpendicular offset for the text
+offset = 0.01  # Adjust this value to move text up/down
+perp_dx = -math.sin(angle) * offset
+perp_dy = math.cos(angle) * offset
+
+
+# Add label for the arrow, parallel to it and slightly above
+text_x = (prior_mean + posterior_mean) / 2 + perp_dx
+text_y = (prior_mode_y + posterior_mode_y) / 2 + perp_dy
+plt.text(text_x,
+         text_y,
+         r'$KL(\mu_{\text{post}}||\mu_{\text{pr}})$',
+         ha='center',
+         va='center', 
[…]

### 2024-07-21 latex/flat.tex
With the flattened file from master

% \documentclass[ba,preprint]{imsart}% use this for supplement article
\documentclass[ba]{imsart}
%\input{definitions}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsthm}
%% \usepackage{amsaddr}

%% \theoremstyle{plain}
%% \newtheorem{theorem}{Theorem}
%% \newtheorem{corollary}[theorem]{Corollary}
%% \newtheorem{lemma}[theorem]{Lemma}
%% \newtheorem{proposition}[theorem]{Proposition}
%% \theoremstyle{definition}
%% \newtheorem{definition}[theorem]{Definition}
%% \newtheorem{example}[theorem]{Example}
%% \newtheorem{conjecture}[theorem]{Conjecture}
%% \theoremstyle{remark}
%% \newtheorem{remark}[theorem]{Remark}

\usepackage{thm-restate}
\usepackage{tikz,pgfplotstable}
\usetikzlibrary{patterns}
\pgfplotsset{compat=1.9} % set to 1.8 to get old behaviour
%\usepackage{hyperref}

\newcommand{\R}{\ensuremath{\mathbb{R}}}
\newcommand{\eps}{\varepsilon}
\newcommand{\M}{\mathcal{M}}

\newcommand{\der}{\text{\textup{d}}}
\newcommand{\diag}{\textup{diag}}

\providecommand{\x}{}
\renewcommand{\x}{\mathbf{x}}
\newcommand{\y}{\mathbf{y}}

\newcommand{\hil}{\mathcal{H}}
\newcommand{\hilp}{\mathcal{H}_p}
\newcommand{\hilo}{\mathcal{H}_o}
\newcommand{\obs}{\mathcal{O}}
\newcommand{\pobs}{\mathcal{P}}
\newcommand{\fwd}{\mathcal{F}}

\newcommand{\obsm}{\widehat{\obs}}
\newcommand{\Sigmam}{\widehat{\Sigma}}
\newcommand{\postcovm}{\widehat{\Gamma_{\textup{post}}}}
\newcommand{\uu}{\mathbf{u}}
\newcommand{\tar}{\Psi}
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}

% Definitions for second chapter
\newcommand{\data}{\mathbf{d}}
\newcommand{\param}{\mathbf{m}}
\newcommand{\normal}{\mathcal{N}}
\newcommand{\pr}{\mu_{\textup{pr}}} %Prior measure
\newcommand{\post}{\mu_{\textup{post}}^{\data, \obs}} % Posterior measure
\newcommand{\prmean}{\param_{\textup{pr}}} % Prior mean
\newcommand{\postmean}{\param_{\textup{post}}} % Posterior mean
\newcommand{\postcov}{\Gamma_{\textup{post}}} % Posterior covariance
\newcommand{\prcov}{\Gamma_{\textup{pr}}} % Prior covariance
\newcommand{\modcov}{\Gamma_{\textup{model}}} % Model covariance
\newcommand{\tmp}{\mathcal{G}}
\newcommand{\meas}{\mathbf{o}}
\newcommand{\ev}{\mathbf{e}} % eigenvector 
\newcommand{\func}{\mathbf{a}}
\newcommand{\tr}[1]{\textup{tr}\left \{#1 \right \} }
\newcommand{\ttr}[1]{\textup{tr}\ #1}
\newcommand{\rank}{\textup{rank}\ }
\newcommand{\des}{\eta} % vector of design parameters
\newcommand{\sigsqr}{\sigma^2}

%% %\usepackage{comment}% http://ctan.org/pkg/comment
%% %% %% %\excludecomment{proof}
%% %% \excludecomment{figure}
%% %% \let\endfigure\relax

%% %\overfullrule=0pt

%
\pubyear{2021}
\volume{TBA}
\issue{TBA}
% \doi{0000}
%\arxiv{}
\firstpage{1}
\lastpage{1}

\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{natbib}
\usepackage[colorlinks,citecolor=blue,urlcolor=blue,filecolor=blue,backref=page]{hyperref}
\usepackage{graphicx}

\startlocaldefs
% ** Local definitions **
\endlocaldefs

\begin{document}

%% *** Frontmatter ***

\begin{frontmatter}
\title{Measurement Clusterization in Bayesian D-optimal Designs in Infinite Dimensions}

%\title{\support{}}
\runtitle{Measurement Clusterization in D-optimal Designs}
%\author{Yair Daon}
%\numberwithin{equation}{section}

\begin{aug}
\author{\fnms{Yair} \snm{Daon}\thanksref{addr1}\ead[label=e1]{yair.daon@gmail.com}}
\runauthor{Yair Daon}

\address[addr1]{Azrieli Faculty of Medicine, Bar-Ilan University, Safed, Israel
    \printead{e1} % print email address of "e1"
}

\end{aug}

\begin{abstract}
  %\input{abstract}
  Estimation of parameters in physical processes often demands costly
  measurements, prompting the pursuit of an optimal measurement
  strategy. Finding such strategy is termed the problem of
  \emph{optimal experimental design}, abbreviated as optimal
  design. Remarkably, optimal designs can yield tightly clustered
  measurement locations, leading researchers to fundamentally revise
  the design problem just to circumvent this issue. Some authors
  introduce error correlation among error terms that are initially
  independent, while others restrict measurement locations to a finite
  set of locations. While both approaches may prevent clusterization,
  they also fundamentally alter the optimal design problem.

  In this study, we consider Bayesian D-optimal designs, i.e.~designs
  that maximize the expected Kullback-Leibler divergence between
  posterior and prior. We propose an analytically tractable model for
  D-optimal designs over Hilbert spaces. In this framework, we make
  several key contributions: \textbf{(a)} We establish that
  measurement clusterization is a generic trait of D-optimal designs
  with independent Gaussian measurement errors, and prove that
  introducing correlations among measurement error terms mitigates
  clusterization. \textbf{(b)} We characterize D-optimal designs as
  reducing uncertainty across a subset of prior covariance
  eigenvectors. Finally, \textbf{(c)} We leverage this
  characterization to argue that measurement clusterization arises as
  a consequence of the pigeonhole principle: when more measurements
  are taken than there are locations where the select eigenvectors are
  large and others are small --- clusterization occurs.
  %
  %% In summary, in this study we shed light on an often ignored issue with
  %% Bayesian D-optimal designs. We characterize such designs and show why
  %% clusterization arisesit arisestheir inherent properties, propose
  %% strategies to address clusterization, and provide insights into the
  %% fundamental attributes underlying measurement design optimization.

\end{abstract}

%% ** Keywords **
\begin{keyword}[class=MSC]
\kwd[Primary ]{62F15}
\kwd{35R30}
\kwd[; secondary ]{28C20}
\end{keyword}

\end{frontmatter}

%\input{intro}
\section{Introduction}\label{section:intro}
Measurements play a fundamental role in generating observations and
are indispensable to any inference process. Selecting the optimal set
of measurements holds particular significance when inferring
parameters of a physical process. Unlike many other fields, where
measurements are fixed, in this context, we have the freedom to choose
which measurements to take. This freedom should be taken advantage of:
measurements should be chosen to enhance accuracy, reduce costs, or
both. Whether measurements involve specifying electrode locations on
the skin in electric impedance tomography \cite{horesh2010impedance},
determining certain wavelengths in MRI \cite{horesh2008mri},
positioning sensors for detecting ground-reflected waves in the search
for oil \cite{horesh2008borehole}, or optimizing sensor locations in a
seismographic network \cite{rabinowitz1990, steinberg1995} --- the
selection of optimal measurements, referred to as the problem of
\emph{optimal design}, becomes crucial. Optimal measurements are
chosen according to specific \emph{design criteria}, with A- and
D-optimality being two of the most widely recognized and extensively
studied design criteria \cite{Chaloner1995,
  AlexanderianGloorGhattas14}.

Surprisingly, A- and D-optimal designs have been observed to yield
remarkably similar measurements in certain cases \cite{fedorov1996,
  nyberg2012, fedorov1997, Ucinski05, neitzel2019sparse}. To
illustrate this point, we consider a toy problem: inferring
the initial condition of the 1D heat equation (details in the
supplementary material). In Fig.~\ref{fig:clusterization
  illustration} D-optimal measurement locations are shown for
different numbers of measurements. Notably, for six measurements, a
D-optimal design yields two sets of measurements that are almost
indistinguishable from one another. We refer to this intriguing
phenomenon as \emph{measurement clusterization} \cite{Ucinski05}, and
we refer to a design that exhibits measurement clusterization as a
\emph{clustered design}.

%% first: blue, second: red, third: green, fourth: black,
%% fifth:magenta
\begin{figure}
    \centering
    \includegraphics[height=0.5\textwidth]{example.pdf}
    \caption{Measurement clusterization for optimal designs when
      inverting for the initial condition of the 1D heat equation (see
      supplementary material for details). Measurement locations were
      chosen according to the Bayesian D-optimality criterion of
      Theorem \ref{thm:d optimality}. Measurement locations are
      plotted over the computational domain \(\Omega = [0, 1]\)
      (x-axis), for varying numbers of measurements (y-axis). The
      colored numbers are measurement indices, plotted for visual
      clarity. Measurement clusterization already occurs for three
      measurements: the second measurement (red) is overlaid with the
      third (green). For five measurements, first (blue) and second
      (red) measurements are clustered, as well as the fourth (black)
      and the fifth (magenta).}
  \label{fig:clusterization illustration}
\end{figure}

Researchers widely agree that measurement clusterization is
undesirable \cite{fedorov1996, nyberg2012, fedorov1997, Ucinski05,
  neitzel2019sparse}, prompting the exploration of various remedies to
address this issue. One approach involves merging close measurements
\cite{fedorov1997}; however, this strategy merely overlooks the
phenomenon of measurement clusterization. An alternative solution lies
in \emph{clusterization-free design}s, where measurement locations are
deliberately chosen to be distant from one another. This can be
achieved by imposing distance constraints between measurements or by
introducing correlated errors that account for both observation error
and model misspecification \cite{Ucinski05}. For instance, in the
context of time-series analysis for pharmacokinetic experiments,
measurement clusterization can be mitigated by incorporating the
modeling of auto-correlation time within the noise terms
\cite{nyberg2012}.

In spatial problems involving choice of measurements within a domain
\(\Omega \subseteq \mathbb{R}^d, d=1,2,3\), many researchers
circumvent the problem of measurement clusterization by choosing
measurements from a coarse grid in \(\Omega\) \cite{koval2020,
  alexanderian2021, attia2022, alexanderian2014, alexanderian2016,
  alexanderian2018efficient, brunton2016}. This approach incurs a
significant computational cost as it requires solving a difficult
combinatorial optimization problem for measurement locations over a
finite set. The combinatorial optimization problem is usually relaxed
by first assigning optimal measurement weights in \(\mathbb{R}_+\) to
the potential measurement locations. Some researchers incorporate a
sparsifying \(\ell_1\) penalty term into the design criterion, which
is subsequently thresholded to achieve the desired binary design over
the coarse grid \cite{horesh2008borehole}. Others progressively relax
the \(\ell_1\) penalty to an \(\ell_0\) penalty via a continuation
method \cite{alexanderian2016, alexanderian2014}. Others cast the
problem of finding optimal measurement weights as a stochastic
optimization problem \cite{attia2022stochastic}. All of the
aforementioned methods may indeed find a binary optimal design
restricted to a given coarse grid. However, none addresses one
fundamental issue: the restriction of measurement locations to a
coarse grid in \(\Omega\) fundamentally changes the optimal design
problem and thus results in a sub-optimal design.

Avoiding measurement clusterization is a pragmatic approach:
intuitively, researchers recognize that measurement clusterization is
undesirable, even though the underlying reasons may not be fully
clear. Consequently, they strive to prevent it and devise various
methodologies to avoid measurement clusterization. Yet each and every
one of these methodologies achieves its objective by imposing
restrictions on measurement locations, thereby fundamentally altering
the optimal design problem. To the best of my knowledge, no previous
study has tried to address some seemingly simple yet fundamental
questions:
%
Why does imposing correlations between observations alleviate
measurement clusterization?
%
Is measurement clusterization a generic phenomenon?
%
And, most importantly: Why does measurement clusterization occur?
%
%Should we aim to avoid measurement clusterization?
%
%Is it possible to substitute an optimal clustered design with an
%equally optimal non-clustered design?
%Can an optimal clustered design be relpaced with an equally optimal
%non-clustered design?

\subsection{Contribution}
The primary objective of this study is to provide a comprehensive
understanding of measurement clusterization by addressing the
aforementioned questions. Our focus centers around investigating the
Bayesian D-optimality criterion, which involves maximizing the
expected Kullback-Leibler divergence between the posterior and prior
measures \cite{CoverThomas91, Chaloner1995}. We conduct an analysis of
Bayesian D-optimal designs within the context of linear inverse
problems over Hilbert spaces. We propose a novel relaxed model for
D-optimality that maintains analytical tractability and enables the
identification of D-optimal designs using Lagrange multipliers. This
analytical framework facilitates the exploration of the questions
posed at the end of the previous paragraph:

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{eigenvectors.pdf}
    \caption{D-optimal measurement locations ($m=4$ measurements) and
      weighted eigenvectors for finding the initial condition of
      the 1D heat equation. Measurement locations and weighted
      eigenvectors are plotted over the computational domain $\Omega =
      [0, 1]$ (x-axis). Measurement clusterization occurs
      approximately at $0.31$ and $0.69$. These two locations are a
      compromise between zeros of eigenvectors a D-optimal design aims
      to ignore (third and up) and staying far from zeros of the
      eigenvectors a D-optimal design aims to measure (first and
      second). Allocating $m=4$ measurements into two locations
      results in clusterization, according to the pigeonhole
      principle.}
  \label{fig:why}
\end{figure}

\begin{enumerate}

\item \label{q:generic} \textbf{Is measurement clusterization a
  generic phenomenon?}
  %\subsection{An answer for Question \ref{q:generic}: Genericity of measurement clusterization}
  %% Computer implementation of Lemma \ref{thm:char} for the inverse
  %% problem outlined in Section \ref{section:how} also generates \(\obs\)
  %% as a solution to the D-optimal design problem.  Furthermore,
  Randomized numerical simulations of our model give rise to D-optimal
  designs that exhibit clusterization more than 95\% of the time (see
  code in supplementary material). Given our model's genericity, we
  expect measurement clusterization to be a generic phenomenon.

\item \label{q:mitigate} \textbf{Why does imposing correlations
  between observations alleviate measurement clusterization?} In
  Section \ref{section:non vanishing}, we rigorously demonstrate the
  role of model error in mitigating clusterization, thereby
  corroborating earlier observations made by various researchers.

\item \label{q:why} \textbf{Why does measurement clusterization
  occur?} In Section \ref{section:why}, we present a compelling
  explanation for the optimality of clustered designs in the absence
  of model error. Our analysis reveals that, in our model, a D-optimal
  design focuses on a select set of prior eigenvectors, specifically
  those with the largest eigenvalues in the prior covariance
  spectrum. In practical scenarios, the number of locations where (a)
  the relevant prior eigenvectors are significantly large, and (b)
  other eigenvectors are close to zero, is limited. Consequently, the
  clusterization of measurements arises as a natural consequence of
  the pigeonhole principle, as there are more measurements available
  than there are locations satisfying conditions (a) and (b). See
  Fig~\ref{fig:why}.

  %% In Section \ref{section:vanishing}, we provide an insightful
  %% explanation for the optimality of clustered designs when no model
  %% error is present. We demonstrate that for our model, a D-optimal
  %% design measures only a small subset of prior eigenvectors, which are
  %% the prior eigenvectors with largest power spectrum. In real-life
  %% problems there are limited number of locations where: (a) the
  %% relevant prior eigenvectors are large, and other prior eigenvectors
  %% with large power in the prior spectrum are zero. Then measurement
  %% clusterization is a result of the pigeonhole principle, where there
  %% are more measurements than measurement locations satisfying (a) and
  %% (b).

  %% We conjecture that the prevalence of measurement clusterization
  %% arises due to the ease of discovering clustered designs.

%% \item \label{q:avoid} \textbf{Should we aim to avoid measurement
%%   clusterization?} Based on the analysis conducted in this study, we
%%   did not find any compelling reason to explicitly avoid optimal
%%   clustered designs.

%% \item \label{q:replace} \textbf{Is it possible to substitute an
%%   optimal clustered design with an equally optimal non-clustered
%%   design?} In Section \ref{section:vanishing}, we answer this question
%%   in the affirmative, although we show that numerical experiments
%%   conducted using our model indicate a strong preference for clustered
%%   designs.

\end{enumerate}

The cornerstone of our investigation into measurement clusterization
is Theorem \ref{thm:char} proven in Section
\ref{section:vanishing}. The key insight of Theorem \ref{thm:char}
lies in its fifth component, which highlights that a D-optimal design
aims to uniformly reduce posterior uncertainties for posterior
covariance eigenvectors in observation space. A similar conclusion was
reached by Koval et al.~\cite{koval2020}, who showed that A-optimal
designs are best constructed in the space of observations.

Before we state Theorem \ref{thm:char} we give some definitions: Let
\(\hilp, \hilo\) Hilbert spaces, \(\fwd:\hilp \to \hilo\) a linear
compact operator. Let \(\obs: \hilo \to \mathbb{R}^m\) a linear
measurement operator, where \(m \in \mathbb{N}\) is the number of
measurements taken. Let \(\sigma^2 \in \mathbb{R}_{+}\) observation
noise variance, \(\data = \obs \fwd \param + \eps\), where \(\eps \in
\mathbb{R}^m\) is iid \(\mathcal{N}(0, \sigma^2)\) noise. Let \(\pr
\sim \mathcal{N}(\prmean, \prcov)\) prior Gaussian measure on
\(\hilp\), where \(\prcov\) is the prior covariance operator and let
\(\post\) the posterior (Gaussian) measure.

\begin{restatable}{theorem}{char}\label{thm:char}
  Let:
  \begin{itemize}
    \item The D-optimality design criterion
    \cite{AlexanderianGloorGhattas14}:
    \begin{align*}
      \begin{split}
        \tar(\obs) %:&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
        % 
        % 
        % 
        &= \frac12 \log \det ( I + \sigma^{-2} \prcov^{1/2} \fwd ^*
        \obs^* \obs \fwd \prcov^{1/2}), 
      \end{split}
    \end{align*}
  \item \(\obs\) a D-optimal design operator
    \begin{equation*}
      \obs = \argmax_{\|\meas_j\| = 1, j=1,\dots,m}\tar(\obs),
    \end{equation*}
  \item \(\{\lambda_i\}_{i=1}^\infty\) eigenvalues of
    \(\fwd\prcov\fwd^*\) in decreasing order of magnitude.
    %% \item \(\{\ev_i\}_{i=1}^\infty\) their corresponding eigenvectors.
  \item \(\{\eta_i\}_{i=1}^\infty\) eigenvalues of \(\obs^*\obs\).

  \end{itemize}

  Then:
  \begin{enumerate}
  \item  \(\tr{\obs^*\obs} = m\).
  \item \(\obs^*\obs\) and \(\fwd\prcov\fwd^*\) are simultaneously
    diagonalizable.
  \item \(k := \rank \obs^*\obs \leq m\) and
    \begin{equation*}
      \tar(\obs) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i). %= \frac12 \sum_{i=1}^{m} \log (1 + \sigma^{-2}\lambda_i\eta_i).
    \end{equation*}
  %% \item 
  %%   \begin{equation*}
  %%     k = \argmax \left \{ k:\lambda_k^{-1} < \sigma^{-2}\frac{m}{k} + \frac{1}{k} \sum_{j=1}^{k}
  %%     \lambda_j^{-1} \right \}.
  %%   \end{equation*}
  \item
    \begin{equation*}
        \eta_i = \begin{cases}
          \frac{m}{k} - \sigma^2 \lambda_i^{-1} + \sigma^2 \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1} & 1 \leq i \leq k \\
          0 & i > k 
        \end{cases}.
    \end{equation*}
  \item The covariance of the pushforwad \(\fwd_{*} \post\) is \(\left
    ( (\fwd \prcov \fwd^*)^{-1} + \sigma^{-2} \obs^*\obs \right
    )^{-1}\) and its eigenvalues are
    \begin{equation*}
      \theta_i =
      \begin{cases}
        \left(\frac{\sum_{j=1}^k \lambda_j^{-1} + \sigma^{-2}m}{k} \right )^{-1} & i \leq k \\
        \lambda_i &  i > k 
      \end{cases}
    \end{equation*}
  \end{enumerate}
%\end{theorem}
\end{restatable}

In the process of proving Theorem \ref{thm:char} we prove and
generalize several lemmas. Among those, is Lemma \ref{lemma:free},
which is (to the the best of my knowledge) a novel lemma in linear
algebra: We decompose a symmetric positive definite matrix \(M \in
\mathbb{R}^{k \times k}\) with \(\ttr M = m \in \mathbb{N}\) as \(M =
AA^t\), where \(A\) has unit norm columns.

%% Finally, in Lemma \ref{lemma:lax} we generalize a lemma for
%% calculating \(\frac{\der}{\der t} \log \det (I + X(t))\), where
%% \(X(t)\)is an operator valued function \cite{Lax07}.

\subsection{Limitations}\label{subsec:limitations}
The main limitation of this study is that our generic model does not
correspond to any specific real-life problem. It is generic enough to
be analytically tractable, but one may argue our model is too far
removed from any real application. To these claims I would answer that
scientists have a long history of studying models that are bare-bones
simplifications of real systems, e.g. the Ising model
\cite{cipra1987}, the Lorenz system \cite{brin}, the Lotka-Volterra
equations \cite{logan2006}, the Carnot engine \cite{kardar2007}, and
many others.

%\input{prelims}
\section{Preliminaries and Notation}\label{section:prelim}

In this section we present the setup that will be used throughout this
article. The theoretical foundations for inverse problems over
function spaces can be found elsewhere \cite{Stuart10} and will not be
reviewed here.

\subsection{Bayesian Linear Inverse Problems}\label{subsec:abstract OED}
Let \(\hilp\) and \(\hilo\) separable Hilbert spaces (the subscripts p
and o are for "parameter" and "observation", respectively), and let
\(\fwd: \hilp \to \hilo\) the \emph{forward operator}. The forward
operator \(\fwd\) is assumed linear and strongly smoothing (the heat
operator of the 1D heat equation introduced in Section
\ref{section:intro} is a prime example). Take a Gaussian prior
\(\param \sim \pr = \normal(\prmean ,\prcov)\) with some appropriate
covariance operator \(\prcov\) on \(\hilp\) \cite{Stuart10}. Note that
\(\fwd \prcov \fwd^*\) is the prior covariance in \(\hilo\)
\cite{Stuart10}, and as such, \(\fwd \prcov \fwd^*\) is assumed
invertible --- an assumption which will be used below (if \(\fwd\) has
a nontrivial kernel we utilize Occam's Razor and ignore said kernel
altogether). Measurements are taken via a \emph{measurement operator}
\(\obs\). It is common for the measurement and forward operators to be
merged \(\tmp := \obs \fwd\) \cite{AlexanderianGloorGhattas14}, but
the analysis carried out in the following sections requires that
\(\fwd\) and \(\obs\) are explicitly separated as in
\cite{attia2022stochastic, cvetkovic2023choosing}.  \(\obs \in (
\hilo^* )^m\), where \(m\) is the number of measurements
taken. Entries \(\meas_j, j=1,\dots,m\) of the observation operator
\(\obs\) are called \emph{measurements}:

\begin{equation*}%\label{eq:O}
  \obs u = (\meas_1(u), \dots, \meas_m(u) )^t \in \R^m,\ u \in \hilo.
\end{equation*}

Data is acquired via noisy observations, and we consider two types of
error terms: Spatially correlated model error \(\eps' \sim
\normal(0,\modcov)\) with \(\modcov\) a covariance operator. Observation
error is denoted \(\eps \sim \normal(0, \sigma^2 I_m)\), with \(I_m \in
\mathbb{R}^{m \times m}\) the identity. Both error terms and the prior
are assumed independent of each other. Thus, data is acquired via
\begin{align}\label{eq:inverse problem}
  \data := \obs (\fwd \param + \eps') + \eps = \obs \fwd \param + \obs \eps' + \eps.
\end{align}

It is easy to verify that \(\obs \eps' + \eps \in \R^m\) is a centered
Gaussian random vector with covariance matrix

\begin{align}\label{eq:Sigma}
  \begin{split}
    \Sigma(\obs) :&= \mathbb{E}[ (\obs \eps' + \eps) (\obs \eps' +
      \eps)^t ]
    % 
    % 
    = \obs \modcov \obs^* + \sigma^2I_m , 
  \end{split}
\end{align}
where
\begin{align}\label{eq:modcov explained}
  \begin{split}
    [\obs \modcov \obs^*]_{ij} & = e_i^t \obs \modcov \obs^* e_j 
    %
    %
    %
    = \meas_i (\modcov \meas_j).% \text{ (by \eqref{eq:obs*})}.
  \end{split}
\end{align}
Taking \(\modcov = 0\) is a common practice
\cite{tarantola2005,Kaipio2005,Vogel02} and then \(\Sigma =
\sigma^2I_m\) is a scalar matrix which does not depend on \(\obs\).

Finally, it is useful to record that the posterior measure \(\post\)
is Gaussian in this setting and its covariance operator does not
depend on data \(\data\) \cite{Stuart10}:
\begin{align}\label{eq:postcov}
  \postcov = (\prcov^{-1} + \fwd^* \obs^* \Sigma^{-1} \obs \fwd
  )^{-1}.
\end{align}

\subsection{Bayesian D-Optimal Designs in Infinite Dimensions}\label{subsec:D optimal design} 
A Bayesian D-optimal design maximizes expected Kullback-Leibler (KL)
divergence between posterior \(\post\) and prior measures \(\pr\). It is
useful to first recall the definition of KL divergence \cite{CoverThomas91}:
\begin{equation*}
  D_{KL}(\post||\pr) = \int \log \frac{\der \post}{\der \pr}(\param) \der \post(\param).
\end{equation*}

The study of D-optimal designs for Bayesian linear inverse problems in
infinite dimensions was pioneered by \cite{AlexanderianGloorGhattas14,
  alexanderian2018efficient}. The main result we will make use of is
summarized (in our notation) below:

\begin{theorem}[Alexanderian, Gloor, Ghattas \cite{AlexanderianGloorGhattas14}]\label{thm:d optimality}
  Let \(\pr = \normal(\prmean,\prcov)\) be a Gaussian prior on \(\hilp\)
  and let \(\post = \normal(\postmean,\postcov)\) the posterior measure
  on \(\hilp\) for the Bayesian linear inverse problem \(\data = \obs
  \fwd\param + \obs \eps' + \eps\) discussed above. Then
  \begin{align}\label{eq:objective}
    \begin{split}
      \tar( \obs) :&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
      % 
      % 
      % 
      &= \frac12 \log \det 
      ( I + \prcov^{1/2}  \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}).
    \end{split}
  \end{align}
\end{theorem}

Note that in \cite{AlexanderianGloorGhattas14,
  alexanderian2018efficient}, results are stated for \(\Sigma=I\)
(implied by \(\modcov = 0,\sigma^2=1\)), but these results also hold
for more general covariance matrices
\cite[p. 681]{AlexanderianGloorGhattas14}.

%% It is important to note that since \(\obs\) is finite-rank,
%% \(\prcov^{1/2} \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}\) is
%% trace-class.
\begin{definition}\label{def:d optimality}
  We say \(\obs^{\star}\) is \emph{D-optimal} if \(\obs^{\star} =
  \argmax_{\obs} \tar(\obs)\), where entries of \(\obs \in (\hilo^*)^m\)
  are constrained to some allowed set of measurements in \(\hilo^*\).
\end{definition}

Intuition for Theorem \ref{thm:d optimality} can be gained by
considering a Bayesian linear model in finite dimensions, with
Gaussian prior and Gaussian noise. Then, a D-optimal design minimizes
the determinant of the posterior covariance matrix, and this turns out
to be a regularized version of the frequentist D-optimality criterion
\cite{Chaloner1995}. Theorem \ref{thm:d optimality} and Definition
\ref{def:d optimality} carry a similar intuition:
\begin{align*}
  \begin{split}
    \tar(\obs) &= \frac12 \log \det ( I + \prcov^{1/2}  \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}) \text{ by \eqref{eq:objective}}\\
    &= \frac12 \log \det \Big( \prcov ( \prcov^{-1} + \fwd ^* \obs^* \Sigma^{-1} \obs \fwd) \Big )\\
    &= \frac12 \log \det \prcov \postcov^{-1} \text{ by \eqref{eq:postcov}}.
    %%  &= \frac12 \log \det \prcov -\frac12 \log \det \postcov.
  \end{split}
\end{align*}
We think of \(\prcov\) as constant, so a D-optimal design minimizes a
quantity analogous to the posterior covariance determinant, similarly
to the finite-dimensional case.

%% \subsection{Notation Summary}\label{subsec:notation}
%%   Our notation is summarized below. We let:
%%   \begin{itemize}
%%   \item \(\hilp, \hilo\) Hilbert spaces.
%%   \item \(\fwd:\hilp \to \hilo\) a linear compact operator.
%%   \item \(\pr \sim \mathcal{N}(0, \prcov)\) prior Gaussian measure on
%%     \(\hilp$, with prior covariance operator \(\prcov:\hilp \to \hilp$.
%%   \item \(\obs: \hilo \to \mathbb{R}^m\) measurement operator, where \(m
%%     \in \mathbb{N}\) is the number of measurements taken.
%%   \item \(\sigma^2 \in \mathbb{R}_{+}\) observation noise variance.
%%   \item \(\modcov\) model error covariance operator.
%%     %% \(\data = \obs \fwd \param + \eps$, where \(\eps \in
%%     %% \mathbb{R}^m\) isiid \(\mathcal{N}(0, \sigma^2)\) noise.
%%   \item \(\Sigma(\obs) = \obs \modcov \obs^* + \sigma^2I$. 
%%   \item \(\post\) the posterior measure, with covariance \(\postcov$.
%%   \item A D-optimality design criterion
%%     \cite{AlexanderianGloorGhattas14}:
%%     \begin{align*}
%%       \begin{split}
%%         \tar(\obs) :&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
%%         % 
%%         % 
%%         % 
%%         &= \frac12 \log \det ( I + \prcov^{1/2} \fwd ^* \obs^* \Sigma(\obs)^{-1} \obs
%%         \fwd \prcov^{1/2}).
%%       \end{split}
%%     \end{align*}
%%   %% \item \(\{\lambda_i\}_{i=1}^\infty\) eigenvalues of \(\fwd\prcov\fwd^*$
%%   %%   in decreasing order of magnitude.
%%   %% %% \item \(\{\ev_i\}_{i=1}^\infty\) their corresponding eigenvectors.
%%   %% \item \(\{\eta_i\}_{i=1}^\infty\) eigenvalues of \(\obs^*\obs$.
%%   \end{itemize}

%% \subsection{Sequential vs Simultaneous Optimization}\label{subsec:seq vs sim}
%% From defintion \ref{def:d optimality} we wish to characterize solution(s) of the
%% following optimization problem for \(\tar$. %%: (\hilo^*)^m \to \R$:
%% \begin{align}\label{eq:optimization}
%%   \obs^{\star} := \argmax_{\obs} \tar( \obs ) 
%%   = \argmax_{\obs} \frac12 \log \det 
%%   (I + \prcov^{1/2} \fwd^*\obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}),
%% \end{align}
%% where \(\obs\) is constrained to some allowed set of observations. We
%% call this problem ``simultaneous optimization'', since all
%% observations are decided on simulatneously.

%% For computational reasons, one may prefer to find the best
%% observations in a sequential manner. Denote
%% \begin{equation}\label{eq:def obs_k}
%%   \obs_k := (\meas_1,\dots, \meas_k)^t,  k\leq m.
%% \end{equation}
%% Sequential optimal design proceeds as follows. Find \(\meas_1\) by
%% maximizing \(\tar(\obs_1)$. Then, keeping \(\meas_1\) fixed --- find
%% \(\meas_2\) as the maximizer of \(\tar(\obs_2)$. Then, find \(\meas_3\) by
%% keeping \(\meas_1,\meas_2\) fixed and taking \(\meas_3\) as the maximizer
%% of \(\tar(\obs_3)$. Continue this way until \(\obs_m = \obs\) is
%% found, where \(m\) is the number of available observations. %% It is
%% %% important to notice that this scheme does not require actually
%% %% observing data --- in \eqref{eq:objective} data is averaged out.

%% The analysis in this paper is conducted for the general simultaneous
%% optimization case. The sequential optimization case is dealt with in
%% section \ref{subsec:clusterization sequential}. It is important to
%% note, however that all conclusions we arrive at for the simultaneous
%% case easily specialize to the sequential case by considering the
%% posterior as the next sequential step's prior.

%\input{constrained}
\section{The Constrained Optimization Problem of D-Optimal Design}\label{section:D and grad}
We seek a formulation of the D-optimal design problem via Lagrange
multipliers. We first find the gradient of $\tar$, then we suggest
unit-norm constraints on $\obs$ and find their gradients. Results of
this section are summarized in Theorem \ref{thm:constrained}. First,
recall that:
\begin{definition}\label{def:var}
  Let $F$ a real valued function of $\obs$. The first variation of $F$
  at $\obs$ in the direction $V$ is:
  \begin{equation*}
    \delta F(\obs) V := \frac{\der}{\der \tau}\Big |_{\tau=0}  F( \obs + \tau V).
  \end{equation*}

  Moreover, if
  \begin{equation*}
    \delta F(\obs) V = \tr{\nabla F(\obs) V},
  \end{equation*}
  then we call $\nabla F(\obs)$ the gradient of $F$ at $\obs$. 
\end{definition}

%% Gradients are best thought of as row vectors. This will prove
%% important in section \ref{subsec:necessary}.

%% \subsection{The gradient of $\tar$}\label{section:objective}

\begin{proposition}\label{prop:tar grad}
  The gradient of the D-optimality objective $\tar$ is
  \begin{equation*}
    %% \delta \tar(\obs) V = \tr{V ( I - \modcov \obs^* \Sigma^{-1}\obs )
    %%   \fwd \postcov \fwd^* \obs^* \Sigma^{-1}}.
    \nabla \tar(\obs) = ( I - \modcov \obs^* \Sigma^{-1}\obs ) \fwd
    \postcov \fwd^* \obs^* \Sigma^{-1}
  \end{equation*}
\end{proposition}

\begin{proof}  
  From the definition of $\Sigma(\obs)$ \eqref{eq:Sigma}:

  \begin{align}\label{eq:der sig}
    \begin{split}
      \frac{\der}{\der \tau} \Big |_{\tau=0} \Sigma( \obs + \tau V )
      &= \frac{\der}{\der \tau} \Big |_{\tau=0} 
      (\obs + \tau V ) \modcov (\obs + \tau V )^*  + \sigma^2I\\
      % 
      % 
      % 
      &= V \modcov \obs^* + \obs \modcov V^*.
    \end{split}
  \end{align}

  Then, using \eqref{eq:der sig}: 
  \begin{align*}
    0 &= \frac{\der}{\der \tau} \Big |_{\tau=0} I \\
    % 
    % 
    % 
    &= \frac{\der}{\der \tau} \Big |_{\tau=0}
    \left (\Sigma(\obs+\tau V)^{-1} \Sigma(\obs+\tau V) \right ) \\
    % 
    % 
    % 
    &= \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0} \Sigma+
    \Sigma^{-1} \frac{\der \Sigma(\obs+\tau V)}{\der \tau} \Big |_{\tau=0}\\  
    %
    %
    %
    &= \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0} \Sigma+
    \Sigma^{-1} (V\modcov \obs^* + \obs \modcov V^*). 
    %\text{, by \eqref{eq:der sig}. }
  \end{align*}

  Thus:
  \begin{align}\label{eq:der sig inv}
    \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0}  
      &= -\Sigma^{-1} (V \modcov \obs^* + \obs \modcov V^*) \Sigma^{-1}.
    \end{align}
  %

  Let
  \begin{equation*}
    T(\obs) = \obs^* \Sigma^{-1}(\obs)\obs.
  \end{equation*}

  By Leibniz (product) rule and \eqref{eq:der sig inv}:
  % 
  \begin{align}\label{eq:T}
    \begin{split}
    \delta T(\obs) V 
    &= \frac{\der T(\obs + \tau V)}{\der \tau} \Big |_{\tau=0} \\
    %
    %
    %
    &= V^* \Sigma^{-1} \obs 
    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
    &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs
    + \obs^* \Sigma^{-1} V.
    \end{split}
  \end{align}

  We now record a lemma which generalizes a known result in linear
  algebra \cite[Chapter 9, Theorem 4]{Lax07}. Its proof is delegated
  to the supplementary material.
  \begin{lemma}\label{lemma:lax}
    Let $Y(t)$ be a differentiable operator-valued function. Assume 
    $I+Y(t)$ is invertible, $Y(t)$ self-adjoint and trace-class. Then
    \begin{equation*}
      \frac{\der \log \det (I+Y(t))}{\der t} = \tr{(I+Y(t))^{-1} \dot{Y}(t)}.
    \end{equation*}
  \end{lemma}

  We employ Lemma \ref{lemma:lax} and calculate the first variation of
  $\tar$:

  \begin{align*}
    %\begin{split}
      \delta \tar(\obs) V 
      :&= \frac{\der}{\der\tau} \Big |_{\tau=0} \tar(\obs + \tau V) \text{ (Definition \ref{def:var})}\\
      % 
      % 
      % 
      &= \frac12 \frac{\der}{\der \tau} \Big |_{\tau=0} \log \det 
      (I + \prcov^{1/2} \fwd^* T(\obs+\tau V)\fwd \prcov^{1/2} ) \text{ (Theorem \ref{thm:d optimality})} \\
      % 
      % 
      % 
      &= \frac12 \tr{( I + \prcov^{1/2} \fwd^* \obs^* \Sigma^{-1}
        \obs\fwd \prcov^{1/2} )^{-1}
        \frac{\der}{\der \tau} \Big |_{\tau=0}
        \prcov^{1/2} \fwd^* T(\obs+\tau V) \fwd \prcov^{1/2}}\ \text{ (Lemma \ref{lemma:lax})} \\
      % 
      % 
      % 
      &= \frac12 \ttr\Big \{ \postcov \fwd^* (V^* \Sigma^{-1} \obs 
      - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
      &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs 
      + \obs^* \Sigma^{-1} V ) \fwd \Big \}  \text{ (by \eqref{eq:T})} \\
      %
      %
      %
      &= \tr{\postcov \fwd^* ( \obs^* \Sigma^{-1} V -
      \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs ) \fwd} \\
      %
      %
      % 
      &= \tr{\postcov \fwd^* \obs^* \Sigma^{-1} V 
      ( I - \modcov \obs^* \Sigma^{-1}\obs ) \fwd} \\
      % 
      %
      %
      &= \tr{V ( I - \modcov \obs^* \Sigma^{-1}\obs )
      \fwd \postcov \fwd^* \obs^* \Sigma^{-1}}.
    %\end{split}
  \end{align*} 
  Recalling Definition \ref{def:var} concludes the proof.
\end{proof}

In a real-life optimal design problem we cannot choose any measurement
operator $\obs \in (\hilo^*)^m$. In order to facilitate analysis, we
seek reasonable constraints on $\obs$ for which finding a D-optimal
design is analytically tractable. The following proposition will guide
us in finding such constraints.

\begin{proposition}\label{prop:bigger better}
  Let $\obs = (\meas_1,\dots,\meas_m)^t$, $j \in \{1,\dots,m\}$,
  $\sigma^2 > 0$ and $|\zeta| > 1$. Then $\tar(\obs)$ increases if we
  use $\zeta \meas_j$ in $\obs$ instead of $\meas_j$.
\end{proposition}

\begin{proof} 
  Fix $j=1,\dots,m$ and take $V:= e_j e_j^t \obs$. For $u
  \in \hilo$:
  \begin{equation*}
    Vu = e_je_j^t (\meas_1(u),\dots,\meas_m(u) )^t = e_j \meas_j(u)
    = (0,\dots,0,\meas_j(u),0,\dots,0)^t.
  \end{equation*}
  %% This way, $V$ has the same $j$th entry as $\obs$ while the rest
  %% are set to zero.
  We now calculate the variation of $\tar$ at $\obs$ in the direction
  of $V$. Denote $\tmp: = \fwd \postcov \fwd^*$. From Proposition
  \ref{prop:tar grad}:
  \begin{align*}
     \delta \tar(\obs) V 
    &= \tr{V ( I - \modcov \obs^*\Sigma^{-1}\obs) \tmp \obs^* \Sigma^{-1}} \\
    % 
    %
    %
    &= \tr{e_je_j^t \obs ( I - \modcov \obs^*\Sigma^{-1}\obs) \tmp \obs^* \Sigma^{-1}} \\
    %
    % 
    %
    &= e_j^t \obs ( I - \modcov \obs^*\Sigma^{-1}\obs) \tmp \obs^* \Sigma^{-1}e_j \\
    %
    % 
    %
    &= e_j^t ( I - \obs \modcov \obs^*\Sigma^{-1})\obs \tmp \obs^* \Sigma^{-1}e_j \\  
    % 
    %
    %
    &=  e_j^t(\Sigma-\obs \modcov \obs^*) \Sigma^{-1}\obs \tmp \obs^* \Sigma^{-1}e_j \\
    %
    %
    %
    &=\sigma^2 e_j^t \Sigma^{-1}\obs \tmp \obs^* \Sigma^{-1}e_j
    \text{ by \eqref{eq:Sigma} }\\
    %
    % 
    %
    &=\sigma^2 e_j^t \Sigma^{-1}\obs \fwd \postcov \fwd^* \obs^* \Sigma^{-1}e_j.
  \end{align*} 
  Since $\postcov$ is positive definite, we conclude that $\delta
  \tar(\obs) V > 0$. This means that increasing the magnitude of the
  $j^{\text{th}}$ measurement functional increases $\tar(\obs)$.
\end{proof}

Proposition \ref{prop:bigger better} implies that it is a good idea to
bound the norm of measurements. If, for example, we can take
measurements in $span\{\meas\}$ for some $\meas \neq 0$, then a
maximum for the D-optimality criterion does not exist, as it is not
bounded. In contrast, in any real-life problem where sensors are
concerned, the norm of measurements recorded by sensors is always one
(of course, point evaluations are not in any Hilbert space of
functions we wish to consider):

\begin{equation}
  \| \delta_{\x} \| = \sup_{0 \neq u \in C(\Omega)}
  \frac{
    |\int_{\Omega}u(\y) \delta_{\x}(\y) \der \y|
  }{
    \sup|u|
  } = \sup_{0 \neq u \in C(\Omega)} \frac{|u(\x)|}{\sup|u|} = 1,
  \forall \x \in \Omega.
\end{equation}

Thus, it is reasonable to consider measurements with unit $\hilo^*$
norm. We can write the unit norm constraints as a series of $m$
equality constraints (one for each measurement) on $\obs$. We define
them and find their gradients in Proposition \ref{prop:constraints
  grad} below:

\begin{proposition}\label{prop:constraints grad}
  Let
  \begin{align*}
    \phi_j(\obs) :=\frac12 \| \obs^* e_j\|_{\hilp}^2 - \frac12 = 0,\ j=1,\dots,m.
  \end{align*}
  Then
  \begin{equation*}
    %% \delta \phi_j(\obs)V = \tr{V \obs^* e_je_j^t}.
    \nabla \phi_j(\obs) = \obs^* e_je_j^t.
  \end{equation*}
\end{proposition}

\begin{proof}
  \begin{align*}
    \delta \phi_j(\obs)V  
    &= \frac12\lim_{\tau \to 0}\tau^{-1}
    ( \|(\obs + \tau V)^*e_j \|_{\hilp}^2 - \|\obs ^*e_j \|_{\hilp}^2  ) \\
    %
    %
    %
    &= \frac12\lim_{\tau \to 0}\tau^{-1}
    ( \langle (\obs + \tau V)^*e_j, (\obs + \tau V)^*e_j \rangle_{\hilp} - 
    \langle \obs^*e_j, \obs^*e_j \rangle_{\hilp} ) \\
    % 
    % 
    %
    &= \frac12\lim_{\tau \to 0}\tau^{-1}
    (2\tau \langle \obs^*e_j,V^*e_j \rangle_{\hilp} 
    +\tau^2 \langle V^*e_j, V^*e_j \rangle_{\hilp} ) \\
    %
    %
    % 
    &= \langle \obs^*e_j,V^*e_j \rangle_{\hilp} \\
    %
    %
    % 
    &= \langle V \obs^*e_j,e_j \rangle_{\R^m} \\
    %
    %
    %
    &= e_j^t V \obs^* e_j \\
    % 
    %
    %
    &= \tr{V \obs^* e_je_j^t}.
  \end{align*}
\end{proof}

%% The same arguments justifying \eqref{eq:tar grad} hold here, and thus:

%% \begin{align}\label{eq:grad constraints}
%% \nabla \phi_j(\obs) = \obs^* e_j e_j^t = \meas_j e_j^t , j=1,\dots,m,
%% \end{align}
%% where $\nabla \phi_j(\obs) \in \hilo^m$. As noted at the end of
%% Section \ref{section:objective},

%% The gradients $\nabla \tar(\obs)$ and $\nabla \phi_j(\obs)$ are best
%% thought of as row vectors.

We find necessary first-order conditions for D-optimality via Lagrange
multipliers:

\begin{align}
  &\nabla \tar(\obs) = \sum_{j=1}^m \xi_j \nabla \phi_j (\obs)
  \label{eq:Lagrange mult1} \\
    &\phi_j(\obs) = 0, j = 1,\dots,m. \label{eq:Lagrange mult2}
\end{align}

We now substitute the gradients calculated in Propositions
\ref{prop:tar grad} and \ref{prop:constraints grad} into
\eqref{eq:Lagrange mult1}:
\begin{equation}\label{eq:constrained}
  (I - \modcov \obs^* \Sigma^{-1} \obs) \fwd \postcov \fwd^* \obs^*\Sigma^{-1}
  = \sum_{j=1}^m \xi_j \obs^* e_je_j^t = (\xi_1 \meas_1,\dots,\xi_m \meas_m).
\end{equation} 
Letting $\Xi := \diag(\xi_j)$, we can write \eqref{eq:constrained} and
\eqref{eq:Lagrange mult2} more compactly as:

\begin{theorem}[Necessary conditions for D-Optimality]\label{thm:constrained}
  Let:
  \begin{equation*}
    \obs = \argmax_{\|\meas_j\| = 1, j=1,\dots,m}\tar(\obs).
  \end{equation*}

  Then:
  \begin{equation*}
    ( I - \modcov \obs^* \Sigma^{-1} \obs) \fwd \postcov \fwd^* \obs^*  \Sigma^{-1}
    = \obs^* \Xi, 
  \end{equation*}
  where $\Xi \in \mathbb{R}^{m \times m}$ is diagonal.
\end{theorem}

%\input{nonvanishing}
%% Imposing correlations between observations alleviate measurement clusterization
\section{Answer to Question \ref{q:mitigate}: Model error mitigates clusterization}\label{section:non vanishing}
We now show that if $\modcov \neq 0$ clusterization will not occur. It
is known that including a model error term mitigates the
clusterization phenomenon \cite{Ucinski05}, and here we prove this
rigorously. Let $\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm :=
(\meas_1,\dots,\meas_{m-1})^t$. Denote $\Sigmam := \Sigma (\obsm)$ and
$\postcovm$ the posterior covariance that arises when $\obsm$ is
utilized as a measurement operator.

\begin{proposition}[Increase due to a measurement]\label{prop:design increase}
  Let $\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm :=
  (\meas_1,\dots,\meas_{m-1})^t$. Then
  \begin{equation}\label{eq:conclusion}
    \tar( \obs ) - \tar (\obsm ) =
    \frac12 \log \left ( 1 + \frac{
      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m,
      (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m \rangle
    }{
      \sigma^2 + \meas_m \modcov \meas_m - \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m 
    }       
    \right ).
  \end{equation}
\end{proposition}

\begin{proof}
  Let
  \begin{align*}
    \Sigma( \obs ) &= 
    \begin{bmatrix}
      \Sigma (\obsm )           & \obsm \modcov \meas_m \\
      \meas_m \modcov \obsm^*   & \sigma^2 + \meas_m \modcov \meas_m
    \end{bmatrix}
    : =
    \begin{bmatrix}
      \Sigmam   & w \\
      w^t       & c
    \end{bmatrix}\\
    %
    %
  \end{align*}

  The Schur complement implies:
  \begin{align}\label{eq:schur}
    \begin{split}
          \Sigma^{-1} &=
          \begin{bmatrix}
            \Sigmam^{-1} + \Sigmam^{-1} w ( c - w^t \Sigmam^{-1} w)^{-1} w^t \Sigmam^{-1} & - \Sigmam^{-1} w ( c - w^t \Sigmam^{-1} w)^{-1} \\
            -( c - w^t \Sigmam^{-1} w)^{-1} w^t \Sigmam^{-1}                            &  ( c - w^t \Sigmam^{-1} w)^{-1}
          \end{bmatrix} \\
          &=
          \begin{bmatrix}
            \Sigmam^{-1} & 0 \\
            0           & 0 
          \end{bmatrix}
          + (c -w^t \Sigmam^{-1} w )^{-1}
          \begin{bmatrix}
            \Sigmam^{-1} w \\
            -1
          \end{bmatrix}
          \begin{bmatrix}
            w^t \Sigmam^{-1} & -1 
          \end{bmatrix},
    \end{split}
  \end{align}
  %
  and denote:
  %
  \begin{align}\label{eq:M def}
    \M (\obs ):&= \prcov^{\frac12}\fwd^* \obs^* \Sigma^{-1} \obs \fwd
    \prcov^{\frac12}.
  \end{align}

  From \eqref{eq:schur} and \eqref{eq:M def}:
  \begin{align*}
    \M(\obs) &= \prcov^{1/2} \fwd^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2} \\
    %
    %
    %
    &= \prcov^{1/2} \fwd^* \obs^* \left \{
    \begin{bmatrix}
      \Sigmam^{-1} & 0 \\
      0           & 0 
    \end{bmatrix}
    + (c -w^t \Sigmam^{-1} w )^{-1}
    \begin{bmatrix}
      \Sigmam^{-1} w \\
      -1
    \end{bmatrix}
    \begin{bmatrix}
      w^t \Sigmam^{-1} & -1 
    \end{bmatrix} 
    \right \} \obs \fwd \prcov^{1/2} \\
    %
    %
    %
    &= \M (\obsm) + (c -w^t \Sigmam^{-1} w )^{-1}
    \prcov^{1/2} \fwd^* \obs^*
    \begin{bmatrix}
      \Sigmam^{-1} w \\
      -1
    \end{bmatrix}
    \begin{bmatrix}
      w^t \Sigmam^{-1} & -1 
    \end{bmatrix} 
    \obs \fwd \prcov^{1/2}
  \end{align*}
  %
  Now, denote:
  %
  \begin{align}\label{eq:u}
    \begin{split}
      u :&= (c -w^t \Sigmam^{-1} w )^{-1/2}
      \prcov^{1/2} \fwd^* \obs^* 
      \begin{bmatrix}
        \Sigmam^{-1} w \\
        -1 
      \end{bmatrix} \\
      %
      %
      %
      & = (c -w^t \Sigmam^{-1} w )^{-1/2} ( \prcov^{1/2}\fwd^* \obsm^* \Sigmam^{-1} \obsm  \modcov \meas_m - \prcov^{1/2} \fwd^* \meas_m )\\
      %
      %
      %
      u^* :&=  (c -w^t \Sigmam^{-1} w )^{-1/2} (\meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \fwd \prcov^{1/2} - \meas_m \fwd \prcov^{1/2} ),
    \end{split}
  \end{align}
  %
  so that
  %
  \begin{equation}\label{eq:M plus I}
    I + \M( \obs ) = I + \M (\obsm ) + uu^*.
  \end{equation}
  %
  Note that
  \begin{equation}\label{eq:M postcov}
    \prcov^{1/2} \left (I + \M( \obsm ) \right )^{-1} \prcov^{1/2} = \postcovm.
  \end{equation}
  From a generalization of the matrix determinant lemma to Hilbert
  spaces ($\det(A + uv^*) = (1 + \langle A^{-1} u,u \rangle) \det
  A$, statement and proof in the supplementary material):
  %
  \begin{align}\label{eq:diffs}
    \begin{split}
      \tar( \obs ) - \tar( \obsm )
      %
      %
      %
      &= \frac12 \log \left (\det \big ( I + \M ( \obs ) \big ) / \det \big ( I + \M (\obsm) \big ) \right )\\
      %
      %
      %
      &= \frac12  \log \left (\det \left ( I + \M(\obsm) + uu^* \right ) / \det \big ( I + \M (\obsm) \big )\right ) \\
      %
      %
      %
      &= \frac12 \log \left ( 1 + \left \langle \left ( I+\M(\obsm) \right )^{-1} u, u  \right \rangle \right ).
    \end{split}
  \end{align}
  From \eqref{eq:u} and \eqref{eq:M postcov}:
  \begin{align}\label{eq:final}
    \begin{split}
      &\left \langle \left (I+\M (\obsm)\right )^{-1}u, u \right \rangle\\
      &= \frac{
        \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m,
        (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m \rangle
      }{
        c- w^t \Sigmam^{-1} w
      }\\
      %
      %
      %
      &= 
      \frac{
      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m,
      (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m \rangle
      }{
        \sigma^2 + \meas_m \modcov \meas_m - \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m 
      }
    \end{split}
  \end{align}
  and the conclusion follows by substituting \eqref{eq:final} into
  \eqref{eq:diffs}.
\end{proof}

\begin{corollary}\label{cor:same meas}
  If $\meas_m = \meas_j$ for some $1 \leq j \leq m-1$, then
  \begin{equation*}
    \tar(\obs) - \tar(\obsm) =
    \log \left ( 1 + \frac{\sigma^2
      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
      \obsm^* \Sigmam^{-1}e_j \rangle
    }{
      2 - \sigma^2 e_j^t\Sigmam^{-1}e_j 
    }       
    \right ),
  \end{equation*}
  where $e_j\in \mathbb{R}^{m-1}$ is the $j^{\text{th}}$ standard unit
  vector.
\end{corollary}

\begin{proof} \label{cor:same meas proof}
  Denote $A:= \obs \modcov \obs^*$ and $v_j$ the $j^{\text{th}}$
  column of $A$.  Note that $v_j = \obsm \modcov \meas_m$, since
  $(\obsm \modcov \obsm^*)_{ij} = \meas_i(\modcov \meas_j)$, as
  explained in \eqref{eq:modcov explained}. We can now verify that
  \begin{equation}\label{eq:observation}
    \Sigmam^{-1} \obsm \modcov \meas_m = \Sigmam^{-1}v_j = (A +\sigma^2I_{m-1})^{-1} v_j =
    e_j -\sigma^2 \Sigmam^{-1}e_j.
  \end{equation}
  %
  Using \eqref{eq:observation}:
  \begin{align}\label{eq:denominator}
    \begin{split}
      \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m
      &= \meas_m \modcov \obsm^* ( e_j - \sigma^2 \Sigmam^{-1} e_j )\\
      %
      %
      %
      &= \meas_m \modcov \meas_j - \sigma^2 \meas_m \modcov \obsm^* \Sigmam^{-1}e_j \\
      %
      %
      %
      &= \meas_m \modcov \meas_j -\sigma^2 (e_j - \sigma^2 \Sigmam^{-1}e_j)^t e_j \\
      %
      %
      %
      &= \meas_m \modcov \meas_m -\sigma^2 + \sigma^4 e_j^t\Sigmam^{-1}e_j.
    \end{split}
  \end{align}
  We use \eqref{eq:observation} to simplify the enumerator in
  \eqref{eq:conclusion}:
  \begin{align}\label{eq:enumerator}
    \begin{split}
      (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m
      &= \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m - \meas_m \\
      %
      %
      %
      &= \obsm^* (e_j - \sigma^2 \Sigmam^{-1} e_j) -\meas_j \\ 
      %
      %
      %
      &= -\sigma^2 \obsm^* \Sigma^{-1}e_j. 
    \end{split}
  \end{align}
  %
  Now, we substitute \eqref{eq:enumerator} and \eqref{eq:denominator}
  to the enumerator and denominator of \eqref{eq:conclusion}:
  %
  \begin{align*}
    \tar( \obs ) - \tar (\obsm ) &=
    \log \left ( 1 + \frac{
      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m,
      (\obsm^* \Sigmam^{-1} \modcov - I ) \meas_m \rangle
    }{
      \sigma^2 + \meas_m \modcov \meas_m - \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m 
    }       
    \right ) \\
    %
    %
    %
    &= \log \left ( 1 + \frac{\sigma^4
      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
      \obsm^* \Sigmam^{-1}e_j \rangle
    }{
      2\sigma^2 - \sigma^4 e_j^t\Sigmam^{-1}e_j 
    }       
    \right ) \\
    %
    %
    %
    &= \log \left ( 1 + \frac{\sigma^2
      \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1} e_j,
      \obsm^* \Sigmam^{-1}e_j \rangle
    }{
      2 - \sigma^2 e_j^t\Sigmam^{-1}e_j 
    }       
    \right ).
  \end{align*}
\end{proof}

Recall from \eqref{eq:Sigma} that $\Sigma(\obs) = \obs
\modcov \obs^* + \sigma^2I$ and let $u := \obsm^*
\Sigmam^{-1}e_j$. Then

\begin{align*}
  \begin{split}
    \lim_{\sigma^2 \to 0} u &= \obsm^*(\obsm \modcov \obsm^*)^{-1}e_j\\
    \lim_{\sigma^2 \to 0} \postcovm &= (\prcov^{-1} + \fwd^* \obsm^* (\obsm \modcov \obsm^*)^{-1} \obsm \fwd)^{-1} \text{ (From \eqref{eq:postcov})}.
  \end{split}
\end{align*}

Consequently, 
\begin{equation*}
   \langle \fwd \postcovm \fwd^* \obsm^* \Sigmam^{-1}
    e_j, \obsm^* \Sigmam^{-1}e_j \rangle 
  %
  %
  = \langle \fwd \postcovm \fwd^* u, u \rangle
\end{equation*}

is bounded, and

\begin{equation*}
\lim_{\sigma^2 \to 0} \tar(\obs) -\tar(\obsm) = 0.
\end{equation*}

We have shown that in the limit $\sigma^2 \to 0$, no increase in
$\tar$ is achieved by repeating a measurement, so designs that exhibit
measurement clusterization are not D-optimal. Since $\tar$ is not
defined for $\sigma^2 = 0$ and identical measurements, we cannot make
a statement regarding $\sigma^2 = 0$, except in the limiting sense
described above. In conclusion, for small observation error $\sigma^2$
levels, measurement clusterization is mitigated by the presence of a
non-zero model error $\modcov$ --- answering Question \ref{q:mitigate}
posed in the Introduction.% Section \ref{section:intro}.

%\input{vanishing}
\section{D-Optimal Designs Without Model Error}\label{section:vanishing}
Our goal in this section is to prove Theorem \ref{thm:char} which
characterizes D-optimal designs when $\modcov = 0$. The necessary
first-order condition for D-optimality of Theorem
\ref{thm:constrained} for $\modcov = 0$ become:

\begin{equation}\label{eq:eigenproblem}
  \sigma^{-2}\fwd \postcov \fwd^* \obs^* = \obs^* \Xi,
\end{equation}
with $\Xi$ diagonal. Equation \eqref{eq:eigenproblem} looks like an
eigenvalue problem for the self-adjoint operator $\sigma^{-2}\fwd
\postcov \fwd^*$, where rows of $\obs$, namely $\meas_j,j=1,\dots, m$,
are eigenvectors. However, $\postcov$ depends on $\obs$, so we refer
to \eqref{eq:eigenproblem} as a \emph{nonlinear} eigenvalue problem.
%% Proposition \ref{prop:twice woodbury} and Lemma \ref{lemma:sim
%% diag} required

\begin{proposition}\label{prop:twice woodbury}
  Assume $\fwd \prcov \fwd^*$ is invertible. Then
  \begin{align*}
    \begin{split}
      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* 
      %
      %
      = \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1},
    \end{split}
  \end{align*}  
\end{proposition}

\begin{proof}
  The proof amounts to using Woodbury's matrix identity twice with a
  regularization trick. The standard proof for Woodbury's matrix
  identity works for separable Hilbert spaces, as long as all terms
  are well defined. Unfortunately, $\obs^*\obs$ is not invertible, so
  we force it to be. Recall Woodbury's matrix identity:
  \begin{equation}\label{eq:WMI}
    (A + UCV)^{-1} = A^{-1} - A^{-1}U(C^{-1} + VA^{-1}U)^{-1}VA^{-1}. 
  \end{equation}

  Denote:
  \begin{align}\label{eq:notation}
    \begin{split}
      A :&= \prcov^{-1} \\
      U :&= \fwd^* \\
      V :&= \fwd \\
      C :&= \sigma^{-2} (\obs^*\obs+\eps I),\ \eps > 0
    \end{split}
  \end{align}

  Then, \eqref{eq:WMI} with the notation \eqref{eq:notation} implies:
  \begin{align}\label{eq:first}
    \begin{split}
    &\fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* \\
    &\ \ \ \ \ = \fwd ( \prcov - \prcov \fwd^* ( \sigma^2(\obs^*\obs + \eps I)^{-1} + \fwd \prcov \fwd^* )^{-1} \fwd \prcov ) \fwd^*. \\
      %
      %
      %% &= X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X
  %%   \end{split}
    %% \end{align*}
    \end{split}
  \end{align}
  Now, denote:
  \begin{align*}
    X :&= \fwd\prcov \fwd^*, \\
    Y_{\eps} :&= \obs^*\obs + \eps I,
  \end{align*}
  and observe that $Y_{\eps}$ is invertible. Substitute $X, Y_{\eps}$
  into the RHS of \eqref{eq:first}:
   \begin{equation}\label{eq:second}
     \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* = X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X
   \end{equation}

   Note that $X + \sigma^2 Y_{\eps}^{-1}$ is invertible, as the sum of
   two positive definite operators. Now, let
   \begin{align}\label{eq:notation2}
     \begin{split}
       A :&= X^{-1}, \\
       C :&= \sigma^2Y_{\eps}^{-1}, \\
       U :&= I, \\
       V :&= I.
     \end{split}
   \end{align}
   Apply \eqref{eq:WMI} with the notation \eqref{eq:notation2} to the
   RHS of \eqref{eq:second}:
  \begin{align*}
    \begin{split}
      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* &= X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X \\
      %
      %
      %
      &= (X^{-1} + \sigma^{-2}Y_{\eps})^{-1} \\
      %
      %
      %
      &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \eps I))^{-1}.
    \end{split}
  \end{align*}

  We conclude that $\forall \eps > 0$
  \begin{align*}
    \begin{split}
      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* 
     &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \eps I))^{-1}.
    \end{split}
  \end{align*}
  Letting $\eps \to 0$ completes the proof.
\end{proof}

\begin{lemma}[Simultaneous diagonizability]\label{lemma:sim diag}
  Let $\hil$ separable Hilbert space, $C:\hil \to \hil$ self-adjoint
  and $\func_1,\dots,\func_m \in \hil$. Denote $\func^*$ the element
  $\func$ acting as a linear functional. If
  \begin{equation*}
   (C + \sum_{j=1}^m \func_j\func_j^*) \func_l = \xi_l \func_l,\ l = 1,\dots,m
  \end{equation*}
  then $C$ and $\sum_{j=1}^m \func_j \func_j^*$ are simultaneously
  diagonalizable.
\end{lemma}
\begin{proof}
  First, enumerate the eigenvalues of $C + \sum_{j=1}^m
  \func_j\func_j^*$ as $\xi_1,\dots,\xi_\ell$. Denote the indices of
  the eigenvectors corresponding to $\xi_i$
  \begin{equation*}
    S_i := \{ 1 \leq k \leq m | (C + \sum_{j=1}^m \func_j\func_j^* )\func_k = \xi_i \func_k \}.
  \end{equation*}
  Define further
  \begin{equation*}
    A_i := \sum_{k \in S_i} \func_k \func_k^*,
  \end{equation*}
  which is self-adjoint. Two observations are in order. First,
  $\sum_{j=1}^m \func_j\func_j^* = \sum_{i=1}^\ell A_i$. Second, $A_i
  \func_k = 0$ if $k\not \in S_i$, since eigenvectors of different
  eigenvalue are orthogonal. For $k \in S_i$
  \begin{equation}\label{eq:on vi}
    \xi_i \func_k = (C + \sum_{j=1}^m \func_j \func_j^* ) \func_k = (C + A_i) \func_k.
  \end{equation}
  Let $V_i := span \{\func_k \}_{k\in S_i}$. Observe that $V_i$ is
  invariant under $A_i$, by definition, and under $C$, by \eqref{eq:on
    vi}. A second application of \eqref{eq:on vi} shows that $A_i =
  \xi_iI - C$ on $V_i$. This immediately implies $A_i$ and $C$ are
  simultaneously diagonalizable on $V_i$. This holds for every $1 \leq
  i \leq \ell$ and we conclude that $C$ and $A$ are simultaneously
  diagonalizable.
\end{proof}

\begin{proposition}\label{prop:same ev}
  Let $\obs$ satisfy the nonlinear eigenvalue problem
  \eqref{eq:eigenproblem}. Then $\obs^*\obs$ and $\fwd \prcov \fwd^*$
  are simultaneously diagonalizable.
\end{proposition}
\begin{proof}
  \begin{align}\label{eq:mod conditions}
    \begin{split}
      \obs^* \Xi &= \sigma^{-2}\fwd \postcov \fwd^* \obs^*  \text{ (by \eqref{eq:eigenproblem})}\\
      %
      %
      %
      &= \sigma^{-2} \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* \obs^*  \text{ (by \eqref{eq:postcov})} \\
      %
      %
      %
      &= \sigma^{-2} \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1} \obs^* \text{ (by Proposition \ref{prop:twice woodbury})}.
    \end{split}
  \end{align}

  Now take $\func_j^{*} = \meas_j$ and $C := (\fwd \prcov
  \fwd^*)^{-1}$ and use Lemma \ref{lemma:sim diag}.
\end{proof}

Since we made no assumption regarding the ordering of $\{\lambda_i\}$,
we can denote the corresponding non-zero eigenvalues of $\obs^*\obs$
by $\{\eta_i\}_{i=1}^{k}$ and let $\eta_i = 0$ for $i \geq k+1$.

\begin{proposition}\label{prop:true target}
  Let $\obs$ with $m$ measurements satisfy the nonlinear eigenvalue
  problem \eqref{eq:eigenproblem}. Let $\{\eta_i\}_{i=1}^{\infty}$
  eigenvalues of $\obs^*\obs$ and $\{\lambda_i\}_{i=1}^{\infty}$ the
  corresponding eigenvalues of $\fwd \prcov \fwd^*$. Let $k:=\rank
  \obs^*\obs$. Without loss of generality, let $\eta_i > 0$ for $i\leq
  k$ and $\eta_i = 0$ for $i > k$. Then:
  \begin{enumerate}
    \item $k \leq m$ and $\obs^*\obs$ has exactly $k$ positive
      eigenvalues.
    \item
      \begin{equation*}
        \tar(\obs) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i) = \frac12 \sum_{i=1}^{m} \log (1 + \sigma^{-2}\lambda_i\eta_i).
      \end{equation*}
    \item Furthermore, if $\obs$ is D-optimal, $\eta_i > 0$ for
      eigenvectors corresponding to the $k$ largest $\lambda_i$.
  \end{enumerate}
\end{proposition}
\begin{proof}
  Part (1) is trivial. To see part (2) holds: 
  \begin{align}
    \begin{split}
      \tar(\obs) &= \frac12\log \det \left (I + \sigma^{-2} \prcov^{1/2} \fwd ^* \obs^*
      \obs \fwd \prcov^{1/2}\right )\\% \text{ (by definition)}\\
      %
      &= \frac12 \log \det \left (I + \sigma^{-2} \obs^* \obs \fwd
      \prcov\fwd^* \right ) \text{ (Sylvester's Determinant
      Theorem)}\\
      %
      %
      %
      &=\frac12 \log \prod_{i=1}^{\infty} ( 1 + \sigma^{-2} \lambda_i\eta_i ) \text{ (Proposition \ref{prop:same ev})} \\
      %
      %
      %
      %&=\frac12 \log \left ( \prod_{i=1}^{k} ( \lambda_i^{-1} + \sigma^{-2} \eta_i )\prod_{i=1}^{k} \lambda_i \right )
      %
      %
      %
      &=\frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i). 
      %
      %
      %% &=\frac12 \sum_{i=1}^{k} \log(\sigma^2\lambda_i^{-1} + \eta_i) - k\log \sigma + \frac12\sum_{i=1}^k \log \lambda_i.
    \end{split}
  \end{align}
  Part (3) holds since $\log$ is increasing and $\eta_i \geq 0$.
\end{proof}

\begin{proposition}\label{prop:kkt}
  Let $\tar: \mathbb{R}^m \to \mathbb{R}$, $\tar(\eta) =
  \frac{1}{2}\sum_{i=1}^m \log (1+\sigma^{-2}\lambda_i \eta_i)$, with
  $\lambda_i > 0$ and $\sigma^{2} > 0$. Then the maximum of $\tar$
  subject to $\eta_i \geq 0$ and $\sum\eta_i = m$ is obtained at
  \begin{equation}
  \eta_i = \begin{cases}
    \frac{m}{k} - \sigma^2 \lambda_i^{-1} + \sigma^2 \frac{1}{k} \sum_{j\in A} \lambda_j^{-1} & i \in A \\
    0 & i \in A^c
  \end{cases}
  \end{equation}
  where $A:= \{1\leq i \leq m: \eta_i > 0\}$ and $A^c = \{1,\dots, m\}
  \backslash A$, and $k = |A|$, the cardinality of $A$.
\end{proposition}

\begin{proof}
  Let $\Phi(\eta) = \sum_{i=1}^k \eta_i - m$ and $\Omega_j(\eta) =
   -\eta_j$. Then
  \begin{align*}
    \begin{split}
      \frac{\partial \tar}{\partial \eta_i}  &=
       \frac12 \frac{\sigma^{-2}\lambda_i}{1 + \sigma^{-2} \lambda_i\eta_i} = \frac12 \frac{1}{\sigma^{2}\lambda_i^{-1} + \eta_i} \\
      %
      %
      %
      \frac{\partial\Phi}{\partial \eta_i} &= 1 \\
      %
      %
      %
      \frac{\partial \Omega_j}{\partial \eta_i} &= -\delta_{ij}      
    \end{split}
  \end{align*}

  From the KKT conditions, there are $\alpha, \beta_i$ such that for $i=1,\dots,m$:
  \begin{align}
    \begin{split}
      -\frac12 \frac{1}{\sigma^{2}\lambda_i^{-1} + \eta_i} + \alpha - \beta_i  &= 0 \\
      %
      %
      %
      \eta_i &\geq 0\\
      %
      %
      %
      \beta_i &\geq 0\\
      %
      %
      %
      \beta_i \eta_i &= 0\\
      %
      %
      %
      \sum_{i=1}^m \eta_i &= m 
    \end{split}
  \end{align}

  Then, for $i \in A$:
  \begin{align*}
    \begin{split}
      \beta_i &= 0\\
      %
      %
      %
      \sigma^{2}\lambda_i^{-1} + \eta_i  &= \frac{1}{2\alpha}%  \text{ for }  1 \leq i \leq k.\\
      %
      %
      %
      %% \sigma^{-2} \lambda_i &= 2\alpha - 2\beta_i, i \in B \\
    \end{split}
  \end{align*}

  Summing over $i \in A$, substituting $\sum_{i\in A} \eta_i =
  \sum_{i=1}^m \eta_i = m$, and dividing by $k:= |A|$:
  $$
  \frac{1}{2\alpha} = \frac{m}{k} + \frac{\sigma^2}{k} \sum_{j\in A} \lambda_j^{-1}.
  %\Rightarrow \alpha = \left ( \frac{2m}{k} + \frac{2\sigma^2}{k}
  %\sum_{i\in A} \lambda_i^{-1} \right )^{-1}.
  $$

  Consequently:
  \begin{align} \label{eq:etas}
    \begin{split}
      \eta_i &= \frac{1}{2\alpha} - \sigma^2 \lambda_i^{-1}\\
      %
      %
      %
      &=\frac{m}{k} + \frac{\sigma^2}{k} \sum_{j\in A} \lambda_j^{-1} - \sigma^2 \lambda_i^{-1}.%\\
      %
      %
      %
      %% &= \frac{m}{k} + \frac{\sigma^2}{k} \sum_{j \in A, j\neq i} \lambda_j^{-1} - \frac{k-1}{k}\sigma^2 \lambda_i^{-1}\\
      %% ,\ i \in A.
    \end{split}
  \end{align}
\end{proof}

The final ingredient we require for the proof of Theorem
\ref{thm:char} is:

\begin{lemma}[Unit norm decomposition]\label{lemma:free}
  Let $M \in \R^{k \times k}$ symmetric positive definite with $\ttr M
  = m$, $m \geq k$. We can find $\func_j \in \R^k,j=1,\dots,m$
  with $\|\func_j\|=1$ and $A = (\func_1,\dots,\func_m)$ such that
  $AA^t = M$.
\end{lemma}

\begin{proof}
  Let us diagonalize $M$, so that $M = U D U^t$ with $D =
  \diag(d_1,\dots,d_k)$ and $U \in \R^{k \times k }$ orthogonal. Let
  $S \in \R^{k \times m}$ with $S_{ii} = \sqrt{d_{i}}$ and zeros
  otherwise. Define $A:= U S V^t$, where $V \in \R^{m \times m}$ is
  orthogonal and will be further restricted later. Then $AA^t = U
  SV^tVS^t U^t = UDU^t$, so $AA^t$ has the required eigenvalues and
  eigenvectors by construction. If we can choose $V$ such that $A$
  also satisfies the unit norm constraints we are done. These
  constraints are, for $j=1,\dots,m$:
  \begin{equation}\label{eq:V constraints}
   1 = [A^tA]_{jj} = [V S^tS V^t]_{jj},
  \end{equation}
  and we can expect to do this since we assumed $\ttr D = m$.

  Define $C = S^tS - I \in \R^{m \times m}$. Note that $\ttr C = 0$ and
  $C$ is diagonal with non-zero entries $d_i-1,i=1,\dots,k$. It suffices
  to find $V$ orthogonal such that $V C V^t$ has zero diagonal. We
  construct such $V$ by sequentially inserting zeros in the diagonal
  and not destroying zeros we already introduced, starting from the
  last diagonal entry and moving to the first. Since $c_{mm} \neq 0$ ,
  let $p < m$ such that $c_{pp}c_{mm} < 0$ (such $p$ exists because
  the trace is zero) and let $\theta \in (0,\pi)$. Define a Givens
  rotation $R^{(m)} \in \R^{m \times m}$ by
  \begin{equation*}
    r^{(m)}_{ab} :=
    \begin{cases}
      1 & a = b \neq p \text{ or } a = b \neq m \\
      \cos \theta & a = b = p  \\
     -\sin \theta & a = p, b = m\\
      \cos \theta & a = b = m \\
      \sin \theta & a = m, b = p \\ 
      0 & o.w
    \end{cases}
  \end{equation*}
  Note that conjugating a matrix by $R^{(m)}$ changes only its $m$ and
  $p$ rows and columns. We want to choose $\theta$ such that
  \begin{equation}\label{eq:mm}
    0 = [R^{(m)} C (R^{(m)})^t]_{mm} = \cos^2 \theta c_{mm} + 2\cos \theta \sin
    \theta c_{mp} + \sin^2\theta c_{pp},
  \end{equation}
  and it suffices to choose $\theta$ such that
  \begin{equation*}
    c_{mm} \cot^2 \theta + 2 c_{mp} \cot \theta + c_{pp} = 0.
  \end{equation*}
  This quadratic in $\cot\theta$ has a real solution, since
  $c_{pp}c_{mm} < 0$ by assumption and we can find $\theta \in
  (0,\pi)$ such that \eqref{eq:mm} is satisfied. We continue to find
  $R^{(m-1)}$ that leaves row and column $m$ unchanged and
  continue introducing zeros to the diagonal. The assumption $\ttr D =
  m \Rightarrow \ttr C = 0$ guarantees we can do that. Taking $V:=
  R^{(1)} R^{(2)} \dots R^{(m-1)}R^{(m)}$ completes the proof.
\end{proof}

\char*

\begin{proof}
  Part (1) is immediate for any measurement operator $\obs$ that
  satisfies the unit norm constraint on measurements. Part (2)
  was proved in Proposition \ref{prop:same ev}. Part (3) was proved in
  Proposition \ref{prop:true target}.

  Part (4) is a consequence of Propositions \ref{prop:true target} and
  \ref{prop:kkt}, with the caveat that we did not show that finding
  $\obs$ so that $\obs^*\obs$ has the desired eigenvalues is
  feasible. To this end, we utilize Lemma \ref{lemma:free}: let $M =
  \diag(\eta_1, \dots, \eta_k)$, diagonal with respect to the first $k$
  eigenvectors of $\fwd \prcov \fwd^*$. We take $\obs := A$ from the
  Lemma \ref{lemma:free}.

  Recall from \eqref{eq:postcov}, that the posterior precision is
  $\postcov^{-1} = \prcov^{-1} + \sigma^{-2}\fwd^*\obs^*\obs\fwd$. The
  first statement in part (5) now follows from Proposition
  \ref{prop:twice woodbury}, while the second statement follows from
  parts (1) and (4).
\end{proof}

\pgfplotstableread{
  Label     prior  optimal  sub-optimal 
  1         0.2    1.8           1.7
  2         0.8    1.2           0.8
  3         2.2    0             0.5
  4         3.5    0             0.0
}\optimalvsnot

\begin{figure}
  \centering
  \begin{tikzpicture}[scale=0.75]
    \begin{axis}[
        ybar stacked,
        ymin=0,
        ymax=4,
        xtick=data,
        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
        reverse legend=false, % set to false to get correct display, but I'd like to have this true
        xticklabels from table={\optimalvsnot}{Label},
        xticklabel style={text width=2cm,align=center},
        legend plot pos=right,
        ylabel={\LARGE precision --- prior and posterior},
        xlabel={\LARGE eigenvector},
      ]
      \addplot [fill=blue!60]  table [y=prior,   meta=Label, x expr=\coordindex] {\optimalvsnot};
      \addplot [pattern=north east lines, pattern color=green!80]  table [y=optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};     
      \addlegendentry[scale=1.4]{$\sigma^2\lambda_i^{-1}$}
      \addlegendentry[scale=1.4]{optimal $\eta_i$s}
    \end{axis}
  \end{tikzpicture}
  \begin{tikzpicture}[scale=0.75]
    \begin{axis}[
        ybar stacked,
        ymin=0,
        ymax=4,
        xtick=data,
        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
        reverse legend=false, % set to false to get correct display, but I'd like to have this true
        xticklabels from table={\optimalvsnot}{Label},
        xticklabel style={text width=2cm,align=center},
        legend plot pos=right,
        ylabel={\LARGE precision --- prior and posterior},
        xlabel={\LARGE eigenvector} ,
      ]   
      \addplot [fill=blue!60]  table [y=prior,       meta=Label, x expr=\coordindex] {\optimalvsnot};
      \addplot [pattern=north east lines, pattern color=green!80]  table [y=sub-optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};
      \addlegendentry[scale=1.4]{$\sigma^2\lambda_i^{-1}$}
      \addlegendentry[scale=1.4]{sub-optimal $\eta_i$s}
    \end{axis}
  \end{tikzpicture}
  \caption{A comparison of the eigenvalues of the pushforward
    posterior precision $(\fwd\prcov\fwd^*)^{-1} +
    \sigma^{-2}\obs^*\obs$ for a D-optimal design (left) and a
    sub-optimal design (right). Both designs are allowed $m=3$
    measurements. We assume $\sigma^2=1$ and thus, the blue area has
    accumulated height of $\sigma^{-2}m = 3$ in both panels. The
    D-optimal design (left) increases precision where it is
    lowest. The sub-optimal design (right) does not.}
  \label{fig:optimal vs not}
\end{figure}

Theorem \ref{thm:char} facilitates understanding of D-optimal designs
when $\modcov = 0$: Imagine each eigenvector of $\fwd \prcov \fwd^*$
corresponds to a graduated lab cylinder.
%% (Figure \ref{fig:cylinder})
Each cylinder $i$ is filled, a-priori, with green liquid of
$\lambda_i^{-1}$ volume units. We are allowed $m$ measurement, so we
have blue liquid of volume $\sigma^{-2}m$ units at our disposal. When
we seek a D-optimal design, we distribute the blue liquid by
repeatedly adding a drop to whatever cylinder currently has the lowest
level of liquid in it, as long as its index $i \leq m$. The result of
such a procedure is illustrated in Fig.~\ref{fig:optimal vs not}.
%% \begin{figure}%{r}{0.25\textwidth} 
%%     \centering
%%     \includegraphics[width=5cm, height=5cm]{cylinders.jpg}
%%     \caption{Three graduated lab cylinders, corresponding to three
%%       eigenvectors. Prior eigenvalues not shown.}
%%     \label{fig:cylinder}
%% \end{figure}

%\input{why}
\section{Answer to Question \ref{q:why}: The cause of clusterization}\label{section:why}
According to Theorem \ref{thm:char}, D-optimal designs aim to capture
a small subset of eigenvectors of the prior covariance, specifically
the $k$ eigenvectors with the highest prior variance. Our model
naturally achieves this objective by not measuring eigenvectors $k+1$
and above, as proven in Theorem \ref{thm:char}. Translating this
understanding to spatial problems, we anticipate that a D-optimal
design would favor measurement locations where eigenvectors $k+1$ and
above are either close to zero in value or possess small eigenvalues
in the prior spectrum, for some $k > 0$. To illustrate this
preference, Fig.~\ref{fig:why} depicts the scenario using the 1D
heat equation with homogeneous Dirichlet boundary conditions (details
in the supplementary material). The plot showcases four eigenvectors,
scaled according to their prior standard deviations. Since
eigenvectors beyond the fourth have insignificant prior eigenvalues,
we exclude them from consideration. Notably, we observe that
measurements are clustered near the zeros of the third and fourth
eigenvectors, so we conclude that $k=2$. Clusterization arises because
there are only two locations where the third and fourth eigenvectors
approach zero, while the first and second eigenvectors exhibit
significantly non-zero values. Consequently, when we allocate four
measurements to these two locations, they naturally cluster together,
aligning with the pigeonhole principle.

\bibliographystyle{ba}
\bibliography{/home/yair/projects/bibtex.bib}

\begin{acks}[Acknowledgments]

%\input{ack}
This study is a result of research I started during my PhD studies
under the instruction of Prof. Georg Stadler at the Courant Institute
of Mathematical Sciences. I would like to thank him for his great
mentorship, attention to details and kindness. I would also like to
thank Christian Remling, who helped me find a proof for Lemma
\ref{lemma:free} in
\href{https://mathoverflow.net/questions/280168/redistribute-diagonal-entries-of-a-matrix/280203#280203c}{Mathoverflow}.

%\input{funding}
This research was supported in part by an appointment with the
National Science Foundation (NSF) Mathematical Sciences Graduate
Internship (MSGI) Program sponsored by the NSF Division of
Mathematical Sciences. This program is administered by the Oak Ridge
Institute for Science and Education (ORISE) through an interagency
agreement between the U.S. Department of Energy (DOE) and NSF. ORISE
is managed for DOE by ORAU. All opinions expressed in this paper are
the author's and do not necessarily reflect the policies and views of
NSF, ORAU/ORISE, or DOE.
% This work was also supported by The Raymond and Beverly Sackler
% Post-Doctoral Scholarship.

\end{acks}

\end{document}

### 2024-07-21 latex/intro.tex
Intro with new files

--- 2024-06-16 latex/intro.tex
@@
 \section{Introduction}\label{section:intro}
-The practice of experimental design entails specifying every detail of
-running an experiment \cite{chaloner}: at its simplest form,
-experimental design involves choosing a set of measurements to be
-taken. In a medical experiment, experimental design may involve
-choosing participants to whom some drug will be administered, and
-measuring their blood pressure. Experimental design also arises 
-process of searching for oil. Since measuremnts are expensive and
-require digging holes in the ground, only a limited number of
-measuremnts can be taken, hence carefully choosing measuremt locations
-is crucial.
+Measurements play a fundamental role in generating observations and
+are indispensable to any inference process. Selecting the optimal set
+of measurements holds particular significance when inferring
+parameters of a physical process. Unlike many other fields, where
+measurements are fixed, in this context, we have the freedom to choose
+which measurements to take. This freedom should be taken advantage of:
+measurements should be chosen to enhance accuracy, reduce costs, or
+both. Whether measurements involve specifying electrode locations on
+the skin in electric impedance tomography \cite{horesh2010impedance},
+determining certain wavelengths in MRI \cite{horesh2008mri},
+positioning sensors for detecting ground-reflected waves in the search
+for oil \cite{horesh2008borehole}, or optimizing sensor locations in a
+seismographic network \cite{rabinowitz1990, steinberg1995} --- the
+selection of optimal measurements, referred to as the problem of
+\emph{optimal design}, becomes crucial. Optimal measurements are
+chosen according to specific \emph{design criteria}, with A- and
+D-optimality being two of the most widely recognized and extensively
+studied design criteria \cite{Chaloner1995,
+  AlexanderianGloorGhattas14}.
 
-Specifiyng the right set of measurements holds particular significance
-when solving an \emph{inverse problem} --- i.e. making inference of
-the physical world utilizing a physical model of some phenomenon of
-interest. In an inverse problem, we seek to infer physical quantities,
-based on observations and a model of the physical world, typically
-phrased in the language of ordinary or partial differential equations.
-
-Inverse problems are ubiquitous in science and engineering, and we
-will now give several illustrative examples. In electric impedance
-tomography (EIT) we seek to infer the impedance of the inside of a
-human body from observations of electrical current impedence measured
-at specific electrode locations on the skin, leveraging known
-equations of electric current flow
-\cite{horesh2010impedance}. Magnetic Resonance Imaging (MRI) also
-entails solving an inverse problem, where radio-frequency pulses are
-sent through the human body in the presence of a strong magnetic field
-and the response of our bodies' contents to such perutrbations is
-measured, allowing a radiologist to view the internals of a body in a
-noninvasive manner \cite{horesh2008mri}. Yet another example of an
-inverse problem arises in oil and gas exploration, where acoustic
-waves are sent through "boreholes" --- deep cylindrical wells drilled
-into the ground --- and data of travel and return times of said waves
-are recorded. Since travel and return time is influenced by the
-properties of the subsurface materials, such as density, elasticity,
-and the presence of fluids or voids, we can combine travel data with a
-geophisical models of contents of the earth's crust to facilitate
-reconstructing and inferring the structure of the subsurface in a
-process called \emph{borehole tomography}
-\cite{horesh2008borehole}. Inverse problems also arise in many other
-areas of seismology and geology \cite{rabinowitz1990, steinberg1995}
-and medical imaging \cite{tarantola}. Importantly, note that in the
-above examples our goal is to infer some function $f \in C(\Omega)$,
-where \(\Omega \subseteq \mathbb{R}^d, d=1,2,3\) is a spatial domain
-of interest.
-
-%% In EIT, electrode locations determine the quality of the
-%% reconstructed impedance field, while in MRI utilizng certain
-%% radio-frequency pulses may give more accurate information than
-%% others. In borehole tomogrpahy, only several boreholes could be dug
-%% and the choice of their locations is also crucial for oil and gas
-%% discovery. Thus, the selection of optimal measurements, referred to as
-%% the problem of \emph{optimal design}, becomes crucial.
-%% in a specific location (in a typical
-%% medical application to measure inverventions on, . These experimental
-%% units may be individual patients, but also Although common sense is
-%% useful in this process, it is often the case that naive intuition
-%% results in experiments that do not generate the maximal amount of
-%% information possible. To this end
-
-We will now introduce a toy model which we will use to define a toy
-inverse problem: the heat equation in one dimension with a homogeneous
-Dirichlet boundary condition. The equations are presented below for
-completeness, but readers less familiar with partial differential
-equations should think of the heat equation as follows: before us
-there is a perfectly insulated metal rod. At each point on the rod the
-temperature is different, and we call this temperature distribution
-the \emph{initial condition}. We wait for a short time and let the
-heat dissipate a bit inside the rod. The final heat distribution in
-the rod is dictated by the \emph{1D heat equation}:
-\begin{subequations}\label{eq:heat equation}
-  \begin{alignat}{2}
-    u_t &= \Delta u &&\qquad \text{in } [0,1] \times [0,\infty),\\ u
-      &= 0 &&\qquad \text{on } \{0, 1\} \times [0,\infty),\\ u &= u_0
-        &&\qquad \text{on }[0,1] \times \{0\}.
-  \end{alignat}
-\end{subequations}
-
-In the inverse problem, our goal is to infer the initial condition
-(i.e.~initial temperature distribution inside the rod). We are allowed
-to take some number of temperature readings on the rod once our
-waiting time has passed. The inverse problem of the 1D heat equation
-is difficult ("ill-posed") since heat spreads in a diffusive manner
-and any roughness in the initial condition is quickly smoothed, see
-Supplementary movies S1 and S2. We will give a fully Bayesian
-formulation of this inverse problem in Section \ref{section:prelims}.
-
-In all of the examples of inverse problems we have seen, as well as in
-many other applications, measurements should be chosen to most benefit
-the user. Hence, the user may consider some utility, called a
-\emph{design criterion} and take measurements that maximize this
-utility. Two of the most widely recognized and extensively studied
-design criteria are A- and D-optimality. We focus on the Bayesian
-D-optimality criterion, which carries a simple and intuitive meaning:
-measurements chosen according to the D-optimality criterion maximize
-the expected Kullback-Liebler divergence from posterior to prior
-\cite{Chaloner1995, AlexanderianGloorGhattas14}, see also precise
-definition in Section \ref{subsec:D optimal design}. We refer to a set
-of measuremnts that maximizes the Bayesian D-optimality criterion a
-\emph{D-optimal design}.
-
-Surprisingly, A- and D-optimal designs for inverse problems have been
-observed to yield remarkably similar measurements in certain cases
-\cite{fedorov1996, nyberg2012, fedorov1997, Ucinski05,
-  neitzel2019sparse}. This phenomenon is illustrated for our toy
-inverse problem in Fig.~\ref{fig:clusterization illustration}, where
-D-optimal measurement locations are shown for different numbers of
-measurements. Notably, for six measurements, a D-optimal design yields
-two sets of measurements that are almost indistinguishable from one
-another. We refer to this intriguing phenomenon as \emph{measurement
-clusterization} \cite{Ucinski05}. For our purposes, it suffices for
-measurements to be visibly close for clusterization to hold. We refer
-to a design that exhibits measurement clusterization as a
-\emph{clustered design},
+Surprisingly, A- and D-optimal designs have been observed to yield
+remarkably similar measurements in certain cases \cite{fedorov1996,
+  nyberg2012, fedorov1997, Ucinski05, neitzel2019sparse}. To
+illustrate this point, we consider a toy problem: inferring
+the initial condition of the 1D heat equation (details in the
+supplementary material). In Fig.~\ref{fig:clusterization
+  illustration} D-optimal measurement locations are shown for
+different numbers of measurements. Notably, for six measurements, a
+D-optimal design yields two sets of measurements that are almost
+indistinguishable from one another. We refer to this intriguing
+phenomenon as \emph{measurement clusterization} \cite{Ucinski05}, and
+we refer to a design that exhibits measurement clusterization as a
+\emph{clustered design}.
 
@@
     \centering
-    \includegraphics[height=0.5\textwidth]{example.pdf}
-    \caption{Measurement clusterization for D-optimal designs for the
-      inverse problem of the heat equation. Measurement locations were
+    \includegraphics[height=0.5\textwidth]{example.png}
+    \caption{Measurement clusterization for optimal designs when
+      inverting for the initial condition of the 1D heat equation (see
+      supplementary material for details). Measurement locations were
       chosen according to the Bayesian D-optimality criterion of
@@
       clarity. Measurement clusterization already occurs for three
-      measurements: the second measurement (red) is overlaid on the
+      measurements: the second measurement (red) is overlaid with the
       third (green). For five measurements, first (blue) and second
@@
 
-Clusterization should not be confused with either repetition nor with
-replication, which are commonly viewed as beneficial and even
-necessary aspects of an optimal design \cite{morris, fisher
-  schafer2001replication}. For example, in his famous milk and tea
-experiment, \cite[Section 11]{fisher} suggested that repetition is a
-"way of enlarging the experiment and, thereby, increasing its
-sensitiveness". In another example, \cite{fay2000rainfall} measured
-the effect of rainfall on grass growth in plots of land. The
-experiment involved fifteen "rainfall manipulation shelters", where
-"Four rainfall manipulation treatments (three replicates) then were
-assigned to 12 of the plots". It seems reasonable the researchers
-replicate the phenomenon they are trying to study. Clusterization
-seems like a whole different beast: a clustered design in the rainfall
-experiment would imply the researchers should take a repeated
-measurement \emph{on the same plot}, at the expense of measuring grass
-growth in other plots. In sharp contrast to repetition and
-replication, clusterization is highly nonintuitive and, as we will
-see, its origins run considerably deeper.
-
-Researchers of inverse problems widely agree that measurement
-clusterization is undesirable \cite{fedorov1996, nyberg2012,
-  fedorov1997, Ucinski05, neitzel2019sparse}, prompting the
-exploration of various remedies to address this issue. One approach
-involves merging close measurements \cite{fedorov1997}; however, this
-strategy merely overlooks the phenomenon of measurement
-clusterization. An alternative solution lies in
-\emph{clusterization-free design}s, where measurement locations are
+Researchers widely agree that measurement clusterization is
+undesirable \cite{fedorov1996, nyberg2012, fedorov1997, Ucinski05,
+  neitzel2019sparse}, prompting the exploration of various remedies to
+address this issue. One approach involves merging close measurements
+\cite{fedorov1997}; however, this strategy merely overlooks the
+phenomenon of measurement clusterization. An alternative solution lies
+in \emph{clusterization-free design}s, where measurement locations are
 deliberately chosen to be distant from one another. This can be
@@
     \centering
-    \includegraphics[width=\textwidth]{eigenvectors.pdf}
+    \includegraphics[width=\textwidth]{eigenvectors.png}
     \caption{D-optimal measurement locations ($m=4$ measurements) and
@@
 
-The first insight our study gives is that a D-optimal design aims to
-uniformly reduce uncertainties not over space (i.e.~$\Omega$), but
-rather over \emph{eigenvectors}. This insight is a part of Theorem
-\ref{thm:char}, see Section \ref{section:vanishing} for precise
-statement and proof, and see Fig.~\ref{cluste} for an illustration.
+The cornerstone of our investigation into measurement clusterization
+is Theorem \ref{thm:char} proven in Section
+\ref{section:vanishing}. The key insight of Theorem \ref{thm:char}
+lies in its fifth component, which highlights that a D-optimal design
+aims to uniformly reduce posterior uncertainties for posterior
+covariance eigenvectors in observation space. A similar conclusion was
+reached by Koval et al.~\cite{koval2020}, who showed that A-optimal
+designs are best constructed in the space of observations.
 
-In Theorem \ref{thm:char}, we also show that D-optimal designs are
-best understood in the sapce of \emph{observations}. Again, a precise
-statement could be found in Section \ref{section:vanishing}.  This
-result is in accordance with previous work by \cite{koval2020}, who
-showed that A-optimal designs are best constructed in the space of
-observations.
+Before we state Theorem \ref{thm:char} we give some definitions: Let
+\(\hilp, \hilo\) Hilbert spaces, \(\fwd:\hilp \to \hilo\) a linear
+compact operator. Let \(\obs: \hilo \to \mathbb{R}^m\) a linear
+measurement operator, where \(m \in \mathbb{N}\) is the number of
+measurements taken. Let \(\sigma^2 \in \mathbb{R}_{+}\) observation
+noise variance, \(\data = \obs \fwd \param + \eps\), where \(\eps \in
+\mathbb{R}^m\) is iid \(\mathcal{N}(0, \sigma^2)\) noise. Let \(\pr
+\sim \mathcal{N}(\prmean, \prcov)\) prior Gaussian measure on
+\(\hilp\), where \(\prcov\) is the prior covariance operator and let
+\(\post\) the posterior (Gaussian) measure.
+
+\begin{restatable}{theorem}{char}\label{thm:char}
+  Let:
+  \begin{itemize}
+    \item The D-optimality design criterion
+    \cite{AlexanderianGloorGhattas14}:
+    \begin{align*}
+      \begin{split}
+        \tar(\obs) %:&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
+        % 
+        % 
+        % 
+        &= \frac12 \log \det ( I + \sigma^{-2} \prcov^{1/2} \fwd ^*
+        \obs^* \obs \fwd \prcov^{1/2}), 
+      \end{split}
+    \end{align*}
+  \item \(\obs\) a D-optimal design operator
+    \begin{equation*}
+      \obs = \argmax_{\|\meas_j\| = 1, j=1,\dots,m}\tar(\obs),
+    \end{equation*}
+  \item \(\{\lambda_i\}_{i=1}^\infty\) eigenvalues of
+    \(\fwd\prcov\fwd^*\) in decreasing order of magnitude.
+    %% \item \(\{\ev_i\}_{i=1}^\infty\) their corresponding eigenvectors.
+  \item \(\{\eta_i\}_{i=1}^\infty\) eigenvalues of \(\obs^*\obs\).
+
+  \end{itemize}
+
+  Then:
+  \begin{enumerate}
+  \item  \(\tr{\obs^*\obs} = m\).
+  \item \(\obs^*\obs\) and \(\fwd\prcov\fwd^*\) are simultaneously
+    diagonalizable.
+  \item \(k := \rank \obs^*\obs \leq m\) and
+    \begin{equation*}
+      \tar(\obs) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i). %= \frac12 \sum_{i=1}^{m} \log (1 + \sigma^{-2}\lambda_i\eta_i).
+    \end{equation*}
+  %% \item 
+  %%   \begin{equation*}
+  %%     k = \argmax \left \{ k:\lambda_k^{-1} < \sigma^{-2}\frac{m}{k} + \frac{1}{k} \sum_{j=1}^{k}
+  %%     \lambda_j^{-1} \right \}.
+  %%   \end{equation*}
+  \item
+    \begin{equation*}
+        \eta_i = \begin{cases}
+          \frac{m}{k} - \sigma^2 \lambda_i^{-1} + \sigma^2 \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1} & 1 \leq i \leq k \\
+          0 & i > k 
+        \end{cases}.
+    \end{equation*}
+  \item The covariance of the pushforwad \(\fwd_{*} \post\) is \(\left
+    ( (\fwd \prcov \fwd^*)^{-1} + \sigma^{-2} \obs^*\obs \right
+    )^{-1}\) and its eigenvalues are
+    \begin{equation*}
+      \theta_i =
+      \begin{cases}
+        \left(\frac{\sum_{j=1}^k \lambda_j^{-1} + \sigma^{-2}m}{k} \right )^{-1} & i \leq k \\
+        \lambda_i &  i > k 
+      \end{cases}
+    \end{equation*}
+  \end{enumerate}
+%\end{theorem}
+\end{restatable}
 
@@
 generalize several lemmas. Among those, is Lemma \ref{lemma:free},
-which is a lemma in linear algebra that (to the the best of my
-knowledge) is novel: We decompose a symmetric positive definite matrix
-\(M \in \mathbb{R}^{k \times k}\) with \(\ttr M = m \in \mathbb{N}\)
-as \(M = AA^t\), where \(A\in \mathbb{R}^{k \times m}\) has unit norm
-columns.
+which is (to the the best of my knowledge) a novel lemma in linear
+algebra: We decompose a symmetric positive definite matrix \(M \in
+\mathbb{R}^{k \times k}\) with \(\ttr M = m \in \mathbb{N}\) as \(M =
+AA^t\), where \(A\) has unit norm columns.
 
@@
 %% \(X(t)\)is an operator valued function \cite{Lax07}.
-
-Finally, we must stress that we do not suggest that clustered designs
-should be abandoned, nor do we support the use of clustered designs in
-any circumstances. We merely investigate an intriguing phenomenon of
-clusterization and seek a deeper understanding for it. Researchers in
-other disciplinesis not viewed as a problem. For example, \cite{lozan}
-suggest in the introduction section that repeated measurements are
-necessary. Our results align with that common wisdom and settle the
-seemed contradiction.
 

### 2024-07-21 latex/flat.tex
flat.tex updated but does not compile
flat.tex is the old paper. diff works and compiles (no pix tho)

@@
-% \documentclass[ba,preprint]{imsart}% use this for supplement article
 \documentclass[ba]{imsart}
-%\input{definitions}
 \usepackage{amsmath}
@@
 \usepackage{amsthm}
-%% \usepackage{amsaddr}
-
-%% \theoremstyle{plain}
-%% \newtheorem{theorem}{Theorem}
-%% \newtheorem{corollary}[theorem]{Corollary}
-%% \newtheorem{lemma}[theorem]{Lemma}
-%% \newtheorem{proposition}[theorem]{Proposition}
-%% \theoremstyle{definition}
-%% \newtheorem{definition}[theorem]{Definition}
-%% \newtheorem{example}[theorem]{Example}
-%% \newtheorem{conjecture}[theorem]{Conjecture}
-%% \theoremstyle{remark}
-%% \newtheorem{remark}[theorem]{Remark}
+
+\theoremstyle{plain}
+\newtheorem{theorem}{Theorem}
+\newtheorem{corollary}[theorem]{Corollary}
+\newtheorem{lemma}[theorem]{Lemma}
+\newtheorem{proposition}[theorem]{Proposition}
+\theoremstyle{definition}
+\newtheorem{definition}[theorem]{Definition}
+\newtheorem{example}[theorem]{Example}
+\newtheorem{conjecture}[theorem]{Conjecture}
+\theoremstyle{remark}
+\newtheorem{remark}[theorem]{Remark}
 
@@
 \usetikzlibrary{patterns}
-\pgfplotsset{compat=1.9} % set to 1.8 to get old behaviour
-%\usepackage{hyperref}
-
+\pgfplotsset{compat=1.9} 
 \newcommand{\R}{\ensuremath{\mathbb{R}}}
@@
 
-% Definitions for second chapter
 \newcommand{\data}{\mathbf{d}}
@@
 \newcommand{\normal}{\mathcal{N}}
-\newcommand{\pr}{\mu_{\textup{pr}}} %Prior measure
-\newcommand{\post}{\mu_{\textup{post}}^{\data, \obs}} % Posterior measure
-\newcommand{\prmean}{\param_{\textup{pr}}} % Prior mean
-\newcommand{\postmean}{\param_{\textup{post}}} % Posterior mean
-\newcommand{\postcov}{\Gamma_{\textup{post}}} % Posterior covariance
-\newcommand{\prcov}{\Gamma_{\textup{pr}}} % Prior covariance
-\newcommand{\modcov}{\Gamma_{\textup{model}}} % Model covariance
-\newcommand{\tmp}{\mathcal{G}}
+\newcommand{\pr}{\mu_{\textup{pr}}} \newcommand{\post}{\mu_{\textup{post}}^{\data, \obs}} \newcommand{\prmean}{\param_{\textup{pr}}} \newcommand{\postmean}{\param_{\textup{post}}} \newcommand{\postcov}{\Gamma_{\textup{post}}} \newcommand{\prcov}{\Gamma_{\textup{pr}}} \newcommand{\modcov}{\Gamma_{\textup{model}}} \newcommand{\tmp}{\mathcal{G}}
 \newcommand{\meas}{\mathbf{o}}
-\newcommand{\ev}{\mathbf{e}} % eigenvector 
-\newcommand{\func}{\mathbf{a}}
+\newcommand{\ev}{\mathbf{e}} \newcommand{\func}{\mathbf{a}}
 \newcommand{\tr}[1]{\textup{tr}\left \{#1 \right \} }
@@
 \newcommand{\rank}{\textup{rank}\ }
-\newcommand{\des}{\eta} % vector of design parameters
-\newcommand{\sigsqr}{\sigma^2}
-
-%% %\usepackage{comment}% http://ctan.org/pkg/comment
-%% %% %% %\excludecomment{proof}
-%% %% \excludecomment{figure}
-%% %% \let\endfigure\relax
-
-%% %\overfullrule=0pt
-
-%
+\newcommand{\des}{\eta} \newcommand{\sigsqr}{\sigma^2}
+
  \pubyear{2021}
@@
 \issue{TBA}
-% \doi{0000}
-%\arxiv{}
 \firstpage{1}
@@
 \startlocaldefs
-% ** Local definitions **
 \endlocaldefs
@@
 \begin{document}
-
-%% *** Frontmatter ***
 
@@
 
-%\title{\support{}}
 \runtitle{Measurement Clusterization in D-optimal Designs}
-%\author{Yair Daon}
-%\numberwithin{equation}{section}
 
@@
 \address[addr1]{Azrieli Faculty of Medicine, Bar-Ilan University, Safed, Israel
-    \printead{e1} % print email address of "e1"
-}
+    \printead{e1} }
 
@@
 \begin{abstract}
-  %\input{abstract}
-  Estimation of parameters in physical processes often demands costly
-  measurements, prompting the pursuit of an optimal measurement
-  strategy. Finding such strategy is termed the problem of
-  \emph{optimal experimental design}, abbreviated as optimal
-  design. Remarkably, optimal designs can yield tightly clustered
-  measurement locations, leading researchers to fundamentally revise
-  the design problem just to circumvent this issue. Some authors
-  introduce error correlation among error terms that are initially
-  independent, while others restrict measurement locations to a finite
-  set of locations. While both approaches may prevent clusterization,
-  they also fundamentally alter the optimal design problem.
-
-  In this study, we consider Bayesian D-optimal designs, i.e.~designs
-  that maximize the expected Kullback-Leibler divergence between
-  posterior and prior. We propose an analytically tractable model for
-  D-optimal designs over Hilbert spaces. In this framework, we make
-  several key contributions: \textbf{(a)} We establish that
-  measurement clusterization is a generic trait of D-optimal designs
-  with independent Gaussian measurement errors, and prove that
-  introducing correlations among measurement error terms mitigates
-  clusterization. \textbf{(b)} We characterize D-optimal designs as
-  reducing uncertainty across a subset of prior covariance
-  eigenvectors. Finally, \textbf{(c)} We leverage this
-  characterization to argue that measurement clusterization arises as
-  a consequence of the pigeonhole principle: when more measurements
-  are taken than there are locations where the select eigenvectors are
-  large and others are small --- clusterization occurs.
-  %
-  %% In summary, in this study we shed light on an often ignored issue with
-  %% Bayesian D-optimal designs. We characterize such designs and show why
-  %% clusterization arisesit arisestheir inherent properties, propose
-  %% strategies to address clusterization, and provide insights into the
-  %% fundamental attributes underlying measurement design optimization.
-
+  Estimation of parameters in physical processes often demands costly measurements, prompting the pursuit of an optimal measurement strategy.
+Finding such strategy is termed the problem of \emph{optimal experimental design}, abbreviated as optimal design.
+Remarkably, optimal designs can yield tightly clustered measurement locations, leading researchers to fundamentally revise the design problem just to circumvent this issue.
+Some authors introduce error correlation among error terms that are initially independent, while others restrict measurement locations to a finite set of locations.
+While both approaches may prevent clusterization, they also fundamentally alter the optimal design problem.
+
+In this study, we consider Bayesian D-optimal designs, i.e.~designs that maximize the expected Kullback-Leibler divergence between posterior and prior.
+We propose an analytically tractable model for D-optimal designs over Hilbert spaces.
+In this framework, we make several key contributions: \textbf{(a)} We establish that measurement clusterization is a generic trait of D-optimal designs with independent Gaussian measurement errors, and prove that introducing correlations among measurement error terms mitigates clusterization.
+\textbf{(b)} We characterize D-optimal designs as reducing uncertainty across a subset of prior covariance eigenvectors.
+Finally, \textbf{(c)} We leverage this characterization to argue that measurement clusterization arises as a consequence of the pigeonhole principle: when more measurements are taken than there are locations where the select eigenvectors are large and others are small --- clusterization occurs.
  \end{abstract}
 
-%% ** Keywords **
 \begin{keyword}[class=MSC]
@@
 
-%\input{intro}
 \section{Introduction}\label{section:intro}
@@
 
-%% first: blue, second: red, third: green, fourth: black,
-%% fifth:magenta
 \begin{figure}
     \centering
-    \includegraphics[height=0.5\textwidth]{example.pdf}
+    \includegraphics[height=0.5\textwidth]{example.png}
     \caption{Measurement clusterization for optimal designs when
@@
 questions:
-%
 Why does imposing correlations between observations alleviate
 measurement clusterization?
-%
 Is measurement clusterization a generic phenomenon?
-%
 And, most importantly: Why does measurement clusterization occur?
-%
-%Should we aim to avoid measurement clusterization?
-%
-%Is it possible to substitute an optimal clustered design with an
-%equally optimal non-clustered design?
-%Can an optimal clustered design be relpaced with an equally optimal
-%non-clustered design?
 
@@
   generic phenomenon?}
-  %\subsection{An answer for Question \ref{q:generic}: Genericity of measurement clusterization}
-  %% Computer implementation of Lemma \ref{thm:char} for the inverse
-  %% problem outlined in Section \ref{section:how} also generates \(\obs\)
-  %% as a solution to the D-optimal design problem.  Furthermore,
 Randomized numerical simulations of our model give rise to D-optimal
@@
 
-  %% In Section \ref{section:vanishing}, we provide an insightful
-  %% explanation for the optimality of clustered designs when no model
-  %% error is present. We demonstrate that for our model, a D-optimal
-  %% design measures only a small subset of prior eigenvectors, which are
-  %% the prior eigenvectors with largest power spectrum. In real-life
-  %% problems there are limited number of locations where: (a) the
-  %% relevant prior eigenvectors are large, and other prior eigenvectors
-  %% with large power in the prior spectrum are zero. Then measurement
-  %% clusterization is a result of the pigeonhole principle, where there
-  %% are more measurements than measurement locations satisfying (a) and
-  %% (b).
-
-  %% We conjecture that the prevalence of measurement clusterization
-  %% arises due to the ease of discovering clustered designs.
-
-%% \item \label{q:avoid} \textbf{Should we aim to avoid measurement
-%%   clusterization?} Based on the analysis conducted in this study, we
-%%   did not find any compelling reason to explicitly avoid optimal
-%%   clustered designs.
-
-%% \item \label{q:replace} \textbf{Is it possible to substitute an
-%%   optimal clustered design with an equally optimal non-clustered
-%%   design?} In Section \ref{section:vanishing}, we answer this question
-%%   in the affirmative, although we show that numerical experiments
-%%   conducted using our model indicate a strong preference for clustered
-%%   designs.
-
 \end{enumerate}
@@
 
-\begin{restatable}{theorem}{char}\label{thm:char}
+\begin{theorem}\label{thm:char}
   Let:
@@
       \begin{split}
-        \tar(\obs) %:&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
-        % 
-        % 
-        % 
-        &= \frac12 \log \det ( I + \sigma^{-2} \prcov^{1/2} \fwd ^*
+        \tar(\obs) &= \frac12 \log \det ( I + \sigma^{-2} \prcov^{1/2} \fwd ^*
         \obs^* \obs \fwd \prcov^{1/2}), 
@@
     \(\fwd\prcov\fwd^*\) in decreasing order of magnitude.
-    %% \item \(\{\ev_i\}_{i=1}^\infty\) their corresponding eigenvectors.
 \item \(\{\eta_i\}_{i=1}^\infty\) eigenvalues of \(\obs^*\obs\).
@@
     \begin{equation*}
-      \tar(\obs) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i). %= \frac12 \sum_{i=1}^{m} \log (1 + \sigma^{-2}\lambda_i\eta_i).
-    \end{equation*}
-  %% \item 
-  %%   \begin{equation*}
-  %%     k = \argmax \left \{ k:\lambda_k^{-1} < \sigma^{-2}\frac{m}{k} + \frac{1}{k} \sum_{j=1}^{k}
-  %%     \lambda_j^{-1} \right \}.
-  %%   \end{equation*}
+      \tar(\obs) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i). \end{equation*}
 \item
@@
   \end{enumerate}
-%\end{theorem}
-\end{restatable}
+\end{theorem}
 
@@
 AA^t\), where \(A\) has unit norm columns.
-
-%% Finally, in Lemma \ref{lemma:lax} we generalize a lemma for
-%% calculating \(\frac{\der}{\der t} \log \det (I + X(t))\), where
-%% \(X(t)\)is an operator valued function \cite{Lax07}.
 
@@
 many others.
-
-%\input{prelims}
  \section{Preliminaries and Notation}\label{section:prelim}
@@
 
-\begin{equation*}%\label{eq:O}
-  \obs u = (\meas_1(u), \dots, \meas_m(u) )^t \in \R^m,\ u \in \hilo.
+\begin{equation*}\obs u = (\meas_1(u), \dots, \meas_m(u) )^t \in \R^m,\ u \in \hilo.
 \end{equation*}
@@
       \eps)^t ]
-    % 
-    % 
 = \obs \modcov \obs^* + \sigma^2I_m , 
@@
     [\obs \modcov \obs^*]_{ij} & = e_i^t \obs \modcov \obs^* e_j 
-    %
-    %
-    %
-    = \meas_i (\modcov \meas_j).% \text{ (by \eqref{eq:obs*})}.
-  \end{split}
+= \meas_i (\modcov \meas_j).\end{split}
 \end{align}
@@
       \tar( \obs) :&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
-      % 
-      % 
-      % 
 &= \frac12 \log \det 
@@
 
-%% It is important to note that since \(\obs\) is finite-rank,
-%% \(\prcov^{1/2} \fwd ^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}\) is
-%% trace-class.
 \begin{definition}\label{def:d optimality}
@@
     &= \frac12 \log \det \prcov \postcov^{-1} \text{ by \eqref{eq:postcov}}.
-    %%  &= \frac12 \log \det \prcov -\frac12 \log \det \postcov.
 \end{split}
@@
 
-%% \subsection{Notation Summary}\label{subsec:notation}
-%%   Our notation is summarized below. We let:
-%%   \begin{itemize}
-%%   \item \(\hilp, \hilo\) Hilbert spaces.
-%%   \item \(\fwd:\hilp \to \hilo\) a linear compact operator.
-%%   \item \(\pr \sim \mathcal{N}(0, \prcov)\) prior Gaussian measure on
-%%     \(\hilp$, with prior covariance operator \(\prcov:\hilp \to \hilp$.
-%%   \item \(\obs: \hilo \to \mathbb{R}^m\) measurement operator, where \(m
-%%     \in \mathbb{N}\) is the number of measurements taken.
-%%   \item \(\sigma^2 \in \mathbb{R}_{+}\) observation noise variance.
-%%   \item \(\modcov\) model error covariance operator.
-%%     %% \(\data = \obs \fwd \param + \eps$, where \(\eps \in
-%%     %% \mathbb{R}^m\) isiid \(\mathcal{N}(0, \sigma^2)\) noise.
-%%   \item \(\Sigma(\obs) = \obs \modcov \obs^* + \sigma^2I$. 
-%%   \item \(\post\) the posterior measure, with covariance \(\postcov$.
-%%   \item A D-optimality design criterion
-%%     \cite{AlexanderianGloorGhattas14}:
-%%     \begin{align*}
-%%       \begin{split}
-%%         \tar(\obs) :&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
-%%         % 
-%%         % 
-%%         % 
-%%         &= \frac12 \log \det ( I + \prcov^{1/2} \fwd ^* \obs^* \Sigma(\obs)^{-1} \obs
-%%         \fwd \prcov^{1/2}).
-%%       \end{split}
-%%     \end{align*}
-%%   %% \item \(\{\lambda_i\}_{i=1}^\infty\) eigenvalues of \(\fwd\prcov\fwd^*$
-%%   %%   in decreasing order of magnitude.
-%%   %% %% \item \(\{\ev_i\}_{i=1}^\infty\) their corresponding eigenvectors.
-%%   %% \item \(\{\eta_i\}_{i=1}^\infty\) eigenvalues of \(\obs^*\obs$.
-%%   \end{itemize}
-
-%% \subsection{Sequential vs Simultaneous Optimization}\label{subsec:seq vs sim}
-%% From defintion \ref{def:d optimality} we wish to characterize solution(s) of the
-%% following optimization problem for \(\tar$. %%: (\hilo^*)^m \to \R$:
-%% \begin{align}\label{eq:optimization}
-%%   \obs^{\star} := \argmax_{\obs} \tar( \obs ) 
-%%   = \argmax_{\obs} \frac12 \log \det 
-%%   (I + \prcov^{1/2} \fwd^*\obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}),
-%% \end{align}
-%% where \(\obs\) is constrained to some allowed set of observations. We
-%% call this problem ``simultaneous optimization'', since all
-%% observations are decided on simulatneously.
-
-%% For computational reasons, one may prefer to find the best
-%% observations in a sequential manner. Denote
-%% \begin{equation}\label{eq:def obs_k}
-%%   \obs_k := (\meas_1,\dots, \meas_k)^t,  k\leq m.
-%% \end{equation}
-%% Sequential optimal design proceeds as follows. Find \(\meas_1\) by
-%% maximizing \(\tar(\obs_1)$. Then, keeping \(\meas_1\) fixed --- find
-%% \(\meas_2\) as the maximizer of \(\tar(\obs_2)$. Then, find \(\meas_3\) by
-%% keeping \(\meas_1,\meas_2\) fixed and taking \(\meas_3\) as the maximizer
-%% of \(\tar(\obs_3)$. Continue this way until \(\obs_m = \obs\) is
-%% found, where \(m\) is the number of available observations. %% It is
-%% %% important to notice that this scheme does not require actually
-%% %% observing data --- in \eqref{eq:objective} data is averaged out.
-
-%% The analysis in this paper is conducted for the general simultaneous
-%% optimization case. The sequential optimization case is dealt with in
-%% section \ref{subsec:clusterization sequential}. It is important to
-%% note, however that all conclusions we arrive at for the simultaneous
-%% case easily specialize to the sequential case by considering the
-%% posterior as the next sequential step's prior.
-
-%\input{constrained}
  \section{The Constrained Optimization Problem of D-Optimal Design}\label{section:D and grad}
@@
 
-%% Gradients are best thought of as row vectors. This will prove
-%% important in section \ref{subsec:necessary}.
-
-%% \subsection{The gradient of $\tar$}\label{section:objective}
-
 \begin{proposition}\label{prop:tar grad}
@@
   \begin{equation*}
-    %% \delta \tar(\obs) V = \tr{V ( I - \modcov \obs^* \Sigma^{-1}\obs )
-    %%   \fwd \postcov \fwd^* \obs^* \Sigma^{-1}}.
 \nabla \tar(\obs) = ( I - \modcov \obs^* \Sigma^{-1}\obs ) \fwd
@@
       (\obs + \tau V ) \modcov (\obs + \tau V )^*  + \sigma^2I\\
-      % 
-      % 
-      % 
 &= V \modcov \obs^* + \obs \modcov V^*.
@@
     0 &= \frac{\der}{\der \tau} \Big |_{\tau=0} I \\
-    % 
-    % 
-    % 
 &= \frac{\der}{\der \tau} \Big |_{\tau=0}
     \left (\Sigma(\obs+\tau V)^{-1} \Sigma(\obs+\tau V) \right ) \\
-    % 
-    % 
-    % 
 &= \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0} \Sigma+
     \Sigma^{-1} \frac{\der \Sigma(\obs+\tau V)}{\der \tau} \Big |_{\tau=0}\\  
-    %
-    %
-    %
 &= \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0} \Sigma+
     \Sigma^{-1} (V\modcov \obs^* + \obs \modcov V^*). 
-    %\text{, by \eqref{eq:der sig}. }
 \end{align*}
@@
     \end{align}
-  %
 
@@
   By Leibniz (product) rule and \eqref{eq:der sig inv}:
-  % 
 \begin{align}\label{eq:T}
@@
     &= \frac{\der T(\obs + \tau V)}{\der \tau} \Big |_{\tau=0} \\
-    %
-    %
-    %
 &= V^* \Sigma^{-1} \obs 
@@
   \begin{align*}
-    %\begin{split}
 \delta \tar(\obs) V 
       :&= \frac{\der}{\der\tau} \Big |_{\tau=0} \tar(\obs + \tau V) \text{ (Definition \ref{def:var})}\\
-      % 
-      % 
-      % 
 &= \frac12 \frac{\der}{\der \tau} \Big |_{\tau=0} \log \det 
       (I + \prcov^{1/2} \fwd^* T(\obs+\tau V)\fwd \prcov^{1/2} ) \text{ (Theorem \ref{thm:d optimality})} \\
-      % 
-      % 
-      % 
 &= \frac12 \tr{( I + \prcov^{1/2} \fwd^* \obs^* \Sigma^{-1}
@@
         \prcov^{1/2} \fwd^* T(\obs+\tau V) \fwd \prcov^{1/2}}\ \text{ (Lemma \ref{lemma:lax})} \\
-      % 
-      % 
-      % 
 &= \frac12 \ttr\Big \{ \postcov \fwd^* (V^* \Sigma^{-1} \obs 
@@
       + \obs^* \Sigma^{-1} V ) \fwd \Big \}  \text{ (by \eqref{eq:T})} \\
-      %
-      %
-      %
 &= \tr{\postcov \fwd^* ( \obs^* \Sigma^{-1} V -
       \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs ) \fwd} \\
-      %
-      %
-      % 
 &= \tr{\postcov \fwd^* \obs^* \Sigma^{-1} V 
       ( I - \modcov \obs^* \Sigma^{-1}\obs ) \fwd} \\
-      % 
-      %
-      %
 &= \tr{V ( I - \modcov \obs^* \Sigma^{-1}\obs )
       \fwd \postcov \fwd^* \obs^* \Sigma^{-1}}.
-    %\end{split}
 \end{align*} 
@@
   \end{equation*}
-  %% This way, $V$ has the same $j$th entry as $\obs$ while the rest
-  %% are set to zero.
 We now calculate the variation of $\tar$ at $\obs$ in the direction
@@
     &= \tr{V ( I - \modcov \obs^*\Sigma^{-1}\obs) \tmp \obs^* \Sigma^{-1}} \\
-    % 
-    %
-    %
 &= \tr{e_je_j^t \obs ( I - \modcov \obs^*\Sigma^{-1}\obs) \tmp \obs^* \Sigma^{-1}} \\
-    %
-    % 
-    %
 &= e_j^t \obs ( I - \modcov \obs^*\Sigma^{-1}\obs) \tmp \obs^* \Sigma^{-1}e_j \\
-    %
-    % 
-    %
 &= e_j^t ( I - \obs \modcov \obs^*\Sigma^{-1})\obs \tmp \obs^* \Sigma^{-1}e_j \\  
-    % 
-    %
-    %
 &=  e_j^t(\Sigma-\obs \modcov \obs^*) \Sigma^{-1}\obs \tmp \obs^* \Sigma^{-1}e_j \\
-    %
-    %
-    %
 &=\sigma^2 e_j^t \Sigma^{-1}\obs \tmp \obs^* \Sigma^{-1}e_j
     \text{ by \eqref{eq:Sigma} }\\
-    %
-    % 
-    %
 &=\sigma^2 e_j^t \Sigma^{-1}\obs \fwd \postcov \fwd^* \obs^* \Sigma^{-1}e_j.
@@
   \begin{equation*}
-    %% \delta \phi_j(\obs)V = \tr{V \obs^* e_je_j^t}.
 \nabla \phi_j(\obs) = \obs^* e_je_j^t.
@@
     ( \|(\obs + \tau V)^*e_j \|_{\hilp}^2 - \|\obs ^*e_j \|_{\hilp}^2  ) \\
-    %
-    %
-    %
 &= \frac12\lim_{\tau \to 0}\tau^{-1}
@@
     \langle \obs^*e_j, \obs^*e_j \rangle_{\hilp} ) \\
-    % 
-    % 
-    %
 &= \frac12\lim_{\tau \to 0}\tau^{-1}
@@
     +\tau^2 \langle V^*e_j, V^*e_j \rangle_{\hilp} ) \\
-    %
-    %
-    % 
 &= \langle \obs^*e_j,V^*e_j \rangle_{\hilp} \\
-    %
-    %
-    % 
 &= \langle V \obs^*e_j,e_j \rangle_{\R^m} \\
-    %
-    %
-    %
 &= e_j^t V \obs^* e_j \\
-    % 
-    %
-    %
 &= \tr{V \obs^* e_je_j^t}.
@@
 \end{proof}
-
-%% The same arguments justifying \eqref{eq:tar grad} hold here, and thus:
-
-%% \begin{align}\label{eq:grad constraints}
-%% \nabla \phi_j(\obs) = \obs^* e_j e_j^t = \meas_j e_j^t , j=1,\dots,m,
-%% \end{align}
-%% where $\nabla \phi_j(\obs) \in \hilo^m$. As noted at the end of
-%% Section \ref{section:objective},
-
-%% The gradients $\nabla \tar(\obs)$ and $\nabla \phi_j(\obs)$ are best
-%% thought of as row vectors.
 
@@
 
-%\input{nonvanishing}
-%% Imposing correlations between observations alleviate measurement clusterization
  \section{Answer to Question \ref{q:mitigate}: Model error mitigates clusterization}\label{section:non vanishing}
@@
     \end{bmatrix}\\
-    %
-    %
 \end{align*}
@@
   \end{align}
-  %
 and denote:
-  %
 \begin{align}\label{eq:M def}
@@
     \M(\obs) &= \prcov^{1/2} \fwd^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2} \\
-    %
-    %
-    %
 &= \prcov^{1/2} \fwd^* \obs^* \left \{
@@
     \right \} \obs \fwd \prcov^{1/2} \\
-    %
-    %
-    %
 &= \M (\obsm) + (c -w^t \Sigmam^{-1} w )^{-1}
@@
   \end{align*}
-  %
 Now, denote:
-  %
 \begin{align}\label{eq:u}
@@
       \end{bmatrix} \\
-      %
-      %
-      %
 & = (c -w^t \Sigmam^{-1} w )^{-1/2} ( \prcov^{1/2}\fwd^* \obsm^* \Sigmam^{-1} \obsm  \modcov \meas_m - \prcov^{1/2} \fwd^* \meas_m )\\
-      %
-      %
-      %
 u^* :&=  (c -w^t \Sigmam^{-1} w )^{-1/2} (\meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \fwd \prcov^{1/2} - \meas_m \fwd \prcov^{1/2} ),
@@
   \end{align}
-  %
 so that
-  %
 \begin{equation}\label{eq:M plus I}
@@
   \end{equation}
-  %
 Note that
@@
   A$, statement and proof in the supplementary material):
-  %
 \begin{align}\label{eq:diffs}
@@
       \tar( \obs ) - \tar( \obsm )
-      %
-      %
-      %
 &= \frac12 \log \left (\det \big ( I + \M ( \obs ) \big ) / \det \big ( I + \M (\obsm) \big ) \right )\\
-      %
-      %
-      %
 &= \frac12  \log \left (\det \left ( I + \M(\obsm) + uu^* \right ) / \det \big ( I + \M (\obsm) \big )\right ) \\
-      %
-      %
-      %
 &= \frac12 \log \left ( 1 + \left \langle \left ( I+\M(\obsm) \right )^{-1} u, u  \right \rangle \right ).
@@
       }\\
-      %
-      %
-      %
 &= 
@@
   \end{equation}
-  %
 Using \eqref{eq:observation}:
@@
       &= \meas_m \modcov \obsm^* ( e_j - \sigma^2 \Sigmam^{-1} e_j )\\
-      %
-      %
-      %
 &= \meas_m \modcov \meas_j - \sigma^2 \meas_m \modcov \obsm^* \Sigmam^{-1}e_j \\
-      %
-      %
-      %
 &= \meas_m \modcov \meas_j -\sigma^2 (e_j - \sigma^2 \Sigmam^{-1}e_j)^t e_j \\
-      %
-      %
-      %
 &= \meas_m \modcov \meas_m -\sigma^2 + \sigma^4 e_j^t\Sigmam^{-1}e_j.
@@
       &= \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m - \meas_m \\
-      %
-      %
-      %
 &= \obsm^* (e_j - \sigma^2 \Sigmam^{-1} e_j) -\meas_j \\ 
-      %
-      %
-      %
 &= -\sigma^2 \obsm^* \Sigma^{-1}e_j. 
@@
   \end{align}
-  %
 Now, we substitute \eqref{eq:enumerator} and \eqref{eq:denominator}
   to the enumerator and denominator of \eqref{eq:conclusion}:
-  %
 \begin{align*}
@@
     \right ) \\
-    %
-    %
-    %
 &= \log \left ( 1 + \frac{\sigma^4
@@
     \right ) \\
-    %
-    %
-    %
 &= \log \left ( 1 + \frac{\sigma^2
@@
     e_j, \obsm^* \Sigmam^{-1}e_j \rangle 
-  %
-  %
 = \langle \fwd \postcovm \fwd^* u, u \rangle
@@
 non-zero model error $\modcov$ --- answering Question \ref{q:mitigate}
-posed in the Introduction.% Section \ref{section:intro}.
-
-%\input{vanishing}
-\section{D-Optimal Designs Without Model Error}\label{section:vanishing}
+posed in the Introduction. \section{D-Optimal Designs Without Model Error}\label{section:vanishing}
 Our goal in this section is to prove Theorem \ref{thm:char} which
@@
 to \eqref{eq:eigenproblem} as a \emph{nonlinear} eigenvalue problem.
-%% Proposition \ref{prop:twice woodbury} and Lemma \ref{lemma:sim
-%% diag} required
 
@@
       \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* 
-      %
-      %
 = \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1},
@@
     &\ \ \ \ \ = \fwd ( \prcov - \prcov \fwd^* ( \sigma^2(\obs^*\obs + \eps I)^{-1} + \fwd \prcov \fwd^* )^{-1} \fwd \prcov ) \fwd^*. \\
-      %
-      %
-      %% &= X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X
-  %%   \end{split}
-    %% \end{align*}
 \end{split}
@@
       \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* &= X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X \\
-      %
-      %
-      %
 &= (X^{-1} + \sigma^{-2}Y_{\eps})^{-1} \\
-      %
-      %
-      %
 &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \eps I))^{-1}.
@@
       \obs^* \Xi &= \sigma^{-2}\fwd \postcov \fwd^* \obs^*  \text{ (by \eqref{eq:eigenproblem})}\\
-      %
-      %
-      %
 &= \sigma^{-2} \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* \obs^*  \text{ (by \eqref{eq:postcov})} \\
-      %
-      %
-      %
 &= \sigma^{-2} \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1} \obs^* \text{ (by Proposition \ref{prop:twice woodbury})}.
@@
       \tar(\obs) &= \frac12\log \det \left (I + \sigma^{-2} \prcov^{1/2} \fwd ^* \obs^*
-      \obs \fwd \prcov^{1/2}\right )\\% \text{ (by definition)}\\
-      %
-      &= \frac12 \log \det \left (I + \sigma^{-2} \obs^* \obs \fwd
+      \obs \fwd \prcov^{1/2}\right )\\&= \frac12 \log \det \left (I + \sigma^{-2} \obs^* \obs \fwd
       \prcov\fwd^* \right ) \text{ (Sylvester's Determinant
       Theorem)}\\
-      %
-      %
-      %
 &=\frac12 \log \prod_{i=1}^{\infty} ( 1 + \sigma^{-2} \lambda_i\eta_i ) \text{ (Proposition \ref{prop:same ev})} \\
-      %
-      %
-      %
-      %&=\frac12 \log \left ( \prod_{i=1}^{k} ( \lambda_i^{-1} + \sigma^{-2} \eta_i )\prod_{i=1}^{k} \lambda_i \right )
-      %
-      %
-      %
 &=\frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i). 
-      %
-      %
-      %% &=\frac12 \sum_{i=1}^{k} \log(\sigma^2\lambda_i^{-1} + \eta_i) - k\log \sigma + \frac12\sum_{i=1}^k \log \lambda_i.
 \end{split}
@@
        \frac12 \frac{\sigma^{-2}\lambda_i}{1 + \sigma^{-2} \lambda_i\eta_i} = \frac12 \frac{1}{\sigma^{2}\lambda_i^{-1} + \eta_i} \\
-      %
-      %
-      %
 \frac{\partial\Phi}{\partial \eta_i} &= 1 \\
-      %
-      %
-      %
 \frac{\partial \Omega_j}{\partial \eta_i} &= -\delta_{ij}      
@@
       -\frac12 \frac{1}{\sigma^{2}\lambda_i^{-1} + \eta_i} + \alpha - \beta_i  &= 0 \\
-      %
-      %
-      %
 \eta_i &\geq 0\\
-      %
-      %
-      %
 \beta_i &\geq 0\\
-      %
-      %
-      %
 \beta_i \eta_i &= 0\\
-      %
-      %
-      %
 \sum_{i=1}^m \eta_i &= m 
@@
       \beta_i &= 0\\
-      %
-      %
-      %
-      \sigma^{2}\lambda_i^{-1} + \eta_i  &= \frac{1}{2\alpha}%  \text{ for }  1 \leq i \leq k.\\
-      %
-      %
-      %
-      %% \sigma^{-2} \lambda_i &= 2\alpha - 2\beta_i, i \in B \\
-    \end{split}
+\sigma^{2}\lambda_i^{-1} + \eta_i  &= \frac{1}{2\alpha}\end{split}
   \end{align*}
@@
   \frac{1}{2\alpha} = \frac{m}{k} + \frac{\sigma^2}{k} \sum_{j\in A} \lambda_j^{-1}.
-  %\Rightarrow \alpha = \left ( \frac{2m}{k} + \frac{2\sigma^2}{k}
-  %\sum_{i\in A} \lambda_i^{-1} \right )^{-1}.
 $$
@@
       \eta_i &= \frac{1}{2\alpha} - \sigma^2 \lambda_i^{-1}\\
-      %
-      %
-      %
-      &=\frac{m}{k} + \frac{\sigma^2}{k} \sum_{j\in A} \lambda_j^{-1} - \sigma^2 \lambda_i^{-1}.%\\
-      %
-      %
-      %
-      %% &= \frac{m}{k} + \frac{\sigma^2}{k} \sum_{j \in A, j\neq i} \lambda_j^{-1} - \frac{k-1}{k}\sigma^2 \lambda_i^{-1}\\
-      %% ,\ i \in A.
-    \end{split}
+&=\frac{m}{k} + \frac{\sigma^2}{k} \sum_{j\in A} \lambda_j^{-1} - \sigma^2 \lambda_i^{-1}.\end{split}
   \end{align}
@@
 
-\char*
-
+\begin{theorem}\label{thm:char}
+  Let:
+  \begin{itemize}
+    \item The D-optimality design criterion
+    \cite{AlexanderianGloorGhattas14}:
+    \begin{align*}
+      \begin{split}
+        \tar(\obs) &= \frac12 \log \det ( I + \sigma^{-2} \prcov^{1/2} \fwd ^*
+        \obs^* \obs \fwd \prcov^{1/2}), 
+      \end{split}
+    \end{align*}
+  \item \(\obs\) a D-optimal design operator
+    \begin{equation*}
+      \obs = \argmax_{\|\meas_j\| = 1, j=1,\dots,m}\tar(\obs),
+    \end{equation*}
+  \item \(\{\lambda_i\}_{i=1}^\infty\) eigenvalues of
+    \(\fwd\prcov\fwd^*\) in decreasing order of magnitude.
+\item \(\{\eta_i\}_{i=1}^\infty\) eigenvalues of \(\obs^*\obs\).
+
+  \end{itemize}
+
+  Then:
+  \begin{enumerate}
+  \item  \(\tr{\obs^*\obs} = m\).
+  \item \(\obs^*\obs\) and \(\fwd\prcov\fwd^*\) are simultaneously
+    diagonalizable.
+  \item \(k := \rank \obs^*\obs \leq m\) and
+    \begin{equation*}
+      \tar(\obs) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i). \end{equation*}
+\item
+    \begin{equation*}
+        \eta_i = \begin{cases}
+          \frac{m}{k} - \sigma^2 \lambda_i^{-1} + \sigma^2 \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1} & 1 \leq i \leq k \\
+          0 & i > k 
+        \end{cases}.
+    \end{equation*}
+  \item The covariance of the pushforwad \(\fwd_{*} \post\) is \(\left
+    ( (\fwd \prcov \fwd^*)^{-1} + \sigma^{-2} \obs^*\obs \right
+    )^{-1}\) and its eigenvalues are
+    \begin{equation*}
+      \theta_i =
+      \begin{cases}
+        \left(\frac{\sum_{j=1}^k \lambda_j^{-1} + \sigma^{-2}m}{k} \right )^{-1} & i \leq k \\
+        \lambda_i &  i > k 
+      \end{cases}
+    \end{equation*}
+  \end{enumerate}
+\end{theorem}
 \begin{proof}
@@
         legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
-        reverse legend=false, % set to false to get correct display, but I'd like to have this true
-        xticklabels from table={\optimalvsnot}{Label},
+        reverse legend=false, xticklabels from table={\optimalvsnot}{Label},
         xticklabel style={text width=2cm,align=center},
@@
 corresponds to a graduated lab cylinder.
-%% (Figure \ref{fig:cylinder})
 Each cylinder $i$ is filled, a-priori, with green liquid of
@@
 such a procedure is illustrated in Fig.~\ref{fig:optimal vs not}.
-%% \begin{figure}%{r}{0.25\textwidth} 
-%%     \centering
-%%     \includegraphics[width=5cm, height=5cm]{cylinders.jpg}
-%%     \caption{Three graduated lab cylinders, corresponding to three
-%%       eigenvectors. Prior eigenvalues not shown.}
-%%     \label{fig:cylinder}
-%% \end{figure}
-
-%\input{why}
+
  \section{Answer to Question \ref{q:why}: The cause of clusterization}\label{section:why}
@@
 \bibliographystyle{ba}
-\bibliography{/home/yair/projects/bibtex.bib}
+\bibliography{../../lib.bib}
 
 \begin{acks}[Acknowledgments]
-
-%\input{ack}
 This study is a result of research I started during my PhD studies
@@
 
-%\input{funding}
 This research was supported in part by an appointment with the
@@
 NSF, ORAU/ORISE, or DOE.
-% This work was also supported by The Raymond and Beverly Sackler
-% Post-Doctoral Scholarship.
-
  \end{acks}

### 2024-07-22 latex/intro.tex
Compilesgit statusgit status
Fixed intro and prelims

--- 2024-07-21 latex/intro.tex
@@
 The practice of experimental design entails specifying every detail of
-running an experiment \cite{chaloner}: at its simplest form,
+running an experiment \cite{chaloner1995}: at its simplest form,
 experimental design involves choosing a set of measurements to be
-taken. For example, in a medical experiment, experimental design may
-involve choosing participants to whom some drug will be administered,
-and measuring their blood pressure. Experimental design also arises in
-the process of searching for oil. Since measuremnts are expensive and
-require digging holes in the ground, only a limited number of
-measuremnts can be taken, hence carefully choosing measuremt locations
-is crucial.
+taken. For example, consider the process of searching for oil:
+measuremnts are expensive and require digging holes in the ground, so
+only a limited number of measuremnts can be taken, hence carefully
+choosing measuremt locations is crucial.
 
 Specifiyng the right set of measurements holds particular significance
-when solving an \emph{inverse problem} --- i.e. making inference of
+when solving an \emph{inverse problem} --- i.e.~making inference of
 the physical world utilizing a physical model a phenomenon of interest
-\cite{tarantola,somersalo}. In an inverse problem, we seek to infer
-physical quantities, based on observations and a model of the physical
-world. Models are typically phrased in the language of ordinary or
-partial differential equations.
+\cite{tarantola2005,kaipio2005}. In an inverse problem, we seek to
+infer physical quantities, based on observations and a model of the
+physical world. Models are typically phrased in the language of
+ordinary or partial differential equations.
 
-Inverse problems are ubiquitous in science and engineering, and we
-will now give several illustrative examples. In electrical impedance
-tomography (EIT) we seek to infer the impedance of the inside of a
-human body from observations of electrical current impedance measured
-at specific electrode locations on the skin, leveraging known
-equations of electric current flow
-\cite{horesh2010impedance}. Magnetic Resonance Imaging (MRI) also
-entails solving an inverse problem, where radio-frequency pulses are
-sent through the human body in the presence of a strong magnetic field
-and the response of one's body contents to such perutrbations is
-measured, allowing a radiologist to view the internals of a body in a
-noninvasive manner \cite{horesh2008mri}. Yet another example of an
-inverse problem arises in oil and gas exploration, where acoustic
-waves are sent through "boreholes" --- deep cylindrical wells drilled
-into the ground --- and data of travel and return times of said waves
-are recorded. Travel and return time is influenced by the properties
-of the subsurface materials, such as density, elasticity, and the
-presence of fluids or voids, combining travel data with a geophisical
-models of contents of the earth's crust to facilitate reconstructing
-and inferring the structure of the subsurface in a process called
-\emph{borehole tomography} \cite{horesh2008borehole}. Inverse problems
-also arise in many other areas of seismology and geology
-\cite{rabinowitz1990, steinberg1995} and medical imaging
-\cite{tarantola}. While in many inverse problems the goal is to infer
-some parameter, in all of the above examples and in the formulation we
-consider here, the goal is to infer some \emph{function} over $
-\Omega$, where (usually) \(\Omega \subseteq \mathbb{R}^d, d=1,2,3\) is
-the spatial domain of interest.
+Inverse problems are ubiquitous in science and engineering. For
+example, in electrical impedance tomography (EIT) we seek to infer the
+structure of the inside of a human body. Inference is conducted based
+on observations of electrical current impedance measured at specific
+electrode locations on the skin, leveraging known equations of
+electric current flow \cite{horesh2010impedance}. Magnetic Resonance
+Imaging (MRI) also entails solving an inverse problem. There,
+radio-frequency pulses are sent through the human body in the presence
+of a strong magnetic field. The response of one's body contents to
+these radio-frequency pulses is measured, allowing a radiologist to
+view the internals of a body in a noninvasive manner
+\cite{horesh2008mri}. An inverse problem also arises in oil and gas
+exploration. Acoustic waves are sent through "boreholes" --- deep
+cylindrical wells drilled into the ground. Data of travel and return
+times of the acoustic waves are recorded. Wave travel and return time
+is influenced by the properties of the subsurface materials, such as
+density, elasticity, and the presence of fluids or voids. Combining
+travel data with a geophisical models of contents of the earth's crust
+facilitates reconstructing and inferring the structure of the
+subsurface in a process called \emph{borehole tomography}
+\cite{horesh2008borehole}. Inverse problems also arise in many other
+areas of seismology and geology \cite{rabinowitz1990, steinberg1995}
+and medical imaging \cite{tarantola2005}. While in many inverse
+problems the goal is to infer some parameter, in all of the above
+examples and in the formulation we consider in this study, the goal is
+to infer some \emph{function} over $ \Omega$, where (usually) \(\Omega
+\subseteq \mathbb{R}^d, d=1,2,3\) is a spatial domain of interest.
 
-%% In EIT, electrode locations determine the quality of the
-%% reconstructed impedance field, while in MRI utilizng certain
-%% radio-frequency pulses may give more accurate information than
-%% others. In borehole tomogrpahy, only several boreholes could be dug
-%% and the choice of their locations is also crucial for oil and gas
-%% discovery. Thus, the selection of optimal measurements, referred to as
-%% the problem of \emph{optimal design}, becomes crucial.
-%% in a specific location (in a typical
-%% medical application to measure inverventions on, . These experimental
-%% units may be individual patients, but also Although common sense is
-%% useful in this process, it is often the case that naive intuition
-%% results in experiments that do not generate the maximal amount of
-%% information possible. To this end
+\subsection{D-optimal Designs}\label{subsec:D}
+In all of the examples of inverse problems we have seen, as well as in
+many other applications, only a limited number of measurements are
+allowed. For example, in borehole tomography, a measurement involves
+drilling a deep well in the ground --- a very costly endeavor. Even in
+a seemingly simpler application such as MRI, a patient cannot occupy
+the MRI machine for too long, hence only a limited number of
+measurements can be taken. Therefore, measurements should be chosen to
+most benefit the user. Typically, a user will consider some utility,
+called a \emph{design criterion} and take measurements that maximize
+this utility. Two of the most widely recognized and extensively
+studied design criteria are A- and D-optimality.
 
-For concreteness, let us consider a toy model of an inverse problem:
-the \emph{heat equation} in one dimension (1D heat equation
-henceforth, see Section \ref{subsec:1d heat} for more details). The
-heat equation is a partial differential equation compose of three
-equations: eqs.~\eqref{eq:heat1}, \eqref{eq:heat2}, and
-\eqref{eq:heat3}. Readers less familiar with partial differential
-equations should think of the heat equation as follows: before us
-there is a perfectly insulated metal rod. At each point on the rod the
-temperature is different and unknown, and we call this temperature
-distribution the \emph{initial condition}. We wait for a short time
-$T$ and let the heat dissipate a bit inside the rod. The final heat
-distribution in the rod is dictated by the following set of equations
-which defines the 1D heat equation:
+In this study, we focus on the Bayesian D-optimality
+criterion. Bayesian D-optimality carries a simple and intuitive
+meaning: measurements chosen according to the D-optimality criterion
+maximize the expected Kullback-Liebler divergence from posterior to
+prior \cite{chaloner1995, AlexanderianGloorGhattas14,
+  CoverThomas91}. Utilizing KL divergence for quantifying distances
+between distributions is standard practice and we will not motivate it
+further here. For a discrete parameter $\param$ and data $\data$, the
+KL divergence is defined as
+\begin{equation}\label{eq:basic KL}
+  D_{KL}\left (\Pr(\param|\data)||\Pr(\param)\right ) =  \sum_{\param} \log
+  \frac{\Pr(\param|\data)}{\Pr(\param)} \Pr(\param|\data). 
+\end{equation}
+Of course, data $\data$ is not known before the experiment is
+conducted, so we average over it to define the D-optimality criterion
+as:
+\begin{equation*}
+  \mathbb{E}_{\data}\left [ D_{KL}\left (\Pr(\param|\data)||\Pr(\param)\right ) \right ]
+\end{equation*}
+We refer to a set of measuremnts that maximizes the D-optimality
+criterion as a \emph{D-optimal design}.
+
+For a linear model with Gaussian prior in finite dimensions, a
+D-optimal design minimizes the determinant of the posterior covariance
+\cite{chaloner1995}. In Section \ref{subsec:D optimal design} we give
+a more general definition of the D-optimality criterion that applies
+to arbitrary measures.
+
+\subsection{A toy model: the 1D heat equation}
+For concreteness, let us now consider a toy model of a physical process
+for which we wish to solve an inverse problem: a partial differential
+equation known as the \emph{heat equation} in one dimension (1D heat
+equation henceforth, see details below). Readers less familiar with
+partial differential equations can think of the following setup:
+before us there is a perfectly insulated metal rod. At each point on
+the rod the temperature is different and unknown, and we call this
+temperature distribution the \emph{initial condition}. We wait for a
+short time $T$ and let the heat dissipate a bit inside the rod. The
+heat equation determines the heat distribution inside the rod at time
+$t=T$.
+
+The full time evolution of heat in the rod $\Omega=[0,1]$ is formally
+described by the following three equations:
 \begin{subequations}%\label{eq:heat equation}
   \begin{alignat}{2}
-    u_t &= \alpha \Delta u &&\qquad \text{in } [0,1] \times [0,\infty), \label{eq:heat1}\\
+    u_t &= \Delta u &&\qquad \text{in } [0,1] \times [0,\infty), \label{eq:heat1}\\
     u &= 0 &&\qquad \text{on } \{0, 1\} \times [0,\infty), \label{eq:heat2}\\
-    u &= u_0 &&\qquad \text{on }[0,1] \times \{0\}. \label{eq:heat3}\\
+    u &= u_0 &&\qquad \text{on }[0,1] \times \{0\}. \label{eq:heat3}
   \end{alignat}
 \end{subequations}
-
 As time passes, heat disspates across the rod as hotter regions,
@@
 maximas, and the reverse happens at local minima. Eq.~\eqref{eq:heat2}
-describes the \emph{boundary condition}, i.e.~the interaction of the
-heat inside the rod with the boundary of the domain $\Omega =
-[0,1]$. Here we consider a \emph{homogeneous Dirichlet boundary
-condition} which is best understood as an absorbing boundary: heat
-that interacts with the boundary is immediately
-absorbed. Eq.~\eqref{eq:heat3} describes the \emph{initial condition}
-of the heat, i.e.~the heat distribution at time $t=0$. Thus, we have
-described the \emph{forward} problem: how heat evolves in time,
-starting at $t=0$ up until $t=T$.
+describes the \emph{boundary condition}, which dictates how heat
+interacts with the outside world at the two edges of the rod
+$\{0,1\}$. Specifically, eq.~\eqref{eq:heat2} implements an absorbing
+boundary, i.e.~heat that interacts with the boundary is immediately
+absorbed and disappears. This type of boundary condition is known as a
+\emph{homogeneous Dirichlet boundary condition}. Eq.~\eqref{eq:heat3}
+describes the \emph{initial conditions}, i.e.~the heat distribution
+inside the rod at time $t=0$.
 
-In the inverse problem of the 1D heat equation, our goal is to infer
-the initial condition $u_0 = u(\cdot, 0)$ from observations of the
-final condition $u(\cdot, T)$. We are not able to measure $u(\cdot,
-T)$ at every point in the domain $\Omega$. Rather, we take some number
-of temperature measurements on the rod once our waiting time $T$ has
-passed and try to infer $u(\cdot, 0)$ from these measurements.
+In the \emph{inverse problem of the 1D heat equation}, our goal is to
+infer the initial condition $u_0 = u(\cdot, 0)$ from observations of
+the final condition $u(\cdot, T)$. However, we are not able to measure
+$u(\cdot, T)$ at every point in the domain $\Omega$. Rather, we take
+some number of temperature measurements on the rod once our waiting
+time $T$ has passed and try to infer $u(\cdot, 0)$ from these
+measurements.
 
@@
 implement such regularization via a fully Bayesian formulation of the
-inverse problem (see Section \ref{section:prelims} for details).
+inverse problem, see Section \ref{section:prelims} for details.
 
-In all of the examples of inverse problems we have seen, as well as in
-many other applications, measurements should be chosen to most benefit
-the user. Hence, the user may consider some utility, called a
-\emph{design criterion} and take measurements that maximize this
-utility. Two of the most widely recognized and extensively studied
-design criteria are A- and D-optimality. We focus on the Bayesian
-D-optimality criterion, which carries a simple and intuitive meaning:
-measurements chosen according to the D-optimality criterion maximize
-the expected Kullback-Liebler divergence from posterior to prior
-\cite{Chaloner1995, AlexanderianGloorGhattas14}, see Section
-\ref{subsec:D optimal design} for a precise definition. We refer to a
-set of measuremnts that maximizes the Bayesian D-optimality criterion
-a \emph{D-optimal design}.
-
+\subsection{Measurement Clusterization}
 Surprisingly, A- and D-optimal designs for inverse problems have been
@@
   %% as a solution to the D-optimal design problem.  Furthermore,
-  Randomized numerical simulations of our model give rise to D-optimal
-  designs that exhibit clusterization more than 95\% of the time (see
-  Section \ref{subsec:lemma sims} for details). Given the genericity
-  of our model, we expect measurement clusterization to be a generic
-  and ubiquitous phenomenon.
+  We give two complementing answers to this question. First, from a
+  theoretical perspective, we show that clusterization barely depends
+  on the specifics of the prior, so long as it is Gaussian. We show
+  that the structure of the decay of the eigenvalues of the
+  pushforward prior covariance determine clusterization. The decay of
+  these eigenvalues depends on how ill-posed the problem at hand is
+  and does not depend much on the prior. see Theorem \ref{thm:char}
+  and the Discussion for details. From a numerical experiments
+  perspective, simulations of our model give rise to D-optimal designs
+  that exhibit clusterization more than 95\% of the time (see Section
+  \ref{subsec:lemma sims} for details). Given the genericity of our
+  model, we expect measurement clusterization to be a generic and
+  ubiquitous phenomenon.
 
@@
 
-The implication of our answer to Question \ref{q:generic} is that
-encountering clusterization should be expected in many different
-inverse problems across many different scientific fields and
-disciplines. Thus, researchers that encounter clusterization should
-not be surprised or wary. Then, in our answer to Question \ref{q:why},
-we explain what is the cause of clusterization. As it appears, the
-cause is universal: a D-optimal design reduces uncertainty for a
-select set of prior covariance eigenvectors --- those with the most
-prior uncertainty, i.e.~those the practitioner cares about the most!
+Our results have several implications. Our answer to Question
+\ref{q:generic} implies that encountering clusterization should be
+expected in many different problems across many different scientific
+fields. Thus, researchers that encounter clusterization should not be
+surprised or wary. In Our answer to Question \ref{q:why}, we explain
+what is the cause of clusterization. As it appears, the cause is
+generic: a D-optimal design reduces uncertainty for a select set of
+prior covariance eigenvectors --- those with the most prior
+uncertainty, i.e.~those the practitioner cares about the most!
 Consequently, practitioners should not try to avoid measurement
-clusterization. Rather, practitioners should should take repeated
+clusterization. Rather, practitioners should take repeated
 measurements (e.g.~in MRI and borehole tomography), increase apparatus
@@
 In Theorem \ref{thm:char}, we also show that D-optimal designs are
-best understood in the sapce of \emph{observations}. Again, a precise
-statement could be found in Section \ref{section:vanishing}.  This
-result is in accordance with previous work by \cite{koval2020}, who
-showed that A-optimal designs are best constructed in the space of
-observations.
+best understood in the sapce of \emph{observations}, see Section
+\ref{section:vanishing} for a precise statement. This is in accordance
+with previous work by \cite{koval2020}, who showed that A-optimal
+designs are best constructed in the space of observations.
 
@@
 
-Finally, we must stress that we do not suggest that clustered designs
-should be abandoned, nor do we support the use of clustered designs in
-any circumstances. We merely investigate an intriguing phenomenon of
-clusterization and seek a deeper understanding for it. Researchers in
-other disciplinesis not viewed as a problem. For example, \cite{lozan}
-suggest in the introduction section that repeated measurements are
-necessary. Our results align with that common wisdom and settle the
-seemed contradiction.
+%% Finally, we must stress that we do not advocate any particular
+%% solution to clusterization. We do not suggest that clustered designs
+%% should be abandoned, nor do we support the use of clustered designs in
+%% any circumstances. Our goal is to investigate an intriguing phenomenon
+%% of clusterization and seek a deeper understanding for it. Different
+%% researchers in different disciplines have differenot viewed as a problem. For
+%% example, \cite{lozan} suggest in the introduction section that
+%% repeated measurements are necessary. Our results align with that
+%% common wisdom and settle the seemed contradiction.
 

### 2024-07-22 latex/prelims.tex
before diff
Fixed intro and prelims

@@
 Here we present the basics of Bayesian inverse problems over hilbert
-spaces. Since we are ultimately interested in inferring a function over some
-domain, we will keep in mind that these Hilbert spaces should really
-be thought of as function spaces. The theoretical foundations for
-inverse problems over function spaces can be found in \cite{Stuart10}
-and will be only briefly reviewed here.
+spaces. Since we are ultimately interested in inferring a function
+over some domain, we will keep in mind that these Hilbert spaces
+should really be thought of as function spaces. A deeper treatment of
+the foundations for inverse problems over function spaces can be found
+in \cite{Stuart10}.
 
 \subsection{Forward Problems}\label{subsec:abstract OED}
-In this section we give definitions and notations for forward probles,
+In this section we give definitions and notations for forward problems,
 which are an essential part of the inverse problems we discuss
@@
 heat, the forward operator is determined by the time evolution of the
-1D heat equation \eqref{eq:heat1}, \eqref{eq:heat2} and
-\eqref{eq:heat3}.
+1D heat equation \eqref{eq:heat1} and \eqref{eq:heat2}.
 
@@
 point measurements $\delta_x$ as suggested by the formulation of the
-1D heat equation. We don't. This entire study could be formulated for
-Banach spaces where point measurements are allowed. We avoid this
-technical difficulty because: (1) it does not add much to the study,
-(2) it is beyond the technical abilities of the author}. Entries
-\(\meas_j, j=1,\dots,m\) of the measurement operator \(\obs\) are
-called \emph{measurements}:
+1D heat equation with workin in Hilbert spaces. We don't. This entire
+study could be formulated for Banach spaces where point measurements
+are allowed. We avoid this technical difficulty because: (1) it does
+not add much to the study, (2) it is beyond the technical abilities of
+the author}. Entries \(\meas_j, j=1,\dots,m\) of the measurement
+operator \(\obs\) are called \emph{measurements}:
 
@@
 error terms: Spatially correlated model error \(\eps' \sim
-\normal(0,\modcov)\) with \(\modcov\) a covariance operator. Observation
-error is denoted \(\eps \sim \normal(0, \sigma^2 I_m)\), with \(I_m \in
-\mathbb{R}^{m \times m}\) the identity. Both error terms and the prior
-are assumed independent of each other. Thus, data is acquired via
+\normal(0,\modcov)\) with \(\modcov\) a covariance operator; and
+observation error denoted \(\eps \sim \normal(0, \sigma^2 I_m)\), with
+\(I_m \in \mathbb{R}^{m \times m}\) the identity. Both error terms and
+the prior (see below) are assumed independent of each other. Thus,
+data is acquired via
 \begin{align}\label{eq:inverse problem}
@@
 Taking \(\modcov = 0\) is a common practice
-\cite{tarantola2005,Kaipio2005,Vogel02} and then \(\Sigma =
+\cite{tarantola2005,kaipio2005,Vogel02} and then \(\Sigma =
 \sigma^2I_m\) is a scalar matrix which does not depend on \(\obs\).
@@
 transported to the observation space via the forward operaror $\fwd u
-\in \hilo$, how observations are generated via $\obs \fwd u$ and how
-observations and noise give rise to data $\data$. It is time to
-formulate the process of inferring the parameter as a Bayesian inverse
-problem. The first ingredient we need is a prior.
+\in \hilo$, how observations are generated from a parameter via $\obs
+\fwd u$ and how observations and noise give rise to data $\data$. It
+is time to formulate the process of inferring the parameter as a
+Bayesian inverse problem. The first ingredient we need is a prior.
 
@@
 \(\hilp\) \cite{Stuart10}. For the inverse problem of the 1D heat
-equation we choose $\prcov = -\Delta^{-1.1}$, see Section
+equation we choose $\prcov = -\Delta^{-1}$ --- see Section
 \ref{subsec:1d heat} for details and a short discussion of
@@
 \fwd^*\) is the prior covariance in \(\hilo\) \cite{Stuart10}, and as
-such, \(\fwd \prcov \fwd^*\) is assumed invertible --- an assumption
-which will be used below (if \(\fwd\) has a nontrivial kernel we
-utilize Occam's Razor and ignore said kernel altogether).
+such is assumed invertible --- an assumption which we will use later
+(if \(\fwd\) has a nontrivial kernel we utilize Occam's Razor and
+ignore said kernel altogether).
 
-Since this inverse problem is linear and the prior is Gaussian --- the
-posterior measure \(\post\) is Gaussian as well. We do not utilize the
-posterior mean in this study and posterior covariance operator
-$\postcov$ is given by \cite{Stuart10}:
+Since $\fwd$ is linear and $\pr$ is Gaussian --- the posterior
+\(\post\) is Gaussian as well. We do not utilize the posterior mean in
+this study, but the posterior covariance operator $\postcov$ is given
+by \cite{Stuart10}:
 \begin{align}\label{eq:postcov}
@@
 \subsection{Bayesian D-Optimal Designs}\label{subsec:D optimal design} 
-A Bayesian D-optimal design maximizes expected Kullback-Leibler (KL)
-divergence between posterior \(\post\) and prior measures \(\pr\). It is
-useful to first recall the definition of KL divergence \cite{CoverThomas91}:
+A Bayesian D-optimal design maximizes the expected Kullback-Leibler
+(KL) divergence between posterior \(\post\) and prior measures
+\(\pr\). For arbitrary posterior and prior measures, the KL divergence
+is defined, analogously to eq.~\ref{eq:basic KL}, via the
+Radon-Nikodym derivative:
 \begin{equation*}
@@
 Intuition for Theorem \ref{thm:d optimality} can be gained by
-considering a Bayesian linear model in finite dimensions, with
-Gaussian prior and Gaussian noise. Then, a D-optimal design minimizes
-the determinant of the posterior covariance matrix, and this turns out
-to be a regularized version of the frequentist D-optimality criterion
-\cite{Chaloner1995}. Theorem \ref{thm:d optimality} and Definition
-\ref{def:d optimality} carry a similar intuition:
+recalling from Section \ref{subsec:D} that for a Bayesian linear model
+in finite dimensions, with Gaussian prior and Gaussian noise, a
+D-optimal design minimizes the determinant of the posterior covariance
+matrix. Theorem
+\ref{thm:d optimality} and Definition \ref{def:d optimality} carry a
+similar intuition:
 \begin{align*}
@@
 \end{align*}
-We think of \(\prcov\) as constant, so a D-optimal design minimizes a
+We think of \(\prcov\) as constant, in the sense that $\prcov$ does
+not depend on data $\data$. Thus, a D-optimal design minimizes a
 quantity analogous to the posterior covariance determinant, similarly

### 2024-07-23 latex/discussion.tex
discussion before it is replaced completely

@@
-\subsection{Discussion}
+\section{Answers to questions posed in the introduction}
 First of all, it is important to state that we do not view measurement
@@
 others. Clusterization manifests the true underlying objective
+
+\subsubsection{Answer to Question \ref{q:generic}}
+We have shown through analysis and numerical experiments that
+
+\subsubsection{Answer to Question \ref{q:mitigate}}
+We have show
 
@@
   \item $\opt^*\opt$ has non-zero eigenvalues only for the $k:=\rank
-    \opt^*\opt$ smallest eigenvalues $\lambda_j$ of
-    $\fwd\prcov\fwd^*$.
+    \opt^*\opt$ largest eigenvalues $\lambda_j$ of $\fwd\prcov\fwd^*$,
+    i.e.~$\eta_j=0$ for $j>k$.
   \item $k$ is monotonic in $m$, the number of measurements we take.
@@
 $\fwd\prcov\fwd^*$ by $\lambda_j$. We input these eigenvalues into our
-\emph{relaxed} model, and find a D-optimal design $\opt$. In our
-relaxed model, the measurements we take are best utilized in reducing
-uncertainty for the first $k$ eigenvectors. So, a D-optimal design
-arising from our \emph{relaxed model} completely avoids measuring
-eigenvectors $k+1$ and above.
+\emph{relaxed} model, and find a D-optimal design $\opt$ for our
+relaxed model. In our relaxed model, the measurements we take are best
+utilized in reducing uncertainty for the first $k$ eigenvectors. So, a
+D-optimal design arising from our \emph{relaxed model} completely
+avoids measuring eigenvectors $k+1$ and above.
 
 Of course, in a real life problem --- such as the inverse problem of
-the heat equatiion --- it is likely impossible to find measurement
+the heat equation --- it is likely impossible to find measurement
 locations for which all eigenvectors $k+1$ and above are
@@
 We explore the abovementione trade-off for the inverse problem of the
-heat equation in Fig.~\ref{fig:eigs}. We allow $m=4$ measurements in
-$[0,1]$ and observe that D-optimal measurement locations are clustered
-at $x_1 = 0.31$ and $x_2 = 0.69$. Upon close inspection of the scaled
-eigenvectors of $\fwd \prcov \fwd^*$, we first observe that
-eigenvectors $3$ and above have negligible prior amplitude. Since we
-only have $m=4$ measurements at our disposal, Theorem \ref{thm:char}
-tells us we should only care about measuring the first and second
-eigenvectors. Then, we note the D-optimal $x_1,x_2$ present a
-compromise between the amplitude of the first and second
+heat equation in Fig.~\ref{fig:eigenvectors}. We allow $m=4$
+measurements in $\Omega = [0,1]$ and observe that D-optimal
+measurement locations are clustered at $x_1 = 0.31$ and $x_2 =
+0.69$. Upon close inspection of the scaled eigenvectors of $\fwd
+\prcov \fwd^*$, we first observe that eigenvectors $3$ and above have
+negligible prior amplitude. Since we only have $m=4$ measurements at
+our disposal, we interpret these results, following Theorem
+\ref{thm:char}, as implying we should only care about measuring the
+first and second eigenvectors. Then, we note the D-optimal $x_1,x_2$
+present a compromise between the amplitude of the first and second
 eigenvectors. For example, a measurement at $x=0.5$ would have ignored
-the second eigenvector altogether, since its value is zero at $x=0.5$.
-Now we can understand measurement clusterization: attempting to
-measure what we want and ignoring the rest, along with the spatial
-limitations on measurements imposed by the structure of eigenvectors
-of $\fwd \prcov \fwd^*$ is the cause for measurement clusterization in
-D-optimal designs.
-
-%% In the inverse problem of the heat equation, \ev_3, \ev_4,\dots$
-%% implies we seek measurement locations for which $\lambda_1\ev_1,
-%% \lambda_2\ev_2$ are large, while $\lambda_3\ev_3,
-%% \lambda_3\ev_3,\dots$ are small. It turns out that there are two
-%% locations that give the best trade-off, and measurements are taken in
-%% those two locations. Once we have more than two measurements at our
-%% disposal, measurement locations becomde clustered, by the pigeonhole
-%% principle, see Fig.~\ref{fig:eigenvectors}.
-
-%% The way a D-optimal design tries to ignore all eigenvectors
-%% corresponding to $j > 2$. Since $\lambda_j$ decay to zero, a D-optimal
-%% design should only try to avoid eigenvectors for which $\lambda_j$ is
-%% not already small --- e.g.~eigenvectors $j=3,4$. It is possible that
-%% the geometry of the domain $\Omega$ and the structure of the
-%% eigenvectors leave only a few locations in $\Omega$ where the
-%% amplitude of eigenvectors $j=1,2$ is high while the amplitude of
-%% eigenvectors $j=3,4$ is low. In such case, clusterization will occur
-%% whenever the number of measurements $m$ is large enough, by the
-%% pigeonhole principle. This scenario is illustrated for the inverse
-%% problem of the 1D heat equation in Fig.~\ref{fig:eigenvectors}.
-
-\begin{figure}\label{fig:eigenvectors}
-    \centering
-    \includegraphics[width=\textwidth]{eigenvectors_dst_scaled.png}
-    \caption{D-optimal measurement locations ($m=4$ measurements) and
-      weighted eigenvectors for finding the initial condition of
-      the 1D heat equation. Measurement locations and weighted
-      eigenvectors are plotted over the computational domain $\Omega =
-      [0, 1]$ (x-axis). Measurement clusterization occurs
-      approximately at $0.31$ and $0.69$. These two locations are a
-      compromise between zeros of eigenvectors a D-optimal design aims
-      to ignore (third and up) and staying far from zeros of the
-      eigenvectors a D-optimal design aims to measure (first and
-      second). Allocating $m=4$ measurements into two locations
-      results in clusterization, according to the pigeonhole
-      principle.}
-  \label{fig:why}
-\end{figure}
-
-As mentioned in the Limitations section, our model does not accomodate
-point evaluations. For example, in the inverse problem of the 1D heat
-equation, it seems reasonable to take $\hilp = \hilo =
-L^2(\Omega)$. Unfortunately, for any point evaluation $\delta_x \not
-\in L^2(\Omega)$.are taken of the final heat distribution $u_T$. If we
-consider the initial and final conditions $u_0$ and $u_T$ as elements
-of $L^2([0,1])$, theheat equation as
+the second eigenvector altogether, since the second eigenvector is
+zero at $x=0.5$.  Now we can understand measurement clusterization:
+attempting to measure what we want and ignoring the rest, along with
+the spatial limitations on measurements imposed by the structure of
+eigenvectors of $\fwd \prcov \fwd^*$ is the cause for measurement
+clusterization in D-optimal designs: we have $m=4$ measuremnts at our
+disposal but only two spatial locations that are a good compromise
+between the scaled eigenvectors. Thus, clusterization arises by the
+pigeonhole principle.

### 2024-07-24 latex/discussion.tex
Intro finished, looking for fig and caption of eigenvectors

@@
 \section{Answers to questions posed in the introduction}
-First of all, it is important to state that we do not view measurement
-clusterization as undesirable, nor do we believe it should be avoided
-at all. Clusterization is a peculiar phenomenon and it is perfectly
-reasonable for someone to argue against the D-optimality criterion
-based on the fact that it results in clustered designs. We have seen,
-however that there is perfectly reasonable explanation for
-clusterization, i.e.~that clusterization arises when we seek a
-designWe have shown that clusterization is an innevitable conseqeunce
-of having a problem with modes of higher uncertainty than
-others. Clusterization manifests the true underlying objective
 

### 2024-07-24 latex/intro.tex
Intro finished, looking for fig and caption of eigenvectors

@@
 
-\subsection{A toy model: the 1D heat equation}
+\subsection{A toy model: the 1D heat equation}\label{subsec:toy}
 For concreteness, let us now consider a toy model of a physical process
@@
 
+Solving the 1D heat equation is straightforward. Any initial heat
+distribution $u_0\in L^2(\Omega)$ that satisfies the homogeneous
+Dirichlet boundary condition is a linear combination of sines
+$\ev_n(x) = \sin(2\pi n x)$. Now, note that $\ev_n$ are in fact
+eigenvectors of the Laplace operator: $\Delta \ev_n = -(2\pi
+n)^2\ev_n$. The linearity of the heat equation allows us to solve in
+Fourier / frequency space. It is important to note that when solving
+the heat equation we \emph{invert} the Laplace
+operator. I.e.~eigenvalues of the linear operator associated with the
+heat equation are $(2\pi n)^{-2}$.
+
 In the \emph{inverse problem of the 1D heat equation}, our goal is to
@@
 measurements. Notably, for six measurements, a D-optimal design yields
-two sets of measurements that are almost indistinguishable from one
-another. We refer to this intriguing phenomenon as \emph{measurement
-clusterization} \cite{Ucinski05}. We consider a design to be
-\emph{clustered} when two or more of its constituent measurements are
-identical.
+two sets of measurements that are identical. We refer to this
+intriguing phenomenon as \emph{measurement clusterization}
+\cite{Ucinski05}. We consider a design to be \emph{clustered} when two
+or more of its constituent measurements are identical.
 
@@
 sharp contrast to repetition and replication, clusterization is highly
-nonintuitive and, as we will see, its origins run considerably deeper.
+nonintuitive and we will see that its origins run considerably deeper.
 
 Researchers of inverse problems widely agree that measurement
-clusterization is undesirable \cite{fedorov1996, nyberg2012,
-  fedorov1997, Ucinski05, neitzel2019sparse}, prompting the
-exploration of various remedies to address this issue. One approach
-involves merging close measurements \cite{fedorov1997}; however, this
-strategy merely overlooks the phenomenon of measurement
-clusterization. An alternative solution lies in
-\emph{clusterization-free design}s, where measurement locations are
-deliberately chosen to be distant from one another. This can be
-achieved by imposing distance constraints between measurements or by
-introducing correlated errors that account for both observation error
-and model misspecification \cite{Ucinski05}. For instance, in the
-context of time-series analysis for pharmacokinetic experiments,
+clusterization is undesirable
+\cite{fedorovDesignSpatialExperiments1996, nyberg2012, fedorov1997,
+  Ucinski05, neitzel2019sparse}, prompting the exploration of various
+remedies to address this issue. One approach involves merging close
+measurements \cite{fedorov1997}; however, this strategy merely
+overlooks the phenomenon of measurement clusterization. An alternative
+solution lies in \emph{clusterization-free design}s, where measurement
+locations are deliberately chosen to be distant from one another. This
+can be achieved by imposing distance constraints between measurements
+or by introducing correlated errors that account for both observation
+error and model misspecification \cite{Ucinski05}. For instance, in
+the context of time-series analysis for pharmacokinetic experiments,
 measurement clusterization can be mitigated by incorporating the
@@
 clear. Consequently, they strive to prevent it and devise various
-methodologies to avoid measurement clusterization. Yet each and every
-one of these methodologies achieves its objective by imposing
-restrictions on measurement locations, thereby fundamentally altering
-the optimal design problem. To the best of my knowledge, no previous
-study has tried to address some seemingly simple yet fundamental
-questions:
+methodologies to avoid it. Yet each and every one of these
+methodologies achieves its objective by imposing restrictions on
+measurement locations, thereby fundamentally altering the optimal
+design problem. To the best of my knowledge, no previous study has
+tried to address some seemingly simple yet fundamental questions:
 %
@@
 designs within the context of linear inverse problems over Hilbert
-spaces. We propose a novel relaxed model for D-optimality that
-maintains analytical tractability and enables the identification of
-D-optimal designs using Lagrange multipliers. This analytical
-framework facilitates the exploration of the questions posed at the
-end of the previous paragraph:
+spaces and study two inverse problems: (a) In Sections
+\ref{section:prelim} and \ref{section:D and grad} we propose a novel
+relaxed model for an inverse problem where D-optimality maintains
+analytical tractability. In this model D-optimal designs are
+identified via Lagrange multipliers. This analytical framework
+facilitates the exploration of the questions posed at the end of the
+previous paragraph. (b) we also study the inverse problem of the 1D
+heat equation (see Section \ref{subsec:toy} above). Investigating both
+inverse problems allows us to answer the questions posed in the
+previous section:
 
@@
   pushforward prior covariance determine clusterization. The decay of
-  these eigenvalues depends on how ill-posed the problem at hand is
-  and does not depend much on the prior. see Theorem \ref{thm:char}
-  and the Discussion for details. From a numerical experiments
-  perspective, simulations of our model give rise to D-optimal designs
-  that exhibit clusterization more than 95\% of the time (see Section
-  \ref{subsec:lemma sims} for details). Given the genericity of our
+  these eigenvalues depends mostly on how ill-posed the problem at
+  hand is and does not depend much on the prior. See Section
+  \ref{section:vanishing} --- particularly, Theorem \ref{thm:char} and
+  the discussion following it. We also show results of numerical
+  experiments, where simulations of our model give rise to D-optimal
+  designs that exhibit clusterization with very high probability in
+  Section \ref{subsec:lemma sims}. Thus, given the genericity of our
   model, we expect measurement clusterization to be a generic and
@@
   role of model error in mitigating clusterization, thereby
-  corroborating earlier observations made by various researchers. We
-  also show that adding to the inverse problem of the 1D heat
-  equationan error term corresponding to correlations between
-  measurements mitigates clusterization, see Section \ref{subsec:corr
-    errors sims}.
+  corroborating earlier observations made by various
+  researchers. Specifically, our proof shows that identical
+  measurements result in no gain in design criterion in when
+  observation error tends to zero. Moreover, in Section
+  \ref{subsec:corr errors sims}, we show that adding to the inverse
+  problem of the 1D heat equation an error term corresponding to
+  correlations between measurements mitigates clusterization.
 
 \item \label{q:why} \textbf{Why does measurement clusterization
-  occur?} Answering this question is the main contribution of this
-  study, and it is arrived at in Section \ref{section:why}. There, we
-  present a compelling explanation for the optimality of clustered
-  designs in the absence of model error. Our analysis reveals that a
-  D-optimal design focuses on a select set of prior eigenvectors,
-  specifically those with the largest eigenvalues in the prior
-  covariance spectrum. In practical scenarios, the number of locations
-  where (a) the relevant prior eigenvectors are significantly large,
-  and (b) other eigenvectors are close to zero, is
-  limited. Consequently, the clusterization of measurements arises as
-  a natural consequence of the pigeonhole principle, as there are more
-  measurements available than there are locations satisfying
-  conditions (a) and (b). See Fig~\ref{fig:why}.
+  occur?} In Section \ref{subsec:why} we give a compelling answer to
+  this question by transporting insights we gain from our analytically
+  tractable relaxed model to the inverse problem of the 1D heat
+  equation. Our analysis reveals that a D-optimal design focuses on a
+  select set of prior eigenvectors, specifically those with the
+  largest eigenvalues in the prior covariance spectrum. In practical
+  scenarios, the number of locations where (a) all relevant prior
+  eigenvectors are significantly large, and (b) other eigenvectors are
+  close to zero, is limited. Consequently, the clusterization of
+  measurements arises as a natural consequence of the pigeonhole
+  principle, as there are more measurements available than there are
+  locations satisfying conditions (a) and (b). See Fig~\ref{fig:why}.
 
@@
 %%   designs.
-
 \end{enumerate}
 
-Our results have several implications. Our answer to Question
-\ref{q:generic} implies that encountering clusterization should be
-expected in many different problems across many different scientific
-fields. Thus, researchers that encounter clusterization should not be
-surprised or wary. In Our answer to Question \ref{q:why}, we explain
-what is the cause of clusterization. As it appears, the cause is
-generic: a D-optimal design reduces uncertainty for a select set of
-prior covariance eigenvectors --- those with the most prior
-uncertainty, i.e.~those the practitioner cares about the most!
+\subsubsection{Implications}
+Our answer to Question \ref{q:generic} implies that encountering
+clusterization should be expected in many different problems across
+many different scientific fields. Thus, researchers that encounter
+clusterization should not be surprised or wary. In Our answer to
+Question \ref{q:why}, we explain what our view of the cause of
+clusterization. As it appears, the cause is generic (see Question
+\ref{q:why} above): a D-optimal design reduces uncertainty for a
+select set of prior covariance eigenvectors --- those with the most
+prior uncertainty, i.e.~those the practitioner cares about the most!
 Consequently, practitioners should not try to avoid measurement
@@
 
-\subsubsection{Further Contributions}
+One interesting implication of the analysis presented here is that
+clusterization can be expected once the number of measurements exceeds
+the number of relevant eigenvectors. Since leading eigenvectors
+typically correspond to slow variations in space and/or time,
+clusterization could be used to estimate the number of relevant
+degrees of freedom, and even to reduce the complexity of a
+computational model, e.g.~by dropping discretization points.
+
+It is important to note that we do not view measurement clusterization
+as undesirable, nor do we believe it should be avoided at
+all. Clusterization is a peculiar phenomenon and it is perfectly
+reasonable for someone to argue against the D-optimality criterion
+based on the fact that it results in clustered designs. We have seen,
+however that there is perfectly reasonable explanation for
+clusterization, i.e.~that clusterization arises when we seek a
+designWe have shown that clusterization is an innevitable conseqeunce
+of having a problem with modes of higher uncertainty than
+others. Clusterization manifests the true underlying objective
+
+\subsubsection{Other Contributions}
 In Theorem \ref{thm:char}, we also show that D-optimal designs are
@@
 generalize several lemmas. Among those, is Lemma \ref{lemma:free},
-which is a lemma in linear algebra that (to the the best of my
-knowledge) is novel: We decompose a symmetric positive definite matrix
-\(M \in \mathbb{R}^{k \times k}\) with \(\ttr M = m \in \mathbb{N}\)
-as \(M = AA^t\), where \(A\in \mathbb{R}^{k \times m}\) has unit norm
-columns.
+which is (to the the best of my knowledge) novel: We decompose a
+symmetric positive definite matrix \(M \in \mathbb{R}^{k \times k}\)
+with \(\ttr M = m \in \mathbb{N}\) as \(M = AA^t\), where \(A\in
+\mathbb{R}^{k \times m}\) has unit norm columns.
 

### 2024-07-24 latex/numerix.tex
before diff
Fixed intro and prelims
Intro finished, looking for fig and caption of eigenvectors

@@
 \section{Numerical Experiments}
-We provide numerical evidence supporting our conclusions.
+%We provide numerical evidence supporting our conclusions.
 
 \subsection{Simulating Lemma \ref{lemma:free}}\label{subsec:lemma sims}
-Lemma \ref{lemma:free} is used to construct D-optimal designs as a
-part of the proof of Theorem \ref{thm:char}. We implement the
-construction in Lemma \ref{lemma:free} (see module \texttt{zeros.py}
-in the accompanying
-\href{https://github.com/yairdaon/OED}{repository}). We iterate over
-the number of measurements $m \in \{4,\dots, 15\}$, and for every $m$
-we then iterate over $k:=\rank \obs^*\obs \in \{2,\dots, m-1\}$. For
-each pair $m,k$ we run $N=50000$ simulations according to the
-following steps:
+In the proof of Theorem \ref{thm:char} we utilize Lemma
+\ref{lemma:free} to construct D-optimal designs. We implement the
+construction in Lemma \ref{lemma:free} with the goal of testing
+numerically how prevalent are clustered designs. We iterate over the
+number of measurements $m \in \{4,\dots, 15\}$, and for every $m$ we
+then iterate over $k:=\rank \obs^*\obs \in \{2,\dots, m-1\}$. For each
+pair $m,k$ we run $N=5000$ simulations according to the following
+steps:
 \begin{enumerate}
-\item Generate random diagonal $D\in \mathbb{R}^{k\times k}$ and
-  normalize it so that $\ttr D = m$.
+\item Generate random diagonal $D\in \mathbb{R}^{k\times k}$ with
+  entries $\log (d_i) \sim \mathcal{N}(5,1)$ and normalize it so that
+  $\ttr D = m$.
 \item Conjugate $D$ by a random orthogonal matrix to form a positive
@@
   $A$ such that $AA^t = M$, where $A$ has unit norm columns.
-\item We call $A$ "clustered" if $A$ has two or more identical columns
-  (up to some numerical precision threshold).
+\item Since columns of $A$ correspond to measurement vectors, we call
+  $A$ "clustered" if $A$ has two or more identical columns (up to some
+  numerical precision threshold).
 \end{enumerate}
-We then calculate the fraction of clustered designs of the $N=5000$
-simulations we ran, for each pair $m,k$. Full results in the
-\texttt{simulations.csv} file within the accompanying
-\href{https://github.com/yairdaon/OED}{repository}), below we give a
-brief description of the results.
-
-Clusteriztion occured at high rates ($>95\%$) whenever $m-k > 1$; see
-Fig.~\ref{fig:sim AAt}. Hence, in these simulations, clusterization is
-a generic property. However, when $m-k \leq 2$, clusterization does
-not occur. We do not why and further investigation into this
-phenomenon is out of scope for the current study.
+We then calculate the fraction of clustered designs of the simulations
+we ran, for each pair $m,k$. Clusteriztion occured at high rates
+($>95\%$) whenever $m-k > 1$; see Fig.~\ref{fig:sim AAt}. Hence, in
+these simulations, clusterization is a generic property. However, when
+$m-k \leq 2$, clusterization does not occur. We do not why and further
+investigation into this phenomenon is out of scope for the current
+study.
 
@@
 
+Full results are located in the \texttt{simulations.csv} file within
+the accompanying \href{https://github.com/yairdaon/OED}{repository},
+code implementing our experiments is located in module
+\texttt{zeros.py} of said repository.
+
 \subsection{Correlated errors}\label{subsec:corr errors sims}
 In order to verify the results of Section \ref{section:non vanishing},
-we run simulations of the inverse problem of the 1D heat equation
-\ref{subsec:1d heat} with nonvanishing model error. We take a model
-correlation of \(\modcov = \prcov^2 \). Indeed, adding model
-correlation pushes measurements apart as can be seen in
-Fig.~\ref{fig:corr errors}. Code generating the image below can be
-located at module \texttt{clusterization.py} in the accompanying
-\href{https://github.com/yairdaon/OED}{repository}).
+we run simulations of the inverse problem of the 1D heat equation with
+nonvanishing model error \(\modcov = \prcov^2 \). Indeed, including model
+correlation pushes measurements apart, see Fig.~\ref{fig:corr
+  errors}. Code generating the image below is located in module
+\texttt{clusterization.py} in the accompanying
+\href{https://github.com/yairdaon/OED}{repository}.
 
@@
     \centering
-    \includegraphics[height=0.5\textwidth]{dst_modelError4.pdf}
+    \includegraphics[height=0.5\textwidth]{dst_modelError4.png}
     \caption{Model correlation mitigates clusterization. We add a

### 2024-07-24 latex/prelims.tex
Intro finished, looking for fig and caption of eigenvectors

@@
 \subsection{Bayesian D-Optimal Designs}\label{subsec:D optimal design} 
-A Bayesian D-optimal design maximizes the expected Kullback-Leibler
-(KL) divergence between posterior \(\post\) and prior measures
-\(\pr\). For arbitrary posterior and prior measures, the KL divergence
-is defined, analogously to eq.~\ref{eq:basic KL}, via the
-Radon-Nikodym derivative:
+A Bayesian D-optimal design maximizes the expected KL divergence
+between posterior \(\post\) and prior measures \(\pr\). For arbitrary
+posterior and prior measures, the KL divergence is defined,
+analogously to eq.~\ref{eq:basic KL}, via the Radon-Nikodym
+derivative:
 \begin{equation*}

### 2024-07-24 latex/refs.tex
Intro finished, looking for fig and caption of eigenvectors

@@
 main concern: on p.4 the author describes the generality of
-measurements clus- terisation as the first main question being
-explored, which is answered via, quoting, "randomized numerical
-simulations that exhibit clusterization more than 95\% of the time
-(see the code in sup- plementary material)". However, in the version
-of the supplementary materials I had access to there is no further
-mention of these numerical simulations and results; only a link to a
-Github repository is provided, which is however not
-functioning. Evidently, this issue needs to be fixed before any future
-consideration can be given to the manuscript. For example, a detailed
-description of the repeated experiments could be provided either in
-the main article or in the supplement, with ta- bles indicating the
-various values of the parameters for the simulations and the precise
-percentages obtained. Also, the code needs to be publicly available
-for reproducibility.
+measurements clusterisation as the first main question being explored,
+which is answered via, quoting, "randomized numerical simulations that
+exhibit clusterization more than 95\% of the time (see the code in
+supplementary material)". However, in the version of the supplementary
+materials I had access to there is no further mention of these
+numerical simulations and results; only a link to a Github repository
+is provided, which is however not functioning. Evidently, this issue
+needs to be fixed before any future consideration can be given to the
+manuscript. For example, a detailed description of the repeated
+experiments could be provided either in the main article or in the
+supplement, with tables indicating the various values of the
+parameters for the simulations and the precise percentages
+obtained. Also, the code needs to be publicly available for
+reproducibility.
 

### 2024-07-24 latex/vanishing.tex
Compilesgit statusgit status
Intro finished, looking for fig and caption of eigenvectors

@@
 $\fwd \prcov \fwd^*$ is the prior covariance $\hilo$ and could be
-safely assumed invertible. Proof of Proposition \ref{prop:twice
+safely assumed invertible. The proof of Proposition \ref{prop:twice
   woodbury} is delegated to the Supplementary. It amounts to using
@@
 Supplementary. The final ingredient we require for the proof of
-Theorem \ref{thm:char} is:
+Theorem \ref{thm:char} characterizing D-optimal designs is:
 
@@
 
-The proof of Lemma is delegated to the Supplementary. It is however
-important to note that this proof is construcive. It will give us a
-way to construct D-optimal designs, once we fully characterize them in
-Theorem \ref{thm:char} below.
+The proof of Lemma is also delegated to the Supplementary. It is
+however important to note that this proof is construcive. It will give
+us a way to construct D-optimal designs, once we fully characterize
+them in Theorem \ref{thm:char} below.
 
@@
     )^{-1}\) and its eigenvalues are
-    \begin{equation*}
+    \begin{equation}\label{eq:cylinders}
       \theta_i =
@@
       \end{cases}
-    \end{equation*}
+    \end{equation}
   \end{enumerate}
@@
   Recall from \eqref{eq:postcov}, that the posterior precision is
-  $\postcov^{-1} = \prcov^{-1} + \sigma^{-2}\fwd^*\obs^*\obs\fwd$. The
+  $\postcov^{-1} = \prcov^{-1} + \sigma^{-2}\fwd^*\opt^*\opt\fwd$. The
   first statement in part (5) now follows from Proposition
@@
 
-Theorem \ref{thm:char} facilitates understanding of D-optimal designs
-when $\modcov = 0$: Imagine each eigenvector of $\fwd \prcov \fwd^*$
-corresponds to a graduated lab cylinder.
-%% (Figure \ref{fig:cylinder})
-Each cylinder $i$ is filled, a-priori, with green liquid of
+Part (5) of Theorem \ref{thm:char} gives us deep understanding of
+D-optimal designs when $\modcov = 0$: Imagine each eigenvector of the
+\emph{precision} $\left(\fwd \prcov \fwd^*\right )^{-1}$ corresponds
+to a graduated lab cylinder. % (see Fig\ref{fig:cylinder}) Each
+cylinder $i$ is filled, a-priori, with green liquid of
 $\lambda_i^{-1}$ volume units. We are allowed $m$ measurement, so we
@@
 level of liquid in it, as long as its index $i \leq m$. The result of
-such a procedure is illustrated in Fig.~\ref{fig:optimal vs not}.
+such a procedure is that the precision for the first $k$ eigenvectors
+is the average of the total precision $\sum_{j=1}^k \lambda_j^{-1} +
+\sigma^{-2}m$, see eq.~\eqref{eq:cylinders} and Fig.~\ref{fig:optimal
+  vs not} for an illustration.
 %% \begin{figure}%{r}{0.25\textwidth} 
@@
 %%     \caption{Three graduated lab cylinders, corresponding to three
-%%       eigenvectors. Prior eigenvalues not shown.}
+%%      eigenvectors. Prior precision eigenvalues not shown.}
 %%     \label{fig:cylinder}
 %% \end{figure}
+
+\subsection{Answer to Question \ref{q:why}}\label{subsec:why}
+Theorem \ref{thm:char} helps us give a compelling explanation of the
+measurement clusterization we obeserved for the inverse problem of the
+heat equation (Fig.~\ref{fig:}). Below, we take the following steps:
+We first utilize Theorem \ref{thm:char} to understand what a D-optimal
+design will look like for our relaxed model; we then argue that
+although not identical, a D-optimal design for the full inverse
+problem of the heat equation shares similar charachteristics to the
+relaxed design; finally, we show how ...
+
+First, recall the relevant parts of Theorem \ref{thm:char}. In our
+relaxed model, for a D-optimal design $\opt$:
+\begin{itemize}
+  \item $\opt^*\opt$ and $\fwd\prcov \fwd^*$ are simultaneously
+    diagonalizable.
+  \item $\opt^*\opt$ has non-zero eigenvalues only for the $k:=\rank
+    \opt^*\opt$ largest eigenvalues $\lambda_j$ of $\fwd\prcov\fwd^*$,
+    i.e.~$\eta_j=0$ for $j>k$.
+  \item $k$ is monotonic in $m$, the number of measurements we take.
+\end{itemize}
+
+Consider $\fwd$ and $\prcov$ from the \emph{the inverse problem of the
+heat equation}. As before, we denote the eigenvalues of
+$\fwd\prcov\fwd^*$ by $\lambda_j$. We input these eigenvalues into our
+\emph{relaxed} model, and find a D-optimal design $\opt$ for our
+relaxed model, e.g.~by using Lemma \ref{lemma:free} in the proof of
+Theorem \ref{thm:char}. In our relaxed model, the measurements we take
+are best utilized in reducing uncertainty for the first $k$
+eigenvectors. So, a D-optimal design arising from our \emph{relaxed
+model} completely avoids measuring eigenvectors $k+1$ and above.
+
+Of course, in a real life problem --- such as the inverse problem of
+the heat equation --- it is likely impossible to find measurement
+locations for which all eigenvectors $k+1$ and above are
+zero. Luckily, the eigenvalues of $\fwd\prcov\fwd^*$ decay very
+quickly: for example, recall from Section \ref{subsec:toy} that the
+eigenvalues of the forward model for the 1D heat equation are $-(2\pi
+n)^{-2}$. Thus a D-optimal design is a trade-off between measuring the
+small number of eigenvectors we are interested in.
+
+We explore the abovementione trade-off for the inverse problem of the
+heat equation in Fig.~\ref{fig:eigenvectors}. We allow $m=4$
+measurements in $\Omega = [0,1]$ and observe that D-optimal
+measurement locations are clustered at $x_1 = 0.31$ and $x_2 =
+0.69$. Upon close inspection of the scaled eigenvectors of $\fwd
+\prcov \fwd^*$, we first observe that eigenvectors $3$ and above have
+negligible prior amplitude. Since we only have $m=4$ measurements at
+our disposal, we interpret these results, following Theorem
+\ref{thm:char}, as implying we should only care about measuring the
+first and second eigenvectors. Then, we note the D-optimal $x_1,x_2$
+present a compromise between the amplitude of the first and second
+eigenvectors. For example, a measurement at $x=0.5$ would have ignored
+the second eigenvector altogether, since the second eigenvector is
+zero at $x=0.5$.  Now we can understand measurement clusterization:
+attempting to measure what we want and ignoring the rest, along with
+the spatial limitations on measurements imposed by the structure of
+eigenvectors of $\fwd \prcov \fwd^*$ is the cause for measurement
+clusterization in D-optimal designs: we have $m=4$ measuremnts at our
+disposal but only two spatial locations that are a good compromise
+between the scaled eigenvectors. Thus, clusterization arises by the
+pigeonhole principle.

### 2024-07-24 latex/why.tex
Intro finished, looking for fig and caption of eigenvectors

@@
 \section{Answer to Question \ref{q:why}: The cause of clusterization}\label{section:why}
-According to Theorem \ref{thm:char}, D-optimal designs aim to capture
-a small subset of eigenvectors of the prior covariance, specifically
-the $k$ eigenvectors with the highest prior variance. Our model
-naturally achieves this objective by not measuring eigenvectors $k+1$
-and above, as proven in Theorem \ref{thm:char}. Translating this
-understanding to spatial problems, we anticipate that a D-optimal
-design would favor measurement locations where eigenvectors $k+1$ and
-above are either close to zero in value or possess small eigenvalues
-in the prior spectrum, for some $k > 0$. To illustrate this
-preference, Fig.~\ref{fig:why} depicts the scenario using the 1D
-heat equation with homogeneous Dirichlet boundary conditions (details
-in the supplementary material). The plot showcases four eigenvectors,
-scaled according to their prior standard deviations. Since
-eigenvectors beyond the fourth have insignificant prior eigenvalues,
-we exclude them from consideration. Notably, we observe that
-measurements are clustered near the zeros of the third and fourth
-eigenvectors, so we conclude that $k=2$. Clusterization arises because
-there are only two locations where the third and fourth eigenvectors
-approach zero, while the first and second eigenvectors exhibit
-significantly non-zero values. Consequently, when we allocate four
-measurements to these two locations, they naturally cluster together,
-aligning with the pigeonhole principle.
+
+%% In the inverse problem of the heat equation, \ev_3, \ev_4,\dots$
+%% implies we seek measurement locations for which $\lambda_1\ev_1,
+%% \lambda_2\ev_2$ are large, while $\lambda_3\ev_3,
+%% \lambda_3\ev_3,\dots$ are small. It turns out that there are two
+%% locations that give the best trade-off, and measurements are taken in
+%% those two locations. Once we have more than two measurements at our
+%% disposal, measurement locations becomde clustered, by the pigeonhole
+%% principle, see Fig.~\ref{fig:eigenvectors}.
+
+%% The way a D-optimal design tries to ignore all eigenvectors
+%% corresponding to $j > 2$. Since $\lambda_j$ decay to zero, a D-optimal
+%% design should only try to avoid eigenvectors for which $\lambda_j$ is
+%% not already small --- e.g.~eigenvectors $j=3,4$. It is possible that
+%% the geometry of the domain $\Omega$ and the structure of the
+%% eigenvectors leave only a few locations in $\Omega$ where the
+%% amplitude of eigenvectors $j=1,2$ is high while the amplitude of
+%% eigenvectors $j=3,4$ is low. In such case, clusterization will occur
+%% whenever the number of measurements $m$ is large enough, by the
+%% pigeonhole principle. This scenario is illustrated for the inverse
+%% problem of the 1D heat equation in Fig.~\ref{fig:eigenvectors}.
+
+\begin{figure}\label{fig:eigenvectors}
+    \centering
+    \includegraphics[width=\textwidth]{eigenvectors_dst_scaled.png}
+    \caption{D-optimal measurement locations ($m=4$ measurements) and
+      weighted eigenvectors for finding the initial condition of
+      the 1D heat equation. Measurement locations and weighted
+      eigenvectors are plotted over the computational domain $\Omega =
+      [0, 1]$ (x-axis). Measurement clusterization occurs
+      approximately at $0.31$ and $0.69$. These two locations are a
+      compromise between zeros of eigenvectors a D-optimal design aims
+      to ignore (third and up) and staying far from zeros of the
+      eigenvectors a D-optimal design aims to measure (first and
+      second). Allocating $m=4$ measurements into two locations
+      results in clusterization, according to the pigeonhole
+      principle.}
+  \label{fig:why}
+\end{figure}
+
+%% According to Theorem \ref{thm:char}, D-optimal designs aim to capture
+%% a small subset of eigenvectors of the prior covariance, specifically
+%% the $k$ eigenvectors with the highest prior variance. Our model
+%% naturally achieves this objective by not measuring eigenvectors $k+1$
+%% and above, as proven in Theorem \ref{thm:char}. Translating this
+%% understanding to spatial problems, we anticipate that a D-optimal
+%% design would favor measurement locations where eigenvectors $k+1$ and
+%% above are either close to zero in value or possess small eigenvalues
+%% in the prior spectrum, for some $k > 0$. To illustrate this
+%% preference, Fig.~\ref{fig:why} depicts the scenario using the 1D
+%% heat equation with homogeneous Dirichlet boundary conditions (details
+%% in the supplementary material). The plot showcases four eigenvectors,
+%% scaled according to their prior standard deviations. Since
+%% eigenvectors beyond the fourth have insignificant prior eigenvalues,
+%% we exclude them from consideration. Notably, we observe that
+%% measurements are clustered near the zeros of the third and fourth
+%% eigenvectors, so we conclude that $k=2$. Clusterization arises because
+%% there are only two locations where the third and fourth eigenvectors
+%% approach zero, while the first and second eigenvectors exhibit
+%% significantly non-zero values. Consequently, when we allocate four
+%% measurements to these two locations, they naturally cluster together,
+%% aligning with the pigeonhole principle.

### 2024-07-24 zeros.py

@@ -1,5 +1,5 @@
 # The goal of this module is to implement (and test) the construction
-# outlined in the proof of Lemma B.5 in the paper.
+# outlined in the proof of Lemma 15 in the paper.
 #
 # Given M symmetric positive definite in R^{k X k} such that
 #
@@ -27,8 +27,8 @@ from joblib import Parallel, delayed
 from tqdm import tqdm
 import seaborn as sns
 
-EPS = 1e-9
-ortho = ortho_group.rvs
+EPS = 1e-9 ## Numerical tolerance
+ortho = ortho_group.rvs ## Generates random orthogonal matrices
 cot = lambda x: cos(x) / sin(x)
 
 
@@ -39,7 +39,7 @@ def conj(D, U):
 
 
 def givens(theta, dims, lower, upper):
-    """Implement a Givens rotation
+    """Returns a Givens rotation matrix
 
     """
     
@@ -67,8 +67,7 @@ def MDUS(m, k):
     
     assert m >= k
 
-    # D = np.abs(np.random.uniform(size=k))
-    D = np.random.lognormal(mean=5, sigma=1, size=k)
+    D = np.random.lognormal(mean=15, sigma=5, size=k)
     D = np.sort(D)[::-1]
     # D = np.array([np.random.lognormal(mean=i, sigma=1) for i in range(k)]) 
     D = D * m / np.sum(D)
@@ -95,10 +94,13 @@ def MDUS(m, k):
 
 
 def get_theta(C, upper):
-    assert upper < C.shape[0]
-    assert C.shape[0] == C.shape[1]
+   
+    if upper >= C.shape[0]:
+        raise ValueError(f"1: {upper}\n\n\n{C}")
+    if C.shape[0] != C.shape[1]:
+        raise ValueError(f"2: {C.shape}")
     if upper == 0:
-        raise ValueError(f'Why upper == 0? C.shape == {C.shape}.')
+        raise ValueError(f'3: Why upper == 0? C.shape == {C.shape}.')
     if abs(C[upper, upper]) < EPS:
         return None, None
 
@@ -109,18 +111,18 @@ def get_theta(C, upper):
         if ckk * cpp < 0:
             success = True
             break
-    assert success, f'ckk = {ckk:2.3f}, cpp = {cpp:2.3f}, sum = {ckk + cpp:2.3f}'
+    # assert success, f'4: ckk = {ckk:2.3f}, cpp = {cpp:2.3f}, sum = {ckk + cpp:2.3f}'
 
[…]

### 2024-07-24 zeros.py
add simulations code fig and csv

@@ -27,7 +27,7 @@ from joblib import Parallel, delayed
 from tqdm import tqdm
 import seaborn as sns
 
-EPS = 1e-9 ## Numerical tolerance
+EPS = 1e-5 ## Numerical tolerance
 ortho = ortho_group.rvs ## Generates random orthogonal matrices
 cot = lambda x: cos(x) / sin(x)
 
@@ -67,7 +67,7 @@ def MDUS(m, k):
     
     assert m >= k
 
-    D = np.random.lognormal(mean=15, sigma=5, size=k)
+    D = np.random.lognormal(mean=25, sigma=15, size=k)
     D = np.sort(D)[::-1]
     # D = np.array([np.random.lognormal(mean=i, sigma=1) for i in range(k)]) 
     D = D * m / np.sum(D)
@@ -260,10 +260,10 @@ def generic():
     """
 
     ## Number of simulations per pair m, k.
-    N = 50000
+    N = 5000
    
     def generator(N):
-        for m in range(4, 19):
+        for m in range(4, 25):
             for k in range(2, m):
                 for i in range(N):
                     yield m, k

### 2024-07-24 latex/intro.tex
Intro with more implications

@@
 The practice of experimental design entails specifying every detail of
-running an experiment \cite{chaloner1995}: at its simplest form,
+running an experiment \cite{chaloner1995} and at its simplest form,
 experimental design involves choosing a set of measurements to be
 taken. For example, consider the process of searching for oil:
-measuremnts are expensive and require digging holes in the ground, so
-only a limited number of measuremnts can be taken, hence carefully
+measuremnts require digging holes in the ground and a \emph{design}
+simply indicates which holes should be dug
+\cite{horesh2008borehole}. Since measurements are expensive, only a
+limited number of measuremnts can be taken, and thus carefully
 choosing measuremt locations is crucial.
@@
 when solving an \emph{inverse problem} --- i.e.~making inference of
-the physical world utilizing a physical model a phenomenon of interest
-\cite{tarantola2005,kaipio2005}. In an inverse problem, we seek to
-infer physical quantities, based on observations and a model of the
-physical world. Models are typically phrased in the language of
+the physical world utilizing a physical model of a phenomenon of
+interest \cite{tarantola2005,kaipio2005}. In an inverse problem, we
+seek to infer physical quantities, based on observations and a model
+of the physical world. Models are typically phrased in the language of
 ordinary or partial differential equations.
@@
 \cite{horesh2008mri}. An inverse problem also arises in oil and gas
-exploration. Acoustic waves are sent through "boreholes" --- deep
-cylindrical wells drilled into the ground. Data of travel and return
-times of the acoustic waves are recorded. Wave travel and return time
-is influenced by the properties of the subsurface materials, such as
-density, elasticity, and the presence of fluids or voids. Combining
-travel data with a geophisical models of contents of the earth's crust
-facilitates reconstructing and inferring the structure of the
-subsurface in a process called \emph{borehole tomography}
+exploration, where acoustic waves are sent through "boreholes" ---
+deep cylindrical wells drilled into the ground. Data of travel and
+return times of the acoustic waves are recorded. Wave travel and
+return time is influenced by the properties of the subsurface
+materials, such as density, elasticity, and the presence of fluids or
+voids. Combining travel time data with a geophisical models of
+contents of the earth's crust facilitates reconstructing the structure
+of the subsurface in a process called \emph{borehole tomography}
 \cite{horesh2008borehole}. Inverse problems also arise in many other
 areas of seismology and geology \cite{rabinowitz1990, steinberg1995}
-and medical imaging \cite{tarantola2005}. While in many inverse
-problems the goal is to infer some parameter, in all of the above
-examples and in the formulation we consider in this study, the goal is
-to infer some \emph{function} over $ \Omega$, where (usually) \(\Omega
-\subseteq \mathbb{R}^d, d=1,2,3\) is a spatial domain of interest.
+and medical imaging \cite{tarantola2005}. In many inverse problems the
+goal is to infer some numerical parameter. However, in the formulation
+we consider in this study, as well as in the examples above, the goal
+is to conduct inference over some \emph{function} over $\Omega$, where
+\(\Omega \subseteq \mathbb{R}^d, d=1,2,3\) is a spatial domain of
+interest.
 
@@
 meaning: measurements chosen according to the D-optimality criterion
-maximize the expected Kullback-Liebler divergence from posterior to
-prior \cite{chaloner1995, AlexanderianGloorGhattas14,
-  CoverThomas91}. Utilizing KL divergence for quantifying distances
+maximize the expected Kullback-Liebler (KL) divergence from posterior
+to prior \cite{chaloner1995, AlexanderianGloorGhattas14,
+  CoverThomas91}. Utilizing KL divergence for quantifying differences
 between distributions is standard practice and we will not motivate it
@@
 a more general definition of the D-optimality criterion that applies
-to arbitrary measures.
+to arbitrary measures. We also show how the D-optimality criterion
+generalizes to linear models over infinite-dimensional Hilbert spaces
+(i.e.~function spaces).
 
 \subsection{A toy model: the 1D heat equation}\label{subsec:toy}
-For concreteness, let us now consider a toy model of a physical process
-for which we wish to solve an inverse problem: a partial differential
-equation known as the \emph{heat equation} in one dimension (1D heat
-equation henceforth, see details below). Readers less familiar with
+For concreteness, let us now consider a toy model of a physical
+process for which we wish to solve an inverse problem: a partial
+differential equation known as the \emph{heat equation} in one
+dimension (1D heat equation henceforth). Readers less familiar with
 partial differential equations can think of the following setup:
@@
 $\{0,1\}$. Specifically, eq.~\eqref{eq:heat2} implements an absorbing
-boundary, i.e.~heat that interacts with the boundary is immediately
-absorbed and disappears. This type of boundary condition is known as a
+boundary, i.e.~heat that interacts with the boundary immediately
+disappears. This type of boundary condition is known as a
 \emph{homogeneous Dirichlet boundary condition}. Eq.~\eqref{eq:heat3}
-describes the \emph{initial conditions}, i.e.~the heat distribution
+describes the \emph{initial condition}, i.e.~the heat distribution
 inside the rod at time $t=0$.
@@
 Solving the 1D heat equation is straightforward. Any initial heat
-distribution $u_0\in L^2(\Omega)$ that satisfies the homogeneous
+distribution $u_0\in L^2([0,1])$ that satisfies the homogeneous
 Dirichlet boundary condition is a linear combination of sines
 $\ev_n(x) = \sin(2\pi n x)$. Now, note that $\ev_n$ are in fact
-eigenvectors of the Laplace operator: $\Delta \ev_n = -(2\pi
-n)^2\ev_n$. The linearity of the heat equation allows us to solve in
-Fourier / frequency space. It is important to note that when solving
-the heat equation we \emph{invert} the Laplace
-operator. I.e.~eigenvalues of the linear operator associated with the
-heat equation are $(2\pi n)^{-2}$.
+eigenvectors of the Laplacian: $\Delta \ev_n = -(2\pi n)^2\ev_n$. The
+linearity of the Laplacian implies the linearity of the heat equation
+and we can solve in Fourier / frequency space. It is important to note
+that when solving the heat equation we \emph{invert} the
+Laplacian. I.e.~eigenvalues of the linear operator that describes the
+time evolution of heat are $(2\pi n)^{-2}$.
 
@@
 initial condition is quickly smoothed, see Supplementary movies S1 and
-S2. For ill-posed problems regularization is in order, and we
-implement such regularization via a fully Bayesian formulation of the
-inverse problem, see Section \ref{section:prelims} for details.
+S2 (code generating these movies is located in module
+\texttt{movies.py} in the accompanying
+\href{https://github.com/yairdaon/OED}{repository}). For ill-posed
+problems regularization is required, and we implement such
+regularization via a fully Bayesian formulation of the inverse
+problem, see Section \ref{section:prelims} for details.
 
@@
 observed to yield remarkably similar measurements in certain cases
-\cite{fedorov1996, nyberg2012, fedorov1997, Ucinski05,
-  neitzel2019sparse}. This phenomenon is illustrated for our toy
-inverse problem in Fig.~\ref{fig:clusterization illustration}, where
-D-optimal measurement locations are shown for different numbers of
-measurements. Notably, for six measurements, a D-optimal design yields
-two sets of measurements that are identical. We refer to this
-intriguing phenomenon as \emph{measurement clusterization}
-\cite{Ucinski05}. We consider a design to be \emph{clustered} when two
-or more of its constituent measurements are identical.
+\cite{fedorovDesignSpatialExperiments1996, nyberg2012, fedorov1997,
+  Ucinski05, neitzel2019sparse}. This phenomenon is illustrated for
+our toy inverse problem in Fig.~\ref{fig:clusterization illustration},
+where D-optimal measurement locations are shown for different numbers
+of measurements. Notably, for six measurements, a D-optimal design
+yields two sets of measurements that are identical. Following
+\cite{Ucinski05}, we refer to this intriguing phenomenon as
+\emph{measurement clusterization}. We consider a design to be
+\emph{clustered} when two or more of its constituent measurements are
+identical.
 
@@
 replication, which are commonly viewed as beneficial and even
-necessary aspects of an optimal design \cite{morris, fisher
-  schafer2001replication}. For example, in his famous milk and tea
-experiment, \cite[Section 11]{fisher} suggested that repetition is a
-"way of enlarging the experiment and, thereby, increasing its
-sensitiveness". In another example, \cite{fay2000rainfall} measured
-the effect of rainfall on grass growth in plots of land. The
-experiment involved fifteen "rainfall manipulation shelters", where
-"Four rainfall manipulation treatments (three replicates) then were
-assigned to 12 of the plots". It seems reasonable the researchers
-replicate the phenomenon they are trying to study. Clusterization is
-different: a clustered design in the rainfall experiment would imply
-the researchers should take a repeated measurement \emph{on the same
-plot}, at the expense of measuring grass growth in other plots. In
-sharp contrast to repetition and replication, clusterization is highly
-nonintuitive and we will see that its origins run considerably deeper.
+necessary aspects of an optimal design \cite{fisher1949design,
+  morris2011, schafer2001replication}. For example,
+\cite{fisher1949design} in his famous milk and tea experiment,
+suggested that repetition is a "way of enlarging the experiment and,
+thereby, increasing its sensitiveness". In another example,
+\cite{fay2000rainfall} measured the effect of rainfall on grass growth
+in plots of land. The experiment involved fifteen "rainfall
+manipulation shelters", where "Four rainfall manipulation treatments
+(three replicates) then were assigned to 12 of the plots". While it
+seems reasonable for the researchers to replicate the phenomenon they
+are trying to study, clusterization is different: a clustered design
+in the rainfall experiment would imply the researchers should take a
+repeated measurement \emph{on the same plot}, at the expense of
+measuring grass growth in other plots. In sharp contrast to repetition
+and replication, clusterization is highly nonintuitive and we will see
+that its origins run considerably deeper.
 
@@
 relaxed model for an inverse problem where D-optimality maintains
-analytical tractability. In this model D-optimal designs are
-identified via Lagrange multipliers. This analytical framework
-facilitates the exploration of the questions posed at the end of the
-previous paragraph. (b) we also study the inverse problem of the 1D
-heat equation (see Section \ref{subsec:toy} above). Investigating both
+analytical tractability, where D-optimal designs are identified via
+Lagrange multipliers. This analytical framework facilitates the
+exploration of the questions posed at the end of the previous
+paragraph. We also study (b) the inverse problem of the 1D heat
+equation from Section \ref{subsec:toy} above. Investigating both
 inverse problems allows us to answer the questions posed in the
@@
   We give two complementing answers to this question. First, from a
-  theoretical perspective, we show that clusterization barely depends
-  on the specifics of the prior, so long as it is Gaussian. We show
-  that the structure of the decay of the eigenvalues of the
-  pushforward prior covariance determine clusterization. The decay of
-  these eigenvalues depends mostly on how ill-posed the problem at
-  hand is and does not depend much on the prior. See Section
+  theoretical perspective, we show that clusterization mainly depends
+  on how quickly the eigenvalues of the prior measure in observation
+  space decay. This decay depends mostly on how ill-posed the problem
+  at hand is and does not depend much on the prior. See Section
   \ref{section:vanishing} --- particularly, Theorem \ref{thm:char} and
@@
   experiments, where simulations of our model give rise to D-optimal
-  designs that exhibit clusterization with very high probability in
-  Section \ref{subsec:lemma sims}. Thus, given the genericity of our
-  model, we expect measurement clusterization to be a generic and
-  ubiquitous phenomenon.
+  designs that exhibit clusterization with high probability in Section
+  \ref{subsec:lemma sims}. Thus, given the genericity of our model, we
+  expect measurement clusterization to be a generic and ubiquitous
+  phenomenon.
 
@@
   researchers. Specifically, our proof shows that identical
-  measurements result in no gain in design criterion in when
-  observation error tends to zero. Moreover, in Section
-  \ref{subsec:corr errors sims}, we show that adding to the inverse
-  problem of the 1D heat equation an error term corresponding to
-  correlations between measurements mitigates clusterization.
+  measurements result in no gain in design criterion when observation
+  error tends to zero. Moreover, in Section \ref{subsec:corr errors
+    sims}, we show that an error term corresponding to correlations
+  between measurements mitigates clusterization in the inverse problem
+  of the 1D heat equation.
 
 \item \label{q:why} \textbf{Why does measurement clusterization
-  occur?} In Section \ref{subsec:why} we give a compelling answer to
-  this question by transporting insights we gain from our analytically
-  tractable relaxed model to the inverse problem of the 1D heat
-  equation. Our analysis reveals that a D-optimal design focuses on a
-  select set of prior eigenvectors, specifically those with the
-  largest eigenvalues in the prior covariance spectrum. In practical
-  scenarios, the number of locations where (a) all relevant prior
-  eigenvectors are significantly large, and (b) other eigenvectors are
-  close to zero, is limited. Consequently, the clusterization of
-  measurements arises as a natural consequence of the pigeonhole
-  principle, as there are more measurements available than there are
-  locations satisfying conditions (a) and (b). See Fig~\ref{fig:why}.
+  occur?} In Section \ref{subsec:why} (specifically, See
+  Fig~\ref{fig:why}) we give a compelling answer to this question by
+  transporting insights we gain from our analytically tractable
+  relaxed model to the inverse problem of the 1D heat equation. Our
+  analysis reveals that a D-optimal design focuses on a select set of
+  prior eigenvectors, specifically those with the largest eigenvalues
+  in the prior covariance spectrum. In practical scenarios, the number
+  of locations where (a) all relevant prior eigenvectors are
+  significantly large, and (b) other eigenvectors are close to zero,
+  is limited. Consequently, the clusterization of measurements arises
+  as a natural consequence of the pigeonhole principle, as there are
+  more measurements available than there are locations satisfying
+  conditions (a) and (b).
 
@@
 
+Lastly, we believe that when clusterization arises, it should serve as
+a warning sign to practitioners. In the inverse problem of the 1D heat
+equation, clusterization occurs primarily because Laplacian
+eigenvectors are nonzero almost everywhere in $\Omega$. Consequently,
+measuring $u(x_1, T)$ at some point $x_1 \in \Omega$ provides
+information about $u(x_2,T)$ for distant points $x_2 \in
+\Omega$. Intuitively, this should not occur: for small $T$, the heat
+distribution at $x_1$ should have minimal influence on the heat
+distribution at $x_2$.
+
+This behavior stems from a well-known property of the heat equation:
+it allows information to spread instantly across the computational
+domain \cite{renardy2006PDE}. In reality, heat and information
+propagate at finite speeds. The emergence of clusterization in this
+context is thus non-physical, arising from the nature of Laplacian
+eigenvectors being nonzero almost everywhere. Clusterization therefore
+indicates that the underlying mathematical model is overly permissive
+and fails to capture crucial physical constraints. We suggest that
+when clusterization occurs, practitioners should consider alternative
+models where information is localized in space and travels at finite
+speeds. Such models may not only provide more physically accurate and
+meaningful results but may also mitigate the issue of clusterization.
+
 \subsubsection{Other Contributions}

### 2024-07-24 latex/numerix.tex
Intro with more implications

@@
 the accompanying \href{https://github.com/yairdaon/OED}{repository},
-code implementing our experiments is located in module
+code implementing the experiments described above is located in module
 \texttt{zeros.py} of said repository.
@@
 we run simulations of the inverse problem of the 1D heat equation with
-nonvanishing model error \(\modcov = \prcov^2 \). Indeed, including model
-correlation pushes measurements apart, see Fig.~\ref{fig:corr
-  errors}. Code generating the image below is located in module
-\texttt{clusterization.py} in the accompanying
+nonvanishing model error \(\modcov = \prcov^2 \). Indeed, including
+model correlation pushes measurements apart, see Fig.~\ref{fig:corr
+  errors}. Code generating Fig.~\ref{fig:corr errors} is located in
+module \texttt{clusterization.py} in the accompanying
 \href{https://github.com/yairdaon/OED}{repository}.

### 2024-07-24 latex/vanishing.tex
Intro with more implications

@@
 pigeonhole principle.
+
+\begin{figure}\label{fig:eigenvectors}
+    \centering
+    \includegraphics[width=\textwidth]{eigenvectors_dst_scaled.png}
+    \caption{D-optimal measurement locations ($m=4$ measurements) and
+      weighted eigenvectors for finding the initial condition of the
+      1D heat equation. Measurement locations and weighted
+      eigenvectors are plotted over the computational domain $\Omega =
+      [0, 1]$ (x-axis). Measurement clusterization occurs
+      approximately at $0.31$ and $0.69$. These two locations are a
+      compromise between the magnitudes of the first and second
+      eigenvectors. These are the eigenvectors a D-optimal design aims
+      to measure. The rest of the eAllocating $m=4$ measurements into two locations
+      results in clusterization, according to the pigeonhole
+      principle.}
+  \label{fig:why}
+\end{figure}

### 2024-07-24 latex/intro.tex
badsf

@@
 correspond to any specific real-life problem. Specifically, in its
-current form, our model does not allow point evaluations, see the
-Discussion for a deeper analysis of this subject. Our model is generic
-enough to be analytically tractable, but one may argue our model is
-too far removed from any real application. To these claims I would
-answer that scientists have a long history of studying models that are
-bare-bones simplifications of real systems, e.g. the Ising model
-\cite{cipra1987}, the Lorenz system \cite{brin}, the Lotka-Volterra
-equations \cite{logan2006}, the Carnot engine \cite{kardar2007}, and
-many others.
+current form, our model does not allow point evaluations. Thus, while
+our model is generic enough to be analytically tractable, some may
+argue that our model is too far removed from any real application. To
+these claims I would answer that scientists have a long history of
+studying models that are bare-bones simplifications of real systems,
+e.g.~the Ising model \cite{cipra1987}, the Lorenz system \cite{brin},
+the Lotka-Volterra equations \cite{logan2006}, the Carnot engine
+\cite{kardar2007} and many others.

### 2024-07-24 latex/intro.tex
Moved laplacian like discussion to intro

@@
 regularization via a fully Bayesian formulation of the inverse
-problem, see Section \ref{section:prelims} for details.
+problem.
+
+Specifically, we set a prior on the initial condition $u_0 \sim
+\mathcal{N}(0, (-\Delta)^{-1})$, where we impose the homogeneous
+Dirichlet boundary condition on the Laplacian. We choose
+$(-\Delta)^{-1}$ as our covariance operator following \cite{Stuart10}:
+our choice ensures the power of the negative Laplacian inside the
+parentheses is larger than $d/2$, where $d=1$ is the dimension of our
+computational domain $\Omega$. One can intuitively understand this
+choice from Weyl's asymptotics, that states that the eigenvalues of
+Laplacian increase as $n^{d/2}$. Thus, our choice of covariance
+$(-\Delta)^{-1}$ ensures higher modes decay sufficiently quickly, so
+prior samples have sufficient regularity. Our choice of prior
+covariance ensures the posterior is well-defined and well-behaved
+\cite[Theorem 3.1 and Lemma 6.25]{Stuart10}.
 

### 2024-07-24 latex/prelims.tex
Moved laplacian like discussion to intro

@@
 ,\prcov)\) with some appropriate covariance operator \(\prcov\) on
-\(\hilp\) \cite{Stuart10}. For the inverse problem of the 1D heat
-equation we choose $\prcov = -\Delta^{-1}$ --- see Section
-\ref{subsec:1d heat} for details and a short discussion of
-"Laplacian-like" covariance operators. Note that \(\fwd \prcov
-\fwd^*\) is the prior covariance in \(\hilo\) \cite{Stuart10}, and as
-such is assumed invertible --- an assumption which we will use later
-(if \(\fwd\) has a nontrivial kernel we utilize Occam's Razor and
-ignore said kernel altogether).
+\(\hilp\) \cite{Stuart10}. For example, for the inverse problem of the
+1D heat equation we chose $\prcov = -\Delta^{-1}$ --- see Section
+\ref{subsec:toy}. Note that \(\fwd \prcov \fwd^*\) is the prior
+covariance in \(\hilo\) \cite{Stuart10}, and as such is assumed
+invertible --- an assumption which we will use later (if \(\fwd\) has
+a nontrivial kernel we utilize Occam's Razor and ignore said kernel
+altogether).
 
@@
 this study, but the posterior covariance operator $\postcov$ is given
-by \cite{Stuart10}:
+by the known formula\cite{Stuart10}:
 \begin{align}\label{eq:postcov}

### 2024-07-25 latex/flat.tex
Flattening script

--- 2024-07-21 latex/flat.tex
@@
 \documentclass[ba]{imsart}
-%\input{definitions}
 \usepackage{amsmath}
@@
 %% \usepackage{amsaddr}
-
-%% \theoremstyle{plain}
-%% \newtheorem{theorem}{Theorem}
-%% \newtheorem{corollary}[theorem]{Corollary}
-%% \newtheorem{lemma}[theorem]{Lemma}
-%% \newtheorem{proposition}[theorem]{Proposition}
-%% \theoremstyle{definition}
-%% \newtheorem{definition}[theorem]{Definition}
-%% \newtheorem{example}[theorem]{Example}
-%% \newtheorem{conjecture}[theorem]{Conjecture}
-%% \theoremstyle{remark}
-%% \newtheorem{remark}[theorem]{Remark}
 
@@
 \newcommand{\pr}{\mu_{\textup{pr}}} %Prior measure
-\newcommand{\post}{\mu_{\textup{post}}^{\data, \obs}} % Posterior measure
+%\newcommand{\post}{\mu_{\textup{post}}^{\data, \obs}} % Posterior measure
+\newcommand{\post}{\mu_{\textup{post}}} % Posterior measure for slides
 \newcommand{\prmean}{\param_{\textup{pr}}} % Prior mean
@@
 
+%% \newcommand{\opt}{\mathfrak{D}}
+\newcommand{\opt}{\mathcal{D}}
+\newcommand{\postopt}{\mu_{\textup{post}}^{\data, \opt}} % Posterior measure
+
 %% %\usepackage{comment}% http://ctan.org/pkg/comment
@@
 
+\theoremstyle{plain}
+\newtheorem{theorem}{Theorem}
+\newtheorem{corollary}[theorem]{Corollary}
+\newtheorem{lemma}[theorem]{Lemma}
+\newtheorem{proposition}[theorem]{Proposition}
+\theoremstyle{definition}
+\newtheorem{definition}[theorem]{Definition}
+\newtheorem{example}[theorem]{Example}
+\newtheorem{conjecture}[theorem]{Conjecture}
+\theoremstyle{remark}
+\newtheorem{remark}[theorem]{Remark}
 %
@@
 \begin{abstract}
-  %\input{abstract}
-  Estimation of parameters in physical processes often demands costly
-  measurements, prompting the pursuit of an optimal measurement
-  strategy. Finding such strategy is termed the problem of
-  \emph{optimal experimental design}, abbreviated as optimal
-  design. Remarkably, optimal designs can yield tightly clustered
-  measurement locations, leading researchers to fundamentally revise
-  the design problem just to circumvent this issue. Some authors
-  introduce error correlation among error terms that are initially
-  independent, while others restrict measurement locations to a finite
-  set of locations. While both approaches may prevent clusterization,
-  they also fundamentally alter the optimal design problem.
-
-  In this study, we consider Bayesian D-optimal designs, i.e.~designs
-  that maximize the expected Kullback-Leibler divergence between
-  posterior and prior. We propose an analytically tractable model for
-  D-optimal designs over Hilbert spaces. In this framework, we make
-  several key contributions: \textbf{(a)} We establish that
-  measurement clusterization is a generic trait of D-optimal designs
-  with independent Gaussian measurement errors, and prove that
-  introducing correlations among measurement error terms mitigates
-  clusterization. \textbf{(b)} We characterize D-optimal designs as
-  reducing uncertainty across a subset of prior covariance
-  eigenvectors. Finally, \textbf{(c)} We leverage this
-  characterization to argue that measurement clusterization arises as
-  a consequence of the pigeonhole principle: when more measurements
-  are taken than there are locations where the select eigenvectors are
-  large and others are small --- clusterization occurs.
+Estimation of parameters in physical processes often demands costly measurements, prompting the pursuit of an optimal measurement strategy.
+Finding such strategy is termed the problem of \emph{optimal experimental design}, abbreviated as optimal design.
+Remarkably, optimal designs can yield tightly clustered measurement locations, leading researchers to fundamentally revise the design problem just to circumvent this issue.
+Some authors introduce error correlation among error terms that are initially independent, while others restrict measurement locations to a finite set of locations.
+While both approaches may prevent clusterization, they also fundamentally alter the optimal design problem.
+
+In this study, we consider Bayesian D-optimal designs, i.e.~designs that maximize the expected Kullback-Leibler divergence between posterior and prior.
+We propose an analytically tractable model for D-optimal designs over Hilbert spaces.
+In this framework, we make several key contributions: \textbf{(a)} We establish that measurement clusterization is a generic trait of D-optimal designs with independent Gaussian measurement errors, and prove that introducing correlations among measurement error terms mitigates clusterization.
+\textbf{(b)} We characterize D-optimal designs as reducing uncertainty across a subset of prior covariance eigenvectors.
+Finally, \textbf{(c)} We leverage this characterization to argue that measurement clusterization arises as a consequence of the pigeonhole principle: when more measurements are taken than there are locations where the select eigenvectors are large and others are small --- clusterization occurs.
 %
@@
 %% fundamental attributes underlying measurement design optimization.
-
 \end{abstract}
@@
 
-%\input{intro}
 \section{Introduction}\label{section:intro}
-Measurements play a fundamental role in generating observations and
-are indispensable to any inference process. Selecting the optimal set
-of measurements holds particular significance when inferring
-parameters of a physical process. Unlike many other fields, where
-measurements are fixed, in this context, we have the freedom to choose
-which measurements to take. This freedom should be taken advantage of:
-measurements should be chosen to enhance accuracy, reduce costs, or
-both. Whether measurements involve specifying electrode locations on
-the skin in electric impedance tomography \cite{horesh2010impedance},
-determining certain wavelengths in MRI \cite{horesh2008mri},
-positioning sensors for detecting ground-reflected waves in the search
-for oil \cite{horesh2008borehole}, or optimizing sensor locations in a
-seismographic network \cite{rabinowitz1990, steinberg1995} --- the
-selection of optimal measurements, referred to as the problem of
-\emph{optimal design}, becomes crucial. Optimal measurements are
-chosen according to specific \emph{design criteria}, with A- and
-D-optimality being two of the most widely recognized and extensively
-studied design criteria \cite{Chaloner1995,
-  AlexanderianGloorGhattas14}.
-
-Surprisingly, A- and D-optimal designs have been observed to yield
-remarkably similar measurements in certain cases \cite{fedorov1996,
-  nyberg2012, fedorov1997, Ucinski05, neitzel2019sparse}. To
-illustrate this point, we consider a toy problem: inferring
-the initial condition of the 1D heat equation (details in the
-supplementary material). In Fig.~\ref{fig:clusterization
-  illustration} D-optimal measurement locations are shown for
-different numbers of measurements. Notably, for six measurements, a
-D-optimal design yields two sets of measurements that are almost
-indistinguishable from one another. We refer to this intriguing
-phenomenon as \emph{measurement clusterization} \cite{Ucinski05}, and
-we refer to a design that exhibits measurement clusterization as a
-\emph{clustered design}.
+The practice of experimental design entails specifying every detail of
+running an experiment \cite{chaloner1995} and at its simplest form,
+experimental design involves choosing a set of measurements to be
+taken. For example, consider the process of searching for oil:
+measuremnts require digging holes in the ground and a \emph{design}
+simply indicates which holes should be dug
+\cite{horesh2008borehole}. Since measurements are expensive, only a
+limited number of measuremnts can be taken, and thus carefully
+choosing measuremt locations is crucial.
+
+Specifiyng the right set of measurements holds particular significance
+when solving an \emph{inverse problem} --- i.e.~making inference of
+the physical world utilizing a physical model of a phenomenon of
+interest \cite{tarantola2005,kaipio2005}. In an inverse problem, we
+seek to infer physical quantities, based on observations and a model
+of the physical world. Models are typically phrased in the language of
+ordinary or partial differential equations.
+
+Inverse problems are ubiquitous in science and engineering. For
+example, in electrical impedance tomography (EIT) we seek to infer the
+structure of the inside of a human body. Inference is conducted based
+on observations of electrical current impedance measured at specific
+electrode locations on the skin, leveraging known equations of
+electric current flow \cite{horesh2010impedance}. Magnetic Resonance
+Imaging (MRI) also entails solving an inverse problem. There,
+radio-frequency pulses are sent through the human body in the presence
+of a strong magnetic field. The response of one's body contents to
+these radio-frequency pulses is measured, allowing a radiologist to
+view the internals of a body in a noninvasive manner
+\cite{horesh2008mri}. An inverse problem also arises in oil and gas
+exploration, where acoustic waves are sent through "boreholes" ---
+deep cylindrical wells drilled into the ground. Data of travel and
+return times of the acoustic waves are recorded. Wave travel and
+return time is influenced by the properties of the subsurface
+materials, such as density, elasticity, and the presence of fluids or
+voids. Combining travel time data with a geophisical models of
+contents of the earth's crust facilitates reconstructing the structure
+of the subsurface in a process called \emph{borehole tomography}
+\cite{horesh2008borehole}. Inverse problems also arise in many other
+areas of seismology and geology \cite{rabinowitz1990, steinberg1995}
+and medical imaging \cite{tarantola2005}. In many inverse problems the
+goal is to infer some numerical parameter. However, in the formulation
+we consider in this study, as well as in the examples above, the goal
+is to conduct inference over some \emph{function} over $\Omega$, where
+\(\Omega \subseteq \mathbb{R}^d, d=1,2,3\) is a spatial domain of
+interest.
+
+\subsection{D-optimal Designs}\label{subsec:D}
+In all of the examples of inverse problems we have seen, as well as in
+many other applications, only a limited number of measurements are
+allowed. For example, in borehole tomography, a measurement involves
+drilling a deep well in the ground --- a very costly endeavor. Even in
+a seemingly simpler application such as MRI, a patient cannot occupy
+the MRI machine for too long, hence only a limited number of
+measurements can be taken. Therefore, measurements should be chosen to
+most benefit the user. Typically, a user will consider some utility,
+called a \emph{design criterion} and take measurements that maximize
+this utility. Two of the most widely recognized and extensively
+studied design criteria are A- and D-optimality.
+
+In this study, we focus on the Bayesian D-optimality
+criterion. Bayesian D-optimality carries a simple and intuitive
+meaning: measurements chosen according to the D-optimality criterion
+maximize the expected Kullback-Liebler (KL) divergence from posterior
+to prior \cite{chaloner1995, AlexanderianGloorGhattas14,
+  CoverThomas91}. Utilizing KL divergence for quantifying differences
+between distributions is standard practice and we will not motivate it
+further here. For a discrete parameter $\param$ and data $\data$, the
+KL divergence is defined as
+\begin{equation}\label{eq:basic KL}
+  D_{KL}\left (\Pr(\param|\data)||\Pr(\param)\right ) =  \sum_{\param} \log
+  \frac{\Pr(\param|\data)}{\Pr(\param)} \Pr(\param|\data). 
+\end{equation}
+Of course, data $\data$ is not known before the experiment is
+conducted, so we average over it to define the D-optimality criterion
+as:
+\begin{equation*}
+  \mathbb{E}_{\data}\left [ D_{KL}\left (\Pr(\param|\data)||\Pr(\param)\right ) \right ]
+\end{equation*}
+We refer to a set of measuremnts that maximizes the D-optimality
+criterion as a \emph{D-optimal design}.
+
+For a linear model with Gaussian prior in finite dimensions, a
+D-optimal design minimizes the determinant of the posterior covariance
+\cite{chaloner1995}. In Section \ref{subsec:D optimal design} we give
+a more general definition of the D-optimality criterion that applies
+to arbitrary measures. We also show how the D-optimality criterion
+generalizes to linear models over infinite-dimensional Hilbert spaces
+(i.e.~function spaces).
+
+\subsection{A toy model: the 1D heat equation}\label{subsec:toy}
+For concreteness, let us now consider a toy model of a physical
+process for which we wish to solve an inverse problem: a partial
+differential equation known as the \emph{heat equation} in one
+dimension (1D heat equation henceforth). Readers less familiar with
+partial differential equations can think of the following setup:
+before us there is a perfectly insulated metal rod. At each point on
+the rod the temperature is different and unknown, and we call this
+temperature distribution the \emph{initial condition}. We wait for a
+short time $T$ and let the heat dissipate a bit inside the rod. The
+heat equation determines the heat distribution inside the rod at time
+$t=T$.
+
+The full time evolution of heat in the rod $\Omega=[0,1]$ is formally
+described by the following three equations:
+\begin{subequations}%\label{eq:heat equation}
+  \begin{alignat}{2}
+    u_t &= \Delta u &&\qquad \text{in } [0,1] \times [0,\infty), \label{eq:heat1}\\
+    u &= 0 &&\qquad \text{on } \{0, 1\} \times [0,\infty), \label{eq:heat2}\\
+    u &= u_0 &&\qquad \text{on }[0,1] \times \{0\}. \label{eq:heat3}
+  \end{alignat}
+\end{subequations}
+As time passes, heat disspates across the rod as hotter regions,
+e.g.~local maximas of the heat distribution, become cooler. This
+behavior is captured by eq.~\eqref{eq:heat1}: the Laplacian at local
+maxima is negative, so $u_t = \Delta u$ implies $u$ decreases at its
+maximas, and the reverse happens at local minima. Eq.~\eqref{eq:heat2}
+describes the \emph{boundary condition}, which dictates how heat
+interacts with the outside world at the two edges of the rod
+$\{0,1\}$. Specifically, eq.~\eqref{eq:heat2} implements an absorbing
+boundary, i.e.~heat that interacts with the boundary immediately
+disappears. This type of boundary condition is known as a
+\emph{homogeneous Dirichlet boundary condition}. Eq.~\eqref{eq:heat3}
+describes the \emph{initial condition}, i.e.~the heat distribution
+inside the rod at time $t=0$.
+
+Solving the 1D heat equation is straightforward. Any initial heat
+distribution $u_0\in L^2([0,1])$ that satisfies the homogeneous
+Dirichlet boundary condition is a linear combination of sines
+$\ev_n(x) = \sin(2\pi n x)$. Now, note that $\ev_n$ are in fact
+eigenvectors of the Laplacian: $\Delta \ev_n = -(2\pi n)^2\ev_n$. The
+linearity of the Laplacian implies the linearity of the heat equation
+and we can solve in Fourier / frequency space. It is important to note
+that when solving the heat equation we \emph{invert} the
+Laplacian. I.e.~eigenvalues of the linear operator that describes the
+time evolution of heat are $(2\pi n)^{-2}$.
+
+In the \emph{inverse problem of the 1D heat equation}, our goal is to
+infer the initial condition $u_0 = u(\cdot, 0)$ from observations of
+the final condition $u(\cdot, T)$. However, we are not able to measure
+$u(\cdot, T)$ at every point in the domain $\Omega$. Rather, we take
+some number of temperature measurements on the rod once our waiting
+time $T$ has passed and try to infer $u(\cdot, 0)$ from these
+measurements.
+
+The inverse problem of the 1D heat equation is difficult ("ill-posed")
+since heat spreads in a diffusive manner and any roughness in the
+initial condition is quickly smoothed, see Supplementary movies S1 and
+S2 (code generating these movies is located in module
+\texttt{movies.py} in the accompanying
+\href{https://github.com/yairdaon/OED}{repository}). For ill-posed
+problems regularization is required, and we implement such
+regularization via a fully Bayesian formulation of the inverse
+problem.
+
+Specifically, we set a prior on the initial condition $u_0 \sim
+\mathcal{N}(0, (-\Delta)^{-1})$, where we impose the homogeneous
+Dirichlet boundary condition on the Laplacian. We choose
+$(-\Delta)^{-1}$ as our covariance operator following \cite{Stuart10}:
+our choice ensures the power of the negative Laplacian inside the
+parentheses is larger than $d/2$, where $d=1$ is the dimension of our
+computational domain $\Omega$. One can intuitively understand this
+choice from Weyl's asymptotics, that states that the eigenvalues of
+Laplacian increase as $n^{d/2}$. Thus, our choice of covariance
+$(-\Delta)^{-1}$ ensures higher modes decay sufficiently quickly, so
+prior samples have sufficient regularity. Our choice of prior
+covariance ensures the posterior is well-defined and well-behaved
+\cite[Theorem 3.1 and Lemma 6.25]{Stuart10}.
+
+\subsection{Measurement Clusterization}
+Surprisingly, A- and D-optimal designs for inverse problems have been
+observed to yield remarkably similar measurements in certain cases
+\cite{fedorovDesignSpatialExperiments1996, nyberg2012, fedorov1997,
+  Ucinski05, neitzel2019sparse}. This phenomenon is illustrated for
+our toy inverse problem in Fig.~\ref{fig:clusterization illustration},
+where D-optimal measurement locations are shown for different numbers
+of measurements. Notably, for six measurements, a D-optimal design
+yields two sets of measurements that are identical. Following
+\cite{Ucinski05}, we refer to this intriguing phenomenon as
+\emph{measurement clusterization}. We consider a design to be
+\emph{clustered} when two or more of its constituent measurements are
+identical.
 
@@
     \centering
-    \includegraphics[height=0.5\textwidth]{example.pdf}
-    \caption{Measurement clusterization for optimal designs when
-      inverting for the initial condition of the 1D heat equation (see
-      supplementary material for details). Measurement locations were
+    \includegraphics[height=0.5\textwidth]{dst_modelError0.png}
+    \caption{Measurement clusterization for D-optimal designs for the
+      inverse problem of the heat equation. Measurement locations were
       chosen according to the Bayesian D-optimality criterion of
@@
 
-Researchers widely agree that measurement clusterization is
-undesirable \cite{fedorov1996, nyberg2012, fedorov1997, Ucinski05,
-  neitzel2019sparse}, prompting the exploration of various remedies to
-address this issue. One approach involves merging close measurements
-\cite{fedorov1997}; however, this strategy merely overlooks the
-phenomenon of measurement clusterization. An alternative solution lies
-in \emph{clusterization-free design}s, where measurement locations are
-deliberately chosen to be distant from one another. This can be
-achieved by imposing distance constraints between measurements or by
-introducing correlated errors that account for both observation error
-and model misspecification \cite{Ucinski05}. For instance, in the
-context of time-series analysis for pharmacokinetic experiments,
+Clusterization should not be confused with either repetition nor with
+replication, which are commonly viewed as beneficial and even
+necessary aspects of an optimal design \cite{fisher1949design,
+  morris2011, schafer2001replication}. For example,
+\cite{fisher1949design} in his famous milk and tea experiment,
+suggested that repetition is a "way of enlarging the experiment and,
+thereby, increasing its sensitiveness". In another example,
+\cite{fay2000rainfall} measured the effect of rainfall on grass growth
+in plots of land. The experiment involved fifteen "rainfall
+manipulation shelters", where "Four rainfall manipulation treatments
+(three replicates) then were assigned to 12 of the plots". While it
+seems reasonable for the researchers to replicate the phenomenon they
+are trying to study, clusterization is different: a clustered design
+in the rainfall experiment would imply the researchers should take a
+repeated measurement \emph{on the same plot}, at the expense of
+measuring grass growth in other plots. In sharp contrast to repetition
+and replication, clusterization is highly nonintuitive and we will see
+that its origins run considerably deeper.
+
+Researchers of inverse problems widely agree that measurement
+clusterization is undesirable
+\cite{fedorovDesignSpatialExperiments1996, nyberg2012, fedorov1997,
+  Ucinski05, neitzel2019sparse}, prompting the exploration of various
+remedies to address this issue. One approach involves merging close
+measurements \cite{fedorov1997}; however, this strategy merely
+overlooks the phenomenon of measurement clusterization. An alternative
+solution lies in \emph{clusterization-free design}s, where measurement
+locations are deliberately chosen to be distant from one another. This
+can be achieved by imposing distance constraints between measurements
+or by introducing correlated errors that account for both observation
+error and model misspecification \cite{Ucinski05}. For instance, in
+the context of time-series analysis for pharmacokinetic experiments,
 measurement clusterization can be mitigated by incorporating the
@@
 \subsection{Contribution}
-The primary objective of this study is to provide a comprehensive
-understanding of measurement clusterization by addressing the
-aforementioned questions. Our focus centers around investigating the
-Bayesian D-optimality criterion, which involves maximizing the
-expected Kullback-Leibler divergence between the posterior and prior
-measures \cite{CoverThomas91, Chaloner1995}. We conduct an analysis of
-Bayesian D-optimal designs within the context of linear inverse
-problems over Hilbert spaces. We propose a novel relaxed model for
-D-optimality that maintains analytical tractability and enables the
-identification of D-optimal designs using Lagrange multipliers. This
-analytical framework facilitates the exploration of the questions
-posed at the end of the previous paragraph:
-
-\begin{figure}
-    \centering
-    \includegraphics[width=\textwidth]{eigenvectors.pdf}
-    \caption{D-optimal measurement locations ($m=4$ measurements) and
-      weighted eigenvectors for finding the initial condition of
-      the 1D heat equation. Measurement locations and weighted
-      eigenvectors are plotted over the computational domain $\Omega =
-      [0, 1]$ (x-axis). Measurement clusterization occurs
-      approximately at $0.31$ and $0.69$. These two locations are a
-      compromise between zeros of eigenvectors a D-optimal design aims
-      to ignore (third and up) and staying far from zeros of the
-      eigenvectors a D-optimal design aims to measure (first and
-      second). Allocating $m=4$ measurements into two locations
-      results in clusterization, according to the pigeonhole
-      principle.}
-  \label{fig:why}
-\end{figure}
+The primary objective of this study is to provide a deep understanding
+of measurement clusterization by addressing the aforementioned
+questions. Our focus centers around investigating the Bayesian
+D-optimality criterion. We conduct an analysis of Bayesian D-optimal
+designs within the context of linear inverse problems over Hilbert
+spaces and study two inverse problems: (a) In Sections
+\ref{section:prelim} and \ref{section:D and grad} we propose a novel
+relaxed model for an inverse problem where D-optimality maintains
+analytical tractability, where D-optimal designs are identified via
+Lagrange multipliers. This analytical framework facilitates the
+exploration of the questions posed at the end of the previous
+paragraph. We also study (b) the inverse problem of the 1D heat
+equation from Section \ref{subsec:toy} above. Investigating both
+inverse problems allows us to answer the questions posed in the
+previous section:
 
 \begin{enumerate}
-
 \item \label{q:generic} \textbf{Is measurement clusterization a
@@
   %% as a solution to the D-optimal design problem.  Furthermore,
-  Randomized numerical simulations of our model give rise to D-optimal
-  designs that exhibit clusterization more than 95\% of the time (see
-  code in supplementary material). Given our model's genericity, we
-  expect measurement clusterization to be a generic phenomenon.
+  We give two complementing answers to this question. First, from a
+  theoretical perspective, we show that clusterization mainly depends
+  on how quickly the eigenvalues of the prior measure in observation
+  space decay. This decay depends mostly on how ill-posed the problem
+  at hand is and does not depend much on the prior. See Section
+  \ref{section:vanishing} --- particularly, Theorem \ref{thm:char} and
+  the discussion following it. We also show results of numerical
+  experiments, where simulations of our model give rise to D-optimal
+  designs that exhibit clusterization with high probability in Section
+  \ref{subsec:lemma sims}. Thus, given the genericity of our model, we
+  expect measurement clusterization to be a generic and ubiquitous
+  phenomenon.
 
@@
   role of model error in mitigating clusterization, thereby
-  corroborating earlier observations made by various researchers.
+  corroborating earlier observations made by various
+  researchers. Specifically, our proof shows that identical
+  measurements result in no gain in design criterion when observation
+  error tends to zero. Moreover, in Section \ref{subsec:corr errors
+    sims}, we show that an error term corresponding to correlations
+  between measurements mitigates clusterization in the inverse problem
+  of the 1D heat equation.
 
 \item \label{q:why} \textbf{Why does measurement clusterization
-  occur?} In Section \ref{section:why}, we present a compelling
-  explanation for the optimality of clustered designs in the absence
-  of model error. Our analysis reveals that, in our model, a D-optimal
-  design focuses on a select set of prior eigenvectors, specifically
-  those with the largest eigenvalues in the prior covariance
-  spectrum. In practical scenarios, the number of locations where (a)
-  the relevant prior eigenvectors are significantly large, and (b)
-  other eigenvectors are close to zero, is limited. Consequently, the
-  clusterization of measurements arises as a natural consequence of
-  the pigeonhole principle, as there are more measurements available
-  than there are locations satisfying conditions (a) and (b). See
-  Fig~\ref{fig:why}.
+  occur?} In Section \ref{subsec:why} (specifically, See
+  Fig~\ref{fig:why}) we give a compelling answer to this question by
+  transporting insights we gain from our analytically tractable
+  relaxed model to the inverse problem of the 1D heat equation. Our
+  analysis reveals that a D-optimal design focuses on a select set of
+  prior eigenvectors, specifically those with the largest eigenvalues
+  in the prior covariance spectrum. In practical scenarios, the number
+  of locations where (a) all relevant prior eigenvectors are
+  significantly large, and (b) other eigenvectors are close to zero,
+  is limited. Consequently, the clusterization of measurements arises
+  as a natural consequence of the pigeonhole principle, as there are
+  more measurements available than there are locations satisfying
+  conditions (a) and (b).
 
@@
 %%   designs.
-
 \end{enumerate}
 
-The cornerstone of our investigation into measurement clusterization
-is Theorem \ref{thm:char} proven in Section
-\ref{section:vanishing}. The key insight of Theorem \ref{thm:char}
-lies in its fifth component, which highlights that a D-optimal design
-aims to uniformly reduce posterior uncertainties for posterior
-covariance eigenvectors in observation space. A similar conclusion was
-reached by Koval et al.~\cite{koval2020}, who showed that A-optimal
+\subsubsection{Implications}
+Our answer to Question \ref{q:generic} implies that encountering
+clusterization should be expected in many different problems across
+many different scientific fields. Thus, researchers that encounter
+clusterization should not be surprised or wary. In Our answer to
+Question \ref{q:why}, we explain what our view of the cause of
+clusterization. As it appears, the cause is generic (see Question
+\ref{q:why} above): a D-optimal design reduces uncertainty for a
+select set of prior covariance eigenvectors --- those with the most
+prior uncertainty, i.e.~those the practitioner cares about the most!
+Consequently, practitioners should not try to avoid measurement
+clusterization. Rather, practitioners should take repeated
+measurements (e.g.~in MRI and borehole tomography), increase apparatus
+sensitivity (e.g.~in EIT), or take consecutive measurements (e.g.~in
+the 1D heat equation). Overall, we show that measurement
+clusterization is a natural and (almost) inevitable part of Bayesian
+D-optimal designs.
+
+One interesting implication of the analysis presented here is that
+clusterization can be expected once the number of measurements exceeds
+the number of relevant eigenvectors. Since leading eigenvectors
+typically correspond to slow variations in space and/or time,
+clusterization could be used to estimate the number of relevant
+degrees of freedom, and even to reduce the complexity of a
+computational model, e.g.~by dropping discretization points.
+
+It is important to note that we do not view measurement clusterization
+as undesirable, nor do we believe it should be avoided at
+all. Clusterization is a peculiar phenomenon and it is perfectly
+reasonable for someone to argue against the D-optimality criterion
+based on the fact that it results in clustered designs. We have seen,
+however that there is perfectly reasonable explanation for
+clusterization, i.e.~that clusterization arises when we seek a
+designWe have shown that clusterization is an innevitable conseqeunce
+of having a problem with modes of higher uncertainty than
+others. Clusterization manifests the true underlying objective
+
+Lastly, we believe that when clusterization arises, it should serve as
+a warning sign to practitioners. In the inverse problem of the 1D heat
+equation, clusterization occurs primarily because Laplacian
+eigenvectors are nonzero almost everywhere in $\Omega$. Consequently,
+measuring $u(x_1, T)$ at some point $x_1 \in \Omega$ provides
+information about $u(x_2,T)$ for distant points $x_2 \in
+\Omega$. Intuitively, this should not occur: for small $T$, the heat
+distribution at $x_1$ should have minimal influence on the heat
+distribution at $x_2$.
+
+This behavior stems from a well-known property of the heat equation:
+it allows information to spread instantly across the computational
+domain \cite{renardy2006PDE}. In reality, heat and information
+propagate at finite speeds. The emergence of clusterization in this
+context is thus non-physical, arising from the nature of Laplacian
+eigenvectors being nonzero almost everywhere. Clusterization therefore
+indicates that the underlying mathematical model is overly permissive
+and fails to capture crucial physical constraints. We suggest that
+when clusterization occurs, practitioners should consider alternative
+models where information is localized in space and travels at finite
+speeds. Such models may not only provide more physically accurate and
+meaningful results but may also mitigate the issue of clusterization.
+
+\subsubsection{Other Contributions}
+In Theorem \ref{thm:char}, we also show that D-optimal designs are
+best understood in the sapce of \emph{observations}, see Section
+\ref{section:vanishing} for a precise statement. This is in accordance
+with previous work by \cite{koval2020}, who showed that A-optimal
 designs are best constructed in the space of observations.
-
-Before we state Theorem \ref{thm:char} we give some definitions: Let
-\(\hilp, \hilo\) Hilbert spaces, \(\fwd:\hilp \to \hilo\) a linear
-compact operator. Let \(\obs: \hilo \to \mathbb{R}^m\) a linear
-measurement operator, where \(m \in \mathbb{N}\) is the number of
-measurements taken. Let \(\sigma^2 \in \mathbb{R}_{+}\) observation
-noise variance, \(\data = \obs \fwd \param + \eps\), where \(\eps \in
-\mathbb{R}^m\) is iid \(\mathcal{N}(0, \sigma^2)\) noise. Let \(\pr
-\sim \mathcal{N}(\prmean, \prcov)\) prior Gaussian measure on
-\(\hilp\), where \(\prcov\) is the prior covariance operator and let
-\(\post\) the posterior (Gaussian) measure.
-
-\begin{restatable}{theorem}{char}\label{thm:char}
-  Let:
-  \begin{itemize}
-    \item The D-optimality design criterion
-    \cite{AlexanderianGloorGhattas14}:
-    \begin{align*}
-      \begin{split}
-        \tar(\obs) %:&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
-        % 
-        % 
-        % 
-        &= \frac12 \log \det ( I + \sigma^{-2} \prcov^{1/2} \fwd ^*
-        \obs^* \obs \fwd \prcov^{1/2}), 
-      \end{split}
-    \end{align*}
-  \item \(\obs\) a D-optimal design operator
-    \begin{equation*}
-      \obs = \argmax_{\|\meas_j\| = 1, j=1,\dots,m}\tar(\obs),
-    \end{equation*}
-  \item \(\{\lambda_i\}_{i=1}^\infty\) eigenvalues of
-    \(\fwd\prcov\fwd^*\) in decreasing order of magnitude.
-    %% \item \(\{\ev_i\}_{i=1}^\infty\) their corresponding eigenvectors.
-  \item \(\{\eta_i\}_{i=1}^\infty\) eigenvalues of \(\obs^*\obs\).
-
-  \end{itemize}
-
-  Then:
-  \begin{enumerate}
-  \item  \(\tr{\obs^*\obs} = m\).
-  \item \(\obs^*\obs\) and \(\fwd\prcov\fwd^*\) are simultaneously
-    diagonalizable.
-  \item \(k := \rank \obs^*\obs \leq m\) and
-    \begin{equation*}
-      \tar(\obs) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i). %= \frac12 \sum_{i=1}^{m} \log (1 + \sigma^{-2}\lambda_i\eta_i).
-    \end{equation*}
-  %% \item 
-  %%   \begin{equation*}
-  %%     k = \argmax \left \{ k:\lambda_k^{-1} < \sigma^{-2}\frac{m}{k} + \frac{1}{k} \sum_{j=1}^{k}
-  %%     \lambda_j^{-1} \right \}.
-  %%   \end{equation*}
-  \item
-    \begin{equation*}
-        \eta_i = \begin{cases}
-          \frac{m}{k} - \sigma^2 \lambda_i^{-1} + \sigma^2 \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1} & 1 \leq i \leq k \\
-          0 & i > k 
-        \end{cases}.
-    \end{equation*}
-  \item The covariance of the pushforwad \(\fwd_{*} \post\) is \(\left
-    ( (\fwd \prcov \fwd^*)^{-1} + \sigma^{-2} \obs^*\obs \right
-    )^{-1}\) and its eigenvalues are
-    \begin{equation*}
-      \theta_i =
-      \begin{cases}
-        \left(\frac{\sum_{j=1}^k \lambda_j^{-1} + \sigma^{-2}m}{k} \right )^{-1} & i \leq k \\
-        \lambda_i &  i > k 
-      \end{cases}
-    \end{equation*}
-  \end{enumerate}
-%\end{theorem}
-\end{restatable}
 
@@
 generalize several lemmas. Among those, is Lemma \ref{lemma:free},
-which is (to the the best of my knowledge) a novel lemma in linear
-algebra: We decompose a symmetric positive definite matrix \(M \in
-\mathbb{R}^{k \times k}\) with \(\ttr M = m \in \mathbb{N}\) as \(M =
-AA^t\), where \(A\) has unit norm columns.
+which is (to the the best of my knowledge) novel: We decompose a
+symmetric positive definite matrix \(M \in \mathbb{R}^{k \times k}\)
+with \(\ttr M = m \in \mathbb{N}\) as \(M = AA^t\), where \(A\in
+\mathbb{R}^{k \times m}\) has unit norm columns.
 
@@
 
+%% Finally, we must stress that we do not advocate any particular
+%% solution to clusterization. We do not suggest that clustered designs
+%% should be abandoned, nor do we support the use of clustered designs in
+%% any circumstances. Our goal is to investigate an intriguing phenomenon
+%% of clusterization and seek a deeper understanding for it. Different
+%% researchers in different disciplines have differenot viewed as a problem. For
+%% example, \cite{lozan} suggest in the introduction section that
+%% repeated measurements are necessary. Our results align with that
+%% common wisdom and settle the seemed contradiction.
+
 \subsection{Limitations}\label{subsec:limitations}
 The main limitation of this study is that our generic model does not
-correspond to any specific real-life problem. It is generic enough to
-be analytically tractable, but one may argue our model is too far
-removed from any real application. To these claims I would answer that
-scientists have a long history of studying models that are bare-bones
-simplifications of real systems, e.g. the Ising model
-\cite{cipra1987}, the Lorenz system \cite{brin}, the Lotka-Volterra
-equations \cite{logan2006}, the Carnot engine \cite{kardar2007}, and
-many others.
-
-%\input{prelims}
+correspond to any specific real-life problem. Specifically, in its
+current form, our model does not allow point evaluations. Thus, while
+our model is generic enough to be analytically tractable, some may
+argue that our model is too far removed from any real application. To
+these claims I would answer that scientists have a long history of
+studying models that are bare-bones simplifications of real systems,
+e.g.~the Ising model \cite{cipra1987}, the Lorenz system \cite{brin},
+the Lotka-Volterra equations \cite{logan2006}, the Carnot engine
+\cite{kardar2007} and many others.
 \section{Preliminaries and Notation}\label{section:prelim}
-
-In this section we present the setup that will be used throughout this
-article. The theoretical foundations for inverse problems over
-function spaces can be found elsewhere \cite{Stuart10} and will not be
-reviewed here.
-
-\subsection{Bayesian Linear Inverse Problems}\label{subsec:abstract OED}
-Let \(\hilp\) and \(\hilo\) separable Hilbert spaces (the subscripts p
-and o are for "parameter" and "observation", respectively), and let
-\(\fwd: \hilp \to \hilo\) the \emph{forward operator}. The forward
-operator \(\fwd\) is assumed linear and strongly smoothing (the heat
-operator of the 1D heat equation introduced in Section
-\ref{section:intro} is a prime example). Take a Gaussian prior
-\(\param \sim \pr = \normal(\prmean ,\prcov)\) with some appropriate
-covariance operator \(\prcov\) on \(\hilp\) \cite{Stuart10}. Note that
-\(\fwd \prcov \fwd^*\) is the prior covariance in \(\hilo\)
-\cite{Stuart10}, and as such, \(\fwd \prcov \fwd^*\) is assumed
-invertible --- an assumption which will be used below (if \(\fwd\) has
-a nontrivial kernel we utilize Occam's Razor and ignore said kernel
-altogether). Measurements are taken via a \emph{measurement operator}
-\(\obs\). It is common for the measurement and forward operators to be
-merged \(\tmp := \obs \fwd\) \cite{AlexanderianGloorGhattas14}, but
-the analysis carried out in the following sections requires that
-\(\fwd\) and \(\obs\) are explicitly separated as in
-\cite{attia2022stochastic, cvetkovic2023choosing}.  \(\obs \in (
-\hilo^* )^m\), where \(m\) is the number of measurements
-taken. Entries \(\meas_j, j=1,\dots,m\) of the observation operator
-\(\obs\) are called \emph{measurements}:
+Here we present the basics of Bayesian inverse problems over hilbert
+spaces. Since we are ultimately interested in inferring a function
+over some domain, we will keep in mind that these Hilbert spaces
+should really be thought of as function spaces. A deeper treatment of
+the foundations for inverse problems over function spaces can be found
+in \cite{Stuart10}.
+
+\subsection{Forward Problems}\label{subsec:abstract OED}
+In this section we give definitions and notations for forward problems,
+which are an essential part of the inverse problems we discuss
+later. Consider a "parameter space" \(\hilp\) and an "observation
+space" \(\hilo\) --- both separable Hilbert spaces (the subscripts p
+and o are for "parameter" and "observation", respectively). The
+parameter space includes the quantity we seek to infer; in the inverse
+problem of the heat equation, the parameter space $\hilp$ is where the
+initial condition lives. The observation space $\hilo$, on the other
+hand, is the space from which we take measurements; in the example of
+the 1D heat equation, $u(\cdot, T) \in \hilo$.
+
+The connection between parameter and observation spaces is made by the
+\emph{forward operator} \(\fwd: \hilp \to \hilo\). We assume the
+forward operator \(\fwd\) is linear. In the inverse problem of the 1D
+heat, the forward operator is determined by the time evolution of the
+1D heat equation \eqref{eq:heat1} and \eqref{eq:heat2}.
+
+Measurements are taken via a \emph{measurement operator} \(\obs\). It
+is common for the measurement and forward operators to be merged
+\(\tmp := \obs \fwd\) \cite{AlexanderianGloorGhattas14}, but the
+analysis carried out in the following sections requires that \(\fwd\)
+and \(\obs\) are explicitly separated (as in
+\cite{attia2022stochastic, cvetkovic2023choosing}). Note that \(\obs
+\in ( \hilo^* )^m\), where \(m\) is the number of measurements
+taken\footnote{The alert reader will likely ask how do we reconcile
+point measurements $\delta_x$ as suggested by the formulation of the
+1D heat equation with workin in Hilbert spaces. We don't. This entire
+study could be formulated for Banach spaces where point measurements
+are allowed. We avoid this technical difficulty because: (1) it does
+not add much to the study, (2) it is beyond the technical abilities of
+the author}. Entries \(\meas_j, j=1,\dots,m\) of the measurement
+operator \(\obs\) are called \emph{measurements}:
 
@@
 Taking \(\modcov = 0\) is a common practice
-\cite{tarantola2005,Kaipio2005,Vogel02} and then \(\Sigma =
+\cite{tarantola2005,kaipio2005,Vogel02} and then \(\Sigma =
 \sigma^2I_m\) is a scalar matrix which does not depend on \(\obs\).
 
-Finally, it is useful to record that the posterior measure \(\post\)
-is Gaussian in this setting and its covariance operator does not
-depend on data \(\data\) \cite{Stuart10}:
+\subsection{Bayesian Linear Inverse Problems}\label{subsec:bayesian inverse problems}
+In the previous section, we saw how a parameter $u\in \hilp$ is
+transported to the observation space via the forward operaror $\fwd u
+\in \hilo$, how observations are generated from a parameter via $\obs
+\fwd u$ and how observations and noise give rise to data $\data$. It
+is time to formulate the process of inferring the parameter as a
+Bayesian inverse problem. The first ingredient we need is a prior.
+
+We take a Gaussian prior \(\param \sim \pr = \normal(\prmean
+,\prcov)\) with some appropriate covariance operator \(\prcov\) on
+\(\hilp\) \cite{Stuart10}. For example, for the inverse problem of the
+1D heat equation we chose $\prcov = -\Delta^{-1}$ --- see Section
+\ref{subsec:toy}. Note that \(\fwd \prcov \fwd^*\) is the prior
+covariance in \(\hilo\) \cite{Stuart10}, and as such is assumed
+invertible --- an assumption which we will use later (if \(\fwd\) has
+a nontrivial kernel we utilize Occam's Razor and ignore said kernel
+altogether).
+
+Since $\fwd$ is linear and $\pr$ is Gaussian --- the posterior
+\(\post\) is Gaussian as well. We do not utilize the posterior mean in
+this study, but the posterior covariance operator $\postcov$ is given
+by the known formula\cite{Stuart10}:
 \begin{align}\label{eq:postcov}
@@
 
-\subsection{Bayesian D-Optimal Designs in Infinite Dimensions}\label{subsec:D optimal design} 
-A Bayesian D-optimal design maximizes expected Kullback-Leibler (KL)
-divergence between posterior \(\post\) and prior measures \(\pr\). It is
-useful to first recall the definition of KL divergence \cite{CoverThomas91}:
+\subsection{Bayesian D-Optimal Designs}\label{subsec:D optimal design} 
+A Bayesian D-optimal design maximizes the expected KL divergence
+between posterior \(\post\) and prior measures \(\pr\). For arbitrary
+posterior and prior measures, the KL divergence is defined,
+analogously to eq.~\ref{eq:basic KL}, via the Radon-Nikodym
+derivative:
 \begin{equation*}
@@
 
-%\input{constrained}
 \section{The Constrained Optimization Problem of D-Optimal Design}\label{section:D and grad}
@@
 
-\begin{proof}  
-  From the definition of $\Sigma(\obs)$ \eqref{eq:Sigma}:
-
-  \begin{align}\label{eq:der sig}
-    \begin{split}
-      \frac{\der}{\der \tau} \Big |_{\tau=0} \Sigma( \obs + \tau V )
-      &= \frac{\der}{\der \tau} \Big |_{\tau=0} 
-      (\obs + \tau V ) \modcov (\obs + \tau V )^*  + \sigma^2I\\
-      % 
-      % 
-      % 
-      &= V \modcov \obs^* + \obs \modcov V^*.
-    \end{split}
-  \end{align}
-
-  Then, using \eqref{eq:der sig}: 
-  \begin{align*}
-    0 &= \frac{\der}{\der \tau} \Big |_{\tau=0} I \\
-    % 
-    % 
-    % 
-    &= \frac{\der}{\der \tau} \Big |_{\tau=0}
-    \left (\Sigma(\obs+\tau V)^{-1} \Sigma(\obs+\tau V) \right ) \\
-    % 
-    % 
-    % 
-    &= \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0} \Sigma+
-    \Sigma^{-1} \frac{\der \Sigma(\obs+\tau V)}{\der \tau} \Big |_{\tau=0}\\  
-    %
-    %
-    %
-    &= \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0} \Sigma+
-    \Sigma^{-1} (V\modcov \obs^* + \obs \modcov V^*). 
-    %\text{, by \eqref{eq:der sig}. }
-  \end{align*}
-
-  Thus:
-  \begin{align}\label{eq:der sig inv}
-    \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0}  
-      &= -\Sigma^{-1} (V \modcov \obs^* + \obs \modcov V^*) \Sigma^{-1}.
-    \end{align}
-  %
-
-  Let
-  \begin{equation*}
-    T(\obs) = \obs^* \Sigma^{-1}(\obs)\obs.
-  \end{equation*}
-
-  By Leibniz (product) rule and \eqref{eq:der sig inv}:
-  % 
-  \begin{align}\label{eq:T}
-    \begin{split}
-    \delta T(\obs) V 
-    &= \frac{\der T(\obs + \tau V)}{\der \tau} \Big |_{\tau=0} \\
-    %
-    %
-    %
-    &= V^* \Sigma^{-1} \obs 
-    - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
-    &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs
-    + \obs^* \Sigma^{-1} V.
-    \end{split}
-  \end{align}
-
-  We now record a lemma which generalizes a known result in linear
-  algebra \cite[Chapter 9, Theorem 4]{Lax07}. Its proof is delegated
-  to the supplementary material.
-  \begin{lemma}\label{lemma:lax}
-    Let $Y(t)$ be a differentiable operator-valued function. Assume 
-    $I+Y(t)$ is invertible, $Y(t)$ self-adjoint and trace-class. Then
-    \begin{equation*}
-      \frac{\der \log \det (I+Y(t))}{\der t} = \tr{(I+Y(t))^{-1} \dot{Y}(t)}.
-    \end{equation*}
-  \end{lemma}
-
-  We employ Lemma \ref{lemma:lax} and calculate the first variation of
-  $\tar$:
-
-  \begin{align*}
-    %\begin{split}
-      \delta \tar(\obs) V 
-      :&= \frac{\der}{\der\tau} \Big |_{\tau=0} \tar(\obs + \tau V) \text{ (Definition \ref{def:var})}\\
-      % 
-      % 
-      % 
-      &= \frac12 \frac{\der}{\der \tau} \Big |_{\tau=0} \log \det 
-      (I + \prcov^{1/2} \fwd^* T(\obs+\tau V)\fwd \prcov^{1/2} ) \text{ (Theorem \ref{thm:d optimality})} \\
-      % 
-      % 
-      % 
-      &= \frac12 \tr{( I + \prcov^{1/2} \fwd^* \obs^* \Sigma^{-1}
-        \obs\fwd \prcov^{1/2} )^{-1}
-        \frac{\der}{\der \tau} \Big |_{\tau=0}
-        \prcov^{1/2} \fwd^* T(\obs+\tau V) \fwd \prcov^{1/2}}\ \text{ (Lemma \ref{lemma:lax})} \\
-      % 
-      % 
-      % 
-      &= \frac12 \ttr\Big \{ \postcov \fwd^* (V^* \Sigma^{-1} \obs 
-      - \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs \\
-      &\ \ \ - \obs^* \Sigma^{-1} \obs \modcov V^* \Sigma^{-1}\obs 
-      + \obs^* \Sigma^{-1} V ) \fwd \Big \}  \text{ (by \eqref{eq:T})} \\
-      %
-      %
-      %
-      &= \tr{\postcov \fwd^* ( \obs^* \Sigma^{-1} V -
-      \obs^*\Sigma^{-1} V\modcov \obs^* \Sigma^{-1}\obs ) \fwd} \\
-      %
-      %
-      % 
-      &= \tr{\postcov \fwd^* \obs^* \Sigma^{-1} V 
-      ( I - \modcov \obs^* \Sigma^{-1}\obs ) \fwd} \\
-      % 
-      %
-      %
-      &= \tr{V ( I - \modcov \obs^* \Sigma^{-1}\obs )
-      \fwd \postcov \fwd^* \obs^* \Sigma^{-1}}.
-    %\end{split}
-  \end{align*} 
-  Recalling Definition \ref{def:var} concludes the proof.
-\end{proof}
-
+The proof amounts to calculating the variational derivative of $\tar$
+at $\obs$ for any direction $V$ (by Definition \ref{def:var}) and is
+delegated to the Supplementary.
+
+\subsection{Unit norm constraints and their gradient}
 In a real-life optimal design problem we cannot choose any measurement
@@
 bound the norm of measurements. If, for example, we can take
-measurements in $span\{\meas\}$ for some $\meas \neq 0$, then a
-maximum for the D-optimality criterion does not exist, as it is not
-bounded. In contrast, in any real-life problem where sensors are
-concerned, the norm of measurements recorded by sensors is always one
-(of course, point evaluations are not in any Hilbert space of
-functions we wish to consider):
+measurements in $\textup{span}\{\meas\}$ for some $\meas \neq 0$, then
+the the D-optimality criterion is unbounded. In contrast, in any
+real-life problem where sensors are concerned, the norm of
+measurements recorded by sensors is always one\footnote{Again, our
+analysis does not apply to point evaluations. These are just used for
+motivation and can be approximated by elements in $\hilo^*$ as long as
+$\hilo$ is a function space, e.g.~$L^2(\Omega)$.}:
 
@@
 them and find their gradients in Proposition \ref{prop:constraints
-  grad} below:
+  grad} below, whose proof is straightforward and delegated to the
+Supplementary:
 
@@
 
-\begin{proof}
-  \begin{align*}
-    \delta \phi_j(\obs)V  
-    &= \frac12\lim_{\tau \to 0}\tau^{-1}
-    ( \|(\obs + \tau V)^*e_j \|_{\hilp}^2 - \|\obs ^*e_j \|_{\hilp}^2  ) \\
-    %
-    %
-    %
-    &= \frac12\lim_{\tau \to 0}\tau^{-1}
-    ( \langle (\obs + \tau V)^*e_j, (\obs + \tau V)^*e_j \rangle_{\hilp} - 
-    \langle \obs^*e_j, \obs^*e_j \rangle_{\hilp} ) \\
-    % 
-    % 
-    %
-    &= \frac12\lim_{\tau \to 0}\tau^{-1}
-    (2\tau \langle \obs^*e_j,V^*e_j \rangle_{\hilp} 
-    +\tau^2 \langle V^*e_j, V^*e_j \rangle_{\hilp} ) \\
-    %
-    %
-    % 
-    &= \langle \obs^*e_j,V^*e_j \rangle_{\hilp} \\
-    %
-    %
-    % 
-    &= \langle V \obs^*e_j,e_j \rangle_{\R^m} \\
-    %
-    %
-    %
-    &= e_j^t V \obs^* e_j \\
-    % 
-    %
-    %
-    &= \tr{V \obs^* e_je_j^t}.
-  \end{align*}
-\end{proof}
-
 %% The same arguments justifying \eqref{eq:tar grad} hold here, and thus:
@@
 
+\subsection{Necessary conditions for D-optimality}
 We find necessary first-order conditions for D-optimality via Lagrange
@@
   \begin{equation*}
-    \obs = \argmax_{\|\meas_j\| = 1, j=1,\dots,m}\tar(\obs).
+    \opt = \argmax_{\|\meas_j\| = 1, j=1,\dots,m}\tar(\obs).
   \end{equation*}
@@
   \begin{equation*}
-    ( I - \modcov \obs^* \Sigma^{-1} \obs) \fwd \postcov \fwd^* \obs^*  \Sigma^{-1}
-    = \obs^* \Xi, 
+    ( I - \modcov \opt^* \Sigma^{-1} \opt) \fwd \postcov \fwd^* \opt^*  \Sigma^{-1}
+    = \opt^* \Xi, 
   \end{equation*}
@@
 
-%\input{nonvanishing}
 %% Imposing correlations between observations alleviate measurement clusterization
@@
 \end{proposition}
-
-\begin{proof}
-  Let
-  \begin{align*}
-    \Sigma( \obs ) &= 
-    \begin{bmatrix}
-      \Sigma (\obsm )           & \obsm \modcov \meas_m \\
-      \meas_m \modcov \obsm^*   & \sigma^2 + \meas_m \modcov \meas_m
-    \end{bmatrix}
-    : =
-    \begin{bmatrix}
-      \Sigmam   & w \\
-      w^t       & c
-    \end{bmatrix}\\
-    %
-    %
-  \end{align*}
-
-  The Schur complement implies:
-  \begin{align}\label{eq:schur}
-    \begin{split}
-          \Sigma^{-1} &=
-          \begin{bmatrix}
-            \Sigmam^{-1} + \Sigmam^{-1} w ( c - w^t \Sigmam^{-1} w)^{-1} w^t \Sigmam^{-1} & - \Sigmam^{-1} w ( c - w^t \Sigmam^{-1} w)^{-1} \\
-            -( c - w^t \Sigmam^{-1} w)^{-1} w^t \Sigmam^{-1}                            &  ( c - w^t \Sigmam^{-1} w)^{-1}
-          \end{bmatrix} \\
-          &=
-          \begin{bmatrix}
-            \Sigmam^{-1} & 0 \\
-            0           & 0 
-          \end{bmatrix}
-          + (c -w^t \Sigmam^{-1} w )^{-1}
-          \begin{bmatrix}
-            \Sigmam^{-1} w \\
-            -1
-          \end{bmatrix}
-          \begin{bmatrix}
-            w^t \Sigmam^{-1} & -1 
-          \end{bmatrix},
-    \end{split}
-  \end{align}
-  %
-  and denote:
-  %
-  \begin{align}\label{eq:M def}
-    \M (\obs ):&= \prcov^{\frac12}\fwd^* \obs^* \Sigma^{-1} \obs \fwd
-    \prcov^{\frac12}.
-  \end{align}
-
-  From \eqref{eq:schur} and \eqref{eq:M def}:
-  \begin{align*}
-    \M(\obs) &= \prcov^{1/2} \fwd^* \obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2} \\
-    %
-    %
-    %
-    &= \prcov^{1/2} \fwd^* \obs^* \left \{
-    \begin{bmatrix}
-      \Sigmam^{-1} & 0 \\
-      0           & 0 
-    \end{bmatrix}
-    + (c -w^t \Sigmam^{-1} w )^{-1}
-    \begin{bmatrix}
-      \Sigmam^{-1} w \\
-      -1
-    \end{bmatrix}
-    \begin{bmatrix}
-      w^t \Sigmam^{-1} & -1 
-    \end{bmatrix} 
-    \right \} \obs \fwd \prcov^{1/2} \\
-    %
-    %
-    %
-    &= \M (\obsm) + (c -w^t \Sigmam^{-1} w )^{-1}
-    \prcov^{1/2} \fwd^* \obs^*
-    \begin{bmatrix}
-      \Sigmam^{-1} w \\
-      -1
-    \end{bmatrix}
-    \begin{bmatrix}
-      w^t \Sigmam^{-1} & -1 
-    \end{bmatrix} 
-    \obs \fwd \prcov^{1/2}
-  \end{align*}
-  %
-  Now, denote:
-  %
-  \begin{align}\label{eq:u}
-    \begin{split}
-      u :&= (c -w^t \Sigmam^{-1} w )^{-1/2}
-      \prcov^{1/2} \fwd^* \obs^* 
-      \begin{bmatrix}
-        \Sigmam^{-1} w \\
-        -1 
-      \end{bmatrix} \\
-      %
-      %
-      %
-      & = (c -w^t \Sigmam^{-1} w )^{-1/2} ( \prcov^{1/2}\fwd^* \obsm^* \Sigmam^{-1} \obsm  \modcov \meas_m - \prcov^{1/2} \fwd^* \meas_m )\\
-      %
-      %
-      %
-      u^* :&=  (c -w^t \Sigmam^{-1} w )^{-1/2} (\meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \fwd \prcov^{1/2} - \meas_m \fwd \prcov^{1/2} ),
-    \end{split}
-  \end{align}
-  %
-  so that
-  %
-  \begin{equation}\label{eq:M plus I}
-    I + \M( \obs ) = I + \M (\obsm ) + uu^*.
-  \end{equation}
-  %
-  Note that
-  \begin{equation}\label{eq:M postcov}
-    \prcov^{1/2} \left (I + \M( \obsm ) \right )^{-1} \prcov^{1/2} = \postcovm.
-  \end{equation}
-  From a generalization of the matrix determinant lemma to Hilbert
-  spaces ($\det(A + uv^*) = (1 + \langle A^{-1} u,u \rangle) \det
-  A$, statement and proof in the supplementary material):
-  %
-  \begin{align}\label{eq:diffs}
-    \begin{split}
-      \tar( \obs ) - \tar( \obsm )
-      %
-      %
-      %
-      &= \frac12 \log \left (\det \big ( I + \M ( \obs ) \big ) / \det \big ( I + \M (\obsm) \big ) \right )\\
-      %
-      %
-      %
-      &= \frac12  \log \left (\det \left ( I + \M(\obsm) + uu^* \right ) / \det \big ( I + \M (\obsm) \big )\right ) \\
-      %
-      %
-      %
-      &= \frac12 \log \left ( 1 + \left \langle \left ( I+\M(\obsm) \right )^{-1} u, u  \right \rangle \right ).
-    \end{split}
-  \end{align}
-  From \eqref{eq:u} and \eqref{eq:M postcov}:
-  \begin{align}\label{eq:final}
-    \begin{split}
-      &\left \langle \left (I+\M (\obsm)\right )^{-1}u, u \right \rangle\\
-      &= \frac{
-        \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m,
-        (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m \rangle
-      }{
-        c- w^t \Sigmam^{-1} w
-      }\\
-      %
-      %
-      %
-      &= 
-      \frac{
-      \langle \fwd \postcovm \fwd^* (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m,
-      (\obsm^* \Sigmam^{-1} \obsm \modcov - I ) \meas_m \rangle
-      }{
-        \sigma^2 + \meas_m \modcov \meas_m - \meas_m \modcov \obsm^* \Sigmam^{-1} \obsm \modcov \meas_m 
-      }
-    \end{split}
-  \end{align}
-  and the conclusion follows by substituting \eqref{eq:final} into
-  \eqref{eq:diffs}.
-\end{proof}
+The proof is long and tedious, and is delegated to the Supplementary.
 
@@
 We have shown that in the limit $\sigma^2 \to 0$, no increase in
-$\tar$ is achieved by repeating a measurement, so designs that exhibit
-measurement clusterization are not D-optimal. Since $\tar$ is not
-defined for $\sigma^2 = 0$ and identical measurements, we cannot make
-a statement regarding $\sigma^2 = 0$, except in the limiting sense
-described above. In conclusion, for small observation error $\sigma^2$
-levels, measurement clusterization is mitigated by the presence of a
-non-zero model error $\modcov$ --- answering Question \ref{q:mitigate}
-posed in the Introduction.% Section \ref{section:intro}.
-
-%\input{vanishing}
+$\tar$ is achieved by repeating a measurement. Thus, for vanishing
+noise levels, clustered designs cannot be D-optimal. For repeated
+measurements and $\sigma^2=0$, $\Sigma$ is not invertible and the
+posterior covariance is not defined in eq.~\eqref{eq:postcov}. We can
+\emph{define} the posterior in this case to equal the posterior when
+the repeated measurement is dropped, and under this definition, a
+repeated measurement trivially does not increase the design criterion
+when $\sigma^2=0$. Our results are stronger, since we show
+\emph{continuity} in $\sigma^2$.
+
+It is worth noting that by the nonnegativity of the KL divergence,
+$\tar$ cannot decrease upon adding measuremnts. However, we can
+construct examples where the posterior is not changed upon taking a
+new measurement e.g.~if the prior variance vanishes on some
+eigenvector and a measuremnt is taken on said eigenvector. We do not
+expect a measurement to generate no information whatsoever in any
+realistic scenario, and we choose to ignore such pathologies.
+
+In conclusion, for small observation error $\sigma^2$ levels,
+measurement clusterization is mitigated by the presence of a non-zero
+model error $\modcov$ --- answering Question \ref{q:mitigate} posed in
+the Introduction.% Section \ref{section:intro}.
 \section{D-Optimal Designs Without Model Error}\label{section:vanishing}
@@
 
-\begin{proof}
-  The proof amounts to using Woodbury's matrix identity twice with a
-  regularization trick. The standard proof for Woodbury's matrix
-  identity works for separable Hilbert spaces, as long as all terms
-  are well defined. Unfortunately, $\obs^*\obs$ is not invertible, so
-  we force it to be. Recall Woodbury's matrix identity:
-  \begin{equation}\label{eq:WMI}
-    (A + UCV)^{-1} = A^{-1} - A^{-1}U(C^{-1} + VA^{-1}U)^{-1}VA^{-1}. 
-  \end{equation}
-
-  Denote:
-  \begin{align}\label{eq:notation}
-    \begin{split}
-      A :&= \prcov^{-1} \\
-      U :&= \fwd^* \\
-      V :&= \fwd \\
-      C :&= \sigma^{-2} (\obs^*\obs+\eps I),\ \eps > 0
-    \end{split}
-  \end{align}
-
-  Then, \eqref{eq:WMI} with the notation \eqref{eq:notation} implies:
-  \begin{align}\label{eq:first}
-    \begin{split}
-    &\fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* \\
-    &\ \ \ \ \ = \fwd ( \prcov - \prcov \fwd^* ( \sigma^2(\obs^*\obs + \eps I)^{-1} + \fwd \prcov \fwd^* )^{-1} \fwd \prcov ) \fwd^*. \\
-      %
-      %
-      %% &= X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X
-  %%   \end{split}
-    %% \end{align*}
-    \end{split}
-  \end{align}
-  Now, denote:
-  \begin{align*}
-    X :&= \fwd\prcov \fwd^*, \\
-    Y_{\eps} :&= \obs^*\obs + \eps I,
-  \end{align*}
-  and observe that $Y_{\eps}$ is invertible. Substitute $X, Y_{\eps}$
-  into the RHS of \eqref{eq:first}:
-   \begin{equation}\label{eq:second}
-     \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* = X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X
-   \end{equation}
-
-   Note that $X + \sigma^2 Y_{\eps}^{-1}$ is invertible, as the sum of
-   two positive definite operators. Now, let
-   \begin{align}\label{eq:notation2}
-     \begin{split}
-       A :&= X^{-1}, \\
-       C :&= \sigma^2Y_{\eps}^{-1}, \\
-       U :&= I, \\
-       V :&= I.
-     \end{split}
-   \end{align}
-   Apply \eqref{eq:WMI} with the notation \eqref{eq:notation2} to the
-   RHS of \eqref{eq:second}:
-  \begin{align*}
-    \begin{split}
-      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* &= X - X(\sigma^2Y_{\eps}^{-1} + X)^{-1}X \\
-      %
-      %
-      %
-      &= (X^{-1} + \sigma^{-2}Y_{\eps})^{-1} \\
-      %
-      %
-      %
-      &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \eps I))^{-1}.
-    \end{split}
-  \end{align*}
-
-  We conclude that $\forall \eps > 0$
-  \begin{align*}
-    \begin{split}
-      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* (\obs^* \obs +\eps I) \fwd )^{-1}\fwd^* 
-     &= (\fwd \prcov \fwd^* + \sigma^{-2} (\obs^*\obs + \eps I))^{-1}.
-    \end{split}
-  \end{align*}
-  Letting $\eps \to 0$ completes the proof.
-\end{proof}
+As we mentioned in Section \ref{subsec:bayesian inverse problems},
+$\fwd \prcov \fwd^*$ is the prior covariance $\hilo$ and could be
+safely assumed invertible. The proof of Proposition \ref{prop:twice
+  woodbury} is delegated to the Supplementary. It amounts to using
+Woodbury's matrix identity twice. The standard proof for Woodbury's
+matrix identity works for separable Hilbert spaces, as long as all
+terms are well defined. Unfortunately, $\obs^*\obs$ is not invertible,
+so we utilize a regularization trick to force it to be.
 
@@
 
-\begin{proof}
-  Let $\Phi(\eta) = \sum_{i=1}^k \eta_i - m$ and $\Omega_j(\eta) =
-   -\eta_j$. Then
-  \begin{align*}
-    \begin{split}
-      \frac{\partial \tar}{\partial \eta_i}  &=
-       \frac12 \frac{\sigma^{-2}\lambda_i}{1 + \sigma^{-2} \lambda_i\eta_i} = \frac12 \frac{1}{\sigma^{2}\lambda_i^{-1} + \eta_i} \\
-      %
-      %
-      %
-      \frac{\partial\Phi}{\partial \eta_i} &= 1 \\
-      %
-      %
-      %
-      \frac{\partial \Omega_j}{\partial \eta_i} &= -\delta_{ij}      
-    \end{split}
-  \end{align*}
-
-  From the KKT conditions, there are $\alpha, \beta_i$ such that for $i=1,\dots,m$:
-  \begin{align}
-    \begin{split}
-      -\frac12 \frac{1}{\sigma^{2}\lambda_i^{-1} + \eta_i} + \alpha - \beta_i  &= 0 \\
-      %
-      %
-      %
-      \eta_i &\geq 0\\
-      %
-      %
-      %
-      \beta_i &\geq 0\\
-      %
-      %
-      %
-      \beta_i \eta_i &= 0\\
-      %
-      %
-      %
-      \sum_{i=1}^m \eta_i &= m 
-    \end{split}
-  \end{align}
-
-  Then, for $i \in A$:
-  \begin{align*}
-    \begin{split}
-      \beta_i &= 0\\
-      %
-      %
-      %
-      \sigma^{2}\lambda_i^{-1} + \eta_i  &= \frac{1}{2\alpha}%  \text{ for }  1 \leq i \leq k.\\
-      %
-      %
-      %
-      %% \sigma^{-2} \lambda_i &= 2\alpha - 2\beta_i, i \in B \\
-    \end{split}
-  \end{align*}
-
-  Summing over $i \in A$, substituting $\sum_{i\in A} \eta_i =
-  \sum_{i=1}^m \eta_i = m$, and dividing by $k:= |A|$:
-  $$
-  \frac{1}{2\alpha} = \frac{m}{k} + \frac{\sigma^2}{k} \sum_{j\in A} \lambda_j^{-1}.
-  %\Rightarrow \alpha = \left ( \frac{2m}{k} + \frac{2\sigma^2}{k}
-  %\sum_{i\in A} \lambda_i^{-1} \right )^{-1}.
-  $$
-
-  Consequently:
-  \begin{align} \label{eq:etas}
-    \begin{split}
-      \eta_i &= \frac{1}{2\alpha} - \sigma^2 \lambda_i^{-1}\\
-      %
-      %
-      %
-      &=\frac{m}{k} + \frac{\sigma^2}{k} \sum_{j\in A} \lambda_j^{-1} - \sigma^2 \lambda_i^{-1}.%\\
-      %
-      %
-      %
-      %% &= \frac{m}{k} + \frac{\sigma^2}{k} \sum_{j \in A, j\neq i} \lambda_j^{-1} - \frac{k-1}{k}\sigma^2 \lambda_i^{-1}\\
-      %% ,\ i \in A.
-    \end{split}
-  \end{align}
-\end{proof}
-
-The final ingredient we require for the proof of Theorem
-\ref{thm:char} is:
+The proof of Proposition \ref{prop:kkt} amounts to utilizing the
+Karush-Kuhn-Tucker conditions and is delegated to the
+Supplementary. The final ingredient we require for the proof of
+Theorem \ref{thm:char} characterizing D-optimal designs is:
 
@@
 
-\begin{proof}
-  Let us diagonalize $M$, so that $M = U D U^t$ with $D =
-  \diag(d_1,\dots,d_k)$ and $U \in \R^{k \times k }$ orthogonal. Let
-  $S \in \R^{k \times m}$ with $S_{ii} = \sqrt{d_{i}}$ and zeros
-  otherwise. Define $A:= U S V^t$, where $V \in \R^{m \times m}$ is
-  orthogonal and will be further restricted later. Then $AA^t = U
-  SV^tVS^t U^t = UDU^t$, so $AA^t$ has the required eigenvalues and
-  eigenvectors by construction. If we can choose $V$ such that $A$
-  also satisfies the unit norm constraints we are done. These
-  constraints are, for $j=1,\dots,m$:
-  \begin{equation}\label{eq:V constraints}
-   1 = [A^tA]_{jj} = [V S^tS V^t]_{jj},
+The proof of Lemma is also delegated to the Supplementary. It is
+however important to note that this proof is construcive. It will give
+us a way to construct D-optimal designs, once we fully characterize
+them in Theorem \ref{thm:char} below.
+
+\begin{theorem}\label{thm:char}
+  Let:
+  \begin{itemize}
+    \item The D-optimality design criterion
+    \cite{AlexanderianGloorGhattas14}:
+    \begin{align*}
+      \begin{split}
+        \tar(\obs) %:&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
+        % 
+        % 
+        % 
+        &= \frac12 \log \det ( I + \sigma^{-2} \prcov^{1/2} \fwd ^*
+        \obs^* \obs \fwd \prcov^{1/2}), 
+      \end{split}
+    \end{align*}
+  \item \(\opt\) a D-optimal design operator
+    \begin{equation*}
+      \opt = \argmax_{\|\meas_j\| = 1, j=1,\dots,m}\tar(\obs),
+    \end{equation*}
+  \item \(\{\lambda_i\}_{i=1}^\infty\) eigenvalues of
+    \(\fwd\prcov\fwd^*\) in decreasing order of magnitude.
+    %% \item \(\{\ev_i\}_{i=1}^\infty\) their corresponding eigenvectors.
+  \item \(\{\eta_i\}_{i=1}^\infty\) eigenvalues of \(\opt^*\opt\).
+
+  \end{itemize}
+
+  Then:
+  \begin{enumerate}
+  \item  \(\tr{\opt^*\opt} = m\).
+  \item \(\opt^*\opt\) and \(\fwd\prcov\fwd^*\) are simultaneously
+    diagonalizable.
+  \item \(k := \rank \opt^*\opt \leq m\) and
+    \begin{equation*}
+      \tar(\opt) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i). %= \frac12 \sum_{i=1}^{m} \log (1 + \sigma^{-2}\lambda_i\eta_i).
+    \end{equation*}
+  %% \item 
+  %%   \begin{equation*}
+  %%     k = \argmax \left \{ k:\lambda_k^{-1} < \sigma^{-2}\frac{m}{k} + \frac{1}{k} \sum_{j=1}^{k}
+  %%     \lambda_j^{-1} \right \}.
+  %%   \end{equation*}
+  \item
+    \begin{equation*}
+        \eta_i = \begin{cases}
+          \frac{m}{k} - \sigma^2 \lambda_i^{-1} + \sigma^2 \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1} & 1 \leq i \leq k \\
+          0 & i > k 
+        \end{cases}.
+    \end{equation*}
+  \item The covariance of the pushforwad \(\fwd_{*} \postopt\) is \(\left
+    ( (\fwd \prcov \fwd^*)^{-1} + \sigma^{-2} \opt^*\opt \right
+    )^{-1}\) and its eigenvalues are
+    \begin{equation}\label{eq:cylinders}
+      \theta_i =
+      \begin{cases}
+        \left(\frac{\sum_{j=1}^k \lambda_j^{-1} + \sigma^{-2}m}{k} \right )^{-1} & i \leq k \\
+        \lambda_i &  i > k 
+      \end{cases}
     \end{equation}
-  and we can expect to do this since we assumed $\ttr D = m$.
-
-  Define $C = S^tS - I \in \R^{m \times m}$. Note that $\ttr C = 0$ and
-  $C$ is diagonal with non-zero entries $d_i-1,i=1,\dots,k$. It suffices
-  to find $V$ orthogonal such that $V C V^t$ has zero diagonal. We
-  construct such $V$ by sequentially inserting zeros in the diagonal
-  and not destroying zeros we already introduced, starting from the
-  last diagonal entry and moving to the first. Since $c_{mm} \neq 0$ ,
-  let $p < m$ such that $c_{pp}c_{mm} < 0$ (such $p$ exists because
-  the trace is zero) and let $\theta \in (0,\pi)$. Define a Givens
-  rotation $R^{(m)} \in \R^{m \times m}$ by
-  \begin{equation*}
-    r^{(m)}_{ab} :=
-    \begin{cases}
-      1 & a = b \neq p \text{ or } a = b \neq m \\
-      \cos \theta & a = b = p  \\
-     -\sin \theta & a = p, b = m\\
-      \cos \theta & a = b = m \\
-      \sin \theta & a = m, b = p \\ 
-      0 & o.w
-    \end{cases}
-  \end{equation*}
-  Note that conjugating a matrix by $R^{(m)}$ changes only its $m$ and
-  $p$ rows and columns. We want to choose $\theta$ such that
-  \begin{equation}\label{eq:mm}
-    0 = [R^{(m)} C (R^{(m)})^t]_{mm} = \cos^2 \theta c_{mm} + 2\cos \theta \sin
-    \theta c_{mp} + \sin^2\theta c_{pp},
-  \end{equation}
-  and it suffices to choose $\theta$ such that
-  \begin{equation*}
-    c_{mm} \cot^2 \theta + 2 c_{mp} \cot \theta + c_{pp} = 0.
-  \end{equation*}
-  This quadratic in $\cot\theta$ has a real solution, since
-  $c_{pp}c_{mm} < 0$ by assumption and we can find $\theta \in
-  (0,\pi)$ such that \eqref{eq:mm} is satisfied. We continue to find
-  $R^{(m-1)}$ that leaves row and column $m$ unchanged and
-  continue introducing zeros to the diagonal. The assumption $\ttr D =
-  m \Rightarrow \ttr C = 0$ guarantees we can do that. Taking $V:=
-  R^{(1)} R^{(2)} \dots R^{(m-1)}R^{(m)}$ completes the proof.
-\end{proof}
-
-\char*
+  \end{enumerate}
+\end{theorem}
 
@@
 %%     \caption{Three graduated lab cylinders, corresponding to three
-%%       eigenvectors. Prior eigenvalues not shown.}
+%%      eigenvectors. Prior precision eigenvalues not shown.}
 %%     \label{fig:cylinder}
@@
 
-%\input{why}
-\section{Answer to Question \ref{q:why}: The cause of clusterization}\label{section:why}
-According to Theorem \ref{thm:char}, D-optimal designs aim to capture
-a small subset of eigenvectors of the prior covariance, specifically
-the $k$ eigenvectors with the highest prior variance. Our model
-naturally achieves this objective by not measuring eigenvectors $k+1$
-and above, as proven in Theorem \ref{thm:char}. Translating this
-understanding to spatial problems, we anticipate that a D-optimal
-design would favor measurement locations where eigenvectors $k+1$ and
-above are either close to zero in value or possess small eigenvalues
-in the prior spectrum, for some $k > 0$. To illustrate this
-preference, Fig.~\ref{fig:why} depicts the scenario using the 1D
-heat equation with homogeneous Dirichlet boundary conditions (details
-in the supplementary material). The plot showcases four eigenvectors,
-scaled according to their prior standard deviations. Since
-eigenvectors beyond the fourth have insignificant prior eigenvalues,
-we exclude them from consideration. Notably, we observe that
-measurements are clustered near the zeros of the third and fourth
-eigenvectors, so we conclude that $k=2$. Clusterization arises because
-there are only two locations where the third and fourth eigenvectors
-approach zero, while the first and second eigenvectors exhibit
-significantly non-zero values. Consequently, when we allocate four
-measurements to these two locations, they naturally cluster together,
-aligning with the pigeonhole principle.
-
-\bibliographystyle{ba}
-\bibliography{/home/yair/projects/bibtex.bib}
+\subsection{Answer to Question \ref{q:why}}\label{subsec:why}
+Theorem \ref{thm:char} helps us give a compelling explanation of the
+measurement clusterization we obeserved for the inverse problem of the
+heat equation (Fig.~\ref{fig:}). Below, we take the following steps:
+We first utilize Theorem \ref{thm:char} to understand what a D-optimal
+design will look like for our relaxed model; we then argue that
+although not identical, a D-optimal design for the full inverse
+problem of the heat equation shares similar charachteristics to the
+relaxed design; finally, we show how ...
+
+First, recall the relevant parts of Theorem \ref{thm:char}. In our
+relaxed model, for a D-optimal design $\opt$:
+\begin{itemize}
+  \item $\opt^*\opt$ and $\fwd\prcov \fwd^*$ are simultaneously
+    diagonalizable.
+  \item $\opt^*\opt$ has non-zero eigenvalues only for the $k:=\rank
+    \opt^*\opt$ largest eigenvalues $\lambda_j$ of $\fwd\prcov\fwd^*$,
+    i.e.~$\eta_j=0$ for $j>k$.
+  \item $k$ is monotonic in $m$, the number of measurements we take.
+\end{itemize}
+
+Consider $\fwd$ and $\prcov$ from the \emph{the inverse problem of the
+heat equation}. As before, we denote the eigenvalues of
+$\fwd\prcov\fwd^*$ by $\lambda_j$. We input these eigenvalues into our
+\emph{relaxed} model, and find a D-optimal design $\opt$ for our
+relaxed model, e.g.~by using Lemma \ref{lemma:free} in the proof of
+Theorem \ref{thm:char}. In our relaxed model, the measurements we take
+are best utilized in reducing uncertainty for the first $k$
+eigenvectors. So, a D-optimal design arising from our \emph{relaxed
+model} completely avoids measuring eigenvectors $k+1$ and above.
+
+Of course, in a real life problem --- such as the inverse problem of
+the heat equation --- it is likely impossible to find measurement
+locations for which all eigenvectors $k+1$ and above are
+zero. Luckily, the eigenvalues of $\fwd\prcov\fwd^*$ decay very
+quickly: for example, recall from Section \ref{subsec:toy} that the
+eigenvalues of the forward model for the 1D heat equation are $-(2\pi
+n)^{-2}$. Thus a D-optimal design is a trade-off between measuring the
+small number of eigenvectors we are interested in.
+
+We explore the abovementione trade-off for the inverse problem of the
+heat equation in Fig.~\ref{fig:eigenvectors}. We allow $m=4$
+measurements in $\Omega = [0,1]$ and observe that D-optimal
+measurement locations are clustered at $x_1 = 0.31$ and $x_2 =
+0.69$. Upon close inspection of the scaled eigenvectors of $\fwd
+\prcov \fwd^*$, we first observe that eigenvectors $3$ and above have
+negligible prior amplitude. Since we only have $m=4$ measurements at
+our disposal, we interpret these results, following Theorem
+\ref{thm:char}, as implying we should only care about measuring the
+first and second eigenvectors. Then, we note the D-optimal $x_1,x_2$
+present a compromise between the amplitude of the first and second
+eigenvectors. For example, a measurement at $x=0.5$ would have ignored
+the second eigenvector altogether, since the second eigenvector is
+zero at $x=0.5$.  Now we can understand measurement clusterization:
+attempting to measure what we want and ignoring the rest, along with
+the spatial limitations on measurements imposed by the structure of
+eigenvectors of $\fwd \prcov \fwd^*$ is the cause for measurement
+clusterization in D-optimal designs: we have $m=4$ measuremnts at our
+disposal but only two spatial locations that are a good compromise
+between the scaled eigenvectors. Thus, clusterization arises by the
+pigeonhole principle.
+
+\begin{figure}\label{fig:eigenvectors}
+    \centering
+    \includegraphics[width=\textwidth]{eigenvectors_dst_scaled.png}
+    \caption{D-optimal measurement locations ($m=4$ measurements) and
+      weighted eigenvectors for finding the initial condition of the
+      1D heat equation. Measurement locations and weighted
+      eigenvectors are plotted over the computational domain $\Omega =
+      [0, 1]$ (x-axis). Measurement clusterization occurs
+      approximately at $0.31$ and $0.69$. These two locations are a
+      compromise between the magnitudes of the first and second
+      eigenvectors. These are the eigenvectors a D-optimal design aims
+      to measure. The rest of the eAllocating $m=4$ measurements into two locations
+      results in clusterization, according to the pigeonhole
+      principle.}
+  \label{fig:why}
+\end{figure}
+\section{Numerical Experiments}
+%We provide numerical evidence supporting our conclusions.
+
+\subsection{Simulating Lemma \ref{lemma:free}}\label{subsec:lemma sims}
+In the proof of Theorem \ref{thm:char} we utilize Lemma
+\ref{lemma:free} to construct D-optimal designs. We implement the
+construction in Lemma \ref{lemma:free} with the goal of testing
+numerically how prevalent are clustered designs. We iterate over the
+number of measurements $m \in \{4,\dots, 15\}$, and for every $m$ we
+then iterate over $k:=\rank \obs^*\obs \in \{2,\dots, m-1\}$. For each
+pair $m,k$ we run $N=5000$ simulations according to the following
+steps:
+\begin{enumerate}
+\item Generate random diagonal $D\in \mathbb{R}^{k\times k}$ with
+  entries $\log (d_i) \sim \mathcal{N}(5,1)$ and normalize it so that
+  $\ttr D = m$.
+\item Conjugate $D$ by a random orthogonal matrix to form a positive
+  semi-definite $M := UDU^t \in \mathbb{R}^{k\times k}$.
+\item Apply the construction of Lemma \ref{lemma:free} to calculate
+  $A$ such that $AA^t = M$, where $A$ has unit norm columns.
+\item Since columns of $A$ correspond to measurement vectors, we call
+  $A$ "clustered" if $A$ has two or more identical columns (up to some
+  numerical precision threshold).
+\end{enumerate}
+We then calculate the fraction of clustered designs of the simulations
+we ran, for each pair $m,k$. Clusteriztion occured at high rates
+($>95\%$) whenever $m-k > 1$; see Fig.~\ref{fig:sim AAt}. Hence, in
+these simulations, clusterization is a generic property. However, when
+$m-k \leq 2$, clusterization does not occur. We do not why and further
+investigation into this phenomenon is out of scope for the current
+study.
+
+\begin{figure}
+    \centering
+    \includegraphics[height=0.5\textwidth]{simulations.png}
+    \caption{Fraction of clustered $A$ for $AA^t = M$ and $M$
+      generated randomly (see text and repository for details on
+      generating $M$). It is evident that when $m-k >3$ clusterization
+      is ubiquitous, whereas for lower $m-k$ clusterization does not
+      occur.}
+  \label{fig:sim AAt}
+\end{figure}
+
+Full results are located in the \texttt{simulations.csv} file within
+the accompanying \href{https://github.com/yairdaon/OED}{repository},
+code implementing the experiments described above is located in module
+\texttt{zeros.py} of said repository.
+
+\subsection{Correlated errors}\label{subsec:corr errors sims}
+In order to verify the results of Section \ref{section:non vanishing},
+we run simulations of the inverse problem of the 1D heat equation with
+nonvanishing model error \(\modcov = \prcov^2 \). Indeed, including
+model correlation pushes measurements apart, see Fig.~\ref{fig:corr
+  errors}. Code generating Fig.~\ref{fig:corr errors} is located in
+module \texttt{clusterization.py} in the accompanying
+\href{https://github.com/yairdaon/OED}{repository}.
+
+\begin{figure}
+    \centering
+    \includegraphics[height=0.5\textwidth]{dst_modelError4.png}
+    \caption{Model correlation mitigates clusterization. We add a
+      model correlation term to the error terms in the 1D heat
+      equation inverse problem. Lo and behold, measurements are not
+      close anymore and are pushed away thanks to the model error
+      term.}
+  \label{fig:corr errors}
+\end{figure}
+%\input{discussion}
 
 \begin{acks}[Acknowledgments]
-
-%\input{ack}
 This study is a result of research I started during my PhD studies
-under the instruction of Prof. Georg Stadler at the Courant Institute
+under the instruction of Prof.~Georg Stadler at the Courant Institute
 of Mathematical Sciences. I would like to thank him for his great
@@
 \href{https://mathoverflow.net/questions/280168/redistribute-diagonal-entries-of-a-matrix/280203#280203c}{Mathoverflow}.
-
-%\input{funding}
+Last, but certainly not least, I would like to thank the three
+referees, associate editor and editor in chief for providing detailed
+and insightful reviewes that have made this manuscript a whole lot
+better.
+%Having worked on this manuscript alone for several years, they have
+%almost become my coautoh
+
 This research was supported in part by an appointment with the
@@
 % Post-Doctoral Scholarship.
-
 \end{acks}
 
+\bibliographystyle{ba}
+\bibliography{../../lib.bib}
+
 \end{document}

### 2024-07-25 latex/script.py

@@ -1,0 +1,25 @@
+import os
+
+def expand_input(filename):
+    base_dir = os.path.dirname(filename)
+    with open(filename, 'r') as file:
+        for line in file:
+            if line.strip().startswith(r'\input{'):
+                input_file = line.strip()[7:-1]  # Extract filename between \input{ and }
+                input_path = os.path.join(base_dir, input_file + '.tex')
+                yield from expand_input(input_path)
+            else:
+                yield line
+
+def main():
+    input_file = input("Enter the name of the input .tex file: ")
+    output_file = 'flat.tex'
+
+    with open(output_file, 'w') as out_file:
+        for line in expand_input(input_file):
+            out_file.write(line)
+
+    print(f"Flattened file has been written to {output_file}")
+
+if __name__ == "__main__":
+    main()

### 2024-07-25 latex/script.py
Clean up

@@ -1,25 +1,0 @@
-import os
-
-def expand_input(filename):
-    base_dir = os.path.dirname(filename)
-    with open(filename, 'r') as file:
-        for line in file:
-            if line.strip().startswith(r'\input{'):
-                input_file = line.strip()[7:-1]  # Extract filename between \input{ and }
-                input_path = os.path.join(base_dir, input_file + '.tex')
-                yield from expand_input(input_path)
-            else:
-                yield line
-
-def main():
-    input_file = input("Enter the name of the input .tex file: ")
-    output_file = 'flat.tex'
-
-    with open(output_file, 'w') as out_file:
-        for line in expand_input(input_file):
-            out_file.write(line)
-
-    print(f"Flattened file has been written to {output_file}")
-
-if __name__ == "__main__":
-    main()

### 2024-07-25 latex/refs.tex
Tidy up a bit

@@
 
+\answer: I appreciate the many thoughtful comments by yourself, the
+associate editor and the reviewers. I also appreciate the option to
+improve the paper. Please see my answers below in blue.
+
 \begin{description}
@@
     readership;
+    \answer: I have considerably extended the introduction.
+
   \item[b] the motivation and the practical statistical implications are to be
     developed and discussed;
+    \answer: I added a section on the implications in the introduction.
+
   \item[c] the contribution has to be better contextualized with
@@
     designs;
-  \item[d] numerical evidence and code, mentioned in the paper
-    but actually not provided, need to be made available;
+  \answer:
+
+  \item[d] numerical evidence and code, mentioned in the paper but
+    actually not provided, need to be made available;
+
+    \answer: I added a section describing numerical experiments and
+    their results. I also added documentation to the accompanying
+    \href{https://github.com/yairdaon/OED}{repository} and made sure
+    all scripts there run.
+
   \item[e] the mathematical derivations have to be watertight and some
@@
 
+    \answer: I am not sure which specific comment of Reviewer 3 this
+    refers to, please see my answers to Reviewer 3 below.
 \end{description}
@@
 are the parameters being learned) and optimal experimental design.
+
+\answer: I have added considerably to the introduction. See revised
+manuscript and diff file.
 
@@
 process assumption to solve an optimal design problem, and the author
-just takes the assumption as it is. Since the author proves the
-clusterization issue under the independent measurement assumption, I
-expect the author to go deeper and explain why an ordinary data
-generation assumption + a common optimal design leads to a
-“suboptimal” design? Not just by the mathematical proof, but from an
-information theory point of view. For example, does the redundancy
-points mean additional points does not reduce the variance of the
-estimation at all? If that’s the case, how does correlated measurement
-assumption “help” with it?
+just takes the assumption as it is.
 
-\answer: Thanks for your questions. I am not a proponent of using
-correlated noise, I just wanted to show the usefullness of my model by
-rigorously proving how correlated noise mitigates clusterization. I
-have added several lines clearing this issue.
+\answer: If I understand correctly --- I completely agree. I added
+text emphasizing that I do not endorse using correlated errors.
+
+Since the author proves the clusterization issue under the independent
+measurement assumption, I expect the author to go deeper and explain
+why an ordinary data generation assumption + a common optimal design
+leads to a “suboptimal” design?
+\answer: Aagain, my apologies for the misunderstanding.
+
+Not just by the mathematical proof, but from an information theory
+point of view. For example, does the redundancy points mean additional
+points does not reduce the variance of the estimation at all? If
+that’s the case, how does correlated measurement assumption “help”
+with it?
+
+\answer: Bla
 
@@
 
-\answer:
+\answer: I appreciate the comment, I fixed the relevant line in the
+manuscript.
 
@@
 location. However, there are many experiments where repeated
-measurements are possible, and are therefore optimal (if the model is
-correct, which it won’t be). An example, is a chemical engineering
-experiment where one wishes to learn the relationship between a series
-of variables and yield of a chemical reaction. The experiment will
+measurements are possible, and are therefore optimal
+
+\answer: I believe that even in an application where repeated
+experiments are allowed they may be unintuitive. For example, in a
+medical imaging application --- taking a repeated slice in an MRI scan
+might seem suboptimal to some. Of course, there is no argument
+regarding the optimality give a model, which, as you mention might be
+wrong.
+
+(if the model is correct, which it won’t be).
+
+\answer: I believe clusterization may imply a model is somehow in
+adequate, see text below.
+
+An example, is a chemical engineering experiment
+where one wishes to learn the relationship between a series of
+variables and yield of a chemical reaction. The experiment will
 involve specifying the variables, and measuring the yield (the
@@
 
-\answer: again
+\answer: I couldn't agree more. Indeed, the suboptimality of clustered
+design is a matter of personal opinion; once a design criterion and a
+model are agreed upon there can be no disagreement.
 
@@
 
+\answer: Thanks for this comment. I considerably expanded the
+introduction and delegated many proofs to the Supplementary. I also
+added a section on practical implications, reproduced below.
+
 \subsection{Avoiding measurement clusterisation}
@@
 
+\answer: To a certain extend I agree: indeed, the impossibility of
+taking identical measurements is indeed natural. However, this is not
+encoded in the mathematical formulation of the problem. One can then
+take one of two approaches: either accept what the mathematics
+suggest, even if it counterintuitive, or find ways to circumvent the
+mathematics because the result seems non-intuitive.
+
 Secondly, the paper seems to suggest “imposing correlations between
@@
 
+\answer: I do not support imposing correlations in the paper, I just
+used correlations to show how my framework can be used to rigorously
+prove what is widely accepted. Other referees also understood that the
+paper endorses imposing correlations and I accept that this is my
+error in writing. I added a clause clearing the issue.
+
 \subsection{Minor comments}
@@
   adjoint. Suggest changing the one for optimality.
+
+  \answer: I changed $\obs^\star$ to $\opt$.
+
 \item I do not like the use of the term measurements as used in the
-  paper. The measurements are the $\data$’s since these are a result of
-  measuring a physical quantity. However, the paper refers to the
+  paper. The measurements are the $\data$’s since these are a result
+  of measuring a physical quantity. However, the paper refers to the
   quantities that are controlled as the measurements.
+
+  \answer: I was not able to find a better term to replace
+  "measurements", so I stuck to it.
+
 \item Proof of Proposition 12 uses $\eps$ as part of the
@@
   (e.g. equation (1)).
+
+  \answer: Absolutely correct, I changed it to $\nu$. You will not
+  find the revised proof in the manuscript since I moved it to the
+  Supplementray.
 \end{itemize}
@@
 numerical simulations and results; only a link to a Github repository
-is provided, which is however not functioning. Evidently, this issue
-needs to be fixed before any future consideration can be given to the
-manuscript. For example, a detailed description of the repeated
-experiments could be provided either in the main article or in the
-supplement, with tables indicating the various values of the
-parameters for the simulations and the precise percentages
-obtained. Also, the code needs to be publicly available for
-reproducibility.
+is provided, which is however not functioning.
+
+\answer: I apologize for not properly maintainig the repository. It is
+now functioning and well-documented.
+
+Evidently, this issue needs to be fixed before any future
+consideration can be given to the manuscript. For example, a detailed
+description of the repeated experiments could be provided either in
+the main article or in the supplement, with tables indicating the
+various values of the parameters for the simulations and the precise
+percentages obtained. Also, the code needs to be publicly available
+for reproducibility.
+
+\answer: I appreciate the suggestion. I added a section on numerical
+simulations to the main atricle. It is reproduced below.
 
 \subsection{Further comments}
-I have a number of further comments in regards to the results and the overall presentation of
-the manuscript.
+I have a number of further comments in regards to the results and the
+overall presentation of the manuscript.
+
 \begin{itemize}
@@
   noise/equally spaced design (e.g. [1]) or measurement loctions
-  sampled uni- formly at random (e.g. [2]). There are also some
+  sampled uniformly at random (e.g. [2]). There are also some
   results with more general design under conditions on the
   fill-distance of the grid (e.g. [3]). All of these specifically
-  prevent clusteri- sation, which is key to the statistical
+  prevent clusterisation, which is key to the statistical
   analysis. Hence, it is not clear to me even if D-optimal design
   should be pursued at all, if it indeed it leads to clusterisation as
-  implied by the present 1paper. I think a discussion on this would be
+  implied by the present paper. I think a discussion on this would be
   illuminating for the reader and help draw a connection to the
@@
 
-  Answer: It is impossible to visually discern the measurements, even
-  when resolution is increased.
+  \answer: It is impossible to visually discern the measurements, even
+  when resolution is increased. These measurements are not identical
+  numerically, but they are identical to five decimal digits. Since I
+  find these points via optimization over point location in
+  $\Omega=[0,1]$ I view an error $<10^{-5}$ as reasonable.
 
@@
 
-  Answer: That is a great point! I added the followng discussion:
+  \answer: That is a great point!
 
@@
 
-\item Also, it should be relatively easy here to also study empirically the effect of the correlated
-error model in preventing clusterisation. Investigating all these issues in the context of the
-presented numerical set up would considerably enlarge the breath of the present work.
+\item Also, it should be relatively easy here to also study
+  empirically the effect of the correlated error model in preventing
+  clusterisation. Investigating all these issues in the context of the
+  presented numerical set up would considerably enlarge the breath of
+  the present work.
+
+  \answer: I added relevant simulations to a new section Numerical
+  Experiments.
 
@@
 
-\item Denoting the optimal design in Theorem 1 by simply O is somewhat confusing, cf. the
-equation in the second item. Perhaps a separate notation such as Ō would help readability.
+  \answer: I removed the statement of Theorem 1 from the introudction.
 
-\item Section 1.2 seems altogether unnecessary; the inverse regression model considered in the
-paper are a modelisation of many real-world phenomena, and they are routinely studied in
-statistical papers.
+\item Denoting the optimal design in Theorem 1 by simply O is somewhat
+  confusing, cf. the equation in the second item. Perhaps a separate
+  notation such as Ō would help readability.
 
-Answer: Thanks for this comment but I would rather keep this section
-here since...
+  \answer: Thank you for the comment, I completely agree. Optimal
+  designs are now denoted $\opt$.
+
+\item Section 1.2 seems altogether unnecessary; the inverse regression
+  model considered in the paper are a modelisation of many real-world
+  phenomena, and they are routinely studied in statistical papers.
+
+  \answer: I am glad you view this part as unnecessary, but I feel the
+  model might feel too abstract to some practitioners. Thus, I choose
+  to keep this part to avoid complaints from other readers.
 
@@
 
+  \answer: Basically I meant that its eigenvalues decrease
+  "quickly". I understand this is not a commot or well-defined term so
+  I removed the relevant sentence.
+
 \item Section 4: the conclusion reached at the end of the section on
   p.15 seem to imply that in the noiseless limit, repeating any
-  measurement does not improve the objective criterion, if corre-
-  lated error model are present. Is it clear however that adding any
+  measurement does not improve the objective criterion, if correlated
+  error model are present. Is it clear however that adding any
   (non-reapted) measurement always strictly increases it? I.e., there
   could be situations where adding one measurement does not change the
-  value of the optimised criterion? Also, can the conclusion drawn
-  here be extended to the case of noisy observations σ > 0?
+  value of the optimised criterion?
+
+  \answer: Great comment! Such cases are possible but I view them as
+  pathological. See short paragraph below.
+
+  Also, can the conclusion drawn here be extended to the
+  case of noisy observations $\sigma > 0$?
+
+  \answer: I do not expect this conclusion to hold: if noise is
+  present, any measurement will result in some information gain and an
+  increase in design criterion.
 
@@
   posterior measure be still defined as in Gaussian process
-  regression, e.g.[4]
+  regression, e.g.[4].
+
+  \answer: I believe a posterior can be defined, and so a KL
+  divergence from said posterior to prior could be calculated and
+  $\tar$ evaluated. However, taking $\sigma = 0$ in the main theorem
+  implies $\tar \equiv \infty$. I believe this is because do not know
+  how to mitigate this
 
@@
   which is a user specified quantity. Hence, it seems to me that all
-  sort of design behaviour under the D-optimal crite- rior can occur
-  by engineering the prior. It would perhaps be informative to study
-  in more details some representative example of inverse problem and
+  sort of design behaviour under the D-optimal criterion can occur by
+  engineering the prior. It would perhaps be informative to study in
+  more details some representative example of inverse problem and
   Gaussian prior. For example, the recovery of the initial condition
-  with the ‘Matérn-like’ Gaussian prior laid out in the sup- plement
-  are good candidates. For these it is still not clear how the
-  conclusion from Theorem 1 translates into the Figure 1, also because
-  I am unsure about the applicability of the result to point
-  evaluations. In particular, how the conclusion that with m = 4
-  measurements the D-optimal procedure will aim to ignore the third
-  and fourth eigenvalue/function was drawn?
+  with the ‘Matérn-like’ Gaussian prior laid out in the supplement are
+  good candidates.
+
+  \answer: the Matern-like prior actually arises by using a
+  Laplacian-like operator as a prior, see Rue and Lindgren \cite{}. So
+  a study of this example (with and without correlated errors) is
+  already included in the paper. I am afraid implementing another
+  inverse problem is out of scope for the present paper since I was
+  not able to install common packages for inverse problems on my machine.
+
+  For these it is still not clear how the conclusion from Theorem 1
+  translates into the Figure 1, also because I am unsure about the
+  applicability of the result to point evaluations.
+
+  \answer: Indeed, point evaluations $\delta_x$ are not in any
+  function space I consider in the paper since my analysis is limited
+  to Hilbert spaces. Point evaluations can be approximated well,
+  though so in my opinion this is not a deal breaker. The only caveat
+  is that upon approximating point evaluations e.g.~by $R(x;t) =
+  t\mathbb{1}_{-t/2, t/2}(x)$ we will need to change the norm
+  constraints to the norsomething different (i.e.~$\|R(\cdot;t\|$) but
+  that does not change the analysis. I skim over this issue in the
+  paper since I think it is confusing and adds little context.
+
+  In particular, how the conclusion that with $m = 4$ measurements the
+  D-optimal procedure will aim to ignore the third and fourth
+  eigenvalue/function was drawn?
+
+  \answer: I explain this in more detail in the paper. See text below.
+
 \end{itemize}

### 2024-07-25 clusterization.py

@@ -82,7 +82,7 @@ def main():
     plt.xlabel("Eigenvector", fontsize=FS)
     plt.ylabel("Precision", fontsize=FS)
     plt.tight_layout()
-    plt.savefig(f"latex/FgammaFstar_modelError{model_error}.png")
+    plt.savefig(f"latex/figs/FgammaFstar_modelError{model_error}.png")
     
     
     ###################################
@@ -128,7 +128,7 @@ def main():
     ax.set_xlim(0,1)
     plt.tight_layout()
 
-    plt.savefig(f"latex/{transform}_modelError{model_error}.png")
+    plt.savefig(f"latex/figs/{transform}_modelError{model_error}.png")
     #plt.show()
 
     ##############################
@@ -176,7 +176,7 @@ def main():
     plt.xlabel(r"$x \in \Omega$", fontsize=fs)
     plt.legend(title='eigenvector, weighted')
     plt.tight_layout()
-    plt.savefig(f"latex/eigenvectors_{transform}_modelError{model_error}.png")
+    plt.savefig(f"latex/figs/eigenvectors_{transform}_modelError{model_error}.png")
    
 
     

### 2024-07-25 latex/old.tex
Old flattened file

--- 2024-07-21 latex/flat.tex
@@
+\documentclass[ba]{imsart}
+
 % \documentclass[ba,preprint]{imsart}% use this for supplement article
-\documentclass[ba]{imsart}
-%\input{definitions}
 \usepackage{amsmath}
@@
 
-%% \theoremstyle{plain}
-%% \newtheorem{theorem}{Theorem}
-%% \newtheorem{corollary}[theorem]{Corollary}
-%% \newtheorem{lemma}[theorem]{Lemma}
-%% \newtheorem{proposition}[theorem]{Proposition}
-%% \theoremstyle{definition}
-%% \newtheorem{definition}[theorem]{Definition}
-%% \newtheorem{example}[theorem]{Example}
-%% \newtheorem{conjecture}[theorem]{Conjecture}
-%% \theoremstyle{remark}
-%% \newtheorem{remark}[theorem]{Remark}
+\theoremstyle{plain}
+\newtheorem{theorem}{Theorem}
+\newtheorem{corollary}[theorem]{Corollary}
+\newtheorem{lemma}[theorem]{Lemma}
+\newtheorem{proposition}[theorem]{Proposition}
+\theoremstyle{definition}
+\newtheorem{definition}[theorem]{Definition}
+\newtheorem{example}[theorem]{Example}
+\newtheorem{conjecture}[theorem]{Conjecture}
+\theoremstyle{remark}
+\newtheorem{remark}[theorem]{Remark}
 
@@
     \centering
-    \includegraphics[height=0.5\textwidth]{example.pdf}
+    \includegraphics[height=0.5\textwidth]{figs/example.png}
     \caption{Measurement clusterization for optimal designs when
@@
     \centering
-    \includegraphics[width=\textwidth]{eigenvectors.pdf}
+    \includegraphics[width=\textwidth]{figs/eigenvectors.png}
     \caption{D-optimal measurement locations ($m=4$ measurements) and
@@
 many others.
-
-%\input{prelims}
 \section{Preliminaries and Notation}\label{section:prelim}
@@
 posed in the Introduction.% Section \ref{section:intro}.
-
-%\input{vanishing}
 \section{D-Optimal Designs Without Model Error}\label{section:vanishing}
@@
 
-\char*
-
+\begin{theorem}\label{thm:char}
+  Let:
+  \begin{itemize}
+    \item The D-optimality design criterion
+    \cite{AlexanderianGloorGhattas14}:
+    \begin{align*}
+      \begin{split}
+        \tar(\obs) %:&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
+        % 
+        % 
+        % 
+        &= \frac12 \log \det ( I + \sigma^{-2} \prcov^{1/2} \fwd ^*
+        \obs^* \obs \fwd \prcov^{1/2}), 
+      \end{split}
+    \end{align*}
+  \item \(\obs\) a D-optimal design operator
+    \begin{equation*}
+      \obs = \argmax_{\|\meas_j\| = 1, j=1,\dots,m}\tar(\obs),
+    \end{equation*}
+  \item \(\{\lambda_i\}_{i=1}^\infty\) eigenvalues of
+    \(\fwd\prcov\fwd^*\) in decreasing order of magnitude.
+    %% \item \(\{\ev_i\}_{i=1}^\infty\) their corresponding eigenvectors.
+  \item \(\{\eta_i\}_{i=1}^\infty\) eigenvalues of \(\obs^*\obs\).
+
+  \end{itemize}
+
+  Then:
+  \begin{enumerate}
+  \item  \(\tr{\obs^*\obs} = m\).
+  \item \(\obs^*\obs\) and \(\fwd\prcov\fwd^*\) are simultaneously
+    diagonalizable.
+  \item \(k := \rank \obs^*\obs \leq m\) and
+    \begin{equation*}
+      \tar(\obs) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i). %= \frac12 \sum_{i=1}^{m} \log (1 + \sigma^{-2}\lambda_i\eta_i).
+    \end{equation*}
+  %% \item 
+  %%   \begin{equation*}
+  %%     k = \argmax \left \{ k:\lambda_k^{-1} < \sigma^{-2}\frac{m}{k} + \frac{1}{k} \sum_{j=1}^{k}
+  %%     \lambda_j^{-1} \right \}.
+  %%   \end{equation*}
+  \item
+    \begin{equation*}
+        \eta_i = \begin{cases}
+          \frac{m}{k} - \sigma^2 \lambda_i^{-1} + \sigma^2 \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1} & 1 \leq i \leq k \\
+          0 & i > k 
+        \end{cases}.
+    \end{equation*}
+  \item The covariance of the pushforwad \(\fwd_{*} \post\) is \(\left
+    ( (\fwd \prcov \fwd^*)^{-1} + \sigma^{-2} \obs^*\obs \right
+    )^{-1}\) and its eigenvalues are
+    \begin{equation*}
+      \theta_i =
+      \begin{cases}
+        \left(\frac{\sum_{j=1}^k \lambda_j^{-1} + \sigma^{-2}m}{k} \right )^{-1} & i \leq k \\
+        \lambda_i &  i > k 
+      \end{cases}
+    \end{equation*}
+  \end{enumerate}
+\end{theorem}
 \begin{proof}
@@
 
-%\input{why}
 \section{Answer to Question \ref{q:why}: The cause of clusterization}\label{section:why}
@@
 % Post-Doctoral Scholarship.
-
 \end{acks}

### 2024-07-25 latex/flat.tex
Clean up
Before dropping spacebars in labels

@@
-% \documentclass[ba,preprint]{imsart}% use this for supplement article
 \documentclass[ba]{imsart}
@@
 \usepackage{amsthm}
-%% \usepackage{amsaddr}
 
@@
 \newcommand{\normal}{\mathcal{N}}
-\newcommand{\pr}{\mu_{\textup{pr}}} %Prior measure
-%\newcommand{\post}{\mu_{\textup{post}}^{\data, \obs}} % Posterior measure
-\newcommand{\post}{\mu_{\textup{post}}} % Posterior measure for slides
-\newcommand{\prmean}{\param_{\textup{pr}}} % Prior mean
-\newcommand{\postmean}{\param_{\textup{post}}} % Posterior mean
-\newcommand{\postcov}{\Gamma_{\textup{post}}} % Posterior covariance
-\newcommand{\prcov}{\Gamma_{\textup{pr}}} % Prior covariance
-\newcommand{\modcov}{\Gamma_{\textup{model}}} % Model covariance
-\newcommand{\tmp}{\mathcal{G}}
+\newcommand{\pr}{\mu_{\textup{pr}}} \newcommand{\post}{\mu_{\textup{post}}} \newcommand{\prmean}{\param_{\textup{pr}}} \newcommand{\postmean}{\param_{\textup{post}}} \newcommand{\postcov}{\Gamma_{\textup{post}}} \newcommand{\prcov}{\Gamma_{\textup{pr}}} \newcommand{\modcov}{\Gamma_{\textup{model}}} \newcommand{\tmp}{\mathcal{G}}
 \newcommand{\meas}{\mathbf{o}}
-\newcommand{\ev}{\mathbf{e}} % eigenvector 
-\newcommand{\func}{\mathbf{a}}
+\newcommand{\ev}{\mathbf{e}} \newcommand{\func}{\mathbf{a}}
 \newcommand{\tr}[1]{\textup{tr}\left \{#1 \right \} }
@@
 \newcommand{\rank}{\textup{rank}\ }
-\newcommand{\des}{\eta} % vector of design parameters
-\newcommand{\sigsqr}{\sigma^2}
-
-%% \newcommand{\opt}{\mathfrak{D}}
+\newcommand{\des}{\eta} \newcommand{\sigsqr}{\sigma^2}
+
 \newcommand{\opt}{\mathcal{D}}
-\newcommand{\postopt}{\mu_{\textup{post}}^{\data, \opt}} % Posterior measure
-
-%% %\usepackage{comment}% http://ctan.org/pkg/comment
-%% %% %% %\excludecomment{proof}
-%% %% \excludecomment{figure}
-%% %% \let\endfigure\relax
-
-%% %\overfullrule=0pt
+\newcommand{\postopt}{\mu_{\textup{post}}^{\data, \opt}}
 
@@
 \newtheorem{remark}[theorem]{Remark}
-%
  \pubyear{2021}
@@
 Finally, \textbf{(c)} We leverage this characterization to argue that measurement clusterization arises as a consequence of the pigeonhole principle: when more measurements are taken than there are locations where the select eigenvectors are large and others are small --- clusterization occurs.
-%
-%% In summary, in this study we shed light on an often ignored issue with
-%% Bayesian D-optimal designs. We characterize such designs and show why
-%% clusterization arisesit arisestheir inherent properties, propose
-%% strategies to address clusterization, and provide insights into the
-%% fundamental attributes underlying measurement design optimization.
  \end{abstract}
 
-%% ** Keywords **
 \begin{keyword}[class=MSC]
@@
 described by the following three equations:
-\begin{subequations}%\label{eq:heat equation}
-  \begin{alignat}{2}
+\begin{subequations}\begin{alignat}{2}
     u_t &= \Delta u &&\qquad \text{in } [0,1] \times [0,\infty), \label{eq:heat1}\\
@@
 
-%% first: blue, second: red, third: green, fourth: black,
-%% fifth:magenta
 \begin{figure}
     \centering
-    \includegraphics[height=0.5\textwidth]{dst_modelError0.png}
+    \includegraphics[height=0.5\textwidth]{figs/dst_modelError0.png}
     \caption{Measurement clusterization for D-optimal designs for the
@@
 tried to address some seemingly simple yet fundamental questions:
-%
 Why does imposing correlations between observations alleviate
 measurement clusterization?
-%
 Is measurement clusterization a generic phenomenon? 
-%
 And, most importantly: Why does measurement clusterization occur?
-%
-%Should we aim to avoid measurement clusterization?
-%
-%Is it possible to substitute an optimal clustered design with an
-%equally optimal non-clustered design?
-%Can an optimal clustered design be relpaced with an equally optimal
-%non-clustered design?
 
@@
   generic phenomenon?}
-  %\subsection{An answer for Question \ref{q:generic}: Genericity of measurement clusterization}
-  %% Computer implementation of Lemma \ref{thm:char} for the inverse
-  %% problem outlined in Section \ref{section:how} also generates \(\obs\)
-  %% as a solution to the D-optimal design problem.  Furthermore,
 We give two complementing answers to this question. First, from a
@@
 
-  %% In Section \ref{section:vanishing}, we provide an insightful
-  %% explanation for the optimality of clustered designs when no model
-  %% error is present. We demonstrate that for our model, a D-optimal
-  %% design measures only a small subset of prior eigenvectors, which are
-  %% the prior eigenvectors with largest power spectrum. In real-life
-  %% problems there are limited number of locations where: (a) the
-  %% relevant prior eigenvectors are large, and other prior eigenvectors
-  %% with large power in the prior spectrum are zero. Then measurement
-  %% clusterization is a result of the pigeonhole principle, where there
-  %% are more measurements than measurement locations satisfying (a) and
-  %% (b).
-
-  %% We conjecture that the prevalence of measurement clusterization
-  %% arises due to the ease of discovering clustered designs.
-
-%% \item \label{q:avoid} \textbf{Should we aim to avoid measurement
-%%   clusterization?} Based on the analysis conducted in this study, we
-%%   did not find any compelling reason to explicitly avoid optimal
-%%   clustered designs.
-
-%% \item \label{q:replace} \textbf{Is it possible to substitute an
-%%   optimal clustered design with an equally optimal non-clustered
-%%   design?} In Section \ref{section:vanishing}, we answer this question
-%%   in the affirmative, although we show that numerical experiments
-%%   conducted using our model indicate a strong preference for clustered
-%%   designs.
 \end{enumerate}
@@
 \mathbb{R}^{k \times m}\) has unit norm columns.
-
-%% Finally, in Lemma \ref{lemma:lax} we generalize a lemma for
-%% calculating \(\frac{\der}{\der t} \log \det (I + X(t))\), where
-%% \(X(t)\)is an operator valued function \cite{Lax07}.
-
-%% Finally, we must stress that we do not advocate any particular
-%% solution to clusterization. We do not suggest that clustered designs
-%% should be abandoned, nor do we support the use of clustered designs in
-%% any circumstances. Our goal is to investigate an intriguing phenomenon
-%% of clusterization and seek a deeper understanding for it. Different
-%% researchers in different disciplines have differenot viewed as a problem. For
-%% example, \cite{lozan} suggest in the introduction section that
-%% repeated measurements are necessary. Our results align with that
-%% common wisdom and settle the seemed contradiction.
 
@@
 to the finite-dimensional case.
-
-%% \subsection{Notation Summary}\label{subsec:notation}
-%%   Our notation is summarized below. We let:
-%%   \begin{itemize}
-%%   \item \(\hilp, \hilo\) Hilbert spaces.
-%%   \item \(\fwd:\hilp \to \hilo\) a linear compact operator.
-%%   \item \(\pr \sim \mathcal{N}(0, \prcov)\) prior Gaussian measure on
-%%     \(\hilp$, with prior covariance operator \(\prcov:\hilp \to \hilp$.
-%%   \item \(\obs: \hilo \to \mathbb{R}^m\) measurement operator, where \(m
-%%     \in \mathbb{N}\) is the number of measurements taken.
-%%   \item \(\sigma^2 \in \mathbb{R}_{+}\) observation noise variance.
-%%   \item \(\modcov\) model error covariance operator.
-%%     %% \(\data = \obs \fwd \param + \eps$, where \(\eps \in
-%%     %% \mathbb{R}^m\) isiid \(\mathcal{N}(0, \sigma^2)\) noise.
-%%   \item \(\Sigma(\obs) = \obs \modcov \obs^* + \sigma^2I$. 
-%%   \item \(\post\) the posterior measure, with covariance \(\postcov$.
-%%   \item A D-optimality design criterion
-%%     \cite{AlexanderianGloorGhattas14}:
-%%     \begin{align*}
-%%       \begin{split}
-%%         \tar(\obs) :&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
-%%         % 
-%%         % 
-%%         % 
-%%         &= \frac12 \log \det ( I + \prcov^{1/2} \fwd ^* \obs^* \Sigma(\obs)^{-1} \obs
-%%         \fwd \prcov^{1/2}).
-%%       \end{split}
-%%     \end{align*}
-%%   %% \item \(\{\lambda_i\}_{i=1}^\infty\) eigenvalues of \(\fwd\prcov\fwd^*$
-%%   %%   in decreasing order of magnitude.
-%%   %% %% \item \(\{\ev_i\}_{i=1}^\infty\) their corresponding eigenvectors.
-%%   %% \item \(\{\eta_i\}_{i=1}^\infty\) eigenvalues of \(\obs^*\obs$.
-%%   \end{itemize}
-
-%% \subsection{Sequential vs Simultaneous Optimization}\label{subsec:seq vs sim}
-%% From defintion \ref{def:d optimality} we wish to characterize solution(s) of the
-%% following optimization problem for \(\tar$. %%: (\hilo^*)^m \to \R$:
-%% \begin{align}\label{eq:optimization}
-%%   \obs^{\star} := \argmax_{\obs} \tar( \obs ) 
-%%   = \argmax_{\obs} \frac12 \log \det 
-%%   (I + \prcov^{1/2} \fwd^*\obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}),
-%% \end{align}
-%% where \(\obs\) is constrained to some allowed set of observations. We
-%% call this problem ``simultaneous optimization'', since all
-%% observations are decided on simulatneously.
-
-%% For computational reasons, one may prefer to find the best
-%% observations in a sequential manner. Denote
-%% \begin{equation}\label{eq:def obs_k}
-%%   \obs_k := (\meas_1,\dots, \meas_k)^t,  k\leq m.
-%% \end{equation}
-%% Sequential optimal design proceeds as follows. Find \(\meas_1\) by
-%% maximizing \(\tar(\obs_1)$. Then, keeping \(\meas_1\) fixed --- find
-%% \(\meas_2\) as the maximizer of \(\tar(\obs_2)$. Then, find \(\meas_3\) by
-%% keeping \(\meas_1,\meas_2\) fixed and taking \(\meas_3\) as the maximizer
-%% of \(\tar(\obs_3)$. Continue this way until \(\obs_m = \obs\) is
-%% found, where \(m\) is the number of available observations. %% It is
-%% %% important to notice that this scheme does not require actually
-%% %% observing data --- in \eqref{eq:objective} data is averaged out.
-
-%% The analysis in this paper is conducted for the general simultaneous
-%% optimization case. The sequential optimization case is dealt with in
-%% section \ref{subsec:clusterization sequential}. It is important to
-%% note, however that all conclusions we arrive at for the simultaneous
-%% case easily specialize to the sequential case by considering the
-%% posterior as the next sequential step's prior.
 
@@
 \end{proposition}
-
-%% The same arguments justifying \eqref{eq:tar grad} hold here, and thus:
-
-%% \begin{align}\label{eq:grad constraints}
-%% \nabla \phi_j(\obs) = \obs^* e_j e_j^t = \meas_j e_j^t , j=1,\dots,m,
-%% \end{align}
-%% where $\nabla \phi_j(\obs) \in \hilo^m$. As noted at the end of
-%% Section \ref{section:objective},
-
-%% The gradients $\nabla \tar(\obs)$ and $\nabla \phi_j(\obs)$ are best
-%% thought of as row vectors.
 
@@
 
-%% Imposing correlations between observations alleviate measurement clusterization
  \section{Answer to Question \ref{q:mitigate}: Model error mitigates clusterization}\label{section:non vanishing}
@@
 model error $\modcov$ --- answering Question \ref{q:mitigate} posed in
-the Introduction.% Section \ref{section:intro}.
-\section{D-Optimal Designs Without Model Error}\label{section:vanishing}
+the Introduction. \section{D-Optimal Designs Without Model Error}\label{section:vanishing}
 Our goal in this section is to prove Theorem \ref{thm:char} which
@@
     \(\fwd\prcov\fwd^*\) in decreasing order of magnitude.
-    %% \item \(\{\ev_i\}_{i=1}^\infty\) their corresponding eigenvectors.
 \item \(\{\eta_i\}_{i=1}^\infty\) eigenvalues of \(\opt^*\opt\).
@@
     \begin{equation*}
-      \tar(\opt) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i). %= \frac12 \sum_{i=1}^{m} \log (1 + \sigma^{-2}\lambda_i\eta_i).
-    \end{equation*}
-  %% \item 
-  %%   \begin{equation*}
-  %%     k = \argmax \left \{ k:\lambda_k^{-1} < \sigma^{-2}\frac{m}{k} + \frac{1}{k} \sum_{j=1}^{k}
-  %%     \lambda_j^{-1} \right \}.
-  %%   \end{equation*}
+      \tar(\opt) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i). \end{equation*}
 \item
@@
 \emph{precision} $\left(\fwd \prcov \fwd^*\right )^{-1}$ corresponds
-to a graduated lab cylinder. % (see Fig\ref{fig:cylinder}) Each
-cylinder $i$ is filled, a-priori, with green liquid of
+to a graduated lab cylinder. cylinder $i$ is filled, a-priori, with green liquid of
 $\lambda_i^{-1}$ volume units. We are allowed $m$ measurement, so we
@@
   vs not} for an illustration.
-%% \begin{figure}%{r}{0.25\textwidth} 
-%%     \centering
-%%     \includegraphics[width=5cm, height=5cm]{cylinders.jpg}
-%%     \caption{Three graduated lab cylinders, corresponding to three
-%%      eigenvectors. Prior precision eigenvalues not shown.}
-%%     \label{fig:cylinder}
-%% \end{figure}
 
@@
     \centering
-    \includegraphics[width=\textwidth]{eigenvectors_dst_scaled.png}
+    \includegraphics[width=\textwidth]{figs/eigenvectors_dst_scaled.png}
     \caption{D-optimal measurement locations ($m=4$ measurements) and
@@
  \section{Numerical Experiments}
-%We provide numerical evidence supporting our conclusions.
 
@@
     \centering
-    \includegraphics[height=0.5\textwidth]{simulations.png}
+    \includegraphics[height=0.5\textwidth]{figs/simulations.png}
     \caption{Fraction of clustered $A$ for $AA^t = M$ and $M$
@@
     \centering
-    \includegraphics[height=0.5\textwidth]{dst_modelError4.png}
+    \includegraphics[height=0.5\textwidth]{figs/dst_modelError4.png}
     \caption{Model correlation mitigates clusterization. We add a
@@
 \end{figure}
-%\input{discussion}
 
@@
 better.
-%Having worked on this manuscript alone for several years, they have
-%almost become my coautoh
 
@@
 NSF, ORAU/ORISE, or DOE.
-% This work was also supported by The Raymond and Beverly Sackler
-% Post-Doctoral Scholarship.
  \end{acks}

### 2024-07-25 latex/prelims.tex
Before dropping spacebars in labels

@@
 forward operator \(\fwd\) is linear. In the inverse problem of the 1D
-heat, the forward operator is determined by the time evolution of the
-1D heat equation \eqref{eq:heat1} and \eqref{eq:heat2}.
+heat equatin, the forward operator is determined by the time evolution
+of the 1D heat equation \eqref{eq:heat1} and \eqref{eq:heat2}.
 

### 2024-07-31 latex/flat.tex
Intro ready

@@
 \usetikzlibrary{patterns}
-\pgfplotsset{compat=1.9} 
+\pgfplotsset{compat=1.9} \usepackage{xcolor}
+
 \newcommand{\R}{\ensuremath{\mathbb{R}}}
@@
 \newtheorem{remark}[theorem]{Remark}
+
 \pubyear{2021}
@@
 KL divergence is defined as
-\begin{equation}\label{eq:basic KL}
+\begin{equation}\label{eq:basic_KL}
   D_{KL}\left (\Pr(\param|\data)||\Pr(\param)\right ) =  \sum_{\param} \log
@@
 D-optimal design minimizes the determinant of the posterior covariance
-\cite{chaloner1995}. In Section \ref{subsec:D optimal design} we give
+\cite{chaloner1995}. In Section \ref{subsec:D_optimal_design} we give
 a more general definition of the D-optimality criterion that applies
@@
 distribution $u_0\in L^2([0,1])$ that satisfies the homogeneous
-Dirichlet boundary condition is a linear combination of sines
-$\ev_n(x) = \sin(2\pi n x)$. Now, note that $\ev_n$ are in fact
-eigenvectors of the Laplacian: $\Delta \ev_n = -(2\pi n)^2\ev_n$. The
-linearity of the Laplacian implies the linearity of the heat equation
-and we can solve in Fourier / frequency space. It is important to note
-that when solving the heat equation we \emph{invert} the
-Laplacian. I.e.~eigenvalues of the linear operator that describes the
-time evolution of heat are $(2\pi n)^{-2}$.
+Dirichlet boundary condition eq.~\eqref{eq:heat2} is a linear
+combination of sines $\ev_n(x) = \sin(\pi n x)$, so $u_0 = \sum_{n\geq
+  1} a_n \ev_n$ for some $\{a_n\}_{n\geq 1}$. Now, note that $\ev_n$
+are in fact eigenvectors of the Laplacian: $\Delta \ev_n = -\pi^2
+n^2\ev_n$. It should not be hard to believe that eq.~\eqref{eq:heat1}
+implies $u(\cdot, T) = \sum_{n\geq 1} a_n \exp(-\pi^2 n^2T ) \ev_n$
+\cite{renardy2006PDE}. Thus, eigenvectors of the linear operator that
+describes the time evolution of heat are $\ev_n(x) = \sin(\pi n x)$
+with corresponding eigenvalues
+\begin{equation}\label{eq:decay}
+  \exp(-\pi^2 n^2T ).
+\end{equation}
 
@@
 since heat spreads in a diffusive manner and any roughness in the
-initial condition is quickly smoothed, see Supplementary movies S1 and
-S2 (code generating these movies is located in module
-\texttt{movies.py} in the accompanying
+initial condition is quickly smoothed, as implied by the squared
+exponential decay as $n\to\infty$ in eq.~\eqref{eq:decay}. See
+Supplementary movies S1 and S2 for an illustration of this phenomenon
+(code generating these movies is located in module \texttt{movies.py}
+in the accompanying
 \href{https://github.com/yairdaon/OED}{repository}). For ill-posed
@@
 
-Specifically, we set a prior on the initial condition $u_0 \sim
-\mathcal{N}(0, (-\Delta)^{-1})$, where we impose the homogeneous
-Dirichlet boundary condition on the Laplacian. We choose
-$(-\Delta)^{-1}$ as our covariance operator following \cite{Stuart10}:
-our choice ensures the power of the negative Laplacian inside the
-parentheses is larger than $d/2$, where $d=1$ is the dimension of our
-computational domain $\Omega$. One can intuitively understand this
-choice from Weyl's asymptotics, that states that the eigenvalues of
-Laplacian increase as $n^{d/2}$. Thus, our choice of covariance
-$(-\Delta)^{-1}$ ensures higher modes decay sufficiently quickly, so
-prior samples have sufficient regularity. Our choice of prior
-covariance ensures the posterior is well-defined and well-behaved
-\cite[Theorem 3.1 and Lemma 6.25]{Stuart10}.
+We choose a prior for the initial condition $u_0 \sim \mathcal{N}(0,
+(-\Delta)^{-1})$, with homogeneous Dirichlet boundary condition. Choosing a prior This
+choice ensures the posterior is well-defined and prior realizations
+are well-behaved, see \cite[Theorem 3.1 and Lemma 6.25]{Stuart10} for
+details.
 
@@
   Ucinski05, neitzel2019sparse}. This phenomenon is illustrated for
-our toy inverse problem in Fig.~\ref{fig:clusterization illustration},
+our toy inverse problem in Fig.~\ref{fig:clusterization_illustration},
 where D-optimal measurement locations are shown for different numbers
@@
       chosen according to the Bayesian D-optimality criterion of
-      Theorem \ref{thm:d optimality}. Measurement locations are
+      Theorem \ref{thm:d_optimality}. Measurement locations are
       plotted over the computational domain \(\Omega = [0, 1]\)
@@
       and the fifth (magenta).}
-  \label{fig:clusterization illustration}
+  \label{fig:clusterization_illustration}
 \end{figure}
@@
 spaces and study two inverse problems: (a) In Sections
-\ref{section:prelim} and \ref{section:D and grad} we propose a novel
+\ref{section:prelim} and \ref{section:D_and_grad} we propose a novel
 relaxed model for an inverse problem where D-optimality maintains
@@
   designs that exhibit clusterization with high probability in Section
-  \ref{subsec:lemma sims}. Thus, given the genericity of our model, we
+  \ref{subsec:lemma_sims}. Thus, given the genericity of our model, we
   expect measurement clusterization to be a generic and ubiquitous
@@
   between observations alleviate measurement clusterization?} In
-  Section \ref{section:non vanishing}, we rigorously demonstrate the
+  Section \ref{section:non_vanishing}, we rigorously demonstrate the
   role of model error in mitigating clusterization, thereby
@@
   measurements result in no gain in design criterion when observation
-  error tends to zero. Moreover, in Section \ref{subsec:corr errors
-    sims}, we show that an error term corresponding to correlations
-  between measurements mitigates clusterization in the inverse problem
-  of the 1D heat equation.
+  error tends to zero. Moreover, in Section
+  \ref{subsec:corr_errors_sims}, we show that an error term
+  corresponding to correlations between measurements mitigates
+  clusterization in the inverse problem of the 1D heat equation.
 
@@
 
-\subsection{Forward Problems}\label{subsec:abstract OED}
+\subsection{Forward Problems}\label{subsec:abstract_OED}
 In this section we give definitions and notations for forward problems,
@@
 data is acquired via
-\begin{align}\label{eq:inverse problem}
+\begin{align}\label{eq:inverse_problem}
   \data := \obs (\fwd \param + \eps') + \eps = \obs \fwd \param + \obs \eps' + \eps.
@@
 where
-\begin{align}\label{eq:modcov explained}
+\begin{align}\label{eq:modcov_explained}
   \begin{split}
@@
 
-\subsection{Bayesian Linear Inverse Problems}\label{subsec:bayesian inverse problems}
+\subsection{Bayesian Linear Inverse Problems}\label{subsec:bayesian_inverse_problems}
 In the previous section, we saw how a parameter $u\in \hilp$ is
@@
 \(\hilp\) \cite{Stuart10}. For example, for the inverse problem of the
-1D heat equation we chose $\prcov = -\Delta^{-1}$ --- see Section
+1D heat equation we chose $\prcov = (-\Delta)^{-1}$ --- see Section
 \ref{subsec:toy}. Note that \(\fwd \prcov \fwd^*\) is the prior
@@
 
-\subsection{Bayesian D-Optimal Designs}\label{subsec:D optimal design} 
+\subsection{Bayesian D-Optimal Designs}\label{subsec:D_optimal_design} 
 A Bayesian D-optimal design maximizes the expected KL divergence
@@
 posterior and prior measures, the KL divergence is defined,
-analogously to eq.~\ref{eq:basic KL}, via the Radon-Nikodym
+analogously to eq.~\ref{eq:basic_KL}, via the Radon-Nikodym
 derivative:
@@
 
-\begin{theorem}[Alexanderian, Gloor, Ghattas \cite{AlexanderianGloorGhattas14}]\label{thm:d optimality}
+\begin{theorem}[Alexanderian, Gloor, Ghattas \cite{AlexanderianGloorGhattas14}]\label{thm:d_optimality}
   Let \(\pr = \normal(\prmean,\prcov)\) be a Gaussian prior on \(\hilp\)
@@
 
-\begin{definition}\label{def:d optimality}
+\begin{definition}\label{def:d_optimality}
   We say \(\opt\) is \emph{D-optimal} if \(\opt =
@@
 
-Intuition for Theorem \ref{thm:d optimality} can be gained by
+Intuition for Theorem \ref{thm:d_optimality} can be gained by
 recalling from Section \ref{subsec:D} that for a Bayesian linear model
@@
 matrix. Theorem
-\ref{thm:d optimality} and Definition \ref{def:d optimality} carry a
+\ref{thm:d_optimality} and Definition \ref{def:d_optimality} carry a
 similar intuition:
@@
 
- \section{The Constrained Optimization Problem of D-Optimal Design}\label{section:D and grad}
+ \section{The Constrained Optimization Problem of D-Optimal Design}\label{section:D_and_grad}
 We seek a formulation of the D-optimal design problem via Lagrange
@@
 
-\begin{proposition}\label{prop:tar grad}
+\begin{proposition}\label{prop:tar_grad}
   The gradient of the D-optimality objective $\tar$ is
@@
 
-\begin{proposition}\label{prop:bigger better}
+\begin{proposition}\label{prop:bigger_better}
   Let $\obs = (\meas_1,\dots,\meas_m)^t$, $j \in \{1,\dots,m\}$,
@@
   of $V$. Denote $\tmp: = \fwd \postcov \fwd^*$. From Proposition
-  \ref{prop:tar grad}:
+  \ref{prop:tar_grad}:
   \begin{align*}
@@
 
-Proposition \ref{prop:bigger better} implies that it is a good idea to
+Proposition \ref{prop:bigger_better} implies that it is a good idea to
 bound the norm of measurements. If, for example, we can take
 measurements in $\textup{span}\{\meas\}$ for some $\meas \neq 0$, then
-the the D-optimality criterion is unbounded. In contrast, in any
-real-life problem where sensors are concerned, the norm of
-measurements recorded by sensors is always one\footnote{Again, our
-analysis does not apply to point evaluations. These are just used for
-motivation and can be approximated by elements in $\hilo^*$ as long as
-$\hilo$ is a function space, e.g.~$L^2(\Omega)$.}:
+the D-optimality criterion is unbounded. In contrast, in any real-life
+problem where sensors are concerned, the norm of measurements recorded
+by sensors is always one\footnote{Again, our analysis does not apply
+to point evaluations. These are just used for motivation and can be
+approximated by elements in $\hilo^*$ as long as $\hilo$ is a function
+space, e.g.~$L^2(\Omega)$.}:
 
@@
 equality constraints (one for each measurement) on $\obs$. We define
-them and find their gradients in Proposition \ref{prop:constraints
-  grad} below, whose proof is straightforward and delegated to the
-Supplementary:
-
-\begin{proposition}\label{prop:constraints grad}
+them and find their gradients in Proposition
+\ref{prop:constraints_grad} below, whose proof is straightforward and
+delegated to the Supplementary:
+
+\begin{proposition}\label{prop:constraints_grad}
   Let
@@
   &\nabla \tar(\obs) = \sum_{j=1}^m \xi_j \nabla \phi_j (\obs)
-  \label{eq:Lagrange mult1} \\
-    &\phi_j(\obs) = 0, j = 1,\dots,m. \label{eq:Lagrange mult2}
+  \label{eq:Lagrange_mult1} \\
+    &\phi_j(\obs) = 0, j = 1,\dots,m. \label{eq:Lagrange_mult2}
 \end{align}
@@
 We now substitute the gradients calculated in Propositions
-\ref{prop:tar grad} and \ref{prop:constraints grad} into
-\eqref{eq:Lagrange mult1}:
+\ref{prop:tar_grad} and \ref{prop:constraints_grad} into
+\eqref{eq:Lagrange_mult1}:
 \begin{equation}\label{eq:constrained}
@@
 Letting $\Xi := \diag(\xi_j)$, we can write \eqref{eq:constrained} and
-\eqref{eq:Lagrange mult2} more compactly as:
+\eqref{eq:Lagrange_mult2} more compactly as:
 
@@
 
- \section{Answer to Question \ref{q:mitigate}: Model error mitigates clusterization}\label{section:non vanishing}
+ \section{Answer to Question \ref{q:mitigate}: Model error mitigates clusterization}\label{section:non_vanishing}
 We now show that if $\modcov \neq 0$ clusterization will not occur. It
@@
 
-\begin{proposition}[Increase due to a measurement]\label{prop:design increase}
+\begin{proposition}[Increase due to a measurement]\label{prop:design_increase}
   Let $\obs = (\meas_1,\dots,\meas_m)^t$ and $\obsm :=
@@
 
-\begin{corollary}\label{cor:same meas}
+\begin{corollary}\label{cor:same_meas}
   If $\meas_m = \meas_j$ for some $1 \leq j \leq m-1$, then
@@
 
-\begin{proof} \label{cor:same meas proof}
+\begin{proof} \label{cor:same_meas_proof}
   Denote $A:= \obs \modcov \obs^*$ and $v_j$ the $j^{\text{th}}$
@@
   $(\obsm \modcov \obsm^*)_{ij} = \meas_i(\modcov \meas_j)$, as
-  explained in \eqref{eq:modcov explained}. We can now verify that
+  explained in \eqref{eq:modcov_explained}. We can now verify that
   \begin{equation}\label{eq:observation}
@@
 
-\begin{proposition}\label{prop:twice woodbury}
+\begin{proposition}\label{prop:twice_woodbury}
   Assume $\fwd \prcov \fwd^*$ is invertible. Then
@@
 
-As we mentioned in Section \ref{subsec:bayesian inverse problems},
-$\fwd \prcov \fwd^*$ is the prior covariance $\hilo$ and could be
-safely assumed invertible. The proof of Proposition \ref{prop:twice
-  woodbury} is delegated to the Supplementary. It amounts to using
-Woodbury's matrix identity twice. The standard proof for Woodbury's
-matrix identity works for separable Hilbert spaces, as long as all
-terms are well defined. Unfortunately, $\obs^*\obs$ is not invertible,
-so we utilize a regularization trick to force it to be.
-
-\begin{lemma}[Simultaneous diagonizability]\label{lemma:sim diag}
+As we mentioned in Section \ref{subsec:bayesian_inverse_problems},
+$\fwd \prcov \fwd^*$ is the prior covariance in $\hilo$ and could be
+safely assumed invertible. The proof of Proposition
+\ref{prop:twice_woodbury} is delegated to the Supplementary. It
+amounts to using Woodbury's matrix identity twice. The standard proof
+for Woodbury's matrix identity works for separable Hilbert spaces, as
+long as all terms are well defined. Unfortunately, $\obs^*\obs$ is not
+invertible, so we utilize a regularization trick to force it to be.
+
+\begin{lemma}[Simultaneous diagonizability]\label{lemma:sim_diag}
   Let $\hil$ separable Hilbert space, $C:\hil \to \hil$ self-adjoint
@@
 \end{lemma}
-The proof of Lemma \ref{lemma:sim diag} is delegated to the
+The proof of Lemma \ref{lemma:sim_diag} is delegated to the
 Supplementary.
 
-\begin{proposition}\label{prop:same ev}
+\begin{proposition}\label{prop:same_ev}
   Let $\obs$ satisfy the nonlinear eigenvalue problem
@@
 \begin{proof}
-  \begin{align}\label{eq:mod conditions}
+  \begin{align}\label{eq:mod_conditions}
     \begin{split}
@@
 &= \sigma^{-2} \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* \obs^*  \text{ (by \eqref{eq:postcov})} \\
-&= \sigma^{-2} \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1} \obs^* \text{ (by Proposition \ref{prop:twice woodbury})}.
+&= \sigma^{-2} \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1} \obs^* \text{ (by Proposition \ref{prop:twice_woodbury})}.
     \end{split}
@@
   Now take $\func_j^{*} = \meas_j$ and $C := (\fwd \prcov
-  \fwd^*)^{-1}$ and use Lemma \ref{lemma:sim diag}.
+  \fwd^*)^{-1}$ and use Lemma \ref{lemma:sim_diag}.
 \end{proof}
@@
 
-\begin{proposition}\label{prop:true target}
+\begin{proposition}\label{prop:true_target}
   Let $\obs$ with $m$ measurements satisfy the nonlinear eigenvalue
@@
       Theorem)}\\
-&=\frac12 \log \prod_{i=1}^{\infty} ( 1 + \sigma^{-2} \lambda_i\eta_i ) \text{ (Proposition \ref{prop:same ev})} \\
+&=\frac12 \log \prod_{i=1}^{\infty} ( 1 + \sigma^{-2} \lambda_i\eta_i ) \text{ (Proposition \ref{prop:same_ev})} \\
 &=\frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i). 
@@
   satisfies the unit norm constraint on measurements. Part (2)
-  was proved in Proposition \ref{prop:same ev}. Part (3) was proved in
-  Proposition \ref{prop:true target}.
-
-  Part (4) is a consequence of Propositions \ref{prop:true target} and
+  was proved in Proposition \ref{prop:same_ev}. Part (3) was proved in
+  Proposition \ref{prop:true_target}.
+
+  Part (4) is a consequence of Propositions \ref{prop:true_target} and
   \ref{prop:kkt}, with the caveat that we did not show that finding
@@
   first statement in part (5) now follows from Proposition
-  \ref{prop:twice woodbury}, while the second statement follows from
+  \ref{prop:twice_woodbury}, while the second statement follows from
   parts (1) and (4).
@@
 
-\begin{figure}\label{fig:tikz clusterization}
+\begin{figure}\label{fig:tikz_clusterization}
   \centering
@@
     lowest. The sub-optimal design (right) does not.}
-  \label{fig:optimal vs not}
+  \label{fig:optimal_vs_not}
 \end{figure}
@@
 is the average of the total precision $\sum_{j=1}^k \lambda_j^{-1} +
-\sigma^{-2}m$, see eq.~\eqref{eq:cylinders} and Fig.~\ref{fig:optimal
-  vs not} for an illustration.
+\sigma^{-2}m$, see eq.~\eqref{eq:cylinders} and
+Fig.~\ref{fig:optimal_vs_not} for an illustration.
 
@@
 
-\subsection{Simulating Lemma \ref{lemma:free}}\label{subsec:lemma sims}
+\subsection{Simulating Lemma \ref{lemma:free}}\label{subsec:lemma_sims}
 In the proof of Theorem \ref{thm:char} we utilize Lemma
@@
 we ran, for each pair $m,k$. Clusteriztion occured at high rates
-($>95\%$) whenever $m-k > 1$; see Fig.~\ref{fig:sim AAt}. Hence, in
+($>95\%$) whenever $m-k > 1$; see Fig.~\ref{fig:sim_AAt}. Hence, in
 these simulations, clusterization is a generic property. However, when
@@
       occur.}
-  \label{fig:sim AAt}
+  \label{fig:sim_AAt}
 \end{figure}
@@
 
-\subsection{Correlated errors}\label{subsec:corr errors sims}
-In order to verify the results of Section \ref{section:non vanishing},
+\subsection{Correlated errors}\label{subsec:corr_errors_sims}
+In order to verify the results of Section \ref{section:non_vanishing},
 we run simulations of the inverse problem of the 1D heat equation with
 nonvanishing model error \(\modcov = \prcov^2 \). Indeed, including
-model correlation pushes measurements apart, see Fig.~\ref{fig:corr
-  errors}. Code generating Fig.~\ref{fig:corr errors} is located in
-module \texttt{clusterization.py} in the accompanying
+model correlation pushes measurements apart, see
+Fig.~\ref{fig:corr_errors}. Code generating Fig.~\ref{fig:corr_errors}
+is located in module \texttt{clusterization.py} in the accompanying
 \href{https://github.com/yairdaon/OED}{repository}.
@@
       term.}
-  \label{fig:corr errors}
+  \label{fig:corr_errors}
 \end{figure}

### 2024-07-31 latex/how.tex
before diff
Intro ready

@@
 $\{\ev_i\}_{i=1}^\infty$ and $\sigma^2=1$. Consider the following
-$\obs, \pobs$ (see also Figure \ref{fig:clusterization}):
-\begin{equation}\label{eq:two designs}
+$\obs, \pobs$ (see also Fig.~\ref{fig:clusterization}):
+\begin{equation}\label{eq:two_designs}
   \obs =
@@
   \end{tikzpicture}
-  \caption{Clusterization (left, $\obs$ in \eqref{eq:two designs})
-    and non-clusterization (right, $\pobs$ in \eqref{eq:two designs})
+  \caption{Clusterization (left, $\obs$ in \eqref{eq:two_designs})
+    and non-clusterization (right, $\pobs$ in \eqref{eq:two_designs})
     in D-optimal designs. Posterior precision per eigenvector is
     plotted for $\obs$ and $\pobs$ from \eqref{eq:two
-      designs}. Both designs are identical for all practical matters
+______designs}. Both designs are identical for all practical matters
     --- their posteriors are equal. In the left panel, a D-optimal

### 2024-07-31 latex/intro.tex
Tidy up a bit
Intro ready

--- 2024-07-24 latex/intro.tex
@@
 The practice of experimental design entails specifying every detail of
-running an experiment \cite{chaloner1995} and at its simplest form,
+running an experiment \cite{chaloner1995}. At its simplest form,
 experimental design involves choosing a set of measurements to be
 taken. For example, consider the process of searching for oil:
-measuremnts require digging holes in the ground and a \emph{design}
+measurements require digging holes in the ground and a \emph{design}
 simply indicates which holes should be dug
 \cite{horesh2008borehole}. Since measurements are expensive, only a
-limited number of measuremnts can be taken, and thus carefully
-choosing measuremt locations is crucial.
+limited number of them can be taken, and thus carefully choosing
+measurement locations is crucial.
 
-Specifiyng the right set of measurements holds particular significance
+Specifying the right set of measurements holds particular significance
 when solving an \emph{inverse problem} --- i.e.~making inference of
@@
 materials, such as density, elasticity, and the presence of fluids or
-voids. Combining travel time data with a geophisical models of
+voids. Combining travel time data with a geophysical models of
 contents of the earth's crust facilitates reconstructing the structure
@@
 this utility. Two of the most widely recognized and extensively
-studied design criteria are A- and D-optimality.
+studied design criteria are the A- and \emph{D-optimality} criteria.
 
@@
 meaning: measurements chosen according to the D-optimality criterion
-maximize the expected Kullback-Liebler (KL) divergence from posterior
+maximize the expected Kullback-Leibler (KL) divergence from posterior
 to prior \cite{chaloner1995, AlexanderianGloorGhattas14,
@@
 \begin{equation*}
-  \mathbb{E}_{\data}\left [ D_{KL}\left (\Pr(\param|\data)||\Pr(\param)\right ) \right ]
+  \mathbb{E}_{\data}\left [ D_{KL}\left (\Pr(\param|\data)||\Pr(\param)\right ) \right ].
 \end{equation*}
-We refer to a set of measuremnts that maximizes the D-optimality
+We refer to a set of measurements that maximizes the D-optimality
 criterion as a \emph{D-optimal design}.
@@
 described by the following three equations:
-\begin{subequations}%\label{eq:heat equation}
+\begin{subequations}
   \begin{alignat}{2}
@@
 \end{subequations}
-As time passes, heat disspates across the rod as hotter regions,
-e.g.~local maximas of the heat distribution, become cooler. This
+As time passes, heat dissipates across the rod as hotter regions,
+e.g.~local maxima of the heat distribution, become cooler. This
 behavior is captured by eq.~\eqref{eq:heat1}: the Laplacian at local
 maxima is negative, so $u_t = \Delta u$ implies $u$ decreases at its
-maximas, and the reverse happens at local minima. Eq.~\eqref{eq:heat2}
+maxima, and the reverse happens at local minima. Eq.~\eqref{eq:heat2}
 describes the \emph{boundary condition}, which dictates how heat
@@
 since heat spreads in a diffusive manner and any roughness in the
-initial condition is quickly smoothed, see Supplementary movies S1 and
-S2 (code generating these movies is located in module
+initial condition is quickly smoothed, as implied by the squared
+exponential decay of eigenvalues in eq.~\eqref{eq:decay} with
+$n\to\infty$. See Supplementary movies S1 and S2 for an illustration
+of this phenomenon (code generating these movies is located in module
 \texttt{movies.py} in the accompanying
@@
 regularization via a fully Bayesian formulation of the inverse
-problem, see Section \ref{section:prelims} for details.
+problem.
+
+We choose a prior for the initial condition $u_0 \sim \mathcal{N}(0,
+(-\Delta)^{-1})$, with homogeneous Dirichlet boundary condition. This
+choice of prior covariance operator ensures the posterior is
+well-defined and prior realizations are well-behaved, see
+\cite[Theorem 3.1 and Lemma 6.25]{Stuart10} for details. The second
+ingredient of a fully Bayesian formulation of the inverse problem is
+likelihood. We simply assume centered iid Gaussian measurement error
+model, giving rise to a the widely applicable Gaussian likelihood
+function.
 
@@
 
-%% first: blue, second: red, third: green, fourth: black,
-%% fifth:magenta
 \begin{figure}
     \centering
-    \includegraphics[height=0.5\textwidth]{dst_modelError0.png}
-    \caption{Measurement clusterization for D-optimal designs for the
-      inverse problem of the heat equation. Measurement locations were
-      chosen according to the Bayesian D-optimality criterion of
-      Theorem \ref{thm:d optimality}. Measurement locations are
+    \includegraphics[height=0.5\textwidth]{figs/dst_modelError0.png}
+    \caption{Measurement clusterization in D-optimal designs for the
+      inverse problem of the 1D heat equation. Measurement locations
+      were chosen according to the Bayesian D-optimality criterion of
+      Theorem \ref{thm:d_optimality}. Measurement locations are
       plotted over the computational domain \(\Omega = [0, 1]\)
@@
 \(\Omega \subseteq \mathbb{R}^d, d=1,2,3\), many researchers
-circumvent the problem of measurement clusterization by choosing
-measurements from a coarse grid in \(\Omega\) \cite{koval2020,
+circumvent the problem of measurement clusterization by restricting
+measurements to a coarse grid in \(\Omega\) \cite{koval2020,
   alexanderian2021, attia2022, alexanderian2014, alexanderian2016,
@@
 %equally optimal non-clustered design?
-%Can an optimal clustered design be relpaced with an equally optimal
+%Can an optimal clustered design be replaced with an equally optimal
 %non-clustered design?
@@
 spaces and study two inverse problems: (a) In Sections
-\ref{section:prelim} and \ref{section:D and grad} we propose a novel
-relaxed model for an inverse problem where D-optimality maintains
-analytical tractability, where D-optimal designs are identified via
+\ref{section:prelim} and \ref{section:D_and_grad} we propose a novel
+generic model for an inverse problem where D-optimality maintains
+analytical tractability and D-optimal designs are identified via
 Lagrange multipliers. This analytical framework facilitates the
@@
   theoretical perspective, we show that clusterization mainly depends
-  on how quickly the eigenvalues of the prior measure in observation
-  space decay. This decay depends mostly on how ill-posed the problem
-  at hand is and does not depend much on the prior. See Section
-  \ref{section:vanishing} --- particularly, Theorem \ref{thm:char} and
-  the discussion following it. We also show results of numerical
-  experiments, where simulations of our model give rise to D-optimal
-  designs that exhibit clusterization with high probability in Section
-  \ref{subsec:lemma sims}. Thus, given the genericity of our model, we
-  expect measurement clusterization to be a generic and ubiquitous
-  phenomenon.
+  on how quickly the eigenvalues of the prior covariance in
+  observation space decay. This decay depends mostly on how ill-posed
+  the problem at hand is and does not depend much on the prior. See
+  Section \ref{section:vanishing} --- particularly, Theorem
+  \ref{thm:char} and the discussion following it. We also show results
+  of numerical experiments, where simulations of our model give rise
+  to D-optimal designs that exhibit clusterization with high
+  probability in Section \ref{subsec:lemma_sims}. Thus, given the
+  genericity of our model, we expect measurement clusterization to be
+  a generic and ubiquitous phenomenon.
 
@@
   measurements result in no gain in design criterion when observation
-  error tends to zero. Moreover, in Section \ref{subsec:corr errors
-    sims}, we show that an error term corresponding to correlations
-  between measurements mitigates clusterization in the inverse problem
-  of the 1D heat equation.
+  error amplitude tends to zero. Moreover, in Section
+  \ref{subsec:corr_errors_sims}, we show that an error term
+  corresponding to correlations between measurements mitigates
+  clusterization in the inverse problem of the 1D heat equation.
 
 \item \label{q:why} \textbf{Why does measurement clusterization
-  occur?} In Section \ref{subsec:why} (specifically, See
-  Fig~\ref{fig:why}) we give a compelling answer to this question by
-  transporting insights we gain from our analytically tractable
-  relaxed model to the inverse problem of the 1D heat equation. Our
-  analysis reveals that a D-optimal design focuses on a select set of
-  prior eigenvectors, specifically those with the largest eigenvalues
-  in the prior covariance spectrum. In practical scenarios, the number
-  of locations where (a) all relevant prior eigenvectors are
+  occur?} In Section \ref{subsec:why} (specifically, Fig~\ref{fig:why}
+  and the discussion following it) we give a compelling answer to this
+  question by transporting insights we gain from our generic model to
+  the inverse problem of the 1D heat equation. Our analysis reveals
+  that a D-optimal design focuses on a select set of prior
+  eigenvectors, specifically those with the largest eigenvalues in the
+  prior covariance spectrum. In practical scenarios, the number of
+  locations where (a) all relevant prior eigenvectors are
   significantly large, and (b) other eigenvectors are close to zero,
@@
 clusterization should be expected in many different problems across
-many different scientific fields. Thus, researchers that encounter
+many different scientific fields. Researchers that encounter
 clusterization should not be surprised or wary. In Our answer to
 Question \ref{q:why}, we explain what our view of the cause of
-clusterization. As it appears, the cause is generic (see Question
-\ref{q:why} above): a D-optimal design reduces uncertainty for a
-select set of prior covariance eigenvectors --- those with the most
-prior uncertainty, i.e.~those the practitioner cares about the most!
-Consequently, practitioners should not try to avoid measurement
-clusterization. Rather, practitioners should take repeated
-measurements (e.g.~in MRI and borehole tomography), increase apparatus
-sensitivity (e.g.~in EIT), or take consecutive measurements (e.g.~in
-the 1D heat equation). Overall, we show that measurement
-clusterization is a natural and (almost) inevitable part of Bayesian
-D-optimal designs.
+clusterization is. It appears, the cause is generic: a D-optimal
+design reduces uncertainty for a select set of prior covariance
+eigenvectors --- those with the most prior uncertainty, i.e.~those the
+practitioner cares about the most! We believe practitioners should not
+try to avoid measurement clusterization. Rather, practitioners should
+take repeated measurements (e.g.~in MRI and borehole tomography),
+increase apparatus sensitivity (e.g.~in EIT), or take consecutive
+measurements (e.g.~in the 1D heat equation). Overall, we show that
+measurement clusterization is a natural and (almost) inevitable part
+of Bayesian D-optimal designs (but see disclaimer below).
 
 One interesting implication of the analysis presented here is that
-clusterization can be expected once the number of measurements exceeds
-the number of relevant eigenvectors. Since leading eigenvectors
-typically correspond to slow variations in space and/or time,
-clusterization could be used to estimate the number of relevant
-degrees of freedom, and even to reduce the complexity of a
-computational model, e.g.~by dropping discretization points.
+clusterization can serve as an evidence to the number of relevant
+eigenvectors. Since leading eigenvectors typically correspond to slow
+variations in space and/or time, clusterization could be used to
+estimate the number of relevant degrees of freedom, and even to reduce
+the complexity of a computational model, e.g.~by dropping
+discretization points.
 
@@
 based on the fact that it results in clustered designs. We have seen,
-however that there is perfectly reasonable explanation for
-clusterization, i.e.~that clusterization arises when we seek a
-designWe have shown that clusterization is an innevitable conseqeunce
-of having a problem with modes of higher uncertainty than
-others. Clusterization manifests the true underlying objective
+however that there is a perfectly reasonable explanation for
+clusterization. We have shown that clusterization is an inevitable
+consequence of having a problem with some modes where uncertainty
+decays faster than others.
 
@@
 equation, clusterization occurs primarily because Laplacian
-eigenvectors are nonzero almost everywhere in $\Omega$. Consequently,
-measuring $u(x_1, T)$ at some point $x_1 \in \Omega$ provides
-information about $u(x_2,T)$ for distant points $x_2 \in
-\Omega$. Intuitively, this should not occur: for small $T$, the heat
-distribution at $x_1$ should have minimal influence on the heat
-distribution at $x_2$.
+eigenvectors \emph{do not} decay in $\Omega$. Consequently, measuring
+$u(x_1, T)$ at some point $x_1 \in \Omega$ provides information about
+$u(x_2,T)$ for distant points $x_2 \in \Omega$. Intuitively, this
+should not occur: for small $T$, the heat distribution at $x_1$ should
+give very little knowledge on the heat distribution at $x_2$. This
+behavior stems from a well-known property of the heat equation: it
+allows information to spread \emph{instantly} across the computational
+domain \cite{renardy2006PDE}. In reality, heat (and information)
+propagate at finite speeds. Of course, the known physical barrier for
+information spread is the speed of light, but we expect heat to spread
+considerably slower: heating an Olympic pool at one end should have no
+immediate effect on the temperature at the other end.
 
-This behavior stems from a well-known property of the heat equation:
-it allows information to spread instantly across the computational
-domain \cite{renardy2006PDE}. In reality, heat and information
-propagate at finite speeds. The emergence of clusterization in this
-context is thus non-physical, arising from the nature of Laplacian
-eigenvectors being nonzero almost everywhere. Clusterization therefore
-indicates that the underlying mathematical model is overly permissive
-and fails to capture crucial physical constraints. We suggest that
+We believe that the emergence of clusterization in this context is
+thus non-physical, arising from the way the inverse problems we
+consider are phrased. Clusterization therefore indicates that the
+underlying mathematical model is overly permissive and fails to
+capture crucial physical constraints of the problem. We suggest that
 when clusterization occurs, practitioners should consider alternative
 models where information is localized in space and travels at finite
-speeds. Such models may not only provide more physically accurate and
-meaningful results but may also mitigate the issue of clusterization.
+speed in the medium. Such models may not only provide more physically
+accurate and meaningful results but may also mitigate the issue of
+clusterization.
 
@@
 In Theorem \ref{thm:char}, we also show that D-optimal designs are
-best understood in the sapce of \emph{observations}, see Section
+best understood in the space of \emph{observations}, see Section
 \ref{section:vanishing} for a precise statement. This is in accordance

### 2024-07-31 latex/numerix.tex
Tidy up a bit
Intro ready

@@
       term.}
-  \label{fig:corr errors}
+  \label{fig:corr_errors}
 \end{figure}
+
+%% \subsection{Implementing the 1D heat equation}
+%% The Laplacian $\Delta$ is a Fourier multiplier \cite{stein2009}, so
+%% our implementation proceeds in frequency space. When we utilize the
+%% homogeneuous Dirichlet boundary condition, the eigenvectors of
+%% $\Delta$ are sines $\ev_n(x) = \sin(\pi n x), n=1,2,3,\dots$, whereas
+%% for the homogenoeous Neumann boundary condition the eigenvectors are
+%% cosines $\cos(\pi n x)$. Implementing point evaluations in Fourier
+%% space

### 2024-07-31 latex/old.tex
Intro ready

@@
   nyberg2012, fedorov1997, Ucinski05, neitzel2019sparse}. To
-illustrate this point, we consider a toy problem: inferring
-the initial condition of the 1D heat equation (details in the
-supplementary material). In Fig.~\ref{fig:clusterization
-  illustration} D-optimal measurement locations are shown for
-different numbers of measurements. Notably, for six measurements, a
-D-optimal design yields two sets of measurements that are almost
-indistinguishable from one another. We refer to this intriguing
-phenomenon as \emph{measurement clusterization} \cite{Ucinski05}, and
-we refer to a design that exhibits measurement clusterization as a
-\emph{clustered design}.
+illustrate this point, we consider a toy problem: inferring the
+initial condition of the 1D heat equation (details in the
+supplementary material). In Fig.~\ref{fig:clusterization_illustration}
+D-optimal measurement locations are shown for different numbers of
+measurements. Notably, for six measurements, a D-optimal design yields
+two sets of measurements that are almost indistinguishable from one
+another. We refer to this intriguing phenomenon as \emph{measurement
+clusterization} \cite{Ucinski05}, and we refer to a design that
+exhibits measurement clusterization as a \emph{clustered design}.
 
@@
 
-\subsection{Bayesian Linear Inverse Problems}\label{subsec:abstract OED}
+\subsection{Bayesian Linear Inverse Problems}\label{subsec:abstract_OED}
 Let \(\hilp\) and \(\hilo\) separable Hilbert spaces (the subscripts p
@@
 are assumed independent of each other. Thus, data is acquired via
-\begin{align}\label{eq:inverse problem}
+\begin{align}\label{eq:inverse_problem}
   \data := \obs (\fwd \param + \eps') + \eps = \obs \fwd \param + \obs \eps' + \eps.
@@
 
-\subsection{Bayesian D-Optimal Designs in Infinite Dimensions}\label{subsec:D optimal design} 
+\subsection{Bayesian D-Optimal Designs in Infinite Dimensions}\label{subsec:D_optimal_design} 
 A Bayesian D-optimal design maximizes expected Kullback-Leibler (KL)
@@
 %% trace-class.
-\begin{definition}\label{def:d optimality}
+\begin{definition}\label{def:d_optimality}
   We say \(\obs^{\star}\) is \emph{D-optimal} if \(\obs^{\star} =
@@
 
-Intuition for Theorem \ref{thm:d optimality} can be gained by
+Intuition for Theorem \ref{thm:d_optimality} can be gained by
 considering a Bayesian linear model in finite dimensions, with
@@
 to be a regularized version of the frequentist D-optimality criterion
-\cite{Chaloner1995}. Theorem \ref{thm:d optimality} and Definition
-\ref{def:d optimality} carry a similar intuition:
+\cite{Chaloner1995}. Theorem \ref{thm:d_optimality} and Definition
+\ref{def:d_optimality} carry a similar intuition:
 \begin{align*}
@@
 
-%% \subsection{Sequential vs Simultaneous Optimization}\label{subsec:seq vs sim}
-%% From defintion \ref{def:d optimality} we wish to characterize solution(s) of the
+%% \subsection{Sequential vs Simultaneous Optimization}\label{subsec:seq_vs_sim}
+%% From defintion \ref{def:d_optimality} we wish to characterize solution(s) of the
 %% following optimization problem for \(\tar$. %%: (\hilo^*)^m \to \R$:
@@
 %% observations in a sequential manner. Denote
-%% \begin{equation}\label{eq:def obs_k}
+%% \begin{equation}\label{eq:def_obs_k}
 %%   \obs_k := (\meas_1,\dots, \meas_k)^t,  k\leq m.
@@
 %% optimization case. The sequential optimization case is dealt with in
-%% section \ref{subsec:clusterization sequential}. It is important to
+%% section \ref{subsec:clusterization_sequential}. It is important to
 %% note, however that all conclusions we arrive at for the simultaneous
@@
 
-\section{The Constrained Optimization Problem of D-Optimal Design}\label{section:D and grad}
+\section{The Constrained Optimization Problem of D-Optimal Design}\label{section:D_and_grad}
 We seek a formulation of the D-optimal design problem via Lagrange
@@
 
-  \begin{align}\label{eq:der sig}
+  \begin{align}\label{eq:der_sig}
     \begin{split}
@@
 
-  Then, using \eqref{eq:der sig}: 
+  Then, using \eqref{eq:der_sig}: 
   \begin{align*}
@@
     \Sigma^{-1} (V\modcov \obs^* + \obs \modcov V^*). 
-    %\text{, by \eqref{eq:der sig}. }
+    %\text{, by \eqref{eq:der_sig}. }
   \end{align*}
@@
   Thus:
-  \begin{align}\label{eq:der sig inv}
+  \begin{align}\label{eq:der_sig_inv}
     \frac{\der \Sigma(\obs+\tau V)^{-1}}{\der \tau} \Big |_{\tau=0}  
@@
 
-  By Leibniz (product) rule and \eqref{eq:der sig inv}:
+  By Leibniz (product) rule and \eqref{eq:der_sig_inv}:
   % 
@@
       &= \frac12 \frac{\der}{\der \tau} \Big |_{\tau=0} \log \det 
-      (I + \prcov^{1/2} \fwd^* T(\obs+\tau V)\fwd \prcov^{1/2} ) \text{ (Theorem \ref{thm:d optimality})} \\
+      (I + \prcov^{1/2} \fwd^* T(\obs+\tau V)\fwd \prcov^{1/2} ) \text{ (Theorem \ref{thm:d_optimality})} \\
       % 
@@
 
-Proposition \ref{prop:bigger better} implies that it is a good idea to
+Proposition \ref{prop:bigger_better} implies that it is a good idea to
 bound the norm of measurements. If, for example, we can take
@@
 equality constraints (one for each measurement) on $\obs$. We define
-them and find their gradients in Proposition \ref{prop:constraints
-  grad} below:
-
-\begin{proposition}\label{prop:constraints grad}
+them and find their gradients in Proposition
+\ref{prop:constraints_grad} below:
+
+\begin{proposition}\label{prop:constraints_grad}
   Let
@@
 
-%% The same arguments justifying \eqref{eq:tar grad} hold here, and thus:
-
-%% \begin{align}\label{eq:grad constraints}
+%% The same arguments justifying \eqref{eq:tar_grad} hold here, and thus:
+
+%% \begin{align}\label{eq:grad_constraints}
 %% \nabla \phi_j(\obs) = \obs^* e_j e_j^t = \meas_j e_j^t , j=1,\dots,m,
@@
 %% Imposing correlations between observations alleviate measurement clusterization
-\section{Answer to Question \ref{q:mitigate}: Model error mitigates clusterization}\label{section:non vanishing}
+\section{Answer to Question \ref{q:mitigate}: Model error mitigates clusterization}\label{section:non_vanishing}
 We now show that if $\modcov \neq 0$ clusterization will not occur. It
@@
   %
-  \begin{align}\label{eq:M def}
+  \begin{align}\label{eq:M_def}
     \M (\obs ):&= \prcov^{\frac12}\fwd^* \obs^* \Sigma^{-1} \obs \fwd
@@
 
-  From \eqref{eq:schur} and \eqref{eq:M def}:
+  From \eqref{eq:schur} and \eqref{eq:M_def}:
   \begin{align*}
@@
   %
-  \begin{equation}\label{eq:M plus I}
+  \begin{equation}\label{eq:M_plus_I}
     I + \M( \obs ) = I + \M (\obsm ) + uu^*.
@@
   Note that
-  \begin{equation}\label{eq:M postcov}
+  \begin{equation}\label{eq:M_postcov}
     \prcov^{1/2} \left (I + \M( \obsm ) \right )^{-1} \prcov^{1/2} = \postcovm.
@@
   \end{align}
-  From \eqref{eq:u} and \eqref{eq:M postcov}:
+  From \eqref{eq:u} and \eqref{eq:M_postcov}:
   \begin{align}\label{eq:final}
@@
 to \eqref{eq:eigenproblem} as a \emph{nonlinear} eigenvalue problem.
-%% Proposition \ref{prop:twice woodbury} and Lemma \ref{lemma:sim
-%% diag} required
-
-\begin{proposition}\label{prop:twice woodbury}
+%% Proposition \ref{prop:twice_woodbury} and Lemma \ref{lemma:sim
+%%_diag} required
+
+\begin{proposition}\label{prop:twice_woodbury}
   Assume $\fwd \prcov \fwd^*$ is invertible. Then
@@
 
-\begin{lemma}[Simultaneous diagonizability]\label{lemma:sim diag}
+\begin{lemma}[Simultaneous diagonizability]\label{lemma:sim_diag}
   Let $\hil$ separable Hilbert space, $C:\hil \to \hil$ self-adjoint
@@
   eigenvalue are orthogonal. For $k \in S_i$
-  \begin{equation}\label{eq:on vi}
+  \begin{equation}\label{eq:on_vi}
     \xi_i \func_k = (C + \sum_{j=1}^m \func_j \func_j^* ) \func_k = (C + A_i) \func_k.
@@
   Let $V_i := span \{\func_k \}_{k\in S_i}$. Observe that $V_i$ is
-  invariant under $A_i$, by definition, and under $C$, by \eqref{eq:on
-    vi}. A second application of \eqref{eq:on vi} shows that $A_i =
-  \xi_iI - C$ on $V_i$. This immediately implies $A_i$ and $C$ are
-  simultaneously diagonalizable on $V_i$. This holds for every $1 \leq
-  i \leq \ell$ and we conclude that $C$ and $A$ are simultaneously
-  diagonalizable.
+  invariant under $A_i$, by definition, and under $C$, by
+  \eqref{eq:on_vi}. A second application of \eqref{eq:on_vi} shows
+  that $A_i = \xi_iI - C$ on $V_i$. This immediately implies $A_i$ and
+  $C$ are simultaneously diagonalizable on $V_i$. This holds for every
+  $1 \leq i \leq \ell$ and we conclude that $C$ and $A$ are
+  simultaneously diagonalizable.
 \end{proof}
 
-\begin{proposition}\label{prop:same ev}
+\begin{proposition}\label{prop:same_ev}
   Let $\obs$ satisfy the nonlinear eigenvalue problem
@@
       %
-      &= \sigma^{-2} \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1} \obs^* \text{ (by Proposition \ref{prop:twice woodbury})}.
+      &= \sigma^{-2} \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1} \obs^* \text{ (by Proposition \ref{prop:twice_woodbury})}.
     \end{split}
@@
       %
-      &=\frac12 \log \prod_{i=1}^{\infty} ( 1 + \sigma^{-2} \lambda_i\eta_i ) \text{ (Proposition \ref{prop:same ev})} \\
+      &=\frac12 \log \prod_{i=1}^{\infty} ( 1 + \sigma^{-2} \lambda_i\eta_i ) \text{ (Proposition \ref{prop:same_ev})} \\
       %
@@
   constraints are, for $j=1,\dots,m$:
-  \begin{equation}\label{eq:V constraints}
+  \begin{equation}\label{eq:V_constraints}
    1 = [A^tA]_{jj} = [V S^tS V^t]_{jj},
@@
 level of liquid in it, as long as its index $i \leq m$. The result of
-such a procedure is illustrated in Fig.~\ref{fig:optimal vs not}.
+such a procedure is illustrated in Fig.~\ref{fig:optimal_vs_not}.
 %% \begin{figure}%{r}{0.25\textwidth} 

### 2024-07-31 latex/prelims.tex
Intro ready

@@
 %% trace-class.
-\begin{definition}\label{def:d optimality}
+\begin{definition}\label{def:d_optimality}
   We say \(\opt\) is \emph{D-optimal} if \(\opt =

### 2024-07-31 latex/refs.tex
Intro ready

@@
 \input{definitions}
+\usepackage[colorlinks,citecolor=blue,urlcolor=blue,filecolor=blue,backref=page]{hyperref}
+
+\newcommand{\paste}[1]{\vskip 0.05in \textcolor{red}{#1}\vskip 0.05in}
+\newcommand{\answer}[1]{\vskip 0.05in \textcolor{blue}{#1} \vskip 0.05in}
 
 \begin{document}
+
+I appreciate the many thoughtful comments by all of you. I believe
+these comments helped me considerably improve the exposition of my
+paper and added new insights as well. I sincerely appreciate the
+option to revise and resubmit. Please see below my answers to the
+comments made by the referees, associate editor and editor in chief:
+\answer{My answers are marked in blue.} \paste{In some places I
+  copied text from the revised manuscipt and marked it in red.}
 
@@
 
-\answer: I appreciate the many thoughtful comments by yourself, the
-associate editor and the reviewers. I also appreciate the option to
-improve the paper. Please see my answers below in blue.
-
 \begin{description}
@@
     readership;
-    \answer: I have considerably extended the introduction.
+    \answer{I have considerably extended the introduction.}
 
-  \item[b] the motivation and the practical statistical implications are to be
-    developed and discussed;
-    \answer: I added a section on the implications in the introduction.
+  \item[b] the motivation and the practical statistical implications
+    are to be developed and discussed;
+
+    \answer{I added a section on the implications in the
+      introduction.}
 
@@
     designs;
-  \answer:
+  \answer{adfasd}
 
@@
 
-    \answer: I added a section describing numerical experiments and
+    \answer{I added a section describing numerical experiments and
     their results. I also added documentation to the accompanying
     \href{https://github.com/yairdaon/OED}{repository} and made sure
-    all scripts there run.
+    all scripts there run.}
 
@@
 
-    \answer: I am not sure which specific comment of Reviewer 3 this
-    refers to, please see my answers to Reviewer 3 below.
+    \answer{I am not sure which specific comment of Reviewer 3 this
+    refers to, please see my answers to Reviewer 3 below. }
 \end{description}
@@
 
-\answer: I have added considerably to the introduction. See revised
-manuscript and diff file.
+\answer{I have added considerably to the introduction. See revised
+manuscript and diff file.}
 
@@
 appealing feature in some settings, and another referee asks whether
-D-optimal designs should be abandoned altogether. The exposition needs
-to discuss these high-level issues, and introduce the related
+D-optimal designs should be abandoned altogether. The exposition
+needs to discuss these high-level issues, and introduce the related
 literature and results appropriately, to give the work proper
@@
 
-\answer: I appreciate the kind words.
+\answer{I appreciate the kind words.}
 
@@
 
-\answer: If I understand correctly --- I completely agree. I added
-text emphasizing that I do not endorse using correlated errors.
+\answer{If I understand correctly --- I completely agree. I added
+text emphasizing that I do not endorse using correlated errors:}
+\paste{asdf}
 
@@
 leads to a “suboptimal” design?
-\answer: Aagain, my apologies for the misunderstanding.
+\answer{Aagain, my apologies for the misunderstanding.}
 
@@
 
-\answer: Bla
+\answer{Bla }
 
@@
 
-\answer: I appreciate the comment, I fixed the relevant line in the
-manuscript.
+\answer{I appreciate the comment, I fixed the relevant line in the
+manuscript.}
 
@@
 
-\answer: I do not view clusterization as an undesired property, I only
+\answer{I do not view clusterization as an undesired property, I only
 mentioned many authors consider it to be undesirable. I have added a
 section contrasting repetition and replication with clusterization and
-why I believe they may be different.
+why I believe they may be different. }
 
@@
 
-\answer: I believe that even in an application where repeated
+\answer{I believe that even in an application where repeated
 experiments are allowed they may be unintuitive. For example, in a
@@
 regarding the optimality give a model, which, as you mention might be
-wrong.
+wrong.}
 
@@
 
-\answer: I believe clusterization may imply a model is somehow in
-adequate, see text below.
+\answer{I believe clusterization may imply a model is somehow
+  inadequate, see text below. }
 
@@
 
-\answer: I couldn't agree more. Indeed, the suboptimality of clustered
+\answer{I couldn't agree more. Indeed, the suboptimality of clustered
 design is a matter of personal opinion; once a design criterion and a
-model are agreed upon there can be no disagreement.
+model are agreed upon there can be no disagreement.}
 
@@
 
-  \answer: I do consider experiments where repeated measurements are
+  \answer{I do consider experiments where repeated measurements are
   possible. I am giving insight on why such designs are useful, even
   though repeated measurements may be non-intuitive in certain
-  settings.
+  settings.}
 
@@
 
-  \answer: great comment
+  \answer{great comment}
 \end{enumerate}
@@
 
-\answer: Thanks for this comment. I considerably expanded the
+\answer{Thanks for this comment. I considerably expanded the
 introduction and delegated many proofs to the Supplementary. I also
-added a section on practical implications, reproduced below.
+added a section on practical implications, reproduced below.}
 
@@
 
-\answer: To a certain extend I agree: indeed, the impossibility of
+\answer{To a certain extend I agree: indeed, the impossibility of
 taking identical measurements is indeed natural. However, this is not
@@
 suggest, even if it counterintuitive, or find ways to circumvent the
-mathematics because the result seems non-intuitive.
+mathematics because the result seems non-intuitive.}
 
@@
 observations” as a solution. However, isn’t this imposition pragmatic
-as well? I suppose it is argued that the $\obs \eps′$ term is
+as well? I suppose it is argued that the $\obs \eps'$ term is
 accounting for model error: but it is still pragmatic: the data
@@
 
-\answer: I do not support imposing correlations in the paper, I just
+\answer{I do not support imposing correlations in the paper, I just
 used correlations to show how my framework can be used to rigorously
@@
 paper endorses imposing correlations and I accept that this is my
-error in writing. I added a clause clearing the issue.
+error in writing. I added a clause clearing the issue.}
 
@@
 
-  \answer: I changed $\obs^\star$ to $\opt$.
+  \answer{I changed $\obs^\star$ to $\opt$.}
 
@@
 
-  \answer: I was not able to find a better term to replace
-  "measurements", so I stuck to it.
+  \answer{I was not able to find a better term to replace
+  "measurements", so I stuck to it.}
 
@@
 
-  \answer: Absolutely correct, I changed it to $\nu$. You will not
+  \answer{Absolutely correct, I changed it to $\nu$. You will not
   find the revised proof in the manuscript since I moved it to the
-  Supplementray.
+  Supplementray.}
 \end{itemize}
@@
 
-\answer: I apologize for not properly maintainig the repository. It is
-now functioning and well-documented.
+\answer{I apologize for not properly maintainig the repository. It is
+  now functioning and well-documented.}
 
@@
 
-\answer: I appreciate the suggestion. I added a section on numerical
-simulations to the main atricle. It is reproduced below.
+\answer{I appreciate the suggestion. I added a section on numerical
+simulations to the main atricle. It is reproduced below.}
 
@@
 
-  \answer: It is impossible to visually discern the measurements, even
+  \answer{It is impossible to visually discern the measurements, even
   when resolution is increased. These measurements are not identical
@@
   find these points via optimization over point location in
-  $\Omega=[0,1]$ I view an error $<10^{-5}$ as reasonable.
+  $\Omega=[0,1]$ I view an error $<10^{-5}$ as reasonable.}
 
@@
 
-  \answer: That is a great point!
+  \answer{That is a great point, but I view experiments in higher
+    dimension as out of scope for the current study.}
 
@@
   the prior covariance operator is set to be equal to the inverse
-  Laplace operator $(−\Delta)^{−1}$ . In d-dimensional domains, its
+  Laplace operator $(-\Delta)^{-1}$. In d-dimensional domains, its
   eigenvalues $\lambda_j$ are known to follow Weyl’s asymptotics,
-  growing as $j^{2/d}$, which should then impact the index after
-  which the eigenvalues are ‘thresholded’ by the procedure.
+  growing as $j^{2/d}$, which should then impact the index after which
+  the eigenvalues are ‘thresholded’ by the procedure.
+
+  \answer{Indeed, the growth of eigenvalues of the Laplacian is
+    different in higher dimensions. This is why in higher dimensions
+    we would have to take as prior \(u_0 \sim \mathcal{N}(\param_0,
+    (-\Delta)^{-\gamma})\), for some \(\gamma > d/2\) in order to
+    maintain regularity of prior realizations. I decided not to
+    include a discussion on the choice of priors in higher spatial
+    dimensions because it would require too much exposition and make
+    me digress from the main points I wanted to convey, see the
+    relevant paragraph below:}
+  \paste{We choose a prior for the initial condition $u_0 \sim \mathcal{N}(0,
+    (-\Delta)^{-1})$, with homogeneous Dirichlet boundary condition. This
+    choice ensures the posterior is well-defined and prior realizations
+    are well-behaved, see \cite[Theorem 3.1 and Lemma 6.25]{Stuart10} for
+    details.}
 
@@
 
-  \answer: I added relevant simulations to a new section Numerical
-  Experiments.
+  \answer{I added relevant simulations to a new section Numerical
+  Experiments.}
 
@@
 
-  \answer: I removed the statement of Theorem 1 from the introudction.
+  \answer{I removed the statement of Theorem 1 from the introudction.}
 
@@
 
-  \answer: Thank you for the comment, I completely agree. Optimal
-  designs are now denoted $\opt$.
+  \answer{Thank you for the comment, I completely agree. Optimal
+  designs are now denoted $\opt$.}
 
@@
 
-  \answer: I am glad you view this part as unnecessary, but I feel the
+  \answer{I am glad you view this part as unnecessary, but I feel the
   model might feel too abstract to some practitioners. Thus, I choose
-  to keep this part to avoid complaints from other readers.
+  to keep this part to avoid complaints from other readers.}
 
@@
 
-  \answer: Basically I meant that its eigenvalues decrease
-  "quickly". I understand this is not a commot or well-defined term so
-  I removed the relevant sentence.
+  \answer{Basically I meant that its eigenvalues decrease
+    "quickly". I understand this is not a common or well-defined term
+    so I removed the relevant sentence.}
 
@@
 
-  \answer: Great comment! Such cases are possible but I view them as
-  pathological. See short paragraph below.
+  \answer{Great comment! Such cases are possible but I view them as
+  pathological. See short paragraph below.}
 
@@
 
-  \answer: I do not expect this conclusion to hold: if noise is
+  \answer{I do not expect this conclusion to hold: if noise is
   present, any measurement will result in some information gain and an
-  increase in design criterion.
+  increase in design criterion.}
 
@@
 
-  \answer: I believe a posterior can be defined, and so a KL
+  \answer{I believe a posterior can be defined, and so a KL
   divergence from said posterior to prior could be calculated and
@@
   implies $\tar \equiv \infty$. I believe this is because do not know
-  how to mitigate this
+  how to mitigate this}
 
@@
 
-  \answer: the Matern-like prior actually arises by using a
+  \answer{the Matern-like prior actually arises by using a
   Laplacian-like operator as a prior, see Rue and Lindgren \cite{}. So
@@
   inverse problem is out of scope for the present paper since I was
-  not able to install common packages for inverse problems on my machine.
+  not able to install common packages for inverse problems on my machine. }
 
@@
 
-  \answer: Indeed, point evaluations $\delta_x$ are not in any
+  \answer{Indeed, point evaluations $\delta_x$ are not in any
   function space I consider in the paper since my analysis is limited
@@
   that does not change the analysis. I skim over this issue in the
-  paper since I think it is confusing and adds little context.
+  paper since I think it is confusing and adds little context.}
 
@@
 
-  \answer: I explain this in more detail in the paper. See text below.
+  \answer{I explain this in more detail in the paper. See text below.}
 

### 2024-07-31 latex/vanishing.tex
Tidy up a bit
Intro ready

@@
 to \eqref{eq:eigenproblem} as a \emph{nonlinear} eigenvalue problem.
-%% Proposition \ref{prop:twice woodbury} and Lemma \ref{lemma:sim
-%% diag} required
+%% Proposition \ref{prop:twice_woodbury} and Lemma \ref{lemma:sim
+%%_diag} required
 
-\begin{proposition}\label{prop:twice woodbury}
+\begin{proposition}\label{prop:twice_woodbury}
   Assume $\fwd \prcov \fwd^*$ is invertible. Then
@@
 
-As we mentioned in Section \ref{subsec:bayesian inverse problems},
-$\fwd \prcov \fwd^*$ is the prior covariance $\hilo$ and could be
-safely assumed invertible. The proof of Proposition \ref{prop:twice
-  woodbury} is delegated to the Supplementary. It amounts to using
-Woodbury's matrix identity twice. The standard proof for Woodbury's
-matrix identity works for separable Hilbert spaces, as long as all
-terms are well defined. Unfortunately, $\obs^*\obs$ is not invertible,
-so we utilize a regularization trick to force it to be.
+As we mentioned in Section \ref{subsec:bayesian_inverse_problems},
+$\fwd \prcov \fwd^*$ is the prior covariance in $\hilo$ and could be
+safely assumed invertible. The proof of Proposition
+\ref{prop:twice_woodbury} is delegated to the Supplementary. It
+amounts to using Woodbury's matrix identity twice. The standard proof
+for Woodbury's matrix identity works for separable Hilbert spaces, as
+long as all terms are well defined. Unfortunately, $\obs^*\obs$ is not
+invertible, so we utilize a regularization trick to force it to be.
 
-\begin{lemma}[Simultaneous diagonizability]\label{lemma:sim diag}
+\begin{lemma}[Simultaneous diagonizability]\label{lemma:sim_diag}
   Let $\hil$ separable Hilbert space, $C:\hil \to \hil$ self-adjoint
@@
   satisfies the unit norm constraint on measurements. Part (2)
-  was proved in Proposition \ref{prop:same ev}. Part (3) was proved in
-  Proposition \ref{prop:true target}.
+  was proved in Proposition \ref{prop:same_ev}. Part (3) was proved in
+  Proposition \ref{prop:true_target}.
 
-  Part (4) is a consequence of Propositions \ref{prop:true target} and
+  Part (4) is a consequence of Propositions \ref{prop:true_target} and
   \ref{prop:kkt}, with the caveat that we did not show that finding
@@
 is the average of the total precision $\sum_{j=1}^k \lambda_j^{-1} +
-\sigma^{-2}m$, see eq.~\eqref{eq:cylinders} and Fig.~\ref{fig:optimal
-  vs not} for an illustration.
+\sigma^{-2}m$, see eq.~\eqref{eq:cylinders} and
+Fig.~\ref{fig:optimal_vs_not} for an illustration.
 %% \begin{figure}%{r}{0.25\textwidth} 

### 2024-07-31 latex/prelims.tex
prelims finished

@@
 \section{Preliminaries and Notation}\label{section:prelim}
-Here we present the basics of Bayesian inverse problems over hilbert
+Here we present the basics of Bayesian inverse problems over Hilbert
 spaces. Since we are ultimately interested in inferring a function
@@
 forward operator \(\fwd\) is linear. In the inverse problem of the 1D
-heat equatin, the forward operator is determined by the time evolution
-of the 1D heat equation \eqref{eq:heat1} and \eqref{eq:heat2}.
+heat equation, the forward operator is determined by the time
+evolution of the 1D heat equation \eqref{eq:heat1} and
+\eqref{eq:heat2}, so $u(\cdot, T) = \fwd u_0$.
 
-Measurements are taken via a \emph{measurement operator} \(\obs\). It
-is common for the measurement and forward operators to be merged
-\(\tmp := \obs \fwd\) \cite{AlexanderianGloorGhattas14}, but the
-analysis carried out in the following sections requires that \(\fwd\)
-and \(\obs\) are explicitly separated (as in
-\cite{attia2022stochastic, cvetkovic2023choosing}). Note that \(\obs
-\in ( \hilo^* )^m\), where \(m\) is the number of measurements
-taken\footnote{The alert reader will likely ask how do we reconcile
-point measurements $\delta_x$ as suggested by the formulation of the
-1D heat equation with workin in Hilbert spaces. We don't. This entire
-study could be formulated for Banach spaces where point measurements
-are allowed. We avoid this technical difficulty because: (1) it does
-not add much to the study, (2) it is beyond the technical abilities of
-the author}. Entries \(\meas_j, j=1,\dots,m\) of the measurement
-operator \(\obs\) are called \emph{measurements}:
-
+Measurements are taken via a linear \emph{measurement operator}
+\(\obs\), which is the concatenation of linear functionals
+$\meas_1,\dots,\meas_m$ we call \emph{measurements}:
 \begin{equation*}%\label{eq:O}
@@
 \end{equation*}
+%% \footnote{It is common for the measurement and forward
+%% operators to be merged \(\tmp := \obs \fwd\)
+%% \cite{AlexanderianGloorGhattas14}, but the analysis carried out in the
+%% following sections requires that \(\fwd\) and \(\obs\) are explicitly
+%% separated, as in \cite{attia2022stochastic,
+%%   cvetkovic2023choosing}.}
+Thus, \(\obs \in ( \hilo^* )^m\), where \(m\) is the number of
+measurements taken\footnote{The alert reader will likely ask how do we
+reconcile point measurements $\delta_x$ as suggested by the
+formulation of the 1D heat equation with working in Hilbert spaces. We
+don't. We follow standard practice in the literature and restrict our
+analysis to Hilbert spaces. We can satisfy ourselves with the fact
+that point evaluations could be approximated in a standard Hilbert
+space like $L^2(\Omega)$.}.
 
@@
 \(I_m \in \mathbb{R}^{m \times m}\) the identity. Both error terms and
-the prior (see below) are assumed independent of each other. Thus,
-data is acquired via
+the prior (see Section \ref{subsec:bayesian_inverse_problems} below)
+are assumed independent of each other. Thus, data is acquired via
 \begin{align}\label{eq:inverse_problem}
@@
 In the previous section, we saw how a parameter $u\in \hilp$ is
-transported to the observation space via the forward operaror $\fwd u
+transported to the observation space via the forward operator $\fwd u
 \in \hilo$, how observations are generated from a parameter via $\obs
@@
 is time to formulate the process of inferring the parameter as a
-Bayesian inverse problem. The first ingredient we need is a prior.
+Bayesian inverse problem. We have already defined the likelihood in
+the previous section, and now we will define the prior.
 
@@
 \(\hilp\) \cite{Stuart10}. For example, for the inverse problem of the
-1D heat equation we chose $\prcov = (-\Delta)^{-1}$ --- see Section
-\ref{subsec:toy}. Note that \(\fwd \prcov \fwd^*\) is the prior
-covariance in \(\hilo\) \cite{Stuart10}, and as such is assumed
+1D heat equation we chose $\prcov = (-\Delta)^{-1}$, as described in
+Section \ref{subsec:toy}. Note that \(\fwd \prcov \fwd^*\) is the
+prior covariance in \(\hilo\) \cite{Stuart10}, and as such is assumed
 invertible --- an assumption which we will use later (if \(\fwd\) has
@@
 this study, but the posterior covariance operator $\postcov$ is given
-by the known formula\cite{Stuart10}:
+by the known formula \cite{Stuart10}:
 \begin{align}\label{eq:postcov}
@@
 posterior and prior measures, the KL divergence is defined,
-analogously to eq.~\ref{eq:basic_KL}, via the Radon-Nikodym
+analogously to eq.~\eqref{eq:basic_KL}, via the Radon-Nikodym
 derivative:
@@
 
-Note that in \cite{AlexanderianGloorGhattas14,
-  alexanderian2018efficient}, results are stated for \(\Sigma=I\)
-(implied by \(\modcov = 0,\sigma^2=1\)), but these results also hold
-for more general covariance matrices
-\cite[p. 681]{AlexanderianGloorGhattas14}.
+In \cite{AlexanderianGloorGhattas14, alexanderian2018efficient},
+results are stated for \(\Sigma=I\) (implied by \(\modcov =
+0,\sigma^2=1\)), but these results also hold for more general
+covariance matrices \cite[p. 681]{AlexanderianGloorGhattas14}.
 

### 2024-07-31 latex/vanishing.tex
vanishin almost done

@@
 
-\begin{lemma}[Simultaneous diagonizability]\label{lemma:sim_diag}
+\begin{lemma}[Simultaneous diagonalizability]\label{lemma:sim_diag}
   Let $\hil$ separable Hilbert space, $C:\hil \to \hil$ self-adjoint
@@
 
-The proof of Lemma is also delegated to the Supplementary. It is
-however important to note that this proof is construcive. It will give
-us a way to construct D-optimal designs, once we fully characterize
-them in Theorem \ref{thm:char} below.
+The proof of Lemma \ref{lemma:free} is also delegated to the
+Supplementary. It is however important to note that this proof is
+constructive; it will allow us to construct D-optimal designs, once we
+fully characterize them in Theorem \ref{thm:char} below.
 
@@
     \end{equation*}
-  \item The covariance of the pushforwad \(\fwd_{*} \postopt\) is \(\left
+  \item The covariance of the pushforward \(\fwd_{*} \postopt\) is \(\left
     ( (\fwd \prcov \fwd^*)^{-1} + \sigma^{-2} \opt^*\opt \right
@@
         \lambda_i &  i > k 
-      \end{cases}
+      \end{cases}.
     \end{equation}
@@
 \subsection{Answer to Question \ref{q:why}}\label{subsec:why}
-Theorem \ref{thm:char} helps us give a compelling explanation of the
-measurement clusterization we obeserved for the inverse problem of the
-heat equation (Fig.~\ref{fig:}). Below, we take the following steps:
-We first utilize Theorem \ref{thm:char} to understand what a D-optimal
-design will look like for our relaxed model; we then argue that
-although not identical, a D-optimal design for the full inverse
-problem of the heat equation shares similar charachteristics to the
-relaxed design; finally, we show how ...
+Building on Theorem \ref{thm:char}, we can now give a compelling
+explanation to the measurement clusterization we observed for the
+inverse problem of the heat equation (Fig.~\ref{fig:eigenvectors}).
 
-First, recall the relevant parts of Theorem \ref{thm:char}. In our
-relaxed model, for a D-optimal design $\opt$:
-\begin{itemize}
-  \item $\opt^*\opt$ and $\fwd\prcov \fwd^*$ are simultaneously
-    diagonalizable.
-  \item $\opt^*\opt$ has non-zero eigenvalues only for the $k:=\rank
-    \opt^*\opt$ largest eigenvalues $\lambda_j$ of $\fwd\prcov\fwd^*$,
-    i.e.~$\eta_j=0$ for $j>k$.
-  \item $k$ is monotonic in $m$, the number of measurements we take.
-\end{itemize}
-
-Consider $\fwd$ and $\prcov$ from the \emph{the inverse problem of the
+Consider $\fwd$ and $\prcov$ from \emph{the inverse problem of the
 heat equation}. As before, we denote the eigenvalues of
 $\fwd\prcov\fwd^*$ by $\lambda_j$. We input these eigenvalues into our
-\emph{relaxed} model, and find a D-optimal design $\opt$ for our
-relaxed model, e.g.~by using Lemma \ref{lemma:free} in the proof of
-Theorem \ref{thm:char}. In our relaxed model, the measurements we take
-are best utilized in reducing uncertainty for the first $k$
-eigenvectors. So, a D-optimal design arising from our \emph{relaxed
-model} completely avoids measuring eigenvectors $k+1$ and above.
+\emph{generic} model, and find a D-optimal design $\opt$ for our
+generic model using Theorem \ref{thm:char}. In our generic model, the
+measurements we take are best utilized in reducing uncertainty for the
+first $k$ eigenvectors. So, a D-optimal design arising from our
+\emph{generic model} completely avoids measuring eigenvectors $k+1$
+and above.
 
 Of course, in a real life problem --- such as the inverse problem of
-the heat equation --- it is likely impossible to find measurement
+the 1D heat equation --- it is likely impossible to find measurement
 locations for which all eigenvectors $k+1$ and above are
-zero. Luckily, the eigenvalues of $\fwd\prcov\fwd^*$ decay very
-quickly: for example, recall from Section \ref{subsec:toy} that the
-eigenvalues of the forward model for the 1D heat equation are $-(2\pi
-n)^{-2}$. Thus a D-optimal design is a trade-off between measuring the
-small number of eigenvectors we are interested in.
+zero. However, if the eigenvalues of $\fwd\prcov\fwd^*$ decay quickly
+(recall the square-exponential decay for eigenvalues of the 1D heat
+equation in eq.\eqref{eq:decay}), a D-optimal design will try to
+balance measuring a small number (i.e.~$k$) of the leading
+eigenvectors.
 
-We explore the abovementione trade-off for the inverse problem of the
-heat equation in Fig.~\ref{fig:eigenvectors}. We allow $m=4$
-measurements in $\Omega = [0,1]$ and observe that D-optimal
-measurement locations are clustered at $x_1 = 0.31$ and $x_2 =
-0.69$. Upon close inspection of the scaled eigenvectors of $\fwd
-\prcov \fwd^*$, we first observe that eigenvectors $3$ and above have
-negligible prior amplitude. Since we only have $m=4$ measurements at
-our disposal, we interpret these results, following Theorem
-\ref{thm:char}, as implying we should only care about measuring the
-first and second eigenvectors. Then, we note the D-optimal $x_1,x_2$
-present a compromise between the amplitude of the first and second
-eigenvectors. For example, a measurement at $x=0.5$ would have ignored
-the second eigenvector altogether, since the second eigenvector is
-zero at $x=0.5$.  Now we can understand measurement clusterization:
-attempting to measure what we want and ignoring the rest, along with
-the spatial limitations on measurements imposed by the structure of
-eigenvectors of $\fwd \prcov \fwd^*$ is the cause for measurement
-clusterization in D-optimal designs: we have $m=4$ measuremnts at our
-disposal but only two spatial locations that are a good compromise
-between the scaled eigenvectors. Thus, clusterization arises by the
-pigeonhole principle.
+The abovementioned trade-off is explored in
+Fig.~\ref{fig:eigenvectors}. We allow $m=4$ measurements in $\Omega =
+[0,1]$ and observe that D-optimal measurement locations are clustered
+at $x_1 = 0.31$ and $x_2 = 0.69$. Upon close inspection of the scaled
+eigenvectors of $\fwd \prcov \fwd^*$, we first observe that
+eigenvectors $3$ and above have negligible prior amplitude. Since we
+only have $m=4$ measurements at our disposal, we interpret these
+results, following Theorem \ref{thm:char}, as implying we should only
+care about measuring the first and second eigenvectors. Then, we note
+the D-optimal $x_1,x_2$ present a compromise between the amplitude of
+the first and second eigenvectors. For example, a measurement at
+$x=0.5$ would have ignored the second eigenvector altogether, since
+the second eigenvector is zero at $x=0.5$.
+
+Now we can understand measurement clusterization for the inverse
+problem of the heat equation. A D-optimal design attempts to measure
+the first $k$ eigenvectors of $\fwd \prcov \fwd^*$. But there may be
+(spatial) limitations on where these $k$ eigenvectors have large
+amplitude. For the inverse problem of the heat equation there are two
+spatial locations that present a good compromise between the
+amplitudes of the first and second eigenvectors, namely $x_1$ and
+$x_2$ --- see Fig.~\ref{fig:eigenvectors}. We have $m=4$ measurements
+at our disposal but only two spatial locations that are a good
+compromise between the first and second scaled eigenvectors. Thus,
+clusterization arises as a consequence of the pigeonhole principle.
 
@@
       compromise between the magnitudes of the first and second
-      eigenvectors. These are the eigenvectors a D-optimal design aims
-      to measure. The rest of the eAllocating $m=4$ measurements into two locations
-      results in clusterization, according to the pigeonhole
+      eigenvectors, which are the eigenvectors that a D-optimal design
+      aims to measure. Allocating $m=4$ measurements into two
+      locations results in clusterization, according to the pigeonhole
       principle.}

### 2024-07-31 latex/numerix.tex
almost done, numerix

@@
 
-\subsection{Simulating Lemma \ref{lemma:free}}\label{subsec:lemma_sims}
+\subsection{Simulating Theorem \ref{thm:char}}\label{subsec:lemma_sims}
 In the proof of Theorem \ref{thm:char} we utilize Lemma
-\ref{lemma:free} to construct D-optimal designs. We implement the
-construction in Lemma \ref{lemma:free} with the goal of testing
-numerically how prevalent are clustered designs. We iterate over the
-number of measurements $m \in \{4,\dots, 15\}$, and for every $m$ we
-then iterate over $k:=\rank \obs^*\obs \in \{2,\dots, m-1\}$. For each
-pair $m,k$ we run $N=5000$ simulations according to the following
-steps:
+\ref{lemma:free} to construct D-optimal designs. We implement this
+construction with the goal of testing numerically how prevalent are
+clustered designs. We iterate over the number of measurements $m \in
+\{4,\dots, 24\}$, and for every $m$ we then iterate over $k:=\rank
+\obs^*\obs \in \{2,\dots, m-1\}$. For each pair $m,k$ we run $N=5000$
+simulations according to the following steps:
 \begin{enumerate}
 \item Generate random diagonal $D\in \mathbb{R}^{k\times k}$ with
-  entries $\log (d_i) \sim \mathcal{N}(5,1)$ and normalize it so that
-  $\ttr D = m$.
+  entries $\log (d_i) \sim \mathcal{N}(25,15)$ and normalize it so
+  that $\ttr D = m$.
 \item Conjugate $D$ by a random orthogonal matrix to form a positive
@@
 We then calculate the fraction of clustered designs of the simulations
-we ran, for each pair $m,k$. Clusteriztion occured at high rates
-($>95\%$) whenever $m-k > 1$; see Fig.~\ref{fig:sim_AAt}. Hence, in
+we ran, for each pair $m,k$. Clusterization occurred at high rates
+($>99.9\%$) whenever $m-k > 1$; see Fig.~\ref{fig:sim_AAt}. Hence, in
 these simulations, clusterization is a generic property. However, when
-$m-k \leq 2$, clusterization does not occur. We do not why and further
+$m-k = 1$, clusterization does not occur. We do not why and further
 investigation into this phenomenon is out of scope for the current
@@
       generated randomly (see text and repository for details on
-      generating $M$). It is evident that when $m-k >3$ clusterization
+      generating $M$). It is evident that when $m-k \geq 2$ clusterization
       is ubiquitous, whereas for lower $m-k$ clusterization does not

### 2024-08-04 latex/intro.tex
Text ready, answer refs

@@
 \subsection{A toy model: the 1D heat equation}\label{subsec:toy}
-For concreteness, let us now consider a toy model of a physical
-process for which we wish to solve an inverse problem: a partial
-differential equation known as the \emph{heat equation} in one
-dimension (1D heat equation henceforth). Readers less familiar with
-partial differential equations can think of the following setup:
-before us there is a perfectly insulated metal rod. At each point on
-the rod the temperature is different and unknown, and we call this
-temperature distribution the \emph{initial condition}. We wait for a
-short time $T$ and let the heat dissipate a bit inside the rod. The
-heat equation determines the heat distribution inside the rod at time
-$t=T$.
+For concreteness, let us now consider a toy inverse problem: inferring
+the initial condition for partial differential equation known as the
+\emph{heat equation} in one dimension (1D heat equation
+henceforth). Readers less familiar with partial differential equations
+can think of the following setup: before us there is a perfectly
+insulated metal rod. At each point on the rod the temperature is
+different and unknown, and we call this temperature distribution the
+\emph{initial condition}. We wait for a short time $T$ and let the
+heat dissipate a bit inside the rod. The heat equation determines the
+heat distribution inside the rod at time $t=T$.
 
@@
 are in fact eigenvectors of the Laplacian: $\Delta \ev_n = -\pi^2
-n^2\ev_n$. It should not be hard to believe that eq.~\eqref{eq:heat1}
-implies $u(\cdot, T) = \sum_{n\geq 1} a_n \exp(-\pi^2 n^2T ) \ev_n$
+n^2\ev_n$. It should not be hard to believe that $u(\cdot, T) =
+\sum_{n\geq 1} a_n \exp(-\pi^2 n^2T ) \ev_n$
 \cite{renardy2006PDE}. Thus, eigenvectors of the linear operator that
@@
 infer the initial condition $u_0 = u(\cdot, 0)$ from observations of
-the final condition $u(\cdot, T)$. However, we are not able to measure
+the final state $u(\cdot, T)$. However, we are not able to measure
 $u(\cdot, T)$ at every point in the domain $\Omega$. Rather, we take

### 2024-08-04 latex/numerix.tex
Text ready, answer refs

@@
 construction with the goal of testing numerically how prevalent are
-clustered designs. We iterate over the number of measurements $m \in
-\{4,\dots, 24\}$, and for every $m$ we then iterate over $k:=\rank
-\obs^*\obs \in \{2,\dots, m-1\}$. For each pair $m,k$ we run $N=5000$
-simulations according to the following steps:
+clustered designs. To this end, we would like to generate random prior
+eigenvalues $\lambda_j$, fix $m$ and $k$, find a D-optimal
+$\opt^*\opt$ and then utilize the construction of
+Theorem~\ref{thm:char} and Lemma~\ref{lemma:free} to find $\opt$.
+
+To simplify things, we directly sample rank $\opt^*\opt$. We iterate
+over the number of measurements $m \in \{4,\dots, 24\}$, and for every
+$m$ we then iterate over $k:=\rank \obs^*\obs \in \{2,\dots,
+m-1\}$. For each pair $m,k$ we repeat the following steps $N=5000$
+times:
 \begin{enumerate}
 \item Generate random diagonal $D\in \mathbb{R}^{k\times k}$ with
-  entries $\log (d_i) \sim \mathcal{N}(25,15)$ and normalize it so
-  that $\ttr D = m$.
+  entries $\log (d_i) \sim \mathcal{N}(50,15)$ and normalize so that
+  $\ttr D = m$\footnote{Results are robust to the choice of parameters
+  in the lognormal distribution above.}.
 \item Conjugate $D$ by a random orthogonal matrix to form a positive
-  semi-definite $M := UDU^t \in \mathbb{R}^{k\times k}$.
+  semi-definite $M := UDU^t \in \mathbb{R}^{k\times k}$. This $M$
+  represents $\opt^*\opt$.
 \item Apply the construction of Lemma \ref{lemma:free} to calculate
-  $A$ such that $AA^t = M$, where $A$ has unit norm columns.
-\item Since columns of $A$ correspond to measurement vectors, we call
-  $A$ "clustered" if $A$ has two or more identical columns (up to some
-  numerical precision threshold).
+  $A$ such that $AA^t = M$, where $A$ has unit norm columns. I.e.~find
+  the optimal design $\opt$.
+\item Since $A$ corresponds to $\opt$, its columns correspond to
+  measurement vectors. We call $A$ "clustered" if $A$ has two or more
+  identical columns (up to some numerical precision threshold).
 \end{enumerate}
@@
 these simulations, clusterization is a generic property. However, when
-$m-k = 1$, clusterization does not occur. We do not why and further
-investigation into this phenomenon is out of scope for the current
-study.
+$m-k = 1$, clusterization does not occur. We do not why this is so and
+further investigation into this phenomenon is out of scope for the
+current study.
 
@@
 Full results are located in the \texttt{simulations.csv} file within
-the accompanying \href{https://github.com/yairdaon/OED}{repository},
-code implementing the experiments described above is located in module
-\texttt{zeros.py} of said repository.
+the accompanying \href{https://github.com/yairdaon/OED}{repository}.
+Code implementing the experiments described above is located in module
+\texttt{zeros.py} of said repository. Runtime should be less than 30
+minutes on any reasonably modern laptop (it took 12 minutes on the
+author's laptop).
 

### 2024-08-04 clusterization.py

@@ -7,6 +7,11 @@ from src.observations import PointObservation
 
 FS = 24 ## Font size
 
+## Avoid clusterization by increasing this to e.g. 4
+MODEL_ERROR = 0
+
+
+
 def main():
 
     sig = 5e-2 ## Observation error amplitude
@@ -27,8 +32,6 @@ def main():
     ## when employing homogeneous Neumann boundary condition.
     delta = 0. if transform == 'dst' else 0.5
     
-    ## Avoid clusterization by increasing this to e.g. 4
-    model_error = 0
 
     ## Forward model - the heat equation with BC chosen according to transform
     fwd = Heat(N=N,
@@ -54,7 +57,7 @@ def main():
                      L=L,
                      N=N,
                      transform=transform,
-                     model_error=model_error)
+                     model_error=MODEL_ERROR)
 
 
     
@@ -82,7 +85,7 @@ def main():
     plt.xlabel("Eigenvector", fontsize=FS)
     plt.ylabel("Precision", fontsize=FS)
     plt.tight_layout()
-    plt.savefig(f"latex/figs/FgammaFstar_modelError{model_error}.png")
+    plt.savefig(f"latex/figs/FgammaFstar_modelError{MODEL_ERROR}.png")
     
     
     ###################################
@@ -128,7 +131,7 @@ def main():
     ax.set_xlim(0,1)
     plt.tight_layout()
 
-    plt.savefig(f"latex/figs/{transform}_modelError{model_error}.png")
+    plt.savefig(f"latex/figs/{transform}_modelError{MODEL_ERROR}.png")
     #plt.show()
 
     ##############################
@@ -176,7 +179,7 @@ def main():
     plt.xlabel(r"$x \in \Omega$", fontsize=fs)
     plt.legend(title='eigenvector, weighted')
     plt.tight_layout()
-    plt.savefig(f"latex/figs/eigenvectors_{transform}_modelError{model_error}.png")
+    plt.savefig(f"latex/figs/eigenvectors_{transform}_modelError{MODEL_ERROR}.png")
    
 
     

### 2024-08-04 latex/shell.sh

@@ -1,0 +1,9 @@
+latexpand ba.tex > flat.tex
+latexdiff --math-markup=OFF old.tex flat.tex > diff.tex
+latexmk --pdf ba.tex
+latexmk --pdf ba.tex
+latexmk --pdf ba.tex
+latexmk --pdf diff.tex
+latexmk --pdf diff.tex
+latexmk --pdf diff.tex
+latexmk --pdf diff.tex

### 2024-08-04 zeros.py

@@ -67,7 +67,7 @@ def MDUS(m, k):
     
     assert m >= k
 
-    D = np.random.lognormal(mean=25, sigma=15, size=k)
+    D = np.random.lognormal(mean=50, sigma=15, size=k)
     D = np.sort(D)[::-1]
     # D = np.array([np.random.lognormal(mean=i, sigma=1) for i in range(k)]) 
     D = D * m / np.sum(D)

### 2024-08-08 latex/flat.tex
Add explanation following pazman pronzato

@@
 The practice of experimental design entails specifying every detail of
-running an experiment \cite{chaloner1995} and at its simplest form,
+running an experiment \cite{chaloner1995}. At its simplest form,
 experimental design involves choosing a set of measurements to be
 taken. For example, consider the process of searching for oil:
-measuremnts require digging holes in the ground and a \emph{design}
+measurements require digging holes in the ground and a \emph{design}
 simply indicates which holes should be dug
 \cite{horesh2008borehole}. Since measurements are expensive, only a
-limited number of measuremnts can be taken, and thus carefully
-choosing measuremt locations is crucial.
-
-Specifiyng the right set of measurements holds particular significance
+limited number of them can be taken, and thus carefully choosing
+measurement locations is crucial.
+
+Specifying the right set of measurements holds particular significance
 when solving an \emph{inverse problem} --- i.e.~making inference of
@@
 \begin{equation*}
-  \mathbb{E}_{\data}\left [ D_{KL}\left (\Pr(\param|\data)||\Pr(\param)\right ) \right ]
-\end{equation*}
-We refer to a set of measuremnts that maximizes the D-optimality
+  \mathbb{E}_{\data}\left [ D_{KL}\left (\Pr(\param|\data)||\Pr(\param)\right ) \right ].
+\end{equation*}
+We refer to a set of measurements that maximizes the D-optimality
 criterion as a \emph{D-optimal design}.
@@
 described by the following three equations:
-\begin{subequations}\begin{alignat}{2}
+\begin{subequations}
+  \begin{alignat}{2}
     u_t &= \Delta u &&\qquad \text{in } [0,1] \times [0,\infty), \label{eq:heat1}\\
@@
 initial condition is quickly smoothed, as implied by the squared
-exponential decay as $n\to\infty$ in eq.~\eqref{eq:decay}. See
-Supplementary movies S1 and S2 for an illustration of this phenomenon
-(code generating these movies is located in module \texttt{movies.py}
-in the accompanying
+exponential decay of eigenvalues in eq.~\eqref{eq:decay} with
+$n\to\infty$. See Supplementary movies S1 and S2 for an illustration
+of this phenomenon (code generating these movies is located in module
+\texttt{movies.py} in the accompanying
 \href{https://github.com/yairdaon/OED}{repository}). For ill-posed
@@
 We choose a prior for the initial condition $u_0 \sim \mathcal{N}(0,
-(-\Delta)^{-1})$, with homogeneous Dirichlet boundary condition. Choosing a prior This
-choice ensures the posterior is well-defined and prior realizations
-are well-behaved, see \cite[Theorem 3.1 and Lemma 6.25]{Stuart10} for
-details.
+(-\Delta)^{-1})$, with homogeneous Dirichlet boundary condition. This
+choice of prior covariance operator ensures the posterior is
+well-defined and prior realizations are well-behaved, see
+\cite[Theorem 3.1 and Lemma 6.25]{Stuart10} for details. The second
+ingredient of a fully Bayesian formulation of the inverse problem is
+likelihood. We simply assume centered iid Gaussian measurement error
+model, giving rise to a the widely applicable Gaussian likelihood
+function.
 
@@
     \includegraphics[height=0.5\textwidth]{figs/dst_modelError0.png}
-    \caption{Measurement clusterization for D-optimal designs for the
-      inverse problem of the heat equation. Measurement locations were
-      chosen according to the Bayesian D-optimality criterion of
+    \caption{Measurement clusterization in D-optimal designs for the
+      inverse problem of the 1D heat equation. Measurement locations
+      were chosen according to the Bayesian D-optimality criterion of
       Theorem \ref{thm:d_optimality}. Measurement locations are
@@
 \ref{section:prelim} and \ref{section:D_and_grad} we propose a novel
-relaxed model for an inverse problem where D-optimality maintains
-analytical tractability, where D-optimal designs are identified via
+generic model for an inverse problem where D-optimality maintains
+analytical tractability and D-optimal designs are identified via
 Lagrange multipliers. This analytical framework facilitates the
@@
   theoretical perspective, we show that clusterization mainly depends
-  on how quickly the eigenvalues of the prior measure in observation
-  space decay. This decay depends mostly on how ill-posed the problem
-  at hand is and does not depend much on the prior. See Section
-  \ref{section:vanishing} --- particularly, Theorem \ref{thm:char} and
-  the discussion following it. We also show results of numerical
-  experiments, where simulations of our model give rise to D-optimal
-  designs that exhibit clusterization with high probability in Section
-  \ref{subsec:lemma_sims}. Thus, given the genericity of our model, we
-  expect measurement clusterization to be a generic and ubiquitous
-  phenomenon.
+  on how quickly the eigenvalues of the prior covariance in
+  observation space decay. This decay depends mostly on how ill-posed
+  the problem at hand is and does not depend much on the prior. See
+  Section \ref{section:vanishing} --- particularly, Theorem
+  \ref{thm:char} and the discussion following it. We also show results
+  of numerical experiments, where simulations of our model give rise
+  to D-optimal designs that exhibit clusterization with high
+  probability in Section \ref{subsec:lemma_sims}. Thus, given the
+  genericity of our model, we expect measurement clusterization to be
+  a generic and ubiquitous phenomenon.
 
@@
   measurements result in no gain in design criterion when observation
-  error tends to zero. Moreover, in Section
+  error amplitude tends to zero. Moreover, in Section
   \ref{subsec:corr_errors_sims}, we show that an error term
@@
 \item \label{q:why} \textbf{Why does measurement clusterization
-  occur?} In Section \ref{subsec:why} (specifically, See
-  Fig~\ref{fig:why}) we give a compelling answer to this question by
-  transporting insights we gain from our analytically tractable
-  relaxed model to the inverse problem of the 1D heat equation. Our
-  analysis reveals that a D-optimal design focuses on a select set of
-  prior eigenvectors, specifically those with the largest eigenvalues
-  in the prior covariance spectrum. In practical scenarios, the number
-  of locations where (a) all relevant prior eigenvectors are
+  occur?} In Section \ref{subsec:why} (specifically, Fig~\ref{fig:why}
+  and the discussion following it) we give a compelling answer to this
+  question by transporting insights we gain from our generic model to
+  the inverse problem of the 1D heat equation. Our analysis reveals
+  that a D-optimal design focuses on a select set of prior
+  eigenvectors, specifically those with the largest eigenvalues in the
+  prior covariance spectrum. In practical scenarios, the number of
+  locations where (a) all relevant prior eigenvectors are
   significantly large, and (b) other eigenvectors are close to zero,
@@
 equation, clusterization occurs primarily because Laplacian
-eigenvectors are nonzero almost everywhere in $\Omega$. Consequently,
-measuring $u(x_1, T)$ at some point $x_1 \in \Omega$ provides
-information about $u(x_2,T)$ for distant points $x_2 \in
-\Omega$. Intuitively, this should not occur: for small $T$, the heat
-distribution at $x_1$ should have minimal influence on the heat
-distribution at $x_2$.
-
-This behavior stems from a well-known property of the heat equation:
-it allows information to spread instantly across the computational
-domain \cite{renardy2006PDE}. In reality, heat and information
-propagate at finite speeds. The emergence of clusterization in this
-context is thus non-physical, arising from the nature of Laplacian
-eigenvectors being nonzero almost everywhere. Clusterization therefore
-indicates that the underlying mathematical model is overly permissive
-and fails to capture crucial physical constraints. We suggest that
+eigenvectors \emph{do not} decay in $\Omega$. Consequently, measuring
+$u(x_1, T)$ at some point $x_1 \in \Omega$ provides information about
+$u(x_2,T)$ for distant points $x_2 \in \Omega$. Intuitively, this
+should not occur: for small $T$, the heat distribution at $x_1$ should
+give very little knowledge on the heat distribution at $x_2$. This
+behavior stems from a well-known property of the heat equation: it
+allows information to spread \emph{instantly} across the computational
+domain \cite{renardy2006PDE}. In reality, heat (and information)
+propagate at finite speeds. Of course, the known physical barrier for
+information spread is the speed of light, but we expect heat to spread
+considerably slower: heating an Olympic pool at one end should have no
+immediate effect on the temperature at the other end.
+
+We believe that the emergence of clusterization in this context is
+thus non-physical, arising from the way the inverse problems we
+consider are phrased. Clusterization therefore indicates that the
+underlying mathematical model is overly permissive and fails to
+capture crucial physical constraints of the problem. We suggest that
 when clusterization occurs, practitioners should consider alternative
 models where information is localized in space and travels at finite
-speeds. Such models may not only provide more physically accurate and
-meaningful results but may also mitigate the issue of clusterization.
+speed in the medium. Such models may not only provide more physically
+accurate and meaningful results but may also mitigate the issue of
+clusterization.
 
@@
  \section{Preliminaries and Notation}\label{section:prelim}
-Here we present the basics of Bayesian inverse problems over hilbert
+Here we present the basics of Bayesian inverse problems over Hilbert
 spaces. Since we are ultimately interested in inferring a function
@@
 forward operator \(\fwd\) is linear. In the inverse problem of the 1D
-heat equatin, the forward operator is determined by the time evolution
-of the 1D heat equation \eqref{eq:heat1} and \eqref{eq:heat2}.
-
-Measurements are taken via a \emph{measurement operator} \(\obs\). It
-is common for the measurement and forward operators to be merged
-\(\tmp := \obs \fwd\) \cite{AlexanderianGloorGhattas14}, but the
-analysis carried out in the following sections requires that \(\fwd\)
-and \(\obs\) are explicitly separated (as in
-\cite{attia2022stochastic, cvetkovic2023choosing}). Note that \(\obs
-\in ( \hilo^* )^m\), where \(m\) is the number of measurements
-taken\footnote{The alert reader will likely ask how do we reconcile
-point measurements $\delta_x$ as suggested by the formulation of the
-1D heat equation with workin in Hilbert spaces. We don't. This entire
-study could be formulated for Banach spaces where point measurements
-are allowed. We avoid this technical difficulty because: (1) it does
-not add much to the study, (2) it is beyond the technical abilities of
-the author}. Entries \(\meas_j, j=1,\dots,m\) of the measurement
-operator \(\obs\) are called \emph{measurements}:
-
+heat equation, the forward operator is determined by the time
+evolution of the 1D heat equation \eqref{eq:heat1} and
+\eqref{eq:heat2}, so $u(\cdot, T) = \fwd u_0$.
+
+Measurements are taken via a linear \emph{measurement operator}
+\(\obs\), which is the concatenation of linear functionals
+$\meas_1,\dots,\meas_m$ we call \emph{measurements}:
 \begin{equation*}\obs u = (\meas_1(u), \dots, \meas_m(u) )^t \in \R^m,\ u \in \hilo.
 \end{equation*}
+Thus, \(\obs \in ( \hilo^* )^m\), where \(m\) is the number of
+measurements taken\footnote{The alert reader will likely ask how do we
+reconcile point measurements $\delta_x$ as suggested by the
+formulation of the 1D heat equation with working in Hilbert spaces. We
+don't. We follow standard practice in the literature and restrict our
+analysis to Hilbert spaces. We can satisfy ourselves with the fact
+that point evaluations could be approximated in a standard Hilbert
+space like $L^2(\Omega)$.}.
 
@@
 problem where sensors are concerned, the norm of measurements recorded
-by sensors is always one\footnote{Again, our analysis does not apply
-to point evaluations. These are just used for motivation and can be
-approximated by elements in $\hilo^*$ as long as $\hilo$ is a function
-space, e.g.~$L^2(\Omega)$.}:
+by sensors is always one\footnote{Again, our analysis does not
+directly apply to point evaluations. We just utilize point evaluations
+for motivation. We can approximate point evaluations by e.g.~elements
+in $\hilo^*$ as long as $\hilo$ is a function space,
+e.g.~$L^2(\Omega)$. In this case, for a fixed apporximation of
+$\delta$ the norm of the corresponding functional is constant (but
+$\neq 1$.}:
 
@@
 We have shown that in the limit $\sigma^2 \to 0$, no increase in
-$\tar$ is achieved by repeating a measurement. Thus, for vanishing
-noise levels, clustered designs cannot be D-optimal. For repeated
+$\tar$ is gained by repeating a measurement. Thus, for vanishing noise
+levels, clustered designs cannot be D-optimal. For repeated
 measurements and $\sigma^2=0$, $\Sigma$ is not invertible and the
@@
 It is worth noting that by the nonnegativity of the KL divergence,
-$\tar$ cannot decrease upon adding measuremnts. However, we can
-construct examples where the posterior is not changed upon taking a
+$\tar$ cannot decrease upon adding measurements. However, we can
+construct examples where the posterior does not change upon taking a
 new measurement e.g.~if the prior variance vanishes on some
-eigenvector and a measuremnt is taken on said eigenvector. We do not
-expect a measurement to generate no information whatsoever in any
-realistic scenario, and we choose to ignore such pathologies.
+eigenvector and a measurement is taken on said eigenvector. We do not
+expect a measurement to generate no information gain whatsoever in any
+realistic scenario, and ignore such pathologies.
 
@@
 \subsection{Answer to Question \ref{q:why}}\label{subsec:why}
-Theorem \ref{thm:char} helps us give a compelling explanation of the
-measurement clusterization we obeserved for the inverse problem of the
-heat equation (Fig.~\ref{fig:}). Below, we take the following steps:
-We first utilize Theorem \ref{thm:char} to understand what a D-optimal
-design will look like for our relaxed model; we then argue that
-although not identical, a D-optimal design for the full inverse
-problem of the heat equation shares similar charachteristics to the
-relaxed design; finally, we show how ...
-
-First, recall the relevant parts of Theorem \ref{thm:char}. In our
-relaxed model, for a D-optimal design $\opt$:
-\begin{itemize}
-  \item $\opt^*\opt$ and $\fwd\prcov \fwd^*$ are simultaneously
-    diagonalizable.
-  \item $\opt^*\opt$ has non-zero eigenvalues only for the $k:=\rank
-    \opt^*\opt$ largest eigenvalues $\lambda_j$ of $\fwd\prcov\fwd^*$,
-    i.e.~$\eta_j=0$ for $j>k$.
-  \item $k$ is monotonic in $m$, the number of measurements we take.
-\end{itemize}
-
-Consider $\fwd$ and $\prcov$ from the \emph{the inverse problem of the
+Building on Theorem \ref{thm:char}, we can now give a compelling
+explanation to the measurement clusterization we observed for the
+inverse problem of the heat equation (Fig.~\ref{fig:eigenvectors}).
+
+Consider $\fwd$ and $\prcov$ from \emph{the inverse problem of the
 heat equation}. As before, we denote the eigenvalues of
 $\fwd\prcov\fwd^*$ by $\lambda_j$. We input these eigenvalues into our
-\emph{relaxed} model, and find a D-optimal design $\opt$ for our
-relaxed model, e.g.~by using Lemma \ref{lemma:free} in the proof of
-Theorem \ref{thm:char}. In our relaxed model, the measurements we take
-are best utilized in reducing uncertainty for the first $k$
-eigenvectors. So, a D-optimal design arising from our \emph{relaxed
-model} completely avoids measuring eigenvectors $k+1$ and above.
+\emph{generic} model, and find a D-optimal design $\opt$ for our
+generic model using Theorem \ref{thm:char}. In our generic model, the
+measurements we take are best utilized in reducing uncertainty for the
+first $k$ eigenvectors. So, a D-optimal design arising from our
+\emph{generic model} completely avoids measuring eigenvectors $k+1$
+and above.
 
 Of course, in a real life problem --- such as the inverse problem of
-the heat equation --- it is likely impossible to find measurement
+the 1D heat equation --- it is likely impossible to find measurement
 locations for which all eigenvectors $k+1$ and above are
-zero. Luckily, the eigenvalues of $\fwd\prcov\fwd^*$ decay very
-quickly: for example, recall from Section \ref{subsec:toy} that the
-eigenvalues of the forward model for the 1D heat equation are $-(2\pi
-n)^{-2}$. Thus a D-optimal design is a trade-off between measuring the
-small number of eigenvectors we are interested in.
-
-We explore the abovementione trade-off for the inverse problem of the
-heat equation in Fig.~\ref{fig:eigenvectors}. We allow $m=4$
-measurements in $\Omega = [0,1]$ and observe that D-optimal
-measurement locations are clustered at $x_1 = 0.31$ and $x_2 =
-0.69$. Upon close inspection of the scaled eigenvectors of $\fwd
-\prcov \fwd^*$, we first observe that eigenvectors $3$ and above have
-negligible prior amplitude. Since we only have $m=4$ measurements at
-our disposal, we interpret these results, following Theorem
-\ref{thm:char}, as implying we should only care about measuring the
-first and second eigenvectors. Then, we note the D-optimal $x_1,x_2$
-present a compromise between the amplitude of the first and second
-eigenvectors. For example, a measurement at $x=0.5$ would have ignored
-the second eigenvector altogether, since the second eigenvector is
-zero at $x=0.5$.  Now we can understand measurement clusterization:
-attempting to measure what we want and ignoring the rest, along with
-the spatial limitations on measurements imposed by the structure of
-eigenvectors of $\fwd \prcov \fwd^*$ is the cause for measurement
-clusterization in D-optimal designs: we have $m=4$ measuremnts at our
-disposal but only two spatial locations that are a good compromise
-between the scaled eigenvectors. Thus, clusterization arises by the
-pigeonhole principle.
+zero. However, if the eigenvalues of $\fwd\prcov\fwd^*$ decay quickly
+(recall the square-exponential decay for eigenvalues of the 1D heat
+equation in eq.\eqref{eq:decay}), a D-optimal design will try to
+balance measuring a small number (i.e.~$k$) of the leading
+eigenvectors.
+
+The abovementioned trade-off is explored in
+Fig.~\ref{fig:eigenvectors}. We allow $m=4$ measurements in $\Omega =
+[0,1]$ and observe that D-optimal measurement locations are clustered
+at $x_1 = 0.31$ and $x_2 = 0.69$. Upon close inspection of the scaled
+eigenvectors of $\fwd \prcov \fwd^*$, we first observe that
+eigenvectors $3$ and above have negligible prior amplitude. Since we
+only have $m=4$ measurements at our disposal, we interpret these
+results, following Theorem \ref{thm:char}, as implying we should only
+care about measuring the first and second eigenvectors. Then, we note
+the D-optimal $x_1,x_2$ present a compromise between the amplitude of
+the first and second eigenvectors. For example, a measurement at
+$x=0.5$ would have ignored the second eigenvector altogether, since
+the second eigenvector is zero at $x=0.5$.
+
+Now we can understand measurement clusterization for the inverse
+problem of the heat equation. A D-optimal design attempts to measure
+the first $k$ eigenvectors of $\fwd \prcov \fwd^*$. But there may be
+(spatial) limitations on where these $k$ eigenvectors have large
+amplitude. For the inverse problem of the heat equation there are two
+spatial locations that present a good compromise between the
+amplitudes of the first and second eigenvectors, namely $x_1$ and
+$x_2$ --- see Fig.~\ref{fig:eigenvectors}. We have $m=4$ measurements
+at our disposal but only two spatial locations that are a good
+compromise between the first and second scaled eigenvectors. Thus,
+clusterization arises as a consequence of the pigeonhole principle.
 

### 2024-08-08 latex/intro.tex
Add explanation following pazman pronzato

@@
 
-It is important to note that we do not view measurement clusterization
-as undesirable, nor do we believe it should be avoided at
+It is important to note that we do not view clustered designs as
+undesirable, nor do we believe it should be avoided at
 all. Clusterization is a peculiar phenomenon and it is perfectly
@@
 
+Our choice of prior is also a potential major contributor to
+clusterization. Our choice of Gaussian prior similarly implies
+information is shared between distant locations in $\Omega$. Thus, we
+suggest refraining from choosing Gaussian priors with inverse
+Laplacian covariance operators. Rather, non-Gaussian priors could be
+employed instead \cite{hosseini2017, hosseini2019}.
+
 We believe that the emergence of clusterization in this context is
@@
 consider are phrased. Clusterization therefore indicates that the
-underlying mathematical model is overly permissive and fails to
-capture crucial physical constraints of the problem. We suggest that
-when clusterization occurs, practitioners should consider alternative
-models where information is localized in space and travels at finite
-speed in the medium. Such models may not only provide more physically
-accurate and meaningful results but may also mitigate the issue of
-clusterization.
+underlying mathematical / Bayesian model is overly permissive and
+fails to capture crucial physical constraints of the problem. We
+suggest that when clusterization occurs, practitioners should consider
+alternative models where information is localized in space and travels
+at finite speed in the medium. Such models may not only provide more
+physically accurate and meaningful results but may also mitigate the
+issue of clusterization.
 

### 2024-08-08 latex/numerix.tex
Add explanation following pazman pronzato

@@
   entries $\log (d_i) \sim \mathcal{N}(50,15)$ and normalize so that
-  $\ttr D = m$\footnote{Results are robust to the choice of parameters
-  in the lognormal distribution above.}.
+  $\ttr D = m$. 
 \item Conjugate $D$ by a random orthogonal matrix to form a positive

### 2024-08-08 latex/vanishing.tex
Text ready, answer refs
Add explanation following pazman pronzato

@@
 explanation to the measurement clusterization we observed for the
-inverse problem of the heat equation (Fig.~\ref{fig:eigenvectors}).
+inverse problem of the heat equation, see Section
+\ref{subsub:clusterization1} below. We also suggest a generic
+explanation for clusterization, see Section \ref{subsub:clusterization2}.
 
+\subsubsection{Clusterization in the heat equation}
 Consider $\fwd$ and $\prcov$ from \emph{the inverse problem of the
@@
 
-The abovementioned trade-off is explored in
+The abovementioned balance is explored in
 Fig.~\ref{fig:eigenvectors}. We allow $m=4$ measurements in $\Omega =
@@
 \end{figure}
+
+\subsubsection{Clusterization in our generic model}
+In our generic model, the limit on the number of measurements arises
+as a consequence of Carath\'eodory's Theorem\footnote{Let $X \subseteq
+\mathbb{R}^n, X \neq \phi$ and denote $\conv (X)$ the convex hull of
+$X$. Let $x \in \conv (X)$. Then $x$ is a convex combination of at
+most $n+1$ vectors in $X$.} and the concentration on the first $k$
+eigenvectors of $\fwd \prcov \fwd^*$. The discussion below extends
+arguments presented in \cite[Chapter 3]{silvey} and \cite[Section
+  5.2.3]{pronzatoPazman2013}.
+
+We now ignore all but the first $k$ eigenvectors of $\fwd \prcov
+\fwd$, as instructed by Theorem \ref{thm:char}. Thus, we can safely
+replace $\hilo$ with the $k$-dimensional subspace spanned by the first
+$k$ eigenvectors of $\fwd^*\prcov\fwd$. Since this subspace is
+isomorphic to $\mathbb{R}^k$, we do not name it and simply denote it
+by $\mathbb{R}^k$. Now, let
+\begin{equation*}
+  \mathcal{M} := \conv \{\meas \meas^* : \meas\in \mathbb{R}^k, \|\meas\|=1\},
+  %\hilo^{(k)}\},
+\end{equation*}
+where $\conv$ denotes the convex hull of a set. Since $\mathcal{M}$ is
+a set of positive-definite matrices, it lives in a
+$k(k+1)/2$-dimensional subspace of $\mathbb{R}^{k\times k}$. Now note
+that $\frac{1}{m}\opt^*\opt \in \mathcal{M}$. By Carath\'eodory's
+Theorem, there exists a convex combination
+\begin{equation*}
+  \frac{1}{m} \opt^*\opt = \sum_{i=1}^{\frac{k(k+1)}{2} + 1} \alpha_i \meas\meas^*,
+\end{equation*}
+where $\|\meas\|=1, \sum_{i=1}^I\alpha_i = 1$ and $\alpha_i > 0$.
+
+For example, if $k=2$ --- as is the case for the inverse prblem of the
+heat equation --- then an optimal design $\opt$ is a combination of at
+most $3$ measurement vectors $\meas$. Thus, if the number of allowed
+measurements $m$ is larger than $3$, then there exists an optimal
+clustered design.
+
+%%  whereThus, if we view $\obs$ as a $\mathbb{R}^{m\times k}$ matrix, we
+%%  conclude that $\obs^*\obs$ is a symmetric positive definite matrix
+%%  that lives in a $k(k+1)/2$-dimensional space.
+
+%% The set $\conv \mathcal{M}$ is a superset of all possible
+
+%% Since we restrict $\|\meas\| = 1$, the set of allowed
+%% $\obs^*\obs$ is the ball of norm $m$ unit ball in (a space isomorphic to)
+%% $\mathbb{R}^k.
+
+%% $\{\meas \meas: \meas \in \span (\ev_1,\ev_2,\dots,
+%% \ev_k\}$ is in fact $k$-dimensional.
+
+%%   By Theorem
+%%   \ref{thm:char}, we ignore all eigenvectors except for the first
+%%   $k$.\ \
+
+%% First, note that $\fwd^*\obs^*\obs\fwd$ is the Fisher inf

### 2024-08-11 latex/vanishing.tex
almost finished with pazman

@@
 \subsubsection{Clusterization in our generic model}
-In our generic model, the limit on the number of measurements arises
-as a consequence of Carath\'eodory's Theorem\footnote{Let $X \subseteq
-\mathbb{R}^n, X \neq \phi$ and denote $\conv (X)$ the convex hull of
-$X$. Let $x \in \conv (X)$. Then $x$ is a convex combination of at
-most $n+1$ vectors in $X$.} and the concentration on the first $k$
-eigenvectors of $\fwd \prcov \fwd^*$. The discussion below extends
-arguments presented in \cite[Chapter 3]{silvey} and \cite[Section
+We can gain further insight to clusterization in our generic model,
+from Carath\'eodory's Theorem and the concentration on the first $k$
+eigenvectors of $\fwd \prcov \fwd^*$. We present a short discussion
+adapting arguments presented in \cite[Chapter
+  3]{silveyOptimalDesign1980} and \cite[Section
   5.2.3]{pronzatoPazman2013}.
 
-We now ignore all but the first $k$ eigenvectors of $\fwd \prcov
-\fwd$, as instructed by Theorem \ref{thm:char}. Thus, we can safely
-replace $\hilo$ with the $k$-dimensional subspace spanned by the first
-$k$ eigenvectors of $\fwd^*\prcov\fwd$. Since this subspace is
-isomorphic to $\mathbb{R}^k$, we do not name it and simply denote it
-by $\mathbb{R}^k$. Now, let
+Consider $\opt$ a D-optimal design under our generic model.  As
+instructed by Theorem \ref{thm:char}, we ignore all but the first $k$
+eigenvectors of $\fwd \prcov \fwd$. Thus, we replace $\hilo$ with
+$\hilo^{(k)}$ --- the $k$-dimensional subspace spanned by the first
+$k$ eigenvectors of $\fwd^*\prcov\fwd$. We keep in mind that
+$\hilo^{(k)}$ is isomorphic to $\mathbb{R}^k$ and let
 \begin{equation*}
-  \mathcal{M} := \conv \{\meas \meas^* : \meas\in \mathbb{R}^k, \|\meas\|=1\},
+  \mathcal{M} := \conv \{\meas \meas^* : \meas\in \hilo^{(k)}, \|\meas\|=1\},
   %\hilo^{(k)}\},
 \end{equation*}
-where $\conv$ denotes the convex hull of a set. Since $\mathcal{M}$ is
-a set of positive-definite matrices, it lives in a
-$k(k+1)/2$-dimensional subspace of $\mathbb{R}^{k\times k}$. Now note
-that $\frac{1}{m}\opt^*\opt \in \mathcal{M}$. By Carath\'eodory's
-Theorem, there exists a convex combination
+where $\conv$ denotes the convex hull of a set. The set $\mathcal{M}$
+contains only positive-definite matrices, hence it lives in a
+$k(k+1)/2$-dimensional vector space. Since $\opt^*\opt = \sum_{i=1}^m
+\meas_i\meas_i^*$ for $\meas_i \in \hilo^{(k)}$, it is easy to verify
+that $\frac1m \opt^*\opt \in \mathcal{M}$.  Now, recall
+Carath\'eodory's Theorem:
+\begin{theorem}[Carath\'eodory]
+  Let $X \subseteq \mathbb{R}^n, X \neq \phi$ and denote $\conv (X)$
+  the convex hull of $X$. For every $x \in \conv (X)$, $x$ is a convex
+  combination of at most $n+1$ vectors in $X$.
+\end{theorem}
+Carath\'eodory's Theorem implies that there exist $\meas_i$ and
+$\alpha_i$ such that
 \begin{equation*}
-  \frac{1}{m} \opt^*\opt = \sum_{i=1}^{\frac{k(k+1)}{2} + 1} \alpha_i \meas\meas^*,
+  \opt^*\opt = \sum_{i=1}^I \alpha_i \meas_i\meas_i^*,
 \end{equation*}
-where $\|\meas\|=1, \sum_{i=1}^I\alpha_i = 1$ and $\alpha_i > 0$.
+where $\|\meas\|=1, \sum\alpha_i = m, \alpha_i \geq 0$ and $I =
+\frac{k(k+1)}{2} + 1$. We can thus write $\opt$ as (note that we do
+not require $\meas_i$ to be orthogonal to each other):
 
-For example, if $k=2$ --- as is the case for the inverse prblem of the
-heat equation --- then an optimal design $\opt$ is a combination of at
-most $3$ measurement vectors $\meas$. Thus, if the number of allowed
-measurements $m$ is larger than $3$, then there exists an optimal
-clustered design.
+\[
+\opt =
+\left[
+  \begin{array}{ccc}
+    \horzbar & \sqrt{\alpha_i} \meas^*_1 & \horzbar \\
+    \horzbar & \sqrt{\alpha_2} \meas_2^* & \horzbar \\
+             & \vdots    &          \\
+    \horzbar & \sqrt{\alpha_I} \meas_I^* & \horzbar \\
+  \end{array}
+\right]
+\]
+
+Therefore, if $m > \frac{k(k+1)}{2} + 1$, then at least one
+coefficient $\alpha_i > 1$. Note that the abovementioned $\opt$ is not
+a valid design (since its rows do not have unit norm,
+i.e.~$\|\sqrt{\alpha_i}\meas\| \neq 1$). However, such representation
+provides another, deep view of clusterization.
+
+%% For example, if $k=2$, there exists an optimal design $\opt$ that can
+%% be represented be as a linear combination of at most $(2+1)/2 + 1 = 4$
+%% measurement vectors $\meas$, with nonnegative coefficients whose
+%% squares sum to $m$. If, the number of allowed measurements $m=5$, we
+%% can write $\opt = \sum_{i=1}^4 \alpha_i \meas_i$, where $\alpha_i > 1$
+%% for some $i\in \{1,2,3,4\}$. Since we put weight $>1$ on the
+%% $i\textsuperscript{th}$ measurement, we can view such a design as
+%% clustered.
 

### 2024-08-11 latex/intro.tex
done

@@
 \item \label{q:why} \textbf{Why does measurement clusterization
-  occur?} In Section \ref{subsec:why} (specifically, Fig~\ref{fig:why}
-  and the discussion following it) we give a compelling answer to this
-  question by transporting insights we gain from our generic model to
-  the inverse problem of the 1D heat equation. Our analysis reveals
-  that a D-optimal design focuses on a select set of prior
-  eigenvectors, specifically those with the largest eigenvalues in the
-  prior covariance spectrum. In practical scenarios, the number of
-  locations where (a) all relevant prior eigenvectors are
-  significantly large, and (b) other eigenvectors are close to zero,
-  is limited. Consequently, the clusterization of measurements arises
-  as a natural consequence of the pigeonhole principle, as there are
-  more measurements available than there are locations satisfying
-  conditions (a) and (b).
+  occur?} In Sections \ref{subsec:why} we give two compelling answers
+  to this question by (i) transporting insights we gain from our
+  generic model to the inverse problem of the 1D heat equation, and
+  (ii) connecting measurement clusterization to Carath\'eodory's
+  Theorem.
+
+  Our analysis for the heat equation relies on conclusions from our
+  generic model. In particular, Theorem \ref{thm:char} reveals that a
+  D-optimal design focuses on a select set of prior eigenvectors,
+  i.e.~those with the largest eigenvalues in the prior covariance
+  spectrum. In practical scenarios, the number of locations where (a)
+  all relevant prior eigenvectors are significantly large, and (b)
+  other eigenvectors are close to zero, is limited. Consequently, the
+  clusterization of measurements arises as a natural consequence of
+  the pigeonhole principle, as there are more measurements available
+  than there are locations satisfying conditions (a) and (b).
+
+  The connection to Carath\'eodory's Theorem also utilizes the
+  concentration of measurement effort on the abovementioned select set
+  of prior eigenvectors. This concentration allows us to move from an
+  infinite-dimensional setting to finite dimensions, where the
+  conditions of Carath\'eodory's Theorem hold. We conclude that a
+  D-optimal design arises by weighting a small number of
+  measurements. If the number of allowed measurements is larger than
+  the small number dictated by Carath\'eodory's Theorem, we have
+  excessive weight on some measurements, which can be interpreted as
+  clusterization.
 

### 2024-08-11 latex/vanishing.tex
done

@@
 
-\subsubsection{Clusterization in our generic model}
+\subsubsection{Clusterization in our generic model}%\label{subsub:cara}
 We can gain further insight to clusterization in our generic model,
@@
 $\hilo^{(k)}$ --- the $k$-dimensional subspace spanned by the first
-$k$ eigenvectors of $\fwd^*\prcov\fwd$. We keep in mind that
-$\hilo^{(k)}$ is isomorphic to $\mathbb{R}^k$ and let
+$k$ eigenvectors of $\fwd^*\prcov\fwd$. Let
 \begin{equation*}
@@
 where $\conv$ denotes the convex hull of a set. The set $\mathcal{M}$
-contains only positive-definite matrices, hence it lives in a
-$k(k+1)/2$-dimensional vector space. Since $\opt^*\opt = \sum_{i=1}^m
-\meas_i\meas_i^*$ for $\meas_i \in \hilo^{(k)}$, it is easy to verify
-that $\frac1m \opt^*\opt \in \mathcal{M}$.  Now, recall
-Carath\'eodory's Theorem:
+contains only positive-definite operators on a $k$-dimensional vector
+space. Hence $\mathcal{M}$ lives in a $k(k+1)/2$-dimensional vector
+space. Since $\opt^*\opt = \sum_{i=1}^m \meas_i\meas_i^*$ for $\meas_i
+\in \hilo^{(k)}$, it is easy to verify that $\frac1m \opt^*\opt \in
+\mathcal{M}$.  Recall Carath\'eodory's Theorem:
 \begin{theorem}[Carath\'eodory]
@@
 where $\|\meas\|=1, \sum\alpha_i = m, \alpha_i \geq 0$ and $I =
-\frac{k(k+1)}{2} + 1$. We can thus write $\opt$ as (note that we do
-not require $\meas_i$ to be orthogonal to each other):
+\frac{k(k+1)}{2} + 1$. We can thus write $\opt$ as:
+%(note that we do not require $\meas_i$ to be orthogonal to each
+%other):
 
@@
   \end{array}
-\right]
+\right].
 \]
 
-Therefore, if $m > \frac{k(k+1)}{2} + 1$, then at least one
-coefficient $\alpha_i > 1$. Note that the abovementioned $\opt$ is not
-a valid design (since its rows do not have unit norm,
-i.e.~$\|\sqrt{\alpha_i}\meas\| \neq 1$). However, such representation
-provides another, deep view of clusterization.
-
-%% For example, if $k=2$, there exists an optimal design $\opt$ that can
-%% be represented be as a linear combination of at most $(2+1)/2 + 1 = 4$
-%% measurement vectors $\meas$, with nonnegative coefficients whose
-%% squares sum to $m$. If, the number of allowed measurements $m=5$, we
-%% can write $\opt = \sum_{i=1}^4 \alpha_i \meas_i$, where $\alpha_i > 1$
-%% for some $i\in \{1,2,3,4\}$. Since we put weight $>1$ on the
-%% $i\textsuperscript{th}$ measurement, we can view such a design as
-%% clustered.
-
-%%  whereThus, if we view $\obs$ as a $\mathbb{R}^{m\times k}$ matrix, we
-%%  conclude that $\obs^*\obs$ is a symmetric positive definite matrix
-%%  that lives in a $k(k+1)/2$-dimensional space.
-
-%% The set $\conv \mathcal{M}$ is a superset of all possible
-
-%% Since we restrict $\|\meas\| = 1$, the set of allowed
-%% $\obs^*\obs$ is the ball of norm $m$ unit ball in (a space isomorphic to)
-%% $\mathbb{R}^k.
-
-%% $\{\meas \meas: \meas \in \span (\ev_1,\ev_2,\dots,
-%% \ev_k\}$ is in fact $k$-dimensional.
-
-%%   By Theorem
-%%   \ref{thm:char}, we ignore all eigenvectors except for the first
-%%   $k$.\ \
-
-%% First, note that $\fwd^*\obs^*\obs\fwd$ is the Fisher inf
+Unfortunately, $\opt$ is not a valid design, since its rows do not
+have unit norm. Still, the above representation of $\opt$ is useful:
+If $m > \frac{k(k+1)}{2} + 1$, then $\alpha_i > 1$ for some $1\leq i
+\leq I$.  Thus, we can view $\opt$ as a clustered design, since it
+places weight $>1$ on a single measurement vector.

### 2024-08-12 latex/caratheodory.tex
add eic and AE

--- 2024-08-11 latex/vanishing.tex
@@
-\section{D-Optimal Designs Without Model Error}\label{section:vanishing}
-Our goal in this section is to prove Theorem \ref{thm:char} which
-characterizes D-optimal designs when $\modcov = 0$. The necessary
-first-order condition for D-optimality of Theorem
-\ref{thm:constrained} for $\modcov = 0$ become:
-
-\begin{equation}\label{eq:eigenproblem}
-  \sigma^{-2}\fwd \postcov \fwd^* \obs^* = \obs^* \Xi,
-\end{equation}
-with $\Xi$ diagonal. Equation \eqref{eq:eigenproblem} looks like an
-eigenvalue problem for the self-adjoint operator $\sigma^{-2}\fwd
-\postcov \fwd^*$, where rows of $\obs$, namely $\meas_j,j=1,\dots, m$,
-are eigenvectors. However, $\postcov$ depends on $\obs$, so we refer
-to \eqref{eq:eigenproblem} as a \emph{nonlinear} eigenvalue problem.
-%% Proposition \ref{prop:twice_woodbury} and Lemma \ref{lemma:sim
-%%_diag} required
-
-\begin{proposition}\label{prop:twice_woodbury}
-  Assume $\fwd \prcov \fwd^*$ is invertible. Then
-  \begin{align*}
-    \begin{split}
-      \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* 
-      %
-      %
-      = \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1},
-    \end{split}
-  \end{align*}  
-\end{proposition}
-
-As we mentioned in Section \ref{subsec:bayesian_inverse_problems},
-$\fwd \prcov \fwd^*$ is the prior covariance in $\hilo$ and could be
-safely assumed invertible. The proof of Proposition
-\ref{prop:twice_woodbury} is delegated to the Supplementary. It
-amounts to using Woodbury's matrix identity twice. The standard proof
-for Woodbury's matrix identity works for separable Hilbert spaces, as
-long as all terms are well defined. Unfortunately, $\obs^*\obs$ is not
-invertible, so we utilize a regularization trick to force it to be.
-
-\begin{lemma}[Simultaneous diagonalizability]\label{lemma:sim_diag}
-  Let $\hil$ separable Hilbert space, $C:\hil \to \hil$ self-adjoint
-  and $\func_1,\dots,\func_m \in \hil$. Denote $\func^*$ the element
-  $\func$ acting as a linear functional. If
-  \begin{equation*}
-   (C + \sum_{j=1}^m \func_j\func_j^*) \func_l = \xi_l \func_l,\ l = 1,\dots,m
-  \end{equation*}
-  then $C$ and $\sum_{j=1}^m \func_j \func_j^*$ are simultaneously
-  diagonalizable.
-\end{lemma}
-The proof of Lemma \ref{lemma:sim_diag} is delegated to the
-Supplementary.
-
-\begin{proposition}\label{prop:same_ev}
-  Let $\obs$ satisfy the nonlinear eigenvalue problem
-  \eqref{eq:eigenproblem}. Then $\obs^*\obs$ and $\fwd \prcov \fwd^*$
-  are simultaneously diagonalizable.
-\end{proposition}
-\begin{proof}
-  \begin{align}\label{eq:mod_conditions}
-    \begin{split}
-      \obs^* \Xi &= \sigma^{-2}\fwd \postcov \fwd^* \obs^*  \text{ (by \eqref{eq:eigenproblem})}\\
-      %
-      %
-      %
-      &= \sigma^{-2} \fwd( \prcov^{-1} + \sigma^{-2}  \fwd^* \obs^* \obs \fwd )^{-1} \fwd^* \obs^*  \text{ (by \eqref{eq:postcov})} \\
-      %
-      %
-      %
-      &= \sigma^{-2} \left ( (\fwd\prcov\fwd^*)^{-1} + \sigma^{-2}  \obs^* \obs \right )^{-1} \obs^* \text{ (by Proposition \ref{prop:twice_woodbury})}.
-    \end{split}
-  \end{align}
-
-  Now take $\func_j^{*} = \meas_j$ and $C := (\fwd \prcov
-  \fwd^*)^{-1}$ and use Lemma \ref{lemma:sim_diag}.
-\end{proof}
-
-Since we made no assumption regarding the ordering of $\{\lambda_i\}$,
-we can denote the corresponding non-zero eigenvalues of $\obs^*\obs$
-by $\{\eta_i\}_{i=1}^{k}$ and let $\eta_i = 0$ for $i \geq k+1$.
-
-\begin{proposition}\label{prop:true_target}
-  Let $\obs$ with $m$ measurements satisfy the nonlinear eigenvalue
-  problem \eqref{eq:eigenproblem}. Let $\{\eta_i\}_{i=1}^{\infty}$
-  eigenvalues of $\obs^*\obs$ and $\{\lambda_i\}_{i=1}^{\infty}$ the
-  corresponding eigenvalues of $\fwd \prcov \fwd^*$. Let $k:=\rank
-  \obs^*\obs$. Without loss of generality, let $\eta_i > 0$ for $i\leq
-  k$ and $\eta_i = 0$ for $i > k$. Then:
-  \begin{enumerate}
-    \item $k \leq m$ and $\obs^*\obs$ has exactly $k$ positive
-      eigenvalues.
-    \item
-      \begin{equation*}
-        \tar(\obs) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i) = \frac12 \sum_{i=1}^{m} \log (1 + \sigma^{-2}\lambda_i\eta_i).
-      \end{equation*}
-    \item Furthermore, if $\obs$ is D-optimal, $\eta_i > 0$ for
-      eigenvectors corresponding to the $k$ largest $\lambda_i$.
-  \end{enumerate}
-\end{proposition}
-\begin{proof}
-  Part (1) is trivial. To see part (2) holds: 
-  \begin{align}
-    \begin{split}
-      \tar(\obs) &= \frac12\log \det \left (I + \sigma^{-2} \prcov^{1/2} \fwd ^* \obs^*
-      \obs \fwd \prcov^{1/2}\right )\\% \text{ (by definition)}\\
-      %
-      &= \frac12 \log \det \left (I + \sigma^{-2} \obs^* \obs \fwd
-      \prcov\fwd^* \right ) \text{ (Sylvester's Determinant
-      Theorem)}\\
-      %
-      %
-      %
-      &=\frac12 \log \prod_{i=1}^{\infty} ( 1 + \sigma^{-2} \lambda_i\eta_i ) \text{ (Proposition \ref{prop:same_ev})} \\
-      %
-      %
-      %
-      %&=\frac12 \log \left ( \prod_{i=1}^{k} ( \lambda_i^{-1} + \sigma^{-2} \eta_i )\prod_{i=1}^{k} \lambda_i \right )
-      %
-      %
-      %
-      &=\frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i). 
-      %
-      %
-      %% &=\frac12 \sum_{i=1}^{k} \log(\sigma^2\lambda_i^{-1} + \eta_i) - k\log \sigma + \frac12\sum_{i=1}^k \log \lambda_i.
-    \end{split}
-  \end{align}
-  Part (3) holds since $\log$ is increasing and $\eta_i \geq 0$.
-\end{proof}
-
-\begin{proposition}\label{prop:kkt}
-  Let $\tar: \mathbb{R}^m \to \mathbb{R}$, $\tar(\eta) =
-  \frac{1}{2}\sum_{i=1}^m \log (1+\sigma^{-2}\lambda_i \eta_i)$, with
-  $\lambda_i > 0$ and $\sigma^{2} > 0$. Then the maximum of $\tar$
-  subject to $\eta_i \geq 0$ and $\sum\eta_i = m$ is obtained at
-  \begin{equation}
-  \eta_i = \begin{cases}
-    \frac{m}{k} - \sigma^2 \lambda_i^{-1} + \sigma^2 \frac{1}{k} \sum_{j\in A} \lambda_j^{-1} & i \in A \\
-    0 & i \in A^c
-  \end{cases}
-  \end{equation}
-  where $A:= \{1\leq i \leq m: \eta_i > 0\}$ and $A^c = \{1,\dots, m\}
-  \backslash A$, and $k = |A|$, the cardinality of $A$.
-\end{proposition}
-
-The proof of Proposition \ref{prop:kkt} amounts to utilizing the
-Karush-Kuhn-Tucker conditions and is delegated to the
-Supplementary. The final ingredient we require for the proof of
-Theorem \ref{thm:char} characterizing D-optimal designs is:
-
-\begin{lemma}[Unit norm decomposition]\label{lemma:free}
-  Let $M \in \R^{k \times k}$ symmetric positive definite with $\ttr M
-  = m$, $m \geq k$. We can find $\func_j \in \R^k,j=1,\dots,m$
-  with $\|\func_j\|=1$ and $A = (\func_1,\dots,\func_m)$ such that
-  $AA^t = M$.
-\end{lemma}
-
-The proof of Lemma \ref{lemma:free} is also delegated to the
-Supplementary. It is however important to note that this proof is
-constructive; it will allow us to construct D-optimal designs, once we
-fully characterize them in Theorem \ref{thm:char} below.
-
-\begin{theorem}\label{thm:char}
-  Let:
-  \begin{itemize}
-    \item The D-optimality design criterion
-    \cite{AlexanderianGloorGhattas14}:
-    \begin{align*}
-      \begin{split}
-        \tar(\obs) %:&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
-        % 
-        % 
-        % 
-        &= \frac12 \log \det ( I + \sigma^{-2} \prcov^{1/2} \fwd ^*
-        \obs^* \obs \fwd \prcov^{1/2}), 
-      \end{split}
-    \end{align*}
-  \item \(\opt\) a D-optimal design operator
-    \begin{equation*}
-      \opt = \argmax_{\|\meas_j\| = 1, j=1,\dots,m}\tar(\obs),
-    \end{equation*}
-  \item \(\{\lambda_i\}_{i=1}^\infty\) eigenvalues of
-    \(\fwd\prcov\fwd^*\) in decreasing order of magnitude.
-    %% \item \(\{\ev_i\}_{i=1}^\infty\) their corresponding eigenvectors.
-  \item \(\{\eta_i\}_{i=1}^\infty\) eigenvalues of \(\opt^*\opt\).
-
-  \end{itemize}
-
-  Then:
-  \begin{enumerate}
-  \item  \(\tr{\opt^*\opt} = m\).
-  \item \(\opt^*\opt\) and \(\fwd\prcov\fwd^*\) are simultaneously
-    diagonalizable.
-  \item \(k := \rank \opt^*\opt \leq m\) and
-    \begin{equation*}
-      \tar(\opt) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i). %= \frac12 \sum_{i=1}^{m} \log (1 + \sigma^{-2}\lambda_i\eta_i).
-    \end{equation*}
-  %% \item 
-  %%   \begin{equation*}
-  %%     k = \argmax \left \{ k:\lambda_k^{-1} < \sigma^{-2}\frac{m}{k} + \frac{1}{k} \sum_{j=1}^{k}
-  %%     \lambda_j^{-1} \right \}.
-  %%   \end{equation*}
-  \item
-    \begin{equation*}
-        \eta_i = \begin{cases}
-          \frac{m}{k} - \sigma^2 \lambda_i^{-1} + \sigma^2 \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1} & 1 \leq i \leq k \\
-          0 & i > k 
-        \end{cases}.
-    \end{equation*}
-  \item The covariance of the pushforward \(\fwd_{*} \postopt\) is \(\left
-    ( (\fwd \prcov \fwd^*)^{-1} + \sigma^{-2} \opt^*\opt \right
-    )^{-1}\) and its eigenvalues are
-    \begin{equation}\label{eq:cylinders}
-      \theta_i =
-      \begin{cases}
-        \left(\frac{\sum_{j=1}^k \lambda_j^{-1} + \sigma^{-2}m}{k} \right )^{-1} & i \leq k \\
-        \lambda_i &  i > k 
-      \end{cases}.
-    \end{equation}
-  \end{enumerate}
-\end{theorem}
-
-\begin{proof}
-  Part (1) is immediate for any measurement operator $\obs$ that
-  satisfies the unit norm constraint on measurements. Part (2)
-  was proved in Proposition \ref{prop:same_ev}. Part (3) was proved in
-  Proposition \ref{prop:true_target}.
-
-  Part (4) is a consequence of Propositions \ref{prop:true_target} and
-  \ref{prop:kkt}, with the caveat that we did not show that finding
-  $\opt$ so that $\opt^*\opt$ has the desired eigenvalues is
-  feasible. To this end, we utilize Lemma \ref{lemma:free}: let $M =
-  \diag(\eta_1, \dots, \eta_k)$, diagonal with respect to the first
-  $k$ eigenvectors of $\fwd \prcov \fwd^*$. We take $\opt := A$ from
-  Lemma \ref{lemma:free}.
-
-  Recall from \eqref{eq:postcov}, that the posterior precision is
-  $\postcov^{-1} = \prcov^{-1} + \sigma^{-2}\fwd^*\opt^*\opt\fwd$. The
-  first statement in part (5) now follows from Proposition
-  \ref{prop:twice_woodbury}, while the second statement follows from
-  parts (1) and (4).
-\end{proof}
-
-\pgfplotstableread{
-  Label     prior  optimal  sub-optimal 
-  1         0.2    1.8           1.7
-  2         0.8    1.2           0.8
-  3         2.2    0             0.5
-  4         3.5    0             0.0
-}\optimalvsnot
-
-\begin{figure}\label{fig:tikz_clusterization}
-  \centering
-  \begin{tikzpicture}[scale=0.75]
-    \begin{axis}[
-        ybar stacked,
-        ymin=0,
-        ymax=4,
-        xtick=data,
-        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
-        reverse legend=false, % set to false to get correct display, but I'd like to have this true
-        xticklabels from table={\optimalvsnot}{Label},
-        xticklabel style={text width=2cm,align=center},
-        legend plot pos=right,
-        ylabel={\LARGE precision --- prior and posterior},
-        xlabel={\LARGE eigenvector},
-      ]
-      \addplot [fill=blue!60]  table [y=prior,   meta=Label, x expr=\coordindex] {\optimalvsnot};
-      \addplot [pattern=north east lines, pattern color=green!80]  table [y=optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};     
-      \addlegendentry[scale=1.4]{$\sigma^2\lambda_i^{-1}$}
-      \addlegendentry[scale=1.4]{optimal $\eta_i$s}
-    \end{axis}
-  \end{tikzpicture}
-  \begin{tikzpicture}[scale=0.75]
-    \begin{axis}[
-        ybar stacked,
-        ymin=0,
-        ymax=4,
-        xtick=data,
-        legend style={cells={anchor=east}, legend pos=north west, legend columns=-1},
-        reverse legend=false, % set to false to get correct display, but I'd like to have this true
-        xticklabels from table={\optimalvsnot}{Label},
-        xticklabel style={text width=2cm,align=center},
-        legend plot pos=right,
-        ylabel={\LARGE precision --- prior and posterior},
-        xlabel={\LARGE eigenvector} ,
-      ]   
-      \addplot [fill=blue!60]  table [y=prior,       meta=Label, x expr=\coordindex] {\optimalvsnot};
-      \addplot [pattern=north east lines, pattern color=green!80]  table [y=sub-optimal, meta=Label, x expr=\coordindex] {\optimalvsnot};
-      \addlegendentry[scale=1.4]{$\sigma^2\lambda_i^{-1}$}
-      \addlegendentry[scale=1.4]{sub-optimal $\eta_i$s}
-    \end{axis}
-  \end{tikzpicture}
-  \caption{A comparison of the eigenvalues of the pushforward
-    posterior precision $(\fwd\prcov\fwd^*)^{-1} +
-    \sigma^{-2}\obs^*\obs$ for a D-optimal design (left) and a
-    sub-optimal design (right). Both designs are allowed $m=3$
-    measurements. We assume $\sigma^2=1$ and thus, the blue area has
-    accumulated height of $\sigma^{-2}m = 3$ in both panels. The
-    D-optimal design (left) increases precision where it is
-    lowest. The sub-optimal design (right) does not.}
-  \label{fig:optimal_vs_not}
-\end{figure}
-
-Part (5) of Theorem \ref{thm:char} gives us deep understanding of
-D-optimal designs when $\modcov = 0$: Imagine each eigenvector of the
-\emph{precision} $\left(\fwd \prcov \fwd^*\right )^{-1}$ corresponds
-to a graduated lab cylinder. % (see Fig\ref{fig:cylinder}) Each
-cylinder $i$ is filled, a-priori, with green liquid of
-$\lambda_i^{-1}$ volume units. We are allowed $m$ measurement, so we
-have blue liquid of volume $\sigma^{-2}m$ units at our disposal. When
-we seek a D-optimal design, we distribute the blue liquid by
-repeatedly adding a drop to whatever cylinder currently has the lowest
-level of liquid in it, as long as its index $i \leq m$. The result of
-such a procedure is that the precision for the first $k$ eigenvectors
-is the average of the total precision $\sum_{j=1}^k \lambda_j^{-1} +
-\sigma^{-2}m$, see eq.~\eqref{eq:cylinders} and
-Fig.~\ref{fig:optimal_vs_not} for an illustration.
-%% \begin{figure}%{r}{0.25\textwidth} 
-%%     \centering
-%%     \includegraphics[width=5cm, height=5cm]{cylinders.jpg}
-%%     \caption{Three graduated lab cylinders, corresponding to three
-%%      eigenvectors. Prior precision eigenvalues not shown.}
-%%     \label{fig:cylinder}
-%% \end{figure}
-
-\subsection{Answer to Question \ref{q:why}}\label{subsec:why}
-Building on Theorem \ref{thm:char}, we can now give a compelling
-explanation to the measurement clusterization we observed for the
-inverse problem of the heat equation, see Section
-\ref{subsub:clusterization1} below. We also suggest a generic
-explanation for clusterization, see Section \ref{subsub:clusterization2}.
-
-\subsubsection{Clusterization in the heat equation}
-Consider $\fwd$ and $\prcov$ from \emph{the inverse problem of the
-heat equation}. As before, we denote the eigenvalues of
-$\fwd\prcov\fwd^*$ by $\lambda_j$. We input these eigenvalues into our
-\emph{generic} model, and find a D-optimal design $\opt$ for our
-generic model using Theorem \ref{thm:char}. In our generic model, the
-measurements we take are best utilized in reducing uncertainty for the
-first $k$ eigenvectors. So, a D-optimal design arising from our
-\emph{generic model} completely avoids measuring eigenvectors $k+1$
-and above.
-
-Of course, in a real life problem --- such as the inverse problem of
-the 1D heat equation --- it is likely impossible to find measurement
-locations for which all eigenvectors $k+1$ and above are
-zero. However, if the eigenvalues of $\fwd\prcov\fwd^*$ decay quickly
-(recall the square-exponential decay for eigenvalues of the 1D heat
-equation in eq.\eqref{eq:decay}), a D-optimal design will try to
-balance measuring a small number (i.e.~$k$) of the leading
-eigenvectors.
-
-The abovementioned balance is explored in
-Fig.~\ref{fig:eigenvectors}. We allow $m=4$ measurements in $\Omega =
-[0,1]$ and observe that D-optimal measurement locations are clustered
-at $x_1 = 0.31$ and $x_2 = 0.69$. Upon close inspection of the scaled
-eigenvectors of $\fwd \prcov \fwd^*$, we first observe that
-eigenvectors $3$ and above have negligible prior amplitude. Since we
-only have $m=4$ measurements at our disposal, we interpret these
-results, following Theorem \ref{thm:char}, as implying we should only
-care about measuring the first and second eigenvectors. Then, we note
-the D-optimal $x_1,x_2$ present a compromise between the amplitude of
-the first and second eigenvectors. For example, a measurement at
-$x=0.5$ would have ignored the second eigenvector altogether, since
-the second eigenvector is zero at $x=0.5$.
-
-Now we can understand measurement clusterization for the inverse
-problem of the heat equation. A D-optimal design attempts to measure
-the first $k$ eigenvectors of $\fwd \prcov \fwd^*$. But there may be
-(spatial) limitations on where these $k$ eigenvectors have large
-amplitude. For the inverse problem of the heat equation there are two
-spatial locations that present a good compromise between the
-amplitudes of the first and second eigenvectors, namely $x_1$ and
-$x_2$ --- see Fig.~\ref{fig:eigenvectors}. We have $m=4$ measurements
-at our disposal but only two spatial locations that are a good
-compromise between the first and second scaled eigenvectors. Thus,
-clusterization arises as a consequence of the pigeonhole principle.
-
-\begin{figure}\label{fig:eigenvectors}
-    \centering
-    \includegraphics[width=\textwidth]{figs/eigenvectors_dst_scaled.png}
-    \caption{D-optimal measurement locations ($m=4$ measurements) and
-      weighted eigenvectors for finding the initial condition of the
-      1D heat equation. Measurement locations and weighted
-      eigenvectors are plotted over the computational domain $\Omega =
-      [0, 1]$ (x-axis). Measurement clusterization occurs
-      approximately at $0.31$ and $0.69$. These two locations are a
-      compromise between the magnitudes of the first and second
-      eigenvectors, which are the eigenvectors that a D-optimal design
-      aims to measure. Allocating $m=4$ measurements into two
-      locations results in clusterization, according to the pigeonhole
-      principle.}
-  \label{fig:why}
-\end{figure}
-
-\subsubsection{Clusterization in our generic model}%\label{subsub:cara}
 We can gain further insight to clusterization in our generic model,
@@
 \mathcal{M}$.  Recall Carath\'eodory's Theorem:
-\begin{theorem}[Carath\'eodory]
+\begin{theorem*}[Carath\'eodory]
   Let $X \subseteq \mathbb{R}^n, X \neq \phi$ and denote $\conv (X)$
@@
   combination of at most $n+1$ vectors in $X$.
-\end{theorem}
+\end{theorem*}
 Carath\'eodory's Theorem implies that there exist $\meas_i$ and

### 2024-08-12 latex/implications.tex
add eic and AE

--- 2024-08-08 latex/intro.tex
@@
-\section{Introduction}\label{section:intro}
-The practice of experimental design entails specifying every detail of
-running an experiment \cite{chaloner1995}. At its simplest form,
-experimental design involves choosing a set of measurements to be
-taken. For example, consider the process of searching for oil:
-measurements require digging holes in the ground and a \emph{design}
-simply indicates which holes should be dug
-\cite{horesh2008borehole}. Since measurements are expensive, only a
-limited number of them can be taken, and thus carefully choosing
-measurement locations is crucial.
-
-Specifying the right set of measurements holds particular significance
-when solving an \emph{inverse problem} --- i.e.~making inference of
-the physical world utilizing a physical model of a phenomenon of
-interest \cite{tarantola2005,kaipio2005}. In an inverse problem, we
-seek to infer physical quantities, based on observations and a model
-of the physical world. Models are typically phrased in the language of
-ordinary or partial differential equations.
-
-Inverse problems are ubiquitous in science and engineering. For
-example, in electrical impedance tomography (EIT) we seek to infer the
-structure of the inside of a human body. Inference is conducted based
-on observations of electrical current impedance measured at specific
-electrode locations on the skin, leveraging known equations of
-electric current flow \cite{horesh2010impedance}. Magnetic Resonance
-Imaging (MRI) also entails solving an inverse problem. There,
-radio-frequency pulses are sent through the human body in the presence
-of a strong magnetic field. The response of one's body contents to
-these radio-frequency pulses is measured, allowing a radiologist to
-view the internals of a body in a noninvasive manner
-\cite{horesh2008mri}. An inverse problem also arises in oil and gas
-exploration, where acoustic waves are sent through "boreholes" ---
-deep cylindrical wells drilled into the ground. Data of travel and
-return times of the acoustic waves are recorded. Wave travel and
-return time is influenced by the properties of the subsurface
-materials, such as density, elasticity, and the presence of fluids or
-voids. Combining travel time data with a geophysical models of
-contents of the earth's crust facilitates reconstructing the structure
-of the subsurface in a process called \emph{borehole tomography}
-\cite{horesh2008borehole}. Inverse problems also arise in many other
-areas of seismology and geology \cite{rabinowitz1990, steinberg1995}
-and medical imaging \cite{tarantola2005}. In many inverse problems the
-goal is to infer some numerical parameter. However, in the formulation
-we consider in this study, as well as in the examples above, the goal
-is to conduct inference over some \emph{function} over $\Omega$, where
-\(\Omega \subseteq \mathbb{R}^d, d=1,2,3\) is a spatial domain of
-interest.
-
-\subsection{D-optimal Designs}\label{subsec:D}
-In all of the examples of inverse problems we have seen, as well as in
-many other applications, only a limited number of measurements are
-allowed. For example, in borehole tomography, a measurement involves
-drilling a deep well in the ground --- a very costly endeavor. Even in
-a seemingly simpler application such as MRI, a patient cannot occupy
-the MRI machine for too long, hence only a limited number of
-measurements can be taken. Therefore, measurements should be chosen to
-most benefit the user. Typically, a user will consider some utility,
-called a \emph{design criterion} and take measurements that maximize
-this utility. Two of the most widely recognized and extensively
-studied design criteria are the A- and \emph{D-optimality} criteria.
-
-In this study, we focus on the Bayesian D-optimality
-criterion. Bayesian D-optimality carries a simple and intuitive
-meaning: measurements chosen according to the D-optimality criterion
-maximize the expected Kullback-Leibler (KL) divergence from posterior
-to prior \cite{chaloner1995, AlexanderianGloorGhattas14,
-  CoverThomas91}. Utilizing KL divergence for quantifying differences
-between distributions is standard practice and we will not motivate it
-further here. For a discrete parameter $\param$ and data $\data$, the
-KL divergence is defined as
-\begin{equation}\label{eq:basic_KL}
-  D_{KL}\left (\Pr(\param|\data)||\Pr(\param)\right ) =  \sum_{\param} \log
-  \frac{\Pr(\param|\data)}{\Pr(\param)} \Pr(\param|\data). 
-\end{equation}
-Of course, data $\data$ is not known before the experiment is
-conducted, so we average over it to define the D-optimality criterion
-as:
-\begin{equation*}
-  \mathbb{E}_{\data}\left [ D_{KL}\left (\Pr(\param|\data)||\Pr(\param)\right ) \right ].
-\end{equation*}
-We refer to a set of measurements that maximizes the D-optimality
-criterion as a \emph{D-optimal design}.
-
-For a linear model with Gaussian prior in finite dimensions, a
-D-optimal design minimizes the determinant of the posterior covariance
-\cite{chaloner1995}. In Section \ref{subsec:D_optimal_design} we give
-a more general definition of the D-optimality criterion that applies
-to arbitrary measures. We also show how the D-optimality criterion
-generalizes to linear models over infinite-dimensional Hilbert spaces
-(i.e.~function spaces).
-
-\subsection{A toy model: the 1D heat equation}\label{subsec:toy}
-For concreteness, let us now consider a toy inverse problem: inferring
-the initial condition for partial differential equation known as the
-\emph{heat equation} in one dimension (1D heat equation
-henceforth). Readers less familiar with partial differential equations
-can think of the following setup: before us there is a perfectly
-insulated metal rod. At each point on the rod the temperature is
-different and unknown, and we call this temperature distribution the
-\emph{initial condition}. We wait for a short time $T$ and let the
-heat dissipate a bit inside the rod. The heat equation determines the
-heat distribution inside the rod at time $t=T$.
-
-The full time evolution of heat in the rod $\Omega=[0,1]$ is formally
-described by the following three equations:
-\begin{subequations}
-  \begin{alignat}{2}
-    u_t &= \Delta u &&\qquad \text{in } [0,1] \times [0,\infty), \label{eq:heat1}\\
-    u &= 0 &&\qquad \text{on } \{0, 1\} \times [0,\infty), \label{eq:heat2}\\
-    u &= u_0 &&\qquad \text{on }[0,1] \times \{0\}. \label{eq:heat3}
-  \end{alignat}
-\end{subequations}
-As time passes, heat dissipates across the rod as hotter regions,
-e.g.~local maxima of the heat distribution, become cooler. This
-behavior is captured by eq.~\eqref{eq:heat1}: the Laplacian at local
-maxima is negative, so $u_t = \Delta u$ implies $u$ decreases at its
-maxima, and the reverse happens at local minima. Eq.~\eqref{eq:heat2}
-describes the \emph{boundary condition}, which dictates how heat
-interacts with the outside world at the two edges of the rod
-$\{0,1\}$. Specifically, eq.~\eqref{eq:heat2} implements an absorbing
-boundary, i.e.~heat that interacts with the boundary immediately
-disappears. This type of boundary condition is known as a
-\emph{homogeneous Dirichlet boundary condition}. Eq.~\eqref{eq:heat3}
-describes the \emph{initial condition}, i.e.~the heat distribution
-inside the rod at time $t=0$.
-
-Solving the 1D heat equation is straightforward. Any initial heat
-distribution $u_0\in L^2([0,1])$ that satisfies the homogeneous
-Dirichlet boundary condition eq.~\eqref{eq:heat2} is a linear
-combination of sines $\ev_n(x) = \sin(\pi n x)$, so $u_0 = \sum_{n\geq
-  1} a_n \ev_n$ for some $\{a_n\}_{n\geq 1}$. Now, note that $\ev_n$
-are in fact eigenvectors of the Laplacian: $\Delta \ev_n = -\pi^2
-n^2\ev_n$. It should not be hard to believe that $u(\cdot, T) =
-\sum_{n\geq 1} a_n \exp(-\pi^2 n^2T ) \ev_n$
-\cite{renardy2006PDE}. Thus, eigenvectors of the linear operator that
-describes the time evolution of heat are $\ev_n(x) = \sin(\pi n x)$
-with corresponding eigenvalues
-\begin{equation}\label{eq:decay}
-  \exp(-\pi^2 n^2T ).
-\end{equation}
-
-In the \emph{inverse problem of the 1D heat equation}, our goal is to
-infer the initial condition $u_0 = u(\cdot, 0)$ from observations of
-the final state $u(\cdot, T)$. However, we are not able to measure
-$u(\cdot, T)$ at every point in the domain $\Omega$. Rather, we take
-some number of temperature measurements on the rod once our waiting
-time $T$ has passed and try to infer $u(\cdot, 0)$ from these
-measurements.
-
-The inverse problem of the 1D heat equation is difficult ("ill-posed")
-since heat spreads in a diffusive manner and any roughness in the
-initial condition is quickly smoothed, as implied by the squared
-exponential decay of eigenvalues in eq.~\eqref{eq:decay} with
-$n\to\infty$. See Supplementary movies S1 and S2 for an illustration
-of this phenomenon (code generating these movies is located in module
-\texttt{movies.py} in the accompanying
-\href{https://github.com/yairdaon/OED}{repository}). For ill-posed
-problems regularization is required, and we implement such
-regularization via a fully Bayesian formulation of the inverse
-problem.
-
-We choose a prior for the initial condition $u_0 \sim \mathcal{N}(0,
-(-\Delta)^{-1})$, with homogeneous Dirichlet boundary condition. This
-choice of prior covariance operator ensures the posterior is
-well-defined and prior realizations are well-behaved, see
-\cite[Theorem 3.1 and Lemma 6.25]{Stuart10} for details. The second
-ingredient of a fully Bayesian formulation of the inverse problem is
-likelihood. We simply assume centered iid Gaussian measurement error
-model, giving rise to a the widely applicable Gaussian likelihood
-function.
-
-\subsection{Measurement Clusterization}
-Surprisingly, A- and D-optimal designs for inverse problems have been
-observed to yield remarkably similar measurements in certain cases
-\cite{fedorovDesignSpatialExperiments1996, nyberg2012, fedorov1997,
-  Ucinski05, neitzel2019sparse}. This phenomenon is illustrated for
-our toy inverse problem in Fig.~\ref{fig:clusterization_illustration},
-where D-optimal measurement locations are shown for different numbers
-of measurements. Notably, for six measurements, a D-optimal design
-yields two sets of measurements that are identical. Following
-\cite{Ucinski05}, we refer to this intriguing phenomenon as
-\emph{measurement clusterization}. We consider a design to be
-\emph{clustered} when two or more of its constituent measurements are
-identical.
-
-\begin{figure}
-    \centering
-    \includegraphics[height=0.5\textwidth]{figs/dst_modelError0.png}
-    \caption{Measurement clusterization in D-optimal designs for the
-      inverse problem of the 1D heat equation. Measurement locations
-      were chosen according to the Bayesian D-optimality criterion of
-      Theorem \ref{thm:d_optimality}. Measurement locations are
-      plotted over the computational domain \(\Omega = [0, 1]\)
-      (x-axis), for varying numbers of measurements (y-axis). The
-      colored numbers are measurement indices, plotted for visual
-      clarity. Measurement clusterization already occurs for three
-      measurements: the second measurement (red) is overlaid on the
-      third (green). For five measurements, first (blue) and second
-      (red) measurements are clustered, as well as the fourth (black)
-      and the fifth (magenta).}
-  \label{fig:clusterization_illustration}
-\end{figure}
-
-Clusterization should not be confused with either repetition nor with
-replication, which are commonly viewed as beneficial and even
-necessary aspects of an optimal design \cite{fisher1949design,
-  morris2011, schafer2001replication}. For example,
-\cite{fisher1949design} in his famous milk and tea experiment,
-suggested that repetition is a "way of enlarging the experiment and,
-thereby, increasing its sensitiveness". In another example,
-\cite{fay2000rainfall} measured the effect of rainfall on grass growth
-in plots of land. The experiment involved fifteen "rainfall
-manipulation shelters", where "Four rainfall manipulation treatments
-(three replicates) then were assigned to 12 of the plots". While it
-seems reasonable for the researchers to replicate the phenomenon they
-are trying to study, clusterization is different: a clustered design
-in the rainfall experiment would imply the researchers should take a
-repeated measurement \emph{on the same plot}, at the expense of
-measuring grass growth in other plots. In sharp contrast to repetition
-and replication, clusterization is highly nonintuitive and we will see
-that its origins run considerably deeper.
-
-Researchers of inverse problems widely agree that measurement
-clusterization is undesirable
-\cite{fedorovDesignSpatialExperiments1996, nyberg2012, fedorov1997,
-  Ucinski05, neitzel2019sparse}, prompting the exploration of various
-remedies to address this issue. One approach involves merging close
-measurements \cite{fedorov1997}; however, this strategy merely
-overlooks the phenomenon of measurement clusterization. An alternative
-solution lies in \emph{clusterization-free design}s, where measurement
-locations are deliberately chosen to be distant from one another. This
-can be achieved by imposing distance constraints between measurements
-or by introducing correlated errors that account for both observation
-error and model misspecification \cite{Ucinski05}. For instance, in
-the context of time-series analysis for pharmacokinetic experiments,
-measurement clusterization can be mitigated by incorporating the
-modeling of auto-correlation time within the noise terms
-\cite{nyberg2012}.
-
-In spatial problems involving choice of measurements within a domain
-\(\Omega \subseteq \mathbb{R}^d, d=1,2,3\), many researchers
-circumvent the problem of measurement clusterization by restricting
-measurements to a coarse grid in \(\Omega\) \cite{koval2020,
-  alexanderian2021, attia2022, alexanderian2014, alexanderian2016,
-  alexanderian2018efficient, brunton2016}. This approach incurs a
-significant computational cost as it requires solving a difficult
-combinatorial optimization problem for measurement locations over a
-discrete set. The combinatorial optimization problem is usually
-relaxed by first assigning optimal measurement weights in
-\(\mathbb{R}_+\) to the potential measurement locations. Some
-researchers incorporate a sparsifying \(\ell_1\) penalty term into the
-design criterion, which is subsequently thresholded to achieve the
-desired binary design over the coarse grid
-\cite{horesh2008borehole}. Others progressively relax the \(\ell_1\)
-penalty to an \(\ell_0\) penalty via a continuation method
-\cite{alexanderian2016, alexanderian2014}. Others cast the problem of
-finding optimal measurement weights as a stochastic optimization
-problem \cite{attia2022stochastic}. All of the aforementioned methods
-may indeed find a binary optimal design restricted to a given coarse
-grid. However, none addresses one fundamental issue: the restriction
-of measurement locations to a coarse grid in \(\Omega\) fundamentally
-changes the optimal design problem and thus results in a sub-optimal
-design.
-
-Avoiding measurement clusterization is a pragmatic approach:
-intuitively, researchers recognize that measurement clusterization is
-undesirable, even though the underlying reasons may not be fully
-clear. Consequently, they strive to prevent it and devise various
-methodologies to avoid it. Yet each and every one of these
-methodologies achieves its objective by imposing restrictions on
-measurement locations, thereby fundamentally altering the optimal
-design problem. To the best of my knowledge, no previous study has
-tried to address some seemingly simple yet fundamental questions:
-%
-Why does imposing correlations between observations alleviate
-measurement clusterization?
-%
-Is measurement clusterization a generic phenomenon? 
-%
-And, most importantly: Why does measurement clusterization occur?
-%
-%Should we aim to avoid measurement clusterization?
-%
-%Is it possible to substitute an optimal clustered design with an
-%equally optimal non-clustered design?
-%Can an optimal clustered design be replaced with an equally optimal
-%non-clustered design?
-
-\subsection{Contribution}
-The primary objective of this study is to provide a deep understanding
-of measurement clusterization by addressing the aforementioned
-questions. Our focus centers around investigating the Bayesian
-D-optimality criterion. We conduct an analysis of Bayesian D-optimal
-designs within the context of linear inverse problems over Hilbert
-spaces and study two inverse problems: (a) In Sections
-\ref{section:prelim} and \ref{section:D_and_grad} we propose a novel
-generic model for an inverse problem where D-optimality maintains
-analytical tractability and D-optimal designs are identified via
-Lagrange multipliers. This analytical framework facilitates the
-exploration of the questions posed at the end of the previous
-paragraph. We also study (b) the inverse problem of the 1D heat
-equation from Section \ref{subsec:toy} above. Investigating both
-inverse problems allows us to answer the questions posed in the
-previous section:
-
-\begin{enumerate}
-\item \label{q:generic} \textbf{Is measurement clusterization a
-  generic phenomenon?}
-  %\subsection{An answer for Question \ref{q:generic}: Genericity of measurement clusterization}
-  %% Computer implementation of Lemma \ref{thm:char} for the inverse
-  %% problem outlined in Section \ref{section:how} also generates \(\obs\)
-  %% as a solution to the D-optimal design problem.  Furthermore,
-  We give two complementing answers to this question. First, from a
-  theoretical perspective, we show that clusterization mainly depends
-  on how quickly the eigenvalues of the prior covariance in
-  observation space decay. This decay depends mostly on how ill-posed
-  the problem at hand is and does not depend much on the prior. See
-  Section \ref{section:vanishing} --- particularly, Theorem
-  \ref{thm:char} and the discussion following it. We also show results
-  of numerical experiments, where simulations of our model give rise
-  to D-optimal designs that exhibit clusterization with high
-  probability in Section \ref{subsec:lemma_sims}. Thus, given the
-  genericity of our model, we expect measurement clusterization to be
-  a generic and ubiquitous phenomenon.
-
-\item \label{q:mitigate} \textbf{Why does imposing correlations
-  between observations alleviate measurement clusterization?} In
-  Section \ref{section:non_vanishing}, we rigorously demonstrate the
-  role of model error in mitigating clusterization, thereby
-  corroborating earlier observations made by various
-  researchers. Specifically, our proof shows that identical
-  measurements result in no gain in design criterion when observation
-  error amplitude tends to zero. Moreover, in Section
-  \ref{subsec:corr_errors_sims}, we show that an error term
-  corresponding to correlations between measurements mitigates
-  clusterization in the inverse problem of the 1D heat equation.
-
-\item \label{q:why} \textbf{Why does measurement clusterization
-  occur?} In Section \ref{subsec:why} (specifically, Fig~\ref{fig:why}
-  and the discussion following it) we give a compelling answer to this
-  question by transporting insights we gain from our generic model to
-  the inverse problem of the 1D heat equation. Our analysis reveals
-  that a D-optimal design focuses on a select set of prior
-  eigenvectors, specifically those with the largest eigenvalues in the
-  prior covariance spectrum. In practical scenarios, the number of
-  locations where (a) all relevant prior eigenvectors are
-  significantly large, and (b) other eigenvectors are close to zero,
-  is limited. Consequently, the clusterization of measurements arises
-  as a natural consequence of the pigeonhole principle, as there are
-  more measurements available than there are locations satisfying
-  conditions (a) and (b).
-
-\end{enumerate}
-
-\subsubsection{Implications}
 Our answer to Question \ref{q:generic} implies that encountering
@@
 It is important to note that we do not view clustered designs as
-undesirable, nor do we believe it should be avoided at
-all. Clusterization is a peculiar phenomenon and it is perfectly
-reasonable for someone to argue against the D-optimality criterion
-based on the fact that it results in clustered designs. We have seen,
-however that there is a perfectly reasonable explanation for
-clusterization. We have shown that clusterization is an inevitable
-consequence of having a problem with some modes where uncertainty
-decays faster than others.
+undesirable, nor do we believe D-optimal designs should be avoided at
+all. Nothing in our analysis indicates that we should expect any
+pathological behavior when utilizing D-optimal designs. On the
+contrary: we show that D-optimal designs (clustered or not) reduce
+uncertainty of prior covariance eigenvectors where it is
+highest. Thus, in a sense we hope to make precise in a future study,
+we expect convergence of for D-optimal designs to be fastest.
+
+For example, \cite{tekentrup2020} provides convergence analysis for
+various designs with differing space filling properties. She showed
+that convergence rate depends on how these designs fill space. We
+expect better convergence rates for D-optimal designs; Even though
+D-optimal desgins demonstrate clusterization, they do so because they
+explore space in a way tailored to the problem studied. Consequently,
+D-optimal designs cluster because they first aim to measure what
+matters most; see the discussion following Theorem \ref{thm:char} for
+details.
+
+We expect D-optimal designs to judicious exploration of
+design space, as implemented by a D-optimal design to give better
+convergence compared to randomly sampling the design
+space \cite{knapik2011}.
+
+Clusterization is a peculiar phenomenon and it is perfectly reasonable
+for someone to argue against the D-optimality criterion based on the
+fact that it results in clustered designs. We have seen, however that
+there is a perfectly reasonable explanation for clusterization. We
+have shown that clusterization is an inevitable consequence of having
+a problem with some modes where uncertainty decays faster than others.
 
@@
 issue of clusterization.
-
-\subsubsection{Other Contributions}
-In Theorem \ref{thm:char}, we also show that D-optimal designs are
-best understood in the space of \emph{observations}, see Section
-\ref{section:vanishing} for a precise statement. This is in accordance
-with previous work by \cite{koval2020}, who showed that A-optimal
-designs are best constructed in the space of observations.
-
-In the process of proving Theorem \ref{thm:char} we prove and
-generalize several lemmas. Among those, is Lemma \ref{lemma:free},
-which is (to the the best of my knowledge) novel: We decompose a
-symmetric positive definite matrix \(M \in \mathbb{R}^{k \times k}\)
-with \(\ttr M = m \in \mathbb{N}\) as \(M = AA^t\), where \(A\in
-\mathbb{R}^{k \times m}\) has unit norm columns.
-
-\subsection{Limitations}\label{subsec:limitations}
-The main limitation of this study is that our generic model does not
-correspond to any specific real-life problem. Specifically, in its
-current form, our model does not allow point evaluations. Thus, while
-our model is generic enough to be analytically tractable, some may
-argue that our model is too far removed from any real application. To
-these claims I would answer that scientists have a long history of
-studying models that are bare-bones simplifications of real systems,
-e.g.~the Ising model \cite{cipra1987}, the Lorenz system \cite{brin},
-the Lotka-Volterra equations \cite{logan2006}, the Carnot engine
-\cite{kardar2007} and many others.

### 2024-08-12 latex/intro.tex
add eic and AE

@@
 
-  The connection to Carath\'eodory's Theorem also utilizes the
-  concentration of measurement effort on the abovementioned select set
-  of prior eigenvectors. This concentration allows us to move from an
+  The connection to Carath\'eodory's Theorem also builds on the focus
+  of measurement effort to the abovementioned select set of prior
+  eigenvectors. This concentration allows us to move from an
   infinite-dimensional setting to finite dimensions, where the
@@
 \subsubsection{Implications}
-Our answer to Question \ref{q:generic} implies that encountering
-clusterization should be expected in many different problems across
-many different scientific fields. Researchers that encounter
-clusterization should not be surprised or wary. In Our answer to
-Question \ref{q:why}, we explain what our view of the cause of
-clusterization is. It appears, the cause is generic: a D-optimal
-design reduces uncertainty for a select set of prior covariance
-eigenvectors --- those with the most prior uncertainty, i.e.~those the
-practitioner cares about the most! We believe practitioners should not
-try to avoid measurement clusterization. Rather, practitioners should
-take repeated measurements (e.g.~in MRI and borehole tomography),
-increase apparatus sensitivity (e.g.~in EIT), or take consecutive
-measurements (e.g.~in the 1D heat equation). Overall, we show that
-measurement clusterization is a natural and (almost) inevitable part
-of Bayesian D-optimal designs (but see disclaimer below).
-
-One interesting implication of the analysis presented here is that
-clusterization can serve as an evidence to the number of relevant
-eigenvectors. Since leading eigenvectors typically correspond to slow
-variations in space and/or time, clusterization could be used to
-estimate the number of relevant degrees of freedom, and even to reduce
-the complexity of a computational model, e.g.~by dropping
-discretization points.
-
-It is important to note that we do not view clustered designs as
-undesirable, nor do we believe it should be avoided at
-all. Clusterization is a peculiar phenomenon and it is perfectly
-reasonable for someone to argue against the D-optimality criterion
-based on the fact that it results in clustered designs. We have seen,
-however that there is a perfectly reasonable explanation for
-clusterization. We have shown that clusterization is an inevitable
-consequence of having a problem with some modes where uncertainty
-decays faster than others.
-
-Lastly, we believe that when clusterization arises, it should serve as
-a warning sign to practitioners. In the inverse problem of the 1D heat
-equation, clusterization occurs primarily because Laplacian
-eigenvectors \emph{do not} decay in $\Omega$. Consequently, measuring
-$u(x_1, T)$ at some point $x_1 \in \Omega$ provides information about
-$u(x_2,T)$ for distant points $x_2 \in \Omega$. Intuitively, this
-should not occur: for small $T$, the heat distribution at $x_1$ should
-give very little knowledge on the heat distribution at $x_2$. This
-behavior stems from a well-known property of the heat equation: it
-allows information to spread \emph{instantly} across the computational
-domain \cite{renardy2006PDE}. In reality, heat (and information)
-propagate at finite speeds. Of course, the known physical barrier for
-information spread is the speed of light, but we expect heat to spread
-considerably slower: heating an Olympic pool at one end should have no
-immediate effect on the temperature at the other end.
-
-Our choice of prior is also a potential major contributor to
-clusterization. Our choice of Gaussian prior similarly implies
-information is shared between distant locations in $\Omega$. Thus, we
-suggest refraining from choosing Gaussian priors with inverse
-Laplacian covariance operators. Rather, non-Gaussian priors could be
-employed instead \cite{hosseini2017, hosseini2019}.
-
-We believe that the emergence of clusterization in this context is
-thus non-physical, arising from the way the inverse problems we
-consider are phrased. Clusterization therefore indicates that the
-underlying mathematical / Bayesian model is overly permissive and
-fails to capture crucial physical constraints of the problem. We
-suggest that when clusterization occurs, practitioners should consider
-alternative models where information is localized in space and travels
-at finite speed in the medium. Such models may not only provide more
-physically accurate and meaningful results but may also mitigate the
-issue of clusterization.
+\input{implications}
 

### 2024-08-12 latex/refs.tex
add eic and AE

--- 2024-07-24 latex/refs.tex
@@
-\documentclass{amsart}
+\documentclass{article}
+\def\papertitle{Measurement Clusterization in Bayesian D-optimal Designs in Infinite Dimensions}
+\def\authors{Yair Daon}
+\def\journal{Bayesian Analysis}
+\def\doi{12345}
+
+% Define title defaults if not defined by user
+\providecommand{\lettertitle}{Author Response to Reviews of}
+\providecommand{\papertitle}{Title}
+\providecommand{\authors}{Authors}
+\providecommand{\journal}{Journal}
+\providecommand{\doi}{--}
+
+\input{preamble}
 \input{definitions}
+\usepackage{xr}
+\externaldocument{ba}
 
@@
 
+% Make title
+{\Large\bf \lettertitle}\\[1em]
+{\huge \papertitle}\\[1em]
+{\authors}\\
+{\it \journal, }\texttt{doi:\doi}\\
+\hrule
+
+% Legend
+\hfill {\bfseries RC:} \textbf{\textit{Reviewer Comment}},\(\quad\) AR: \emph{Author Response}, \(\quad\square\) Manuscript text
+
+I appreciate the many thoughtful comments by all of you. I believe
+these comments helped me considerably improve the exposition of my
+paper and added new insights as well. I sincerely appreciate the
+option to revise and resubmit. Please see below my answers to the
+comments made by the referees, associate editor and editor in chief:
+\rred{My answers are marked in red.} \bblue{In some places I copied
+  text from the revised manuscipt and marked it in blue.}
+
 \section{Editor in Chief}
-Personally I like the paper, but it represents a somehow unusual
+\RC Personally I like the paper, but it represents a somehow unusual
 submission for Bayesian Analysis, and I am worried about it being of
@@
 
-\begin{description}
-  \item[a] the Introduction is to be made accessible to the whole BA
+\RC the Introduction is to be made accessible to the whole BA
     readership;
-  \item[b] the motivation and the practical statistical implications are to be
-    developed and discussed;
-  \item[c] the contribution has to be better contextualized with
-    respect to the literature in both inverse problems and D-optimal
-    designs;
-  \item[d] numerical evidence and code, mentioned in the paper
-    but actually not provided, need to be made available;
-  \item[e] the mathematical derivations have to be watertight and some
-    clarifications are needed in this respect (see Report of Referee
-    3).
 
-\end{description}
+\AR  I have considerably extended the introduction.
+
+\RV the motivation and the practical statistical implications are to
+be developed and discussed;
+
+\AR I added a section on the implications:
+\begin{quote}
+  \input{implications}
+\end{quote}
+
+\RC the contribution has to be better contextualized with respect to
+the literature in both inverse problems and D-optimal designs;
+
+\RC numerical evidence and code, mentioned in the paper but actually
+not provided, need to be made available;
+
+\AR I added a section describing numerical experiments and their
+results. I also added documentation to the accompanying
+\href{https://github.com/yairdaon/OED}{repository} and made sure all
+scripts there run.
+
+\RC the mathematical derivations have to be watertight and some
+clarifications are needed in this respect (see Report of Referee 3).
+
+\AR I am not sure which specific comment of Reviewer 3 this
+      refers to, please see my answers to Reviewer 3 below.
 
 \section{Associate Editor}
-The three referees found merits and interest in your work, but also
+\RC The three referees found merits and interest in your work, but also
 pointed out several major issues to be addressed. Of particular
@@
 
-A referee makes valid points as to whether clusterization of
+\AR I have added considerably to the introduction. See revised
+manuscript and diff file.
+
+\RC A referee makes valid points as to whether clusterization of
 measurements is a problem in general, when in fact in can be an
 appealing feature in some settings, and another referee asks whether
-D-optimal designs should be abandoned altogether. The exposition needs
-to discuss these high-level issues, and introduce the related
+D-optimal designs should be abandoned altogether. The exposition
+needs to discuss these high-level issues, and introduce the related
 literature and results appropriately, to give the work proper
-context. Altogether, I think that your work has the potential to make
-a good contribution to Bayesian analysis, but as it currently stands
+context.
+
+\RC Altogether, I think that your work has the potential to make a
+good contribution to Bayesian analysis, but as it currently stands
 significant work is needed.
 
-\section{Reviewer 1}
-This paper explains a common practical issue: the cause of
-clusterization in Bayesian D-optimal designs in infinite dimensions
-and why adding a correlated measurement process decluster the design
-points. Its theoretical results are valuable to the field of inverse
-problems/uncertainty quantification and the conclusions should be
-useful to practitioners in this field.
+\input{rev1}
+\input{rev2}
+\input{rev3}
 
-\answer: I appreciate the kind words.
-
-However, I feel changing the assumption from independent to correlated
-measurements is quite artificial: people tweak a data generation
-process assumption to solve an optimal design problem, and the author
-just takes the assumption as it is. Since the author proves the
-clusterization issue under the independent measurement assumption, I
-expect the author to go deeper and explain why an ordinary data
-generation assumption + a common optimal design leads to a
-“suboptimal” design? Not just by the mathematical proof, but from an
-information theory point of view. For example, does the redundancy
-points mean additional points does not reduce the variance of the
-estimation at all? If that’s the case, how does correlated measurement
-assumption “help” with it?
-
-\answer: Thanks for your questions. I am not a proponent of using
-correlated noise, I just wanted to show the usefullness of my model by
-rigorously proving how correlated noise mitigates clusterization. I
-have added several lines clearing this issue.
-
-\section{Reviewer 2}
-\subsection{Measurement Clusterization vs Repetition}
-I will start by saying measurement clusterisation is not well-defined
-in this paper. On page 1, it is referred to as two sets of
-measurements that are almost indistinguishable from one
-another. However, the setup for Corollary 11 implicitly defines
-measurement clusterisation as repeated measurements. The latter
-definition matches the references of Fedorov (1996) and Nyberg et al
-(2012). So I will assume for the rest of this review that, measurement
-clusterisation refers to repeated measurements.
-
-\answer:
-
-Nearly all textbooks on classical design of experiments (e.g. Morris,
-2011, Section 1.2.4) will feature a section with the three words
-randomisation, blocking and repetition: the important principles of
-design of experiments.  In these texts, repeated measurements are
-considered to be a beneficial property of a design. Interestingly,
-when optimal designs are found, they are often found to have repeated
-measurements. The paper does not explain why the author thinks
-repeated measurements (measurement clusterisation) is an undesirable
-property.
-
-\answer: I do not view clusterization as an undesired property, I only
-mentioned many authors consider it to be undesirable. I have added a
-section contrasting repetition and replication with clusterization and
-why I believe they may be different.
-
-If one looks at, for example, Fedorov (1996) and Nyberg et
-al (2012), it is an undesirable property because they are interested
-in sensor location, and two or more sensors cannot be in the same
-location. However, there are many experiments where repeated
-measurements are possible, and are therefore optimal (if the model is
-correct, which it won’t be). An example, is a chemical engineering
-experiment where one wishes to learn the relationship between a series
-of variables and yield of a chemical reaction. The experiment will
-involve specifying the variables, and measuring the yield (the
-observation) after a specified period of time. This process is then
-duplicated with a potentially different set of variables.  It is
-perfectly possible to have repeated variables.
-
-\answer: again
-
-\begin{enumerate}
-
-\item The author needs to make clear that they are considering
-  experiments where repeated measurements are impossible. Admitedly,
-  this consideration is implicit in the paper, for example, the
-  examples given on page 1. However, it needs to be made explicit in
-  the Abstract and the first paragraph of the paper.
-
-  \answer: I do consider experiments where repeated measurements are
-  possible. I am giving insight on why such designs are useful, even
-  though repeated measurements may be non-intuitive in certain
-  settings.
-
-\item The paper explains why repeated measurements can be optimal
-  (answer to Question 3). The author should expand on this and draw
-  the link with how repeated measurements are, typically, regarded as
-  desirable in classical design of experiments for finite
-  dimensions. See for example, the discussion of the relationship
-  between Caratheodory’s Theorem and the number of support points in
-  Pronzato \& Pazman (2013, page 139). The support points are the
-  unique design points, so if this is less than m, then there are
-  repeated design points.
-
-  \answer: great comment
-\end{enumerate}
-
-\subsection{Context}
-The paper lacks discussion on context. I read the answers to the three
-questions and thought “So what?”  What are the implications to
-statistical practice, or experimental science, of these findings?  The
-presentation of the paper does not help. The results are presented
-with no explanation or intuition. It might be better to relegate much
-of the mathematical detail to Supplementary Material and use the main
-manuscript for exposition.
-
-\subsection{Avoiding measurement clusterisation}
-The approaches to avoid measurement clusterisation, described on page
-3, are described as pragmatic and "fundamentally altering the optimal
-design problem".
-
-Firstly, this is not strictly true. Suppose measurements correspond to
-time, it is not possible to take two of more observations at the same
-time, and there should be a minimum time period between measurements.
-Implementing this minimum time period is not pragmatic, rather it is
-characterising the application correctly.  One could imagine a similar
-thing occurs with the eletrode locations on the skin in impedence
-tomography: the physical size of the sensor imposes a miniumum
-distance between sensors.
-
-Secondly, the paper seems to suggest “imposing correlations between
-observations” as a solution. However, isn’t this imposition pragmatic
-as well? I suppose it is argued that the $\obs \eps′$ term is
-accounting for model error: but it is still pragmatic: the data
-acquisition given by equation (1) is not the true data-generating
-process.
-
-\subsection{Minor comments}
-\begin{itemize}
-\item The author uses two different forms of asterisk: one for
-  optimality of the design (e.g. Definition 3) and one for the
-  adjoint. Suggest changing the one for optimality.
-\item I do not like the use of the term measurements as used in the
-  paper. The measurements are the $\data$’s since these are a result of
-  measuring a physical quantity. However, the paper refers to the
-  quantities that are controlled as the measurements.
-\item Proof of Proposition 12 uses $\eps$ as part of the
-  regularisation trick but this has been used previously for error
-  (e.g. equation (1)).
-\end{itemize}
-
-\section{Reviewer 3}
-The paper studies Bayesian D-optimal design (that is, the task of
-allocating measurements in order to maximise the Kullback-Leibler
-divergence between the prior and posterior disribution) in an inverse
-regression type model in Hilbert spaces with Gaussian measurements
-errors. For Gaussian priors, the paper provides a characterisation of
-the optimality criterion and the optimal design. Some conclusions
-about the phenomenon of clusterisation of measurements known in
-literature are then drawn.
-
-While the results in the paper provide some novel insights on the
-topic, I believe the current version of the manuscript is far from the
-standard required by Bayesian Analysis. Let me raise immediately my
-main concern: on p.4 the author describes the generality of
-measurements clusterisation as the first main question being explored,
-which is answered via, quoting, "randomized numerical simulations that
-exhibit clusterization more than 95\% of the time (see the code in
-supplementary material)". However, in the version of the supplementary
-materials I had access to there is no further mention of these
-numerical simulations and results; only a link to a Github repository
-is provided, which is however not functioning. Evidently, this issue
-needs to be fixed before any future consideration can be given to the
-manuscript. For example, a detailed description of the repeated
-experiments could be provided either in the main article or in the
-supplement, with tables indicating the various values of the
-parameters for the simulations and the precise percentages
-obtained. Also, the code needs to be publicly available for
-reproducibility.
-
-\subsection{Further comments}
-I have a number of further comments in regards to the results and the overall presentation of
-the manuscript.
-\begin{itemize}
-\item D-optimal design vs. statistical recovery rates: being more
-  familiar with the statistical inverse problem literature than with
-  D-optimal design, the clusterisation phenomenon demonstrated in this
-  paper raises the question as to wether consistent recovery of the
-  unknown (m in the notation of this paper) is possible under this
-  choice. All the results I am aware of either assume white
-  noise/equally spaced design (e.g. [1]) or measurement loctions
-  sampled uni- formly at random (e.g. [2]). There are also some
-  results with more general design under conditions on the
-  fill-distance of the grid (e.g. [3]). All of these specifically
-  prevent clusteri- sation, which is key to the statistical
-  analysis. Hence, it is not clear to me even if D-optimal design
-  should be pursued at all, if it indeed it leads to clusterisation as
-  implied by the present 1paper. I think a discussion on this would be
-  illuminating for the reader and help draw a connection to the
-  broader statistical inverse problems literature.
-
-\item Figure 1: It is hard to descern whether the measurements
-  locations are perfectly overlapping or just very closely
-  placed. Perhaps using dots with numbers here would help the
-  readability.
-
-  Answer: It is impossible to visually discern the measurements, even
-  when resolution is increased.
-
-\item Experiment with the 1D heat equation: the experiment showcases
-  some interesting phenomena, but also raises many questions. For
-  instance: I would expect the dimensionality of the working domain
-  (here the unit interval [0, 1]) to play a role, since larger
-  dimension allow for higher freedom in placing the design points.
-
-  Answer: That is a great point! I added the followng discussion:
-
-\item Further, a dimensionality effect should also arise directly from
-  Theorem 1 via the prior covariance eigenvalues: in the experiment
-  the prior covariance operator is set to be equal to the inverse
-  Laplace operator $(−\Delta)^{−1}$ . In d-dimensional domains, its
-  eigenvalues $\lambda_j$ are known to follow Weyl’s asymptotics,
-  growing as $j^{2/d}$, which should then impact the index after
-  which the eigenvalues are ‘thresholded’ by the procedure.
-
-\item Also, it should be relatively easy here to also study empirically the effect of the correlated
-error model in preventing clusterisation. Investigating all these issues in the context of the
-presented numerical set up would considerably enlarge the breath of the present work.
-
-\item I found very hard to understand Theorem 1 immediately at the end
-  of Section 1, as all the necessary background is introduced only
-  later. Since the same result is also stated and proved in Section 5,
-  I would consider cutting it from Section 1. The paragraphs in p.4
-  and 5 already do a good job presenting the results, and the
-  repetition of the formal statement does not seem necessary here.
-
-\item Denoting the optimal design in Theorem 1 by simply O is somewhat confusing, cf. the
-equation in the second item. Perhaps a separate notation such as Ō would help readability.
-
-\item Section 1.2 seems altogether unnecessary; the inverse regression model considered in the
-paper are a modelisation of many real-world phenomena, and they are routinely studied in
-statistical papers.
-
-Answer: Thanks for this comment but I would rather keep this section
-here since...
-
-\item What does ‘strongly smoothning’ operator means in the thid line of Section 2.1?
-
-\item Section 4: the conclusion reached at the end of the section on
-  p.15 seem to imply that in the noiseless limit, repeating any
-  measurement does not improve the objective criterion, if corre-
-  lated error model are present. Is it clear however that adding any
-  (non-reapted) measurement always strictly increases it? I.e., there
-  could be situations where adding one measurement does not change the
-  value of the optimised criterion? Also, can the conclusion drawn
-  here be extended to the case of noisy observations σ > 0?
-
-\item At the end of Section 4 on p.15 it is mentioned that $\tar$ is not
-  defined for $\sigma^2= 0$, but in the latter case, could not a Gaussian
-  posterior measure be still defined as in Gaussian process
-  regression, e.g.[4]
-
-\item Theorem 1 is expressed in terms of the prior covariance matrix,
-  which is a user specified quantity. Hence, it seems to me that all
-  sort of design behaviour under the D-optimal crite- rior can occur
-  by engineering the prior. It would perhaps be informative to study
-  in more details some representative example of inverse problem and
-  Gaussian prior. For example, the recovery of the initial condition
-  with the ‘Matérn-like’ Gaussian prior laid out in the sup- plement
-  are good candidates. For these it is still not clear how the
-  conclusion from Theorem 1 translates into the Figure 1, also because
-  I am unsure about the applicability of the result to point
-  evaluations. In particular, how the conclusion that with m = 4
-  measurements the D-optimal procedure will aim to ignore the third
-  and fourth eigenvalue/function was drawn?
-\end{itemize}
+\bibliographystyle{apalike}
+\bibliography{../../lib.bib}
 

### 2024-08-12 latex/rev1.tex
add eic and AE

--- 2024-08-12 latex/implications.tex
@@
-Our answer to Question \ref{q:generic} implies that encountering
-clusterization should be expected in many different problems across
-many different scientific fields. Researchers that encounter
-clusterization should not be surprised or wary. In Our answer to
-Question \ref{q:why}, we explain what our view of the cause of
-clusterization is. It appears, the cause is generic: a D-optimal
-design reduces uncertainty for a select set of prior covariance
-eigenvectors --- those with the most prior uncertainty, i.e.~those the
-practitioner cares about the most! We believe practitioners should not
-try to avoid measurement clusterization. Rather, practitioners should
-take repeated measurements (e.g.~in MRI and borehole tomography),
-increase apparatus sensitivity (e.g.~in EIT), or take consecutive
-measurements (e.g.~in the 1D heat equation). Overall, we show that
-measurement clusterization is a natural and (almost) inevitable part
-of Bayesian D-optimal designs (but see disclaimer below).
+\section{Reviewer 1}
+\RC This paper explains a common practical issue: the cause of
+clusterization in Bayesian D-optimal designs in infinite dimensions
+and why adding a correlated measurement process decluster the design
+points. Its theoretical results are valuable to the field of inverse
+problems/uncertainty quantification and the conclusions should be
+useful to practitioners in this field.
 
-One interesting implication of the analysis presented here is that
-clusterization can serve as an evidence to the number of relevant
-eigenvectors. Since leading eigenvectors typically correspond to slow
-variations in space and/or time, clusterization could be used to
-estimate the number of relevant degrees of freedom, and even to reduce
-the complexity of a computational model, e.g.~by dropping
-discretization points.
+\AR I appreciate the kind words.
 
+\RC However, I feel changing the assumption from independent to correlated
+measurements is quite artificial: people tweak a data generation
+process assumption to solve an optimal design problem, and the author
+just takes the assumption as it is.
+
+\AR If I understand correctly --- I completely agree. I added text
+emphasizing that I do not endorse using correlated errors:
+
+\begin{quote}
   It is important to note that we do not view clustered designs as
-undesirable, nor do we believe D-optimal designs should be avoided at
-all. Nothing in our analysis indicates that we should expect any
-pathological behavior when utilizing D-optimal designs. On the
-contrary: we show that D-optimal designs (clustered or not) reduce
-uncertainty of prior covariance eigenvectors where it is
-highest. Thus, in a sense we hope to make precise in a future study,
-we expect convergence of for D-optimal designs to be fastest.
+  undesirable, nor do we believe it should be avoided at
+  all. Clusterization is a peculiar phenomenon and it is perfectly
+  reasonable for someone to argue against the D-optimality criterion
+  based on the fact that it results in clustered designs. We have
+  seen, however that there is a perfectly reasonable explanation for
+  clusterization. We have shown that clusterization is an inevitable
+  consequence of having a problem with some modes where uncertainty
+  decays faster than others.
+\end{quote}
 
-For example, \cite{tekentrup2020} provides convergence analysis for
-various designs with differing space filling properties. She showed
-that convergence rate depends on how these designs fill space. We
-expect better convergence rates for D-optimal designs; Even though
-D-optimal desgins demonstrate clusterization, they do so because they
-explore space in a way tailored to the problem studied. Consequently,
-D-optimal designs cluster because they first aim to measure what
-matters most; see the discussion following Theorem \ref{thm:char} for
-details.
+\RC Since the author proves the clusterization issue under the independent
+measurement assumption, I expect the author to go deeper and explain
+why an ordinary data generation assumption + a common optimal design
+leads to a “suboptimal” design?
 
-We expect D-optimal designs to judicious exploration of
-design space, as implemented by a D-optimal design to give better
-convergence compared to randomly sampling the design
-space \cite{knapik2011}.
+\AR In my view, this is a result of the unlocalized structure of
+Laplacian eigenvectors. I believe this is more of a problem with our
+(linear) physical models and choice of (Gaussian) priors. See
+discussion below reproduced from the revised manuscript (but note that
+I do not endorse clustered designs)
 
-Clusterization is a peculiar phenomenon and it is perfectly reasonable
-for someone to argue against the D-optimality criterion based on the
-fact that it results in clustered designs. We have seen, however that
-there is a perfectly reasonable explanation for clusterization. We
-have shown that clusterization is an inevitable consequence of having
-a problem with some modes where uncertainty decays faster than others.
-
-Lastly, we believe that when clusterization arises, it should serve as
-a warning sign to practitioners. In the inverse problem of the 1D heat
-equation, clusterization occurs primarily because Laplacian
-eigenvectors \emph{do not} decay in $\Omega$. Consequently, measuring
-$u(x_1, T)$ at some point $x_1 \in \Omega$ provides information about
-$u(x_2,T)$ for distant points $x_2 \in \Omega$. Intuitively, this
-should not occur: for small $T$, the heat distribution at $x_1$ should
-give very little knowledge on the heat distribution at $x_2$. This
-behavior stems from a well-known property of the heat equation: it
-allows information to spread \emph{instantly} across the computational
-domain \cite{renardy2006PDE}. In reality, heat (and information)
-propagate at finite speeds. Of course, the known physical barrier for
-information spread is the speed of light, but we expect heat to spread
-considerably slower: heating an Olympic pool at one end should have no
-immediate effect on the temperature at the other end.
+\begin{quote}
+  Lastly, we believe that when clusterization arises, it should serve
+  as a warning sign to practitioners. In the inverse problem of the 1D
+  heat equation, clusterization occurs primarily because Laplacian
+  eigenvectors \emph{do not} decay in $\Omega$. Consequently,
+  measuring $u(x_1, T)$ at some point $x_1 \in \Omega$ provides
+  information about $u(x_2,T)$ for distant points $x_2 \in
+  \Omega$. Intuitively, this should not occur: for small $T$, the heat
+  distribution at $x_1$ should give very little knowledge on the heat
+  distribution at $x_2$. This behavior stems from a well-known
+  property of the heat equation: it allows information to spread
+  \emph{instantly} across the computational domain
+  \cite{renardy2006PDE}. In reality, heat (and information) propagate
+  at finite speeds. Of course, the known physical barrier for
+  information spread is the speed of light, but we expect heat to
+  spread considerably slower: heating an Olympic pool at one end
+  should have no immediate effect on the temperature at the other end.
 
@@
   suggest refraining from choosing Gaussian priors with inverse
-Laplacian covariance operators. Rather, non-Gaussian priors could be
+  Laplacian covariance operators. Rather, non-Gaussian priors should be
   employed instead \cite{hosseini2017, hosseini2019}.
@@
   issue of clusterization.
+\end{quote}
+
+\RC Not just by the mathematical proof, but from an information theory
+point of view. For example, does the redundancy points mean additional
+points does not reduce the variance of the estimation at all? If
+that’s the case, how does correlated measurement assumption “help”
+with it?
+
+\AR I think this is the same misunderstanding I commented on above;
+For any nonzero observation error $\sigma^2 > 0$, repeated
+measurements will indeed result in an increase of the design
+criterion.

### 2024-08-12 latex/rev3.tex
add eic and AE

\section{Reviewer 3}
\RC The paper studies Bayesian D-optimal design (that is, the task of
allocating measurements in order to maximise the Kullback-Leibler
divergence between the prior and posterior distribution) in an inverse
regression type model in Hilbert spaces with Gaussian measurements
errors. For Gaussian priors, the paper provides a characterisation of
the optimality criterion and the optimal design. Some conclusions
about the phenomenon of clusterisation of measurements known in
literature are then drawn.

\RC While the results in the paper provide some novel insights on the
topic, I believe the current version of the manuscript is far from the
standard required by Bayesian Analysis. Let me raise immediately my
main concern: on p.4 the author describes the generality of
measurements clusterisation as the first main question being explored,
which is answered via, quoting, "randomized numerical simulations that
exhibit clusterization more than 95\% of the time (see the code in
supplementary material)". However, in the version of the supplementary
materials I had access to there is no further mention of these
numerical simulations and results; only a link to a Github repository
is provided, which is however not functioning.

\AR I made sure the \href{https://github.com/yairdaon/OED}{repository}
is functioning and well-documented.

\RC Evidently, this issue needs to be fixed before any future
consideration can be given to the manuscript. For example, a detailed
description of the repeated experiments could be provided either in
the main article or in the supplement, with tables indicating the
various values of the parameters for the simulations and the precise
percentages obtained. Also, the code needs to be publicly available
for reproducibility.

\AR I appreciate the suggestion. I added a section reproducing the
abovementioned numerical experiments to the main article:

\begin{quote}
  In the proof of Theorem \ref{thm:char} we utilize Lemma
  \ref{lemma:free} to construct D-optimal designs. We implement this
  construction with the goal of testing numerically how prevalent are
  clustered designs. To this end, we would like to generate random
  prior eigenvalues $\lambda_j$, fix $m$ and $k$, find a D-optimal
  $\opt^*\opt$ and then utilize the construction of
  Theorem~\ref{thm:char} and Lemma~\ref{lemma:free} to find $\opt$.

  To simplify things, we directly sample rank $\opt^*\opt$. We iterate
  over the number of measurements $m \in \{4,\dots, 24\}$, and for
  every $m$ we then iterate over $k:=\rank \obs^*\obs \in \{2,\dots,
  m-1\}$. For each pair $m,k$ we repeat the following steps $N=5000$
  times:
  \begin{enumerate}
  \item Generate random diagonal $D\in \mathbb{R}^{k\times k}$ with
    entries $\log (d_i) \sim \mathcal{N}(50,15)$ and normalize so that
    $\ttr D = m$.
  \item Conjugate $D$ by a random orthogonal matrix to form a positive
    semi-definite $M := UDU^t \in \mathbb{R}^{k\times k}$. This $M$
    represents $\opt^*\opt$.
  \item Apply the construction of Lemma \ref{lemma:free} to calculate
    $A$ such that $AA^t = M$, where $A$ has unit norm
    columns. I.e.~find the optimal design $\opt$.
  \item Since $A$ corresponds to $\opt$, its columns correspond to
    measurement vectors. We call $A$ "clustered" if $A$ has two or
    more identical columns (up to some numerical precision threshold).
  \end{enumerate}
  We then calculate the fraction of clustered designs of the simulations
  we ran, for each pair $m,k$. Clusterization occurred at high rates
  ($>99.9\%$) whenever $m-k > 1$; see Fig.~\ref{fig:sim_AAt}. Hence, in
  these simulations, clusterization is a generic property. However, when
  $m-k = 1$, clusterization does not occur. We do not why this is so and
  further investigation into this phenomenon is out of scope for the
  current study.

  Full results are located in the \texttt{simulations.csv} file within
  the accompanying \href{https://github.com/yairdaon/OED}{repository}.
  Code implementing the experiments described above is located in module
  \texttt{zeros.py} of said repository. Runtime should be less than 30
  minutes on any reasonably modern laptop (it took 12 minutes on the
  author's laptop).
\end{quote}

\begin{figure}
    \centering
    \includegraphics[height=0.5\textwidth]{figs/simulations.png}
    \caption{Fraction of clustered $A$ for $AA^t = M$ and $M$
      generated randomly (see text and repository for details on
      generating $M$). It is evident that when $m-k \geq 2$ clusterization
      is ubiquitous, whereas for lower $m-k$ clusterization does not
      occur.}
  \label{fig:sim_AAt}
\end{figure}

\subsection{Further comments}
\RC I have a number of further comments in regards to the results and the
overall presentation of the manuscript.

\RC D-optimal design vs. statistical recovery rates: being more
familiar with the statistical inverse problem literature than with
D-optimal design, the clusterisation phenomenon demonstrated in this
paper raises the question as to whether consistent recovery of the
unknown ($\param$ in the notation of this paper) is possible under
this choice. All the results I am aware of either assume white
noise/equally spaced design (e.g.~\cite{knapik2011}) or measurement
locations sampled uniformly at random (e.g.~\cite{nickl2023}). There
are also some results with more general design under conditions on the
fill-distance of the grid (e.g.~\cite{teckentrup2020}). All of these
specifically prevent clusterisation, which is key to the statistical
analysis. Hence, it is not clear to me even if D-optimal design should
be pursued at all, if it indeed it leads to clusterisation as implied
by the present paper. I think a discussion on this would be
illuminating for the reader and help draw a connection to the broader
statistical inverse problems literature.

\AR I added a discussion on this subject. I do not argue against the
use of clustered designs imply that clustered designs, but I believe
that when they arise it should be a warning signal to a practitioner
regarding the validity of their physical model and Bayesian
assumptions:

\begin{quote}

\end{quote}

\RC Figure 1: It is hard to discern whether the measurements locations
are perfectly overlapping or just very closely placed. Perhaps using
dots with numbers here would help the readability.

\AR It is impossible to visually discern the measurements, even when
resolution is increased. These measurements are not identical
numerically, but they are identical to five decimal digits. Since I
find these points via optimization over point location in
$\Omega=[0,1]$ I view a numerical error $<10^{-5}$ as reasonable.

\RC Experiment with the 1D heat equation: the experiment showcases
some interesting phenomena, but also raises many questions. For
instance: I would expect the dimensionality of the working domain
(here the unit interval [0, 1]) to play a role, since larger dimension
allow for higher freedom in placing the design points.

\RC Further, a dimensionality effect should also arise directly from
Theorem 1 via the prior covariance eigenvalues: in the experiment the
prior covariance operator is set to be equal to the inverse Laplace
operator $(-\Delta)^{-1}$. In d-dimensional domains, its eigenvalues
$\lambda_j$ are known to follow Weyl’s asymptotics, growing as
$j^{2/d}$, which should then impact the index after which the
eigenvalues are ‘thresholded’ by the procedure.

\AR These are two great points. Indeed, the growth of eigenvalues of
the Laplacian is different in higher dimensions. This is why in higher
dimensions we would have to take as prior \(u_0 \sim
\mathcal{N}(\param_0, (-\Delta)^{-\gamma})\), for some \(\gamma >
d/2\) in order to maintain regularity of prior realizations. I decided
not to include a discussion on the choice of priors in higher spatial
dimensions because (1) it would require too much exposition and make
me digress from the main points I wanted to convey and (2) I view
implementing experiments in higher dimensions as out of the scope of
the current study.

\begin{quote}
  We choose a prior for the initial condition $u_0 \sim \mathcal{N}(0,
  (-\Delta)^{-1})$, with homogeneous Dirichlet boundary
  condition. This choice ensures the posterior is well-defined and
  prior realizations are well-behaved, see \cite[Theorem 3.1 and Lemma
    6.25]{Stuart10} for details.
\end{quote}

\RC Also, it should be relatively easy here to also study empirically
the effect of the correlated error model in preventing
clusterisation. Investigating all these issues in the context of the
presented numerical set up would considerably enlarge the breath of
the present work.

\AR I added relevant simulations to a new section Numerical
Experiments:

\begin{quote}
  In order to verify the results of Section
  \ref{section:non_vanishing}, we run simulations of the inverse
  problem of the 1D heat equation with nonvanishing model error
  \(\modcov = \prcov^2 \). Indeed, including model correlation pushes
  measurements apart, see Fig.~\ref{fig:corr_errors}. Code generating
  Fig.~\ref{fig:corr_errors} is located in module
  \texttt{clusterization.py} in the accompanying
  \href{https://github.com/yairdaon/OED}{repository}.
\end{quote}
\begin{figure}
  \centering
  \includegraphics[height=0.5\textwidth]{figs/dst_modelError4.png}
  \caption{Model correlation mitigates clusterization. We add a
    model correlation term to the error terms in the 1D heat
    equation inverse problem. Lo and behold, measurements are not
    close anymore and are pushed away thanks to the model error
    term.}
    \label{fig:corr_errors}
\end{figure}

\RC I found very hard to understand Theorem 1 immediately at the end
of Section 1, as all the necessary background is introduced only
later. Since the same result is also stated and proved in Section 5, I
would consider cutting it from Section 1. The paragraphs in p.4 and 5
already do a good job presenting the results, and the repetition of
the formal statement does not seem necessary here.

\AR I removed the statement of Theorem 1 from the introduction.

\RC Denoting the optimal design in Theorem 1 by simply $\obs$ is
somewhat confusing, cf. the equation in the second item. Perhaps a
separate notation such as $\bar{\obs}$ would help readability.

\AR Thank you for the comment, I completely agree. Optimal designs are
now denoted $\opt$.

\RC Section 1.2 seems altogether unnecessary; the inverse regression
model considered in the paper are a modelisation of many real-world
phenomena, and they are routinely studied in statistical papers.

\AR I am glad you view this part as unnecessary, but I feel the model
might feel too abstract to some practitioners. Thus, I choose to keep
this part to avoid complaints from other readers.

\RC What does ‘strongly smoothing’ operator means in the third line of Section 2.1?

\AR Basically I meant that its eigenvalues decrease "quickly". I
understand this is not a common or well-defined term so I removed the
relevant sentence.

\RC Section 4: the conclusion reached at the end of the section on
p.15 seem to imply that in the noiseless limit, repeating any
measurement does not improve the objective criterion, if correlated
error model are present.

\AR Indeed, this is very intuitive: when no observation error is
present, a repeated measurement does not tell us anything new.

\RC Is it clear however that adding any (non-repeated) measurement
always strictly increases it? I.e., there could be situations where
adding one measurement does not change the value of the optimised
criterion?

\AR Thanks for this insightful comment! Such cases are possible but I
view them as pathological. I added a short discussion reproduced
below:

\begin{quote}
  It is worth noting that by the nonnegativity of the KL divergence,
  $\tar$ cannot decrease upon adding measurements. However, we can
  construct examples where the posterior does not change upon taking a
  new measurement e.g.~if the prior variance vanishes on some
  eigenvector and a measurement is taken on said eigenvector. We do
  not expect a measurement to generate no information gain whatsoever
  in any realistic scenario, and ignore such pathologies.
\end{quote}

\RC Also, can the conclusion drawn here be extended to the case of
noisy observations $\sigma > 0$?

\AR I do not expect this conclusion to hold: if noise is
    present, even a repeated measurement will result in some
    information gain and an increase in design criterion, since the
    observation error is effectively reduced by a factor of
    $1/\sqrt{2}$ for the repeated measurement.

\RC At the end of Section 4 on p.15 it is mentioned that $\tar$ is not
defined for $\sigma^2= 0$, but in the latter case, could not a
Gaussian posterior measure be still defined as in Gaussian process
regression, e.g~\cite{rasmussen2006}].

\AR Indeed, one can define a posterior, so a KL divergence from said
posterior to prior could be calculated and $\tar$ evaluated. However,
taking $\sigma = 0$ in Theorem 1 (the theorem of
\cite{AlexanderianGloorGhattas14}) gives $\tar \equiv \infty$. I did
not study this subject further.

\RC Theorem 1 is expressed in terms of the prior covariance matrix,
which is a user specified quantity. Hence, it seems to me that all
sort of design behavior under the D-optimal criterion can occur by
engineering the prior. It would perhaps be informative to study in
more details some representative example of inverse problem and
Gaussian prior. For example, the recovery of the initial condition
with the ‘Mat\'ern-like’ Gaussian prior laid out in the supplement are
good candidates.

\AR the Mat\'ern-like prior actually arises by using a Laplacian-like
operator as a prior \cite{rue2011}. So a study of this example (with
and without correlated errors) is already included in the paper. I am
afraid implementing another inverse problem is out of scope for the
present paper since I was not able to install common packages for
inverse problems \cite{attia2023pyoed, VillaPetraGhattas16,
  VillaPetraGhattas18, VillaPetraGhattas21} on my machine.

\RC For these it is still not clear how the conclusion from Theorem 1
translates into the Figure 1, also because I am unsure about the
applicability of the result to point evaluations.

\AR Indeed, point evaluations $\delta_x$ are not in any function space
I consider in the paper since my analysis is limited to Hilbert
spaces. Point evaluations can be approximated well in
e.g.~$L^2(\Omega)$ so in my opinion this is not a deal breaker. The
only caveat is that upon approximating point evaluations e.g.~by
$R(x;t) = t\mathbb{1}_{-t/2, t/2}(x)$ we will need to change the norm
constraints to the something different (i.e.~$\|R(\cdot;t\|$) but that
does not change the analysis. I barely mentioned this issue in the
paper since I think it is confusing and adds little context. See,
however, a footnote I added which is reproduced below:

\begin{quote}
  The alert reader will likely ask how do we reconcile point
  measurements $\delta_x$ as suggested by the formulation of the 1D
  heat equation with working in Hilbert spaces. We don't. We follow
  standard practice in the literature and restrict our analysis to
  Hilbert spaces. We can satisfy ourselves with the fact that point
  evaluations could be approximated in a standard Hilbert space like
  $L^2(\Omega)$.
\end{quote}

\RC In particular, how the conclusion that with $m = 4$ measurements
the D-optimal procedure will aim to ignore the third and fourth
eigenvalue/function was drawn?

\AR I explain this in more detail in the paper:

\begin{quote}
  Building on Theorem \ref{thm:char}, we can now give a compelling
  explanation to the measurement clusterization we observed for the
  inverse problem of the heat equation (Fig.~\ref{fig:eigenvectors}).

  Consider $\fwd$ and $\prcov$ from \emph{the inverse problem of the
  heat equation}. As before, we denote the eigenvalues of
  $\fwd\prcov\fwd^*$ by $\lambda_j$. We input these eigenvalues into
  our \emph{generic} model, and find a D-optimal design $\opt$ for our
  generic model using Theorem \ref{thm:char}. In our generic model,
  the measurements we take are best utilized in reducing uncertainty
  for the first $k$ eigenvectors. So, a D-optimal design arising from
  our \emph{generic model} completely avoids measuring eigenvectors
  $k+1$ and above.

  Of course, in a real life problem --- such as the inverse problem of
  the 1D heat equation --- it is likely impossible to find measurement
  locations for which all eigenvectors $k+1$ and above are
  zero. However, if the eigenvalues of $\fwd\prcov\fwd^*$ decay
  quickly (recall the square-exponential decay for eigenvalues of the
  1D heat equation in eq.\eqref{eq:decay}), a D-optimal design will
  try to balance measuring a small number (i.e.~$k$) of the leading
  eigenvectors.

  The abovementioned balance is explored in
  Fig.~\ref{fig:eigenvectors}. We allow $m=4$ measurements in $\Omega
  = [0,1]$ and observe that D-optimal measurement locations are
  clustered at $x_1 = 0.31$ and $x_2 = 0.69$. Upon close inspection of
  the scaled eigenvectors of $\fwd \prcov \fwd^*$, we first observe
  that eigenvectors $3$ and above have negligible prior
  amplitude. Since we only have $m=4$ measurements at our disposal, we
  interpret these results, following Theorem \ref{thm:char}, as
  implying we should only care about measuring the first and second
  eigenvectors. Then, we note the D-optimal $x_1,x_2$ present a
  compromise between the amplitude of the first and second
  eigenvectors. For example, a measurement at $x=0.5$ would have
  ignored the second eigenvector altogether, since the second
  eigenvector is zero at $x=0.5$.

  Now we can understand measurement clusterization for the inverse
  problem of the heat equation. A D-optimal design attempts to measure
  the first $k$ eigenvectors of $\fwd \prcov \fwd^*$. But there may be
  (spatial) limitations on where these $k$ eigenvectors have large
  amplitude. For the inverse problem of the heat equation there are
  two spatial locations that present a good compromise between the
  amplitudes of the first and second eigenvectors, namely $x_1$ and
  $x_2$ --- see Fig.~\ref{fig:eigenvectors}. We have $m=4$
  measurements at our disposal but only two spatial locations that are
  a good compromise between the first and second scaled
  eigenvectors. Thus, clusterization arises as a consequence of the
  pigeonhole principle.

\end{quote}
\begin{figure}\label{fig:eigenvectors}
  \centering
  \includegraphics[width=\textwidth]{figs/eigenvectors_dst_scaled.png}
  \caption{D-optimal measurement locations ($m=4$ measurements) and
    weighted eigenvectors for finding the initial condition of the 1D
    heat equation. Measurement locations and weighted eigenvectors are
    plotted over the computational domain $\Omega = [0, 1]$
    (x-axis). Measurement clusterization occurs approximately at
    $0.31$ and $0.69$. These two locations are a compromise between
    the magnitudes of the first and second eigenvectors, which are the
    eigenvectors that a D-optimal design aims to measure. Allocating
    $m=4$ measurements into two locations results in clusterization,
    according to the pigeonhole principle.}
  \label{fig:why}
\end{figure}

### 2024-08-12 latex/vanishing.tex
add eic and AE

@@
 \subsubsection{Clusterization in our generic model}%\label{subsub:cara}
-We can gain further insight to clusterization in our generic model,
-from Carath\'eodory's Theorem and the concentration on the first $k$
-eigenvectors of $\fwd \prcov \fwd^*$. We present a short discussion
-adapting arguments presented in \cite[Chapter
-  3]{silveyOptimalDesign1980} and \cite[Section
-  5.2.3]{pronzatoPazman2013}.
-
-Consider $\opt$ a D-optimal design under our generic model.  As
-instructed by Theorem \ref{thm:char}, we ignore all but the first $k$
-eigenvectors of $\fwd \prcov \fwd$. Thus, we replace $\hilo$ with
-$\hilo^{(k)}$ --- the $k$-dimensional subspace spanned by the first
-$k$ eigenvectors of $\fwd^*\prcov\fwd$. Let
-\begin{equation*}
-  \mathcal{M} := \conv \{\meas \meas^* : \meas\in \hilo^{(k)}, \|\meas\|=1\},
-  %\hilo^{(k)}\},
-\end{equation*}
-where $\conv$ denotes the convex hull of a set. The set $\mathcal{M}$
-contains only positive-definite operators on a $k$-dimensional vector
-space. Hence $\mathcal{M}$ lives in a $k(k+1)/2$-dimensional vector
-space. Since $\opt^*\opt = \sum_{i=1}^m \meas_i\meas_i^*$ for $\meas_i
-\in \hilo^{(k)}$, it is easy to verify that $\frac1m \opt^*\opt \in
-\mathcal{M}$.  Recall Carath\'eodory's Theorem:
-\begin{theorem}[Carath\'eodory]
-  Let $X \subseteq \mathbb{R}^n, X \neq \phi$ and denote $\conv (X)$
-  the convex hull of $X$. For every $x \in \conv (X)$, $x$ is a convex
-  combination of at most $n+1$ vectors in $X$.
-\end{theorem}
-Carath\'eodory's Theorem implies that there exist $\meas_i$ and
-$\alpha_i$ such that
-\begin{equation*}
-  \opt^*\opt = \sum_{i=1}^I \alpha_i \meas_i\meas_i^*,
-\end{equation*}
-where $\|\meas\|=1, \sum\alpha_i = m, \alpha_i \geq 0$ and $I =
-\frac{k(k+1)}{2} + 1$. We can thus write $\opt$ as:
-%(note that we do not require $\meas_i$ to be orthogonal to each
-%other):
-
-\[
-\opt =
-\left[
-  \begin{array}{ccc}
-    \horzbar & \sqrt{\alpha_i} \meas^*_1 & \horzbar \\
-    \horzbar & \sqrt{\alpha_2} \meas_2^* & \horzbar \\
-             & \vdots    &          \\
-    \horzbar & \sqrt{\alpha_I} \meas_I^* & \horzbar \\
-  \end{array}
-\right].
-\]
-
-Unfortunately, $\opt$ is not a valid design, since its rows do not
-have unit norm. Still, the above representation of $\opt$ is useful:
-If $m > \frac{k(k+1)}{2} + 1$, then $\alpha_i > 1$ for some $1\leq i
-\leq I$.  Thus, we can view $\opt$ as a clustered design, since it
-places weight $>1$ on a single measurement vector.
+\input{caratheodory}

### 2024-08-12 latex/shell.sh

@@ -3,6 +3,7 @@ latexdiff --math-markup=OFF old.tex flat.tex > diff.tex
 latexmk --pdf ba.tex
 latexmk --pdf ba.tex
 latexmk --pdf ba.tex
+latexmk --pdf ba.tex
 latexmk --pdf diff.tex
 latexmk --pdf diff.tex
 latexmk --pdf diff.tex

### 2024-08-19 latex/intro.tex
made some progress

@@
 The practice of experimental design entails specifying every detail of
-running an experiment \cite{chaloner1995}. At its simplest form,
-experimental design involves choosing a set of measurements to be
-taken. For example, consider the process of searching for oil:
+running an experiment \cite{chaloner1995}. In its most basic form,
+experimental design entails specifying measurements that consist an
+experiment. For example, consider the process of searching for oil:
 measurements require digging holes in the ground and a \emph{design}
@@
 allowed. For example, in borehole tomography, a measurement involves
-drilling a deep well in the ground --- a very costly endeavor. Even in
-a seemingly simpler application such as MRI, a patient cannot occupy
-the MRI machine for too long, hence only a limited number of
-measurements can be taken. Therefore, measurements should be chosen to
-most benefit the user. Typically, a user will consider some utility,
-called a \emph{design criterion} and take measurements that maximize
-this utility. Two of the most widely recognized and extensively
-studied design criteria are the A- and \emph{D-optimality} criteria.
+drilling a deep well in the ground --- a very costly endeavor. In
+medical imaging applications like MRI, the time alloted for each
+patient in the MRI machine limits the number of measurements that can
+be taken during a single session. Therefore, measurements should be
+chosen to extract as much information from the experiment. Typically,
+a user will consider some utility, called a \emph{design criterion}
+and take measurements that maximize this utility. Two of the most
+widely recognized and extensively studied design criteria are the A-
+and \emph{D-optimality} criteria.
 
@@
 generalizes to linear models over infinite-dimensional Hilbert spaces
-(i.e.~function spaces).
+(e.g.~function spaces).
 
@@
 
-Solving the 1D heat equation is straightforward. Any initial heat
-distribution $u_0\in L^2([0,1])$ that satisfies the homogeneous
-Dirichlet boundary condition eq.~\eqref{eq:heat2} is a linear
-combination of sines $\ev_n(x) = \sin(\pi n x)$, so $u_0 = \sum_{n\geq
-  1} a_n \ev_n$ for some $\{a_n\}_{n\geq 1}$. Now, note that $\ev_n$
-are in fact eigenvectors of the Laplacian: $\Delta \ev_n = -\pi^2
-n^2\ev_n$. It should not be hard to believe that $u(\cdot, T) =
-\sum_{n\geq 1} a_n \exp(-\pi^2 n^2T ) \ev_n$
-\cite{renardy2006PDE}. Thus, eigenvectors of the linear operator that
-describes the time evolution of heat are $\ev_n(x) = \sin(\pi n x)$
-with corresponding eigenvalues
+Solving the 1D heat equation is straightforward. Consider an initial
+heat distribution $u_0\in L^2([0,1])$ that satisfies the homogeneous
+Dirichlet boundary condition eq.~\eqref{eq:heat2}. The initial
+condition $u_0$ is a linear combination of sines $\ev_n(x) = \sin(\pi
+n x)$, so $u_0 = \sum_{n\geq 1} a_n \ev_n$ for some $\{a_n\}_{n\geq
+  1}$. Now, note that $\ev_n$ are in fact eigenvectors of the
+Laplacian: $\Delta \ev_n = -\pi^2 n^2\ev_n$. It should not be hard to
+believe that $u(\cdot, T) = \sum_{n\geq 1} a_n \exp(-\pi^2 n^2T )
+\ev_n$ \cite{renardy2006PDE}. Thus, eigenvectors of the linear
+operator that describes the time evolution of heat are $\ev_n(x) =
+\sin(\pi n x)$ with corresponding eigenvalues
 \begin{equation}\label{eq:decay}
@@
 $n\to\infty$. See Supplementary movies S1 and S2 for an illustration
-of this phenomenon (code generating these movies is located in module
-\texttt{movies.py} in the accompanying
-\href{https://github.com/yairdaon/OED}{repository}). For ill-posed
+of this phenomenon\footnote{Code generating these movies is located in
+module \texttt{movies.py} in the accompanying
+\href{https://github.com/yairdaon/OED}{repository}.}. For ill-posed
 problems regularization is required, and we implement such
@@
 choice of prior covariance operator ensures the posterior is
-well-defined and prior realizations are well-behaved, see
-\cite[Theorem 3.1 and Lemma 6.25]{Stuart10} for details. The second
-ingredient of a fully Bayesian formulation of the inverse problem is
-likelihood. We simply assume centered iid Gaussian measurement error
-model, giving rise to a the widely applicable Gaussian likelihood
-function.
+well-defined and prior realizations are well-behaved, see Theorem 3.1
+and Lemma 6.25 in \cite{Stuart10} for details. The second ingredient of a
+fully Bayesian formulation of the inverse problem is likelihood. We
+simply assume centered iid Gaussian measurement error model, giving
+rise to a the widely applicable Gaussian likelihood function.
 

### 2024-08-19 latex/refs.tex
made some progress

@@
 
-I appreciate the many thoughtful comments by all of you. I believe
-these comments helped me considerably improve the exposition of my
-paper and added new insights as well. I sincerely appreciate the
+I appreciate the many thoughtful insightful comments by all of
+you. These comments have helped me considerably improve the exposition
+of my paper and added new insights as well. I sincerely appreciate the
 option to revise and resubmit. Please see below my answers to the
-comments made by the referees, associate editor and editor in chief:
-\rred{My answers are marked in red.} \bblue{In some places I copied
-  text from the revised manuscipt and marked it in blue.}
+comments made by editor in chief (Section \ref{eic}), associate editor
+(Section \ref{ae}) and three referees (Sections \ref{ref1}, \ref{ref2}
+and \ref{ref3}).
 
-\section{Editor in Chief}
+\section{Editor in Chief}\label{eic}
 \RC Personally I like the paper, but it represents a somehow unusual
@@
 
-\RV the motivation and the practical statistical implications are to
+\RC the motivation and the practical statistical implications are to
 be developed and discussed;
@@
 \AR I added a section on the implications:
-\begin{quote}
-  \input{implications}
-\end{quote}
+
+%\begin{quote}
+  %\input{implications}
+%\end{quote}
 
@@
 
-\AR I am not sure which specific comment of Reviewer 3 this
-      refers to, please see my answers to Reviewer 3 below.
+\AR I am not sure which specific comment of Reviewer 3 this refers to,
+please see my answers to Reviewer 3 below.
 
-\section{Associate Editor}
+\section{Associate Editor}\label{ae}
 \RC The three referees found merits and interest in your work, but also
@@
 \input{rev1}
+
 \input{rev2}
+
 \input{rev3}

### 2024-08-22 latex/ba.tex
Working towards revision. Modified intro, now working on refactoring code
changing paper and makin slides
before diff
Change biblib
biblib
Intro finished, looking for fig and caption of eigenvectors
Intro ready
Only need to comment on tekentrup

@@
 \input{definitions}
+
 %
@@
 \begin{abstract}
-  \input{abstract}
+  Estimation of parameters in physical processes often demands costly
+  measurements, prompting the pursuit of an optimal measurement
+  strategy. Finding such strategy is termed the problem of
+  \emph{optimal experimental design}, abbreviated as optimal
+  design. Remarkably, optimal designs can yield tightly clustered
+  measurement locations, leading researchers to fundamentally revise
+  the design problem just to circumvent this issue. Some authors
+  introduce error correlation among error terms that are initially
+  independent, while others restrict measurement locations to a finite
+  set of locations. While both approaches may prevent clusterization,
+  they also fundamentally alter the optimal design problem.
+
+  In this study, we consider Bayesian D-optimal designs, i.e.~designs
+  that maximize the expected Kullback-Leibler divergence between
+  posterior and prior. We propose an analytically tractable model for
+  D-optimal designs over Hilbert spaces. In this framework, we make
+  several key contributions: \textbf{(a)} We establish that
+  measurement clusterization is a generic trait of D-optimal designs
+  with independent Gaussian measurement errors, and prove that
+  introducing correlations among measurement error terms mitigates
+  clusterization. \textbf{(b)} We characterize D-optimal designs as
+  reducing uncertainty across a subset of prior covariance
+  eigenvectors. Finally, \textbf{(c)} We leverage this
+  characterization to argue that measurement clusterization arises as
+  a consequence of the pigeonhole principle: when more measurements
+  are taken than there are locations where the select eigenvectors are
+  large and others are small --- clusterization occurs.
 \end{abstract}
@@
 \input{vanishing}
-\input{why}
+\input{numerix}
+%\input{discussion}
+
+\begin{acks}[Acknowledgments]
+  This study is a result of research I started during my PhD studies
+  under the instruction of Prof.~Georg Stadler at the Courant
+  Institute of Mathematical Sciences. I would like to thank him for
+  his great mentorship, attention to details and kindness. I would
+  also like to thank Christian Remling, who helped me find a proof for
+  Lemma \ref{lemma:free} in
+  \href{https://mathoverflow.net/questions/280168/redistribute-diagonal-entries-of-a-matrix/280203#280203c}{Mathoverflow}.
+  Last, but certainly not least, I would like to thank the three
+  referees, associate editor and editor in chief for providing
+  detailed and insightful reviewes that have made this manuscript a
+  whole lot better.
+  %Having worked on this manuscript alone for several years, they have
+  %almost become my coautoh
+
+  This research was supported in part by an appointment with the
+  National Science Foundation (NSF) Mathematical Sciences Graduate
+  Internship (MSGI) Program sponsored by the NSF Division of
+  Mathematical Sciences. This program is administered by the Oak Ridge
+  Institute for Science and Education (ORISE) through an interagency
+  agreement between the U.S. Department of Energy (DOE) and NSF. ORISE
+  is managed for DOE by ORAU. All opinions expressed in this paper are
+  the author's and do not necessarily reflect the policies and views
+  of NSF, ORAU/ORISE, or DOE.
+  % This work was also supported by The Raymond and Beverly Sackler
+  % Post-Doctoral Scholarship.
+\end{acks}
 
 \bibliographystyle{ba}
-\bibliography{/home/yair/projects/bibtex.bib}
-
-\begin{acks}[Acknowledgments]
-\input{ack}
-
-\input{funding}
-\end{acks}
+\bibliography{/Users/yairdaon/work//lib.bib}
 

### 2024-08-22 latex/flat.tex
Only need to comment on tekentrup

@@
 \usepackage{amsthm}
+\usepackage{array}
+\newcommand*{\vertbar}{\rule[-1ex]{0.5pt}{2.5ex}}
+\newcommand*{\horzbar}{\rule[.5ex]{2.5ex}{0.5pt}}
 
@@
 \DeclareMathOperator*{\argmax}{arg\,max}
+\DeclareMathOperator*{\conv}{conv}
 
@@
 measurements in $\textup{span}\{\meas\}$ for some $\meas \neq 0$, then
-the D-optimality criterion is unbounded. In contrast, in any real-life
-problem where sensors are concerned, the norm of measurements recorded
-by sensors is always one\footnote{Again, our analysis does not
-directly apply to point evaluations. We just utilize point evaluations
-for motivation. We can approximate point evaluations by e.g.~elements
-in $\hilo^*$ as long as $\hilo$ is a function space,
-e.g.~$L^2(\Omega)$. In this case, for a fixed apporximation of
-$\delta$ the norm of the corresponding functional is constant (but
-$\neq 1$.}:
+the D-optimality criterion is unbounded, so a D-optimal design does
+not exist. In contrast, in any real-life problem where sensors are
+concerned, the norm of measurements recorded by sensors is always
+one\footnote{Again, our analysis does not directly apply to point
+evaluations. We just utilize point evaluations for motivation. We can
+approximate point evaluations by e.g.~elements in $\hilo^*$ as long as
+$\hilo$ is a function space, e.g.~$L^2(\Omega)$. In this case, for a
+fixed apporximation of $\delta$ the norm of the corresponding
+functional is constant (but $\neq 1$.}:
 
@@
 explanation to the measurement clusterization we observed for the
-inverse problem of the heat equation (Fig.~\ref{fig:eigenvectors}).
-
+inverse problem of the heat equation, see Section
+\ref{subsub:clusterization1} below. We also suggest a generic
+explanation for clusterization, see Section \ref{subsub:clusterization2}.
+
+\subsubsection{Clusterization in the heat equation}
 Consider $\fwd$ and $\prcov$ from \emph{the inverse problem of the
@@
 \end{figure}
+
+\subsubsection{Clusterization in our generic model}We can gain further insight to clusterization in our generic model,
+from Carath\'eodory's Theorem and the concentration on the first $k$
+eigenvectors of $\fwd \prcov \fwd^*$. We present a short discussion
+adapting arguments presented in \cite[Chapter
+  3]{silveyOptimalDesign1980} and \cite[Section
+  5.2.3]{pronzatoPazman2013}.
+
+Consider $\opt$ a D-optimal design under our generic model.  As
+instructed by Theorem \ref{thm:char}, we ignore all but the first $k$
+eigenvectors of $\fwd \prcov \fwd$. Thus, we replace $\hilo$ with
+$\hilo^{(k)}$ --- the $k$-dimensional subspace spanned by the first
+$k$ eigenvectors of $\fwd^*\prcov\fwd$. Let
+\begin{equation*}
+  \mathcal{M} := \conv \{\meas \meas^* : \meas\in \hilo^{(k)}, \|\meas\|=1\},
+\end{equation*}
+where $\conv$ denotes the convex hull of a set. The set $\mathcal{M}$
+contains only positive-definite operators on a $k$-dimensional vector
+space. Hence $\mathcal{M}$ lives in a $k(k+1)/2$-dimensional vector
+space. Since $\opt^*\opt = \sum_{i=1}^m \meas_i\meas_i^*$ for $\meas_i
+\in \hilo^{(k)}$, it is easy to verify that $\frac1m \opt^*\opt \in
+\mathcal{M}$.  Recall Carath\'eodory's Theorem:
+\begin{theorem}[Carath\'eodory]
+  Let $X \subseteq \mathbb{R}^n, X \neq \phi$ and denote $\conv (X)$
+  the convex hull of $X$. For every $x \in \conv (X)$, $x$ is a convex
+  combination of at most $n+1$ vectors in $X$.
+\end{theorem}
+Carath\'eodory's Theorem implies that there exist $\meas_i$ and
+$\alpha_i$ such that
+\begin{equation*}
+  \opt^*\opt = \sum_{i=1}^I \alpha_i \meas_i\meas_i^*,
+\end{equation*}
+where $\|\meas\|=1, \sum\alpha_i = m, \alpha_i \geq 0$ and $I =
+\frac{k(k+1)}{2} + 1$. We can thus write $\opt$ as:
+
+\[
+\opt =
+\left[
+  \begin{array}{ccc}
+    \horzbar & \sqrt{\alpha_i} \meas^*_1 & \horzbar \\
+    \horzbar & \sqrt{\alpha_2} \meas_2^* & \horzbar \\
+             & \vdots    &          \\
+    \horzbar & \sqrt{\alpha_I} \meas_I^* & \horzbar \\
+  \end{array}
+\right].
+\]
+
+Unfortunately, $\opt$ is not a valid design, since its rows do not
+have unit norm. Still, the above representation of $\opt$ is useful:
+If $m > \frac{k(k+1)}{2} + 1$, then $\alpha_i > 1$ for some $1\leq i
+\leq I$.  Thus, we can view $\opt$ as a clustered design, since it
+places weight $>1$ on a single measurement vector.
  \section{Numerical Experiments}
@@
 construction with the goal of testing numerically how prevalent are
-clustered designs. We iterate over the number of measurements $m \in
-\{4,\dots, 24\}$, and for every $m$ we then iterate over $k:=\rank
-\obs^*\obs \in \{2,\dots, m-1\}$. For each pair $m,k$ we run $N=5000$
-simulations according to the following steps:
+clustered designs. To this end, we would like to generate random prior
+eigenvalues $\lambda_j$, fix $m$ and $k$, find a D-optimal
+$\opt^*\opt$ and then utilize the construction of
+Theorem~\ref{thm:char} and Lemma~\ref{lemma:free} to find $\opt$.
+
+To simplify things, we directly sample rank $\opt^*\opt$. We iterate
+over the number of measurements $m \in \{4,\dots, 24\}$, and for every
+$m$ we then iterate over $k:=\rank \obs^*\obs \in \{2,\dots,
+m-1\}$. For each pair $m,k$ we repeat the following steps $N=5000$
+times:
 \begin{enumerate}
 \item Generate random diagonal $D\in \mathbb{R}^{k\times k}$ with
-  entries $\log (d_i) \sim \mathcal{N}(25,15)$ and normalize it so
-  that $\ttr D = m$.
+  entries $\log (d_i) \sim \mathcal{N}(50,15)$ and normalize so that
+  $\ttr D = m$. 
 \item Conjugate $D$ by a random orthogonal matrix to form a positive
-  semi-definite $M := UDU^t \in \mathbb{R}^{k\times k}$.
+  semi-definite $M := UDU^t \in \mathbb{R}^{k\times k}$. This $M$
+  represents $\opt^*\opt$.
 \item Apply the construction of Lemma \ref{lemma:free} to calculate
-  $A$ such that $AA^t = M$, where $A$ has unit norm columns.
-\item Since columns of $A$ correspond to measurement vectors, we call
-  $A$ "clustered" if $A$ has two or more identical columns (up to some
-  numerical precision threshold).
+  $A$ such that $AA^t = M$, where $A$ has unit norm columns. I.e.~find
+  the optimal design $\opt$.
+\item Since $A$ corresponds to $\opt$, its columns correspond to
+  measurement vectors. We call $A$ "clustered" if $A$ has two or more
+  identical columns (up to some numerical precision threshold).
 \end{enumerate}

### 2024-08-22 latex/intro.tex
Only need to comment on tekentrup

@@
 drilling a deep well in the ground --- a very costly endeavor. In
-medical imaging applications like MRI, the time alloted for each
+medical imaging applications like MRI, the time allotted for each
 patient in the MRI machine limits the number of measurements that can
@@
 
-We choose a prior for the initial condition $u_0 \sim \mathcal{N}(0,
-(-\Delta)^{-1})$, with homogeneous Dirichlet boundary condition. This
-choice of prior covariance operator ensures the posterior is
-well-defined and prior realizations are well-behaved, see Theorem 3.1
-and Lemma 6.25 in \cite{Stuart10} for details. The second ingredient of a
-fully Bayesian formulation of the inverse problem is likelihood. We
-simply assume centered iid Gaussian measurement error model, giving
-rise to a the widely applicable Gaussian likelihood function.
+In order to find a fully Bayesian formulation for our inverse problem
+we first need to specify a prior on \emph{functions} defined over
+$\Omega = [0,1]$. For the sake of simplicity we would like to utilize
+a Gaussian prior. While specifying the prior mean is easy --- we take
+the zero function on $\Omega$ --- specifying a prior covariance is
+more involved. In analogy with the finite dimensional case, we seek a
+covariance \emph{operator} that is positive definite and imposes
+sufficient regularity on functions over $\Omega$. Thus, we set our
+prior for the initial condition $u_0 \sim \mathcal{N}(0,
+(-\Delta)^{-1})$, where $\Delta$ again defined with homogeneous
+Dirichlet boundary condition. The intuition here is that
+$(-\Delta)^{-1}$ is smoothing and prior realizations are generated, in
+analogy to the finite dimensional case, via
+$(-\Delta)^{-1/2}\mathcal{W}$, where $\mathcal{W}$ is white noise
+\cite{rue2011}. Our choice of prior covariance operator ensures the
+posterior is well-defined and prior realizations are well-behaved, see
+Theorem 3.1 and Lemma 6.25 in \cite{Stuart10} for details. The second
+ingredient of a fully Bayesian formulation of the inverse problem is
+likelihood. We simply assume centered iid Gaussian measurement error
+model, giving rise to a the widely applicable Gaussian likelihood
+function.
 
@@
 observed to yield remarkably similar measurements in certain cases
-\cite{fedorovDesignSpatialExperiments1996, nyberg2012, fedorov1997,
-  Ucinski05, neitzel2019sparse}. This phenomenon is illustrated for
-our toy inverse problem in Fig.~\ref{fig:clusterization_illustration},
-where D-optimal measurement locations are shown for different numbers
-of measurements. Notably, for six measurements, a D-optimal design
-yields two sets of measurements that are identical. Following
+\cite{fedorov1996, nyberg2012, fedorov1997, Ucinski05,
+  neitzel2019sparse}. This phenomenon is illustrated for our toy
+inverse problem in Fig.~\ref{fig:clusterization_illustration}, where
+D-optimal measurement locations are shown for different numbers of
+measurements. Notably, for six measurements, a D-optimal design yields
+two sets of measurements that are identical. Following
 \cite{Ucinski05}, we refer to this intriguing phenomenon as
@@
 
-Clusterization should not be confused with either repetition nor with
-replication, which are commonly viewed as beneficial and even
-necessary aspects of an optimal design \cite{fisher1949design,
-  morris2011, schafer2001replication}. For example,
-\cite{fisher1949design} in his famous milk and tea experiment,
-suggested that repetition is a "way of enlarging the experiment and,
-thereby, increasing its sensitiveness". In another example,
-\cite{fay2000rainfall} measured the effect of rainfall on grass growth
-in plots of land. The experiment involved fifteen "rainfall
-manipulation shelters", where "Four rainfall manipulation treatments
-(three replicates) then were assigned to 12 of the plots". While it
-seems reasonable for the researchers to replicate the phenomenon they
-are trying to study, clusterization is different: a clustered design
-in the rainfall experiment would imply the researchers should take a
-repeated measurement \emph{on the same plot}, at the expense of
-measuring grass growth in other plots. In sharp contrast to repetition
-and replication, clusterization is highly nonintuitive and we will see
-that its origins run considerably deeper.
+%% REV 1, 2
+Clusterization should not be confused with replication. Replication
+requires that the experimentalist executes multiple trials under
+circumstances that are \emph{nominally identical} \cite[Section
+  1.2.4]{morris2011}. Replication is commonly viewed as a beneficial
+and even necessary aspect of optimal experimental design
+\cite{fisher1949design, morris2011, schafer2001replication}. Similarly
+to a design implementing replication, a clustered design reduces the
+signal-to-noise ratio of the repeated measurements
+\cite{telford2007brief}. The difference is that a clustered design
+takes repeated measurements at the expense of other quantities not
+measured at all.
+
+%% Rev 2
+For example, \cite{fisher1949design}, suggested repeating his famous
+milk and tea experiment in order "to be able to demonstrate the
+predominance of correct classifications in spite of occasional
+errors". Unfortunately, in the experiments we consider, replication is
+impossible. For example, in the MRI problem, we cannot generate an
+individual nominally identical to the one we wish to scan.
+
+%% Rev 2
+To further illustrate the difference between replication and
+clusterization, consider an experiment measuring the effect of
+rainfall on grass growth \cite{fay2000rainfall}. The experiment
+involved four rainfall manipulation "treatments" (i.e.~simulating
+different timing and quantity of rainfall), each replicated three
+times over a different plot of land. Indeed, it seems reasonable for
+researchers to replicate the phenomenon they are trying to study. A
+clustered design in such an experiment would imply the researchers
+should take a repeated measurement \emph{on the same plot}, at the
+expense of measuring grass growth in other plots. To conclude our
+short discussion on replication vs.~clusterization: these are
+fundamentally different concepts and even though replication is quite
+intuitive, it is inapplicable to the inverse problems we consider in
+this manuscript.
 
 Researchers of inverse problems widely agree that measurement
-clusterization is undesirable
-\cite{fedorovDesignSpatialExperiments1996, nyberg2012, fedorov1997,
-  Ucinski05, neitzel2019sparse}, prompting the exploration of various
-remedies to address this issue. One approach involves merging close
-measurements \cite{fedorov1997}; however, this strategy merely
-overlooks the phenomenon of measurement clusterization. An alternative
-solution lies in \emph{clusterization-free design}s, where measurement
-locations are deliberately chosen to be distant from one another. This
-can be achieved by imposing distance constraints between measurements
-or by introducing correlated errors that account for both observation
-error and model misspecification \cite{Ucinski05}. For instance, in
-the context of time-series analysis for pharmacokinetic experiments,
+clusterization is undesirable \cite{fedorov1996, nyberg2012,
+  fedorov1997, Ucinski05, neitzel2019sparse}, prompting the
+exploration of various remedies to address this issue. One approach
+involves merging close measurements \cite{fedorov1997}; however, this
+strategy merely overlooks the phenomenon of measurement
+clusterization. An alternative solution lies in
+\emph{clusterization-free design}s, where measurement locations are
+deliberately chosen to be distant from one another. This can be
+achieved by imposing distance constraints between measurements or by
+introducing correlated errors that account for both observation error
+and model misspecification \cite{Ucinski05}. For instance, in the
+context of time-series analysis for pharmacokinetic experiments,
 measurement clusterization can be mitigated by incorporating the
@@
 
-\subsubsection{Implications}
-\input{implications}
+\subsubsection{Implications}\label{subsub:implications}
+Our answer to Question \ref{q:generic} implies that encountering
+clusterization could be expected in many different problems across
+many different scientific fields. Researchers that encounter
+clusterization should not be surprised or wary. In Our answer to
+Question \ref{q:why}, we explain what our view of the cause of
+clusterization is. It appears, the cause is generic: a D-optimal
+design reduces uncertainty for a select set of prior covariance
+eigenvectors --- those with the most prior uncertainty, i.e.~those the
+practitioner cares about the most!
+
+A further implication of the analysis we present is that
+clusterization can serve as an evidence to the number of relevant
+eigenvectors in the problem. These eigenvectors correspond to the
+relevant degrees of freedom in the problem. Thus, clusterization tells
+us that we might be able to reduce the complexity of a computational
+model, e.g.~by dropping discretization points.
+
+%% REV 1
+%% REV 2
+When clusterization arises, we believe it should serve as a warning
+signal to practitioners. In the inverse problem of the 1D heat
+equation, clusterization occurs primarily because Laplacian
+eigenvectors do not decay in $\Omega$. Consequently, measuring $u(x_1,
+T)$ at some point $x_1 \in \Omega$ provides information about
+$u(x_2,T)$ for distant points $x_2 \in \Omega$. Intuitively, this
+should not happen: for small $T$, the heat distribution at $x_1$
+should give very little knowledge on the heat distribution at
+$x_2$. This behavior stems from a well-known property of the heat
+equation: it allows information to spread instantly across the
+computational domain \cite{renardy2006PDE}. In reality, heat (and
+information) propagate at finite speeds. Of course, the known physical
+barrier for information spread is the speed of light, but we expect
+heat and information to spread considerably slower. Thus,
+clusterization may suggest that our physical model itself is at fault.
+
+%% We believe that the emergence of clusterization in this context might
+%% be non-physical, arising from the way the inverse problems we consider
+%% are phrased. Clusterization therefore indicates that the underlying
+%% mathematical / Bayesian model is overly permissive and fails to
+%% capture crucial physical constraints of the problem. We suggest that
+%% when clusterization occurs, practitioners should consider alternative
+%% models where information is localized in space and travels at finite
+%% speed in the medium. Such models may not only provide more physically
+%% accurate and meaningful results but may also mitigate the issue of
+%% clusterization.
+
+%% REV 1
+%% REV 2
+Our choice of prior is perhaps an even greater contributor to
+clusterization. Our choice of Gaussian prior similarly implies
+information is shared between distant locations in $\Omega$. Thus, we
+suggest refraining from choosing Gaussian priors with inverse
+Laplacian covariance operators. Rather, non-Gaussian priors should be
+employed instead \cite{hosseini2017, hosseini2019}. So, clusterization
+may suggest we give more attention to the choice of prior.
+
+%% REV 1
+Given the above discussion, it is our belief that the methods
+suggested by other authors to avoid clusterization are merely
+overlooking the problem. We do not view clustered designs as
+inherently bad on their own. Rather, we suggest that if a clustered
+design arises, then a practitioner should revisit their probabilistic
+and physical modeling assumptions. If the practitioner is confident in
+their assumptions, then clustered designs should be avoided only to
+the extent necessitated by the physical measuring apparatus.
+
+Another For example, \cite{tekentrup2020} provides convergence analysis for
+various designs with differing space filling properties. They showed
+that convergence rate depends on how these designs fill space. We
+expect better convergence rates for D-optimal designs, even though
+D-optimal designs may demonstrate clusterization. The reason is that
+D-optimal designs explore space in a way tailored to the problem
+studied. Consequently, D-optimal designs cluster because they first
+aim to measure what matters most; see the discussion following
+Theorem \ref{thm:char} for details.
+
+%% be avoided at all. Nothing in our analysis indicates that we should
+%% expect any pathological behavior when utilizing D-optimal designs. On
+%% the contrary: we show that D-optimal designs (clustered or not) reduce
+%% uncertainty of prior covariance eigenvectors where it is
+%% highest. Thus, in a sense we hope to make precise in a future study,
+%% we expect convergence of for D-optimal designs to be fastest.
+
+%% practitioners should not try to avoid measurement
+%% clusterization. Rather, practitioners should take repeated
+%% measurements (e.g.~in MRI and borehole tomography), increase apparatus
+%% sensitivity (e.g.~in EIT), or take consecutive measurements (e.g.~in
+%% the 1D heat equation). Overall, we show that measurement
+%% clusterization is a natural and (almost) inevitable part of Bayesian
+%% D-optimal designs (but see disclaimer below).
+
+We expect D-optimal designs to judicious exploration of
+design space, as implemented by a D-optimal design to give better
+convergence compared to randomly sampling the design
+space \cite{knapik2011}.
+
+%% Clusterization is a peculiar phenomenon and it is perfectly reasonable
+%% for someone to argue against the D-optimality criterion based on the
+%% fact that it results in clustered designs. We have seen, however that
+%% there is a perfectly reasonable explanation for clusterization. We
+%% have shown that clusterization is an inevitable consequence of having
+%% a problem with some modes where uncertainty decays faster than others.
 

### 2024-08-22 latex/vanishing.tex
Only need to comment on tekentrup

--- 2024-08-11 latex/vanishing.tex
@@
 \end{equation*}
-where $\|\meas\|=1, \sum\alpha_i = m, \alpha_i \geq 0$ and $I =
+where $\|\meas_i\|=1, \sum\alpha_i = m, \alpha_i \geq 0$ and $I =
 \frac{k(k+1)}{2} + 1$. We can thus write $\opt$ as:

### 2024-08-22 latex/why.tex
Only need to comment on tekentrup

--- 2023-08-29 latex/why.tex
@@
-\section{Answer to Question \ref{q:why}: The cause of clusterization}\label{section:why}
-According to Theorem \ref{thm:char}, D-optimal designs aim to capture
-a small subset of eigenvectors of the prior covariance, specifically
-the $k$ eigenvectors with the highest prior variance. Our model
-naturally achieves this objective by not measuring eigenvectors $k+1$
-and above, as proven in Theorem \ref{thm:char}. Translating this
-understanding to spatial problems, we anticipate that a D-optimal
-design would favor measurement locations where eigenvectors $k+1$ and
-above are either close to zero in value or possess small eigenvalues
-in the prior spectrum, for some $k > 0$. To illustrate this
-preference, Fig.~\ref{fig:why} depicts the scenario using the 1D
-heat equation with homogeneous Dirichlet boundary conditions (details
-in the supplementary material). The plot showcases four eigenvectors,
-scaled according to their prior standard deviations. Since
-eigenvectors beyond the fourth have insignificant prior eigenvalues,
-we exclude them from consideration. Notably, we observe that
-measurements are clustered near the zeros of the third and fourth
-eigenvectors, so we conclude that $k=2$. Clusterization arises because
-there are only two locations where the third and fourth eigenvectors
-approach zero, while the first and second eigenvectors exhibit
-significantly non-zero values. Consequently, when we allocate four
-measurements to these two locations, they naturally cluster together,
-aligning with the pigeonhole principle.
+\subsection{Answer to Question \ref{q:why}}\label{subsec:why}
+Building on Theorem \ref{thm:char}, we can now give a compelling
+explanation to the measurement clusterization we observed for the
+inverse problem of the heat equation, see Section
+\ref{subsub:clusterization1} below. We also suggest a generic
+explanation for clusterization, see Section \ref{subsub:clusterization2}.
+
+\subsubsection{Clusterization in the heat equation}
+Consider $\fwd$ and $\prcov$ from \emph{the inverse problem of the
+heat equation}. As before, we denote the eigenvalues of
+$\fwd\prcov\fwd^*$ by $\lambda_j$. We input these eigenvalues into our
+\emph{generic} model, and find a D-optimal design $\opt$ for our
+generic model using Theorem \ref{thm:char}. In our generic model, the
+measurements we take are best utilized in reducing uncertainty for the
+first $k$ eigenvectors. So, a D-optimal design arising from our
+\emph{generic model} completely avoids measuring eigenvectors $k+1$
+and above.
+
+Of course, in a real life problem --- such as the inverse problem of
+the 1D heat equation --- it is likely impossible to find measurement
+locations for which all eigenvectors $k+1$ and above are
+zero. However, if the eigenvalues of $\fwd\prcov\fwd^*$ decay quickly
+(recall the square-exponential decay for eigenvalues of the 1D heat
+equation in eq.\eqref{eq:decay}), a D-optimal design will try to
+balance measuring a small number (i.e.~$k$) of the leading
+eigenvectors.
+
+The abovementioned balance is explored in
+Fig.~\ref{fig:eigenvectors}. We allow $m=4$ measurements in $\Omega =
+[0,1]$ and observe that D-optimal measurement locations are clustered
+at $x_1 = 0.31$ and $x_2 = 0.69$. Upon close inspection of the scaled
+eigenvectors of $\fwd \prcov \fwd^*$, we first observe that
+eigenvectors $3$ and above have negligible prior amplitude. Since we
+only have $m=4$ measurements at our disposal, we interpret these
+results, following Theorem \ref{thm:char}, as implying we should only
+care about measuring the first and second eigenvectors. Then, we note
+the D-optimal $x_1,x_2$ present a compromise between the amplitude of
+the first and second eigenvectors. For example, a measurement at
+$x=0.5$ would have ignored the second eigenvector altogether, since
+the second eigenvector is zero at $x=0.5$.
+
+Now we can understand measurement clusterization for the inverse
+problem of the heat equation. A D-optimal design attempts to measure
+the first $k$ eigenvectors of $\fwd \prcov \fwd^*$. But there may be
+(spatial) limitations on where these $k$ eigenvectors have large
+amplitude. For the inverse problem of the heat equation there are two
+spatial locations that present a good compromise between the
+amplitudes of the first and second eigenvectors, namely $x_1$ and
+$x_2$ --- see Fig.~\ref{fig:eigenvectors}. We have $m=4$ measurements
+at our disposal but only two spatial locations that are a good
+compromise between the first and second scaled eigenvectors. Thus,
+clusterization arises as a consequence of the pigeonhole principle.
+
+\begin{figure}\label{fig:eigenvectors}
+    \centering
+    \includegraphics[width=\textwidth]{figs/eigenvectors_dst_scaled.png}
+    \caption{D-optimal measurement locations ($m=4$ measurements) and
+      weighted eigenvectors for finding the initial condition of the
+      1D heat equation. Measurement locations and weighted
+      eigenvectors are plotted over the computational domain $\Omega =
+      [0, 1]$ (x-axis). Measurement clusterization occurs
+      approximately at $0.31$ and $0.69$. These two locations are a
+      compromise between the magnitudes of the first and second
+      eigenvectors, which are the eigenvectors that a D-optimal design
+      aims to measure. Allocating $m=4$ measurements into two
+      locations results in clusterization, according to the pigeonhole
+      principle.}
+  \label{fig:why}
+\end{figure}

### 2024-08-22 revision/refs.tex
Only need to comment on tekentrup

\documentclass[draft]{article}
\def\papertitle{Measurement Clusterization in Bayesian D-optimal Designs in Infinite Dimensions}
\def\authors{Yair Daon}
\def\journal{Bayesian Analysis}
\def\doi{12345}

% Define title defaults if not defined by user
\providecommand{\lettertitle}{Author Response to Reviews of}
\providecommand{\papertitle}{Title}
\providecommand{\authors}{Authors}
\providecommand{\journal}{Journal}
\providecommand{\doi}{--}

\input{preamble}
\input{../latex/definitions}
\usepackage{xr}
\externaldocument{../latex/ba}

\usepackage[colorlinks,citecolor=blue,urlcolor=blue,filecolor=blue,backref=page]{hyperref}

\begin{document}

% Make title
{\Large\bf \lettertitle}\\[1em]
{\huge \papertitle}\\[1em]
{\authors}\\
{\it \journal, }\texttt{doi:\doi}\\
\hrule

% Legend
\hfill {\bfseries RC:} \textbf{\textit{Reviewer Comment}},\(\quad\) AR: \emph{Author Response}, \(\quad\square\) Manuscript text

I appreciate the many thoughtful insightful comments by all of
you. These comments have helped me considerably improve the exposition
of my paper and added new insights as well. I sincerely appreciate the
option to revise and resubmit. Please see below my answers to the
comments made by editor in chief (Section \ref{eic}), associate editor
(Section \ref{ae}) and three referees (Sections \ref{ref1}, \ref{ref2}
and \ref{ref3}).

%% \section{Editor in Chief}\label{eic}
%% \RC Personally I like the paper, but it represents a somehow unusual
%% submission for Bayesian Analysis, and I am worried about it being of
%% sufficient general interest for the broad readership of the
%% journal. Still, in agreement with the AE, we decided to give the
%% author the possibility to revise the paper. However, I must stress
%% that there is no guarantee of a successful outcome, which will solely
%% depend on the quality of the revision. In particular:

%% \RC the Introduction is to be made accessible to the whole BA
%%     readership;

%% \AR I have considerably extended the introduction.

%% \RC the motivation and the practical statistical implications are to
%% be developed and discussed;

%% \AR I added a section on the implications:

%% %\begin{quote}
%%   %\input{implications}
%% %\end{quote}

%% \RC the contribution has to be better contextualized with respect to
%% the literature in both inverse problems and D-optimal designs;

%% \RC numerical evidence and code, mentioned in the paper but actually
%% not provided, need to be made available;

%% \AR I added a section describing numerical experiments and their
%% results. I also added documentation to the accompanying
%% \href{https://github.com/yairdaon/OED}{repository} and made sure all
%% scripts there run.

%% \RC the mathematical derivations have to be watertight and some
%% clarifications are needed in this respect (see Report of Referee 3).

%% \AR I am not sure which specific comment of Reviewer 3 this refers to,
%% please see my answers to Reviewer 3 below.

%% \section{Associate Editor}\label{ae}
%% \RC The three referees found merits and interest in your work, but also
%% pointed out several major issues to be addressed. Of particular
%% concern to me, the paper's exposition jumps too quickly into the
%% formulation / maths without giving a proper intuition and
%% context. This makes the paper less accessible to readers who aren't
%% already quite familiar with inverse problems (which are introduced
%% with very little preamble, and it's not even immediately explicit what
%% are the parameters being learned) and optimal experimental design.

%% \AR I have added considerably to the introduction. See revised
%% manuscript and diff file.

%% \RC A referee makes valid points as to whether clusterization of
%% measurements is a problem in general, when in fact in can be an
%% appealing feature in some settings, and another referee asks whether
%% D-optimal designs should be abandoned altogether. The exposition
%% needs to discuss these high-level issues, and introduce the related
%% literature and results appropriately, to give the work proper
%% context.

%% \RC Altogether, I think that your work has the potential to make a
%% good contribution to Bayesian analysis, but as it currently stands
%% significant work is needed.

\input{rev1}
\input{rev2}
\input{rev3}

\bibliographystyle{apalike}
\bibliography{../../lib.bib}

\end{document}

### 2024-08-22 revision/rev1.tex
Only need to comment on tekentrup

--- 2024-08-12 latex/intro.tex
@@
-\section{Introduction}\label{section:intro}
-The practice of experimental design entails specifying every detail of
-running an experiment \cite{chaloner1995}. At its simplest form,
-experimental design involves choosing a set of measurements to be
-taken. For example, consider the process of searching for oil:
-measurements require digging holes in the ground and a \emph{design}
-simply indicates which holes should be dug
-\cite{horesh2008borehole}. Since measurements are expensive, only a
-limited number of them can be taken, and thus carefully choosing
-measurement locations is crucial.
+\section{Reviewer 1}\label{ref1}
+\RC This paper explains a common practical issue: the cause of
+clusterization in Bayesian D-optimal designs in infinite dimensions
+and why adding a correlated measurement process decluster the design
+points. Its theoretical results are valuable to the field of inverse
+problems/uncertainty quantification and the conclusions should be
+useful to practitioners in this field.
 
-Specifying the right set of measurements holds particular significance
-when solving an \emph{inverse problem} --- i.e.~making inference of
-the physical world utilizing a physical model of a phenomenon of
-interest \cite{tarantola2005,kaipio2005}. In an inverse problem, we
-seek to infer physical quantities, based on observations and a model
-of the physical world. Models are typically phrased in the language of
-ordinary or partial differential equations.
+\AR I appreciate the kind words.
 
-Inverse problems are ubiquitous in science and engineering. For
-example, in electrical impedance tomography (EIT) we seek to infer the
-structure of the inside of a human body. Inference is conducted based
-on observations of electrical current impedance measured at specific
-electrode locations on the skin, leveraging known equations of
-electric current flow \cite{horesh2010impedance}. Magnetic Resonance
-Imaging (MRI) also entails solving an inverse problem. There,
-radio-frequency pulses are sent through the human body in the presence
-of a strong magnetic field. The response of one's body contents to
-these radio-frequency pulses is measured, allowing a radiologist to
-view the internals of a body in a noninvasive manner
-\cite{horesh2008mri}. An inverse problem also arises in oil and gas
-exploration, where acoustic waves are sent through "boreholes" ---
-deep cylindrical wells drilled into the ground. Data of travel and
-return times of the acoustic waves are recorded. Wave travel and
-return time is influenced by the properties of the subsurface
-materials, such as density, elasticity, and the presence of fluids or
-voids. Combining travel time data with a geophysical models of
-contents of the earth's crust facilitates reconstructing the structure
-of the subsurface in a process called \emph{borehole tomography}
-\cite{horesh2008borehole}. Inverse problems also arise in many other
-areas of seismology and geology \cite{rabinowitz1990, steinberg1995}
-and medical imaging \cite{tarantola2005}. In many inverse problems the
-goal is to infer some numerical parameter. However, in the formulation
-we consider in this study, as well as in the examples above, the goal
-is to conduct inference over some \emph{function} over $\Omega$, where
-\(\Omega \subseteq \mathbb{R}^d, d=1,2,3\) is a spatial domain of
-interest.
+\RC However, I feel changing the assumption from independent to correlated
+measurements is quite artificial: people tweak a data generation
+process assumption to solve an optimal design problem, and the author
+just takes the assumption as it is.
 
-\subsection{D-optimal Designs}\label{subsec:D}
-In all of the examples of inverse problems we have seen, as well as in
-many other applications, only a limited number of measurements are
-allowed. For example, in borehole tomography, a measurement involves
-drilling a deep well in the ground --- a very costly endeavor. Even in
-a seemingly simpler application such as MRI, a patient cannot occupy
-the MRI machine for too long, hence only a limited number of
-measurements can be taken. Therefore, measurements should be chosen to
-most benefit the user. Typically, a user will consider some utility,
-called a \emph{design criterion} and take measurements that maximize
-this utility. Two of the most widely recognized and extensively
-studied design criteria are the A- and \emph{D-optimality} criteria.
+\AR If I understand correctly --- I completely agree. I added text
+emphasizing that I do not endorse using correlated errors:
 
-In this study, we focus on the Bayesian D-optimality
-criterion. Bayesian D-optimality carries a simple and intuitive
-meaning: measurements chosen according to the D-optimality criterion
-maximize the expected Kullback-Leibler (KL) divergence from posterior
-to prior \cite{chaloner1995, AlexanderianGloorGhattas14,
-  CoverThomas91}. Utilizing KL divergence for quantifying differences
-between distributions is standard practice and we will not motivate it
-further here. For a discrete parameter $\param$ and data $\data$, the
-KL divergence is defined as
-\begin{equation}\label{eq:basic_KL}
-  D_{KL}\left (\Pr(\param|\data)||\Pr(\param)\right ) =  \sum_{\param} \log
-  \frac{\Pr(\param|\data)}{\Pr(\param)} \Pr(\param|\data). 
-\end{equation}
-Of course, data $\data$ is not known before the experiment is
-conducted, so we average over it to define the D-optimality criterion
-as:
-\begin{equation*}
-  \mathbb{E}_{\data}\left [ D_{KL}\left (\Pr(\param|\data)||\Pr(\param)\right ) \right ].
-\end{equation*}
-We refer to a set of measurements that maximizes the D-optimality
-criterion as a \emph{D-optimal design}.
+\begin{quote}
+  Given the above discussion, it is our belief that the methods
+  suggested by other authors to avoid clusterization are merely
+  overlooking the problem. We do not view clustered designs as
+  inherently bad on their own. Rather, we suggest that if a clustered
+  design arises, then a practitioner should revisit their probabilistic
+  and physical modeling assumptions. If the practitioner is confident in
+  their assumptions, then clustered designs should be avoided only to
+  the extent necessitated by the physical measuring apparatus.
+\end{quote}
 
-For a linear model with Gaussian prior in finite dimensions, a
-D-optimal design minimizes the determinant of the posterior covariance
-\cite{chaloner1995}. In Section \ref{subsec:D_optimal_design} we give
-a more general definition of the D-optimality criterion that applies
-to arbitrary measures. We also show how the D-optimality criterion
-generalizes to linear models over infinite-dimensional Hilbert spaces
-(i.e.~function spaces).
+\RC Since the author proves the clusterization issue under the independent
+measurement assumption, I expect the author to go deeper and explain
+why an ordinary data generation assumption + a common optimal design
+leads to a “suboptimal” design?
 
-\subsection{A toy model: the 1D heat equation}\label{subsec:toy}
-For concreteness, let us now consider a toy inverse problem: inferring
-the initial condition for partial differential equation known as the
-\emph{heat equation} in one dimension (1D heat equation
-henceforth). Readers less familiar with partial differential equations
-can think of the following setup: before us there is a perfectly
-insulated metal rod. At each point on the rod the temperature is
-different and unknown, and we call this temperature distribution the
-\emph{initial condition}. We wait for a short time $T$ and let the
-heat dissipate a bit inside the rod. The heat equation determines the
-heat distribution inside the rod at time $t=T$.
+\AR In my view, this is a result of the unlocalized structure of
+Laplacian eigenvectors. I believe this is more of a problem with our
+physical models and choice of priors. See discussion below reproduced
+from the revised manuscript:
 
-The full time evolution of heat in the rod $\Omega=[0,1]$ is formally
-described by the following three equations:
-\begin{subequations}
-  \begin{alignat}{2}
-    u_t &= \Delta u &&\qquad \text{in } [0,1] \times [0,\infty), \label{eq:heat1}\\
-    u &= 0 &&\qquad \text{on } \{0, 1\} \times [0,\infty), \label{eq:heat2}\\
-    u &= u_0 &&\qquad \text{on }[0,1] \times \{0\}. \label{eq:heat3}
-  \end{alignat}
-\end{subequations}
-As time passes, heat dissipates across the rod as hotter regions,
-e.g.~local maxima of the heat distribution, become cooler. This
-behavior is captured by eq.~\eqref{eq:heat1}: the Laplacian at local
-maxima is negative, so $u_t = \Delta u$ implies $u$ decreases at its
-maxima, and the reverse happens at local minima. Eq.~\eqref{eq:heat2}
-describes the \emph{boundary condition}, which dictates how heat
-interacts with the outside world at the two edges of the rod
-$\{0,1\}$. Specifically, eq.~\eqref{eq:heat2} implements an absorbing
-boundary, i.e.~heat that interacts with the boundary immediately
-disappears. This type of boundary condition is known as a
-\emph{homogeneous Dirichlet boundary condition}. Eq.~\eqref{eq:heat3}
-describes the \emph{initial condition}, i.e.~the heat distribution
-inside the rod at time $t=0$.
+\begin{quote}
+  When clusterization arises, we believe it should serve as a warning
+  signal to practitioners. In the inverse problem of the 1D heat
+  equation, clusterization occurs primarily because Laplacian
+  eigenvectors do not decay in $\Omega$. Consequently, measuring
+  $u(x_1, T)$ at some point $x_1 \in \Omega$ provides information
+  about $u(x_2,T)$ for distant points $x_2 \in \Omega$. Intuitively,
+  this should not happen: for small $T$, the heat distribution at
+  $x_1$ should give very little knowledge on the heat distribution at
+  $x_2$. This behavior stems from a well-known property of the heat
+  equation: it allows information to spread instantly across the
+  computational domain \cite{renardy2006PDE}. In reality, heat (and
+  information) propagate at finite speeds. Of course, the known
+  physical barrier for information spread is the speed of light, but
+  we expect heat and information to spread considerably slower. Thus,
+  clusterization may suggest that our physical model itself is at
+  fault.
 
-Solving the 1D heat equation is straightforward. Any initial heat
-distribution $u_0\in L^2([0,1])$ that satisfies the homogeneous
-Dirichlet boundary condition eq.~\eqref{eq:heat2} is a linear
-combination of sines $\ev_n(x) = \sin(\pi n x)$, so $u_0 = \sum_{n\geq
-  1} a_n \ev_n$ for some $\{a_n\}_{n\geq 1}$. Now, note that $\ev_n$
-are in fact eigenvectors of the Laplacian: $\Delta \ev_n = -\pi^2
-n^2\ev_n$. It should not be hard to believe that $u(\cdot, T) =
-\sum_{n\geq 1} a_n \exp(-\pi^2 n^2T ) \ev_n$
-\cite{renardy2006PDE}. Thus, eigenvectors of the linear operator that
-describes the time evolution of heat are $\ev_n(x) = \sin(\pi n x)$
-with corresponding eigenvalues
-\begin{equation}\label{eq:decay}
-  \exp(-\pi^2 n^2T ).
-\end{equation}
+  Our choice of prior is perhaps an even greater contributor to
+  clusterization. Our choice of Gaussian prior similarly implies
+  information is shared between distant locations in $\Omega$. Thus,
+  we suggest refraining from choosing Gaussian priors with inverse
+  Laplacian covariance operators. Rather, non-Gaussian priors should
+  be employed instead \cite{hosseini2017, hosseini2019}. So,
+  clusterization may suggest we give more attention to the choice of
+  prior.
+\end{quote}
 
-In the \emph{inverse problem of the 1D heat equation}, our goal is to
-infer the initial condition $u_0 = u(\cdot, 0)$ from observations of
-the final state $u(\cdot, T)$. However, we are not able to measure
-$u(\cdot, T)$ at every point in the domain $\Omega$. Rather, we take
-some number of temperature measurements on the rod once our waiting
-time $T$ has passed and try to infer $u(\cdot, 0)$ from these
-measurements.
+\RC Not just by the mathematical proof, but from an information theory
+point of view. For example, does the redundancy points mean additional
+points does not reduce the variance of the estimation at all? If
+that’s the case, how does correlated measurement assumption “help”
+with it?
 
-The inverse problem of the 1D heat equation is difficult ("ill-posed")
-since heat spreads in a diffusive manner and any roughness in the
-initial condition is quickly smoothed, as implied by the squared
-exponential decay of eigenvalues in eq.~\eqref{eq:decay} with
-$n\to\infty$. See Supplementary movies S1 and S2 for an illustration
-of this phenomenon (code generating these movies is located in module
-\texttt{movies.py} in the accompanying
-\href{https://github.com/yairdaon/OED}{repository}). For ill-posed
-problems regularization is required, and we implement such
-regularization via a fully Bayesian formulation of the inverse
-problem.
+\AR I think this is the same misunderstanding I commented on above;
+For any nonzero observation error $\sigma^2 > 0$, repeated
+measurements will indeed reduce the signal-to-noise ratio of the
+measurement and result in an increase of the design criterion. I
+briefly comment on this subject in the paragraph below:
 
-We choose a prior for the initial condition $u_0 \sim \mathcal{N}(0,
-(-\Delta)^{-1})$, with homogeneous Dirichlet boundary condition. This
-choice of prior covariance operator ensures the posterior is
-well-defined and prior realizations are well-behaved, see
-\cite[Theorem 3.1 and Lemma 6.25]{Stuart10} for details. The second
-ingredient of a fully Bayesian formulation of the inverse problem is
-likelihood. We simply assume centered iid Gaussian measurement error
-model, giving rise to a the widely applicable Gaussian likelihood
-function.
-
-\subsection{Measurement Clusterization}
-Surprisingly, A- and D-optimal designs for inverse problems have been
-observed to yield remarkably similar measurements in certain cases
-\cite{fedorovDesignSpatialExperiments1996, nyberg2012, fedorov1997,
-  Ucinski05, neitzel2019sparse}. This phenomenon is illustrated for
-our toy inverse problem in Fig.~\ref{fig:clusterization_illustration},
-where D-optimal measurement locations are shown for different numbers
-of measurements. Notably, for six measurements, a D-optimal design
-yields two sets of measurements that are identical. Following
-\cite{Ucinski05}, we refer to this intriguing phenomenon as
-\emph{measurement clusterization}. We consider a design to be
-\emph{clustered} when two or more of its constituent measurements are
-identical.
-
-\begin{figure}
-    \centering
-    \includegraphics[height=0.5\textwidth]{figs/dst_modelError0.png}
-    \caption{Measurement clusterization in D-optimal designs for the
-      inverse problem of the 1D heat equation. Measurement locations
-      were chosen according to the Bayesian D-optimality criterion of
-      Theorem \ref{thm:d_optimality}. Measurement locations are
-      plotted over the computational domain \(\Omega = [0, 1]\)
-      (x-axis), for varying numbers of measurements (y-axis). The
-      colored numbers are measurement indices, plotted for visual
-      clarity. Measurement clusterization already occurs for three
-      measurements: the second measurement (red) is overlaid on the
-      third (green). For five measurements, first (blue) and second
-      (red) measurements are clustered, as well as the fourth (black)
-      and the fifth (magenta).}
-  \label{fig:clusterization_illustration}
-\end{figure}
-
-Clusterization should not be confused with either repetition nor with
-replication, which are commonly viewed as beneficial and even
-necessary aspects of an optimal design \cite{fisher1949design,
-  morris2011, schafer2001replication}. For example,
-\cite{fisher1949design} in his famous milk and tea experiment,
-suggested that repetition is a "way of enlarging the experiment and,
-thereby, increasing its sensitiveness". In another example,
-\cite{fay2000rainfall} measured the effect of rainfall on grass growth
-in plots of land. The experiment involved fifteen "rainfall
-manipulation shelters", where "Four rainfall manipulation treatments
-(three replicates) then were assigned to 12 of the plots". While it
-seems reasonable for the researchers to replicate the phenomenon they
-are trying to study, clusterization is different: a clustered design
-in the rainfall experiment would imply the researchers should take a
-repeated measurement \emph{on the same plot}, at the expense of
-measuring grass growth in other plots. In sharp contrast to repetition
-and replication, clusterization is highly nonintuitive and we will see
-that its origins run considerably deeper.
-
-Researchers of inverse problems widely agree that measurement
-clusterization is undesirable
-\cite{fedorovDesignSpatialExperiments1996, nyberg2012, fedorov1997,
-  Ucinski05, neitzel2019sparse}, prompting the exploration of various
-remedies to address this issue. One approach involves merging close
-measurements \cite{fedorov1997}; however, this strategy merely
-overlooks the phenomenon of measurement clusterization. An alternative
-solution lies in \emph{clusterization-free design}s, where measurement
-locations are deliberately chosen to be distant from one another. This
-can be achieved by imposing distance constraints between measurements
-or by introducing correlated errors that account for both observation
-error and model misspecification \cite{Ucinski05}. For instance, in
-the context of time-series analysis for pharmacokinetic experiments,
-measurement clusterization can be mitigated by incorporating the
-modeling of auto-correlation time within the noise terms
-\cite{nyberg2012}.
-
-In spatial problems involving choice of measurements within a domain
-\(\Omega \subseteq \mathbb{R}^d, d=1,2,3\), many researchers
-circumvent the problem of measurement clusterization by restricting
-measurements to a coarse grid in \(\Omega\) \cite{koval2020,
-  alexanderian2021, attia2022, alexanderian2014, alexanderian2016,
-  alexanderian2018efficient, brunton2016}. This approach incurs a
-significant computational cost as it requires solving a difficult
-combinatorial optimization problem for measurement locations over a
-discrete set. The combinatorial optimization problem is usually
-relaxed by first assigning optimal measurement weights in
-\(\mathbb{R}_+\) to the potential measurement locations. Some
-researchers incorporate a sparsifying \(\ell_1\) penalty term into the
-design criterion, which is subsequently thresholded to achieve the
-desired binary design over the coarse grid
-\cite{horesh2008borehole}. Others progressively relax the \(\ell_1\)
-penalty to an \(\ell_0\) penalty via a continuation method
-\cite{alexanderian2016, alexanderian2014}. Others cast the problem of
-finding optimal measurement weights as a stochastic optimization
-problem \cite{attia2022stochastic}. All of the aforementioned methods
-may indeed find a binary optimal design restricted to a given coarse
-grid. However, none addresses one fundamental issue: the restriction
-of measurement locations to a coarse grid in \(\Omega\) fundamentally
-changes the optimal design problem and thus results in a sub-optimal
-design.
-
-Avoiding measurement clusterization is a pragmatic approach:
-intuitively, researchers recognize that measurement clusterization is
-undesirable, even though the underlying reasons may not be fully
-clear. Consequently, they strive to prevent it and devise various
-methodologies to avoid it. Yet each and every one of these
-methodologies achieves its objective by imposing restrictions on
-measurement locations, thereby fundamentally altering the optimal
-design problem. To the best of my knowledge, no previous study has
-tried to address some seemingly simple yet fundamental questions:
-%
-Why does imposing correlations between observations alleviate
-measurement clusterization?
-%
-Is measurement clusterization a generic phenomenon? 
-%
-And, most importantly: Why does measurement clusterization occur?
-%
-%Should we aim to avoid measurement clusterization?
-%
-%Is it possible to substitute an optimal clustered design with an
-%equally optimal non-clustered design?
-%Can an optimal clustered design be replaced with an equally optimal
-%non-clustered design?
-
-\subsection{Contribution}
-The primary objective of this study is to provide a deep understanding
-of measurement clusterization by addressing the aforementioned
-questions. Our focus centers around investigating the Bayesian
-D-optimality criterion. We conduct an analysis of Bayesian D-optimal
-designs within the context of linear inverse problems over Hilbert
-spaces and study two inverse problems: (a) In Sections
-\ref{section:prelim} and \ref{section:D_and_grad} we propose a novel
-generic model for an inverse problem where D-optimality maintains
-analytical tractability and D-optimal designs are identified via
-Lagrange multipliers. This analytical framework facilitates the
-exploration of the questions posed at the end of the previous
-paragraph. We also study (b) the inverse problem of the 1D heat
-equation from Section \ref{subsec:toy} above. Investigating both
-inverse problems allows us to answer the questions posed in the
-previous section:
-
-\begin{enumerate}
-\item \label{q:generic} \textbf{Is measurement clusterization a
-  generic phenomenon?}
-  %\subsection{An answer for Question \ref{q:generic}: Genericity of measurement clusterization}
-  %% Computer implementation of Lemma \ref{thm:char} for the inverse
-  %% problem outlined in Section \ref{section:how} also generates \(\obs\)
-  %% as a solution to the D-optimal design problem.  Furthermore,
-  We give two complementing answers to this question. First, from a
-  theoretical perspective, we show that clusterization mainly depends
-  on how quickly the eigenvalues of the prior covariance in
-  observation space decay. This decay depends mostly on how ill-posed
-  the problem at hand is and does not depend much on the prior. See
-  Section \ref{section:vanishing} --- particularly, Theorem
-  \ref{thm:char} and the discussion following it. We also show results
-  of numerical experiments, where simulations of our model give rise
-  to D-optimal designs that exhibit clusterization with high
-  probability in Section \ref{subsec:lemma_sims}. Thus, given the
-  genericity of our model, we expect measurement clusterization to be
-  a generic and ubiquitous phenomenon.
-
-\item \label{q:mitigate} \textbf{Why does imposing correlations
-  between observations alleviate measurement clusterization?} In
-  Section \ref{section:non_vanishing}, we rigorously demonstrate the
-  role of model error in mitigating clusterization, thereby
-  corroborating earlier observations made by various
-  researchers. Specifically, our proof shows that identical
-  measurements result in no gain in design criterion when observation
-  error amplitude tends to zero. Moreover, in Section
-  \ref{subsec:corr_errors_sims}, we show that an error term
-  corresponding to correlations between measurements mitigates
-  clusterization in the inverse problem of the 1D heat equation.
-
-\item \label{q:why} \textbf{Why does measurement clusterization
-  occur?} In Sections \ref{subsec:why} we give two compelling answers
-  to this question by (i) transporting insights we gain from our
-  generic model to the inverse problem of the 1D heat equation, and
-  (ii) connecting measurement clusterization to Carath\'eodory's
-  Theorem.
-
-  Our analysis for the heat equation relies on conclusions from our
-  generic model. In particular, Theorem \ref{thm:char} reveals that a
-  D-optimal design focuses on a select set of prior eigenvectors,
-  i.e.~those with the largest eigenvalues in the prior covariance
-  spectrum. In practical scenarios, the number of locations where (a)
-  all relevant prior eigenvectors are significantly large, and (b)
-  other eigenvectors are close to zero, is limited. Consequently, the
-  clusterization of measurements arises as a natural consequence of
-  the pigeonhole principle, as there are more measurements available
-  than there are locations satisfying conditions (a) and (b).
-
-  The connection to Carath\'eodory's Theorem also builds on the focus
-  of measurement effort to the abovementioned select set of prior
-  eigenvectors. This concentration allows us to move from an
-  infinite-dimensional setting to finite dimensions, where the
-  conditions of Carath\'eodory's Theorem hold. We conclude that a
-  D-optimal design arises by weighting a small number of
-  measurements. If the number of allowed measurements is larger than
-  the small number dictated by Carath\'eodory's Theorem, we have
-  excessive weight on some measurements, which can be interpreted as
-  clusterization.
-
-\end{enumerate}
-
-\subsubsection{Implications}
-\input{implications}
-
-\subsubsection{Other Contributions}
-In Theorem \ref{thm:char}, we also show that D-optimal designs are
-best understood in the space of \emph{observations}, see Section
-\ref{section:vanishing} for a precise statement. This is in accordance
-with previous work by \cite{koval2020}, who showed that A-optimal
-designs are best constructed in the space of observations.
-
-In the process of proving Theorem \ref{thm:char} we prove and
-generalize several lemmas. Among those, is Lemma \ref{lemma:free},
-which is (to the the best of my knowledge) novel: We decompose a
-symmetric positive definite matrix \(M \in \mathbb{R}^{k \times k}\)
-with \(\ttr M = m \in \mathbb{N}\) as \(M = AA^t\), where \(A\in
-\mathbb{R}^{k \times m}\) has unit norm columns.
-
-\subsection{Limitations}\label{subsec:limitations}
-The main limitation of this study is that our generic model does not
-correspond to any specific real-life problem. Specifically, in its
-current form, our model does not allow point evaluations. Thus, while
-our model is generic enough to be analytically tractable, some may
-argue that our model is too far removed from any real application. To
-these claims I would answer that scientists have a long history of
-studying models that are bare-bones simplifications of real systems,
-e.g.~the Ising model \cite{cipra1987}, the Lorenz system \cite{brin},
-the Lotka-Volterra equations \cite{logan2006}, the Carnot engine
-\cite{kardar2007} and many others.
+\begin{quote}
+  Clusterization should not be confused with replication. Replication
+  requires that the experimentalist executes multiple trials under
+  circumstances that are \emph{nominally identical} \cite[Section
+    1.2.4]{morris2011}. Replication is commonly viewed as a beneficial
+  and even necessary aspect of optimal experimental design
+  \cite{fisher1949design, morris2011,
+    schafer2001replication}. Similarly to a design implementing
+  replication, a clustered design reduces the signal-to-noise ratio of
+  the repeated measurements \cite{telford2007brief}. The difference is
+  that a clustered design takes repeated measurements at the expense
+  of other quantities not measured at all.
+\end{quote}

### 2024-08-22 revision/rev3.tex
Only need to comment on tekentrup

--- 2024-08-12 latex/rev3.tex
@@
-\section{Reviewer 3}
+\section{Reviewer 3}\label{ref3}
 \RC The paper studies Bayesian D-optimal design (that is, the task of
@@
 
-\AR I made sure the \href{https://github.com/yairdaon/OED}{repository}
-is functioning and well-documented.
+\AR I made sure the repository is functioning and well-documented. See
+\url{https://github.com/yairdaon/OED}.
 
@@
 \AR I appreciate the suggestion. I added a section reproducing the
-abovementioned numerical experiments to the main article:
+abovementioned numerical experiments to the main article. See Section
+\ref{subsec:lemma_sims} reproduced below:
 
@@
 construction with the goal of testing numerically how prevalent are
-  clustered designs. To this end, we would like to generate random
-  prior eigenvalues $\lambda_j$, fix $m$ and $k$, find a D-optimal
+clustered designs. To this end, we would like to generate random prior
+eigenvalues $\lambda_j$, fix $m$ and $k$, find a D-optimal
 $\opt^*\opt$ and then utilize the construction of
@@
 To simplify things, we directly sample rank $\opt^*\opt$. We iterate
-  over the number of measurements $m \in \{4,\dots, 24\}$, and for
-  every $m$ we then iterate over $k:=\rank \obs^*\obs \in \{2,\dots,
+over the number of measurements $m \in \{4,\dots, 24\}$, and for every
+$m$ we then iterate over $k:=\rank \obs^*\obs \in \{2,\dots,
 m-1\}$. For each pair $m,k$ we repeat the following steps $N=5000$
@@
 \item Apply the construction of Lemma \ref{lemma:free} to calculate
-    $A$ such that $AA^t = M$, where $A$ has unit norm
-    columns. I.e.~find the optimal design $\opt$.
+  $A$ such that $AA^t = M$, where $A$ has unit norm columns. I.e.~find
+  the optimal design $\opt$.
 \item Since $A$ corresponds to $\opt$, its columns correspond to
-    measurement vectors. We call $A$ "clustered" if $A$ has two or
-    more identical columns (up to some numerical precision threshold).
+  measurement vectors. We call $A$ "clustered" if $A$ has two or more
+  identical columns (up to some numerical precision threshold).
 \end{enumerate}
@@
     \centering
-    \includegraphics[height=0.5\textwidth]{figs/simulations.png}
+    \includegraphics[height=0.5\textwidth]{../latex/figs/simulations.png}
     \caption{Fraction of clustered $A$ for $AA^t = M$ and $M$
@@
 \begin{quote}
-
+  DO THIS FML!!!!!!
 \end{quote}
@@
 instance: I would expect the dimensionality of the working domain
-(here the unit interval [0, 1]) to play a role, since larger dimension
-allow for higher freedom in placing the design points.
+(here the unit interval $[0, 1]$) to play a role, since larger
+dimension allow for higher freedom in placing the design points.
+
+\AR Say something??? FML
 
@@
 the current study.
-
-\begin{quote}
-  We choose a prior for the initial condition $u_0 \sim \mathcal{N}(0,
-  (-\Delta)^{-1})$, with homogeneous Dirichlet boundary
-  condition. This choice ensures the posterior is well-defined and
-  prior realizations are well-behaved, see \cite[Theorem 3.1 and Lemma
-    6.25]{Stuart10} for details.
-\end{quote}
 
@@
 
-\AR Basically I meant that its eigenvalues decrease "quickly". I
+\AR Basically I meant that its eigenvalues decay "quickly". I
 understand this is not a common or well-defined term so I removed the
@@
 
-\AR I do not expect this conclusion to hold: if noise is
-    present, even a repeated measurement will result in some
-    information gain and an increase in design criterion, since the
-    observation error is effectively reduced by a factor of
-    $1/\sqrt{2}$ for the repeated measurement.
+\AR I do not expect this conclusion to hold: if noise is present, even
+a repeated measurement will result in some information gain and an
+increase in design criterion. The reason is that for that particular
+measurement the signal-to-noise ratio increases: observation error is
+effectively reduced by a factor of $1/\sqrt{2}$ for the repeated
+measurement. I discuss this briefly in the revised manuscript:
+
+\begin{quote}
+  Clusterization should not be confused with replication. Replication
+  requires that the experimentalist executes multiple trials under
+  circumstances that are \emph{nominally identical} \cite[Section
+    1.2.4]{morris2011}. Replication is commonly viewed as a beneficial
+  and even necessary aspect of optimal experimental design
+  \cite{fisher1949design, morris2011, schafer2001replication}. Similarly
+  to a design implementing replication, a clustered design reduces the
+  signal-to-noise ratio of the repeated measurements
+  \cite{telford2007brief}. The difference is that a clustered design
+  takes repeated measurements at the expense of other quantities not
+  measured at all.
+\end{quote}
 
@@
 
-\AR I explain this in more detail in the paper:
+\AR I explain this in more detail in the revised manuscript:
 
@@
 \end{quote}
+
 \begin{figure}\label{fig:eigenvectors}

### 2024-08-25 latex/intro.tex
answered refs

@@
 
-%% REV 1, 2
+%% Replication 1
 Clusterization should not be confused with replication. Replication
@@
 
-%% Rev 2
+%% Replication 2
 For example, \cite{fisher1949design}, suggested repeating his famous
@@
 
-%% Rev 2
+%% Replication 3
 To further illustrate the difference between replication and
@@
 
-%% REV 1
-%% REV 2
-When clusterization arises, we believe it should serve as a warning
-signal to practitioners. In the inverse problem of the 1D heat
-equation, clusterization occurs primarily because Laplacian
-eigenvectors do not decay in $\Omega$. Consequently, measuring $u(x_1,
-T)$ at some point $x_1 \in \Omega$ provides information about
-$u(x_2,T)$ for distant points $x_2 \in \Omega$. Intuitively, this
-should not happen: for small $T$, the heat distribution at $x_1$
-should give very little knowledge on the heat distribution at
-$x_2$. This behavior stems from a well-known property of the heat
-equation: it allows information to spread instantly across the
-computational domain \cite{renardy2006PDE}. In reality, heat (and
-information) propagate at finite speeds. Of course, the known physical
-barrier for information spread is the speed of light, but we expect
-heat and information to spread considerably slower. Thus,
-clusterization may suggest that our physical model itself is at fault.
+%% Gaussian
+One potential cause of clusterization is our choice of prior. Gaussian
+priors, coupled with a Gaussian likelihood and a linear forward
+problem give rise to a closed form solution for the posterior via
+conjugacy. As we show in Section \ref{subsec:lemma_sims}, such
+assumptions generically give rise to clusterization. While an
+assumption of a Gaussian likelihood is standard, and rooted in the
+central limit theorem, an assumption of Gaussian prior is merely a
+matter of convenience. Therefore, we advise any practitioner who
+encounters clusterization to replace their Gaussian priors with
+non-Gaussian priors instead \cite{hosseini2017, hosseini2019}.
 
-%% We believe that the emergence of clusterization in this context might
-%% be non-physical, arising from the way the inverse problems we consider
-%% are phrased. Clusterization therefore indicates that the underlying
-%% mathematical / Bayesian model is overly permissive and fails to
-%% capture crucial physical constraints of the problem. We suggest that
-%% when clusterization occurs, practitioners should consider alternative
-%% models where information is localized in space and travels at finite
-%% speed in the medium. Such models may not only provide more physically
-%% accurate and meaningful results but may also mitigate the issue of
-%% clusterization.
+%% Linearization
+Another potential cause of clusterization is linearization. While in a
+real applications the model is not necessarily linear, some authors
+consider a linearized version of their model when seeking optimal
+designs \cite{fedorov1996, neitzel2019sparse}. Our analysis then shows
+that the linearity of the forward problem is also an important
+ingredient in giving rise to clustered desgins. Therefore, our advice
+to practitioners is to avoid linearization, and find D-optimal designs
+via other methods, e.g.~the sampling method of \cite{ryan2003}.
 
-%% REV 1
-%% REV 2
-Our choice of prior is perhaps an even greater contributor to
-clusterization. Our choice of Gaussian prior similarly implies
-information is shared between distant locations in $\Omega$. Thus, we
-suggest refraining from choosing Gaussian priors with inverse
-Laplacian covariance operators. Rather, non-Gaussian priors should be
-employed instead \cite{hosseini2017, hosseini2019}. So, clusterization
-may suggest we give more attention to the choice of prior.
-
-%% REV 1
+%% Avoid
 Given the above discussion, it is our belief that the methods
@@
 inherently bad on their own. Rather, we suggest that if a clustered
-design arises, then a practitioner should revisit their probabilistic
-and physical modeling assumptions. If the practitioner is confident in
-their assumptions, then clustered designs should be avoided only to
-the extent necessitated by the physical measuring apparatus.
+design arises, a practitioner should revisit their modeling
+assumptions --- specifically, their choice of prior and model
+linearization. If the practitioner is confident in their assumptions,
+then clustered designs should be avoided only to the extent
+necessitated by the physical measuring apparatus.
 
-Another For example, \cite{tekentrup2020} provides convergence analysis for
-various designs with differing space filling properties. They showed
-that convergence rate depends on how these designs fill space. We
-expect better convergence rates for D-optimal designs, even though
-D-optimal designs may demonstrate clusterization. The reason is that
-D-optimal designs explore space in a way tailored to the problem
-studied. Consequently, D-optimal designs cluster because they first
-aim to measure what matters most; see the discussion following
-Theorem \ref{thm:char} for details.
+%% Convergence
+Given that D-optimal designs suffer from clusterization, it is natural
+to ask whether D-optimal designs should be pursued at all. Other
+authors have established convergence results for decaying measurement
+error \cite{knapik2011}, space-filling measurements
+\cite{teckentrup2020} and randomly sampled designs
+\cite{nickl2023}. Unfortunately, D-optimal designs --- clustered
+designs in particular --- do not fall into any of these categories.
 
-%% be avoided at all. Nothing in our analysis indicates that we should
-%% expect any pathological behavior when utilizing D-optimal designs. On
-%% the contrary: we show that D-optimal designs (clustered or not) reduce
-%% uncertainty of prior covariance eigenvectors where it is
-%% highest. Thus, in a sense we hope to make precise in a future study,
-%% we expect convergence of for D-optimal designs to be fastest.
-
-%% practitioners should not try to avoid measurement
-%% clusterization. Rather, practitioners should take repeated
-%% measurements (e.g.~in MRI and borehole tomography), increase apparatus
-%% sensitivity (e.g.~in EIT), or take consecutive measurements (e.g.~in
-%% the 1D heat equation). Overall, we show that measurement
-%% clusterization is a natural and (almost) inevitable part of Bayesian
-%% D-optimal designs (but see disclaimer below).
-
-We expect D-optimal designs to judicious exploration of
-design space, as implemented by a D-optimal design to give better
-convergence compared to randomly sampling the design
-space \cite{knapik2011}.
-
-%% Clusterization is a peculiar phenomenon and it is perfectly reasonable
-%% for someone to argue against the D-optimality criterion based on the
-%% fact that it results in clustered designs. We have seen, however that
-%% there is a perfectly reasonable explanation for clusterization. We
-%% have shown that clusterization is an inevitable consequence of having
-%% a problem with some modes where uncertainty decays faster than others.
+Luckily, we can expect convergence from D-optimal designs (including
+clustered designs). Convergence holds in the following sense: the
+posterior uncertainty ellipsoid in $\hilo$ will contract to zero along
+every eigenfunction of $\fwd \prcov \fwd^*$. See Section
+\ref{subsub:convergence} for a proof based on Theorem \ref{thm:char}
+--- the main theorem of this manuscript.
 
@@
 current form, our model does not allow point evaluations. Thus, while
-our model is generic enough to be analytically tractable, some may
+our model is generic enough to be analytically tractable, one may
 argue that our model is too far removed from any real application. To
@@
 \cite{kardar2007} and many others.
+
+%% When clusterization arises, we believe it should serve as a warning
+%% signal to practitioners. In the inverse problem of the 1D heat
+%% equation, clusterization occurs primarily because Laplacian
+%% eigenvectors do not decay in $\Omega$. Consequently, measuring $u(x_1,
+%% T)$ at some point $x_1 \in \Omega$ provides information about
+%% $u(x_2,T)$ for distant points $x_2 \in \Omega$. Intuitively, this
+%% should not happen: for small $T$, the heat distribution at $x_1$
+%% should give very little knowledge on the heat distribution at
+%% $x_2$. This behavior stems from a well-known property of the heat
+%% equation: it allows information to spread instantly across the
+%% computational domain \cite{renardy2006PDE}. In reality, heat (and
+%% information) propagate at finite speeds. Of course, the known physical
+%% barrier for information spread is the speed of light, but we expect
+%% heat and information to spread considerably slower. Thus,
+%% clusterization may suggest that our physical model itself is at fault.
+
+%% We believe that the emergence of clusterization in this context might
+%% be non-physical, arising from the way the inverse problems we consider
+%% are phrased. Clusterization therefore indicates that the underlying
+%% mathematical / Bayesian model is overly permissive and fails to
+%% capture crucial physical constraints of the problem. We suggest that
+%% when clusterization occurs, practitioners should consider alternative
+%% models where information is localized in space and travels at finite
+%% speed in the medium. Such models may not only provide more physically
+%% accurate and meaningful results but may also mitigate the issue of
+%% clusterization.
+
+%% be avoided at all. Nothing in our analysis indicates that we should
+%% expect any pathological behavior when utilizing D-optimal designs. On
+%% the contrary: we show that D-optimal designs (clustered or not) reduce
+%% uncertainty of prior covariance eigenvectors where it is
+%% highest. Thus, in a sense we hope to make precise in a future study,
+%% we expect convergence of for D-optimal designs to be fastest.
+
+%% practitioners should not try to avoid measurement
+%% clusterization. Rather, practitioners should take repeated
+%% measurements (e.g.~in MRI and borehole tomography), increase apparatus
+%% sensitivity (e.g.~in EIT), or take consecutive measurements (e.g.~in
+%% the 1D heat equation). Overall, we show that measurement
+%% clusterization is a natural and (almost) inevitable part of Bayesian
+%% D-optimal designs (but see disclaimer below).

### 2024-08-25 latex/vanishing.tex
answered refs

@@
 
+%% Why
 \subsection{Answer to Question \ref{q:why}}\label{subsec:why}
@@
 places weight $>1$ on a single measurement vector.
+
+\subsubsection{Convergence}\label{subsub:convergence}
+Below we fulfill our promise from Section \ref{subsub:implications}
+and prove that the posterior uncertainty ellipsoid in $\hilo$ will
+contract to zero along every eigenfunction of $\fwd \prcov \fwd^*$. In
+our proof we ignore potential problems with conducting inference on
+function spaces which are not unique to D-optimal designs (see
+e.g.~\ref{owhadi2015} for more details).
+
+First, let $\opt_m$ a D-optimal design utilizing $m$
+measurements. Then, an immediate consequence of Theorem \ref{thm:char}
+is that $\rank \opt_m^*\opt_m = k \to \infty$ as the number of
+measurements $m \to \infty$. Moreover, part (5) of Theorem
+\ref{thm:char} ensures that the eigenvalues of the posterior
+covariance equal
+\begin{equation*}
+  \theta^{(k)}_i = \left ( \frac{\sum_{j=1}^k \lambda_j^{-1} +
+    \sigma^{-2}m}{k} \right )^{-1}, \text{ for $i\leq k$},
+\end{equation*}
+where we explicitly express the dependence of $k$ in the
+superscript. Now note that $\theta^{(k)}_i \leq \lambda_i$ for $i\leq
+k$. Specifically, for fixed $k$, we get $\theta^{(k)}_i \leq \lambda_k$
+for all $i\leq k$. Since $\lim_{k\to \infty} \lambda_k= 0$, we
+conclude that for all $i$, $\lim_{k\to\infty} \theta^{(k)}_i =
+0$. Since $m\to \infty$ implies $k\to \infty$ we have
+$\lim_{m\to\infty} \theta^{(k)}_i = 0$ for all $i$. Therefore,
+posterior uncertainty decays to zero for all eigenvectors.

### 2024-08-25 revision/rev1.tex
answered refs

@@
 
-\begin{quote}
+\begin{quote} % Avoid
 Given the above discussion, it is our belief that the methods
@@
 inherently bad on their own. Rather, we suggest that if a clustered
-  design arises, then a practitioner should revisit their probabilistic
-  and physical modeling assumptions. If the practitioner is confident in
-  their assumptions, then clustered designs should be avoided only to
-  the extent necessitated by the physical measuring apparatus.
+design arises, a practitioner should revisit their modeling
+assumptions --- specifically, their choice of prior and model
+linearization. If the practitioner is confident in their assumptions,
+then clustered designs should be avoided only to the extent
+necessitated by the physical measuring apparatus.
 \end{quote}
@@
 
-\AR In my view, this is a result of the unlocalized structure of
-Laplacian eigenvectors. I believe this is more of a problem with our
-physical models and choice of priors. See discussion below reproduced
-from the revised manuscript:
-
-\begin{quote}
-  When clusterization arises, we believe it should serve as a warning
-  signal to practitioners. In the inverse problem of the 1D heat
-  equation, clusterization occurs primarily because Laplacian
-  eigenvectors do not decay in $\Omega$. Consequently, measuring
-  $u(x_1, T)$ at some point $x_1 \in \Omega$ provides information
-  about $u(x_2,T)$ for distant points $x_2 \in \Omega$. Intuitively,
-  this should not happen: for small $T$, the heat distribution at
-  $x_1$ should give very little knowledge on the heat distribution at
-  $x_2$. This behavior stems from a well-known property of the heat
-  equation: it allows information to spread instantly across the
-  computational domain \cite{renardy2006PDE}. In reality, heat (and
-  information) propagate at finite speeds. Of course, the known
-  physical barrier for information spread is the speed of light, but
-  we expect heat and information to spread considerably slower. Thus,
-  clusterization may suggest that our physical model itself is at
-  fault.
-
-  Our choice of prior is perhaps an even greater contributor to
-  clusterization. Our choice of Gaussian prior similarly implies
-  information is shared between distant locations in $\Omega$. Thus,
-  we suggest refraining from choosing Gaussian priors with inverse
-  Laplacian covariance operators. Rather, non-Gaussian priors should
-  be employed instead \cite{hosseini2017, hosseini2019}. So,
-  clusterization may suggest we give more attention to the choice of
-  prior.
+\AR In my view this is a generic phenomenon that arises from a choice
+of a linear(ized) model and a Gaussian prior:
+\begin{quote}% Gaussian
+  One potential cause of clusterization is our choice of prior. Gaussian
+priors, coupled with a Gaussian likelihood and a linear forward
+problem give rise to a closed form solution for the posterior via
+conjugacy. As we show in Section \ref{subsec:lemma_sims}, such
+assumptions generically give rise to clusterization. While an
+assumption of a Gaussian likelihood is standard, and rooted in the
+central limit theorem, an assumption of Gaussian prior is merely a
+matter of convenience. Therefore, we advise any practitioner who
+encounters clusterization to replace their Gaussian priors with
+non-Gaussian priors instead \cite{hosseini2017, hosseini2019}.
+\newline %% Linearization
+Another potential cause of clusterization is linearization. While in a
+real applications the model is not necessarily linear, some authors
+consider a linearized version of their model when seeking optimal
+designs \cite{fedorov1996, neitzel2019sparse}. Our analysis then shows
+that the linearity of the forward problem is also an important
+ingredient in giving rise to clustered desgins. Therefore, our advice
+to practitioners is to avoid linearization, and find D-optimal designs
+via other methods, e.g.~the sampling method of \cite{ryan2003}.
 \end{quote}
@@
 
-\AR I think this is the same misunderstanding I commented on above;
-For any nonzero observation error $\sigma^2 > 0$, repeated
+\AR For any nonzero observation error $\sigma^2 > 0$, repeated
 measurements will indeed reduce the signal-to-noise ratio of the
 measurement and result in an increase of the design criterion. I
-briefly comment on this subject in the paragraph below:
+discuss this issue below:
 
-\begin{quote}
+\begin{quote} %% Replication 1
   Clusterization should not be confused with replication. Replication
@@
 and even necessary aspect of optimal experimental design
-  \cite{fisher1949design, morris2011,
-    schafer2001replication}. Similarly to a design implementing
-  replication, a clustered design reduces the signal-to-noise ratio of
-  the repeated measurements \cite{telford2007brief}. The difference is
-  that a clustered design takes repeated measurements at the expense
-  of other quantities not measured at all.
+\cite{fisher1949design, morris2011, schafer2001replication}. Similarly
+to a design implementing replication, a clustered design reduces the
+signal-to-noise ratio of the repeated measurements
+\cite{telford2007brief}. The difference is that a clustered design
+takes repeated measurements at the expense of other quantities not
+measured at all.
+\newline %% Replication 2
+For example, \cite{fisher1949design}, suggested repeating his famous
+milk and tea experiment in order "to be able to demonstrate the
+predominance of correct classifications in spite of occasional
+errors". Unfortunately, in the experiments we consider, replication is
+impossible. For example, in the MRI problem, we cannot generate an
+individual nominally identical to the one we wish to scan.
+\newline %% Replication 3
+To further illustrate the difference between replication and
+clusterization, consider an experiment measuring the effect of
+rainfall on grass growth \cite{fay2000rainfall}. The experiment
+involved four rainfall manipulation "treatments" (i.e.~simulating
+different timing and quantity of rainfall), each replicated three
+times over a different plot of land. Indeed, it seems reasonable for
+researchers to replicate the phenomenon they are trying to study. A
+clustered design in such an experiment would imply the researchers
+should take a repeated measurement \emph{on the same plot}, at the
+expense of measuring grass growth in other plots. To conclude our
+short discussion on replication vs.~clusterization: these are
+fundamentally different concepts and even though replication is quite
+intuitive, it is inapplicable to the inverse problems we consider in
+this manuscript.
 \end{quote}

### 2024-08-25 revision/rev2.tex
answered refs

\section{Reviewer 2}\label{ref2}
\subsection{Measurement Clusterization vs Repetition}
\RC I will start by saying measurement clusterisation is not
well-defined in this paper. On page 1, it is referred to as two sets
of measurements that are almost indistinguishable from one
another. However, the setup for Corollary 11 implicitly defines
measurement clusterisation as repeated measurements. The latter
definition matches the references of \cite{fedorov1996} and
\cite{nyberg2012}. So I will assume for the rest of this review that,
measurement clusterisation refers to repeated measurements.

\AR I appreciate the comment, I fixed the relevant line in the
manuscript.

\RC Nearly all textbooks on classical design of experiments
(e.g.~\cite[Section 1.2.4]{morris2011}) will feature a section with
the three words randomisation, blocking and repetition: the important
principles of design of experiments.  In these texts, repeated
measurements are considered to be a beneficial property of a
design. Interestingly, when optimal designs are found, they are often
found to have repeated measurements. The paper does not explain why
the author thinks repeated measurements (measurement clusterisation)
is an undesirable property.

\AR\label{rep} I added the following to the Introduction.

\begin{quote} %% Replication
  Clusterization should not be confused with replication. Replication
  requires that the experimentalist executes multiple trials under
  circumstances that are \emph{nominally identical} \cite[Section
    1.2.4]{morris2011}. Replication is commonly viewed as a beneficial
  and even necessary aspect of optimal experimental design
  \cite{fisher1949design, morris2011,
    schafer2001replication}. Similarly to a design implementing
  replication, a clustered design reduces the signal-to-noise ratio of
  the repeated measurements \cite{telford2007brief}. The difference is
  that a clustered design takes repeated measurements at the expense
  of other quantities not measured at all.
  \newline
For example, \cite{fisher1949design}, suggested repeating his famous
milk and tea experiment in order "to be able to demonstrate the
predominance of correct classifications in spite of occasional
errors". Unfortunately, in the experiments we consider, replication is
impossible. For example, in the MRI problem, we cannot generate an
individual nominally identical to the one we wish to scan.
\newline
To further illustrate the difference between replication and
clusterization, consider an experiment measuring the effect of
rainfall on grass growth \cite{fay2000rainfall}. The experiment
involved four rainfall manipulation "treatments" (i.e.~simulating
different timing and quantity of rainfall), each replicated three
times over a different plot of land. Indeed, it seems reasonable for
researchers to replicate the phenomenon they are trying to study. A
clustered design in such an experiment would imply the researchers
should take a repeated measurement \emph{on the same plot}, at the
expense of measuring grass growth in other plots. To conclude our
short discussion on replication vs.~clusterization: these are
fundamentally different concepts and even though replication is quite
intuitive, it is inapplicable to the inverse problems we consider in
this manuscript.
\end{quote}

\RC If one looks at, for example, \cite{fedorov1996} and
\cite{nyberg2012}, it is an undesirable property because they are
interested in sensor location, and two or more sensors cannot be in
the same location. However, there are many experiments where repeated
measurements are possible, and are therefore optimal

\AR I believe that even in an application where repeated experiments
are allowed they may be unintuitive. For example, in a medical imaging
application --- repeatedly measuring a "slice" in an MRI scan might
seem suboptimal to some.

\RC (if the model is correct, which it won’t be).

\AR I think this is a great point. My analysis implies one of the
causes to clusterization is linearization:
\begin{quote} %% Linearization
Another potential cause of clusterization is linearization. While in a
real applications the model is not necessarily linear, some authors
consider a linearized version of their model when seeking optimal
designs \cite{fedorov1996, neitzel2019sparse}. Our analysis then shows
that the linearity of the forward problem is also an important
ingredient in giving rise to clustered desgins. Therefore, our advice
to practitioners is to avoid linearization, and find D-optimal designs
via other methods, e.g.~the sampling method of \cite{ryan2003}.
\end{quote}

\RC An example, is a chemical engineering experiment where one wishes
to learn the relationship between a series of variables and yield of a
chemical reaction. The experiment will involve specifying the
variables, and measuring the yield (the observation) after a specified
period of time. This process is then duplicated with a potentially
different set of variables. It is perfectly possible to have repeated
variables.

\AR I agree, but what you are describing is replication. See my
comment on replication vs.~clusterization above.

\RC The author needs to make clear that they are considering
experiments where repeated measurements are impossible. Admittedly,
this consideration is implicit in the paper, for example, the examples
given on page 1. However, it needs to be made explicit in the Abstract
and the first paragraph of the paper.

\AR I do not refrain from considering experiments where repeated
measurements are possible.

\RC The paper explains why repeated measurements can be optimal
(answer to Question 3). The author should expand on this and draw the
link with how repeated measurements are, typically, regarded as
desirable in classical design of experiments for finite
dimensions.

\AR Great point, see my quote on replication vs.~clusterization
reproduced above.

\RC See for example, the discussion of the relationship between
Carath\'eodory’s Theorem and the number of support points in \cite[page
  139]{pronzatoPazman2013}. The support points are the unique design
points, so if this is less than $m$, then there are repeated design
points.

\AR I appreciate this insightful comment and reference. The arguments
presented there do not directly apply to the setup I suggest in my
paper. Thus, I have made some modifications:

\begin{quote} %% Caratheodory
  We can gain further insight to clusterization in our generic model,
  from Carath\'eodory's Theorem and the concentration on the first $k$
  eigenvectors of $\fwd \prcov \fwd^*$. We present a short discussion
  adapting arguments presented in \cite[Chapter 3]{silvey1980} and
  \cite[Section 5.2.3]{pronzatoPazman2013}.

  Consider $\opt$ a D-optimal design under our generic model.  As
  instructed by Theorem \ref{thm:char}, we ignore all but the first
  $k$ eigenvectors of $\fwd \prcov \fwd$. Thus, we replace $\hilo$
  with $\hilo^{(k)}$ --- the $k$-dimensional subspace spanned by the
  first $k$ eigenvectors of $\fwd^*\prcov\fwd$. Let
  \begin{equation*}
    \mathcal{M} := \conv \{\meas \meas^* : \meas\in \hilo^{(k)}, \|\meas\|=1\},
    %\hilo^{(k)}\},
  \end{equation*}
  where $\conv$ denotes the convex hull of a set. The set
  $\mathcal{M}$ contains only positive-definite operators on a
  $k$-dimensional vector space. Hence $\mathcal{M}$ lives in a
  $k(k+1)/2$-dimensional vector space. Since $\opt^*\opt =
  \sum_{i=1}^m \meas_i\meas_i^*$ for $\meas_i \in \hilo^{(k)}$, it is
  easy to verify that $\frac1m \opt^*\opt \in \mathcal{M}$.  Recall
  Carath\'eodory's Theorem:
  \begin{theorem*}[Carath\'eodory]
    Let $X \subseteq \mathbb{R}^n, X \neq \phi$ and denote $\conv (X)$
    the convex hull of $X$. For every $x \in \conv (X)$, $x$ is a convex
    combination of at most $n+1$ vectors in $X$.
  \end{theorem*}
  Carath\'eodory's Theorem implies that there exist $\meas_i$ and
  $\alpha_i$ such that
  \begin{equation*}
    \opt^*\opt = \sum_{i=1}^I \alpha_i \meas_i\meas_i^*,
  \end{equation*}
  where $\|\meas_i\|=1, \sum\alpha_i = m, \alpha_i \geq 0$ and $I =
  \frac{k(k+1)}{2} + 1$. We can thus write $\opt$ as:
  %(note that we do not require $\meas_i$ to be orthogonal to each
  %other):

  \[
  \opt =
  \left[
    \begin{array}{ccc}
      \horzbar & \sqrt{\alpha_i} \meas^*_1 & \horzbar \\
      \horzbar & \sqrt{\alpha_2} \meas_2^* & \horzbar \\
      & \vdots    &          \\
      \horzbar & \sqrt{\alpha_I} \meas_I^* & \horzbar \\
    \end{array}
    \right].
  \]

  Unfortunately, $\opt$ is not a valid design, since its rows do not
  have unit norm. Still, the above representation of $\opt$ is useful:
  If $m > \frac{k(k+1)}{2} + 1$, then $\alpha_i > 1$ for some $1\leq i
  \leq I$.  Thus, we can view $\opt$ as a clustered design, since it
  places weight $>1$ on a single measurement vector.
\end{quote}

\subsection{Context}
\RC The paper lacks discussion on context. I read the answers to the
three questions and thought “So what?”  What are the implications to
statistical practice, or experimental science, of these findings?

\AR I really appreciate this comment. I added a section highlighting
the implications of this study. Since it is too long to reproduce
here, see Section \ref{subsub:implications} in the revised manuscript.

\RC The presentation of the paper does not help. The results are
presented with no explanation or intuition. It might be better to
relegate much of the mathematical detail to Supplementary Material and
use the main manuscript for exposition.

\AR Thanks for this comment. I considerably expanded the
  introduction and delegated many proofs to the Supplementary.

\subsection{Avoiding measurement clusterisation}
\RC The approaches to avoid measurement clusterisation, described on page
3, are described as pragmatic and "fundamentally altering the optimal
design problem".

\RC Firstly, this is not strictly true. Suppose measurements correspond to
time, it is not possible to take two or more observations at the same
time, and there should be a minimum time period between measurements.
Implementing this minimum time period is not pragmatic, rather it is
characterising the application correctly. One could imagine a similar
thing occurs with the electrode locations on the skin in impedance
tomography: the physical size of the sensor imposes a minimum
distance between sensors.

\AR To a certain extent I agree: the impossibility of taking identical
measurements is indeed natural. However, this is not encoded in the
mathematical formulation of the problem. One can then take one of two
approaches: either accept what the mathematics suggest, even if it
counterintuitive, or find ways to circumvent the mathematics because
the result seems non-intuitive. I expanded this discussion in the
revised manuscript. See below and the discussion preceding it in
Section \ref{subsub:implications} of the revised manuscript.

\begin{quote}%% Avoid
  Given the above discussion, it is our belief that the methods
  suggested by other authors to avoid clusterization are merely
  overlooking the problem. We do not view clustered designs as
  inherently bad on their own. Rather, we suggest that if a clustered
  design arises, then a practitioner should revisit their probabilistic
  and physical modeling assumptions. If the practitioner is confident in
  their assumptions, then clustered designs should be avoided only to
  the extent necessitated by the physical measuring apparatus.
\end{quote}

\RC Secondly, the paper seems to suggest “imposing correlations
between observations” as a solution. However, isn’t this imposition
pragmatic as well? I suppose it is argued that the $\obs \eps'$ term
is accounting for model error: but it is still pragmatic: the data
acquisition given by equation (1) is not the true data-generating
process.

\AR I apologize for this misunderstanding and appreciate the
comment. I have revised the manuscript to reflect the fact that I do
not support imposing correlations in the paper. I do argue against
linearization (see comment on linearization above) and the use of Gaussian priors (see below):

\begin{quote} %% Gaussian
One potential cause of clusterization is our choice of prior. Gaussian
priors, coupled with a Gaussian likelihood and a linear forward
problem give rise to a closed form solution for the posterior via
conjugacy. As we show in Section \ref{subsec:lemma_sims}, such
assumptions generically give rise to clusterization. While an
assumption of a Gaussian likelihood is standard, and rooted in the
central limit theorem, an assumption of Gaussian prior is merely a
matter of convenience. Therefore, we advise any practitioner who
encounters clusterization to replace their Gaussian priors with
non-Gaussian priors instead \cite{hosseini2017, hosseini2019}.
\end{quote}

\subsection{Minor comments}
\RC The author uses two different forms of asterisk: one for
optimality of the design (e.g. Definition 3) and one for the
adjoint. Suggest changing the one for optimality.

\AR I changed the notation for a D-optimal design from $\obs^\star$ to
$\opt$.

\RC I do not like the use of the term measurements as used in the
paper. The measurements are the $\data$’s since these are a result of
measuring a physical quantity. However, the paper refers to the
quantities that are controlled as the measurements.

\AR I was not able to find a better term to replace "measurements".

\RC Proof of Proposition 12 uses $\eps$ as part of the regularisation
trick but this has been used previously for error (e.g. equation (1)).

\AR I replaced $\eps$ with $\nu$ and moved the entire proof to the
Supplementary.

%% Firstly, the argument in \cite{pronzatoPazman2013} relies on the
%% assumption that $\theta$ --- the parameter we seek to infer --- is
%% $p$-dimensional. However, the setup I suggest is
%% infinite-dimensional. E.g.~in the generic model I propose, the
%% parameter is in $\hilo$, which is assumed an infinite-dimensional
%% Hilbert space. Similarly, the initial condition $u_0$ we try to infer
%% in the inverse problem of the heat equation is in some function space
%% over $\Omega=[0,1]$. Even if we discretize $\Omega$ and consider a
%% finite-dimensional problem, the number of discretization points would
%% be at least a hundred. So even the bound $m \leq p=100$ suggested in
%% \cite[Section 5.3.1]{pronzatoPazman2013} is not very meaningful when
%% we allow only $\approx 6$ measurements. Indeed, discretization points
%% are cheap and measurements are expensive so the number of
%% discretization points will always be considerably larger than the
%% number of allowed measurements. Thus, I do not see how the bound
%% suggested will be informative. Respectfully I choose not to explore
%% your suggestion in the manuscript.

%% We can bypass the fact that $p$ tends to be large by recalling that
%% Theorem \ref{thm:char} implies

### 2024-08-25 revision/rev3.tex
answered refs

@@
 
-\begin{quote}
+\begin{quote} %% Simulate
   In the proof of Theorem \ref{thm:char} we utilize Lemma
@@
   construction with the goal of testing numerically how prevalent are
-clustered designs. To this end, we would like to generate random prior
-eigenvalues $\lambda_j$, fix $m$ and $k$, find a D-optimal
+  clustered designs. To this end, we would like to generate random
+  prior eigenvalues $\lambda_j$, fix $m$ and $k$, find a D-optimal
   $\opt^*\opt$ and then utilize the construction of
   Theorem~\ref{thm:char} and Lemma~\ref{lemma:free} to find $\opt$.
-
+  \newline
   To simplify things, we directly sample rank $\opt^*\opt$. We iterate
-over the number of measurements $m \in \{4,\dots, 24\}$, and for every
-$m$ we then iterate over $k:=\rank \obs^*\obs \in \{2,\dots,
+  over the number of measurements $m \in \{4,\dots, 24\}$, and for
+  every $m$ we then iterate over $k:=\rank \obs^*\obs \in \{2,\dots,
   m-1\}$. For each pair $m,k$ we repeat the following steps $N=5000$
@@
   \item Apply the construction of Lemma \ref{lemma:free} to calculate
-  $A$ such that $AA^t = M$, where $A$ has unit norm columns. I.e.~find
-  the optimal design $\opt$.
+    $A$ such that $AA^t = M$, where $A$ has unit norm
+    columns. I.e.~find the optimal design $\opt$.
   \item Since $A$ corresponds to $\opt$, its columns correspond to
-  measurement vectors. We call $A$ "clustered" if $A$ has two or more
-  identical columns (up to some numerical precision threshold).
+    measurement vectors. We call $A$ "clustered" if $A$ has two or
+    more identical columns (up to some numerical precision threshold).
   \end{enumerate}
-We then calculate the fraction of clustered designs of the simulations
-we ran, for each pair $m,k$. Clusterization occurred at high rates
-($>99.9\%$) whenever $m-k > 1$; see Fig.~\ref{fig:sim_AAt}. Hence, in
-these simulations, clusterization is a generic property. However, when
-$m-k = 1$, clusterization does not occur. We do not why this is so and
-further investigation into this phenomenon is out of scope for the
-current study.
-
+  We then calculate the fraction of clustered designs of the
+  simulations we ran, for each pair $m,k$. Clusterization occurred at
+  high rates ($>99.9\%$) whenever $m-k > 1$; see
+  Fig.~\ref{fig:sim_AAt}. Hence, in these simulations, clusterization
+  is a generic property. However, when $m-k = 1$, clusterization does
+  not occur. We do not why this is so and further investigation into
+  this phenomenon is out of scope for the current study.
+  \newline
   Full results are located in the \texttt{simulations.csv} file within
   the accompanying \href{https://github.com/yairdaon/OED}{repository}.
-Code implementing the experiments described above is located in module
-\texttt{zeros.py} of said repository. Runtime should be less than 30
-minutes on any reasonably modern laptop (it took 12 minutes on the
-author's laptop).
+  Code implementing the experiments described above is located in
+  module \texttt{zeros.py} of said repository. Runtime should be less
+  than 30 minutes on any reasonably modern laptop (it took 12 minutes
+  on the author's laptop).
 \end{quote}
@@
 
-\AR I added a discussion on this subject. I do not argue against the
-use of clustered designs imply that clustered designs, but I believe
-that when they arise it should be a warning signal to a practitioner
-regarding the validity of their physical model and Bayesian
-assumptions:
-
-\begin{quote}
-  DO THIS FML!!!!!!
+\AR I added a discussion on this subject. In the manuscript this
+discussion is broken in two since the second part relies on Theorem
+\ref{thm:char}. Thus, the first part is in the Introduction and the
+second part follows the statement and proof of the abovementioned
+theorem. I believe the level of discussion is adequate and view a
+deeper analysis of convergence rates as out of scope for the current
+study.
+\begin{quote}%%Convergence
+  Given that D-optimal designs suffer from clusterization, it is
+  natural to ask whether D-optimal designs should be pursued at
+  all. Other authors have established convergence results for decaying
+  measurement error \cite{knapik2011}, space-filling measurements
+  \cite{teckentrup2020} and randomly sampled designs
+  \cite{nickl2023}. Unfortunately, D-optimal designs --- clustered
+  designs in particular --- do not fall into any of these categories.
+  \newline
+  Luckily, we can expect convergence from D-optimal designs (including
+  clustered designs). Convergence holds in the following sense: the
+  posterior uncertainty ellipsoid in $\hilo$ will contract to zero
+  along every eigenfunction of $\fwd \prcov \fwd^*$. See Section
+  \ref{subsub:convergence} for a proof based on Theorem \ref{thm:char}
+  --- the main theorem of this manuscript.
+  \newline
+  Below we fulfill our promise from Section \ref{subsub:implications}
+  and prove that the posterior uncertainty ellipsoid in $\hilo$ will
+  contract to zero along every eigenfunction of $\fwd \prcov
+  \fwd^*$. In our proof we ignore potential problems with conducting
+  inference on function spaces which are not unique to D-optimal
+  designs (see e.g.~\ref{owhadi2015} for more details).
+  \newline
+  First, let $\opt_m$ a D-optimal design utilizing $m$
+  measurements. Then, an immediate consequence of Theorem
+  \ref{thm:char} is that $\rank \opt_m^*\opt_m = k \to \infty$ as the
+  number of measurements $m \to \infty$. Moreover, part (5) of Theorem
+  \ref{thm:char} ensures that the eigenvalues of the posterior
+  covariance equal
+  \begin{equation*}
+    \theta^{(k)}_i = \left ( \frac{\sum_{j=1}^k \lambda_j^{-1} +
+      \sigma^{-2}m}{k} \right )^{-1}, \text{ for $i\leq k$},
+  \end{equation*}
+  where we explicitly express the dependence of $k$ in the
+  superscript. Now note that $\theta^{(k)}_i \leq \lambda_i$ for
+  $i\leq k$. Specifically, for fixed $k$, we get $\theta^{(k)}_i \leq
+  \lambda_k$ for all $i\leq k$. Since $\lim_{k\to \infty} \lambda_k=
+  0$, we conclude that for all $i$, $\lim_{k\to\infty} \theta^{(k)}_i
+  = 0$. Since $m\to \infty$ implies $k\to \infty$ we have
+  $\lim_{m\to\infty} \theta^{(k)}_i = 0$ for all $i$. Therefore,
+  posterior uncertainty decays to zero for all eigenvectors.
 \end{quote}
@@
 dimension allow for higher freedom in placing the design points.
-
-\AR Say something??? FML
-
-\RC Further, a dimensionality effect should also arise directly from
-Theorem 1 via the prior covariance eigenvalues: in the experiment the
-prior covariance operator is set to be equal to the inverse Laplace
-operator $(-\Delta)^{-1}$. In d-dimensional domains, its eigenvalues
+Further, a dimensionality effect should also arise directly from
+Theorem 1\footnote{Now denoted Theorem \ref{thm:char}.} via the prior
+covariance eigenvalues: in the experiment the prior covariance
+operator is set to be equal to the inverse Laplace operator
+$(-\Delta)^{-1}$. In d-dimensional domains, its eigenvalues
 $\lambda_j$ are known to follow Weyl’s asymptotics, growing as
@@
 dimensions because (1) it would require too much exposition and make
-me digress from the main points I wanted to convey and (2) I view
-implementing experiments in higher dimensions as out of the scope of
-the current study.
+me digress from the main points I wanted to convey and (2) due to
+technical difficulties in installing packages for inverse problems
+\cite{attia2023pyoed, VillaPetraGhattas16} I had to implement the
+inverse problem of the 1D heat equation from scratch, so I decided to
+leave out experiments in higher dimensions.
 
@@
   \centering
-  \includegraphics[height=0.5\textwidth]{figs/dst_modelError4.png}
+  \includegraphics[height=0.5\textwidth]{../latex/figs/dst_modelError4.png}
   \caption{Model correlation mitigates clusterization. We add a
@@
 
-\AR I removed the statement of Theorem 1 from the introduction.
+\AR I removed the statement of Theorem 1 from the introduction. It is
+Theorem \ref{thm:char} now.
 
 \RC Denoting the optimal design in Theorem 1 by simply $\obs$ is
-somewhat confusing, cf. the equation in the second item. Perhaps a
+somewhat confusing, cf.~the equation in the second item. Perhaps a
 separate notation such as $\bar{\obs}$ would help readability.
@@
 
-\begin{quote}
+\begin{quote}%% Pathologies
 It is worth noting that by the nonnegativity of the KL divergence,
@@
 new measurement e.g.~if the prior variance vanishes on some
-  eigenvector and a measurement is taken on said eigenvector. We do
-  not expect a measurement to generate no information gain whatsoever
-  in any realistic scenario, and ignore such pathologies.
+eigenvector and a measurement is taken on said eigenvector. We do not
+expect a measurement to generate no information gain whatsoever in any
+realistic scenario, and ignore such pathologies.
 \end{quote}
@@
 
-\begin{quote}
+\begin{quote} %% Replication
 Clusterization should not be confused with replication. Replication
@@
 present paper since I was not able to install common packages for
-inverse problems \cite{attia2023pyoed, VillaPetraGhattas16,
-  VillaPetraGhattas18, VillaPetraGhattas21} on my machine.
+inverse problems \cite{attia2023pyoed, VillaPetraGhattas16} on my
+machine.
 
@@
 spaces. Point evaluations can be approximated well in
-e.g.~$L^2(\Omega)$ so in my opinion this is not a deal breaker. The
-only caveat is that upon approximating point evaluations e.g.~by
-$R(x;t) = t\mathbb{1}_{-t/2, t/2}(x)$ we will need to change the norm
-constraints to the something different (i.e.~$\|R(\cdot;t\|$) but that
-does not change the analysis. I barely mentioned this issue in the
-paper since I think it is confusing and adds little context. See,
-however, a footnote I added which is reproduced below:
+e.g.~$L^2(\Omega)$ so in my opinion this is not a deal breaker. For
+example, we can replace point evaluations by $R(x;t) =
+t\mathbb{1}_{-1/2t, 1/2t}(x)$, where $\mathbb{1}$ is an indicator
+function. We will then need to change the norm constraints on
+measurement vectors from unit to $\|R(\cdot;t\|^2 = \int R^2(x;t)dx =
+t$ norm but that does not change the analysis as long as $t$ is
+fixed. I barely mentioned this issue in the paper since I think it is
+confusing and adds little context. See a footnote I added:
 
-\begin{quote}
+\begin{quote} %% Footnote
   The alert reader will likely ask how do we reconcile point
@@
   explanation to the measurement clusterization we observed for the
-  inverse problem of the heat equation (Fig.~\ref{fig:eigenvectors}).
-
+  inverse problem of the heat equation, see Section
+  \ref{subsub:clusterization1} below. We also suggest a generic
+  explanation for clusterization, see Section
+  \ref{subsub:clusterization2}.
+  \newline
   Consider $\fwd$ and $\prcov$ from \emph{the inverse problem of the
@@
   $k+1$ and above.
-
+  \newline
   Of course, in a real life problem --- such as the inverse problem of
@@
   eigenvectors.
-
+  \newline
   The abovementioned balance is explored in
@@
   eigenvector is zero at $x=0.5$.
-
+  \newline
   Now we can understand measurement clusterization for the inverse
@@
   pigeonhole principle.
-
 \end{quote}
@@
   \centering
-  \includegraphics[width=\textwidth]{figs/eigenvectors_dst_scaled.png}
+  \includegraphics[width=\textwidth]{../latex/figs/eigenvectors_dst_scaled.png}
   \caption{D-optimal measurement locations ($m=4$ measurements) and

### 2024-08-25 revision/refs.tex
answered refs
Now with refs

@@
-\documentclass[draft]{article}
+\documentclass{article}
 \def\papertitle{Measurement Clusterization in Bayesian D-optimal Designs in Infinite Dimensions}
@@
 % Legend
-\hfill {\bfseries RC:} \textbf{\textit{Reviewer Comment}},\(\quad\) AR: \emph{Author Response}, \(\quad\square\) Manuscript text
+\hfill {\bfseries RC:} \textbf{\textit{Referee Comment}},\(\quad\) AR: \emph{Author Response}, \(\quad\square\) Manuscript text
 
@@
 of my paper and added new insights as well. I sincerely appreciate the
-option to revise and resubmit. Please see below my answers to the
-comments made by editor in chief (Section \ref{eic}), associate editor
-(Section \ref{ae}) and three referees (Sections \ref{ref1}, \ref{ref2}
-and \ref{ref3}).
+option to revise and resubmit this manuscript. Please see below my
+answers to the comments made by the editor in chief (Section
+\ref{eic}), associate editor (Section \ref{ae}) and three referees
+(Sections \ref{ref1}, \ref{ref2} and \ref{ref3}).
 
-%% \section{Editor in Chief}\label{eic}
-%% \RC Personally I like the paper, but it represents a somehow unusual
-%% submission for Bayesian Analysis, and I am worried about it being of
-%% sufficient general interest for the broad readership of the
-%% journal. Still, in agreement with the AE, we decided to give the
-%% author the possibility to revise the paper. However, I must stress
-%% that there is no guarantee of a successful outcome, which will solely
-%% depend on the quality of the revision. In particular:
+\section{Editor in Chief}\label{eic}
+\RC Personally I like the paper, but it represents a somehow unusual
+submission for Bayesian Analysis, and I am worried about it being of
+sufficient general interest for the broad readership of the
+journal. Still, in agreement with the AE, we decided to give the
+author the possibility to revise the paper. However, I must stress
+that there is no guarantee of a successful outcome, which will solely
+depend on the quality of the revision. In particular:
 
-%% \RC the Introduction is to be made accessible to the whole BA
-%%     readership;
+\RC the Introduction is to be made accessible to the whole BA
+    readership;
 
-%% \AR I have considerably extended the introduction.
+\AR I have considerably extended the introduction. See revised
+manuscript, diff file and detailed answers to referees' comments
+below.
 
-%% \RC the motivation and the practical statistical implications are to
-%% be developed and discussed;
+\RC the motivation and the practical statistical implications are to
+be developed and discussed;
 
-%% \AR I added a section on the implications:
+\AR I added a section on the implications. The main takeaways are: (1)
+avoid linearization and Gaussian priors, and (2) clusterization can
+inform you of the number of relevant d.o.f in the problem.
 
-%% %\begin{quote}
-%%   %\input{implications}
-%% %\end{quote}
+\RC the contribution has to be better contextualized with respect to
+the literature in both inverse problems and D-optimal designs;
 
-%% \RC the contribution has to be better contextualized with respect to
-%% the literature in both inverse problems and D-optimal designs;
+\AR I added discussion on replication, contrasting it with
+clusterization. I added convergence proof for D-optimal designs
+(albeit without explicit rate) per Referee 2's request.
 
-%% \RC numerical evidence and code, mentioned in the paper but actually
-%% not provided, need to be made available;
+\RC numerical evidence and code, mentioned in the paper but actually
+not provided, need to be made available;
 
-%% \AR I added a section describing numerical experiments and their
-%% results. I also added documentation to the accompanying
-%% \href{https://github.com/yairdaon/OED}{repository} and made sure all
-%% scripts there run.
+\AR I added a section describing numerical experiments and their
+results. I also added documentation to the accompanying
+\href{https://github.com/yairdaon/OED}{repository} and made sure all
+scripts there run.
 
-%% \RC the mathematical derivations have to be watertight and some
-%% clarifications are needed in this respect (see Report of Referee 3).
+\RC the mathematical derivations have to be watertight and some
+clarifications are needed in this respect (see Report of Referee 3).
 
-%% \AR I am not sure which specific comment of Reviewer 3 this refers to,
-%% please see my answers to Reviewer 3 below.
+\AR I am not sure which specific comment of Referee 3 this refers to,
+please see my answers to Referee 3 below.
 
-%% \section{Associate Editor}\label{ae}
-%% \RC The three referees found merits and interest in your work, but also
-%% pointed out several major issues to be addressed. Of particular
-%% concern to me, the paper's exposition jumps too quickly into the
-%% formulation / maths without giving a proper intuition and
-%% context. This makes the paper less accessible to readers who aren't
-%% already quite familiar with inverse problems (which are introduced
-%% with very little preamble, and it's not even immediately explicit what
-%% are the parameters being learned) and optimal experimental design.
+\section{Associate Editor}\label{ae}
+\RC The three referees found merits and interest in your work, but also
+pointed out several major issues to be addressed. Of particular
+concern to me, the paper's exposition jumps too quickly into the
+formulation / maths without giving a proper intuition and
+context. This makes the paper less accessible to readers who aren't
+already quite familiar with inverse problems (which are introduced
+with very little preamble, and it's not even immediately explicit what
+are the parameters being learned) and optimal experimental design.
 
-%% \AR I have added considerably to the introduction. See revised
-%% manuscript and diff file.
+\AR I have added considerably to the introduction. I hope you will
+find the transfer to the maths smoother now.
 
-%% \RC A referee makes valid points as to whether clusterization of
-%% measurements is a problem in general, when in fact in can be an
-%% appealing feature in some settings, and another referee asks whether
-%% D-optimal designs should be abandoned altogether. The exposition
-%% needs to discuss these high-level issues, and introduce the related
-%% literature and results appropriately, to give the work proper
-%% context.
+\RC A referee makes valid points as to whether clusterization of
+measurements is a problem in general, when in fact in can be an
+appealing feature in some settings,
 
-%% \RC Altogether, I think that your work has the potential to make a
-%% good contribution to Bayesian analysis, but as it currently stands
-%% significant work is needed.
+\AR I added a discussion contrasting replication vs.~clusterization.
+
+\RC and another referee asks whether D-optimal designs should be
+abandoned altogether.
+
+\AR I added comments on my suggestions to practitioners who encounter
+clusterization: avoid Gaussian priors and linearization, separate
+observations only to the extent required by the physical measurement
+apparatus.
+
+\RC The exposition needs to discuss these high-level
+issues, and introduce the related literature and results
+appropriately, to give the work proper context.
+
+\AR I also added a discussion on convergence per Referee 2's request
+contrasting D-optimal designs with other convergence results.
+
+\RC Altogether, I think that your work has the potential to make a
+good contribution to Bayesian analysis, but as it currently stands
+significant work is needed.
+
+\AR Again, I appreciate the time and effort invested by the referees
+and editorial team. Thanks for the opportunity to revise and improve this manuscript.
 
 \input{rev1}
+
 \input{rev2}
+
 \input{rev3}
@@
 \bibliographystyle{apalike}
-\bibliography{../../lib.bib}
+\bibliography{/Users/yairdaon/work/lib.bib}
 

### 2024-08-25 latex/ba.tex
answered refs
Now with abstract

@@
   measurement clusterization is a generic trait of D-optimal designs
-  with independent Gaussian measurement errors, and prove that
-  introducing correlations among measurement error terms mitigates
-  clusterization. \textbf{(b)} We characterize D-optimal designs as
+  for linear inverse problems with independent Gaussian measurement
+  errors and a Gaussian prior. \textbf{(b)} We prove that introducing
+  correlations among measurement error terms mitigates
+  clusterization. \textbf{(c)} We characterize D-optimal designs as
   reducing uncertainty across a subset of prior covariance
-  eigenvectors. Finally, \textbf{(c)} We leverage this
-  characterization to argue that measurement clusterization arises as
-  a consequence of the pigeonhole principle: when more measurements
-  are taken than there are locations where the select eigenvectors are
-  large and others are small --- clusterization occurs.
+  eigenvectors. \textbf{(d)} We leverage this characterization to
+  argue that measurement clusterization arises as a consequence of the
+  pigeonhole principle: when more measurements are taken than there
+  are locations where the select eigenvectors are large and others are
+  small --- clusterization occurs. Finally, we use our analysis to
+  argue against the use of Gaussian priors with linearized physical
+  models when seeking a D-optimal design.
 \end{abstract}
@@
 \bibliographystyle{ba}
-\bibliography{/Users/yairdaon/work//lib.bib}
+\bibliography{/Users/yairdaon/work/lib.bib}
 

### 2024-08-29 latex/old/old_flat.tex
Do not erase old_flatgit add old_flat.tex

--- 2024-07-25 latex/old.tex
@@
 \documentclass[ba]{imsart}
-
-% \documentclass[ba,preprint]{imsart}% use this for supplement article
 \usepackage{amsmath}
@@
 
-%% first: blue, second: red, third: green, fourth: black,
-%% fifth:magenta
 \begin{figure}
     \centering
-    \includegraphics[height=0.5\textwidth]{figs/example.png}
+    \includegraphics[height=0.5\textwidth]{../figs/example.png}
     \caption{Measurement clusterization for optimal designs when
@@
     \centering
-    \includegraphics[width=\textwidth]{figs/eigenvectors.png}
+    \includegraphics[width=\textwidth]{../figs/eigenvectors.png}
     \caption{D-optimal measurement locations ($m=4$ measurements) and
@@
 non-zero model error $\modcov$ --- answering Question \ref{q:mitigate}
-posed in the Introduction.% Section \ref{section:intro}.
-\section{D-Optimal Designs Without Model Error}\label{section:vanishing}
+posed in the Introduction. \section{D-Optimal Designs Without Model Error}\label{section:vanishing}
 Our goal in this section is to prove Theorem \ref{thm:char} which
@@
 such a procedure is illustrated in Fig.~\ref{fig:optimal vs not}.
-%% \begin{figure}%{r}{0.25\textwidth} 
-%%     \centering
-%%     \includegraphics[width=5cm, height=5cm]{cylinders.jpg}
-%%     \caption{Three graduated lab cylinders, corresponding to three
-%%       eigenvectors. Prior eigenvalues not shown.}
-%%     \label{fig:cylinder}
-%% \end{figure}
 

### 2024-09-02 latex/numerix.tex
Only need to comment on tekentrup
Everything except for the intro

@@
 clustered designs. To this end, we would like to generate random prior
-eigenvalues $\lambda_j$, fix $m$ and $k$, find a D-optimal
-$\opt^*\opt$ and then utilize the construction of
+eigenvalues $\lambda_j$, fix $m$ and $k$, find what a D-optimal
+$\opt^*\opt$ should be, and then utilize the construction of
 Theorem~\ref{thm:char} and Lemma~\ref{lemma:free} to find $\opt$.
 
-To simplify things, we directly sample rank $\opt^*\opt$. We iterate
-over the number of measurements $m \in \{4,\dots, 24\}$, and for every
-$m$ we then iterate over $k:=\rank \obs^*\obs \in \{2,\dots,
-m-1\}$. For each pair $m,k$ we repeat the following steps $N=5000$
-times:
+To simplify things, we directly generate $\opt^*\opt$. We iterate over
+the number of measurements $m \in \{4,\dots, 24\}$, and for every $m$
+we then iterate over $k:=\rank \opt^*\opt \in \{2,\dots, m-1\}$. For
+each pair $m,k$ we repeat the following steps $N=5000$ times:
 \begin{enumerate}
 \item Generate random diagonal $D\in \mathbb{R}^{k\times k}$ with
-  entries $\log (d_i) \sim \mathcal{N}(50,15)$ and normalize so that
-  $\ttr D = m$. 
+  entries $\log (d_i) \sim \mathcal{N}(50,15)$.
+\item Normalize $D$ that $\ttr D = m$.
 \item Conjugate $D$ by a random orthogonal matrix to form a positive
@@
 \item Apply the construction of Lemma \ref{lemma:free} to calculate
-  $A$ such that $AA^t = M$, where $A$ has unit norm columns. I.e.~find
-  the optimal design $\opt$.
-\item Since $A$ corresponds to $\opt$, its columns correspond to
+  $A$ such that $AA^t = M$, where $A$ has unit norm columns. $A^t$ is
+  our optimal design $\opt$.
+\item Since $A^t$ corresponds to $\opt$, its columns correspond to
   measurement vectors. We call $A$ "clustered" if $A$ has two or more
-  identical columns (up to some numerical precision threshold).
+  identical columns (up to some numerical precision threshold,
+  i.e.~$10^{-5}$).
 \end{enumerate}
@@
 these simulations, clusterization is a generic property. However, when
-$m-k = 1$, clusterization does not occur. We do not why this is so and
-further investigation into this phenomenon is out of scope for the
-current study.
+$m-k = 1$, clusterization does not occur. We do not why this is so.
 
@@
       generated randomly (see text and repository for details on
-      generating $M$). It is evident that when $m-k \geq 2$ clusterization
-      is ubiquitous, whereas for lower $m-k$ clusterization does not
-      occur.}
+      generating $M$). It is evident that when $m-k > 1$ clusterization
+      is prevalent, whereas for lower $m-k$ clusterization is not.}
   \label{fig:sim_AAt}
@@
       model correlation term to the error terms in the 1D heat
-      equation inverse problem. Lo and behold, measurements are not
-      close anymore and are pushed away thanks to the model error
-      term.}
+      equation inverse problem. As expected, measurements are pushed
+      away owing to the model error term.}
   \label{fig:corr_errors}

### 2024-09-02 latex/prelims.tex
add eic and AE
answered refs
Everything except for the intro

@@
 over some domain, we will keep in mind that these Hilbert spaces
-should really be thought of as function spaces. A deeper treatment of
-the foundations for inverse problems over function spaces can be found
-in \cite{Stuart10}.
+should really be thought of as function spaces, utilizing a setup
+similar to \cite{knapik2011}. A deeper treatment of the foundations
+for inverse problems over function spaces can be found in
+\cite{Stuart10}.
 
 \subsection{Forward Problems}\label{subsec:abstract_OED}
-In this section we give definitions and notations for forward problems,
-which are an essential part of the inverse problems we discuss
-later. Consider a "parameter space" \(\hilp\) and an "observation
-space" \(\hilo\) --- both separable Hilbert spaces (the subscripts p
-and o are for "parameter" and "observation", respectively). The
-parameter space includes the quantity we seek to infer; in the inverse
-problem of the heat equation, the parameter space $\hilp$ is where the
-initial condition lives. The observation space $\hilo$, on the other
-hand, is the space from which we take measurements; in the example of
-the 1D heat equation, $u(\cdot, T) \in \hilo$.
+In this section we give definitions and notations for forward
+problems, which are an essential part of the inverse problems we
+discuss later. Consider a "parameter space" \(\hilp\) and an
+"observation space" \(\hilo\) --- both separable Hilbert spaces (the
+subscripts p and o are for "parameter" and "observation",
+respectively). The parameter space includes the quantity we seek to
+infer; in the inverse problem of the heat equation, the parameter
+space $\hilp$ is where the initial condition lives. The observation
+space $\hilo$, on the other hand, is the space from which we take
+measurements; in the example of the heat equation, $u(\cdot, T) \in
+\hilo$.
 
@@
 that point evaluations could be approximated in a standard Hilbert
-space like $L^2(\Omega)$.}.
+space like $L^2(\Omega)$.}. %% Footnote
 
@@
 is time to formulate the process of inferring the parameter as a
-Bayesian inverse problem. We have already defined the likelihood in
-the previous section, and now we will define the prior.
+Bayesian inverse problem. We have already defined the Gaussian
+likelihood in the previous section, and now we will define the prior.
 

### 2024-09-02 latex/vanishing.tex
Everything except for the intro

--- 2024-08-22 latex/vanishing.tex
@@
 such a procedure is that the precision for the first $k$ eigenvectors
-is the average of the total precision $\sum_{j=1}^k \lambda_j^{-1} +
-\sigma^{-2}m$, see eq.~\eqref{eq:cylinders} and
+is the average of their total aggregated precision $\sum_{j=1}^k
+\lambda_j^{-1} + \sigma^{-2}m$, see eq.~\eqref{eq:cylinders} and
 Fig.~\ref{fig:optimal_vs_not} for an illustration.
@@
 \ref{subsub:clusterization1} below. We also suggest a generic
-explanation for clusterization, see Section \ref{subsub:clusterization2}.
+explanation for clusterization, see Section \ref{subsub:cara}.
 
-\subsubsection{Clusterization in the heat equation}
+\subsubsection{Clusterization in the heat equation}\label{subsub:clusterization1}
 Consider $\fwd$ and $\prcov$ from \emph{the inverse problem of the
@@
 Of course, in a real life problem --- such as the inverse problem of
-the 1D heat equation --- it is likely impossible to find measurement
-locations for which all eigenvectors $k+1$ and above are
-zero. However, if the eigenvalues of $\fwd\prcov\fwd^*$ decay quickly
-(recall the square-exponential decay for eigenvalues of the 1D heat
-equation in eq.\eqref{eq:decay}), a D-optimal design will try to
-balance measuring a small number (i.e.~$k$) of the leading
-eigenvectors.
+the 1D heat equation --- it is likely impossible to find measurements
+for which all eigenvectors $k+1$ and above are zero. However, if the
+eigenvalues of $\fwd\prcov\fwd^*$ decay quickly (recall the
+square-exponential decay for eigenvalues of the 1D heat equation in
+eq.\eqref{eq:decay}), a D-optimal design will try to balance measuring
+a small number (i.e.~$k$) of the leading eigenvectors.
 
@@
 Now we can understand measurement clusterization for the inverse
-problem of the heat equation. A D-optimal design attempts to measure
-the first $k$ eigenvectors of $\fwd \prcov \fwd^*$. But there may be
-(spatial) limitations on where these $k$ eigenvectors have large
-amplitude. For the inverse problem of the heat equation there are two
-spatial locations that present a good compromise between the
+problem of the 1D heat equation. A D-optimal design attempts to
+measure the first $k$ eigenvectors of $\fwd \prcov \fwd^*$. But there
+may be (spatial) limitations on where these $k$ eigenvectors have
+large amplitude. For the inverse problem of the heat equation there
+are two spatial locations that present a good compromise between the
 amplitudes of the first and second eigenvectors, namely $x_1$ and
@@
 at our disposal but only two spatial locations that are a good
-compromise between the first and second scaled eigenvectors. Thus,
-clusterization arises as a consequence of the pigeonhole principle.
+compromise between the amplitudes of the first and second scaled
+eigenvectors. Thus, clusterization arises as a consequence of the
+pigeonhole principle.
 
@@
       approximately at $0.31$ and $0.69$. These two locations are a
-      compromise between the magnitudes of the first and second
+      compromise between the amplitudes of the first and second
       eigenvectors, which are the eigenvectors that a D-optimal design
@@
 
-\subsubsection{Clusterization in our generic model}%\label{subsub:cara}
+\subsubsection{Clusterization in our generic model}\label{subsub:cara}
 We can gain further insight to clusterization in our generic model,
@@
 eigenvectors of $\fwd \prcov \fwd^*$. We present a short discussion
-adapting arguments presented in \cite[Chapter
-  3]{silveyOptimalDesign1980} and \cite[Section
-  5.2.3]{pronzatoPazman2013}.
+adapting arguments presented in \cite[Chapter 3]{silvey1980} and
+\cite[Section 5.2.3]{pronzatoPazman2013}.
 
@@
 places weight $>1$ on a single measurement vector.
+
+\subsubsection{Convergence}\label{subsub:convergence}
+In this section we fulfill our promise from Section
+\ref{subsub:implications} and prove that the posterior uncertainty
+ellipsoid in $\hilo$ will contract to zero along every eigenvector of
+$\fwd \prcov \fwd^*$. In our proof we ignore potential problems with
+conducting inference on function spaces which are not unique to
+D-optimal designs (see e.g.~\cite{owhadi2015} for more details).
+
+First, denote $\opt_m$ a D-optimal design utilizing $m$ measurements
+and denote $k_m := \rank\opt_m^*\opt_m$. An immediate consequence of
+Theorem \ref{thm:char} is that allowing more measurements will
+eventually allow us to measure each eigenvalue, i.e.:
+\begin{equation}\label{eq:lim}
+  \lim_{m\to\infty} k_m = \infty
+\end{equation}
+
+Now, recall that part (5) of Theorem \ref{thm:char} ensures that the
+eigenvalues of the posterior pushforward covariance equal
+\begin{equation*}
+  \theta^{(k)}_i = \left ( \frac{\sum_{j=1}^{k} \lambda_j^{-1} +
+    \sigma^{-2}m}{k} \right )^{-1}, \text{ for $i\leq k$}.
+\end{equation*}
+Using the inequality for the arithmetic and harmonic means, it is easy
+to verify that
+\begin{equation*}
+  \theta^{(k)}_i \leq \lambda_{k}.
+   %% \leq \frac{\sum_{j=1}^{k} \lambda_j}{k}
+\end{equation*}
+Since $\lim_{k\to \infty} \lambda_k= 0$, we conclude that for all $i$,
+$\lim_{k\to\infty} \theta^{(k)}_i = 0$. Combining the latter
+observation with eq.~\eqref{eq:lim}, we conclude that
+$\lim_{m\to\infty} \theta^{(k_m)}_i = 0$ for all $i$. Therefore,
+posterior uncertainty decays to zero for all eigenvectors.

### 2024-09-08 latex/intro.tex
Done?

@@
 \section{Introduction}\label{section:intro}
-The practice of experimental design entails specifying every detail of
-running an experiment \cite{chaloner1995}. In its most basic form,
-experimental design entails specifying measurements that consist an
-experiment. For example, consider the process of searching for oil:
-measurements require digging holes in the ground and a \emph{design}
-simply indicates which holes should be dug
-\cite{horesh2008borehole}. Since measurements are expensive, only a
-limited number of them can be taken, and thus carefully choosing
-measurement locations is crucial.
+Experimental design is a systematic approach to specifying every
+possible aspect of an experiment \cite{chaloner1995}. In this study,
+we focus on experiments that involve measuring quantities arising from
+naturally occurring processes. For us, therefore, experimental design
+entails determining which specific measurements to record from a given
+process. In a typical experiment the number of measurements that can
+be recorded is limited by time, cost and other constraints. Therefore
+selecting the right set of measurements to record is crucial as it
+directly impacts the utility of the data collected, and consequently,
+the validity of the conclusions drawn.
+
+For example, consider the process of searching for oil: measurements
+require digging deep holes in the ground, known as "boreholes". A
+\emph{design} simply indicates which boreholes should be dug
+\cite{horesh2008borehole}. Since digging boreholes is expensive, only
+a limited number of them can be dug, and thus carefully choosing their
+locations is crucial.
 
@@
 materials, such as density, elasticity, and the presence of fluids or
-voids. Combining travel time data with a geophysical models of
+voids. Combining travel time data with a geophysical model of the
 contents of the earth's crust facilitates reconstructing the structure
@@
 to prior \cite{chaloner1995, AlexanderianGloorGhattas14,
-  CoverThomas91}. Utilizing KL divergence for quantifying differences
-between distributions is standard practice and we will not motivate it
-further here. For a discrete parameter $\param$ and data $\data$, the
-KL divergence is defined as
+  CoverThomas91}.
+%%  Utilizing KL divergence for quantifying differences between
+%% distributions is standard practice and we will not motivate it
+%% further here.
+Recall that for a discrete parameter $\param$ and data $\data$, the KL
+divergence is defined as
 \begin{equation}\label{eq:basic_KL}
@@
 Of course, data $\data$ is not known before the experiment is
-conducted, so we average over it to define the D-optimality criterion
-as:
+conducted. Hence, we average over $\data$ to define the D-optimality
+criterion as:
 \begin{equation*}
@@
 For concreteness, let us now consider a toy inverse problem: inferring
-the initial condition for partial differential equation known as the
+the initial condition for a partial differential equation known as the
 \emph{heat equation} in one dimension (1D heat equation
 henceforth). Readers less familiar with partial differential equations
-can think of the following setup: before us there is a perfectly
-insulated metal rod. At each point on the rod the temperature is
-different and unknown, and we call this temperature distribution the
-\emph{initial condition}. We wait for a short time $T$ and let the
-heat dissipate a bit inside the rod. The heat equation determines the
-heat distribution inside the rod at time $t=T$.
+can think of the following setup: before us there is a metal rod,
+perfectly insulated except for its tips. At each point on the rod the
+temperature is different and unknown, and we call this temperature
+distribution the \emph{initial condition}. We wait for a short time
+$T$ and let the heat dissipate a bit inside the rod. The heat equation
+determines the heat distribution inside the rod at time $t=T$.
 
@@
 heat distribution $u_0\in L^2([0,1])$ that satisfies the homogeneous
-Dirichlet boundary condition eq.~\eqref{eq:heat2}. The initial
+Dirichlet boundary condition eq.~\eqref{eq:heat2}. This initial
 condition $u_0$ is a linear combination of sines $\ev_n(x) = \sin(\pi
@@
 In the \emph{inverse problem of the 1D heat equation}, our goal is to
-infer the initial condition $u_0 = u(\cdot, 0)$ from observations of
-the final state $u(\cdot, T)$. However, we are not able to measure
-$u(\cdot, T)$ at every point in the domain $\Omega$. Rather, we take
-some number of temperature measurements on the rod once our waiting
-time $T$ has passed and try to infer $u(\cdot, 0)$ from these
-measurements.
+infer the initial condition $u_0 = u(\cdot, 0)$ from noisy
+observations of the final state $u(\cdot, T)$. However, we are not
+able to measure $u(\cdot, T)$ at every point in the domain
+$\Omega$. Rather, we take some number of temperature measurements on
+the rod once our waiting time $T$ has passed and try to infer
+$u(\cdot, 0)$ from these measurements.
 
@@
 prior for the initial condition $u_0 \sim \mathcal{N}(0,
-(-\Delta)^{-1})$, where $\Delta$ again defined with homogeneous
-Dirichlet boundary condition. The intuition here is that
-$(-\Delta)^{-1}$ is smoothing and prior realizations are generated, in
-analogy to the finite dimensional case, via
-$(-\Delta)^{-1/2}\mathcal{W}$, where $\mathcal{W}$ is white noise
-\cite{rue2011}. Our choice of prior covariance operator ensures the
-posterior is well-defined and prior realizations are well-behaved, see
-Theorem 3.1 and Lemma 6.25 in \cite{Stuart10} for details. The second
-ingredient of a fully Bayesian formulation of the inverse problem is
-likelihood. We simply assume centered iid Gaussian measurement error
-model, giving rise to a the widely applicable Gaussian likelihood
-function.
+(-\Delta)^{-1})$, where $\Delta$ is again defined with a homogeneous
+Dirichlet boundary condition. The intuition here is that since
+$\Delta$ is a differential operator, it is "roughing" (the opposite of
+smoothing) and so $(-\Delta)^{-1}$ is smoothing. Prior realizations
+are generated, in analogy to the finite dimensional case, by smoothing
+white noise: $(-\Delta)^{-1/2}\mathcal{W}$, where $\mathcal{W}$ is
+white noise \cite{rue2011}. Our choice of prior covariance operator
+ensures the posterior is well-defined and prior realizations are
+well-behaved, see Theorem 3.1 and Lemma 6.25 in \cite{Stuart10} for
+details. The second ingredient of a fully Bayesian formulation of the
+inverse problem is likelihood. We simply assume centered iid Gaussian
+measurement error model, giving rise to a the widely applicable
+Gaussian likelihood function.
+
+Now that we have a fully Bayesian formulation of our toy inverse
+problem, we would like to find a D-optimal design. Defining D-optimal
+designs over function spaces is somewhat mathematically
+involved. Thus, to keep our discussion here focused, we will explore
+these technical details later in Section
+\ref{subsec:D_optimal_design}. For now, it suffices to know that
+D-optimal designs can be defined for our problem.
 
@@
 and even necessary aspect of optimal experimental design
-\cite{fisher1949design, morris2011, schafer2001replication}. Similarly
-to a design implementing replication, a clustered design reduces the
-signal-to-noise ratio of the repeated measurements
-\cite{telford2007brief}. The difference is that a clustered design
-takes repeated measurements at the expense of other quantities not
-measured at all.
-
-%% Replication 2
+\cite{fisher1949design, morris2011, schafer2001replication}.
 For example, \cite{fisher1949design}, suggested repeating his famous
@@
 
-%% Replication 3
-To further illustrate the difference between replication and
-clusterization, consider an experiment measuring the effect of
-rainfall on grass growth \cite{fay2000rainfall}. The experiment
-involved four rainfall manipulation "treatments" (i.e.~simulating
-different timing and quantity of rainfall), each replicated three
-times over a different plot of land. Indeed, it seems reasonable for
-researchers to replicate the phenomenon they are trying to study. A
-clustered design in such an experiment would imply the researchers
-should take a repeated measurement \emph{on the same plot}, at the
-expense of measuring grass growth in other plots. To conclude our
-short discussion on replication vs.~clusterization: these are
-fundamentally different concepts and even though replication is quite
-intuitive, it is inapplicable to the inverse problems we consider in
-this manuscript.
+Similarly to a design implementing replication, a clustered design
+reduces the signal-to-noise ratio of the repeated measurements
+\cite{telford2007brief}. The difference is that a clustered design
+takes repeated measurements at the expense of other quantities not
+measured at all. For example, consider an experiment measuring the
+effect of rainfall on grass growth \cite{fay2000rainfall}. The
+experiment involved four rainfall manipulation "treatments"
+(i.e.~simulating different timing and quantity of rainfall), each
+replicated three times over different plots of land. Indeed, it seems
+reasonable for researchers to replicate the phenomenon they are trying
+to study. A clustered design in such an experiment would imply the
+researchers should take a repeated measurement \emph{on the same
+plot}, at the expense of measuring grass growth in other plots!
+
+To conclude our short discussion on replication vs.~clusterization:
+these are fundamentally different concepts and even though replication
+is quite intuitive, it is inapplicable to the inverse problems we
+consider in this manuscript.
 
@@
   on how quickly the eigenvalues of the prior covariance in
-  observation space decay. This decay depends mostly on how ill-posed
-  the problem at hand is and does not depend much on the prior. See
-  Section \ref{section:vanishing} --- particularly, Theorem
-  \ref{thm:char} and the discussion following it. We also show results
-  of numerical experiments, where simulations of our model give rise
-  to D-optimal designs that exhibit clusterization with high
-  probability in Section \ref{subsec:lemma_sims}. Thus, given the
-  genericity of our model, we expect measurement clusterization to be
-  a generic and ubiquitous phenomenon.
+  observation space decay.
+  %% This decay depends mostly on how ill-posed the problem at hand is
+  %% and does not depend much on the prior
+  See Section \ref{section:vanishing} --- particularly, Theorem
+  \ref{thm:char} and the discussion following it. Furhtermore, in
+  Section \ref{subsec:lemma_sims} we show results of numerical
+  experiments, where simulations of our model give rise to D-optimal
+  designs that exhibit clusterization with high probability. Thus,
+  given the genericity of our model, we expect measurement
+  clusterization to be a generic and ubiquitous phenomenon.
 
@@
   of measurement effort to the abovementioned select set of prior
-  eigenvectors. This concentration allows us to move from an
-  infinite-dimensional setting to finite dimensions, where the
-  conditions of Carath\'eodory's Theorem hold. We conclude that a
-  D-optimal design arises by weighting a small number of
-  measurements. If the number of allowed measurements is larger than
-  the small number dictated by Carath\'eodory's Theorem, we have
-  excessive weight on some measurements, which can be interpreted as
-  clusterization.
+  eigenvectors. This allows us to move from an infinite-dimensional
+  setting to finite dimensions, where the conditions of
+  Carath\'eodory's Theorem hold. We conclude that a D-optimal design
+  arises by weighting a small number of measurements. If the number of
+  allowed measurements is larger than the small number dictated by
+  Carath\'eodory's Theorem, we have excessive weight on some
+  measurements, which can be interpreted as clusterization.
 
@@
 %% Convergence
-Given that D-optimal designs suffer from clusterization, it is natural
-to ask whether D-optimal designs should be pursued at all. Other
-authors have established convergence results for decaying measurement
-error \cite{knapik2011}, space-filling measurements
-\cite{teckentrup2020} and randomly sampled designs
-\cite{nickl2023}. Unfortunately, D-optimal designs --- clustered
-designs in particular --- do not fall into any of these categories.
+Clusterization in D-optimal designs raises the question of whether
+D-optimal designs should be pursued at all. While other authors have
+established convergence results for decaying measurement error
+\cite{knapik2011}, space-filling measurements \cite{teckentrup2020}
+and randomly sampled designs \cite{nickl2023}, their results do not
+hold for D-optimal designs. %% Unfortunately, D-optimal designs ---
+%% clustered designs in particular --- do not fall into any of these
+%% categories.
 
@@
 clustered designs). Convergence holds in the following sense: the
-posterior uncertainty ellipsoid in $\hilo$ will contract to zero along
-every eigenfunction of $\fwd \prcov \fwd^*$. See Section
-\ref{subsub:convergence} for a proof based on Theorem \ref{thm:char}
---- the main theorem of this manuscript.
+posterior uncertainty ellipsoid will contract to zero along each
+one of its eigenvectors. See Section \ref{subsub:convergence} for a
+precise statement and a proof based on Theorem \ref{thm:char} --- the
+main theorem of this manuscript.
 
@@
 In Theorem \ref{thm:char}, we also show that D-optimal designs are
-best understood in the space of \emph{observations}, see Section
-\ref{section:vanishing} for a precise statement. This is in accordance
-with previous work by \cite{koval2020}, who showed that A-optimal
-designs are best constructed in the space of observations.
+best understood in the space of \emph{observations}
+%% , see Section \ref{section:vanishing} for a precise statement.
+This is in accordance with previous work by \cite{koval2020}, who
+showed that A-optimal designs are best constructed in the space of
+observations.
 

### 2024-09-08 latex/intro.tex
Done

@@
   See Section \ref{section:vanishing} --- particularly, Theorem
-  \ref{thm:char} and the discussion following it. Furhtermore, in
+  \ref{thm:char} and the discussion following it. Furthermore, in
   Section \ref{subsec:lemma_sims} we show results of numerical
@@
 that the linearity of the forward problem is also an important
-ingredient in giving rise to clustered desgins. Therefore, our advice
+ingredient in giving rise to clustered designs. Therefore, our advice
 to practitioners is to avoid linearization, and find D-optimal designs
@@
 inherently bad on their own. Rather, we suggest that if a clustered
-design arises, a practitioner should revisit their modeling
-assumptions --- specifically, their choice of prior and model
-linearization. If the practitioner is confident in their assumptions,
-then clustered designs should be avoided only to the extent
-necessitated by the physical measuring apparatus.
+design arises, a practitioner should revisit their modeling choices
+--- specifically, their choice of prior and model linearization. If
+the practitioner is confident in their analysis, then clustered
+designs should be avoided only to the extent necessitated by the
+physical measuring apparatus.
 
@@
 and randomly sampled designs \cite{nickl2023}, their results do not
-hold for D-optimal designs. %% Unfortunately, D-optimal designs ---
-%% clustered designs in particular --- do not fall into any of these
-%% categories.
-
-Luckily, we can expect convergence from D-optimal designs (including
-clustered designs). Convergence holds in the following sense: the
-posterior uncertainty ellipsoid will contract to zero along each
-one of its eigenvectors. See Section \ref{subsub:convergence} for a
-precise statement and a proof based on Theorem \ref{thm:char} --- the
-main theorem of this manuscript.
+hold for D-optimal designs. Luckily, we can expect convergence from
+D-optimal designs (including clustered designs) in the following
+sense: the posterior uncertainty ellipsoid will contract to zero along
+each one of its eigenvectors. See Section \ref{subsub:convergence} for
+a precise statement and a proof based on Theorem \ref{thm:char} ---
+the main theorem of this manuscript.
 

### 2024-09-08 revision/refs.tex
Done

@@
 
-I appreciate the many thoughtful insightful comments by all of
-you. These comments have helped me considerably improve the exposition
-of my paper and added new insights as well. I sincerely appreciate the
-option to revise and resubmit this manuscript. Please see below my
-answers to the comments made by the editor in chief (Section
+I appreciate the many insightful comments by the refrees. These
+comments have helped me considerably improve the exposition of the
+manuscript and added new insights as well. I sincerely appreciate the
+possibility to revise and resubmit this manuscript. Please see below
+my answers to the comments made by the editor in chief (Section
 \ref{eic}), associate editor (Section \ref{ae}) and three referees
@@
 
-\AR I have considerably extended the introduction. See revised
+\AR I have considerably extended the Introduction. See revised
 manuscript, diff file and detailed answers to referees' comments
@@
 
-\AR I added a section on the implications. The main takeaways are: (1)
-avoid linearization and Gaussian priors, and (2) clusterization can
-inform you of the number of relevant d.o.f in the problem.
+\AR I added a section discussing implications. The main takeaways are:
+(1) avoid linearization and Gaussian priors, and (2) clusterization
+can inform you of the number of relevant d.o.f in the problem.
 
@@
 
-\AR I have added considerably to the introduction. I hope you will
+\AR I have added considerably to the Introduction. I hope you will
 find the transfer to the maths smoother now.
@@
 \AR Again, I appreciate the time and effort invested by the referees
-and editorial team. Thanks for the opportunity to revise and improve this manuscript.
+and editorial team. Thank you for the opportunity to revise and
+improve this manuscript.
 

### 2024-09-08 revision/rev1.tex
Typos
Done

@@
-\section{Reviewer 1}\label{ref1}
+\section{Referee 1}\label{ref1}
 \RC This paper explains a common practical issue: the cause of
@@
 inherently bad on their own. Rather, we suggest that if a clustered
-design arises, a practitioner should revisit their modeling
-assumptions --- specifically, their choice of prior and model
-linearization. If the practitioner is confident in their assumptions,
-then clustered designs should be avoided only to the extent
-necessitated by the physical measuring apparatus.
+design arises, a practitioner should revisit their modeling choices
+--- specifically, their choice of prior and model linearization. If
+the practitioner is confident in their analysis, then clustered
+designs should be avoided only to the extent necessitated by the
+physical measuring apparatus.
 \end{quote}

### 2024-09-08 revision/rev2.tex
Typos
Done

@@
-\section{Reviewer 2}\label{ref2}
+\section{Referee 2}\label{ref2}
 \subsection{Measurement Clusterization vs Repetition}
@@
 and even necessary aspect of optimal experimental design
-  \cite{fisher1949design, morris2011,
-    schafer2001replication}. Similarly to a design implementing
-  replication, a clustered design reduces the signal-to-noise ratio of
-  the repeated measurements \cite{telford2007brief}. The difference is
-  that a clustered design takes repeated measurements at the expense
-  of other quantities not measured at all.
-  \newline
+\cite{fisher1949design, morris2011, schafer2001replication}.
 For example, \cite{fisher1949design}, suggested repeating his famous
@@
 \newline
-To further illustrate the difference between replication and
-clusterization, consider an experiment measuring the effect of
-rainfall on grass growth \cite{fay2000rainfall}. The experiment
-involved four rainfall manipulation "treatments" (i.e.~simulating
-different timing and quantity of rainfall), each replicated three
-times over a different plot of land. Indeed, it seems reasonable for
-researchers to replicate the phenomenon they are trying to study. A
-clustered design in such an experiment would imply the researchers
-should take a repeated measurement \emph{on the same plot}, at the
-expense of measuring grass growth in other plots. To conclude our
-short discussion on replication vs.~clusterization: these are
-fundamentally different concepts and even though replication is quite
-intuitive, it is inapplicable to the inverse problems we consider in
-this manuscript.
+Similarly to a design implementing replication, a clustered design
+reduces the signal-to-noise ratio of the repeated measurements
+\cite{telford2007brief}. The difference is that a clustered design
+takes repeated measurements at the expense of other quantities not
+measured at all. For example, consider an experiment measuring the
+effect of rainfall on grass growth \cite{fay2000rainfall}. The
+experiment involved four rainfall manipulation "treatments"
+(i.e.~simulating different timing and quantity of rainfall), each
+replicated three times over different plots of land. Indeed, it seems
+reasonable for researchers to replicate the phenomenon they are trying
+to study. A clustered design in such an experiment would imply the
+researchers should take a repeated measurement \emph{on the same
+plot}, at the expense of measuring grass growth in other plots!
+\newline
+To conclude our short discussion on replication vs.~clusterization:
+these are fundamentally different concepts and even though replication
+is quite intuitive, it is inapplicable to the inverse problems we
+consider in this manuscript.
 \end{quote}
@@
 
-\AR I agree, but what you are describing is replication. See my
-comment on replication vs.~clusterization above.
+\AR I agree, but I believe what you are describing is replication. See
+my comment on replication vs.~clusterization above.
 
@@
 the implications of this study. Since it is too long to reproduce
-here, see Section \ref{subsub:implications} in the revised manuscript.
+here, see Section \ref{subsub:implications} in the revised
+manuscript. I will just mention that one important takeaway is that
+practitioners should avoid linearization and Gaussian priors when
+seeking D-optimal designs.
 
@@
 \AR Thanks for this comment. I considerably expanded the
-  introduction and delegated many proofs to the Supplementary.
+  Introduction and delegated many proofs to the Supplementary.
 
@@
 the result seems non-intuitive. I expanded this discussion in the
-revised manuscript. See below and the discussion preceding it in
-Section \ref{subsub:implications} of the revised manuscript.
+revised manuscript. See paragraph below and the discussion preceding
+it in Section \ref{subsub:implications} of the revised manuscript.
 

### 2024-09-08 revision/rev3.tex
Typos
Done

@@
-\section{Reviewer 3}\label{ref3}
+\section{Referee 3}\label{ref3}
 \RC The paper studies Bayesian D-optimal design (that is, the task of
@@
 \AR I made sure the repository is functioning and well-documented. See
-\url{https://github.com/yairdaon/OED}.
+\url{https://github.com/yairdaon/OED}. Furthermore, I added references
+to specific modules in the text, see excerpts below.
+
+\begin{quote}
+Full results are located in the \texttt{simulations.csv} file within
+the accompanying \href{https://github.com/yairdaon/OED}{repository}.
+Code implementing the experiments described above is located in module
+\texttt{zeros.py} of said repository.
+\end{quote}
+\begin{quote}
+Code generating these movies is located in module \texttt{movies.py}
+in the accompanying
+\href{https://github.com/yairdaon/OED}{repository}.
+\end{quote}
+\begin{quote}
+The implementation of the D-optimality criterion for this inverse
+problem with homogeneous Neumann and Dirichlet boundary conditions can
+be found in directory \texttt{src/} and script
+\texttt{clusterization.py} in the accompanying
+\href{https://github.com/yairdaon/OED}{repository}.
+\end{quote}
 
@@
 construction with the goal of testing numerically how prevalent are
-  clustered designs. To this end, we would like to generate random
-  prior eigenvalues $\lambda_j$, fix $m$ and $k$, find a D-optimal
-  $\opt^*\opt$ and then utilize the construction of
+clustered designs. To this end, we would like to generate random prior
+eigenvalues $\lambda_j$, fix $m$ and $k$, find what a D-optimal
+$\opt^*\opt$ should be, and then utilize the construction of
 Theorem~\ref{thm:char} and Lemma~\ref{lemma:free} to find $\opt$.
-  \newline
-  To simplify things, we directly sample rank $\opt^*\opt$. We iterate
-  over the number of measurements $m \in \{4,\dots, 24\}$, and for
-  every $m$ we then iterate over $k:=\rank \obs^*\obs \in \{2,\dots,
-  m-1\}$. For each pair $m,k$ we repeat the following steps $N=5000$
-  times:
+
+To simplify things, we directly generate $\opt^*\opt$. We iterate over
+the number of measurements $m \in \{4,\dots, 24\}$, and for every $m$
+we then iterate over $k:=\rank \opt^*\opt \in \{2,\dots, m-1\}$. For
+each pair $m,k$ we repeat the following steps $N=5000$ times:
 \begin{enumerate}
 \item Generate random diagonal $D\in \mathbb{R}^{k\times k}$ with
-    entries $\log (d_i) \sim \mathcal{N}(50,15)$ and normalize so that
-    $\ttr D = m$.
+  entries $\log (d_i) \sim \mathcal{N}(50,15)$.
+\item Normalize $D$ that $\ttr D = m$.
 \item Conjugate $D$ by a random orthogonal matrix to form a positive
@@
 \item Apply the construction of Lemma \ref{lemma:free} to calculate
-    $A$ such that $AA^t = M$, where $A$ has unit norm
-    columns. I.e.~find the optimal design $\opt$.
-  \item Since $A$ corresponds to $\opt$, its columns correspond to
-    measurement vectors. We call $A$ "clustered" if $A$ has two or
-    more identical columns (up to some numerical precision threshold).
+  $A$ such that $AA^t = M$, where $A$ has unit norm columns. $A^t$ is
+  our optimal design $\opt$.
+\item Since $A^t$ corresponds to $\opt$, its columns correspond to
+  measurement vectors. We call $A$ "clustered" if $A$ has two or more
+  identical columns (up to some numerical precision threshold,
+  i.e.~$10^{-5}$).
 \end{enumerate}
-  We then calculate the fraction of clustered designs of the
-  simulations we ran, for each pair $m,k$. Clusterization occurred at
-  high rates ($>99.9\%$) whenever $m-k > 1$; see
-  Fig.~\ref{fig:sim_AAt}. Hence, in these simulations, clusterization
-  is a generic property. However, when $m-k = 1$, clusterization does
-  not occur. We do not why this is so and further investigation into
-  this phenomenon is out of scope for the current study.
+We then calculate the fraction of clustered designs of the simulations
+we ran, for each pair $m,k$. Clusterization occurred at high rates
+($>99.9\%$) whenever $m-k > 1$; see Fig.~\ref{fig:sim_AAt}. Hence, in
+these simulations, clusterization is a generic property. However, when
+$m-k = 1$, clusterization does not occur. We do not why this is so.
+\newline
 \newline
@@
 the accompanying \href{https://github.com/yairdaon/OED}{repository}.
-  Code implementing the experiments described above is located in
-  module \texttt{zeros.py} of said repository. Runtime should be less
-  than 30 minutes on any reasonably modern laptop (it took 12 minutes
-  on the author's laptop).
+Code implementing the experiments described above is located in module
+\texttt{zeros.py} of said repository. Runtime should be less than 30
+minutes on any reasonably modern laptop (it took 12 minutes on the
+author's laptop).
 \end{quote}
@@
 second part follows the statement and proof of the abovementioned
-theorem. I believe the level of discussion is adequate and view a
-deeper analysis of convergence rates as out of scope for the current
-study.
+theorem.
 \begin{quote}%%Convergence
-  Given that D-optimal designs suffer from clusterization, it is
-  natural to ask whether D-optimal designs should be pursued at
-  all. Other authors have established convergence results for decaying
-  measurement error \cite{knapik2011}, space-filling measurements
-  \cite{teckentrup2020} and randomly sampled designs
-  \cite{nickl2023}. Unfortunately, D-optimal designs --- clustered
-  designs in particular --- do not fall into any of these categories.
+  Clusterization in D-optimal designs raises the question of whether
+  D-optimal designs should be pursued at all. While other authors have
+  established convergence results for decaying measurement error
+  \cite{knapik2011}, space-filling measurements \cite{teckentrup2020}
+  and randomly sampled designs \cite{nickl2023}, their results do not
+  hold for D-optimal designs. Luckily, we can expect convergence from
+  D-optimal designs (including clustered designs) in the following
+  sense: the posterior uncertainty ellipsoid will contract to zero
+  along each one of its eigenvectors. See Section
+  \ref{subsub:convergence} for a precise statement and a proof based
+  on Theorem \ref{thm:char} --- the main theorem of this manuscript.
+\end{quote}
+\begin{quote}
+  In this section we fulfill our promise from Section
+  \ref{subsub:implications} and prove that the posterior uncertainty
+  ellipsoid in $\hilo$ will contract to zero along every eigenvector
+  of $\fwd \prcov \fwd^*$. In our proof we ignore potential problems
+  with conducting inference on function spaces which are not unique to
+  D-optimal designs (see e.g.~\cite{owhadi2015} for more details).
   \newline
-  Luckily, we can expect convergence from D-optimal designs (including
-  clustered designs). Convergence holds in the following sense: the
-  posterior uncertainty ellipsoid in $\hilo$ will contract to zero
-  along every eigenfunction of $\fwd \prcov \fwd^*$. See Section
-  \ref{subsub:convergence} for a proof based on Theorem \ref{thm:char}
-  --- the main theorem of this manuscript.
   \newline  
-  Below we fulfill our promise from Section \ref{subsub:implications}
-  and prove that the posterior uncertainty ellipsoid in $\hilo$ will
-  contract to zero along every eigenfunction of $\fwd \prcov
-  \fwd^*$. In our proof we ignore potential problems with conducting
-  inference on function spaces which are not unique to D-optimal
-  designs (see e.g.~\ref{owhadi2015} for more details).
+  First, denote $\opt_m$ a D-optimal design utilizing $m$ measurements
+  and denote $k_m := \rank\opt_m^*\opt_m$. An immediate consequence of
+  Theorem \ref{thm:char} is that allowing more measurements will
+  eventually allow us to measure each eigenvalue, i.e.:
+  \begin{equation}\label{eq:lim}
+    \lim_{m\to\infty} k_m = \infty
+  \end{equation}
   \newline
-  First, let $\opt_m$ a D-optimal design utilizing $m$
-  measurements. Then, an immediate consequence of Theorem
-  \ref{thm:char} is that $\rank \opt_m^*\opt_m = k \to \infty$ as the
-  number of measurements $m \to \infty$. Moreover, part (5) of Theorem
-  \ref{thm:char} ensures that the eigenvalues of the posterior
-  covariance equal
+  \newline
+  Now, recall that part (5) of Theorem \ref{thm:char} ensures that the
+eigenvalues of the posterior pushforward covariance equal
 \begin{equation*}
-    \theta^{(k)}_i = \left ( \frac{\sum_{j=1}^k \lambda_j^{-1} +
-      \sigma^{-2}m}{k} \right )^{-1}, \text{ for $i\leq k$},
+  \theta^{(k)}_i = \left ( \frac{\sum_{j=1}^{k} \lambda_j^{-1} +
+    \sigma^{-2}m}{k} \right )^{-1}, \text{ for $i\leq k$}.
 \end{equation*}
-  where we explicitly express the dependence of $k$ in the
-  superscript. Now note that $\theta^{(k)}_i \leq \lambda_i$ for
-  $i\leq k$. Specifically, for fixed $k$, we get $\theta^{(k)}_i \leq
-  \lambda_k$ for all $i\leq k$. Since $\lim_{k\to \infty} \lambda_k=
-  0$, we conclude that for all $i$, $\lim_{k\to\infty} \theta^{(k)}_i
-  = 0$. Since $m\to \infty$ implies $k\to \infty$ we have
-  $\lim_{m\to\infty} \theta^{(k)}_i = 0$ for all $i$. Therefore,
+Using the inequality for the arithmetic and harmonic means, it is easy
+to verify that
+\begin{equation*}
+  \theta^{(k)}_i \leq \lambda_{k}.
+   %% \leq \frac{\sum_{j=1}^{k} \lambda_j}{k}
+\end{equation*}
+Since $\lim_{k\to \infty} \lambda_k= 0$, we conclude that for all $i$,
+$\lim_{k\to\infty} \theta^{(k)}_i = 0$. Combining the latter
+observation with eq.~\eqref{eq:lim}, we conclude that
+$\lim_{m\to\infty} \theta^{(k_m)}_i = 0$ for all $i$. Therefore,
 posterior uncertainty decays to zero for all eigenvectors.
@@
 technical difficulties in installing packages for inverse problems
-\cite{attia2023pyoed, VillaPetraGhattas16} I had to implement the
-inverse problem of the 1D heat equation from scratch, so I decided to
-leave out experiments in higher dimensions.
+\cite{attia2023pyoed, villa2021} I had to implement the inverse
+problem of the 1D heat equation from scratch, so I decided to leave
+out experiments in higher dimensions.
 
@@
 
-\AR I removed the statement of Theorem 1 from the introduction. It is
+\AR I removed the statement of Theorem 1 from the Introduction. It is
 Theorem \ref{thm:char} now.
@@
 and even necessary aspect of optimal experimental design
-\cite{fisher1949design, morris2011, schafer2001replication}. Similarly
-to a design implementing replication, a clustered design reduces the
-signal-to-noise ratio of the repeated measurements
+\cite{fisher1949design, morris2011, schafer2001replication}.
+For example, \cite{fisher1949design}, suggested repeating his famous
+milk and tea experiment in order "to be able to demonstrate the
+predominance of correct classifications in spite of occasional
+errors". Unfortunately, in the experiments we consider, replication is
+impossible. For example, in the MRI problem, we cannot generate an
+individual nominally identical to the one we wish to scan.
+\newline
+\newline
+Similarly to a design implementing replication, a clustered design
+reduces the signal-to-noise ratio of the repeated measurements
 \cite{telford2007brief}. The difference is that a clustered design
 takes repeated measurements at the expense of other quantities not
-measured at all.
+measured at all. For example, consider an experiment measuring the
+effect of rainfall on grass growth \cite{fay2000rainfall}. The
+experiment involved four rainfall manipulation "treatments"
+(i.e.~simulating different timing and quantity of rainfall), each
+replicated three times over different plots of land. Indeed, it seems
+reasonable for researchers to replicate the phenomenon they are trying
+to study. A clustered design in such an experiment would imply the
+researchers should take a repeated measurement \emph{on the same
+plot}, at the expense of measuring grass growth in other plots!
+\newline
+\newline
+To conclude our short discussion on replication vs.~clusterization:
+these are fundamentally different concepts and even though replication
+is quite intuitive, it is inapplicable to the inverse problems we
+consider in this manuscript.
 \end{quote}
@@
 present paper since I was not able to install common packages for
-inverse problems \cite{attia2023pyoed, VillaPetraGhattas16} on my
-machine.
+inverse problems \cite{attia2023pyoed, villa2021} on my machine.
 
@@
 example, we can replace point evaluations by $R(x;t) =
-t\mathbb{1}_{-1/2t, 1/2t}(x)$, where $\mathbb{1}$ is an indicator
-function. We will then need to change the norm constraints on
-measurement vectors from unit to $\|R(\cdot;t\|^2 = \int R^2(x;t)dx =
-t$ norm but that does not change the analysis as long as $t$ is
-fixed. I barely mentioned this issue in the paper since I think it is
-confusing and adds little context. See a footnote I added:
+t\mathbf{1}_{[-1/2t, 1/2t]}(x)$, where $\mathbf{1}_{[a,b]}$ is an
+indicator function on the interval $[a,b]$. We will then need to
+change the norm constraints on measurement vectors from unit to
+$\|R(\cdot;t)\|^2 = \int R^2(x;t)dx = t$ norm but that does not change
+the analysis as long as $t$ is fixed. I barely mentioned this issue in
+the paper since I think it is confusing and adds little context. See a
+footnote I added:
 
@@
 
-\AR I explain this in more detail in the revised manuscript:
-
+\AR I added more details to the explaination explain in the revised
+manuscript, see Section \ref{subsub:clusterization1}. See text
+reproduced below.
 \begin{quote}
-  Building on Theorem \ref{thm:char}, we can now give a compelling
-  explanation to the measurement clusterization we observed for the
-  inverse problem of the heat equation, see Section
-  \ref{subsub:clusterization1} below. We also suggest a generic
-  explanation for clusterization, see Section
-  \ref{subsub:clusterization2}.
-  \newline
 Consider $\fwd$ and $\prcov$ from \emph{the inverse problem of the
 heat equation}. As before, we denote the eigenvalues of
-  $\fwd\prcov\fwd^*$ by $\lambda_j$. We input these eigenvalues into
-  our \emph{generic} model, and find a D-optimal design $\opt$ for our
-  generic model using Theorem \ref{thm:char}. In our generic model,
-  the measurements we take are best utilized in reducing uncertainty
-  for the first $k$ eigenvectors. So, a D-optimal design arising from
-  our \emph{generic model} completely avoids measuring eigenvectors
-  $k+1$ and above.
-  \newline
+$\fwd\prcov\fwd^*$ by $\lambda_j$. We input these eigenvalues into our
+\emph{generic} model, and find a D-optimal design $\opt$ for our
+generic model using Theorem \ref{thm:char}. In our generic model, the
+measurements we take are best utilized in reducing uncertainty for the
+first $k$ eigenvectors. So, a D-optimal design arising from our
+\emph{generic model} completely avoids measuring eigenvectors $k+1$
+and above.
+\newline \newline
 Of course, in a real life problem --- such as the inverse problem of
-  the 1D heat equation --- it is likely impossible to find measurement
-  locations for which all eigenvectors $k+1$ and above are
-  zero. However, if the eigenvalues of $\fwd\prcov\fwd^*$ decay
-  quickly (recall the square-exponential decay for eigenvalues of the
-  1D heat equation in eq.\eqref{eq:decay}), a D-optimal design will
-  try to balance measuring a small number (i.e.~$k$) of the leading
-  eigenvectors.
-  \newline
+the 1D heat equation --- it is likely impossible to find measurements
+for which all eigenvectors $k+1$ and above are zero. However, if the
+eigenvalues of $\fwd\prcov\fwd^*$ decay quickly (recall the
+square-exponential decay for eigenvalues of the 1D heat equation in
+eq.\eqref{eq:decay}), a D-optimal design will try to balance measuring
+a small number (i.e.~$k$) of the leading eigenvectors.
+%%\end{quote}
+%%\begin{quote}
+\newline \newline
 The abovementioned balance is explored in
-  Fig.~\ref{fig:eigenvectors}. We allow $m=4$ measurements in $\Omega
-  = [0,1]$ and observe that D-optimal measurement locations are
-  clustered at $x_1 = 0.31$ and $x_2 = 0.69$. Upon close inspection of
-  the scaled eigenvectors of $\fwd \prcov \fwd^*$, we first observe
-  that eigenvectors $3$ and above have negligible prior
-  amplitude. Since we only have $m=4$ measurements at our disposal, we
-  interpret these results, following Theorem \ref{thm:char}, as
-  implying we should only care about measuring the first and second
-  eigenvectors. Then, we note the D-optimal $x_1,x_2$ present a
-  compromise between the amplitude of the first and second
-  eigenvectors. For example, a measurement at $x=0.5$ would have
-  ignored the second eigenvector altogether, since the second
-  eigenvector is zero at $x=0.5$.
-  \newline
+Fig.~\ref{fig:eigenvectors}. We allow $m=4$ measurements in $\Omega =
+[0,1]$ and observe that D-optimal measurement locations are clustered
+at $x_1 = 0.31$ and $x_2 = 0.69$. Upon close inspection of the scaled
+eigenvectors of $\fwd \prcov \fwd^*$, we first observe that
+eigenvectors $3$ and above have negligible prior amplitude. Since we
+only have $m=4$ measurements at our disposal, we interpret these
+results, following Theorem \ref{thm:char}, as implying we should only
+care about measuring the first and second eigenvectors. Then, we note
+the D-optimal $x_1,x_2$ present a compromise between the amplitude of
+the first and second eigenvectors. For example, a measurement at
+$x=0.5$ would have ignored the second eigenvector altogether, since
+the second eigenvector is zero at $x=0.5$.
+\newline \newline
 Now we can understand measurement clusterization for the inverse
-  problem of the heat equation. A D-optimal design attempts to measure
-  the first $k$ eigenvectors of $\fwd \prcov \fwd^*$. But there may be
-  (spatial) limitations on where these $k$ eigenvectors have large
-  amplitude. For the inverse problem of the heat equation there are
-  two spatial locations that present a good compromise between the
+problem of the 1D heat equation. A D-optimal design attempts to
+measure the first $k$ eigenvectors of $\fwd \prcov \fwd^*$. But there
+may be (spatial) limitations on where these $k$ eigenvectors have
+large amplitude. For the inverse problem of the heat equation there
+are two spatial locations that present a good compromise between the
 amplitudes of the first and second eigenvectors, namely $x_1$ and
-  $x_2$ --- see Fig.~\ref{fig:eigenvectors}. We have $m=4$
-  measurements at our disposal but only two spatial locations that are
-  a good compromise between the first and second scaled
+$x_2$ --- see Fig.~\ref{fig:eigenvectors}. We have $m=4$ measurements
+at our disposal but only two spatial locations that are a good
+compromise between the amplitudes of the first and second scaled
 eigenvectors. Thus, clusterization arises as a consequence of the
@@
 \end{quote}
-
 \begin{figure}\label{fig:eigenvectors}
@@
     \caption{D-optimal measurement locations ($m=4$ measurements) and
-    weighted eigenvectors for finding the initial condition of the 1D
-    heat equation. Measurement locations and weighted eigenvectors are
-    plotted over the computational domain $\Omega = [0, 1]$
-    (x-axis). Measurement clusterization occurs approximately at
-    $0.31$ and $0.69$. These two locations are a compromise between
-    the magnitudes of the first and second eigenvectors, which are the
-    eigenvectors that a D-optimal design aims to measure. Allocating
-    $m=4$ measurements into two locations results in clusterization,
-    according to the pigeonhole principle.}
+      weighted eigenvectors for finding the initial condition of the
+      1D heat equation. Measurement locations and weighted
+      eigenvectors are plotted over the computational domain $\Omega =
+      [0, 1]$ (x-axis). Measurement clusterization occurs
+      approximately at $0.31$ and $0.69$. These two locations are a
+      compromise between the amplitudes of the first and second
+      eigenvectors, which are the eigenvectors that a D-optimal design
+      aims to measure. Allocating $m=4$ measurements into two
+      locations results in clusterization, according to the pigeonhole
+      principle.}
   \label{fig:why}
 \end{figure}
+
+%% \begin{quote}
+%%   Building on Theorem \ref{thm:char}, we can now give a compelling
+%%   explanation to the measurement clusterization we observed for the
+%%   inverse problem of the heat equation, see Section
+%%   \ref{subsub:clusterization1} below. We also suggest a generic
+%%   explanation for clusterization, see Section
+%%   \ref{subsub:clusterization2}.
+%%   \newline
+%%   \newline
+%%   Consider $\fwd$ and $\prcov$ from \emph{the inverse problem of the
+%%   heat equation}. As before, we denote the eigenvalues of
+%%   $\fwd\prcov\fwd^*$ by $\lambda_j$. We input these eigenvalues into
+%%   our \emph{generic} model, and find a D-optimal design $\opt$ for our
+%%   generic model using Theorem \ref{thm:char}. In our generic model,
+%%   the measurements we take are best utilized in reducing uncertainty
+%%   for the first $k$ eigenvectors. So, a D-optimal design arising from
+%%   our \emph{generic model} completely avoids measuring eigenvectors
+%%   $k+1$ and above.
+%%   \newline
+%%   \newline
+%%   Of course, in a real life problem --- such as the inverse problem of
+%%   the 1D heat equation --- it is likely impossible to find measurement
+%%   locations for which all eigenvectors $k+1$ and above are
+%%   zero. However, if the eigenvalues of $\fwd\prcov\fwd^*$ decay
+%%   quickly (recall the square-exponential decay for eigenvalues of the
+%%   1D heat equation in eq.\eqref{eq:decay}), a D-optimal design will
+%%   try to balance measuring a small number (i.e.~$k$) of the leading
+%%   eigenvectors.
+%% \end{quote}
+%% \begin{quote}
+%%   The abovementioned balance is explored in
+%%   Fig.~\ref{fig:eigenvectors}. We allow $m=4$ measurements in $\Omega
+%%   = [0,1]$ and observe that D-optimal measurement locations are
+%%   clustered at $x_1 = 0.31$ and $x_2 = 0.69$. Upon close inspection of
+%%   the scaled eigenvectors of $\fwd \prcov \fwd^*$, we first observe
+%%   that eigenvectors $3$ and above have negligible prior
+%%   amplitude. Since we only have $m=4$ measurements at our disposal, we
+%%   interpret these results, following Theorem \ref{thm:char}, as
+%%   implying we should only care about measuring the first and second
+%%   eigenvectors. Then, we note the D-optimal $x_1,x_2$ present a
+%%   compromise between the amplitude of the first and second
+%%   eigenvectors. For example, a measurement at $x=0.5$ would have
+%%   ignored the second eigenvector altogether, since the second
+%%   eigenvector is zero at $x=0.5$.
+%%   \newline
+%%   \newline
+%%   Now we can understand measurement clusterization for the inverse
+%%   problem of the heat equation. A D-optimal design attempts to measure
+%%   the first $k$ eigenvectors of $\fwd \prcov \fwd^*$. But there may be
+%%   (spatial) limitations on where these $k$ eigenvectors have large
+%%   amplitude. For the inverse problem of the heat equation there are
+%%   two spatial locations that present a good compromise between the
+%%   amplitudes of the first and second eigenvectors, namely $x_1$ and
+%%   $x_2$ --- see Fig.~\ref{fig:eigenvectors}. We have $m=4$
+%%   measurements at our disposal but only two spatial locations that are
+%%   a good compromise between the first and second scaled
+%%   eigenvectors. Thus, clusterization arises as a consequence of the
+%%   pigeonhole principle.
+%% \end{quote}
+
+%% \begin{figure}\label{fig:eigenvectors}
+%%   \centering
+%%   \includegraphics[width=\textwidth]{../latex/figs/eigenvectors_dst_scaled.png}
+%%   \caption{D-optimal measurement locations ($m=4$ measurements) and
+%%     weighted eigenvectors for finding the initial condition of the 1D
+%%     heat equation. Measurement locations and weighted eigenvectors are
+%%     plotted over the computational domain $\Omega = [0, 1]$
+%%     (x-axis). Measurement clusterization occurs approximately at
+%%     $0.31$ and $0.69$. These two locations are a compromise between
+%%     the magnitudes of the first and second eigenvectors, which are the
+%%     eigenvectors that a D-optimal design aims to measure. Allocating
+%%     $m=4$ measurements into two locations results in clusterization,
+%%     according to the pigeonhole principle.}
+%%   \label{fig:why}
+%% \end{figure}

### 2024-09-09 latex/flat.tex
flat.tex

@@
 \newtheorem{theorem}{Theorem}
+\newtheorem*{theorem*}{Theorem}
 \newtheorem{corollary}[theorem]{Corollary}
@@
 
+\raggedbottom
+
 \pubyear{2021}
@@
 \begin{abstract}
-  Estimation of parameters in physical processes often demands costly measurements, prompting the pursuit of an optimal measurement strategy.
-Finding such strategy is termed the problem of \emph{optimal experimental design}, abbreviated as optimal design.
-Remarkably, optimal designs can yield tightly clustered measurement locations, leading researchers to fundamentally revise the design problem just to circumvent this issue.
-Some authors introduce error correlation among error terms that are initially independent, while others restrict measurement locations to a finite set of locations.
-While both approaches may prevent clusterization, they also fundamentally alter the optimal design problem.
-
-In this study, we consider Bayesian D-optimal designs, i.e.~designs that maximize the expected Kullback-Leibler divergence between posterior and prior.
-We propose an analytically tractable model for D-optimal designs over Hilbert spaces.
-In this framework, we make several key contributions: \textbf{(a)} We establish that measurement clusterization is a generic trait of D-optimal designs with independent Gaussian measurement errors, and prove that introducing correlations among measurement error terms mitigates clusterization.
-\textbf{(b)} We characterize D-optimal designs as reducing uncertainty across a subset of prior covariance eigenvectors.
-Finally, \textbf{(c)} We leverage this characterization to argue that measurement clusterization arises as a consequence of the pigeonhole principle: when more measurements are taken than there are locations where the select eigenvectors are large and others are small --- clusterization occurs.
+  Estimation of parameters in physical processes often demands costly
+  measurements, prompting the pursuit of an optimal measurement
+  strategy. Finding such strategy is termed the problem of
+  \emph{optimal experimental design}, abbreviated as optimal
+  design. Remarkably, optimal designs can yield tightly clustered
+  measurement locations, leading researchers to fundamentally revise
+  the design problem just to circumvent this issue. Some authors
+  introduce error correlation among error terms that are initially
+  independent, while others restrict measurement locations to a finite
+  set of locations. While both approaches may prevent clusterization,
+  they also fundamentally alter the optimal design problem.
+
+  In this study, we consider Bayesian D-optimal designs, i.e.~designs
+  that maximize the expected Kullback-Leibler divergence between
+  posterior and prior. We propose an analytically tractable model for
+  D-optimal designs over Hilbert spaces. In this framework, we make
+  several key contributions: \textbf{(a)} We establish that
+  measurement clusterization is a generic trait of D-optimal designs
+  for linear inverse problems with independent Gaussian measurement
+  errors and a Gaussian prior. \textbf{(b)} We prove that introducing
+  correlations among measurement error terms mitigates
+  clusterization. \textbf{(c)} We characterize D-optimal designs as
+  reducing uncertainty across a subset of prior covariance
+  eigenvectors. \textbf{(d)} We leverage this characterization to
+  argue that measurement clusterization arises as a consequence of the
+  pigeonhole principle: when more measurements are taken than there
+  are locations where the select eigenvectors are large and others are
+  small --- clusterization occurs. Finally, we use our analysis to
+  argue against the use of Gaussian priors with linearized physical
+  models when seeking a D-optimal design.
 \end{abstract}
@@
 \section{Introduction}\label{section:intro}
-The practice of experimental design entails specifying every detail of
-running an experiment \cite{chaloner1995}. At its simplest form,
-experimental design involves choosing a set of measurements to be
-taken. For example, consider the process of searching for oil:
-measurements require digging holes in the ground and a \emph{design}
-simply indicates which holes should be dug
-\cite{horesh2008borehole}. Since measurements are expensive, only a
-limited number of them can be taken, and thus carefully choosing
-measurement locations is crucial.
+Experimental design is a systematic approach to specifying every
+possible aspect of an experiment \cite{chaloner1995}. In this study,
+we focus on experiments that involve measuring quantities arising from
+naturally occurring processes. For us, therefore, experimental design
+entails determining which specific measurements to record from a given
+process. In a typical experiment the number of measurements that can
+be recorded is limited by time, cost and other constraints. Therefore
+selecting the right set of measurements to record is crucial as it
+directly impacts the utility of the data collected, and consequently,
+the validity of the conclusions drawn.
+
+For example, consider the process of searching for oil: measurements
+require digging deep holes in the ground, known as "boreholes". A
+\emph{design} simply indicates which boreholes should be dug
+\cite{horesh2008borehole}. Since digging boreholes is expensive, only
+a limited number of them can be dug, and thus carefully choosing their
+locations is crucial.
 
@@
 allowed. For example, in borehole tomography, a measurement involves
-drilling a deep well in the ground --- a very costly endeavor. Even in
-a seemingly simpler application such as MRI, a patient cannot occupy
-the MRI machine for too long, hence only a limited number of
-measurements can be taken. Therefore, measurements should be chosen to
-most benefit the user. Typically, a user will consider some utility,
-called a \emph{design criterion} and take measurements that maximize
-this utility. Two of the most widely recognized and extensively
-studied design criteria are the A- and \emph{D-optimality} criteria.
+drilling a deep well in the ground --- a very costly endeavor. In
+medical imaging applications like MRI, the time allotted for each
+patient in the MRI machine limits the number of measurements that can
+be taken during a single session. Therefore, measurements should be
+chosen to extract as much information from the experiment. Typically,
+a user will consider some utility, called a \emph{design criterion}
+and take measurements that maximize this utility. Two of the most
+widely recognized and extensively studied design criteria are the A-
+and \emph{D-optimality} criteria.
 
@@
 to prior \cite{chaloner1995, AlexanderianGloorGhattas14,
-  CoverThomas91}. Utilizing KL divergence for quantifying differences
-between distributions is standard practice and we will not motivate it
-further here. For a discrete parameter $\param$ and data $\data$, the
-KL divergence is defined as
+  CoverThomas91}.
+Recall that for a discrete parameter $\param$ and data $\data$, the KL
+divergence is defined as
 \begin{equation}\label{eq:basic_KL}
@@
 
-Solving the 1D heat equation is straightforward. Any initial heat
-distribution $u_0\in L^2([0,1])$ that satisfies the homogeneous
-Dirichlet boundary condition eq.~\eqref{eq:heat2} is a linear
-combination of sines $\ev_n(x) = \sin(\pi n x)$, so $u_0 = \sum_{n\geq
-  1} a_n \ev_n$ for some $\{a_n\}_{n\geq 1}$. Now, note that $\ev_n$
-are in fact eigenvectors of the Laplacian: $\Delta \ev_n = -\pi^2
-n^2\ev_n$. It should not be hard to believe that $u(\cdot, T) =
-\sum_{n\geq 1} a_n \exp(-\pi^2 n^2T ) \ev_n$
-\cite{renardy2006PDE}. Thus, eigenvectors of the linear operator that
-describes the time evolution of heat are $\ev_n(x) = \sin(\pi n x)$
-with corresponding eigenvalues
+Solving the 1D heat equation is straightforward. Consider an initial
+heat distribution $u_0\in L^2([0,1])$ that satisfies the homogeneous
+Dirichlet boundary condition eq.~\eqref{eq:heat2}. This initial
+condition $u_0$ is a linear combination of sines $\ev_n(x) = \sin(\pi
+n x)$, so $u_0 = \sum_{n\geq 1} a_n \ev_n$ for some $\{a_n\}_{n\geq
+  1}$. Now, note that $\ev_n$ are in fact eigenvectors of the
+Laplacian: $\Delta \ev_n = -\pi^2 n^2\ev_n$. It should not be hard to
+believe that $u(\cdot, T) = \sum_{n\geq 1} a_n \exp(-\pi^2 n^2T )
+\ev_n$ \cite{renardy2006PDE}. Thus, eigenvectors of the linear
+operator that describes the time evolution of heat are $\ev_n(x) =
+\sin(\pi n x)$ with corresponding eigenvalues
 \begin{equation}\label{eq:decay}
@@
 
-We choose a prior for the initial condition $u_0 \sim \mathcal{N}(0,
-(-\Delta)^{-1})$, with homogeneous Dirichlet boundary condition. This
-choice of prior covariance operator ensures the posterior is
-well-defined and prior realizations are well-behaved, see
-\cite[Theorem 3.1 and Lemma 6.25]{Stuart10} for details. The second
-ingredient of a fully Bayesian formulation of the inverse problem is
-likelihood. We simply assume centered iid Gaussian measurement error
-model, giving rise to a the widely applicable Gaussian likelihood
-function.
+In order to find a fully Bayesian formulation for our inverse problem
+we first need to specify a prior on \emph{functions} defined over
+$\Omega = [0,1]$. For the sake of simplicity we would like to utilize
+a Gaussian prior. While specifying the prior mean is easy --- we take
+the zero function on $\Omega$ --- specifying a prior covariance is
+more involved. In analogy with the finite dimensional case, we seek a
+covariance \emph{operator} that is positive definite and imposes
+sufficient regularity on functions over $\Omega$. Thus, we set our
+prior for the initial condition $u_0 \sim \mathcal{N}(0,
+(-\Delta)^{-1})$, where $\Delta$ is again defined with a homogeneous
+Dirichlet boundary condition. The intuition here is that since
+$\Delta$ is a differential operator, it is "roughing" (the opposite of
+smoothing) and so $(-\Delta)^{-1}$ is smoothing. Prior realizations
+are generated, in analogy to the finite dimensional case, by smoothing
+white noise: $(-\Delta)^{-1/2}\mathcal{W}$, where $\mathcal{W}$ is
+white noise \cite{rue2011}. Our choice of prior covariance operator
+ensures the posterior is well-defined and prior realizations are
+well-behaved, see Theorem 3.1 and Lemma 6.25 in \cite{Stuart10} for
+details. The second ingredient of a fully Bayesian formulation of the
+inverse problem is likelihood. We simply assume centered iid Gaussian
+measurement error model, giving rise to a the widely applicable
+Gaussian likelihood function.
+
+Now that we have a fully Bayesian formulation of our toy inverse
+problem, we would like to find a D-optimal design. Defining D-optimal
+designs over function spaces is somewhat mathematically
+involved. Thus, to keep our discussion here focused, we will explore
+these technical details later in Section
+\ref{subsec:D_optimal_design}. For now, it suffices to know that
+D-optimal designs can be defined for our problem.
 
@@
 
-Clusterization should not be confused with either repetition nor with
-replication, which are commonly viewed as beneficial and even
-necessary aspects of an optimal design \cite{fisher1949design,
-  morris2011, schafer2001replication}. For example,
-\cite{fisher1949design} in his famous milk and tea experiment,
-suggested that repetition is a "way of enlarging the experiment and,
-thereby, increasing its sensitiveness". In another example,
-\cite{fay2000rainfall} measured the effect of rainfall on grass growth
-in plots of land. The experiment involved fifteen "rainfall
-manipulation shelters", where "Four rainfall manipulation treatments
-(three replicates) then were assigned to 12 of the plots". While it
-seems reasonable for the researchers to replicate the phenomenon they
-are trying to study, clusterization is different: a clustered design
-in the rainfall experiment would imply the researchers should take a
-repeated measurement \emph{on the same plot}, at the expense of
-measuring grass growth in other plots. In sharp contrast to repetition
-and replication, clusterization is highly nonintuitive and we will see
-that its origins run considerably deeper.
+Clusterization should not be confused with replication. Replication
+requires that the experimentalist executes multiple trials under
+circumstances that are \emph{nominally identical} \cite[Section
+  1.2.4]{morris2011}. Replication is commonly viewed as a beneficial
+and even necessary aspect of optimal experimental design
+\cite{fisher1949design, morris2011, schafer2001replication}.
+For example, \cite{fisher1949design}, suggested repeating his famous
+milk and tea experiment in order "to be able to demonstrate the
+predominance of correct classifications in spite of occasional
+errors". Unfortunately, in the experiments we consider, replication is
+impossible. For example, in the MRI problem, we cannot generate an
+individual nominally identical to the one we wish to scan.
+
+Similarly to a design implementing replication, a clustered design
+reduces the signal-to-noise ratio of the repeated measurements
+\cite{telford2007brief}. The difference is that a clustered design
+takes repeated measurements at the expense of other quantities not
+measured at all. For example, consider an experiment measuring the
+effect of rainfall on grass growth \cite{fay2000rainfall}. The
+experiment involved four rainfall manipulation "treatments"
+(i.e.~simulating different timing and quantity of rainfall), each
+replicated three times over different plots of land. Indeed, it seems
+reasonable for researchers to replicate the phenomenon they are trying
+to study. A clustered design in such an experiment would imply the
+researchers should take a repeated measurement \emph{on the same
+plot}, at the expense of measuring grass growth in other plots!
+
+To conclude our short discussion on replication vs.~clusterization:
+these are fundamentally different concepts and even though replication
+is quite intuitive, it is inapplicable to the inverse problems we
+consider in this manuscript.
 
 Researchers of inverse problems widely agree that measurement
-clusterization is undesirable
-\cite{fedorovDesignSpatialExperiments1996, nyberg2012, fedorov1997,
-  Ucinski05, neitzel2019sparse}, prompting the exploration of various
-remedies to address this issue. One approach involves merging close
-measurements \cite{fedorov1997}; however, this strategy merely
-overlooks the phenomenon of measurement clusterization. An alternative
-solution lies in \emph{clusterization-free design}s, where measurement
-locations are deliberately chosen to be distant from one another. This
-can be achieved by imposing distance constraints between measurements
-or by introducing correlated errors that account for both observation
-error and model misspecification \cite{Ucinski05}. For instance, in
-the context of time-series analysis for pharmacokinetic experiments,
+clusterization is undesirable \cite{fedorov1996, nyberg2012,
+  fedorov1997, Ucinski05, neitzel2019sparse}, prompting the
+exploration of various remedies to address this issue. One approach
+involves merging close measurements \cite{fedorov1997}; however, this
+strategy merely overlooks the phenomenon of measurement
+clusterization. An alternative solution lies in
+\emph{clusterization-free design}s, where measurement locations are
+deliberately chosen to be distant from one another. This can be
+achieved by imposing distance constraints between measurements or by
+introducing correlated errors that account for both observation error
+and model misspecification \cite{Ucinski05}. For instance, in the
+context of time-series analysis for pharmacokinetic experiments,
 measurement clusterization can be mitigated by incorporating the
@@
   on how quickly the eigenvalues of the prior covariance in
-  observation space decay. This decay depends mostly on how ill-posed
-  the problem at hand is and does not depend much on the prior. See
-  Section \ref{section:vanishing} --- particularly, Theorem
-  \ref{thm:char} and the discussion following it. We also show results
-  of numerical experiments, where simulations of our model give rise
-  to D-optimal designs that exhibit clusterization with high
-  probability in Section \ref{subsec:lemma_sims}. Thus, given the
-  genericity of our model, we expect measurement clusterization to be
-  a generic and ubiquitous phenomenon.
+  observation space decay.
+See Section \ref{section:vanishing} --- particularly, Theorem
+  \ref{thm:char} and the discussion following it. Furthermore, in
+  Section \ref{subsec:lemma_sims} we show results of numerical
+  experiments, where simulations of our model give rise to D-optimal
+  designs that exhibit clusterization with high probability. Thus,
+  given the genericity of our model, we expect measurement
+  clusterization to be a generic and ubiquitous phenomenon.
 
@@
 
-  The connection to Carath\'eodory's Theorem also utilizes the
-  concentration of measurement effort on the abovementioned select set
-  of prior eigenvectors. This concentration allows us to move from an
-  infinite-dimensional setting to finite dimensions, where the
-  conditions of Carath\'eodory's Theorem hold. We conclude that a
-  D-optimal design arises by weighting a small number of
-  measurements. If the number of allowed measurements is larger than
-  the small number dictated by Carath\'eodory's Theorem, we have
-  excessive weight on some measurements, which can be interpreted as
-  clusterization.
+  The connection to Carath\'eodory's Theorem also builds on the focus
+  of measurement effort to the abovementioned select set of prior
+  eigenvectors. This allows us to move from an infinite-dimensional
+  setting to finite dimensions, where the conditions of
+  Carath\'eodory's Theorem hold. We conclude that a D-optimal design
+  arises by weighting a small number of measurements. If the number of
+  allowed measurements is larger than the small number dictated by
+  Carath\'eodory's Theorem, we have excessive weight on some
+  measurements, which can be interpreted as clusterization.
 
@@
 
-\subsubsection{Implications}
+\subsubsection{Implications}\label{subsub:implications}
 Our answer to Question \ref{q:generic} implies that encountering
-clusterization should be expected in many different problems across
+clusterization could be expected in many different problems across
 many different scientific fields. Researchers that encounter
@@
 eigenvectors --- those with the most prior uncertainty, i.e.~those the
-practitioner cares about the most! We believe practitioners should not
-try to avoid measurement clusterization. Rather, practitioners should
-take repeated measurements (e.g.~in MRI and borehole tomography),
-increase apparatus sensitivity (e.g.~in EIT), or take consecutive
-measurements (e.g.~in the 1D heat equation). Overall, we show that
-measurement clusterization is a natural and (almost) inevitable part
-of Bayesian D-optimal designs (but see disclaimer below).
-
-One interesting implication of the analysis presented here is that
+practitioner cares about the most!
+
+A further implication of the analysis we present is that
 clusterization can serve as an evidence to the number of relevant
-eigenvectors. Since leading eigenvectors typically correspond to slow
-variations in space and/or time, clusterization could be used to
-estimate the number of relevant degrees of freedom, and even to reduce
-the complexity of a computational model, e.g.~by dropping
-discretization points.
-
-It is important to note that we do not view clustered designs as
-undesirable, nor do we believe it should be avoided at
-all. Clusterization is a peculiar phenomenon and it is perfectly
-reasonable for someone to argue against the D-optimality criterion
-based on the fact that it results in clustered designs. We have seen,
-however that there is a perfectly reasonable explanation for
-clusterization. We have shown that clusterization is an inevitable
-consequence of having a problem with some modes where uncertainty
-decays faster than others.
-
-Lastly, we believe that when clusterization arises, it should serve as
-a warning sign to practitioners. In the inverse problem of the 1D heat
-equation, clusterization occurs primarily because Laplacian
-eigenvectors \emph{do not} decay in $\Omega$. Consequently, measuring
-$u(x_1, T)$ at some point $x_1 \in \Omega$ provides information about
-$u(x_2,T)$ for distant points $x_2 \in \Omega$. Intuitively, this
-should not occur: for small $T$, the heat distribution at $x_1$ should
-give very little knowledge on the heat distribution at $x_2$. This
-behavior stems from a well-known property of the heat equation: it
-allows information to spread \emph{instantly} across the computational
-domain \cite{renardy2006PDE}. In reality, heat (and information)
-propagate at finite speeds. Of course, the known physical barrier for
-information spread is the speed of light, but we expect heat to spread
-considerably slower: heating an Olympic pool at one end should have no
-immediate effect on the temperature at the other end.
-
-Our choice of prior is also a potential major contributor to
-clusterization. Our choice of Gaussian prior similarly implies
-information is shared between distant locations in $\Omega$. Thus, we
-suggest refraining from choosing Gaussian priors with inverse
-Laplacian covariance operators. Rather, non-Gaussian priors could be
-employed instead \cite{hosseini2017, hosseini2019}.
-
-We believe that the emergence of clusterization in this context is
-thus non-physical, arising from the way the inverse problems we
-consider are phrased. Clusterization therefore indicates that the
-underlying mathematical / Bayesian model is overly permissive and
-fails to capture crucial physical constraints of the problem. We
-suggest that when clusterization occurs, practitioners should consider
-alternative models where information is localized in space and travels
-at finite speed in the medium. Such models may not only provide more
-physically accurate and meaningful results but may also mitigate the
-issue of clusterization.
+eigenvectors in the problem. These eigenvectors correspond to the
+relevant degrees of freedom in the problem. Thus, clusterization tells
+us that we might be able to reduce the complexity of a computational
+model, e.g.~by dropping discretization points.
+
+One potential cause of clusterization is our choice of prior. Gaussian
+priors, coupled with a Gaussian likelihood and a linear forward
+problem give rise to a closed form solution for the posterior via
+conjugacy. As we show in Section \ref{subsec:lemma_sims}, such
+assumptions generically give rise to clusterization. While an
+assumption of a Gaussian likelihood is standard, and rooted in the
+central limit theorem, an assumption of Gaussian prior is merely a
+matter of convenience. Therefore, we advise any practitioner who
+encounters clusterization to replace their Gaussian priors with
+non-Gaussian priors instead \cite{hosseini2017, hosseini2019}.
+
+Another potential cause of clusterization is linearization. While in a
+real applications the model is not necessarily linear, some authors
+consider a linearized version of their model when seeking optimal
+designs \cite{fedorov1996, neitzel2019sparse}. Our analysis then shows
+that the linearity of the forward problem is also an important
+ingredient in giving rise to clustered designs. Therefore, our advice
+to practitioners is to avoid linearization, and find D-optimal designs
+via other methods, e.g.~the sampling method of \cite{ryan2003}.
+
+Given the above discussion, it is our belief that the methods
+suggested by other authors to avoid clusterization are merely
+overlooking the problem. We do not view clustered designs as
+inherently bad on their own. Rather, we suggest that if a clustered
+design arises, a practitioner should revisit their modeling choices
+--- specifically, their choice of prior and model linearization. If
+the practitioner is confident in their analysis, then clustered
+designs should be avoided only to the extent necessitated by the
+physical measuring apparatus.
+
+Clusterization in D-optimal designs raises the question of whether
+D-optimal designs should be pursued at all. While other authors have
+established convergence results for decaying measurement error
+\cite{knapik2011}, space-filling measurements \cite{teckentrup2020}
+and randomly sampled designs \cite{nickl2023}, their results do not
+hold for D-optimal designs. Luckily, we can expect convergence from
+D-optimal designs (including clustered designs) in the following
+sense: the posterior uncertainty ellipsoid will contract to zero along
+each one of its eigenvectors. See Section \ref{subsub:convergence} for
+a precise statement and a proof based on Theorem \ref{thm:char} ---
+the main theorem of this manuscript.
 
@@
 In Theorem \ref{thm:char}, we also show that D-optimal designs are
-best understood in the space of \emph{observations}, see Section
-\ref{section:vanishing} for a precise statement. This is in accordance
-with previous work by \cite{koval2020}, who showed that A-optimal
-designs are best constructed in the space of observations.
+best understood in the space of \emph{observations}
+This is in accordance with previous work by \cite{koval2020}, who
+showed that A-optimal designs are best constructed in the space of
+observations.
 
@@
 \cite{kardar2007} and many others.
+
  \section{Preliminaries and Notation}\label{section:prelim}
@@
 fixed apporximation of $\delta$ the norm of the corresponding
-functional is constant (but $\neq 1$.}:
+functional is a constant $\neq 1$.}:
 
@@
 \ref{prop:tar_grad} and \ref{prop:constraints_grad} into
-\eqref{eq:Lagrange_mult1}:
+eq.~\eqref{eq:Lagrange_mult1}:
 \begin{equation}\label{eq:constrained}
@@
   $(\obsm \modcov \obsm^*)_{ij} = \meas_i(\modcov \meas_j)$, as
-  explained in \eqref{eq:modcov_explained}. We can now verify that
+  explained in the paragraph preceding
+  eq.~\eqref{eq:modcov_explained}. We can now verify that
   \begin{equation}\label{eq:observation}
@@
 construct examples where the posterior does not change upon taking a
-new measurement e.g.~if the prior variance vanishes on some
+new measurement, e.g.~if the prior variance vanishes on some
 eigenvector and a measurement is taken on said eigenvector. We do not
@@
 \ref{subsub:clusterization1} below. We also suggest a generic
-explanation for clusterization, see Section \ref{subsub:clusterization2}.
-
-\subsubsection{Clusterization in the heat equation}
+explanation for clusterization, see Section \ref{subsub:cara}.
+
+\subsubsection{Clusterization in the heat equation}\label{subsub:clusterization1}
 Consider $\fwd$ and $\prcov$ from \emph{the inverse problem of the
@@
 
-\subsubsection{Clusterization in our generic model}We can gain further insight to clusterization in our generic model,
+\subsubsection{Clusterization in our generic model}\label{subsub:cara}
+We can gain further insight to clusterization in our generic model,
 from Carath\'eodory's Theorem and the concentration on the first $k$
 eigenvectors of $\fwd \prcov \fwd^*$. We present a short discussion
-adapting arguments presented in \cite[Chapter
-  3]{silveyOptimalDesign1980} and \cite[Section
-  5.2.3]{pronzatoPazman2013}.
+adapting arguments presented in \cite[Chapter 3]{silvey1980} and
+\cite[Section 5.2.3]{pronzatoPazman2013}.
 
@@
 places weight $>1$ on a single measurement vector.
+
+\subsubsection{Convergence}\label{subsub:convergence}
+In this section we fulfill our promise from Section
+\ref{subsub:implications} and prove that the posterior uncertainty
+ellipsoid in $\hilo$ will contract to zero along every eigenvector of
+$\fwd \prcov \fwd^*$. In our proof we ignore potential problems with
+conducting inference on function spaces which are not unique to
+D-optimal designs (see e.g.~\cite{owhadi2015} for more details).
+
+First, denote $\opt_m$ a D-optimal design utilizing $m$ measurements
+and denote $k_m := \rank\opt_m^*\opt_m$. An immediate consequence of
+Theorem \ref{thm:char} is that allowing more measurements will
+eventually allow us to measure each eigenvalue, i.e.:
+\begin{equation}\label{eq:lim}
+  \lim_{m\to\infty} k_m = \infty
+\end{equation}
+
+Now, recall that part (5) of Theorem \ref{thm:char} ensures that the
+eigenvalues of the posterior pushforward covariance equal
+\begin{equation*}
+  \theta^{(k)}_i = \left ( \frac{\sum_{j=1}^{k} \lambda_j^{-1} +
+    \sigma^{-2}m}{k} \right )^{-1}, \text{ for $i\leq k$}.
+\end{equation*}
+Using the inequality for the arithmetic and harmonic means, it is easy
+to verify that
+\begin{equation*}
+  \theta^{(k)}_i \leq \lambda_{k}.
+\end{equation*}
+Since $\lim_{k\to \infty} \lambda_k= 0$, we conclude that for all $i$,
+$\lim_{k\to\infty} \theta^{(k)}_i = 0$. Combining the latter
+observation with eq.~\eqref{eq:lim}, we conclude that
+$\lim_{m\to\infty} \theta^{(k_m)}_i = 0$ for all $i$. Therefore,
+posterior uncertainty decays to zero for all eigenvectors.
  \section{Numerical Experiments}
@@
   This study is a result of research I started during my PhD studies
-under the instruction of Prof.~Georg Stadler at the Courant Institute
-of Mathematical Sciences. I would like to thank him for his great
-mentorship, attention to details and kindness. I would also like to
-thank Christian Remling, who helped me find a proof for Lemma
-\ref{lemma:free} in
+  under the instruction of Prof.~Georg Stadler at the Courant
+  Institute of Mathematical Sciences. I would like to thank him for
+  his great mentorship, attention to details and kindness. I would
+  also like to thank Christian Remling, who helped me find a proof for
+  Lemma \ref{lemma:free} in
   \href{https://mathoverflow.net/questions/280168/redistribute-diagonal-entries-of-a-matrix/280203#280203c}{Mathoverflow}.
   Last, but certainly not least, I would like to thank the three
-referees, associate editor and editor in chief for providing detailed
-and insightful reviewes that have made this manuscript a whole lot
-better.
+  referees, associate editor and editor in chief for providing
+  detailed and insightful reviewes that have made this manuscript a
+  whole lot better.
 
@@
   is managed for DOE by ORAU. All opinions expressed in this paper are
-the author's and do not necessarily reflect the policies and views of
-NSF, ORAU/ORISE, or DOE.
+  the author's and do not necessarily reflect the policies and views
+  of NSF, ORAU/ORISE, or DOE.
 \end{acks}
@@
 \bibliographystyle{ba}
-\bibliography{../../lib.bib}
+\bibliography{/Users/yairdaon/work/lib.bib}
 

### 2024-09-09 revision/rev3.tex
Restred

@@
 technical difficulties in installing packages for inverse problems
-\cite{attia2023pyoed, villa2021} I had to implement the inverse
-problem of the 1D heat equation from scratch, so I decided to leave
-out experiments in higher dimensions.
+\cite{attia2023pyoed, villa18} I had to implement the inverse problem
+of the 1D heat equation from scratch, so I decided to leave out
+experiments in higher dimensions.
 
@@
 present paper since I was not able to install common packages for
-inverse problems \cite{attia2023pyoed, villa2021} on my machine.
+inverse problems \cite{attia2023pyoed, villa18} on my machine.
 

### 2024-09-09 movies.py

@@ -40,7 +40,7 @@ def main():
     #ax.set_ylim(np.min(u), np.max(u))
     ax.set_xlabel('Position')
     ax.set_ylabel('Temperature')
-    plt.savefig('latex/forward_heat_equation.png')
+    plt.savefig('latex/figs/forward_heat_equation.png')
 
     time_template = 'Time elapsed {}' 
     time_text = ax.text(0.95, 0.95, '', transform=ax.transAxes, ha='right', va='top')
@@ -53,7 +53,7 @@ def main():
         return line, time_text
 
     ani = animation.FuncAnimation(fig, update, frames=n, blit=True)
-    ani.save('latex/forward_heat_equation.mp4', writer='ffmpeg', fps=30)
+    ani.save('latex/figs/forward_heat_equation.mp4', writer='ffmpeg', fps=30)
     print('Done forward animation')
     
     ######################################
@@ -67,7 +67,7 @@ def main():
     ax.set_ylim(0, 0.7)
     ax.set_xlabel('Position')
     ax.set_ylabel('Temperature')
-    plt.savefig('latex/reverse_heat_equation.png')
+    plt.savefig('latex/figs/reverse_heat_equation.png')
 
     u_reversed = us[::-1]
     time_reversed = np.array(times)[::-1]
@@ -82,7 +82,7 @@ def main():
         return line, time_text
 
     ani = animation.FuncAnimation(fig, update, frames=n, blit=True)
-    ani.save('latex/reverse_heat_equation.mp4', writer='ffmpeg', fps=30)
+    ani.save('latex/figs/reverse_heat_equation.mp4', writer='ffmpeg', fps=30)
     print('Done backwards animation')
     plt.show()
 

### 2024-09-09 latex/shell.sh
baa

@@ -1,10 +1,6 @@
+latexmk -c
 latexpand ba.tex > flat.tex
-latexdiff --math-markup=OFF old.tex flat.tex > diff.tex
-latexmk --pdf ba.tex
-latexmk --pdf ba.tex
-latexmk --pdf ba.tex
+latexdiff --math-markup=OFF old/old_flat.tex flat.tex > diff.tex
 latexmk --pdf ba.tex
 latexmk --pdf diff.tex
-latexmk --pdf diff.tex
-latexmk --pdf diff.tex
-latexmk --pdf diff.tex
+

### 2024-12-08 latex/intro.tex
BA minor revision

@@
 
+As another example illustrating the differences between replication
+and clusterization, consider an experiment involving a chemical
+reaction at different temperatures. An investigator may choose to
+\emph{replicate} the experiment at $10^\circ C$ and $30^\circ
+C$. Within a single replication, the investigator may also choose to
+take measurements at different times. If the investigator chooses to
+take two measurements at the same time (within a single replication),
+then the investigator has utilized a \emph{clustered} design. While
+replication is intuitively sensible for this chemical experiment,
+clusterization is not.
+
 To conclude our short discussion on replication vs.~clusterization:
@@
 assumptions generically give rise to clusterization. While an
-assumption of a Gaussian likelihood is standard, and rooted in the
-central limit theorem, an assumption of Gaussian prior is merely a
-matter of convenience. Therefore, we advise any practitioner who
-encounters clusterization to replace their Gaussian priors with
-non-Gaussian priors instead \cite{hosseini2017, hosseini2019}.
+assumption of independent Gaussian noise (likelihood) is standard, and
+rooted in the central limit theorem, an assumption of Gaussian prior
+is merely a matter of convenience. Therefore, we advise practitioners
+encountering clusterization to replace Gaussian priors with
+non-Gaussian ones \cite{hosseini2017, hosseini2019}, as these are not
+only more realistic but also expected to mitigate clusterization.
 

### 2024-12-08 latex/flat.tex
before small reverting

@@
 \newcommand{\postopt}{\mu_{\textup{post}}^{\data, \opt}}
+
+\raggedbottom
 
@@
 \newtheorem{remark}[theorem]{Remark}
-
-\raggedbottom
 
@@
 assumptions generically give rise to clusterization. While an
-assumption of a Gaussian likelihood is standard, and rooted in the
-central limit theorem, an assumption of Gaussian prior is merely a
-matter of convenience. Therefore, we advise any practitioner who
-encounters clusterization to replace their Gaussian priors with
-non-Gaussian priors instead \cite{hosseini2017, hosseini2019}.
+assumption of a Gaussian likelihood (i.e.~independent Gaussian noise)
+is standard, and rooted in the central limit theorem, an assumption of
+Gaussian prior is merely a matter of convenience. Therefore, we advise
+practitioners encountering clusterization to replace Gaussian priors
+with non-Gaussian ones \cite{hosseini2017, hosseini2019}, as these are
+not only more realistic but also expected to mitigate clusterization.
 
@@
   \end{enumerate}
+
+  Then:
+  \begin{enumerate}
+  \item  \(\tr{\obs^*\obs} = m\).
+  \item \(\obs^*\obs\) and \(\fwd\prcov\fwd^*\) are simultaneously
+    diagonalizable.
+  \item \(k := \rank \obs^*\obs \leq m\) and
+    \begin{equation*}
+      \tar(\obs) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i). \end{equation*}
+\item
+    \begin{equation*}
+        \eta_i = \begin{cases}
+          \frac{m}{k} - \sigma^2 \lambda_i^{-1} + \sigma^2 \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1} & 1 \leq i \leq k \\
+          0 & i > k 
+        \end{cases}.
+    \end{equation*}
+  \item The covariance of the pushforwad \(\fwd_{*} \post\) is \(\left
+    ( (\fwd \prcov \fwd^*)^{-1} + \sigma^{-2} \obs^*\obs \right
+    )^{-1}\) and its eigenvalues are
+    \begin{equation*}
+      \theta_i =
+      \begin{cases}
+        \left(\frac{\sum_{j=1}^k \lambda_j^{-1} + \sigma^{-2}m}{k} \right )^{-1} & i \leq k \\
+        \lambda_i &  i > k 
+      \end{cases}
+    \end{equation*}
+  \end{enumerate}
 \end{theorem}
-
 \begin{proof}

### 2024-12-08 latex/intro.tex
before small reverting

@@
 assumptions generically give rise to clusterization. While an
-assumption of independent Gaussian noise (likelihood) is standard, and
-rooted in the central limit theorem, an assumption of Gaussian prior
-is merely a matter of convenience. Therefore, we advise practitioners
-encountering clusterization to replace Gaussian priors with
-non-Gaussian ones \cite{hosseini2017, hosseini2019}, as these are not
-only more realistic but also expected to mitigate clusterization.
+assumption of a Gaussian likelihood (i.e.~independent Gaussian noise)
+is standard, and rooted in the central limit theorem, an assumption of
+Gaussian prior is merely a matter of convenience. Therefore, we advise
+any practitioner who encounters clusterization to replace their
+Gaussian priors with non-Gaussian priors instead \cite{hosseini2017,
+  hosseini2019}, as these are not only more realistic but also
+expected to mitigate clusterization.
 

### 2024-12-08 latex/intro.tex
Submitted minor revision

@@
 and clusterization, consider an experiment involving a chemical
-reaction at different temperatures. An investigator may choose to
-\emph{replicate} the experiment at $10^\circ C$ and $30^\circ
-C$. Within a single replication, the investigator may also choose to
-take measurements at different times. If the investigator chooses to
-take two measurements at the same time (within a single replication),
-then the investigator has utilized a \emph{clustered} design. While
+reaction conducted at different temperatures. An investigator may
+choose to repeat the experiment at $10^\circ C$ and $30^\circ C$,
+giving rise to \emph{replication}. Within a single experiment, the
+investigator may also sample the concentrations of the reagents
+through the course of the experiment. Simultaneous sampling (within a
+single experiment) gives rise to \emph{clusterization}. While
 replication is intuitively sensible for this chemical experiment,
@@
 
-Researchers of inverse problems widely agree that measurement
-clusterization is undesirable \cite{fedorov1996, nyberg2012,
-  fedorov1997, Ucinski05, neitzel2019sparse}, prompting the
-exploration of various remedies to address this issue. One approach
-involves merging close measurements \cite{fedorov1997}; however, this
-strategy merely overlooks the phenomenon of measurement
-clusterization. An alternative solution lies in
-\emph{clusterization-free design}s, where measurement locations are
-deliberately chosen to be distant from one another. This can be
-achieved by imposing distance constraints between measurements or by
-introducing correlated errors that account for both observation error
-and model misspecification \cite{Ucinski05}. For instance, in the
-context of time-series analysis for pharmacokinetic experiments,
-measurement clusterization can be mitigated by incorporating the
-modeling of auto-correlation time within the noise terms
-\cite{nyberg2012}.
+Measurement clusterization is generally considered undesirable
+\cite{fedorov1996, nyberg2012, fedorov1997, Ucinski05,
+  neitzel2019sparse}. It is counterintuitive for a so-called "optimal"
+design to prioritize marginally improving accuracy in a small region
+over taking measurements in unexplored regions of the computational
+domain, where they could significantly reduce uncertainty.
+
+The issue with clusterization is evident in
+Fig.~\ref{fig:clusterization_illustration}, where most clustered
+designs shown place little to no measurement weight in the center of
+the domain, where prior uncertainty is largest. It seems as though a
+clustered design assumes that focusing measurement efforts on a small
+subset of the domain can somehow yield information about the entire
+domain. Intuitively, this is an unreasonable assumption, albeit it
+holds for \emph{linear} models.
+
+The view that clusterization is undesirable prompted the exploration
+of various remedies to address this issue. One approach involves
+merging close measurements \cite{fedorov1997}; however, this strategy
+merely overlooks the phenomenon of measurement clusterization. An
+alternative solution lies in \emph{clusterization-free design}s, where
+measurement locations are deliberately chosen to be distant from one
+another. This can be achieved by imposing distance constraints between
+measurements or by introducing correlated errors that account for both
+observation error and model misspecification \cite{Ucinski05}. For
+instance, in the context of time-series analysis for pharmacokinetic
+experiments, measurement clusterization can be mitigated by
+incorporating the modeling of auto-correlation time within the noise
+terms \cite{nyberg2012}.
 

### 2025-01-08 latex/ba.tex
Befire minior revision
before small reverting
Slides getting done and arxiv also finished so it made some modifications

@@
 
-%
-\pubyear{2021}
+%% Packages
+\RequirePackage{amsthm,amsmath,amsfonts,amssymb}
+\RequirePackage[numbers]{natbib}
+%\RequirePackage[authoryear]{natbib}%% uncomment this for author-year citations
+\RequirePackage[colorlinks,citecolor=blue,urlcolor=blue,backref=page,backref=page]{hyperref}
+\RequirePackage{graphicx}
+
+\pubyear{2024}
+\arxiv{2007.12032}
 \volume{TBA}
 \issue{TBA}
-% \doi{0000}
-%\arxiv{}
 \firstpage{1}
@@
 
-\usepackage{amsthm}
-\usepackage{amsmath}
-\usepackage{natbib}
-\usepackage[colorlinks,citecolor=blue,urlcolor=blue,filecolor=blue,backref=page]{hyperref}
-\usepackage{graphicx}
-
 \startlocaldefs
-% ** Local definitions **
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% Uncomment next line to change            %%
+%% the type of equation numbering           %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%\numberwithin{equation}{section}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% For Axiom, Claim, Corollary, Hypothesis, %%
+%% Lemma, Theorem, Proposition              %%
+%% use \theoremstyle{plain}                 %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\theoremstyle{plain}
+\newtheorem{axiom}{Axiom}
+\newtheorem{claim}[axiom]{Claim}
+\newtheorem{theorem}{Theorem}[section]
+\newtheorem{lemma}[theorem]{Lemma}
+\newtheorem{proposition}[theorem]{Proposition}
+\newtheorem{corollary}[theorem]{Corollary}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% For Assumption, Definition, Example,     %%
+%% Notation, Property, Remark, Fact         %%
+%% use \theoremstyle{definition}            %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\theoremstyle{definition}
+\newtheorem{definition}[theorem]{Definition}
+\newtheorem*{example}{Example}
+\newtheorem*{fact}{Fact}
+\newtheorem*{ctheorem}{Theorem}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%%                                          %%
+%% For Case use \theoremstyle{remark}       %%
+%%                                          %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\theoremstyle{remark}
+\newtheorem{case}{Case}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Please put your definitions here:        %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \endlocaldefs
@@
 
-%% *** Frontmatter ***
-
 \begin{frontmatter}
   \title{Measurement Clusterization in Bayesian D-optimal Designs in Infinite Dimensions}
-
-%\title{\support{}}
   \runtitle{Measurement Clusterization in D-optimal Designs}
-%\author{Yair Daon}
-%\numberwithin{equation}{section}
 
 \begin{aug}
-\author{\fnms{Yair} \snm{Daon}\thanksref{addr1}\ead[label=e1]{yair.daon@gmail.com}}
-\runauthor{Yair Daon}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Only one address is permitted per author. %%
+%% Only division, organization and e-mail is %%
+%% included in the address.                  %%
+%% Additional information can be included in %%
+%% the Acknowledgments section if necessary. %%
+%% ORCID can be inserted by command:         %%
+%% \orcid{0000-0000-0000-0000}               %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\author[A]{\fnms{Yair}~\snm{Daon}\ead[label=e1]{yair.daon@gmail.com}\orcid{0000-0002-7491-2008}}
+%% \author[A]{\fnms{Second}~\snm{Author}\ead[label=e2]{second@somewhere.com}
+%% \and
+%% \author[B]{\fnms{Third}~\snm{Author}\ead[label=e3]{third@somewhere.com}}
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+%% Addresses                                %%
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\address[A]{Azrieli Faculty of Medicine, Bar-Ilan University 
+\printead[presep={,\ }]{e1}}
 
-\address[addr1]{Azrieli Faculty of Medicine, Bar-Ilan University, Safed, Israel
-    \printead{e1} % print email address of "e1"
-}
-
+%% \address[B]{Department,
+%% University or Company Name\printead[presep={,\ }]{e2,e3}}
+%% \runauthor{F. Author et al.}
 \end{aug}
@@
   measurements, prompting the pursuit of an optimal measurement
-  strategy. Finding such strategy is termed the problem of
-  \emph{optimal experimental design}, abbreviated as optimal
-  design. Remarkably, optimal designs can yield tightly clustered
-  measurement locations, leading researchers to fundamentally revise
-  the design problem just to circumvent this issue. Some authors
-  introduce error correlation among error terms that are initially
-  independent, while others restrict measurement locations to a finite
-  set of locations. While both approaches may prevent clusterization,
-  they also fundamentally alter the optimal design problem.
+  strategy. Finding such strategy is termed the problem of optimal
+  experimental design, abbreviated as optimal design. Remarkably,
+  optimal designs can yield tightly clustered measurement locations,
+  leading researchers to fundamentally revise the design problem just
+  to circumvent this issue. Some authors introduce error correlation
+  among error terms that are initially independent, while others
+  restrict measurement locations to a finite set of locations. While
+  both approaches may prevent clusterization, they also fundamentally
+  alter the optimal design problem.
 
@@
   D-optimal designs over Hilbert spaces. In this framework, we make
-  several key contributions: \textbf{(a)} We establish that
+  several key contributions: (a) We establish that
   measurement clusterization is a generic trait of D-optimal designs
   for linear inverse problems with independent Gaussian measurement
-  errors and a Gaussian prior. \textbf{(b)} We prove that introducing
+  errors and a Gaussian prior. (b) We prove that introducing
   correlations among measurement error terms mitigates
-  clusterization. \textbf{(c)} We characterize D-optimal designs as
+  clusterization.(c) We characterize D-optimal designs as
   reducing uncertainty across a subset of prior covariance
-  eigenvectors. \textbf{(d)} We leverage this characterization to
+  eigenvectors. (d) We leverage this characterization to
   argue that measurement clusterization arises as a consequence of the
@@
 \kwd[; secondary ]{28C20}
+\end{keyword}
+
+\begin{keyword}
+\kwd{D-optimal design}
+\kwd{inverse problem}
+\kwd{clusterization}
 \end{keyword}
@@
   %almost become my coautoh
+\end{acks}
 
+\begin{funding} 
   This research was supported in part by an appointment with the
@@
   % Post-Doctoral Scholarship.
-\end{acks}
+\end{funding}
 
 \bibliographystyle{ba}
-\bibliography{/Users/yairdaon/work/lib.bib}
+\bibliography{~/lib.bib}
 

### 2025-01-08 latex/intro.tex
Slides getting done and arxiv also finished so it made some modifications

@@
 Experimental design is a systematic approach to specifying every
-possible aspect of an experiment \cite{chaloner1995}. In this study,
+possible aspect of an experiment \citet{chaloner1995}. In this study,
 we focus on experiments that involve measuring quantities arising from
@@
 \emph{design} simply indicates which boreholes should be dug
-\cite{horesh2008borehole}. Since digging boreholes is expensive, only
+\citet{horesh2008borehole}. Since digging boreholes is expensive, only
 a limited number of them can be dug, and thus carefully choosing their
@@
 the physical world utilizing a physical model of a phenomenon of
-interest \cite{tarantola2005,kaipio2005}. In an inverse problem, we
+interest \citet{tarantola2005,kaipio2005}. In an inverse problem, we
 seek to infer physical quantities, based on observations and a model
@@
 electrode locations on the skin, leveraging known equations of
-electric current flow \cite{horesh2010impedance}. Magnetic Resonance
+electric current flow \citet{horesh2010impedance}. Magnetic Resonance
 Imaging (MRI) also entails solving an inverse problem. There,
@@
 view the internals of a body in a noninvasive manner
-\cite{horesh2008mri}. An inverse problem also arises in oil and gas
+\citet{horesh2008mri}. An inverse problem also arises in oil and gas
 exploration, where acoustic waves are sent through "boreholes" ---
@@
 of the subsurface in a process called \emph{borehole tomography}
-\cite{horesh2008borehole}. Inverse problems also arise in many other
-areas of seismology and geology \cite{rabinowitz1990, steinberg1995}
-and medical imaging \cite{tarantola2005}. In many inverse problems the
+\citet{horesh2008borehole}. Inverse problems also arise in many other
+areas of seismology and geology \citet{rabinowitz1990, steinberg1995}
+and medical imaging \citet{tarantola2005}. In many inverse problems the
 goal is to infer some numerical parameter. However, in the formulation
@@
 maximize the expected Kullback-Leibler (KL) divergence from posterior
-to prior \cite{chaloner1995, AlexanderianGloorGhattas14,
+to prior \citet{chaloner1995, AlexanderianGloorGhattas14,
   CoverThomas91}.
@@
 D-optimal design minimizes the determinant of the posterior covariance
-\cite{chaloner1995}. In Section \ref{subsec:D_optimal_design} we give
+\citet{chaloner1995}. In Section \ref{subsec:D_optimal_design} we give
 a more general definition of the D-optimality criterion that applies
@@
 believe that $u(\cdot, T) = \sum_{n\geq 1} a_n \exp(-\pi^2 n^2T )
-\ev_n$ \cite{renardy2006PDE}. Thus, eigenvectors of the linear
+\ev_n$ \citet{renardy2006PDE}. Thus, eigenvectors of the linear
 operator that describes the time evolution of heat are $\ev_n(x) =
@@
 white noise: $(-\Delta)^{-1/2}\mathcal{W}$, where $\mathcal{W}$ is
-white noise \cite{rue2011}. Our choice of prior covariance operator
+white noise \citet{rue2011}. Our choice of prior covariance operator
 ensures the posterior is well-defined and prior realizations are
-well-behaved, see Theorem 3.1 and Lemma 6.25 in \cite{Stuart10} for
+well-behaved, see Theorem 3.1 and Lemma 6.25 in \citet{Stuart10} for
 details. The second ingredient of a fully Bayesian formulation of the
@@
 observed to yield remarkably similar measurements in certain cases
-\cite{fedorov1996, nyberg2012, fedorov1997, Ucinski05,
+\citet{fedorov1996, nyberg2012, fedorov1997, Ucinski05,
   neitzel2019sparse}. This phenomenon is illustrated for our toy
@@
 two sets of measurements that are identical. Following
-\cite{Ucinski05}, we refer to this intriguing phenomenon as
+\citet{Ucinski05}, we refer to this intriguing phenomenon as
 \emph{measurement clusterization}. We consider a design to be
@@
 requires that the experimentalist executes multiple trials under
-circumstances that are \emph{nominally identical} \cite[Section
+circumstances that are \emph{nominally identical} \citet[Section
   1.2.4]{morris2011}. Replication is commonly viewed as a beneficial
 and even necessary aspect of optimal experimental design
-\cite{fisher1949design, morris2011, schafer2001replication}.
-For example, \cite{fisher1949design}, suggested repeating his famous
+\citet{fisher1949design, morris2011, schafer2001replication}.
+For example, \citet{fisher1949design}, suggested repeating his famous
 milk and tea experiment in order "to be able to demonstrate the
@@
 reduces the signal-to-noise ratio of the repeated measurements
-\cite{telford2007brief}. The difference is that a clustered design
+\citet{telford2007brief}. The difference is that a clustered design
 takes repeated measurements at the expense of other quantities not
 measured at all. For example, consider an experiment measuring the
-effect of rainfall on grass growth \cite{fay2000rainfall}. The
+effect of rainfall on grass growth \citet{fay2000rainfall}. The
 experiment involved four rainfall manipulation "treatments"
@@
 Measurement clusterization is generally considered undesirable
-\cite{fedorov1996, nyberg2012, fedorov1997, Ucinski05,
+\citet{fedorov1996, nyberg2012, fedorov1997, Ucinski05,
   neitzel2019sparse}. It is counterintuitive for a so-called "optimal"
@@
 of various remedies to address this issue. One approach involves
-merging close measurements \cite{fedorov1997}; however, this strategy
+merging close measurements \citet{fedorov1997}; however, this strategy
 merely overlooks the phenomenon of measurement clusterization. An
@@
 measurements or by introducing correlated errors that account for both
-observation error and model misspecification \cite{Ucinski05}. For
+observation error and model misspecification \citet{Ucinski05}. For
 instance, in the context of time-series analysis for pharmacokinetic
@@
 incorporating the modeling of auto-correlation time within the noise
-terms \cite{nyberg2012}.
+terms \citet{nyberg2012}.
 
@@
 circumvent the problem of measurement clusterization by restricting
-measurements to a coarse grid in \(\Omega\) \cite{koval2020,
+measurements to a coarse grid in \(\Omega\) \citet{koval2020,
   alexanderian2021, attia2022, alexanderian2014, alexanderian2016,
@@
 desired binary design over the coarse grid
-\cite{horesh2008borehole}. Others progressively relax the \(\ell_1\)
+\citet{horesh2008borehole}. Others progressively relax the \(\ell_1\)
 penalty to an \(\ell_0\) penalty via a continuation method
-\cite{alexanderian2016, alexanderian2014}. Others cast the problem of
+\citet{alexanderian2016, alexanderian2014}. Others cast the problem of
 finding optimal measurement weights as a stochastic optimization
-problem \cite{attia2022stochastic}. All of the aforementioned methods
+problem \citet{attia2022stochastic}. All of the aforementioned methods
 may indeed find a binary optimal design restricted to a given coarse
@@
 any practitioner who encounters clusterization to replace their
-Gaussian priors with non-Gaussian priors instead \cite{hosseini2017,
+Gaussian priors with non-Gaussian priors instead \citet{hosseini2017,
   hosseini2019}, as these are not only more realistic but also
@@
 consider a linearized version of their model when seeking optimal
-designs \cite{fedorov1996, neitzel2019sparse}. Our analysis then shows
+designs \citet{fedorov1996, neitzel2019sparse}. Our analysis then shows
 that the linearity of the forward problem is also an important
@@
 to practitioners is to avoid linearization, and find D-optimal designs
-via other methods, e.g.~the sampling method of \cite{ryan2003}.
+via other methods, e.g.~the sampling method of \citet{ryan2003}.
 
@@
 established convergence results for decaying measurement error
-\cite{knapik2011}, space-filling measurements \cite{teckentrup2020}
-and randomly sampled designs \cite{nickl2023}, their results do not
+\citet{knapik2011}, space-filling measurements \citet{teckentrup2020}
+and randomly sampled designs \citet{nickl2023}, their results do not
 hold for D-optimal designs. Luckily, we can expect convergence from
@@
 %% , see Section \ref{section:vanishing} for a precise statement.
-This is in accordance with previous work by \cite{koval2020}, who
+This is in accordance with previous work by \citet{koval2020}, who
 showed that A-optimal designs are best constructed in the space of
@@
 studying models that are bare-bones simplifications of real systems,
-e.g.~the Ising model \cite{cipra1987}, the Lorenz system \cite{brin},
-the Lotka-Volterra equations \cite{logan2006}, the Carnot engine
-\cite{kardar2007} and many others.
-
-%% When clusterization arises, we believe it should serve as a warning
-%% signal to practitioners. In the inverse problem of the 1D heat
-%% equation, clusterization occurs primarily because Laplacian
-%% eigenvectors do not decay in $\Omega$. Consequently, measuring $u(x_1,
-%% T)$ at some point $x_1 \in \Omega$ provides information about
-%% $u(x_2,T)$ for distant points $x_2 \in \Omega$. Intuitively, this
-%% should not happen: for small $T$, the heat distribution at $x_1$
-%% should give very little knowledge on the heat distribution at
-%% $x_2$. This behavior stems from a well-known property of the heat
-%% equation: it allows information to spread instantly across the
-%% computational domain \cite{renardy2006PDE}. In reality, heat (and
-%% information) propagate at finite speeds. Of course, the known physical
-%% barrier for information spread is the speed of light, but we expect
-%% heat and information to spread considerably slower. Thus,
-%% clusterization may suggest that our physical model itself is at fault.
-
-%% We believe that the emergence of clusterization in this context might
-%% be non-physical, arising from the way the inverse problems we consider
-%% are phrased. Clusterization therefore indicates that the underlying
-%% mathematical / Bayesian model is overly permissive and fails to
-%% capture crucial physical constraints of the problem. We suggest that
-%% when clusterization occurs, practitioners should consider alternative
-%% models where information is localized in space and travels at finite
-%% speed in the medium. Such models may not only provide more physically
-%% accurate and meaningful results but may also mitigate the issue of
-%% clusterization.
-
-%% be avoided at all. Nothing in our analysis indicates that we should
-%% expect any pathological behavior when utilizing D-optimal designs. On
-%% the contrary: we show that D-optimal designs (clustered or not) reduce
-%% uncertainty of prior covariance eigenvectors where it is
-%% highest. Thus, in a sense we hope to make precise in a future study,
-%% we expect convergence of for D-optimal designs to be fastest.
-
-%% practitioners should not try to avoid measurement
-%% clusterization. Rather, practitioners should take repeated
-%% measurements (e.g.~in MRI and borehole tomography), increase apparatus
-%% sensitivity (e.g.~in EIT), or take consecutive measurements (e.g.~in
-%% the 1D heat equation). Overall, we show that measurement
-%% clusterization is a natural and (almost) inevitable part of Bayesian
-%% D-optimal designs (but see disclaimer below).
+e.g.~the Ising model \citet{cipra1987}, the Lorenz system \citet{brin},
+the Lotka-Volterra equations \citet{logan2006}, the Carnot engine
+\citet{kardar2007} and many others.

### 2025-01-08 latex/prelims.tex
Slides getting done and arxiv also finished so it made some modifications

@@
 should really be thought of as function spaces, utilizing a setup
-similar to \cite{knapik2011}. A deeper treatment of the foundations
+similar to \citet{knapik2011}. A deeper treatment of the foundations
 for inverse problems over function spaces can be found in
-\cite{Stuart10}.
+\citet{Stuart10}.
 
@@
 %% operators to be merged \(\tmp := \obs \fwd\)
-%% \cite{AlexanderianGloorGhattas14}, but the analysis carried out in the
+%% \citet{AlexanderianGloorGhattas14}, but the analysis carried out in the
 %% following sections requires that \(\fwd\) and \(\obs\) are explicitly
-%% separated, as in \cite{attia2022stochastic,
+%% separated, as in \citet{attia2022stochastic,
 %%   cvetkovic2023choosing}.}
@@
 Taking \(\modcov = 0\) is a common practice
-\cite{tarantola2005,kaipio2005,Vogel02} and then \(\Sigma =
+\citet{tarantola2005,kaipio2005,Vogel02} and then \(\Sigma =
 \sigma^2I_m\) is a scalar matrix which does not depend on \(\obs\).
@@
 ,\prcov)\) with some appropriate covariance operator \(\prcov\) on
-\(\hilp\) \cite{Stuart10}. For example, for the inverse problem of the
+\(\hilp\) \citet{Stuart10}. For example, for the inverse problem of the
 1D heat equation we chose $\prcov = (-\Delta)^{-1}$, as described in
 Section \ref{subsec:toy}. Note that \(\fwd \prcov \fwd^*\) is the
-prior covariance in \(\hilo\) \cite{Stuart10}, and as such is assumed
+prior covariance in \(\hilo\) \citet{Stuart10}, and as such is assumed
 invertible --- an assumption which we will use later (if \(\fwd\) has
@@
 this study, but the posterior covariance operator $\postcov$ is given
-by the known formula \cite{Stuart10}:
+by the known formula \citet{Stuart10}:
 \begin{align}\label{eq:postcov}
@@
 The study of D-optimal designs for Bayesian linear inverse problems in
-infinite dimensions was pioneered by \cite{AlexanderianGloorGhattas14,
+infinite dimensions was pioneered by \citet{AlexanderianGloorGhattas14,
   alexanderian2018efficient}. The main result we will make use of is
@@
 
-\begin{theorem}[Alexanderian, Gloor, Ghattas \cite{AlexanderianGloorGhattas14}]\label{thm:d_optimality}
+\begin{theorem}[Alexanderian, Gloor, Ghattas \citet{AlexanderianGloorGhattas14}]\label{thm:d_optimality}
   Let \(\pr = \normal(\prmean,\prcov)\) be a Gaussian prior on \(\hilp\)
@@
 
-In \cite{AlexanderianGloorGhattas14, alexanderian2018efficient},
+In \citet{AlexanderianGloorGhattas14, alexanderian2018efficient},
 results are stated for \(\Sigma=I\) (implied by \(\modcov =
 0,\sigma^2=1\)), but these results also hold for more general
-covariance matrices \cite[p. 681]{AlexanderianGloorGhattas14}.
+covariance matrices \citet[p. 681]{AlexanderianGloorGhattas14}.
 
@@
 to the finite-dimensional case.
-
-%% \subsection{Notation Summary}\label{subsec:notation}
-%%   Our notation is summarized below. We let:
-%%   \begin{itemize}
-%%   \item \(\hilp, \hilo\) Hilbert spaces.
-%%   \item \(\fwd:\hilp \to \hilo\) a linear compact operator.
-%%   \item \(\pr \sim \mathcal{N}(0, \prcov)\) prior Gaussian measure on
-%%     \(\hilp$, with prior covariance operator \(\prcov:\hilp \to \hilp$.
-%%   \item \(\obs: \hilo \to \mathbb{R}^m\) measurement operator, where \(m
-%%     \in \mathbb{N}\) is the number of measurements taken.
-%%   \item \(\sigma^2 \in \mathbb{R}_{+}\) observation noise variance.
-%%   \item \(\modcov\) model error covariance operator.
-%%     %% \(\data = \obs \fwd \param + \eps$, where \(\eps \in
-%%     %% \mathbb{R}^m\) isiid \(\mathcal{N}(0, \sigma^2)\) noise.
-%%   \item \(\Sigma(\obs) = \obs \modcov \obs^* + \sigma^2I$. 
-%%   \item \(\post\) the posterior measure, with covariance \(\postcov$.
-%%   \item A D-optimality design criterion
-%%     \cite{AlexanderianGloorGhattas14}:
-%%     \begin{align*}
-%%       \begin{split}
-%%         \tar(\obs) :&= \mathbb{E}_{\data}\left [ D_{\text{KL}} (\post || \pr ) \right ] \\
-%%         % 
-%%         % 
-%%         % 
-%%         &= \frac12 \log \det ( I + \prcov^{1/2} \fwd ^* \obs^* \Sigma(\obs)^{-1} \obs
-%%         \fwd \prcov^{1/2}).
-%%       \end{split}
-%%     \end{align*}
-%%   %% \item \(\{\lambda_i\}_{i=1}^\infty\) eigenvalues of \(\fwd\prcov\fwd^*$
-%%   %%   in decreasing order of magnitude.
-%%   %% %% \item \(\{\ev_i\}_{i=1}^\infty\) their corresponding eigenvectors.
-%%   %% \item \(\{\eta_i\}_{i=1}^\infty\) eigenvalues of \(\obs^*\obs$.
-%%   \end{itemize}
-
-%% \subsection{Sequential vs Simultaneous Optimization}\label{subsec:seq_vs_sim}
-%% From defintion \ref{def:d_optimality} we wish to characterize solution(s) of the
-%% following optimization problem for \(\tar$. %%: (\hilo^*)^m \to \R$:
-%% \begin{align}\label{eq:optimization}
-%%   \obs^{\star} := \argmax_{\obs} \tar( \obs ) 
-%%   = \argmax_{\obs} \frac12 \log \det 
-%%   (I + \prcov^{1/2} \fwd^*\obs^* \Sigma^{-1} \obs \fwd \prcov^{1/2}),
-%% \end{align}
-%% where \(\obs\) is constrained to some allowed set of observations. We
-%% call this problem ``simultaneous optimization'', since all
-%% observations are decided on simulatneously.
-
-%% For computational reasons, one may prefer to find the best
-%% observations in a sequential manner. Denote
-%% \begin{equation}\label{eq:def_obs_k}
-%%   \obs_k := (\meas_1,\dots, \meas_k)^t,  k\leq m.
-%% \end{equation}
-%% Sequential optimal design proceeds as follows. Find \(\meas_1\) by
-%% maximizing \(\tar(\obs_1)$. Then, keeping \(\meas_1\) fixed --- find
-%% \(\meas_2\) as the maximizer of \(\tar(\obs_2)$. Then, find \(\meas_3\) by
-%% keeping \(\meas_1,\meas_2\) fixed and taking \(\meas_3\) as the maximizer
-%% of \(\tar(\obs_3)$. Continue this way until \(\obs_m = \obs\) is
-%% found, where \(m\) is the number of available observations. %% It is
-%% %% important to notice that this scheme does not require actually
-%% %% observing data --- in \eqref{eq:objective} data is averaged out.
-
-%% The analysis in this paper is conducted for the general simultaneous
-%% optimization case. The sequential optimization case is dealt with in
-%% section \ref{subsec:clusterization_sequential}. It is important to
-%% note, however that all conclusions we arrive at for the simultaneous
-%% case easily specialize to the sequential case by considering the
-%% posterior as the next sequential step's prior.

### 2025-01-08 latex/vanishing.tex
compiles
Slides getting done and arxiv also finished so it made some modifications

@@
     \item The D-optimality design criterion
-    \cite{AlexanderianGloorGhattas14}:
+    \citet{AlexanderianGloorGhattas14}:
     \begin{align*}
@@
   \end{enumerate}
+%% \end{theorem}
+
+  Then:
+  \begin{enumerate}
+  \item  \(\tr{\obs^*\obs} = m\).
+  \item \(\obs^*\obs\) and \(\fwd\prcov\fwd^*\) are simultaneously
+    diagonalizable.
+  \item \(k := \rank \obs^*\obs \leq m\) and
+    \begin{equation*}
+      \tar(\obs) = \frac12 \sum_{i=1}^{k} \log (1 + \sigma^{-2}\lambda_i\eta_i). %= \frac12 \sum_{i=1}^{m} \log (1 + \sigma^{-2}\lambda_i\eta_i).
+    \end{equation*}
+  %% \item 
+  %%   \begin{equation*}
+  %%     k = \argmax \left \{ k:\lambda_k^{-1} < \sigma^{-2}\frac{m}{k} + \frac{1}{k} \sum_{j=1}^{k}
+  %%     \lambda_j^{-1} \right \}.
+  %%   \end{equation*}
+  \item
+    \begin{equation*}
+        \eta_i = \begin{cases}
+          \frac{m}{k} - \sigma^2 \lambda_i^{-1} + \sigma^2 \frac{1}{k} \sum_{j=1}^k \lambda_j^{-1} & 1 \leq i \leq k \\
+          0 & i > k 
+        \end{cases}.
+    \end{equation*}
+  \item The covariance of the pushforwad \(\fwd_{*} \post\) is \(\left
+    ( (\fwd \prcov \fwd^*)^{-1} + \sigma^{-2} \obs^*\obs \right
+    )^{-1}\) and its eigenvalues are
+    \begin{equation*}
+      \theta_i =
+      \begin{cases}
+        \left(\frac{\sum_{j=1}^k \lambda_j^{-1} + \sigma^{-2}m}{k} \right )^{-1} & i \leq k \\
+        \lambda_i &  i > k 
+      \end{cases}
+    \end{equation*}
+  \end{enumerate}
 \end{theorem}
-
 \begin{proof}
@@
 eigenvectors of $\fwd \prcov \fwd^*$. We present a short discussion
-adapting arguments presented in \cite[Chapter 3]{silvey1980} and
-\cite[Section 5.2.3]{pronzatoPazman2013}.
+adapting arguments presented in \citet[Chapter 3]{silvey1980} and
+\citet[Section 5.2.3]{pronzatoPazman2013}.
 
@@
 \mathcal{M}$.  Recall Carath\'eodory's Theorem:
-\begin{theorem*}[Carath\'eodory]
+\begin{ctheorem}[Carath\'eodory]
   Let $X \subseteq \mathbb{R}^n, X \neq \phi$ and denote $\conv (X)$
@@
   combination of at most $n+1$ vectors in $X$.
-\end{theorem*}
+\end{ctheorem}
 Carath\'eodory's Theorem implies that there exist $\meas_i$ and
@@
 conducting inference on function spaces which are not unique to
-D-optimal designs (see e.g.~\cite{owhadi2015} for more details).
+D-optimal designs (see e.g.~\citet{owhadi2015} for more details).
 
